---
title: "PCMIbackgroundDistributions"
author: "Martin Gordon"
date: "2023-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PCMI get background distributions

goal here to to detemrine the range of the prey-prey distributions
We need to subsample the prey-prey set to ~100 iterations

```{r}
library(data.table)
library(ggplot2)
library(magrittr)


source('~/Documents/utils/bp_utils/ManageScriptData.R')
```
read in the datasets and view

```{r}
af.summary <- fread('./data/2023_11_08_AlphaFoldSummaryScores')
pcmi.ppi <-  fread('./data/D_Final_v3.txt')
```


summary of the data
for the af.scores

```{r}
af.scores %>%  str()
pcmi.ppi %>%  str()

# so, we have bait-prey (~1.6k)
# same number of bait-random prey
# also a prey-prey link (we want to basically subset these)
af.scores[,.N, by=.(cleanedGroup)]

```

```{r}
af.scores %>%  head()

# from here we want Bait and PrwyGene, 
pcmi.baits <- pcmi.ppi$Bait %>%  unique() #100
pcmi.preys <-  pcmi.ppi$PreyGene %>%  unique() # > 1000

# which baits are also preys?
pcmi.preyBaits<- pcmi.baits[which(pcmi.baits %in% pcmi.preys)]
```

prey-prey subset; want to avoid also selecting baits in the prey-prey set?
For now proceed with just prey..

pairs are named alphabetically 

First thought was to just plot for each gene, but the bait=prey set is a sample of 100 

What about randomly selecting 100 genes from gene 1 per subsampling? repeat this 100 times and then draw these plots for each subsampling
Start here first...

```{r}
# have a set of 100 baits, prob with different sizes etc.

prey.oi <- pcmi.preys[!pcmi.preys %in% pcmi.preyBaits]
pp.dt <- af.summary[gene1 %in% prey.oi & gene2 %in% prey.oi,]

# just a santiy check
af.summary[cleanedGroup == 'asd_bait_prey', unique(gene1)] %in% pcmi.baits
af.summary[cleanedGroup == 'asd_bait_prey' & (gene1 %in% pcmi.baits | gene2 %in% pcmi.baits),] # sanity check

# very different distiribtutions
par(mfrow = c(1,2)) # Create a 2 x 2 plotting matrix
af.summary[cleanedGroup == 'asd_bait_prey', .N, by=gene1]$N %>% 
  hist()
af.summary[cleanedGroup == 'asd_prey_prey', .N, by=gene1]$N %>% 
  hist()


# jsut checking number of unique prey genes.. ~20 less did some of the runs fail?
af.summary[cleanedGroup == 'asd_prey_prey', unique(c(gene1,gene2)) %>%  length()]
```
# for first attempt, subsample 100 random prey to use as (pseudo)Baits  and extract the alphafold scores for all their interactions

```{r}
samplePreys <- function(input.dt=pp.dt, af.dt=af.summary, subsample.size=100, ntimes=1, input.colOI='gene1', withReplacement=F) {
  
  sample.dt <- data.table()
  
  for (i in 1:ntimes) {
    # get keys for filtering
    Preys <- sample(input.dt[,get(input.colOI)], subsample.size, replace=withReplacement)
    # sample the preyset 
    s.dt <-  af.dt[cleanedGroup == 'asd_prey_prey' & gene1 %in% Preys, ] 
    #extract cols OI
    s.dt <-  s.dt[, cleanedGroup := paste0(cleanedGroup, '_sample.', i) ] %>% 
      .[, .(pair, cleanedGroup, gene1, gene2, mean_iptm)]
    
    sample.dt <-  rbind(sample.dt, s.dt)
  }
  
  return(sample.dt)
}
```

run 1000 times sampling 100 genes and all their AF interactions
```{r}
# just a test take 100 psuedobaits 10 times and extract all the af predictions for that set
subsampled.pcmi <-  samplePreys(pp.dt, af.summary, subsample.size=50, ntimes=100, input.colOI='gene1', withReplacement=F)


# interesting, in generale each group of 100 genes has many more interactors that the pcmi baits
subsampled.pcmi[,.N,by=cleanedGroup] %>%  summary()
af.summary[,.N,by=cleanedGroup]
```
created subsampled DT for plotting
```{r}
# get all iptm values in the dataset
sampMeanIptm <-  subsampled.pcmi$mean_iptm %>%  unique()

sample.expanded <-  subsampled.pcmi[data.table(cleanedGroup = unique(subsampled.pcmi$cleanedGroup) #dt with 3 cols
                                  )[, .(mean_iptm = sampMeanIptm), by = cleanedGroup], # all cleanedGroups by all mean_confidence
                       , on = c("cleanedGroup", "mean_iptm")]

sample.binned <- sample.expanded[, .(binnedCount =  sum(!is.na(unique(pair)))), by = .(cleanedGroup, mean_iptm)]
sample.binned[order(-mean_iptm), cp := cumsum(binnedCount)/sum(binnedCount), by = .(cleanedGroup)]
sample.binned[cp == 0.0, cp := NA]
```

std plot of prey-prey, bait-prey and bait-random_prey comparison iptm vals

```{r}
allMeanIptm <- af.summary$mean_iptm |> unique()  


expanded <- af.summary[data.table(cleanedGroup = unique(af.summary$cleanedGroup) #dt with 3 cols
                                  )[, .(mean_iptm = allMeanIptm), by = cleanedGroup], # all cleanedGroups by all mean_confidence
                       , on = c("cleanedGroup", "mean_iptm")]


binned <- expanded[, .(binnedCount =  sum(!is.na(unique(pair)))), by = .(cleanedGroup, mean_iptm)]

binned[order(-mean_iptm), cp := cumsum(binnedCount)/sum(binnedCount), by = .(cleanedGroup)]

binned[cp == 0.0, cp := NA]
ggplot (binned, aes(x = mean_iptm, y = cp, color = cleanedGroup)) + geom_line() + scale_y_log10(name = "proportion > threshold")

```
ggplot with the subsampled plots added

```{r}
g <-  ggplot (binned, aes(x = mean_iptm, y = cp, color = cleanedGroup)) + 
  geom_line(data=sample.binned, aes(x = mean_iptm, y = cp, group=cleanedGroup), alpha=0.01, color='black') +
  geom_line() + 
  scale_y_log10(name = "proportion > threshold")
g
BackupAsPDF(g, 'afscores.mean_iptm.linechart.sample100prots.1000.iters.')
```

lets try with 30 genes and 1000 iters for testing
more similiar to the other

```{r}
subsampled.pcmi <-  samplePreys(pp.dt, af.summary, subsample.size=30, ntimes=1000, input.colOI='gene1', withReplacement=F)
subsampled.pcmi[, .N, by=cleanedGroup]

subsampled.pcmi[,.N,by=cleanedGroup] %>%  summary()
af.summary[,.N,by=cleanedGroup]
```
```{r}
# get all iptm values in the dataset
sampMeanIptm <-  subsampled.pcmi$mean_iptm %>%  unique()

sample.expanded <-  subsampled.pcmi[data.table(cleanedGroup = unique(subsampled.pcmi$cleanedGroup) #dt with 3 cols
                                  )[, .(mean_iptm = sampMeanIptm), by = cleanedGroup], # all cleanedGroups by all mean_confidence
                       , on = c("cleanedGroup", "mean_iptm")]

sample.binned <- sample.expanded[, .(binnedCount =  sum(!is.na(unique(pair)))), by = .(cleanedGroup, mean_iptm)]
sample.binned[order(-mean_iptm), cp := cumsum(binnedCount)/sum(binnedCount), by = .(cleanedGroup)]
sample.binned[cp == 0.0, cp := NA]
```

plot the 100 iters sample

```{r}
g <-  ggplot (binned, aes(x = mean_iptm, y = cp, color = cleanedGroup)) + 
  geom_line(data=sample.binned, aes(x = mean_iptm, y = cp, group=cleanedGroup), alpha=0.01, color='black') +
  geom_line() + 
  scale_y_log10(name = "proportion > threshold")
g
BackupAsPDF(g, 'afscores.mean_iptm.linechart.sample30prots.1000.iters.')
```

another sub-sampling approach
rather than taking a random prey-prey interaction, look at the preys for each bait, take one from each and then create a list of 100

update: modified to generate random bait-prey interactions, previously taking first element..
```{r}
af.summary[cleanedGroup == 'asd_bait_prey',] 

# within each bait, select another prey to act as bait (Pseudobait)
# want to order sets by size (reducing size) to 

# first identify the individual Baits and sort based on N prey (this will be the loop)
pcmi.ppi[,.N, by=Bait]
hist(pcmi.ppi[,.N, by=Bait]$N)


# sort Baits based on N prey (this will be our loop iteration)
bait.iter <- pcmi.ppi[,.N, by=Bait] %>% 
  .[order(N), Bait]


sampleUniquePairsfromSets <-  function(bait.vec, input.dt, colKey='Bait', colVal='PreyGene', n.samples=1, verbose=T){
  
  # initialise vec to track Prey values
  prey.id <- character(0)
  # initialise results table
  bp.dt <- data.table()

  # iterate through this list of baits and sample from their sets of preys
  for (bait in bait.vec){
  
    if (verbose %in% c(T, TRUE)){
      print(paste0('Sampling ', bait,'..'))
    }
    # take a random bait
    preySample <- input.dt[get(colKey) == bait, sample(get(colVal), size=n.samples)] 
  
    #check the bait is not already detected
    if (!preySample %in% prey.id) {
      
      if (verbose %in% c(T, TRUE)){
        print(paste0('Prey ', preySample, ' found...'))
      }
      prey.id <- c(prey.id, preySample)
      bp.dt <- rbind(bp.dt, data.table(Key.id = bait, Val.id = preySample))
    
    # if already present resample...
    } else {
      if (verbose %in% c(T, TRUE)){
        print(paste0(preySample, ' already recorded. Resampling ', bait, '...'))
      }
      while (preySample %in% prey.id) {
        # keep sampling until unique prey found
        preySample <- input.dt[get(colKey) == bait, sample(get(colVal), size=n.samples)]
      }
      if (verbose %in% c(T, TRUE)){
        print(paste0('Prey ', preySample, ' found...'))
      }
      prey.id <- c(prey.id, preySample)
      bp.dt <- rbind(bp.dt, data.table(Key.id = bait, Val.id = preySample))
    }
  }
  
  return(bp.dt)
}

# testing
bp.dt <- sampleUniquePairsfromSets(bait.iter, input.dt=pcmi.ppi, n.samples = 1, colKey = 'Bait', colVal='PreyGene', verbose = F)
bp.dt$Val.id %>%  unique()
any(duplicated(bp.dt$Val.id))
```

Now that you have a unique sample of bait-prey pairs, use these prey as (pseudo)bait and pull out interactions
(remove duplicate edges from final output)

```{r}
pseudo.bp.dt <-  data.table()

for (p in bp.dt$Val.id){
  #extract rows where the 
  sub.dt <- af.summary[cleanedGroup == 'asd_prey_prey' & (p == gene1 | p == gene2),]
  pseudo.bp.dt <- rbind(pseudo.bp.dt,sub.dt)
}

pseudo.bp.dt[, cleanedGroup := 'asd_pseudobait_prey']

# remove duplicate rows
pseudo.bp.dt <- pseudo.bp.dt[!duplicated(pair)]
```
captured all prey-prey interactions involving preys that bait targets (pseudoPrey)

```{r}

# get all iptm values in the dataset
sampMeanIptm <-  pseudo.bp.dt$mean_iptm %>%  unique()

sample.expanded <-  pseudo.bp.dt[data.table(cleanedGroup = unique(pseudo.bp.dt$cleanedGroup) #dt with 3 cols
                                  )[, .(mean_iptm = sampMeanIptm), by = cleanedGroup], # all cleanedGroups by all mean_confidence
                       , on = c("cleanedGroup", "mean_iptm")]

sample.binned <- sample.expanded[, .(binnedCount =  sum(!is.na(unique(pair)))), by = .(cleanedGroup, mean_iptm)]
sample.binned[order(-mean_iptm), cp := cumsum(binnedCount)/sum(binnedCount), by = .(cleanedGroup)]
sample.binned[cp == 0.0, cp := NA]


new.binned <-  rbind(binned, sample.binned)
```

plot the new linechart w pseudobait

```{r}
g <-  ggplot (new.binned, aes(x = mean_iptm, y = cp, color = cleanedGroup)) + 
  #geom_line(data=sample.binned, aes(x = mean_iptm, y = cp, group=cleanedGroup), color='purple') +
  geom_line() + 
  scale_y_log10(name = "proportion > threshold")
g
BackupAsPDF(g, 'afscores.mean_iptm.linechart.pseudobait.')
```
repeat the sampling event many times (150x) and plot the distribution of this
Do the repeated resamplings overlap with the other groups, or are they distinct?
```{r}

# can we improve speed here? V slow to subsample all 
ntimes <- 200

afSample.dt <-  data.table()

for (i in 1:ntimes){
  
  print(paste0('Run ',i,' sampling bait-prey dt for unique combos...'))
  bpSample.dt <- sampleUniquePairsfromSets(bait.iter, input.dt=pcmi.ppi, n.samples = 1, colKey = 'Bait', colVal='PreyGene', verbose=F)
  
  #looping through (pseudoprey)list to extract interactions.. faster way to do this?
  for (p in bpSample.dt$Val.id){
    #sampling f
    sub.dt <- af.summary[cleanedGroup == 'asd_prey_prey' & (p == gene1 | p == gene2),]
    sub.dt[, cleanedGroup := paste0('asd_pseudobait_prey_', i)]
    
    afSample.dt <- rbind(afSample.dt,sub.dt)
  }
}

#remove any duplicated rows per sampling
subsamp.dt <- afSample.dt[, .SD[!duplicated(pair),], by=cleanedGroup] 
```
Todo
----
as pair-pair, check for duplicates per sampling event 

```{r}
#backup.dt <- afSample.dt
subsamp.dt[,.N, by=(cleanedGroup)]

# get all iptm values in the dataset
sampMeanIptm <-  subsamp.dt$mean_iptm %>%  unique()

sample.expanded <-  subsamp.dt[data.table(cleanedGroup = unique(subsamp.dt$cleanedGroup) #dt with 3 cols
                                  )[, .(mean_iptm = sampMeanIptm), by = cleanedGroup], # all cleanedGroups by all mean_confidence
                       , on = c("cleanedGroup", "mean_iptm")]

sample.binned <- sample.expanded[, .(binnedCount =  sum(!is.na(unique(pair)))), by = .(cleanedGroup, mean_iptm)]
sample.binned[order(-mean_iptm), cp := cumsum(binnedCount)/sum(binnedCount), by = .(cleanedGroup)]
sample.binned[cp == 0.0, cp := NA]
```


```{r}
g <-  ggplot (binned, aes(x = mean_iptm, y = cp, color = cleanedGroup)) + 
  geom_line(data=sample.binned, aes(x = mean_iptm, y = cp, group=cleanedGroup), color='purple', alpha=0.05) +
  geom_line() + 
  scale_y_log10(name = "proportion > threshold")
g
BackupAsPDF(g, 'afscores.mean_iptm.linechart.pseudobait.200samples.')
```

not used
-----
```{r}
ntimes <- 50

afSample.dt <-  data.table()

for (i in 1:ntimes){
  
  print(paste0('Run ',i,' sampling bait-prey dt for unique combos...'))
  bpSample.dt <- sampleUniquePairsfromSets(bait.iter, input.dt=pcmi.ppi, n.samples = 1, colKey = 'Bait', colVal='PreyGene', verbose=F)

  keyProts <- unique(bpSample.dt$Val.id)
  
  #subset all pseudoBait-prey interactions for this run
  sub.dt <- af.summary[cleanedGroup == 'asd_prey_prey' & (gene1 == keyProts | gene2 == keyProts),] %>% 
    .[, cleanedGroup := paste0('asd_pseudobait_prey_', i)]
  
  afSample.dt <- rbind(afSample.dt,sub.dt)
}

afSample.dt

prac.dt <- data.table()
for (i in bp.dt$Val.id){
  sub.dt <- (af.summary[cleanedGroup == 'asd_prey_prey' & (i == gene1 | i == gene2),])
  prac.dt <- rbind(prac.dt,sub.dt)
}
prac.dt



af.summary[cleanedGroup == 'asd_prey_prey',] %>% 
  .["NUP155" %in% gene1]
```



