---
title: "260923_QLi_DAR34.11.locations"
author: "Martin Gordon"
date: "2023-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

packages

```{r}

library (data.table)
library (ComplexHeatmap)
library (ggplot2)
library (emmeans)
library(stringr)
library(circlize)


rotate.x.axis.text <- theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")

#source Bens utility scripts
source ("/Users/martingordon/Documents/utils/bp_utils/MS_ModelFitting_TimeSeries.R")
source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/MSstats_Helper_Functions.R")
source("/Users/martingordon/Documents/utils/bp_utils/MSstats_V4_Functions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")


source("../../utils/mg_utils/helperFunctions.R") #functions to xtract enriched clusters from the Heatmaps

source("../../utils/bp_utils/Localization_CellMap.R") # functions to enrich for cell pos localisation

WriteSessionInfo()
WriteInstalledPackages()
```


## SAR34.11 location analysis

look at our clusters recovered from the 34.11 clustering of the DEG genes and check for enrichment in each of the different cellular localisations using CellMAP 

read in the interaction analysis data
```{r}
mrx2.drug <- fread('~/Documents/projects/070923_DAR34.11_contrasts/mrx2.drug.anova.csv.gz')
receptor.cz <- fread('~/Documents/projects/070923_DAR34.11_contrasts/receptor.CP.anova.csv.gz')

#inital ts data. Filter hits for the sig proteins
ts.res <- fread('~/Documents/projects/082423_QLi_DARPA.34.11/DARAP.34.11_data/2023_08_23_allFits.csv')

```
protein expression data

```{r}
# including the dropped sample for now
protQuant <- fread ("/Users/martingordon/Documents/projects/082423_QLi_DARPA.34.11/DARAP.34.11_data/2023_08_21_ProteinLevelData.tmp.2round.csv")

# parse GROUP names to define the actual columns that the linear models will work on
# this work depends on how your columsn are defined
protQuant[, receptor   := str_extract(GROUP,"MRX2|NK1R")]
protQuant[, drug := 'SP'] %>% .[GROUP %like% 'MRX2_R[0-9]{2}', drug := 'CZ']
protQuant[, receptor.drug := interaction(receptor,drug)]

#protQuant <- protQuant[receptor != 'NK1R']

# fix subject info to account for batch
protQuant[ drug != 'SP', SUBJECT := SUBJECT + 4 ]
protQuant[, SUBJECT := as.factor(SUBJECT)]
#protQuant[, SUBJECT     := paste0(drug, '.', SUBJECT)] # remember superbatch with the treatment lvl so control for this


protQuant[, timeStr := str_extract(GROUP, "[0-9]{2}$")]
protQuant[, rankTime := as.integer(as.factor(timeStr))] #timeranked to 0-5
protQuant[, .(timeStr, rankTime)] %>% unique() #timepoint to time rank 

protQuant[, .N, by=.(receptor,drug, timeStr,SUBJECT)] %>% 
  .[order(timeStr,drug)]


#vs T0 
protQuant[, vsTimeZero := LogIntensities - mean(LogIntensities[timeStr == timeStr[!is.na(LogIntensities)][1] ]), by = .(Protein, receptor.drug)]

# num of identified proteins per condition
protQuant[, prots.per.condition := .N, by=.(Protein,receptor.drug)] 
```

remove proteins with only one feature as likely miss-identification

```{r}
# remove prots with many na vals per group
na.prots <- protQuant[prots.per.condition <= 6, unique(Protein)] 
#protQuant <- protQuant[NumMeasuredFeature != 1,] # remove proteins with only 1 measured feature
```


```{r}
sigProts <- mrx2.drug[term %in% c('drug:rankTime.2','drug:rankTime.3') & fdr < 0.05, unique(Protein)] 

#just take proteins that as significantly differentially expressed in 
dea.sigProts <- ts.res[ sig != 'not' & receptor.drug %in% c('MRX2.CZ', 'MRX2.SP'), unique(Protein)]
```

only plot these hits for the genes
```{r}
prot.mat <- dcast (protQuant[Protein %in% sigProts & receptor != 'NK1R' & !Protein %in% na.prots, ], Protein~GROUP+SUBJECT, value.var = "vsTimeZero") |> as.matrix(rownames = "Protein")
rownames(prot.mat) <-multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
prot.mat

hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints. vs T0',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 4),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("SP", colnames(prot.mat)), "SP", "CZ")))
hm
BackupAsPDF(hm, 'mrx2.treatment.rmNAsInteractions.heatmap', dimensions=c(8,14))
```
drops to 99 features when removing NAs
Clustering these values w pearson distance into 4 and 5 clusters  and enrich

```{r}
prot.mat <- dcast (protQuant[Protein %in% sigProts & receptor != 'NK1R' & !Protein %in% na.prots, ], Protein~GROUP+SUBJECT, value.var = "vsTimeZero") |> as.matrix(rownames = "Protein")
rownames(prot.mat) <-multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")

set.seed(123)
hm.5clust <- Heatmap(prot.mat, 
              cluster_columns = FALSE, 
              name = 'Log2 Ints. vs T0',
              clustering_distance_rows = 'pearson',
              row_split = 5,
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
             # row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 4),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("SP", colnames(prot.mat)), "SP", "CZ")))

BackupAsPDF(draw(hm.5clust, column_title='Different response to treatment (Interactions fdr < 0.05)'), "pearsonDist.5clusters.heatmap", dimensions = c(8,12))
```
```{r}
set.seed(123)
hm.4clust <- Heatmap(prot.mat, 
              cluster_columns = FALSE, 
              name = 'Log2 Ints. vs T0',
              clustering_distance_rows = 'pearson',
              row_split = 4,
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
             # row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 4),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("SP", colnames(prot.mat)), "SP", "CZ")))

BackupAsPDF(draw(hm.4clust, column_title='Different response to treatment (Interactions fdr < 0.05)'), "pearsonDist.4clusters.heatmap", dimensions = c(8,12))
```
Extract the genes belonging to both clusters and enrich

```{r}
cluster.grps.4 <- extractClustersfromHeatmap(draw(hm.4clust), prot.mat)
cluster.grps.5 <- extractClustersfromHeatmap(draw(hm.5clust), prot.mat)
```

enrich for 4 & 5 clusters
```{r}
gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")
universe <- ts.res$gene %>%  unique()

enrich.dt <- enricherOnGroups(cluster.grps.4, groupColumns = 'cluster', geneColumn = "feature", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'cluster')

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, groupColumn = 'cluster.x', topN = 20, title='GO enrichment clusters', 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 6), upperThreshold =6)

BackupAsPDF(ht, 'go.enrichment.4clusters.')

enrich.dt <- enricherOnGroups(cluster.grps.5, groupColumns = 'cluster', geneColumn = "feature", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'cluster')


ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, groupColumn = 'cluster.x', topN = 20, title='GO enrichment clusters', 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 6), upperThreshold =6)

BackupAsPDF(ht, 'go.enrichment.5clusters.')

View(loadCellMap)
```


Look at all our genes and look for cell localization: based on CellMap HEK cells
This plot will just tell us if the DEGs for each contrast are enriched for different cell locations

```{r}

# create a label col from receptor.drug to work with BP functions 
ts.res[, Label := receptor.drug]

# drop the other receptor as not interested atm
#dea.res <- ts.res[receptor.drug !=  "NK1R.SP",]
dea.res <- ts.res

results.list <- cellMapLocalizationScores(dea.res) # instead of calling loadCellMap function, just read from file 

p <- violinsAndScatterLocations(dea.res, xlimits=c(-4,4))
p
#BackupAsPDF(p, 'allrecptor.drug.allsig.localizationEnrichment')


# generate our own plot of the sig enrichment results (maybe also think of adding NES score for plotting)

# get a list of sig locations in either condition
sig.locations <- dea.res[!is.na(location) & abs(sea.sigScore) > 1.3, unique(location)]


p <- ggplot (dea.res[!is.na(location) & location %in% sig.locations,],
               aes(y = location, x = log2FC, fill = sea.sigScore )) +
    geom_vline(xintercept = 0) +
    geom_jitter(alpha = 0.1, width = 0.0, height = 0.1) +
    geom_violin(lwd = 0.2) + 
    facet_grid(~receptor.drug) +
    #scale_x_continuous(limits = c(-2,2)) +
    coord_cartesian(xlim = c(-4,4)) + 
    scale_fill_gradient2(low = "blue", high = "red") +

    theme_bw()
p
BackupAsPDF(p, 'mrx2.drug.siglocations.localizationEnrichment')
```
Visualise these results as a heatmap

```{r}
# drop the other receptor as not interested atm
#dea.res <- ts.res[receptor.drug !=  "NK1R.SP",]
dea.res <- ts.res

sig.locations <- dea.res[!is.na(location) & abs(sea.sigScore) > 1.3, unique(location)]


sea.mat <- summarizesSEAMatrixFromScoresDT(dea.res)

sea.mat[rownames(sea.mat) %in% sig.locations,]
  
hm <- Heatmap(sea.mat,
        name = 'SEA enrichment score',
        cluster_columns = F,
        na_col = 'grey',
        row_names_gp  = gpar(fontsize= 8),
         col = colorRamp2(c(-10,0,10), c('blue', 'white', 'red')),
         #column_split = ifelse(grepl('SP',colnames(sea.mat)), "SP", "CZ")
        column_split = colnames(sea.mat)
        )
hm
BackupAsPDF(hm, 'allreceptors.location.enrichment.scores.heatmap')


# only plot any locations with sig results in one of the datasets

```

now generate heatmaps of only sig locations

```{r}
# drop the other receptor as not interested atm
dea.res <- ts.res[receptor.drug !=  "NK1R.SP",]
#dea.res <- ts.res


sig.locations <- dea.res[!is.na(location) & abs(sea.sigScore) > 1.3, unique(location)]

sea.mat <- summarizesSEAMatrixFromScoresDT(dea.res)

sea.mat <- sea.mat[rownames(sea.mat) %in% sig.locations,]
  
hm <- Heatmap(sea.mat,
        name = 'SEA enrichment score',
        cluster_columns = F,
        na_col = 'grey',
        row_names_gp  = gpar(fontsize= 8),
         col = colorRamp2(c(-10,0,10), c('blue', 'white', 'red')),
         #column_split = ifelse(grepl('SP',colnames(sea.mat)), "SP", "CZ")
        column_split = colnames(sea.mat)
        )
hm
BackupAsPDF(hm, 'mrx2.siglocations.enrichment.scores.heatmap')


# only plot any locations with sig results in one of the datasets

```
replot violin plots; color by NES and adjust the x-axis by subtracting NES from log2FC
color by NES or median log2FC per group/cell location
then for each value plotted subtract NES from log2FC  (why?)

```{r}
# change col names in sea output to allow merging
setnames(results.list$sea.dt, old=colnames(results.list$sea.dt), new=c("receptor.drug", "location", "pval", "padj", "log2err","ES","NES","size","leadingEdge","sigScore"  ))

# merge the cols in the dt
results.list$scores.dt[results.list$sea.dt, NES := i.NES, on=.(receptor.drug, location)]


#calculate log2FC shift vs location NES value per prot/label/location
sea.dt <- results.list$scores.dt[!is.na(location), sum.log2FC.NES := NES+log2FC, by=.(location,Label,Protein)] %>% 
  .[!is.na(location), .(Label, Protein, gene, log2FC, pvalue, adj.pvalue, location, sea.sigScore, NES, sum.log2FC.NES)]

# c
sea.dt[, median.log2FC := median(log2FC, na.rm=T), by=.(location,Label)]

```
plot the results 

```{r}
p <- ggplot(sea.dt, aes(x=log2FC,y=location, fill=sea.sigScore)) + 
         geom_jitter(alpha = 0.1, width = 0.0, height = 0.1) +
         geom_violin(lwd= 0.2) +
         geom_vline(xintercept = 0) +
         coord_cartesian(xlim = c(-5,5)) + 
         scale_fill_gradient2(low = "blue", high = "red") +
         facet_wrap(~Label) +
         theme_bw() +
         theme(axis.text.y = element_text(size=6))

BackupAsPDF(p, 'cellMap.location.enrichment.allGroups.allLocaitons', dimensions=c(8,10))


p <- ggplot(sea.dt[Label != 'NK1R.SP',], aes(x=log2FC,y=location, fill=sea.sigScore)) + 
         geom_jitter(alpha = 0.1, width = 0.0, height = 0.1) +
         geom_violin(lwd= 0.2) +
         geom_vline(xintercept = 0) +
         coord_cartesian(xlim = c(-5,5)) + 
         scale_fill_gradient2(low = "blue", high = "red") +
         facet_wrap(~Label) +
         theme_bw() +
         theme(axis.text.y = element_text(size=6))


BackupAsPDF(p, 'cellMap.location.enrichment.MRX2.allLocations', dimensions=c(8,10))
```
plot the log2fc x axis and sea.sig scores for the significant results for both

```{r}
sig.locations <- ts.res[!is.na(location) & abs(sea.sigScore) > 1.3, unique(location)]


p <- ggplot(sea.dt[location %in% sig.locations], aes(x=log2FC,y=location, fill=sea.sigScore)) + 
         geom_jitter(alpha = 0.1, width = 0.0, height = 0.1) +
         geom_violin(lwd= 0.2) +
         geom_vline(xintercept = 0) +
         coord_cartesian(xlim = c(-5,5)) + 
         scale_fill_gradient2(low = "blue", high = "red") +
         facet_wrap(~Label) +
         theme_bw() +
         theme(axis.text.y = element_text(size=6))

BackupAsPDF(p, 'cellMap.location.enrichment.allGroups.sigLocations', dimensions=c(9,6))

sig.locations <- ts.res[!is.na(location) &  Label != 'NK1R.SP' & abs(sea.sigScore) > 1.3, unique(location)]

p <- ggplot(sea.dt[location %in% sig.locations & Label != 'NK1R.SP',], aes(x=log2FC,y=location, fill=sea.sigScore)) + 
         geom_jitter(alpha = 0.1, width = 0.0, height = 0.1) +
         geom_violin(lwd= 0.2) +
         geom_vline(xintercept = 0) +
         coord_cartesian(xlim = c(-5,5)) + 
         scale_fill_gradient2(low = "blue", high = "red") +
         facet_wrap(~Label) +
         theme_bw() +
         theme(axis.text.y = element_text(size=6))

BackupAsPDF(p, 'cellMap.location.enrichment.MRX2.sigLocations', dimensions=c(9,6))


```


now only plot the significant results 
what is this value telling us?  NES (per grp) - log2FC

```{r}
sig.locations <- ts.res[!is.na(location) & abs(sea.sigScore) > 1.3, unique(location)]


p <- ggplot(sea.dt[location %in% sig.locations], aes(x=sum.log2FC.NES,y=location, fill=sea.sigScore)) +
         geom_jitter(alpha = 0.1, width = 0.0, height = 0.1) +
         geom_violin(lwd= 0.2) +
         geom_vline(xintercept = 0, alpha=0.5) +
         #coord_cartesian(xlim = c(-5,5)) + 
         scale_fill_gradient2(low = "blue", high = "red") +
         xlab('Log2FC + NES') +
         facet_wrap(~Label) +
         theme_bw() +
         theme(axis.text.y = element_text(size=6))
p
BackupAsPDF(p, 'cellMap.location.enrichment.allGroups.FC.NES.sigLocations', dimensions=c(10,5))

sig.locations <- ts.res[!is.na(location) &  Label != 'NK1R.SP' & abs(sea.sigScore) > 1.3, unique(location)]


p <- ggplot(sea.dt[Label != 'NK1R.SP' & location %in% sig.locations,], aes(x=sum.log2FC.NES,y=location, fill=sea.sigScore)) + 
         geom_jitter(alpha = 0.1, width = 0.0, height = 0.1) +
         geom_violin(lwd= 0.2) +
         geom_vline(xintercept = 0, alpha=0.5) +
        # coord_cartesian(xlim = c(-5,5)) + 
         scale_fill_gradient2(low = "blue", high = "red") +
         facet_wrap(~Label) +
         xlab('Log2FC + NES') +
         theme_bw() +
         theme(axis.text.y = element_text(size=6))
p

BackupAsPDF(p, 'cellMap.location.enrichment.MRX2.FC.NES.sigLocations', dimensions=c(10,5))

```
write out table of the sig hits in each 

```{r}
#fwrite(results.list$sea.dt, './output/cellMaplocationenrichment.csv')
sea.dt <- fread('./output/cellMaplocationenrichment.csv')
```

03-03-23
----
Redo the gsea enrichment analysis plot faceting by location instead of treatment

```{r}
# create a new colored value of threshold
sea.dt[, SEA.sig := sea.sigScore]
sea.dt[SEA.sig > 10 | SEA.sig < -10, SEA.sig := ifelse(SEA.sig > 0, 10, -10)]

p <- ggplot(sea.dt[location %in% sig.locations], aes(x=Label,y=sum.log2FC.NES, fill=SEA.sig)) +
         geom_jitter(alpha = 0.1, width = 0.0, height = 0.1) +
         geom_violin(lwd= 0.2) +
         geom_hline(yintercept = 0, alpha=0.5, linetype=2) +
        # coord_cartesian(ylim = c(-5,5)) + 
         scale_fill_gradient2(low = "blue", mid='white', high = "red") +
         xlab('Group') +
         ylab('Log2FC + NES') +
         facet_wrap(~location, scales = 'free') +
         theme_bw() +
         theme(axis.text.y = element_text(size=6)) 
p
BackupAsPDF(p, 'cellMap.location.enrichment.allGroups.sigLocationsFacet.FC.NES.colthreshold', dimensions=c(10,5))

p <- ggplot(sea.dt[location %in% sig.locations & Label != 'NK1R.SP'], aes(x=Label,y=sum.log2FC.NES, fill=SEA.sig)) +
         geom_jitter(alpha = 0.1, width = 0.0, height = 0.1) +
         geom_violin(lwd= 0.2) +
         geom_hline(yintercept = 0, alpha=0.5, linetype=2) +
         scale_fill_gradient2(low = "blue", mid='white', high = "red") +
         xlab('Group') +
         ylab('Log2FC + NES') +
         facet_wrap(~location, scales = 'free') +
         theme_bw() +
         theme(axis.text.y = element_text(size=6)) 
p
BackupAsPDF(p, 'cellMap.location.enrichment.MRX2Groups.sigLocationsFacet.FC.NES.coltrheshold', dimensions=c(10,5))
```




