---
title: "061724_AViDD_WGCNA_rerun"
author: "Martin Gordon"
date: "2024-06-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## AViDD dataset rerun WGCNA processing

```{r}
library(magrittr)
library(data.table)
library(circlize)
library(ComplexHeatmap)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(ggrepel)
library(readxl)
library(viridis)
library(viridis)
library(usedist) #package for working with distances
library(ggh4x) # additional functionality for ggplot2 obj
library(scales)
library(seqinr)
library(WGCNA)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

source ("../../utils/mg_utils/r_utils/IDmapping.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```

Read in the normalised data and ms out results

```{r}
mss.dt  <- fread('~/Documents/projects/052924_ARichards_AViDD_Murine/061324_AViDD_AB_rerun_data/2024_06_18_mss.pwcontrasts.cleanDat.csv')
p.quant <- fread('~/Documents/projects/052924_ARichards_AViDD_Murine/061324_AViDD_AB_rerun_data/2024_06_17_svaCorrected.pquant.csv')
setnames(p.quant, new=c('Protein', 'sample', 'LogIntensities'))
```

Map to gene IDs
```{r}
# want to tidy this function to automatically map IDs to datatable
fetchSomethingFromBioMart(ids = p.quant$Protein, 
                          species='MOUSE',
                          attributes=c("ensembl_gene_id", "ensembl_peptide_id_version", "external_gene_name"),
                          filters="ensembl_peptide_id_version")

# so this works to filter the datatable 
IDmapping.dt <- fetchSomethingFromBioMart(ids = p.quant$Protein, species='MOUSE', attributes=c('external_gene_name', 'ensembl_peptide_id_version'), filters = 'ensembl_peptide_id_version')

p.quant <- merge(x=p.quant, y=IDmapping.dt, by.x='Protein', by.y='ensembl_peptide_id_version', all.x=T)
p.quant[, gene := ifelse(is.na(external_gene_name), Protein, external_gene_name)]
```

drop samples not used in the DE analysis due to noise
convert to matrix 
5222 proteins retained in the matrix
Dropped samples: "Nirmatrelvir_4d.1","Vehicle_7d.2","x4516_7d.1","x4516_7d.3"
```{r}
p.mat <- dcast(p.quant[!sample %in% c("Nirmatrelvir_4d.1","Vehicle_7d.2","x4516_7d.1","x4516_7d.3"),], gene~sample, value.var = 'LogIntensities') %>% 
  as.matrix(rownames='gene')
```

# subtract Mock expression values and look at some other contrasts
```{r}
p.quant[,c("status", "rep") := tstrsplit(sample, "[.]", keep = c(1,2)) ]
p.quant[, c("treatment", "timepoint") := tstrsplit(status, '[_]', keep=c(1,2)) ]
p.quant[status == 'Mock', timepoint := 'NA']

# calculate difference vsMock
p.quant[, vsMock := LogIntensities - mean(LogIntensities[status == 'Mock'], na.rm=T), by=.(Protein)]

fwrite(p.quant, ScriptAndDatedFileName('input.ProteinInts.clean.csv'))
```

```{r}
p.quant <- fread('/Users/martingordon/Documents/projects/052924_ARichards_AViDD_Murine/061724_AViDD_AB_WGCNA_rerun_data/2024_06_18_input.ProteinInts.clean.csv')

# drop outlier samples
p.quant <- p.quant[!sample %in% c("Nirmatrelvir_4d.1","Vehicle_7d.2","x4516_7d.1","x4516_7d.3"),]
```
Looks ok.. proceed as is 
```{r}
ggplot(p.quant, aes(x=paste0(status,'.',rep), y=LogIntensities, fill=status)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=90))
```
Get matrix of protein intensity values
```{r}
p.mat <- dcast(p.quant, Protein~paste0(status,'.',rep), value.var = 'vsMock') %>% 
  as.matrix(rownames='Protein')
```

Question: should we filter only to genes that are DE? 
maybe leave as is for now and can retry

```{r}
hm <- Heatmap(p.mat, 
              cluster_rows = T, 
              cluster_columns = F,
              name='vsMock',
              na_col = 'darkgrey',
              column_split = gsub('.[0-9]+$', '', colnames(p.mat)), 
              row_title=sprintf('%s proteins', nrow(p.mat)), 
              show_row_names = F, 
              column_names_gp = gpar(fontsize=5), 
              column_title_gp = gpar(fontsize = 8))

BackupAsPDF(draw(hm, column_title='WGCNA input'), 'wgcna.input.mat')
```
filter the p.quant dataset by the set of nominally significant results
```{r}
prots.oi <- mss.dt[pvalue < 0.05, unique(Protein)]
length(prots.oi)
```

```{r}
p.mat <- dcast(p.quant[Protein %in% prots.oi,], Protein~paste0(status,'.',rep), value.var = 'vsMock') %>% 
  as.matrix(rownames='Protein')
```

Question: should we filter only to genes that are DE? 
maybe leave as is for now and can retry

```{r}
hm <- Heatmap(p.mat, 
              cluster_rows = T, 
              cluster_columns = F,
              name='vsMock',
              na_col = 'darkgrey',
              column_split = gsub('.[0-9]+$', '', colnames(p.mat)), 
              row_title=sprintf('%s proteins', nrow(p.mat)), 
              show_row_names = F, 
              column_names_gp = gpar(fontsize=5), 
              column_title_gp = gpar(fontsize = 8))

BackupAsPDF(draw(hm, column_title='WGCNA input'), 'wgcna.input.mat', dimensions = c(12,8))
```
Try another round of tmp normalisation first and lets look at the output
Looks better, but worth the extra round of correction

```{r}
tmp <- medpolish(p.mat)

tmp.dt <- data.table(sample = names(tmp$col),
                     offset = tmp$col)

p.quant <- merge(p.quant, y=tmp.dt, by.x='sample', by.y='sample', all.x=T)
p.quant[,tmp.LogIntensities := LogIntensities - offset]
p.quant[, tmp.vsMock := tmp.LogIntensities - mean(tmp.LogIntensities[status == 'Mock'], na.rm=T), by=.(Protein)]


ggplot(p.quant, aes(x=paste0(status, '.', rep), y=tmp.LogIntensities, fill=status)) +
  geom_boxplot()
```

```{r}
p.mat <-dcast(p.quant[Protein %in% prots.oi,], Protein~sample, value.var = 'tmp.vsMock') %>% 
  as.matrix(rownames='Protein')

#p.mat <- sweep(p.mat, 1, apply(p.mat, 1, median, na.rm=T))

hm <- Heatmap(p.mat, 
              cluster_rows = T, 
              cluster_columns = F,
              name='vsMock',
              na_col = 'darkgrey',
              column_split = gsub('.[0-9]+$', '', colnames(p.mat)), 
              row_title=sprintf('%s proteins', nrow(p.mat)), 
              show_row_names = F, 
              column_names_gp = gpar(fontsize=5), 
              column_title_gp = gpar(fontsize = 8))

BackupAsPDF(draw(hm, column_title='WGCNA input'), 'wgcna.input.tmpNorm.mat', dimensions = c(12,8))
```
I think the tmp norm set looks much better.. .rerun the WGCNA on this then return to the DE analysis and run before executing the workflow

```{r}
submat <- p.mat[, grep('Mock', colnames(p.mat), invert=T)]

hm <- Heatmap(submat, 
              cluster_rows = T, 
              cluster_columns = F,
              name='vsMock',
              na_col = 'darkgrey',
              column_split = gsub('.[0-9]+$', '', colnames(submat)), 
              row_title=sprintf('%s proteins', nrow(submat)), 
              show_row_names = F, 
              column_names_gp = gpar(fontsize=5), 
              column_title_gp = gpar(fontsize = 8))

BackupAsPDF(draw(hm, column_title='WGCNA input'), 'wgcna.input.noMock.mat', dimensions=c(14,12))
```

Look at the sample clustering 
```{r}
#Group data in a dendogram to check outliers
sampleTree = hclust(dist(t(submat)), method = "average")

pdf(file = "/Users/martingordon/Documents/projects/052924_ARichards_AViDD_Murine/061724_AViDD_AB_WGCNA_rerun_data/pdfs/sampleClustering.pdf", width = 12, height = 9); 
par(cex = 0.6)
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering", sub="", xlab="", cex.lab = 1.5, 
     cex.axis = 1.5, cex.main = 2)
#Plot a line showing the cut-off
abline(h = 56, col = "red") #This value of 31000 was chosen based on my data, you need to check the best value to your data

dev.off()
```
Now that we have looked at the samples, lets checkout the genes we have selected
```{r}
gsg <- goodSamplesGenes(t(submat))
summary(gsg)
gsg$allOK
```
```{r}
net.in <- t(submat)

s.thres <- data.table(pickSoftThreshold(net.in)$fitIndices)

s.thres %>%  
  ggplot(aes(x=Power, y=SFT.R.sq)) + 
  geom_point() +
  geom_hline(yintercept = 0.8, color='darkred', linetype=2) +
  geom_hline(yintercept = 0.9, color='darkred', linetype=2) +
  labs(y="Scale Free Topology Model Fit, signed R^2") +
  theme_classic()

s.thres %>%  
  ggplot(aes(x=Power, y=mean.k.)) + 
  geom_point() +
  labs(y='Mean Connectivity') +
  theme_classic() 
```
softThreshold 4-5 seems like a decent choice for this..

```{r}
softPower = 4 #Chosen in the graphs before
adj.mat = adjacency(net.in, power = softPower, corFnc = 'bicor', type = "signed") #Calculating the adjacency matrix
#help(adjacency )
```

Define our topological overlap matrix
```{r}
TOM <- TOMsimilarity(adj.mat)
TOM.dissimilarity <- 1-TOM

#creating the dendrogram 
geneTree <- hclust(as.dist(TOM.dissimilarity), method = "average") 
#plotting the dendrogram
sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity", 
labels = FALSE, hang = 0.04)
```
At a quick snapshot seems like some nice clustering visible and also quite distinct to other 
```{r}
hm <- Heatmap(TOM[1:2000,1:2000], cluster_rows = T, cluster_columns=T, name='similarity', column_title = 'TOM similarity matrix', show_row_names = F, show_column_names = F)
BackupAsPDF(hm, 'tom.mat.st4.')
```

Detected 20 modules
Lets look at these and plot the average expression of each of these
```{r}
Modules <- cutreeDynamic(dendro = geneTree, distM = TOM.dissimilarity, deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = 30)
table(Modules)
# make a datatable of the module assignments
modules.dt <- data.table(Protein=rownames(adj.mat),
                         Module=as.character(Modules))

modules.dt <- merge(modules.dt, IDmapping.dt, by.x='Protein', by.y='ensembl_peptide_id_version', all.x=T)
setnames(modules.dt, new=c('Protein', 'Module', 'gene'))

fwrite(modules.dt, ScriptAndDatedFileName('tmpNorm.sp4.module.assignments.deepSplit2.csv'))
```

```{r}

# get color palette for the modules
col.pal <- c(
  "dodgerblue2", 
  "#E31A1C", # red
  "green4",
  "#6A3D9A", # purple
  "#FF7F00", # orange
  "black", "gold1",
  "skyblue2", "#FB9A99", # lt pink
  "palegreen2",
  "#CAB2D6", # lt purple
  "#FDBF6F", # lt orange
  "gray70", "khaki2",
  "maroon", "orchid1", "deeppink1", "blue1", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3",
  "darkorange4", "brown"
)

modules.tom <- as.character(Modules)

# add wgcna module annotation
row_ha <- rowAnnotation(wgcna.modules=modules.tom, col=list(modules.tom = c("1"=col.pal[1], '2'=col.pal[2], '3'=col.pal[3], '4'=col.pal[4], '5'=col.pal[5], 
                                                                              '6'=col.pal[6], '7'=col.pal[7], '8'=col.pal[8], '9'=col.pal[9], '10'=col.pal[10],
                                                                              '11'=col.pal[11], '12'=col.pal[12], '13'=col.pal[13])))

# add split order
split <- factor(modules.tom, levels=c('1','2','3','4','5','6','7','8','9','10','11','12','13'))

hm <- Heatmap(submat, 
              cluster_columns = FALSE, 
              show_row_names = FALSE,
              split = split,
              right_annotation = row_ha,
              name = "Ints vs Mock",
              column_split = tstrsplit(colnames(submat), "[.]")[[1]],
              na_col = 'darkgrey',
              row_title =sprintf("%d Proteins",
                            nrow(submat)),
              row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize= 3),
              cluster_row_slices = F,
              column_title_gp = gpar (fontsize= 7))
hm
hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))

BackupAsPDF(hm, 'heatmap.wgcna.modules.ordered',dimensions=c(12,16))
```
```{r}
# safe color pallete from https://stackoverflow.com/questions/9563711/r-color-palettes-for-many-data-classes

col.subset <- c(col.pal[1], col.pal[2], col.pal[3], col.pal[4], col.pal[5], col.pal[15], col.pal[7], col.pal[8], col.pal[9], col.pal[10], col.pal[11],col.pal[12],col.pal[13])

modules.tom <- as.character(Modules)

# add wgcna module annotation
row_ha <- rowAnnotation(wgcna.modules=modules.tom, col=list(wgcna.modules = c("1"=col.pal[1], '2'=col.pal[2], '3'=col.pal[3], '4'=col.pal[4], '5'=col.pal[5], 
                                                                              '6'=col.pal[15], '7'=col.pal[7], '8'=col.pal[8], '9'=col.pal[9], '10'=col.pal[10],
                                                                              '11'=col.pal[11], '12'=col.pal[12], '13'=col.pal[13])))

# add split order
split <- factor(modules.tom, levels=c('"1','2','3','4','5','6','7','8','9','10','11','12','13'))

hm <- Heatmap(submat, 
              cluster_columns = FALSE, 
              split = split,
              right_annotation = row_ha,
              name = "Ints vs Mock",
              column_split = tstrsplit(colnames(submat), "[.]")[[1]],
              na_col = 'darkgrey',
              row_title =sprintf("%d Proteins",
                            nrow(submat)),
              row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize= 3),
              cluster_row_slices = T,
              show_row_names = F,
              column_title_gp = gpar (fontsize= 7))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))

BackupAsPDF(hm, 'heatmap.wgcna.modules.sigGnes.ordered',dimensions=c(12,16))
```

Lets subset to modules we are interested in 
Run GO enrichment to identify interesting modules 

```{r}
# load the GO table
gmt.go <- loadGmtFromBioconductor(dbName = 'org.Mm.eg.db', ontology = "ALL", keyType = "SYMBOL")
```

```{r}
# our background? only genes used for WGCNA analysis
universe <- unique(modules.dt$gene)

enrich.dt <- enricherOnGroups(modules.dt, groupColumns = 'Module', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'Module')

simp.enrich$simplified[, Module := factor(Module, levels=c('1','2','3','4','5','6','7','8','9','10','11','12','13'))]

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, groupColumn = 'Module', topN = 8, title='GO enrichment WGCNA modules', 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 6), upperThreshold = 8, cluster_columns =F)

BackupAsPDF(ht, 'go.enrichment.wgcna.modules.sigGenes.ordered', dimensions=c(14,12))
```
merge the module and p.quant tables 

```{r}
p.quant <- merge(p.quant[Protein %in% prots.oi,], modules.dt[,-c('gene')], by='Protein')
```

Create a top annotation for the modules
```{r}

p.quant[, moduleMean := mean(tmp.vsMock), by=.(sample,Module)]

# get the average per module/sample
sumMat <- dcast(p.quant, Module~sample, value.var = 'tmp.vsMock', fun.aggregate = mean) %>% 
  as.matrix(rownames='Module')

rownames(sumMat) <- factor(rownames(sumMat), levels=c('1','2','3','4','5','6','7','8','9','10','11','12','13'))

mod.order <- c('1','2','3','4','5','6','7','8','9','10','11','12','13')

sumMat <- sumMat[mod.order, grep('Mock', colnames(sumMat), invert = T, value = T)]


topAnnotation <- HeatmapAnnotation(moduleAverage = anno_lines(t(sumMat),
                                   gp = gpar(col = col.subset),
                                   add_points = TRUE, 
                                   pt_gp = gpar(col = col.subset)),  
                                   which = "column",
                                   height = unit(5, "cm"))

hm <- Heatmap(submat, 
              cluster_columns = FALSE, 
              right_annotation = row_ha,
              top_annotation = topAnnotation,
              name = "Ints vs Mock",
              na_col = 'darkgrey',
              split = split,
              row_title =sprintf("%d Proteins",
                            nrow(submat)),
              row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize= 6),
              cluster_row_slices = F,
              show_row_names = F,
              column_title_gp = gpar (fontsize= 7))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))

BackupAsPDF(hm, 'heatmap.wgcna.modules.sigGnes.ordered',dimensions=c(22,22))   
```
Name the clusters 
```{r}

# now need to name these module genes according to their enrichment profiles
modules.dt[, enrichModules := factor(ifelse(Module %in% c('9','10','13'), Module, 
                                      ifelse(Module == '1', '1 (Organic acid metabolic processes)',
                                             ifelse(Module == '2', '2 (Transmembrane receptor protein kinase activity)',
                                                    ifelse(Module == '3', '3 (Translation)',
                                                      ifelse(Module == '4', '4 (MHC protein complex)',       
                                                           ifelse(Module == '5', '5 (Immunoglobulin complex)',
                                                                ifelse(Module == '6', '6 (Response to interferon-beta)',
                                                                  ifelse(Module == '7', '7 (Oxidative phosphorylation)',
                                                                         ifelse(Module == '8', '8 (Endopeptidase inhibitor activity)',
                                                                                ifelse(Module == '11', '11 (Endosome)',
                                                                                       ifelse(Module == '12', '12 (ATP-dependent activity)',
                                                                                       ''))))))))))) )]
# sanity check
modules.dt[,.N, by=.(Module,enrichModules)]


modules.tom <- as.character(modules.dt$enrichModules)

# add split order
split <- factor(modules.tom, levels=c('1 (Organic acid metabolic processes)', '2 (Transmembrane receptor protein kinase activity)', '3 (Translation)',
                                      '4 (MHC protein complex)','5 (Immunoglobulin complex)', '6 (Response to interferon-beta)', '7 (Oxidative phosphorylation)',
                                      '8 (Endopeptidase inhibitor activity)', '9', '10', '11 (Endosome)','12 (ATP-dependent activity)', '13'))

# add wgcna module annotation
row_ha <- rowAnnotation(wgcna.modules=modules.tom, 
                        col=list(wgcna.modules = c('1 (Organic acid metabolic processes)'=col.pal[1], '2 (Transmembrane receptor protein kinase activity)'=col.pal[2], '3 (Translation)'=col.pal[3],
                                      '4 (MHC protein complex)'=col.pal[4],'5 (Immunoglobulin complex)'=col.pal[5], '6 (Response to interferon-beta)'=col.pal[6], '7 (Oxidative phosphorylation)'=col.pal[7],
                                      '8 (Endopeptidase inhibitor activity)'=col.pal[8], '9'=col.pal[9], '10'=col.pal[10], '11 (Endosome)'=col.pal[11],'12 (ATP-dependent activity)'=col.pal[12], '13'=col.pal[13])))

hm <- Heatmap(submat, 
              cluster_columns = FALSE, 
              split = split,
              cluster_row_slices = F,
              right_annotation = row_ha,
              name = "Ints vs Mock",
              column_split = tstrsplit(colnames(submat), "[.]")[[1]],
              na_col = 'darkgrey',
              row_title =sprintf("%d Proteins",
                            nrow(submat)),
              row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize= 3),
              show_row_names = F,
              column_title_gp = gpar (fontsize= 7))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))

hm
BackupAsPDF(hm, 'heatmap.wgcna.modules.sigGnes.anno.ordered',dimensions=c(22,22))  
```
Calculate the moduleEigengenes/PC1 of the data
```{r}
# Get Module Eigengenes per cluster
MEs0 <- moduleEigengenes(t(submat), modules.dt$Module)$eigengenes
colnames(MEs0) <- paste0('ME',levels(modules.dt$enrichModules))
```

Heatmap of the module eigenProteins
```{r}
# create annotation
ann <- data.table(treatment=sub("_.+", "", colnames(t(MEs0))),
                  timepoint=ifelse(grepl("[247]d", colnames(t(MEs0))), str_extract(colnames(t(MEs0)), "[247]d"), 'na'))

colours <- list('timepoint' = c('na' = '#4477AA', '2d'='#66CCEE', '4d'="#228833", '7d'="#CCBB44"),
                'treatment' =c('Nirmatrelvir' = col.pal[1], 'Vehicle'=col.pal[2], 'x4516'=col.pal[3], 'x4673'=col.pal[4]))

colAnn <- HeatmapAnnotation(df = ann, col = colours)

hm <- Heatmap(t(MEs0),
        name='module PCs',
        top_annotation = colAnn,
        column_names_gp = gpar(fontsize=6),
        row_names_gp = gpar(fontsize=7),
        column_split = gsub('_.+','',colnames(t(MEs0))),
        #column_split = gsub('.[1-6]$','',colnames(t(MEs0))),
        cluster_columns = FALSE)
hm
BackupAsPDF(hm, 'ME.PC1.heatmap',dimensions=c(12,6))
```
Lets collapse our input matrix to mean per group and then plot boxplots of the distributions of the genes in that module

```{r}
fmat.dt <- setDT(reshape2::melt(submat, value.name = 'LogInts.vs.Mock'))
setnames(fmat.dt, new=c('Protein', 'sample', 'LogInts.vs.Mock'))

fmat.dt[, GROUP := gsub('[.][1-6]$','',sample)]

# convert to aggregated matrix
agg.mat <- dcast(fmat.dt, Protein~GROUP, value.var = 'LogInts.vs.Mock', fun.aggregate = mean, na.rm=T) %>% 
  as.matrix(rownames='Protein')
```
ok, now want to seperate by timepoint (or order by timepoint) and then look at the distributions of the interesting modules as boxplots
Also, share the files with the enrichment results for each module and the protein assignments

```{r}
# add boxplot annotation of each of module ME
topAnnotation <- HeatmapAnnotation('2 (Transmembrane receptor protein kinase activity)' = anno_boxplot(agg.mat[rownames(agg.mat) %in% modules.dt[Module == '2', Protein],],
                                   gp = gpar(fill = col.pal[2])),
                                   '5 (Immunoglobulin complex)' = anno_boxplot(agg.mat[rownames(agg.mat) %in% modules.dt[Module == '5', Protein],],
                                   gp = gpar(fill = col.pal[5])),
                                  '8 (Endopeptidase inhibitor activity)' = anno_boxplot(agg.mat[rownames(agg.mat) %in% modules.dt[Module == '5', Protein],],
                                   gp = gpar(fill = col.pal[8])),
                                   treatment = gsub('_.+$','', colnames(agg.mat)),
                                   which = "column",
                                   height = unit(15, "cm"))

hm <- Heatmap(agg.mat, 
        column_split = str_extract(colnames(agg.mat), '[247]d'),
        show_row_names = F,
        name='Ints. vs Mock',
        top_annotation = topAnnotation,
        split = split,
        cluster_row_slices = F,
        right_annotation = row_ha,
        row_title = sprintf('%s Proteins', nrow(agg.mat)),
        cluster_columns = F)

hm
BackupAsPDF(hm, 'heatmap.wgcna.modules.sigGnes.moduleBoxplots.anno.ordered',dimensions=c(23,22)) 


# treatment split
hm <- Heatmap(agg.mat, 
        column_split = gsub('_[247]d','', colnames(agg.mat)),
        show_row_names = F,
        name='Ints. vs Mock',
        top_annotation = topAnnotation,
        split = split,
        cluster_row_slices = F,
        right_annotation = row_ha,
        row_title = sprintf('%s Proteins', nrow(agg.mat)),
        cluster_columns = F)

hm
BackupAsPDF(hm, 'heatmap.wgcna.modules.sigGnes.moduleBoxplots.treatSplit.anno.ordered',dimensions=c(23,22)) 

MEs0
```

Lets split this heatmap differently and plot the MEs of the interesting modules at the top

```{r}
# just subset to the interesting modules
modules.oi <- c('ME2 (Transmembrane receptor protein kinase activity)', 'ME5 (Immunoglobulin complex)','ME8 (Endopeptidase inhibitor activity)')

subME <- MEs0[, modules.oi]
subME
ME.subset <- c(col.pal[2], col.pal[5], col.pal[8])

# add linechart annotation of each of module ME
topAnnotation <- HeatmapAnnotation(moduleEigenProtein = anno_lines(subME,
                                   gp = gpar(col = ME.subset),
                                   add_points = TRUE, 
                                   pt_gp = gpar(col = ME.subset)), 
                                   timepoint = str_extract(rownames(subME), "[247]d"),
                                   gp = gpar(col = c('na' = '#4477AA', '2d'='#66CCEE', '4d'="#228833", '7d'="#CCBB44")),
                                   which = "column",
                                   height = unit(5, "cm"))

hm <- Heatmap(submat, 
              cluster_columns = FALSE, 
              split = split,
              cluster_row_slices = F,
              show_row_names = F,
              right_annotation = row_ha,
              top_annotation = topAnnotation,
              name = "Ints vs Mock",
              column_split = tstrsplit(colnames(submat), "_")[[1]],
              na_col = 'darkgrey',
              row_title =sprintf("%d Proteins",
                            nrow(submat)),
              row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize=9),
              #show_row_names = F,
              column_title_gp = gpar (fontsize= 12))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))

BackupAsPDF(hm,  'heatmap.wgcna.modules.sigGnes.ME.anno.ordered',dimensions=c(23,22)) 
```
# Write out the modules for the group to inspect

```{r}
fwrite(modules.dt, ScriptAndDatedFileName('tmpNorm.sp4.deepSplit2.modules.csv'))
```



## Not useds
------

```{r}
output <- EnsemblIDsToSymbol(unique(p.quant$Protein), seqType = 'PROTEIN', species='MOUSE', attributes=c('external_gene_name', 'ensembl_peptide_id_version'), filters = 'ensembl_peptide_id_version')

test <- EnsemblIDsToSymbol(p.quant$Protein, seqType = 'PROTEIN', species='MOUSE', attributes=c('external_gene_name', 'ensembl_peptide_id_version'), filters = 'ensembl_peptide_id_version')
```

```{r}

tmp <- medpolish(p.mat)

tmp.dt <- data.table(sample=names(tmp$col),
                     offset=tmp$col)

p.quant <- merge(x=p.quant, y=tmp.dt, by='sample')
p.quant[, newLogIntensities := LogIntensities - offset]
```