---
title: "AViDD_AB_murine_WGCNA"
author: "Martin Gordon"
date: "2024-06-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

PW comparisons of the groups did not return many interesting results..
I think we need to take the sig hits vs Mock for all samples and then perform a WGCNA clustering

*Notes* 
I've joined clusters based on WARD distance; advantage of ward is it tried to reduce intra-cluster variance (idea is to have smaller, tight clusters)

## packages

```{r packages}
library(magrittr)
library(data.table)
library(circlize)
library(ComplexHeatmap)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(MSstats)
library(ggrepel)
library(readxl)
library(viridis)
library(ggvenn)
library(hrbrthemes)
library(viridis)
library(ggbeeswarm)
library(usedist) #package for working with distances
library(ggh4x) # additional functionality for ggplot2 obj
library(scales)
library(seqinr)
library(WGCNA)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

source ("../../utils/mg_utils/r_utils/IDmapping.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```

Read in the normalized count data
```{r}
p.quant <-  fread('~/Documents/projects/052924_ARichards_AViDD_Murine/AViDD_AB_murine_DE_data/2024_05_31_AB.ProteinLevelData.tmpNormNewCol.csv')
mss.dt <- fread('~/Documents/projects/052924_ARichards_AViDD_Murine/AViDD_AB_murine_DE_data/2024_06_03_mss.pwcontrasts.unfiltered.tmpNorm.csv')

contrasts.oi <- unique(grep('-Mock', mss.dt$Label, value=T))

#For the input, we will subtract the Mock counts from the others and input this to the contrast matrix
prots.oi <- mss.dt[Label %in% contrasts.oi & pvalue < 0.05, unique(Protein)]
```

get the heatmap of the proteins 
Lets try PAM clustering (more robust to noise)
```{r}
library(cluster)

p.mat <- dcast(p.quant, Protein~GROUP+SUBJECT, value.var = 'newLogIntensities') %>% 
  as.matrix(rownames='Protein')

# remove rows with lots of NA
subMat <- p.mat[rowSums(is.na(p.mat)) < 10,]


# euclidean as we want the strong signal to stand alone in a cluster
pamOut <- pam(subMat, k=4, diss=FALSE, metric='euclidean', stand=FALSE)

subMat <-  sweep(subMat, 1, STAT=apply(subMat, 1 , median, na.rm=T))

hm <- Heatmap(subMat, 
        row_split = pamOut$clustering,
        row_gap=unit(2, "mm"),
        show_row_names = F,
        column_title_gp = gpar(fontsize=8),
        row_names_gp = gpar(fontsize=2),
        border=TRUE,
        column_split = gsub('.[1-6]$', '', colnames(subMat)))

BackupAsPDF(hm, 'pamClust.cluster medianScaled.heatmap', dimensions=c(14,20))

#just plot the subset of the interesting clusters 
# yeah looks like we have recovered. cluster4 from the data 

# subset to cluster 4 and plot
subMat <- subMat[rownames(subMat) %in% names(pamOut$clustering[pamOut$clustering %in% c(4)]),]
rownames(subMat) <- multiUniprots2multiGenes(rownames(subMat), species='MOUSE')

hm <- Heatmap(subMat,
        row_gap=unit(2, "mm"),
        name = 'Clust4 Vs rowMedian',
        show_row_names = T,
        column_title_gp = gpar(fontsize=8),
        row_names_gp = gpar(fontsize=2),
        border=TRUE,
        column_split = gsub('.[1-6]$', '', colnames(subMat)))
hm

BackupAsPDF(hm, 'pamClust.cluster4.medianScaled.heatmap', dimensions=c(14,17))



#subdivide cluster 4
hm <- Heatmap(subMat[complete.cases(subMat),],
        row_gap=unit(2, "mm"),
        row_km = 2,
        name = 'Clust4 Vs rowMedian',
        show_row_names = T,
        column_title_gp = gpar(fontsize=8),
        row_names_gp = gpar(fontsize=3),
        border=TRUE,
        column_split = gsub('.[1-6]$', '', colnames(subMat)))
hm
BackupAsPDF(hm, 'pamClust.cluster4.subSplit.medianScaled.heatmap', dimensions=c(14,17))


#fwrite(data.table(gene=multiUniprots2multiGenes(names(pamOut$clustering[pamOut$clustering == 4]), species='MOUSE')), ScriptAndDatedFileName('batchAffectedCluster.txt'))
```

Calculate the difference of logIntensities vs Mock
how do we want to do this? Take the average value of mock samples and substract?

```{r}
# calculate difference vsMock
p.quant[, vsMock := newLogIntensities - mean(newLogIntensities[GROUP == 'Mock'], na.rm=T), by=.(Protein)]

p.mat <- dcast(p.quant, Protein~paste0(GROUP,'.',SUBJECT), value.var = 'vsMock') %>% 
  as.matrix(rownames='Protein')

p.mat <-  p.mat[rownames(p.mat) %in% prots.oi,]
```
5.5k proteins in input with this minimal filtering 
Now filter for missing values, want at least 80% completeness

4.5k after filtering out missing rows
```{r}
rm.cols <- apply(p.mat, 1, function(x) {sum(is.na(x)) })

submat <- p.mat[rm.cols < 14,]
dim(submat)
```
Heatmap of the input data

```{r}
colnames(submat)

hm <- Heatmap(submat, 
              cluster_rows = T, 
              cluster_columns = F,
              name='vsMock',
              na_col = 'darkgrey',
              column_split = gsub('.[0-9]+$', '', colnames(submat)), 
              row_title=sprintf('%s proteins', nrow(submat)), 
              show_row_names = F, 
              column_names_gp = gpar(fontsize=5), 
              column_title_gp = gpar(fontsize = 8))

hm
BackupAsPDF(draw(hm, column_title='WGCNA input'), 'wgcna.input.mat')
```

Now filter out mock and review

Clustering on WGCNA output

```{r}
submat <- p.mat[rm.cols < 14, grep('Mock', colnames(submat), invert=T)]

pamOut <- pam(submat, k=4, diss=FALSE, metric='euclidean', stand=FALSE)


# add annotation for cluster 4
# bracket enclosed is the filter
batch_anno <- rownames(submat) %in% names(pamOut$clustering[pamOut$clustering == 4]) # creates a filter; use this to extract the rownames
batch_genes <- rownames(submat)[batch_anno]

names_bar = rownames(submat) %in% batch_genes

hm <- Heatmap(submat, 
              cluster_rows = T, 
              cluster_columns = F,
              row_split = pamOut$clustering,
              na_col = 'darkgrey',
              name = 'Ints. vs Mock',
              column_split = gsub('.[0-9]+$', '', colnames(submat)), 
             # row_title=sprintf('%s proteins', nrow(submat)), 
              show_row_names = F, 
              border = T,
              column_names_gp = gpar(fontsize=5), 
              column_title_gp = gpar(fontsize = 8))

hm
BackupAsPDF(hm, 'wgnca.input.pamClust.heatmap', dimensions=c(12,19))

clust.dt <- extractClustersfromHeatmap(draw(hm), submat)
clust.dt[, gene := multiUniprots2multiGenes(feature, species='MOUSE')]


genes.oi <- multiUniprots2multiGenes(names(pamOut$clustering[pamOut$clustering == 4]), species='MOUSE')

rownames(submat) <-  multiUniprots2multiGenes(rownames(submat), species='MOUSE')

hm <- Heatmap(submat[rownames(submat) %in% genes.oi,], 
              cluster_rows = T, 
              cluster_columns = F,
            #  row_split = pamOut$clustering,
              na_col = 'darkgrey',
              name = 'Ints. vs Mock',
              column_split = gsub('.[0-9]+$', '', colnames(submat)), 
             # row_title=sprintf('%s proteins', nrow(submat)), 
              show_row_names = T, 
              border = T,
              row_names_gp = gpar(fontsize=2),
              column_names_gp = gpar(fontsize=5), 
              column_title_gp = gpar(fontsize = 8))

hm
BackupAsPDF(hm, 'wgcnaInput.pamClust.Cluster4.heatmap', dimensions = c(15,20))

fwrite(clust.dt[cluster ==4,], ScriptAndDatedFileName('wgcnaInput.cluster4.csv'))
```

Look at the sample clustering 
```{r}
#Group data in a dendogram to check outliers
sampleTree = hclust(dist(t(submat)), method = "average")

pdf(file = "./AViDD_AB_murine_WGCNA_data/sampleClustering.pdf", width = 12, height = 9); 
par(cex = 0.6)
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering", sub="", xlab="", cex.lab = 1.5, 
     cex.axis = 1.5, cex.main = 2)
#Plot a line showing the cut-off
abline(h = 56, col = "red") #This value of 31000 was chosen based on my data, you need to check the best value to your data

dev.off()
```

Now that we have looked at the samples, lets checkout the genes we have selected
```{r}
gsg <- goodSamplesGenes(t(submat))
summary(gsg)
gsg$allOK
```

Lets go with 5 as it is the last signficant jump in R2 and reaches R > .9

```{r}
net.in <- t(submat)

s.thres <- data.table(pickSoftThreshold(net.in)$fitIndices)
s.thres
s.thres %>%  
  ggplot(aes(x=Power, y=SFT.R.sq)) + 
  geom_point() +
  geom_hline(yintercept = 0.8, color='darkred') +
  labs(y="Scale Free Topology Model Fit, signed R^2") +
  theme_classic()

s.thres %>%  
  ggplot(aes(x=Power, y=mean.k.)) + 
  geom_point() +
  labs(y='Mean Connectivity') +
  theme_classic() 
```

```{r}
softPower = 5 #Chosen in the graphs before
adj.mat = adjacency(net.in, power = softPower, corFnc = 'bicor', type = "signed") #Calculating the adjacency matrix
#help(adjacency )
```

Define our topological overlap matrix
```{r}
TOM <- TOMsimilarity(adj.mat)
TOM.dissimilarity <- 1-TOM

#creating the dendrogram 
geneTree <- hclust(as.dist(TOM.dissimilarity), method = "average") 
#plotting the dendrogram
sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity", 
labels = FALSE, hang = 0.04)
```

I dont think these clusters look too promising, but lets run the analysis and see
```{r}
hm <- Heatmap(TOM[1:1000,1:1000], cluster_rows = T, cluster_columns=T, column_title = 'adjacency matrix', show_row_names = F, show_column_names = F)
hm
#BackupAsPDF(hm, 'tom.mat.st5.')
hm <- Heatmap(adj.mat[1:1000,1:1000], cluster_rows = T, cluster_columns=T, column_title = 'TOM similiary', show_row_names = F, show_column_names = F)
#BackupAsPDF(hm, 'tom.mat.st4.')
```

```{r}
#detected six modules; want to correlate these 
Modules <- cutreeDynamic(dendro = geneTree, distM = TOM.dissimilarity, deepSplit = 4, pamRespectsDendro = FALSE, minClusterSize = 20)
table(Modules)
```

Only three modules detected... try with the adjacency matrix as this tends to avoid merging larger modules

```{r}
adj.dissimilarity <-  1 - adj.mat

geneTree <- hclust(as.dist(TOM.dissimilarity), method = "ward.D2") 
plot(geneTree, xlab='', sub='', labels = F)

help(hclust)
```
```{r}
# detected 13 modules
Modules <- cutreeDynamic(dendro = geneTree, distM = TOM.dissimilarity, deepSplit = 4, pamRespectsDendro = FALSE, minClusterSize = 30)
table(Modules)
```

```{r}
modules.tom <- as.character(Modules)
```

```{r}
col.pal <- randomcoloR::distinctColorPalette(length(unique(Modules)))

# add wgcna module annotation
row_ha <- rowAnnotation(wgcna.modules=modules.tom, col=list(wgcna.modules = c("1"=col.pal[1], '2'=col.pal[2], '3'=col.pal[3], '4'=col.pal[4], '5'=col.pal[5], 
                                                                              '6'=col.pal[6], '7'=col.pal[7], '8'=col.pal[8], '9'=col.pal[9], '10'=col.pal[10],
                                                                              '11'=col.pal[11], '12'=col.pal[12], '13'=col.pal[13]
                                                                              )))

# add split order
split <- factor(modules.tom, levels=c('1','2','3','4','5','6','7','8','9','10','11','12','13'))

hm <- Heatmap(submat, 
              cluster_columns = FALSE, 
              #show_row_names = FALSE,
              split = split,
              right_annotation = row_ha,
              name = "Ints vs Mock",
              column_split = tstrsplit(colnames(submat), "[.]")[[1]],
              na_col = 'darkgrey',
              row_title =sprintf("%d Proteins",
                            nrow(submat)),
              row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize= 3),
              cluster_row_slices = T,
              column_title_gp = gpar (fontsize= 7))
hm
hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))

BackupAsPDF(hm, 'heatmap.wgcna.modules.ordered',dimensions=c(12,16))
```
Reorder the column clustering based on the PCA



Poor results with the TOM distance matrix
As this merges different ones together, lets try the adj distance matrix
```{r}
Modules <- cutreeDynamic(dendro = geneTree, distM = TOM.dissimilarity, deepSplit = 4, pamRespectsDendro = FALSE, minClusterSize = 20)
table(Modules)

Modules <- cutreeDynamic(dendro = geneTree, distM = adj.dissimilarity, deepSplit = 4, pamRespectsDendro = FALSE, minClusterSize = 20)
table(Modules)
```
