---
title: "AViDD_AB_murine_WGCNA_smoothMatrix"
author: "Martin Gordon"
date: "2024-06-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## WGCNA network analysis with batch-corrected and tmp normalized matrix

Now that we have adjusted for batch, rerun the WGCNA analysis and see if we can detect non-linear trends with the batch corrected data

## packages

```{r packages}
library(magrittr)
library(data.table)
library(circlize)
library(ComplexHeatmap)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(ggrepel)
library(readxl)
library(viridis)
library(viridis)
library(usedist) #package for working with distances
library(ggh4x) # additional functionality for ggplot2 obj
library(scales)
library(seqinr)
library(WGCNA)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

source ("../../utils/mg_utils/r_utils/IDmapping.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```

Read in the normalised data and ms out results

```{r}
mss.dt <- fread('~/Documents/projects/052924_ARichards_AViDD_Murine/AViDD_AB_murine_DE_smoothMatrix_data/2024_06_07_mss.pwcontrasts.tmpNorm.cleanDat.csv')
p.quant <- fread('~/Documents/projects/052924_ARichards_AViDD_Murine/PC1_batchRemoval_data/2024_06_04_ProteinLevelQuant.PC1SmoothedValues.csv')

p.mat <- dcast(p.quant, Protein~sample, value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')

tmp <- medpolish(p.mat)

tmp.dt <- data.table(sample=names(tmp$col),
                     offset=tmp$col)

p.quant <- merge(x=p.quant, y=tmp.dt, by='sample')
p.quant[, newLogIntensities := LogIntensities - offset]
```
 plot of the adjusted intensity distributions
 
```{r}
ggplot(p.quant, aes(x=interaction(GROUP,SUBJECT), y=newLogIntensities)) +
  geom_boxplot()
```

Calculate the difference of logIntensities vs Mock
how do we want to do this? Take the average value of mock samples and substract?

```{r}
# calculate difference vsMock
p.quant[, vsMock := newLogIntensities - mean(newLogIntensities[GROUP == 'Mock'], na.rm=T), by=.(Protein)]

#fwrite(p.quant, ScriptAndDatedFileName('input.ProteinInts.clean.tmpNorm.csv'))
p.quant <- fread('~/Documents/projects/052924_ARichards_AViDD_Murine/AViDD_AB_murine_WGCNA_smoothMatrix_data/2024_06_10_input.ProteinInts.clean.tmpNorm.csv')
```

Get matrix of protein intensity values
```{r}
p.mat <- dcast(p.quant, Protein~paste0(GROUP,'.',SUBJECT), value.var = 'vsMock') %>% 
  as.matrix(rownames='Protein')
```

Question: should we filter only to genes that are DE? 
maybe leave as is for now and can retry

```{r}
hm <- Heatmap(p.mat, 
              cluster_rows = T, 
              cluster_columns = F,
              name='vsMock',
              na_col = 'darkgrey',
              column_split = gsub('.[0-9]+$', '', colnames(p.mat)), 
              row_title=sprintf('%s proteins', nrow(p.mat)), 
              show_row_names = F, 
              column_names_gp = gpar(fontsize=5), 
              column_title_gp = gpar(fontsize = 8))

hm
BackupAsPDF(draw(hm, column_title='WGCNA input'), 'wgcna.input.mat')

```
filter out the Mock samples from the input as we are already accounting for their expression in the Ints values

```{r}
submat <- p.mat[, grep('Mock', colnames(p.mat), invert=T)]

hm <- Heatmap(submat, 
              cluster_rows = T, 
              cluster_columns = F,
              name='vsMock',
              na_col = 'darkgrey',
              column_split = gsub('.[0-9]+$', '', colnames(submat)), 
              row_title=sprintf('%s proteins', nrow(submat)), 
              show_row_names = F, 
              column_names_gp = gpar(fontsize=5), 
              column_title_gp = gpar(fontsize = 8))

hm
BackupAsPDF(draw(hm, column_title='WGCNA input'), 'wgcna.input.noMock.mat', dimensions=c(14,12))
```

Look at the sample clustering 
```{r}
#Group data in a dendogram to check outliers
sampleTree = hclust(dist(t(submat)), method = "average")

pdf(file = "./AViDD_AB_murine_WGCNA_smoothMatrix_data/sampleClustering.pdf", width = 12, height = 9); 
par(cex = 0.6)
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering", sub="", xlab="", cex.lab = 1.5, 
     cex.axis = 1.5, cex.main = 2)
#Plot a line showing the cut-off
abline(h = 56, col = "red") #This value of 31000 was chosen based on my data, you need to check the best value to your data

dev.off()
```

Now that we have looked at the samples, lets checkout the genes we have selected
```{r}
gsg <- goodSamplesGenes(t(submat))
summary(gsg)
gsg$allOK
```
Create matrix by selecting a power used for thresholding edges

Lets go with 5 as it is the last signficant jump in R2 and reaches R > .9

```{r}
net.in <- t(submat)

s.thres <- data.table(pickSoftThreshold(net.in)$fitIndices)

s.thres %>%  
  ggplot(aes(x=Power, y=SFT.R.sq)) + 
  geom_point() +
  geom_hline(yintercept = 0.8, color='darkred', linetype=2) +
  geom_hline(yintercept = 0.9, color='darkred', linetype=2) +
  labs(y="Scale Free Topology Model Fit, signed R^2") +
  theme_classic()

s.thres %>%  
  ggplot(aes(x=Power, y=mean.k.)) + 
  geom_point() +
  labs(y='Mean Connectivity') +
  theme_classic() 
```

```{r}
softPower = 5 #Chosen in the graphs before
adj.mat = adjacency(net.in, power = softPower, corFnc = 'bicor', type = "signed") #Calculating the adjacency matrix
#help(adjacency )
```

Define our topological overlap matrix
```{r}
TOM <- TOMsimilarity(adj.mat)
TOM.dissimilarity <- 1-TOM

#creating the dendrogram 
geneTree <- hclust(as.dist(TOM.dissimilarity), method = "average") 
#plotting the dendrogram
sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity", 
labels = FALSE, hang = 0.04)
```

I dont think these clusters look too promising, but lets run the analysis and see
```{r}
hm <- Heatmap(TOM[1:2000,1:2000], cluster_rows = T, cluster_columns=T, column_title = 'adjacency matrix', show_row_names = F, show_column_names = F)
hm
BackupAsPDF(hm, 'tom.mat.st5.')
hm <- Heatmap(adj.mat[1:1000,1:1000], cluster_rows = T, cluster_columns=T, column_title = 'TOM similiary', show_row_names = F, show_column_names = F)
hm
#BackupAsPDF(hm, 'tom.mat.st4.')
```

Detected 20 modules
Lets look at these and plot the average expression of each of these
```{r}
Modules <- cutreeDynamic(dendro = geneTree, distM = TOM.dissimilarity, deepSplit = 1, pamRespectsDendro = FALSE, minClusterSize = 30)
# make a datatable of the module assignments
modules.dt <- data.table(Protein=rownames(adj.mat),
                         Module=as.character(Modules))
```
plot the heatmap of the genes


```{r}

col.pal <- randomcoloR::distinctColorPalette(length(unique(Modules)))
length(unique(Modules))


modules.tom <- as.character(Modules)

# add wgcna module annotation
row_ha <- rowAnnotation(wgcna.modules=modules.tom, col=list(modules.tom = c("1"=col.pal[1], '2'=col.pal[2], '3'=col.pal[3], '4'=col.pal[4], '5'=col.pal[5], 
                                                                              '6'=col.pal[6], '7'=col.pal[7], '8'=col.pal[8], '9'=col.pal[9], '10'=col.pal[10],
                                                                              '11'=col.pal[11], '12'=col.pal[12], '13'=col.pal[13])))

# add split order
split <- factor(modules.tom, levels=c('1','2','3','4','5','6','7','8','9','10','11','12','13'))

hm <- Heatmap(submat, 
              cluster_columns = FALSE, 
              #show_row_names = FALSE,
              split = split,
              right_annotation = row_ha,
              name = "Ints vs Mock",
              column_split = tstrsplit(colnames(submat), "[.]")[[1]],
              na_col = 'darkgrey',
              row_title =sprintf("%d Proteins",
                            nrow(submat)),
              row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize= 3),
              cluster_row_slices = T,
              column_title_gp = gpar (fontsize= 7))
hm
hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))

BackupAsPDF(hm, 'heatmap.wgcna.modules.ordered',dimensions=c(12,16))
```
Save results
```{r}
#fwrite(modules.dt, ScriptAndDatedFileName('moduleAssignments.deepsplit2.csv'))
```
Not sure about this.. these clusters dont look great.... lets try filter to DEG and rerun 

```{r}
# filter to prots that are DE w/o correction 
prots.oi <- mss.dt[Label %in% grep('Mock', Label, value=T) & pvalue < 0.05, unique(Protein)]

# filter the submat to DEGs
f.mat <- submat[rownames(submat) %in% prots.oi,]
```

Lets go with 5 again
```{r}
net.in <- t(f.mat)

s.thres <- data.table(pickSoftThreshold(net.in)$fitIndices)

s.thres %>%  
  ggplot(aes(x=Power, y=SFT.R.sq)) + 
  geom_point() +
  geom_hline(yintercept = 0.8, color='darkred', linetype=2) +
  geom_hline(yintercept = 0.9, color='darkred', linetype=2) +
  labs(y="Scale Free Topology Model Fit, signed R^2") +
  theme_classic()

s.thres %>%  
  ggplot(aes(x=Power, y=mean.k.)) + 
  geom_point() +
  labs(y='Mean Connectivity') +
  theme_classic() 
```

```{r}
softPower = 5 #Chosen in the graphs before
adj.mat = adjacency(net.in, power = softPower, corFnc = 'bicor', type = "signed") #Calculating the adjacency matrix
#help(adjacency )
```

Define our topological overlap matrix
```{r}
TOM <- TOMsimilarity(adj.mat)
TOM.dissimilarity <- 1-TOM

#creating the dendrogram 
geneTree <- hclust(as.dist(TOM.dissimilarity), method = "average") 
#plotting the dendrogram
sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity", 
labels = FALSE, hang = 0.04)
```
```{r}
Modules <- cutreeDynamic(dendro = geneTree, distM = TOM.dissimilarity, deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = 30)
table(Modules)

modules.dt <- data.table(Protein=rownames(adj.mat),
                         Module=as.character(Modules))

#fwrite(modules.dt, ScriptAndDatedFileName('sigGenes.moduleAssignments.deesplit2.csv'))
```
11 modules.. lets plot these and see how they look
```{r}
#col.pal <- randomcoloR::distinctColorPalette(length(unique(Modules)))

# safe color pallete from https://stackoverflow.com/questions/9563711/r-color-palettes-for-many-data-classes
col.pal <- c(
  "dodgerblue2", 
  "#E31A1C", # red
  "green4",
  "#6A3D9A", # purple
  "#FF7F00", # orange
  "black", "gold1",
  "skyblue2", "#FB9A99", # lt pink
  "palegreen2",
  "#CAB2D6", # lt purple
  "#FDBF6F", # lt orange
  "gray70", "khaki2",
  "maroon", "orchid1", "deeppink1", "blue1", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3",
  "darkorange4", "brown"
)

col.subset <- c(col.pal[1], col.pal[2], col.pal[3], col.pal[4], col.pal[5], col.pal[15], col.pal[7], col.pal[8], col.pal[9], col.pal[10], col.pal[11],col.pal[12],col.pal[13])

modules.tom <- as.character(Modules)

# add wgcna module annotation
row_ha <- rowAnnotation(wgcna.modules=modules.tom, col=list(wgcna.modules = c("1"=col.pal[1], '2'=col.pal[2], '3'=col.pal[3], '4'=col.pal[4], '5'=col.pal[5], 
                                                                              '6'=col.pal[15], '7'=col.pal[7], '8'=col.pal[8], '9'=col.pal[9], '10'=col.pal[10],
                                                                              '11'=col.pal[11], '12'=col.pal[12], '13'=col.pal[13])))


# add split order
split <- factor(modules.tom, levels=c('"1','2','3','4','5','6','7','8','9','10','11','12','13'))

hm <- Heatmap(f.mat, 
              cluster_columns = FALSE, 
              split = split,
              right_annotation = row_ha,
              name = "Ints vs Mock",
              column_split = tstrsplit(colnames(f.mat), "[.]")[[1]],
              na_col = 'darkgrey',
              row_title =sprintf("%d Proteins",
                            nrow(f.mat)),
              row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize= 3),
              cluster_row_slices = T,
              show_row_names = F,
              column_title_gp = gpar (fontsize= 7))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))

BackupAsPDF(hm, 'heatmap.wgcna.modules.sigGnes.ordered',dimensions=c(12,16))
```
Lets check out the enrichment of the different clusters and see if any seem driven by bio processes

```{r}
# load the GO table
gmt.go <- loadGmtFromBioconductor(dbName = 'org.Mm.eg.db', ontology = "ALL", keyType = "SYMBOL")
```

```{r}
# our background? only genes used for WGCNA analysis
modules.dt[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]

universe <- unique(modules.dt$gene)

enrich.dt <- enricherOnGroups(modules.dt, groupColumns = 'Module', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'Module')

simp.enrich$simplified[, Module := factor(Module, levels=c('1','2','3','4','5','6','7','8','9','10','11'))]

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, groupColumn = 'Module', topN = 10, title='GO enrichment WGCNA modules', 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 6), upperThreshold = 6, cluster_columns =F)

BackupAsPDF(ht, 'go.enrichment.wgcna.modules.sigGenes.ordered', dimensions=c(8,12))
```
Create a top annotation

```{r}

p.quant[, moduleMean := mean(vsMock), by=.(sample,Module)]

# get the average per module/sample
sumMat <- dcast(p.quant, Module~sample, value.var = 'vsMock', fun.aggregate = mean) %>% 
  as.matrix(rownames='Module')

rownames(sumMat) <- factor(rownames(sumMat), levels=c('1','2','3','4','5','6','7','8','9','10','11'))

mod.order <- c('1','2','3','4','5','6','7','8','9','10','11')

sumMat <- sumMat[mod.order, grep('Mock', colnames(sumMat), invert = T, value = T)]


topAnnotation <- HeatmapAnnotation(moduleAverage = anno_lines(t(sumMat),
                                   gp = gpar(col = col.subset),
                                   add_points = TRUE, 
                                   pt_gp = gpar(col = col.subset)),  
                                   which = "column",
                                   height = unit(5, "cm"))

hm <- Heatmap(f.mat, 
              cluster_columns = FALSE, 
              split = split,
              right_annotation = row_ha,
              top_annotation = topAnnotation,
              name = "Ints vs Mock",
              column_split = tstrsplit(colnames(f.mat), "[.]")[[1]],
              na_col = 'darkgrey',
              row_title =sprintf("%d Proteins",
                            nrow(f.mat)),
              row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize= 6),
              cluster_row_slices = T,
              show_row_names = F,
              column_title_gp = gpar (fontsize= 7))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))


BackupAsPDF(hm, 'heatmap.wgcna.modules.sigGnes.ordered',dimensions=c(22,22))   
```

06-11-24
-----
Read in the results from the sigGene clustering 

```{r}
clust.genes <- fread('./AViDD_AB_murine_WGCNA_smoothMatrix_data/2024_06_10_sigGenes.moduleAssignments.deesplit2.csv')
```
```{r}
# now need to name these module genes according to their enrichment profiles
clust.genes[, enrichModules := ifelse(Module %in% c('4','7','9','11'), Module, 
                                      ifelse(Module == '1', '1 (Response to External Stimuli)',
                                             ifelse(Module == '2', '2 (Macromolecule Catabolism)',
                                                    ifelse(Module == '3', '3 (Translation)',
                                                           ifelse(Module == '5', '5 (Immune Response Regulation)',
                                                                  ifelse(Module == '6', '6 (TCA Cycle)',
                                                                         ifelse(Module == '8', '8 (adenyl nucleotide binding)',
                                                                                ifelse(Module == '10', '10 (Protein Depolymerization)', '')))))))) ]
# sanity check
clust.genes[,.N, by=.(Module,enrichModules)]

modules.tom <- as.character(clust.genes$enrichModules)

# add split order
split <- factor(modules.tom, levels=c('1 (Response to External Stimuli)','2 (Macromolecule Catabolism)','3 (Translation)',
                                      '4','5 (Immune Response Regulation)','6 (TCA Cycle)','7','8 (adenyl nucleotide binding)',
                                      '9','10 (Protein Depolymerization)','11'))


unique(split)
# add wgcna module annotation
row_ha <- rowAnnotation(wgcna.modules=modules.tom, 
                        col=list(wgcna.modules = c('1 (Response to External Stimuli)'=col.pal[1], '2 (Macromolecule Catabolism)'=col.pal[2], 
                                                                              '3 (Translation)'=col.pal[3], '4'=col.pal[4], '5 (Immune Response Regulation)'=col.pal[5], 
                                                                              '6 (TCA Cycle)'=col.pal[15], '7'=col.pal[7], '8 (adenyl nucleotide binding)'=col.pal[8], 
                                                                              '9'=col.pal[9], '10 (Protein Depolymerization)'=col.pal[10], '11'=col.pal[11])))

hm <- Heatmap(f.mat, 
              cluster_columns = FALSE, 
              split = split,
              cluster_row_slices = F,
              right_annotation = row_ha,
              name = "Ints vs Mock",
              column_split = tstrsplit(colnames(f.mat), "[.]")[[1]],
              na_col = 'darkgrey',
              row_title =sprintf("%d Proteins",
                            nrow(f.mat)),
              row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize= 3),
              show_row_names = F,
              column_title_gp = gpar (fontsize= 7))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))

hm
BackupAsPDF(hm, 'heatmap.wgcna.modules.sigGnes.anno.ordered',dimensions=c(22,22))  
```
Lets split this heatmap differently and plot the MEs of the interesting modules at the top


```{r}
colnames(MEs0)

# just subset to the interesting modules
modules.oi <- c("ME1 (Response to External Stimuli)", 'ME4', "ME5 (Immune Response Regulation)","ME8 (adenyl nucleotide binding)")

subME <- MEs0[, modules.oi]

subME
ME.subset <- c(col.pal[1], col.pal[4], col.pal[5], col.pal[8])

# add linechart annotation of each of module ME
topAnnotation <- HeatmapAnnotation(moduleAverage = anno_lines(subME,
                                   gp = gpar(col = ME.subset),
                                   add_points = TRUE, 
                                   pt_gp = gpar(col = ME.subset)), 
                                   timepoint = str_extract(rownames(subME), "[247]d"),
                                   which = "column",
                                   height = unit(5, "cm"))

hm <- Heatmap(f.mat, 
              cluster_columns = FALSE, 
              split = split,
              cluster_row_slices = F,
              right_annotation = row_ha,
              top_annotation = topAnnotation,
              name = "Ints vs Mock",
              column_split = tstrsplit(colnames(f.mat), "_")[[1]],
              na_col = 'darkgrey',
              row_title =sprintf("%d Proteins",
                            nrow(f.mat)),
              row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize=9),
              show_row_names = F,
              column_title_gp = gpar (fontsize= 12))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))

hm
BackupAsPDF(hm,  'heatmap.wgcna.modules.sigGnes.ME.anno.ordered',dimensions=c(23,22)) 
```
Need to collapse these visualisations a little more.. maybe take the moduleME and plot boxplots of these values? Or else distributions of

```{r}
# our background? only genes used for WGCNA analysis
clust.genes[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]
clust.genes

universe <- unique(clust.genes$gene)

enrich.dt <- enricherOnGroups(clust.genes, groupColumns = 'Module', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)

simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'Module')
simp.enrich$simplified[, Module := factor(Module, levels=c('1','2','3','4','5','6','7','8','9','10','11'))]

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, groupColumn = 'Module', topN = 10, title='GO enrichment WGCNA modules', 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 6), upperThreshold = 6, cluster_columns =F)

ht
```
Module eigengene/First PC
---
Identify the key trends in each of the modules
```{r}
# Get Module Eigengenes per cluster
MEs0 <- moduleEigengenes(t(f.mat), clust.genes$Module)$eigengenes
```
Heatmap of Module eigengenes
```{r}
# create annotation
ann <- data.table(treatment=sub("_.+", "", colnames(t(MEs0))),
                  timepoint=ifelse(grepl("[247]d", colnames(t(MEs0))), str_extract(colnames(t(MEs0)), "[247]d"), 'na'))

colours <- list('timepoint' = c('na' = '#4477AA', '2d'='#66CCEE', '4d'="#228833", '7d'="#CCBB44"),
                'treatment' =c('Nirm' = col.pal[1], 'Vehicle'=col.pal[2], 'x4516'=col.pal[3], 'x4673'=col.pal[4]))

colAnn <- HeatmapAnnotation(df = ann, col = colours)


hm <- Heatmap(t(MEs0),
        name='module PC1',
        top_annotation = colAnn,
        column_names_gp = gpar(fontsize=8),
        column_split = gsub('_.+','',colnames(t(MEs0))),
        #column_split = gsub('.[1-6]$','',colnames(t(MEs0))),
        cluster_columns = FALSE)

BackupAsPDF(hm, 'ME.PC1.heatmap',dimensions=c(9,6))
```
I think we should use the GO enrichment terms to relabel the MEs

```{r}
colnames(MEs0) #"ME1"  "ME2"  "ME3"  "ME4"  "ME5"  "ME6"  "ME7"  "ME8"  "ME9"  "ME10" "ME11"
colnames(MEs0) <- c('ME1 (Response to External Stimuli)', 'ME2 (Macromolecule Catabolism)', 'ME3 (Translation)', 'ME4',
                    'ME5 (Immune Response Regulation)', 'ME6 (TCA Cycle)', 'ME7', 'ME8 (adenyl nucleotide binding)', 'ME9', 'ME10 (Protein Depolymerization)', 'ME11')
```

```{r}
hm <- Heatmap(t(MEs0),
        name='module PC1',
        top_annotation = colAnn,
        column_names_gp = gpar(fontsize=8),
        row_names_gp = gpar(fontsize=8),
        column_split = gsub('_.+','',colnames(t(MEs0))),
        #column_split = gsub('.[1-6]$','',colnames(t(MEs0))),
        cluster_columns = FALSE)
hm
BackupAsPDF(hm, 'ME.PC1.heatmap',dimensions=c(10,6))
```
Lets collapse our input matrix to mean per group and then plot boxplots of the distributions of the genes in that module

```{r}
fmat.dt <- setDT(reshape2::melt(f.mat, value.name = 'LogInts.vs.Mock'))
setnames(fmat.dt, new=c('Protein', 'sample', 'LogInts.vs.Mock'))

fmat.dt[, GROUP := gsub('[.][1-6]$','',sample)]

# convert to aggregated matrix
agg.mat <- dcast(fmat.dt, Protein~GROUP, value.var = 'LogInts.vs.Mock', fun.aggregate = mean, na.rm=T) %>% 
  as.matrix(rownames='Protein')

```

ok, now want to seperate by timepoint (or order by timepoint) and then look at the distributions of the interesting modules as boxplots
Also, share the files with the enrichment results for each module and the protein assignments

```{r}
# just subset to the interesting modules
modules.oi <- c("ME1 (Response to External Stimuli)", "ME5 (Immune Response Regulation)")







# add boxplot annotation of each of module ME
topAnnotation <- HeatmapAnnotation('ME1 (Response to Stimuli)' = anno_boxplot(agg.mat[rownames(agg.mat) %in% clust.genes[Module == '1', Protein],],
                                   gp = gpar(fill = col.pal[1])),
                                   'ME5 (Immune Response)' = anno_boxplot(agg.mat[rownames(agg.mat) %in% clust.genes[Module == '5', Protein],],
                                   gp = gpar(fill = col.pal[5])),
                                   treatment = gsub('_.+$','', colnames(agg.mat)),
                                   which = "column",
                                   height = unit(8, "cm"))

hm <- Heatmap(agg.mat, 
        column_split = str_extract(colnames(agg.mat), '[247]d'),
        show_row_names = F,
        name='Ints. vs Mock',
        top_annotation = topAnnotation,
        split = split,
        cluster_row_slices = F,
        right_annotation = row_ha,
        row_title = sprintf('%s Proteins', nrow(agg.mat)),
        cluster_columns = F)


BackupAsPDF(hm, 'heatmap.wgcna.modules.sigGnes.moduleBoxplots.anno.ordered',dimensions=c(23,22)) 
```
These plots dont make sense... why is module 5 genes down regulated in mock vs treatment?
Extract these genes and look at the box plots

```{r}
prots.oi <- clust.genes[Module == '5', Protein]

ggplot(p.quant[Protein %in% prots.oi,], aes(x=sample, y=newLogIntensities, fill=GROUP)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=90))

ggplot(p.quant, aes(x=sample, y=newLogIntensities, fill=GROUP)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=90))
```
Share the results 

```{r}
fwrite(clust.genes, ScriptAndDatedFileName('wgcna.moduleGenes.csv'))

```












Not used
----

```{r}
f.mat

p.quant[, moduleMean := mean(vsMock), by=.(sample,Module)]

# get the average per module/sample
sumMat <- dcast(p.quant, Module~sample, value.var = 'vsMock', fun.aggregate = mean) %>% 
  as.matrix(rownames='Module')

rownames(sumMat) <- factor(rownames(sumMat), levels=c('1','2','3','4','5','6','7','8','9','10','11'))

mod.order <- c('1','2','3','4','5','6','7','8','9','10','11')

sumMat <- sumMat[mod.order, grep('Mock', colnames(sumMat), invert = T, value = T)]


topAnnotation <- HeatmapAnnotation(moduleAverage = anno_lines(t(sumMat),
                                   gp = gpar(col = col.subset),
                                   add_points = TRUE, 
                                   pt_gp = gpar(col = col.subset)),  
                                   which = "column",
                                   height = unit(5, "cm"))

hm <- Heatmap(f.mat, 
              cluster_columns = FALSE, 
              split = split,
              right_annotation = row_ha,
              top_annotation = topAnnotation,
              name = "Ints vs Mock",
              column_split = tstrsplit(colnames(f.mat), "[.]")[[1]],
              na_col = 'darkgrey',
              row_title =sprintf("%d Proteins",
                            nrow(f.mat)),
              row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize= 6),
              cluster_row_slices = T,
              show_row_names = F,
              column_title_gp = gpar (fontsize= 7))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))


BackupAsPDF(hm, 'heatmap.wgcna.modules.sigGnes.ordered',dimensions=c(22,22))   
```





Correlations as are show best performance/clustering
lets extract these hits 

Lets retry the correlation using pearson/spearman and see if this recovers more uniform clusters

```{r}
softPower = 5 #Chosen in the graphs before
adj.mat = adjacency(net.in, 
                    power = softPower, 
                    corFnc = 'cor', #pearson by default
                    corOptions = list(use = 'p', method = 'spearman'),
                    type = "signed") #Calculating the adjacency matrix

```

```{r}
TOM <- TOMsimilarity(adj.mat)
TOM.dissimilarity <- 1-TOM
```

```{r}
Modules <- cutreeDynamic(dendro = geneTree, distM = TOM.dissimilarity, deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = 30)
table(Modules)

modules.dt <- data.table(Protein=rownames(adj.mat),
                         Module=as.character(Modules))

#fwrite(modules.dt, ScriptAndDatedFileName('sigGenes.moduleAssignments.deesplit2.csv'))
```

```{r}
#col.pal <- randomcoloR::distinctColorPalette(length(unique(Modules)))

# safe color pallete from https://stackoverflow.com/questions/9563711/r-color-palettes-for-many-data-classes
col.pal <- c(
  "dodgerblue2", 
  "#E31A1C", # red
  "green4",
  "#6A3D9A", # purple
  "#FF7F00", # orange
  "black", "gold1",
  "skyblue2", "#FB9A99", # lt pink
  "palegreen2",
  "#CAB2D6", # lt purple
  "#FDBF6F", # lt orange
  "gray70", "khaki2",
  "maroon", "orchid1", "deeppink1", "blue1", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3",
  "darkorange4", "brown"
)

col.subset <- c(col.pal[1], col.pal[2], col.pal[3], col.pal[4], col.pal[5], col.pal[15], col.pal[7], col.pal[8], col.pal[9], col.pal[10], col.pal[11])

modules.tom <- as.character(Modules)

# add wgcna module annotation
row_ha <- rowAnnotation(wgcna.modules=modules.tom, col=list(wgcna.modules = c("1"=col.pal[1], '2'=col.pal[2], '3'=col.pal[3], '4'=col.pal[4], '5'=col.pal[5], 
                                                                              '6'=col.pal[15], '7'=col.pal[7], '8'=col.pal[8], '9'=col.pal[9], '10'=col.pal[10])))


# add split order
split <- factor(modules.tom, levels=c('"1','2','3','4','5','6','7','8','9','10'))


hm <- Heatmap(f.mat, 
              cluster_columns = FALSE, 
              split = split,
              right_annotation = row_ha,
              name = "Ints vs Mock",
              column_split = tstrsplit(colnames(f.mat), "[.]")[[1]],
              na_col = 'darkgrey',
              row_title =sprintf("%d Proteins",
                            nrow(f.mat)),
              row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize= 3),
              cluster_row_slices = T,
              show_row_names = F,
              column_title_gp = gpar (fontsize= 7))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))

BackupAsPDF(hm, 'heatmap.wgcna.modules.sigGnes.ordered',dimensions=c(12,16))

```

Not used
----
Filter for genes that are de in the different comparisons

```{r}
contrasts.oi <- unique(grep('-Mock', mss.dt$Label, value=T))

#For the input, we will subtract the Mock counts from the others and input this to the contrast matrix
prots.oi <- mss.dt[pvalue < 0.05, unique(Protein)]
```
