---
title: "AB_PC1_batchRemoval"
author: "Martin Gordon"
date: "2024-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Goal here is to remove the batch effect
Take the rotation matrix and multiply by the loadings and subtract from the inital log2FC values
Then run the differential expression

```{r packages}
library(magrittr)
library(data.table)
library(circlize)
library(ComplexHeatmap)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(MSstats)
library(ggrepel)
library(readxl)
library(viridis)
library(ggvenn)
library(hrbrthemes)
library(viridis)
library(ggbeeswarm)
library(usedist) #package for working with distances
library(ggh4x) # additional functionality for ggplot2 obj
library(scales)
library(seqinr)
library(WGCNA)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

source ("../../utils/mg_utils/r_utils/IDmapping.R")
source ("/Users/martingordon/Documents/utils/mg_utils/r_utils/HelperFunctions.R") # extract clusters from heatmaps


# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```

Read in the p.quant data

```{r}
p.quant <-  fread('~/Documents/projects/052924_ARichards_AViDD_Murine/AViDD_AB_murine_DE_data/2024_05_31_AB.ProteinLevelData.tmpNormNewCol.csv')

prot.mat <- dcast(p.quant, Protein~paste0(GROUP,'.',SUBJECT), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')

prot.mat <- prot.mat[complete.cases(prot.mat),]
```
 Run PCA on the matrix
 
```{r}
col.pal <- randomcoloR::distinctColorPalette(k=5)
col.pal
pcaOut <- prcomp(t(prot.mat))

colInfo <- data.table(colname = colnames(prot.mat))
colInfo[,c("status", "rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]
colInfo[, c("treatment", "timepoint") := tstrsplit(status, '[_]', keep=c(1,2)) ]
colInfo[status == 'Mock', timepoint := '0d']

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)

#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = treatment, shape = timepoint)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  #scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = c(21:26)) +
  scale_fill_manual(values = col.pal) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

BackupAsPDF(p, 'PCA.proteins.treatandtimepoint')


pcaDT[order(PC1), rn]
```
 
 
 
 First lets plot the heatmap and order by the x-axis of PCA
 
```{r}
col.order <- pcaDT[order(PC1), rn]

prot.mat <- prot.mat[, col.order]
colnames(prot.mat)
```
Plot the heatmap
---
As expected, seems that the top cluster drive pc1

Plot a simple protein heatmap
```{r}
submat <- sweep(prot.mat,1, STATS=apply(prot.mat, 1, median, na.rm=T))

# create annotation
ann <- data.table(treatment=sub("[_.].+", "", colnames(prot.mat)),
                  timepoint=ifelse(grepl("[0247]d", colnames(prot.mat)), str_extract(colnames(prot.mat), "[0247]d"), '0d'))

colours <- list('timepoint' = c('0d' = '#4477AA', '2d'="#228833", '4d'="#CCBB44", '7d'='#EE6677'),
                'treatment' =c('Mock' = col.pal[1], 'Vehicle'=col.pal[2], 'x4516'=col.pal[3], 'x4673'=col.pal[4], 'Nirm'=col.pal[5])
                )

colAnn <- HeatmapAnnotation(df = ann, col = colours)

hm <- Heatmap(submat, 
              cluster_rows=clusterWNA(submat), 
              show_row_names = F, 
              cluster_columns = F,
              cluster_column_slices = F,
              row_title = paste0(nrow(submat), ' proteins detected in all samples'),
              column_split = gsub('.[1-6]$','', colnames(submat)),
              column
              #top_annotation = colAnn, 
              name='LogIntensities/Median',  
              column_title_gp = gpar(fontsize=7))
hm
BackupAsPDF(hm, 'prots.medianScaled.colSplit.heatmap', dimensions=c(14,9))
```

```{r}
submat <- sweep(prot.mat,1, STATS=apply(prot.mat, 1, median, na.rm=T))

# create annotation
ann <- data.table(treatment=sub("[_.].+", "", colnames(prot.mat)),
                  timepoint=ifelse(grepl("[0247]d", colnames(prot.mat)), str_extract(colnames(prot.mat), "[0247]d"), '0d'))

colours <- list('timepoint' = c('0d' = '#4477AA', '2d'="#228833", '4d'="#CCBB44", '7d'='#EE6677'),
                'treatment' =c('Mock' = col.pal[1], 'Vehicle'=col.pal[2], 'x4516'=col.pal[3], 'x4673'=col.pal[4], 'Nirm'=col.pal[5])
                )

colAnn <- HeatmapAnnotation(df = ann, col = colours)

hm <- Heatmap(submat, 
              cluster_rows=clusterWNA(submat), 
              show_row_names = F, 
              cluster_columns = F,
              row_title = paste0(nrow(submat), ' proteins detected in all samples'),
              top_annotation = colAnn, 
              name='LogIntensities/Median',  
              column_names_gp = gpar(fontsize=6))
hm
BackupAsPDF(hm, 'proteins.orderbyPC1.medianscaled.complete.cases.heatmap', dimensions=c(10,9))
```
Out of curiosity, what are these proteins involved in? 
Lets cluster the heatmap and extract names of the proteins in the top section

```{r}
hm <- Heatmap(submat, 
              cluster_rows=clusterWNA(submat), 
              show_row_names = F,
              row_split = 3,
              cluster_columns = F,
              #row_title = paste0(nrow(submat), ' proteins detected in all samples'),
              top_annotation = colAnn, 
              name='LogIntensities/Median',  
              column_names_gp = gpar(fontsize=6))
hm <- draw(hm)
```

```{r}
clusters.dt <- extractClustersfromHeatmap(hm, submat)
clusters.dt[,.N, by=cluster] # want cluster 2 for the enrichment

clusters.dt[, gene := multiUniprots2multiGenes(feature, species='MOUSE')]
```
Run enrichment on the 

```{r}

9E10-3
0.009
# load the GO table
gmt.go <- loadGmtFromBioconductor(dbName = 'org.Mm.eg.db', ontology = "ALL", keyType = "SYMBOL")
```

Define universe and perform the enrichment
```{r}
# define the universe, the total set of identified genes in our study
universe <- multiUniprots2multiGenes(unique(rownames(submat)), species='MOUSE')

# now want to run enrichment on each 
mss.dt[,enrich.grp := interaction(Label,sig)]

enrich.dt <- enricherOnGroups(clusters.dt[cluster == 2,], 
                              groupColumns = 'cluster', 
                              geneColumn = "gene", 
                              term2gene.gmt = gmt.go, 
                              universe = universe)


fwrite(enrich.dt, ScriptAndDatedFileName('GOenrichmentsOfCluster2.csv'))

enrich.dt
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'cluster')
fwrite(simp.enrich$simplified, ScriptAndDatedFileName('GOenrichmentsOfCluster2.simplified.csv'))

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, groupColumn = 'cluster.x', topN = 20, title='GO term enrichment', 
                           row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 6), upperThreshold = 6)
ht
BackupAsPDF(ht, 'go.heatmap.cluster2.')
```
# what are we doing here?
# x is the principal components and rotation is the loadings for each of the different PCs
# matrix multiplication
pc1.mat <- pcaOut$x[,1] %*% t(pcaOut$rotation[,1])

```{r}
pcaOut <- prcomp(t(prot.mat))

# what are we doing here?
# x is the principal components and rotation is the loadings for each of the different
pc1.mat <- pcaOut$x[,1] %*% t(pcaOut$rotation[,1])

# this is PCs (pc1) for each of the sample
pcaOut$x[,1:2]
# loadings for each gene (contribution per PC) to each PC
pcaOut$rotation[,1:2]
```


```{r}
pc1.mat <- t(pc1.mat) 

Heatmap(pc1.mat, show_row_names = F)

colnames(pc1.mat) <- col.order

ht <- Heatmap(pc1.mat, 
        cluster_columns = F, 
        top_annotation = colAnn, 
        name='PC1 mat', 
        show_row_names = F,
        column_names_gp = gpar(fontsize=6))

ht
draw(hm + ht)

BackupAsPDF(draw(hm + ht), 'OriAndPC1.heatmap', dimensions=c(12,10))
```
Lets subtract the PC1 matrix from the Ori
Looks much cleaner... this seems like a better example to use as input for the clustering and differential expression

```{r}
smooth.mat <- prot.mat - pc1.mat

subSmooth.mat <- sweep(smooth.mat, 1, STATS=apply(smooth.mat,1, median, na.rm=T))

ht <- Heatmap(subSmooth.mat, 
        cluster_columns = F, 
        top_annotation = colAnn, 
        name='Smooth Matrix', 
        show_row_names = F,
        column_names_gp = gpar(fontsize=6))
ht
BackupAsPDF(draw(ht), 'Smooth.heatmap', dimensions=c(10,8))


BackupAsPDF(draw(hm + ht), 'OriAndSmooth.heatmap', dimensions=c(12,10))
```
Clustering of smooth matrix
Not much more evidence of group clustering... lets regenerate the PCA and visualise

```{r}
Heatmap(smooth.mat, 
        cluster_columns = T, 
        top_annotation = colAnn, 
        name='Smooth Matrix', 
        show_row_names = F,
        column_names_gp = gpar(fontsize=6))

```
plot the heatmap and col split to identify the different groups 

```{r}
hm <- Heatmap(smooth.mat, 
        cluster_columns = F,
        column_split = gsub('[.][1-6]$', '', colnames(smooth.mat)),
        name='Smooth Matrix', 
        show_row_names = F,
        column_title_gp = gpar(fontsize=8),
        column_names_gp = gpar(fontsize=6))

BackupAsPDF(hm, 'smooth.colSplit.heatmap', dimensions=c(12,9))
```
Convert the smooth matrix to longformat and write to file

```{r}
smooth.exp.dt <- reshape2::melt(smooth.mat, value.name ='LogIntensities', varnames =c('Protein','sample'))
setDT(smooth.exp.dt)
smooth.exp.dt[, `:=`(SUBJECT = str_extract(sample, '[1-6]$'),
                     GROUP = gsub('.[1-6]$','', sample)
                     )]

fwrite(smooth.exp.dt, ScriptAndDatedFileName('ProteinLevelQuant.PC1SmoothedValues.csv'))
```

Run GSEA on the principal component loadings

```{r}
library(fgsea)

# conver tthe GO genesets to named list
go.list <- as.character(gmt.go$gene) %>% 
split(., gmt.go$ont)

pc1.rnk <- sort(pcaOut$rotation[,1], decreasing = T)
names(pc1.rnk) <- multiUniprots2multiGenes(names(pc1.rnk), species='MOUSE')

gsea.res <- fgsea(go.list, stats=pc1.rnk, minSize = 10)

#fwrite(gsea.res[padj < 0.05,], ScriptAndDatedFileName('gsea.sigTerms.GOenrichment'))

```

```{r}
gsea.res <- fread('~/Documents/projects/052924_ARichards_AViDD_Murine/PC1_batchRemoval_data/2024_06_04_gsea.sigTerms.GOenrichment')

head(gsea.res)
# plot the gsea results; 

g <- ggplot(gsea.res[padj < 1e-05,], aes(x=pathway, y=NES, size=size, color=-log10(padj))) +
  geom_point() +
  coord_flip() +
  scale_color_viridis(option = "C") +
  ggtitle('PC1 loadings GO GSEA enrichment') +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90))
g
```

```{r}
#BackupAsPDF(g, 'pc1.GO.gsea.dotplot', dimensions=c(9,9))
```

```{r}
pcaOut <- prcomp(t(smooth.mat))

colInfo <- data.table(colname = colnames(smooth.mat))
colInfo[,c("status", "rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]
colInfo[, c("treatment", "timepoint") := tstrsplit(status, '[_]', keep=c(1,2)) ]
colInfo[status == 'Mock', timepoint := '0d']

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)

#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = treatment, shape = timepoint)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  #scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = c(21:26)) +
  scale_fill_manual(values = col.pal) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
#BackupAsPDF(p, 'PCA.proteins.rmPC1.treatandtimepoint')

```
Lets try identify batch from the data; 

Can we pull out genes associated with IM or viral response

```{r}
viral.res.genes <- go.list[which(names(go.list) %like% 'viral')] %>%  
  unlist() %>%  
  as.character()

rownames(submat) <- multiUniprots2multiGenes(rownames(submat), species='MOUSE')

submat <- submat[rownames(submat) %in% viral.res.genes,]
                 
Heatmap(submat, 
        row_title = sprintf('%s viral response proteins', nrow(submat)),
        column_split = gsub('.[1-6]$','', colnames(submat)))
```
07-06-24
---
Retry the batch-removal process with the MSS normalized data, then run tmp if needed
I dont understand why this reordering is occuring.. ask BP Monday.. for now just rerun TMP on the output from the first run

```{r}
p.quant <-  fread('~/Documents/projects/052924_ARichards_AViDD_Murine/AViDD_AB_murine_DE_data/2024_05_31_AB.ProteinLevelData.tmpNormNewCol.csv')

prot.mat <- dcast(p.quant, Protein~paste0(GROUP,'.',SUBJECT), value.var = 'newLogIntensities') %>% 
  as.matrix(rownames='Protein')

prot.mat <- prot.mat[complete.cases(prot.mat),]
```

Rerun PCA

```{r}
col.pal <- randomcoloR::distinctColorPalette(k=5)

pcaOut <- prcomp(t(prot.mat))

colInfo <- data.table(colname = colnames(prot.mat))
colInfo[,c("status", "rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]
colInfo[, c("treatment", "timepoint") := tstrsplit(status, '[_]', keep=c(1,2)) ]
colInfo[status == 'Mock', timepoint := '0d']

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)

#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = treatment, shape = timepoint)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  #scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = c(21:26)) +
  scale_fill_manual(values = col.pal) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, 'PCA.proteins.treatandtimepoint')
```
Get the order by PC1 and reorder the input matrix 

```{r}
col.order <- pcaDT[order(PC1), rn]
prot.mat <- prot.mat[, col.order]
```

```{r}
submat <- sweep(prot.mat,1, STATS=apply(prot.mat, 1, median, na.rm=T))

# create annotation
ann <- data.table(treatment=sub("[_.].+", "", colnames(prot.mat)),
                  timepoint=ifelse(grepl("[0247]d", colnames(prot.mat)), str_extract(colnames(prot.mat), "[0247]d"), '0d'))

colours <- list('timepoint' = c('0d' = '#4477AA', '2d'="#228833", '4d'="#CCBB44", '7d'='#EE6677'),
                'treatment' =c('Mock' = col.pal[1], 'Vehicle'=col.pal[2], 'x4516'=col.pal[3], 'x4673'=col.pal[4], 'Nirm'=col.pal[5])
                )

colAnn <- HeatmapAnnotation(df = ann, col = colours)

hm <- Heatmap(submat, 
              cluster_rows=clusterWNA(submat), 
              show_row_names = F, 
              cluster_columns = F,
              top_annotation = colAnn,
              row_title = paste0(nrow(submat), ' proteins detected in all samples'),
              name='LogIntensities/Median',  
              column_names_gp = gpar(fontsize=7),
              column_title_gp = gpar(fontsize=7))
hm
BackupAsPDF(hm, 'prots.medianScaled.PC1order.annotated.heatmap', dimensions=c(10,9))

```



