---
title: "geneSetEnrichments"
author: "Martin Gordon"
date: "2024-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## z-score 

Gene set enrichmemt for ISG and proinflammatory genes
Get the median/mean z-score, sample randomly same number of genes and calculate median z-score, acquire p-val for how many times they have been found

```{r}
library(magrittr)
library(data.table)
library(circlize)
library(ComplexHeatmap)
library(circlize)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(MSstats)
library(ggrepel)
library(readxl)
library(nichenetr)
library(patchwork) # combine plots




source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")
source("../../utils/bp_utils/Localization_CellMap.R") # functions to enrich for cell pos localisation
source("../../utils/mg_utils/r_utils/HelperFunctions.R") #

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```

Read in the data
```{r}

piGenes <- read_xlsx('docs/1-s2.0-S0092867423009157-mmc4.xlsx', col_names = T, sheet=7) %>% 
  setDT(.) %>% 
  .[category == 'Proinflammatory'] %>% 
  .[,unique(gene)]

isGenes <- fread('./docs/ISGs.txt', header=F) %>% 
  .[,V1]

# p.quant 
p.quant <- fread ("./output/AB.ProteinLevelData.csv")
p.quant[, REP := SUBJECT]
p.quant[, SUBJECT := interaction(GROUP,SUBJECT)]
# for plotting group averages
p.quant[, mean.LogIntensities := mean(LogIntensities, na.rm=T), by=.(Protein,GROUP)]

#msstats output
mss <- fread('./output/mss.contrasts.ab.csv')

#clean the res
mss <- mss[!is.infinite(abs(log2FC)) & !issue %in% c("oneConditionMissing","completeMissing"), ] %>% 
  .[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')] %>% 
  .[, sig := 'not'] %>% 
  .[abs(log2FC) > 0.58 & adj.pvalue < 0.05, sig := ifelse(log2FC > 0, 'up', 'down')]


mss %>%  colnames()
```

Map to human homologs using db on file and remove dups
```{r}
# lets map the homologs to the mss output and write the complete and tidied data to file 
mh_mappings <- fread('~/Documents/utils/mg_utils/data/mouseHumanIDConversion.txt')

new.mss <- merge.data.table(x=mss, y=mh_mappings[,.(Symbol.mouse, Symbol.human, SWISS_PROT_IDs.mouse, SWISS_PROT_IDs.human)], by.x='Protein', by.y='SWISS_PROT_IDs.mouse', all.x=T)
# nearly 7k more rows added... need to filter these out dups.
# just take unqiue rows of DT based on Protein Label combo
mss <- unique(new.mss, by=c("Protein", "Label"))

new.p.quant <- merge.data.table(x=p.quant, y=mh_mappings[,.(Symbol.mouse, Symbol.human, SWISS_PROT_IDs.mouse, SWISS_PROT_IDs.human)], by.x='Protein', by.y='SWISS_PROT_IDs.mouse', all.x=T)
p.quant <- unique(new.p.quant, by=c('Protein', 'GROUP', 'SUBJECT'))
```
 look at histograms of the log2FCs per label
 
```{r}

for (i in unique(mss$Label)){
  g <- ggplot(mss[Label == i,], aes(x=log2FC)) +
    ggtitle(paste0(i, ' log2FC ISG distribution')) +
    geom_histogram() +
    geom_bar(data=mss[Label == i & Symbol.human %in% isGenes,], aes(x=log2FC, y=3000), color='red', fill='red', stat='Identity') +
    theme_classic()
  print(g)
}

# save the plot with all ISG 
g <-  ggplot(mss, aes(x=log2FC)) +
    ggtitle(paste0('log2FC ISG distribution')) +
    geom_histogram() +
    geom_bar(data=mss[Symbol.human %in% isGenes,], aes(x=log2FC, y=3000), color='red', fill='red', stat='Identity') +
    theme_classic() +
    facet_wrap(~Label, ncol=3, scales = 'free')

BackupAsPDF(g, 'log2FC.ISG.distribution', dimensions=c(12,16))



# save the plot with all PI genes
g <-  ggplot(mss, aes(x=log2FC)) +
    ggtitle(paste0('log2FC Pro-Inflamatory gene distribution')) +
    geom_histogram() +
    geom_bar(data=mss[Symbol.human %in% piGenes,], aes(x=log2FC, y=3000), color='blue', fill='blue', stat='Identity') +
    theme_classic() +
    facet_wrap(~Label, ncol=3, scales = 'free')

BackupAsPDF(g, 'log2FC.PIG.distribution', dimensions=c(12,16))

```
 
```{r}
# calculate geom mean of log2FC and pval; we will use this for our computations
#mss[, magSig := magSigScore(log2FC, pvalue)]

# convert this to wide
#mag.mat <- dcast(mss, Protein~Label, value.var = 'magSig') %>% 
#  as.matrix(rownames='Protein')

# how do we consider sign here? Just use log2FC so they dont cancel each other out...
# convert this to wide
lfc.mat <- dcast(mss, Protein~Label, value.var = 'log2FC') %>% 
  as.matrix(rownames='Protein')


#' ColwiseZScorePermutation Function
#'
#' Calculate z-score for a given subset of genes in a matrix column, assess significance with permutations.
#'
#' @param colName The name of the column/sample for which z-scores are calculated.
#' @param mat The input matrix.
#' @param set A set of genes for which the z-score is assessed.
#' @param npermutations Number of permutations.
#' @return A data.table with results.
#'
#' @examples
#' ColwiseZScorePermutation("GeneExpression", myMatrix, geneSet, npermutations = 100)
#'
ColwiseZScorePermutation <- function(colName, mat, set, npermutations=10, center = TRUE, scale = TRUE){
  
  if (!colName %in% colnames(mat)) {
    stop("The specified column name does not exist in the matrix.")
  }
  message(paste0('Working on ', colName, '...'))
  # named vector
  scores <- mat[,colName]

  # remove infinite and na vals 
  scores <- scores[!is.na(scores)]
  if (any (is.infinite(scores))){
    message (sprintf ("Removing %d infinite values from scores for %s", sum(is.infinite(scores)), colName))
     scores <- scores[is.finite(scores)]
  }
  
  # calcultate z-score of the genes 
  scores.z <- scale(scores, center=center, scale=scale)
  names(scores.z) <- names(scores)
  
  sample_size <- sum(names(scores.z) %in% set)
  if ( sample_size < 2){
    message('check geneset..')
    return(NULL)
  }

  # get median value of LFC
  set_score <- scores.z[names(scores.z) %in% set] %>% 
    median(na.rm=T)
  # 
  message(paste0('permuting scores for ', npermutations, ' permutations....'))
   
  perm_cols <-  do.call(cbind, (lapply(1:npermutations, function(x){
    
    # sample length of geneSet
    perm_sample <- sample(names(scores.z), sample_size)
    perm_scores <- scores.z[names(scores.z) %in% perm_sample]
  }))) 
  
  #calculate summary statistics
  message("Summarizing Z statistics")
  permutedMeans <- apply(perm_cols, 2, mean)
  permutedMedians <- apply(perm_cols,2,  median)
  permutedSD  <- apply(perm_cols,2, sd) 
  
  
  # by chance, how often is median LFC exceeded?
  # take abs value of FC 
  deltaLFC <- abs(permutedMedians) - abs(set_score)
  pValue <- sum(deltaLFC > 0)/ncol(perm_cols)
  
  results <- data.table (Label = colName,
                         set.log2FC = set_score,
                         pvalue = pValue,
                         mean.perm.lfc = permutedMeans,
                         median.perm.lfc = permutedMedians,
                         delta.lfc = deltaLFC,
                         sd.perm.lfc = permutedSD)
  return(results)
}

ISset <- mss[Symbol.human %in% isGenes, unique(Protein)]

test <- ColwiseZScorePermutation(colName = "N1062D_7D-N1062D_4D", lfc.mat, set=ISset, npermutations = 10000)
test

```

calculate GS enrichment 

```{r}
## subset mss to contrasts we are interested in
contrasts.oi <- grep('-Mock|x4052|x3769', unique(mss$Label), invert = T, value=T)



ISset <- mss[Symbol.human %in% isGenes, unique(Protein)]

contrasts.zscores <-lapply(colnames(lfc.mat), function(x){
  print(paste0('permuting label ', x ))
  ColwiseZScorePermutation(colName = x, mat = lfc.mat, set=ISset, npermutations = 10000)
})

perm.dt <- rbindlist(contrasts.zscores)
perm.dt %>%  head()

perm.dt[, sig := ifelse(pvalue < 0.05, 'yes', 'not')]

g <- ggplot(perm.dt[Label %in% contrasts.oi,], aes(x=Label, y=median.perm.heat)) +
  geom_boxplot() +
  geom_point(data=perm.dt[Label %in% contrasts.oi,], aes(x=Label, y=set.log2FC, color='median log2FC z-score'), size=2) + 
  ylab('permuted log2FC z-scores') +
  xlab('Contrast') +
  #labs(color='median log2FC (z-score)') +
  scale_color_manual(values=c('median log2FC z-score' = 'darkred')) +
  #ggrepel::geom_text_repel(data=perm.dt[sig != 'not',], show.legend = FALSE, size = 2, max.overlaps = 20) +
  #scale_color_manual(values=c('not'='darkgrey', 'yes'='#440154FF')) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90))
BackupAsPDF(g, 'isg.zscore.distributions.boxplot')

# plot the histograms
g <-  ggplot(perm.dt[Label %in% contrasts.oi,], aes(x=median.perm.lfc)) +
  geom_histogram() +
  geom_vline(data=perm.dt[Label %in% contrasts.oi], aes(xintercept = set.log2FC), color='darkred') +
  theme_classic() +
  facet_wrap(~Label, ncol=3, scales = 'free')

BackupAsPDF(g, 'isg.zscore.distributions.histograms') 
```
look at the heatmap... does this look correct?

```{r}
subset <- lfc.mat[rownames(lfc.mat) %in% ISset, colnames(lfc.mat) %in% contrasts.oi]
rownames(subset) <- multiUniprots2multiGenes(rownames(subset), species='MOUSE')

BackupAsPDF(Heatmap(subset), 'isg.lfc.heatmap', dimensions=c(8,8))
```
now run for the PI genes 

```{r}
PIset <- mss[Symbol.human %in% piGenes, unique(Protein)]

contrasts.zscores <- lapply(colnames(lfc.mat), function(x){
  print(paste0('permuting label ', x ))
  ColwiseZScorePermutation(colName = x, mat = lfc.mat, set=PIset, npermutations = 10000)
})

perm.dt <- rbindlist(contrasts.zscores)
perm.dt %>%  head()

perm.dt[, sig := ifelse(pvalue < 0.05, 'yes', 'not')]

g <- ggplot(perm.dt[Label %in% contrasts.oi,], aes(x=Label, y=median.perm.lfc)) +
  geom_boxplot() +
  geom_point(data=perm.dt[Label %in% contrasts.oi,], aes(x=Label, y=set.log2FC, color='median log2FC z-score'), size=2) + 
  ylab('permuted log2FC z-scores') +
  xlab('Contrast') +
  #labs(color='median log2FC (z-score)') +
  scale_color_manual(values=c('median log2FC z-score' = 'darkred')) +
  #ggrepel::geom_text_repel(data=perm.dt[sig != 'not',], show.legend = FALSE, size = 2, max.overlaps = 20) +
  #scale_color_manual(values=c('not'='darkgrey', 'yes'='#440154FF')) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90))
BackupAsPDF(g, 'proinflammatorygenes.zscore.distributions.boxplot')


#show the distirbutions of the medians and overlay with recorded

# plot the histograms
g <-  ggplot(perm.dt[Label %in% contrasts.oi,], aes(x=median.perm.lfc)) +
  geom_histogram() +
  geom_vline(data=perm.dt[Label %in% contrasts.oi], aes(xintercept = set.log2FC), color='darkred') +
  theme_classic() +
  facet_wrap(~Label, ncol=3, scales = 'free')

BackupAsPDF(g, 'proinflammatroygenes.zscore.distributions.histograms') 

```





