---
title: "WGCNA.exploratory"
author: "Martin Gordon"
date: "2024-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## WGCNA EDA;
INterested in discovering the mechanisms involved in treatment group


First lets, just read
```{r}
library(data.table)
library(ggplot2)
library(ComplexHeatmap)
library(WGCNA)
library(magrittr)
library(preprocessCore) # quantile normalisation
library(stringr)
library(matrixStats)
library(simplifyEnrichment)
library(fmsb)
library(readxl)
library(ggbeeswarm)
library(ggrepel)
library(matrixStats)

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")

```
read in the normalised data 
since we have a complete matrix and many


```{r}
p.quant <-fread('./output/AB.ProteinLevelData.csv')

p.quant[GROUP == 'Mock', GROUP := 'Mock_ND']

p.mat <-  dcast(p.quant,Protein~interaction(GROUP,SUBJECT), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')

ncol(p.mat)
# find rows with many NA values
na.counts <- apply(p.mat, 1, function(x){ 
  vec <- sum(is.na(x))
  return(vec)})


# drop proteins with more than half missing values
# only keep rows with 80% completeness
drop.rows <- names(na.counts)[na.counts > 10]
p.mat <- p.mat[!rownames(p.mat) %in% drop.rows,]

```


```{r}
hm <- Heatmap(p.mat, show_row_names = F, cluster_columns = F,
              column_split = list(sub('_[74N]D[.][1-9]$', '', colnames(p.mat)),
                                                       str_extract(colnames(p.mat), '[74N]D')
                                                       ))
BackupAsPDF(hm, 'wgcna.inputmat')


# look at the data with the median subtracted
submat <- sweep(p.mat, 1, apply(p.mat,1, median, na.rm=T))

hm <- Heatmap(submat, show_row_names = F, cluster_columns = F,
              name='Ints/Median',
              column_split = list(sub('_[74N]D[.][1-9]$', '', colnames(submat)),
                                                       str_extract(colnames(submat), '[74N]D')
                                                       ))

hm
BackupAsPDF(hm, 'wgcna.inputmat.mediansweep.')

```
visualise mock vs all, this is what we will use for our clustering and drop the mock samples


```{r}
# subtract the Veh4 median from the other samples
submat <- sweep(p.mat, 1, apply(p.mat[,grepl('Vehicle_4D', colnames(p.mat))], 1, median, na.rm=T))

hm <- Heatmap(submat, show_row_names = F, cluster_columns = F,
              name='Ints/Median',
              cluster_rows = clusterWNA(submat),
              column_split = list(sub('_[74N]D[.][1-9]$', '', colnames(submat)),
                                                       str_extract(colnames(submat), '[74N]D')
                                                       ))
hm
BackupAsPDF(hm, 'wgcna.inputmat.VehD4sweep.', dimensions=c(12,12))
```
Lets look at variance of each gene
This would 
```{r}
var.dt <-  data.table(Protein = rownames(submat),
                      Var = apply(submat, 1, var, na.rm=T))[order(-Var)]
var.dt[, id := .I]

g <- ggplot(var.dt, aes(x=id, y=log2(Var))) +
  geom_line()
g
BackupAsPDF(g, 'ordered.Variance.decreasing')
```
```{r}
submat <- submat[, grep('Vehicle_4D|Mock', colnames(submat), invert=T)]

hm <- Heatmap(submat, show_row_names = F, cluster_columns = F,
              name='Ints/Median',
              cluster_rows = clusterWNA(submat),
              column_split = list(sub('_[74N]D[.][1-9]$', '', colnames(submat)),
                                                       str_extract(colnames(submat), '[74N]D')
                                                       ))
hm
BackupAsPDF(hm, 'wgcna.inputmat.VehD4sweep.noMock.', dimensions=c(12,12))
```
QC
----
Check sample clustering based on our input gene list 
One of the 4206 treatments is quite different but cant drop as important..

```{r}
#Group data in a dendogram to check outliers
sampleTree = hclust(dist(t(submat)), method = "average")

#pdf(file = "./WGCNA_data/sampleClustering.pdf", width = 12, height = 9); 
par(cex = 0.6)
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering", sub="", xlab="", cex.lab = 1.5, 
     cex.axis = 1.5, cex.main = 2)

#dev.off()
```


```{r}
gsg <- goodSamplesGenes(t(submat))
summary(gsg)
gsg$allOK
```

All genes look good; lets cluster 

```{r}
net.in <- t(submat)

s.thres <- data.table(pickSoftThreshold(net.in)$fitIndices)

s.thres %>%  
  ggplot(aes(x=Power, y=SFT.R.sq)) + 
  geom_point() +
  geom_hline(yintercept = 0.8, color='darkred') +
  labs(y="Scale Free Topology Model Fit, signed R^2") +
  theme_classic()

s.thres %>%  
  ggplot(aes(x=Power, y=mean.k.)) + 
  geom_point() +
  labs(y='Mean Connectivity') +
  theme_classic() 

s.thres
```
Creating adjacency matrix
Our network is signed; 0 anti-correlated,1 is strongly correlated

Adjusting to soft threshold of 4
A balance between good R2 and a decent K score
```{r}
softPower = 6 #Chosen in the graphs before
adj.mat = adjacency(net.in, power = softPower, corFnc = 'bicor', type = "signed") #Calculating the adjacency matrix
#help(adjacency )

subset <- sample(rownames(adj.mat), 3000)

dim(adj.mat[rownames(adj.mat) %in% subset, colnames(adj.mat) %in% subset])

# plot a subset of the adjacency matrix
#Heatmap(adj.mat[rownames(adj.mat) %in% subset, colnames(adj.mat) %in% subset], show_row_names = F, show_column_names = F)
```
Create a topological oveerlpa matrix

```{r}
TOM <- TOMsimilarity(adj.mat)
TOM.dissimilarity <- 1 - TOM
```

Identify modules in our data
Perform hierarchical clustering of the TOM dissimiliarity measure to identify modules

```{r}
#TOM.dissimilarity <- 1-TOM
adj.dissimilarity <- 1 - adj.mat

#creating the dendrogram 
geneTree <- hclust(as.dist(TOM.dissimilarity), method = "average") 
#plotting the dendrogram
sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity", 
labels = FALSE, hang = 0.04)
```


```{r}
#only detected 4 modules... not enough variability in the data 
Modules <- cutreeDynamic(dendro = geneTree, distM = TOM.dissimilarity, deepSplit = 4, pamRespectsDendro = FALSE, minClusterSize = 30)
table(Modules)
```

```{r}
Modules
# convert to char vector
modules <- as.character(Modules)
```
Heatmap of the modules 
---
```{r}
col.pal <- randomcoloR::distinctColorPalette(length(unique(Modules)))


# add wgcna module annotation
row_ha <- rowAnnotation(wgcna.modules=modules, col=list(wgcna.modules = c("1"=col.pal[1], '2'=col.pal[2], '3'=col.pal[3], '4'=col.pal[4], '5'=col.pal[5], 
                                                                              '6'=col.pal[6], '7'=col.pal[7], '8'=col.pal[8], '9'=col.pal[9])))
# add split order
split <- factor(modules, levels=c('1','2','3','4','5','6', '7','8','9'))

hm <- Heatmap(submat, 
              cluster_columns = FALSE, 
              show_row_names = FALSE,
              split = split, # split rows on module assignment
              right_annotation = row_ha,
              name = "log2(intensity/\nmean(Vehicle_D4))",
              column_split = list(tstrsplit(colnames(submat), "[_.]")[[1]],
                                  tstrsplit(colnames(submat), "[_.]")[[2]]),
              na_col = 'darkgrey',
              row_title =sprintf("%d Proteins",nrow(submat)),
              # row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize= 3),
              cluster_row_slices = F,
              column_title_gp = gpar (fontsize= 7))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))

BackupAsPDF(hm, 'heatmap.wgcna.modules.ordered', dimensions=c(10,14))
```

Might be something in these modules.. 
lets try summarise the groups and visualise medians

```{r}
# find the 
prots.oi <- rownames(submat)

sum.p.mat <- dcast(p.quant, Protein~GROUP, value.var='LogIntensities', fun.aggregate = median, na.rm=T) %>% 
  as.matrix(rownames='Protein')
sum.submat <- sum.p.mat[rownames(sum.p.mat) %in% prots.oi,]

# subtract the Veh4 median and rm the unwanted cols
sum.submat <- sweep(sum.submat,1, sum.submat[, grepl('Vehicle_4D', colnames(sum.submat))])
sum.submat <- sum.submat[, grep('Vehicle_4D|Mock', colnames(sum.submat), invert=T)]

#sanity check...
all(rownames(sum.submat) == rownames(submat))

# create the plot
hm <- Heatmap(sum.submat, 
              cluster_columns = TRUE, 
              cluster_rows = T,
              show_row_names = FALSE,
              #split = split, # split rows on module assignment
              #right_annotation = row_ha,
              name = "log2(intensity/\nmean(Vehicle_D4))",
              na_col = 'black',
              row_title =sprintf("%d Proteins",nrow(sum.submat)),
              # row_names_gp = gpar (fontsize= 1),
              column_names_gp = gpar (fontsize= 7),
             # cluster_row_slices = T,
              column_title_gp = gpar (fontsize= 7))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))


BackupAsPDF(hm, 'heatmap.wgcna.modules.ordered', dimensions=c(10,14))
```
Clustering doesnt look great... not enough variation in the dataset.. lets try quickly enrich to find anything interesting, else just leave as is..

```{r}
# helper functions including script to extract heatmap genes
source("/Users/martingordon/Documents/utils/mg_utils/r_utils/HelperFunctions.R")

clust.genes <- extractClustersfromHeatmap(hm, submat)
clust.genes[,.N, by=cluster]

# sanity check.. 
all(clust.genes[cluster==7, feature] %in% rownames(submat)[Modules==7])
clust.genes[, gene := multiUniprots2multiGenes(feature, species='MOUSE')]

# load the GO table
gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL", dbName = 'org.Mm.eg.db')

# our background? only proteins used for WGCNA analysis
# maybe some interesting things in the smaller modules
universe <- clust.genes[, unique(gene)]

enrich.dt <- enricherOnGroups(clust.genes, groupColumns = 'cluster', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)

enrich.dt[p.adjust < 0.05 & !cluster %in% c('1','2')]
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'Cluster')

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt, groupColumn = 'cluster', topN = 10, title='GO enrichment WGCNA modules', 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 6), upperThreshold = 8, cluster_columns =F)

BackupAsPDF(ht, 'go.enrichment.wgcna.modules.adj.mat.ordered', dimensions=c(8,12))
```
lets look at a heatmap of Module4; how does the expression profile of these proteins look?
Actually this is kind of interesting.... seems that the response is much more strongly upregulated in our treatment OI than the other groups

```{r}
mod4.mat <-  submat[rownames(submat)[Modules==4], ]

rownames(mod4.mat) <- multiUniprots2multiGenes(rownames(mod4.mat), species='MOUSE')

hm <- Heatmap(mod4.mat, 
        cluster_rows = T,
        name='Ints. vs Veh D4',
        row_names_gp = gpar(fontsize=5),
        column_split = list(gsub("[_].+", "", colnames(mod4.mat)),
                            str_extract(colnames(mod4.mat), "[47]D"))
                            )

hm <- draw(hm, column_title="WGCNA cluster 4",
           column_title_gp=grid::gpar(fontsize=10))

BackupAsPDF(hm, 'heatmap.wgcna.module4', dimensions=c(12,14))


```




