---
title: "011325_ProteinSummarisation"
author: "Martin Gordon"
date: "2025-01-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To compare the different lysis methods, we want to first summarise the features to proteins 
3 lysis methods being compared, Spectronaut used for the search
In this case dont want a combined search, as we don't want to 'impute' intensity values predicted in one run but not the other

```{r}
library(MSstats)
library(data.table)
library(ggplot2)
library(magrittr)
library(stringr)


# useful utility scripts
source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")


source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("../../utils/mg_utils/r_utils/msstats_helperFunctions.R")

clusterWNA <-  function(mat, na.value=0,...){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}

customTheme <-  theme_ipsum_rc() +
  theme(axis.text.x = element_text(angle=90),
        panel.border = element_rect(color = "lightgrey",
                                    fill = NA,
                                    size = 1))
```

Read in the runs; we should have 3 reps in each msstats file
```{r}
lysisPrep.list <- list('sonic'=fread('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/data/Spectronaut_Final_010224/20241227_104957_MM_Sonic_Lysis_122724_PEP_Report.tsv'),
                       'FT'=fread('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/data/Spectronaut_Final_010224/20241227_104610_MM_FT_Lysis_122724_PEP_Report.tsv'),
                       'CL'=fread('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/data/Spectronaut_Final_010224/20241227_104259_MM_CL_Lysis_122724_PEP_Report.tsv')
                       )


lysisPrep.dt[Run == '[1] Sonic-1_Frac-1.PG.MS1Quantity',.N by=.(P)]
```

convert files to long format

```{r}
lysisPrep.list <- lapply(lysisPrep.list, function(x){
  
  x.long <- reshape2::melt(x, id.vars=c('PG.MolecularWeight', 'PG.ProteinAccessions', "PG.Genes", "PG.ProteinDescriptions", "EG.PrecursorId"))
  head(x.long)
  return(x.long)
})
```

look at an example MSS input file 
```{r}
test.f <- fread('../121924_JBatra_SARS2_HuBatComparison_ABPHProteomics/data/010425_ViralProtsInDBSearch/MRC5_AB/MSstats_20250104_141826_JB_MRC5_AB_rerun_010425_Report.tsv')
cols.oi <- colnames(test.f)
```

Read a MSStats in example file format so we know what to use for the summarisation
```{r}
lysisPrep.list <- lapply(lysisPrep.list, setDT)

lapply(lysisPrep.list, function(x){
  setnames(x, old=c('variable', 'value'), new=c('sample', 'Intensity'))
})

lapply(lysisPrep.list, head)
```

Now add an identifier to the table, and r bind to make it easier to 

```{r}
lysisPrep.list[['sonic']][, Condition := 'sonic']
lysisPrep.list[['FT']][, Condition := 'FT']
lysisPrep.list[['CL']][, Condition := 'CL']

lapply(lysisPrep.list, function(x){
  # add necessary MSS columns
  x[, BioReplicate := ifelse(grepl('Sonic-1_Frac|FT-1_Frac-|CL-1_Frac-', sample), 1, 2)]
})
```
bind the data tables together to make the processing easier 

```{r}
lysisPrep.dt <- rbindlist(lysisPrep.list)
rm(lysisPrep.list)

lysisPrep.dt[, PG.MolecularWeight := NULL]
lysisPrep.dt[, PrecursorCharge :=  as.numeric(str_extract(EG.PrecursorId, '[0-9]{1,2}$'))]
lysisPrep.dt[, BioReplicate := ifelse(grepl('Sonic-1_Frac|FT-1-Frac-|CL-1-Frac-', sample), 1, 2)]
lysisPrep.dt[, FragmentIon := NA]
lysisPrep.dt[, PG.ProteinDescriptions := NULL]
lysisPrep.dt[, fraction := str_extract(sample, '(?<=Frac-)[0-9]{1,2}')] #pos lookbehind to extract the sample id
lysisPrep.dt[,  IsotopeLabelType := 'L']

setnames(lysisPrep.dt, old=c('PG.ProteinAccessions',  'PG.Genes'), new=c('ProteinName', 'gene'))

lysisPrep.dt[, PeptideSequence := gsub('[.][0-9]+$', '', EG.PrecursorId)]
lysisPrep.dt[,.N, by=.(Condition, BioReplicate)]


#(?<=a)b # only matches b preceded by an a
lysisPrep.dt[, fraction := str_extract(sample, '(?<=Frac-)[0-9]{1,2}')]
lysisPrep.dt[,.N,by=.(Condition, BioReplicate, fraction)][, .SD[which.min(fraction)], by=.(Condition,BioReplicate)]
```

For the protein level summarisation, I think we want to consider each fraction as its own condition, with two replicates each,
do we want to normalize all these searches together?
```{r}
#fwrite(lysisPrep.dt[, .(Condition,Run=sample,BioReplicate, ProteinName, PeptideSequence, FragmentIon, PrecursorCharge,Intensity, ProductCharge=NA, fraction, IsotopeLabelType)], ScriptAndDatedFileName('combined.secms.input.csv'))

lysisPrep.dt <- fread('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/011324_proteinSummarization_data/2025_01_13_combined.secms.input.csv')
```



```{r}
# split the dt into its conditioanal components
spec.list <- split(lysisPrep.dt, lysisPrep.dt$Condition)

names(spec.list)

spec.list <- lapply(spec.list, function(x){
  
  x[, Condition :=  paste0(Condition, '.', fraction)]
  x[, fraction := NULL]
  x[, Run := paste0(Condition, '.', BioReplicate, '.raw')]
  return(x)
})

```
Question here is what do we consider a sample. I think a sample should be the fraction/condition combo


```{r}

spec.list[['sonic']][!is.na(Intensity)]

lapply(spec.list, str)
lapply(spec.list, head)

dp.out <- lapply(names(spec.list), function(x){

  message('Running MSStats dataProcess on ', x)
  
  dp.out <- MSstats::dataProcess(spec.list[[x]], 
                                MBimpute =  FALSE, 
                                normalization = FALSE, # turn off for SECMS
                                summaryMethod = "TMP",
                                featureSubset = 'highQuality',
                                remove_uninformative_feature_outlier=T)

  saveRDS(dp.out, paste0('./output/', x, '.mss.dp.rds'))
  return(dp.out)
})
```


```{r}
CL.ms <- readRDS('./output/CL.mss.dp.rds')
f.quant <- setDT(CL.ms$FeatureLevelData)
p.quant <- setDT(CL.ms$ProteinLevelData)

f.quant[!is.na(ABUNDANCE), .N, by=.(PROTEIN, RUN)][N > 10][order(N)]
CL.ms$FeatureLevelData[PROTEIN == 'O00213' & RUN == 76]


CL.ms$FeatureLevelData


f.quant <- f.quant[!is.na(INTENSITY),.(RUN, PROTEIN, F_INTENSITY=log2(INTENSITY))] %>% unique()
p.quant <- p.quant[!is.na(LogIntensities),.(RUN, PROTEIN=Protein, INTENSITY=LogIntensities)]

p.quant <- merge(p.quant, f.quant, by=c('RUN', 'PROTEIN'))

```

```{r}
test[,.N, by=.(PROTEIN,RUN)]
test[PROTEIN == 'O00213' & RUN == 10]
```


read in the new dataproc output and check out the data
```{r}
dp.out <- list('sonic' = readRDS('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/011324_proteinSummarization_data/2025_01_13_sonic.mss.dp.rds'),
               'FT'= readRDS('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/011324_proteinSummarization_data/2025_01_13_FT.mss.dp.rds'),
               'CL' = readRDS('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/011324_proteinSummarization_data/2025_01_13_CL.mss.dp.rds'))
```
Filter out the proteins with only feature used in the identification
Use the script to identify them

Worrying that the subject field is different in all 3? Maybe try rerunnning this overnight.. for now fix, merge and write output
```{r}
dp.out <- lapply(dp.out, function(x){
  
  x$FeatureLevelData <- setDT(x$FeatureLevelData)
  x$ProteinLevelData <- setDT(x$ProteinLevelData)
  
  keys.dt <- unique(x$ProteinLevelData[, .(originalRUN, newGROUP=GROUP)])
  
  x$FeatureLevelData <- merge(x$FeatureLevelData, keys.dt, by='originalRUN', all.x=T)
  return(x)
})
```

Seems a lot of feat level data exists in this with NA values
```{r}
dp.out[[1]]$FeatureLevelData[,.(originalRUN,RUN, GROUP,newGROUP,SUBJECT)] %>% unique()
dp.out[[2]]$FeatureLevelData[,.(originalRUN,RUN, GROUP,newGROUP,SUBJECT)] %>% unique()
dp.out[[3]]$FeatureLevelData[,.(originalRUN,RUN,GROUP,newGROUP,SUBJECT)] %>% unique()

dp.out[[1]]$FeatureLevelData[, GROUP := newGROUP]
dp.out[[2]]$FeatureLevelData[, GROUP := newGROUP]
dp.out[[3]]$FeatureLevelData[, GROUP := newGROUP]

dp.out[[1]]$ProteinLevelData[,.N, by=.(GROUP,SUBJECT)][order(GROUP)]
dp.out[[2]]$ProteinLevelData[,.N, by=.(GROUP,SUBJECT)][order(GROUP)]
dp.out[[3]]$ProteinLevelData[,.N, by=.(GROUP,SUBJECT)][order(GROUP)]
# fix labelling in 3rd set
dp.out[[3]]$ProteinLevelData[, GROUP := gsub('[.][0-9]{1,2}$', '', GROUP)]
dp.out[[3]]$FeatureLevelData[, GROUP := gsub('[.][0-9]{1,2}$', '', GROUP)]
```

```{r}
dp.out[[3]]$ProteinLevelData[,.N, by=.(GROUP,SUBJECT)][order(GROUP)]
dp.out[[2]]$ProteinLevelData[,.N, by=.(GROUP,SUBJECT)][order(GROUP)]
dp.out[[1]]$ProteinLevelData[,.N, by=.(GROUP,SUBJECT)][order(GROUP)] # sonic.1 biorep1 and sonic.2 rep 1, sonic.4.rep 1, sonic.5 rep 1, sonic6.rep1,

```
ok, maybe just proceed as is for now with this...
Write out the 6 files as is

```{r}
dp.out <- lapply(dp.out, function(x){
  
  x$FeatureLevelData <- setDT(x$FeatureLevelData)
  x$ProteinLevelData <- setDT(x$ProteinLevelData)
  
  x$ProteinLevelData <- getNPeptidesPerProtein(proteinlvlDT = x$ProteinLevelData, x$FeatureLevelData, peptideThreshold = 2)
  return(x)
})

# filter to just keep the set with two peptides
dp.out[[1]]$ProteinLevelData <- dp.out[[1]]$ProteinLevelData[nPeptides >= 2,]
dp.out[[2]]$ProteinLevelData <- dp.out[[2]]$ProteinLevelData[nPeptides >= 2,]
dp.out[[3]]$ProteinLevelData <- dp.out[[3]]$ProteinLevelData[nPeptides >= 2,]


dp.out[[1]]$ProteinLevelData
dp.out[[1]]$FeatureLevelData[PROTEIN == 'A0A024RBG1' ]



dp.out[[2]]$FeatureLevelData[PROTEIN == 'O60232']


dp.out[[2]]$ProteinLevelData[nPeptides > 10,][order(NumMeasuredFeature)]

dp.out[[3]]$FeatureLevelData[PROTEIN == 'A1KXE4']
```
look at another test rds obj

```{r}
test <- setDT(readRDS('../121924_JBatra_SARS2_HuBatComparison_ABPHProteomics/080125_AB_PWcomparisons_data/MRC5.rds')$FeatureLevelData)
test[,.N, by=.(originalRUN,PROTEIN)]

test[PROTEIN == 'A0JNW5' & RUN == 10,]
```

```{r}
ggplot(dp.out[[1]]$ProteinLevelData, aes(x=GROUP, y=LogIntensities)) +
  geom_boxplot() +
  facet_wrap(~SUBJECT) +
  theme(axis.text.x = element_text(angle=90))
```

```{r}
secms.in <- fread('011324_proteinSummarization_data/2025_01_13_combined.secms.input.csv')
secms.in[,.N, by=.(Condition,BioReplicate,fraction)][order(fraction)][,.N, by=fraction]
secms.in
```



```{r}
dp.out <- lapply(dp.out, function(x){
  
  x$ProteinLevelData <- setDT(x$ProteinLevelData)
  x$FeatureLevelData <- setDT(x$FeatureLevelData)
  
  return(x)
  #getNPeptidesPerProtein(proteinlvlDT = x$ProteinLevelData, petidelvlDT = x$FeatureLevelData, peptideThreshold=2)
})
```
011425- read in reprocessed data

```{r}
test <- fread('~/Library/CloudStorage/Box-Box/Spectronaut_Final_010224/20250114_135820_MM_CL_Lysis_122724_MSStat_Report.tsv')
test.fill <- fread('~/Library/CloudStorage/Box-Box/Spectronaut_Final_010224/20250114_135820_MM_CL_Lysis_122724_MSStat_Report.tsv', fill=T)
```
