---
title: "012125_proteinSummarization"
author: "Martin Gordon"
date: "2025-01-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(MSstats)
library(data.table)
library(ggplot2)
library(magrittr)
library(stringr)
library(ComplexHeatmap)
library(viridis)
library(ggrepel)
library(stringr)
library(ComplexHeatmap)
library(circlize)
library(RColorBrewer)
library(hrbrthemes) # visually nice set of themes
library(patchwork)
library(showtext)
library(readxl)



# fonts needed for hrbrthemes
library(extrafont)
font_import()
loadfonts()

# directly from google font
sysfonts::font_add_google("Roboto Condensed")
showtext_auto()


# useful utility scripts
source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")


# fucntions to remove proteins with less than N peptides
source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("../../utils/mg_utils/r_utils/msstats_helperFunctions.R")

clusterWNA <-  function(mat, na.value=0,...){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```
read in the spectronaut reports and share the summarized data
```{r}
lysisPrep.list <- list('sonic'=fread('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/data/Spectronaut_Final_010224/20250115_092735_MM_Sonic_Lysis_122724_MSStats_Report.tsv'),
                       'FT'=fread('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/data/Spectronaut_Final_010224/20250114_172438_MM_FT_Lysis_122724_MSStats_Report.tsv'),
                       'CL'=fread('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/data/Spectronaut_Final_010224/20250114_135820_MM_CL_Lysis_122724_MSStat_Report.tsv'))
```

Check that these proteins have more than one feature

```{r}
lapply(lysisPrep.list, head)

lapply(lysisPrep.list, function(x){
  x[,.N, by=.(Run,BioReplicate,Condition)]
})
```
Ok, lets create some new col ids 
```{r}
lapply(lysisPrep.list, function(x){
  
  x[, Fraction := str_extract(Run, '(?<=Frac-)[0-9]{1,2}')]
  x[, Condition :=  paste0(Condition, '.', Fraction)] 
  x[, IsotopeLabelType := 'L']
})

# looking at some missing fractions in the two groups.. lets summarize first then remove
lapply(lysisPrep.list, function(x){
  x[,.N, by=.(BioReplicate,Condition)][,.N, by=Condition][N < 2,]
})
```

```{r}
lapply(lysisPrep.list, head)

lapply(lysisPrep.list, function(x){

  x[, Fraction := NULL]
  return(x)
})
```

Look at the distribution of intensity values; no -ve values so keep all

```{r}
lapply(lysisPrep.list, function(x){
  
  hist(log2(x$Intensity))
})
```

Look at the intensity distributions

```{r}
lysisPrep.dt[, Fraction := factor(Fraction, levels=as.character(seq(1,72,1)))]
lysisPrep.dt[, Group := paste0(prep, '.', BioReplicate)]


g <- ggplot(lysisPrep.dt, aes(x=Fraction, y=log2(Intensity), fill=prep)) +
  geom_boxplot() +
  facet_wrap(~Group, ncol=2, scales='free') +
  scale_fill_ipsum() +
  theme_ipsum() +
  customTheme
g
BackupAsPDF(g, 'log2Ints.hitogram', dimensions=c(20,8), format='png')
```


Sanity check qc plots

```{r}
lysisPrep.dt <- rbindlist(lysisPrep.list, idcol='prep')
```


```{r}
lysisPrep.dt[, Fraction := factor(Fraction, levels=as.character(seq(1,72,1)))]
lysisPrep.dt[, Group := paste0(prep, '.', BioReplicate)]


g <- ggplot(lysisPrep.dt, aes(x=Fraction, y=log2(Intensity), fill=prep)) +
  geom_boxplot() +
  facet_wrap(~Group, ncol=2, scales='free') +
  scale_fill_ipsum() +
  theme_ipsum() +
  customTheme
g
BackupAsPDF(g, 'log2Ints.hitogram', dimensions=c(20,8), format='png')
```

plot of N peptides detected per run

```{r}
# true; single feautre (pepseq + charge) elution per peak, so we can just count the number of rows
lysisPrep.dt[,.N, by=.(prep,Run,Fraction, PeptideSequence, PrecursorCharge)]$N %>% unique() == 1

g <- ggplot(lysisPrep.dt[,.(nFeaturesPerRun=.N), by=.(Group,prep,Fraction)], aes(x=Fraction, y=nFeaturesPerRun, fill=prep)) +
  geom_bar(stat='Identity') +
  facet_wrap(~Group, ncol=2, scales='free_x') +
  scale_fill_ipsum() +
  theme_ipsum() +
  customTheme
g
BackupAsPDF(g, 'nFeaturesperRun.barplot', dimensions=c(20,8), format='png')
```


```{r,eval=FALSE}
dp.out <- lapply(names(lysisPrep.list), function(x){

  message('Running MSStats dataProcess on ', x)
  
  dp.out <- MSstats::dataProcess(lysisPrep.list[[x]], 
                                MBimpute =  FALSE, 
                                normalization = FALSE, # turn off for SECMS
                                summaryMethod = "TMP",
                                featureSubset = 'highQuality',
                                remove_uninformative_feature_outlier=T)

  saveRDS(dp.out, paste0('./output/', x, '.mss.dp.rds'))
  return(dp.out)
})
```

load in the completed runs

```{r}
dp.out <- list('sonic'= readRDS('./output/sonic.mss.dp.rds'),
               'FT' = readRDS('./output/FT.mss.dp.rds'),
               'CL' = readRDS('./output/CL.mss.dp.rds'))

dp.dt <- rbindlist(dp.out)
# a large number of proteins have very low intensity; want to include these as do
dp.dt[LogIntensities < 0, .SD[which.min(LogIntensities)], by=.N]
```
Find  proteins with 1 feature measured across all fractions and remove 
```{r}
oneCountPeptides <- lapply(dp.out, function(x){
  
  dt <- setDT(x$ProteinLevelData)
  print(dt)
  mat <- dcast(dt, Protein~paste(GROUP, '.', SUBJECT), value.var='NumMeasuredFeature') %>% 
    as.matrix(rownames='Protein')
  
  # convert NA to zero
  mat[is.na(mat)] <-  0
  
  # identify proteins with max 1 feature ID per row
  singlePeps <- apply(mat[1:10,],1, function(x){ max(x) == 1 })
  return(singlePeps[singlePeps == TRUE])
  
})

# filter out the 1 feature prots 
prot.list <- lapply(names(dp.out), function(x){

  dp.out[[x]]$ProteinLevelData <- dp.out[[x]]$ProteinLevelData[!Protein %in% names(oneCountPeptides[[x]]),]
  return(dp.out[[x]]$ProteinLevelData)
})
names(prot.list) <- names(dp.out)
```
convert from long format to matrices
```{r}
p.mat.list <- lapply(prot.list,  function(x){
  
  p.quant <- setDT(x)
  
  mat.1 <- dcast(p.quant[SUBJECT == 1, ], Protein~GROUP, value.var='LogIntensities') %>% 
    as.matrix(rownames='Protein')
  mat.2 <- dcast(p.quant[SUBJECT == 2, ], Protein~GROUP, value.var='LogIntensities') %>% 
    as.matrix(rownames='Protein')
  
  colnames(mat.1) <- gsub('[A-Za-z]+[.]', '', colnames(mat.1))
  colnames(mat.2) <- gsub('[A-Za-z]+[.]', '', colnames(mat.2))

  colorder <- as.character(seq(1,72,1))
  
  message('reordering matrix columns and filling in missing fractions...') 
  # fills in missing fractions with NA so easy to populate
  mat.1 <- mat.1[,  match(colorder, colnames(mat.1))]
  mat.2 <- mat.2[,  match(colorder, colnames(mat.2))]

  colnames(mat.1) <- colorder
  colnames(mat.2) <- colorder
  
  mat.1 <- 2^mat.1
  mat.2 <- 2^mat.2
  
  # convert to 0
  mat.1[is.na(mat.1)] <- 0
  mat.2[is.na(mat.2)] <- 0
  
  return(list(biorep1=mat.1, biorep2=mat.2))
})

```

flatten list to one level (sample=prep x biorep)
```{r}
# unlist 1 level
p.mat.list <- unlist(p.mat.list, recursive = F)
```

Format to PCP specifications
```{r}
lapply(names(p.mat.list), function(x,i){

  p.mat <- data.table(x[[i]], keep.rownames = T)
  setnames(p.mat, old='rn', new='ID')
  p.mat[, GN := multiUniprots2multiGenes(ID, species='HUMAN')]
  
  col.order <- c('GN', 'ID', as.character(seq(1,72,1)))
  p.mat <-  p.mat[, ..col.order]
  
  fwrite(p.mat, sep='\t', ScriptAndDatedFileName(paste0(i, 'pcp.in.mat.txt')))
  
},x=p.mat.list)
```

**not used**

Convert to log2 scale
```{r}
# plot heatmaps to quickly view the sample quality
lapply((names(p.mat.list)), function(x,i) {
  
  print(i)
  hm <- Heatmap(log2(x[[i]]), 
         cluster_columns = F,
         name = i,
         col=viridis(20),
         show_row_names = F,
         cluster_rows = clusterWNA(x[[i]]))
  
  BackupAsPDF(draw(hm, column_title=i), paste0(i, 'Ints.heatmap'), format = 'png')
  
  }, x=p.mat.list)
```

Sanity check of results before pushing to Wynton for run
```{r}
clean.list <- list(CL.1 =fread('./012125_proteinSummarization_data/2025_01_23_CL.biorep1pcp.in.mat.txt', header = T),
                   CL.2= fread('./012125_proteinSummarization_data/2025_01_23_CL.biorep2pcp.in.mat.txt', header=T),
                   FT.1= fread('./012125_proteinSummarization_data/2025_01_23_FT.biorep1pcp.in.mat.txt', header=T),
                   FT.2=fread('./012125_proteinSummarization_data/2025_01_23_FT.biorep2pcp.in.mat.txt', header=T),
                   Sonic.1=fread('./012125_proteinSummarization_data/2025_01_23_sonic.biorep1pcp.in.mat.txt', header=T),
                   Sonic.2=fread('./012125_proteinSummarization_data/2025_01_23_sonic.biorep2pcp.in.mat.txt', header=T))


apply(clean.list[[1]][GN == 'DYRK1A',], 1,print)
apply(clean.list[[5]][GN == 'DYRK1A',], 1,print)

setDT(dp.out[[1]]$ProteinLevelData)[SUBJECT==1 & Protein =='Q13627', .(GROUP, 2^LogIntensities)]
setDT(dp.out[[3]]$ProteinLevelData)[SUBJECT==1 & Protein =='Q13627', .(GROUP, 2^LogIntensities)]
```

# not using this filtered set for now to preserve proteins 
After summary, read in the protein set, remove the proteins with less than 2 peptides, then convert to wide format, convert NA to 

```{r}
dp.out <- lapply(dp.out, function(x){
  
  x$FeatureLevelData <- setDT(x$FeatureLevelData)
  x$ProteinLevelData <- setDT(x$ProteinLevelData)
  
  x$ProteinLevelData <- getNPeptidesPerProtein(proteinlvlDT = x$ProteinLevelData, x$FeatureLevelData, peptideThreshold = 2)
  return(x)
})

names(dp.out) <- names(lysisPrep.list)
```
Remove proteins from each of the runs with just a single peptide

```{r}
# filter to just keep the set with two peptides
dp.out[['sonic']]$ProteinLevelData <- dp.out[['sonic']]$ProteinLevelData[nPeptides >= 2,]
dp.out[['FT']]$ProteinLevelData <- dp.out[['FT']]$ProteinLevelData[nPeptides >= 2,]
dp.out[['CL']]$ProteinLevelData <- dp.out[['CL']]$ProteinLevelData[nPeptides >= 2,]
```
