---
title: "050925_preparePCPInput"
author: "Martin Gordon"
date: "2025-05-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

For benchmarking our SEC_MS pipeline, lets compare it to CCprofiler performance and to PCprophet
PCprophet expects as input a protein matrix; read in this data (see if its alreay been prepared) and submit it to the pipeline on Wynton
We just want to feed these tools the raw data; no preprocessing/smoothing as this is not what is expected by the tool. Confirm this is how the matrix has been supplied previously

```{r}
library(MSstats)
library(data.table)
library(ggplot2)
library(magrittr)
library(stringr)


# useful utility scripts
source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")


source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("../../utils/mg_utils/r_utils/msstats_helperFunctions.R")

clusterWNA <-  function(mat, na.value=0,...){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}

customTheme <-  theme_bw() +
  theme(axis.text.x = element_text(angle=90),
        panel.border = element_rect(color = "lightgrey",
                                    fill = NA,
                                    size = 1))
```
read in the protein data
```{r}
lysisPrep.list <- list('sonic'=fread('./data/050925_newBoxFolder/2025_02_SEC_Lysis_Test/Spectronaut/2.\ NP40-Lysis\ buffer\ +\ Sonication/20241227_104957_MM_Sonic_Lysis_122724_PG_Report.tsv'),
                       'FT'=fread('./data/050925_newBoxFolder/2025_02_SEC_Lysis_Test/Spectronaut/3.\ NP40-Lysis\ buffer\ +\ Freeze-Thaw/20241227_104610_MM_FT_Lysis_122724_PG_Report.tsv'),
                       'CL'=fread('./data/050925_newBoxFolder/2025_02_SEC_Lysis_Test/Spectronaut/1.\ CellLytic\ M\ Lysis\ buffer/20241227_104259_MM_CL_Lysis_122724_PG_Report.tsv')
                       )
```

Functions

First split the matrix into two samples


```{r}
rmSpectronautCols <- function(dt, cols.noi=c("PG.MolecularWeight", "PG.ProteinDescriptions")){
  return(dt[, !..cols.noi])
}

checkSpectronautCols <- function(dt, colpattern='\\[[0-9]{1,2}]'){
  fractionCols <- grep(colpattern, colnames(dt), value=T)
  print(fractionCols)
  fractID <-  stringr::str_extract(fractionCols,'(?<=Frac-)[0-9]{1,2}')
  print(fractID)
  leadingID <-  str_extract(fractionCols,"(?<=[)[0-9]{1,2}(?=<])")
  print(leadingID)
}

lysisPrep.list <- lapply(lysisPrep.list, rmSpectronautCols)

# split out the rep 1 and rep2 cols
# could write. a niceer function to do this but ignore for now
lysisSplit.list <- list('sonic_1'=lysisPrep.list[['sonic']][, .SD, .SDcols = patterns('^PG.|Sonic-1')],
                        'sonic_2'=lysisPrep.list[['sonic']][, .SD, .SDcols = patterns('^PG.|Sonic-2')],
                        'FT_1'=lysisPrep.list[['FT']][, .SD, .SDcols = patterns('^PG.|FT-1')],
                        'FT_2'=lysisPrep.list[['FT']][, .SD, .SDcols = patterns('^PG.|FT-2')],
                        'CL_1'=lysisPrep.list[['CL']][, .SD, .SDcols = patterns('^PG.|CL-1')],
                        'CL_2'=lysisPrep.list[['CL']][, .SD, .SDcols = patterns('^PG.|CL-2')]
                       )

# looks good
lapply(lysisSplit.list, colnames)
```

```{r}
# confirm we have 1-72 in columns
checkSpectronautCols <- function(dt, colpattern="(?<=Frac-)[0-9]{1,3}"){
  
  
  fractionCols <- grep(colpattern, colnames(dt), perl=T, value=T)
  print(length(fractionCols))
  # dont use this as the replicates were searched together
  leadingID <-  str_extract(fractionCols,"(?<=\\[)[0-9]{1,3}(?=\\])")
  fractID <-  stringr::str_extract(fractionCols, colpattern)
  print(fractID == as.character(seq(1,72,1)))
  print(all(fractID == as.character(seq(1,72,1))))
}

```

```{r}
lapply(lysisSplit.list, checkSpectronautCols)
```

Coonvert to matrix, upload to Wynton and run

```{r}
convertToPCprophetMatrix <- function(dt, colpattern="(?<=Frac-)[0-9]{1,3}"){
  
  fractionCols <- grep(colpattern, colnames(dt), perl=T, value=T)
  fractID <-  stringr::str_extract(fractionCols, colpattern)
  fractID <- as.character(order(as.numeric(fractID))) # ensure order is numeric
  
  stopifnot(all(as.numeric(fractID) == seq(1,72,1)))
  colorder <- c('GN', 'ID', fractID)
  setnames(dt, old=c('PG.ProteinAccessions', 'PG.Genes', fractionCols), new=c('ID', 'GN', fractID)) 
  
  message('converting to matrix and removing zeros')
  dt[, matID := paste0(GN, '__', ID)]
  mat <- as.matrix(dt[, -c('GN','ID')], rownames='matID')
  mat[is.na(mat)] <- 0
  
  # convert bsck to DT
  dt.mat <- as.data.table(mat, keep.rownames = T)
  dt.mat[, c('GN', 'ID') := tstrsplit(rn, '__', keep=c(1,2))]
  dt.mat[, rn := NULL]
  return(dt.mat[,..colorder])
}

```

```{r}
pcp.list <- lapply(lysisSplit.list, convertToPCprophetMatrix)

```

Inspect, if looks good upload and run. Do the same prep for the Donor data and then try CCprofiler

```{r}
lapply(names(pcp.list), function(x){
  
  fwrite(pcp.list[[x]], sep ='\t', ScriptAndDatedFileName(paste0(x, '.spectronautPG.pcpinput.txt')))
})
```
```{r}
fread('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/050925_preparePCPInput_data/2025_05_09_sonic_2.spectronautPG.pcpinput.txt', col.names=T) %>% 
  colnames() #looks fine double chekc one or two vs what is on the Box folder
```


**not used**

```{r}


lysisPrep.list[[1]][, .SD, .SDcols=patterns('Sonic-1|PG')]
paste0(names(lysisPrep.list), '-1')

splitSamples <- function(dt, pattern=paste0('PG.|',paste0(names(lysisPrep.list), '-1'))){
  rep1Cols <- pattern
  rep2Cols <- gsub('-1', '-2', pattern)
                      
  return(list('1'=dt[, .SD, .SDcols=rep1Cols],
              '2'=dt[, .SD, .SDcols=rep2Cols]))

}


splitSamples(lysisPrep.list[[1]])

lysisPrep.list
checkSpectronautCols <- function(dt, colpattern='\\[[0-9]{1,2}]'){
  
  fractionCols <- grep(colpattern, colnames(dt), value=T)
  print(fractionCols)
  fractID <-  stringr::str_extract(fractionCols,'(?<=Frac-)[0-9]{1,2}')
  print(fractID
  leadingID <-  str_extract(fractionCols,"(?<=[)[0-9]{1,2}(?=<])")
  print(leadingID)
  
  
  #print(leadingID); print(fractID)
  
}
```

the takes are combined... why?
```{r}

lysisPrep.list[['sonic']] %>% colnames()
cols.oi <- 
  

  
lysisPrep.list[['sonic_1']] <- lysisPrep.list[['sonic']][]
```

