---
title: "051225_lysisPrep_CCprofiler"
author: "Martin Gordon"
date: "2025-05-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview 

Run `CCprofiler` on the lysisPrep datasets to benchmark vs bp pipeline
Starting from protein level info, no pre-filtering/ normalization etc prior to importing
Using default thresholds where possible to allow comparison 

```{r}
library(MSstats)
library(data.table)
library(ggplot2)
library(magrittr)
library(stringr)
library(CCprofiler)


# useful utility scripts
source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")


source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("../../utils/mg_utils/r_utils/msstats_helperFunctions.R")

clusterWNA <-  function(mat, na.value=0,...){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}

customTheme <-  theme_bw() +
  theme(axis.text.x = element_text(angle=90),
        panel.border = element_rect(color = "lightgrey",
                                    fill = NA,
                                    size = 1))
```

```{r}
lysisPrep.list <- list('sonic'=fread('./data/050925_newBoxFolder/2025_02_SEC_Lysis_Test/Spectronaut/2.\ NP40-Lysis\ buffer\ +\ Sonication/20241227_104957_MM_Sonic_Lysis_122724_PG_Report.tsv'),
                       'FT'=fread('./data/050925_newBoxFolder/2025_02_SEC_Lysis_Test/Spectronaut/3.\ NP40-Lysis\ buffer\ +\ Freeze-Thaw/20241227_104610_MM_FT_Lysis_122724_PG_Report.tsv'),
                       'CL'=fread('./data/050925_newBoxFolder/2025_02_SEC_Lysis_Test/Spectronaut/1.\ CellLytic\ M\ Lysis\ buffer/20241227_104259_MM_CL_Lysis_122724_PG_Report.tsv')
                       )
```

```{r}
lysisSplit.list <- list('sonic_1'=lysisPrep.list[['sonic']][, .SD, .SDcols = patterns('^PG.|Sonic-1')],
                        'sonic_2'=lysisPrep.list[['sonic']][, .SD, .SDcols = patterns('^PG.|Sonic-2')],
                        'FT_1'=lysisPrep.list[['FT']][, .SD, .SDcols = patterns('^PG.|FT-1')],
                        'FT_2'=lysisPrep.list[['FT']][, .SD, .SDcols = patterns('^PG.|FT-2')],
                        'CL_1'=lysisPrep.list[['CL']][, .SD, .SDcols = patterns('^PG.|CL-1')],
                        'CL_2'=lysisPrep.list[['CL']][, .SD, .SDcols = patterns('^PG.|CL-2')]
)
```

Use the calibration file to estimate protein monomer mw at elution
```{r}
cal.std <- fread('~/Library/CloudStorage/Box-Box/2025_02_SEC_Lysis_Test/Cal_std.txt',header=F)

cal.std <- rbind(cal.std, data.table(V1=36, V2=1398)) %>% 
  .[order(V1), .(std_weights_kDa=V2, std_elu_fractions=V1)]


cal.std
calibration <-  calibrateMW(cal.std) 

calibration$FractionToMW(30)
```
Need to make these files match the format expected by CCProfiler `traces` object
data structure to replicate
```{r}
example.prot.traces <- readRDS('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/050925_CCprofiler_testRun_data/2025_05_12_protTraces.example.rds')
example.prot.traces$trace_annotation
```
object type
```{r}
# obj type
trace.type <- 'protein'
```

Fraction annotation
```{r}
keys.list <- lapply(lysisSplit.list, function(x){
  
  colnames.oi <- grep('Frac-[0-9]', colnames(x), value=T)
  fraction.oi <- str_extract(colnames.oi, '(?<=[_-]Frac[_-])[0-9]{1,2}')
  print(colnames.oi); print(fraction.oi)
  dt <- data.table(filename=colnames.oi, id=as.numeric(fraction.oi))
  dt[, molecular_weight := calibration$FractionToMW(id)]
})

# sanity check; all look good
sapply(keys.list, function(x){ x$fraction == as.character(seq(1,72,1))}) %>% 
  summary() 
```
trace annotation
First prep the annotation info
```{r}
uniprot.txt <- fread('./docs/uniprotkb_reviewed_true_2025_05_13.tsv.gz')
# tidy col names 
uniprot.txt <- uniprot.txt[Organism == 'Homo sapiens (Human)', .(id=Entry, protein_id=Entry, Entry_name=`Entry Name`, Status='reviewed', Protein_names=`Protein names`, Gene_names=`Gene Names`, Organism, Length, Mass, GO_ID=`Gene Ontology IDs`, protein_mw=(Mass/1000), Decoy=0)]

```

remove NA
```{r}
anno.list <- lapply(lysisSplit.list, function(x){
  prots.oi <- unique(x$PG.ProteinAccessions)
  # first, lets subset this to proteins with at least one measurement 
  dt.long <- melt(x, id.vars=c(c('PG.MolecularWeight', 'PG.ProteinAccessions', 'PG.Genes', 'PG.ProteinDescriptions')))
  keep.prots <- dt.long[!is.na(value), .N, by=.(PG.ProteinAccessions)]$PG.ProteinAccessions
  # subset it to keep prots; dont want empty rows in the data
  return(uniprot.txt[id %in% keep.prots,])
})
```

```{r}
test <- copy(lysisSplit.list)

#' function to generate traces object from input matrix
convertToCCprofilerMatrix <- function(dt, colpattern="(?<=Frac[-_])[0-9]{1,2}"){
  
  fractionCols <- grep(colpattern, colnames(dt), perl=T, value=T)
  fractID <-  stringr::str_extract(fractionCols, colpattern)
  #fractID <- as.character(order(as.numeric(fractID))) # ensure order is numeric
  
  stopifnot(all((fractID) == seq(1,72,1)))
  colorder <- c(fractID, 'protein_id')

  setnames(dt, old=c('PG.ProteinAccessions', fractionCols), new=c('protein_id', fractID)) 

  message('converting to matrix and removing zeros')
  mat <- as.matrix(dt[, ..colorder], rownames='protein_id')
  mat[is.na(mat)] <- 0
  dt.mat <- as.data.table(mat, keep.rownames = T)
  
  colorder <- c(fractID, 'id')
  setnames(dt.mat, old='rn', new='id')
  return(dt.mat[,..colorder])
}
# looks good!
convertToCCprofilerMatrix(test[[5]])
```
matrix list
```{r}
mat.list <- lapply(lysisSplit.list, function(x){
  # first, lets subset this to proteins with at least one measurement 
  dt.long <- melt(x, id.vars=c(c('PG.MolecularWeight', 'PG.ProteinAccessions', 'PG.Genes', 'PG.ProteinDescriptions')))
  keep.prots <- dt.long[!is.na(value), .N, by=.(PG.ProteinAccessions)]$PG.ProteinAccessions
  # subset it to keep prots; dont want empty rows in the data
  convertToCCprofilerMatrix(x[PG.ProteinAccessions %in% keep.prots, ])
})

names(mat.list)
```
Now assmeble these objects into one data format

```{r}
traces.list <- lapply(names(lysisSplit.list), function(br){
  
  traces.obj <- list(traces=mat.list[[br]],
                     trace_type=trace.type,
                     trace_annotation=anno.list[[br]],
                     fraction_annotation=keys.list[[br]]
                     )
  
  return(traces.obj)
})

# set the class type as traces
trace.list <- lapply(traces.list, function(x){ class(x) <- 'traces'; return(x)})

#looks good... lets test a single one, then save and run later 
trace.list
```
now make sure that the annotation mat matches the list of proteins in matrix and the order matches
filter to the same subset and match row orders between mat and anno obj
```{r}
trace.list <- lapply(trace.list, function(x){
  
  mat <- x$traces
  anno <- x$trace_annotation
  
  # subset the mat to the set of proteins in the anno.dt
  submat <- mat[id %in% anno$id,]
  print(dim(submat))
  
  #match the row ordering between anno and ints max
  x$traces <- submat[match(anno$id, submat$id)]
  return(x)
})
```

test on one sample
```{r}
test <- trace.list[[1]]
names(trace.list) <- names(lysisSplit.list)
saveRDS(trace.list, ScriptAndDatedFileName('trace.class.input.rds'))

```

Complex recovery; load CORUM

```{r}
corum.dt <- fread('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/050925_CCprofiler_testRun_data/2025_05_12_corum.5.1.hsComplexes.long.csv.gz')
# filter to complexes with at least  two proteins or throws error
corum.dt <- corum.dt[complex_id %in% corum.dt[,.N, by=complex_id][N>1, complex_id]]
```

Convert the complex db to binary matrix, get path lengths between interactors, and then 
```{r,eval=TRUE}
# complex to  ppi format
binaryHypotheses <- generateBinaryNetwork(corum.dt)
pathLength <- calculatePathlength(binaryHypotheses) # shortest path

# generating corum complex + decoys 
corumTargetsPlusDecoys <- generateComplexDecoys(target_hypotheses=corum.dt,
                                                dist_info=pathLength,
                                                min_distance = 2,
                                                append=TRUE)

corumTargetsPlusDecoys[, isDecoy := ifelse(grepl('DECOY', complex_name, ignore.case = T), 'yes', 'no')]
corumTargetsPlusDecoys[,.N, by=isDecoy] # even split of true and false sets
```

Finding complex features in our data.. test run on single sample
```{r}
# each complex_id is considered a hyptohesis
# imputes matrix missing values with random sample from bottom 5% of the distribution (per row)
# loops over each complex and runs a sliding window, subsets imputed ints mat to these protein subunits
# findComplexFeatureSW; sliding window correlation for all features in complex_id
# findComplexFeaturesPP: peak picking; 3 up and 3 down to declare a peak, left and right ar ethe margins of hte picked peaks
# selects peaks within the sliding window defined above to define features
# collapses features, then correlates features across proteins to get a cor score

complexFeatures <- findComplexFeatures(traces=test,
                                       parallelized = TRUE,
                                       n_cores = 3,
                                       complex_hypothesis = corumTargetsPlusDecoys)

```
Run on all samples, filter complexes, save output
```{r}
complexes.list <- lapply(names(trace.list), function(dat){
  res <- complexFeatures <- findComplexFeatures(traces=trace.list[[dat]],
                                                parallelized = T,
                                                n_cores = 6,
                                                complex_hypothesis = corumTargetsPlusDecoys)
  return(res)
})

```
Attach names and save file
```{r}
names(complexes.list) <- names(trace.list)
saveRDS(complexes.list, ScriptAndDatedFileName('complexes.list.rds'))
```

Functions for filtering detected complexes protein, complex ids, 
Also filter using mw; complex APEX should be >=  <min_monoomer_distance_fractor> x largest monomer mw to be considered a complex
Expect APEX peak to be larger than this
```{r}
test <- complexes.list[[1]]

filterFeatures(complexFeatures,
               complex_ids = NULL,
               protein_ids = NULL,
               min_feature_completeness = NULL,
               min_hypothesis_completeness = NULL,
               min_subunits = NULL,
               min_peak_corr = NULL,
               min_monomer_distance_factor = 2)
```


## Protein summarization, scoring, FDR estimation and filtering 

The resulting protein complex features can be scored based on the co-elution characteristics (average protein subunit correlation) of the proteins within each protein complex feature. 
The co-elution score is further statistcally corrected for the number of correlating proteins out of all the proteins in the protein complex hypothesis. 
The statistically corrected co-elution score can be used to estimate q-values for each protein complex peak group. 

Filtering the protein complex peak groups based on a specified q-value cutoff then achieves a final result table at specified FDR (a q-value cutoff of 0.05 results in a final table with 5% FDR). 
First try with test dataset

```{r}
# this adds an overall co-elution score: what is this?
complexFeaturesScored <- calculateCoelutionScore(test)
# adds a pvalue and a q-value for the score (corrects for muliple testing)
qvalueComplexFeaturesScored <- calculateQvalue(complexFeaturesScored)
# not great below 0.05 is a safer cutoff
qvalueComplexFeaturesScoredStats <- qvaluePositivesPlot(qvalueComplexFeaturesScored)

complexFeaturesFiltered <- subset(qvalueComplexFeaturesScored, qvalue <= 0.05)
summarizeFeatures(complexFeaturesFiltered, PDF=TRUE, name=ScriptAndDatedFileName('test'))
```

 loop through each of the files, filter using standard thresholds and share with Ben

```{r}
filtered.complexes <- lapply(names(complexes.list), function(x){
  
  # get co-elution score 
  complexFeaturesScored <- calculateCoelutionScore(complexes.list[[x]])
  # assign p/q values
  qvalueComplexFeaturesScored <- calculateQvalue(complexFeaturesScored, PDF=TRUE, name=ScriptAndDatedFileName(paste0(x, '.pvalplots')))
  
  # n
  qvalueComplexFeaturesScoredStats <- qvaluePositivesPlot(qvalueComplexFeaturesScored, PDF=TRUE, name=ScriptAndDatedFileName(paste0(x, '.TPvsQval')))
  complexFeaturesFiltered <- subset(qvalueComplexFeaturesScored, qvalue <= 0.05)
  
  summarizeFeatures(complexFeaturesFiltered, PDF=TRUE, name=ScriptAndDatedFileName(paste0(x, '.complexSummaryplots.filtered.qval0.05')))
  return(qvalueComplexFeaturesScored)
  })
```

filtered.complexes
```{r}
names(filtered.complexes) <- names(complexes.list)
allComplex.features <- rbindlist(filtered.complexes, idcol='sample')

# number of unique complex features detected
fwrite(allComplex.features, ScriptAndDatedFileName('lysisPrep.ccprofiler.allComplexFeatures.csv.gz'))

# take the best peak per complex, sample
bestComplexFeature <- allComplex.features[,.SD[which.min(pvalue)], by=.(sample, complex_id)]
fwrite(bestComplexFeature, ScriptAndDatedFileName('lysisPrep.ccprofiler.bestScoringFeaturePerComplex.csv.gz'))

# number of complexes with at least one significant feature 
bestComplexFeature[qvalue < 0.05, .N, by=sample]
```


## collapse complexes

 *CCprofiler* is able to retrieve co-elution evidence from full protein complex queries but also subsets thereof.
Due to partial overlap of the protein sets grouped as complex queries, the same co-elution signal can be retrieved as evidence for multiple diﬀerent queries. 
collapse the initial results to truly unique signals with regard to (i), subunit composition and (ii), resolution in the chromatographic dimension.
collapse complex features and save
```{r}
collapsed.complexes <- lapply(complexes.list, function(x){
  
  complexFeaturesUnique <- getUniqueFeatureGroups(x,
                                                  rt_height = 0,
                                                  distance_cutoff = 1.25)

  complexFeaturesCollapsed <- callapseByUniqueFeatureGroups(complexFeaturesUnique,
                                                            rm_decoys = TRUE)
  complexFeaturesCollapsed
})
```

```{r}
fwrite(rbindlist(collapsed.complexes, idcol = 'sample'), ScriptAndDatedFileName('lysisPrep.ccprofiler.collapsedComplexes'))
```


complex subfeature summary is useful; most of the recovered complexes are based on a single elution window... maybe not super confident? hard to say...
```{r}
complexFeaturesFiltered[,.N, by=complex_id]

View(complexClustering)
```


```{r}
lapply(complexes.list, function(x){
  
  complexFeaturesScored <- calculateCoelutionScore(complexFeatures)
  qvalueComplexFeaturesScored <- calculateQvalue(complexFeaturesScored)
  qvalueComplexFeaturesScoredStats <- qvaluePositivesPlot(qvalueComplexFeaturesScored)
  complexFeaturesFiltered <- subset(qvalueComplexFeaturesScored, qvalue <= 0.05)
summarizeFeatures(complexFeaturesFiltered)
  
})
```

