---
title: "051225_plotsForReview"
author: "Martin Gordon"
date: "2025-05-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Prepare heatmaps/linecharts for Qiongyu for meeting with Robyn and Kirsten tomorrow

```{r packages}
library(ComplexHeatmap)
library(circlize)
library(data.table)
library(ggplot2)
library(magrittr)
library(stringr)

source ("../../utils/bp_utils/MSstats_Helper_Functions.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
source("/Users/martingordon/Documents/utils/mg_utils/r_utils/plottingHelperFunctions.R")


col.pal <- getQualitativePalette(n=15)
```
read in the msstats dt
```{r}
dp.dt <- fread('/Users/martingordon/Documents/projects/071123_QLi_DARPA.GC01.GC02/output/QRFPR.mss.pwcomparisons.proc.csv')
#MRX2.mss <- fread('/Users/martingordon/Documents/projects/071123_QLi_DARPA.GC01.GC02/output/MRX2.mss.pwcomparisons.proc.csv')

mss.out <-  rbind(MRX2=MRX2.mss,
                 QRFPR=QRFPR.mss)
```

proteins OI 

```{r}
prots.oi <- 'ARPC3
ABI1
CTTN
EFHD2
DBNL
ACTR3
FNBP1L
TRIP10
ABI2
PRKCA
WASF1
KCTD12
PRKD2
RAPH1
BRK1
WASF3
ARPC2
GNAQ
ARPC5
ARPC1A
GNA11
BCORL1
WASF2
ING3'
```

```{r}
prots.oi <- strsplit(prots.oi, '\n')[[1]]


dp.dt$Label %>% unique()

# looking at impactof YM here
conditions.oi <- c("RF05-RF00", "RF10-RF00",  "RF05_YM-RF00_YM", "RF10_YM-RF00_YM")
reduced.conditions.oi <- c("RF10-RF00", "RF10_YM-RF00_YM")
expanded.conditions.oi <- c("RF05-RF00", "RF10-RF00",  "RF05_YM-RF00_YM", "RF10_YM-RF00_YM", # basically looks at time and YM pre-treatment
                            "RF05_YM-RF05", "RF10_YM-RF10",
                            "RF05_PTX-RF00_PTX", "RF10_PTX-RF00_PTX") 
unique(dp.dt$Label)


grep('PTX', unique(dp.dt$Label), invert=T,value=T)
```

this is the output of the pw contrasts... we want the counts
```{r}
dp.dt[, gene := multiUniprots2multiGenes(Protein, species='HUMAN')]
subdt <- dp.dt[gene %in% prots.oi]
```

read in the LFC values/counts values for this QRFPR analysis
```{r, lfc-heatmap, fig.height=3.5, fig.width=3.5}
mat <- dcast(subdt, gene~Label, value.var='log2FC') %>% 
  as.matrix(rownames='gene')

submat <- mat[,  grepl(paste0(conditions.oi, collapse = '|'), colnames(mat))]


hm <- Heatmap(submat,
        name='log2FC',
        border=T,
        show_column_dend = FALSE,
        col=colorRamp2(breaks=c(-1,0,1), colors = c(col.pal[1], 'white', col.pal[2])),
        row_names_gp = gpar(fontsize=6),
        column_title_gp = gpar(fontsize=8),
        column_names_gp = gpar(fontsize=7),
        column_split = ifelse(grepl('YM-', colnames(submat)), 'YM pre-treated', 'Not pre-treated')
        )
hm
BackupAsPDF(hm, 'YM.pretreatment.lfc.heatmap')


submat <- mat[,  grepl(paste0(reduced.conditions.oi, collapse = '|'), colnames(mat))]


hm <- Heatmap(submat,
        name='log2FC',
        border=T,
        show_column_dend = FALSE,
        col=colorRamp2(breaks=c(-1,0,1), colors = c(col.pal[1], 'white', col.pal[2])),
        row_names_gp = gpar(fontsize=6),
        column_title_gp = gpar(fontsize=8),
        column_names_gp = gpar(fontsize=7),
        column_split = ifelse(grepl('YM-', colnames(submat)), 'YM pre-treated', 'Not pre-treated')
        )
hm
BackupAsPDF(hm, 'YM.pretreatment.reduced.lfc.heatmap')
```

get the intensity heatmap and plot these

```{r}
ints.dt <- fread('output/QRFPR.ProteinlvlData.tmpNormalized.csv.gz')
ints.dt[, gene := multiUniprots2multiGenes(Protein, species='HUMAN')]

ints.sub <- ints.dt[gene %in% prots.oi,]
```

Heatmap; get the groups of interest

```{r, ints-heatmap, fig.height=3.5, fig.width=8}
groups.oi <- ints.sub

ints.mat <- dcast(ints.sub, gene~paste0(GROUP, '.', SUBJECT), value.var='LogIntensities') %>% 
  as.matrix(rownames='gene')


groups.oi
groups.oi <- grep('PTX', colnames(ints.mat), invert=T, value=T)
groups.subset.oi <- grep('RF00_YM', groups.oi, invert=T, value=T)

submat <- ints.mat[,  grepl(paste0(groups.oi, collapse = '|'), colnames(ints.mat))]
# sweep outt he median intensity
submat <- sweep(submat, 1, apply(submat, 1, median, na.rm=T))

# first sweep out the median, then maybe try some baseline (T0 no treat or pretreat)
hm <- Heatmap(submat,
        name='ints/median',
        border=T,
        cluster_columns=FALSE,
        show_column_dend = FALSE,
        show_column_names = FALSE,
        column_split = list(ifelse(grepl('RF05', colnames(submat)), '05 min',
                              ifelse(grepl('RF10', colnames(submat)), '10 min',
                                     '0 min')),
                            ifelse(grepl('YM', colnames(submat)), ' YM pretreatment', ' No pretreatment')),
        col=colorRamp2(breaks=c(-1,0,1), colors = c(col.pal[1], 'white', col.pal[2])),
        row_names_gp = gpar(fontsize=6),
        column_title_gp = gpar(fontsize=8, fontface='bold'),
        column_names_gp = gpar(fontsize=7),
        )
hm
BackupAsPDF(hm, 'YM.pretreatment.ints.heatmap')


# plot the reduced number of groups; and subtract t0 from them 
submat <- ints.mat[,  grepl(paste0(groups.subset.oi, collapse = '|'), colnames(ints.mat))]
# sweep outt he median intensity
submat <- sweep(submat, 1, apply(submat[, grepl('RF00', colnames(submat))], 1, mean, na.rm=T))
submat <- submat[, !grepl('RF00', colnames(submat))]

# first sweep out the median, then maybe try some baseline (T0 no treat or pretreat)
hm <- Heatmap(submat,
        name='ints/mean[0 min]',
        border=T,
        cluster_columns=FALSE,
        show_column_dend = FALSE,
        show_column_names = FALSE,
        column_split = list(ifelse(grepl('RF05', colnames(submat)), '05 min',
                              ifelse(grepl('RF10', colnames(submat)), '10 min',
                                     '0 min')),
                            ifelse(grepl('YM', colnames(submat)), ' YM pretreatment', ' No pretreatment')),
        col=colorRamp2(breaks=c(-1,0,1), colors = c(col.pal[1], 'white', col.pal[2])),
        row_names_gp = gpar(fontsize=6),
        column_title_gp = gpar(fontsize=8, fontface='bold'),
        column_names_gp = gpar(fontsize=7),
        )
hm
BackupAsPDF(hm, 'YM.pretreatment.ints.subset.heatmap')

```

linechart plot of GNAQ and GNA11 across time 0, 5, 10 4 lines in total, or two if faceted

```{r, linechart,  fig.height=3.5, fig.width=4}
gna.set <- c('GNAQ', 'GNA11')

str_extract()

ints.dt[, timepoint := factor(stringr::str_extract(GROUP, '00|05|10'), levels=c('00', '05','10'))]
ints.dt[, pretreatment := ifelse(grepl('YM|PTX', GROUP), str_extract(GROUP, 'YM|PTX'), 'None')]

g <- ggplot(ints.dt[gene %in% gna.set & GROUP %in% unique(gsub('[.][0-9]$','', groups.oi)),], aes(x=timepoint, color=pretreatment, group=pretreatment, y=LogIntensities)) +
  geom_smooth(stat = 'summary', fun.y = mean, se =TRUE, show.legend = T, aes(fill=pretreatment), alpha=0.4) +
  geom_point() +
  scale_color_manual(values = (col.pal)) + 
  scale_x_discrete(expand=c(0.04,0.04)) +
  scale_fill_manual(values = (col.pal)) + 
  ggtitle('GNAQ & GNA11 response') +
 # geom_hline(yintercept = 0, linetype=2, alpha=0.6) +
  labs(x='timepoint (mins)', y="Intensity (log2)") +
  facet_grid(gene~., scales='free_y') +
  theme_bw() +
  theme(strip.text.y=element_text(size=10),
          strip.text=element_text(color='white'),
          strip.background = element_rect(colour = "white", fill = "grey80"))


BackupAsPDF(g, 'gnaQ.gna11.linecharts')
```

