---
title: "DARPA.GC01.GC02.InteractionAnalysis"
author: "Martin Gordon"
date: "2023-11-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Interaction Analysis
Question: how does the addition of PTX/YM (the Gprotein inhibitors impact the expression profile of the data?)

SO want to compare (ZC05-ZC00) - (ZC05_YM/PTX - ZC00_YM/PTX) etc
Need to recode the data table for this 

```{r}
library (data.table)
library (ComplexHeatmap)
library (ggplot2)
library (emmeans)
library(stringr)

rotate.x.axis.text <- theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")

```

read in the protein quant data 
For now, just proceed with the one round normalised data... speak to Ben
```{r}
MRX2.dp  <- readRDS('./output/MRX2.rds') 
QRFPR.dp <- readRDS('./output/QRFPR.rds') 


MRX2 <- data.table(MRX2.dp$ProteinLevelData)
QRFPR <-data.table(QRFPR.dp$ProteinLevelData)

p.quant.list <-  list(MRX2 = MRX2,
                      QRFPR = QRFPR)

#p.quant.list$QRFPR <- fread('./output/QRFPR.ProteinlvlData.tmpNormalized.csv.gz')
```

Create new dummy variables
```{r}

lapply(p.quant.list, function(x){
  
  x[ !GROUP %like% "_", GROUP := paste0(GROUP,'_CTRL')] %>% 
  .[, batch := paste0('batch.',SUBJECT)] %>% 
  .[, receptor := ifelse(GROUP %like%  'ZC', 'MRX2', 'QRFPR')] %>% 
  .[, pretreat := str_extract(GROUP, 'PTX|CTRL|YM')] %>% 
  .[, timeStr := sub('RF|ZC','', str_extract(GROUP, 'ZC[0-9]{2}|RF[0-9]{2}'))]
})


lapply(p.quant.list, str)
```
For now, leave the contrasts as is
w/o running second round of correction, all terms are significant
```{r}

lapply(p.quant.list, function(x){
  
 model <- aov(LogIntensities ~ batch+pretreat+timeStr+pretreat:timeStr, data=x)  
 summary(model)
})
```
Interaction analysis check how pretreatment impacts response
```{r}

# linearModelsAllProteins can this function which computes statistic for a contrast of two different log2FC
# how does the treatment effect change per antagonist level? basically compare treatment effect within each antagonist level
emmeans.contrastOfContrasts <- function (l, factorFormula = ~timeStr|pretreat){  # how does response overe time change with pretreatment status
  emm <- emmeans::emmeans(l, factorFormula)
  contrast1 <- pairs(emm)
  contrast2 <- pairs(contrast1, by = NULL, adjust = "none")
  return (as.data.table(contrast2))
}

# postProcessFunction = emmeans.contrastOfContrasts tells it to use the function above after doing anova.
# it is optional, adds significant time to the computation (~5x), but is nice to
# put a magnitude and p value on specific differences of differences
#
# cl = 6 parameter says to use 6 processes.  Set to NULL if multiprocess is failing, or higher or lower as you wish.
# Numbers near the number of processors available in your machine is probably counterproductive.

lm.out <- lapply(p.quant.list, function(x){
  
  # Numbers near the number of processors available in your machine is probably counterproductive.
  x.lm <- linearModelsAllProteins(x, 
                                  formulaList = list(full   = LogIntensities~batch + timeStr + pretreat + pretreat:timeStr), # full model; 4 way ANOV
                                  splitColumn = "Protein", postProcessFunction = emmeans.contrastOfContrasts, cl = 6) |> suppressWarnings()
  
  return(x.lm)
  
})

```
```{r}
lapply(seq_along(lm.out), function(x,n,i){

  dt <- x[[i]][['anova']]
  str(dt)

 g <- ggplot(dt, aes(x=dt$p.value)) +
   geom_histogram(boundary=0.0, binwidth = 0.005) +
   xlim(c(0,0.05)) +
   facet_wrap(~dt$term)
 
 #BackupAsPDF(g, paste0(n[[i]], '.aov.term.pval.historgrams'))
 
},x=lm.out, n=names(lm.out))

```
## adjust pvalues 

```{r}
lapply(lm.out, function(x){
  
  x[['anova']][, fdr := p.adjust(p.value, method = "BH"), by= term]
  
})

```
Write out files to dir
```{r}
lapply(seq_along(lm.out), function(x,n,i){
  
  fwrite(x[[i]][['anova']], paste0('./output/',n[[i]], '.anova.contrasts.csv.gz'))
  
},x=lm.out, n=names(lm.out))

```

plot the interaction effects

## interaction effects
different response given pretreatment

```{r}
clusterwNA <- function(x, na.val=0){
  
  x[is.na(x)] <- na.val
  dist(x) %>%  hclust(method='complete')
}

#play with thresholds here
lapply(seq_along(lm.out), function(x,n,i){
  dt <- x[[i]][['anova']]
  
  receptor <- n[[i]]
  
  p.quant <- p.quant.list[[receptor]]
  
  sigProtein <- dt[term %in% c("timeStr:pretreat") & p.value < 0.001,] %>% 
   #  .[order(p.value), head(.SD, 100), .SDcols = c('p.value', 'Protein')] %>% 
    .[,unique(Protein)]
 

  # filter just to our data
   prot.mat <- dcast (p.quant[Protein %in% sigProtein,], Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "Protein")


  rownames(prot.mat) <- multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
  
  prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))
  
  # vector for col splitting
  column.vec <- sub('_[1-6]$', '', colnames(prot.mat))
  
  hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'log Ints/Median',
                cluster_rows = clusterwNA(prot.mat),
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (p.val < 0.001)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = column.vec,
              column_title_gp = gpar(fontsize=7))
  
#BackupAsPDF(hm, paste0(n[[i]], '.npff:cm101.pval<0.001.heatmap'), dimensions = c(8,9))
  BackupAsPDF(draw(hm, column_title=paste0(n[[i]], ' PTX/YM pretreat interaction')), paste0(n[[i]], '.ptx.ym.response.pval<0.001.heatmap'), dimensions = c(8,6))
  
},x=lm.out, n=names(lm.out))
```
contrast of contrast
----
emmeans package allows us to extract a magnitude change and a pvalue for the specific contrast per protein
magnitude difficult to interpret in interaction analysis; 

```{r}
# pick out the contrassts we are interested in basically CTRL vs YM/PTX treat
contrasts.oi <- c( "(timeStr00 - timeStr05 CTRL) - (timeStr00 - timeStr05 YM)", "(timeStr00 - timeStr05 CTRL) - (timeStr00 - timeStr05 PTX)", 
                   "(timeStr00 - timeStr10 CTRL) - (timeStr00 - timeStr10 YM)", "(timeStr00 - timeStr10 CTRL) - (timeStr00 - timeStr10 PTX)", 
                   "(timeStr05 - timeStr10 CTRL) - (timeStr05 - timeStr10 YM)", "(timeStr05 - timeStr10 CTRL) - (timeStr05 - timeStr10 PTX)")

contrasts.list <- lapply(seq_along(lm.out), function(x,n,i){
  dt <- x[[i]][['postProcess']]
  print(dt[contrast %in% contrasts.oi,]) }, x=lm.out, n=names(lm.out))

names(contrasts.list) <-  names(lm.out)

lapply(seq_along(contrasts.list), function(x,n,i){
  fwrite(x[[i]], paste0('./output/', n[[i]], '.emmeans.contrasts.oi.csv'))
},x=contrasts.list, n=names(contrasts.list))


lapply(contrasts.list, print)
```
Value here is we can require a magnitude of difference. 
Signs are confusing here because the loss of a negative effect will show as  a positive effect, and vice versa. 
Just plot all the sig movers 

```{r}

#play with thresholds here
lapply(seq_along(contrasts.list), function(x,n,i){
  dt <- x[[i]]

  receptor <- n[[i]]
  
  p.quant <- p.quant.list[[receptor]]
  
  sigProtein <- dt[contrast %in% contrasts.oi & p.value < 0.001,] %>% 
    .[,unique(Protein)]
  
  print(sigProtein)
 
  # filter just to our data
   prot.mat <- dcast (p.quant[Protein %in% sigProtein,], Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "Protein")


  rownames(prot.mat) <- multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
  
  prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))
  
  # vector for col splitting
  column.vec <- sub('_[1-6]$', '', colnames(prot.mat))
  
  hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'log Ints/Median',
                cluster_rows = clusterwNA(prot.mat),
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (p.val < 0.001)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = column.vec,
              column_title_gp = gpar(fontsize=7))
  
  BackupAsPDF(draw(hm, column_title=paste0(n[[i]], ' PTX/YM pretreat interaction')), paste0(n[[i]], '.ptx.ym.emmeans.contrastofcontrasts.pval<0.001.heatmap'), dimensions = c(9,8))
  
},x=contrasts.list, n=names(contrasts.list))
```
