---
title: "DARPA.GC01.GC02.PWcontrats"
author: "Martin Gordon"
date: "2023-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MSstats PW contrasts

Two sets of data:
ZC conditions: these are for the receptor MRX2.
RF conditions: these are for the receptor QRFPR.

Notes for conditions:
ZCxx: means the time for Zinc compound treatment for MRX2. So ZC00 means 0 min (ctrl), and ZC10
means 10 min of Zinc compound treatment.
YM: means the cells were pretreated with YM compound for Gq signaling inhibition.
PTX: means the cells were pretreated with pertussis toxin for Gi signaling inhibition.
So for example, the condition of ZC05_YM means Zinc compound treatment for 5 min and with YM
pretreatment for Gq inhibition.
The conditions for QRFPR follows the same logic


Packages
```{r}
library(data.table)
library(ggplot2)
library(MSstats)
library(magrittr)
library(stringr)
library(ComplexHeatmap)
library(randomcoloR)

source ("../../utils/bp_utils/MSstats_Helper_Functions.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
```
Read in data and get overview
```{r}
spec <- fread('./data/GC01_GC02/20231102_155913_GC01_GC02_EXc_dirDIA_APEX_MSStatsReport.tsv')


spec[,.N,.(Run,Condition,BioReplicate)]
dim(spec)
summary(spec)
str(spec)
```

```{r}
spec[,.N,by=.(BioReplicate,Condition)]
spec %>%  colnames()
```
histogram of intensites

```{r}
ggplot(spec, aes(x=log2(Intensity))) + geom_histogram()


filter <- 2^4
spec <- spec[Intensity > filter]
  
ggplot(spec, aes(x=log2(Intensity))) + geom_histogram()
```
N peptides per run 
Clearly samples ZC00_YM2 and ZC00_PTX2 biorep 2 have lower counts than other groups, biorep in general seems to have lower counts
```{r}
colnames(spec)
head(spec)

# sanity check; make sure each peptide has a unique elution profile
any(spec[,.N,by=.(PeptideSequence,PrecursorCharge, Condition, BioReplicate)]$N > 1)


col.pal <- randomcoloR::randomColor(count=length(unique(spec$Condition)))

g <- ggplot(spec[,.N,by=.(Condition, BioReplicate,PrecursorCharge)], aes(x=reorder(interaction(Condition, BioReplicate)), y=N, fill=Condition)) + 
  geom_bar(stat='Identity') +
  scale_fill_manual(values=col.pal)  +
  theme_classic() +
  ggtitle('N peptides per run') +
  theme(
    axis.text.x = element_text(angle=90)
  )
g

BackupAsPDF(g, 'NpeptidesPerRun.barplot.')
```

Counts in general seem lower for the second biorep

```{r}
g <- ggplot(spec[,.N,by=.(Condition, BioReplicate,PrecursorCharge)], aes(x=reorder(interaction(Condition, BioReplicate)), y=N, fill=as.factor(BioReplicate))) + 
  geom_bar(stat='Identity') + 
  scale_fill_manual(values=col.pal)  +
  theme_classic() +
  ggtitle('N peptides per run') +
  theme(
    axis.text.x = element_text(angle=90)
  )
g
```
boxplots of peptide intensities per sample
samples  ZC00_YM2 and ZC00_PTX2 biorep 2 clear outliers; remove

```{r}
g <- ggplot(spec, aes(x=reorder(interaction(Condition,BioReplicate)), y=log2(Intensity), fill=Condition))  +
  geom_boxplot() +
   scale_fill_manual(values=col.pal)  +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle=90)
  )

BackupAsPDF(g, 'raw.intensities.boxplotrmSamples.')
```

Looks to me the RF treated QRFPR samples are much more variable
Split the dataset into two and process seperately to avoid over-correcting during normalisation
```{r}
# drop the problematic samples
spec <- spec[!interaction(Condition,BioReplicate) %in% c('ZC00_YM.2', 'ZC00_PTX.2'), ] 

ggplot(spec, aes(x=reorder(interaction(Condition,BioReplicate)), y=log2(Intensity), fill=Condition))  +
  geom_boxplot() +
   scale_fill_manual(values=col.pal)  +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle=90)
  )
summary(spec)
```
```{r}
spec[, receptor := ifelse(grepl('ZC', Condition), 'MRX2', 'QRFPR')]
# clear batch effect to me in spec: renmae the bioreps 4-6
spec[receptor == 'QRFPR', BioReplicate := BioReplicate +3]

```

Before splitting, view PCA of samples 
```{r}
clusterwNA <- function(x, na.val=0){
  
  x[is.na(x)] <- na.val
  dist(x) %>%  hclust(method='complete')
}


f.mat <-dcast(spec, interaction(PeptideSequence,PrecursorCharge)~interaction(Condition,BioReplicate), value.var = 'Intensity') %>% 
  as.matrix(rownames='PeptideSequence')

f.mat <- log2(f.mat)

f.mat <- f.mat[complete.cases(f.mat),]

submat <- f.mat[sample(rownames(f.mat), 3000),]
Heatmap(submat, cluster_rows = clusterwNA(submat), show_row_names = F)
```
PCA of the samples; two large global batches depending on receptor
Also clear outlier in sample ZC10.2; drop this sample also

```{r}
pcaOut <- prcomp(t(f.mat))

colInfo <- data.table(colname = colnames(f.mat))
colInfo[,c("treat","rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ] # v useful fucntion

title <- "PCA"

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = treat, shape=rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values=c(19, 21:26)) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,#legend settings
         color = guide_legend(override.aes = list(shape =21) ) )


BackupAsPDF(p, "PCA.Complete.Features")

```

Drop ZC10.2 intensites very different to other members of group and replot

```{r}
spec <- spec[!interaction(Condition,BioReplicate) %in% c('ZC10.2'), ] 
spec.list <- split(spec, spec$receptor)


f.mat <-dcast(spec, interaction(PeptideSequence,PrecursorCharge)~interaction(Condition,BioReplicate), value.var = 'Intensity') %>% 
  as.matrix(rownames='PeptideSequence')

f.mat <- log2(f.mat)

f.mat <- f.mat[complete.cases(f.mat),]

submat <- f.mat[sample(rownames(f.mat), 3000),]
Heatmap(submat, cluster_rows = clusterwNA(submat), show_row_names = F)
```


Now process the two spec list elements seperately

```{r}

spec.list <- lapply(spec.list, function(x){
  
  spec.mss <- specFileToCompleteMSstats(x)
  spec.mss[, IsotopeLabelType := 'L']
  setDF(spec.mss)  # MSstats likes data.frame not data.table, so convert to class data.frame

})

names(spec.list)
```
Run MSstats preprocessing function

```{r}
dp.out.list <- lapply(spec.list, function(x){
  
           MSstats::dataProcess(x,
                               featureSubset = "highQuality",
                               remove_uninformative_feature_outlier = TRUE,
                               MBimpute = FALSE ) 
  
})
```
save output
```{r}
#saveRDS(dp.out.list[['MRX2']], './output/MRX2.rds')
#saveRDS(dp.out.list[['QRFPR']], './output/QRFPR.rds')

MRX2.dp  <- readRDS('./output/MRX2.rds')
QRFPR.dp <- readRDS('./output/QRFPR.rds')

dp.out.list <- list(MRX2 = MRX2.dp,
                    QRFPR = QRFPR.dp)

```

boxplots of normalised logIntensites, 

```{r}
lapply(seq_along(dp.out.list), function(x,n,i){
  
  g <- ggplot(x[[i]]$ProteinLevelData, aes(x=reorder(interaction(GROUP,SUBJECT)), y=LogIntensities, fill=GROUP)) + 
    geom_boxplot() +
    scale_fill_manual(values=col.pal) +
    ggtitle(paste0(n[[i]], ' protein intensities')) +
    theme_classic() +
    theme(axis.text.x = element_text(angle=90))

  BackupAsPDF(g, paste0(n[[i]], '.protein.intensity.tmpnormalised.boxplot.'))
},x=dp.out.list, n=names(dp.out.list))
```
Protein counts per sample

```{r}
lapply(seq_along(dp.out.list), function(x,n,i){
  
  dt <- data.table(x[[i]]$ProteinLevelData)
  print(dt)
  g <- ggplot(dt[, .N, by=.(Protein,GROUP,SUBJECT)], aes(x=reorder(interaction(GROUP,SUBJECT)), y=N, fill=GROUP)) + 
    geom_bar(stat='Identity') +
    scale_fill_manual(values=col.pal) +
    ggtitle(paste0(n[[i]], ' Protein counts')) +
    theme_classic() +
    theme(axis.text.x = element_text(angle=90))
 g
  BackupAsPDF(g, paste0(n[[i]], '.protein.counts.'))
},x=dp.out.list, n=names(dp.out.list))
```



MRX2 data looks good, but issues with QRFPR.
For the QRFPR, run a second round of normalisation

```{r}
dt <- data.table(dp.out.list$QRFPR$ProteinLevelData) 

pmat <- data.table(dp.out.list$QRFPR$ProteinLevelData) %>% 
  dcast(., Protein~interaction(GROUP,SUBJECT), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')

# tukey median polish proceedure
tmp <- medpolish(pmat, na.rm = T)

tmp.dt <- data.table(sample = names(tmp$col),
                     offset = tmp$col
                     )

dt[, sample := interaction(GROUP,SUBJECT)]

dt[tmp.dt, offset := i.offset, on=c("sample")]
dt[, new.LogIntensities := LogIntensities - offset,]

# looks much better save this to the data
ggplot(dt, aes(x=interaction(GROUP,SUBJECT), y=new.LogIntensities, fill=GROUP)) + geom_boxplot()


dt[, LogIntensities := new.LogIntensities]

# use this if rerunning
#fwrite(dt, './output/QRFPR.ProteinlvlData.tmpNormalized.csv.gz')

dp.out.list$QRFPR$ProteinLevelData <- dt
```

heatmaps of 3K subsample of proteins

```{r}
# renmae the variables to add a control flag
dp.out.list$QRFPR$ProteinLevelData <- setDF(data.table(dp.out.list$QRFPR$ProteinLevelData)[!GROUP %like% ('YM|PTX'), GROUP := paste0(GROUP,'_CTRL')])
dp.out.list$MRX2$ProteinLevelData <- setDF(data.table(dp.out.list$MRX2$ProteinLevelData)[!GROUP %like% ('YM|PTX'), GROUP := paste0(GROUP,'_CTRL')])
```

```{r}
dp.out.list$QRFPR$ProteinLevelData$GROUP %>%  unique()

lapply(seq_along(dp.out.list), function(x,n,i){
  
  p.mat <-  data.table(x[[i]]$ProteinLevelData) %>% 
    dcast(., Protein~interaction(GROUP,SUBJECT), value.var = 'LogIntensities') %>% 
    as.matrix(rownames='Protein')
  
  subMat <- p.mat[sample(rownames(p.mat),2000), ]

  #heatmap annotation 
  grp <- sub(".[1-6]$", "", colnames(subMat))
  time <-  str_extract(colnames(subMat), "00|05|10")
  treat <- str_extract(colnames(subMat), "CTRL|PTX|YM_PTX|YM")
  batch <- sub(".","",str_extract(colnames(subMat), ".[1-6]$"))

  
  ha <- HeatmapAnnotation(group = grp, treatment=treat, time=time, batch=batch, col=list(time=c('00'='red', '05'='blue', '10'='green'),
                                                                                         treatment=c('PTX'='orange', 'YM_PTX'='yellow','YM'='palegreen', 'CTRL'='grey'),
                                                                                         batch=c('1'='pink','2'='cyan', '3'='palegreen', '4'='purple', '5'='orange', '6'='lightyellow')))
  
  hm <-  Heatmap(subMat, 
          cluster_rows = clusterwNA(subMat),
          show_row_names = F,
          name = 'Log2 Intensities',
          row_title = sprintf('%s sampled proteins', nrow(subMat)),
          top_annotation = ha)
  
  BackupAsPDF(draw(hm, column_title=n[[i]]), paste0(n[[i]], '.intensity.heatmaps.'),  dimensions=c(8,8))
  
}, x=dp.out.list, n=names(dp.out.list))

```
heatmaps of median swept plots

```{r}
lapply(seq_along(dp.out.list), function(x,n,i){
  
  p.mat <-  data.table(x[[i]]$ProteinLevelData) %>% 
    dcast(., Protein~interaction(GROUP,SUBJECT), value.var = 'LogIntensities') %>% 
    as.matrix(rownames='Protein')
  
  # subtract median vals from rows
  subMat <- sweep(p.mat, 1, FUN='-', STATS=apply(p.mat, 1, median, na.rm=T))
  
  subMat <- subMat[sample(rownames(subMat),2000), ]

  #heatmap annotation 
  grp <- sub(".[1-6]$", "", colnames(subMat))
  time <-  str_extract(colnames(subMat), "00|05|10")
  treat <- str_extract(colnames(subMat), "CTRL|PTX|YM_PTX|YM")
  batch <- sub(".","",str_extract(colnames(subMat), ".[1-6]$"))

  
  ha <- HeatmapAnnotation(group = grp, treatment=treat, time=time, batch=batch, col=list(time=c('00'='red', '05'='blue', '10'='green'),
                                                                                         treatment=c('PTX'='orange', 'YM_PTX'='yellow','YM'='palegreen', 'CTRL'='grey'),
                                                                                         batch=c('1'='pink','2'='cyan', '3'='palegreen', '4'='purple', '5'='orange', '6'='lightyellow')))
  
  hm <-  Heatmap(subMat, 
          cluster_rows = clusterwNA(subMat),
          show_row_names = F,
          name = 'Intensities vs. Median',
          row_title = sprintf('%s sampled proteins', nrow(subMat)),
          column_names_gp = gpar(fontsize=6),
          top_annotation = ha,
          heatmap_legend_param = list(
            legend_direction='vertical',
            legend_width = unit(2, "cm"))
            
          )
  
  #draw(hm, column_title=n[[i]], heatmap_legend_side="bottom")
BackupAsPDF(draw(hm, column_title=n[[i]]), paste0(n[[i]], '.mediansweep.heatmaps.'), dimensions=c(8,8))
  
}, x=dp.out.list, n=names(dp.out.list))
```

Still not seeing a lot of group clustering in the samples
PCAs of each of the normalised samples
Clear issue with batch so want to include this in our model
```{r}
lapply(seq_along(dp.out.list), function(x,n,i){
  
  p.mat <-  data.table(x[[i]]$ProteinLevelData) %>% 
    dcast(., Protein~interaction(GROUP,SUBJECT), value.var = 'LogIntensities') %>% 
    as.matrix(rownames='Protein')
  
  p.mat <- p.mat[complete.cases(p.mat),]
  
  pcaOut <- prcomp(t(p.mat))

  colInfo <- data.table(colname = colnames(p.mat))
  colInfo[,c("treat","rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ] # v useful fucntion


#PCA
  pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
  pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
  pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
  #plot first two components
  p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = treat, shape=rep)) + 
    geom_point(alpha=1.0, size=4) + 
    ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
    ggtitle(n[[i]]) +
    theme_bw() + 
    xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
    ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
    ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_shape_manual(values=c(21:26)) +
    guides(fill = guide_legend(override.aes = list(shape =21) ) ,#legend settings
         color = guide_legend(override.aes = list(shape =21) ) )

 p
  BackupAsPDF(p, paste0(n[[i]], '.pca.complete.proteins.'))

  
}, x=dp.out.list, n=names(dp.out.list))
```
APEX works by labeling neighboring/interacting proteins with biotin which is then used to purify labeled proteins. There are also proteins that are endogenously biotinylated which will co-purify with the APEX-labeled proteins.  Here we look at a subset of these endogenous biotin proteins, and we inspect their post-normalization background levels, which are inversely related to the labeling-efficiency of APEX.  More background after normalization implies there is less APEX-labeled signal.

possible issue here; big differences in labeling efficiency for QRFPR rep 6 there is a drastic drop compared to the other groups
```{r}
biotin.carboxylases.up <- c("O00763","P05165","P11498","Q13085","Q96RQ3")

lapply(seq_along(dp.out.list), function(x,n,i){
  dt <- data.table(x[[i]]$ProteinLevelData)
  
  p <- ggplot(dt[Protein %in% biotin.carboxylases.up], aes (x = interaction(GROUP,SUBJECT), y = LogIntensities, color = Protein)) + 
    geom_line(aes(group = Protein)) + 
    geom_point(aes(shape = as.factor(SUBJECT))) +
    theme_classic() +
    ggtitle(n[[i]]) +
    theme(axis.text.x= element_text(angle=90))
 p
 BackupAsPDF(p, paste0(n[[i]], '.endoBiotinylatedProts.linechart.'))
  },x=dp.out.list,n=names(dp.out.list))
```


differential expression
-----

First things; create two seperate list of contrasts, one for each dataset and add to list
Give lists same name as the dp.out objects and process

```{r}
# first MRX2 
mat.dimnames <- list(rnames='ZC05-ZC00',
                     cnames= as.character(dp.out.list$MRX2$ProteinLevelData$GROUP %>% unique()))


mrx2.contrasts <- matrix(data= c(-1,0,0,1,0,0,0,0,0,0), nrow = 1, dimnames=mat.dimnames)
mrx2.contrasts <- rbind(mrx2.contrasts, c(-1,0,0,0,0,0,1,0,0,0))
rownames(mrx2.contrasts)[2] <- 'ZC10-ZC00' 
mrx2.contrasts <- rbind(mrx2.contrasts, c(0,0,-1,0,0,1,0,0,0,0))
rownames(mrx2.contrasts)[3] <- 'ZC05_YM-ZC00_YM'
mrx2.contrasts <- rbind(mrx2.contrasts, c(0,0,-1,0,0,0,0,0,1,0))
rownames(mrx2.contrasts)[4] <- 'ZC10_YM-ZC00_YM'
mrx2.contrasts <- rbind(mrx2.contrasts, c(0, -1,0,0,1,0,0,0,0,0))
rownames(mrx2.contrasts)[5] <- 'ZC05_PTX-ZC00_PTX'
mrx2.contrasts <- rbind(mrx2.contrasts, c(0, -1,0,0,0,0,0,1,0,0))
rownames(mrx2.contrasts)[6] <- 'ZC10_PTX-ZC00_PTX'
mrx2.contrasts <- rbind(mrx2.contrasts, c(-1,0,1,0,0,0,0,0,0,0))
rownames(mrx2.contrasts)[7] <- 'ZC00_YM-ZC00'
mrx2.contrasts <- rbind(mrx2.contrasts, c(-1,1,0,0,0,0,0,0,0,0))
rownames(mrx2.contrasts)[8] <- 'ZC00_PTX-ZC00'
mrx2.contrasts <- rbind(mrx2.contrasts, c(0,0,0,-1,0,1,0,0,0,0))
rownames(mrx2.contrasts)[9] <- 'ZC05_YM-ZC05'
mrx2.contrasts <- rbind(mrx2.contrasts, c(0,0,0,-1,1,0,0,0,0,0))
rownames(mrx2.contrasts)[10] <- 'ZC05_PTX-ZC05' 
mrx2.contrasts <- rbind(mrx2.contrasts, c(0,0,0,0,0,0,-1,0,1,0))
rownames(mrx2.contrasts)[11] <- 'ZC10_YM-ZC10' 
mrx2.contrasts <- rbind(mrx2.contrasts, c(0,0,0,0,0,0,-1,1,0,0))
rownames(mrx2.contrasts)[12] <- 'ZC10_PTX-ZC10'
mrx2.contrasts <- rbind(mrx2.contrasts, c(0,0,0,0,0,0,-1,0,0,1))
rownames(mrx2.contrasts)[13] <- 'ZC10_YM_PTX-ZC10'

mrx2.contrasts
```
Now do the NPFFR1 contrasts 

```{r}
# first MRX2 
mat.dimnames <- list(rnames='RF05-RF00',
                     cnames= as.character(dp.out.list$QRFPR$ProteinLevelData$GROUP %>% unique()))

 
qrfpr.contrasts <- matrix(data= c(-1,0,0,1,0,0,0,0,0,0), nrow = 1, dimnames=mat.dimnames)
qrfpr.contrasts <- rbind(qrfpr.contrasts, c(-1,0,0,0,0,0,1,0,0,0))
rownames(qrfpr.contrasts)[2] <- 'RF10-RF00' 
qrfpr.contrasts <- rbind(qrfpr.contrasts, c(0,0,-1,0,0,1,0,0,0,0))
rownames(qrfpr.contrasts)[3] <- 'RF05_YM-RF00_YM'
qrfpr.contrasts <- rbind(qrfpr.contrasts, c(0,0,-1,0,0,0,0,0,1,0))
rownames(qrfpr.contrasts)[4] <- 'RF10_YM-RF00_YM'
qrfpr.contrasts <- rbind(qrfpr.contrasts, c(0, -1,0,0,1,0,0,0,0,0))
rownames(qrfpr.contrasts)[5] <- 'RF05_PTX-RF00_PTX'
qrfpr.contrasts <- rbind(qrfpr.contrasts, c(0, -1,0,0,0,0,0,1,0,0))
rownames(qrfpr.contrasts)[6] <- 'RF10_PTX-RF00_PTX'
qrfpr.contrasts <- rbind(qrfpr.contrasts, c(-1,0,1,0,0,0,0,0,0,0))
rownames(qrfpr.contrasts)[7] <- 'RF00_YM-RF00'
qrfpr.contrasts <- rbind(qrfpr.contrasts, c(-1,1,0,0,0,0,0,0,0,0))
rownames(qrfpr.contrasts)[8] <- 'RF00_PTX-RF00'
qrfpr.contrasts <- rbind(qrfpr.contrasts, c(0,0,0,-1,0,1,0,0,0,0))
rownames(qrfpr.contrasts)[9] <- 'RF05_YM-RF05'
qrfpr.contrasts <- rbind(qrfpr.contrasts, c(0,0,0,-1,1,0,0,0,0,0))
rownames(qrfpr.contrasts)[10] <- 'RF05_PTX-RF05' 
qrfpr.contrasts <- rbind(qrfpr.contrasts, c(0,0,0,0,0,0,-1,0,1,0))
rownames(qrfpr.contrasts)[11] <- 'RF10_YM-RF10' 
qrfpr.contrasts <- rbind(qrfpr.contrasts, c(0,0,0,0,0,0,-1,1,0,0))
rownames(qrfpr.contrasts)[12] <- 'RF10_PTX-RF10'
qrfpr.contrasts <- rbind(qrfpr.contrasts, c(0,0,0,0,0,0,-1,0,0,1))
rownames(qrfpr.contrasts)[13] <- 'RF10_YM_PTX-RF10'

qrfpr.contrasts
```
now combine the two contrast matrices as a list

perform the pairwise comparisons for each; may need to control for correcting so many at once


Not recovering anything in the comparisons of interest
Read back in the dp.processed data 

```{r}
MRX2.dp  <- readRDS('./output/MRX2.rds')
QRFPR.dp <- readRDS('./output/QRFPR.rds')

dp.out.list <- list(MRX2 = MRX2.dp,
                    QRFPR = QRFPR.dp)
```

check out the MRX2 data first. Looks fine when adjusting names so run for both datasets

```{r}
colnames(mrx2.contrasts) <-  gsub("_CTRL","", colnames(mrx2.contrasts))
colnames(qrfpr.contrasts) <-  gsub("_CTRL","", colnames(qrfpr.contrasts))

contrast.list <- list(MRX2=mrx2.contrasts,
                      QRFPR=qrfpr.contrasts)


```



```{r}
# add batch term for linear modelling and correct the feature levle data grp names
mss.in.list  <- lapply(dp.out.list, function(x){
  x$ProteinLevelData$SUBJECT <- paste0('batch.', x$ProteinLevelData$SUBJECT)
  x$FeatureLevelData$SUBJECT <- paste0('batch.', x$FeatureLevelData$SUBJECT)
  return(x)
  })


# run the contrasts

mss.out.list <- lapply(seq_along(mss.in.list), function(x,n,i){
  
  recpt <- n[[i]]
  print(paste0('running comparison on ', recpt))
  
  # run msstats correcting for batch 
  mss <- groupComparison(contrast.matrix=contrast.list[[recpt]], data=x[[recpt]])
  mss.dt <- setDT(mss$ComparisonResult)
  
  return(mss.dt)
  
},x=mss.in.list, n=names(mss.in.list))


names(mss.out.list) <- names(mss.in.list)
```

Identify the number of degs in each condition

```{r eval=FALSE}
lapply(seq_along(mss.out.list), function(x,n,i){
  fwrite(x[[i]], paste0('./output/', n[[i]], '.mss.pwcomparisons.raw.csv'))
  
},x=mss.out.list,n=names(mss.out.list))

lapply(mss.out.list, head)
```

count number of DEGs
More DEGs seem to be found in the QRFPR samples
```{r}
lapply(seq_along(mss.out.list), function(x,n,i){
  dt <- x[[i]]
  
  dt[,gene := multiUniprots2multiGenes(as.character(Protein))] %>% #needed as in factor format
  .[, sig := 'not'] %>% 
  .[!issue %in% c('completeMissing','oneConditionMissing') & abs(log2FC) > 0.58 & adj.pvalue < 0.05, sig := ifelse(log2FC > 0, 'up', 'down')]
  
   fwrite(dt, paste0('./output/', n[[i]], '.mss.pwcomparisons.proc.csv'))
  }, x=mss.out.list, n=names(mss.out.list))
```

Volcano plots of all the comparisons

```{r}
lapply(seq_along(mss.out.list), function(x,n,i){
  dt <- x[[i]]
  
  # vecotr to iterate through
  contrasts.oi <- dt$Label %>%  unique()
  
  # nested loop w 
  lapply(contrasts.oi, function(cont){
    
      g <- ggplot(dt[Label == cont,], aes(x=log2FC, y=-log10(adj.pvalue), color=sig, label=gene)) + 
      geom_point() +
      geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
      geom_hline(yintercept = -log10(0.05), linetype=2, alpha=0.4) +
      ggrepel::geom_text_repel(data=dt[Label == cont & sig != 'not',],  size = 2, max.overlaps = 20) +
      scale_color_manual(values=c('up'='red', 'down'='blue', 'not'='grey')) +
      ggtitle(paste0(n[[i]], ' ', cont, ' (FC +/ 50% & p.adj < 0.05)')) +
      # facet_wrap(~Label, scales = 'free') +
      theme_bw() +
      theme(strip.text.x = element_text(size = 5))
      
      BackupAsPDF(g, paste0(n[[i]],'.',cont,'.volcanoplot.'))
    
  })
  
  }, x=mss.out.list, n=names(mss.out.list))
```

Generate heat maps of sig genes in each comparison
Looks like a very fast interaction; already less sig hits by 10 mins

```{r}
lapply(seq_along(mss.out.list), function(x,y,n,i){
  
  mss <- x[[i]]
  dt <- data.table(y[[i]]$ProteinLevelData)
  
  # get the receptor expres matrix
  p.mat <-  dt %>% 
    dcast(., Protein~interaction(GROUP,SUBJECT), value.var = 'LogIntensities') %>% 
    as.matrix(rownames='Protein')

  # vecotor to iterate through
  contrasts.oi <- mss[,unique(Label)]

  # nested loop w 
  lapply(contrasts.oi, function(cont){
    
    # extract the sig genes for each contrast
    sig.prots <- mss[sig != 'not' & !issue %in% c("oneConditionMissing","completeMissing") & Label == cont, unique(Protein)]
    
    if (length(sig.prots) > 2) {

      # extract the submat of sig prots
      subMat <- p.mat[rownames(p.mat) %in% sig.prots, ]

      #heatmap annotation 
      grp <- sub(".[1-6]$", "", colnames(subMat))
      time <-  str_extract(colnames(subMat), "00|05|10")
      treat <- str_extract(colnames(subMat), "CTRL|PTX|YM_PTX|YM")
      batch <- sub(".","",str_extract(colnames(subMat), ".[1-6]$"))

  
      ha <- HeatmapAnnotation(group = grp, treatment=treat, time=time, batch=batch, col=list(time=c('00'='red', '05'='blue', '10'='green'),
                                                                                         treatment=c('PTX'='orange', 'YM_PTX'='yellow','YM'='palegreen', 'CTRL'='grey'),
                                                                                     batch=c('1'='pink','2'='cyan', '3'='palegreen', '4'='purple', '5'='orange', '6'='lightyellow')))
      hm <-  Heatmap(subMat, 
                     cluster_rows = clusterwNA(subMat), 
                     top_annotation = ha, 
                     name='Log2 Intensities', 
                     row_names_gp = gpar(fontsize=3), 
                     column_names_gp = gpar(fontsize=6),
                     row_title = sprintf("%s sig proteins", nrow(subMat)) )
    
    
    BackupAsPDF(draw(hm, column_title=cont), paste0(n[[i]],'.', cont, '.sig.prots.heatmap.'), dimensions=c(8,8))
    #draw(hm, column_title=cont)
    }
  })
  
  }, x=mss.out.list, y=dp.out.list, n=names(mss.out.list))
```
median scaled heatmap

```{r}
lapply(seq_along(mss.out.list), function(x,y,n,i){
  
  mss <- x[[i]]
  dt <- data.table(y[[i]]$ProteinLevelData)
  
  # get the receptor expres matrix
  p.mat <-  dt %>% 
    dcast(., Protein~interaction(GROUP,SUBJECT), value.var = 'LogIntensities') %>% 
    as.matrix(rownames='Protein')
  
  subMat <- sweep(p.mat, MARGIN = 1, FUN = '-', STAT=apply(p.mat, 1, median, na.rm=T))

  # vecotor to iterate through
  contrasts.oi <- mss[,unique(Label)]

  # nested loop w 
  lapply(contrasts.oi, function(cont){
    
    # extract the sig genes for each contrast
    sig.prots <- mss[sig != 'not' & !issue %in% c("oneConditionMissing","completeMissing") & Label == cont, unique(Protein)]
    
    if (length(sig.prots) > 2) {

      # extract the submat of sig prots
      subMat <- subMat[rownames(subMat) %in% sig.prots, ]

      #heatmap annotation 
      grp <- sub(".[1-6]$", "", colnames(subMat))
      time <-  str_extract(colnames(subMat), "00|05|10")
      treat <- str_extract(colnames(subMat), "CTRL|PTX|YM_PTX|YM")
      batch <- sub(".","",str_extract(colnames(subMat), ".[1-6]$"))

  
      ha <- HeatmapAnnotation(group = grp, treatment=treat, time=time, batch=batch, col=list(time=c('00'='red', '05'='blue', '10'='green'),
                                                                                         treatment=c('PTX'='orange', 'YM_PTX'='yellow','YM'='palegreen', 'CTRL'='grey'),
                                                                                     batch=c('1'='pink','2'='cyan', '3'='palegreen', '4'='purple', '5'='orange', '6'='lightyellow')))
      hm <-  Heatmap(subMat, 
                     cluster_rows = clusterwNA(subMat), 
                     top_annotation = ha, 
                     name='Log2 Intensities - Median', 
                     row_names_gp = gpar(fontsize=3), 
                     column_names_gp = gpar(fontsize=6),
                     row_title = sprintf("%s sig proteins", nrow(subMat)) )
    
    
    BackupAsPDF(draw(hm, column_title=cont), paste0(n[[i]],'.', cont, '.sig.prots.medianScaled.heatmap.'), dimensions=c(8,8))
    #draw(hm, column_title=cont)
    }
  })
  
  }, x=mss.out.list, y=dp.out.list, n=names(mss.out.list))
```



generate heatmap of detected sig proteins in the data
Only plotting all proteins not missing in one of the comparisons

```{r}
lapply(seq_along(mss.out.list), function(x,y,n,i){
  
  mss <- x[[i]]
  #get sig proteins across all comparisons
  sig.prots <- mss[sig != 'not' & !issue %in% c("oneConditionMissing","completeMissing") , unique(Protein)]

  
  dt <- data.table(y[[i]]$ProteinLevelData)
  
  # get the receptor expres matrix
  p.mat <-  dt %>% 
    dcast(., Protein~interaction(GROUP,SUBJECT), value.var = 'LogIntensities') %>% 
    as.matrix(rownames='Protein')

    # extract the submat of sig prots
    subMat <- p.mat[rownames(p.mat) %in% sig.prots, ]
    
    rownames(subMat) <- multiUniprots2multiGenes(rownames(subMat))

    #heatmap annotation 
    grp <- sub(".[1-6]$", "", colnames(subMat))
    time <-  str_extract(colnames(subMat), "00|05|10")
    treat <- str_extract(colnames(subMat), "CTRL|PTX|YM_PTX|YM")
    batch <- sub(".","",str_extract(colnames(subMat), ".[1-6]$"))

  
    ha <- HeatmapAnnotation(group = grp, treatment=treat, time=time, batch=batch, col=list(time=c('00'='red', '05'='blue', '10'='green'),
                                                                                         treatment=c('PTX'='orange', 'YM_PTX'='yellow','YM'='palegreen', 'CTRL'='grey'),
                                                                                        batch=c('1'='pink','2'='cyan', '3'='palegreen', '4'='purple', '5'='orange', '6'='lightyellow')))
  
    hm <-  Heatmap(subMat, 
                   cluster_rows = clusterwNA(subMat), 
                   top_annotation = ha,
                   row_names_gp = gpar(fontsize=2),
                   column_names_gp = gpar(fontsize=6),
                   name = 'Log2 Intensities',
                   row_title = paste0(sprintf('%s sig proteins', nrow(subMat)), ' ', n[[i]]) )

    BackupAsPDF(hm, paste0(n[[i]], '.sigProteins.noCondMissing.heatmap.'), dimensions=c(8,12))
  
  }, x=mss.out.list, y=dp.out.list, n=names(mss.out.list))

```
median sweep plot on the data 
```{r}
lapply(seq_along(mss.out.list), function(x,y,n,i){
  
  mss <- x[[i]]
  #get sig proteins across all comparisons
  sig.prots <- mss[sig != 'not' & !issue %in% c("oneConditionMissing","completeMissing") , unique(Protein)]
  
  print(length(sig.prots))
  
  dt <- data.table(y[[i]]$ProteinLevelData)
  
  # get the receptor expres matrix
  p.mat <-  dt %>% 
    dcast(., Protein~interaction(GROUP,SUBJECT), value.var = 'LogIntensities') %>% 
    as.matrix(rownames='Protein')
  
   subMat <- sweep(p.mat, MARGIN = 1, FUN = '-', STATS=apply(p.mat, 1, median, na.rm=T))

    # extract the submat of sig prots
    subMat <- subMat[rownames(subMat) %in% sig.prots, ]
    
    rownames(subMat) <- multiUniprots2multiGenes(rownames(subMat))

    #heatmap annotation 
    grp <- sub(".[1-6]$", "", colnames(subMat))
    time <-  str_extract(colnames(subMat), "00|05|10")
    treat <- str_extract(colnames(subMat), "CTRL|PTX|YM_PTX|YM")
    batch <- sub(".","",str_extract(colnames(subMat), ".[1-6]$"))

  
    ha <- HeatmapAnnotation(group = grp, treatment=treat, time=time, batch=batch, col=list(time=c('00'='red', '05'='blue', '10'='green'),
                                                                                         treatment=c('PTX'='orange', 'YM_PTX'='yellow','YM'='palegreen', 'CTRL'='grey'),
                                                                                        batch=c('1'='pink','2'='cyan', '3'='palegreen', '4'='purple', '5'='orange', '6'='lightyellow')))
  
    hm <-  Heatmap(subMat, 
                   cluster_rows = clusterwNA(subMat), 
                   top_annotation = ha,
                   row_names_gp = gpar(fontsize=2),
                   column_names_gp = gpar(fontsize=6),
                   name = 'Log2 Intensities',
                   row_title = paste0(sprintf('%s sig proteins', nrow(subMat)), ' ', n[[i]]) )

    BackupAsPDF(hm, paste0(n[[i]], '.sigProteins.noCondMissing.medianSweep.heatmap.'), dimensions=c(8,12))
  
  }, x=mss.out.list, y=dp.out.list, n=names(mss.out.list))

```
Go enrichment for each of the differentially expressed gene groups across comparisons
Not very strong enrichment across groups; could 

```{r}

gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "UNIPROT")

#universe: all proteins detected in exp
universe <- sapply(dp.out.list, function(x){
  
 as.character(x$ProteinLevelData$Protein)
  #x$ProteinLevelData[,unique(Protein)]
}) %>%  unlist() %>%  unique()


lapply(seq_along(mss.out.list), function(x,y,n,i){
  
  mss <- x[[i]]
  mss[,enrich_grp := interaction(Label,sig)]
  
  enrich.dt <- enricherOnGroups(mss[sig != 'not',], groupColumns = 'enrich_grp', geneColumn = "Protein", term2gene.gmt = gmt.go, universe = universe)
  
  simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')
  
  hm <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, NULL, groupColumn = 'enrich_grp', topN = 8,
                                  negCols=unique(simp.enrich$simplified$enrich_grp[grep('down', simp.enrich$simplified$enrich_grp)]), 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 5), upperThreshold = 8) 
  
  BackupAsPDF(hm, paste0(n[[i]], '.GOenrichment.heatmap.'), dimensions=c(8,8))

  }, x=mss.out.list, y=dp.out.list, n=names(mss.out.list))


# run enrichment on non-missing hits

lapply(seq_along(mss.out.list), function(x,y,n,i){
  
  mss <- x[[i]]
  mss[,enrich_grp := interaction(Label,sig)]
  
  enrich.dt <- enricherOnGroups(mss[sig != 'not' & !issue %in% c('completeMissing','oneConditionMissing'),], groupColumns = 'enrich_grp', geneColumn = "Protein", term2gene.gmt = gmt.go, universe = universe)
  
  simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')
  
  hm <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, NULL, groupColumn = 'enrich_grp', topN = 8,
                                  negCols=unique(simp.enrich$simplified$enrich_grp[grep('down', simp.enrich$simplified$enrich_grp)]), 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 5), upperThreshold = 8) 
  
  BackupAsPDF(hm, paste0(n[[i]], '.GOenrichment.noNAConditions.heatmap.'), dimensions=c(8,8))

  }, x=mss.out.list, y=dp.out.list, n=names(mss.out.list))


```

TODO 
----
Check heatmaps & volcano pmatch
Question: these are only for prots w/o missing conditions. What are the proportion of sig genes in each that are infinite?
Is the proportion of NA increasing? all RF10 treatments have more missing values, same with ZC10

Maybe isolate the 

```{r}
lapply(seq_along(mss.out.list), function(x,n,i){
  
  x[[i]][!issue %in% c('completeMissing','oneConditionMissing'), issue := 'noConditionsMissing']
  dt <- x[[i]][sig != 'not' & issue != 'completeMissing',.N, by=.(sig,Label,issue)]

  
 g <-  ggplot(dt, aes(x=Label, y=N, fill=issue)) + 
    geom_bar(stat = 'Identity') + 
    scale_fill_manual(values=c('noConditionsMissing'='orange','oneConditionMissing'='grey')) + 
    ggtitle(paste0('Breakdown of sig hits ', n[[i]])) +
    theme_classic() + 
    theme(axis.text.x=element_text(angle=90))
 
 
 #BackupAsPDF(g, 'sig.hits.breakdown.NAvals.barchart')
 
  g <-  ggplot(dt[issue == 'noConditionsMissing',], aes(x=Label, y=N, fill=sig)) + 
    geom_bar(stat = 'Identity') + 
    scale_fill_manual(values=c('up'='red','down'='blue')) + 
    ggtitle(paste0('Breakdown of sig hits ', n[[i]])) +
    theme_classic() + 
    theme(axis.text.x=element_text(angle=90))
  
  BackupAsPDF(g, 'sig.hits.breakdown.up.vs.down.barchart')
  
},x=mss.out.list, n=names(mss.out.list))
```

Lots of missing values in the dataset. Rerun the enrichment using only the sig hits detected in both conditions compared

```{r}
MRX2.mss <-  fread('./output/MRX2.mss.pwcomparisons.proc.csv')
QRFPR.mss <- fread('./output/QRFPR.mss.pwcomparisons.proc.csv')

mss.out.list <- list(MRX2 = MRX2.mss,
                     QRFPR = QRFPR.mss)


# run enrichment on non-missing hits
lapply(seq_along(mss.out.list), function(x,y,n,i){
  
  mss <- x[[i]]
  mss[,enrich_grp := interaction(Label,sig)]
  
  enrich.dt <- enricherOnGroups(mss[sig != 'not' & !issue %in% c('completeMissing','oneConditionMissing'),], groupColumns = 'enrich_grp', geneColumn = "Protein", term2gene.gmt = gmt.go, universe = universe)
  
  simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')
  
  hm <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, NULL, groupColumn = 'enrich_grp', topN = 8,
                                  negCols=unique(simp.enrich$simplified$enrich_grp[grep('down', simp.enrich$simplified$enrich_grp)]), 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 5), upperThreshold = 8) 
  
  BackupAsPDF(hm, paste0(n[[i]], '.GOenrichment.noNAConditions.heatmap.'), dimensions=c(8,8))

  }, x=mss.out.list, y=dp.out.list, n=names(mss.out.list))

```
TODO
--
Maybe try renormalise the QRFPR data,  but results look good

