---
title: "DE.ph.291123"
author: "Martin Gordon"
date: "2023-11-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis of the PH DARPA data
Differential expression analysis of PH data
Murine origin, so be wary of this when converting between identifiers

Simple experimental set-up: 4 conditions; control (veh), compound (Nbon), antagonist (M100), M100_Nbon (antagonist + compound)
4 replicates per condition

Have both phospho and abundance data available (abundance data inital analysis)
Goal is to present by Dec 4 DARPA II meeting

```{r}
```{r}
library(ggplot2)
library(data.table)
library(magrittr)
library(ComplexHeatmap)
library(randomcoloR)
library(stringr)


#source BP utility scripts
source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/MSstats_Helper_Functions.R")
source("/Users/martingordon/Documents/utils/bp_utils/MSstats_V4_Functions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/linearModels.R")

# mg utility functions
source("/Users/martingordon/Documents/utils/mg_utils/HelperFunctions.R")
```

