---
title: "WGCNA.paperFig5.correlationPlots"
author: "Martin Gordon"
date: "2023-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Generating plots of the correlation scores from figure 5 in the DARPA proteomics paper

```{r}
library(data.table)
library(ggplot2)
library(ComplexHeatmap)
library(WGCNA)
library(magrittr)
library(preprocessCore) # quantile normalisation
library(stringr)
library(matrixStats)
library(simplifyEnrichment)
library(fmsb)
library(readxl)
library(ggbeeswarm)
library(ggrepel)
library(curl)
library(dendextend) # extension for dendogram

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")


# helper scripts for module/network export etc
source ("/Users/martingordon/Documents/utils/mg_utils/HelperFunctions.R")
source ("/Users/martingordon/Documents/utils/mg_utils/plottingHelperFunctions.R")

```

plots we wish to generate
- rank plot of the correlation scores
- heatmaps of the top ranking genes [maybe take top 10-20 per group] and plot log2FC and correlation values
- a kind of scatterplot of the correlation scores

lets first look at the data we have; are these cor scores correctly calculated?

06/09/23
-----
Correlations with the IUPHAR classification

lets first check the raw data 

```{r}
pred <- fread("./data/predictionScores.txt")
exp <- fread ("./data/experimentalScores.txt")
iupar <- data.table(read_xlsx("./data/IUPHAR_11GPCRs_Gprotein_class_primary.xlsx"))

iupar

exp.long <- melt(exp, id.vars = "GPCR", measure.vars = grep ("GNA", colnames(pred)), variable.name = "GProtein", value.name = "score")[, score := as.numeric(score)][]
pred.long <- melt(pred, id.vars = "GPCR", measure.vars = grep ("GNA", colnames(pred)), variable.name = "GProtein", value.name = "score")[, score := as.numeric(score)][]

# convert to long format
iupar.long <- melt(iupar, id.vars = 'GPCR', measure.vars = grep('GPCR', colnames(iupar), invert=T),  variable.name = "GProtein", value.name = "score") %>% 
  .[, score := as.numeric(score)] 
```


```{r scores.list}
scores.list <- list(exp = exp.long,
                    pred = pred.long,
                    iupar = iupar.long)
```
Convert the scores to a matrix

```{r}
scores.mat <- lapply(scores.list, function(x){
  dcast(x, GProtein~GPCR, value.var = 'score') %>% 
    as.matrix(rownames = 'GProtein')
})

lapply(scores.mat, colnames)
```
apply the scores mat to the data
```{r}
scores.mat <- lapply(scores.mat, function(x){
  colnames(x) <- c("MRX2","MRX4","NPFFR1","NPY1R","DOR","KOR","MOR","PTGER1","PTGER4","QRFPR","NK1R")
  return(x)
})
```

reorder the mss columns to follow the order above 

Restrict the analysis to the sig genes from the APEX experiment (pval < 0.005 & abs log2FC(1.5))

```{r}
col.order <-  c("MRX2","MRX4","NPFFR1","NPY1R","DOR","KOR","MOR","PTGER1","PTGER4","QRFPR","NK1R")

mss.out <- fread('./data/2022_08_31_NiceFitsPower3.csv')

exp.mat <- dcast(mss.out,gene~receptor, value.var = 'log2FC') %>% 
  .[, MOR.LG94 := NULL] %>% 
  as.matrix(rownames ='gene')


# filter to sig genes and reorder
sig.genes <- mss.out[pvalue < 0.005 & abs(log2FC) > log2(1.5) & receptor != 'MOR.LG94', unique(gene) ]
exp.mat <- exp.mat[rownames(exp.mat) %in% sig.genes,]


# renanme MOR DAMGO to MOR
colnames(exp.mat) <- c('DOR', 'KOR', 'MOR', 'MRX2', 'MRX4', 'NK1R', 'NPFFR1', 'NPY1R','PTGER1', 'PTGER4','QRFPR')

# reorder the cols to match the matrices
col.order <-  c("MRX2","MRX4","NPFFR1","NPY1R","DOR","KOR","MOR","PTGER1","PTGER4","QRFPR","NK1R")

exp.mat[1:10,1:10]
exp.mat <- exp.mat[, col.order]
exp.mat[1:10,1:10]


# sanit check; looks fine
lapply(scores.mat, colnames)
```

run the correlations for each of the datasets

```{r}
# cors run colwise with the two vectors
cor.list <- lapply(scores.mat, function(x){
  cor(t(exp.mat), t(x), use= "pairwise.complete.obs")
})

cor.long <- lapply(cor.list, function(x){
  x <- melt(as.data.table(x, keep.rownames = TRUE), id.vars = "rn", variable.name  = "GProtein", value.name = "R")
  setnames(x, old = "rn", new = "gene")
  return(x)
})
```

count number of observations, only want to take results with vals observed for all receptors to reduce FPs

```{r}
cor.long$iupar[, .SD[which.max(R)], by=GProtein]

# most of these have values for the 11 receptors
obs.counts <- mss.out[!is.na(log2FC) & receptor != 'MOR.LG94' & gene %in% sig.genes, .N, by = gene]
hist(obs.counts$N, breaks=seq(1,11))

```

add this info to the correlation res

```{r}
lapply(cor.long, function(x){
  
  x[obs.counts, numObservations := N, on = 'gene']
  setorder(x, -numObservations, -R, na.last = TRUE) #strongest associations at top
})

```
Now that we have the correlations, what are the top hits per gprotein 

```{r}
#col.pal <-  randomcoloR::distinctColorPalette(k=length(unique(cor.long$iupar$GProtein)))

# strong +/- correlation to IUPHAR classification
g <-  cor.long$iupar[abs(R) > .7, .N, by=GProtein] %>% 
  ggplot(., aes(x=reorder(GProtein,-N), y=N, fill=GProtein)) +
  geom_bar(stat='Identity') + 
  ggtitle('Strong correlations abs(r) > 0.7') +
  xlab('IUPHAR classification') +
  ylab('N') +
  scale_fill_manual(values=col.pal) +
  theme_classic()

g

BackupAsPDF(g, 'apex.sigHits.pearsonR.0.7.barplot.')
```
```{r}
# strong +/- correlation to IUPHAR classification
g <-  cor.long$iupar[abs(R) > .6, .N, by=GProtein] %>% 
  ggplot(., aes(x=reorder(GProtein,-N), y=N, fill=GProtein)) +
  geom_bar(stat='Identity') + 
  ggtitle('Moderate correlations abs(r) > 0.6') +
  xlab('IUPHAR classification') +
  ylab('N') +
  scale_fill_manual(values=col.pal) +
  theme_classic()

g

BackupAsPDF(g, 'apex.sigHits.pearsonR.0.6.barplot.')
```


As a start, maybe we should check if these APEX hits are enriched for high cor scores relative to bg?
One way we could look at promising hits is to create a background, (maybe random geneset the same size as our sig hits, find a 

```{r}
apply(iupar[,-1],2, sum)
nrow(iupar)
```

What does the distribution of r scores look like for each of the GProteins?
Is this a concern? Why the 0 inflation for Gs? is it just bc the binary vecotr only has two 1's?
```{r}
col.pal <-  c("#AFC4C4","#C472C9" ,"#B7D878")

g <- ggplot(cor.long$iupar, aes(x=R, fill=GProtein)) + 
  geom_histogram() +
  scale_fill_manual(values=col.pal) +
  theme_classic() +
  facet_wrap(~GProtein)

BackupAsPDF(g, 'apex.sigHits.pearsonCor.histogram')

```

rank plot of the
# 
```{r}

cor.long$iupar[, lab:= '']
cor.long$iupar[abs(R) > 0.7, lab:= gene]

lapply(seq_along(gprots), function(x,i){
  
  g <- ggplot(data=cor.long$iupar[GProtein == x[[i]]], aes(x=reorder(gene,R), y=R, col=GProtein, label=lab)) +
    geom_point(size=1) +
    geom_hline(aes(alpha=0.6), yintercept = 0, linetype='dotdash') +
    geom_text_repel(data=cor.long$iupar[GProtein == x[[i]]], aes(label=lab),segment.linetype=3, segement.color='grey', max.overlaps = 30,size=1.5) +
    xlab('genes (Pearson r ordered)') +
    theme_classic() +
    scale_x_discrete(expand=expansion(add=50)) +
    scale_color_manual(values=c(col.pal[[i]])) +
    ggtitle(paste0(x[[i]], ' Pearson r')) +
    theme(axis.text.x = element_blank()
        )
  
 BackupAsPDF(g, 'rankplot.Gprots.R0.7.')
},x=gprots)

```
another rankplot with many of the median values removed
What about only plotting the top 20 percentile in each direction?

```{r}
cor.dt <- cor.long$iupar[numObservations ==11,]
cor.dt
```

```{r}

lapply(seq_along(gprots), function(x,i){
  
g <- ggplot(data=cor.dt[GProtein == x[[i]]], aes(x=reorder(gene,R), y=R, col=GProtein, label=lab)) +
    geom_point(size=1) +
    geom_hline(aes(alpha=0.6), yintercept = 0, linetype='dotdash') +
    geom_text_repel(data=cor.dt[GProtein == x[[i]]], aes(label=lab),segment.linetype=3, segement.color='grey', max.overlaps = 30,size=1.5) +
    xlab('genes (Pearson r ordered)') +
    theme_classic() +
    scale_x_discrete(expand=expansion(add=50)) +
    scale_color_manual(values=c(col.pal[[i]])) +
    ggtitle(paste0(x[[i]], ' Pearson r')) +
    theme(axis.text.x = element_blank()
        )
  
 BackupAsPDF(g, 'rankplot.Gprots.all.R0.7.')
},x=gprots)

```

lets try generate dendograms of the correlation scores
do the requested plots first and come back to this tomorrow maybe..
show BP tomorrow morning and refine clusters maybe before inputting to heatmap

too many nodes to visualise effectively; drop the nodes with no real correlation to GProteins
only plot nodes with a weak cor score (> 0.5)


```{r}
gprot.cor.list

lapply(seq_along(gprot.cor.list), function(x,n,i){
  x[[i]] <-  x[[i]][abs(R) > 0.5,]
  
  mat <- as.matrix(x[[i]][,.(gene,R)], rownames='gene')
  hc <- hclust(dist(mat))
  
  # create dendogram object
  dend <- as.dendrogram(hc)
  cut_height <- 0.3 # modify to set cut height
  cut_tree <- dendextend::cutree(dend, h = cut_height, order_clusters_as_data = FALSE)
  
  # at the set height, extract the cluster members and calculate their mean cor score to the GProtein
  # so, need to extract genes of each cluster (names)  calculate the mean/median R value of this cluster and then use this as the lables inn color_branches
  clust.summary.vec <- c()
  
  for (clust  in unique(cut_tree)){
    genes.oi <-  names(cut_tree[cut_tree == clust])
    mean.val <-  x[[i]][gene %in% genes.oi, mean(R)]
    clust.summary.vec <- c(clust.summary.vec, mean.val)
  }
  print(clust.summary.vec)
  clust.summary.vec <- round(clust.summary.vec,2)
  
  d.plot <- dend %>%
    dendextend::color_branches(., h = cut_height, groupLabels = clust.summary.vec) %>%
    dendextend::set(., "labels_cex", 0.2)

  
  # saving output 
  #change names
  outDir <-  '310723_QLi_GPCR.correlations_data/'
  fName <-  gsub('/', '_',n[[i]])
  
  timeStp <- format(Sys.time(),  "%Y_%m_%d__%H_%M__%S")
  pdf(paste0(outDir,fName, '.R0.5.dendogram.', timeStp,'.pdf'), width=10, height=7)
  plot(d.plot, main=paste0(n[[i]], ' Pearson r (>0.5)'))
  dev.off()

    #ggsave(paste0(n[[i]], 'Ro.5.dendogram'), plot=plot(d.plot, main=paste0(n[[i]], ' Pearson r (>0.5)'))
   # BackupAsPDF(plot(d.plot), paste0(n[[i]], '.R0.5.dendo.'))
    #plot(main=paste0(n[[i]], ' Pearson r (>0.5)'))
},x=gprot.cor.list, n=names(gprot.cor.list))

```
what about subsampling 

```{r}
#subset to top values either side of the 
lapply(seq_along(gprots), function(x,i){
  
g <- ggplot(data=cor.dt[GProtein == x[[i]]], aes(x=reorder(gene,R), y=R, col=GProtein, label=lab)) +
    geom_point(size=1) +
    geom_hline(aes(alpha=0.6), yintercept = 0, linetype='dotdash') +
    geom_text_repel(data=cor.dt[GProtein == x[[i]]], aes(label=lab),segment.linetype=3, segement.color='grey', max.overlaps = 30,size=1.5) +
    xlab('genes (Pearson r ordered)') +
    theme_classic() +
    scale_x_discrete(expand=expansion(add=50)) +
    scale_color_manual(values=c(col.pal[[i]])) +
    ggtitle(paste0(x[[i]], ' Pearson r')) +
    theme(axis.text.x = element_blank()
        )

g
 BackupAsPDF(g, 'rankplot.Gprots.all.R0.7.')
},x=gprots)
```
geenrate plot of cor scores vs log2FC
can use this to filter out genes that have a small log2FC 

we have 11 scores, 1 per receptor
x axis cor scores, y axis log2FC


combine the mss_subset with cor scores
keep only things with 11 observations 

```{r}
# subset to the genes with 11 observations
mss.subset <- mss.out[receptor != 'MOR.LG94' & gene %in% cor.dt$gene, .(receptor,gene,log2FC,pvalue)]

# calculate some summary statistics per gene; mean, median,SD, variance mad...
mss.subset[, `:=`(Mean=mean(log2FC), Median=median(log2FC), sd=sd(log2FC), Variance=var(log2FC), MAD=mad(log2FC)), by=gene]


# want all possible combos per key so use allow.cartesian
merged.dt <- merge(x=mss.subset, y=cor.dt[,.(gene,GProtein,R)], all.x=T, all.y=T, by.x='gene', by.y='gene', allow.cartesian = T)

#sanitry check
#merged.dt[, .N, by=gene]$N %>%  range()
```

now lets create a 'volcano plot' of correlation and log2FC scores
These are R scores vs log2FC values

mean-median log2FC + variance?
label the most significant genes
geometric mean of log2FC and correlation (need to just take mean of cor as otherwise will be outweighted by log2FC vals)

```{r}
# geo mean of the abs R and summary values per gene
merged.dt[, sigScore := sqrt(abs(R) * abs(Mean)), by=gene]

# just label thethings in the top 25% 
merged.dt[, lab := ifelse(sigScore >= quantile(merged.dt$sigScore, .75), gene, '')]

unique(merged.dt)

g <- ggplot(merged.dt, aes(x=R, y=Median, col=MAD)) +
  geom_point() +
  facet_wrap(~GProtein, nrow=3) +
  scale_color_gradient_colorRamp2(breaks = c(0, 2), colors = c('black', 'red'), name='Median Absolute Deviation') +  
  xlab('Pearson r') +
  ylab('Median Log2FC (all receptors)') +
  geom_vline(xintercept=0, linetype=3, alpha=0.6) +
  geom_hline(yintercept=0, linetype=3, alpha=0.6) +
  theme_bw() + 
  theme(legend.text=element_text(size=8),
        legend.title =element_text(size=8))

g
BackupAsPDF(g, 'medianLog2FC.Rscore.unlabeled.scatterplot.', dimensions=c(8,10)) 


g <- ggplot(merged.dt, aes(x=R, y=Mean, col=sd)) +
  geom_point() +
  facet_wrap(~GProtein, nrow=3) +
  scale_color_gradient_colorRamp2(breaks = c(0, 2), colors = c('black', 'red'), name='Standard Deviation') +  
  xlab('Pearson r') +
  ylab('Mean Log2FC (all receptors)') +
  geom_vline(xintercept=0, linetype=3, alpha=0.6) +
  geom_hline(yintercept=0, linetype=3, alpha=0.6) +
  theme_bw() + 
  theme(legend.text=element_text(size=8),
        legend.title =element_text(size=8))
g
BackupAsPDF(g, 'meanLog2FC.Rscore.unlabeled.scatterplot.', dimensions=c(8,10)) 

```
Just too many things to plot;
iterate through the GProteins and plot and label the extreme values in each

```{r}
# just take summary scores dt
merged.dt <- merge(x=unique(mss.subset[,.(gene,Mean,Median,sd,Variance,MAD)]), y=cor.dt[,.(gene,GProtein,R)], all.x=T, all.y=T, by.x='gene', by.y='gene', allow.cartesian = T)

#subset to top values either side of the 
lapply(seq_along(gprots), function(x,i){
  
  gprot.dt <-  merged.dt[GProtein == x[[i]], ]

  gprot.dt[, sigScore := sqrt(abs(R) * abs(Mean)) ]

  # just label the things in the top 10% 
  gprot.dt[, lab := ifelse(sigScore >= quantile(gprot.dt$sigScore, .9), gene, '')]
  
  g <- ggplot(gprot.dt, aes(x=R, y=Mean, col=Variance), label=lab) +
    geom_point() +
    geom_text_repel(aes(label=lab),segment.linetype=3, max.overlaps = 20,size=1.5) +
    scale_color_gradient_colorRamp2(breaks = c(0, 2,4), colors = c('black', 'darkred', 'red'), name='Variance') +  
    xlab('Pearson r') +
    ylab('Mean Log2FC (all receptors)') +
    geom_vline(xintercept=0, linetype=3, alpha=0.6) +
    geom_hline(yintercept=0, linetype=3, alpha=0.6) +
    ggtitle(paste0(x[[i]], ' r vs log2FC')) +
    theme_bw() + 
    theme(legend.text=element_text(size=8),
        legend.title =element_text(size=8))

 BackupAsPDF(g, 'log2FC.R.scatterplot.', dimensions=c(8,6))
},x=gprots)
```



tried geom_violin, but labelling the individual top hits in each would require splitting the plot
trouble plotting with jitter; maybe if we plot the 3 individually this will be more manageable

plot with jitter to show individual points
```{r}
cor.dt[, lab := ifelse(abs(R) >= 0.6, gene, '')]


g <- ggplot(cor.dt, aes(x=GProtein, y=R, col=GProtein, label=lab)) + 
  geom_violin(position = position_dodge(1), width = 1) +
  geom_boxplot(width=0.1) +
  geom_jitter(data=cor.dt[abs(R) > 0.6,], alpha=0.6, position=position_jitter(seed = 1)) +
  geom_text_repel(data=cor.dt[abs(R) > 0.6,], aes(label=lab), segment.linetype=3, size=3, 
                  position = position_jitter(seed = 1)) +
  #scale_fill_manual(values = c(col.pal)) +
  scale_color_manual(values = c(col.pal)) +
  theme_bw()

g
BackupAsPDF(g, 'r0.6.labelled.violinwJitter.', dimensions=c(10,8))



g <- ggplot(cor.dt, aes(x=GProtein, y=R, col=GProtein, label=lab)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(data=cor.dt[abs(R) > 0.6,], alpha=0.6, position=position_jitter(seed = 1)) +
  geom_text_repel(data=cor.dt[abs(R) > 0.6,], aes(label=lab), segment.linetype=3, size=3, 
                  position = position_jitter(seed = 1)) +
  scale_color_manual(values = c(col.pal)) +
  theme_bw()

g
BackupAsPDF(g, 'r0.6.labelled.boxplotwJitter.', dimensions=c(10,8))
```
now that we have these, take R scores > 0.6 and cluster using their correlation scores and their log2FC values and see how the output looks

Cluster all things with a (abs) correlation score over 0.6

```{r}
genes.oi <- cor.dt[abs(R) > 0.6, unique(gene)]

cor.mat <- dcast(cor.dt, gene~GProtein, value.var = 'R') %>% 
  as.matrix(rownames='gene')

subMat <-  cor.mat[rownames(cor.mat) %in% genes.oi,]


hm <- Heatmap(subMat,
        cluster_columns = F, 
        show_column_names = F,
        name='Pearson r',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 5),
         column_names_gp = gpar (fontsize= 8),
         column_split = colnames(subMat),
         column_title_gp = gpar (fontsize= 9))

hm <- draw(hm, column_title='abs Pearson r > 0.6')

BackupAsPDF(hm, 'R.0.6.gprotein.heatmap', dimensions = c(6,8))
```
plot the same genes, except using log2FC and per receptor

```{r}
# change name for plotting
mss.subset[receptor == 'MOR.DAMGO', receptor := 'MOR']

lfc.mat <- dcast(mss.subset, gene~receptor, value.var = 'log2FC') %>% 
  as.matrix(rownames='gene')

# extract the sig genes and reorder columns to match the scoring mat
subMat <-  lfc.mat[rownames(lfc.mat) %in% genes.oi, colnames(scores.mat$iupar)]


ha <- HeatmapAnnotation(Gi.Go  = as.logical(scores.mat$iupar[1,]),
                        Gq.G11 = as.logical(scores.mat$iupar[2,]),
                        Gs     = as.logical(scores.mat$iupar[3,]),
                        
                        col=list(Gi.Go=c('TRUE'='green', 'FALSE'='lightgrey'),
                                 Gq.G11=c('TRUE'='red', 'FALSE'='lightgrey'),
                                 Gs=c('TRUE'='blue', 'FALSE'='lightgrey')),
                        
                        show_legend = c(F,F,F))
hm <- Heatmap(subMat,
              top_annotation = ha,
        cluster_columns = T, 
        show_column_names = T,
        name='log2FC',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 4),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))
hm <- draw(hm, column_title='abs Pearson r > 0.6')

BackupAsPDF(hm, 'R.0.6.receptor.log2FC.heatmap', dimensions = c(7,11))
```
dominated by higher cor scores in GqG11 (looking at distributions, more likely to achieve high cor score?)
instead, lets take the 5 and 95% quantiles and plot these

```{r}
genes.oi <- sapply(gprots, function(x){
  sub.dt <- cor.dt[GProtein == x,]
  return(sub.dt[R >= quantile(R,0.95) | R <= quantile(R,0.05), unique(gene)])
}) %>% c() %>% unique()

# each GProtein cor has returned many unique values. Perhaps a good sign!
genes.oi %>%  length()

```
cor score plot of the quantiles
```{r}
subMat <-  cor.mat[rownames(cor.mat) %in% genes.oi,]

hm <- Heatmap(subMat,
        cluster_columns = F, 
        show_column_names = F,
        name='Pearson r',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 3),
         column_names_gp = gpar (fontsize= 8),
         column_split = colnames(subMat),
         column_title_gp = gpar (fontsize= 9))

hm <- draw(hm, column_title='5% and 95% quantiles')

BackupAsPDF(hm, 'R.05.95.quantiles.gprotein.heatmap', dimensions = c(6,11))
```
```{r}
# extract the sig genes and reorder columns to match the scoring mat
subMat <-  lfc.mat[rownames(lfc.mat) %in% genes.oi, colnames(scores.mat$iupar)]

hm <- Heatmap(subMat,
              top_annotation = ha,
        cluster_columns = T, 
        show_column_names = T,
        name='log2FC',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 3),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))
hm <- draw(hm, column_title='5% and 95% quantiles')

BackupAsPDF(hm, 'R.05.95.quantiles.log2FC.heatmap', dimensions = c(7,12.5))
```
I think we might be overplotting; take the 2 and 98% quantiles
150 proteins in total
```{r}
genes.oi <- sapply(gprots, function(x){
  sub.dt <- cor.dt[GProtein == x,]
  return(sub.dt[R >= quantile(R,0.98) | R <= quantile(R,0.02), unique(gene)])
}) %>% c() %>% unique()

# each GProtein cor has returned many unique values. Perhaps a good sign!
genes.oi %>%  length()
```

after plotting both these views, we can look at the distribution of log2FC vs cor scores and also the 

```{r}
subMat <-  cor.mat[rownames(cor.mat) %in% genes.oi,]

hm <- Heatmap(subMat,
        cluster_columns = F, 
        show_column_names = F,
        name='Pearson r',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 4.5),
         column_names_gp = gpar (fontsize= 8),
         column_split = colnames(subMat),
         column_title_gp = gpar (fontsize= 9))

hm <- draw(hm, column_title='2% and 98% quantiles')


BackupAsPDF(hm, 'R.02.98.quantiles.gprotein.heatmap', dimensions = c(6,8))
```
and log2FC scores
```{r}
# extract the sig genes and reorder columns to match the scoring mat
subMat <-  lfc.mat[rownames(lfc.mat) %in% genes.oi, colnames(scores.mat$iupar)]

hm <- Heatmap(subMat,
              top_annotation = ha,
        cluster_columns = T, 
        show_column_names = T,
        name='log2FC',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 5),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))
hm <- draw(hm, column_title='2% and 98% quantiles')

BackupAsPDF(hm, 'R.02.98.quantiles.log2FC.heatmap', dimensions = c(8,11))
```
plot the 5 and 10% quantile genes for each dataset 
Reduce overplotting and help focus on specific results

```{r}
genes.oi <- sapply(gprots, function(x){
  sub.dt <- cor.dt[GProtein == x,]
  return(sub.dt[R >= quantile(R,0.95) | R <= quantile(R,0.05), unique(gene)])
}) %>% c() %>% unique()


g <- ggplot(merged.dt[gene %in% genes.oi,], aes(x=R, y=Mean, col=sd)) +
  geom_point() +
  facet_wrap(~GProtein, nrow=3) +
  scale_color_gradient_colorRamp2(breaks = c(0, 2), colors = c('black', 'red'), name='Standard Deviation') +  
  xlab('Pearson r') +
  ylab('Mean Log2FC (all receptors)') +
  geom_vline(xintercept=0, linetype=3, alpha=0.6) +
  geom_hline(yintercept=0, linetype=3, alpha=0.6) +
  theme_bw() + 
  theme(legend.text=element_text(size=8),
        legend.title =element_text(size=8))
g
BackupAsPDF(g, 'meanLog2FC.Rscore.unlabeled.quantile.05.95.scatterplot.', dimensions=c(8,10)) 
```
top quantile (5% and 95%) quantile genes plotted individually per GProtein and label
maybe more helpful to overlay these top scoring correlated proteins on the plot and col red? Hopefully then people can see if the scores match
Plot by receptor and color? Instead of taking average over alll log2FC values, we only want to toake average over 
```{r}
#subset to top values either side of the 
lapply(seq_along(gprots), function(x,i){
  
  gprot.dt <-  merged.dt[GProtein == x[[i]], ]
  genes.oi <- gprot.dt[R >= quantile(R,0.95) | R <= quantile(R,0.05), unique(gene)]
  gprot.dt <- gprot.dt[gene %in% genes.oi,]

  gprot.dt[, sigScore := sqrt(abs(R) * abs(Mean)) ]

  # just label the things in the top 10% 
 # gprot.dt[, lab := ifelse(sigScore >= quantile(gprot.dt$sigScore, .7), gene, '')]
  
  g <- ggplot(gprot.dt, aes(x=R, y=Mean, col=Variance), label=gene) +
    geom_point() +
    geom_text_repel(aes(label=gene),segment.linetype=3, max.overlaps = 20,size=1.5) +
    scale_color_gradient_colorRamp2(breaks = c(0, 1.5,3), colors = c('black', 'darkred', 'red'), name='Variance') +  
    xlab('Pearson r') +
    ylab('Mean Log2FC (all receptors)') +
    geom_vline(xintercept=0, linetype=3, alpha=0.6) +
    geom_hline(yintercept=0, linetype=3, alpha=0.6) +
    ggtitle(paste0(x[[i]], ' r vs log2FC 5% and 95% quantiles')) +
    theme_bw() + 
    theme(legend.text=element_text(size=8),
        legend.title =element_text(size=8))

 BackupAsPDF(g, 'log2FC.R.quantile.05.95.scatterplot.', dimensions=c(8,6))
},x=gprots)

```
28-11-23
----
Generate plots with only the log2FC values used per GProtein group

```{r}
# just take summary scores dt
merged.dt <- merge(x=unique(mss.subset[,.(gene,receptor,log2FC,pvalue)]), y=cor.dt[,.(gene,GProtein,R)], all.x=T, all.y=T, by.x='gene', by.y='gene', allow.cartesian = T)

#subset to top values either side of the 
lapply(seq_along(gprots), function(x,i){
  
  # extract subDT with relevant genes
  gprot.dt <-  merged.dt[GProtein == x[[i]], ]
  genes.oi <- gprot.dt[R >= quantile(R,0.95) | R <= quantile(R,0.05), unique(gene)]
  gprot.dt <- gprot.dt[gene %in% genes.oi,]
  
  
  # calculate summary stats per gprotein group
  if (x[[i]] == 'Gs'){
   sub.dt <- gprot.dt[receptor %in% c('PTGER4', 'NK1R'),] %>% 
     .[, `:=`(Mean=mean(log2FC), Median=median(log2FC), sd=sd(log2FC), Variance=var(log2FC), MAD=mad(log2FC)), by=gene]

  } else if (x[[i]] == 'Gq/G11'){
   sub.dt <-  gprot.dt[receptor %in% c('MRX2', 'MRX4', 'PTGER1', 'QRFPR','NK1R'),] %>% 
     .[, `:=`(Mean=mean(log2FC), Median=median(log2FC), sd=sd(log2FC), Variance=var(log2FC), MAD=mad(log2FC)), by=gene]

  } else if (x[[i]] == 'Gi/Go'){
   sub.dt <-  gprot.dt[receptor %in% c('MRX2', 'NPFFR1', 'NPY1R', 'DOR','KOR', 'MOR', 'QRFPR'),] %>% 
     .[, `:=`(Mean=mean(log2FC), Median=median(log2FC), sd=sd(log2FC), Variance=var(log2FC), MAD=mad(log2FC)), by=gene]
  }
  
  sub.dt[, sigScore := sqrt(abs(R) * abs(Mean)) ]
  # extract cols we want and remove the duplicate rows
  sub.dt <- sub.dt[, .(gene, R, Mean, Median, sd, Variance, MAD)] %>% unique()

  # just label the things in the top 10% 
 # gprot.dt[, lab := ifelse(sigScore >= quantile(gprot.dt$sigScore, .7), gene, '')]
  
  g <- ggplot(sub.dt, aes(x=R, y=Mean, col=Variance), label=gene) +
       geom_point() +
       geom_text_repel(aes(label=gene),segment.linetype=3, max.overlaps = 20,size=1.5) +
       scale_color_gradient_colorRamp2(breaks = c(0, 1.5,3), colors = c('black', 'darkred', 'red'), name='Variance') +  
       xlab('Pearson r') +
       ylab('Mean Log2FC (IUPAR group receptors)') +
       geom_vline(xintercept=0, linetype=3, alpha=0.6) +
       geom_hline(yintercept=0, linetype=3, alpha=0.6) +
       ggtitle(paste0(x[[i]], ' r vs log2FC 5% and 95% quantiles')) +
       theme_bw() + 
       theme(legend.text=element_text(size=8),
        legend.title =element_text(size=8))

 BackupAsPDF(g, 'log2FCGProtSubset.R.quantile.05.95.scatterplot.', dimensions=c(8,6))
},x=gprots)
```
Generate the same plot with the full dataset

```{r}
#subset to top values either side of the 
lapply(seq_along(gprots), function(x,i){
  
  gprot.dt <-  merged.dt[GProtein == x[[i]], ]
  
  # calculate summary stats per gprotein group
  if (x[[i]] == 'Gs'){
   sub.dt <- gprot.dt[receptor %in% c('PTGER4', 'NK1R'),] %>% 
     .[, `:=`(Mean=mean(log2FC), Median=median(log2FC), sd=sd(log2FC), Variance=var(log2FC), MAD=mad(log2FC)), by=gene]

  } else if (x[[i]] == 'Gq/G11'){
   sub.dt <-  gprot.dt[receptor %in% c('MRX2', 'MRX4', 'PTGER1', 'QRFPR','NK1R'),] %>% 
     .[, `:=`(Mean=mean(log2FC), Median=median(log2FC), sd=sd(log2FC), Variance=var(log2FC), MAD=mad(log2FC)), by=gene]

  } else if (x[[i]] == 'Gi/Go'){
   sub.dt <-  gprot.dt[receptor %in% c('MRX2', 'NPFFR1', 'NPY1R', 'DOR','KOR', 'MOR', 'QRFPR'),] %>% 
     .[, `:=`(Mean=mean(log2FC), Median=median(log2FC), sd=sd(log2FC), Variance=var(log2FC), MAD=mad(log2FC)), by=gene]
  }
  
  sub.dt[, sigScore := sqrt(abs(R) * abs(Mean)) ]
  # extract cols we want and remove the duplicate rows
  sub.dt <- sub.dt[, .(gene, R, Mean, Median, sd, Variance, MAD,sigScore)] %>% unique()
  
  
  # just label the things in the top 10% 
  sub.dt[, lab := ifelse(sigScore >= quantile(sub.dt$sigScore, .9), gene, '')]
  
  g <- ggplot(sub.dt, aes(x=R, y=Mean, col=Variance), label=lab) +
    geom_point() +
    geom_text_repel(aes(label=lab),segment.linetype=3, max.overlaps = 20,size=1.5) +
    scale_color_gradient_colorRamp2(breaks = c(0, 2,4), colors = c('black', 'darkred', 'red'), name='Variance') +  
    xlab('Pearson r') +
    ylab('Mean Log2FC (IUPAR group receptors)') +
    geom_vline(xintercept=0, linetype=3, alpha=0.6) +
    geom_hline(yintercept=0, linetype=3, alpha=0.6) +
    ggtitle(paste0(x[[i]], ' r vs log2FC')) +
    theme_bw() + 
    theme(legend.text=element_text(size=8),
        legend.title =element_text(size=8))

 BackupAsPDF(g, 'log2FC.GProtSubset.R.scatterplot.', dimensions=c(8,6))
},x=gprots)
```



write out these results for the scores to the WGCNA directory

```{r}
#lapply(seq_along(cor.long), function(x,n,i){
#  fwrite(x[[i]][numObservations == 11], paste0('output/', n[[i]], '.proteinClassificationCorrelations.csv'))
#},x=cor.long, n=names(cor.long))
```

read in this data and plot the correlation ranks, then a heatmap of the top scores per gene




```{r}
cor.long$iupar$GProtein

lapply(cor.long$iupar$GProtein, function(x){
  
  ggplot(cor.long$iupar[ GProtein == x,], aes(x=reorder(gene,-R), y=R)) + geom_point()
  
})


```


