---
title: "AFscores_plots"
author: "Martin Gordon"
date: "2024-04-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

packages
```{r}
library(data.table)
library(ggplot2)
library(stringr)
library(magrittr)
library(ComplexHeatmap)

source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source("../../utils/bp_utils/manageScriptData.R")
```



read in the data

```{r}
scores.dt <- fread('./data/allScores.csv')
setnames(scores.dt, new=c('file', 'ptm', 'iptm'))

scores.dt[, pair := tstrsplit(file, '[/]', keep=1)]
scores.dt[, model :=  sub("model_","", str_extract(file, 'model_[1-5]'))]
scores.dt[, c('GPCR', 'GProtein') :=  tstrsplit(pair, '__', keep=c(1,2))]
```
All runs produced 5 models
```{r}
scores.dt[, .N, by=pair][N < 5]
```
get gene names for gprotein and gpcrs

```{r}
scores.dt[, GPCR_gene := multiUniprots2multiGenes(GPCR, species='HUMAN')]
scores.dt[, GProtein_gene := multiUniprots2multiGenes(GProtein, species='HUMAN')]
scores.dt[, gene_pair := paste0(GPCR_gene, "__", GProtein_gene)]

scores.dt[, mean.iptm := mean(iptm), by=pair]
scores.dt[, above.threshold := ifelse(iptm > .5, 'yes', 'no')]
```
boxplots of the scores

```{r}
g <- ggplot(scores.dt, aes(x=gene_pair, y=iptm, color=above.threshold)) +
  #geom_boxplot(alpha=0.4, color='darkgrey') +
  geom_point(alpha=0.8) +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  scale_color_manual(values=c('no'='black', 'yes'='red')) +
  labs(title='AF-multimer GPCR-GProtein pair scores') +
  ylim(c(0,1)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, size=5))

g
BackupAsPDF(g, 'AFscores.boxplot', dimensions=c(14,10))
```
Order by mean IPTM
```{r}
g <- ggplot(scores.dt, aes(x=reorder(gene_pair, -mean.iptm), y=iptm, color=above.threshold)) +
  #geom_boxplot(alpha=0.4, color='darkgrey') +
  geom_point(alpha=0.8) +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  scale_color_manual(values=c('no'='black', 'yes'='red')) +
  labs(title='AF-multimer GPCR-GProtein pair scores') +
  ylim(c(0,1)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, size=5))

g
BackupAsPDF(g, 'AFscores.boxplot', dimensions=c(14,10))
```
All the scores seem pretty high...
facet by GPCR and GProtein

```{r}
g <- ggplot(scores.dt, aes(x=reorder(gene_pair, -mean.iptm), y=iptm, color=above.threshold)) +
  #geom_boxplot(alpha=0.4, color='darkgrey') +
  geom_point(alpha=0.8) +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  scale_color_manual(values=c('no'='black', 'yes'='red')) +
  labs(title='AF-multimer GPCR-GProtein pair scores') +
  ylim(c(0,1)) +
  facet_wrap(~GProtein_gene,scales = 'free_x') +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, size=5))

g
BackupAsPDF(g, 'AFscores.GProteinfacet', dimensions=c(14,14))


g <- ggplot(scores.dt, aes(x=gene_pair, y=iptm, color=above.threshold)) +
  #geom_boxplot(alpha=0.4, color='darkgrey') +
  geom_point(alpha=0.8) +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  scale_color_manual(values=c('no'='black', 'yes'='red')) +
  labs(title='AF-multimer GPCR-GProtein pair scores') +
  ylim(c(0,1)) +
  facet_wrap(~GPCR_gene,scales = 'free_x') +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, size=5))

g
BackupAsPDF(g, 'AFscores.GPCRfacet', dimensions=c(14,14))
```
Barplot of mean IPTM scores

```{r}
dt <- scores.dt[,.(above.threshold = ifelse(mean.iptm > 0.5, 'yes', 'no'), gene_pair, mean.iptm,GProtein_gene, GPCR_gene)] %>% 
  unique()


g <- ggplot(dt, aes(x=reorder(gene_pair,-mean.iptm), y=mean.iptm, fill=above.threshold)) +
  geom_bar(alpha=0.8, stat='Identity') +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  scale_fill_manual(values=c('no'='black', 'yes'='red')) +
  labs(title='AF-multimer GPCR-GProtein pair scores') +
  ylim(c(0,1)) +
  facet_wrap(~GPCR_gene,scales = 'free_x') +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, size=5))

g
BackupAsPDF(g, 'AFscores.GPCRfacet.barplot', dimensions=c(14,14))
```
What models contribute to the high scores

```{r}
g <- ggplot(scores.dt, aes(x=gene_pair, y=iptm, shape=model, color=above.threshold)) +
  geom_point(alpha=0.8) +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  labs(title='AF-multimer Sacsin pair scores') +
  ylim(c(0,1)) +
  scale_color_manual(values=c('no'='black', 'yes'='red')) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90))

g
BackupAsPDF(g, 'AFscores.boxplot', dimensions=c(14,10))
```

Looks like there might be a GPCR-GProtein bias
First lets visualise these scores as a heatmap
Input all the results, and cluster; then take the averages of the models

```{r}
# just get the mean iptm from each of the groups
scores.mat <- dcast(scores.dt, GProtein_gene~GPCR_gene, value.var = 'iptm', fun.aggregate=mean) %>%  
  as.matrix(rownames='GProtein_gene')

colnames(scores.mat)

# create annotation
ann <- data.table('Gi/Go' = c('Yes','No', 'Yes', 'Yes', 'Yes','Yes', 'Yes', 'No', 'No', 'Yes', 'No'),
                  'Gq/G11' = c('Yes', 'Yes', 'No', 'No', 'No', 'No', 'No', 'Yes', 'No', 'Yes', 'Yes'),
                  'Gs' = c('No', 'No', 'No', 'No', 'No', 'No', 'No', 'No', 'Yes', 'No', 'Yes'))

colours <- list('Gi/Go'= c('Yes' = '#4477AA', 'No'="grey"),
                'Gq/G11'= c('Yes' = '#CCBB44','No'="grey"),
                'Gs'= c('Yes' = 'darkred', 'No'="grey")
                )

colAnn <- HeatmapAnnotation(df = ann, col = colours)

# draw heatmap
hm <- Heatmap(scores.mat, 
              cluster_rows=T,
              show_row_names = T,
              cluster_columns = T, 
              top_annotation = colAnn, 
              name='Mean iPTM')


BackupAsPDF(hm, 'mean.iPTM.heatmap')
```
Make a heatmap of 'all the models' maybe interesting to see if models with templates tend to have better scores

```{r}
# just get the mean iptm from each of the groups
scores.mat <- dcast(unique(scores.dt), GProtein_gene~paste0(GPCR_gene,'.model', model), value.var = 'iptm') %>%  
  as.matrix(rownames='GProtein_gene')

# create annotation
ann.new <- data.table('model' = gsub('*[.]model', '', colnames(scores.mat)),
                  'receptor'= gsub('[.]model[1-5]$', '', colnames(scores.mat)))

ann.new[, `Gi/Go` := ifelse(receptor %in% c('QRFPR','OPRM1', 'ORPD1', 'OPRK1', 'NPFFR1','NPY1R', 'MRGPRX2'), 'Yes', "No")]
ann.new[, `Gq/G11` := ifelse(receptor %in% c('TACR1','QRFPR', 'PTGER1','MRGPRX2', 'MRGPRX4'), 'Yes', "No")]
ann.new[, `Gs` := ifelse(receptor %in% c('PTGER4','TACR1'), 'Yes', "No")]

ann.new$receptor %>%  unique()
# draw heatmap
hm <- Heatmap(scores.mat, 
              cluster_rows=T,
              show_row_names = T,
              cluster_columns = T, 
              column_names_gp = gpar(fontsize=6),
              row_names_gp = gpar(fontsize=6),
              name='iPTM')

BackupAsPDF(hm, 'ipTM.heatmap')
```
```{r}
scores.dt$GProtein_gene %>%  unique()
scores.dt$GPCR_gene %>%  unique()
```


Take the inital scores, matrix and do a TMP and look at the residuals
Color by predicted to interact score
```{r}
scores.dt[, knownInteraction := ifelse( GProtein_gene %in%  c('GNAQ', 'GNA11') & GPCR_gene %in% c('TACR1', 'QRFPR', 'PTGER1', 'MRGPRX2', 'MRGPRX4'), 'Gq.G11',
                                        ifelse(GProtein_gene %in%  c("GNAI1", "GNAI2", "GNAI3","GNAO1") & GPCR_gene %in% c('QRFPR', "OPRM1","OPRD1","OPRK1", 'NPY1R', 'MRGPRX2','NPFFR1'), 'Gi.Go',
                                               ifelse(GProtein_gene %in%  c("GNAS") & GPCR_gene %in% c('PTGER4', 'TACR1'), 'Gs','unknown')) )]


g <- ggplot(scores.dt, aes(x=reorder(gene_pair, -mean.iptm), y=iptm, color=knownInteraction)) +
  #geom_boxplot(alpha=0.4, color='darkgrey') +
  geom_point(alpha=0.8) +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  scale_color_manual(values=c('Gq.G11'='red', 'Gi.Go'='blue', 'Gs'='green', 'unknown'='grey')) +
  labs(title='AF-multimer GPCR-GProtein pair scores') +
  ylim(c(0,1)) +
  facet_wrap(~GProtein_gene,scales = 'free_x') +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, size=5))

g

BackupAsPDF(g,'AFscores.GProteinfacet.colInteractions', dimensions=c(14,14))
```

Need to attempt TMP on this dataset; subtract column and row effects get residuals and see how the interactions look

```{r}
scores.mat <- dcast(scores.dt, GProtein_gene~GPCR_gene, value.var = 'iptm', fun.aggregate=mean) %>%  
  as.matrix(rownames='GProtein_gene')

tmp.out <- medpolish(scores.mat)

tmp.out$residuals %>% str()

        # create annotation
ann <- data.table('Gi/Go' = c('Yes','No', 'Yes', 'Yes', 'Yes','Yes', 'Yes', 'No', 'No', 'Yes', 'No'),
                  'Gq/G11' = c('Yes', 'Yes', 'No', 'No', 'No', 'No', 'No', 'Yes', 'No', 'Yes', 'Yes'),
                  'Gs' = c('No', 'No', 'No', 'No', 'No', 'No', 'No', 'No', 'Yes', 'No', 'Yes'))

colours <- list('Gi/Go'= c('Yes' = '#4477AA', 'No'="grey"),
                'Gq/G11'= c('Yes' = '#CCBB44','No'="grey"),
                'Gs'= c('Yes' = 'darkred', 'No'="grey")
                )

colAnn <- HeatmapAnnotation(df = ann, col = colours)

# draw heatmap
hm <- Heatmap(tmp.out$residuals, 
              cluster_rows=T,
              show_row_names = T,
              cluster_columns = T, 
              top_annotation = colAnn, 
              name='Mean iPTM')
hm
```
plot the avg iPTM from the GPRotein and GPCR respectively

```{r}
g <- ggplot(scores.dt, aes(x=GPCR_gene, y=iptm, fill=GPCR_gene)) +
  geom_boxplot() +
  ggtitle('Distributions of iPTM scores') +
  theme_classic() +
  theme(axis.text.x = element_text(angle=90))
g
BackupAsPDF(g, 'GPCRDistributions.boxplots.')

g <- ggplot(scores.dt, aes(x=GProtein_gene, y=iptm, fill=GProtein_gene)) +
  geom_boxplot() +
  ggtitle('Distributions of iPTM scores') +
  theme_classic() +
  theme(axis.text.x = element_text(angle=90))
  
BackupAsPDF(g, 'GProteinDistributions.boxplots.')
```

Can we just subtract the row effect?
```{r}
# remove row effects
submat <- sweep(scores.mat, 1, FUN='-', tmp.out$row)
# remove col effects
#submat <- sweep(submat, 2, FUN='-', tmp.out$col)

# draw heatmap
hm <- Heatmap(submat, 
              cluster_rows=T,
            ##  clustering_distance_rows = "pearson",
          ##    clustering_distance_columns = 'pearson',
              show_row_names = T,
              cluster_columns = T, 
              top_annotation = colAnn, 
              name='Mean iPTM')
hm
BackupAsPDF(hm, 'GProtein.tmp.rmRowEffects.heatmap')
```





Problem with tmp method for this is we are assuming the effect is additive. Not the case here as these values are constrained to 1.
What we want to do is convert each cell to rank
We want weigthing by both rows and columns (think of if one row doesnt contribute much but column does, then that row will be driven by the column effect)

```{r}

```


```{r}
help("medpolish")
```

