---
title: "DARPA35.1"
author: "Martin Gordon"
date: "2023-08-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DARPA 35.1 

Time series analysis MOR receptor
(T0,T1,T15,T30)
Quick QC of the data, run the timeseries analysis, return heatmaps and enrichment etc (enrichment TODO - does it make sense to do enrichment with a timeseries analysis?)
Compare with DAMGO (DAR30)? Speak to Ben; maybe just an UPset or venn diagram of sig hits
Check with only 4 timepoints for significant changes? (ask ben)
Touch base with Ben Monday; for now just complete ananlysis


```{r packages}
library(data.table)
library(MSstats)
library(ggplot2)
library(magrittr)
library(circlize)
library(ComplexHeatmap)
library(parallelDist) #needed for some custom enrichment functions
library(ggvenn)
library(randomcoloR)
library(RColorBrewer)
library(stringr)

#source Bens utility scripts
source ("/Users/martingordon/Documents/utils/bp_utils/MS_ModelFitting_TimeSeries.R")
source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/MSstats_Helper_Functions.R")
source("/Users/martingordon/Documents/utils/bp_utils/MSstats_V4_Functions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")

rotate.x.axis.text <- theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

```



# Loaad spec

```{r}
#spec <- fread ("/Users/ben/Box/DAR031/20230322_DAR31_MSstats_Report.xls")
spec <- fread ("./data/20230802_212344_DAR35.1_EXc_dirDIA_AllRuns_MSStatsFormatReport.tsv")
```

```{r}
head(spec)
mean (log2(spec$Intensity))
sd(log2(spec$Intensity))

unique(spec$Condition)
```

Again a left hand side shoulder, seems there is imputation for this data

```{r}
hist(log2(spec$Intensity), breaks = 100)


g <- ggplot(spec, aes(x=log2(Intensity), fill=Condition)) + 
  geom_histogram(bins=100) + 
  facet_wrap(~Condition) 

g

BackupAsPDF(g, '.conditions.histogram.')


```
Still seems to be a linear relationship between data 
Remove noise
```{r}
wide <- dcast(spec, ProteinName+PeptideSequence+PrecursorCharge~Condition+BioReplicate, value.var = "Intensity")

ggplot (wide, aes (x =log2(MOR_00_1), y = log2(Z28_01_1))) + geom_point(shape = ".") + geom_density_2d() + coord_fixed()
```

```{r}
hist(log2(spec$Intensity), breaks = 100)
spec <- spec[Intensity > 2^4,]
```
```{r}
g <- ggplot(spec, aes(x=log2(Intensity), fill=Condition)) + 
  geom_histogram(bins=100) + 
  facet_wrap(~Condition) 

g

BackupAsPDF(g, '.conditions.clean.histogram.')
```
What are these shoulders? Seems imputation is turned on
Not a lot of proteins recovered it seems..

```{r}
wide <- dcast(spec, ProteinName+PeptideSequence+PrecursorCharge~Condition+BioReplicate, value.var = "Intensity")

ggplot (wide, aes (x =log2(MOR_00_1), y = log2(Z28_01_2))) + geom_point(shape = ".") + geom_density_2d() 

```

Counts per condition
need to drop Z28_01.1 & Z28_05.3

```{r, fig.width = 10, fig.height = 5}
numColors <- length(unique(spec$Condition))
rColors <- randomcoloR::distinctColorPalette(numColors) 
p <- ggplot (spec, aes (x = interaction(BioReplicate, Condition), fill = Condition)) + geom_bar()  + scale_fill_manual(values = rColors) + theme_bw() + rotate.x.axis.text
BackupAsPDF(p, "SpectronautRowCounts.")
```

```{r, fig.width = 10, fig.height = 4}
p <- ggplot (spec[, .(numFeatures = length(unique(paste0(PeptideSequence, PrecursorCharge)))), by =.(BioReplicate, Condition)], 
        aes (x = interaction(BioReplicate, Condition), y= numFeatures, fill = Condition)) +
  geom_col() +
  theme_bw() +
  scale_fill_manual(values = rColors)  +
  rotate.x.axis.text
p
BackupAsPDF(p, "FeatureCounts.")

```
Just drop these two samples

```{r}
drop.cols <- c('Z28_01.1', 'Z28_05.3')
spec <- spec[!paste0(Condition,'.',BioReplicate) %chin% drop.cols,]


g <- ggplot(spec, aes(x=paste0(Condition, '.', BioReplicate), y=log2(Intensity), fill=Condition)) +
  geom_boxplot()  +
  scale_fill_manual(values = rColors)  +
  rotate.x.axis.text

BackupAsPDF(g, 'feature.intensity.boxplots.')
```



## PCA, features
```{r}
featureMat <- dcast (spec[!Condition %in% c("Cyto")],
                     paste0(PeptideSequence, PrecursorCharge)~Condition+BioReplicate, value.var = "Intensity") |> as.matrix(rownames = "PeptideSequence")

featureMat <- featureMat[complete.cases(featureMat), ]
```
Festure Intensity boxplots

```{r}

g <- ggplot(spec, aes(x=paste(Condition, BioReplicate), y=log2(Intensity), fill=Condition)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=90))

g

BackupAsPDF(g, '.boxplots.intensity.')
```



```{r}
boxplot(log2(featureMat))
```



```{r}
pcaOut <- prcomp(t(log2(featureMat)))

colInfo <- data.table(colname = colnames(featureMat))
colInfo
colInfo[,c("treat", "time", "rep") := tstrsplit(colname, "[_.]", keep = c(1,2,3)) ] # v useful fucntion

title <- "PCA"

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = time, shape = treat)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,#legend settings
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, "PCA.Complete.Features")

```

Difficult to discern pattern here.. lets color points by timepoint, treatment and bio rep seperately to see if we can identify patterns
No obvious trends with timepoint
Second PC seems to broadly seperate batch (assuming this is bioreplicate)

```{r}

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = time, shape = treat)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
 # scale_color_brewer(type = "seq", palette = "Reds") +
  scale_shape_manual(values = 21:25) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,#legend settings
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, "PCA.complete.Features.time.")
```



# protein summarization:
10/08/23 added cyto back in for merging for location coefficients

```{r}

#mssInput <- specFileToCompleteMSstats(spec[!Condition %in% c("Cyto", "Endo", "GalT", "LAMP1", "LAMTOR1", "PM")])
#mssInput[, IsotopeLabelType := 'L']


mssInput <- specFileToCompleteMSstats(spec)
mssInput[, IsotopeLabelType := 'L']

```

```{r}
dp.out<- MSstats::dataProcess(setDF(mssInput), MBimpute =  FALSE, featureSubset = "highQuality", remove_uninformative_feature_outlier = TRUE)
```

```{r}
fwrite (dp.out$ProteinLevelData, ScriptAndDatedFileName("ProteinLevelData.wCyto.csv"))
fwrite (dp.out$FeatureLevelData, ScriptAndDatedFileName("FeatureLevelData.wCyto.csv.gz"))
```


```{r}
protQuant <- fread ("./DARPA_35.1_data/2023_08_10_ProteinLevelData.wCyto.csv")
```


```{r}
prot.mat <- dcast (protQuant, Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix (rownames = "Protein")
```

PCA at protein level

```{r}
nrow(prot.mat)
prot.mat <- prot.mat[complete.cases(prot.mat),]
nrow(prot.mat)
```

No clear trend on time
```{r}
pcaOut <- prcomp(t(prot.mat))

colInfo <- data.table(colname = colnames(prot.mat))
colInfo[,c("treat", "time", "rep") := tstrsplit(colname, "[_.]", keep = c(1,2,3)) ]

title <- "PCA"

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = time, shape = treat)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )
BackupAsPDF(p, "PCA.complete.proteins.time.")

```
Seems to be some clustering according to biorep... lets look at protein intensity boxplots and clustering to see if renormalisation needed...
```{r}
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = rep, shape = treat)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )
p
BackupAsPDF(p, "PCA.complete.proteins.biorep.")
```

Boxplots. Renormalise the data 
```{r}

ggplot(protQuant, aes(x=paste0(GROUP,'.', SUBJECT), y=LogIntensities, fill=GROUP)) +
  geom_boxplot() + 
  theme(
    axis.text.x = element_text(angle=90)
  )
```
`

# renormalize using TMP
Find the TMP offset  and subtract logIntensities

```{r}
prot.mat <- dcast (protQuant, Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix (rownames = "Protein")

#perform tmp normalisation
tmp <- medpolish(prot.mat, na.rm = TRUE) #iteratively subtract row and col median values 
tmp.dt <- data.table(offset = tmp$col, group_subject = names(tmp$col))

protQuant[,group_subject := paste0(GROUP,'_',SUBJECT)]
protQuant[tmp.dt, normalize.offset := i.offset, on = 'group_subject']
protQuant[, LogIntensities := LogIntensities - normalize.offset]


head(prot.mat)

fwrite(protQuant, ScriptAndDatedFileName("ProteinLevelData.tmpnorm.wCyto.csv"))
```

Boxplots. Renormalise the data 
```{r}

g <- ggplot(protQuant, aes(x=paste0(GROUP,'.', SUBJECT), y=LogIntensities, fill=GROUP)) +
  geom_boxplot() + 
  theme(
    axis.text.x = element_text(angle=90)
  )

BackupAsPDF(g, 'protein.intensity.tmpnorm.boxplots.')
```
Look at PCA again fter renormalisation

```{r}
prot.mat <- dcast (protQuant, Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix (rownames = "Protein")
prot.mat <- prot.mat[complete.cases(prot.mat),]

pcaOut <- prcomp(t(prot.mat))

colInfo <- data.table(colname = colnames(prot.mat))
colInfo[,c("treat", "time", "rep") := tstrsplit(colname, "[_.]", keep = c(1,2,3)) ]

title <- "PCA"

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = time, shape = treat)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, "PCA.complete.proteins.tmpnorm.time")

```
Generate heatmaps
Colored by treatment, timepoint.. more breakdown by timepoint & biorep (batch 3 clusters)

```{r}

clusterwithNA <- function(x, val=0,...){
  x[is.na(x)] <- val
  hclust(dist(x),...)
}


subMat <- prot.mat[sample(rownames(prot.mat), size=3000),]

type <-str_extract(colnames(subMat),"(?<=_)[0-9]{1}$") #pos lookbehind modify as needed 

#ha <- HeatmapAnnotation(grp = type, annotation_name_side = "right", col= list(grp = c('MOR_00'='red', 'Z28_01' = 'blue', 'Z28_05'='green', 'Z28_10'='orange', 'Z28_30'='cyan', 'Z47_01'='purple', 'Z47_05'='yellow',  "Z47_10"='pink', 'Z47_30'='darkgrey')))
#ha <- HeatmapAnnotation(grp = type, annotation_name_side = "right", col= list(grp = c('00'='red', '01' = 'blue', '05'='green', '10'='orange', '30'='cyan')))

ha <- HeatmapAnnotation(grp = type, annotation_name_side = "right", col= list(grp = c('0'='red', '1' = 'blue', '2'='green', '3'='orange')))

  #generate heatmap
hm <- Heatmap(subMat,
        top_annotation = ha,
        cluster_rows = clusterwithNA(subMat),
        show_row_names = FALSE,
        row_title = sprintf("%d subsampled proteins", nrow(subMat)),
        na_col = 'grey',
       # column_title = paste0(' heatmap'),
        name = "Log2 Int.")
hm



BackupAsPDF(hm, '.logints.biorep.heatmap.')
```
median sweep and visualise

```{r}
scaled.mat <- sweep(prot.mat, 1, FUN='-',apply(prot.mat,1, median, na.rm=T))


subMat <- scaled.mat[sample(rownames(scaled.mat), size=3000),]


type <- sub("_[0-9]{1}$", "", colnames(subMat)) #grp
#type <- str_extract(colnames(subMat),"(?<=_)[0-9]{2}") #time
#type <-str_extract(colnames(subMat),"(?<=_)[0-9]{1}$") #rep pos lookbehind modify as needed 
#type <-sub("_[0-9]{2}_[0-9]{1}$", "", colnames(subMat)) #treat 


#ha <- HeatmapAnnotation(grp = type, annotation_name_side = "right", col= list(grp = c('00'='red', '01' = 'blue', '05'='green', '10'='orange', '30'='cyan')))
ha <- HeatmapAnnotation(grp = type, annotation_name_side = "right", col= list(grp = c('MOR_00'='red', 'Z28_01' = 'blue', 'Z28_05'='green', 'Z28_10'='orange', 'Z28_30'='cyan', 'Z47_01'='purple', 'Z47_05'='yellow',  "Z47_10"='pink', 'Z47_30'='darkgrey')))

  #generate heatmap
hm <- Heatmap(subMat,
        top_annotation = ha,
        cluster_rows = clusterwithNA(subMat),
        show_row_names = FALSE,
        row_title = sprintf("%d subsampled proteins", nrow(subMat)),
        na_col = 'grey',
       # column_title = paste0(' heatmap'),
        name = "Log2 Int.")
hm


BackupAsPDF(hm, 'scaled.logints.sample.heatmap.')
```
## what do the known-background protein intensities look like:

This section is only relevant for an APEX dataset.  APEX works by labeling neighboring/interacting proteins with biotin which is then used to purify labeled proteins. There are also proteins that are endogenously biotinylated which will co-purify with the APEX-labeled proteins.  Here we look at a subset of these endogenous biotin proteins, and we inspect their post-normalization background levels, which are inversely related to the labeling-efficiency of APEX.  More background after normalization implies there is less APEX-labeled signal.

```{r, fig.width = 10, fig.height = 4}
biotin.carboxylases.up <- c("O00763","P05165","P11498","Q13085","Q96RQ3")

p <- ggplot(protQuant[Protein %in% biotin.carboxylases.up], aes (x = interaction ( SUBJECT, GROUP ), y = LogIntensities, color = Protein)) + geom_line(aes(group = Protein)) + geom_point(aes(shape = as.factor(SUBJECT))) +
  #geom_text(data = protQuant[Protein %in% c("O00763")], aes(label = paste0(GROUP, ".", SUBJECT), size = paste0(GROUP, "_", SUBJECT) %in% problemRuns)) +
  rotate.x.axis.text

p
BackupAsPDF(p, "BiotinCarboxylaseLevels.")

```
Request genes of interest,
Seems to be some increase in Z47 relative to 

```{r}

protQuant[, gene := multiUniprots2multiGenes(Protein)]

protQuant[, c("drug", "time") := tstrsplit(GROUP, "_")]

p <- ggplot(protQuant[gene %in% c("ARRB2", "EEA1", "EYA4", "SNX3", "FBXW11;BTRC", 'WASHC4') & drug != "MOR"], 
            aes (x = time, y = LogIntensities, color = as.factor(SUBJECT))) + 
  geom_line(aes(group = as.factor(SUBJECT))) + 
  geom_point(aes(shape = as.factor(SUBJECT))) +
  rotate.x.axis.text +
  facet_grid(gene~drug, scales= "free_y") +                                                              # Change font size
  theme(strip.text.x = element_text(size = 8),
        strip.text.y = element_text(size = 5))

BackupAsPDF(p, 'genesOI.linegraph')
```
Plotting MORIPs 
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3695902/
CSN5, AUP1, (SIAH1 and SIAH2),DOK4,DOK5, RanBP9, Kv5.1, VAPA or VAP33, WLS, pACT
Share when we get a better list of things

```{r}
mor.ip <- c('CSN5', 'AUP1', 'SIAH1', 'SIAH2','DOK4','DOK5', 'RanBP9', 'Kv5.1', 'VAPA','VAP33', 'WLS', 'pACT')

p <- ggplot(protQuant[gene %in% mor.ip & drug != "MOR"], 
            aes (x = time, y = LogIntensities, color = as.factor(SUBJECT))) + 
  geom_line(aes(group = as.factor(SUBJECT))) + 
  geom_point(aes(shape = as.factor(SUBJECT))) +
  rotate.x.axis.text +
  facet_grid(gene~drug, scales= "free_y") +                                                              # Change font size
  theme(strip.text.x = element_text(size = 8),
        strip.text.y = element_text(size = 5))

p
#BackupAsPDF(p, 'MORIP.linegraph')
```
Create a new T0 condition for both agonists using 

```{r}
protQuant <- fread('DARPA_35.1_data/2023_08_04_ProteinLevelData.tmpnorm.csv')

# rename and add back to prot.Quant
# duplicating; check that ther eis not an issue with this with the algorithm
Z28_0 <-  protQuant[GROUP == 'MOR_00', ] %>% .[,GROUP := "Z28_00"]
Z47_0 <-  protQuant[GROUP == 'MOR_00',] %>%  .[,GROUP := "Z47_00"]

p.quant.merge <- Reduce(rbind,(list(protQuant, Z28_0,Z47_0))) %>% 
  .[GROUP != 'MOR_00',] %>% 
  .[order(GROUP)]

p.quant.merge[, unique(GROUP) ]
```

# fit cubic time courses

```{r}

p.quant.merge[, c("drug", "timeStr") := tstrsplit(GROUP, "_")]
p.quant.merge[, SUBJECT := sprintf ("batch.%d", SUBJECT)]
p.quant.merge[, rankTime := as.integer(as.factor(timeStr))] #timeranked to 0-5
p.quant.merge[, .(timeStr, rankTime)] %>% unique() #timepoint to time rank 

allDrugs <- sort(unique(p.quant.merge$drug))
names(allDrugs) <- allDrugs

allDrugs

#run timeseires analysis for each drug

out.list <- lapply(allDrugs, function(d) nicePolyFits.fullTable(p.quant.merge[drug == d], splitColumn = "Protein", polyColumn = "rankTime", yColumn = "LogIntensities", otherTerms = "SUBJECT"))
```


```{r}
allFits <- rbindlist(out.list, idcol = "drug")

allFits[, adj.pvalue := p.adjust(pvalue, method = "BH"), by = drug]
allFits[, gene := multiUniprots2multiGenes(Protein)]

```

Volcano plots

```{r}

allFits[, sig := "not"]
allFits[abs(log2FC) > log2(1.5) & adj.pvalue < 0.05, sig := ifelse(log2FC > 0, "up", "down")]

g <- ggplot (allFits, aes(x = log2FC, y = -log10(adj.pvalue), color = sig)) + scale_color_manual(values = c(not = "gray", up = "red", blue = "down")) + geom_point() + facet_wrap(~drug) + coord_cartesian(c(-10,10)) +
  ggrepel::geom_text_repel(aes(label = gene))

BackupAsPDF(g, 'volcanoplots.t0-t30')

range (allFits$log2FC, na.rm = TRUE)
```

```{r}
fwrite(allFits, ScriptAndDatedFileName("allFits.csv"))
```

Not significant...
```{r}
allFits[gene == "ARRB2"]
```



```{r, fig.height = 5, fig.width = 6}
allFits <- fread('./DARPA_35.1_data/2023_08_04_allFits.csv')

allFits[, sig := "not"]
allFits[abs(log2FC) > log2(1.5) & pvalue < 0.01, sig := ifelse(log2FC > 0, "up", "down")]
allFits[, gene := multiUniprots2multiGenes(Protein)]

p <- ggplot (allFits, aes(x = log2FC, y = -log10(pvalue), color = sig)) + scale_color_manual(values = c(not = "gray", up = "red", down = "blue")) + geom_point(show.legend = FALSE) + facet_wrap(~drug, ncol = 1) + coord_cartesian(xlim = c(-10, 10)) +
  ggrepel::geom_text_repel(data = allFits[sig %in% c("up", "down")], aes(label = gene), show.legend = FALSE, size = 2, max.overlaps = 20) +
  theme_bw()
p

BackupAsPDF(p, 'volcanoplots.t0-t30.pval<0.01')


```


```{r, fig.width = 6, fig.height = 10}

sigProteins <- allFits[sig != "not", Protein]
protQuant[, gene := multiUniprots2multiGenes(Protein)]

prot.mat <- dcast (protQuant[Protein %in% sigProteins], gene~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "gene")
hm <- Heatmap (prot.mat, cluster_rows = T, cluster_columns = FALSE, row_names_gp  = gpar(fontsize= 5), column_split = tstrsplit(colnames(prot.mat), "_")[[1]])
hm

BackupAsPDF(hm, 'logIntensitiy.heatmap.t0-t30.pval<0.01', dim=c(7,12))


protQuant
```

```{r, fig.width = 6, fig.height = 6}
numMissing <- rowSums(is.na(prot.mat))
numMissing
prot.mat <- prot.mat[numMissing < ncol(prot.mat)/2,]

prot.mat <- sweep(prot.mat, 1, apply(prot.mat,1, median, na.rm = TRUE))

hm <- Heatmap (prot.mat, cluster_rows = TRUE, cluster_columns = FALSE, row_names_gp  = gpar(fontsize= 4), column_split = tstrsplit(colnames(prot.mat), "_")[[1]], name = "log2 vs Median")
hm
BackupAsPDF(hm, 'log2vssmedian.heatmap.scaled.t0-t30.pval<0.01', dim=c(7,12))
```


Just generate heatmap of the upregulated genes

```{r}
sigProteins <- allFits[sig == "up", Protein]
protQuant[, gene := multiUniprots2multiGenes(Protein)]

prot.mat <- dcast (protQuant[Protein %in% sigProteins], gene~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "gene")
hm <- Heatmap (prot.mat, cluster_rows = T, cluster_columns = FALSE, row_names_gp  = gpar(fontsize= 5), column_split = tstrsplit(colnames(prot.mat), "_")[[1]])
hm

BackupAsPDF(hm, 'logIntensitiy.upreg.heatmap.t0-t30.pval<0.01', dim=c(7,12))

```
```{r, fig.width = 6, fig.height = 6}
numMissing <- rowSums(is.na(prot.mat))

prot.mat <- prot.mat[numMissing < ncol(prot.mat)/2,]

prot.mat <- sweep(prot.mat, 1, apply(prot.mat,1, median, na.rm = TRUE))

hm <- Heatmap (prot.mat, cluster_rows = TRUE, cluster_columns = FALSE, row_names_gp  = gpar(fontsize= 4), column_split = tstrsplit(colnames(prot.mat), "_")[[1]], name = "log2 vs Median")
hm
BackupAsPDF(hm, 'log2vssmedian.upreg.heatmap.scaled.t0-t30.pval<0.01', dim=c(7,12))
```

Log2FC heatmap of genes

```{r}
sig.genes <- allFits[sig != "not",gene]

ms.mat <- allFits %>% 
  dcast(gene~drug, value.var = 'log2FC') %>% 
  as.matrix(rownames='gene')

ms.mat <- ms.mat[rownames(ms.mat) %in% sig.genes,]

hm <- Heatmap (ms.mat, cluster_rows = cluster.w.NA(ms.mat), cluster_columns = FALSE, row_names_gp  = gpar(fontsize= 4), column_split = tstrsplit(colnames(ms.mat), "_")[[1]],  width = ncol(ms.mat)*unit(20, "mm"), name = "log2FC")
hm
BackupAsPDF(hm, 'log2FC.heatmap.pval<0.01', dim=c(7,12))


```
Remove gene names and collapse replicates

```{r}
sigProteins <- allFits[sig == "up", Protein]
protQuant[, gene := multiUniprots2multiGenes(Protein)]


protQuant[, mean.LogIntensities := mean(LogIntensities, na.rm=T), by=.(Protein,GROUP)]
protQuant[is.na(mean.LogIntensities), mean.LogIntensities := mean.LogIntensities, by=.(Protein,GROUP)]


prot.mat <- dcast (protQuant[Protein %in% sigProteins], gene~GROUP+SUBJECT, value.var = "mean.LogIntensities") |> as.matrix(rownames = "gene")
prot.mat <- prot.mat[complete.cases(prot.mat),]

colnames(prot.mat) <- sub("_[0-3]{1}$", '', colnames(prot.mat))
# subset the matrix to remove dup rows
duplicated.columns <- duplicated(t(prot.mat))
prot.mat <- prot.mat[, !duplicated.columns]


hm <- Heatmap (prot.mat, cluster_rows = cluster.w.NA(prot.mat), cluster_columns = FALSE, width = ncol(prot.mat)*unit(6, "mm"), height = nrow(prot.mat)*unit(1.1, "mm"),row_names_gp  = gpar(fontsize= 4), column_split = tstrsplit(colnames(prot.mat), "_")[[1]], name = "log2 Intensity")
hm
BackupAsPDF(hm, 'heatmap.grp.mean.logintensities.long')


```
Also perform a median sweep plot

```{r}
sigProteins <- allFits[sig == "up", Protein]
protQuant[, gene := multiUniprots2multiGenes(Protein)]


protQuant[, mean.LogIntensities := mean(LogIntensities, na.rm=T), by=.(Protein,GROUP)]
protQuant[is.na(mean.LogIntensities), mean.LogIntensities := mean.LogIntensities, by=.(Protein,GROUP)]


prot.mat <- dcast (protQuant[Protein %in% sigProteins], gene~GROUP+SUBJECT, value.var = "mean.LogIntensities") |> as.matrix(rownames = "gene")
prot.mat <- sweep(prot.mat, 1, apply(prot.mat,1, median, na.rm = TRUE))
prot.mat <- prot.mat[complete.cases(prot.mat),]

colnames(prot.mat) <- sub("_[0-3]{1}$", '', colnames(prot.mat))
# subset the matrix to remove dup rows
duplicated.columns <- duplicated(t(prot.mat))
prot.mat <- prot.mat[, !duplicated.columns]


hm <- Heatmap (prot.mat, cluster_rows = cluster.w.NA(prot.mat), cluster_columns = FALSE, width = ncol(prot.mat)*unit(6, "mm"), height = nrow(prot.mat)*unit(1.1, "mm"), row_names_gp  = gpar(fontsize= 4), column_split = tstrsplit(colnames(prot.mat), "_")[[1]], name = "Ints. vs median")
hm
BackupAsPDF(hm, 'heatmap.grp.mean.median.sweep.logintensities.long')
```



```{r}
source ("../../bp_utils/UniprotIDMapping.R")
protQuant[, gene := multiUniprots2multiGenes(Protein)]

protQuant[, c("drug", "time") := tstrsplit(GROUP, "_")]


p <- ggplot(protQuant[gene %in% c("ARRB2", "KCTD12", "EYA4", "PRKCD") & drug != "Cyto"], aes (x = time, y = LogIntensities, color = as.factor(SUBJECT))) + geom_line(aes(group = as.factor(SUBJECT))) + geom_point(aes(shape = as.factor(SUBJECT))) +
  #geom_text(data = protQuant[Protein %in% c("O00763")], aes(label = paste0(GROUP, ".", SUBJECT), size = paste0(GROUP, "_", SUBJECT) %in% problemRuns)) +
  rotate.x.axis.text +
  facet_grid(gene~drug, scales= "free_y")

p

```

Reduce size of the 
```{r, fig.width = 5, fig.height = 6}
source ("../../bp_utils/UniprotIDMapping.R")
protQuant[, gene := multiUniprots2multiGenes(Protein)]

protQuant[, c("drug", "time") := tstrsplit(GROUP, "_")]

p <- ggplot(protQuant[gene %in% c("SNX3", 'KCTD12', "EEA1", "ARRB2", "EYA4", "PALD1", "WASHC4") & drug != "MOR"], aes (x = as.integer(time), y = LogIntensities, color = as.factor(SUBJECT))) + geom_line(aes(group = as.factor(SUBJECT))) + geom_point(aes(shape = as.factor(SUBJECT))) +
  #geom_text(data = protQuant[Protein %in% c("O00763")], aes(label = paste0(GROUP, ".", SUBJECT), size = paste0(GROUP, "_", SUBJECT) %in% problemRuns)) +
  rotate.x.axis.text +
  facet_grid(gene~drug, scales= "free_y") +
  guides(color='none',
         shape='none')
        

p

BackupAsPDF(p, 'log2FC.genesoi.linegraph')
```

Tables in wide format (write function?)

```{r}
allFits[, .(drug,Protein,gene,pvalue,adj.pvalue,log2FC,sig)] %>% 
  dcast(Protein+gene~drug, value.var = c('pvalue','adj.pvalue','log2FC','sig')) %>% 
  fwrite(ScriptAndDatedFileName("allFits.tidy.csv"))

```

09/08/23
Compare these agonists with DAMGO and PZM21 
Enrichment plots for the two anotagonists data (just compare any genes that moved up or down over time series)

```{r}
hekProt <- fread ("./data/2020_10_08_MOR_Full_DataProcess_Normalized_RunlevelData.txt.gz")
hekProt[, gene := multiUniprots2multiGenes(Protein)]
hekProt[, SUBJECT := sprintf("batch.%d", SUBJECT) ]
hekProt[, GROUP := GROUP_ORIGINAL]
hekProt[, c("receptor", "drug", "time") := tstrsplit(GROUP, "_")]
# clean
hekProt[drug == 'P', drug := 'PZM21']
hekProt[drug == 'D', drug := 'DAMGO']

# our data 
protQuant <- fread('DARPA_35.1_data/2023_08_04_ProteinLevelData.tmpnorm.csv')
#allFits <- fread('~/Documents/projects/080423_QLi_DARPA35.1/DARPA_35.1_data/2023_08_04_allFits.csv')
protQuant[, GROUP := paste0('MOR_',GROUP)]
protQuant[, SUBJECT := paste0('batch.', SUBJECT)]
protQuant[, c("receptor", "drug", "time") := tstrsplit(GROUP, "_")]
protQuant[, gene := multiUniprots2multiGenes(Protein)]
protQuant[drug == 'MOR', drug := 'Ctrl']

```
2*2 plot per gene 
```{r}
comb.dt <- rbind(hekProt, protQuant, fill=T) # hekProt[GROUP %like% '_[DP]_' & gene %in% genes.oi,] 

# genes of interest
genes.oi <-c("SNX3", 'KCTD12', "EEA1", "ARRB2", "EYA4", "PALD1", "WASHC4")

lapply(genes.oi,function(x){
  
  p = ggplot(comb.dt[gene %in% x & drug != "Ctrl" & drug != "M"], aes (x = as.integer(time), y = LogIntensities, color = as.factor(SUBJECT))) + 
   geom_line(aes(group = as.factor(SUBJECT))) +
   geom_point(aes(shape= as.factor(SUBJECT))) +
   rotate.x.axis.text +
   ggtitle(x) +
   xlab('time') +
   facet_wrap(~drug, scales= "free_y") +
   
  # guides(color='none',
  #       shape='none') +
  theme_bw()
  BackupAsPDF(p, paste0(x, '.all.timepoints.linegraph.'))
})

```
t0-t30 plots of each gene

```{r}
lapply(genes.oi,function(x){
  
  p = ggplot(comb.dt[gene %in% x & drug != "Ctrl" & drug != "M" & time %in% c("0","01","05","10","30")], aes (x = as.integer(time), y = LogIntensities, color = as.factor(SUBJECT))) + 
   geom_line(aes(group = as.factor(SUBJECT))) +
   geom_point(aes(shape= as.factor(SUBJECT))) +
   rotate.x.axis.text +
   ggtitle(x) +
   xlab('time') +
   facet_wrap(~drug, scales= "free_y") +
  theme_bw()
  p
  BackupAsPDF(p, paste0(x, '.t0-t30.linegraph.'))
})
unique(comb.dt$time)
  
```
GO enrichment heatmaps
Regen sig column for genes threshold: abs(log2FC) > log2(0.58) & pval < 0.01 

```{r}
# read in the ts results
allFits <- fread('~/Documents/projects/080423_QLi_DARPA35.1/DARPA_35.1_data/2023_08_04_allFits.csv')
# regen sig 
allFits[, sig := 'not']
allFits[pvalue < 0.01 & log2FC > log2(.58), sig := ifelse(log2FC > 0, 'up', 'down')]
allFits[,enrich_grp := paste0(drug,'.',sig)]


# read in GO gmt annotation
gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "UNIPROT")

universe <- unique(allFits$Protein)
enrich.dt <- enricherOnGroups(allFits[sig != 'not'], groupColumns = 'enrich_grp', geneColumn = "Protein", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')



ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt, groupColumn = 'enrich_grp', topN = 10, title='GO term enrichment', 
                                  negCols=unique(simp.enrich$simplified$enrich_grp[grep('down', simp.enrich$simplified$enrich_grp)]), 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 4), upperThreshold = 5)

BackupAsPDF(ht, 'go.enrich.heatmap')


View(enricherOnGroups)
```
tidy version of output with sig label for pval < 0.01 & log2FC > 0.58

```{r}
allFits[, .(drug,Protein,gene,pvalue,adj.pvalue,log2FC,sig)] %>% 
  dcast(Protein+gene~drug, value.var = c('pvalue','adj.pvalue','log2FC','sig')) %>% 
  fwrite(ScriptAndDatedFileName("allFits.pval<0.01.tidy.csv"))
```

HEatmap of all significant genes @ pval < 0.01 and abs(log2FC) > 0.58
Read in the results csv 

```{r}
protQuant <- fread('./DARPA_35.1_data/2023_08_04_ProteinLevelData.tmpnorm.csv')
mss.r <- fread('./DARPA_35.1_data/2023_08_04_allFits.csv')

protQuant[]
p.mat <- protQuant %>% 
  dcast(Protein~interaction(GROUP,SUBJECT), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')




rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))
  
# get list of sig genes
mss.r[,sig :=  fcase(
  is.na(pvalue) | is.na(log2FC) |pvalue > 0.01 | abs(log2FC) < .58, 'not',
  pvalue < 0.01 & log2FC < -.58, 'down',
  pvalue < 0.01 & log2FC > .58, 'up') ]

sig.genes <- mss.r[sig != 'not', gene]

```
Heatmap of sig genes

```{r}
subMat <- p.mat[rownames(p.mat) %chin% sig.genes,]

cluster.w.NA <- function(dat, na.val=0,...){
  dat[is.na(dat)] <- na.val
  hclust(dist(dat))
}

Heatmap(subMat, cluster_rows = cluster.w.NA(subMat), cluster_columns = F,row_names_gp  = gpar(fontsize= 5), column_split = tstrsplit(colnames(p.mat), "_")[[1]])
```


```{r}
extra_var = c('yearID', 'teamID', 'G', 'L') # model covariates added to ERA + G

#comb: generate all combinations of n elements ranging from 0 to length of extra_var
#0L,1L etc: handled as integer not double

#lapply produced a list of matrices and unlist flattens into a character vector
models = unlist( 
  lapply(0L:length(extra_var), combn, x = extra_var, simplify = FALSE),
  recursive = FALSE
)
models
```

