---
author: "Martin Gordon"
date: "2023-09-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DARPA 34.11 contrast of contrasts

Want to look at both the drug and receptor response
First look at the drug; drop NK1R and just look 

libraries

```{r}
library (data.table)
library (ComplexHeatmap)
library (ggplot2)
library (emmeans)
library(stringr)

rotate.x.axis.text <- theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")

#source Bens utility scripts
source ("/Users/martingordon/Documents/utils/bp_utils/MS_ModelFitting_TimeSeries.R")
source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/MSstats_Helper_Functions.R")
source("/Users/martingordon/Documents/utils/bp_utils/MSstats_V4_Functions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")

WriteSessionInfo()
WriteInstalledPackages()


```
Prepare the data 

```{r}
# including the dropped sample for now
protQuant <- fread ("/Users/martingordon/Documents/projects/082423_QLi_DARPA.34.11/DARAP.34.11_data/2023_08_21_ProteinLevelData.tmp.2round.csv")

# parse GROUP names to define the actual columns that the linear models will work on
# this work depends on how your columsn are defined
protQuant[, receptor   := str_extract(GROUP,"MRX2|NK1R")]
protQuant[, drug := 'SP'] %>% .[GROUP %like% 'MRX2_R[0-9]{2}', drug := 'CZ']
protQuant[, receptor.drug := interaction(receptor,drug)]

protQuant <- protQuant[receptor != 'NK1R']

# fix subject info to account for batch
protQuant[ drug != 'SP', SUBJECT := SUBJECT + 4 ]
protQuant[, SUBJECT := as.factor(SUBJECT)]
#protQuant[, SUBJECT     := paste0(drug, '.', SUBJECT)] # remember superbatch with the treatment lvl so control for this


protQuant[, timeStr := str_extract(GROUP, "[0-9]{2}$")]
protQuant[, rankTime := as.integer(as.factor(timeStr))] #timeranked to 0-5
protQuant[, .(timeStr, rankTime)] %>% unique() #timepoint to time rank 

protQuant[, .N, by=.(receptor,drug, timeStr,SUBJECT)] %>% 
  .[order(timeStr,drug)]

# need to add rankTime columns
protQuant[, `:=`(rankTime.2 = rankTime^2,
                 rankTime.3 = rankTime^3)
                  ]
```

Summarise ANOVA to assess output
emmeans contrast I think is telling us if magnitude is different between drug levels; not sure if this is what we want
This model comprises fitting, for each drug, a separate linear trend for expression versus rankTime 
Accordingly, we can estimate and compare the slopes of those lines via the emtrends() function:


```{r}
model <- lm(LogIntensities ~ SUBJECT + poly(rankTime,degree=3)*drug, data=protQuant)
emm <- emtrends(model, pairwise ~ drug, var='rankTime')

#emm <- emmeans(model, ~ drug|poly(rankTime,degree=3))

emmip(model,  drug ~ rankTime, cov.reduce= range)


emmip(model, ~ drug|poly(rankTime,degree=3))

test <-data.table(anova(model), keep.rownames = T)
test
```

```{r}
lm.out <- linearModelsAllProteins(protQuant, 
                                  formulaList = list(full  = LogIntensities ~ SUBJECT + drug + rankTime + rankTime.2 + rankTime.3 + drug:rankTime.2 + drug:rankTime.3),
                                  splitColumn = "Protein", cl=6) #use 6 threads to process in parallel

```


Looks like there are some significant interactors. Maybe just pull these out and plot

```{r}
g <- ggplot(lm.out$anova, aes(x=p.value)) +
  geom_histogram(boundary=0.0, binwidth = 0.005) +
  xlim(c(0,0.05)) +
  facet_wrap(~term)
g

BackupAsPDF(g, 'pval.histograms.')
```

correct for multiple testing

```{r}
lm.out$anova[, fdr := p.adjust(p.value, method = "BH"), by= term]
```


Write out files to dir
```{r}
fwrite(lm.out$anova,'./mrx2:drug.anova.csv.gz')
```

## interaction between response & drug type

proteins significant with FDR < 0.05

# when plotting highlight diff between drugs by subtracting T0

```{r}
protQuant[, vsTimeZero := LogIntensities - mean(LogIntensities[timeStr == timeStr[!is.na(LogIntensities)][1] ]), by = .(Protein, receptor.drug)]

sigProtein <- lm.out$anova[term %in%  c("drug:rankTime.2","drug:rankTime.3") & fdr < 0.05,] %>% 
   #  .[order(p.value), head(.SD, 100), .SDcols = c('p.value', 'Protein')] %>% 
    .[,unique(Protein)]

prot.mat <- dcast (protQuant[Protein %in% sigProtein, ], Protein~GROUP+SUBJECT, value.var = "vsTimeZero") |> as.matrix(rownames = "Protein")

rownames(prot.mat) <-multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
  
hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'Log2 Ints. vs T0',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("R[0-9]{2}", colnames(prot.mat)), "CZ", "SP")))


hm
BackupAsPDF(hm, 'mrx2.drug:time.vsT0.fdr<0.05.heatmap', dimensions = c(8,9))
```
expand net to look at any genes with pval < 0.001
again subtracting t0

```{r}
sigProtein <- lm.out$anova[term %in%  c("drug:rankTime.2","drug:rankTime.3") & p.value < 0.001,] %>% 
   #  .[order(p.value), head(.SD, 100), .SDcols = c('p.value', 'Protein')] %>% 
    .[,unique(Protein)]

protQuant[, vsTimeZero := LogIntensities - mean(LogIntensities[timeStr == timeStr[!is.na(LogIntensities)][1] ]), by = .(Protein, receptor.drug)]
prot.mat <- dcast (protQuant[Protein %in% sigProtein, ], Protein~GROUP+SUBJECT, value.var = "vsTimeZero") |> as.matrix(rownames = "Protein")

rownames(prot.mat) <-multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
  

hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'Log2 Ints. vs T0',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (pvalue < 0.001)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("R[0-9]{2}", colnames(prot.mat)), "CZ", "SP")))

hm
BackupAsPDF(hm, 'mrx2.drug:time.vsT0.pval<0.001.heatmap', dimensions = c(8,12))

```

## interaction effects

Running lm model and pull out the contrast of the interaction terms (ie different response to the drugs) but does interaction term give us this anyway?
just for testing the functions... only using complete.cases of proteins for now



070923
------
custom lm model 
for now just test with complete cases and see if there is an overlap in sig results with Ben's analysis
Issue is many significant results as different magnitude from T0..

```{r}

protQuant[, vsTimeZero := LogIntensities - mean(LogIntensities[timeStr == timeStr[!is.na(LogIntensities)][1] ]), by = .(Protein, receptor.drug)]

prot.mat <- protQuant %>% 
  dcast(Protein ~ interaction(receptor.drug,timeStr,SUBJECT), value.var = 'vsTimeZero') %>% 
  as.matrix(rownames='Protein')

prot.mat <- prot.mat[complete.cases(prot.mat),] # over 4k proteins

#run our own analysis;anova per protein
# try again with this filter, will also rm prots with low numbers of reps
test.dt <- protQuant[Protein %in% rownames(prot.mat),] %>% # filter for complete cases for now
split(., .$Protein)


# run lm on each. of the proteins, 
test.out <- lapply(test.dt, function(x,n,i){
     lm.x <- lm(vsTimeZero ~ SUBJECT + poly(rankTime, degree=3) * drug, data=x) # different response to drug over time
     #emm1 = emmeans(lm.x, specs = pairwise ~ poly(rankTime, degree=3):drug) #extract results for all pairwise comparisons for each factor combo
     emm1 = emmeans(lm.x, specs = pairwise ~ drug|poly(rankTime, degree=3)) #compare the CZ SP drug response over time 
    ##emm1$emmeans # emms for each combo. These are averaged over subject
   emm1$contrasts %>%  
   as.data.table()
})


# run lm on each. of the proteins, 
test.out <- lapply(test.dt, function(x){
     lm.x <- lm(vsTimeZero ~ SUBJECT + poly(rankTime, degree=3) * drug, data=x) # different response to drug over time
     emm1 = emmeans(lm.x, specs = pairwise ~ drug|poly(rankTime, degree=3)) #compare the CZ SP drug response over time 
     pairwise_comparisons <- pairs(emm1)
  
    # Conduct hypothesis tests (adjust p-values for multiple testing)
    contrast_tests <- summary(pairwise_comparisons, adjust = "BH")
  
  # Extract and return the results
  return(contrast_tests)
})

test.out <- rbindlist(test.out, idcol = T)
test.out[, fdr:= p.adjust(p.value, 'BH')]


final.out <- test.out


final.out[, gene := multiUniprots2multiGenes(.id)]
fwrite(final.out, './emm.mrx2:drug.csv')

# both of the above proteins look significantly different profile generate a lineplot of expression
# for the ts you need at least the 2 tps with the same number of replicates
```


plot the time0 values of these results to see if there is a significant difference
inital results look quite good

```{r}
sigProtein <- test.out[fdr < 0.05, '.id'] %>% 
   #  .[order(p.value), head(.SD, 100), .SDcols = c('p.value', 'Protein')] %>% 
    .[,unique(.id)]

prot.mat <- dcast (protQuant[Protein %in% sigProtein, ], Protein~GROUP+SUBJECT, value.var = "vsTimeZero") |> as.matrix(rownames = "Protein")
rownames(prot.mat) <-multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
  
hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'Log2 Ints. vs T0',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("R[0-9]{2}", colnames(prot.mat)), "CZ", "SP")))

hm
BackupAsPDF(hm, 'mrx2:drug.fdr<0.05.heatmap', dimensions = c(9,12))
```
```{r}
sigProtein <- test.out[fdr < 0.01, '.id'] %>% 
   #  .[order(p.value), head(.SD, 100), .SDcols = c('p.value', 'Protein')] %>% 
    .[,unique(.id)]

prot.mat <- dcast (protQuant[Protein %in% sigProtein, ], Protein~GROUP+SUBJECT, value.var = "vsTimeZero") |> as.matrix(rownames = "Protein")
rownames(prot.mat) <-multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
  
hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'Log2 Ints. vs T0',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (fdr < 0.01)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("R[0-9]{2}", colnames(prot.mat)), "CZ", "SP")))

hm
BackupAsPDF(hm, 'mrx2:drug.fdr<0.01.heatmap', dimensions = c(9,12))
```
Think we first need to confirm the proteins with a significant interaction. run the lm function again, this time check the anova output and extract specific coefficients

```{r}
# include all proteins 
protQuant[, vsTimeZero := LogIntensities - mean(LogIntensities[timeStr == timeStr[!is.na(LogIntensities)][1] ]), by = .(Protein, receptor.drug)]

prot.mat <- protQuant %>% 
  dcast(Protein ~ interaction(receptor.drug,timeStr,SUBJECT), value.var = 'vsTimeZero') %>% 
  as.matrix(rownames='Protein')

prot.mat <- prot.mat[complete.cases(prot.mat)]

#run our own analysis;anova per protein
# try again with this filter, will also rm prots with low numbers of reps
test.dt <- protQuant[Protein %in% rownames(prot.mat),] %>% # filter for complete cases for now
split(., .$Protein)

anov.out <- lapply(test.dt, function(x){
     lm.x <- lm(LogIntensities ~ SUBJECT + poly(rankTime, degree=3) * drug, data=x) 
     anov.x <- data.table(anova(lm.x), keep.rownames = T) # get the anova table 
     
     # tidy
     setnames(anov.x, old=c("rn", "Pr(>F)"), new=c('term','p.value'))
     return(anov.x[term == 'poly(rankTime, degree = 3):drug',]) # only interactions
}) %>% rbindlist(idcol = TRUE)


anov.out[, fdr := p.adjust(p.value, 'BH'), by=term]
anov.out[fdr < 0.05, .N] # only 44 significant

anov.out
```

```{r}
sigProtein <- anov.out[fdr < 0.05, '.id'] %>% 
   #  .[order(p.value), head(.SD, 100), .SDcols = c('p.value', 'Protein')] %>% 
    .[,unique(.id)]

prot.mat <- dcast (protQuant[Protein %in% sigProtein, ], Protein~GROUP+SUBJECT, value.var = "vsTimeZero") |> as.matrix(rownames = "Protein")
rownames(prot.mat) <-multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
  
hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'Log2 Ints. vs T0',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (fdr < 0.1)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("R[0-9]{2}", colnames(prot.mat)), "CZ", "SP")))

hm
BackupAsPDF(hm, 'mrx2:drug.siginteractions.fdr<0.1.vsT0.heatmap', dimensions = c(9,10))

```

are these proteins found in the list of genes with a different drug response?
about 50% of the significant interaction results in emmeans output..
Can check for overlap with BP approach; good overlap with anova; seems this is due to 
```{r}
# compare results from emmeans with those showing a significant interaction
emm.out <- fread('./emm.mrx2:drug.csv')
sum(sigProtein.BP %in% emm.out[fdr < 0.05, .id]) # only 26 prots overlap between the two

sigProtein.BP <- lm.out$anova[term %in%  c("drug:rankTime.2","drug:rankTime.3") & fdr < 0.05,] %>% 
   #  .[order(p.value), head(.SD, 100), .SDcols = c('p.value', 'Protein')] %>% 
    .[,unique(Protein)]

sum(sigProtein.BP %in% sigProtein) 
# 33 overlap out of 44 sig results, so seems the aov output looks good (modelling approaach I used in BP function doesnt account for colinearity and i also drop incomplete rows so seems ok)

length(sigProtein.BP)
length(sigProtein)
```

repeat for the cell line comparison

### cell-line effect for the same treatment (substance P)

```{r}
protQuant <- fread ("/Users/martingordon/Documents/projects/082423_QLi_DARPA.34.11/DARAP.34.11_data/2023_08_21_ProteinLevelData.tmp.2round.csv")

# parse GROUP names to define the actual columns that the linear models will work on
# this work depends on how your columsn are defined
protQuant[, receptor   := str_extract(GROUP,"MRX2|NK1R")]
protQuant[, drug := 'SP'] %>% .[GROUP %like% 'MRX2_R[0-9]{2}', drug := 'CZ']
protQuant[, receptor.drug := interaction(receptor,drug)]

# filter out the MRX2 compound Z  to allow substance P comparisons
protQuant <- protQuant[drug != 'CZ',]

# fix subject info to account for batch
protQuant[ receptor != 'MRX2', SUBJECT := SUBJECT + 4 ]
protQuant[, SUBJECT := as.factor(SUBJECT)]

#protQuant[, .N, by=.(receptor.drug, SUBJECT)]

protQuant[, timeStr := str_extract(GROUP, "[0-9]{2}$")]
protQuant[, rankTime := as.integer(as.factor(timeStr))] #timeranked to 0-5
protQuant[, .(timeStr, rankTime)] %>% unique() #timepoint to time rank 

protQuant[, .N, by=.(receptor,drug, timeStr,SUBJECT)] %>% 
  .[order(timeStr,drug)]

# need to add rankTime columns
protQuant[, `:=`(rankTime.2 = rankTime^2,
                 rankTime.3 = rankTime^3)
                  ]
```

Summarise ANOVA to assess output

```{r}
model <- aov(LogIntensities ~ SUBJECT + receptor+rankTime+rankTime.2+rankTime.3+receptor:rankTime.2+receptor:rankTime.3, data=protQuant)
summary(model)
```

```{r}

lm.out <- linearModelsAllProteins(protQuant, 
                                  formulaList = list(full  = LogIntensities ~ SUBJECT + receptor + rankTime + I(rankTime^2) + I(rankTime^3) + receptor:I(rankTime^2) + receptor:I(rankTime^3)),
                                  splitColumn = "Protein", cl=6)

```


```{r}
g <- ggplot(lm.out$anova, aes(x=p.value)) +
  geom_histogram(boundary=0.0, binwidth = 0.005) +
  facet_wrap(~term, scales = 'free_y')
g

BackupAsPDF(g, 'pval.histograms')
```

mtc correct and write out the results to file 
# anova are the BP results
```{r}
lm.out$anova[, fdr := p.adjust(p.value, 'BH')]
lm.out$anova[, gene := multiUniprots2multiGenes(Protein)]
#fwrite(lm.out$anova, './receptor:CP.anova.csv.gz')

lm.out <- fread('./receptor:CP.anova.csv.gz')
```

lets plot the significant results from each 

lots of proteins with different response to the treatment, which is good. May even have to make more stringent
Again vs t0 plotting
```{r}
protQuant[, vsTimeZero := LogIntensities - mean(LogIntensities[timeStr == timeStr[!is.na(LogIntensities)][1] ]), by = .(Protein, receptor.drug)]

sigProtein <- lm.out[term %in% c( "receptor:I(rankTime^2)","receptor:I(rankTime^3)") & fdr < 0.05, unique(Protein)] 

prot.mat <- dcast (protQuant[Protein %in% sigProtein, ], Protein~GROUP+SUBJECT, value.var = "vsTimeZero") |> as.matrix(rownames = "Protein")
rownames(prot.mat) <-multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
  

hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'Log2 Ints. vs T0',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 2),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("MRX2", colnames(prot.mat)), "MRX2", "NK1R")))

BackupAsPDF(hm, 'receptor.treatment.fdr<0.05.heatmap', dimensions=c(8,14))
```
Also look at terms significant at fdr < 0.01 as many found here

```{r}
protQuant[, vsTimeZero := LogIntensities - mean(LogIntensities[timeStr == timeStr[!is.na(LogIntensities)][1] ]), by = .(Protein, receptor.drug)]

sigProtein <- lm.out[term %in% c( "receptor:I(rankTime^2)","receptor:I(rankTime^3)") & fdr < 0.01, unique(Protein)] 

prot.mat <- dcast (protQuant[Protein %in% sigProtein, ], Protein~GROUP+SUBJECT, value.var = "vsTimeZero") |> as.matrix(rownames = "Protein")
rownames(prot.mat) <-multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
  

hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'Log2 Ints. vs T0',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (fdr < 0.01)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 4),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("MRX2", colnames(prot.mat)), "MRX2", "NK1R")))

BackupAsPDF(hm, 'receptor.treatment.fdr<0.01.heatmap', dimensions=c(8,14))
```
TODO look at my anova and lm functions also to plot and compare
other options: enrich the significant results? maybe just return to QIongyu, let her pick targets of interest and work from there

12/09/23
---
Issue with previous emmeans is we are extracting the EMMs for facotr/covariate but we want the slopes of the covariate for each level of the factor: 
instead use the `emtrends` function to extract slopes for each drug level
if we don't handle correctly, emmeans will just use the average of each covariate



Perform GO enrichment analysis of the different results and see what pops out

read in the data
```{r}
# look at receptor drug combo first
emm.mrx2.drug <- fread('./emm.mrx2.drug.csv')
anova.mrx2.drug <- fread('./mrx2.drug.anova.csv.gz')
```

re-run lm and emmeans using emtrends to extract estimates of slopes for each covariate for each drug level
```{r}
protQuant[, vsTimeZero := LogIntensities - mean(LogIntensities[timeStr == timeStr[!is.na(LogIntensities)][1] ]), by = .(Protein, receptor.drug)]

prot.mat <- protQuant %>% 
  dcast(Protein ~ interaction(receptor.drug,timeStr,SUBJECT), value.var = 'vsTimeZero') %>% 
  as.matrix(rownames='Protein')

prot.mat <- prot.mat[complete.cases(prot.mat),] # over 4k proteins

#just filter out proteins with more than 25% of values misssing
na.vals <- apply(prot.mat, 1, function(x){
  # use the any function to detect 
  any(sum(is.na(x)) > 9) })

sum(na.vals) # removing over 1k proteins
names(na.vals[na.vals == TRUE])


#run our own analysis;anova per protein
# try again with this filter, will also rm prots with low numbers of reps
test.dt <- protQuant[!Protein %in% names(na.vals[na.vals == TRUE]),] %>% # filter out missing vals
split(., .$Protein)


# run lm on each. of the proteins, 
test.out <- lapply(test.dt, function(x){
     lm.x <- lm(vsTimeZero ~ SUBJECT + poly(rankTime, degree=3) * drug, data=x) # different response to drug over time
     #emm1 = emmeans(lm.x, specs = pairwise ~ drug|poly(rankTime, degree=3)) #compare the CZ SP drug response over time 
     emm1 = emtrends(lm.x, pairwise ~ drug, var='rankTime') #asking is the slope different for each level of the factor
     emm1$contrasts %>%  
     as.data.table()
})

```

Adjust for multiple testing

```{r}
test.out <- rbindlist(test.out, idcol = T)
test.out[, fdr:= p.adjust(p.value, 'BH')]

final.out <- test.out


final.out[, gene := multiUniprots2multiGenes(.id)]

fwrite(final.out, './emm.mrx2.drug.slope.contrast.csv')
```

Lets plot the significant results for this 

```{r}
hist(final.out$p.value)

sigProtein <- final.out[p.value < 0.01, unique(.id)] 
sigProtein

prot.mat <- dcast (protQuant[Protein %in% sigProtein, ], Protein~GROUP+SUBJECT, value.var = "vsTimeZero") |> as.matrix(rownames = "Protein")
rownames(prot.mat) <-multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
  
hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'Log2 Ints. vs T0',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (pval < 0.01)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("SP", colnames(prot.mat)), "SP", "CZ")))

hm
BackupAsPDF(hm, 'emm.drug.pval<0.01.heatmap', dimensions=c(8,12))


```
There is overlapping enriched terms; look at the enrichment table for the 3 and decide which to use
```{r}
sigProtein %in% anova.mrx2.drug[term %in% c('drug:rankTime.2','drug:rankTime.3') & fdr < 0.1, unique(Protein)] 

```

# enrichment for anova drug.mrx2 and other lm approaches
enrichment for BP method looks good; write output to file and then run the other two

```{r}
# load the GO table
gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")

universe <- protQuant$gene %>%  unique()

anova.mrx2.drug[, enrich_grp := 'no']
anova.mrx2.drug[term %in% c('drug:rankTime.2','drug:rankTime.3') & fdr < 0.05, enrich_grp := 'yes']

anova.mrx2.drug[, gene := multiUniprots2multiGenes(Protein)]

enrich.dt <- enricherOnGroups(anova.mrx2.drug[enrich_grp == 'yes'], groupColumns = 'enrich_grp', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt, groupColumn = 'enrich_grp', topN = 10, title='GO term enrichment (proteins FDR < 0.05)', 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 4), upperThreshold = 10)

BackupAsPDF(ht, '/mrx2.drug.go.anova.heatmap.')

fwrite(enrich.dt, 'mrx2.drug.go.anova.enrich.table.csv.gz')

```

Run enrichment for sig results from emmeans slope analysis
Dont trust significant results here as no coordinated enrichment
Look at first GO enrichment for emmeans, otherwise proceed with BP method
Take genes with pval < 0.01 as significant

```{r}
final.out[, enrich_grp := 'no']
final.out[fdr < 0.05, enrich_grp := 'yes']


enrich.dt <- enricherOnGroups(final.out[enrich_grp == 'yes'], groupColumns = 'enrich_grp', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)

#simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')

head(enrich.dt)

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt,groupColumn = 'enrich_grp', topN = 10, title='GO term enrichment (proteins fder < 0.05)', row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 4), upperThreshold = 10)

BackupAsPDF(ht, '/mrx2.drug.go.emmeans.cov.interaction.fdr<0.05.heatmap.')

fwrite(enrich.dt, 'mrx2.drug.go.emmeans.cov.interaction.enrich.table.fdr<0.05.csv.gz')
```

Run enrichment on the original emmeans data


```{r}
emm.mrx2.drug[, enrich_grp := 'no']
emm.mrx2.drug[ fdr < 0.05, enrich_grp := 'yes']

enrich.dt <- enricherOnGroups(emm.mrx2.drug[enrich_grp == 'yes'], groupColumns = 'enrich_grp', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt,groupColumn = 'enrich_grp', topN = 10, title='GO term enrichment (proteins FDR < 0.05)', row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 4), upperThreshold = 10)

BackupAsPDF(ht, '/mrx2.drug.go.emmeans.heatmap.')

fwrite(enrich.dt, 'mrx2.drug.go.emmeans.enrich.table.csv.gz')
```
Strong overlap in enrichment for the emmeans and anova output from BP script; think this is interesting even though they are looking at different things
Now for the receptor, not interested in magnitude as may have different starting points, but we are interested in magnitude, rather just different trends

```{r}
receptor.cp <- fread('./receptor:CP.anova.csv.gz')

receptor.cp[term %in% c('receptor:I(rankTime^2)','receptor:I(rankTime^2)') & fdr < 0.05, .N] # ~450 significant results

gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")
universe <- protQuant$gene %>%  unique()

receptor.cp[, enrich_grp := 'no']
receptor.cp[term %in% c('receptor:I(rankTime^2)','receptor:I(rankTime^2)') & fdr < 0.05, enrich_grp := 'yes']

enrich.dt <- enricherOnGroups(receptor.cp[enrich_grp == 'yes'], groupColumns = 'enrich_grp', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt, groupColumn = 'enrich_grp', topN = 10, title='GO term enrichment (proteins FDR < 0.05)', 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 4), upperThreshold = 10)

BackupAsPDF(ht, '/receptor.cp.go.anova.heatmap.')

fwrite(enrich.dt, 'receptor.cp.go.anova.enrich.table.csv.gz')
```
After this lets use the emmeans package to look at interaction of factor and covariate
using emtrends to estimate the slopes for each receptor level

```{r}
protQuant[, vsTimeZero := LogIntensities - mean(LogIntensities[timeStr == timeStr[!is.na(LogIntensities)][1] ]), by = .(Protein, receptor.drug)]

prot.mat <- protQuant %>% 
  dcast(Protein ~ interaction(receptor.drug,timeStr,SUBJECT), value.var = 'vsTimeZero') %>% 
  as.matrix(rownames='Protein')

#prot.mat <- prot.mat[complete.cases(prot.mat),] # over 4k proteins

#just filter out proteins with more than 25% of values misssing
na.vals <- apply(prot.mat, 1, function(x){
  # use the any function to detect 
  any(sum(is.na(x)) > 9) })

sum(na.vals) # removing over 1k proteins
names(na.vals[na.vals == TRUE])


#run our own analysis;anova per protein
# try again with this filter, will also rm prots with low numbers of reps
test.dt <- protQuant[!Protein %in% names(na.vals[na.vals == TRUE]),] %>% # filter out missing vals
split(., .$Protein)


# run lm on each. of the proteins, 
test.out <- lapply(test.dt, function(x){
     lm.x <- lm(vsTimeZero ~ SUBJECT + receptor*poly(rankTime, degree=3), data=x) # different response to drug over time
     #emm1 = emmeans(lm.x, specs = pairwise ~ drug|poly(rankTime, degree=3)) #compare the CZ SP drug response over time 
     emm1 = emtrends(lm.x, pairwise ~ receptor, var='rankTime') #asking is the slope different for each level of the factor
     emm1$contrasts %>%  
     as.data.table()
     
})

test.out <- rbindlist(test.out, idcol = T)
test.out[, fdr:= p.adjust(p.value, 'BH')]

final.out <- test.out

final.out[, gene := multiUniprots2multiGenes(.id)]

final.out

fwrite(final.out, './emm.receptor.SP.slope.contrast.csv')
```

plot the most significant results for the emmeans contrast

```{r}
protQuant[, vsTimeZero := LogIntensities - mean(LogIntensities[timeStr == timeStr[!is.na(LogIntensities)][1] ]), by = .(Protein, receptor.drug)]

sigProtein <- final.out[p.value < 0.01, unique(.id)] 

prot.mat <- dcast (protQuant[Protein %in% sigProtein, ], Protein~GROUP+SUBJECT, value.var = "vsTimeZero") |> as.matrix(rownames = "Protein")
rownames(prot.mat) <-multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
  

hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'Log2 Ints. vs T0',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (p.value < 0.01)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 4),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("MRX2", colnames(prot.mat)), "MRX2", "NK1R")))

hm
BackupAsPDF(hm, 'receptor.treatment.emmeans.cov.interaction.pvalue<0.01.heatmap', dimensions=c(8,14))


```


Run enrichment on the emmeans output and see if it aligns with what we expect from the GO groups

```{r}
final.out[, enrich_grp := 'no']
final.out[p.value < 0.01, enrich_grp := 'yes']


enrich.dt <- enricherOnGroups(final.out[enrich_grp == 'yes'], groupColumns = 'enrich_grp', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)
#simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')

head(enrich.dt)

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt,groupColumn = 'enrich_grp', topN = 10, title='GO term enrichment (proteins fder < 0.05)', row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 4), upperThreshold = 10)

BackupAsPDF(ht, '/receptor.treatment.go.emmeans.cov.interaction.fdr<0.05.heatmap.')

fwrite(enrich.dt, 'receptor.treatment.go.emmeans.cov.interaction.enrich.table.fdr<0.05.csv.gz')
```


Not used
----

issues so far: unbalanced (only 1 factor level per protein) and timepoints with less values than polynomial degree... need to catch these errors before running

```{r}
# maybe need another sanity check to ensure both treatments exist

# handling missing vlaues for the ts analysis
# count the number of replicates per timeStr/drug.receptor combo per protein
protQuant[, n.reps := .N, by=.(timeStr,receptor.drug,Protein)]

# check for at least one timepoint with 3 reps
protQuant[, na.filter := fcase(
  n.reps >= 3, 'yes',
  n.reps < 3, 'no'), by=.(receptor.drug,Protein)]



#just filter out proteins with more than 25% of values misssing
na.vals <- apply(prot.mat, 1, function(x){
  # use the any function to detect 
  any(sum(is.na(x)) > 9) })

keep <- rownames(prot.mat)

36/4
```

Rerun with emtrends to look for differences in slopes of the two

```{r}
model <- lm(LogIntensities ~ SUBJECT + poly(rankTime,degree=3)*drug, data=protQuant)
emm <- emtrends(model, pairwise ~ drug, var='rankTime')
```

