---
title: "msstats-tutorial-neu2022"
author: "Martin Gordon"
date: "2023-05-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages}
library(data.table)
library(stringr)
library(purrr)
library(MSstats)
library(ggplot2)
```

## MSStats Tutorial North Eastern University orkshop 2022

Using MSstats for DIA workflow
Analyse data produced by Skline peptide identification/quantification tool
Benchmark dataset Selevsek et al 2015
All (Selevsek-all; quantify everything) vs refined dataset (Selevsek-reps; replicate data) vs high-qual (Selevsek-lowcv; detectable and repeatable)
Remember how refined datasets improve the signal to noise ratio


MSStats Annotation emplate
------
Needs three cols: Conditon, Bioreplicate, Run
For time course, bioreplicates need to be repeated, but must be distinct for simple comparisons to model sample independence correctly
For MSstats need output from the processing tools and then the metadata file

MSStats Key Functions
-----
- Import command: preprocessing for MSStats input
- dataProcess function: transformation, normalisation, imputation and normalisation *this step has NB influence on output
- groupComparison: inference & hypthesis testing
- designSampleSize: info on experimental power and sample size requirements for future experiments


MSStats Statistical Inference
---
For statistical inference, MSStats is quite restrictive compared to other tools and essentially assigns a model based on the Annotation table input
Basially belong to case-control design or time-study design; assignment belongs to bioreplicate assignment
*Important to assign the correct linear model*
For timecourse, MSStats uses a linear mixed model fit by REML, for case-control, uses a linear model

Design Future Experiment
---
Can either look at power or sample size; what sample size do you need to achieve X power?
What power does a sample size of x give you?

MSStats Preprocessing
----
Important step
Equilizemedian; default make medians for all peptides equal
Quantile; make intensity distributions the same, maybe more appropriate for genomics as proteomics v skewed at lower abundance and will remove much of this data which is important for determining LOD and imputation; especially not suited to DDA
Global standards; give specific peptide that should be same across runs (spike-ins) and use this to control
Features used; for DIA lots of features detected, but maybe many low quality and not very computationally efficient so can select top N features
*Normalisation step NB* EquilizeMedians assumes you are globally profiling, you should have a lot of data and most features should not change. Not suitable eg for targeted approaches!

Missing values:
How do we treat this? Tools treat missing values differently (NAs or 0s) so handle differently
For this workflow, maybe NA are randomly missing, while censored are v low values so they can be imputed

Imputaion:
What imputation approach to use? We use none at Krogan

Summarisation:
Uses Tukey median polish method. Can also select to drop runs ith 50% NA values
If you are curious about the values you are basing your summarisation on, you can generate profile plots using code below(grey features/peptides, redline is summarised proteins)

Remember MSStats cuts off low qual values (< 3rd percentile)

Questions
- MSStats not v flexible 
- Batch effects? Recommended to use MSstats up to dataprocessing then import to Limma to deal with batch effects
- Is a tool available for DIA-NN import?

Workflow
----

Preprocessing Output
----
Protein level summarisation headers:
Protein; protein id (UNIprot??)
TotalgroupMeasurements; no of measured feature across the group. 
NumMeasuredFeatures; features measured in that biorep;
MissingPercentage;number of features with missing values

```{r}
#input files
path <- getwd()
files <- data.table(path=dir(path = path, recursive=T, full.names=F, pattern= "./*plus*.csv"))

files[, f_name:=str_extract(path, "Selevsek-.*?(?=/)")] #matches "Selevsek-" at the beginning of the input, .*?" matches any character (except newline) lazily, meaning it matches as few characters as possible (?=/)" is a lookahead assertion that checks if the next character is a forward slash (/)

#read in files
input_list <- Map(fread, files$path)
names(input_list) <- files$f_name
```

Prepare annotation dt
----

```{r}
annotation_list <- input_list %>% 
  map( ~{
    unique(.x[,  c('Condition', 'BioReplicate', 'FileName')])
  })

```

Import Skyline data to MSStats format
----

```{r}
quant_list <- input_list %>% 
  map( ~{
    SkylinetoMSstatsFormat(.x,
                          filter_with_Qvalue = TRUE, ## same as default
                          qvalue_cutoff = 0.01, ## same as default
                          fewMeasurements="remove", ## same as default
                          removeProtein_with1Feature = TRUE)
  })

```

EDA
---
Plot histograms of the data to inspect the visual quality
No peak in low abundance proteins; already filtered?
```{r}
quant_list %>% 
  map( ~{
    ggplot(data = .x, mapping=aes(x=log10(Intensity))) +
    geom_histogram()
  })
```

Raw (log2) normalised plots. Much better quality than our runs but it is a benchmark dataset!

```{r}
quant_list %>% 
  map( ~{
    ggplot(data = .x, mapping=aes(x=paste0(Condition, "_", BioReplicate), y=log2(Intensity), color=Condition)) +
    geom_boxplot()
  })
```

Also barplot of raw intensities; looks like high quality filtered dataset
```{r}

quant_list %>% 
  map( ~{
    ggplot(data = .x, mapping=aes(x=paste0(Condition, "_", BioReplicate), fill=Condition)) +
    geom_bar()
  })
```

MSStats data process function
-----

```{r}
quant_list <- quant_list %>% 
  map( ~{
   dataProcess(.x,
              normalization = 'equalizeMedians',
              summaryMethod="TMP",
              featureSubset = "highQuality", #optional; flags uninformative features/outliers
              remove_uninformative_feature_outlier = TRUE, #if above set need this to remove flagged features
              MBimpute=FALSE,
              censoredInt="0", #Skyline censors using 0
              maxQuantileforCensored=0.999)#default
    
  })

MSstatsContrastMatrix(contrasts, conditions, labels = NULL)
```

Quality Inspection -seems good

```{r}

quant_list %>% 
  map( ~{
    ggplot(data = .x$ProteinLevelData, mapping=aes(x=paste0(GROUP, "_", SUBJECT), y=LogIntensities, fill=GROUP)) +
    geom_boxplot()
  })
```

Profile Plot Code below - can use to inspect the summarizarion (not run)

```{r eval=FALSE}
## featureName option
dataProcessPlots(processed.quant, type="Profileplot", 
                 ylimDown=0, 
                 which.Protein = 'YKL096W',
                 featureName = 'Peptide',
                 originalPlot = TRUE,
                 summaryPlot = TRUE,
                 width=7, height=7,  
                 address="Selevsek2015_DIA_Skyline_all_YKL096W_peptide_")

dataProcessPlots(processed.quant, type="Profileplot", 
                 ylimDown=0, 
                 which.Protein = 'YKL096W',
                 featureName = 'Transition',
                 originalPlot = TRUE,
                 summaryPlot = TRUE,
                 width=7, height=7,  
                 address="Selevsek2015_DIA_Skyline_all_YKL096W_feature_")
```

Model-based comparison + adjust p-value
---
- Generate contrast matrix
- Perform groupcomparson
- Can generate volcano plots, comparison plots etc.



