---
title: "5min_talk_plots"
author: "Martin Gordon"
date: "2024-07-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Generate some plots for the 5 minute AViDD talk
Just read in the original data and rename the groups

```{r}
library(magrittr)
library(data.table)
library(circlize)
library(ComplexHeatmap)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(MSstats)
library(ggrepel)
library(hrbrthemes)
library(viridis)
library(ggbeeswarm)
library(usedist) #package for working with distances
library(ggh4x) # additional functionality for ggplot2 obj
library(scales)
library(seqinr) # reading fasta files
library(RColorBrewer)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

source ("../../utils/mg_utils/r_utils/IDmapping.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```

Read in the abundance data

```{r}
spec <- fread('/Users/martingordon/Documents/projects/052924_ARichards_AViDD_Murine/data/MPro_Lung_052324/061224/v2/20240615_143854_AVD3_APdb_061524_Report.tsv')

spec[,.N, by=.(Condition,BioReplicate,Run)] # Nirm 4d 5  seems to be missing

spec <- spec[!Condition %in% c("x4673_2d", "x4673_4d" ,"x4673_7d"),]
```

Need to tidy names in the two files 

Drop junk measurements on the left shoulder
```{r}
# no multiple feature peptide ions detected
spec[,.N, by=.(PeptideSequence,PrecursorCharge,Run)][N >1]

hist(log2(spec$Intensity))
spec.dt <- spec[Intensity > 2^4,]
hist(log2(spec.dt$Intensity))
```

Lets rename the conditions

```{r}
spec.dt[Condition == 'mock', Condition := 'Mock']
spec.dt[Condition %like% 'V_', Condition := gsub('V_', 'Vehicle_', Condition)]
spec.dt[Condition %like% 'Nirm_', Condition := gsub('Nirm_', 'Treatment1_', Condition)]
spec.dt[Condition %like% 'x4516', Condition := gsub('x4516', 'Treatment2_', Condition)]
spec.dt$Condition %>%  unique()
```

Normalize to the protein level and plot

```{r}
mssInput <- spec.dt[,  IsotopeLabelType := 'L']

ggplot(spec.dt, aes(x=paste0(Condition, BioReplicate), y=log2(Intensity), fill=Condition)) +
  geom_boxplot() +
  theme(axis.text.x=element_text(angle=90))
```


```{r}
dp.out <- MSstats::dataProcess(mssInput, 
                              MBimpute =  FALSE, 
                              normalization = 'EQUALIZEMEDIANS',
                              featureSubset = "highQuality", 
                              remove_uninformative_feature_outlier = TRUE)
```

plot protein level data

```{r}
p.quant <-  setDT(dp.out$ProteinLevelData)

p.quant[GROUP %like% 'Treatment2', GROUP := gsub('Treatment2_', 'Treatment2', GROUP)]
p.quant[, GROUP := as.factor(GROUP)]

ggplot(p.quant, aes(x=paste0(GROUP, SUBJECT), y=LogIntensities, fill=GROUP)) +
  geom_boxplot() +
  theme(axis.text.x=element_text(angle=90))

```
Lets add a second round of normalization
Try TMP or median subtraction

```{r}
p.mat <- dcast(p.quant, Protein~interaction(GROUP,SUBJECT), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')

tmp <- medpolish(p.mat, na.rm = T)

tmp.dt <- data.table(sample = names(tmp$col),
                     offSet = tmp$col)

p.quant[,sample := interaction(GROUP,SUBJECT)]

p.quant <- merge(p.quant, tmp.dt, by='sample', all.x=T)

p.quant[, newLogInts := LogIntensities - offSet]
```

```{r}
ggplot(p.quant, aes(x=paste0(GROUP, SUBJECT), y=newLogInts, fill=GROUP)) +
  geom_boxplot() +
  theme(axis.text.x=element_text(angle=90))
```

```{r}
col.pal <- randomcoloR::distinctColorPalette(k=length(unique(p.quant$GROUP)))

p.mat <- dcast(p.quant, Protein~interaction(GROUP,SUBJECT), value.var = 'newLogInts') %>% 
  as.matrix(rownames='Protein')

p.mat <-  p.mat[complete.cases(p.mat),]

submat <- sweep(p.mat, 1, apply(p.mat, 1, median, na.rm=T))

# create annotation
ann <- data.table(treatment=sub("[_.].+", "", colnames(p.mat)),
                  timepoint=ifelse(grepl("[0247]d", colnames(p.mat)), str_extract(colnames(p.mat), "[0247]d"), 'NA'))

colours <- list('timepoint' = c('NA' = '#4477AA', '2d'="#228833", '4d'="#CCBB44", '7d'='#EE6677'),
                'treatment' =c('Mock' = col.pal[1], 'Vehicle'=col.pal[2], 'Treatment1'=col.pal[3],  'Treatment2'=col.pal[5])
                )

colAnn <- HeatmapAnnotation(df = ann, col = colours)

hm <- Heatmap(submat, 
        border=T,
        name='Ints./Median',
        top_annotation = colAnn,
        cluster_column_slices = F,
        cluster_columns=F,
        show_row_names = F,
        column_title_gp = gpar(font='bold', fontsize=8),
        column_names_gp = gpar(font='bold', fontsize=5),
        column_split = gsub('_[247]{1}[0-9]{1}$|[0-9]{1}$', '', colnames(submat)))


BackupAsPDF(hm, 'medianScaled.heatmap', dimensions=c(11,9))

#subtract vs mock values
submat <- sweep(p.mat,1, apply(p.mat[, colnames(p.mat) %like% 'Mock'], 1, median,na.rm=T))


hm <- Heatmap(submat, 
        border=T,
        name='Ints/Mock',
        top_annotation = colAnn,
        cluster_column_slices = F,
        cluster_columns=F,
        show_row_names = F,
        column_title_gp = gpar(font='bold', fontsize=8),
        column_names_gp = gpar(font='bold', fontsize=5),
        column_split = gsub('_[247]{1}[0-9]{1}$|[0-9]{1}$', '', colnames(submat)))

hm
BackupAsPDF(hm, 'vsMock.heatmap', dimensions=c(11,9))
```
Create a PCA of this plot

```{r}
pcaOut <- prcomp(t(p.mat))

ann[, colname := colnames(p.mat)]
ann[, group := paste0(treatment,'.',timepoint)]
#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, ann, by.x = "rn", by.y = "colname", all.x = TRUE)

#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = group, shape = timepoint)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle(sprintf ("PCA using %d features (log intensity)", nrow(p.mat))) +
  #scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = c(21:26)) +
  scale_fill_manual(values = col.pal) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, 'proteins.groupandtimepoint.pca', dimension=c(10,8))
```
Run PCA batch correction and replot the two heatmaps side-by-side

```{r}
p.mat <- dcast(p.quant, Protein~interaction(GROUP,SUBJECT), value.var = 'newLogInts') %>% 
  as.matrix(rownames='Protein')

p.mat <-  p.mat[complete.cases(p.mat),]

# order samples by PC1
col.order <- pcaDT[order(PC1), rn]

p.mat <- p.mat[, col.order]
submat <- sweep(p.mat, 1, apply(p.mat, 1, median, na.rm=T))
```

Visualise the data

```{r}
# create annotation
ann <- data.table(treatment=sub("[_.].+", "", colnames(p.mat)),
                  timepoint=ifelse(grepl("[0247]d", colnames(p.mat)), str_extract(colnames(p.mat), "[0247]d"), 'NA'))

colours <- list('timepoint' = c('NA' = '#4477AA', '2d'="#228833", '4d'="#CCBB44", '7d'='#EE6677'),
                'treatment' =c('Mock' = col.pal[1], 'Vehicle'=col.pal[2], 'Treatment1'=col.pal[3],  'Treatment2'=col.pal[5])
                )

colAnn <- HeatmapAnnotation(df = ann, col = colours)


hm <- Heatmap(submat, 
        border=T,
        name='Ints/Median',
     #   top_annotation = colAnn,
        cluster_column_slices = F,
        cluster_columns=F,
        show_row_names = F,
        column_title_gp = gpar(font='bold', fontsize=8),
        column_names_gp = gpar(font='bold', fontsize=5),
        column_split = gsub('_[247]{1}[0-9]{1}$|[0-9]{1}$', '', colnames(submat)))

hm

```
Subtract the PC1 and visualize side-by-side

```{r}
# recover the first pc matrix by the protein loadings to recover the original matrix
pc1.mat <- pcaOut$x[,1] %*% t(pcaOut$rotation[,1])
pc1.mat <- t(pc1.mat) 


Heatmap(pc1.mat,row_order = row.order, show_row_names = F)

colnames(pc1.mat) <- colnames(p.mat)

smooth.mat <- p.mat - pc1.mat

smooth.clean.mat <- sweep(smooth.mat, 1, apply(smooth.mat, 1, median, na.rm=T))

Heatmap(smooth.clean.mat, row_order = row.order, cluster_columns = F, show_row_names = F)


hm + Heatmap(smooth.clean.mat, 
        border=T,
        name='Smoothed Ints/Median',
        row_order = row.order,
        cluster_column_slices = F,
        cluster_columns=F,
        show_row_names = F,
        column_title_gp = gpar(font='bold', fontsize=8),
        column_names_gp = gpar(font='bold', fontsize=5),
        column_split = gsub('_[247]{1}[0-9]{1}$|[0-9]{1}$', '', colnames(submat)))


hm

```




Run PCA batch correction and replot 

vok, now we have imputed the missing values, lets see if we can detect the latent batch variables using sva
```{r}
library(sva)

metadata <- p.quant[, .(sample = interaction(GROUP,SUBJECT),
                        GROUP = factor(GROUP),
                        status = gsub('[_].+$', '', GROUP), 
                        day=ifelse(GROUP != 'Mock', str_extract(GROUP, '[247]d'), 'NA') )] %>%  unique()


#create full model inc. explanatroy variables and null model with just uninteresting var (unknown so just inlcude intercept model)
mod <- model.matrix(~GROUP, data=metadata)
mod0 <- model.matrix(~1, data=metadata)
```
Run SVA; estimate batch and other artifacts in the expression matrix
Estimated only 1 latent variable that explains variance... 
```{r}
n.sv <- num.sv(p.mat, mod, method='leek')

svobj <- sva(p.mat, mod, mod0, n.sv)


batch.dt <- data.table(sample=colnames(p.mat),
                       GROUP=gsub('[.][0-9]$','', colnames(p.mat)),
                       sva=svobj$sv[1:57,])

## subtracting the latent batch variable from the expression method
## form here: https://www.biostars.org/p/262616/
## solving a least squares problem to find the coefficients (alpha) that quantify the contribution of surrogate variables to the observed data. 
# This involves matrix transposition, multiplication, and inversion to achieve the solution efficiently using linear algebra.

svaBatchCor <- function(dat, mmi, mm0,n.sv=NULL){
    dat <- as.matrix(dat)
    Y <- t(dat)
    library(sva)
    if(is.null(n.sv))   n.sv <- num.sv(dat,mmi,method="leek")
    o <- sva(dat,mmi,mm0,n.sv=n.sv)
    W <- o$sv
    alpha <- solve(t(W) %*% W) %*% t(W) %*% Y
    o$corrected <- t(Y - W %*% alpha)
    return(o)
}

batchCor <- svaBatchCor(p.mat, mmi=mod, mm0=mod0, n.sv=1)

corrected.sub <- sweep(batchCor$corrected, 1, apply(batchCor$corrected, 1, median, na.rm=T))

ht <- Heatmap(corrected.sub, 
        border=T,
        name='Smoothed Ints./Mock',
       # top_annotation = colAnn,
        cluster_column_slices = F,
        cluster_columns=F,
        show_row_names = F,
        column_title_gp = gpar(font='bold', fontsize=8),
        column_names_gp = gpar(font='bold', fontsize=5),
        column_split = gsub('_[247]{1}[0-9]{1}$|[0-9]{1}$', '', colnames(submat)))

ht
hm + ht


BackupAsPDF(hm, 'vsMock.corrected.heatmap', dimensions=c(11,9))
```

```{r}
pcaOut <- prcomp(t(corrected.mat))

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)

pcaDT
pcaDT <- merge (pcaDT, ann, by.x = "rn", by.y = "colname", all.x = TRUE)
pcaDT <- merge(pcaDT, batch.dt, by.x='rn', by.y='sample', all.x=T)


#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = group, shape = timepoint)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  #scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = c(21:26)) +
  scale_fill_manual(values = col.pal) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, 'proteins.svaCol.pca',  dimension=c(10,8))


```

PCA of the corrected matrix

```{r}

```

