---
title: "061625_D3_MSstatsSmmarization"
author: "Martin Gordon"
date: "2025-06-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Prepare the SEC-MS data for input into BP pipeline and CCprofiler
Stil grappling with the different data formats produced by the Spectronaut searches. Now we will try comparing the MSstats format file between CCprofiler and Bens pipeline
Review the code from the previous MSS summarization and implement this for the protein level summarization. Tidy and use the peptide level info for CCprofiler

```{r packages}
library(MSstats)
library(data.table)
library(ggplot2)
library(magrittr)
library(stringr)
library(parallel)
library(pbapply)



# useful utility scripts
source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")


source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("../../utils/mg_utils/r_utils/msstats_helperFunctions.R")

clusterWNA <-  function(mat, na.value=0,...){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}

customTheme <-  theme_bw() +
  theme(axis.text.x = element_text(angle=90),
        panel.border = element_rect(color = "lightgrey",
                                    fill = NA,
                                    size = 1))
```

Using parallel package to deal the lappy functions 

```{r}
cores <- parallel::detectCores() - 2
```


Focus on the msstats reports; read these into file, summarize for input to SEC-MS pipeline 
```{r}
allFiles <- dir('./data/HIV_Infection_CD4T_cells/Spectronaut_Reports_061625', full.names = T, include.dirs = T, recursive = T, pattern="*MSStats_Report.tsv")
allFiles <-  grep('Cell_Lysis', allFiles, invert=T, value=T)

# 
pep.ls <-  pbapply::pblapply(allFiles, fread, cl=9)
names(pep.ls) <-  sapply(allFiles, function(x) unlist(strsplit(x, '[/]'))[7])
names(pep.ls) <- gsub('MSStats_Report.tsv', '', names(pep.ls))
```

naming convention not standardised... you know what? Lets just name manually for simplification
```{r}
names(pep.ls) <- c('Infected_1', 'Challenged_1', 'Uninfected_1', 'Infected_2', 'Challenged_2', 'Uninfected_2', 'Infected_3', 'Challenged_3', 'Uninfected_3')
```

No imputation from the looks of things
```{r}
lapply(pep.ls, function(x) summary(x))

lapply(pep.ls, function(x) {hist(log2(x$Intensity))})

pep.dt <- rbindlist(pep.ls, idcol='sample')
```


Want to run the MSstats summarization for each sample seperately. RN we are using tmp to summarize, but need to avoid more complicated outlier detection, imputation etc.
Do a simplified MS summarization
```{r}
test <- pep.ls[[1]]

#' Given an MSstats long format file, we want to summairze to protein level information
#' Check for essential 
MSSproteinSummarization <- function(mss.in, summary.method='TMP', imputeVals=FALSE, ...){
  
  dt <- copy(mss.in)

  cols.oi <- c("Condition", "Run","BioReplicate","ProteinName","PeptideSequence","FragmentIon","PrecursorCharge","Intensity","ProductCharge")
  
  # summarizing protein level information
  stopifnot(all(cols.oi %in% colnames(dt)))
  stopifnot(length(unique(dt$BioReplicate)) == 1)
  stopifnot(is.logical(imputeVals))
  stopifnot(toupper(summary.method) %in% c('TMP', 'linear'))
  
  message('Adding PeptideModifiedSequence IsotopeLabelType columns...')
  dt[, c('PeptideModifiedSequence', 'IsotopeLabelType') := .(NA, 'L')]
  
  message('Warning: Considering each Run/Condition combo as a seperate fraction...\nPlease ensure Run column matches the fraction before running')
  dt[, Condition := paste0(Condition, '.fraction', Run)]
  
  message('Running MSstats summarization...')
  if (imputeVals)
    message('Imputation enabled. This will significantly increase dataProcess times')
  
  dp.out <- MSstats::dataProcess(dt, 
                                MBimpute =  imputeVals, 
                                normalization = FALSE, # turn off for SECMS
                                summaryMethod = summary.method,
                                ...)
  
  return(dp.out)
}

```

Check out runtimes of various runs; plot

```{r}
# simpleiest version
t1 <- system.time(MSSproteinSummarization(test, summary.method = 'TMP', imputeVals = FALSE))
# subset to only include features >= n features
t2 <- system.time(MSSproteinSummarization(test, summary.method = 'TMP', imputeVals = FALSE, featureSubset='top3'))

# high quality features only 
t3 <- system.time(MSSproteinSummarization(test, summary.method = 'TMP', imputeVals = FALSE, min_feature_count=3, featureSubset='highQuality'))

# imputation enabled
t4 <- system.time(MSSproteinSummarization(test, summary.method = 'TMP', imputeVals = TRUE))

# imputation and high qual feature. Considered most robust, but runtime maybe an issue...
t5 <- system.time(MSSproteinSummarization(test, summary.method = 'TMP', imputeVals = TRUE, min_feature_count=3, featureSubset='highQuality'))
```

```{r runtimes, fig.width=4, fig.height=5}
exec.dt <- data.table(Reduce(rbind, list(t1,t2,t3,t4,t5)))
exec.dt[, rn := c('TMP', 'TMP (top 3 features)', 'TMP (highQual features)', 'TMP + Impute', 'TMP (highQual features) + Impute')]

g <- ggplot(exec.dt, aes(x=reorder(rn, elapsed/60),y=elapsed/60)) +
  geom_bar(stat='Identity') +
  theme(axis.text.x=element_text(angle=90))

BackupAsPDF(g, 'ms.summarization.runtimes.barplot')
```

Lets summarize the samples, we will use the most simple summarization method (use tmp on all features)

```{r prot-summarization}
cores <- parallel::detectCores() - 2

ms.ls <- pbmcapply::pbmclapply(pep.ls[1:9],  function(pep.dt){ 
  tryCatch({
    MSSproteinSummarization(pep.dt,
                            summary.method = 'TMP', 
                            imputeVals = FALSE)
  }, error = function(e){
     message("ERROR on one of the data.tables: ", e$message)
  })
},mc.cores = cores)
```

Save the RDS objects to file, read in the peptide level data, summarize to 

```{r}
ms.ls <- parallel::mclapply(seq_along(pep.ls), function(i) {
  tryCatch({
    message("Processing index: ", i)
    MSSproteinSummarization(pep.ls[[i]], summary.method = 'TMP', imputeVals = FALSE)
  }, error = function(e) {
    message("ERROR at index ", i, ": ", e$message)
    NULL
  })
}, mc.cores = cores)

names(ms.ls) <- names(pep.ls)

```

An issue with the pbmclapply function; just use mclapply or oblapply with `cl` might be the equivalent
# save the protein and feature level data in DT format and use to run the analysis
```{r}
feat.dt <- unlist(ms.ls, recursive=FALSE)[grepl('FeatureLevelData', names(unlist(ms.ls, recursive=FALSE)))] %>% 
  rbindlist(idcol='sample')

prot.dt <- unlist(ms.ls, recursive=FALSE)[grepl('ProteinLevelData', names(unlist(ms.ls, recursive=FALSE)))] %>% 
  rbindlist(idcol='sample') %>% 
  setDT()

prot.dt <- prot.dt[, .(sample = gsub('.ProteinLevelData', '', sample), protein=Protein, intensity=2^LogIntensities, LogIntensities, fraction=originalRUN, NumMeasuredFeature)]
fwrite(prot.dt, ScriptAndDatedFileName('hiv.d123.mss.summarized.csv.gz'))
```
For the feature level data, summarize to protein level and write out

```{r}
feat.dt[, c('peptideSeq', 'precursorCharge') := tstrsplit(PEPTIDE, '__', keep=c(1,2))]
feat.dt <- feat.dt[, .(sample = gsub('.FeatureLevelData', '', sample), fraction=originalRUN, protein=PROTEIN, sequence=gsub('_', '', peptideSeq), precursorCharge, intensity=INTENSITY)]
pep.dt <- feat.dt[, .(intensity=sum(intensity,na.rm=T)), by=.(sample, fraction, protein, sequence)]

feat.dt[,.N, by=.(protein, sequence, sample, fraction)][N>1]
```
```{r}
pep.dt[protein == 'A0AVT1' & sequence == 'DKFESSFSHKPSLFNK' & sample == 'Infected_1' & fraction == 1,]


feat.dt[protein == 'A0AVT1' & sequence == 'GHTEVIVPHLTESYNSHRDPPEEEIPFC[Carbamidomethyl (C)]TLK' & sample == 'Infected_1' & fraction == 1,]
pep.ls[['Infected_1']][ProteinName == 'A0AVT1' & Run == 1,]

# what is interesting is that this lists the peptides/features found across all runs per sample, but will attach an NA values for where that run/fraction is undetected
feat.dt[protein == 'A0AVT1' & sample == 'Infected_1' & fraction == 1,]

fwrite(feat.dt[intensity != 0],  ScriptAndDatedFileName('hiv.d123.mss.feature.csv.gz'))
fwrite(pep.dt[intensity != 0],  ScriptAndDatedFileName('hiv.d123.mss.peptide.csv.gz'))
```

sanity check; looks good!
```{r}
feat.dt[intensity != 0][,.N, by=.(sample,protein, sequence, fraction)][N>1]
feat.dt[sample== 'Infected_1' & protein == 'O00487' & sequence == 'QTTSNLGHLNKPSIQALIHGLNR' & fraction == 1, sum(intensity)]
pep.dt[sample== 'Infected_1' & protein == 'O00487' & sequence == 'QTTSNLGHLNKPSIQALIHGLNR' & fraction == 1]

```

Question here is what do we consider a sample. I think a sample should be the fraction/condition combo

```{r}

spec.list[['sonic']][!is.na(Intensity)]

lapply(spec.list, str)
lapply(spec.list, head)

dp.out <- lapply(names(spec.list), function(x){

  message('Running MSStats dataProcess on ', x)
  
  dp.out <- MSstats::dataProcess(spec.list[[x]], 
                                MBimpute =  FALSE, 
                                normalization = FALSE, # turn off for SECMS
                                summaryMethod = "TMP",
                                featureSubset = 'highQuality',
                                remove_uninformative_feature_outlier=T)

  saveRDS(dp.out, paste0('./output/', x, '.mss.dp.rds'))
  return(dp.out)
})
```



**not used**

```{r}
allFiles <- dir('./data/HIV_Infection_CD4T_cells/Donor-3-061125/D3_Stringent_Spectronaut_Parameters_061225/', full.names = T, include.dirs = T, recursive = T)
pep.ls <- lapply(grep('Peptide', allFiles, value=T), fread)
names(pep.ls) <- sapply(grep('Peptide', allFiles, value=T), function(x) unlist(strsplit(x, '[/]'))[7])
names(pep.ls)
```

# convert to DT and tidy

```{r}
pep.dt <- rbindlist(pep.ls, idcol='sample')
pep.dt[, sample := gsub('_061225', '', sample)]
```

nasty format... lets tidy, remove unnessary colnames and combine in long format and save the files
```{r}
ints.cols <- grep('TargetQuantity', colnames(pep.dt), value=T)
qval.cols <- grep('Qvalue', colnames(pep.dt), value=T)
id.cols <- colnames(pep.dt)[1:9]

ints.long <- pep.dt[, !..qval.cols] %>% 
  melt(., idvars=id.cols, measure.vars = ints.cols)
setnames(ints.long, old=c('variable', 'value'), new=c('run', 'intensity'))

qval.long <- pep.dt[, !..ints.cols] %>% 
  melt(., id.vars = id.cols, measure.vars = qval.cols)
setnames(qval.long, old=c('variable', 'value'), new=c('run', 'qvalue'))
colnames(qval.long)

# tidy the run column; extract the val in square brackets and the follwoing first numberic set
ints.long[, c('fraction', 'fraction2') := tstrsplit(run, '[^0-9]+', keep=c(2,3))]
qval.long[, c('fraction', 'fraction2') := tstrsplit(run, '[^0-9]+', keep=c(2,3))]
# check fraction == fraction2, if so drop fract two and then create a replicate column. we will save these as peptide level measurements for importing into 

ints.long[fraction == fraction2]

#drop ununsed cols
cols.noi <- c('PG.ProteinDescriptions', 'run', 'fraction2')
ints.long[, c(cols.noi) := NULL]
qval.long[, c(cols.noi) := NULL]

# drop fraction and ints
key.cols <- colnames(ints.long)[-9]

# merge the values
pep.dt <- merge(ints.long, qval.long, by=key.cols)

colnames(pep.dt)
pep.dt[, replicate := 3]

```

tr



tidy and save the peptide info 

```{r}
fwrite(pep.dt[, .(sample, replicate, fraction, protein=PG.ProteinAccessions, gene=PG.Genes, precursor=EG.PrecursorId, pg_specific=PEP.IsProteinGroupSpecific,
           decoy=EG.IsDecoy, mw=PG.MolecularWeight, org=PG.Organisms, intensity, qvalue)], ScriptAndDatedFileName('hiv.d3.peptide.ints.long.csv.gz'))
```
read in the data and summarize to the protein level 
```{r}
msin.dt <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061625_D3_MSstatsSmmarization_data/2025_06_16_hiv.d3.peptide.ints.long.csv.gz')

```

# so it looks like the decoys are detected at the protein level; we definitely dont want to include this in our other protein quant estimates;
for safety prefix 'decoy' to each protei/gene name
```{r}
# all of these decoy have an NA qvalue
msin.dt[decoy == TRUE, .(protein, gene,precursor, qvalue)]
msin.dt[gene == 'CIZ1', .N, by=.(decoy,sample)]


# tidy the decoy names for gene and protein
paste0('decoy_',strsplit("Q12959;Q12959-3;Q12959-5;Q12959-6" , ";")[[1]], collapse=';')
msin.dt[decoy == TRUE, unique(protein)]


sdcols <- c('protein', 'gene')
msin.dt[decoy == TRUE, (sdcols) :=  lapply(.SD, function(x){ paste0('decoy_',strsplit(x, ";")[[1]], collapse=';')}), .SDcols = sdcols, by=.I]
```

```{r}
fwrite(msin.dt, ScriptAndDatedFileName('hiv.d3.peptide.ints.long.csv.gz'))
msin.dt <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061625_D3_MSstatsSmmarization_data/2025_06_16_hiv.d3.peptide.ints.long.csv.gz')
```

# prepare the files for MSS input

first prepare an annotation data.table 
```{r}
anno.dt <- msin.dt[, .(protein, gene, mw)] %>% 
  unique()
#fwrite(anno.dt, ScriptAndDatedFileName('prot.id.mapper.csv.gz'))
```

Tidy the input to see which proteins will be input into the analysis
```{r}
# create cols
msin.dt[, PrecursorCharge :=  as.numeric(str_extract(precursor, '[0-9]{1,2}$'))]
msin.dt[, IsotopeLabelType := 'L']
msin.dt[, Condition := ifelse(sample == 'D3-Challenged_Infected', 'Infected',
                              ifelse(sample == "D3-'Challenged_Uninfected", 'Challenged', 'Uninfected'))]
msin.dt[, BioReplicate := replicate]
msin.dt[, PeptideSequence := gsub('[.][0-9]+$', '', precursor)]
msin.dt[, Run := paste0(Condition, '_', fraction, '.raw')]
msin.dt %>% colnames()
setnames(msin.dt, old=c('protein'), new=c('ProteinName'))

fwrite(msin.dt[, .(Run, Condition, BioReplicate, ProteinName, PeptideSequence, PrecursorCharge, IsotopeLabelType, Intensity=intensity)], ScriptAndDatedFileName('d3.hiv.mss.input.csv.gz'))
```
Do a protein summarization; lets see if the pquant groups are maintained
```{r}
msin.dt <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061625_D3_MSstatsSmmarization_data/2025_06_16_d3.hiv.mss.input.csv.gz')
msin.dt[, c('PeptideModifiedSequence', 'FragmentIon', 'ProductCharge') := NA]  #need these cols in the data 
msin.dt[, treatment := Condition]
msin.dt[, Condition :=  Run]
msin.dt[, Run := paste0(Condition,'.', BioReplicate, '.raw')]
```
Run in parallel
```{r}
library(parallel)
nCores <- detectCores() - 2

dp.ls <- mclapply(split(msin.dt, msin.dt$treatment), function(x){

  sname <- unique(x$treatment)
  message('Running MSStats dataProcess on ', sname)
  
  print(x)
  dp.out <- MSstats::dataProcess(x, 
                                MBimpute =  FALSE, 
                                normalization = FALSE, # turn off for SECMS
                                summaryMethod = "TMP",
                                featureSubset = 'highQuality',
                                remove_uninformative_feature_outlier=T)

  saveRDS(dp.out, ScriptAndDatedFileName(sname,'.mss.data.proc.rds'))
  return(dp.out)

}, mc.cores = nCores)
```

```{r}
print('hello')
```

