---
title: "061625_D3HIV_CCprofiler"
author: "Martin Gordon"
date: "2025-06-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Now we have the D3 data, try run CCprofiler on this sample to see the complexes we recover (and the differential analysis results)
Will start from protein traces object, to allow fairest comparability to the other datasets.

```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)
library(CCprofiler)

source("../../utils/bp_utils/ManageScriptData.R")
```
read in the processed peptide level data and generate. a CCprofiler traces file. Look at the example data workup
```{r}
msin.dt <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061625_D3_MSstatsSmmarization_data/2025_06_16_hiv.d3.peptide.ints.long.csv.gz')
msin.dt[!is.na(intensity) & org == "Human immunodeficiency virus type 1" & decoy != TRUE, .N, by=sample]


msin.dt[!is.na(intensity) & org == "Human immunodeficiency virus type 1", ] %>% 
  .[sample == 'D3-Uninfected']
```
Look at histograms of HIV peptide intensities 
```{r}
g <- ggplot(msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE], aes(x=log2(intensity), fill=sample)) +
  geom_density(alpha=0.6)
g

```


```{r}
g <- ggplot(msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE], aes(x=-log10(qvalue), fill=sample)) +
  geom_density(alpha=0.6) +
  geom_vline(xintercept = -log10(0.01)) +
  coord_cartesian(xlim=c(0,20)) +
  facet_wrap(sample~., scales='free', nrow=3) +
  theme_bw()
g
```
```{r}
g <- ggplot(msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE & pg_specific == 'True'], aes(x=-log10(qvalue), fill=sample)) +
  geom_density(alpha=0.6) +
  geom_vline(xintercept = -log10(0.01)) +
  coord_cartesian(xlim=c(0,20)) +
  facet_wrap(sample~., scales='free', nrow=3) +
  theme_bw()
g

g <- ggplot(msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE & pg_specific == 'True' & qvalue < 0.01, .N, by=.(gene,sample)] , aes(x=sample, y=N, fill=gene)) +
  labs(title='Detected donor 3 viral peptides', subtitle = 'Q-value < 0.01') +
  geom_bar(stat='identity', position='stack') +
  scale_fill_brewer(type='qual') +
  theme_bw()
g
BackupAsPDF(g, 'nviralpeptides.qval0.01.barplot')
```
now look at the intensity and qvalue distributions for these peptides

```{r}
ggplot(msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE & pg_specific == 'True' & qvalue < 0.01, ], aes(x=log2(intensity), fill=sample)) +
  geom_density(alpha=0.6)

ggplot(msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE & pg_specific == 'True' & qvalue < 0.01, ], aes(x=-log10(qvalue), fill=sample)) +
  geom_density(alpha=0.6) +
  coord_cartesian(xlim=c(0,10))


g <- ggplot(msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE & pg_specific == 'True' & qvalue < 0.01, ] , aes(x=fraction,y=log2(intensity),color=gene)) +
  geom_point() +
  geom_line(aes(group=precursor)) +
  labs(title='Detected donor 3 viral peptides', subtitle = 'Q-value < 0.01') +
  scale_color_brewer(type='qual') +
  facet_grid(sample~.) +
  theme_bw()

g

-log10(0.0001)
```

```{r, boxplot, fig.width=10, fig.height=7}
# get N unique peptides 
msin.dt[, peptide := gsub('[.][0-9]{1}$', '', precursor)]


g <- ggplot(msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE & pg_specific == 'True' & qvalue < 0.01, ] , aes(x=as.factor(fraction), y=-log10(qvalue), color=sample)) +
  geom_boxplot(outliers=F) +
  geom_point(data=msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE & pg_specific == 'True' & qvalue < 0.01, 
                          .(length(unique(peptide))) , by=.(fraction, sample)], aes(x=as.factor(fraction), y=V1), color='red') +
  scale_color_brewer(type='qual', palette=2) +
  facet_grid(sample~., scales='free') +
  #coord_cartesian(ylim=c(0,100)) +
  theme_bw()

BackupAsPDF(g, 'qval.distributions.npeptides.overlaid.boxplots')



Reduce(intersect, hiv.ids)
```
find the N runs peptides are identified in 
```{r}
plot(msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE & pg_specific == 'True' & qvalue < 0.01, 
             .N, by=.(precursor)][order(-N)], aes(x=order(precursor, -N), y=N)) +
  geom_point()

ggplot(msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE & pg_specific == 'True' & qvalue < 0.01, 
             .N, by=.(precursor)], aes(x=N)) +
  geom_histogram() +
  geom_vline(xintercept=72, color='red') 

```
```{r}

g <- ggplot(msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE & pg_specific == 'True' & qvalue < 0.01, ] , aes(x=as.factor(fraction), y=-log10(qvalue), color=sample)) +
  geom_boxplot(outliers=F) +
  geom_point(data=msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE & pg_specific == 'True' & qvalue < 0.01, 
                          .(length(unique(peptide))) , by=.(fraction, sample)], aes(x=as.factor(fraction), y=V1), color='red') +
  scale_color_brewer(type='qual', palette=2) +
  facet_grid(sample~., scales='free') +
  #coord_cartesian(ylim=c(0,100)) +
  theme_bw()
```



Look at the msstats file and compare intensities to teh 

```{r}
allFiles <- dir('./data/HIV_Infection_CD4T_cells/Donor-3-061125/D3_Stringent_Spectronaut_Parameters_061225/', full.names = T, include.dirs = T, recursive = T)
allFiles
d3.mss <- fread(allFiles[13])
```

```{r}
msin.dt[sample == 'D3-Uninfected' & protein == 'Q9HCS7' & !is.na(intensity)] %>% 
  .[order(fraction, precursor), .(protein, fraction, precursor, intensity)]


ggplot(msin.dt[!is.na(intensity), length(unique(protein)), by=.(sample, fraction)],  aes(x=fraction, y=V1, fill=sample)) +
  geom_bar(stat='identity') +
  facet_wrap(~sample, ncol=1)


ggplot(msin.dt[sample == 'D3-Uninfected' & !is.na(intensity) & qvalue < 0.01, length(unique(protein)), by=.(sample, fraction)],  aes(x=fraction, y=V1, fill=sample)) +
  geom_bar(stat='identity') +
  facet_wrap(~sample, ncol=1)
```
```{r}
ggplot(d3.mss[!is.na(Intensity), length(unique(ProteinName)), by=.(Condition, Run)],  aes(x=Run, y=V1, fill=Condition)) +
  geom_bar(stat='identity')


d3.mss[!is.na(Intensity) & ProteinName == 'A0A075B6N2'][order(Run)][order(PeptideSequence), .(PeptideSequence, Run, Intensity)] 
```
check the intensity values of the peptides
```{r}
msin.dt[sample == 'D3-Uninfected' & !is.na(intensity) & qvalue <= 0.01 & protein == 'A0A075B6N2'][order(fraction)][order(precursor), .(precursor, fraction, intensity)]


ggplot(msin.dt[!is.na(intensity) & qvalue < 0.01, .N, by=.(sample, org)], aes(x=sample, y=log10(N), color=org)) +
  geom_point()
```

```{r}
hq.msin <- msin.dt[!is.na(intensity) & qvalue < 0.01]

# drop m
hq.msin[!grepl(';', protein), length(unique(precursor)) != 1, by=.(gene, org,sample)][V1 == TRUE][, .N, by=.(sample, org)]
```
```{r}
hq.msin[org != 'Homo sapiens', .N, by=precursor]


hq.msin$sample %>% unique(0)
hiv.ids <- list('uninfected' = hq.msin[sample == 'D3-Uninfected' & org == 'Human immunodeficiency virus type 1', unique(precursor)],
                'infected' =hq.msin[sample == 'D3-Challenged_Infected' & org == 'Human immunodeficiency virus type 1', unique(precursor)],
                'challenged' =hq.msin[sample == 'D3-Challenged_Uninfected' & org == 'Human immunodeficiency virus type 1', unique(precursor)])
                


g <- plot(eulerr::venn(hiv.ids), quantities = T)
g
Backup

hq.msin[sample == 'D3-Uninfected' & org == 'Human immunodeficiency virus type 1']
hq.msin[sample == 'D3-Uninfected' & org == 'Human immunodeficiency virus type 1', unique(prec)]
```

replot the intersected proteins and see how they look

```{r}

allsamples <- Reduce(intersect, hiv.ids)

g <- ggplot(msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE & pg_specific == 'True' & qvalue < 0.01 & precursor %in% allsamples, ] , aes(x=as.factor(fraction), y=-log10(qvalue), color=sample)) +
  geom_boxplot(outliers=F) +
  labs(title='high confidence viral peptides found in all samples') +
  geom_point(data=msin.dt[org == "Human immunodeficiency virus type 1" & !is.na(intensity) & decoy != TRUE & pg_specific == 'True' & qvalue < 0.01 & precursor %in% allsamples, 
                          .(length(unique(peptide))) , by=.(fraction, sample)], aes(x=as.factor(fraction), y=V1), color='red') +
  scale_color_brewer(type='qual', palette=2) +
  facet_grid(sample~., scales='free') +
  #coord_cartesian(ylim=c(0,100)) +
  theme_bw()
g
BackupAsPDF(g, 'qval.distributions.npeptides.overlaid.intersectViralprots.boxplots')
```


Now look at the msstats report

```{r}
p.quant <- fread("./data/HIV_Infection_CD4T_cells/Donor-3-061125/D3_Stringent_Spectronaut_Parameters_061225//D3-Uninfected_061225/20250616_115716_D3_CD4_Uninfecetd_061125_Prot_Report_Recent.tsv")

p.quant <- melt(p.quant, id.vars = c("PG.ProteinGroups", "PG.MolecularWeight"))
setnames(p.quant, old=c('variable', 'value'), new=c('sample', 'intensity'))


# test prtoein
p.quant[!is.na(intensity), .N, by=PG.ProteinGroups]
plot(p.quant[PG.ProteinGroups == 'O00299',]$intensity)


plot(p.quant[PG.ProteinGroups == 'O00299' & !is.na(intensity),]$intensity)
plot(p.quant[PG.ProteinGroups == 'O00299' & !is.na(intensity),]$intensity)
length(p.quant[PG.ProteinGroups == 'O00299' & !is.na(intensity),]$intensity)


```
H
```{r}
plot(d3.mss[!is.na(Intensity) & ProteinName == 'O00299', sum(Intensity), by=.(Condition, Run)]$V1) 
length(d3.mss[!is.na(Intensity) & ProteinName == 'O00299', sum(Intensity), by=.(Condition, Run)]$V1)




d3.mss[!is.na(Intensity) & ProteinName == 'O00299', mean(Intensity), by=.(Condition, Run)][order(Run)]
```

Safest comparison between the files is an all by all protein correlation; just sum the intensities inhte peptide file 

```{r}
library(stringr)

p.quant[, c('fraction', 'fraction2') := as.list(str_extract_all(sample, '[0-9]+', simplify = T)), by=.I]
p.quant <- p.quant[, .(protein=PG.ProteinGroups, fraction, intensity)]
pep.sum <- d3.mss[, .(intensity=sum(Intensity, na.rm=T)), by=.(Run, ProteinName, Condition)][,.(protein=ProteinName, fraction=as.character(Run), sum_intensity=intensity)]

p.quant
pep.sum
p.quant[!is.na(intensity)]

comb.dt <- merge(p.quant[!is.na(intensity)], pep.sum, by=c('protein','fraction'))


g <- ggplot(comb.dt[, cor(intensity, sum_intensity), by=.(protein)], aes(V1)) +
  geom_density(fill='cornflowerblue') +
  theme_bw()

g

as.list(str_extract_all('[1] 1.PG.Quantity', '[0-9]+', simplify = T))
str_extract_all()


```



