---
title: "061725_D123_BPSECMS_pipeline"
author: "Martin Gordon"
date: "2025-06-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Analysis of Donor1-3 HIV data; first load in the protein quant data, look at standard QC plots etc first to see sample agreement, then perform the PPI detection and differential analysis
```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)
library(circlize)
library(stringr)
library(viridis)
library(pracma)
library(magick)
library(eulerr)

source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/enrichmentTestFunctions.R")
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")
source("../../utils/bp_utils/UniprotIDMapping.R") # map hu to mm 

customTheme <-  theme_bw() +
  theme(axis.text.x = element_text(angle=90),
        panel.border = element_rect(color = "lightgrey",
                                    fill = NA,
                                    size = 1))
```

mw standards 
same standards for all 3 files
```{r}
mw.standards <- fread('./data/HIV_Infection_CD4T_cells/SEC_Profiles/D1-D3-cal_SRT.txt')
setnames(mw.standards, new=c('fraction', 'mw'))
# want the standards in daltons (right?)
mw.standards[, mw := mw*1000]
mc <- calculateFractionMassConverters(mw.standards)
```

read in the cleaned/normalized interpolated intensity values for D1 and D2
```{r}
# this is the second set of SEC data; now process the donor3 data
sec.long <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_HIV_CD4T_TwoDonors_data/2025_05_06_sec.long.donor1donor2.normalized.interpolated.scaled.csv.gz')

allFiles <- dir(path='./data/HIV_Infection_CD4T_cells/Spectronaut_Reports_061625', full.names = T, recursive = T, pattern="*Report.tsv")
allFiles <- grep('Cell_Lysis', invert=T, allFiles, value=T)
```
Read in all results from file and process in the same way

```{r}
sec.ls <- lapply(grep('Protein_Report.tsv', allFiles, value=T), fread)
names(sec.ls) <- unlist(tstrsplit(grep('Protein_Report.tsv', allFiles, value=T), '/', 4)[6])

sec.ls <- lapply(sec.ls, function(x) {dt <- reshape2::melt(x, id.vars=c('PG.MolecularWeight', 'PG.ProteinGroups')); setnames(dt, old=c('value'), new='intensity'); return(setDT(dt))})
```
tidy the datatables; check the fractions match what we expect

```{r}
lapply(sec.ls, function(x){
  x[, c('fraction', 'test') := as.list(str_extract_all(variable, '[0-9]+', simplify = T)), by=.I]
})


lapply(sec.ls, function(x){
  sdcols <- c('fraction', 'test')
  x[, (sdcols) := lapply(.SD, as.numeric), .SDcols=sdcols]
})

# all looks good
lapply(sec.ls, function(x){
  x[fraction != test,]
  x[, all(order(unique(fraction)) == seq(1,72,1))]
  })

# drop unnecessary cols 
lapply(sec.ls, function(x){
  x[, c('variable', 'test') := NULL]
  setnames(x, old=c('PG.MolecularWeight', 'PG.ProteinGroups'), new=c('mw', 'protein'))
  return(x)
})

sec.dt <- rbindlist(sec.ls, idcol='sample')
sec.dt[, c('replicate', 'treatment') := tstrsplit(sample, '-', keep=c(1,2))]
sec.dt[, replicate := as.numeric(sub('[A-Z]', '', replicate))]

#fwrite(sec.dt, ScriptAndDatedFileName('hiv.donor123.rawinput.csv.gz'))
sec.dt <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_hiv.donor123.rawinput.csv.gz')
nrow(sec.dt)
```
add gene, mw, org annotation
```{r}
idmapper <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061625_D3_MSstatsSmmarization_data/2025_06_16_hiv.d3.peptide.ints.long.csv.gz')[, .(gene, protein,mw, org)] %>% 
  unique()

sec.dt <- merge(sec.dt[, -c('mw')], idmapper, by='protein', all.x=T)

col.order <- c("sample", "treatment", "replicate", "fraction", "protein", "gene", "mw", "org", "intensity")
#fwrite(sec.dt[, ..col.order], ScriptAndDatedFileName('hiv.donor123.rawinput.csv.gz'))

sec.dt <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_hiv.donor123.rawinput.csv.gz')
sec.dt[, sample := factor(sample, levels=c("D1-Uninfected", "D1-Challenged_Infected", "D1-Challenged_Uninfected",
                                           "D2-Uninfected", "D2-Challenged_Uninfected", "D2-Challenged_Infected",
                                           "D3-Uninfected", "D3-Challenged_Uninfected", "D3-Challenged_Infected"))]
```

elutions look quite different? much more proteins detected in the 2nd set than set3.. also the curves look pretty different
```{r nproteins, fig.width=8, fig.height=6}
qc.dt <- qcSummaryTable(sec.dt)

# number of protiens eluting looks good
BackupAsPDF(qcPlotProteinCount(qc.dt), 'Nproteins.linechart') +
  theme_classic()
```
Interesting the shapes are also subtly different instead of just y-axis shift (intensity difference) Need to fit a different curve to each sample I think?
```{r medpolish, fig.width=8, fig.height=6}
g <- qcPlotMedianInt(qc.dt) +
  theme_bw()
g
BackupAsPDF(g, 'medianPolish.linechart')
```
Donor 1 uninfected failed?
Also viral protein profiles are quite different...
```{r, heatmaps, fig.width=20, fig.height=6}
sec.dt[, interpolated := FALSE]

View(scaledIntensityMatrices)
intMats <- scaledIntensityMatrices(sec.dt, useInterpolated = FALSE)

BackupAsPDF(intensityHeatmaps(intMats, 
                              border=T), 'raw.ints.heatmap')

# and viral only 8 detectedand v different profile
BackupAsPDF(intensityHeatmaps(lapply(intMats, function(x){ return(x[rownames(x) %in% viral.prots,])}), 
                              border=T), 'viral.raw.ints.heatmap')
```

Fit loess to each sample (fit an interaction model as there is a clear donor effect visible here between 2 & 3)

```{r, detect-outliers}
allFits <- fitLocalCubics(qc.dt, window = 15, sampleTerm = "interaction")
labelOutliers(qc.dt, allFits, threshold = 2) 

p <-plotNormAndOutlierFits(qc.dt , allFits)
BackupAsPDF(p, 'localRegression.window15.linechart')
```
```{r, normalize-by-residuals}
normalizeByResiduals(secLong.dt = sec.dt, qc.dt)
#fwrite(sec.dt, ScriptAndDatedFileName('hiv.donor123.rawinput.csv.gz'))
sec.dt <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_hiv.donor123.rawinput.csv.gz')
sec.dt[, interpolated := NULL]
```
Visualize the new QC results
Most fractions look pretty good; I still think we can do a quick imputation step to tidy up a couple of fraction in `Challenged_Uninfected_3`, but otherwise things look pretty good
```{r norm-qc}
norm.qc.dt <- qcSummaryTable(sec.dt)

g <- qcPlotMedianInt(norm.qc.dt) +
  theme_bw()
g
BackupAsPDF(g, 'normalized.qc')
```

Normalization looks great! A couple of problematic fractions, but fall outside the informative windows
Turn on interpolation anyway and save the values for (possibly) plotting later...
```{r}
sec.long <- interpolateMissingAndOutlierFractions(sec.dt, qc.dt, fractions = seq(1,72,1), maxGap = 2)
#fwrite(sec.long, ScriptAndDatedFileName('sec.long.normalized.interpolated.csv.gz'))
sec.long <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_sec.long.normalized.interpolated.csv.gz')
```
plot the normalized set; looks to have cleaned issues in later fraction 
```{r nproteins, fig.width=8, fig.height=6}
new.qc.dt <- qcSummaryTable(sec.long)

g <- qcPlotMedianInt(new.qc.dt) +
  theme_bw()
g
BackupAsPDF(g, 'normalized.interpolated.qc')

# number of protiens eluting looks good
BackupAsPDF(qcPlotProteinCount(new.qc.dt), 'Nproteins.interpolated.linechart') +
  theme_classic()
```
The two heatmaps to compare
I think the reason the interpolation 'half works' as these are not flagged as outlier samples based on ints alone, so may need to adjust this for the interpolation function to inlcude fractions with low N proteins relative to neighbours
```{r, heatmaps, fig.width=20, fig.height=6}
intMats <- scaledIntensityMatrices(sec.long, useInterpolated = TRUE)

BackupAsPDF(intensityHeatmaps(intMats, 
                              border=T), 'norm.interpolated.prots.heatmap')

intMats <- scaledIntensityMatrices(sec.long, useInterpolated = FALSE)

BackupAsPDF(intensityHeatmaps(intMats, 
                              border=T), 'norm.prots.heatmap')
```
Remove proteins with < 5 consecutvie measurements
Do this after interpolation on expanded but non-interpolated data to ensure the consecutvie meaurements are consecutive functions...
```{r}
# want to run this on the originalOIntensity as dont want to consider mssing 
sec.long[, ori.intensity := originalIntensity]
sec.long[is.na(ori.intensity), ori.intensity := 0]


sec.long  <- MaxConsecutiveDetections(secLong.dt = sec.long, idcol = 'protein', intsCol = 'ori.intensity', detectionCutoff=0, plot=T)
sec.long[, passDetections := FALSE]
sec.long[consecutiveDetections >=5, passDetections := TRUE]

g <- ggplot(sec.long[, length(unique(protein)), by=.(sample,passDetections)], aes(x=sample, y=V1, fill=passDetections)) +
  geom_bar(stat='identity') +
  theme_bw() +
  scale_fill_brewer(type='qual', palette=3) +
  guides(fill=guide_legend('> 5 consecutive fraction detections')) +
  theme_classic() +
  theme(axis.text.x = element_text(angle=90))
g
BackupAsPDF(g, 'nProteinsConsecDetections.barplot')
```
Remove the filtered proteins and save the results. This will be our starting set for the analysis
```{r}
# keep passing proteins in sec.long 
fwrite(sec.long[passDetections == TRUE,], ScriptAndDatedFileName('sec.long.normalized.interpolated.filtered.csv.gz'))
```

```{r}
sec.long <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_sec.long.normalized.interpolated.filtered.csv.gz')
```


## QC plots of clean data

We basically want to answer how similar/different are the samples
Need to make a call on the whether a third donor is needed; look at pairwise sample correlations

First scale our intensity. use the interpolated values for now, wont really impact results as few fractions impacted
```{r}
scaleByTotalIntensity(sec.long)
scaleByMaxIntensity(sec.long)

corMat <- qcFullSampleCorrelation(sec.long)
corMat
#' plot correlation matrices
#' taken from https://jokergoo.github.io/ComplexHeatmap-reference/book/a-single-heatmap.html#customize-the-heatmap-body

plotSampleCorrelationHeatmap <- function(cor.mat=cor.mat,...){
  
  #enforce same row and column ordering
  od =  hclust(dist(cor.mat))$order
  cm = cor.mat[od, od]

  hm <- Heatmap(cm,
                name='Sample Pearson Corr.',
                rect_gp = gpar(type = "none"), 
                cluster_rows = F, 
                cluster_columns = F,
                column_names_gp = gpar(fontsize=8),
                row_names_gp = gpar(fontsize=8),
	              cell_fun = function(j, i, x, y, w, h, fill) {
		            if(i >= j) {
		              grid.rect(x, y, w, h, gp = gpar(fill = fill, col = fill))
		              grid.text(sprintf("%.2f", cm[i, j]), x, y, gp = gpar(fontsize = 6, col='white'))
		              }
	             },
		            ...)
  draw(hm)
}
BackupAsPDF(plotSampleCorrelationHeatmap(corMat, col=viridis(10)), 'sampleFraction.correlations.heatmap')
```
basically looks at a sliding window to see if fractions become more similar, but not the case
Take away is one of the D1 samples has failed  (challenged uninfected) and D2 and D3 are very different expression patterns 

start comparisons from fraction 20 due to the difficulties with earlier fractions
Not promising; clearly huge donor effect v little correlation between donors 2&3
```{r, cross-corr-barplot, fig.width=12, fig.height=10}
cc.dt <- crossCorrelationFullSample(sec.long, startFraction = 20, endFraction = 60, radius=5)

sampleOrdr <- c('D1-Challenged_Infected','D1-Challenged_Uninfected', 'D1-Uninfected',
                'D2-Challenged_Infected','D2-Challenged_Uninfected', 'D2-Uninfected',
                'D3-Challenged_Infected', 'D3-Challenged_Uninfected', 'D3-Uninfected')

cc.dt[, refSample_fct := factor(refSample, levels=sampleOrdr)]
cc.dt[, otherSample_fct := factor(otherSample, levels=sampleOrdr)]
cc.dt
g <- ggplot(cc.dt, aes(y=pearsonR, x=offset, fill=pearsonR)) +
  geom_bar(stat='identity') +
  facet_grid(refSample_fct~otherSample_fct) +
  scale_fill_viridis_c() +
  theme_bw() +
  theme(strip.text=element_text(size=6)) 

g
BackupAsPDF(g, 'crossCor.barplot')
```
before we do this, lets tidy the sample names

```{r}
sec.long[treatment == 'Challenged_Infected', treatment:= 'Infected']
sec.long[treatment == 'Challenged_Uninfected', treatment := 'Challenged']
sec.long[, sample.old := sample]
sec.long[, sample := paste0(treatment, '_', replicate)]

```

```{r}
full.cor.dt <- qcFractionByFractionCorrelation(sec.long)
```

```{r, fig.width=9, fig.height=5}
g <- qcFractionCorrelationLinePlot(full.cor.dt,
                                   splitByReplicate=T, 
                                   colorFacetStrips = T)
BackupAsPDF(function()grid.draw(g), 'fractionCorrelation.lineplot.splitReps.pdf')

```
I find the collapsed version too confusing, use the split version
```{r, fig.width=6, fig.height=5}
g <- qcFractionCorrelationLinePlot(full.cor.dt,
                                   splitByReplicate=F, 
                                   colorFacetStrips = T)
BackupAsPDF(function()grid.draw(g), 'fractionCorrelation.lineplot.splitReps.collapsed.pdf')
```

Color by donor?

```{r, fig.width=7, fig.height=5}
full.cor.dt[, ref.treatment.old := ref.treatment]
full.cor.dt[, other.treatment.old := other.treatment]
full.cor.dt[, ref.treatment := ref.replicate]
full.cor.dt[, other.treatment := other.replicate]

g <- qcFractionCorrelationLinePlot(full.cor.dt,
                                   splitByReplicate=F, 
                                   colorFacetStrips = T)

BackupAsPDF(function()grid.draw(g), 'fractionCorrelation.donorColor.lineplot.splitReps.collapsed.pdf')
```
# peak detection
```{r}

# if we normalized, we have to catch the post-normalization matrices
normInt.mats <- scaledIntensityMatrices(sec.dt, useInterpolated = TRUE)
peakTables <- lapply(normInt.mats, goodPeaksTableFromIntensityMatrix, minCV = -Inf)

allPeaks <- rbindlist(peakTables, idcol = "sample")
allPeaks[, c('treatment', 'replicate') := tstrsplit(sample, '_', keep=c(1,2))]
```

```{r, fig.width = 8, fig.height = 7}
# udpate height for more samples
p <- ggplot(allPeaks, aes(x= peakLocation, fill = goodPeak)) + 
  geom_bar() + 
  facet_grid(treatment~replicate) +
  theme_bw()
p
BackupAsPDF(p, "peakCounts.perfraction.barplot")
```

# fraction adjustments based on peaks
This will update the peakTables sec.dt with new standardized fraction numbers
```{r}
# I don't trust hte alignment below about fraction 20, so update `startFitAtPeak` to skip it
standardizeAllPeakTablesToStandard(peakTables, sec.dt, standardIdx = "Uninfected_3", fitPortion = 0.65, minPeaksPerFraction = 45, startFitAtPeak = 20)
sec.dt[, standardFraction.global := standardFraction]
```

Donors seem too different to align. lets try another way
Split the DT into, donor lists and process seperately. Then combine the output together

```{r}
peakTables.ls <- lapply(seq(1,3,1), function(x){
  
  mats.subset <- normInt.mats[grepl(paste0(x,'$'), names(normInt.mats))]
  print(names(mats.subset))
  peakTables <- lapply(mats.subset, goodPeaksTableFromIntensityMatrix, minCV = -Inf)
  return(peakTables)
})

# now align peaks within each donor
sec.long <- lapply(seq(1,3,1), function(x){
  subdt <- copy(sec.dt[replicate == x,])
  standardizeAllPeakTablesToStandard(peakTables.ls[[x]], subdt, standardIdx = 3, fitPortion = 0.65, minPeaksPerFraction = 45, startFitAtPeak = 20)
  return(subdt)
}) %>% rbindlist(.)
```
now we have standard fractions and standard fractions global. Lets use the local alignment and interpolate intensities, we can then use these these values fro plotting

```{r}
sec.dt[, intensity_interpolated := approx(standardFraction, intensity, xout = fraction)$y, by = .(sample, protein)]
#fwrite(sec.dt, ScriptAndDatedFileName('hiv.d123.processed.ints.csv.gz'))
sec.dt <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_hiv.d123.processed.ints.csv.gz')
```

Todo:
filter protein peaks using MW
Detect PW PPI 
Perform differential analysis
Upload the new data to the app

```{r}
#standards <- loadStandardsFractionToMW ("~/Downloads/cal_biosep.txt")
mc <- calculateFractionMassConverters(standards)

# label fractions by mass, and fraction/proteins by mass ratio, aka log2MassNumber
sec.dt[, fractionMass := mc$fraction2Mass(standardFraction)]
sec.dt[, log2MassNumber := log2(fractionMass/as.numeric(mw))]


# pay attention to which cofmX.XXXX is used below: cofmN.standardized
scorePeakByMassShift <- function(peak.dt, mw){
  # the 150% fraction
  peak.dt[sec.dt, fiftyPCFraction := mc$mass2Fraction(1.2 * as.numeric(mw)), on = "protein"]
  peak.dt[!is.na(fiftyPCFraction), goodPeak.rm.monomers := goodPeak & cofmN.standardized < fiftyPCFraction]
  peak.dt
}

# collapse to one list (1 level unnesting)
peakTables <- do.call(c,peakTables.ls)

purrr::walk(peakTables, scorePeakByMassShift, mw)

allPeaks <- rbindlist(peakTables, idcol='sample')
allPeaks[is.na(goodPeak.rm.monomers), goodPeak.rm.monomers := FALSE]
#fwrite(allPeaks, ScriptAndDatedFileName('hiv.d123.detectedPeaks.csv.gz'))
```
Actually the filtration looks pretty good 
```{r}
allPeaks <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_hiv.d123.detectedPeaks.csv.gz')
allPeaks[, c('treatment', 'replicate') := tstrsplit(sample, '_', keep=c(1,2))]
# udpate height for more samples
p <- ggplot(allPeaks, aes(x= peakLocation, fill = goodPeak.rm.monomers)) + 
  geom_bar() + 
  facet_grid(treatment~replicate) +
  theme_bw()
p
BackupAsPDF(p, "peakCounts.filtered.perfraction.barplot")
```

```{r prepare-gs}
# id mapper for protein -> string ID
huString.mapper <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_protein.string.idMapper.csv.gz')

#need to merge our protein ids with 
gs.interactors <- goldStandardPairs(huString.mapper[protein %in% sec.long$protein, unique(stringID)], 
                                    corum.path = NULL,
                                    string.links.path = "~/Documents/utils/mg_utils/data/stringDB/9606.protein.physical.links.detailed.v12.0.txt.gz",
                                    string.info.path =  NULL,
                                    stringCombinedScoreThreshold = 400)

gs.decoys <- decoysFromString(huString.mapper[protein %in% sec.long$protein, unique(stringID)], 
                              links.path = "~/Documents/utils/mg_utils/data/stringDB/9606.protein.physical.links.detailed.v12.0.txt.gz",
                              info.path =  NULL)
gs.decoys[, source := 'string']

gs.comb.interactors <- rbind(gs.interactors, gs.decoys)
```
## peak-peak correlations
```{r, peak-correlations}
# just ensure a refresh
intMats <- scaledIntensityMatrices(sec.dt, useInterpolated = FALSE)

View(windowedCorrelation)
peakMatrices <- purrr::map2(intMats, peakTables, goodPeaksMatFromPeaksTable)
corPeaks.ls <- purrr::map2(intMats, peakMatrices, windowedCorrelation, outerRadius = 6)

allPeakCorrs <- rbindlist(purrr::map(corPeaks.ls, "cor"), idcol = "sample")
```

```{r annotate-peakDT}
# read in datatables
allPeakCorrs[huString.mapper, gene1 := i.stringID, on = c(protein1 = "protein")]
allPeakCorrs[huString.mapper, gene2 := i.stringID, on = c(protein2 = "protein")]
allPeakCorrs[idmapper, alias1 := i.gene, on = c(protein1 = "protein")]
allPeakCorrs[idmapper, alias2 := i.gene, on = c(protein2 = "protein")]


# for now use protein as gene for the NA gene
allPeakCorrs[is.na(gene1), gene1 := protein1]
allPeakCorrs[is.na(gene2), gene2 := protein2]

allPeakCorrs[is.na(alias1), alias1 := protein1]
allPeakCorrs[is.na(alias2), alias2 := protein2]
#fwrite(allPeakCorrs, ScriptAndDatedFileName('secms.hiv.d123.allByallPeakCorr.csv.gz'))
allPeakCorrs <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_secms.hiv.d123.allByallPeakCorr.csv.gz')
```

The scores seem to do a good job of distinguishing truth set vs decoys
```{r,plot-correlations, fig.width=9,fig.height=6}
g <-  ggplot(allPeakCorrs[gene1 < gene2 & gs != 'unknown'], aes(x = sample, fill = gs)) +
    geom_bar() + 
    labs(title=paste0('Breakdown of detected interactors (R >= .9)')) +
    theme_bw() + 
    scale_fill_manual(values=c('decoy'='grey70', 'interactor'='gold'))
g
BackupAsPDF(g, 'interactorSet.breakdown.barplot')

g <- ggplot(allPeakCorrs[gene1 < gene2 & pearsonR >= 0.99 & gs != 'unknown'], aes(x = sample, fill = gs)) +
    geom_bar() + 
    labs(title=('Breakdown of detected interactors'), subtitle='pearson R > 0.99') +
    theme_bw() + 
    scale_fill_manual(values=c('decoy'='grey70', 'interactor'='gold'))
g
BackupAsPDF(g, 'interactorSet.breakdown.r99.barplot')
```
```{r}
allGenes <- huString.mapper$stringID %>% 
  unique()
denomInteractor <- nrow(gs.interactors[gene1 %in% allGenes & gene2 %in% allGenes])
denomDecoy <- nrow(gs.decoys[gene1 %in% allGenes & gene2 %in% allGenes])
```
Using log-likihood ratio approach; how more likey is the PPI derived from the truth set than the random set
Looks liek a lot of decoys also correlate with
```{r ppi-scoring}
scoreByGS(allPeakCorrs, denomDecoy, denomInteractor, column = "corScore", groupByVariable = c("sample", "start"))
```
plot the LLR to see how; distance between lines is what is being scored
```{r}
g <- ggplot(allPeakCorrs, aes(x = corScore, y = log10RateRatioID, color = start)) + geom_hline(yintercept = 0.0, lty = "dotted") + 
    geom_line(aes(group = start)) + 
    facet_wrap(~sample, ncol=3) + 
    scale_color_viridis_c(option = "H", name = "fraction") + 
    theme_bw()
g
BackupAsPDF(g, 'LLR.samples.linechart')


allPeakCorrs[corScore > 10]

ggplot(sec.long[sample == 'Infected_3' & protein %in%c('P04233','P01889'),], aes(x=fraction, y=intensity_totalScaled, color=protein)) +
  geom_point() +
  geom_line(aes(group=protein))

# Im super confident a lot of duplicate rows made it into INfected_3, ... that is an insanely high value of perfect cor scores
g <- ggplot(allPeakCorrs[pearsonR == 1, .N, by=sample], aes(x=sample, y=log2(N))) +
  geom_bar(stat='identity') +
  labs(title='Number of PPI with perfect correlation scores')
BackupAsPDF(g, 'N.perfectCor.scores.pdf')
```
Just looking at the elution profiles, I am not sure if I trust anything before fraction 20 in the allpeakCors set for infected_3, 
I think these might just be missing fractions surround by a single measurment
histogram clearly highlights this; drop these measurements
```{r}
ggplot(allPeakCorrs[pearsonR == 1], aes(x=prot1Peak)) +
  geom_histogram(bins=72) +
  facet_wrap(~sample)
```
I think looking at this we want to use non-interpolated values for peak finding, but proceed as is for now as this will be researched anyway..
```{r}
filteredPeakCorrs <- allPeakCorrs[!(sample == 'Infected_3' & (prot1Peak < 20)| prot2Peak < 20),]

g <- ggplot(allPeakCorrs, aes(x = corScore, y = log10RateRatioID, color = start)) + geom_hline(yintercept = 0.0, lty = "dotted") + 
    geom_line(aes(group = start)) + 
    facet_wrap(~sample, ncol=3) + 
    scale_color_viridis_c(option = "H", name = "fraction") + 
    theme_bw()
g
```
save the ppi best scores to file 
repeat peak finding and analysis w/o the using the interpolated values and see if this removes the FP set

```{r}
stoichDevThreshold <- log2(1 + 0.2)

setorder(allPeakCorrs, sample, -log10RateRatioID )

summary <- filteredPeakCorrs[, .SD[1], by= .(sample, protein1, protein2, gene1, gene2, prot1Peak, prot2Peak, gs) # best correlation per pair of peaks
                          ][, .(sumLLRatio = sum(log10RateRatioID[abs(log2Ratio - log2Ratio[1]) < stoichDevThreshold]), # only sum peaks that are within 20% of the first
                                numPeaks = .N,
                                numSameStoichPeaks  = sum (abs(log2Ratio - log2Ratio[1]) < stoichDevThreshold)),
                            by = .(sample, protein1, protein2, gene1, gene2, gs)
                            ]
setorder(summary, sample, -sumLLRatio)
fwrite (summary[gene1 < gene2], ScriptAndDatedFileName("hiv.d123.peaksimilarities.csv.gz"))
```

## cosine similarity
TODO....


## Differential peak analysis

read files
```{r}
# all filtered peaks 
allPeaks <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_hiv.d123.detectedPeaks.csv.gz')
allPeaks <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_hiv.d123.detectedPeaks.csv.gz')

fwrite(allPeaks, ScriptAndDatedFileName('detectedPeaks.csv'))
# dont consider monomeric peaks in th eset
allPeaks[,goodPeak := goodPeak.rm.monomers]

sec.dt <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_hiv.d123.processed.ints.csv.gz')
```

Now, we want to split the datasets into two donors, run seperately and run each group pairwise. Inspect the results in the app
```{r}
allTreatments <- sec.long$treatment %>% 
  unique()


allByAll <- function(items){
  data.table(x = items)[, .(y = items), by = x][]
}

pair.mat <- allByAll(allTreatments)[x < y] %>% 
  as.matrix()

allPeaks[,treatment := tstrsplit(sample, "_", keep=1)]


contrastTreatmentSubset <- function (treatments, sec.dt, allPeaks){
  print (treatments)
  allPeaks <- allPeaks[treatment %in% treatments]
  allPeaks[, proteinPeakCluster := NULL]
  # per protein, join peaks across runs into clusters
  allPeaks[goodPeak == TRUE, proteinPeakCluster := clusterPeaks(cofmN.standardized, maxDistance = 2.5), by = protein]

  # per peak cluster (per protein), summarize to a center (and height etc, but most important is center)
  peakClusters <- allPeaks[!is.na(proteinPeakCluster), 
                           .(.N, meanPeakHeight  = mean(peakHeight), center = mean(cofmN.standardized)),
                           keyby = .(protein, proteinPeakCluster)]
  
  # split data into single-protein chunks
  # peaks in the protein
  peakClusters.subTables <- split(peakClusters, by = "protein")
  # sec data per protein (only those in above table)
  allProteins <- names(peakClusters.subTables)
  sec.subTables <- split(sec.dt[treatment %in% treatments][protein %in% allProteins], by = "protein" )
  
  # iterate over proteins
  names(allProteins) <- allProteins
  anova.ls <- pbapply::pblapply(allProteins, function(proteinOI) anovaPeaksInOneProtein(sec.subTables[[proteinOI]], peakClusters.subTables[[proteinOI]], radius = 5))
  
  # format output
  anova.dt <- rbindlist(anova.ls, use.names = TRUE, fill = TRUE, idcol = "protein")
  print(anova.dt)
  # friendly names
  setnames(anova.dt,
           old = c( "rn",        "Sum.Sq",     "Mean.Sq",    "F.value", "Pr..F."),
           new = c( "modelTerm", "SumSquares", "MeanSquare", "F",       "p.F"))
  
  
  peakClusters[anova.dt[modelTerm == "poly(standardFraction, 4):treatment"], treatmentDiff.p := i.p.F]
  peakClusters[anova.dt[modelTerm == "treatment"], treatmentIntensity.p := i.p.F]

  setorder(peakClusters, treatmentDiff.p, na.last = TRUE)
  
  peakClusters[, gene := multiUniprots2multiGenes(protein)]
  setcolorder(peakClusters, c("protein", "gene"))
  
  return (peakClusters)
}
```

now, run the analysis within each donor
```{r}
d1_pairwiseContrasts <- apply(pair.mat, 1, contrastTreatmentSubset, sec.dt[replicate == 1,], allPeaks[peakHeight > 0.01])
d2_pairwiseContrasts <- apply(pair.mat, 1, contrastTreatmentSubset, sec.dt[replicate == 2,], allPeaks[peakHeight > 0.01])
d3_pairwiseContrasts <- apply(pair.mat, 1, contrastTreatmentSubset, sec.dt[replicate == 3,], allPeaks[peakHeight > 0.01])

names(d1_pairwiseContrasts) <- apply(pair.mat, 1, function(x) paste0(paste0(x, collapse = "_"), '_1'))
names(d2_pairwiseContrasts) <- apply(pair.mat, 1, function(x) paste0(paste0(x, collapse = "_"), '_2'))
names(d3_pairwiseContrasts) <- apply(pair.mat, 1, function(x) paste0(paste0(x, collapse = "_"), '_3'))

allContrasts <- rbindlist(c(d1_pairwiseContrasts, d2_pairwiseContrasts, d3_pairwiseContrasts),idcol = "contrast")
allContrasts[, treatmentDiff.padj := p.adjust(treatmentDiff.p, method='BH')]
allContrasts[, treatmentIntensity.p := p.adjust(treatmentIntensity.p, method='BH')]

fwrite(allContrasts, ScriptAndDatedFileName('hiv.d123.pwContrasts.csv.gz'))
fwrite(allContrasts, ScriptAndDatedFileName('hiv.d123.pwContrasts.csv'))



allContrasts.filtered <- allContrasts[center >= 20, ]
allContrasts.filtered[, treatmentDiff.padj := p.adjust(treatmentDiff.p, method='BH')]
allContrasts.filtered[, treatmentIntensity.p := p.adjust(treatmentIntensity.p, method='BH')]

fwrite(allContrasts.filtered, ScriptAndDatedFileName('hiv.d123.pwContrasts.rmfrac20peaks.csv.gz'))
```
peak similarities; annotate with protein names and share

```{r}

idmapper <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061625_D3_MSstatsSmmarization_data/2025_06_16_hiv.d3.peptide.ints.long.csv.gz')[, .(gene, protein,mw, org)] %>% 
  unique()
peakCor <- fread('./061725_D123_BPSECMS_pipeline_data/2025_06_18_hiv.d123.peaksimilarities.csv.gz')
peakCor <- merge(peakCor, idmapper[, .(protein1=protein, alias1=gene)], by='protein1', all.x=T)
peakCor <- merge(peakCor, idmapper[, .(protein2=protein, alias2=gene)], by='protein2', all.x=T)

peakCor <- peakCor[, .(sample, protein1, protein2, alias1, alias2, gene1, gene2, gs, sumLLRatio, numPeaks, numSameStoichPeaks)]
fwrite(peakCor, ScriptAndDatedFileName('hiv.d123.proteinCorrelations.csv'))
```


