---
title: "062025_HIV.d123.ccprofiler"
author: "Martin Gordon"
date: "2025-06-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Overview

Run `CCprofiler` differential expression on the HIV dataset,so we have a set of PPI for Ben to compare
We will start with the protein data in long format for direct comparability to BP pipeline
Create a traces object from the data to run the CCprofiler pipeline

```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)
library(circlize)
library(stringr)
library(viridis)
library(pracma)
library(magick)
library(eulerr)
library(CCprofiler)

source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/enrichmentTestFunctions.R")
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")
source("../../utils/bp_utils/UniprotIDMapping.R") # map hu to mm 

customTheme <-  theme_bw() +
  theme(axis.text.x = element_text(angle=90),
        panel.border = element_rect(color = "lightgrey",
                                    fill = NA,
                                    size = 1))
```

read in the data in sec.long format
```{r}
mw.standards <- fread('./data/HIV_Infection_CD4T_cells/SEC_Profiles/D1-D3-cal_SRT.txt')
setnames(mw.standards, new=c('fraction', 'mw'))
# want the standards in daltons (right?)
mw.standards[, mw := mw*1000]
mc <- calculateFractionMassConverters(mw.standards)
```

```{r}
sec.long <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_sec.long.normalized.interpolated.filtered.csv.gz')
sec.long[treatment == 'Challenged_Infected', treatment:= 'Infected']
sec.long[treatment == 'Challenged_Uninfected', treatment := 'Challenged']
sec.long[, sample.old := sample]
sec.long[, sample := paste0(treatment, '_', replicate)]
```

```{r}
idmapper <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061625_D3_MSstatsSmmarization_data/2025_06_16_hiv.d3.peptide.ints.long.csv.gz')[, .(gene, protein,mw, org)] %>% 
  unique()
```
example protein traces; use this to perform the PPI comparison
```{r}
example.prot.traces <- readRDS('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/050925_CCprofiler_testRun_data/2025_05_12_protTraces.example.rds')
example.prot.traces$trace_annotation
```
function to prepare trace annotation
```{r}
prepareTracesProteinAnnotation <- function(uniprot_path='~/Documents/utils/mg_utils/data/uniprotkb_reviewed_true_2025_05_13.tsv.gz'){
  
  if (file.exists(uniprot_path)){
    uniprot.txt <- fread(uniprot_path)
  }
  # tidy col names 
  uniprot.txt <- uniprot.txt[Organism == 'Homo sapiens (Human)', .(id=Entry, protein_id=Entry, Entry_name=`Entry Name`, 
                                                                   Status='reviewed', Protein_names=`Protein names`,  
                                                                   Gene_names=`Gene Names`, Organism, Length, Mass, GO_ID=`Gene Ontology IDs`, protein_mw=(Mass/1000), Decoy=0)]
  return(uniprot.txt)
}
prepareTracesProteinAnnotation('~/Documents/utils/mg_utils/data/uniprotkb_reviewed_true_2025_05_13.tsv.gz')
```

```{r}
example.prot.traces$traces

intsMats <- scaledIntensityMatrices(sec.long, useInterpolated = FALSE, scaleDenom='none')


prepareTracesIntsMatrix <- function(mat){
  
  nfracts <- length(colnames(mat))
  
  idcol <- rownames(mat)
  mat <- cbind(mat, rownames(mat))
  colnames(mat)[length(colnames(mat))] <- 'id'
  rownames(mat) <- seq(1,nrow(mat),1)
  return(mat)
}


test <- intsMats[[1]]
test$id <- rownames(test)

prepareTracesIntsMatrix(intsMats[[1]])
```

we can make it directly from sec.long... use this 
```{r}

#' Just extract each sample seperately 
prepareTracesIntensity <- function(sec.dt, splitCol='sample', intsCol='intensity', useInterpolated=FALSE){
  
  dt <- copy(sec.dt)
  
  if (useInterpolated == FALSE){
    dt[interpolated == TRUE, eval(intsCol) := NA]
  } else {
    message('Warning: including interpolated values.\nIf you wish to disable, use the `useInterpolated=FALSE`')
  }
  
  ints.dt <- dcast(dt, protein~fraction, value.var = eval(intsCol))
  # convert NA to 0
  ints.long <- setDT(reshape2::melt(ints.dt, id.vars='protein'))
  ints.long[is.na(value), value := 0]
  
  ints.dt <- dcast(ints.long, protein~variable, value.vars='value')
  ints.dt[, id := protein]
  ints.dt[, protein := NULL]
  
  # check fraction ordering
  if(all(colnames(ints.dt) == c(seq(1, max(dt$fraction),1), 'id')) != TRUE)
    stop('Missing fractions in input datatable...\nPlease handle these missing fractions before running.\nExiting...')
  
  return(ints.dt)
}
ints.dt <- prepareTracesIntensity(sec.dt = sec.long[sample=='Infected_3',], useInterpolated = FALSE)
```

```{r}
example.prot.traces$fraction_annotation
```
```{r}
mw.standards <- fread('./data/HIV_Infection_CD4T_cells/SEC_Profiles/D1-D3-cal_SRT.txt')
setnames(mw.standards, new=c('fraction', 'mw'))
# want the standards in daltons (right?)
mw.standards[, mw := mw*1000]
mc <- calculateFractionMassConverters(mw.standards)
```

```{r}
# use BPs function to annotate the proteins correctly 
prepareTracesFractionAnnotation <-  function(sec.long, mw_path='/Users/martingordon/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/data/HIV_Infection_CD4T_cells/SEC_Profiles/D1-D3-cal_SRT.txt'){
  
  dt <- copy(sec.long)
  
  if (file.exists(mw_path)){
    mw <- fread(mw_path)
    print(mw)
  }
  setnames(mw, new=c('fraction', 'mw'))
  
  mc <- calculateFractionMassConverters(mw)
  mw.dt <- sec.long[, .(filename=paste0(sample,'.fraction',fraction), id=fraction, molecular_weight=mc$fraction2Mass(fraction))] %>% 
    unique()
  return(mw.dt)
}
```

```{r}
mw.traces <- prepareTracesFractionAnnotation(sec.long = sec.long[sample=='Infected_3'])
```
function to ensure the anno and traces names and order match

```{r}
enforceTraceAndAnnotationRowOrder <- function(traces.obj){

  mat  <- traces.obj$traces
  anno <- traces.obj$trace_annotation
  
  # subset the mat to the set of proteins in the anno.dt
  prot.overlaps <- intersect(mat$id, anno$protein_id)
  submat <- mat[id %in% prot.overlaps,]
  message(nrow(submat),  ' out of ', nrow(mat), ' rows match betweem the protein annotation and traces file\nSubsetting to the overlapping proteins')
  
  anno <- anno[id %in% prot.overlaps]
  #match the row ordering between anno and ints
  traces.obj$traces <- submat[match(anno$id, submat$id)]
  traces.obj$trace_annotation <- anno
  return(traces.obj)
}

traces.subset <- enforceTraceAndAnnotationRowOrder(traces.obj = traces.obj)
```


## Create CCprofiler traces object

test on Infected_3
```{r}
ints.dt <- prepareTracesIntensity(sec.dt = sec.long[sample=='Infected_3',], useInterpolated = FALSE)
p.anno.dt <- prepareTracesProteinAnnotation('~/Documents/utils/mg_utils/data/uniprotkb_reviewed_true_2025_05_13.tsv.gz')
f.anno.dt <- mw.traces <- prepareTracesFractionAnnotation(sec.long = sec.long[sample=='Infected_3'], mw_path = '/Users/martingordon/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/data/HIV_Infection_CD4T_cells/SEC_Profiles/D1-D3-cal_SRT.txt')
trace.type <- 'protein'

traces.obj <- list(traces=ints.dt,
                     trace_type=trace.type,
                     trace_annotation=p.anno.dt,
                     fraction_annotation=f.anno.dt
                     )


class(traces.obj) <- 'traces'

traces.obj
```
Now create the CORUM complexes for scoring 

```{r}
corum.dt <- fread('~/Documents/projects/011325_MMuralidharan_SECMS_lysisPrepBenchmark/050925_CCprofiler_testRun_data/2025_05_12_corum.5.1.hsComplexes.long.csv.gz')
# filter to complexes with at least  two proteins or throws error
corum.dt <- corum.dt[complex_id %in% corum.dt[,.N, by=complex_id][N>1, complex_id]]
```

```{r}
# complex to  ppi format
binaryHypotheses <- generateBinaryNetwork(corum.dt)
pathLength <- calculatePathlength(binaryHypotheses) # shortest path

# generating corum complex + decoys 
corumTargetsPlusDecoys <- generateComplexDecoys(target_hypotheses=corum.dt,
                                                dist_info=pathLength,
                                                min_distance = 2,
                                                append=TRUE)

corumTargetsPlusDecoys[, isDecoy := ifelse(grepl('DECOY', complex_name, ignore.case = T), 'yes', 'no')]
corumTargetsPlusDecoys[,.N, by=isDecoy] # even split of true and false sets
```

```{r}
complexFeatures <- findComplexFeatures(traces=traces.subset,
                                       parallelized = TRUE,
                                       n_cores = 3,
                                       complex_hypothesis = corumTargetsPlusDecoys)
```

Ok, now we have the utility functions, need to try the differential expression for each of the analysis steps 

```{r}

```



