---
title: "062425_HIV.d123.ccprofiler.QCandpreprocessing"
author: "Martin Gordon"
date: "2025-06-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

The differential analysis doesnt seem to work well combining the different donors due to huge batch/run differences
Just split into three lists and process seperately

```{r packages}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)
library(circlize)
library(stringr)
library(viridis)
library(pracma)
library(magick)
library(eulerr)
library(CCprofiler)

source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/enrichmentTestFunctions.R")
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")
source("../../utils/bp_utils/UniprotIDMapping.R") # map hu to mm 

customTheme <-  theme_bw() +
  theme(axis.text.x = element_text(angle=90),
        panel.border = element_rect(color = "lightgrey",
                                    fill = NA,
                                    size = 1))
```

Read in the peptide traces obj and split into the different donor sets

Just read it in and drop the subgroups... like below, ugly, but (hopefully) preserves the structure
```{r}
pep.traces <- readRDS('./062325_HIV.d123.ccprofiler.diffAnalysis_data/2025_06_24_hiv.d123.huAnnotated.peptide.traces.rds')

pep.traces$Challenged_2 <- NULL
pep.traces$Challenged_3 <- NULL
pep.traces$Infected_2 <- NULL
pep.traces$Infected_3 <- NULL
pep.traces$Uninfected_2 <- NULL
pep.traces$Uninfected_3 <- NULL

saveRDS(pep.traces, ScriptAndDatedFileName('hiv.d1.ccprofiler.peptide.traces.rds'))
pep.traces <- readRDS('./062325_HIV.d123.ccprofiler.diffAnalysis_data/2025_06_24_hiv.d123.huAnnotated.peptide.traces.rds')

pep.traces$Challenged_1 <- NULL
pep.traces$Challenged_3 <- NULL
pep.traces$Infected_1 <- NULL
pep.traces$Infected_3 <- NULL
pep.traces$Uninfected_1 <- NULL
pep.traces$Uninfected_3 <- NULL
pep.traces

saveRDS(pep.traces, ScriptAndDatedFileName('hiv.d2.ccprofiler.peptide.traces.rds'))

# ditto rep3...
pep.traces <- readRDS('./062325_HIV.d123.ccprofiler.diffAnalysis_data/2025_06_24_hiv.d123.huAnnotated.peptide.traces.rds')

pep.traces$Challenged_2 <- NULL
pep.traces$Challenged_1 <- NULL
pep.traces$Infected_2 <- NULL
pep.traces$Infected_1 <- NULL
pep.traces$Uninfected_2 <- NULL
pep.traces$Uninfected_1 <- NULL
pep.traces

saveRDS(pep.traces, ScriptAndDatedFileName('hiv.d3.ccprofiler.peptide.traces.rds'))
rm(pep.traces)
```
alright drop the original and preprocess the other 3 objects 

```{r}
traces.donors.ls <- list(
  donor1=readRDS(ScriptAndDatedFileName('hiv.d1.ccprofiler.peptide.traces.rds')),
  donor2=readRDS(ScriptAndDatedFileName('hiv.d2.ccprofiler.peptide.traces.rds')),
  donor3=readRDS(ScriptAndDatedFileName('hiv.d3.ccprofiler.peptide.traces.rds')))
```

```{r}
traces.donors.ls[[1]]$Challenged_1


examplePeptideTraces 
```

QC on the donor sets
Within donor groups the alignemts look at cross correlation across fractions
```{r}
lapply(names(traces.donors.ls), function(x) alignTraces(traces.donors.ls[[x]], min_lag = -5, max_lag = 5, plot=T, PDF=T, name = paste0(x, '.crossCorr.plot')))
```
plot global intensites as proxy for extraction efficiency

```{r}
lapply(names(traces.donors.ls), function(x) plotGlobalIntensities(traces.donors.ls[[x]],plot=T, PDF=T, name = paste0(x, '.globalInts.plot')))
```
find missing values (those with values in flanking fractions)
replaces 0 with NA so these values can then be imputed
```{r}
# want vals in left and right borders
traces.donors.ls.mv <- lapply(names(traces.donors.ls), function(x) findMissingValues(traces.donors.ls[[x]], 
                                                              bound_left = 1,
                                                              bound_right = 1, 
                                                              consider_borders = TRUE
                                                              ))
names(traces.donors.ls.mv) <- names(traces.donors.ls)
```

now impute the missing values by fityting a spline

```{r}
# want vals in left and right borders
traces.donors.ls.imp <- lapply(names(traces.donors.ls.mv), function(x) imputeMissingVals(traces.donors.ls.mv[[x]], 
                                                              method = 'spline'
                                                              ))

names(traces.donors.ls.imp) <- names(traces.donors.ls.mv)
names(traces.donors.ls.imp)
```
plot the imputation success; package output of date.
Needs updating...

```{r}
lapply(names(traces.donors.ls.imp), function(x) plotImputationSummary(traces.donors.ls.mv[[x]], 
                                                                      traces.donors.ls.imp[[x]], 
                                                                      PDF=F,
                                                                      plot_traces = T, max_n_traces = 2
                                                                      ))
```
```{r}
pepTracesNormalized <- normalizeByCyclicLoess(pepTracesImp, window = 3, step = 1, plot = TRUE, PDF = TRUE, name = "normalizeByCyclicLoess")
```

Save the imputed values
```{r}
lapply(names(traces.donors.ls.imp), function(x) saveRDS(traces.donors.ls.imp[[x]], ScriptAndDatedFileName(paste0(x, '.hiv.pep.traces.impute.rds'))))
```
read in the imputed values
```{r}
traces.donors.ls.imp <- list(donor1=readRDS('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/062425_HIV.d123.ccprofiler.QCandpreprocessing_data/2025_06_24_donor1.hiv.pep.traces.impute.rds'),
                             donor2=readRDS('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/062425_HIV.d123.ccprofiler.QCandpreprocessing_data/2025_06_24_donor2.hiv.pep.traces.impute.rds'),
                             donor3=readRDS('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/062425_HIV.d123.ccprofiler.QCandpreprocessing_data/2025_06_24_donor3.hiv.pep.traces.impute.rds'))
```

## 06-24-25
Tidy the ints values; just summarize features to peptides and clean the peptide names

```{r}
pep.dt <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/062325_HIV.d123.ccprofiler.diffAnalysis_data/2025_06_24_ccprofiler.peptidesInt.input.csv.gz')

pep.dt[, pep_id := gsub('_|_[.][0-9]{1,2}$', '', peptide_id)]
pep.dt[, pep_ints := sum(intensity, na.rm=T), by=.(pep_id, protein_id, filename)]

sum.dt <- pep.dt[, .(protein_id, peptide_id=pep_id, filename, intensity=pep_ints)] %>% 
  unique()

fwrite(sum.dt, ScriptAndDatedFileName('ccprofiler.sum.peptidesInt.input.csv.gz'))
```
Now input this set and see if it works
prepare the fraction annotation info, include sample, condition and replicate id
```{r}
anno.dt <- pep.dt[, .(filename, fraction_number=filename)] %>% 
  unique() %>% 
  .[, c('sample', 'fraction_number') := tstrsplit(filename, '[.]', keep=c(1,2))]

anno.dt[, c('condition_id', 'replicate_id') := tstrsplit(sample, '_', keep=c(1,2))]
anno.dt[, fraction_number := as.integer(fraction_number)]

```
Design matrix; prepare this for the differential comparisons
Should consist of i)Sample_name (chr), ii)Condition(chr) and iii) Replicate(int)

```{r}
design.dt <- anno.dt[, .(Sample_name=sample,
                         Condition=condition_id,
                         Replicate=as.integer(replicate_id)
                         )] %>% 
  unique()
```

```{r}
mw.dt <- fread("./data/HIV_Infection_CD4T_cells/SEC_Profiles/D1-D3-cal_SRT.txt")
setnames(mw.dt, new=c('std_elu_fractions', 'std_weights_kDa'))
mw.dt <- mw.dt[, .(std_weights_kDa, std_elu_fractions)]


calibration = calibrateMW(mw.dt,
                          PDF=T,
                          plot=TRUE)
```

test CCprofiler input import; these two seem to match? Why is the function failing? it seems to drop a number of the fractions...
```{r}
sum.dt[grepl('Challenged_1', filename)]$filename %>% unique() == 
anno.dt[grepl('Challenged_1', filename), .(filename, fraction_number)]$filename %>% unique()



test.pcp.long <- sum.dt[grepl('Challenged_1', filename)]


setkey(test.pcp.long)
test.anno <- anno.dt[grepl('Challenged_1', filename), .(filename, fraction_number)]
setkey(test.anno, fraction_number)

m.test.pcp.long <- merge(test.pcp.long, test.anno, by='filename')
setkey(m.test.pcp.long,fraction_number)
m.test.pcp.long[, fraction_number:= NULL]

# order of filenames need to match in both tables
test.import <- importPCPdata(input_data=test.pcp.long,
                             fraction_annotation=test.anno)

```
can we annotate the test import? see if works
```{r}
idMapper <- fread('/Users/martingordon/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/062325_HIV.d123.ccprofiler.diffAnalysis_data/2025_06_23_id.mappers.csv.gz')

idMapper[,  peptide_id := gsub('_|_[.][0-9]{1,2}$', '', peptide)]
idMapper <- idMapper[, .(protein_id=protein, gene, mw, org)] %>% 
  unique() %>% 
  .[, head(.SD, 1), by=protein_id]

test.import.anno <- annotateTraces(traces = test.import, 
                                   trace_annotation = idMapper, 
                                   traces_id_column = "protein_id", 
                                   trace_annotation_id_column = "protein_id",
                                   trace_annotation_mass_column = "mw", 
                                   uniprot_mass_format = FALSE, 
                                   replace_whitespace = TRUE)
# annotate MW
test.import.anno <- annotateMolecularWeight(test.import.anno, calibration)
```

Try the imputation and normalization to see if it works as expected
Find missing values
```{r}

test.import.ms <-  findMissingValues(test.import.anno,
                                 bound_left = 1,
                                 bound_right = 1,
                                 consider_borders = TRUE)
```
now impute the missing values
```{r}
test.import.imp <- imputeMissingVals(test.import.ms, method = "spline")
```
fails due to melt function... again.. need to prefix with reshape2::
```{r}
plotImputationSummary(test.import.imp, test.import.imp, PDF = F,
                      plot_traces = T, max_n_traces = 2)
```
```{r}
test.import.norm <- normalizeByCyclicLoess(test.import.imp, window = 3, step = 1, plot = TRUE, PDF = TRUE, name = "normalizeByCyclicLoess")
```
ok, test the pipeline with just two conditions

```{r}

#' # Import traces list
#samples <- unique(anno.dt$sample)
samples <- c('Infected_1', 'Uninfected_1')
# Import data as traces object for each sample
traces_list <- lapply(samples,function(x){
  message(x) #just need these two cols I think
  ann <- anno.dt[grepl(x, filename), .(filename, fraction_number)]
  setkey(ann,fraction_number)
  
  data <- sum.dt[grepl(x, filename),]
  data <- merge(data, ann, by=c("filename"))
  setkey(data,fraction_number)
  data[, fraction_number:= NULL]

  traces <- importPCPdata(input_data=data,fraction_annotation=ann, rm_decoys = FALSE)
  return(traces)
})

names(traces_list) = samples
class(traces_list) <- "tracesList"
```

annotate with IDmapper info

```{r}
test.import.anno <- annotateTraces(traces = traces_list, 
                                   trace_annotation = idMapper, 
                                   traces_id_column = "protein_id", 
                                   trace_annotation_id_column = "protein_id",
                                   trace_annotation_mass_column = "mw", 
                                   uniprot_mass_format = FALSE, 
                                   replace_whitespace = TRUE)
```

find missing and label NA
```{r}
test.import.ms <- findMissingValues(traces_list,
                  bound_left = 1,
                  bound_right = 1, 
                  consider_borders = TRUE
                  )

test.import.ms
```

impute 
```{r}
test.import.imp <-imputeMissingVals(test.import.ms, method = "spline")
```
normalization; only two sample test; failed at same issue, *but* have what looks like a decent rds obj with correct looking peptide ids, so its a start..
continue from here tomorrow after cleaning up 
```{r}

debug(normalizeByCyclicLoess)
test.import.norm <- normalizeByCyclicLoess(test.import.imp, window = 3, step = 1, plot = TRUE, PDF = TRUE, name = "normalizeByCyclicLoess")


readRDS('./combi_table_norm.rds')
```

Issue: mismatch between trace and trace annotation obj. Why are some duplicated?
Pulled these from function:
Issue is; duplicated peptides assigned to different proteins. Will result in different number of records in the peptideTraces object.... how is this resolved?

```{r}
length(trace.ids)
length(trace.anno.ids)
which(trace.ids != trace.anno.ids)

trace.ids.dt <- data.table(trace.ids)[,.N, by=trace.ids]
trace.anno.dt <- data.table(trace.anno.ids)[,.N, by=trace.anno.ids]

trace.anno.dt
trace.ids.dt
trace.anno.dt[, .(id=trace.anno.ids, anno.N=N)]

# these are the problematic peptide IDs.. why are these duplicated? Is this issue only with this sample?
dupPeptides <- merge(trace.anno.dt[, .(id=trace.anno.ids, anno.N=N)], trace.ids.dt[, .(id=trace.ids, trace.N=N)], by='id') %>% 
  .[anno.N != trace.N, id]



traces_list$Infected_1$trace_annotation[id %in% dupPeptides]

```



normalize intensities
method? cyclic loess;
```{r}
library(parallel)
ncores <- parallel::detectCores() - 2

assignInNamespace("melt", reshape2::melt, ns = "CCprofiler")

# lets try run apply in parallel
traces.donors.ls.norm <- lapply(names(traces.donors.ls.imp), function(x) normalizeByCyclicLoess(traces.donors.ls.imp[[x]], 
                                                                                                window = 3, 
                                                                                                step = 1, 
                                                                                                plot = FALSE, 
                                                                                                PDF = FALSE, 
                                                                                                name = paste0(x,".normalization.plots")))
```


Drop the ID col
```{r}
testForTracelist <- function (tracesList, type = NULL, additionalItems = NULL) 
{
    if (!class(tracesList) == "tracesList") {
        stop("Object is not of class tracesList")
    }
    if (is.null(names(tracesList))) 
        stop("TracesList must consist of named traces objects. No names detected.")
    res <- lapply(tracesList, function(traces) {
        if (!all(c("traces", "trace_type", "trace_annotation", 
            "fraction_annotation") %in% names(traces))) {
            stop("At least one traces object doesn't contain all necessary items: traces, trace_type, trace_annotation, and fraction_annotation.")
        }
        if (!is.null(type)) {
            if (type != traces$trace_type) {
                stop("At least one traces object is of wrong type. Please check your input traces.")
            }
        }
        if (!identical(traces$traces$id, traces$trace_annotation$id)) {
            stop("In at least one traces object: IDs in traces and trace_annotation are not identical.")
        }
        if (!identical(names(traces$traces), c(traces$fraction_annotation$id, 
            "id"))) {
            stop("In at least one traces object: Fractions in traces and fraction_annotation are not identical.")
        }
        if (!is.null(additionalItems)) {
            contained <- (additionalItems %in% names(traces))
            if (!all(contained)) {
                stop(paste0("Required entries not found: ", additionalItems[!contained]))
            }
        }
    })
}
```

```{r}
class(test.trace) <- 'tracesList'
testForTracelist(test.trace)
```


Here is the error 

Error in FUN(X[[i]], ...) : 
  At least one traces object doesn't contain all necessary items: traces, trace_type, trace_annotation, and fraction_annotation.
5.
stop("At least one traces object doesn't contain all necessary items: traces, trace_type, trace_annotation, and fraction_annotation.")
4.
FUN(X[[i]], ...)
3.
lapply(tracesList, function(traces) {
if (!all(c("traces", "trace_type", "trace_annotation", "fraction_annotation") %in%
names(traces))) {
stop("At least one traces object doesn't contain all necessary items: traces, trace_type, trace_annotation, and fraction_annotation.") ...
2.
.tracesListTest(traces_list_norm, type = "peptide")
1.
normalizeByCyclicLoess(test.trace, window = 3, step = 1)


test a single list obj to normalize
```{r}
test <- normalizeByCyclicLoess(traces.donors.ls.imp$donor1, window = 3, step = 1, plot=FALSE, PDF=FALSE)

undebug(normalizeByCyclicLoess)
normalizeByCyclicLoess(traces.donors.ls.imp$donor1, window = 3, step = 1, plot=FALSE, PDF=FALSE)


traces.donors.ls.imp$donor1$Challenged_1
normalizeByCyclicLoess()
```

Other option; maybe it is bc we are using traces not peptide lvl info? Sum the intensities across features to arise at peptide lvl

```{r}
test.trace$Infected_1$traces <- head(test.trace$Infected_1$traces, 10)
test.trace$Uninfected_1$traces <- head(test.trace$Uninfected_1$traces, 10)

test <- normalizeByCyclicLoess(traces.donors.ls.imp$donor1, window = 3, step = 1, plot=FALSE, PDF=FALSE)
```


Lets try something different.. overwrite the package use of `melt` to specify reshape2::melt (this might be risky, but lets try otherwise wont work..)
```{r}
devtools::install('/Users/martingordon/modifiedPackages/CCprofiler_localMods')
```

```{r}
pep.dt
melt(traces.donors.ls.imp$donor1$Challenged_1$traces, )
getAnywhere(normalizeByCyclicLoess)
```




