---
title: "070825_CCprofiler_QCandProcessing"
author: "Martin Gordon"
date: "2025-07-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Process the loaded CCprofiler data and see if we can get to the differential analysis section

```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)
library(circlize)
library(stringr)
library(viridis)
library(pracma)
library(magick)
library(eulerr)
library(CCprofiler)
library(parallel)
library(matrixStats)

source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/enrichmentTestFunctions.R")
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")
source("../../utils/bp_utils/UniprotIDMapping.R") # map hu to mm 

customTheme <-  theme_bw() +
  theme(axis.text.x = element_text(angle=90),
        panel.border = element_rect(color = "lightgrey",
                                    fill = NA,
                                    size = 1))
```



```{r}
traces_list <- readRDS('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/070825_CCprofiler_retryPrepInput_data/2025_07_08_hiv.d2.features.traces.annotated.rds')
```

ALign the peptide intensities to see how things look
```{r}
# the first frations are quite noisy; restrict to the >= 15
traces_subset <- subset(traces_list, fraction_ids=seq(20,70,1))

parDir
parDir <- sub(".+[/]", "", getwd())

alignTraces(traces_subset, min_lag = -5, max_lag = 5, plot=T, PDF=T, name =paste0(parDir,'hiv.sample2.allfract.crossCorr.plot'))
```
plot the global peptide intesnities of the 
```{r}
plotGlobalIntensities(traces_list, plot = T, PDF=T, name='hiv.donor2.globalintensities.plot')
```

find missing values and impute by fitting a spline to the data
```{r}
#' Find missing values
#' (defined as having identifications in left and right neigbouring fractions):
traces_wMissing <- findMissingValues(traces_list,
                                 bound_left = 1,
                                 bound_right = 1,
                                 consider_borders = TRUE)

#' Impute NA values by fitting a spline: (what is the boundary used for curve fitting?)
traces_wImp <- imputeMissingVals(traces_wMissing, method = "spline")

#' Plot imputation summary; fails due to melt function need to prefix with reshape2
plotImputationSummary(traces_wMissing, 
                      traces_wImp, 
                      PDF = T,
                      plot_traces = T,
                      name='hiv.donor2.imputationSummary.plot',
                      max_n_traces = 2)

saveRDS(traces_wImp,  ScriptAndDatedFileName('hiv.d2.peptide.traces.allfract.wImpute.rds'))
traces_wImp <- readRDS('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/070825_CCprofiler_QCandProcessing_data/2025_07_09_hiv.d2.peptide.traces.allfract.wImpute.rds')
```

Normalize across donors
Normalize using cyclic loess : windowed normalization across runs; based on MA plot, want to remove systematic differences across runs (valid? are the curves different? If so not appropriate)
Also v slow to run for numerous samples... other approaches to normalize if we dont expect same curve?

```{r}
traces_norm <- normalizeByCyclicLoess(traces_wImp,  window = 3, step = 1, plot = TRUE, PDF = TRUE, name = "d2_normalizeByCyclicLoess")
```

# save the normalized output

```{r}
saveRDS(traces_norm,  ScriptAndDatedFileName('hiv.d2.peptide.traces.allfract.normalized.rds'))
```

```{r}
traces_norm <- readRDS('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/070825_CCprofiler_QCandProcessing_data/2025_07_09_hiv.d2.peptide.traces.allfract.normalized.rds')
```

plot normalized intensities, filter out < consec ids
```{r}
#' # Assess total intensity after normalization
plotGlobalIntensities(traces_norm, plot = T, PDF=T, name = "hivd2.IntensitySummary_postNormalization")

# looks like an issue combining the traces object
traces_combined <- combineTracesMutiCond(traces_norm)

# Filter by consecutive IDs:
traces_consec <- filterConsecutiveIdStretches(traces_combined, min_stretch_length = 5,remove_empty = T)
```
set up test traces dataset 

```{r}
test.traces <- examplePeptideTracesList
class(test.traces) <- "tracesList"

test.traces.comb <-  combineTracesMutiCond(test.traces)
test.traces.comb$trace_annotation %>% str()
test.traces.comb$fraction_annotation

# look ok to compare here..
test.traces.comb %>% print()
traces_combined %>% print()
```


Investigate why they diverge

Too frustrating and I'm not learning anything here... write my own function to calculate sister peptide correlation scores, and use this to do the filtering

```{r}
# use the traces.obj for now; will modify to use with BP pipeline
sec.dt <- fread(traces.obj)

```

Sister peptide cor
```{r}
#' functions for filtering sibling peptides
#' filter by max score and/or average
filterBySisterPeptideCorrelations <- function(traces.obj, score='mean', plot=T, cutoff=0.2){

  stopifnot(class(traces.obj) == 'traces')
  stopifnot(score %in% c('mean', 'max'))
  
  # ncores to process in parallel
  cores <- parallel::detectCores() - 2

  
  .convertTracesToMatrix <- function(traces){
    return(as.matrix(traces, rownames='id'))
  }
  .doPWCorrelations <- function(subMat){
    cor.mat <- cor(subMat, use='everything', method='pearson')
    diag(cor.mat) <- NA
    return(cor.mat)
  }
  .calculateColMeans <- function(corMat){
    matrixStats::colMeans2(corMat, na.rm=T)
  }
  .calculateColMax <- function(corMat){
    matrixStats::colMaxs(corMat, na.rm=T)
  }

  scoreFunc <- if (score == 'mean'){
    .calculateColMeans
    } else {
    .calculateColMax }
  
  intMat <- .convertTracesToMatrix(traces.obj$traces)
  
  proteins.oi <- unique(traces.obj$trace_annotation$protein_id)
  
  corScores <- pbmcapply::pbmclapply(proteins.oi, function(prot){
    subMat <- intMat[rownames(intMat) %in% traces.obj$trace_annotation[protein_id == prot, unique(id)],]
    corMat <- .doPWCorrelations(t(subMat))
    scores <- scoreFunc(corMat)
    names(scores) <- colnames(corMat)
    return(scores)
  }, mc.cores = cores)
  names(corScores) <- proteins.oi
  allcorScores <- unlist(corScores)
  names(allcorScores) <- gsub("^.*?\\.", "", names(allcorScores))
  
  if (plot == TRUE) {
    p <- ggplot(data.table(corScore = allcorScores), aes(x = corScore)) + 
      geom_histogram(bins = 30) + theme_classic()
    plot(p)
  }
  filterpeps <- names(allcorScores)[allcorScores > cutoff]
  traces_filt <- subset(traces.obj, filterpeps, "id")
  return(traces_filt)
}
```
compare the output of the two
```{r}
filterByMaxCorr(traces=test.traces.comb, cutoff = 0.5, plot=T)
filterBySisterPeptideCorrelations(traces=test.traces.comb, cutoff = 0.5, score='max', plot=T)

undebug(filterBySibPepCorr)
filterBySibPepCorr(test.traces.comb,
                                   absolute_spcCutoff = 0.2,
                                   plot = T)
```
test the functions
```{r}
filterByMaxCorr(traces=test.traces.comb, cutoff = 0.5, plot=T)
filterBySibPepCorr(traces=test.traces.comb, plot=T)
debug(filterBySibPepCorr)

filterBySisterPeptideCorrelations(traces=test.traces.comb, cutoff = 0.5, score='max', plot=T)
filterBySisterPeptideCorrelations(traces=test.traces.comb, cutoff = 0.5, score='mean', plot=T)
```
Filter by max peptide cor score and min peptide cor score
```{r}
traces_maxCor <- filterBySisterPeptideCorrelations(traces=traces_consec, cutoff = 0.2, score='max', plot=T)
traces_meanCor <- filterBySisterPeptideCorrelations(traces=traces_consec, cutoff = 0.5, score='mean', plot=T)
```
Now, filter the traces list to this subset of things

```{r}
goodpeptides <- unique(traces_meanCor$trace_annotation$id)

traces_filtered <- lapply(traces_norm, function(x){
  subset(x, trace_subset_ids=goodpeptides)
})

class(traces_filtered) <- "tracesList"
```


```{r}
traces_filtered <- updateTraces(traces_filtered)
summary(traces_filtered)
```
```{r}
saveRDS(traces_filtered, ScriptAndDatedFileName('hiv.d2.peptide.traces.allfract.filtered.rds'))
```

plot some examples?
```{r}
design.dt <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/070825_CCprofiler_retryPrepInput_data/2025_07_11_hiv.d2.ccprofiler.designMat.csv')
```

write a function to plot the elution profiles 
Useful for the other plots we are sharing...


work on htis monday... include the MW function to fit the mw, so we can add a triangle point for the monomer elution
```{r}
convertToLong <- function(traces.obj, logTrans=F){
  long.dt <- setDT(reshape2::melt(traces.obj$traces,idcol='id'))
  setnames(long.dt, new=c('id', 'fraction', 'intensity'))
  long.dt[, fraction := as.character(fraction)]
  
  ori.dt <- copy(long.dt)
  message('annotating traces..')
  long.dt <- long.dt[traces.obj$trace_annotation,, on=.(id)]
  long.dt <- long.dt[traces.obj$fraction_annotation[,.(filename, fraction=as.character(id), molecular_weight)],, on=.(fraction)]
  long.dt[, fraction := as.numeric(fraction)]
  stopifnot(nrow(ori.dt) == nrow(long.dt))
  return(long.dt)
}
plot.dt <- convertToLong(traces_filtered$Infected_2)


cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7","#999999")



p <- ggplot(plot.dt[gene == 'MSL1',]) +
      geom_line(aes(x=fraction, y=intensity, colour=id, group=id)) +
      theme_bw() +
      theme(legend.position="none") + 
      scale_x_continuous(breaks=seq(1,max(plot.dt$fraction))) +
      scale_color_manual(values=cbPalette)
   
p
```
tidy this plot to show what we want
```{r}
generateElutionProfiles <- function(sec.long, log=FALSE, legend = TRUE,name="Traces", colour_by = "id", highlight=NULL, highlight_col=NULL, colorMap=NULL, monomer_MW=TRUE) {


  if(!is.null(highlight)){
    traces.long$outlier <- gsub("\\(.*?\\)","",traces.long$id) %in% gsub("\\(.*?\\)","",highlight)
    if(!any(traces.long$outlier)) highlight <- NULL
  }

  if(colour_by!="id") {
    if(!colour_by %in% names(traces$trace_annotation)){
      stop("colour_by is not availbale in trace_annotation.")
    }
    isoform_annotation <- subset(traces$trace_annotation,select=c("id",colour_by))
    traces.long <- merge(traces.long,isoform_annotation, by.x="id",by.y="id")
    traces.long[,line:=paste0(get(colour_by),id)]
  }

  ## Create a reproducible coloring for the peptides plotted
  if(!is.null(colorMap)){
    if(!all(unique(traces.long$id) %in% names(colorMap))){
      stop("Invalid colorMap specified. Not all traces to be plotted are contained in the colorMap")
    }
  }else{
    cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7","#999999")
    ids <- sort(unique(traces.long[[colour_by]]))
    if (length(ids) <= length(cbPalette)) {
      colorMap <- cbPalette[1:length(unique(traces.long[[colour_by]]))]
      names(colorMap) <- ids
    } else {
      colorMap <- createGGplotColMap(unique(traces.long$id))
    }
  }

  if(colour_by == "id") {
    p <- ggplot(traces.long) +
      geom_line(aes_string(x='fraction', y='intensity', colour='id', group='id'))
  } else {

    p <- ggplot(traces.long) +
      geom_line(aes_string(x='fraction', y='intensity', colour=colour_by, group='line'))
  }

  p <- p + xlab('fraction') +
    ylab('intensity') +
    theme_bw() +
    theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()) +
    theme(plot.margin = unit(c(1,.5,.5,.5),"cm")) +
    ggtitle(name) +
    scale_color_manual(values=colorMap)
  #theme(plot.title = element_text(vjust=19,size=10))
  if (log) {
    p <- p + scale_y_log10('log(intensity)')
  }
  if (!legend) {
    p <- p + theme(legend.position="none")
  } else {
    if (length(unique(traces.long$id)) > 25) {
      p <- p + theme(legend.position="none")
    } else {
      p <- p + theme(legend.position="bottom", legend.text=element_text(size = 5), legend.title = element_blank())
    }
  }

  if(!is.null(highlight)){
    legend_peps <- unique(traces.long[outlier == TRUE, id])
    if(is.null(highlight_col)){
      p <- p +
        geom_line(data = traces.long[outlier == TRUE],
                  aes_string(x='fraction', y='intensity', color='id'), lwd=2) +
        scale_color_manual(values=colorMap, breaks = legend_peps)
      ## scale_color_discrete(breaks = legend_peps)
    }else{
      ## legend_map <- unique(ggplot_build(p)$data[[1]]$colour)
      ## names(legend_map) <- unique(p$data$id)
      ## legend_map[legend_peps] <- highlight_col
      ## legend_vals <- rep(highlight_col, ceiling(length(legend_peps)/ length(highlight_col)))[1:length(legend_peps)]
      p <- p +
        geom_line(data = traces.long[outlier == TRUE],
                  aes_string(x='fraction', y='intensity', lty = 'id'),
                  color = highlight_col, lwd=2)
      # scale_color_discrete(guide = F)
      ## scale_color_manual(values = legend_map, limits = legend_peps)
      # guides(lty = FALSE)
      # scale_color_manual(limits = legend_peps, values = rep(highlight_col, length(legend_peps))) +
      # geom_line(aes_string(x='fraction', y='intensity', color='id'))
    }
  }

  if ("molecular_weight" %in% names(traces$fraction_annotation)) {
    fraction_ann <- traces$fraction_annotation
    tr <- lm(log(fraction_ann$molecular_weight) ~ fraction_ann$id)
    intercept <- as.numeric(tr$coefficients[1])
    slope <- as.numeric(tr$coefficients[2])
    mwtransform <- function(x){exp(slope*x + intercept)}
    MWtoFraction <- function(x){round((log(x)-intercept)/(slope), digits = 0)}
    mw <- round(fraction_ann$molecular_weight, digits = 0)
    breaks_MW <- mw[seq(1,length(mw), length.out = length(seq(min(traces$fraction_annotation$id),
                                                              max(traces$fraction_annotation$id),10)))]
    p <- p + scale_x_continuous(name="fraction",
                                breaks=seq(min(traces$fraction_annotation$id),
                                           max(traces$fraction_annotation$id),10),
                                labels=seq(min(traces$fraction_annotation$id),
                                           max(traces$fraction_annotation$id),10),
                                sec.axis = dup_axis(trans = ~.,
                                                    breaks=seq(min(traces$fraction_annotation$id),
                                                               max(traces$fraction_annotation$id),10),
                                                    labels = breaks_MW,
                                                    name = "MW (kDa)"))
    if (monomer_MW==TRUE){
      if ("protein_mw" %in% names(traces$trace_annotation)) {
        subunitMW.dt <- data.table(id=traces$trace_annotation$id,mw=traces$trace_annotation$protein_mw)
        subunitMW.dt$fraction <- MWtoFraction(subunitMW.dt$mw)
        subunitMW.dt[,boundary:=MWtoFraction(2*mw)]
        if (length(unique(subunitMW.dt$mw)) > 1) {
          p <- p + geom_point(data = subunitMW.dt, mapping = aes(x = fraction, y = Inf, colour=id),shape=18,size=5,alpha=.5)
        } else {
          p <- p + geom_vline(data = unique(subunitMW.dt), aes(xintercept = fraction), colour="red", linetype="dashed", size=.5)
          p <- p + geom_vline(data = unique(subunitMW.dt), aes(xintercept = boundary), colour="red", linetype="dashed", size=.5, alpha=0.5)
        }
      } else {
        message("No molecular weight annotation of the traces. Cannot plot monomer molecular weight.")
      }
    }
  } else {
    p <- p + scale_x_continuous(name="fraction",
                                breaks=seq(min(traces$fraction_annotation$id),
                                           max(traces$fraction_annotation$id),10),
                                labels=seq(min(traces$fraction_annotation$id),
                                           max(traces$fraction_annotation$id),10))
  }
}



generateElutionProfiles(plot.dt[gene =='AK2'])
View(CCprofiler::plot)


help(plot)
examples <- unique(traces_filtered[[1]]$trace_annotation$protein_id)[1:20]
for(test_proteins in examples){
  pepTest <- subset(traces_filtered, trace_subset_ids = test_proteins, trace_subset_type = "protein_id")
  print(pepTest)
  plot(pepTest)
}

proteinTraces=exampleProteinTraces
plot(proteinTraces)
```


```{r}
View(filterByMaxCorr)

filterByMaxCorr(traces = traces_consec, cutoff = 0.5, plot=T)

filterByMaxCorr(traces = test.traces.comb, cutoff = 0.5, plot = T)
filterByMaxCorr(traces = test.traces.comb, cutoff = 0.5, plot = T)

filterByMaxCorr(traces_norm$Uninfected_2, cutoff = 0.5, plot=T)

filterByMaxCorr(traces = test.traces.comb, cutoff = 0.2, plot = T, name='test.data')
undebug(filterByMaxCorr)
undebug(filterBySibPepCorr)

str(test.traces.comb)
str(traces_combined)

traces_combined$trace_annotation[protein_id == 'Q9Y6Y8']
traces_combined$traces[id == 'AHTSSTQLQEELEK.3']
```
```{r}
test.traces.comb
```


Look at sister peptide correlation.. fails.. why? due to protein ID?
```{r}
traces_consec$traces %>% dim()
traces_combined$traces %>% dim()





# Filter by maximum correlation for outlier removal:
traces_maxcor <- filterByMaxCorr(traces_consec,cutoff = 0,plot = T)

traces_combined$fraction_annotation

traces_norm$Infected_2$fraction_annotation %>% str()

traces_maxcor$fraction_annotation %>% str()

traces_consec$fraction_annotation %>% 

# Filter by consecutive SPC:
pepTracesSPC <- filterBySibPepCorr(pepTracesMaxCorr,
                                   absolute_spcCutoff = 0.2,
                                   plot = T, PDF = T)


#' # Subset pepTracesNormalized to valid peptides selected 
#' by consecutive filtering and SPC cutoff
validPeps <- unique(pepTracesSPC$trace_annotation$id)

traces_filtered <- lapply(traces_norm, function(x){
  subset(x, trace_subset_ids=validPeps)
})
class(traces_filtered) <- "tracesList"

#' Update traces with additional metrics for each fraction:
traces_filtered <- updateTraces(traces_filtered)

#' Inspect traces list:
summary(traces_filtered)

saveRDS(traces_filtered, ScriptAndDatedFileName("allsamples.peptide.traces.norm.filtered.rds"))
```
```{r}

```


