---
title: "070725_CRISPR_D2_Overlaps"
author: "Martin Gordon"
date: "2025-07-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Look at the overlap to CRISPR hits from Ujwals study with the donor2 samples
Big probelm the differential results are very different across both analysis... not sure what to do here... reanalyze all together? repeat with the MSstats format?

```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)
library(circlize)
library(stringr)
library(viridis)
library(pracma)
library(magick)
library(eulerr)

source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/enrichmentTestFunctions.R")
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")
source("../../utils/bp_utils/UniprotIDMapping.R") # map hu to mm 

customTheme <-  theme_bw() +
  theme(axis.text.x = element_text(angle=90),
        panel.border = element_rect(color = "lightgrey",
                                    fill = NA,
                                    size = 1))
```

Read in the significant CRISPR results
```{r}
crispr.ko <- setDT(readxl::read_xlsx('~/Library/CloudStorage/Box-Box/HIV_Infection_CD4T_cells/CRISPR/CRISPRko_significant_genes.xlsx'))
crispr.act <- setDT(readxl::read_xlsx('~/Library/CloudStorage/Box-Box/HIV_Infection_CD4T_cells/CRISPR/CRISPRact_significant_genes.xlsx'))


colnames(crispr.ko) <- gsub('\\"', '', colnames(crispr.ko))
colnames(crispr.act) <- gsub('\\"', '', colnames(crispr.act))

crispr.act[, assay := 'CRISPR_act' ]
crispr.ko[, assay := 'CRISPR_ko' ]

crispr.dt <- rbind(crispr.act, crispr.ko) %>% 
  .[, .(assay, gene=id, num_guides=num, neg.p.value, neg.fdr, neg.lfc, neg.goodsgrna, pos.p.value, pos.fdr, pos.lfc, pos.goodsgrna)]
```
What do these scores mean? CRISPR activation overexpress a gene
A negative fold change is a reduction in numerator so negatively impacts growth
A positive fold change increases growth

CRISPR ko we knockout a gene and pos lfc is increased growth, negative lfc is impacted growth

```{r}
#fwrite(crispr.dt,  ScriptAndDatedFileName('hiv.crispr.scores.csv.gz'))
```


Read in the donor 2 data and see if any of these scores overlap 

```{r}
old.pw <- fread('~/Documents/projects/040825_MMuralidharan_PCProphet_HIV_CD4Tcells/060225_differentialAnalysisSeperateDonors_data/2025_06_02_allPWcontrasts.anova.toShare.csv.gz')

pw.contrasts <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_hiv.d123.pwContrasts.csv.gz')

pw.contrasts <- pw.contrasts[grepl('_2$', contrast)]

# get the best score per protein
pw.contrasts <- pw.contrasts[, .SD[which.min(treatmentDiff.p)], by=.(contrast,gene,protein)]

pw.contrasts[, sig := 'not']
pw.contrasts[treatmentDiff.padj < 0.01, sig := 'yes']

pw.contrasts[,.N, by=.(contrast, sig)]
```


```{r}
crispr.dt[,.N, by=gene][N > 1]

# get the best score per protein
old.pw <- old.pw[grepl('_2$', contrast)]
old.pw <- old.pw[, .SD[which.min(treatmentDiff.p)], by=.(contrast,gene,protein)]

old.pw[, sig := 'not']
old.pw[treatmentIntensity.adjP < 0.01, sig := 'yes']

pw.contrasts[,.N, by=.(contrast, sig)]
old.pw[,.N, by=.(contrast, sig)]
```
The reuslts are completely different between the two searches.... what to look at overlap in?
Try both, new first
```{r}
old.pw$gene %>% unique() %>% length()
pw.contrasts$gene %>% unique() %>% length()


intersect(old.pw[contrast == 'infected_uninfected_2' & sig != 'not', gene], 
          pw.contrasts[contrast == 'Infected_Uninfected_2' & sig != 'not', gene]
          )
 

pw.contrasts[contrast == 'Infected_Uninfected_2' & sig != 'not', protein]


pw.contrasts.wCRISPR <- merge(pw.contrasts, dcast(crispr.dt, gene~assay, value.var = c('neg.p.value', 'neg.fdr', 'neg.lfc', 'pos.p.value', 'pos.fdr', 'pos.lfc')), by='gene')
pw.contrasts.wCRISPR
fwrite(pw.contrasts.wCRISPR, ScriptAndDatedFileName('hiv.d2.pw.contrasts.CRISPRsetScores.csv.gz'))


old.pw.wCRISPR <- merge(old.pw, dcast(crispr.dt, gene~assay, value.var = c('neg.p.value', 'neg.fdr', 'neg.lfc', 'pos.p.value', 'pos.fdr', 'pos.lfc')), by='gene')
old.pw.wCRISPR[ sig != 'not',]
```

```{r}
pw.contrasts.wCRISPR[sig != 'not']
```

