---
title: "Untitled"
author: "Martin Gordon"
date: "2023-06-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
library (data.table)
library (ComplexHeatmap)
library (ggplot2)
library(RCy3)
library(ndexr)
rotate.x.axis.text <- theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

source("../../utils/bp_utils/ManageScriptData.R")
WriteSessionInfo()
WriteInstalledPackages()
```

## Downloaded the CB PPI and plugged into Cytoscape. From cytoscape, export network in simple interaction format (*.sif) and also the node and edge table with more annotation information

info on the network: 
nodes: Baits yelow, prey are grey
edges: Coloured based on differential interaction. Pink edges enriched in BC cell lines (unique to either MDA-MB-231 or MCF7) as compared to MCF10A. 
Dotted line represents the physical protein-protein association (validated in other studies) with high Integrated Association Stringency score (think this is identified in another study?. 
Dotted lines are not imported into our network as they are taken from a public PPI. Do we need these? Maybe best to import 
How are these handled by the network? Seems they are treated as baits instead of prey eg XRCC2 & RAD51C interaction

```{r}
bc.ppi <- fread("./data/Breast cancer protein-protein interaction network.sif", header = F) #no header in file
colnames(bc.ppi) <- c('Bait', 'relationship', 'Prey')              

bc.ppi[, baitsPerPrey := .N, by=.(Prey)]


#matches n nodes in cytoscape network
length(unique(c(bc.ppi$Prey,bc.ppi$Bait)))
```

# compare with node and edge table 
Note on the network encoding: the public ppi data (dotted lines) node interaction directionality is not important. This is an issue with our network as it distorts network centrality scores
Example of RAD51C: public ppi interaction

```{r CTNA1 example}
bc.ppi[Bait == 'RAD51C',]
bc.ppi[Bait == 'P40692',]
```
The *.sif file only contains edges between nodes and doesnt distinguish origin (experimental or public ppi) origin of the interactions
Public PPI interactions: directionality not useful (baits vs prey not distingished). Lots of prey-prey interactions which are not of interest(?)
Prepare our own data; Load the node and edge tables from cytoscape

```{r}
bc.ppi.edge <- fread("./data/Breast cancer protein-protein interaction network default edge.csv", header = T, na.strings = "", fill = T)
bc.ppi.nodes <- fread("./data/Breast cancer protein-protein interaction network default node.csv", header = T, na.strings = "", fill = T)
```


Simplify the nodes & edges table. 
Extract names of baits in the experiment for filtering
```{r node-tab}
# rm
bc.ppi.nodes[, c('codelen','longname','shared name','Prey', 'selected','npat', 'nsam', 'p', 'pCL', 'pCV', 'pFN','q') := NULL]

colnames(bc.ppi.nodes) <- c('Bait', 'BaitUniprot', 'Gene', 'PreyUniprot')


#extract baits
measured.baits <- bc.ppi.nodes[!is.na(Bait), Gene]
#measured.baits.prey <- bc.ppi.nodes[!is.na(Bait) & !is.na(PreyUniprot), Gene]

```

Clean the edge table and  identify interacting nodes

```{r edge-tab}
# simplify edge tablee$name)
bc.ppi.edge[, c('selected', 'interaction', 'shared interaction', 'name','info') := NULL]

# split the shared name column into 3 and rm leading/trailing whitespace
bc.ppi.edge[, c('Node1', 'Interaction', 'Node2') := tstrsplit(`shared name`, split='\\(|\\)') ]
bc.ppi.edge[, Node1 := trimws(Node1) ]
bc.ppi.edge[, Node2 := trimws(Node2) ]
#drop shared name
bc.ppi.edge[, `shared name` := NULL]
```

, preys and baits are mixed in the public PPI data and many rows only contain PPIs identified in public datasets. Clean/remove as required
Nodes in the bc.ppi only record the interactions one way; for baits with other baits as preys, we also need to record the interaction in reverse to accurately count

```{r dt-clean}
#remove all rows with no baits as not of interest
bc.ppi.edge <- bc.ppi.edge[!(!Node1 %in% measured.baits & !Node2 %in% measured.baits)]
bc.ppi.edge[Node1 %in% measured.baits,unique(Node1)]
# some baits are in col2 with none in col1; find these and swap the columns they are assigned to  
# swap columns for rows matching i.. useful
bc.ppi.edge[!Node1 %in% measured.baits & Node2 %in% measured.baits, `:=`(Node1 = Node2, Node2 = Node1) ] #swap node1 and node2 places


# where both bait and prey are baits, duplicate the record and swap the cols they are assigned to 
bc.ppi.edge[Node1 %in% measured.baits & Node2 %in% measured.baits,]
# create a new row where both node values are swapped
new.rows <- bc.ppi.edge[Node1 %in% measured.baits & Node2 %in% measured.baits,]
new.rows <- new.rows[, `:=`(Node1=Node2, Node2=Node1)]

bc.ppi.edge <- rbind(bc.ppi.edge,new.rows)
bc.ppi.edge

#sanity check
bc.ppi.edge[Node2 %in% measured.baits, .N] # only 32 cases with bait also a node
bc.ppi.edge[Node1 %in% measured.baits, .N, by=Node1][,length(Node1)]#39 baits detected 
```
Now that dataset is correctly ordered, format & create a baitsPerPrey column

```{r}
bc.ppi.edge[, c('bfdr', 'diff_score', 'Interaction') := NULL]
colnames(bc.ppi.edge) <- c( "Cell","Public.PPI", "Bait","Prey")
bc.ppi.edge[, baitsPerPrey := .N, by = Prey]

#simplify name
bc.ppi <- bc.ppi.edge
```

# Load in the k.ppi and pcmi.ppi networks

```{r}
pcmi.ppi <- fread ("./data/D_Final_v3.txt")
k.ppi <- fread("./data/kaushikPPI.csv")
```

## Create a list of ppi data

```{r}

```



## Calculate number of bait per prey for the PCMI PPI network

```{r}
pcmi.ppi <- pcmi.ppi[, .(Bait, Prey = PreyGene, Clusters)]
pcmi.ppi[, baitsPerPrey := .N, by = Prey]

#k.ppi <- k.ppi[, .(Bait, Prey = PreyGene, Clusters)]
k.ppi[, baitsPerPrey := .N, by = Prey]
```

## What is the bait degree for each of the datasets

```{r}

```


# Create histogram of the prey degree

On first glance the bc network seems sparser; more nodes with fewer connections but likey due to size

```{r}
bc.ppi[,exp := 'bc']
pcmi.ppi[,exp := 'pcmi']

merge.ppi <- rbind(pcmi.ppi, bc.ppi, fill=TRUE)

g <- ggplot(data=merge.ppi, aes(x=baitsPerPrey, fill=exp)) +
  geom_bar(position = "dodge2") +
  ggtitle('Prey Degree') +
  theme_bw()
g

BackupAsPDF(g, 'prey-degree-')

```

```{r}
hists <- rbindlist( list (bc = bc.ppi[, .(baitsPerPrey = unique(baitsPerPrey)), by= Prey][, .N, by= baitsPerPrey], 
                          pcmi = pcmi.ppi[, .(baitsPerPrey = unique(baitsPerPrey)), by= Prey][, .N, by= baitsPerPrey]), idcol = "exp")



hists[exp == "bc", fractionPrey := N/length(unique(bc.ppi$Prey))]
hists[exp == "pcmi", fractionPrey := N/length(unique(pcmi.ppi$Prey))]

p <- ggplot (hists, aes(x = baitsPerPrey, y = N, fill = exp)) + geom_col( position = "dodge2") + theme_classic()
BackupAsPDF(p)


p <- ggplot (hists, aes(x = baitsPerPrey, y = fractionPrey, color = exp, fill=exp)) +
  geom_point() +
  scale_y_log10() +
  #scale_x_log10() +
  geom_smooth(se = FALSE, lty = "dashed", lwd = 0.5, span = 2, data = hists[baitsPerPrey < 15]) +
  xlab ("Prey Degree") +
  ylab ("Fraction of Prey") +
  scale_color_manual(values = c(bc = "orange", pcmi = "cornflowerblue")) +
  theme_bw() + 
  theme(legend.position="top")

BackupAsPDF(p)
```
Pairwise Overlaps
(Creating an adjacency matrix from bait prey relationships)
```{r}
pcmi.mat <- dcast (pcmi.ppi, Prey~Bait, value.var = "Bait", fun.aggregate = length) |> as.matrix(rownames= "Prey")
pcmi.mat[pcmi.mat > 1] <- 1 #constrain to 1 for baits-bait connections
Heatmap(pcmi.mat, col = c("white", "firebrick"), column_names_gp = gpar(fontsize =5))

```
## Pairwise Overlaps

baitPrey matrix

```{r}
bc.mat <- dcast (bc.ppi, Prey~Bait, value.var = "Bait", fun.aggregate = length) |> as.matrix(rownames= "Prey")
bc.mat[bc.mat > 1] <- 1
Heatmap(bc.mat, col = c("white", "firebrick"), column_names_gp = gpar(fontsize =5))
```
Calculates the jaccard distance (dissimilarity between two sets) of the different baits (ratio of size of intersection vs size of union; 0 is identical, 1 no common elements)
1-jaccard so only baits with common elements colored
```{r}

pcmi.jaccard <- dist (t(pcmi.mat), method = "binary") |> as.matrix()
Heatmap (1-pcmi.jaccard, col = circlize::colorRamp2(breaks = c(0, 0.4), colors =  c("white", "firebrick")))

```

```{r}
bc.jaccard <- dist (t(bc.mat), method = "binary") |> as.matrix()
Heatmap (1-bc.jaccard, col = circlize::colorRamp2(breaks = c(0, 0.4), colors =  c("white", "firebrick")))
```

```{r pcmi-heatmap}
#  File from Belinda, all HEK detectable proteins + PCMI prey
#scan("/Users/ben/Downloads/hek293tProteome_BekkerJenwen2017_ASDppiPrey.csv", what = character()) |> unique() |> length()
# 11169, use this number below as the background


#this function calculates p-value from Fishers exact test for the contigency table comparison between bait1 ansd bait2 preys in the matrix
pairwiseFisher <- function (bait1, bait2, matrix){
  both <- sum(rowSums(matrix[, c(bait1, bait2)]) == 2) #totoal N occurances of b1 + b2 #calculates the number of occurrences where both bait1 and bait2 are present by summing the rows where the sum of values for bait1 and bait2 is equal to 2
  firstOnly  <- sum(matrix[, c(bait1)]) - both #b1 only
  secondOnly <- sum(matrix[, c(bait2)]) - both #b2 only
  neither <- 11169 - both - firstOnly  -secondOnly
 # neither <- nrow(matrix) - both- firstOnly- secondOnly
  
  fisher.test (matrix(c(both, firstOnly, secondOnly, neither), nrow = 2), alternative = "greater")$p.value #2*2 contigency table; test to see if there is dependency between the two baits
}

allByAll <- data.table(bait1 = colnames(pcmi.mat))[, .(bait2 = colnames(pcmi.mat)), by = bait1][bait1 < bait2] #only cases where bait 1 < bait 2?

allByAll[ , fisherP := pairwiseFisher(bait1, bait2, pcmi.mat), by = .(bait1, bait2)] #fishers exact test applied to each pair

allByAll[, pP := -log10(fisherP)]
pcmi.fisher.dt <- allByAll

#combine std dt and a modified version with bait1 and bait2 swapped
#takes the allByAll datatable, appends a modified version of it with swapped columns, and then reshapes it into a matrix format where the rows represent the unique values of bait1, the columns represent the unique values of bait2, and the matrix cells pP values
pcmi.fisher.mat <- dcast (rbind (allByAll, allByAll[, .(bait1 = bait2, bait2 = bait1, fisherP, pP)]),
       bait1~bait2, value.var = "pP") |> as.matrix (rownames = "bait1")


baitClusters <- unique(pcmi.ppi[, .(Bait, Clusters)])


hm <- Heatmap (pcmi.fisher.mat,
         #name = "-log10(P) fisher.test",
         row_split = as.factor(baitClusters[rownames(pcmi.fisher.mat), Clusters, on = "Bait"]),
         column_split = as.factor(baitClusters[colnames(pcmi.fisher.mat), Clusters, on = "Bait"]),
         col = circlize::colorRamp2(breaks = c(0, 4), color = c("white", "firebrick")),
         left_annotation = rowAnnotation(cluster = as.factor(baitClusters[rownames(pcmi.fisher.mat), Clusters, on = "Bait"]),
                                         col = list (cluster = setNames(c("#FF3333","#FF9900","#F6D80C","#0CB212","#4A56FD","#AA4AFD","#1CDDEB","black"), 1:8))),
         top_annotation = columnAnnotation(cluster = as.factor(baitClusters[rownames(pcmi.fisher.mat), Clusters, on = "Bait"]),
                                         col = list (cluster = setNames(c("#FF3333","#FF9900","#F6D80C","#0CB212","#4A56FD","#AA4AFD","#1CDDEB","black"), 1:8))),
         border  = TRUE,
         row_names_gp = gpar(fontsize = 2),
         column_names_gp = gpar(fontsize = 2),
         gap = unit(0.3, "mm"),
         column_gap = unit(0.3, "mm"))

BackupAsPDF(hm, 'pcmi.-log10p-fishertest')
```

```{r bc-heatmap}
allByAll <- data.table(bait1 = colnames(bc.mat))[, .(bait2 = colnames(bc.mat)), by = bait1][bait1 < bait2]
allByAll[ , fisherP := pairwiseFisher(bait1, bait2, bc.mat), by = .(bait1, bait2)]

allByAll[, pP := -log10(fisherP)]

bc.fisher.dt <- allByAll

bc.fisher.mat <- dcast (rbind (allByAll, allByAll[, .(bait1 = bait2, bait2 = bait1, fisherP, pP)]),
       bait1~bait2, value.var = "pP") |> as.matrix (rownames = "bait1")


hm <- Heatmap (bc.fisher.mat, 
        # name = "-log10(P)fisher.test",
         col = circlize::colorRamp2(breaks = c(0, 4), color = c("white", "firebrick")))

BackupAsPDF(hm, 'bc.-log10p-fishertest')

hist(pcmi.fisher.mat[pcmi.fisher.mat > 0], breaks = 100)
hist(bc.fisher.mat[bc.fisher.mat > 0], breaks = 100, add = TRUE)


fisher.dt <- rbindlist(list (pcmi= pcmi.fisher.dt, bc = bc.fisher.dt), idcol = "exp")
ggplot(fisher.dt, aes (x = pP, color = exp)) + geom_density(alpha  = 0.5) +theme_bw() 
ggplot(fisher.dt, aes (x = pP, color = exp)) + geom_density(alpha  = 0.5) +theme_bw() + coord_cartesian(ylim = c(0, 0.1)) 
ggplot(fisher.dt, aes (x = pP, color = exp)) + geom_density(alpha  = 0.5) +theme_bw() + coord_cartesian(xlim = c(0, 5), ylim = c(0, 0.1)) 
p <- ggplot (fisher.dt[, 100 *sum(fisherP < 0.05)/.N, by = exp], aes(x = exp, y= V1, fill = exp)) +
  geom_col(show.legend = FALSE, width = 0.5) +
  geom_hline(yintercept = 5, color = "grey") +
  ylab( "% of bait pairs with significant (p < 0.05)\n prey convergence") +
  scale_fill_manual(values = c(bc = "orange", pcmi = "cornflowerblue")) +
  theme_classic() +
  coord_cartesian(ylim = c(5,15))
  p
  
BackupAsPDF(p)

```

```{r}

fisher.dt[baitClusters, cluster1 := i.Clusters, on = c(bait1 = "Bait")]
fisher.dt[baitClusters, cluster2 := i.Clusters, on = c(bait2 = "Bait")]
fisher.dt[!is.na(cluster2) & study == "kaushik"]
p <- ggplot (fisher.dt[study== "kaushik" | (cluster1 !=   cluster2),
                       100 *sum(fisherP < 0.05)/.N, by = study],
             aes(x = study, y= V1, fill = study)) +
  geom_hline(yintercept = 5, color = "grey") +
  geom_col(show.legend = FALSE, width = 0.5) +
  ylab( "% of bait pairs with significant (p < 0.05)\n prey convergence") +
  scale_fill_manual(values = c(kaushik = "orange", pcmi = "cornflowerblue")) +
  theme_classic() 
  
  
BackupAsPDF(p)
ggplot (fisher.dt[study== "kaushik" | (cluster1 !=   cluster2), sum(fisherP < 0.05)/.N, by = study], aes(x = study, y= V1)) + geom_col() + ylab( "% of bait pairs with signfiicant (p < 0.05) white balls") +
  ggtitle("Ball-and-Urn tests between all pairs of baits")


fisher.dt
hist (kppi[, .(baitDegree = length(unique(Prey))), by = Bait]$baitDegree, breaks = 0:120)
hist (pcmi.ppi[, .(baitDegree = length(unique(Prey))), by = Bait]$baitDegree, breaks = 0:120, add = TRUE, col = "red")

baitDegrees <- rbindlist (list (kaushik = kppi[, .(baitDegree = length(unique(Prey))), by = Bait],
                                pcmi = pcmi.ppi[, .(baitDegree = length(unique(Prey))), by = Bait]),
                          idcol = "study")

p <- ggplot (baitDegrees, aes(x = baitDegree, fill = study, color = study))  + 
  geom_dotplot(method = "histodot", binwidth = 1.5) +
  #geom_density(color = "black") +
  facet_wrap(~study, ncol = 1) +
  coord_cartesian(ylim = c(0,0.1)) +
  theme_classic() +
  geom_vline(xintercept = 33.5)
BackupAsPDF(p)

```

















*Not Used*

```{r}
# seperage the dt by public & measured. data
bc.ppi.edge.list <- split(bc.ppi.edge, by='public_ppi', keep.by = TRUE)
names(bc.ppi.edge.list) <- c('public.ppi', 'measured.ppi')

# all baits either in node1 or node2
#bc.ppi.edge.list[["measured.ppi"]][!Node1 %in% measured.baits & !Node2 %in% measured.baits]

# filtered out connections where neither node is a bait (rm 970 connections from the public DB) as not of interest
bc.ppi.edge.list[['public.ppi']] <- bc.ppi.edge.list[['public.ppi']][!(Node1 %in% measured.baits | Node2 %in% measured.baits),]
bc.ppi.edge.list[[1]]

# add the nodes to clean dt where both prey and bait are nodes
bc.ppi.edge.list[['measured.ppi']] <- rbind(bc.ppi.edge.list[['measured.ppi']],bc.ppi.edge.list[['public.ppi']][Node1 %in% measured.baits & Node2 %in% measured.baits])
# rm
bc.ppi.edge.list[['public.ppi']] <- bc.ppi.edge.list[['public.ppi']][!(Node1 %in% measured.baits & Node2 %in% measured.baits)]

# find the rows where the baits are in node1
bc.ppi.edge.list[['measured.ppi']] <-  rbind(bc.ppi.edge.list[['measured.ppi']],bc.ppi.edge.list[['public.ppi']][Node1 %in% measured.baits])
#rm
bc.ppi.edge.list[['public.ppi']] <-  bc.ppi.edge.list[['public.ppi']][!(Node1 %in% measured.baits)]

# find baits in node 2 (could combine with above, could actually merge any rows with baits in node1 or node2)
bc.ppi.edge.list[['measured.ppi']] <-  rbind(bc.ppi.edge.list[['measured.ppi']],bc.ppi.edge.list[['public.ppi']][Node2 %in% measured.baits])
#has 1070 rows
bc.ppi.edge
test
#filering out 940 rows

bc.ppi.edge.list[[1]][Node1 %in% measured.baits | Node2 %in% measured.baits,]


  # interactions between baits only encoded one-way; need to modify later, first lets clean and merge
bc.ppi.edge.list[[2]][Node1 == 'BRIP1' | Node2 == 'BRIP1']
```


```{r}
#these are the measured baits
measured.baits <- bc.ppi.nodes[!is.na(BaitGene), BaitGene]
measured.baits


# most baits are not in our bait list...
bc.ppi.edge.public[!Bait %in% measured.baits,]

bc.ppi.edge.public[Bait == 'BRE']

# sanity check is if Prey

bc.ppi.edge.public
bc.ppi.nodes

bc.ppi.edge[, c('Bait', 'Interaction', 'Prey')
```



Plot of baits per prey for both networks
```{r}
ggplot(data=bc.ppi, aes(x=baitsPerPrey, fill='blue', alpha=0.5)) +
  geom_bar( fill='blue') +
  geom_bar(data=pcmi.ppi, aes(x=baitsPerPrey, fill='red')) +
  ggtitle('Network Baits per Prey (BC=blue, PCMI=red)')
```


