---
title: "PPI.kaushik.BC.PCMI"
author: "Martin Gordon"
date: "2023-08-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## BC, TryKin & PCMI PPIs integration

Issues
What do we want to use in our BC data? We have a network consisting of multiple cell types and public data; high degree genes are bumped up in each dataset.. should we at least filter out public data in this?
Issue also is smaller dataset than others, so how comparable is this?
we have less baits than the other datasets, so sparser counts make sense.. Is this due to my filtering or the data?
Do we include public or just experimentally detected PPI interactions? Including public for now..
How to set the background in the fisher test, 
which baits to exclude from PCMI to make the baitDegree distributions more comparable

```{r}
library (data.table)
library (ComplexHeatmap)
library (ggplot2)
library(magrittr)

rotate.x.axis.text <- theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

source("../../utils/bp_utils/ManageScriptData.R")
WriteSessionInfo()
WriteInstalledPackages()
```

# read in the BC PPI


```{r}
bc.ppi <- fread("./data/Breast cancer protein-protein interaction network.sif", header = F) #no header in file
colnames(bc.ppi) <- c('Bait', 'relationship', 'Prey')              

bc.ppi[, baitsPerPrey := .N, by=.(Prey)]


#matches n nodes in cytoscape network
length(unique(c(bc.ppi$Prey,bc.ppi$Bait)))
```
# compare with node and edge table 
Note on the network encoding: public ppi data overlaid on experimentally detected interacitons and edges not distinguished in the file. Don't know what is bait and what is prey?
Example of RAD51C: public ppi interaction

```{r CTNA1 example}
bc.ppi[Bait == 'RAD51C',]
```
The *.sif file only contains edges between nodes and doesnt distinguish origin (experimental or public ppi) origin of the interactions
Public PPI interactions: directionality not useful (baits vs prey not distingished). Lots of prey-prey interactions which are not of interest(?)
Prepare our own data; Load the node and edge tables from cytoscape

```{r}
bc.ppi.nodes <- fread("./data/Breast cancer protein-protein interaction network default node.csv", header = T, na.strings = "", fill = T)

# only 39 baits in the dataset
bc.ppi.nodes[!is.na(Bait),] %>% nrow()
```


Simplify the nodes & edges table. 
Extract names of baits in the experiment for filtering
```{r node-tab}
# rm uninteresting datapoints
bc.ppi.nodes[, c('codelen','longname','shared name','Prey', 'selected','npat', 'nsam', 'p', 'pCL', 'pCV', 'pFN','q') := NULL]

colnames(bc.ppi.nodes) <- c('Bait', 'BaitUniprot', 'Gene', 'PreyUniprot')

#extract baits
measured.baits <- bc.ppi.nodes[!is.na(Bait), Gene]
```

Clean the edge table and  identify interacting nodes

22/08/23 drop public PPI nodes

```{r edge-tab}
bc.ppi.edge <- fread("./data/Breast cancer protein-protein interaction network default edge.csv", header = T, na.strings = "", fill = T)

# simplify edge tablee$name)
bc.ppi.edge[, c('selected', 'interaction', 'shared interaction', 'name','info', 'bfdr', 'cell', 'diff_score') := NULL]

# split the shared name column into 3 and rm leading/trailing whitespace
bc.ppi.edge[, c('Node1', 'Interaction', 'Node2') := tstrsplit(`shared name`, split='\\(|\\)') ]
bc.ppi.edge[, Node1 := trimws(Node1) ]
bc.ppi.edge[, Node2 := trimws(Node2) ]

#drop shared name
bc.ppi.edge[, `shared name` := NULL]
```

Drop the shared nodes

```{r}
bc.ppi.edge[, .N, by=public_ppi]
#bc.ppi.edge <- bc.ppi.edge[public_ppi == FALSE,]
```


, preys and baits are mixed in the public PPI data and many rows only contain PPIs identified in public datasets. Clean/remove as required
Nodes in the bc.ppi only record the interactions one way; for baits with other baits as preys, we also need to record the interaction in reverse to accurately count

```{r dt-clean}
#remove all rows with no baits as not of interest
bc.ppi.edge <- bc.ppi.edge[Node1 %in% measured.baits | Node2 %in% measured.baits]

# for the public ppi data, bait/prey cols order may be swapped. Find these and reverse order
# swap columns for rows matching i.. useful
# not needed if we are filtering out public ppi data
bc.ppi.edge[!Node1 %in% measured.baits & Node2 %in% measured.baits, `:=`(Node1 = Node2, Node2 = Node1) ] #swap node1 and node2 places


# where both bait and prey are baits, duplicate the record and swap the cols they are assigned to 
# create a new row where both node values are swapped
new.rows <- bc.ppi.edge[Node1 %in% measured.baits & Node2 %in% measured.baits,]
new.rows <- new.rows[, `:=`(Node1=Node2, Node2=Node1)]

bc.ppi.edge <- rbind(bc.ppi.edge,new.rows)

#sanity check
bc.ppi.edge[Node2 %in% measured.baits, .N] # only 32 cases with bait also a node (only 20 when removing public PPI data)
bc.ppi.edge[Node1 %in% measured.baits, .N, by=Node1][,length(Node1)]#39 baits detected 
```

Now that dataset is correctly ordered, format & create a baitsPerPrey column

```{r}
bc.ppi.edge[, c("public_ppi", 'bfdr', 'diff_score', 'Interaction') := NULL]
colnames(bc.ppi.edge) <- c("Bait","Prey")

#simplify name
bc.ppi <- bc.ppi.edge %>% 
  .[, baitsPerPrey := .N, by = Prey]
```

Read in other PPI datasets

```{r}
pcmi.ppi <- fread ("./data/D_Final_v3.txt")
pcmi.ppi <- pcmi.ppi[, .(Bait, Prey = PreyGene, Clusters)]
pcmi.ppi[, baitsPerPrey := .N, by = Prey]

k.ppi <- fread("./data/kaushikPPI.csv")[hc == TRUE]
k.ppi[, baitsPerPrey := .N, by = Prey]
```


# histogram of all datasets baitsPerPrey
V similar distributions 

```{r}

pdf("/Users/martingordon/Documents/projects/060823_BPolacco_TryKPPI/baitsperprey.pdf")
hist ( k.ppi[, .(baitsPerPrey), by= Prey]$baitsPerPrey, breaks = 0:20 + 0.5, col='blue')
hist ( pcmi.ppi[, .(baitsPerPrey), by= Prey]$baitsPerPrey, breaks = 0:20 + 0.5, add=T, col='red')
hist(bc.ppi[, .(baitsPerPrey), by= Prey]$baitsPerPrey, breaks = 0:20 + 0.5, add=T, col='grey')
legend(15, 800, legend=c("k.ppi", "pcmi.ppi", "bc.ppi"), col=c("blue","red", "grey"), lty=1, cex=0.8)
dev.off()


```
Histogram of N baits per prey. Seem to follow a similiar distribtion

```{r}
# smart way to format results and then plot together
hists <- rbindlist( list (kaushik = k.ppi[, .(baitsPerPrey), by= Prey] %>%  .[, .N, by=baitsPerPrey],
                          pcmi = pcmi.ppi[, .(baitsPerPrey), by= Prey]  %>% .[, .N, by=baitsPerPrey], 
                          bc   = bc.ppi[, .(baitsPerPrey), by= Prey] %>%  .[, .N, by=baitsPerPrey]),
                    idcol = "study")



p <- ggplot (hists, aes(x = baitsPerPrey, y = N, fill = study)) + 
  geom_col( position = "dodge2") +
  xlab("Baits Per Prey") + 
  ylab("Counts") +
  theme_classic()

p
BackupAsPDF(p, 'exp.ppi.baitsPerPrey')

```
Plot prey degree breakdown
```{r fig.width = 4, fig.height = 3}
# smart way to format results and then plot together
hists <- rbindlist( list (kaushik = k.ppi[, .(baitsPerPrey = unique(baitsPerPrey)), by= Prey][, .N, by= baitsPerPrey], 
                          pcmi = pcmi.ppi[, .(baitsPerPrey = unique(baitsPerPrey)), by= Prey][, .N, by= baitsPerPrey],
                          bc   = bc.ppi[, .(baitsPerPrey = unique(baitsPerPrey)), by= Prey][, .N, by= baitsPerPrey]),
                    idcol = "study")

hists[study == "kaushik", fractionPrey := N/length(unique(k.ppi$Prey))]
hists[study == "pcmi", fractionPrey := N/length(unique(pcmi.ppi$Prey))]
hists[study == "bc", fractionPrey := N/length(unique(bc.ppi$Prey))]

p <- ggplot (hists, aes(x = baitsPerPrey, y = fractionPrey, fill = study, color = study)) +
  #geom_hline(yintercept = c(1,2), lwd = 0.25, color = "grey") +
  geom_point() +
  scale_y_log10() +
  #scale_x_log10() +
  geom_smooth(se = FALSE, lty = "dashed", lwd = 0.5, span = 2, data = hists[baitsPerPrey < 15]) +
  xlab ("Prey Degree") +
  ylab ("Fraction of Prey") +
  scale_color_manual(values = c(kaushik = "orange", pcmi = "cornflowerblue", bc = "pink")) +
  theme_bw() + 
  theme(legend.position="top")
p

BackupAsPDF(p, 'data.ppi.networks.fractionprey')

```
# Inspect bait degrees for each of the datasets

```{r}

pdf('/Users/martingordon/Documents/projects/060823_BPolacco_TryKPPI/exp.bait.degree.histograms.pdf')
hist (k.ppi[, .(baitDegree = length(unique(Prey))), by = Bait]$baitDegree, breaks = 0:120, col='blue')
hist (pcmi.ppi[, .(baitDegree = length(unique(Prey))), by = Bait]$baitDegree, breaks = 0:120, add = TRUE, col = "red")
hist(bc.ppi[, .(baitDegree = length(unique(Prey))), by = Bait]$baitDegree, breaks = 0:120, add = TRUE, col = "grey")
legend(1, 95, legend=c("k.ppi", "pcmi.ppi", "bc.ppi"), col=c("blue","red", "grey"), lty=1, cex=0.8)
dev.off()
```

```{r fig.width = 4, fig.height = 3}

baitDegrees <- rbindlist (list(kaushik = k.ppi[, .(baitDegree = length(unique(Prey))), by = Bait],
                               pcmi.   = pcmi.ppi[, .(baitDegree = length(unique(Prey))), by = Bait],
                               bc      = bc.ppi[, .(baitDegree = length(unique(Prey))), by = Bait]),
                          idcol = "study")


p <- ggplot(baitDegrees, aes(x = baitDegree, fill=study))  +
  geom_bar() +
  facet_wrap(~ study, ncol = 1) +  # Adjust ncol as needed
  xlab("Bait Degree") +
  ylab("Count")  +
  scale_color_manual(values = c(kaushik = "orange", pcmi = "cornflowerblue", bc = "pink")) +
  theme_classic() +
  geom_vline(xintercept = 33.5) +
  facet_wrap(~ study, ncol = 1) +  # Adjust ncol as needed
  theme_classic() 
p
BackupAsPDF(p, 'baitdegree.barplots.')

```
Do again with the degree
```{r fig.width = 4, fig.height = 3}
p <- ggplot (baitDegrees, aes(x = baitDegree, fill = study, col=study))  + 
  geom_dotplot(method = "histodot", binwidth = 1.5) +
  theme_classic() +
  geom_vline(xintercept = 33.5) +
  xlab ("Bait Degree") +
 # ylab ("Count") +
  facet_wrap(~study, scales = 'free_y')  +
  scale_color_manual(values = c(kaushik = "orange", pcmi = "cornflowerblue", bc = "pink")) +
  theme_bw() + 
  theme(legend.position="top")
p 

BackupAsPDF(p, 'exp.baitdegree.histograms.')
```


Redo excluding those with preys with bait degree above 33 (too broad? why do we do this? exclude so we have a more consistent overlap?

```{r}
#simplify name
bc.ppi <- bc.ppi.edge %>% 
  .[, baitsPerPrey := .N, by = Prey]
```

Read in other PPI datasets

```{r}
pcmi.ppi <- fread ("./data/D_Final_v3.txt")
pcmi.ppi <- pcmi.ppi[, .(Bait, Prey = PreyGene, Clusters)]
pcmi.ppi[, baitsPerPrey := .N, by = Prey]

k.ppi <- fread("./data/kaushikPPI.csv")[hc == TRUE]
k.ppi[, baitsPerPrey := .N, by = Prey]
```

To me this looks a little sparse. Talk to Ben and redo the analysis if needed. Discuss breakdown of the BC PPI data; we have less baits; have I filtered this data correctly? Reinspect the cleaning and 

```{r}
baitDegrees <- rbindlist (list (kaushik = k.ppi[, .(baitDegree = length(unique(Prey))), by = Bait],
                                pcmi = pcmi.ppi[, .(baitDegree = length(unique(Prey))), by = Bait],
                                bc      = bc.ppi[, .(baitDegree = length(unique(Prey))), by = Bait]),
                          idcol = "study")

threshold = 33

excludeBaits <- baitDegrees[baitDegree > threshold, Bait]


p <- ggplot (baitDegrees[!Bait %in% excludeBaits], aes(x = baitDegree, fill = study, color = study))  + 
  geom_dotplot(method = "histodot", binwidth = 0.75) +
  #geom_density(color = "black") +
  facet_wrap(~study, ncol = 1) +
  coord_cartesian(ylim = c(0,0.1)) +
  theme_classic() +
  ggtitle(sprintf ("Only baits with <= %d prey", threshold))
p

BackupAsPDF(p, 'exp.bait.degrees.exclude.n>33')

```

## overview 
PCMI: 100 baits, 1074 prey, 1881 interactions
```{r}
pcmi.ppi[,.(length(unique(Bait)), length(unique(Prey)), .N)]
```
Kaushik: 90 baits, 1122 prey, 1462 interactions
```{r}
k.ppi[, .(length(unique(Bait)), length(unique(Prey)), .N)]
```
BC: 39 baits, 499 prey, 730 interactions
BC exp data : 39 baits, 493 prey, 594 interactions
Seems v low... need to investigate if correct
```{r}
bc.ppi[, .(length(unique(Bait)), length(unique(Prey)), .N)]
```


## Thresholded data overview
Lets look at unique prey and baits in each of the. data sets
PCMI: 82 baits, 659 prey, 933 interactions
```{r}
pcmi.ppi[!Bait %in% excludeBaits, .(length(unique(Bait)), length(unique(Prey)), .N)]
```
Kaushik: 90 baits, 1122 prey, 1462 interactions
```{r}
k.ppi[!Bait %in% excludeBaits, .(length(unique(Bait)), length(unique(Prey)), .N)]
```
BC: 33 baits, 316 prey, 391 interactions
BC exp data: 34 bait, 308 prey, 346 interactions

Seems v low... need to investigate if correct
```{r}
bc.ppi[!Bait %in% excludeBaits, .(length(unique(Bait)), length(unique(Prey)), .N)]
```



# run a pairwise fisher exact test 
Look for significance in overlap between all bait prey sets in the dataset

```{r}
#  File from Belinda, all HEK detectable proteins + PCMI prey
#scan("/Users/ben/Downloads/hek293tProteome_BekkerJenwen2017_ASDppiPrey.csv", what = character()) |> unique() |> length()
# 11169


pairwiseFisher <- function (bait1, bait2, matrix){
  both <- sum(rowSums(matrix[, c(bait1, bait2)]) == 2)
  firstOnly  <- sum(matrix[, c(bait1)]) - both
  secondOnly <- sum(matrix[, c(bait2)]) - both
  neither <- 11169 - both - firstOnly  -secondOnly
  #neither <- nrow(matrix) - both- firstOnly- secondOnly
  
  
  
  fisher.test (matrix(c(both, firstOnly, secondOnly, neither), nrow = 2), alternative = "greater")$p.value
}

```

PCMI fihser res mat

```{r}
# binary mat of prey membership per bait
pcmi.mat <- dcast (pcmi.ppi, Prey~Bait, value.var = "Bait", fun.aggregate = length) |> as.matrix(rownames= "Prey") 
pcmi.mat[pcmi.mat > 1] <- 1
#pcmi.mat

allByAll <- data.table(bait1 = colnames(pcmi.mat))[, .(bait2 = colnames(pcmi.mat)), by = bait1][bait1 < bait2] # v useful with character vecotors; filter out duplicates & self pairings
allByAll <- allByAll[ (!bait1 %in% excludeBaits) & (!bait2 %in% excludeBaits)]


allByAll[ , fisherP := pairwiseFisher(bait1, bait2, pcmi.mat), by = .(bait1, bait2)]

allByAll[, pP := -log10(fisherP)]
pcmi.fisher.dt <- allByAll

pcmi.fisher.mat <- dcast (rbind (allByAll, allByAll[, .(bait1 = bait2, bait2 = bait1, fisherP, pP)]),
       bait1~bait2, value.var = "pP") |> as.matrix (rownames = "bait1")


baitClusters <- unique(pcmi.ppi[, .(Bait, Clusters)])


hm <- Heatmap (pcmi.fisher.mat,
         name = "-log10(P)\nfisher.test",
         row_split = as.factor(baitClusters[rownames(pcmi.fisher.mat), Clusters, on = "Bait"]),
         column_split = as.factor(baitClusters[colnames(pcmi.fisher.mat), Clusters, on = "Bait"]),
         col = circlize::colorRamp2(breaks = c(0, 4), color = c("white", "firebrick")),
         left_annotation = rowAnnotation(cluster = as.factor(baitClusters[rownames(pcmi.fisher.mat), Clusters, on = "Bait"]),
                                         col = list (cluster = setNames(c("#FF3333","#FF9900","#F6D80C","#0CB212","#4A56FD","#AA4AFD","#1CDDEB","black"), 1:8))),
         top_annotation = columnAnnotation(cluster = as.factor(baitClusters[rownames(pcmi.fisher.mat), Clusters, on = "Bait"]),
                                         col = list (cluster = setNames(c("#FF3333","#FF9900","#F6D80C","#0CB212","#4A56FD","#AA4AFD","#1CDDEB","black"), 1:8))),
         border  = TRUE,
         row_names_gp = gpar(fontsize = 2),
         column_names_gp = gpar(fontsize = 2),
         gap = unit(0.3, "mm"),
         column_gap = unit(0.3, "mm"))

hm
BackupAsPDF(hm, 'pcmi.heatmap.fisher.pval.clusters')
```
kaushik mat
```{r, fig.width = 5, fig.height= 4}
kaushik.mat <- dcast (k.ppi, Prey~Bait, value.var = "Bait", fun.aggregate = length) |> as.matrix(rownames= "Prey")

allByAll <- data.table(bait1 = colnames(kaushik.mat))[, .(bait2 = colnames(kaushik.mat)), by = bait1][bait1 < bait2]

allByAll <- allByAll[ (!bait1 %in% excludeBaits) & (!bait2 %in% excludeBaits)]

allByAll[ , fisherP := pairwiseFisher(bait1, bait2, kaushik.mat), by = .(bait1, bait2)]

allByAll[, pP := -log10(fisherP)]

kaushik.fisher.dt <- allByAll

kaushik.fisher.mat <- dcast (rbind (allByAll, allByAll[, .(bait1 = bait2, bait2 = bait1, fisherP, pP)]),
       bait1~bait2, value.var = "pP") |> as.matrix (rownames = "bait1")


hm <- Heatmap (kaushik.fisher.mat, 
         name = "-log10(P)\nfisher.test",
         col = circlize::colorRamp2(breaks = c(0, 8), color = c("white", "firebrick")))

BackupAsPDF(hm, 'kaushik.heatmap.fisher.pval.clusters')
```
Repeat the same test for the BC PPI

```{r}
bc.mat <- dcast (bc.ppi, Prey~Bait, value.var = "Bait", fun.aggregate = length) |> as.matrix(rownames= "Prey")
bc.mat[bc.mat > 1] <- 1

allByAll <- data.table(bait1 = colnames(bc.mat))[, .(bait2 = colnames(bc.mat)), by = bait1][bait1 < bait2]

allByAll <- allByAll[ (!bait1 %in% excludeBaits) & (!bait2 %in% excludeBaits)]

allByAll[ , fisherP := pairwiseFisher(bait1, bait2, bc.mat), by = .(bait1, bait2)]

allByAll[, pP := -log10(fisherP)]

bc.fisher.dt <- allByAll

bc.fisher.mat <- dcast (rbind (allByAll, allByAll[, .(bait1 = bait2, bait2 = bait1, fisherP, pP)]),
       bait1~bait2, value.var = "pP") |> as.matrix (rownames = "bait1")


hm <- Heatmap (bc.fisher.mat, 
         name = "-log10(P)\nfisher.test",
         col = circlize::colorRamp2(breaks = c(0, 8), color = c("white", "firebrick")))


BackupAsPDF(hm, 'bc.heatmap.fisher.pval.clusters')
```

```{r}

pdf("/Users/martingordon/Documents/projects/060823_BPolacco_TryKPPI/exp.fisher.-log10pval.pdf")
hist(pcmi.fisher.mat[pcmi.fisher.mat > 0], breaks = 100)
hist(kaushik.fisher.mat[kaushik.fisher.mat > 0], breaks = 100, add = TRUE, col = "red")
hist(bc.fisher.mat[bc.fisher.mat > 0], breaks = 100, add = TRUE, col = "blue")
dev.off()

```

The pvalue distriubtion for BC a little concerning; no increase in density ~-log10(0.05)
```{r}
fisher.dt <- rbindlist(list (pcmi= pcmi.fisher.dt, kaushik = kaushik.fisher.dt, bc = bc.fisher.dt), idcol = "study")

ggplot(fisher.dt, aes (x = pP, color = study)) + geom_density(alpha  = 0.5) +theme_bw() 
ggplot(fisher.dt, aes (x = pP, color = study)) + geom_density(alpha  = 0.5) +theme_bw() + coord_cartesian(ylim = c(0, 0.1)) 

# save this
g <- ggplot(fisher.dt, aes (x = pP, color = study)) + geom_density(alpha  = 0.5) +theme_bw() + coord_cartesian(xlim = c(0, 5), ylim = c(0, 0.1)) + geom_vline(xintercept= -log10(0.05))
BackupAsPDF(g, 'exp.fisher.pval.densityplot.')
```


```{r fig.width = 4, fig.height= 3.5}
p <- ggplot (fisher.dt[, 100 *sum(fisherP < 0.05)/.N, by = study], aes(x = study, y= V1, fill = study)) +
  geom_hline(yintercept = 5, color = "grey") +
  geom_col(show.legend = FALSE, width = 0.5) +
  ylab( "% of bait pairs with significant (p < 0.05)\n prey convergence") +
  scale_fill_manual(values = c(kaushik = "orange", pcmi = "cornflowerblue", bc = 'pink')) +
  theme_classic() +
  coord_cartesian(ylim = c(2,10))
p
BackupAsPDF(p, 'perc.sig.baitpairs.barplot.')
```

Compare these results to BP
Percentage of significant FET from each of the studies

```{r}
fisher.dt[, 100 * sum(fisherP < 0.05) / .N, by=study]
```


#22/08/23
Threshold PCMI at 33 interactors and look at everything else

kaushik mat
```{r, fig.width = 5, fig.height= 4}
kaushik.mat <- dcast (k.ppi, Prey~Bait, value.var = "Bait", fun.aggregate = length) |> as.matrix(rownames= "Prey")

allByAll <- data.table(bait1 = colnames(kaushik.mat))[, .(bait2 = colnames(kaushik.mat)), by = bait1][bait1 < bait2]

allByAll[ , fisherP := pairwiseFisher(bait1, bait2, kaushik.mat), by = .(bait1, bait2)]

allByAll[, pP := -log10(fisherP)]

kaushik.fisher.dt <- allByAll
```

```{r}
bc.mat <- dcast (bc.ppi, Prey~Bait, value.var = "Bait", fun.aggregate = length) |> as.matrix(rownames= "Prey")
bc.mat[bc.mat > 1] <- 1

allByAll <- data.table(bait1 = colnames(bc.mat))[, .(bait2 = colnames(bc.mat)), by = bait1][bait1 < bait2]

allByAll[ , fisherP := pairwiseFisher(bait1, bait2, bc.mat), by = .(bait1, bait2)]

allByAll[, pP := -log10(fisherP)]

bc.fisher.dt <- allByAll
```

```{r  fig.width = 5, fig.height= 4}
fisher.dt <- rbindlist(list (pcmi= pcmi.fisher.dt, kaushik = kaushik.fisher.dt, bc = bc.fisher.dt), idcol = "study")

ggplot(fisher.dt, aes (x = pP, color = study)) + geom_density(alpha  = 0.5) +theme_bw() 
ggplot(fisher.dt, aes (x = pP, color = study)) + geom_density(alpha  = 0.5) +theme_bw() + coord_cartesian(ylim = c(0, 0.1)) 

# save this
BackupAsPDF(g, 'exp.fisher.pval.densityplot.')

p <- ggplot (fisher.dt[, 100 *sum(fisherP < 0.05)/.N, by = study], aes(x = study, y= V1, fill = study)) +
  geom_hline(yintercept = 5, color = "grey") +
  geom_col(show.legend = FALSE, width = 0.5) +
  ylab( "% of bait pairs with significant (p < 0.05)\n prey convergence") +
  scale_fill_manual(values = c(kaushik = "orange", pcmi = "cornflowerblue", bc = 'pink')) +
  theme_classic() +
  coord_cartesian(ylim = c(2,10))
p
BackupAsPDF(p, 'exp.pcmi.thresh.perc.sig.baitpairs.barplot.')

```



# pairwise overlaps

baitPrey matrix 
similiarity mat based on jaccard distance between bait prey sets
Cluster similiar bait profiles based on their target prey


```{r, fig.width =8, fig.height = 5}
# cast the cols but fun.aggreagte calculates unique counts of each bait prey combo
pcmi.mat <- dcast (pcmi.ppi, Prey~Bait, value.var = "Bait", fun.aggregate = length) |> as.matrix(rownames= "Prey")
# convert to binary
pcmi.mat[pcmi.mat > 1] <- 1
Heatmap(pcmi.mat, col = c("white", "firebrick"), column_names_gp = gpar(fontsize =5))
```

```{r, fig.width =8, fig.height = 5}
kaushik.mat <- dcast (k.ppi, Prey~Bait, value.var = "Bait", fun.aggregate = length) |> as.matrix(rownames= "Prey")
Heatmap(kaushik.mat, col = c("white", "firebrick"), column_names_gp = gpar(fontsize =5))
```

```{r, fig.width =8, fig.height = 5}
bc.mat <- dcast (bc.ppi, Prey~Bait, value.var = "Bait", fun.aggregate = length) |> as.matrix(rownames= "Prey")
bc.mat[bc.mat > 1] <- 1
Heatmap(bc.mat, col = c("white", "firebrick"), column_names_gp = gpar(fontsize =5))
```

Distance matrix' what baits are the most similar? 
```{r, fig.width = 5, fig.height  = 4}
pcmi.jaccard <- dist (t(pcmi.mat), method = "binary") |> as.matrix()
Heatmap (1-pcmi.jaccard, col = circlize::colorRamp2(breaks = c(0, 0.4), colors =  c("white", "firebrick")))
```

```{r, fig.width = 5, fig.height  = 4}
kaushik.jaccard <- dist (t(kaushik.mat), method = "binary") |> as.matrix()
Heatmap (1-kaushik.jaccard, col = circlize::colorRamp2(breaks = c(0, 0.4), colors =  c("white", "firebrick")))
```

```{r, fig.width = 5, fig.height  = 4}
bc.jaccard <- dist (t(bc.mat), method = "binary") |> as.matrix()
Heatmap (1-bc.jaccard, col = circlize::colorRamp2(breaks = c(0, 0.4), colors =  c("white", "firebrick")))
```

