---
title: "062724_AgonistPWComparisons"
author: "Martin Gordon"
date: "2024-06-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In addition to the earlier contrasts, we also need to run a PW comparison of the agonist vs w/o agonist groups
(same background (cells) in this case, so hopefully cleaner signal?)


Maybe todo: attach a linechart of the endogenous biotincarboxylases
Try find more of these 
Tidy the plots to make our point for the differential analysis
Finish regenerating the plots

```{r cars}
library(magrittr)
library(data.table)
library(circlize)
library(ComplexHeatmap)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(MSstats)
library(ggrepel)
library(readxl)
library(RColorBrewer)
library(cluster) # pam clustering of genes
library(eulerr) # eulerr plot 
library(ggvenn)
library(scales)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

source("../../utils/mg_utils/r_utils/IDmapping.R")
source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("../../utils/mg_utils/r_utils/HelperFunctions.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0,...){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```

Run the pw comparisons
```{r}
p.quant <- fread('~/Documents/projects/061324_XFang_Cilia_GPCR/062424_PWComparisons_data/2024_06_26_ProteinLevelQuant.renorm.csv')
dp.out <- readRDS('./062424_PWComparisons_data/2024_06_24_dp.out.rds')

f.quant <- setDT(dp.out$FeatureLevelData)

p.quant[, SUBJECT := paste0('batch.',SUBJECT)]
p.quant[, LogIntensities := newLogInts]
f.quant[, SUBJECT := paste0('batch.',SUBJECT)]

dp.out$FeatureLevelData <- f.quant
dp.out$ProteinLevelData <- p.quant
```

ID mappings file 

```{r}
# lets map the homologs to the mss output and write the complete and tidied data to file 
mh_mappings <- fread('~/Documents/utils/mg_utils/data/mouseHumanIDConversion.txt')
```

Run the comparisons w/wo agonist
```{r}
grep('_In', p.quant$GROUP, value=T) %>%  unique()

contrasts.list <- list('+cilia_SMO_ag vs +cilia_SMO' = data.table(V1='+cilia_SMO_ag', V2='+cilia_SMO'),
                       '+cilia_GPR161_ag vs +cilia_GPR161' = data.table(V1='+cilia_GPR161_ag', V2='+cilia_GPR161'),
                       '+cilia_Pth2r_ag vs +cilia_Pth2r' = data.table(V1='+cilia_Pth2r_ag', V2='+cilia_Pth2r'),
                       '+cilia_Crhr2_ag vs +cilia_Crhr2' = data.table(V1='+cilia_Crhr2_ag', V2='+cilia_Crhr2'),
                       '+cilia_Pde8A_In vs +cilia_Pde8A' = data.table(V1='+cilia_Pde8A_In', V2='+cilia_Pde8A'),
                       '+cilia_Pde8B_In vs +cilia_Pde8B' = data.table(V1='+cilia_Pde8B_In', V2='+cilia_Pde8B'))

contrasts.mat <- MSstats::MSstatsContrastMatrix(contrasts.list, 
                               conditions = unique(p.quant$GROUP), 
                               labels = names(contrasts.list))
```

Run the comparisons w/wo agonists

```{r}
# run msstats correcting for batch 
mss <- groupComparison(contrast.matrix=contrasts.mat, 
                       verbose=T,
                       data=dp.out)

mss.dt <- setDT(mss$ComparisonResult)
```

process the msstats output
```{r}
# write out raw results
mss.dt[, gene := multiUniprots2multiGenes(as.character(Protein), species = 'MOUSE')]
mss.dt[, p.adj := p.adjust(pvalue, method='BH'), by=.(Label)]
mss.dt[, sig := 'not']
mss.dt[abs(log2FC) > 0.58 & pvalue < 0.005, sig := ifelse(log2FC > 0, 'up', 'down')]

mss.dt[abs(log2FC) > 0.58 & pvalue < 0.005,]

# merge with human ID table
anno.dt <- merge(mss.dt, unique(mh_mappings[,.(SWISS_PROT_IDs.mouse, SWISS_PROT_IDs.human,Symbol.human)], by='SWISS_PROT_IDs.mouse'), by.x='Protein', by.y='SWISS_PROT_IDs.mouse', all.x=T, all.y=F)
```

Write out the cilia comparisons to file, plot quickly and generate heatmaps
log2FC > 0.58 * pval < 0.05

```{r}
fwrite(anno.dt, ScriptAndDatedFileName('mss.pwComparisons.agonistComparisons.csv'))
fwrite(dcast(anno.dt, gene+Protein~Label, value.var = c('log2FC','pvalue', 'p.adj')), ScriptAndDatedFileName('mss.pwComparisons.agonistComparisons.wide.csv'))
```
plot of N sig hits
V low number of hits with additon of agonist... not great results...maybe I've washed out the biological signal by renormalizing?
Need to return to the ori output and process

```{r}
g <- ggplot(mss.dt[sig != 'not', .N, by=.(sig,Label)], aes(x=reorder(Label,-N), y=N, fill=sig)) +
  geom_bar(stat='Identity') +
  ggtitle('Significant hits per contrast (log2FC > 0.58 & pvalue < 0.005)') +
  scale_fill_manual(values=c('down'=muted('blue'), 'up'=muted('red'))) +
  ylab('Number of significant hits') +
  xlab('Contrasts') +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90,size=9))
g

BackupAsPDF(g, 'N.sigHits.agonists.barplots',dimensions = c(8,9))
```

Heatmap of the significant hits..
Nothing really of interest I think...
```{r}
col.pal <- randomcoloR::distinctColorPalette(k=length(unique(p.quant$GROUP)))

sig.prots <- mss.dt[sig != 'not', unique(Protein)]

p.mat <- dcast(p.quant, Protein~paste0(GROUP,'.', SUBJECT), value.var='newLogInts') %>% 
  as.matrix(rownames='Protein')

submat <-  sweep(p.mat, 1, apply(p.mat,1 ,median,na.rm=T))

submat <- submat[rownames(submat) %in% sig.prots, grep('[-]cilia', colnames(submat), invert=T)]

colnames(submat) <- gsub('[.]batch','', colnames(submat))
rownames(submat) <- multiUniprots2multiGenes(rownames(submat), species='MOUSE')

# need to adjust the column order 

#make a column annotation to show the differences in the groups
# create annotation
ann <- data.table(group = colnames(submat))
ann[, cilia_present := ifelse(grepl('-cilia', group), 'No', 'Yes')]
ann[, 'agonist/inhibitor' := ifelse(grepl('_In[.][0-9]{1}$|_ag[.][0-9]{1}$', group), 'Yes', 'No')]
ann[, status := gsub("[-+]{1}cilia_|_ag[.][0-9]{1}$|[.][0-9]{1}$|_In[.][0-9]{1}$",'', group)]
ann[, rep := str_extract(group, '[0-9]$')]

# unique col palette
col.pal <- randomcoloR::distinctColorPalette(k=length(unique(ann$status)))
names(col.pal) <- unique(ann$status)   

colours <- list('agonist/inhibitor' = c('Yes' = '#EE6677', 'No'='#228833'),
                'status' = col.pal)
   
colAnn <- HeatmapAnnotation(df = ann[,.(status,`agonist/inhibitor_present`)], col= colours)

hm <- Heatmap(submat,
        cluster_columns=F,
        cluster_column_slices = F,
        border=T,
        name='Ints/Median',
        row_title = sprintf('%s Proteins (FC +/- 50%% & pval < 0.005)', nrow(submat)),
        top_annotation = colAnn,
        column_names_gp = gpar(fontsize=8),
        column_title_gp = gpar(fontsize=10),
        row_names_gp = gpar(fontsize=7),
        column_split = list(ann$status))
                            #gsub('[+-]{1}cilia_|[.][0-9]$','', colnames(submat))))
      
hm
BackupAsPDF(hm, 'sigProts.agonistContrast.FC1.5.pval0.005.heatmap', dimensions=c(14,11))


# subset of just the contrasted pairs in the analysis
hm <- Heatmap(submat,
        cluster_columns=F,
        cluster_column_slices = F,
        border=T,
        name='Ints/Median',
        row_title = sprintf('%s Proteins (FC +/- 50%% & pval < 0.005)', nrow(submat)),
        top_annotation = colAnn,
        column_names_gp = gpar(fontsize=8),
        column_title_gp = gpar(fontsize=10),
        row_names_gp = gpar(fontsize=7),
        column_split = list(ann[grep('PM|ARL13b|GPR124|GPR135', colnames(submat), invert=T), status]))


colAnn
      
BackupAsPDF(hm, 'sigProts.agonistContrast.samplesCompared.FC1.5.pval0.005.heatmap', dimensions=c(14,11))
```
volcanoplots 

```{r}
g <- ggplot(mss.dt, aes(x=log2FC, y=-log10(pvalue), col=sig, label=gene)) +
  geom_point() + 
  ylab('-log10 adjusted p-value') +
  ggrepel::geom_text_repel(data=mss.dt[sig != 'not'], show.legend = FALSE, size = 4, max.overlaps = 20) +
  geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
  geom_hline(yintercept = -log10(0.005), linetype=2, alpha=0.4) +
  ggtitle('Agonists vs No agonist') +
  scale_color_manual(values=c('up'='#990033', 'down'='#336699', 'not'='grey')) +
  facet_wrap(~Label, scales='free_x') +
  theme_bw()
g
BackupAsPDF(g, 'contrasts.agonists.volcano', dimensions=c(14,14))
```
Run enrichment of these proteins 

GO CC enrichment first
```{r}
gmt.go <- loadGmtFromBioconductor(dbName='org.Mm.eg.db', ontology='CC', keyType='UNIPROT')
```

```{r}
# define the universe, the total set of identified genes in our study
universe <- as.character(unique(p.quant$Protein))

mss.dt[, enrich.grp := interaction(Label,sig)]

enrich.dt <- enricherOnGroups(mss.dt[sig != 'not'], 
                              groupColumns = 'enrich.grp', 
                              geneColumn = "Protein", 
                              term2gene.gmt = gmt.go, 
                              universe = universe)

fwrite(enrich.dt, ScriptAndDatedFileName('agonistContrasts.GO.CC.enrichments.csv'))

simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich.grp')
fwrite(simp.enrich$simplified, ScriptAndDatedFileName('agonistContrasts.GO.CC.simplified.enrichments.csv'))

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, 
                                groupColumn = 'enrich.grp', 
                                topN = 8,
                                title='GO Cellular Component', 
                                row_names_gp = gpar(fontsize = 7), 
                                negCols=unique(grep('down', simp.enrich$simplified$enrich.grp, value=T)),
                                column_names_gp= gpar(fontsize = 6), 
                                upperThreshold = 8)
ht
BackupAsPDF(ht, 'agonistContrast.GOCC.enrichments.heatmap', dimensions=c(10,10))
```
Run GO BP enrichment

```{r}
gmt.go <- loadGmtFromBioconductor(dbName='org.Mm.eg.db', ontology='BP', keyType='UNIPROT')
```

```{r}
# define the universe, the total set of identified genes in our study
universe <- as.character(unique(p.quant$Protein))

mss.dt[, enrich.grp := interaction(Label,sig)]

enrich.dt <- enricherOnGroups(mss.dt[sig != 'not'], 
                              groupColumns = 'enrich.grp', 
                              geneColumn = "Protein", 
                              term2gene.gmt = gmt.go, 
                              universe = universe)

enrich.dt
# this is a very small set, just save the output
# cant simplify further...
fwrite(enrich.dt, ScriptAndDatedFileName('agonistContrasts.GO.BP.enrichments.csv'))
```

Plot pka mouse protein 

```{r}
gmt.go <- loadGmtFromBioconductor(dbName='org.Mm.eg.db', ontology='CC', keyType='UNIPROT')

```

