---
title: "062724_PWComparisons_Review"
author: "Martin Gordon"
date: "2024-06-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis Review
Want to review the analysis performed presented at the cilia GPCR meeting 06-27-24
I think the first question is our normalization method... can we adjust this to correct for the effect
i) Do the included samples make sense? Could we take the avg of the techRep groups?
ii) Look at comparisons with more 'gentle' normalization. (check out the SMO group and visualise the PKA interaction; can we see more obvious differences between the groups with the different normalization methods)


Tidy the plots and produce the final set of PW comparisons for Aaron


Sare these plots with the group
Also share overlaps with Max Nauchery annotated cilia proteins (high confidence and 2021 proteins only for now) 

```{r cars}
library(magrittr)
library(data.table)
library(circlize)
library(ComplexHeatmap)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(MSstats)
library(ggrepel)
library(readxl)
library(RColorBrewer)
library(cluster) # pam clustering of genes
library(eulerr) # eulerr plot 
library(ggvenn)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

source("../../utils/mg_utils/r_utils/IDmapping.R")
source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("../../utils/mg_utils/r_utils/HelperFunctions.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0,...){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```

Read in the raw and normalized datasets and look at the output for PKA proteins and peptides

```{r}
dp.out <- readRDS('~/Documents/projects/061324_XFang_Cilia_GPCR/062424_PWComparisons_data/2024_06_24_dp.out.rds')
old.pQuant <- setDT(dp.out$ProteinLevelData)
old.pQuant[, gene := multiUniprots2multiGenes(as.character(Protein), species='MOUSE')]

p.Quant <- fread('./062424_PWComparisons_data/2024_06_26_ProteinLevelQuant.renorm.csv')
p.Quant[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]

spec <-  fread('./data/CR1/Re-run/20240618_164945_CR1_rep5_Report.tsv')
# create some new ID columns 
spec[, agonist := ifelse(grepl('_ag$', Condition), 'Yes','No')]

spec <- spec[Intensity > 2^4,]

spec[, gene := multiUniprots2multiGenes(ProteinName, species='MOUSE')]
```

plot a comparison of the impact of normalization
I think we want to extract to look at 
i) distributions of the cilia annotated proteins in all 3 sets and 
ii) the Pkaca subunit

interactors of Pkaca

## first look at the Pka subunit
```{r}
col.pal <-  randomcoloR::distinctColorPalette(k=length(unique(ints.dt$GROUP)))

genes.oi <- c('Prkaca') #, 'Prkar1a')
```

we want to combine the 3 datasets into one, first subsettting to the things we are interested in

```{r}
spec[gene %in% genes.oi, .(GROUP = Condition, SUBJECT = BioReplicate, gene, LogIntensities = log2(Intensity))]

ints.dt <- list('MSSNorm'= old.pQuant[gene %in% genes.oi, .(Protein, GROUP, SUBJECT, gene, LogIntensities)],
                  'AdditionalRoundNorm'= p.Quant[gene %in% genes.oi, .(Protein, GROUP, SUBJECT, gene, LogIntensities = newLogInts)],
                  'Raw' = spec[gene %in% genes.oi, .(GROUP = Condition, SUBJECT = BioReplicate, gene, Protein = ProteinName, PeptideSequence, LogIntensities = log2(Intensity))]
                  ) %>% rbindlist(idcol='source', fill=T)


#add a status col
ints.dt[, probe := gsub('[+-]{1}cilia_|[.][0-9]{1}','', GROUP)]
ints.dt[, cilia_present := ifelse(grepl('[+]cilia', GROUP), 'yes', 'no')]
ints.dt[, agonist_present := ifelse(grepl('_ag', GROUP), 'yes', 'no')]


#plot the intensity values of the different approaches
g <- ggplot(ints.dt[(gene %in% genes.oi & probe %like% 'SMO'),], aes(x=source, y=LogIntensities, fill=cilia_present)) +
  geom_boxplot(outliers=F) +
  geom_point(aes(fill = cilia_present), size = 1, shape = 21, alpha=0.7, position = position_jitterdodge()) +
  ggtitle('PKA interaction') +
  facet_wrap(~probe, scales = 'free_y') +
  scale_fill_brewer(type='qual', palette=1) +
  scale_y_continuous(breaks=seq(1,20,1)) +
  theme_bw()
g
BackupAsPDF(g, 'smo.posControl.normalizationComparision.boxplot', dimensions=c(10,8))


# now plot the protein on ciliated cells only
#plot the intensity values of the different approaches
g <- ggplot(ints.dt[(gene %in% genes.oi & probe %like% 'SMO' & cilia_present == 'yes'),], aes(x=source, y=LogIntensities, fill=agonist_present)) +
  geom_boxplot(outliers = F) +
  geom_point(aes(fill = agonist_present), size = 1, shape = 21, alpha=0.7, position = position_jitterdodge()) +
  ggtitle('SMO probe PKA interaction') +
  scale_fill_brewer(type='qual', palette=1) +
  scale_y_continuous(breaks=seq(1,20,1)) +
  theme_bw()
g
BackupAsPDF(g, 'smo.posControl.ciliaOnly.normalizationComparision.boxplot', dimensions=c(10,8))

```
Visualize the literature annotated cilia proteins from GO
```{r}
gmt.go <- loadGmtFromBioconductor(dbName='org.Mm.eg.db', ontology='CC', keyType='UNIPROT')
cilium.genes <- gmt.go[ont == 'cilium', gene]

#plot the intensity values of the different approaches
g <- ggplot(ints.dt[Protein %in% cilium.genes,], aes(x=source, y=LogIntensities, fill=cilia_present)) +
  geom_boxplot() +
  geom_point(aes(fill = cilia_present), size = 1, shape = 21, alpha=0.7, position = position_jitterdodge()) +
  ggtitle('PKA interaction') +
  facet_wrap(~probe, scales = 'free_y') +
  scale_fill_manual(values=col.pal[c(3,6)]) +
  theme_bw()
g
BackupAsPDF(g, 'smo.posControl.normalizationComparision.boxplot', dimensions=c(10,8))
```

GO annotated protein distributions 

```{r}
ints.dt <- list('MSSNorm'= old.pQuant[Protein %in% cilium.genes, .(Protein, GROUP, SUBJECT, gene, LogIntensities)],
                  'AdditionalRoundNorm'= p.Quant[Protein %in% cilium.genes, .(Protein, GROUP, SUBJECT, gene, LogIntensities = newLogInts)],
                  'Raw' = spec[ProteinName %in% cilium.genes, .(GROUP = Condition, SUBJECT = BioReplicate, gene, Protein = ProteinName, PeptideSequence, LogIntensities = log2(Intensity))]
                  ) %>% rbindlist(idcol='source', fill=T)

#add a status col
ints.dt[, probe := gsub('[+-]{1}cilia_|[.][0-9]{1}','', GROUP)]
ints.dt[, cilia_present := ifelse(grepl('[+]cilia', GROUP), 'yes', 'no')]
ints.dt[, agonist_present := ifelse(grepl('_ag', GROUP), 'yes', 'no')]



#plot the intensity values of the different approaches
g <- ggplot(ints.dt[Protein %in% cilium.genes,], aes(x=source, y=LogIntensities, fill=cilia_present)) +
  geom_boxplot() +
 # geom_point(aes(fill = cilia_present), size = 1, shape = 21, alpha=0.7, position = position_jitterdodge()) +
  ggtitle(sprintf('GO annotated cilia protein distributions (n=%s)', sum(cilium.genes %in% unique(spec$ProteinName)))) +
  facet_wrap(~probe, ncol=3, scales = 'free_y') +
  scale_fill_manual(values=col.pal[c(3,6)]) +
  theme_bw()
g
BackupAsPDF(g, 'GOcilia.distributions.boxplot', dimensions=c(12,18))
```
Now plot the distributions of Max Nauchery cilia annotated protein distributions (high confidence ones)

```{r}
read_xlsx(path='./docs/ciliaProteins.Nachury.etal.2021.xlsx', sheet=1, skip = 1)
#read in literature sig hits
cilia.Nachury.2021 <- read_xlsx(path='./docs/ciliaProteins.Nachury.etal.2021.xlsx', sheet=1, skip = 1)$`Gene Symbol`
cilia.Nachury.2021 <- cilia.Nachury.2021[!is.na(cilia.Nachury.2021)]

cilia.Nachury.2015.highConf <- read_xlsx(path='./docs/Nachury.etal.2015cilliaProteins.xlsx', sheet=1)$`Uniprot accession number`
cilia.Nachury.2015.highConf <- multiUniprots2multiGenes(cilia.Nachury.2015.highConf, species='MOUSE')

cilia.literature <- list('Nachury et al 2021' = cilia.Nachury.2021,
                         'Nachury et al 2015 (High Conf)' = cilia.Nachury.2015.highConf)
```



```{r}
ints.dt <- list('MSSNorm'= old.pQuant[, .(Protein, GROUP, SUBJECT, gene, LogIntensities)],
                  'AdditionalRoundNorm'= p.Quant[, .(Protein, GROUP, SUBJECT, gene, LogIntensities = newLogInts)],
                  'Raw' = spec[, .(GROUP = Condition, SUBJECT = BioReplicate, gene, Protein = ProteinName, PeptideSequence, LogIntensities = log2(Intensity))]
                  ) %>% rbindlist(idcol='source', fill=T)

#add a status col
ints.dt[, probe := gsub('[+-]{1}cilia_|[.][0-9]{1}','', GROUP)]
ints.dt[, cilia_present := ifelse(grepl('[+]cilia', GROUP), 'yes', 'no')]
ints.dt[, agonist_present := ifelse(grepl('_ag', GROUP), 'yes', 'no')]


lapply(names(cilia.literature), function(i,x){

  g <- ggplot(x[gene %in% cilia.literature[[i]],], aes(x=source, y=LogIntensities, fill=cilia_present)) +
  geom_boxplot() +
  #geom_point(aes(fill = cilia_present), size = 1, shape = 21, alpha=0.7, position = position_jitterdodge()) +
  ggtitle(paste0(i, ' cilla protein distributions')) +
  facet_wrap(~probe, scales = 'free_y', ncol=3) +
  scale_fill_brewer(type='qual', palette='Dark2') +
  theme_bw()
  g
  BackupAsPDF(g, paste0(i, '.curatedCilla.boxplots'), dimensions=c(14,22))
},x=ints.dt)
```

