---
title: "062724_PWComparisons_Review"
author: "Martin Gordon"
date: "2024-06-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis Review
Want to review the analysis performed presented at the cilia GPCR meeting 06-27-24
I think the first question is our normalization method... can we adjust this to correct for the effect
i) Do the included samples make sense? Could we take the avg of the techRep groups?
ii) Look at comparisons with more 'gentle' normalization. (check out the SMO group and visualise the PKA interaction; can we see more obvious differences between the groups with the different normalization methods)


Tidy the plots and produce the final set of PW comparisons for Aaron


Sare these plots with the group
Also share overlaps with Max Nauchery annotated cilia proteins (high confidence and 2021 proteins only for now) 

```{r cars}
library(magrittr)
library(data.table)
library(circlize)
library(ComplexHeatmap)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(MSstats)
library(ggrepel)
library(readxl)
library(RColorBrewer)
library(cluster) # pam clustering of genes
library(eulerr) # eulerr plot 
library(ggvenn)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

#source("../../utils/mg_utils/r_utils/IDmapping.R") functions dependent on dplyr package that has been broken by being uploaded.
source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("../../utils/mg_utils/r_utils/HelperFunctions.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0,...){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```

Read in the raw and normalized datasets and look at the output for PKA proteins and peptides

```{r}
dp.out <- readRDS('~/Documents/projects/061324_XFang_Cilia_GPCR/062424_PWComparisons_data/2024_06_24_dp.out.rds')
old.pQuant <- setDT(dp.out$ProteinLevelData)
old.pQuant[, gene := multiUniprots2multiGenes(as.character(Protein), species='MOUSE')]

p.Quant <- fread('./062424_PWComparisons_data/2024_06_26_ProteinLevelQuant.renorm.csv')
p.Quant[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]

spec <-  fread('./data/CR1/Re-run/20240618_164945_CR1_rep5_Report.tsv')
# create some new ID columns 
spec[, agonist := ifelse(grepl('_ag$', Condition), 'Yes','No')]

spec <- spec[Intensity > 2^4,]

spec[, gene := multiUniprots2multiGenes(ProteinName, species='MOUSE')]
```

plot a comparison of the impact of normalization
I think we want to extract to look at 
i) distributions of the cilia annotated proteins in all 3 sets and 
ii) the Pkaca subunit

interactors of Pkaca

## first look at the Pka subunit
```{r}
genes.oi <- c('Prkaca') #, 'Prkar1a')
```

we want to combine the 3 datasets into one, first subsettting to the things we are interested in

```{r}
spec[gene %in% genes.oi, .(GROUP = Condition, SUBJECT = BioReplicate, gene, LogIntensities = log2(Intensity))]

ints.dt <- list('equilizeMedians'= old.pQuant[gene %in% genes.oi, .(Protein, GROUP, SUBJECT, gene, LogIntensities)],
                  'additionalProteinLvlNormalization'= p.Quant[gene %in% genes.oi, .(Protein, GROUP, SUBJECT, gene, LogIntensities = newLogInts)],
                  'Raw' = spec[gene %in% genes.oi, .(GROUP = Condition, SUBJECT = BioReplicate, gene, Protein = ProteinName, PeptideSequence, LogIntensities = log2(Intensity))]
                  ) %>% rbindlist(idcol='source', fill=T)

col.pal <-  randomcoloR::distinctColorPalette(k=length(unique(ints.dt$GROUP)))


#add a status col
ints.dt[, probe := gsub('[+-]{1}cilia_|[.][0-9]{1}','', GROUP)]
ints.dt[, cilia_present := ifelse(grepl('[+]cilia', GROUP), 'yes', 'no')]
ints.dt[, agonist_present := ifelse(grepl('_ag', GROUP), 'yes', 'no')]

#plot the intensity values of the different approaches
g <- ggplot(ints.dt[(gene %in% genes.oi & probe %like% 'SMO'),], aes(x=source, y=IntsVsMedian, fill=cilia_present)) +
  geom_boxplot(outliers=F) +
  geom_point(aes(fill = cilia_present), size = 1, shape = 21, alpha=0.7, position = position_jitterdodge()) +
  ggtitle('PKA interaction') +
  xlab('Normalization approach') +
  facet_wrap(~probe, scales = 'free_y') +
  scale_fill_brewer(type='qual', palette=1) +
  #scale_y_continuous(breaks=seq(1,20,1)) +
  theme_bw()
g
BackupAsPDF(g, 'smo.posControl.normalizationComparision.boxplot', dimensions=c(10,8))


# now plot the protein on ciliated cells only
#plot the intensity values of the different approaches
g <- ggplot(ints.dt[(gene %in% genes.oi & probe %like% 'SMO' & cilia_present == 'yes'),], aes(x=source, y=LogIntensities, fill=agonist_present)) +
  geom_boxplot(outliers = F) +
  geom_point(aes(fill = agonist_present), size = 1, shape = 21, alpha=0.7, position = position_jitterdodge()) +
  xlab('Normalization approach') +
  ggtitle('SMO probe PKA interaction in ciliated cells') +
  scale_fill_brewer(type='qual', palette=1) +
 scale_y_continuous(breaks=seq(4,18,1)) +
  theme_bw()
g
BackupAsPDF(g, 'smo.posControl.ciliaOnly.normalizationComparision.boxplot', dimensions=c(10,8))
```
Visualize the literature annotated cilia proteins from GO
For SMO, we would expect a shift in the intensity distributions of the ciliated proteins with the addition of agonist to SMO probe

center the distributions as want to see how they differentiate from 0
```{r}
gmt.go <- loadGmtFromBioconductor(dbName='org.Mm.eg.db', ontology='CC', keyType='UNIPROT')
cilium.genes <- gmt.go[ont == 'cilium', gene]


ints.dt <- list('equilizeMedians'= old.pQuant[, .(Protein, GROUP, SUBJECT, gene, LogIntensities)],
                  'additionalProteinLvlNormalization'= p.Quant[, .(Protein, GROUP, SUBJECT, gene, LogIntensities = newLogInts)],
                  'Raw' = spec[, .(GROUP = Condition, SUBJECT = BioReplicate, gene, Protein = ProteinName, PeptideSequence, LogIntensities = log2(Intensity))]
                  ) %>% rbindlist(idcol='source', fill=T)

#add a status col
ints.dt[, probe := gsub('[+-]{1}cilia_|[.][0-9]{1}','', GROUP)]
ints.dt[, cilia_present := ifelse(grepl('[+]cilia', GROUP), 'yes', 'no')]
ints.dt[, agonist_present := ifelse(grepl('_ag|_Ih', GROUP), 'yes', 'no')]


# this isnt what we want... we dont want to normalize by row, we want to normalize by col (maybe plot Zscore of distributions?)
ints.dt[ GROUP %like% '[+]cilia_SMO', IntsVsMean := LogIntensities - median(LogIntensities), by=.(Protein)]

#plot the intensity values of the different approaches
g <- ggplot(ints.dt[(Protein %in% cilium.genes & probe %like% 'SMO' & cilia_present == 'yes'),], aes(x=source, y=IntsVsMean, fill=agonist_present)) +
  geom_boxplot() +
 # geom_point(aes(fill = cilia_present), size = 1, shape = 21, alpha=0.7, position = position_jitterdodge()) +
  ggtitle('SMO probe cilia protein distributions before/after activation') +
  xlab('Normalization') +
  ylab('logIntensities (median subtracted)') +
  scale_y_continuous(n.breaks = 10) +
  scale_fill_manual(values=col.pal[c(3,6)]) +
  theme_bw()
g
BackupAsPDF(g, 'smo.posControl.ciliaOnly.ciliaProteinDistributions.medianScaled.boxplot', dimensions=c(10,8))
```


Generate the same plot with all ligand activated groups; just show our normalization data and show before 

```{r}
ints.dt[, ligand_present := ifelse(grepl('_ag|_In', GROUP), 'yes', 'no')]
ints.dt[, ligand := ifelse(grepl('_ag|_In', GROUP), str_extract(GROUP,'ag$|In$'), 'no')]
ints.dt[, bait := gsub('_ag$|_In$', '', probe)]


ints.dt[ GROUP %in% grep('[+]cilia_SMO|[+]cilia_GPR161|[+]cilia_Pde8A|[+]cilia_Pde8B|[+]cilia_Pth2r|[+]cilia_Crhr2', GROUP, value = T), IntsVsMean := LogIntensities - median(LogIntensities), by=Protein]


#plot the intensity values of the different approaches
g <- ggplot(ints.dt[(Protein %in% cilium.genes & probe %in% grep('Crhr2|GPR161|Pth2r|SMO|Pde8A|Pde8B', probe, value=T) & cilia_present == 'yes'),], aes(x=bait, y=IntsVsMean, fill=ligand_present)) +
  geom_boxplot() +
  scale_y_continuous(breaks=seq(-10,12, by=1)) +
  xlab('Probe') +
  ggtitle('Impact of ligand binding on cilia protein distributions') +
  ylab('logIntensities (median subtracted)') +
  scale_fill_manual(values=col.pal[c(3,5,7)]) +
  theme_bw()
g

BackupAsPDF(g, 'ligandbaits.ciliaOnly.ciliaProteinDistributions.medianScaled.boxplot', dimensions=c(10,8))
```
Regenerate the PCA of the expression profiles with the treatmen conditions dropped

```{r}
p.mat <-dcast(p.Quant, Protein~interaction(GROUP,SUBJECT), value.var = 'newLogInts') %>% 
  as.matrix(rownames='Protein')

p.mat <- p.mat[complete.cases(p.mat), grep('cilia_PM|[-]cilia_ARL13b[.]', colnames(p.mat), invert=T, value = T)]

pca <- prcomp(t(p.mat))

# create annotation dt
ann <- data.table(group = colnames(p.mat))
ann[, cilia_present := ifelse(grepl('-cilia', group), 'No', 'Yes')]
ann[, agonist_present := ifelse(grepl('_ag[.][0-9]{1}$', group), 'Yes', 'No')]
ann[, status := gsub("[-+]{1}cilia_|_ag[.][0-9]{1}$|[.][0-9]{1}$",'', group)]
ann[, rep := str_extract(group, '[0-9]$')]

pca.dt <- data.table(pca$x, keep.rownames = T)
pca.dt <- merge(x=pca.dt, y=ann, by.x='rn', by.y='group')
pca.dt[, group := gsub('.[0-9]$','', rn)]

# calculate variance explained
pcaPercentVar <- round(100 * (pca$sdev^2)/sum(pca$sdev^2), 1)

#plot first two components
p <- ggplot (pca.dt, aes(x=PC1, y=PC2,  fill = status, shape = cilia_present)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab(sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab(sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle(sprintf ("PCA using %d features (log intensity)", nrow(pca$rotation))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = c(21:26)) +
  scale_fill_manual(values = col.pal) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, 'proteins.probCol.noControls.', dimensions=c(12,9))
```



```{r}
ints.dt <- list('MSSNorm'= old.pQuant[Protein %in% cilium.genes, .(Protein, GROUP, SUBJECT, gene, LogIntensities)],
                  'AdditionalRoundNorm'= p.Quant[Protein %in% cilium.genes, .(Protein, GROUP, SUBJECT, gene, LogIntensities = newLogInts)],
                  'Raw' = spec[ProteinName %in% cilium.genes, .(GROUP = Condition, SUBJECT = BioReplicate, gene, Protein = ProteinName, PeptideSequence, LogIntensities = log2(Intensity))]
                  ) %>% rbindlist(idcol='source', fill=T)

#add a status col
ints.dt[, probe := gsub('[+-]{1}cilia_|[.][0-9]{1}','', GROUP)]
ints.dt[, cilia_present := ifelse(grepl('[+]cilia', GROUP), 'yes', 'no')]
ints.dt[, agonist_present := ifelse(grepl('_ag', GROUP), 'yes', 'no')]


#plot the intensity values of the different approaches
g <- ggplot(ints.dt[Protein %in% cilium.genes,], aes(x=source, y=LogIntensities, fill=cilia_present)) +
  geom_boxplot() +
 # geom_point(aes(fill = cilia_present), size = 1, shape = 21, alpha=0.7, position = position_jitterdodge()) +
  ggtitle(sprintf('GO annotated cilia protein distributions (n=%s)', sum(cilium.genes %in% unique(spec$ProteinName)))) +
  facet_wrap(~probe, ncol=3, scales = 'free_y') +
  scale_fill_manual(values=col.pal[c(3,6)]) +
  theme_bw()
g
BackupAsPDF(g, 'GOcilia.distributions.boxplot', dimensions=c(12,18))
```
Now plot the distributions of Max Nauchery cilia annotated protein distributions (high confidence ones)

```{r}
read_xlsx(path='./docs/ciliaProteins.Nachury.etal.2021.xlsx', sheet=1, skip = 1)
#read in literature sig hits
cilia.Nachury.2021 <- read_xlsx(path='./docs/ciliaProteins.Nachury.etal.2021.xlsx', sheet=1, skip = 1)$`Gene Symbol`
cilia.Nachury.2021 <- cilia.Nachury.2021[!is.na(cilia.Nachury.2021)]

cilia.Nachury.2015.highConf <- read_xlsx(path='./docs/Nachury.etal.2015cilliaProteins.xlsx', sheet=1)$`Uniprot accession number`
cilia.Nachury.2015.highConf <- multiUniprots2multiGenes(cilia.Nachury.2015.highConf, species='MOUSE')

cilia.literature <- list('Nachury et al 2021' = cilia.Nachury.2021,
                         'Nachury et al 2015 (High Conf)' = cilia.Nachury.2015.highConf)
```
```{r}
ints.dt <- list('MSSNorm'= old.pQuant[, .(Protein, GROUP, SUBJECT, gene, LogIntensities)],
                  'AdditionalRoundNorm'= p.Quant[, .(Protein, GROUP, SUBJECT, gene, LogIntensities = newLogInts)],
                  'Raw' = spec[, .(GROUP = Condition, SUBJECT = BioReplicate, gene, Protein = ProteinName, PeptideSequence, LogIntensities = log2(Intensity))]
                  ) %>% rbindlist(idcol='source', fill=T)

#add a status col
ints.dt[, probe := gsub('[+-]{1}cilia_|[.][0-9]{1}','', GROUP)]
ints.dt[, cilia_present := ifelse(grepl('[+]cilia', GROUP), 'yes', 'no')]
ints.dt[, agonist_present := ifelse(grepl('_ag', GROUP), 'yes', 'no')]


lapply(names(cilia.literature), function(i,x){

  g <- ggplot(x[gene %in% cilia.literature[[i]],], aes(x=source, y=LogIntensities, fill=cilia_present)) +
  geom_boxplot() +
  ggtitle(paste0(i, ' cilla protein distributions')) +
  facet_wrap(~probe, scales = 'free_y', ncol=3) +
  scale_fill_brewer(type='qual', palette='Dark2') +
  theme_bw()
  g
  BackupAsPDF(g, paste0(i, '.curatedCilla.boxplots'), dimensions=c(14,22))
},x=ints.dt)

```

ToDo
--
Regenerate the control SMO plots to ensure it is correctly calculated

```{r}
col.pal <-  randomcoloR::distinctColorPalette(k=length(unique(p.Quant$GROUP)))

#add a status col
p.Quant[, probe := gsub('[+-]{1}cilia_|[.][0-9]{1}','', GROUP)]
p.Quant[, cilia_present := ifelse(grepl('[+]cilia', GROUP), 'yes', 'no')]
p.Quant[, agonist_present := ifelse(grepl('_ag|_Ih', GROUP), 'yes', 'no')]
p.Quant[, ligand_present := ifelse(grepl('_ag|_In', GROUP), 'yes', 'no')]
p.Quant[, ligand := ifelse(grepl('_ag|_In', GROUP), str_extract(GROUP,'ag$|In$'), 'no')]
p.Quant[, bait := gsub('_ag$|_In$', '', probe)]
p.Quant[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]

```
plot for the ligand data the distributions of all the

```{r}
p.Quant[probe %in% grep('Crhr2|GPR161|Pth2r|SMO|Pde8A|Pde8B', probe, value=T), IntsVsMean := newLogInts - median(newLogInts), by=Protein]

#plot the intensity values of the different approaches
g <- ggplot(p.Quant[(Protein %in% cilium.genes & probe %in% grep('Crhr2|GPR161|Pth2r|SMO|Pde8A|Pde8B', probe, value=T) & cilia_present == 'yes'& IntsVsMean < 7),], aes(x=bait, y=IntsVsMean, fill=ligand_present)) +
  geom_boxplot() +
  scale_y_continuous(breaks=seq(-5,5, by=1)) +
  xlab('Probe') +
  ggtitle('Impact of ligand binding on cilia protein distributions') +
  ylab('logIntensities (median subtracted)') +
  scale_fill_manual(values=col.pal[c(3,5,7)]) +
  theme_bw()
g



# try average and then plot the distribution
p.Quant[, meanInts := mean(newLogInts, na.rm=T), by=.(GROUP,Protein)]


g <- ggplot(p.Quant[(Protein %in% cilium.genes & probe %in% grep('Crhr2|GPR161|Pth2r|SMO|Pde8A|Pde8B', probe, value=T) & cilia_present == 'yes'& IntsVsMean < 7),], aes(x=bait, y=IntsVsMean, fill=ligand_present)) +
  geom_boxplot() +
  scale_y_continuous(breaks=seq(-5,5, by=1)) +
  xlab('Probe') +
  ggtitle('Impact of ligand binding on cilia protein distributions') +
  ylab('logIntensities (median subtracted)') +
  scale_fill_manual(values=col.pal[c(3,5,7)]) +
  theme_bw()


BackupAsPDF(g, 'ligandedProbes.cillaProt.boxplots', dimensions=c(6,5))


# confirm by plotting the median sweep
p.mat <- dcast(p.Quant[Protein %in% cilium.genes,], Protein~paste0(GROUP,'.',SUBJECT), value.var = 'newLogInts') %>% 
  as.matrix(rownames='Protein')


submat <- sweep(p.mat, 1, apply(p.mat, 1, median, na.rm=T))

sub.dt <- setDT(reshape2::melt(submat))
setnames(sub.dt, new=c('Protein', 'sample', 'newLogInts'))
sub.dt[, c('GROUP', 'SUBJECT') :=  tstrsplit(sample, '[.]', keep = c(1,2))]
sub.dt[, probe := gsub('[+-]{1}cilia_|[.][0-9]{1}','', GROUP)]
sub.dt[, cilia_present := ifelse(grepl('[+]cilia', GROUP), 'yes', 'no')]
sub.dt[, agonist_present := ifelse(grepl('_ag|_Ih', GROUP), 'yes', 'no')]
sub.dt[, ligand_present := ifelse(grepl('_ag|_In', GROUP), 'yes', 'no')]
sub.dt[, ligand := ifelse(grepl('_ag|_In', GROUP), str_extract(GROUP,'ag$|In$'), 'no')]
sub.dt[, bait := gsub('_ag$|_In$', '', probe)]
sub.dt[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]


g <- ggplot(sub.dt[(probe %in% grep('Crhr2|GPR161|Pth2r|SMO|Pde8A|Pde8B', probe, value=T) & cilia_present == 'yes' & newLogInts < 7),], aes(x=bait, y=newLogInts, fill=ligand_present)) +
  geom_boxplot() +
  scale_y_continuous(breaks=seq(-10,12, by=1)) +
  xlab('Probe') +
  ggtitle(sprintf('Impact of ligand binding on cilia protein distributions (n=%s)',  sum(cilium.genes %in% unique(p.Quant$Protein)))) +
  ylab('logIntensities (median subtracted)') +
  scale_fill_manual(values=col.pal[c(3,5,7)]) +
  theme_bw()
g
BackupAsPDF(g, 'ligandProbes.cillaProt.boxplots', dimensions=c(6,5))
```
Look at the SMO proteins 

```{r}
#plot the intensity values of the different approaches
g <- ggplot(p.Quant[Protein %in% cilium.genes & bait == 'SMO',], aes(x=GROUP, y=meanInts, fill=GROUP)) +
  geom_boxplot() +
 # geom_point(aes(fill = cilia_present), size = 1, shape = 21, alpha=0.7, position = position_jitterdodge()) +
  ggtitle(sprintf('SMO probe cilia protein distributions (n=%s)', sum(cilium.genes %in% unique(p.Quant$Protein)))) +
  scale_fill_manual(values=col.pal[c(3:7)]) +
  xlab('Probe') +
  ylab('LogIntensities (median subtracted)') +
  theme_bw() +
  theme(axis.text.x = element_blank())
g
BackupAsPDF(g, 'SMOprobes.cillaProt.boxplots', dimensions=c(6,5))
```

```{r}

```




Regenerate the plots to adjust the label sizes



Differences in the log2FC changes between cilia/non ciliated? Lets look at venn diagrams, and maybe better is a correlation plot
Redo the plot with the noCilia denominator enrichment (seems to be the wrong log2FC thresholds)
Redo the no-ligand denominator comparisons plots

The ciliated vs non-ciliated background plots comparison. 
