---
title: "082224_Pth2rSmo_interactionAnalysis"
author: "Martin Gordon"
date: "2024-08-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro 

Goal of the current analysis is to test for interaction between cell-line (ciliated vs non-ciliated) and ligand activation
The question is: does the probe neighbourhood upon activation cilia-independent?

## packages

```{r}
library(magrittr)
library(data.table)
library(circlize)
library(ComplexHeatmap)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(MSstats)
library(ggrepel)
library(readxl)
library(usedist) #package for working with distances
library(ggh4x) # additional functionality for ggplot2 obj
library(scales)
library(patchwork)
library(RColorBrewer)
library(cluster) # pam clusterin∂ƒg of genes
library(eulerr) # eulerr plot 
library(ggvenn)
library(viridis)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

source("../../utils/mg_utils/r_utils/IDmapping.R")
source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("../../utils/mg_utils/r_utils/HelperFunctions.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0,...){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}

redbluColpal <- c('#D01B1B', '#FF4242', '#FFFFFF', '#95D2EC','#47abd8')
```

Lets use msstats to produce the model to test for interacitons (non-additivity) between conditions

```{r}
p.quant <- fread('~/Documents/projects/061324_XFang_Cilia_GPCR/062424_PWComparisons_data/2024_06_26_ProteinLevelQuant.renorm.csv')

# add other annotation information
p.quant[, cilia_present := ifelse(grepl('[-]cilia', GROUP), 'No', 'Yes')]
p.quant[, ag_present := ifelse(grepl('_ag|_In', GROUP), 'Yes', 'No')]
p.quant[, probe := gsub("[-+]{1}cilia_|_ag{1}$|_In{1}$$",'', GROUP)]

p.quant[,.N, by=.(ag_present, cilia_present, probe,GROUP)]

dp.out <- readRDS('./062424_PWComparisons_data/2024_06_24_dp.out.rds')

f.quant <- setDT(dp.out$FeatureLevelData)

p.quant[, SUBJECT := paste0('batch.',SUBJECT)]
p.quant[, LogIntensities := newLogInts]
f.quant[, SUBJECT := paste0('batch.',SUBJECT)]

dp.out$FeatureLevelData <- f.quant
dp.out$ProteinLevelData <- p.quant
```
Create the contrast matrix
```{r}
levels(dp.out$ProteinLevelData$GROUP)
```
Looking at the interactor distributions, it looks like all the samples are very similar

```{r}
mss.dt <- fread('./070224_newPWComparisons_data/2024_07_03_all.mss.pwComparisons.csv')

g <- ggplot(mss.dt[Label %like% '_In|_ag',], aes(x=pvalue)) +
  geom_histogram() +
  facet_wrap(~Label)

BackupAsPDF(g, 'interactor.pval.histograms')
```


Not used... combine reps

```{r, eval=FALSE}
subset <- dp.out$ProteinLevelData[GROUP == '+cilia_SMO_ag',]
subset[, GROUP := '+cilia_SMO_ag_dup']


# add some random noise to the measurements
subset[, LogIntensities := rnorm(mean=LogIntensities, n=1), by=.(Protein,RUN)]

dp.out$ProteinLevelData <- rbind(dp.out$ProteinLevelData, subset[,-c('noisyLogIntensities')])
```


```{r}
grep( '_ag|_In', dp.out$ProteinLevelData$GROUP, value=T) %>%  unique()

#basicaly testing if the coefficient = 0
contrasts.list <- list('Smo cilia:ligand interactions' = data.table(V1="+cilia_SMO_ag", V2="+cilia_SMO", V3='-cilia_SMO_ag', V4='cilia_SMO'),
                       'Pth2r cilia:ligand interactions' = data.table(V1="+cilia_Pth2r_ag", V2="+cilia_Pth2r", V3='-cilia_Pth2r_ag', V4='-cilia_Pth2r'),
                       '+cilia_Pth2r_ag vs -cilia_Pth2r_ag' = data.table(V1="+cilia_Pth2r_ag", V2='-cilia_Pth2r_ag'),
                       '+cilia_SMO_ag vs -cilia_SMO_ag' = data.table(V1="+cilia_SMO_ag", V2='-cilia_SMO_ag'),
                       '+cilia_Crhr2_ag vs -cilia_Crhr2_ag' =  data.table(V1="+cilia_Crhr2_ag", V2='-cilia_Crhr2_ag'),
                       '+cilia_GPR161_ag vs -cilia_GPR161_ag' =  data.table(V1="+cilia_GPR161_ag", V2='-cilia_GPR161_ag'),
                       '+cilia_Pde8A_In vs -cilia_Pde8A_In' = data.table(V1="+cilia_Pde8A_In", V2='-cilia_Pde8A_In'),
                       '+cilia_Pde8B_In vs -cilia_Pde8B_In' = data.table(V1="+cilia_Pde8B_In", V2='-cilia_Pde8B_In')
                       )

contrasts.mat <- MSstats::MSstatsContrastMatrix(contrasts.list, 
                               conditions = unique(dp.out$ProteinLevelData$GROUP), 
                               labels = names(contrasts.list))

# fix the contrast matrix for the interaction analysis
contrasts.mat[1,32] <- -1
contrasts.mat[1,31] <-  1
contrasts.mat[2,30] <-  -1
contrasts.mat[2,29] <-  1
```


```{r}
# run msstats correcting for batch 
mss <- groupComparison(contrast.matrix=contrasts.mat, 
                       verbose=T,
                       data=dp.out)

mss.dt <- setDT(mss$ComparisonResult)
```
Tidy the mss results

```{r}
mss.dt[, p.adj := p.adjust(pvalue, method='BH'), by=Label]
mss.dt[, sig := 'not']
mss.dt[, gene := multiUniprots2multiGenes(as.character(Protein), species = 'MOUSE')]

mss.dt[abs(log2FC) > 0.58 & p.adj < 0.05, sig := ifelse(log2FC > 0, 'up', 'down')]
```

Add protein counts info and write out

```{r}
mss.dt
#table of N detected per feature
nCounts.dt <- p.quant[, .SD[, .(numeratorCounts=.N, denominatorCounts=.N)], by=.(Protein,GROUP)]

# maybe the easiest way is to split label in mss.dt into denominator and numerator, and then match label in this datatable and just take the sum of measurements
mss.dt[, c('numerator', 'denomimator') :=  tstrsplit(Label, ' vs ', keep=c(1,2))]
# fix for the interaction analysis

comb.dt <- merge(x=mss.dt[!Label %in% c('Pth2r cilia:ligand interactions', 'Smo cilia:ligand interactions'),], y=nCounts.dt[,.(Protein, GROUP, numeratorCounts)], by.x=c('numerator','Protein'), by.y=c('GROUP','Protein'), all.x=T)

comb.dt <- merge(x=comb.dt, y=nCounts.dt[,.(Protein, GROUP, denominatorCounts)], by.x=c('denomimator','Protein'), by.y=c('GROUP','Protein'), all.x=T)

comb.dt <- rbind(comb.dt, mss.dt[Label %in% c('Pth2r cilia:ligand interactions', 'Smo cilia:ligand interactions'),], fill=T)

#now, correct for those that have condition missing to only except present in all conditions as a positive result
comb.dt[issue == 'oneConditionMissing', sig := ifelse(numeratorCounts == 3 | denominatorCounts == 3, 'oneConditionMissing', 'not')]

fwrite(comb.dt, ScriptAndDatedFileName('pwComparisons.interactions.ciliaVsnonCilia.csv'))
```



```{r}
g <- ggplot(mss.dt[sig != 'not', .N, by=.(sig,Label)], aes(x=reorder(Label, -N), y=N, fill=sig)) +
  geom_bar(stat='Identity') +
  ggtitle('Significant hits per contrast (log2FC > 1 & p.adjust < 0.05)') +
  xlab('Contrast') +
  ylab('Number sig hits') +
  scale_fill_manual(values=c('up'='#990033', 'down'='#336699', 'oneConditionMissing'='grey')) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90,size=10))
g

BackupAsPDF(g, 'sigHits.allContrasts.barplot', dimensions=c(12,10))
```
cila proteins
```{r}
# read in mouse to hu mapping table; will use this to collect all possible mappings from human to filter our results by
mh_mappings <- fread('~/Documents/utils/mg_utils/data/mouseHumanIDConversion.txt')

cilia.AM <- c('INPP5E', 'LCA5', 'PRKAR1A', 'ARL13B', 'SMO', 'EVC2', 'EVC', 'PKD2', 'TTC21B', 'PKHD1', 'IFT88', 'NPHP4', 'EFCAB7', 'CYS1')

cilia.Consensus <- c('Pdap1','Cep131','Cep41','Ift122','Ift172','Ift52','Ift57','Ift74','Ift81','Ift88','Cfap97','Lca5','Cemip2','Wdr35',
                     'Arl13b','Arl3','Inpp5e','Nme7','Kif3a','Kif3b','Kifap3','Ywhaz','Cys1','Evc2','Ttc21b','Ift56')

# "Kiaa1430" "Ttc26"  not found in my mappings
# alias on ncbi: Cfap97 & IFT56
cilia.Consensus[!cilia.Consensus %in% unique(mh_mappings$Symbol.mouse)]

# all found
cilia.AM.prots <- mh_mappings[Symbol.human %in% cilia.AM, SWISS_PROT_IDs.mouse]

cilia.prots <- c(cilia.AM.prots, cilia.Consensus)
```



```{r}
subDT <- mss.dt
subDT[sig != 'not' & issue == 'oneConditionMissing', sig := ifelse( (!is.na(numeratorCounts) & numeratorCounts == 3)  | (!is.na(denominatorCounts) & denominatorCounts==3), sig, 'not')]

subDT[, lab := sig]
subDT[, lab := ifelse(Protein %in% cilia.prots, 'cilia', lab)]


g <- ggplot(subDT[!is.infinite(-log10(adj.pvalue)),], aes(x=log2FC, y=-log10(p.adj), col=lab, label=gene)) +
  geom_point(alpha=0.5) + 
  geom_point(data=subDT[Protein %in% c(cilia.prots),]) +
  ylab('-log10 p-value') +
  ggrepel::geom_text_repel(data=subDT[sig != 'not' & !is.infinite(-log10(adj.pvalue)),], show.legend = FALSE, size = 2, max.overlaps = 20) +
  ggtitle(paste0('cell-line contrasts + interaction analysis (FC +/- 50% & adj.pval < 0.05)')) +
  geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
  geom_hline(yintercept = -log10(0.05), linetype=2, alpha=0.4) +
  scale_color_manual(values=c('up'='#990033', 'down'='#336699', 'not'='grey', 'cilia'='#5ec962')) +
  facet_wrap(~Label, scales='free', ncol=3) +
  theme_bw()
g

BackupAsPDF(g, 'cellline.contrasts.cilaanno.volcano', dimensions=c(16,12))
```
plot reduced thresholds

```{r}

subDT[,sig := 'not']
subDT[abs(log2FC) > 0.58 & pvalue < 0.005, sig := ifelse(log2FC > 0, 'up', 'down')]
subDT[, lab := sig]
subDT[, lab := ifelse(Protein %in% cilia.prots, 'cilia', lab)]

g <- ggplot(subDT[!is.infinite(-log10(adj.pvalue)),], aes(x=log2FC, y=-log10(pvalue), col=lab, label=gene)) +
  geom_point(alpha=0.5) + 
  geom_point(data=subDT[Protein %in% c(cilia.prots),]) +
  ylab('-log10 p-value') +
  ggrepel::geom_text_repel(data=subDT[sig != 'not' & !is.infinite(-log10(adj.pvalue)),], show.legend = FALSE, size = 2, max.overlaps = 20) +
  ggtitle(paste0('cell-line contrasts + interaction analysis (FC +/- 50% & pval < 0.005)')) +
  geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
  geom_hline(yintercept = -log10(0.005), linetype=2, alpha=0.4) +
  scale_color_manual(values=c('up'='#990033', 'down'='#336699', 'not'='grey', 'cilia'='#5ec962')) +
  facet_wrap(~Label, scales='free', ncol=3) +
  theme_bw()
g

BackupAsPDF(g, 'cellline.contrasts.cilaanno.reducedThresholds.volcano', dimensions=c(16,12))
```

Go enrichment heatmaps

```{r}
gmt.go <- loadGmtFromBioconductor(dbName='org.Mm.eg.db', ontology='ALL', keyType='UNIPROT')
```

GO BP and GO cellular component
---
Issue with the enrichment results... seems many of these proteins are localized to the Golgi... have they been 
```{r}

# define the universe, the total set of identified genes in our study
universe <- as.character(unique(p.quant$Protein))

mss.dt[, enrich.grp := interaction(Label,sig)]


enrich.dt <- enricherOnGroups(mss.dt[sig %in% c("down","up")], 
                              groupColumns = 'enrich.grp', 
                              geneColumn = "Protein", 
                              term2gene.gmt = gmt.go, 
                              universe = universe)

fwrite(enrich.dt, ScriptAndDatedFileName('allContrasts.GO.enrichments.csv'))


simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich.grp')
fwrite(simp.enrich$simplified, ScriptAndDatedFileName('allContrasts.GO.BP.simplified.enrichments.csv'))
```

```{r}

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, 
                                groupColumn = 'enrich.grp', 
                                topN = 8,
                                title='GO enrichment terms',
                                row_names_gp = gpar(fontsize = 7), 
                                upperThreshold = 8,
                                #negCols=unique(grep('down', simp.enrich$simplified, value=T))) # ,
                                column_names_gp= gpar(fontsize = 7))
ht

BackupAsPDF(ht, 'GO.enrich.heatmap', dimensions = c(11,15))
```

# write out the mss results in wide format for Aaron to inspect

```{r}
comb.dt %>%  colnames()

fwrite(dcast(comb.dt, gene+Protein~Label, value.var = c('log2FC', 'pvalue', 'p.adj', 'numerator', 'numeratorCounts', 'denomimator', 'denominatorCounts')), ScriptAndDatedFileName('mssPWcomparisons.celllinesinteractions.wideformat.csv'))
```



Look into the oneCondition missing set more
What is this set of proteins enriched for?

```{r}

```


Lets look at the pvalue distributions
Same as earlier 
Clear systematic deviation from theoretical nulls..
Looking at tstat distirbutions, there is a very clear peak around 0 with narrow tails; little happening
When plotting the tStat distribution there is a sharp peak at 0 and very narrow range
Basically evdience the shift may be very small, or we lack power to detect the shift

Narrow tails mean that extreme test statistic values are less likely than in a standard t-distribution. This reduces the likelihood of observing significant differences (large test statistics), further contributing to higher p-values.

```{r}
# t statis

g <- ggplot(mss.dt, aes(x=pvalue)) +
  geom_histogram(boundary=0) +
  facet_wrap(~Label, scales='free_y')
g
BackupAsPDF(g, 'interactionEffect.histograms')

# construct a histogram of the test statistics going into this
# this is difference in the group means divided by the standard error
g <- ggplot(mss.dt, aes(x=Tvalue)) +
  geom_histogram(boundary=0, binwidth=0.1) +
  ggtitle('test statistic distribution') +
  xlim(c(-5,5)) +
  facet_wrap(~Label, scales='free')
g
BackupAsPDF(g, 'tStat.histogram')

#calculate a datatable of the mean log intensities of each group
g <- ggplot(mss.dt, aes(x=log2FC)) +
  geom_histogram(boundary=0, binwidth=0.1) +
  ggtitle('test statistic distribution') +
  xlim(c(-5,5)) +
  facet_wrap(~Label)
g
BackupAsPDF(g, 'LFC.histogram')


g <- ggplot(mss.dt, aes(x=SE)) +
  geom_histogram(boundary=0, binwidth=0.1) +
  ggtitle('test statistic distribution') +
  facet_wrap(~Label)
g
BackupAsPDF(g, 'SE.histogram')

```




I guess need to contact Xiaofang about details for the experimental design.. should these be paired t-tests?
Key assumption of LM is *independence* of measurements.. are

```{r}

```


I guess one alternative is a permutation test...
Want to see if the effect == 0.. maybe take the t-statistic (which is just a measure of signal(FC)/noise) and permutat

```{r}
# run this over each label seperately

lapply(mss.dt$Label, function(x){
  
  
  
  
  
})
mss.dt
```


```{r}
ggplot(mss.dt, aes(sample=Tvalue)) +
  stat_qq() +
  stat_qq_line(color = "red", linetype = "dashed") +
  labs(title = "Q-Q Plot vs. Normal Distribution", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  facet_wrap(~Label, scales='free') +
  theme_minimal()

qqnorm(mss.dt$Tvalue[mss.dt$Label == "Smo cilia vs Smo no cilia"])
qqline(mss.dt$Tvalue[mss.dt$Label == "Smo cilia vs Smo no cilia"], col = "red", lwd = 2)



mss.dt$Label %>%  unique()
```

```{r}
my_data <- rnorm(100)

# Q-Q plot with ggplot2
df <- data.frame(sample = my_data)

ggplot(df, aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "red") +
  labs(title = "Q-Q Plot of Normally Distributed Data", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()
```

```{r}
mss.dt <- merge(x=mss.dt, y=p.quant[,.(NObservations=.N), by=Protein], by='Protein')
mss.dt[, Var := (2^SE)* NObservations] #caclualte variance for each sample

g <- ggplot(mss.dt, aes(x=Var, y=log2FC)) +
  geom_point() +
  geom_density_2d() +
  facet_wrap(~Label)

g


mss.dt
```
Go back to the protein quant data with the group measurements, calcul
```{r}
summary.dt <- p.quant[probe %in% c("SMO", "Pth2r"), .(meanLogInts = mean(LogIntensities, na.rm=T),
                                        variance = var(LogIntensities, na.rm=T),
                                        stdDev = sd(LogIntensities, na.rm=T),
                                        NObservations = .N), by=.(Protein,GROUP)]



ggplot(summary.dt, aes(x=variance, fill=GROUP)) +
  geom_density(alpha=0.5) +
  xlim(c(0,2))

ggplot(summary.dt, aes(x=meanLogInts, y=2^(stdDev))) +
  geom_point() +
  facet_wrap(~GROUP,scales='free_y') +
  coord_cartesian(expand = F)

ggplot(summary.dt, aes(x=meanLogInts, y=variance)) +
  geom_point() +
  facet_wrap(~GROUP,scales='free_y') +
  coord_cartesian(expand = F)

```
Look at the 

read int he protein quant data and add the new explanatory variables to the data

```{r}
p.quant <- fread('./062424_PWComparisons_data/2024_06_26_ProteinLevelQuant.renorm.csv')

p.quant[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]
# add other annotation information
p.quant[, cilia_present := factor(ifelse(grepl('[-]cilia', GROUP), 'noCilia', 'Cilia'), levels=c('noCilia', 'Cilia'))]
p.quant[, ligand_present := factor(ifelse(grepl('_ag|_In', GROUP), 'Ligand', 'noLigand'), levels=c('Ligand', 'noLigand'))]
p.quant[, probe := factor(gsub("[-+]{1}cilia_|_ag{1}$|_In{1}$$",'', GROUP))]
p.quant[, batch := factor(SUBJECT)]

# subset the p.quant data to 
p.quant <- p.quant[probe %in% c('SMO', 'Pth2r'),]

p.quant[, LogIntensities := newLogInts]
```

'Clean' the p.quant data; try subtract batch?

```{r}
p.mat <- dcast(p.quant, Protein~interaction(GROUP,SUBJECT), value.var = 'LogIntensities') %>% 
  as.matrix(rownames="Protein")

submat <-  sweep(p.mat, 1, apply(p.mat, 1, median, na.rm=T))

# use by for aggregation
ann <- p.quant[,.(sample=paste0(GROUP,'.',SUBJECT),
                  batch=interaction('batch', SUBJECT)), by=.(GROUP,SUBJECT, ligand_present, cilia_present, probe)]

ann$sample == colnames(submat)

# match order in mat columns
ann <- ann[match(colnames(submat), sample),]

ann$sample == colnames(submat)

Heatmap(submat,
        column_split = ann$batch,
        cluster_rows = clusterWNA(submat))



```
```{r}
p.mat
```



read in associated metadata with the cilia annotated proteins
```{r}
# gprotein subunits 
gp.subunits  <- c("Gnai2","Gna11","Gnaq","Gna12","Gna13", 'Gnl1', 'Gnb1', "Gnb2","Gnas", 'Gng5', 'Gnl3l', "Gns","Gnl3", "Gnl2", 'Gng12', 'Gnai3')

# read in mouse to hu mapping table; will use this to collect all possible mappings from human to filter our results by
mh_mappings <- fread('~/Documents/utils/mg_utils/data/mouseHumanIDConversion.txt')

cilia.AM <- c('INPP5E', 'LCA5', 'PRKAR1A', 'ARL13B', 'SMO', 'EVC2', 'EVC', 'PKD2', 'TTC21B', 'PKHD1', 'IFT88', 'NPHP4', 'EFCAB7', 'CYS1')

cilia.Consensus <- c('Pdap1','Cep131','Cep41','Ift122','Ift172','Ift52','Ift57','Ift74','Ift81','Ift88','Cfap97','Lca5','Cemip2','Wdr35',
                     'Arl13b','Arl3','Inpp5e','Nme7','Kif3a','Kif3b','Kifap3','Ywhaz','Cys1','Evc2','Ttc21b','Ift56')

# "Kiaa1430" "Ttc26"  not found in my mappings
# alias on ncbi: Cfap97 & IFT56
cilia.Consensus[!cilia.Consensus %in% unique(mh_mappings$Symbol.mouse)]

# all found
cilia.AM.prots <- mh_mappings[Symbol.human %in% cilia.AM, SWISS_PROT_IDs.mouse]
cilia.Consensus.prots <- mh_mappings[Symbol.mouse %in% cilia.Consensus, SWISS_PROT_IDs.mouse]

```

I think we want to run this per probe, so split probe into a list

```{r}
p.quant.list <- split(p.quant, p.quant$probe)
```


Run the lm with an interaction term for cilia status
What proteins show significant difference in impact depending cilia status?
```{r}
# how does ligand effect change given the cilia status of the cell. IE are the interacitons independent of cilia presence?
emmeans.contrastOfContrasts <- function (l, factorFormula = ~ligand_present|cilia_present){  
  emm <- emmeans::emmeans(l, factorFormula)
  contrast1 <- pairs(emm)
  contrast2 <- pairs(contrast1, by = NULL, adjust = "none")
  return (as.data.table(contrast2))
}


# postProcessFunction = emmeans.contrastOfContrasts tells it to use the function above after doing anova.
# it is optional, adds significant time to the computation (~5x), but is nice to
# put a magnitude and p value on specific differences of differences
#
# cl = 6 parameter says to use 6 processes.  Set to NULL if multiprocess is failing, or higher or lower as you wish.
# Numbers near the number of processors available in your machine is probably counterproductive.


p.quant.list

lm.list <- lapply(p.quant.list, function(x){
  
  lm.out <- linearModelsAllProteins(p.quant, 
                                  formulaList = list(Interaction =  newLogInts ~ batch + ligand_present ),
                                  splitColumn = "Protein", 
                                  postProcessFunction = emmeans.contrastOfContrasts, cl = 6) |> suppressWarnings()
  return(lm.out)
})
```

Now, rbind the anova and the emmeans outputs to 

suggests this is a poor choice of test for the data?

```{r}
lm.list$Pth2r$anova[, probe := 'Pth2r']
lm.list$SMO$anova[, probe := 'SMO']

anova.dt <- rbind(lm.list$SMO$anova,lm.list$Pth2r$anova)

g <- ggplot(anova.dt, aes(x=p.value)) +
  geom_histogram(binwidth=0.01) +
  facet_wrap(~term+probe, scales='free_y')
g
BackupAsPDF(g, 'newRes')
#this us a crazy distribution

anova.dt

prots.oi <- anova.dt[probe =='SMO' & term == 'ligand_present' & p.value > 0.8, unique(Protein)]


p.mat <-dcast(p.quant.list$SMO, Protein~GROUP+SUBJECT, value.var = 'newLogInts') %>% 
  as.matrix(rownames='Protein')

submat <- p.mat[rownames(p.mat) %in% prots.oi,]

submat <- sweep(submat, 1, apply(submat,1,median ,na.rm=T))


# these are clearly cilia seperated
Heatmap(submat,
        column_split = list(grepl('[+]cilia', colnames(submat)),
                            grepl('ag', colnames(submat))
                                  ),
        show_row_names = F,
        col = colorRamp2(breaks=c(-2,0,2), colors = c('blue', 'white', 'red')),
        cluster_rows=clusterWNA(submat))


# whhat aboubt 
```

