---
title: "051623_DSwaney"
author: "Martin Gordon"
date: "2023-05-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(purrr)
library(tidyr)
library(dplyr)
library(data.table)
library(MSstats)
library(ggplot2)
library(magrittr)
library(circlize)
library(ComplexHeatmap)
library(parallelDist) #needed for some custom enrichment functions
library(ggvenn)

#source Bens utility scripts
source("/Users/martingordon/Documents/utils/bp_utils/MSstats_V4_Functions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/MSstats_Helper_Functions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
```

```{r helper-functions}

#function to save plots generated
save_plot <- function(plot_func, plot_args, filename, directory) {
  # Call the plot function with the provided arguments
  plot <- do.call(plot_func, plot_args)
  
   # create the directory if it doesn't exist
  if (!dir.exists(directory)) {
    dir.create(directory, recursive=TRUE)
  }
  
  # save the plot to file in the specified directory
  filepath <- file.path(directory, filename)
  ggsave(plot, file = filepath)
  
  message(paste0("Plot saved to file: ", filepath))
  
}


## Bens Functions

today <- function(){
  format(Sys.time(), "%Y_%m_%d")
}
DateFileName <- function(x){
  name <-   paste0(today(), "_", x)
  print (name)
  return (name)
}

ScriptNamedDir <- function(scriptName = NULL){
  if(is.null(scriptName))
    scriptName <- rstudioapi::getActiveDocumentContext()$path
  if (is.null (scriptName) || scriptName == "")
    stop("No script name found -- you may need to save this file first")
  outDir <- gsub(".R(md)?$", "_data", scriptName, ignore.case = TRUE)
  stopifnot( outDir != scriptName)
  if (!dir.exists(outDir)){
    message ("Creating directory associated with ", scriptName,", at ", outDir)
    dir.create(outDir)
  }
  return(outDir)
}

ScriptAndDatedFileName <- function(x, scriptName = NULL){
  dir <- ScriptNamedDir(scriptName)
  fileName <- DateFileName(x)
  path <- file.path(dir, fileName)
  print (path)
  return (path)
}

GetLatestScriptFile <- function(x, scriptName=NULL){
  stopifnot (length(x) == 1)
  dir <- ScriptNamedDir(scriptName)
  filePattern <- paste0("^\\d{4}_\\d{2}_\\d{2}_", x, "$", collapse = "")
  filesFound <- list.files(dir, filePattern)
  stopifnot (length(filesFound) > 0)
  if (length(filesFound) > 1){
    message ("Multiple files  with matching names found.  Using the last one")
    print (filesFound)
  }
  return (file.path(dir, tail(filesFound, 1)))
} 

PDFBackupFileName <- function(prefix = "", subDir = ""){
  scriptDir <- ScriptNamedDir()
  imageDir <- file.path(scriptDir, "pdfs", subDir)
  if (!dir.exists(imageDir)) dir.create(imageDir, recursive = TRUE)
  now <- format(Sys.time(),  "%Y_%m_%d__%H_%M__%S")
  counter <- 0
  path <- file.path (imageDir, sprintf("%s%s.%02d.pdf", prefix, now, counter))
  while (file.exists(path)){
    counter <- counter + 1
    path <- file.path (imageDir, sprintf("%s%s.%02d.pdf", prefix, now, counter))
  }
  return (path)
}

BackupAsPDF <- function(graphics, prefix = "", subDir = "", dimensions = NULL){
  path <- PDFBackupFileName(prefix, subDir)
  if (is.null(dimensions))
  dimensions <- dev.size(units = "in")

  print (sprintf("Writing image to  %s", path))
  cairo_pdf(path, width = dimensions[1], height = dimensions[2])
  
  # handle functions, my enrichment heatmaps that are part of a list
  if ("function" %in% class(graphics)){
    graphics()
    g <- "finished" # something to print to console instead of graphics to device 
  }else if (! ("ggplot" %in% class(graphics) | "grob" %in% class(graphics) | "Heatmap" %in% class(graphics) | "HeatmapList" %in% class(graphics))){
    g <- graphics$hmList    
  }  else{
    g <- graphics
  }
  print (g)
  
  dev.off()
  return (graphics)
}





WriteSessionInfo <- function(path = NULL){
  if (is.null(path))
    path <- ScriptAndDatedFileName("SessionInfo.txt")
  si <- devtools::session_info()
  fileOut <- file(path, open = "wt")
  writeLines("─ Session info ─────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────",
             con = fileOut)
  write.table(data.frame( value = unlist(si[[1]])), fileOut)
  #writeLines(capture.output(data.frame(value = unlist(si[[1]]))), con = fileOut) # for some mysterious reason, this always goes to the notebook output when run in a notebook. type = "message" is no help...[shrug]
  writeLines(capture.output(data.table(setting = names(si[[1]]), value = unlist(si[[1]]))), con = fileOut) # this too. Maybe an RStudio version issue
  writeLines("─ Packages ─────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────",
             con  = fileOut)
  #write.table(si[[2]], fileOut)
  writeLines(capture.output(si[[2]]), con = fileOut)
  close(fileOut)
}

WriteInstalledPackages <- function (path = NULL){
  if (is.null(path))
    path <- ScriptAndDatedFileName("Installed.Packages.csv")
  package.mat <- installed.packages()
  fwrite (as.data.table(package.mat, keep.rownames = TRUE), path)
}

WriteSessionInfo()
WriteInstalledPackages()

```
 

## 051623 DSwaney Project

Read in spectronaut data & metadata

----
```{r read-in}
path <- getwd()
path
AB_input <- fread(paste0(path, '/docs/MartinGordon_MSAnalysis_forDanielle/AA08/AB/20230516_094910_AA08_AB_Report.csv'))
AB_meta <- fread(paste0(path, '/docs/MartinGordon_MSAnalysis_forDanielle/AA08/AB/AA08_AB_ConditionSetup.xls'))
AB_input$ProteinName <- sub(">", "", AB_input$ProteinName)
```



```{r data-inspection}
summary(AB_input)

AB_input$BioReplicate <- as.factor(AB_input$BioReplicate)
AB_input$Run <- as.factor(AB_input$Run)
AB_input$Condition <- as.factor(AB_input$Condition)
```
Data inspection
---------

```{r data structure}
# order by Condition
setorder(AB_input, Condition)
# One run per condition per replicate
AB_input[, .N, by = .(Run, Condition, BioReplicate)]
```

Remove '>' preceeding some ProteinNames
```{r}
AB_input$ProteinName <- sub(">", "", AB_input$ProteinName)
```



Bioreplicate numbers (1-4) are repeated across conditions. Modified as msstats assumes repeated bioreplicate ids indicate a time-series analysis.
Asigned each cond/biorep combo a unique ID (1-20) for msstats to model as a case-control experiment.

```{r}
# create new ID col for pairwise comparisons
AB_input[, ID:=.GRP, by=.(Condition,BioReplicate)]

#reassign to biorep col and drop ID
AB_input[, BioReplicate:=ID]
AB_input[, ID:=NULL]

#sanity check
AB_input[, .N, by = .(Condition,BioReplicate)]
```

Removed spectronaut peptides with intensities < 2^5 as likely artifacts of the identification/quantification tool.

```{r}
ggplot(data=AB_input,  mapping=aes(x=log2(Intensity))) +
  geom_histogram(binwidth = 0.1)

AB_input <- AB_input[log2(Intensity) > 5] #filter out low abundance 

ggplot(data=AB_input,  mapping=aes(x=log2(Intensity))) +
  geom_histogram(binwidth = 0.1)

```

Identify if there any multiple-feature peptide ions in the runs. We expect one peptide feature per row per run, so identify duplicates.
No multi-feature peptides detected

```{r}
any(AB_input[, .N, by = .(Run, PeptideSequence, PrecursorCharge)]$N) > 1
```
Plot of number of peptide ion measurements per run. Counts seem pretty consistent across replicates & conditions

```{r}
### plot of peptide ions per run
p <- AB_input %>%  
      ggplot(mapping=aes(x=reorder(interaction(Condition,BioReplicate)), fill=Condition)) +
      geom_bar() + 
      theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

BackupAsPDF(p, "BarPlot_NPeptides_")
```
Boxplots of distributions of raw intensity values per condition/replicate. Again abundances are pretty consistent and no real concerns with raw data quality. 
```{r create contrasts}
p <- ggplot(data = AB_input, mapping=aes(x=reorder(interaction(Condition, BioReplicate)), y=log2(Intensity), fill=Condition)) +
      geom_boxplot() +
      theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

BackupAsPDF(p, "BoxPlot_RawIntensities_")
```
Data Processing
----
Convert spectronaut data to msstats format
```{r msstats-formatting}
ab.mss <- specFileToCompleteMSstats(AB_input)
ab.mss[, IsotopeLabelType := 'L'] #label free
setDF(ab.mss)  # convert to DF for msstats processing

```
MSStats DataProcess: Normalise (equilizemedians), filter & summarize peptides to protein intensities

```{r mss-dataproc}

dp.out <- MSstats::dataProcess(ab.mss,
                               normalization = 'equalizeMedians',
                               summaryMethod="TMP",
                               featureSubset = "highQuality",
                               remove_uninformative_feature_outlier = TRUE,
                               MBimpute = FALSE )
```
Write out protein level data to file (provide feature level data also)
```{r writeout-data}
# the main data of interest is this one:
fwrite(dp.out$ProteinLevelData, ScriptAndDatedFileName("ProteinLevelData.csv.gz"))

#this table is mostly a copy of the input with some additional columns for hte post-normalization intensity (labeled ABUNDANCE), and columns for is_outlier and is_informative
fwrite(dp.out$FeatureLevelData, ScriptAndDatedFileName("FeatureLevelData.csv.gz"))
```
Boxplot of protein intensities post-normalization:

```{r}
prot.quant <- setDT(dp.out$ProteinLevelData)

p <- prot.quant %>% 
      ggplot(mapping=aes(x=interaction(GROUP,SUBJECT), y=LogIntensities, fill=GROUP)) +
      geom_boxplot() +
      theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

BackupAsPDF(p, "BoxPlot_NormalizedProteinIntensities_")
```

Convert uniprot ids to gene names
Some protein names contain illegal characers ('_'); convert to '-'
```{r}
prot.quant$Protein <- gsub("_", "-", prot.quant$Protein)
prot.quant[, gene := multiUniprots2multiGenes(Protein)]
```
Clustering
-----

```{r create-matrix}
title = ""
#data reshaping - long to wide format.
#protein vs interaction group and rep
intensity.mat <- as.matrix(dcast(prot.quant, Protein ~GROUP+SUBJECT, value.var = "LogIntensities"),
                           rownames = "Protein")

```

Subsampled 1000 proteins and clustered.
At high level, two distinct clusters formed:PAR (neg & pos) & WT,EV & N10T20.
*Note* Clustering dominated by protein intensites, scale the data and replot

```{r subsample-heatmap}
#sub in 0 for NA val
rowClusterWithNA <- function(mat, na.value = 0, ...){
  mat[is.na(mat)] <- na.value
  hclust(dist(mat), ...) #euclidean
}

subMat <- intensity.mat[sample.int(nrow(intensity.mat), 1000),]

#create heatmap annotation col
type <- gsub("_[0-9]{1,2}", "", colnames(subMat))
ha <- HeatmapAnnotation(type = type, annotation_name_side = "right")

hm <- Heatmap(subMat,
        #clustering_distance_rows = function(m) dist(m, method = 'euclidean'),
        #clustering_method_rows= "complete",
        top_annotation = ha,
        cluster_rows = rowClusterWithNA(subMat),
        show_row_names = FALSE,
        row_title = sprintf("%d subsampled proteins", nrow(subMat)),
        name = "Log2 Int.")

hm
#BackupAsPDF(draw(ha), 'heatmap_subsample_log2intensities')
rm(hm)
```
Above heatmap appears to primarily cluster the low intensity proteins together and then the high intensity proteins, and the run-to-run trends appear secondary. 
Subtracted median protein expression intensities to observe run-to-run trends.
Again, we consistently see a seperation between PAR cluster and other groups.

```{r scaled-subsample-heatmap}
#row median - median prot exp value across experiment

# sweep subtracts a vector of values from either rows (dimension = 1) or columns (dimension = 2)
# apply applies a function (median here) to each row (dimension = 1 )
# used together here we are subtracting row medians
scaled.intensity.mat <- sweep(intensity.mat, 1, apply (intensity.mat, 1, median, na.rm = TRUE))


scaled.subMat <- scaled.intensity.mat[sample.int(nrow(scaled.intensity.mat), 1000),]

hm <- Heatmap(scaled.subMat,
        top_annotation = ha,
        cluster_rows = rowClusterWithNA(scaled.subMat),
        show_row_names = FALSE,
        row_title = sprintf("%d subsampled proteins", nrow(scaled.subMat)),
        name = "Log2 Int.")

hm
#BackupAsPDF(draw(hm), 'heatmap_subsample_scaled')

```
Heatmap of all clustered proteins (proteins with no missing values)

```{r allprot-heatmap}
# pca requires no missing values.  Function complete.cases finds rows with no missing values
complete.mat <- scaled.intensity.mat[complete.cases(scaled.intensity.mat),]

hm <- Heatmap(complete.mat,
       top_annotation = ha,
       row_title = sprintf ("%d complete proteins", nrow(complete.mat)),
       name = "Log2 Int\nvs Median",
       show_row_names = FALSE,
       cluster_columns = FALSE)

BackupAsPDF(draw(hm), 'heatmap_prots_scaled')
```
PCA plot
-----

PCA plots seem pretty consistent with clustering output; first principal component clearly distinguishes PAR groups and others. 
EV, WT and N10T20 groups cluster closely
```{r pca}
pcaOut <- prcomp(t(complete.mat))
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaDT[, mainGroup := tstrsplit(rn, "_")[[1]]] #transpose & split


pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)

p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color = mainGroup )) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA using %d proteins (log intensity)", nrow(complete.mat))) 
p


BackupAsPDF(p, "PCA_Complete_Proteins")
```
Differential Expression Analysis
------

Perform DEA with provided contrasts
Generate a contrast matrix for requested pairwise comparisons:

```{r make contrasts}
#all pairwise; need to customise
dp.contrasts <- MSstatsContrastMatrix("pairwise", unique(dp.out$ProteinLevelData$GROUP))

#make new contrasts
dp.contrasts <- dp.contrasts[c(3,8),]

## option 2: customized comparison
WTpos_N10t20pos <- matrix(c(0,-1,0,0,1),nrow=1)
WTpos_PARpos <- matrix(c(0,0,0,-1,1),nrow=1)
WTpos_EVpos <- matrix(c(-1,0,0,0,1),nrow=1)

dp.contrasts <- rbind(dp.contrasts, WTpos_EVpos, WTpos_N10t20pos, WTpos_PARpos)
row.names(dp.contrasts)<-c("EVpos vs PARpos","PARneg vs PARpos","WTpos vs EVpos","WTpos vs N10t20pos","WTpos vs PARpos")
rm(WTpos_EVpos,WTpos_N10t20pos,WTpos_PARpos)

dp.contrasts
```
Write output to file.

```{r mss-pairwise}
test.mss <- groupComparison(contrast.matrix=dp.contrasts, data=dp.out)
test.mss.all <- setDT(test.mss$ComparisonResult)

# adjust p values using the Benjamini Hochberg method, which estimates False Discovery Rate
test.mss.all[, fdr := p.adjust(pvalue, method = "BH")] 
#get gene names
test.mss.all[, gene := multiUniprots2multiGenes(as.character(Protein), simplify = TRUE, allowDups = FALSE)]


fwrite(test.mss.all, ScriptAndDatedFileName("msstats_groupcomparisons.csv.gz"))
```
```{r rformat test-mms}
View(test.mss.all)

reform.test.mss <- test.mss.all[, .(Protein,Label,log2FC,pvalue,fdr,gene)] # extract required columns
reform.test.mss <- dcast(reform.test.mss, Protein+gene+pvalue + fdr ~ Label, value.var="log2FC")
fwrite(reform.test.mss, ScriptAndDatedFileName("log2fc_groupcomparisons_all.csv"))

# maybe filter this for FDR to see log change
sig.reform.test.mss <- test.mss.all[fdr <= 0.05,]
sig.reform.test.mss <- dcast(sig.reform.test.mss, Protein+gene ~ Label, value.var="log2FC")
rownames(sig.reform.test.mss) <- sig.reform.test.mss$Protein
sig.reform.test.mss[,Protein:=NULL]
fwrite(sig.reform.test.mss, ScriptAndDatedFileName("log2fc_groupcomparisons_siggenes.csv"))
```

```{r}
#Also, a file that rows as proteins, and then columns are gene, and the log2 intensity value in each condition. Basically like what would go into this heatmap, but also include proteins for which there are NA values for some conditions
# use prot.quant file


log2.fc.out <- prot.quant[, Sample:=paste0(GROUP,"_",SUBJECT)] %>% 
  dcast(Protein+gene~Sample, value.var = 'LogIntensities') 

fwrite(log2.fc.out, ScriptAndDatedFileName("log2intensities_groupcomparisons.csv"))

```


Generate Volcano plots for each of the pairwise comparisons

```{r multi-volcanoplot}
## volcano plots

# define significant proteins
test.mss.all[, sig := "Not"]
test.mss.all[fdr < 0.05 & abs(log2FC) > log2(1.5),sig := ifelse (log2FC  > 0, "Up", "Down")]

p <- ggplot (test.mss.all, aes (x = log2FC, y = -log10(fdr), color = sig, label = gene)) +
  geom_point() +
  scale_color_manual(values = c(Not = "gray", Down = "blue", Up = "red")) +
  ggrepel::geom_text_repel(data = test.mss.all[sig != "Not"]  ,  size = 2, max.overlaps = 20) +
  theme_bw() +
  facet_wrap(~ Label)

BackupAsPDF(p, "volcanoplot_multipanel")
```

Over plotting. Generate each volcano plot separately.
```{r split-dt}
#split datatable by comparison

test.mss.list <- split(test.mss.all, test.mss.all$Label)
```

```{r cal-geommean}
#calculate geomemtric mean
test.mss.list <- test.mss.list %>% 
  map( ~{
    dt <- .x
    dt$gm.pval.log2fc <- sqrt(dt[,'fdr']*abs(dt[,'log2FC']))
    
    return(dt)
  })

```


```{r volcanoplots}
# use purr to plot each list element 


  imap(test.mss.list, ~ {
    dt <- .x
    dt.name <- .y
    
    p <- ggplot (dt, aes (x = log2FC, y = -log10(fdr), color = sig, label = gene)) +
      geom_point() +
      scale_color_manual(values = c(Not = "gray", Down = "blue", Up = "red")) +
      ggrepel::geom_text_repel(data = dt[sig != "Not"][order(gm.pval.log2fc)][1:40]  ,  size = 2, max.overlaps = 20) +
      ggtitle(paste0(dt.name, ' Volcano plot')) +
      theme_minimal()
    p
   #BackupAsPDF(p, paste0(dt.name,"_volcanoplot"))
  })
```

Perform enrichment analysis for each of the comparisons
```{r}
#creat new factor col of combined label sig info
test.mss.all[, enrich_grp:= interaction(Label,sig)]

# all go ontology genes (biological process, cellular componetn & metabolic function) for enrichment analysis
gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")

# all identified genes as the background set
universe <- unique(test.mss.all$gene)

# perform enrichment on significant genes for all groups
enrich.dt <- enricherOnGroups(test.mss.all[test.mss.all$sig != 'Not'], groupColumns = 'enrich_grp', term2gene.gmt = gmt.go, universe = universe)

#simplify enrich results to remove redundant GO terms
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')


# failed.. troubleshoot
enrichmentOnGroupsPL(test.mss.all[test.mss.all$sig != 'Not'], geneColumn = 'gene', groupColumns = 'enrich_grp', gmt = gmt.go, universe = universe)

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, NULL, groupColumn = 'enrich_grp', topN = 8, negCols =unique(simp.enrich$simplified$enrich_grp[grep('Down', simp.enrich$simplified$enrich_grp)]), row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 4), upperThreshold = 5)  # compare enrichment in the top 10 terms per group
BackupAsPDF(ht, "heatmap_go_enrichment")
```


