---
title: "230623_QLi_DARPA"
author: "Martin Gordon"
date: "2023-06-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary
Perform Differential Expression Analysis of MS-MS Spectronaut data using MSstats
Data from DARPA project: Single timepoint with 6 conditions in total: 3 ligands vs NPFFR1, treated vs untreated with 30 min exposure to NPFF
Perform pairwise contrasts for each of the conditions

###: Contact: Qiongyu Li
###: Date: 23-06-23

```{r packages}
library(purrr)
library(tidyr)
library(dplyr)
library(data.table)
library(MSstats)
library(ggplot2)
library(magrittr)
library(circlize)
library(ComplexHeatmap)
library(parallelDist) #needed for some custom enrichment functions
library(ggvenn)

#source Bens utility scripts
source("/Users/martingordon/Documents/utils/bp_utils/MSstats_V4_Functions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/MSstats_Helper_Functions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
```

Read in spectronaut data & metadata

----
```{r read-in}
getwd()
AB_input <- fread(paste0(path, '/docs/MartinGordon_MSAnalysis_forDanielle/AA08/AB/20230516_094910_AA08_AB_Report.csv'))
AB_meta <- fread(paste0(path, '/docs/MartinGordon_MSAnalysis_forDanielle/AA08/AB/AA08_AB_ConditionSetup.xls'))
AB_input$ProteinName <- sub(">", "", AB_input$ProteinName)
```



```{r data-inspection}
summary(AB_input)

AB_input$BioReplicate <- as.factor(AB_input$BioReplicate)
AB_input$Run <- as.factor(AB_input$Run)
AB_input$Condition <- as.factor(AB_input$Condition)
```
Data inspection
---------

```{r data structure}
# order by Condition
setorder(AB_input, Condition)
# One run per condition per replicate
AB_input[, .N, by = .(Run, Condition, BioReplicate)]
```

Remove '>' preceeding some ProteinNames
```{r}
AB_input$ProteinName <- sub(">", "", AB_input$ProteinName)
```



Bioreplicate numbers (1-4) are repeated across conditions. Modified as msstats assumes repeated bioreplicate ids indicate a time-series analysis.
Asigned each cond/biorep combo a unique ID (1-20) for msstats to model as a case-control experiment.

```{r}
# create new ID col for pairwise comparisons
AB_input[, ID:=.GRP, by=.(Condition,BioReplicate)]

#reassign to biorep col and drop ID
AB_input[, BioReplicate:=ID]
AB_input[, ID:=NULL]

#sanity check
AB_input[, .N, by = .(Condition,BioReplicate)]
```

Removed spectronaut peptides with intensities < 2^5 as likely artifacts of the identification/quantification tool.

```{r}
ggplot(data=AB_input,  mapping=aes(x=log2(Intensity))) +
  geom_histogram(binwidth = 0.1)

AB_input <- AB_input[log2(Intensity) > 5] #filter out low abundance 

ggplot(data=AB_input,  mapping=aes(x=log2(Intensity))) +
  geom_histogram(binwidth = 0.1)

```

Identify if there any multiple-feature peptide ions in the runs. We expect one peptide feature per row per run, so identify duplicates.
No multi-feature peptides detected

```{r}
any(AB_input[, .N, by = .(Run, PeptideSequence, PrecursorCharge)]$N) > 1
```
Plot of number of peptide ion measurements per run. Counts seem pretty consistent across replicates & conditions

```{r}
### plot of peptide ions per run
p <- AB_input %>%  
      ggplot(mapping=aes(x=reorder(interaction(Condition,BioReplicate)), fill=Condition)) +
      geom_bar() + 
      theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

BackupAsPDF(p, "BarPlot_NPeptides_")
```
Boxplots of distributions of raw intensity values per condition/replicate. Again abundances are pretty consistent and no real concerns with raw data quality. 
```{r create contrasts}
p <- ggplot(data = AB_input, mapping=aes(x=reorder(interaction(Condition, BioReplicate)), y=log2(Intensity), fill=Condition)) +
      geom_boxplot() +
      theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

BackupAsPDF(p, "BoxPlot_RawIntensities_")
```
