---
title: "Alligator RNAEdits"
author: "Martin Gordon"
date: "2024-05-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## packages

```{r}
library(data.table)
library(ggplot2)
library(magrittr)
library(stringr)
source("../../utils/bp_utils/ManageScriptData.R")
```
Combine the individual dts into one large datatable

```{r}
files <- dir(path='./data/filtered/', pattern='*filtered.tsv', recursive = T, full.names = T)
files
files.list <- lapply(files, fread, header=F)
names(files.list) <- gsub("(?<=[.][0-9]{1}).+", "", gsub('./data/filtered//|.reditools.filtered.tsv|_table','',files), perl = T)

# remove data tables with no values
files.list <- files.list[sapply(files.list, nrow) > 0]

edits.dt <- rbindlist(files.list, idcol='sample', fill=T)
```

Read in the output of the redditools command and check the formatting

```{r}
edits.dt[, condition := tstrsplit(sample, '_', keep=1)]
edits.dt
setnames(edits.dt, new=c('sample', 'chr', 'position', 'reference', 'strand', 'coverage_rna', 'meanQuality_rna', 'baseCount_rna[A,C,G,T]', 'allSubs_rna', 'frequency_rna', 
                         'coverage_dna', 'meanQuality_dna', 'baseCount_dna[A,C,G,T]', 'allSubs_dna', 'frequency_dna', 'condition'))

col.oi <- c('condition', 'chr', 'position', 'reference', 'strand', 'coverage_rna', 'meanQuality_rna', 'baseCount_rna[A,C,G,T]', 'allSubs_rna', 'frequency_rna', 
                         'coverage_dna', 'meanQuality_dna', 'baseCount_dna[A,C,G,T]', 'allSubs_dna', 'frequency_dna')


fwrite(edits.dt[frequency_dna <= 0.05, ..col.oi], ScriptAndDatedFileName('reditool.filtered.csv'))


fread('~/Documents/projects/051524_JMoen_AlligatorEditSites/RNAedit_QC_data/2024_05_15_reditool.filtered.csv')

edits.dt[frequency_dna <= 0.05][order(-frequency_rna)][,.N, by=condition]
edits.dt$strand %>%  unique()
```

052124
----
Read in the rerun data with updated REDItools params

```{r}
files <- dir(path='./data/anno_0517', pattern='*anno.tsv', recursive = T, full.names = T)
files.list <- lapply(files, fread, header=F)

names(files.list) <- gsub("(?<=[.][0-9]{1}).+", "", gsub('./data/anno_0517/|.reditools.anno.tsv|_table','',files), perl = T)
names(files.list)
# remove data tables with no values
files.list <- files.list[sapply(files.list, nrow) > 0]

edits.dt <- rbindlist(files.list, idcol='sample', fill=T)

```

Add col names
```{r}
edits.dt[, condition := tstrsplit(sample, '_', keep=1)]

setnames(edits.dt, new=c('sample', 'chr', 'position', 'reference', 'strand', 'coverage_rna', 'meanQuality_rna', 'baseCount_rna[A,C,G,T]', 'allSubs_rna', 'frequency_rna', 
                         'coverage_dna', 'meanQuality_dna', 'baseCount_dna[A,C,G,T]', 'allSubs_dna', 'frequency_dna', 'feature_type', 'gene', 'feature_id','condition'))

edits.f.dt <- edits.dt[frequency_rna >= 0.1 & frequency_dna <= 0.1 & coverage_rna >= 10,]
#
fwrite(edits.f.dt, ScriptAndDatedFileName('editSites.minimalfiltering.annotated.csv'))
```
Filter out the unedited sites from the data

```{r}
edits.f.dt[,.(treatment = ifelse(grepl('GFC', condition), 'GFC', 'GFV'), .N), by=condition]

g <- ggplot(edits.f.dt[,.(treatment = ifelse(grepl('GFC', condition), 'GFC', 'GFV'), .N), by=condition], aes(x=condition, y=N, fill=treatment)) +
  geom_bar(stat = 'Identity') +
  scale_fill_brewer(type='qual', palette=2) +
  ggtitle('N edit sites') +
  theme_bw()

BackupAsPDF(g, 'NeditSites.barplot')

```
Lets look at the number of 

```{r}
edits.f.dt[, editType := ifelse(grepl('AG|TC', allSubs_rna), 'A-I', 
                                ifelse(grepl('CT|GA', allSubs_rna), 'C-U', 'other'))]


g <- ggplot(edits.f.dt[,.(treatment = ifelse(grepl('GFC', condition), 'GFC', 'GFV'), .N, editType), by=condition], aes(x=condition, y=N, fill=editType)) +
  geom_bar(stat = 'Identity') +
  scale_fill_brewer(type='qual', palette=2) +
  ggtitle('N edit sites') +
  theme_bw()
g


edits.f.dt[editType == 'other',]
```

write out the table

```{r}
fwrite(edits.f.dt, ScriptAndDatedFileName('editSites.minimalfiltering.csv'))
edits
```


05-23-24
---
Read in the annotated data and check for the conserved modified sites 
```{r}
edits.dt <- fread('~/Documents/projects/051524_JMoen_AlligatorEditSites/RNAedit_QC_data/2024_05_21_editSites.minimalfiltering.csv')
c.sites.dt <- fread('~/Documents/projects/051524_JMoen_AlligatorEditSites/docs/mammilian.conserved.sites.txt', fill=T, rownames=F)

setnames(c.sites.dt, new=c('N',' chr', 'pos', 'strand', 'gene', 'feature','HGVS'))
```

Looking v quickly, seems two of the genes have multiple conserved sites
TSHZ2 DACT3

```{r}
edits.dt[gene %in% c.sites.dt[, unique(gene)], ]

# none of thes three sites have been found 
edits.dt[gene %in% c('GRIA1','CYFIP','FLNA'),]

edits.dt[grep('CDS', feature_type),]


edits.dt[frequency_rna >= 0.1 & frequency_dna < 0.1 & grepl('CDS', feature_type),]$gene %>%  unique()

grep()
```
Not used...
```{r}
# Step 1: Convert the character column to a list of numeric vectors
edits.dt[, list_column := lapply(strsplit(gsub("\\[|\\]", "", `baseCount_rna[A,C,G,T]`), ","), function(x) as.numeric(x))]
# step 2: Sort each list and extract the second highest value
edits.dt[, variantCount_rna := sapply(list_column, function(x) sort(x, decreasing = TRUE)[2])]
```