---
title: "012225_requestedPlots"
author: "Martin Gordon"
date: "2025-01-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Goal of this analysis is to produce some meeting ready plots for Ronald to share with Nevan ahead of the meeting on Friday
To do: plot showings summary of data (boxplots, scatterplots of replicates, correlation matrix)

```{r}
library(data.table)
library(ggplot2)
library(magrittr)
library(ggrepel)
library(stringr)
library(ComplexHeatmap)
library(circlize)
library(viridis)
library(RColorBrewer)
library(hrbrthemes) # visually nice set of themes
library(patchwork)
library(showtext)
library(seqinr)
library(readxl)
library(tidymodels)
library(ggridges)
library(RColorBrewer)


# fonts needed for hrbrthemes
library(extrafont)
font_import()
loadfonts()

# directly from google font
sysfonts::font_add_google("Roboto Condensed")
showtext_auto()

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/enrichmentTestFunctions.R")
source ("../../utils/mg_utils/r_utils/IDmapping.R")
source ("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")

source("../../utils/mg_utils/r_utils/CRISPR_Functions.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0,...){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}

customTheme <- theme_ipsum_rc() +
  theme(panel.border = element_rect(color = "lightgrey", 
                                    fill = NA, 
                                    size = 1),
        axis.text.x = element_text(angle=90)
        )

#set one
col.pal <- getQualitativePalette(n=13)
#col.pal <- randomcoloR::distinctColorPalette(k=13)
```

Before any of this, lets plot the distributions of the raw counts and ensure we are not 'normalizing out` true differences between the conditions

Take a look at the data

```{r}
# deseq2 contrasts for the original data
deseq.dt <- fread('~/Documents/projects/101224_RBabu_CRISPRe_PIK3CA/111824_plotsForKroganWeekly_data/2025_01_22_deseq.pwcomparisons.allAnno.csv') 


raw.counts <- fread('./output/141024_noNormalization.count.txt')
raw.dt <- melt(raw.counts, id.vars = c('sgRNA', 'Gene'))
#setnames(raw.dt, old=c('variable', 'value'), new=c('sample', 'counts'))

raw.meta <- fread('./output/141024_noNormalization.countsummary.txt')
raw.meta[, sample := gsub('/wd/data/fastq/bclconvert_defaults/|_S[0-9]{1,2}_R1.+', '', File)]
raw.meta[, sample := ifelse(grepl("ABE8ENG_", sample), gsub("ABE8ENG_", "abe8e_", sample),  gsub("BE4MaxNG_", "bemax_", sample))]
raw.meta[, sample := ifelse(grepl("ABE8eNG_", sample), gsub("ABE8eNG_", "abe8e_", sample),  gsub("BE4MaxBG_", "bemax_", sample))]
raw.meta[, sample := ifelse(grepl("_CTRL", sample), gsub("_CTRL", "_Ctrl", sample),  gsub("BE4MaxBG_", "bemax_", sample))]

raw.meta[, c('editor', 'treatment', 'timepoint') := tstrsplit(sample, '_', keep=c(1,2,3))]
raw.meta[is.na(timepoint), timepoint := 'day0']
raw.meta[, timepoint := gsub('day', '', timepoint)]

raw.meta[, replicate := str_extract(treatment, '[1,2]$')]
raw.meta[is.na(replicate), replicate := '1']

raw.meta[, treatment := gsub('[0-9]$', '', treatment)]
raw.meta[grepl('[Ll]ib[-]{0,1}', treatment), treatment := 'LibraryRepresentation']

# check the sample IDs in the nromalized set also match up
raw.meta[, .(Label, sample, editor, treatment, timepoint, replicate)]

raw.dt <- merge(x=raw.dt, y=raw.meta[,.(Label, sample, editor, treatment, timepoint, replicate)], by.x=c('variable'), by.y=c('Label'), all.x=T)
setnames(raw.dt, old=c('variable', 'value'), new=c('Label', 'counts'))
```

Boxplot of the raw counts
```{r}
ggplot(raw.dt, aes(x=sample, fill=sample, y=log2(counts+1))) +
  geom_boxplot() +
  customTheme
```
much lower but trends follow the same as the above sample
the fill is the sgSet which clearly follows trends of the other samples
```{r}
ggplot(raw.dt[grepl('sgINTERGENIC', Gene),], aes(x=sample, fill=sample, y=log2(counts+1))) +
  geom_boxplot() +
  geom_boxplot(data=raw.dt, aes(x=sample, y=log2(counts+1)), color='red', outliers =F) +
  customTheme
```
Try the log2 normalization method

```{r}
raw.mat <- dcast(raw.dt, paste(Gene,sgRNA, sep='.')~paste(editor,treatment, timepoint, replicate, sep  = '_'), value.var='counts') %>% 
  as.matrix(rownames=1)

cpm.mat <- apply(raw.mat, 2, function(x){(x/sum(x, na.rm=T)) * 1000000 + 1})
cpm.mat <- log2(cpm.mat)

cpm.long <- setDT(melt(cpm.mat))

cpm.long[, c('gene', 'sgRNA') := tstrsplit(Var1,'[.]')]
cpm.long[, c('editor', 'treatment', 'timepoint', 'replicate') := tstrsplit(Var2, '_')]
setnames(cpm.long, old=c('value'), new=c('log2_CPM'))

setnames(raw.dt, old='Gene', new='gene')

raw.dt <- merge(raw.dt[treatment != 'Undetermined',], cpm.long[treatment != 'Undetermined',-c('Var1', 'Var2')], by=c('gene', 'sgRNA', 'editor', 'treatment', 'timepoint', 'replicate'))
raw.dt %>%  colnames()
```

Look at the distributions of cpm vs raw counts for each of the samples

```{r}
raw.dt[grepl('sgINTERGENIC', gene), gene :='Intergenic']

raw.dt[treatment == 'Ctrl', treatment := 'DMSO']
raw.dt[gene == 'non-targeting', gene := 'Non-targeting']

# log2FC look much. better, but still not great
ggplot(raw.dt, aes(x=sample, y=log2(counts+1)), color='black') +
  geom_boxplot() +
  geom_boxplot(data=raw.dt, aes(x=sample, y=log2_CPM), color='red', outliers =F) +
  customTheme
```
Read in the other count data for these files 
These may still be in matrix format... reformat like this and add to the col
```{r}
norm.dt <- fread('101524_MAGECKContrasts_firstPass_data/2024_10_15_sgRNAcounts.normalized.txt')
setnames(norm.dt, old=c('Gene', 'normCounts'), new=c('gene', 'deseq.norm.counts'))

norm.dt
raw.dt[sgRNA == 'BE_1',] %>% 
  .[order(sample)]
norm.dt[sgRNA == 'BE_1',] %>% 
  .[order(sample)]

raw.dt[, unique(sample)]

raw.dt[editor == 'Plasmid', editor := 'plasmid']
raw.dt[, sample := paste(editor, treatment, timepoint, replicate, sep='.')]

norm.dt[, sample := paste(editor, treatment, timepoint, rep, sep='.')]
norm.dt[treatment == 'control', treatment := 'DMSO']
norm.dt[treatment == 'librep', treatment := 'LibraryRepresentation']
norm.dt[, timepoint := as.character(timepoint)]
norm.dt[, replicate := as.character(rep)]
norm.dt[grepl('sgINTERGENIC', gene), gene := 'Intergenic']
norm.dt[grepl('non-targeting', gene), gene := 'Non-targeting']

# ok this has all the normalized counts we need, just write the output to file and generate the plots tomorrow
raw.dt <- merge(x=raw.dt, y=norm.dt[,.(gene, sgRNA, editor, treatment, timepoint, replicate, deseq.norm.counts, sgRNAnormCounts)], by=c("gene","sgRNA","editor","treatment","timepoint","replicate"))

#fwrite(test, ScriptAndDatedFileName('sampleCounts.rawAndNormalized.dt'))
#fwrite(raw.dt, ScriptAndDatedFileName('sampleCounts.rawAndNormalized.dt'))
```
Will consider editors seperately as have poor QC of the 
col.pal for treatments/timepoint/editor. keep these the same for all plots

```{r}
brewer.pal(n=9, name = 'Blues')[4:8]
brewer.pal(n=9, name = 'PuRd')[4:8]
brewer.pal(n=9, name = 'Reds')[4:8]
brewer.pal(n=9, name = 'YlOrRd')[4:8]
brewer.pal(n=9, name = 'YlGn')[4:8]
brewer.pal(n=9, name = 'Greys')[4:8]
# blues

group.pal <- c("#BDBDBD", "#DF65B0", "#C994C7", "#6BAED6", "#9ECAE1","#FB6A4A", "#FC9272" )
names(group.pal) <- unique(counts.dt$group)

"#FC9272" "#FB6A4A" "#EF3B2C" "#CB181D" "#A50F15"
"#9ECAE1" "#6BAED6" "#4292C6" "#2171B5" "#08519C"
"#C994C7" "#DF65B0" "#E7298A" "#CE1256" "#980043"
"#ADDD8E" "#78C679" "#41AB5D" "#238443" "#006837"
"#BDBDBD" "#969696" "#737373" "#525252" "#252525"
"#FEB24C" "#FD8D3C" "#FC4E2A" "#E31A1C" "#BD0026"


# timepoint and editor col pal
counts.dt$timepoint %>% unique()

group.pal <- c("#BDBDBD", "#DF65B0", "#C994C7", "#6BAED6", "#9ECAE1","#FB6A4A", "#FC9272" )
names(group.pal) <- unique(counts.dt$group)

timepoint.pal <- c("#ADDD8E" ,"#78C679", "#41AB5D")
names(timepoint.pal) <-c('0','7', '22')

editor.pal <- brewer.pal(n=9, name = 'Set3')[4:6]
names(editor.pal) <- c('abe8e', 'bemax', 'plasmid')
```



read in the counts
```{r}
counts.dt <- fread('~/Documents/projects/101224_RBabu_CRISPRe_PIK3CA/012825_requestedMeetingPlots_data/2025_01_29_sampleCounts.rawAndNormalized.dt')
counts.dt[, group := paste(treatment, timepoint, sep='_')]
```

Find a nice color palette for the different conditions and stick with that throughout plots

scatterplots of day0 replicates to show sample consistency
Lets also color the labels according 
```{r}
counts.wide <- dcast(counts.dt, paste(gene, sgRNA, sep='.')~paste(editor, treatment, timepoint, replicate, sep='_'), value.var = 'sgRNAnormCounts') %>% 
  as.matrix(rownames=1)

counts.wide <- log2(counts.wide +1)

cor.mat <- cor(counts.wide[, grepl('abe8e', colnames(counts.wide))], use='pairwise.complete.obs', method='pearson')
colnames(cor.mat) <- gsub('abe8e_|bemax_','', colnames(cor.mat))
rownames(cor.mat) <- gsub('abe8e_|bemax_','', rownames(cor.mat))
#set ordering for diagonal mat
od =  hclust(dist(cor.mat))$order
cm = cor.mat[od, od]


sample.pal <- sort(rep(group.pal,2))
names(sample.pal) <- paste(names(sample.pal), rep(c('1','2'),time=length(sample.pal)/2), sep='_')
hm <- Heatmap(cm, 

        name='R log2(counts+1)',
        rect_gp = gpar(type = "none"), 
        cluster_rows = F, 
        cluster_columns = F,
        column_names_gp = gpar(fontsize=6),
        row_names_gp = gpar(fontsize=6),
        col=viridis(n=5, option='A'),
      	cell_fun = function(j, i, x, y, w, h, fill) {
	      	if(i >= j) {
			                grid.rect(x, y, w, h, gp = gpar(fill = fill, col = fill))
		                  grid.text(sprintf("%.2f", cm[i, j]), x, y, gp = gpar(fontsize = 6, col='white'))
		}
	})
hm

hm <- draw(hm, column_title='Abe8e sample correlations', column_title_gp=gpar(fontface='bold'))
BackupAsPDF(hm, 'abe8e.sampleCor.normCounts.heatmap')
```
```{r}
counts.wide <- dcast(counts.dt, paste(gene, sgRNA, sep='.')~paste(editor, treatment, timepoint, replicate, sep='_'), value.var = 'sgRNAnormCounts') %>% 
  as.matrix(rownames=1)

counts.wide <- log2(counts.wide +1)

cor.mat <- cor(counts.wide[, grepl('bemax', colnames(counts.wide))], use='pairwise.complete.obs', method='pearson')
colnames(cor.mat) <- gsub('abe8e_|bemax_','', colnames(cor.mat))
rownames(cor.mat) <- gsub('abe8e_|bemax_','', rownames(cor.mat))
#set ordering for diagonal mat
od =  hclust(dist(cor.mat))$order
cm = cor.mat[od, od]


sample.pal <- sort(rep(group.pal,2))
names(sample.pal) <- paste(names(sample.pal), rep(c('1','2'),time=length(sample.pal)/2), sep='_')

hm <- Heatmap(cm, 
        name='R log2(counts+1)',
        rect_gp = gpar(type = "none"), 
        cluster_rows = F, 
        cluster_columns = F,
        column_names_gp = gpar(fontsize=6),
        row_names_gp = gpar(fontsize=6),
        col=viridis(n=5, option='A'),
      	cell_fun = function(j, i, x, y, w, h, fill) {
	      	if(i >= j) {
			                grid.rect(x, y, w, h, gp = gpar(fill = fill, col = fill))
		                  grid.text(sprintf("%.2f", cm[i, j]), x, y, gp = gpar(fontsize = 6, col='white'))
		}
	})
hm

hm <- draw(hm, column_title='Bemax sample correlations', column_title_gp=gpar(fontface='bold'))
BackupAsPDF(hm, 'bemax.sampleCor.normCounts.heatmap')
```

After sample correlations, we want pairwise correlations of the sample replicates
Can then just show a select few to highlight sample agreement
Maybe also plot count distributions 

```{r}
library(tidymodels)
scatterplot.set <- UniqueFactorCombos(counts.dt$sample)
scatterplot.set < scatterplot.set[gsub('[.][12]$' ,'', Var1) == gsub('[.][12]$' ,'', Var2)]


apply(scatterplot.set, 1,  function(x){
  
  #get the editor
  editor <- str_extract(x[1], 'abe8e|bemax')
  
  dt <- dcast(counts.dt[sample %in% x,], paste(gene, sgRNA, sep='.')~sample, value.var = c('sgRNAnormCounts'))
  
  colnames(dt) <- gsub('bemax[.]|abe8e[.]', '', colnames(dt))
  oldNames <- copy(colnames(dt)) # create copy so DT not changed in place

  # print
  setnames(dt, new=c('guide', 'counts.x', 'counts.y'))
  
  r.score <- cor(dt$counts.x, dt$counts.y)
  
  g <- ggplot(dt, aes(x=log2(counts.x+1), y=log2(counts.y+1))) +
    annotate(geom='text', x=1, y=10, label = paste("r = ", round(r.score,2)), fontface = 'bold') +
    geom_hex(bins=50) +
    scale_fill_viridis(option='A', name='N') +
    ggtitle(paste(editor, 'sample correlations'), subtitle = "log2(normCounts+1)") +
    xlab(oldNames[2]) +
    ylab(oldNames[3]) +
    customTheme +
    coord_obs_pred()
  
  g
  BackupAsPDF(g, paste0('scatterplots/',editor,'.',oldNames[2], 'vs', oldNames[3],'.sgNormCounts.scatterplots'), dimensions = c(8,8))
})
```
look at a plot of the distributions 

```{r}
g <- ggplot(counts.dt[editor == 'bemax'], aes(x=sample, y=log2(sgRNAnormCounts+1), fill=group)) +
  geom_boxplot() +
  ggtitle(label=expression('Bemax sample distributions')) +
  labs(x='sample', y=expression(log[2]~'(normCounts+1)')) +
  scale_fill_manual(values=group.pal) +
  customTheme

BackupAsPDF(g, 'bemax.distributions.boxplot', dimensions=c(10,7))


g <- ggplot(counts.dt[editor == 'abe8e'], aes(x=sample, y=log2(sgRNAnormCounts+1), fill=group)) +
  geom_boxplot() +
  ggtitle(label=expression('Abe8e sample distributions')) +
  labs(x='sample', y=expression(log[2]~'(normCounts+1)')) +
  scale_fill_manual(values=group.pal) +
  customTheme

BackupAsPDF(g, 'abe8e.distributions.boxplot', dimensions=c(10,7))
```
Counts distributions in each of the samples?
Maybe overlay the histograms of the counts,faceted by condition
Makes no sense as normalized; leave as is

```{r}
ggplot(counts.dt[editor != 'plasmid'], aes(x=log2(sgRNAnormCounts+1), y=group, fill=group)) +
  geom_density_ridges2() +
  facet_grid(replicate~editor) +
  scale_fill_manual(values=group.pal) +
  customTheme
```
Look at PCA plot of the samples; generate seperately and then together

```{r}
pcaOut <- prcomp(t(counts.wide))


colInfo <- data.table(sample = colnames(counts.wide))
colInfo[, c('editor', 'treatment', 'timepoint', 'replicate') := tstrsplit(sample, '_', keep=c(1,2,3,4))]
colInfo[, group := paste(treatment, timepoint, sep='_')]

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)

pcaDT <- merge(pcaDT, colInfo, by.x = "rn", by.y = "sample", all.x = TRUE)

#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = timepoint, shape = editor)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  customTheme +
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_manual(values=timepoint.pal) +
  scale_shape_manual(values = 21:25) +
  scale_fill_manual(values=timepoint.pal) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, "normalizedCounts.editorCol.pca")
```
Look at distribution of control guides in this set

```{r}
sginfo.dt <- read_xlsx('./docs/PIK3CA pathway base editing.xlsx', sheet=1) %>%  
  as.data.table()
sginfo.dt <- sginfo.dt[, .(gene = toupper(Gene), sgRNA=`Sl No`, guide=Guide)]
sginfo.dt[grepl('SGINTERGENIC', gene), gene := 'Intergenic']

sginfo.dt[gene == 'Intergenic', sgRNA]
```

