---
title: "050225_DESeq2_rerun"
author: "Martin Gordon"
date: "2025-05-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Overview**
Assess if adding batch term to the DE model improves results; confirm we still recover our set of DE genes
Do we find more things? Is there a better aggreement of the annotations and the DE results? Investigate

For filtering, required at least 10 read in 75% of the samples

## packages 

```{r}
library(data.table)
library(ggplot2)
library(magrittr)
library(ggrepel)
library(stringr)
library(DESeq2)
library(ComplexHeatmap)
library(circlize)
library(viridis)
library(RColorBrewer)
library(hrbrthemes) # visually nice set of themes
library(patchwork)
library(showtext)
library(seqinr)
library(readxl)
library(ggridges)
library(RColorBrewer)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/enrichmentTestFunctions.R")
source ("../../utils/mg_utils/r_utils/IDmapping.R")
source ("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("../../utils/mg_utils/r_utils/CRISPR_Functions.R")


# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0,...){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}

customTheme <- theme_bw() +
  theme(panel.border = element_rect(color = "lightgrey", 
                                    fill = NA, 
                                    size = 1),
        axis.text.x = element_text(angle=90)
        )

#set one
col.pal <- getQualitativePalette(n=13)
#col.pal <- randomcoloR::distinctColorPalette(k=13)

# for heatmaps etc use this palette
redBlue <- RedBlueColPal()

# clinvar palette
clinvar.pal <- c(brewer.pal(8, "Blues"), 'grey')
```

Counts etc.
```{r}
counts.dt <- fread('~/Documents/projects/101224_RBabu_CRISPRe_PIK3CA/012825_requestedMeetingPlots_data/2025_01_29_sampleCounts.rawAndNormalized.dt')
counts.dt$group %>% unique()
grp.levels <- c('LibraryRepresentation_0', 'Alpelisib_22', 'Alpelisib_7', 'DMSO_22', 'DMSO_7', 'Paxalisib_22', 'Paxalisib_7')
counts.dt[, group := factor(paste(treatment, timepoint, sep='_'), levels=grp.levels)]

# also nned the PW comparisons and anno; this should be the set of everything we have
de.dt <- fread('~/Documents/projects/101224_RBabu_CRISPRe_PIK3CA/111824_plotsForKroganWeekly_data/2025_01_22_deseq.pwcomparisons.allAnno.csv')

gene.oi <- c('MTOR','PIK3CA', 'PTEN', 'AKT1')

de.dt[, c('numerator', 'denominator') := tstrsplit(contrast, '_vs_', keep=c(1,2))]
de.dt[, numerator := factor(numerator, levels=c('DMSO_7','DMSO_22', 'Paxalisib_7','Paxalisib_22', 'Alpelisib_7','Alpelisib_22'))]
de.dt[, numeratorTreatment := gsub('_[0-9]{1,2}', '', numerator)]
de.dt[, Gene := gene] # just to avoid reediting


selected.targets <- setDT(read_xlsx('./data/ronald_targets.xlsx', col_names =F))
setnames(selected.targets, new=c('guideName', 'sequence','status'))

de.dt <- merge(de.dt, selected.targets[,.(guideName,sequence, status)], by.x=c('guide'), by.y=c('sequence'), all.x=T)

# get the list of IDs to label
guidesOI <- de.dt[status != 'NA', unique(id)]
```

color palette stuff 
```{r}

brewer.pal(n=9, name = 'Blues')[4:8]
brewer.pal(n=9, name = 'PuRd')[4:8]
brewer.pal(n=9, name = 'Reds')[4:8]
brewer.pal(n=9, name = 'YlOrRd')[4:8]
brewer.pal(n=9, name = 'YlGn')[4:8]
brewer.pal(n=9, name = 'Greys')[4:8]
# blues

group.pal <- c("#BDBDBD", "#DF65B0", "#C994C7", "#6BAED6", "#9ECAE1","#FB6A4A", "#FC9272" )
#names(group.pal) <- unique(counts.dt$group)
names(group.pal) <- c('LibraryRepresentation_0', 'Alpelisib_22', 'Alpelisib_7', 'DMSO_22 DMSO_7', 'Paxalisib_22', 'Paxalisib_7')

# "#FC9272" "#FB6A4A" "#EF3B2C" "#CB181D" "#A50F15"
# "#9ECAE1" "#6BAED6" "#4292C6" "#2171B5" "#08519C"
# "#C994C7" "#DF65B0" "#E7298A" "#CE1256" "#980043"
# "#ADDD8E" "#78C679" "#41AB5D" "#238443" "#006837"
# "#BDBDBD" "#969696" "#737373" "#525252" "#252525"
# "#FEB24C" "#FD8D3C" "#FC4E2A" "#E31A1C" "#BD0026"


# timepoint and editor col pal
counts.dt$timepoint %>% unique()

group.pal <- c("#BDBDBD", "#DF65B0", "#C994C7", "#6BAED6", "#9ECAE1","#FB6A4A", "#FC9272" )
names(group.pal) <- unique(counts.dt$group)

timepoint.pal <- c("#ADDD8E" ,"#78C679", "#41AB5D")
names(timepoint.pal) <-c('0','7', '22')

editor.pal <- brewer.pal(n=9, name = 'Set3')[4:6]
names(editor.pal) <- c('abe8e', 'bemax', 'plasmid')

gene.col <- brewer.pal(n=6, name = 'Set2')
# set6 if need non targeting
gene.col <- c("#66C2A5" ,"#FC8D62","#FFD92F", "#8DA0CB") # "#A6D854" "#FFD92F" "#E78AC3""#8DA0CB"

```
Read in the DESeq2 ojects 
Size factors set to 0 as already normalized

```{r}
des.obj.list <- list(abe8e=readRDS('161024_PWComparisons_DESeq2_data/abe8e.dds.obj'),
                     bemax=readRDS('161024_PWComparisons_DESeq2_data/bemax.dds.obj'))
```

Quickly plot the PCA; does it make sense to do this? Are all the samples consistent naming?
Make big image so we can see the seperation

```{r}
summary.dt <- fread('./output/141024_noNormalization.countsummary.txt')

# drop the undetermined file.. may need to dig into this set later; see if we can recover some barcodes from here
# remove this undetermined file
summary.dt <- summary.dt[!grepl('Undetermined', File)]


summary.dt[, newLabel := gsub('/wd/data/fastq/bclconvert_defaults/|_R1_001.fastq.gz', '', File)]

summary.dt[, editor := ifelse(grepl('^ABE8[Ee]', newLabel), 'abe8e', 
                              ifelse(grepl('^BE4Max', newLabel), 'bemax', 'plasmid'))]


summary.dt[, treatment := stringr::str_extract(newLabel, 'Alpelisib|CTRL|Paxalisib|Librep|Lib-rep|librep')]

# lets tidy the treatment names
summary.dt[, treatment ] %>%  unique()
summary.dt[treatment == 'CTRL', treatment := 'control']
summary.dt[treatment %in% c("Librep","Lib-rep","librep"), treatment := 'librep']

# add a timepoint marke r
summary.dt[, timepoint := gsub('day', '', stringr::str_extract(newLabel, 'day[0-9]{1,2}'))]
summary.dt[is.na(timepoint), timepoint := '0']
summary.dt[, timepoint := factor(timepoint, levels=c('0', '7', '22'))]

summary.dt
summary.dt[, .(timepoint, newLabel)]
```

```{r}

```
```{r}
counts.dt <- fread('~/Documents/projects/101224_RBabu_CRISPRe_PIK3CA/012825_requestedMeetingPlots_data/2025_01_29_sampleCounts.rawAndNormalized.dt')
counts.dt[, group := paste(treatment, timepoint, sep='_')]
```


```{r}
counts.wide <- dcast(counts.dt, paste(gene, sgRNA, sep='.')~paste(editor, treatment, timepoint, replicate, sep='_'), value.var = 'sgRNAnormCounts') %>% 
  as.matrix(rownames=1)

counts.wide <- log2(counts.wide +1)
```

```{r}
pcaOut <- prcomp(t(counts.wide))

colInfo <- data.table(sample = colnames(counts.wide))
colInfo[, c('editor', 'treatment', 'timepoint', 'replicate') := tstrsplit(sample, '_', keep=c(1,2,3,4))]
colInfo
colInfo[, group := paste(treatment, timepoint, sep='_')]
colInfo[, group := factor(paste(treatment, timepoint, sep='_'), levels=grp.levels)]

group.pal <- c("#BDBDBD", "#DF65B0", "#C994C7", "#6BAED6", "#9ECAE1","#FB6A4A", "#FC9272" )
#names(group.pal) <- unique(counts.dt$group)
names(group.pal) <- c('LibraryRepresentation_0', 'Alpelisib_22', 'Alpelisib_7', 'DMSO_22', 'DMSO_7', 'Paxalisib_22', 'Paxalisib_7')



#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)

pcaDT <- merge(pcaDT, colInfo, by.x = "rn", by.y = "sample", all.x = TRUE)

#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = group, shape = editor)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3, max.overlaps = 30) +
  customTheme +
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_manual(values=timepoint.pal) +
  scale_shape_manual(values = 21:25) +
  scale_fill_manual(values=group.pal) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, "normalizedCounts.editorCol.pca", dimensions = c(12,12))
```


