---
title: "111824_AdditionalDESeq2Comparisons"
author: "Martin Gordon"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Additional PW comparisons using DESeq; 
Want to run with the plasmid as background, and also do comparisons of 

```{r}
library(data.table)
library(parallel)
library(ggplot2)
library(magrittr)
library(ComplexHeatmap)
library(viridis)
library(dendextend) # for dendogram modifications
library(DESeq2)
library(patchwork)
library(ggrepel)
library(tidymodels)
library(readxl)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/enrichmentTestFunctions.R")
source ("../../utils/mg_utils/r_utils/IDmapping.R")
source ("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")

source("../../utils/mg_utils/r_utils/CRISPR_Functions.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0,...){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```

In this script, we want to read in the DESeq2 objects and also the annotations
When rerunning the PWcomparisons, remove the guides in each set that dont have a target in the edit window; these are just adding noise to our output..

Read in the DESeq2 ojects 

```{r}
des.obj.list <- list(abe8e=readRDS('161024_PWComparisons_DESeq2_data/abe8e.dds.obj'),
                     bemax=readRDS('161024_PWComparisons_DESeq2_data/bemax.dds.obj'))
```

Recreate the metadata
Use both normalization methods and merge to original data.table
```{r}
summary.dt <- fread('~/Documents/projects/101224_RBabu_CRISPRe_PIK3CA/101524_MAGECKContrasts_firstPass_data/2024_10_15_sgCount.summary.csv')

metadata <- summary.dt[, .(sample=paste(editor,treatment,timepoint,rep, sep='_'), group=paste(editor,treatment,timepoint, sep='_'), editor, treatment, timepoint, rep)] %>% 
  as.data.table()

# tidy some of the conditions
metadata[treatment == 'control', treatment := 'DMSO']
metadata[treatment == 'librep', treatment := 'LibraryRep']
metadata[, treatment := factor(treatment, levels=c('LibraryRep', 'DMSO', 'Paxalisib', 'Alpelisib'))]
metadata[, timepoint := factor(timepoint, levels=c('0','7', '22'))]
  
# make a new condition group and lets use that for our comparisons
metadata[, condition := factor(paste0(treatment, '_',timepoint))]
metadata[, condition := relevel(condition, ref='LibraryRep_0')]

abe8e.meta <- metadata[grepl('abe8', sample),] %>% 
    as.data.frame(row.names = .$sample)
bemax.meta <- metadata[grepl('bemax', sample),] %>% 
  as.data.frame(row.names = .$sample)


meta.list <- list('abe8e'= abe8e.meta,
                 'bemax' = bemax.meta)

meta.list[[1]]$condition %>%  unique()
meta.list[[1]]$treatment %>%  unique
```


```{r}
mat.list <- lapply(names(mat.list), function(i){
  
  mat.list[[i]] <- mat.list[[i]][, rownames(meta.list[[i]])]
  return(mat.list[[i]])
  
})
names(mat.list) <- c('abe8e', 'bemax')

# all true
colnames(mat.list[[1]]) == rownames(meta.list[[1]])
colnames(mat.list[[2]]) == rownames(meta.list[[2]])

```


Read in the output from the annotation process to extract guides with targets

```{r}
# this is a selection of the guides with edits found in window; subset the datatables to these guides and other non-targeting set 
#does not include the non targeting set, so create a superset of those and the unique set in each of these
guides.dt <- fread('110724_annotateCRISPRguides_Functions_data/2024_11_15_guides.mutant.anno.collapsed.txt')

# subset to guides here and the non=targeting set 
good.abe8e <- guides.dt[editor == 'abe8e', unique(guide)]
good.bemax <- guides.dt[editor == 'bemax', unique(guide)]
```

```{r}
guide.table <- read_xlsx('./docs/PIK3CA pathway base editing.xlsx', sheet=1) %>%  
  as.data.table()

guide.table[, Gene := toupper(Gene)]
guide.table[Gene %like% "SGINTERGENIC", Gene := 'INTERGENIC']

# I think we only want to run comparisons on the different
sgControls <- guide.table[Gene %in% c('NON-TARGETING', 'INTERGENIC'), unique(`Sl No`)]

abe8e.set <- c(guide.table[Guide %in% c(good.abe8e), unique(`Sl No`)], sgControls)
bemax.set <- c(guide.table[Guide %in% c(good.bemax), unique(`Sl No`)], sgControls) 
```
Filter the DESeq2 objects to only keep the 'good' guides and non-targeting/intergenic set

```{r}
des.obj.list[['abe8e']] <- des.obj.list[['abe8e']][rownames(des.obj.list[['abe8e']]) %in% abe8e.set,]
des.obj.list[['bemax']] <- des.obj.list[['bemax']][rownames(des.obj.list[['bemax']]) %in% bemax.set,]


colData(des.obj.list[['abe8e']])
```

Rerun the contrasts; we also want to run the contrasts of treatment vs control at tp7 and tp22
```{r}
de.list  <- lapply(des.obj.list, function(x){
  
  dds <- DESeq(x)
  
  print(resultsNames(dds))
  contrasts.oi<- grep('Intercept', resultsNames(dds), invert=T, value = T)
  
  all.res <- lapply(contrasts.oi, function(n){
    
    res <- results(dds, name=n) %>% 
      as.data.table(., keep.rownames=T) %>% 
      .[, contrast := n]
  }) %>% rbindlist()
})


de.list[[1]]$contrast %>%  unique()
```
```{r}
colData(des.obj.list[['abe8e']] )

# define the model matrix
mod_mat <- model.matrix(design(des.obj.list[['abe8e']]), colData(des.obj.list[['abe8e']]))
mod_mat
# calculate coefficient vectors for each group
alp22 <- colMeans(mod_mat[des.obj.list[['abe8e']]$condition == "Alpelisib_22", ])
alp7 <- colMeans(mod_mat[des.obj.list[['abe8e']]$condition == "Alpelisib_7", ])
pax22 <- colMeans(mod_mat[des.obj.list[['abe8e']]$condition == "Paxalisib_22", ])
pax7 <- colMeans(mod_mat[des.obj.list[['abe8e']]$condition == "Paxalisib_7", ])
dmso22 <- colMeans(mod_mat[des.obj.list[['abe8e']]$condition == "DMSO_22", ])
dmso7 <- colMeans(mod_mat[des.obj.list[['abe8e']]$condition == "DMSO_7", ])
 

# sanity check this matches our calculations running automatically...

de.additionalcontrasts.list <- lapply(des.obj.list, function(x){
  
  dds <- DESeq(x)
  message('Building design matrix..')
  
  mod_mat <- model.matrix(design(x), colData(x))

  
  alp22 <- colMeans(mod_mat[x$condition == "Alpelisib_22", ])
  alp7 <- colMeans(mod_mat[x$condition == "Alpelisib_7", ])
  pax22 <- colMeans(mod_mat[x$condition == "Paxalisib_22", ])
  pax7 <- colMeans(mod_mat[x$condition == "Paxalisib_7", ])
  dmso22 <- colMeans(mod_mat[x$condition == "DMSO_22", ])
  dmso7 <- colMeans(mod_mat[x$condition == "DMSO_7", ])
  d0 <-  colMeans(mod_mat[x$condition == "LibraryRep_0", ])

  # pulling out individual contrasts
  alp22_d0 <- as.data.table(results(dds, contrast = alp22 - d0), keep.rownames = T)[, contrast := 'Alpelisib_22_vs_LibraryRep_0']
  alp7_d0 <- as.data.table(results(dds, contrast = alp7 - d0), keep.rownames = T)[, contrast := 'Alpelisib_7_vs_LibraryRep_0']
  pax22_d0 <- as.data.table(results(dds, contrast = pax22 - d0), keep.rownames = T)[, contrast := 'Paxalisib_22_vs_LibraryRep_0']
  pax7_d0 <- as.data.table(results(dds, contrast = pax7 - d0), keep.rownames = T)[, contrast := 'Paxalisib_7_vs_LibraryRep_0']
  dmso22_d0 <- as.data.table(results(dds, contrast = dmso22 - d0), keep.rownames = T)[, contrast := 'DMSO_22_vs_LibraryRep_0']
  dmso7_d0 <- as.data.table(results(dds, contrast = dmso7 - d0), keep.rownames = T)[, contrast := 'DMSO_7_vs_LibraryRep_0']

  # now get the individual
  alp22_dmso22 <- as.data.table(results(dds, contrast = alp22 - dmso22), keep.rownames = T)[, contrast := 'Alpelisib_22_vs_DMSO_22']
  alp7_dmso7 <- as.data.table(results(dds, contrast = alp7 - dmso7), keep.rownames = T)[, contrast := 'Alpelisib_7_vs_DMSO_7']
  pax22_dmso22 <- as.data.table(results(dds, contrast = pax22 - dmso22), keep.rownames = T)[, contrast := 'Paxalisib_22_vs_DMSO_22']
  pax7_dmso7 <- as.data.table(results(dds, contrast = pax7 - dmso7), keep.rownames = T)[, contrast := 'Paxalisib_7_vs_DMSO_7']
  alp22_pax22 <- as.data.table(results(dds, contrast = alp22 - pax22), keep.rownames = T)[, contrast := 'Alpelisib_22_vs_Paxalisib_22']
  alp7_pax7 <- as.data.table(results(dds, contrast = alp7 - pax7), keep.rownames = T)[, contrast := 'Alpelisib_7_vs_Paxalisib_7']

  all.contrasts <- rbind(alp22_d0, alp7_d0, pax22_d0, pax7_d0, dmso22_d0, dmso7_d0,
                         alp22_dmso22, alp7_dmso7, pax22_dmso22, pax7_dmso7, alp22_pax22, alp7_pax7)  
  

  return(all.contrasts)
})

```
Save the output of the contrasts and lets combine the two datatables and regenerate the volcanoplots


```{r}
de.newContrasts.dt <- rbindlist(de.additionalcontrasts.list, idcol = 'editor')
de.newContrasts.dt[, adj.pvalue := p.adjust(pvalue, method = 'BH'), by=.(editor,contrast)]

fwrite(de.newContrasts.dt, ScriptAndDatedFileName('DESeq2.additionalPWcontrasts.txt'))
```

