---
title: "112224_readINPDBfiles"
author: "Martin Gordon"
date: "2024-11-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Ask BP about PDB structures; what is the best structure to use?
Using alphafold models for comparability to PPI and also as no MTOR PDB available


```{r}
library(data.table)
library(ggplot2)
library(bio3d)
library(magrittr)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/enrichmentTestFunctions.R")
source ("../../utils/mg_utils/r_utils/IDmapping.R")
source ("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")

source("../../utils/mg_utils/r_utils/CRISPR_Functions.R")
```

Reading in the alphafold pdb list 
```{r}
fpaths <- dir('./data', pattern = "*.pdb", full.names = T)

pdb.list <- lapply(fpaths, function(x){
  bio3d::read.pdb(x)
  })

names(pdb.list) <- c('AKT1', 'PIK3CA', 'MTOR', 'PTEN')

af.pdb.list <- pdb.list

```


Read in the DESeq2 results
```{r}
de.dt <- fread('~/Documents/projects/101224_RBabu_CRISPRe_PIK3CA/112124_mutantHotspots_data/2024_11_22_deseq2.pwcomparisons.wAnnotations.csv')
```
Map the log2FC to the structures for each contrast 
```{r}
de.dt[!gene %in% c('NON-TARGETING','INTERGENIC'),.N, by=.(gene, peptide_position, contrast)][order(-N)]

# check whyy some are returning NA and others nothing
de.dt[!gene %in% c('NON-TARGETING','INTERGENIC') & peptide_position == '']
de.dt[!gene %in% c('NON-TARGETING','INTERGENIC') & grepl('NA', peptide_position)]
```
Good treshold for log2FC? 
Left align the peptide positions (take first residue), calculate the geometric mean and take the average geometric mean of each site
```{r}
de.dt[, gm_lfc.pval := (-log10(pvalue) * abs(log2FoldChange))/2]
```

# take the average of the geometric mean per site contrast
```{r}
de.dt[, average_gm := mean(gm_lfc.pval, na.rm=T), by=.(contrast, gene, first_peptide_idx)]

de.dt[, any_sig := ifelse(any(sig != 'not'), 'sig', 'no'), by=.(contrast, gene, first_peptide_idx)]

lapply(unique(de.dt$contrast), function(x){
  
  subdt  <-de.dt[contrast == x,]
  ggplot(subdt, aes(x=first_peptide_idx, y=average_gm, col=any_sig)) +
    geom_point() +
    facet_wrap(~gene, scales='free') +
    coord_cartesian(ylim=c(0,10))
    
})
```

Now create a data.table that runs from the length of 1 : len(gene) per gene/contrast

```{r}
gene.idx <- list(MTOR = data.table(peptide_position=seq(1,2549, 1)),
     PIK3CA = data.table(peptide_position=seq(1,1068, 1)),
     AKT1 = data.table(peptide_positon = seq(1,480, 1)),
     PTEN = data.table(peptide_position=seq(1,403, 1))) %>% 
  rbindlist(idcol='gene')

# create a list of contrast lfcs with the position

LFCs.list <- lapply(unique(de.dt$contrast), function(x){
  
  subdt <- de.dt[!gene %in% c('NON-TARGETING','INTERGENIC') & contrast == x & !is.na(first_peptide_idx),] %>% 
    .[, .(gene, contrast=x, peptide_idx=first_peptide_idx, residue_GM=average_gm ,any_sig)] %>% 
      .[order(peptide_idx)] %>% 
      unique()
    
    #merge with the index info 
   merge(subdt, gene.idx, by.x=c('gene', 'peptide_idx'), by.y=c('gene', 'peptide_position'), all.x=T, all.y=T)
})

```

So now we merge this info to the PDB; want to create mutliple PDBs 
bioseq only writes out PDB files so write out one at a time

```{r}
names(LFCs.list) <- unique(de.dt$contrast)


de.dt[average_gm == 0,]
```


# overwrite the plDDT score for each of the structures with the geom mean values for each site. O means no evidence of change (not distinguishing no score vs no evidence)

# just look at the proteins for specific contrasts that may have interesting FCs 
```{r}
lapply(names(LFCs.list), function(x){
  
  message('Working on contrast ', x)
  LFC.subset <- LFCs.list[[x]]

  message('Setting missing resiudes to 0... ')
  LFC.subset[is.na(residue_GM), residue_GM := 0]
  
  lapply(names(af.pdb.list), function(y){
    
      gene.pdb <- af.pdb.list[[y]]
      # get the pdb structure as a dt
      gene.pdb.dt <- as.data.table(gene.pdb$atom)
      # get the correct cols 
      cols.oi <- colnames(gene.pdb.dt)
      
      # safer merge as only considers overlpaping resiudes
      gene.pdb.dt <- merge(gene.pdb.dt, LFC.subset[gene == y, .(resno=peptide_idx, residue_GM)], by='resno', all.x=T)
      gene.pdb.dt[, `:=`(b=residue_GM, old.b = b)]
      gene.pdb.dt <- gene.pdb.dt[, ..cols.oi]
      
      # save the new b-factor scores
      gene.pdb$atom <- gene.pdb.dt
      
      # write out the modified pdb
      bio3d::write.pdb(gene.pdb, file=paste0('/Users/martingordon/Documents/projects/101224_RBabu_CRISPRe_PIK3CA/output/pdbs/', interaction(y,x), '.geommean_lfcpval.scores.pdb'))
      
  })
})
```

How do the LFCs overlap with the contact regions?

```{r}

```



*some work using the bio3d package*

Use `attributes` to pull out the attributes of the PDB
Focusing on working with the carbon atom residues
```{r}
# list all attirbutes
lapply(af.pdb.list, attributes)

# pull out individual attributes
lapply(af.pdb.list, function(x) print(x$atom))

# this is just a binary checking if the atom is a Ca carbon alpha atom

lapply(af.pdb.list, function(x) head(x$calpha, n=10))


#plotting the plddt score along the length of the residues
# here we are only plotting the carbon atoms (why is this??)
lapply(af.pdb.list, function(x) plot.bio3d(x$atom, type='l', ylab="B-factor"))
```
Can align PDBs
Strcutre based superimposition only here:
Need MUSCLE installed for this to work...
```{r}
# Align and superpose two or more structures
pdbs <- pdbaln(c("4q21", "521p"), fit=TRUE)

attributes(pdbs)
```


