---
title: "AFmultimer_PIK3CA"
author: "Martin Gordon"
date: "2024-11-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(data.table)
library(magrittr)
library(ggplot2)


source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/UniprotIDMapping.R")
source ("../../utils/mg_utils/r_utils/IDmapping.R")
source ("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
```


```{r}
af.dt <- fread('./output/AFmultimer_runs/allScores.csv')
setnames(af.dt, new=c('filepath', 'ptm', 'iptm'))
```
plot of average iPTM score per bait prey combo
```{r}
af.dt[, interactors := stringr::str_extract(filepath, "[A-Z0-9]+__[A-Z0-9]+")]
af.dt[, c('Bait', 'Prey') := tstrsplit(interactors, '__', keep=c(1,2))]
af.dt[,gene := 'PIK3CA']

af.dt[, `:=`(BaitGene = multiUniprots2multiGenes(Bait, species='HUMAN'),
             PreyGene = multiUniprots2multiGenes(Prey, species='HUMAN')
             )]

# a couple of extra runs..no harm
af.dt[interactors == 'P42336__P08582',] %>% unique()

af.dt[, geneInteractors := paste0(BaitGene, '__', PreyGene)]
af.dt[, avg_iptm := mean(iptm, na.rm=T), by=interactors]

af.dt[, sig := ifelse(iptm >= 0.6, 'yes', 'no')]
```
quick plot of the alphafold2 iptm distributions 

```{r}
g <- ggplot(af.dt, aes(x=reorder(PreyGene, (avg_iptm)),  y=iptm, color=sig)) +
  geom_point() +
  geom_hline(yintercept=0.6
             , color='red', linetype=2) +
  ggtitle('PIK3CA interactors (Swaney et al, 2022)') +
  scale_color_manual(values=c('no'='black', 'yes'='steelblue4')) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, size=5))

g

BackupAsPDF(g, 'PIK3CA.afmultimer.interactor.scores.csv', dimensions=c(8,6))


paste(unique(af.dt[iptm > 0.6, .(PreyGene, Prey)]$Prey), collapse='|')
```

Read in the pten data and plot 

```{r}
pten.scores <-  fread('output/AFmultimer_runs/pten.highScoringModels.allScores.csv')
setnames(pten.scores, new=c('filepath', 'ptm', 'iptm'))
pten.scores[, gene := 'PTEN']


pten.scores[, interactors := stringr::str_extract(filepath, "[A-Z0-9]+[-]+[0-9]+__[A-Z0-9]+")]
pten.scores[, c('Bait', 'Prey') := tstrsplit(interactors, '__', keep=c(1,2))]

pten.scores[, BaitOriginal := Bait]
pten.scores[, Bait := gsub('[-][0-9]{1}','', Bait)]


pten.scores[, `:=`(BaitGene = multiUniprots2multiGenes(Bait, species='HUMAN'),
                   PreyGene = multiUniprots2multiGenes(Prey, species='HUMAN')
             )]



pten.scores[, geneInteractors := paste0(BaitGene, '__', PreyGene)]
pten.scores[, avg_iptm := mean(iptm, na.rm=T), by=interactors]

pten.scores[, sig := ifelse(iptm >= 0.6, 'yes', 'no')]
```

```{r}
g <- ggplot(pten.scores, aes(x=reorder(PreyGene, (avg_iptm)),  y=iptm, color=sig)) +
  geom_point() +
  geom_hline(yintercept=0.6
             , color='red', linetype=2) +
  ggtitle('PTEN interactors') +
  scale_color_manual(values=c('no'='black', 'yes'='steelblue4')) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, size=5))

g
BackupAsPDF(g, 'PTEN.afmultimer.interactor.scores.csv', dimensions=c(8,6))
```
Weird that for USP7 two o fthe scores diverge greatly..
Model 4 & 5 scored very well (rank 0 and rank1 I presume)
```{r}
pten.scores[PreyGene %in% c('NEDD4', 'USP7', 'WWP2')] %>% 
  .[order(-iptm)]
```
**Contact PAE plots**
PTEN & PIK3CA contact plots
```{r}
pten.pae <- fread('./output/AFmultimer_runs/pten.paeContacts.csv')
pten.pae[, gene := 'PTEN']

# rsead in PIK3CA contact plot
pik3ca.pae <- fread('output/AFmultimer_runs/pae_contacts/pik3ca.highScoringModels.PAEcontacts.csv')
pik3ca.pae[, gene := 'PIK3CA']

# adjust col names dfor merging
setnames(pten.pae, old="./112224_ptenPrey/P60484-1__A2Q0Z1/relaxed_model_1_multimer_v3_pred_0.contacts.csv:chainRow", new="model")
setnames(pik3ca.pae, old="P42336__A1L390/relaxed_model_1_multimer_v3_pred_0.contacts.csv:chainRow", new="model")

pae.dt <- rbind(pten.pae, pik3ca.pae)
```

# identify the interacting pairs and filter for the residues that are in the interface. Find out if any of the high quality edits reside in the interface

Ge thte high quality interactors
```{r}
pten.scores[iptm > 0.6, .SD[which.max(iptm)], by=.(geneInteractors)] %>% 
  .[filepath]

# get the model id
pten.scores[, modelID := stringr::str_extract(filepath, 'model_[12345]{1}_multimer_v3_pred_0')]
af.dt[, modelID := stringr::str_extract(filepath, 'model_[12345]{1}_multimer_v3_pred_0')]

pten.interactors <- pten.scores[iptm > 0.6, .SD[which.max(iptm)], by=interactors]
pik3ca.interactors <- af.dt[iptm > 0.6, .SD[which.max(iptm)], by=interactors]
```

Tidy the pae table and pull out the contacts
```{r}
pae.dt[, interactors := stringr::str_extract(model, "[A-Z0-9]+[-]+[0-9]+__[A-Z0-9]+")]
# fix the naming convention for one of the proteins
pae.dt[is.na(interactors), interactors := stringr::str_extract(model, "[A-Z0-9]+__[A-Z0-9]+")]

pae.dt[, c('Bait', 'Prey') := tstrsplit(interactors, '__', keep=c(1,2))]
pae.dt[, BaitOriginal := Bait]
pae.dt[, Bait := gsub('[-][0-9]{1}','', Bait)]

pae.dt[, `:=`(BaitGene = multiUniprots2multiGenes(Bait, species='HUMAN'),
                   PreyGene = multiUniprots2multiGenes(Prey, species='HUMAN')
             )]

pae.dt[, geneInteractors := paste0(BaitGene, '__', PreyGene)]
```

```{r}
pae.dt[, modelID := stringr::str_extract(model, 'model_[12345]{1}_multimer_v3_pred_0')]

hc.contact.dt[BaitGene == 'PIK3CA',]
hc.contact.dt<- pae.dt[interactors %in%  c(pik3ca.interactors$interactors, pten.interactors$interactors) & modelID %in% c(pik3ca.interactors$modelID, pten.interactors$modelID)]

# we want the residue no Row; assuming here the row is the 1st member of the chain (A which is our bait)
hc.contact.dt <- hc.contact.dt[, .(model, interactors, geneInteractors, Bait, BaitGene, PreyGene, Prey, BaitResidue=resnoRow, PreyResidue=resnoCol, pae)][BaitResidue != 'resnoRow',]

af.dt[iptm > 0.6, .SD[which.max(iptm)], by=geneInteractors] %>% 
  .[, .(geneInteractors, filepath)]
```

*Share the table with Ronald*

Read in the DE results and identify if any of the high quality edits are in
```{r}
de.dt <-  fread('112124_mutantHotspots_data/2024_11_22_deseq2.pwcomparisons.wAnnotations.csv')

# for PTEN, something like 22 guides overlap these contact interfaces and are significant
de.dt[gene == 'PTEN' & first_peptide_idx %in% hc.contact.dt[BaitGene == 'PTEN', unique(BaitResidue)] & sig != 'not' & grepl('22_vs', contrast), ][order(first_peptide_idx),]
# get the contacts so we can color these sites in chimeraX
hc.contact.dt[BaitGene == 'PTEN' & unique(BaitResidue) %in% de.dt[gene == 'PTEN' & sig != 'not' & grepl('22_vs_', contrast), unique(first_peptide_idx)],]
hc.contact.dt
# show one of the top scoing ppi models with the PTEN bait sites highlighted
de.dt[gene == 'PIK3CA' & first_peptide_idx %in% hc.contact.dt[BaitGene == 'PIK3CA', unique(BaitResidue)] & sig != 'not' & grepl('22_vs_', contrast), ]
```
Look at the individual ppi
```{r}
hc.contact.dt[BaitGene == 'PTEN' &  PreyGene == 'LDHA' & model == './112224_ptenPrey/P60484-1__P00338/relaxed_model_1_multimer_v3_pred_0.contacts.csv:A', unique(BaitResidue)]

# get the contacts so we can color these sites in chimeraX
pten.usp7.residues <- hc.contact.dt[BaitGene == 'PTEN' &  PreyGene == 'USP7' & model =='./112224_ptenPrey/P60484-1__Q93009/relaxed_model_4_multimer_v3_pred_0.contacts.csv:A',]$BaitResidue
pten.ldha.residues <- hc.contact.dt[BaitGene == 'PTEN' &  PreyGene == 'LDHA' & model == './112224_ptenPrey/P60484-1__P00338/relaxed_model_1_multimer_v3_pred_0.contacts.csv:A', unique(BaitResidue)]
pten.ARF4.residues <- hc.contact.dt[BaitGene == 'PTEN' &  PreyGene == 'ARF4' & model == './112224_ptenPrey/P60484-1__P18085/relaxed_model_2_multimer_v3_pred_0.contacts.csv:A', unique(BaitResidue)]
pten.GSTM2.residues<- hc.contact.dt[BaitGene == 'PTEN' &  PreyGene == 'GSTM2' & model == './112224_ptenPrey/P60484-1__P28161/relaxed_model_3_multimer_v3_pred_0.contacts.csv:A', unique(BaitResidue)]
pten.GTF3C1.residues<- hc.contact.dt[BaitGene == 'PTEN' &  PreyGene == 'GTF3C1' & model == './112224_ptenPrey/P60484-1__Q12789/relaxed_model_4_multimer_v3_pred_0.contacts.csv:A', unique(BaitResidue)]

#definitely something here; consecutive sites and all strongly inhibited
de.dt[gene =='PTEN' & sig != 'not' & first_peptide_idx %in% pten.GTF3C1.residues, unique(first_peptide_idx)]

# look at the pik3ca interactors and see what we consider a good score
pik3ca.pik3r2.residues <- pae.dt[BaitGene == 'PIK3CA' &  PreyGene == 'PIK3R2' & model =='P42336__O00459/relaxed_model_1_multimer_v3_pred_0.contacts.csv:A', unique(resnoRow)]
pik3ca.pik3r1.residues <- pae.dt[BaitGene == 'PIK3CA' &  PreyGene == 'PIK3R1' & model == 'P42336__P27986/relaxed_model_1_multimer_v3_pred_0.contacts.csv:A', unique(resnoRow)]
pik3ca.pik3r3.residues <- pae.dt[BaitGene == 'PIK3CA' &  PreyGene == 'PIK3R3' & model == 'P42336__Q92569/relaxed_model_1_multimer_v3_pred_0.contacts.csv:A', unique(resnoRow)]
pik3ca.sec16a.residues <- pae.dt[BaitGene == 'PIK3CA' &  PreyGene == 'SEC16A' & model == 'P42336__O15027/relaxed_model_3_multimer_v3_pred_0.contacts.csv:A', unique(resnoRow)]
pik3ca.ext2.residues <- pae.dt[BaitGene == 'PIK3CA' &  PreyGene == 'EXT2' & model == 'P42336__Q93063/relaxed_model_4_multimer_v3_pred_0.contacts.csv:A', unique(resnoRow)]
pik3ca.my06.residues <- pae.dt[BaitGene == 'PIK3CA' &  PreyGene == 'MYO6' & model == 'P42336__Q9UM54/relaxed_model_5_multimer_v3_pred_0.contacts.csv:A', unique(resnoRow)]
```

```{r}
af.dt[iptm > 0.6, .SD[which.max(iptm)], by=geneInteractors]

hc.contact.dt[BaitGene == 'PIK3CA',]


pik3ca.pik3r2.residues
```



Look at the PIK3CA interaction; smaller set as not many interactions found model very welll...
```{r}

```

