---
title: "042123_Danielle_S"
author: "Martin Gordon"
date: "2023-04-21"
output: 
  html_document:
    fig_width: 12
    fig_height: 8
    theme: cerulean
    highlight: tango
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Goal: Comparison of 4 MS protocols
Assess:
how many peptides and proteins per sample? #count
reproducibility between replicates of the same prep method? #clustering 
Overlap in peptide/protein IDs between methods #seems like a venn diagram or umap plot
When different prep methods identify the same protein, is there any systematic difference in the abundance of proteins? because if so, I want to pick the method that give us higher abundances

scatterplot at protein level

```{bash eval=FALSE }

git clone https://github.com/kroganlab/bp_utils.git ./scripts

```
```{r load-packages}
library(artMS) #for initial analysis
library(org.Hs.eg.db)
library(viridis)

library(ggpubr)
library(ggplot2)
library(MSstats)
library(circlize)
library(data.table)

library(dtplyr) #data.table backend for dplyr
library(data.table)
library(tidyverse) #should load subsets of pkg..


library(clusterProfiler)
```

```{r mg-util-functions}

#function for converting enrichment results to dec for plotting
fract_to_dec <- function(x) {
  x <- as.numeric(x)
  x <- x[1] / x[2] 
  return(round(x,3))
}

# only works on R df
extract_sig_prots <- function(x, LFC=0.58, pval=0.05) {
  x %>% 
    filter(abs(log2FC) >= LFC & adj.pvalue <= pval) %>% 
    return(select(UNiProt, log2FC, SE, pvalue, adj.pvalue))
}


enrich_bar_fun <- function(x,y){
  
  ego_df <- as.data.frame(x) 
  ego_df$generatio_dec <- map_dbl(strsplit(ego_df$GeneRatio, "/"), ~ fract_to_dec(.) ) #return numeric vector w gene ratio
  
  ggplot(ego_df, aes(x = -log(p.adjust, base=10), 
                          y = fct_reorder(Description, -log(p.adjust, base=10)))) + 
    geom_col(aes(fill = generatio_dec)) +
    scale_fill_viridis(name = 'gene_ratio')  +
    scale_y_discrete(position= 'right') +
    scale_x_continuous(expand = c(0,0.1)) +
    xlab('-log10(P)') +
    ylab(NULL) +
    ggtitle(paste0("GO BP enrichment ", y )) +
    theme_bw() +
     theme(
     axis.title.x = element_text(size=10),
     legend.position = "bottom",
     )
}



```

```{r util-functions}


today <- function(){
  format(Sys.time(), "%Y_%m_%d")
}
DateFileName <- function(x){
  name <-   paste0(today(), "_", x)
  print (name)
  return (name)
}

ScriptNamedDir <- function(scriptName = NULL){
  if(is.null(scriptName))
    scriptName <- rstudioapi::getActiveDocumentContext()$path
  if (is.null (scriptName) || scriptName == "")
    stop("No script name found -- you may need to save this file first")
  outDir <- gsub(".R(md)?$", "_data", scriptName, ignore.case = TRUE)
  stopifnot( outDir != scriptName)
  if (!dir.exists(outDir)){
    message ("Creating directory associated with ", scriptName,", at ", outDir)
    dir.create(outDir)
  }
  return(outDir)
}

ScriptAndDatedFileName <- function(x, scriptName = NULL){
  dir <- ScriptNamedDir(scriptName)
  fileName <- DateFileName(x)
  path <- file.path(dir, fileName)
  print (path)
  return (path)
}

GetLatestScriptFile <- function(x, scriptName=NULL){
  stopifnot (length(x) == 1)
  dir <- ScriptNamedDir(scriptName)
  filePattern <- paste0("^\\d{4}_\\d{2}_\\d{2}_", x, "$", collapse = "")
  filesFound <- list.files(dir, filePattern)
  stopifnot (length(filesFound) > 0)
  if (length(filesFound) > 1){
    message ("Multiple files  with matching names found.  Using the last one")
    print (filesFound)
  }
  return (file.path(dir, tail(filesFound, 1)))
} 

PDFBackupFileName <- function(prefix = "", subDir = ""){
  scriptDir <- ScriptNamedDir()
  imageDir <- file.path(scriptDir, "pdfs", subDir)
  if (!dir.exists(imageDir)) dir.create(imageDir, recursive = TRUE)
  now <- format(Sys.time(),  "%Y_%m_%d__%H_%M__%S")
  counter <- 0
  path <- file.path (imageDir, sprintf("%s%s.%02d.pdf", prefix, now, counter))
  while (file.exists(path)){
    counter <- counter + 1
    path <- file.path (imageDir, sprintf("%s%s.%02d.pdf", prefix, now, counter))
  }
  return (path)
}

BackupAsPDF <- function(graphics, prefix = "", subDir = "", dimensions = NULL){
  path <- PDFBackupFileName(prefix, subDir)
  if (is.null(dimensions))
  dimensions <- dev.size(units = "in")

  print (sprintf("Writing image to  %s", path))
  cairo_pdf(path, width = dimensions[1], height = dimensions[2])
  
  # handle functions, my enrichment heatmaps that are part of a list
  if ("function" %in% class(graphics)){
    graphics()
    g <- "finished" # something to print to console instead of graphics to device 
  }else if (! ("ggplot" %in% class(graphics) | "grob" %in% class(graphics) | "Heatmap" %in% class(graphics) | "HeatmapList" %in% class(graphics))){
    g <- graphics$hmList    
  }  else{
    g <- graphics
  }
  print (g)
  
  dev.off()
  return (graphics)
}





WriteSessionInfo <- function(path = NULL){
  if (is.null(path))
    path <- ScriptAndDatedFileName("SessionInfo.txt")
  si <- devtools::session_info()
  fileOut <- file(path, open = "wt")
  writeLines("─ Session info ─────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────",
             con = fileOut)
  write.table(data.frame( value = unlist(si[[1]])), fileOut)
  #writeLines(capture.output(data.frame(value = unlist(si[[1]]))), con = fileOut) # for some mysterious reason, this always goes to the notebook output when run in a notebook. type = "message" is no help...[shrug]
  writeLines(capture.output(data.table(setting = names(si[[1]]), value = unlist(si[[1]]))), con = fileOut) # this too. Maybe an RStudio version issue
  writeLines("─ Packages ─────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────────",
             con  = fileOut)
  #write.table(si[[2]], fileOut)
  writeLines(capture.output(si[[2]]), con = fileOut)
  close(fileOut)
}

WriteInstalledPackages <- function (path = NULL){
  if (is.null(path))
    path <- ScriptAndDatedFileName("Installed.Packages.csv")
  package.mat <- installed.packages()
  fwrite (as.data.table(package.mat, keep.rownames = TRUE), path)
}

WriteSessionInfo()
WriteInstalledPackages()

```

# ArtMS Analysis

Have a .csv file output from Spectronaut software
Convert to evidence.txt file to use with ArtMS software

Ben has a script to perform this
```{r}
source('./scripts/spectronautFile2ArtMS.R')
```

# Create ArtMS input
Adjust config file for extended ArtMS QC plots
```{r artms, eval = FALSE}
#get our input file
input <- './data/MSstats.csv'
ms_counts <- read_delim(input)


# define some settings that will over-write defaults in artMS config file:
cf<- list()
cf$output_extras$annotate$enabled <- as.integer(0) # turn off any attempt to annotate proteins (only works with HUMAN)
cf$qc$extended <- as.integer(1)  # extended QC plots

globalInput <- spectronautFile2ArtMS(input,   # a path to a spectronaut "evidence" file
                                     outFilePrefix = "artMS/spec_to_mss_input",   # the prefix for file names of generated 
                                     artmsConfig = cf,  # the config object we created above with different-from-default 
                                     # contrastPatterns (below) is a vector of contrasts and/or regular expressions that 
                                     contrastPatterns = c("Urea-UreaSon",
                                                          "Urea-TFA",
                                                          "Urea-TFASon",
                                                          "UreaSon-TFA",
                                                          "UreaSon-TFASon",
                                                          "TFA-TFASon"
                                                          ))

```

Ran extended quality control pipeline to assess results
```{r, eval = FALSE}
# its also possible to just run QC instead:
artmsQualityControlEvidenceExtended(globalInput$config_data$files$evidence, globalInput$config_data$files$keys)
```


```{r artms-filterna, eval = FALSE}
# filter na values prior to ArtMS input
artms_in<- ms_counts %>% 
  filter( ! is.na(Intensity)) %>% 
  write.csv('/Users/martingordon/Documents/projects/041223_Danielle_Swaney/data/MSstats_rmna.csv')


input_rmna <- './data/MSstats_rmna.csv'


globalInput <- spectronautFile2ArtMS(input_rmna,   # a path to a spectronaut "evidence" file
                                     outFilePrefix = "artMS/042423_ds_qccheck_rmna",   # the prefix for file names of generated 
                                     artmsConfig = cf,  # the config object we created above with different-from-default 
                                     # contrastPatterns (below) is a vector of contrasts and/or regular expressions that 
                                     contrastPatterns = c("Urea-UreaSon",
                                                          "Urea-TFA",
                                                          "Urea-TFASon",
                                                          "UreaSon-TFA",
                                                          "UreaSon-TFASon",
                                                          "TFA-TFASon"
                                                          ))
```


#Review Spectronaut to ArtMS output.

```{r sanity-check, eval = FALSE}
evidence.txt_f <- fread('./artMS/042423_ds_qccheck_rmna/evidence.txt')
# sanity check
sum(is.na(evidence.txt_f$Intensity))
```

Standard global abundance assessment using ArtMS with extended QC section.

Rerun Extended QC pipeline to assess data quality

```{r run-extended-narm, eval = FALSE}
# its also possible to just run QC instead:
artmsQualityControlEvidenceExtended(globalInput$config_data$files$evidence, globalInput$config_data$files$keys)
```

Run Differential Expression Analysis using ArtMS
```{r}
artmsQuantification(yaml_config_file='./artMS/042423_ds_qccheck_rmna/config.yaml')
```


Standard Enrichment analysis using ArtMS.. not working.. seems to be an error caused when trying to impute values possibly due to lack of observations

```{r}
#setwd as output for the ArtMS quantification run
setwd('./artMS/042423_ds_qccheck_rmna/output')



artmsAnalysisQuantifications(log2fc_file = "results.txt",
                              modelqc_file = "results_ModelQC.txt",
                              species = "human",
                              mnbr = 2,
                              output_dir = "AnalysisQuantifications")

setwd('/Users/martingordon/Documents/projects/041223_Danielle_Swaney')
```


## ArtMS enrichment Analysis

Pull out `results.txt` file and process 

```{r}
artms_df <- read_delim('/Users/martingordon/Documents/projects/041223_Danielle_Swaney/artMS/042423_ds_qccheck_rmna/output/results.txt')

artms_results_list <- artms_df %>%
  mutate(Protein = as.character(Protein)) %>% #ensure Protein name is character
  mutate(UNiProt = str_split_i(Protein, "\\|",2)) %>% #recover uniprot_ids
  group_by(Label) %>% 
  group_split() %>% 
  setNames(unique(artms_df$Label))
```

```{r artms-filter}
#filter rows with 'issues'
#artms_results_list <- purrr::map(artms_results_list, ~filter(.x, is.na(issue)))
```



```{r artms-get-sig-genes}
# sig_gene_list
artms_sig_prots_list <- artms_results_list %>% 
  map( ~ extract_sig_prots(.x, 0.58, 0.05))
```

```{r}

# first need to drop the empty elements where no sig differential expression
artms_sig_prots_list[['Urea-UreaSon']] <- NULL
#artms_sig_prots_list[['UreaSon-TFASon']] <- NULL
```


```{r split-sig-genes}
artms_up_sig <- artms_sig_prots_list %>% 
  map( ~{
    filter(.x, log2FC > 0) 
  })

artms_down_sig <- artms_sig_prots_list %>% 
  map( ~{
    filter(.x, log2FC < 0) 
  })
```

Check out up regulated genes involved in go_bp

```{r up-enrichment}
artms_go_bp_up <- artms_up_sig %>% 
  imap(~{ 
    sig_mapping = AnnotationDbi::select(org.Hs.eg.db, .x$UNiProt, "ENTREZID", "UNIPROT") #mapping of significant genes from DPA
    bg_mapping = AnnotationDbi::select(org.Hs.eg.db, artms_results_list[[.y]]$UNiProt, "ENTREZID", "UNIPROT" )
    
    #run the enrichGO using
    ego <- enrichGO(gene = sig_mapping$ENTREZID,
                universe = bg_mapping$ENTREZID,
                keyType = 'ENTREZID',
                pvalueCutoff = 0.05,
                qvalueCutoff = 0.05,
                pAdjustMethod = 'BH',
                minGSSize = 10,
                ont='BP', #biological processes
                OrgDb = org.Hs.eg.db)
   return(ego)
    })
```

Check out downregulated genes in go_bp

```{r down-enrichment}
artms_go_bp_down <- artms_down_sig %>% 
  imap(~{ 
    sig_mapping = AnnotationDbi::select(org.Hs.eg.db, .x$UNiProt, "ENTREZID", "UNIPROT") #mapping of significant genes from DPA
    bg_mapping = AnnotationDbi::select(org.Hs.eg.db, artms_results_list[[.y]]$UNiProt, "ENTREZID", "UNIPROT" )
    
    #run the enrichGO using
    ego <- enrichGO(gene = sig_mapping$ENTREZID,
                universe = bg_mapping$ENTREZID,
                keyType = 'ENTREZID',
                pvalueCutoff = 0.05,
                qvalueCutoff = 0.05,
                pAdjustMethod = 'BH',
                ont='BP', #biological processes
                OrgDb = org.Hs.eg.db)
   return(ego)
    })
```

Check out downregulated genes in go_cc


```{r up-enrichment-gocc}
artms_go_cc_up <- artms_up_sig %>% 
  imap(~{ 
    sig_mapping = AnnotationDbi::select(org.Hs.eg.db, .x$UNiProt, "ENTREZID", "UNIPROT") #mapping of significant genes from DPA
    bg_mapping = AnnotationDbi::select(org.Hs.eg.db, artms_results_list[[.y]]$UNiProt, "ENTREZID", "UNIPROT" )
    
    #run the enrichGO using
    ego <- enrichGO(gene = sig_mapping$ENTREZID,
                universe = bg_mapping$ENTREZID,
                keyType = 'ENTREZID',
                pvalueCutoff = 0.05,
                qvalueCutoff = 0.05,
                pAdjustMethod = 'BH',
                ont='CC', #cellular compartment
                OrgDb = org.Hs.eg.db)
   return(ego)
    })
```

Check out downregulated genes in go_cc

```{r down-enrichment-gocc}
artms_cc_bp_down <- artms_down_sig %>% 
  imap(~{ 
    sig_mapping = AnnotationDbi::select(org.Hs.eg.db, .x$UNiProt, "ENTREZID", "UNIPROT") #mapping of significant genes from DPA
    bg_mapping = AnnotationDbi::select(org.Hs.eg.db, artms_results_list[[.y]]$UNiProt, "ENTREZID", "UNIPROT" )
    
    #run the enrichGO using
    ego <- enrichGO(gene = sig_mapping$ENTREZID,
                universe = bg_mapping$ENTREZID,
                keyType = 'ENTREZID',
                pvalueCutoff = 0.05,
                qvalueCutoff = 0.05,
                pAdjustMethod = 'BH',
                ont='CC',#cellular compartment
                OrgDb = org.Hs.eg.db)
   return(ego)
    })
```

No significant enrichment for GO CC down reg genes

```{r-dotplot-of-processes}
dotplot(artms_go_bp_up[['UreaSon-TFASon']], showCategory=30, title='GO BP enrichment results UreaSon-TFASon', font.size=8)


artms_go_bp_down %>% 
  imap( ~{
      dotplot(.x, 
      showCategory=30,
      title=paste0('GO BP enrichment results - ', .y),
      font.size = 8
        )
  })
    
```

```{r}

```


```{r get-enrichment-obj}
artms_go_df <- artms_sig_prots_list %>% 
  
  imap(~{ 
    sig_mapping = AnnotationDbi::select(org.Hs.eg.db, .x$UNiProt, "ENTREZID", "UNIPROT") #mapping of significant genes from DPA
    bg_mapping = AnnotationDbi::select(org.Hs.eg.db, artms_results_list[[.y]]$UNiProt, "ENTREZID", "UNIPROT" )
    
    #run the enrichGO using
    ego <- enrichGO(gene = sig_mapping$ENTREZID,
                universe = bg_mapping$ENTREZID,
                keyType = 'ENTREZID',
                pvalueCutoff = 0.05,
                qvalueCutoff = 0.05,
                pAdjustMethod = 'BH',
                ont='BP', #biological processes
                OrgDb = org.Hs.eg.db)
   return(ego)
    })

#artms_go_df[['Urea-TFA']] <- NULL
#artms_go_df[['TFA-TFASon']] <- NULL

enrichGO(gene = artms_sig_prots_list[["Urea-TFA"]]$UNiProt, universe = artms_results_list[['Urea-TFA']]$UNiProt, keyType = 'UNIPROT',
                         pvalueCutoff = 0.05,
                qvalueCutoff = 0.05,
                pAdjustMethod = 'BH',
                ont='CC', #biological processes
                OrgDb = org.Hs.eg.db)
         
```

```{r artms-plot-enrichment}

artms_go_df %>% 
  purrr::imap(~ { enrich_bar_fun(.x,.y)})

```

```{r artms-}

go_df <- artms_go_df %>% 
  purrr::imap(~ { as.data.frame(.x)})

```


Q. How many peptides and proteins per sample? 

Peptides.pdf/Proteins.pdf: number of unique peptides same for the entire dataset. TFA samples seem to detect both more proteins and peptides 
Number of unique proteins and peptides the same for each experiment?

IntensityStats.pdf: 
Total sum of intensities -Intensity sums highest in URea samples, but plot is misleading as y axis is truncated.  Quite similar numbers for all reps
Peptides counts: TFA samples  recover most

Q. Reproducibility between replicates of the same prep method?
PCA plot.pdf

Indicates broad prep method (Urea vs TFA) are the main drivers of variance in the dataset
TFA's show very strong clustering at both protein & peptide level, TFASon2 slightly removed from the others

Looking at Urea samples, not clusered as tightly at protein or peptide levels. UreaSon1 clearly seperated from other samples

ProteinIntensityCV.pdf/PeptideIntensityCV.pdf
Again TFA Samples seem to have the tightest distribution of CV values at both the protein & peptide level, indicating reduced spread relative to other methods

This is consistent across all protein expression groups (low to high expression)

ProteinDetection.pdf
TFA method shows highest overlap (92.2%)




Q. Overlap in peptide/protein IDs between methods 
ID-overlap.pdf

Up

Clean distinction of two clusters at a high level: TFA & UREA.
TFA/TFASon replicates protein IDs more similiar at both protein & pepetide levels. 
At protein level >90% similiarity across all four samples
TFA replicates 92.2% protein similiarity & 76.5% peptide similiarity

Less reproducibility in Urea-based methods, but still >87% at protein and ~70% at peptide level

Strong(er) overlap in TFA peptide & protein identification


Q. When different prep methods identify the same protein, is there any systematic difference in the abundance of proteins? because if so, I want to pick the method that give us higher abundances #assess mean and sd protein abundances for each method

Globally, it seems Urea-based methods detect higher intensities on average

# Custom Analysis

## High Level Inspection of Data

```{r read-in}

setwd('/Users/martingordon/Documents/projects/041223_Danielle_Swaney')

ms_counts <- read_delim('./data/MSstats.csv')

spec(ms_counts)

#10 cols & 57K rows
dim(ms_counts)

#why are all Fragment Ion & Product Cahrge values NA? Indicates DDA data?
summary(ms_counts)

#data already in long format
ms_counts[1:10,]

summary(ms_counts)

ms_counts %>% 
  distinct(Condition)
```

Histogram of overall intensities (log2)
Cut-off threshold? Don't seem to be any obvious 'garbage' overall

```{r}

ms_counts %>% 
  ggplot(mapping=aes(x=log2(Intensity))) +
  geom_histogram(intervals=100,binwidth = .1)
```

Histograms of each MS protocol. Distributions seem very similiar at high level

```{r}
ms_counts %>% 
  ggplot(mapping=aes(x=log2(Intensity), color=Condition)) + 
  geom_histogram(intervals=100,binwidth = .1) +
  facet_grid(rows=vars(Condition))
```

Looking at summary statistics for each group
Median values all in the range of 10^7
TFA Mean & Median values closest and lowest number of NA's
Also tighter variance around the mean 
Urea dataset has highest number of NA's of the four groups


```{r}
#split tibble into list of tibble on condition
ms_counts_cond <- ms_counts %>%
  split(f = as.factor(.$Condition)) 

map(ms_counts_cond, summary)
```

## balance of peptide identifications per MS run

Each row in the table should be the intensity from one MS "feature" per run (an elution peak from a single peptide ion). It's possible that there are multiple features per peptide ion and run--if the peptide elutes in two different peaks perhaps.

### are there any multiple-feature peptide ions? No

```{r}

ms_counts %>% 
  group_by(Run, PeptideSequence, PrecursorCharge) %>% 
  tally() %>% 
  dim()

  dim(ms_counts) #check if grouped table rows are same as original tibble.. true
  
```

confirmed with DT method
```{r, eval=FALSE}

ms_counts_dt <- fread('./data/MSstats.csv')

any(ms_counts_dt[, .N, by = .(Run, PeptideSequence, PrecursorCharge)]$N) > 1

```

Bar plot of number of number of peptides detected in each biosample

```{r peptide-intensities}
ms_counts_narm <- ms_counts %>% 
  filter( ! is.na(Intensity))

ggplot (data = ms_counts_narm, aes (x = paste0(Condition, "_", BioReplicate), fill = Condition)) + 
  geom_bar() 

```

Peptide intensities distributions for each  biosample. Intensity values are quite consistent between both conditions and replicates, but overall Urea-based samples seem to have slightly higher peptide intensities on average

```{r boxplot-intensities}
ggplot (data = ms_counts_narm, 
        aes (x = paste0(Condition, "_", BioReplicate), y = log2(Intensity), fill = Condition)) + 
  geom_boxplot() 
```


Count number of NA values in each individual biorep
TFA has the smallest number of NA values in df, but counts seem to be fairly close

```{r NA values}
# number of rows with NA values
ms_counts %>% 
  group_by(Condition,BioReplicate) %>% 
  summarise(na_counts = sum(is.na(Intensity))) 

```

Plot of missing values in the individual biosamples. TFA samples seem to have the lowest proportion of missing values, in addition to highest number of detected peptides (above).

```{r plot-NA-values}


missing.values <- ms_counts %>%
  group_by(Condition,BioReplicate) %>% #group by biosample
  mutate(isna = is.na(Intensity)) %>% #boolean if value missing
  mutate(total = n()) %>% #count total n records per sample
  group_by(Condition, BioReplicate, total, isna) %>% 
  summarise(num.isna = n()) %>% #calc na records per sample
  mutate(pct = num.isna / total * 100) # % for plotting



gg <- missing.values %>%
      ggplot() +
        geom_bar(aes(x = paste0(Condition, "_", BioReplicate), 
                     y = pct, fill=isna), 
                 stat = 'identity', alpha=0.8) +
     # scale_x_discrete(limits = levels) +
      scale_fill_manual(name = "", 
                        values = c('steelblue', 'tomato3'), labels = c("Present", "Missing")) +
      coord_flip() +
      labs(title = "% missing values per sample", x =
             'Variable', y = "% of missing values")

gg
```


# Run MSstats summarize peptide to protein intensities

## Remove NA values for MSstats data processing

MSStats likes explicit missing values, so ensure if protein detected in one of runs, it must be present with a NA value in other runs if not detected. Use BP utiliy function for this.

MSStats normalises by adjusting based on median expression counts
```{r}

#sanity check to ensure not odd number of rows for peptide groups. 
#NB to as MSStats likes NA values where no detection rather than missing row

ms_counts_t <- ms_counts %>% 
  group_by(PeptideSequence, PrecursorCharge) %>% 
  mutate(num = n())
unique(ms_counts_t$num)

rm(ms_counts_t)

dp.out <- MSstats::dataProcess(ms_counts,
                               normalization = 'equalizeMedians',
                               featureSubset = "highQuality",
                               summaryMethod = 'TMP',
                               censoredInt = 'NA', #censor NA values
                               remove_uninformative_feature_outlier = TRUE,
                               MBimpute = FALSE )
          
```

The MSstats::dataProcess normalized by adjusting so that median peptide intensity is equal across runs.  Inspect the output with another boxplot to look at ProteinIntensity trends.

```{r, fig.width = 10, fig.height = 4}
gg <- ggplot (dp.out$ProteinLevelData, aes (x = interaction ( SUBJECT, GROUP), y = LogIntensities, fill = GROUP)) + geom_boxplot()

gg
```

### Generating Data Process plots

EDA to investigate the data after MSS data normalisation & processing.
Plots are quite slow to generate..

```{r, eval=FALSE}
# 
# # Profile plot
dataProcessPlots(data=dp.out, type="ProfilePlot")
# 
# # Quality control plot 
dataProcessPlots(data=dp.out, type="QCPlot")	
# 
# # Quantification plot for conditions
dataProcessPlots(data=dp.out, type="ConditionPlot")

```

Differential Expression Analysis of the Different Proteins

First generate a contrast matrix for a pairwise comparison of sample groups
```{r}
# load helper functions to process the data
source ("./scripts/MSStats_V4_Functions.R")

# use helper function to make a contrast matrix  for all vs all comparison
contrasts <- makeContrast.AllByAll(dp.out)

# input contrast matrix for 
testResultComparisons <- groupComparison(contrast.matrix=contrasts, data=dp.out)

  
mss_results_list <- testResultComparisons$ComparisonResult %>%
  mutate(Protein = as.character(Protein)) %>% #ensure Protein name is character
  mutate(UNiProt = str_split_i(Protein, "\\|",2)) %>% #recover uniprot_ids
  group_by(Label) %>% 
  group_split() %>% 
  setNames(unique(testResultComparisons$ComparisonResult$Label))#can't directly do this in tidyverse I think, so take unique names from ori df used for splitting


```

Filter rows where 'issues' were found. Issues relate to missing values and only want to keep 'full-set' comparisons

```{r filter-issues}
#filter rows with 'issues'
mss_results_list <- purrr::map(mss_results_list, ~filter(.x, is.na(issue)))
```


Plot volcano plots for each comparison. For starters, just use the in-built volcano plot package

```{r lmm-volcano}

# # Volcano plot 
purrr::map(mss_results_list, ~groupComparisonPlots(data=.x, type="VolcanoPlot"))
```

```{r lmm-heatmap, eval=FALSE}
# Heatmap plot - not running
#purrr::map(mss_results_list, ~groupComparisonPlots(data=.x, type="Heatmap"))
```

Plots of protein expression across samples. Disabled as takes a long time to run

```{r lmm-comparison-p, eval=FALSE}
# # Comparison Plot
groupComparisonPlots(data=testResultComparisons$ComparisonResult, type="ComparisonPlot")

```

## Enrichment Analysis

Extract the significant genes for each comparison. For current comparisons, let it be proteins expression changes larger than log2FC >= 0.58 and p.adjust <= 0.05

Enrichment tests to identify significantly enriched proteins in GO Biological Processes 

```{r}

# sig_gene_list
sig_prots_list <- mss_results_list %>% 
  map( ~ extract_sig_prots(.x, 1, 0.05))

# first need to drop the empty elements where no sig differential expression
sig_prots_list[['TFA-TFASon']] <- NULL
sig_prots_list[['Urea-UreaSon']] <- NULL
```


```{r get-enrichment-obj}
go_bp_df <- sig_prots_list %>% 
  
  imap(~{ 
    sig_mapping = AnnotationDbi::select(org.Hs.eg.db, .x$UNiProt, "ENTREZID", "UNIPROT") #mapping of significant genes from DPA
    bg_mapping = AnnotationDbi::select(org.Hs.eg.db, mss_results_list[[.y]]$UNiProt, "ENTREZID", "UNIPROT" )
    
    #run the enrichGO using
    ego <- enrichGO(gene = sig_mapping$ENTREZID,
                universe = bg_mapping$ENTREZID,
                keyType = 'ENTREZID',
                pvalueCutoff = 0.05,
                qvalueCutoff = 0.05,
                pAdjustMethod = 'BH',
                ont='BP', #biological processes
                OrgDb = org.Hs.eg.db)
   return(ego)
    })

```

Plot results from the GO_BP analysis

```{r}
sig_mapping = AnnotationDbi::select(org.Hs.eg.db, sig_prots_list[['Urea-TFA']]$UNiProt, "ENTREZID", "UNIPROT") #mapping of significant genes from DPA
    bg_mapping = AnnotationDbi::select(org.Hs.eg.db, mss_results_list[[.y]]$UNiProt, "ENTREZID", "UNIPROT" )
    
    #run the enrichGO using
    ego <- enrichGO(gene = sig_mapping$ENTREZID,
                universe = bg_mapping$ENTREZID,
                keyType = 'ENTREZID',
                pvalueCutoff = 0.05,
                qvalueCutoff = 0.05,
                pAdjustMethod = 'BH',
                ont='BP', #biological processes
                OrgDb = org.Hs.eg.db)
```

```{r}

go_bp_df %>% 
  purrr::imap(~ { enrich_bar_fun(.x,.y)})

```



```{r t-test, eval=FALSE}

#extract the protein expression dataset
mss_df <- dp.out$ProteinLevelData
dp.out

(dp.out$ProcessedData)

#convert to wide format for t-test
mss_df_w <- mss_df %>% 
  as_tibble() %>% 
  unite('BioRep','GROUP':'SUBJECT', remove = T) %>% 
  select(-c(originalRUN,RUN,NumImputedFeature, TotalGroupMeasurements, NumMeasuredFeature, MissingPercentage, more50missing)) %>% 
  spread(BioRep, LogIntensities)

```


 #Create the different datasets (write function to do this)
```{r eval=FALSE}
TFA_TFAS_df <- mss_df_w %>% 
  select(c(Protein, TFA_1, TFA_2, TFASon_1, TFASon_2)) %>% 
  drop_na()


str(TFA_TFAS_df$TFASon_1)

t.test(TFA_TFAS_df[2,2:3],TFA_TFAS_df[2,4:5])

TFA_Urea_all = apply(TFA_Urea_df,1, function(x) { t.test(x[2:3], x[4:5]) $p.value } )

TFA_Urea_df <- mss_df_w %>% 
  select(c(Protein, TFA_1, TFA_2, Urea_1, Urea_2))

TFA_UreaS_df <- mss_df_w %>% 
  select(c(Protein, TFA_1, TFA_2, UreaSon_1, UreaSon_2))

TFAS_Urea_df <- mss_df_w %>% 
  select(c(Protein, TFASon_1, TFASon_2, Urea_1, Urea_2))

TFAS_UreaS_df <- mss_df_w %>% 
  select(c(Protein, TFASon_1, TFASon_2, UreaSon_1, UreaSon_2))

Urea_UreaS_df <- mss_df_w %>% 
  select(c(Protein, Urea_1, Urea_2, UreaSon_1, UreaSon_2))

```




Map Protein names to Gene IDs

```{r}

source ("./scripts/UniprotIDMapping.R")

mss_prot[, gene := multiUniprots2multiGenes(Protein)]


p <- ggplot(protQuant[gene %in% c("ARRB2","PRKCD")], aes (x = interaction ( SUBJECT, GROUP ), y = LogIntensities, color = gene)) + geom_line(aes(group = Protein)) + geom_point(aes(shape = as.factor(SUBJECT))) +  rotate.x.axis.text

```



Generated scatterplots of each pair of samples at peptide level to assess global expression patterns

Plot of peptide level abundances for each sample and compare pairwise

```{r peptide-scatter, eval=FALSE}
# use spread to convert to wide format
ms_counts_w <- ms_counts %>% 
  unite('BioRep', 'Condition':'BioReplicate',remove=FALSE) %>% 
  select(-c(Run,Condition, BioReplicate)) %>% 
  spread(BioRep, Intensity)


ms_counts_w %>% 
  ggplot(aes(log2(Urea_1),log2(Urea_2)), color=BioReplicate) +
  geom_point() +
  geom_smooth(method=lm)

ms_counts_w %>% 
  ggplot(aes(log2(TFA_1),log2(TFA_2))) +
  geom_point() +
  geom_smooth(method=lm)

ms_counts_w %>% 
  ggplot(aes(log2(TFASon_1),log2(TFASon_2))) +
  geom_point()

ms_counts_w %>% 
  ggplot(aes(log2(UreaSon_1),log2(UreaSon_2))) +
  geom_point()

ms_counts_w %>% 
  ggplot(aes(log2(Urea_1),log2(TFA_2)), color=BioReplicate) +
  geom_point()

```

Bioreplicate scatterplots

```{r protein-scatter, eval=FALSE}
# use spread to convert to wide format
ms_counts_w <- ms_counts %>% 
  unite('BioRep', 'Condition':'BioReplicate',remove=FALSE) %>% 


ms_counts_w %>% 
  ggplot(aes(log2(Urea_1),log2(Urea_2)), color=BioReplicate) +
  geom_point() +
  geom_smooth(method=lm)

ms_counts_w %>% 
  ggplot(aes(log2(TFA_1),log2(TFA_2))) +
  geom_point() +
  geom_smooth(method=lm)

ms_counts_w %>% 
  ggplot(aes(log2(TFASon_1),log2(TFASon_2))) +
  geom_point()

ms_counts_w %>% 
  ggplot(aes(log2(UreaSon_1),log2(UreaSon_2))) +
  geom_point()

ms_counts_w %>% 
  ggplot(aes(log2(Urea_1),log2(TFA_2)), color=BioReplicate) +
  geom_point()
  select(-c(Run,Condition, BioReplicate)) %>% 
  spread(BioRep, Intensity)

```

Generating Data Process plots

Perform DEA of the different cell types. This should give a decent indication of library prep protcols with significantly divergent expression patterns
We will perform a two sample t-test for each prep method. As there is only two replicates,

```{r}

```


Perform DEA of the different cell types. This should give a decent indication of library prep protcols with significantly divergent expression patterns





