---
title: "R Notebook"
output: html_notebook
---
This template file located at:
/Applications/RStudio.app/Contents/Resources/resources/templates/notebook_osx.Rmd

```{r}
library (data.table)
library (ComplexHeatmap)
library (ggplot2)
rotate.x.axis.text <- theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

source("../../bp_utils/ManageScriptData.R")
WriteSessionInfo()
WriteInstalledPackages()
```



RSV RSV RSV RSV


# AB
```{r}
AB <- fread ("~/Box/Project_ROADS_Roche/.//2021-08-15_RSV_PIV_Ab_Ph_A549_V1/Ab/dir/output/results.txt")

piv.ab.08 <- AB[Label == "PIV_8h-Mock_8h"]
piv.ab.24 <- AB[Label == "PIV_24h-Mock_24h"]

rsv.ab.08 <- AB[Label == "RSV_8h-Mock_8h"]
rsv.ab.24 <- AB[Label == "RSV_24h-Mock_24h"]

```

# PH
```{r}
PH <- fread ("~/Box/Project_ROADS_Roche/.//2021-08-15_RSV_PIV_Ab_Ph_A549_V1/Ph_V3/dir/output/results.txt" )
PH[, sigScore := sqrt(-log10(pvalue) * abs(log2FC))]
PH[!grepl(";", Protein), singleProtein := tstrsplit(Protein, "_")[[1]]]

piv.ph.08 <- PH[Label == "PIV_8h-Mock_8h"]
piv.ph.24 <- PH[Label == "PIV_24h-Mock_24h"]

rsv.ph.08 <- PH[Label == "RSV_8h-Mock_8h"]
rsv.ph.24 <- PH[Label == "RSV_24h-Mock_24h"]

```


# UB

```{r}
UB <- fread ("~/Box/Project_ROADS_Roche/.//2022-07-27_RSV_PIV_Ab_Ub_A549_Jack/data_for_roche/data/A549_UB_results.txt")
UB[, sigScore := sqrt(-log10(pvalue) * abs(log2FC))]
UB[!grepl(";", Protein), singleProtein := tstrsplit(Protein, "_")[[1]]]


piv.ub.08 <- UB[Label == "PIV_8h-Mock_8h"]
piv.ub.24 <- UB[Label == "PIV_24h-Mock_24h"]

rsv.ub.08 <- UB[Label == "RSV_8h-Mock_8h"]
rsv.ub.24 <- UB[Label == "RSV_24h-Mock_24h"]

```

#RNA

```{r}
rna.rsv <- openxlsx::read.xlsx ("/Users/ben/Box/Project_ROADS_Roche/RNAseq/RNAseq data RSV and PIV3 from Adrian/NHBE_24h_Mock_vs_RSV.xlsx") |> setDT()

ggplot(rna.rsv, aes(x = log2FoldChange, y = lfcSE)) + geom_point() + geom_density_2d()+ geom_hline(yintercept = 0.11)

rna.rsv[, adj.log2FC := ifelse(lfcSE > median(lfcSE), log2FoldChange * median(lfcSE)/lfcSE, log2FoldChange)]

ggplot(rna.rsv, aes(x = adj.log2FC, y = lfcSE)) + geom_point() + geom_density_2d()+ geom_hline(yintercept = 0.11)
ggplot(rna.rsv, aes(y = adj.log2FC, x = log2FoldChange, color = lfcSE)) + geom_point() + geom_density_2d()+ geom_hline(yintercept = 0.11)

median(rna.rsv$lfcSE)
hist(rna.rsv$lfcSE)
```

# PPI
```{r}
rsv.hek <- openxlsx::read.xlsx("/Users/ben/Box/Project_ROADS_Roche/PPI/ROADS/transfection/apms_RSV_HEK293T.xlsx", sheet = "result_filtered", startRow = 2) |> as.data.table()
rsv.a549 <- openxlsx::read.xlsx("/Users/ben/Box/Project_ROADS_Roche/PPI/ROADS/transfection/apms_RSV_A549.xlsx", sheet = "result_filtered", startRow = 2) |> as.data.table()
rsv.a549 <- rsv.a549[grepl("^RSV", Bait)]

rsv.lysate.hek <- openxlsx::read.xlsx("/Users/ben/Box/Project_ROADS_Roche/PPI/ROADS/Lysate Approach/result_HEK293T_PPI_MSstats.xlsx", sheet = "result_filtered", startRow  = 1) |> as.data.table()
rsv.lysate.hek[grepl ("^RSV", Bait.x)]


rsv.lysate.nhbe <- openxlsx::read.xlsx("/Users/ben/Box/Project_ROADS_Roche/PPI/ROADS/Lysate Approach/result_NHBE_PPI_MSstats.xlsx", sheet = "result_filtered") |> as.data.table()
rsv.lysate.nhbe <- rsv.lysate.nhbe[grepl ("^RSV", Bait.x)]



lapply(list (rsv.hek, rsv.a549, rsv.lysate.hek, rsv.lysate.nhbe),
       function(x)grep("Prey", colnames(x),value= TRUE))


allPreys <- unique(c(rsv.hek$Prey, rsv.a549$Prey, rsv.lysate.hek$Prey.x, rsv.lysate.nhbe$Prey.x))
preyUniverse <- unique( c( openxlsx::read.xlsx("/Users/ben/Box/Project_ROADS_Roche/PPI/ROADS/transfection/apms_RSV_HEK293T.xlsx", sheet = "result", startRow = 2)$Prey,
                   openxlsx::read.xlsx("/Users/ben/Box/Project_ROADS_Roche/PPI/ROADS/transfection/apms_RSV_A549.xlsx", sheet = "result", startRow = 2)$Prey,
                   openxlsx::read.xlsx("/Users/ben/Box/Project_ROADS_Roche/PPI/ROADS/Lysate Approach/result_HEK293T_PPI_MSstats.xlsx", sheet = "result_HEK293T_PPI_MSstats")$Prey.x,
                   openxlsx::read.xlsx("/Users/ben/Box/Project_ROADS_Roche/PPI/ROADS/Lysate Approach/result_NHBE_PPI_MSstats.xlsx", sheet = "result_NHBE_PPI_MSstats")$Prey.x,
                   rsv.ab.24$Protein)
                   
)

ppiHeats <- data.table (Protein = unique(c(preyUniverse, allPreys)), heat = 0)
ppiHeats[allPreys, heat := 1, on = "Protein"]

```



```{r}
source ("../../bp_utils/STRING_db_utils.R")
heats <- list(rna = rna.rsv[, .(gene = GetStringIDMapping.inOrder(name), heat = abs(adj.log2FC))][!is.na(gene)][!is.na(heat)],
               ab = rsv.ab.24[, .(gene = GetStringIDMapping.inOrder(Protein), heat = sqrt(abs(log2FC) * -log10(pvalue)))][!is.na(gene)][!is.na(heat)],
               ph = rsv.ph.24[, .(gene = GetStringIDMapping.inOrder(singleProtein), heat = sigScore)][!is.na(gene)][!is.na(heat)],
               ub = rsv.ub.24[, .(gene = GetStringIDMapping.inOrder(singleProtein), heat = sigScore)][!is.na(gene)][!is.na(heat)],
               ppi = ppiHeats[, .(gene = GetStringIDMapping.inOrder(Protein), heat = heat)][!is.na(gene)][!is.na(heat)]
              )
```



```{r}
heatDir <- file.path (ScriptNamedDir(), "heats")
dir.create(heatDir)

for (omics in names(heats)){
  fwrite (heats[[omics]], file.path (heatDir, sprintf("%s.csv", omics))  )
}


```

```{r}
ScriptNamedDir()
```

```{r}
#heatDir <- file.path(ScriptNamedDir(), "heats")

allSubsets <- function(s){
  unlist(sapply(1:length(s), function(m)combn(s, m, simplify = FALSE)), recursive = FALSE)
}
combos <- allSubsets(file.path(basename(heatDir), list.files(path = heatDir)))  |> sapply(paste0, collapse = " ")

fwrite ( data.table(x = combos), file.path(ScriptNamedDir(), "DataCombos.txt"), col.names = FALSE)
```



# DoNP.sh
```{R}

DoNP.string <- "#!/bin/bash

SMatrixPath='../../network_propagation/S_matrix.string11.5.gt900.time0.50.npy'
numPermutations=20000

combosPath='DataCombos.txt'
comboCounts=`grep -c . $combosPath`
echo Will do $comboCounts network propagation. This may take a while... 

# fall back to $1, first argument from command line
#SGE_TASK_ID=${SGE_TASK_ID:-$1}

for (( i=1; i<=$comboCounts; i++ ))
do
  SGE_TASK_ID=$i
  echo SGE_TASK_ID = $SGE_TASK_ID
  npOutPath='npOut.'$SGE_TASK_ID.csv
  # read the row from the combosPath file
  combos=`awk -v taskID=$SGE_TASK_ID 'NR==taskID {print $0}' $combosPath`
  
  echo Rscript DoMultiDataNetworkPropagation.R  $SMatrixPath $numPermutations $npOutPath $combos
  Rscript DoMultiDataNetworkPropagation.R  $SMatrixPath $numPermutations $npOutPath $combos
done
"

cat(DoNP.string, file  = file.path(ScriptNamedDir(), "DoNP.sh"))

```

# 
```{r}
DoMultiDataNetworkPropagation.string <- 'library (data.table)
library (RcppCNPy)

source ("../../../bp_utils/NetworkPropagation.R")


args = commandArgs(trailingOnly=TRUE)

SMatrixPath <- args[1]
numPermutations <- as.integer(args[2])
outFile <- args[3]
inputHeatFiles <- args[4:length(args)]


nodesTablePath <- gsub ("npy$", "nodeNames.csv", SMatrixPath)
message ("Loading S matrix from ", SMatrixPath, " and ", nodesTablePath)
S_matrix <- LoadNumPyS_matrix(SMatrixPath, nodesTablePath)


names(inputHeatFiles) <- tstrsplit(basename(inputHeatFiles), "\\\\.")[[1]]
if(any(duplicated(names(inputHeatFiles))))
  names(inputHeatFiles) <- sprintf ("%s.%d",  names(inputHeatFiles), 1:length(inputHeatFiles) )

heats  <- lapply(inputHeatFiles, fread)


npout <- NetworkPropagate.multiHeats.S_matrix(S_matrix,
                                              geneHeats = heats,
                                              numPermutations = numPermutations,
                                              networkHeatOnly = FALSE, 
                                              permuteOnlyInObserved = TRUE  # optional but important, I think
)


fwrite (npout, outFile)

'

cat(DoMultiDataNetworkPropagation.string, file = file.path(ScriptNamedDir(), "DoMultiDataNetworkPropagation.R"))

```



# modify path and copy to terminal
```{bash, eval = FALSE}
cd ~/UCSF/kroganlab/BenPolacco/roche/2023_06_02_RSV_NetProp_HeatDiffusion_data/ 
bash DoNP.sh
```


# Load NP output


```{r}
allNP <- lapply (list.files("~/UCSF/kroganlab/BenPolacco/roche/2023_06_02_RSV_NetProp_HeatDiffusion_data/", full.names = TRUE, pattern = "npOut.*csv"), fread)

inputFromColumnNames <- function(dt){
  grep ("heat0", colnames(dt), value = TRUE) |> tstrsplit("\\.", keep = 2) |> unlist() |> sort() |> paste0(collapse = "_")
}

names(allNP) <- sapply(allNP, inputFromColumnNames)
allNP.dt <- rbindlist(allNP, use.names = TRUE, fill = TRUE, idcol = "omics")
source ("../../bp_utils/UniprotIDMapping.R")
allNP.dt[,symbol := translateString2Gene(gene, species= "HUMAN")]

setorder(allNP.dt, adj.pvalue)

fwrite (allNP.dt, ScriptAndDatedFileName("NP.allIntegrations.csv.gz"))

```

```{r}
allNP.dt <- fread ("~/UCSF/kroganlab/BenPolacco/roche/2023_06_02_RSV_NetProp_HeatDiffusion_data/2023_06_05_NP.allIntegrations.csv.gz")
```

```{r}
source ("../../bp_utils/enrichmentTestFunctions.R")
gmt <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")
```

```{r}
enrich.dt <- enricherOnGroups(groupTable = allNP.dt[adj.pvalue < 0.05], geneColumn = "symbol", groupColumns = "omics",term2gene.gmt = gmt, universe = unique(allNP.dt$symbol), numProcessors = 8 )

simp <- simplifyEnrichBySimilarUniverseMembership(enrich.dt, gmt, groupColumn = "omics")

enrichHeatmapBestPerGroup(simp[[1]], simp[[2]], upperThreshold = 10, topN = 3, cols = unique(allNP.dt$omics))

#enpl <- enrichmentOnGroupsPL(groupTable = allNP.dt[adj.pvalue < 0.05], geneColumn = "symbol", groupColumns = "omics", gmt = gmt, universe = unique(allNP.dt$symbol), numProcessors = 8)
```

```{r, fig.width=6, fig.height=6}

enrichHeatmapBestPerGroup(simp[[1]], simp[[2]], upperThreshold = 10, topN = 10, cols = unique(allNP.dt$omics))

```


```{r, fig.width = 8.5, fig.height= 3.6}
columns <-  enrich.dt$omics |> unique() |> sort()

hm <- enrichHeatmapBestPerGroup(simp[[1]], simp[[2]], topN = 4, upperThreshold = 12,
                          column_split = stringr::str_count(columns, "_"),
                          cluster_column_slices = FALSE)

BackupAsPDF(hm)

```

```{r}
allNP.dt <- fread ("~/UCSF/kroganlab/BenPolacco/roche/2023_06_02_RSV_NetProp_HeatDiffusion_data/2023_06_05_NP.allIntegrations.csv.gz")


allNP.dt$omics |> unique() |> (function(x)grep("_", x,value = TRUE, invert  = TRUE))()

omicsOI <- allNP.dt$omics |> unique() |> (function(x)grep("_", x,value = TRUE, invert  = TRUE))()
omicsOI <- c(omicsOI,"ab_ph_ppi_rna_ub" )


for (t in c(1,1.5,2,4)){
  
  
  sets.list <- with (allNP.dt[adj.pvalue < 0.05 & log2PvE > log2(t)   & omics %in% omicsOI,],
                     split(symbol, omics))
  
  eulerr::euler(sets.list) |> plot(quantities = TRUE) |> print()
  
}
```



```{r, fig.width = 10, fig.height = 8}

stringIDs <- allNP.dt[omics %in% omicsOI & adj.pvalue < 0.05 & log2PvE > log2(2)][, .(name  = gene, gene = symbol)] |> unique()

stringGraph <- subsetSTRINGasGraph(stringIDs, threshold = 600)


  com <- igraph::walktrap.community(stringGraph)
  V(stringGraph)$walkTrap <- com$membership
  
library(ggraph)
p <- ggraph(stringGraph, layout = "fr")+
    geom_edge_fan(color = "gray") +
    geom_node_point(aes (color = as.factor(walkTrap)), show.legend = FALSE)  +
    geom_node_text(aes(label = gene, color = as.factor(walkTrap)), repel = TRUE, max.overlaps = 15, show.legend = FALSE, size = 2) +
    theme(panel.background = NULL)
  
p

```



# Sequential venn diagrams of integration

## data input

```{r}
AB <- rsv.ab.24[ abs(log2FC) > log2(1.5) & adj.pvalue< 0.05]
AB[, gene := multiUniprots2multiGenes(Protein)]
#nothing

PH <- rsv.ph.24[abs(log2FC) > 1 & adj.pvalue < 0.05 & is.finite(log2FC)]
PH[, gene := multiUniprots2multiGenes(singleProtein)]

UB <- rsv.ub.24[abs(log2FC) > log2(1.5) & adj.pvalue < 0.05 & is.finite(log2FC)]
UB[is.na(singleProtein), singleProtein := tstrsplit(Protein, "_")[[1]]]
UB[, gene := multiUniprots2multiGenes(singleProtein)]
#nothing

PPI <- ppiHeats[heat > 0]
PPI[, gene := multiUniprots2multiGenes(Protein)]

RNA <- rna.rsv[abs(log2FoldChange) > 1 & padj < 0.05]
RNA[, gene := geneAlias2officialGeneSymbol(name)]


p <- plot(eulerr::euler( lapply(list(AB = AB$gene, 
                           PH = PH$gene,
                           UB = UB$gene,
                           PPI = PPI$gene,
                           RNA = RNA$gene)[c("AB", "PH", "PPI", "RNA", "UB")], unique)),
     quantities = TRUE)

png(ScriptAndDatedFileName("1DataInput.png"))
p
dev.off()
p

BackupAsPDF(p)
```


## single propagations

```{r}
allNP.dt <- fread ("~/UCSF/kroganlab/BenPolacco/roche/2023_06_02_RSV_NetProp_HeatDiffusion_data/2023_06_05_NP.allIntegrations.csv.gz")
allNP.dt[,omics := toupper(omics)]
setorder(allNP.dt, omics)

np.singles <- allNP.dt[!grepl("_", omics) & adj.pvalue < 0.05 & log2PvE > log2(1), .(omics, gene = symbol)] |> with(split(gene, omics))

p <- plot(eulerr::euler(np.singles),
              quantities = TRUE)

png(ScriptAndDatedFileName("2SinglePropagations.png"))
p
dev.off()
p

BackupAsPDF(p)

```


## single + integrated


```{r}
allNP.dt <- fread ("~/UCSF/kroganlab/BenPolacco/roche/2023_06_02_RSV_NetProp_HeatDiffusion_data/2023_06_05_NP.allIntegrations.csv.gz")
allNP.dt[,omics := toupper(omics)]
setorder(allNP.dt, omics)

np.singles <- allNP.dt[!grepl("_", omics) & adj.pvalue < 0.05 & log2PvE > log2(1), .(omics, gene = symbol)] |> with(split(gene, omics))
np.plus <- c(np.singles, list("AB_PH_PPI_RNA_UB" = allNP.dt[omics == "AB_PH_PPI_RNA_UB"& adj.pvalue < 0.05 & log2PvE > log2(1),]$symbol))

p <- plot(eulerr::euler(np.plus),
              quantities = TRUE)


png(ScriptAndDatedFileName("3IntegratedPropagations.png"))
p
dev.off()
p


BackupAsPDF(p)

```
## PIV and RSV 
```{r}
PIV.set <- fread ("~/UCSF/kroganlab/BenPolacco/roche/2023_06_05_PIV_NP_HeatDiffusion_data/2023_06_05_NP.allIntegrations.csv.gz")[omics == "ab_ph_ppi_rna_ub"& adj.pvalue < 0.05 & log2PvE > log2(1.5),]$symbol

bothVirus <- list(RSV = allNP.dt[omics == "AB_PH_PPI_RNA_UB"& adj.pvalue < 0.05 & log2PvE > log2(1),]$symbol,
     PIV  = PIV.set)

p <- plot(eulerr::euler(bothVirus),
              quantities = TRUE)

p

png(ScriptAndDatedFileName("3PIVAndRSVCompared.png"), height = 2, width = 3, units = "in", res = 200)
p
dev.off()
p


```


```{r}
sets.dt <- rbindlist(lapply(bothVirus, data.table), idcol = "virus")

enrich.dt <- enricherOnGroups(sets.dt, geneColumn = "V1", groupColumns = "virus", term2gene.gmt = gmt, universe = unique(allNP.dt$symbol))
```

```{r}
simp <- simplifyEnrichBySimilarUniverseMembership(enrich.dt, gmt, groupColumn = "virus")
```

```{r, fig.width = 6, fig.height = 10}
enrichHeatmapBestPerGroup(simp[[1]], simp[[2]], groupColumn = "virus", topN = 50, reduceRedundantsAcrossGroups = TRUE, upperThreshold = 20)
```


# modularize


```{r}
source ("../../bp_utils/STRING_db_utils.R")
sigProteins <- unique(sets.dt$V1)
stringMapping <- GetStringIDMapping (sigProteins)

stringsOI <- c (setdiff(sigProteins, stringMapping$alias), stringMapping$string)  # because some were already string


stringEdges <- GetStringSubNetwork(stringsOI, oneHopConnections = TRUE) #includes 1 hop connectors


fwrite ( stringEdges,
        ScriptAndDatedFileName("StringSubNetwork.Preys.1hop.txt"),
        sep = "\t", col.names=FALSE)
```




### DSD in external python script

format of command is `python2 scriptName edgeFile outputDir`

This takes a little while, especially dependent on STRING subnetwork sizes. I downloaded the script from <http://dsd.cs.tufts.edu/capdsd/>     

```{bash}

python2 /Users/ben/UCSF/biotools/capDSD-src/DSD.py -c -s 20\
      ~/UCSF/kroganlab/BenPolacco/roche/2023_06_02_RSV_NetProp_HeatDiffusion_data/2023_06_05_StringSubNetwork.Preys.1hop.txt  \
      ~/UCSF/kroganlab/BenPolacco/roche/2023_06_02_RSV_NetProp_HeatDiffusion_data/2023_06_05_StringSubNetwork.Preys.1hop 



```








### load DSD results

Change path below to output of DSD above

```{r, fig.width = 12, fig.height = 12}
dsd.dt <- fread ("/Users/ben/UCSF/kroganlab/BenPolacco/roche/2023_06_02_RSV_NetProp_HeatDiffusion_data/2023_06_05_StringSubNetwork.Preys.1hop.dsd")

dsd.mat <- as.matrix(dsd.dt, rownames = "V1")

# remove the 1 hop connectors
stringsInMat <- intersect(rownames(dsd.mat), c(stringsOI))
dsd.mat <- dsd.mat[stringsInMat, stringsInMat]

dsd.dist <- as.dist (dsd.mat)
# ddr <- hclust (dsd.dist) %>% as.dendrogram()
# Heatmap(dsd.mat, show_row_names = FALSE, show_column_names = FALSE, cluster_rows = ddr, cluster_columns = ddr)
# ddr <- hclust (dsd.dist, method= "ward.D") %>% as.dendrogram()
# Heatmap(dsd.mat, show_row_names = FALSE, show_column_names = FALSE, cluster_rows = ddr, cluster_columns = ddr)
# ddr <- hclust (dsd.dist, method= "ward.D2") %>% as.dendrogram()
# Heatmap(dsd.mat, show_row_names = FALSE, show_column_names = FALSE, cluster_rows = ddr, cluster_columns = ddr)
ddr <- hclust (dsd.dist, method= "average") |> as.dendrogram()
Heatmap(dsd.mat, show_row_names = FALSE, show_column_names = FALSE, cluster_rows = ddr, cluster_columns = ddr)

```

### tsne based on DSD

For visualizing the modularization later

```{r, fig.width = 6, fig.height =  6}
t <- Rtsne::Rtsne(dsd.dist, is.distance = TRUE, theta = 0.0 ) # exact tsne when theta = 0.0

tsne.dt <- as.data.table(t$Y)
tsne.dt$string <- attr(dsd.dist, "Labels")
ggplot(tsne.dt, aes (x = V1, y = V2 )) + geom_point()

```

# modularization

```{r, fig.width = 6, fig.height =  6}
source ("../../bp_utils/PPINetworkModularized.R")

# cluster.dt <- defineModulesFromDistance("/Users/ben/UCSF/kroganlab/BenPolacco/roche/2023_06_02_RSV_NetProp_HeatDiffusion_data/2023_06_05_StringSubNetwork.Preys.1hop.dsd",geneNamesOI = stringsOI, deepSplit = 0.8)
# tsne.dt[cluster.dt, c("clusterID", "cluster") := .(i.clusterID, i.cluster), on = c(string = "gene")]


pam.out <- cluster::pam(dsd.mat,diss = TRUE, k = 12)
tsne.dt[names(pam.out$clustering), clusterID := pam.out$clustering, on = "string"]
tsne.dt[, cluster := sprintf ("clust.%02d", clusterID)]

tsne.dt[, c("x", "y") := .(V1, V2)]
p <- ggplot(tsne.dt, aes (x = x, y = y,)) +
  #geom_segment(data = edgeView.dt, aes(xend = xend, yend = yend),  alpha = 0.2, linewidth = 0.1) +
  geom_point(aes( color = cluster)) +
  #scale_size_manual(values = c(bait = 4, prey = 0.5)) +
  #scale_shape_manual(values = c(bait = 23, prey = 1)) +
  coord_fixed() + theme_void()
p
```

## enrich and label the modules
```{r}
tsne.dt[, gene := translateString2Gene(string, species = "HUMAN")]
#uniprots <- stringMapping[cluster.dt$gene, alias, on = "string"]
geneToCluster <- tsne.dt[,.(gene = gene, cluster = cluster)]

#source ("../../bp_utils/enrichmentTestFunctions.R")
# we are in string/uniprot land. Load a gmt that we can map to string
#gmt <- loadGmtFromBioconductor(ontology = "ALL", keyType = "UNIPROT")
#gmt.string <- gmt[stringMapping, .(term = ont, gene = i.string) , on = c(gene = "alias")][!is.na(term)]

enrichNames.out <- labelModulesByEnrichment(cluster.dt = geneToCluster, gmt = gmt, numProcessors = 8)

# choose a naming scheme
clusterNames.dt <- enrichNames.out$clusterNames.dt[, .(cluster, name = termScore.name, go = termScore.go)]
enrich.dt <- enrichNames.out$enrich.dt
```

```{r}

tsne.dt[clusterNames.dt, c("name", "go" ) := .( i.name, i.go), on = "cluster"]

tsne.dt[gmt, nameMatch := TRUE, on = c(gene = "gene", go = "ont") ]

tsne.dt[nameMatch == TRUE, nameMatchCluster := cluster]

```

```{r, fig.width = 10, fig.height = 6}
p <- ggplot(tsne.dt, aes (x = x, y = y,)) +
  #geom_segment(data = edgeView.dt, aes(xend = xend, yend = yend),  alpha = 0.2, linewidth = 0.1) +
  geom_point(aes( color = name  )) +
  #scale_size_manual(values = c(bait = 4, prey = 0.5)) +
  #scale_shape_manual(values = c(bait = 23, prey = 1)) +
  coord_fixed() + theme_void()
p
```



```{r}
sets.dt <- rbindlist(lapply(bothVirus, data.table), idcol = "virus")

sets.dt[stringMapping, string := i.string, on = c(V1 = "alias")]
sets.dt[is.na(string), string := V1]

virusMapped <- sets.dt[, paste(sort(virus), collapse= "_"), by = string]

tsne.dt[virusMapped, virus := i.V1, on = "string"]
```




```{r, fig.width = 8, fig.height = 4}
p <- ggplot(tsne.dt, aes (x = x, y = y,)) +
  #geom_segment(data = edgeView.dt, aes(xend = xend, yend = yend),  alpha = 0.2, linewidth = 0.1) +
  geom_jitter(aes( color = name , shape = virus, alpha = virus), height = 1, width = 1, ) +
  #ggrepel::geom_text_repel(data = tsne.dt[virus == "PIV_RSV"], mapping = aes(label = gene), size = 2, color = "black", show.legend = FALSE) +
  #scale_size_manual(values = c(bait = 4, prey = 0.5)) +
  #scale_shape_manual(values = c(bait = 23, prey = 1)) +
  scale_shape_manual(values = c(PIV  = 0, RSV = 2, PIV_RSV = 19)) +
  scale_alpha_manual(values = c(PIV = 0.5, RSV = 0.5, PIV_RSV = 0.8)) + 
  coord_fixed() + theme_void()
p

BackupAsPDF(p)

#png (ScriptAndDatedFileName("TSNE_RSV_"))

```


```{r}
PIV <- fread ("~/UCSF/kroganlab/BenPolacco/roche/2023_06_05_PIV_NP_HeatDiffusion_data/2023_06_05_NP.allIntegrations.csv.gz")[omics == "ab_ph_ppi_rna_ub"]
RSV <- fread ("~/UCSF/kroganlab/BenPolacco/roche/2023_06_02_RSV_NetProp_HeatDiffusion_data/2023_06_05_NP.allIntegrations.csv.gz")[omics == "ab_ph_ppi_rna_ub"]
alan <- openxlsx::read.xlsx("/Users/ben/Downloads/VIROME_MASTER_TABLE_EXTERNAL.xlsx")
alan.shortlist <- openxlsx::read.xlsx("/Users/ben/Downloads/VIROME_MASTER_TABLE_EXTERNAL.xlsx", sheet = 2)
```

```{r}
PIV[symbol %in% alan.shortlist$X2] |> View()


```
















