---
title: "NP06_interactionAnalysis"
author: "Martin Gordon"
date: "2023-10-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NPO6 new reps interaction analysis

Question is:
Is the efficacy/impact of NPFF treatment impacted by addition of CM101? ie. "(NP00 - NP30 CM00) - (NP00 - NP30 CM15)"

Different response to NPFF treatment based on CM101 treamtment status 

091023
---
redo with the renomralised data
```{r}
library (data.table)
library (ComplexHeatmap)
library (ggplot2)
library (emmeans)
library(stringr)

rotate.x.axis.text <- theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")
```

```{r}
mss <- fread('./output/mss.contrasts.batchCtrl.nob2.csv.gz')
p.quant <- fread('./NPO6.rerun.protein.noB2.lvl.quant.csv.gz')
#p.quant.all <- fread('NPO6.rerun.protein.lvl.quant.rm1outlier.csv.gz')
```

```{r}
p.quant[, `:=`(BTstatus = as.factor(str_extract(GROUP,'BT[0-9]{2}$')),
               CMstatus = as.factor(str_extract(GROUP, 'CM[0-9]{2}')),
               NPstatus = as.factor(str_extract(GROUP, 'NP[0-9]{2}')) )]


p.quant[, batch := as.factor(paste0('batch.',SUBJECT) )]

p.quant[,.N, by=.(GROUP, SUBJECT, BTstatus, CMstatus, NPstatus)]


BT00.p.quant <- p.quant[BTstatus != 'BT02',]
BT02.p.quant <- p.quant[BTstatus == 'BT02',]


p.quant.list <- list(BT00=BT00.p.quant,
                     BT02=BT02.p.quant)
```



run a lm on the data to assess if there is a significantly different..
No strong evidence of CM:NP treatment interaction effect

```{r}
model <- aov(LogIntensities ~ batch+BTstatus+CMstatus+NPstatus+CMstatus:NPstatus, data=p.quant)
BT00.model <- aov(LogIntensities ~ batch+CMstatus+NPstatus+CMstatus:NPstatus, data=BT00.p.quant)
BT02.model <- aov(LogIntensities ~ batch+CMstatus+NPstatus+CMstatus:NPstatus, data=BT02.p.quant)

model %>%  summary()
BT00.model %>%  summary()
BT02.model %>%  summary()
```

contrast formula:

```{r}

# linearModelsAllProteins can this function which computes statistic for a contrast of two different log2FC
#emmeans.contrastOfContrasts <- function (l, factorFormula = ~status|ant){  # how does the treatment effect change per antagonist level? basically compare treatment effect within each antagonist level
emmeans.contrastOfContrasts <- function (l, factorFormula = ~NPstatus|CMstatus){  # how does the NPFF treatment effect change with CM101 treatment status
  emm <- emmeans::emmeans(l, factorFormula)
  contrast1 <- pairs(emm)
  contrast2 <- pairs(contrast1, by = NULL, adjust = "none")
  return (as.data.table(contrast2))
}

# postProcessFunction = emmeans.contrastOfContrasts tells it to use the function above after doing anova.
# it is optional, adds significant time to the computation (~5x), but is nice to
# put a magnitude and p value on specific differences of differences
#
# cl = 6 parameter says to use 6 processes.  Set to NULL if multiprocess is failing, or higher or lower as you wish.
# Numbers near the number of processors available in your machine is probably counterproductive.

lm.out <- lapply(p.quant.list, function(x){
  
  # Numbers near the number of processors available in your machine is probably counterproductive.
  x.lm <- linearModelsAllProteins(x, 
                                  formulaList = list(full   = LogIntensities~batch + NPstatus + CMstatus + CMstatus:NPstatus), # full model; 4 way ANOV
                                  splitColumn = "Protein", postProcessFunction = emmeans.contrastOfContrasts, cl = 6) |> suppressWarnings()
  
  
  return(x.lm)
  
})

```
look at pval histgorams per term
looks to be v little to no. interaction between CM status and NP status
some genes DE with NP status and CM status, but no strong effects
```{r}
lapply(seq_along(lm.out), function(x,n,i){

  dt <- x[[i]][['anova']]
  str(dt)

 g <- ggplot(dt, aes(x=dt$p.value)) +
   geom_histogram(boundary=0.0, binwidth = 0.005) +
   xlim(c(0,0.05)) +
   facet_wrap(~dt$term)
 
 #BackupAsPDF(g, paste0(n[[i]], '.aov.term.pval.historgrams'))
 
},x=lm.out, n=names(lm.out))

```
## adjust pvalues 

```{r}
lapply(lm.out, function(x){
  
  x[['anova']][, fdr := p.adjust(p.value, method = "BH"), by= term]
  
})

```
Write out files to dir
```{r}

lapply(seq_along(lm.out), function(x,n,i){
  
  fwrite(x[[i]][['anova']], paste0(n[[i]], '.anova.contrasts.csv.gz'))
  
},x=lm.out, n=names(lm.out))


lm.out$BT00$anova
```

# CM101 treatment effects
No significant results at FDR < 0.05
Take genes significant at pval < 0.001 

```{r}

rowClusterWithNA <- function(mat, na.value = 0, ...){
    mat[is.na(mat)] <- na.value
    hclust(dist(mat), ...)
  }


#play with thresholds here
lapply(seq_along(lm.out), function(x,n,i){
  dt <- x[[i]][['anova']]
  
  sigProtein <- dt[term %in% c( "CMstatus") & p.value < 0.001 & !is.infinite(p.value),] %>% 
   #  .[order(p.value), head(.SD, 100), .SDcols = c('p.value', 'Protein')] %>% 
    .[,unique(Protein)]

 print(sigProtein)
  # filter just to our data
   prot.mat <- dcast (p.quant[Protein %in% sigProtein & BTstatus == n[[i]], ], Protein~GROUP+SUBJECT, value.var = "LogIntensities") %>% 
     as.matrix(rownames = "Protein")


  rownames(prot.mat) <- multiUniprotSites2multiGeneSites(rownames(prot.mat), species = "HUMAN")
  
  prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))
  
  hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
               cluster_rows = rowClusterWithNA(prot.mat),
                name = 'log2 Ints/Median',
              #column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Sites (p.val < 0.001)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("NP00", colnames(prot.mat)), "NP00", "NP30"),
                                ifelse(grepl("CM00", colnames(prot.mat)), "CM00", "CM15")))
  
# BackupAsPDF(draw(hm, column_title=paste0(n[[i]], ' CM101 treatment effects')), paste0(n[[i]], '.cm101.pval<0.001.heatmap'), dimensions = c(8,6))
  
},x=lm.out, n=names(lm.out))

```
# NPFF effects

```{r}

#play with thresholds here
lapply(seq_along(lm.out), function(x,n,i){
  dt <- x[[i]][['anova']]
  
  sigProtein <- dt[term %in% c( "NPstatus") & fdr < 0.05,] %>% 
   #  .[order(p.value), head(.SD, 100), .SDcols = c('p.value', 'Protein')] %>% 
    .[,unique(Protein)]
  
  print(sigProtein)

  # filter just to our data
   prot.mat <- dcast (p.quant[Protein %in% sigProtein & BTstatus == n[[i]], ], Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "Protein")


  rownames(prot.mat) <- multiUniprotSites2multiGeneSites(rownames(prot.mat), species = "HUMAN")
  
  prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))
  
  hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'log2 Ints/Median',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Sites (adj.pval < 0.05)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("NP00", colnames(prot.mat)), "NP00", "NP30"),
                                ifelse(grepl("CM00", colnames(prot.mat)), "CM00", "CM15")))
  
 #BackupAsPDF(hm, paste0(n[[i]], '.npff.pval<0.001.heatmap'), dimensions = c(8,9))
#BackupAsPDF(draw(hm, column_title=paste0(n[[i]], ' NPFF treatment effects')), paste0(n[[i]], '.npff.adj.val<0.05.heatmap'), dimensions = c(8,6))

  
},x=lm.out, n=names(lm.out))
```
## interaction effects
(different response to NPFF treatment, given CM101 treatment status)
Not v interesting results; many NA values

```{r}

#play with thresholds here
lapply(seq_along(lm.out), function(x,n,i){
  dt <- x[[i]][['anova']]
  
  sigProtein <- dt[term %in% c( "NPstatus:CMstatus") & p.value < 0.001,] %>% 
   #  .[order(p.value), head(.SD, 100), .SDcols = c('p.value', 'Protein')] %>% 
    .[,unique(Protein)]
  
  print(sigProtein)

  # filter just to our data
   prot.mat <- dcast (p.quant[Protein %in% sigProtein & BTstatus == n[[i]], ], Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "Protein")


  rownames(prot.mat) <- multiUniprotSites2multiGeneSites(rownames(prot.mat), species = "HUMAN")
  
  prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))
  
  hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'log Ints/Median',
                cluster_rows = rowClusterWithNA(prot.mat),
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Sites (p.val < 0.001)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              #column_split = list(
                           #     ifelse(grepl("CM00", colnames(prot.mat)), "CM-", "CM+")))
              column_split = list(ifelse(grepl("NP00", colnames(prot.mat)), "NP00", "NP30"),
                                ifelse(grepl("CM00", colnames(prot.mat)), "CM00", "CM15")))
  
#BackupAsPDF(hm, paste0(n[[i]], '.npff:cm101.pval<0.001.heatmap'), dimensions = c(8,9))
BackupAsPDF(draw(hm, column_title=paste0(n[[i]], ' NPFF:CM101 interaction effects')), paste0(n[[i]], '.cm101npff.pval<0.001.heatmap'), dimensions = c(8,6))
hm
  
},x=lm.out, n=names(lm.out))
```
contrast of contrast

## interaction effects observed in contrast of contrasts

Value here is we can require a magnitude of difference. 
Signs are confusing here because the loss of a negative effect will show as  a positive effect, and vice versa. 

```{r, fig.width = 8, fig.height = 8}

#play with thresholds here
lapply(seq_along(lm.out), function(x,n,i){
  dt <- x[[i]][['postProcess']]
  head(dt)
  
   sigProtein <- dt[contrast == "(NP00 - NP30 CM00) - (NP00 - NP30 CM15)" & p.value < 0.001 & abs(estimate) > 1, Protein]# %>% 
  

  # filter just to our data
  prot.mat <- dcast (p.quant[Protein %in% sigProtein & BTstatus == n[[i]], ], Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "Protein")


  rownames(prot.mat) <- multiUniprotSites2multiGeneSites(rownames(prot.mat), species = "HUMAN")
  
  prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))

  hm <- Heatmap(prot.mat, cluster_columns = FALSE,
                name = 'Ints/Median',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (p.val < 0.001)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 7),
              column_names_gp = gpar(fontsize = 8),
              column_split = list(
                                ifelse(grepl("CM00", colnames(prot.mat)), "CM-", "CM+")))

 #BackupAsPDF(hm, paste0(n[[i]], '.npff:cm101.contrastofcontrasts.pval<0.001.heatmap'), dimensions = c(8,9))
BackupAsPDF(draw(hm, column_title=paste0(n[[i]], 'CM+/CM- contrast of contrasts')), paste0(n[[i]], '.cm101npffContrastofcontrasts.pval<0.001.heatmap'), dimensions = c(8,6))

},x=lm.out, n=names(lm.out))
```

