---
title: "121124_AFCRISPR_Integration"
author: "Martin Gordon"
date: "2024-12-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 

```{r}
library(data.table)
library(magrittr)
library(readxl)
library(ggplot2)

source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/UniprotIDMapping.R")
source ("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
```

# read in the CRISPR results (summarised to the gene level) and annotate the data.table with the CRISPR positive and negative selection hits from Ujwals analysis
# also first confirm my table scores match the inital datatable

# CRISPRactivation: upreg (GTF+ pop) are proviral, downreg (GFP -vep pop) are antiviral
# CRISPR inhibition: upreg (GTF+ pop) are antiviral, downreg (GFP -ve pop) are proviral

```{r}

merge.dt <- fread('./201124_AFInteractiorsIntegration_data/2024_11_21_af.hiv.interactors.MISTscores.iMEX.anno.csv')
# this is the HIV interacted set
ori.dt <-  fread('./data/AF2M_HsHIV_interactors.csv')

ori.dt[Protein_1_name == 'PPIAL4F' & Protein_2 %in% c('CA-SF', 'Gag-SF'),]
merge.dt[Protein_1_name == 'PPIAL4F' &  Protein_2 %in% c('CA-SF', 'Gag-SF'),]

# this looks v similar
merge.dt[!is.na(HEK_MiST_score) & Bait =='CA' & Protein_1_name == 'PC']
ori.dt[ Protein_2 =='CA-SF' & Protein_1_name == 'PC']
```
Read in the CRISPR data and visualize


I think maybe we want pvalue, 
```{r}
crispr.ko.dt <- read_xlsx('./data/CRISPRko_significant_genes.xlsx') %>% 
  as.data.table()
crispr.act.dt <- read_xlsx('./data/CRISPRact_significant_genes.xlsx') %>% 
  as.data.table(colnames)


# mageck tests each gene for both positive and negative selection... I think we only want to show the results
# maybe as this is a targeted subset, we should only report
# no geens are both positively and negatively enriched
colnames(crispr.ko.dt) <- gsub("\"", "", colnames(crispr.ko.dt)) 
colnames(crispr.act.dt) <- gsub("\"", "", colnames(crispr.act.dt)) 

# ok, no genes significant at both directions
crispr.ko.dt[neg.fdr < 0.05 | pos.fdr < 0.05,]
crispr.act.dt[neg.fdr < 0.05 | pos.fdr < 0.05,]
```
Lets take the most signficiant result for each crispr screen 
```{r}
# to use for merging 
crispr.ko.dt[, Protein_1_name := id]
crispr.act.dt[, Protein_1_name := id]


#ko so donwreg are GFP+ proviral factors
crispr.ko.dt[, status := 'na']
crispr.ko.dt[, status := ifelse(neg.p.value < pos.p.value, 'proviral', 'antiviral')]

#ko so donwreg are GFP+ proviral factors
crispr.act.dt[, status := 'na']
crispr.act.dt[, status := ifelse(neg.p.value > pos.p.value, 'proviral', 'antiviral')]

# sanity check
crispr.act.dt[id == 'CD4']
crispr.ko.dt[id == 'CD4']
crispr.act.dt[id == 'TRIM5']
crispr.ko.dt[id == 'MYB']
```
 # okcreate new columns based pval and lfc and 
```{r}
crispr.act.dt[, `:=`(CRISPRa_adj.pvalue=ifelse(status == 'proviral', pos.fdr, neg.fdr),
                     CRISPRa_log2FoldChange=ifelse(status == 'proviral', pos.lfc, neg.lfc),
                     CRISPRa_pvalue=ifelse(status == 'proviral', pos.p.value, neg.p.value)
                     )]

crispr.ko.dt[, `:=`(CRISPRko_adj.pvalue=ifelse(status == 'proviral', neg.fdr, pos.fdr),
                    CRISPRko_log2FoldChange=ifelse(status == 'proviral', neg.lfc, pos.lfc),
                    CRISPRko_pvalue=ifelse(status == 'proviral', neg.p.value, pos.p.value)
                     )]

crispr.act.dt[id == 'CD4', .(pos.fdr, pos.lfc, neg.fdr, neg.lfc, CRISPRa_adj.pvalue, CRISPRa_Log2FoldChange)]
crispr.act.dt[id == 'TRIM5', .(pos.fdr, pos.lfc, neg.fdr, neg.lfc, CRISPRa_adj.pvalue, CRISPRa_Log2FoldChange)]

crispr.ko.dt[id == 'CD4', .(pos.fdr, pos.lfc, neg.fdr, neg.lfc, CRISPRko_adjPvalue, CRISPRko_log2FoldChange)]
crispr.ko.dt[id == 'MYB', .(pos.fdr, pos.lfc, neg.fdr, neg.lfc, CRISPRko_adjPvalue, CRISPRko_log2FoldChange)]
```

Ok, merge the CRISPR data  with the full set of results
```{r}
merge.dt <- fread('./201124_AFInteractiorsIntegration_data/2024_11_21_af.hiv.interactors.MISTscores.iMEX.anno.csv')
merge.dt
# ok, now lets merge with the activating set
new.merge.dt  <- merge(x=merge.dt, y=crispr.act.dt[, .(Protein_1_name=id, CRISPRa_log2FoldChange, CRISPRa_pvalue, CRISPRa_adj.pvalue)], by='Protein_1_name', all.x=T)


# now with the inhibiting set
new.merge.dt  <- merge(x=new.merge.dt, y=crispr.ko.dt[, .(Protein_1_name=id, CRISPRko_log2FoldChange, CRISPRko_pvalue, CRISPRko_adj.pvalue)], by='Protein_1_name', all.x=T)


new.merge.dt[!is.na(CRISPRa_log2FoldChange), unique(Protein_1_name)]

# 15 of these records are sig
crispr.act.dt[!id %in% new.merge.dt[!is.na(CRISPRa_log2FoldChange), unique(Protein_1_name)]]
crispr.act.dt[!id %in% new.merge.dt[!is.na(CRISPRa_log2FoldChange), unique(Protein_1_name)] & (neg.fdr < 0.05 | pos.fdr < 0.05) & abs(pos.lfc) > 1, unique(id)]


# 20 hits not found # KMT2A not found
crispr.ko.dt[!id %in% new.merge.dt[!is.na(CRISPRko_log2FoldChange), unique(Protein_1_name)]]
crispr.ko.dt[!id %in% new.merge.dt[!is.na(CRISPRko_log2FoldChange), unique(Protein_1_name)] & (neg.fdr < 0.05 | pos.fdr < 0.05)  & abs(pos.lfc) > 1, unique(id)]

```
May want to share the non-overlaping set, but for now write outpu

```{r}
fwrite(new.merge.dt, ScriptAndDatedFileName('af.hiv.interactors.MISTscores.iMEX.CRISPR.anno.txt'))
```


Prepare our AF3 input to compare results

```{r}
# take this set as we have AF scores for these interactors
ppi.dt <- merge.dt[HEK_MiST_score >= 0.75 | Jurkat_MiST_score >= 0.75, .(Bait, Protein_1)]

# write out the files we need
fwrite(ppi.dt, sep=',', col.names =F, ScriptAndDatedFileName('af3.hiv.jobTable.txt'))
fwrite(data.table(merge.dt[HEK_MiST_score >= 0.75 | Jurkat_MiST_score >= 0.75, unique(Protein_1),]), col.names = F, ScriptAndDatedFileName('af3.hiv.preyuniprots.txt'))
```


```{r}
mist.scores.list <- list(HEK = as.data.table(read_xls('./docs/NIHMS358189-supplement-Supplementary_Data_2.xls', sheet=1)),
                         Jurkat = as.data.table(read_xls('./docs/NIHMS358189-supplement-Supplementary_Data_2.xls', sheet=2)))


mist.scores.dt <- rbindlist(mist.scores.list, idcol='cellline')
mist.scores.dt[, PPI := paste(Bait, '__', PreyGene)]
mist.scores.dt[, PreyGene := multiUniprots2multiGenes(Prey, species='HUMAN')]


mist.scores.dt[`MiST score` > 0.75, .(Bait, Prey)] %>%  unique()
mis
```

ori.dt[Protein_1_name == 'PPIAL4F',]