---
title: "kenrichment_tool"
author: "Martin Gordon"
date: "2023-07-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

TODO

16/08/23
Wary of inflating significance; maybe do the sig gene counting after filtering for sites not in DB? larger denominator than expected? prob adjust
Wrapper to split each contrast and run function on each (necessary? maybe just loop through the contrasts and plot for each. Not urgent atm
Heatmap of the enrichment results

14/08/23
String match using Bens code as template
multi-ph sites: split sites and take single. Replicate with the script used to process msout data

08/08/23
match by string rather than prot name
Lapply for multiple contrasts
heatmap plotting function
add kinase type info to the db file?
mlutiple testing? currently correcting per contrast (n=606, up & down together); correct per kinase instead (up and down seperately)
Considering sites that are both up and down reg?

Find discrepancies in counts in app and our code; numerators different although sig hits the same, but less counts in our data... more sites available on website? 
-  match by AA sequence rather than uniprot_site combo? maybe more sensitive and seems to be how app works..


```{r packages}
library(data.table)
library(ggplot2)
library(magrittr)
library(readxl)
library(ggrepel)


# load bens util scripts for the analysis
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/LocatePeptidePtmInProteins.R")
source ("../..//utils/bp_utils/MSstats_Helper_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")
source ("../../utils/bp_utils/KinaseActivityScores.R")

# script with enrichment functions., plots etc.
source ("../../utils/mg_utils/HelperFunctions.R")
```


## Convert the tool for kinase enrichment analysis from Pyhton to R

Tool available here: https://kinase-library.phosphosite.org/ea?a=de
Goal is to run this in-house per contrast/label 
TODO: annotate with kinase group information? We also don't distinguish from S/Th vs Tyrosine enrichment.. maybe include gorup. info They only considet the S/Thr set

General workflow:
Score all sites for all kinases (think this is complete) - For scoring substrates, the values of the corresponding amino acids in the corresponding positions were multiplied and scaled by the probability of a random peptide (Supplementary Note 2)

1) categorize phosphosites in MS data as  up,down,not-regulated based on log2FC & adj.pval (define our standard threshodls)
2) Find top 15 predicted kinases per site based on percentile
3) Per kinase, compare the number of times it appears in up-regulated prediction set vs non-regulated prediction
set using 
3a) one-sided Fisher’s exact test for a p value. (Repeat for down-regulated vs non-regulated)
3b) Calculate frequency factors (or log odds ratio)
4) Adjust p.values based on number of kinases

Input data?
Based on our other enrichment results we would want the output of the FC estimates per site per contrast (controlling for abundance? Basically a subtraction of AB from the PH)
cols: site, protein, log2FC, pval

For testing, take a  subset of the AViDD PH dataset (significant results for one contrast)


```{r}

# for now just take 50 
mss.ph <- fread('/Users/martingordon/Documents/projects/060123_DSwaney_AshworthPH/060123_DSwaney_AshworthPH_data/2023_06_01_msstats_groupcomparisons.csv.gz') %>% 
  .[Label %chin% c('WT+ vs EV+', 'WT+ vs N10t20+') & abs(log2FC) > log2(1.5) & adj.pvalue < 0.05 & !issue %like% 'Condition' ]

test.ph <- fread('/Users/martingordon/Documents/projects/060123_DSwaney_AshworthPH/060123_DSwaney_AshworthPH_data/2023_06_01_msstats_groupcomparisons.csv.gz') %>% 
  .[Label %chin% c('WT+ vs EV+'),]
  #.[Label %chin% c('WT+ vs EV+') & abs(log2FC) > log2(1.5) & adj.pvalue < 0.05 & !issue %like% 'Condition' ] %>% 
  #.[1:20,] # rem any rows with condition issues


```

Look at the DB file:
Alphabetically ordered for each protein based on UNIPROT ID, phosphosite
Gives the sequence for each site where it is phosphorylated; take just the rank 1:15 %between% data.table function

To do:
- format the input file from our log2 test [protein, site, log2FC, pval, adj.pval]


test the tool by first creating a file format we need:
merge the contrast table with the PH id/count table (need to create this first)

```{r }
#function to read in keys and spec file overwrite spec Condition and BioReplicate cols with keys file values
preprocessSpec <- function(spec_f,keys_f){

  #fread stores as data.table by default
  spec <- fread(spec_f)
  keys <- fread(keys_f)

  # apply Condition, BioReplicate from keys to spec
  spec[keys , Condition := i.Condition , on = c(Run = "RawFile")] #overwrite condition col w values from keys file, match run col in spec to rawfile col in keys
  spec[keys , BioeReplicate := i.BioReplicate , on = c(Run = "RawFile")]
  
  return(spec)
  setnames(spec) = dat_name
}
```


```{r load-data}

specs <- '/Users/martingordon/Documents/projects/060123_DSwaney_AshworthPH/data/MSstats_PH.txt'
keys <- '/Users/martingordon/Documents/projects/060123_DSwaney_AshworthPH/data/keys.txt'

spec <- preprocessSpec(specs,keys)

```

Check each biorep has a unique identifier for MSstats processing
```{r}
#remove keys biorep column
spec[,10] <- NULL
spec[,.N, by=.(Condition)]
spec[, .N, by=.(Condition,BioReplicate)]
```

Remove likely artifacts

```{r}

ggplot(spec, aes(x=log2(Intensity))) +
  geom_histogram(bins = 100)

spec <- spec[log2(Intensity) >5,]

ggplot(spec, aes(x=log2(Intensity))) +
  geom_histogram(bins = 100)

```
Confirm that each detected peptide occurs once per run/charge
False indicates each peptide detected only once per run
```{r n-ids}
any(spec[, .N, by=.(PeptideSequence,PrecursorCharge,Run)]$N) > 1
```

No fasta file specified, so just run with the canonical list of UNIProt proteins prepared by Ben & available on the Krogan github

```{r}
#function belwo takes path to fasta reference
fasta <- '/Users/martingordon/Documents/projects/060123_DSwaney_AshworthPH/data/human_all_proteins_canonical_uniprot-proteome_UP000005640.fasta.gz'

spec.ph <- sitifyProteins_SpectronautFile(spec, fastaFile = fasta, site = "PH") 

#fwrite(spec.ph, './data/spec.ph.csv.gz')

```

Output for the tool: 
For each kinase; it outputs the name, its group , n upregualted sites in the data, n downregulatd sites in the data, background set sites(??), dominant set (upreg/downreg), log2 enrihcment val, p-val, adj-pval

Questions:
- Not all our PH sites are in the kinase db? How likely are FP hits in data we generate?
- multi sites per line, duplicate lines or is this double counting? For simplicity, remove multi PH sites per peptide. Other options? Split each site on (tstrsplit), duplicate rows per site, take one site? If dupliating and multiple peptides, take most significant, or avg of the peptide?
- Need to handle different contrasts also!!! lapply, but for now just keep simple until running


Prepare kinase DB for function use
Top 15 based on percentile (use second approach to break ties)
(top 15 based on percentile below..)

I think I have to match the sites...

```{r}
#kinase.db <- data.table(read_xlsx('./docs/kinase.scores.xlsx', sheet=2))
#kinase.db[,Protein.Site := paste0(`Uniprot Primary Accession`,'_',Phosphosite)]

#neworder <- c(1:8,620,9:619)

#setcolorder(kinase.db,neworder)

#head(kinase.db)

# write out this file to somewhere 
#fwrite(kinase.db, '~/Documents/utils/mg_utils/data/kinase.scores.csv.gz')

kinase.db <- fread('~/Documents/utils/mg_utils/data/kinase.scores.csv.gz')

# drop unnecessary cols form kinase.db for now; just keep IDs, kinase_ranks

cols.noi <- c('Database Uniprot Accession', 'Uniprot Entry','Gene', 'Alternative Gene Names', 'Protein',
              'Description', 'Phosphosite', "median_percentile","promiscuity_index",'SITE_+/-7_AA','Kinase','Database', 'phos_res', grep("*rank*",colnames(kinase.db), value=TRUE))
kinase.db <- kinase.db[, !..cols.noi]

kinase.db <- melt(kinase.db, id.vars = c('Uniprot Primary Accession','Protein.Site'))

colnames(kinase.db) <- c('Uniprot', 'Protein.Site', 'Kinase', 'Percentile')
kinase.db[, Kinase := sub('_percentile', '', Kinase)]


#write out all kinase percentile scores
kinase.db <- fread('~/Documents/utils/mg_utils/data/kinase.percentiles.long.csv.gz')

#enriched kinases -top 15 per site
# fixed as based on available DB

kinase.db.prc <- kinase.db %>% 
  .[order(Protein.Site,-Percentile),] %>% 
  .[, head(.SD, 15), by = Protein.Site] %>% 
  .[,.(Protein.Site,Kinase,Percentile)]

#fwrite(kinase.db.prc, '~/Documents/utils/mg_utils/data/kinase.percentile.top15.long.csv.gz')


kinase.db[, unique(Kinase)]
# go with this
prc.scores.long <- fread('~/Documents/utils/mg_utils/data/kinase.percentile.top15.long.csv.gz')



#kinase.db <- fread('~/Documents/utils/mg_utils/data/kinase.scores.long.csv.gz')

#top 15 per site
prc.scores.long[, .N, by=Protein.Site] %>% 
  .[N != 15,]
      

#some 606 per site; duplicate uniprot Ids...
#need to go by site, but write the app first
kinase.db[, .N, by=Protein.Site] %>% 
  .[N!= 303,]

```

1) categorize phosphosites in MS data as up,down, not-regulated based on log2FC & adj.pval
Our thresholds: log2FC > 0.58, adj p.val < 0.05

```{r}
## lets read in the raw data for this
mss.ph <- fread('/Users/martingordon/Documents/projects/060123_DSwaney_AshworthPH/060123_DSwaney_AshworthPH_data/2023_06_01_msstats_groupcomparisons.csv.gz') 

test.ph <- mss.ph %>% 
  #.[Label %chin% c('WT+ vs EV+', 'WT+ vs N10t20+') & abs(log2FC) > log2(1.5) & adj.pvalue < 0.05 & !issue %like% 'Condition' ] %>% 
  .[Label %chin% c('WT+ vs EV+') & issue == '',] %>%
  .[ Protein %like% "_" & !Protein %like% ';' ,] #keep only sites with mods and rem multi-uniprots brings to 10954 when removing stuff
  #.[1:10000,] # rem any rows with condition issues

#classify sites (set pval = 1 as not looking at p-val threshold)
class.ph <- classifySites(test.ph, log2FC = 0.58, pval = 0.05)
```

2) Identify biochemically favoured sites
Like paper only looking at singly phosphorylated peptides (multi-ph dropped) * create warning for how many are dropped
Find the top 15 kinases per site based on percentile (should we use score this looks to find top candidates per substrate?? I think percentile looka)

Want warning message like this: baiscally break up sites into not reg upreg and down reg
8% (121 out of 1442) upregulated sites are invalid
26% (167 out of 635) downregulated sites are invalid
21% (2807 out of 13287) background sites are invalid

Merge sig and kinase.db site info

Not filtering for sites only in DB at moment as not done by algorithm.. maybe an issue as increasing the denominator..
We also don't distinguish from S/Th vs Tyrosine enrichment..

```{r}
#top15 kinases per all site in DB; give same results in most cases...
kinase.db <- fread('~/Documents/utils/mg_utils/data/kinase.prc.scores.long.csv.gz') #percentiles
#kinase.db <- fread('~/Documents/utils/mg_utils/data/kinase.ranks.long.csv.gz') #ranks 

# calculate set size - NOT adjusting only for proteins present in kinase DB.. ask Ben
# this has all counts of sig sites in the data 
# add to classify sites
class.ph[, `:=`(upregulated_set_size   = sum(sig == 'Up'),
                downregulated_set_size = sum(sig == 'Down'),
                background_set_size    = sum(sig == 'Not'))]

#keep only those in DB... dropping multi-PH sites
message('dropping multisites as not present in DB...')

mapper <- kinase.db[Protein.Site %in% class.ph$Protein,] %>% 
  .[order(Protein.Site)]

# sanity check.. Y when we use rank as there are ties
any(mapper[, .N, by=.(Protein.Site)]$N != 15)

# check loss of sites
start.sites <- class.ph[, .N,by=Protein] %>% nrow()
end.sites <- mapper[, .N, by=.(Protein.Site)] %>% nrow()

sprintf("% s of % s sites detected in kinase DB. This function does not currently handle multi-phosphorlyated peptides. Continuing... ", end.sites, start.sites) 
# TODO break down discarded sites
#maybe include a list of PH sites not analysed ( and seperately the multisites; print to screen as error msg.)


message('merging...')
# merge exp data each site; top 15 kinases per site
# good; 15 kinases per site in data
mapper[class.ph, `:=`(sig = sig, 
                   upregulated_set_size = upregulated_set_size, 
                   downregulated_set_size = downregulated_set_size,
                   background_set_size = background_set_size), on=c(Protein.Site = "Protein")] %>% 
  .[order(Protein.Site,-Percentile)]

```


3) Per kinase, compare the number of times it appears in up-regulated prediction set vs unregulated prediction

```{r}
# group by kinase, count breakdown
cont.tab <-   mapper[, `:=`(upregulated_set_hits   = sum(sig == 'Up'),
                            downregulated_set_hits = sum(sig == 'Down'), 
                            background_set_hits    = sum(sig == 'Not')), by=.(Kinase)] %>%
  .[, dominant_direction := ifelse(upregulated_set_hits > downregulated_set_hits, 'up', 'down')] %>% 
  .[order(Kinase), .(Kinase,upregulated_set_hits,upregulated_set_size,
                     downregulated_set_hits,downregulated_set_size,
                     background_set_hits,background_set_size, dominant_direction)] %>% unique() #not sure if this is v tidy...

cont.tab[order(Kinase)] %>% head()
app.out[order(kinase)] %>% head()


cont.tab
```
3a) one-sided Fisher’s exact test for a p value. (Repeat for down-regulated vs non-regulated) (or log odds ratio)
- Do Fishers Exact Test & extract log odds ratio (frequency factors maybe)

```{r}

cont.tab

# test this function
# fisher test on each row
#               outcome
#              +       -
#          +-------+-------+
#        + |   a   |   b   | a + b
#  group   +-------+-------+
#        - |   c   |   d   | c + d
#          +-------+-------+
#            a + c   b + d
#

row_fisher <- function(counts.tab, countCols = c('upregulated_set_hits','upregulated_set_size','background_set_hits','background_set_size'), # cols to use to create matrix - c(a,b,c,d) ORDER IMPORTANT! See diagram above
                       direction = 'up', return.OR = TRUE,
                       alt = 'greater', cnf = 0.95) {
  
  mat <- matrix(as.numeric(counts.tab[countCols]), ncol=2, byrow=T)
  
  f <- fisher.test(mat, alternative = alt, conf.level = cnf)
  
  if (return.OR == TRUE) {
    return(c(counts.tab,
          assign(paste0(direction, pvalue)) = f$p.value,
          odds.ratio = f$estimate[[1]],
          or.lower = f$conf.int[1],
          or.upper = f$conf.int[2]))
  } else {
    return(c(counts.tab,
          pval = f$p.value))
  }
}




out <- t(apply(cont.tab, 1, row_fisher, return.OR = F))
cont.tab


ftest.out <- rbindlist(apply(cont.tab, 1, 
      function(x){
          
           up.ft <- fisher.test(matrix(as.numeric(x[c('upregulated_set_hits','upregulated_set_size','background_set_hits','background_set_size')]), ncol=2, byrow=T),
                       alternative='greater')
           
           up.x <- data.table(Kinase    = x['Kinase'],
                              test       = 'up',
                              odds.ratio = up.ft$estimate,
                              pval       = up.ft$p.value)
                           
           down.ft <- fisher.test(matrix(as.numeric(x[c('downregulated_set_hits','downregulated_set_size','background_set_hits','background_set_size')]), ncol=2, byrow=T),
                        alternative='greater')
           
           down.x <- data.table(Kinase = x['Kinase'],
                              test       = 'down',
                              odds.ratio = (down.ft$estimate),
                              pval       = down.ft$p.value)
           
           x <- rbind(up.x, down.x)
           x[, adj.pval := p.adjust(pval, method='BH')] # correction for n=606 tests

           return(x)
           }))

```


3b) merge DFs & calculate frequency factors & odds ratio

```{r}
ftest.out <- ftest.out[cont.tab, on=c(Kinase = 'Kinase')] %>%
  #.[test == dominant_direction,] %>% # only keep dominant direction data
  .[, `:=`(frequency_factor = ifelse(test == 'up', upregulated_set_hits/background_set_hits, downregulated_set_hits/background_set_hits),
           odds_ratio       = ifelse(test == 'up', (upregulated_set_hits/upregulated_set_size)/(background_set_hits/background_set_size),
                                     (downregulated_set_hits/downregulated_set_size)/(background_set_hits/background_set_size)) )]

```


Plot to inspect the results
Plot upregulated sites, plot downregulated sites and plot volcano
Generate a volcano plot of the significant results
````{r}

ftest.out[,sig := fcase(
  adj.pval > 0.05, 'not',
  adj.pval < 0.05 & frequency_factor > 0, 'up',
  adj.pval < 0.05 & frequency_factor < 0, 'down'
)]

ggplot(ftest.out[dominant_direction == test,], aes(x=log2(frequency_factor), y=-log10(adj.pval))) +
  geom_point(aes(color=sig)) +
  geom_hline(yintercept = -log10(0.05), linetype="dotdash", alpha=0.4) +
  geom_vline(xintercept = c(-0.58, 0.58), linetype="dotdash", alpha=0.4) +
  geom_text_repel(aes(label=Kinase), size=3) +
  theme_classic() +
  coord_cartesian()

```




Structure below: kinase as explanatory variable

            up-reg not-reg
            ------ --------
kinaseA
All


Difference in how we compare hits; they basically take count of all sig up and sig down regules
Hits v size?? maybe try their way and see if we reocver same results; sig below right but background hits will be up + down and N for both will be all background

k.up and k.down the same, but background up is sum N when sig != 'up'


3a) one-sided Fisher’s exact test for a p value. (Repeat for down-regulated vs non-regulated) (or log odds ratio)
- Do Fishers Exact Test

I think makes more sense here to wirte a function to perform fishers exact test and simplify an output and add a id col ('up' or 'down'), fdr correction, then rbind the two dfs 
Then group_by kinase, and check which direction is larger

merge info from the count table; use this information to calculate the odds ratio (think FET uses MLE to estimate, so not the same..)


Steps
Compared the percentage of phosphorylation sites for which each kinase was predicted among the upregulated/downregulated (increased/decreased, respectively) phosphorylation sites (sites with |log2[fold change]| equal or greater than the log[fold change] threshold), versus the percentage of biochemically favoured phosphorylation sites for that kinase within the set of unregulated (unchanged) sites in this study (sites with |log2[fold change]| less than the log[fold change] threshold). 



The log-transformed fold change threshold was determined to be 1.5 for all panels in Fig. 4, except for Fig. 4e, in which the threshold was set to 0.5 owing to the low range of the log[fold change] in the data. Contingency tables were corrected using Haldane correction (adding 0.5 to the cases with zero in one of the counts). Statistical significance was determined using one-sided Fisher’s exact tests, and the corresponding P values were adjusted using the Benjamini–Hochberg procedure. Kinases that were significantly enriched (adjusted P ≤ 0.1), or depleted (log2[frequency factor] < 0) for both upregulated and downregulated analysis were excluded from downstream analysis. Then, for every kinase, the most significant enrichment side (upregulated or downregulated) was selected on the basis of the adjusted P value and presented in the volcano plots

DB Prep Stuff..
-----

Taking top 15 ranked.. ties an issue so go with percentile..
```{r}
kinase.db <- fread('~/Documents/utils/mg_utils/data/kinase.scores.long.csv.gz')

#kinase.db <- kinase.db %>% 
#  .[Rank %between% c(1,15), .(Kinase,Rank), by=Protein.Site] %>% 
#  .[order(Protein.Site,Rank),]


#fwrite(kinase.rk.db, '~/Documents/utils/mg_utils/data/kinase.ranks.long.csv.gz')
```

testing fishers exact test 
```{r}
FisherbyRow <- function(counts.tab,          # cols below to use to create matrix - c(a,b,c,d) ORDER IMPORTANT! See diagram above
                        countCols = c('upregulated_set_hits','upregulated_set_size','background_set_hits','background_set_size'),
                        return.OR = TRUE,
                        alt = 'greater',
                        cnf = 0.95) {
  
  mat <- matrix(as.numeric(counts.tab[countCols]), ncol=2, byrow=T)
  
  f <- fisher.test(mat, alternative = alt, conf.level = cnf)
  
  if (return.OR == TRUE) {
    return(c(counts.tab,
             pval = f$p.value,
             odds.ratio = f$estimate[[1]],
             or.lower = f$conf.int[1],
             or.upper = f$conf.int[2]))
  } else {
    return(c(counts.tab,
             pval = f$p.value))
  }
}


```


## Testing function
# now recovering about 150 more when using aa rather than site so replace (add info to script), but are these correctly added?
# how to improve? look through the ones we dropped; are they in the database? proteins are but the ph site is different... do we allow error tolerant matchin?


```{r test-functions}
out <- kinaseEnrichmentonMSSResults(test.ph, top.sites = 15)
kinaseEnrichmentVolcano(out, pvalCol = 'adj.pval', plot_dominant_only = T)

```


## Updates

Testing splitting multi-ph sites to one per line

```{r}
test <- test.ph

# create new row for each single ph site
t <- test[, .(Label, log2FC, pvalue, adj.pvalue, singProtein = unlist(strsplit(Protein, ";"))), by=.(Protein)] %>% 
  .[, .SD[which.max(abs(log2FC))], by=singProtein] %>% 
  .[, Protein := singProtein] %>% 
  .[, singProtein := NULL]

t

expandMultiPtmSites <- function(dt, cols=c()){
  
  dt
  
  
}
  
          lapply (unique) %>% #removes any duplicate elements in list
          lapply (paste, collapse = ";") %>% #colapses back to unique prot names
          unlist]


test
unlist(strsplit(test$Protein,";"))
```


Idnetify site and merge on peptide string matching rather than uniprot+site.pos
Losing approx 10% of our data using our method, so modify to mirror app approach; expand size size of flanking region to match substr?

First; read in our data, merge with spec to recover ph site and simplify string (remove mods)

```{r updates}
# read in test data 
## lets read in the raw data for this
mss.ph <- fread('/Users/martingordon/Documents/projects/060123_DSwaney_AshworthPH/060123_DSwaney_AshworthPH_data/2023_06_01_msstats_groupcomparisons.csv.gz') 

test.ph <- mss.ph %>% 
  #.[Label %chin% c('WT+ vs EV+', 'WT+ vs N10t20+') & abs(log2FC) > log2(1.5) & adj.pvalue < 0.05 & !issue %like% 'Condition' ] %>% 
  .[Label %chin% c('WT+ vs EV+'),]
 # .[ Protein %like% "_" & !Protein %like% ';' ,] #

#write function to string match the kinase DB and our sites

#fix error...
test <- preparePHplusTable(test.ph, spec.f = spec.ph)
test.ph

View(test)
```
read in kinase db with AA string info

```{r}
kinase.db <- fread('~/Documents/utils/mg_utils/data/kinase.scores.csv.gz')
colnames(kinase.db)
# drop unnecessary cols form kinase.db for now; just keep IDs, kinase_ranks

cols.noi <- c('Database Uniprot Accession', 'Uniprot Entry','Gene', 'Alternative Gene Names', 'Protein',
              'Description', 'Phosphosite', "median_percentile","promiscuity_index",'Kinase','Database', 'phos_res', grep("*rank*",colnames(kinase.db), value=TRUE))
kinase.db <- kinase.db[, !..cols.noi]


kinase.db <- melt(kinase.db, id.vars = c('Uniprot Primary Accession','Protein.Site','SITE_+/-7_AA'))

head(kinase.db)
colnames(kinase.db) <- c('Uniprot', 'Protein.Site', 'Neg7_Site_Pos7', 'Kinase', 'Percentile')
kinase.db[, Kinase := sub('_percentile', '', Kinase)]


#not neecded
#kinase.db[, site := gsub("^(.{8})(.*)$", "\\1(ph)\\2", 'SITE_+/-7_AA')] # insert ph at phosphorylated site

# write out all kinase results
#fwrite(kinase.db,'~/Documents/utils/mg_utils/data/kinase.percentiles.sites.long.csv.gz')

head(kinase.db)
## not needed....

#kinase.db
#write out all kinase percentile scores
kinase.db <- fread('~/Documents/utils/mg_utils/data/kinase.percentiles.sites.long.csv.gz')
head(kinase.db)
#enriched kinases -top 15 per site
# fixed as based on available DB

#kinase.db.prc <- kinase.db %>% 
#  .[order(Protein.Site,-Percentile),] %>% 
#  .[, head(.SD, 15), by = Protein.Site] %>% 
#  .[,.(Protein.Site,Kinase,Percentile)]


```

Read in the kinase DB and perform string mathching
First we need to acquire a similar site info based on 

Don't think I need spec.ph; if I have uniprot and site I can extract the sequence

```{r}
# read in test data 
## lets read in the raw data for this
mss.ph <- fread('/Users/martingordon/Documents/projects/060123_DSwaney_AshworthPH/060123_DSwaney_AshworthPH_data/2023_06_01_msstats_groupcomparisons.csv.gz') 

test.ph <- mss.ph %>% 
  #.[Label %chin% c('WT+ vs EV+', 'WT+ vs N10t20+') & abs(log2FC) > log2(1.5) & adj.pvalue < 0.05 & !issue %like% 'Condition' ] %>% 
  .[Label %chin% c('WT+ vs EV+'),] %>% 
  .[ Protein %like% "_" & !Protein %like% ';' ,] #

spec.ph <- fread('./data/spec.ph.csv.gz')


# split the sites
test.ph[, c("Uniprot","Site") := tstrsplit(Protein, "_")]
test.ph[, pos := as.numeric(sub("^[STY]{1}", "", Site))]

# lod uniprot mapping
uniprots <- loadUniprots(test.ph$Uniprot, fastaFile = "~/Documents/utils/mg_utils/data/human_all_proteins_canonical_uniprot-proteome_UP000005640.fasta.gz")

# match on our msout file
test.ph[uniprots, sequence := i.sequence, on=c(Uniprot = 'uniprot')]

# create a matching sequence for the uniprots
test.ph[, `:=`(aa.lflank = substr(sequence, pos-7, pos-1),
               aa.rflank = substr(sequence, pos+1, pos+7) )]

test.ph[, `:=`(aa.lflank = sapply(aa.lflank, FUN = function(x) addFlankingSites(x, n.char=7, char='_', side='left')),
               aa.rflank = sapply(aa.rflank, FUN = function(x) addFlankingSites(x, n.char=7, char='_', side='right')) )]


# create the character vector
test.ph[, ph.pos.aa.flank := paste0(aa.lflank, gsub('[0-9]+', '', Site), aa.rflank, sep='')]


test <- classifySites(test.ph)

# using uniprot ID and reference fasta file, extract flanking sequences (7 n characters default)
getFlankingSequence <- function(spec.f, 
                                fastaFile = "~/Documents/utils/mg_utils/data/human_all_proteins_canonical_uniprot-proteome_UP000005640.fasta.gz", 
                                n.char=7){
  
  #get site uniprot id
  spec.f[, c("Uniprot","Site") := tstrsplit(Protein, "_")]
  spec.f[, pos := as.numeric(sub("^[STY]{1}", "", Site))]
  
  uniprots <- loadUniprots(spec.f$Uniprot, fastaFile = "~/Documents/utils/mg_utils/data/human_all_proteins_canonical_uniprot-proteome_UP000005640.fasta.gz")
  
  # match on our msout file
  spec.f[uniprots, sequence := i.sequence, on=c(Uniprot = 'uniprot')]
  
  # create a matching sequence for the uniprots
  spec.f[, `:=`(aa.lflank = substr(sequence, pos-n.char, pos-1),
                aa.rflank = substr(sequence, pos+1, pos+n.char) )]
  # add flanking char for shoter sequences
  spec.f[, `:=`(aa.lflank = sapply(aa.lflank, FUN = function(x) addFlankingSites(x, n.char=n.char, char='_', side='left')),
                aa.rflank = sapply(aa.rflank, FUN = function(x) addFlankingSites(x, n.char=n.char, char='_', side='right')) )]
  
  # concatenate flanking seqs and site
  spec.f[, neg7_ph_pos7 := paste0(aa.lflank, gsub('[0-9]+', '', Site), aa.rflank, sep='')]
  
  message("Checking ph sites...")
  if (any(!spec.f[, substring(neg7_ph_pos7,8,8)] %in% c('S','T','Y'))){
    stop(paste0("Error; wrong aa residue phosphorylated. Exiting.."))} else {
      message('Check complete')
    }
  
  return(spec.f[, .(Label, Protein, neg7_ph_pos7, sig)])
}

test <- getFlankingSequence(test, fastaFile = "~/Documents/utils/mg_utils/data/human_all_proteins_canonical_uniprot-proteome_UP000005640.fasta.gz")

#sanity check
any(!test[, substring(neg7_ph_pos7,8,8)] %in% c('S','T','Y'))

```

Combine our sites with the kinase.db

```{r}
kinase.raw <- fread('~/Documents/utils/mg_utils/data/kinase.scores.csv.gz')
kinase.raw[`Database Uniprot Accession`  ==  'Q8IWC1',] %>%  .[order(Phosphosite)]
colnames(kinase.raw)

Q9UIF9_S137, Q9ULL5_S915, Q8IWC1_S429

kinase.db <- fread('~/Documents/utils/mg_utils/data/kinase.percentiles.sites.long.csv.gz')
head(kinase.db)

top.kinases <- kinase.db %>% 
    .[order(Protein.Site,-Percentile),] %>% 
    .[, head(.SD, 15), by = Protein.Site] %>% #select n kinases per site
    .[,.(Protein.Site,Neg7_Site_Pos7,Kinase,Percentile)]

#keep only those in DB... dropping multi-PH sites
message('dropping multisites as not present in DB...')


mapper <- top.kinases[test, ,on = c("Protein.Site" = "Protein")] %>% 
  .[order(Protein.Site)]


# sanity check.. Y when we use rank as there are ties
mapper[, .N, by=.(Neg7_Site_Pos7)]$N

# check loss of sites
start.sites <- test[, .N,by=Protein] %>% nrow()
end.sites <- mapper[, .N, by=.(Protein.Site)] %>% nrow()

sprintf("% s of % s sites detected in kinase DB. This function does not currently handle multi-phosphorlyated peptides. Continuing... ", end.sites, start.sites) 
# TODO break down discarded sites
#maybe include a list of PH sites not analysed ( and seperately the multisites; print to screen as error msg.)


message('merging...')
# merge exp data each site; top 15 kinases per site
# good; 15 kinases per site in data
mapper[class.ph, `:=`(sig = sig, 
                   upregulated_set_size = upregulated_set_size, 
                   downregulated_set_size = downregulated_set_size,
                   background_set_size = background_set_size), on=c(Protein.Site = "Protein")] %>% 
  .[order(Protein.Site,-Percentile)]

```


















Not needed 
-----

Top hits for total data and subset 
Discrepancies in the counts. More hits in the apps, although trends seem consistent
```{r}
app.out <- fread('./docs/enrichment-analysis-result-table.txt')

head(app.out[order(kinase),])
head(class.ph)

app.sub.out <- fread('./docs/enrichment-analysis-result-table.subset.txt')

app.sub.out[order(kinase)]
```
Create phosphosite per label for testing (output the same for both tools)
Same input for app (processPHTable) & script

```{r}
test.ph <- mss.ph %>% 
  #.[Label %chin% c('WT+ vs EV+', 'WT+ vs N10t20+') & abs(log2FC) > log2(1.5) & adj.pvalue < 0.05 & !issue %like% 'Condition' ] %>% 
  .[Label %chin% c('WT+ vs EV+') & issue == '',] %>%
  .[ Protein %like% "_" & !Protein %like% ';' ,] 

# now fine as long as correct cols specified 
DT <- processPHTable(test.ph,spec.ph, pvalCol = 'adj.pvalue')

#fwrite(DT[, .(site,log2FC,pvalue)], './docs/kinase.enrich.test.tsv', sep = '\t')

# test subset
#fwrite(DT[1:2000, .(site,log2FC,pvalue)], './docs/kinase.enrich.test.subset.tsv', sep = '\t')

sub.ph <- test.ph[1:2000,]
```


Lets look at the processed table to see if we can find the issue
Work back;
first identify number of upregulated sites(total)
then number of upreg sites of a specific kinase (find the upreg sites and count N times that kinase is in top 15 -AAK1 for test)

N uprg sites: 2168 (adj.pval) #for both DT and test
Same number of upreg/downreg sites as our pipeline detected, seems the dataset may be different for app and ours

Our counts look ok, so this may be an issue with our DB setup....
not pval threshold.. investigate the output of app run, then subset dt used as input and we can find the error

```{r}
DT[ abs(log2FC) > log2(1.5),.N]
up.sites <- sub.ph[adj.pvalue <= 0.05 & log2FC >= .58,]$Protein

up.sites # 262 sites

# now for the kinase.db, find out how many times AAK1 is in the top 15 for these
kinase.db[Protein.Site %in% up.sites,]
up.sites

#this is our output.. so for AAK1 that was found in 220 of the upreg sites
kinase.db.prc[Protein.Site %chin% up.sites & Kinase == 'AAK1',]  #& Per %between% c(1,15),.N]

head(kinase.db)


kinase.db[,.N, by=.(Protein.Site)] %>%  View()


kinase.db.prc
```

No needed merging sites in the file to log2FC results...

```{r}

#split each; and unlist, for each combo of log2FC, pval, adj.pval etc
# need to split by label!
#test.ph[, .(Protein = unlist(tstrsplit(Protein, ";"))), by=.(Label, log2FC, pvalue, adj.pvalue, sig)]
test.ph
# for now just drop multi-labels - 7 sites in test
test.ph <- test.ph[ !Protein %like% ';',]

# merge exp data each site (303n per site, 1 per kinase)
mapper <- kinase.db[test.ph, `:=`(log2FC     =    log2FC,
                                  pvalue     =    pvalue,
                                  sig        =    sig,
                                  adj.pvalue = adj.pvalue), on=c(Protein.Site = "Protein")] %>% 
  .[!is.na(log2FC),] %>%  #subset to sites only detected in our exp
  .[order(Protein.Site)]

#extract the top ranked 1-15 kinases (allow user to refine threshold as desired)
# keep rows with top 15 sites
mapper <- mapper[Rank %between% c(1,15), .(Kinase, Rank,log2FC,pvalue,adj.pvalue,sig), by=Protein.Site] %>% 
  .[order(Protein.Site,Rank)]


#extract the top ranked 1-15 kinases (allow user to refine threshold as desired)
# keep rows with top 15 sites
#top.n.K <- mapper[Rank %between% c(1,15), .(Kinase, Rank,log2FC,pvalue,adj.pvalue,sig), by=Protein.Site] %>% 
 # .[order(Protein.Site,Rank)]


###

```

counting frequencies 

```{r}
#%>% 
#  .[, `:=`(upregulated_set_size   = sum(site.sig == 'Up'),
           downregulated_set_size = sum(site.sig == 'Down'),
           background_set_hits    = sum(site.sig != 'Not'),
           background_set_size    = sum(site.sig == 'Not')), by=.(Protein.Site) ]

 # .[, dominant_set := ifelse(k.up > k.down, 'up', 'down')] %>% 
#  .[, `:=`(N = NULL,
 #          sig = NULL)] %>% #drop unnecessary cols
 # unique()

cont.tab[Kinase == 'ZAK',]
# when calculating their background correct odds ratio appears
 # log2((17/1313)/(89/10280))


counts <- cont.tab  <- top.n.K[, .N, by=.(Kinase,sig)] %>% 
  .[, `:=`(k.up   = sum(N[sig == 'Up']),
           k.down = sum(N[sig == 'Down'])
           k.not  = sum(N[sig == 'Not'])), by=.(Kinase) ] %>% 
  .[, `:=`(other.up   = sum(k.up) - k.up,
           other.down = sum(k.down) - k.down,
           other.not  = sum(k.not) - k.not) ] %>% 
  .[, dominant_set := ifelse(k.up > k.down, 'up', 'down')] %>% 
  .[, `:=`(N = NULL,
           sig = NULL)] %>% #drop unnecessary cols
  unique()

```

