---
title: "230623_QLi_DARPA"
author: "Martin Gordon"
date: "2023-06-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary
Perform Differential Expression Analysis of MS-MS Spectronaut data using MSstats
Data from DARPA project: Single timepoint with 6 conditions in total: 3 ligands vs NPFFR1, treated vs untreated with 30 min exposure to NPFF
Perform pairwise contrasts for each of the conditions

###: Contact: Qiongyu Li
###: Date: 23-06-23

```{r packages}
library(purrr)
library(tidyr)
library(stringr)
library(dplyr)
library(data.table)
library(MSstats)
library(ggplot2)
library(magrittr)
library(circlize)
library(ComplexHeatmap)
library(parallelDist) #needed for some custom enrichment functions
library(ggvenn)

#source Bens utility scripts
source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/MSstats_Helper_Functions.R")
source("/Users/martingordon/Documents/utils/bp_utils/MSstats_V4_Functions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
```

Read in spectronaut data & metadata
----
```{r read-in}

input <- fread('./data/20230619_090015_DAR32_dirDIA_EXc_15cm_MSStatsFormatReport.tsv')
meta <- fread('./data/DAR32_dirDIA_EXc_15cm_ConditionSetup.tsv')
```



```{r data-inspection}
input[,BioReplicate := as.factor(BioReplicate)]
input[,Run := as.factor(Run)]
input[,Condition := as.factor(Condition)]
```

Data inspection
---------

```{r data structure}
# order by Condition
setorder(input, Condition)

# One run per condition per replicate
input[, .N, by= .(Run,Condition, BioReplicate)]
```
Bioreplicate numbers (1-3) are repeated across conditions. Modified as msstats assumes repeated bioreplicate are processed as timeseries data
Asigned each cond/biorep combo a unique ID (1-20) for msstats to model as a case-control experiment.

```{r}
# create new ID col for pairwise comparisons
input[, ID:=.GRP, by=.(Condition,BioReplicate)]


#reassign to biorep col and drop ID
input[, BioReplicate:=ID]
input[, ID:=NULL]

#sanity check
input[, .N, by = .(Condition,BioReplicate)]
```

Removed spectronaut peptides with intensities < 2^5 as likely artifacts of the identification/quantification tool.
There is a lot of low intensity junk; might be worth notifying Qiongyu
Looks like a bimodal distribution; what is that on left hand shoulder? Present in all groups

Imputation turned on?

```{r}
ggplot(data=input,  mapping=aes(x=log2(Intensity))) +
  geom_histogram(binwidth = 0.1) + 
  facet_wrap(~Condition)

input <- input[log2(Intensity) > 5] #filter out low abundance 

ggplot(data=input,  mapping=aes(x=log2(Intensity))) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~Condition)


print(input)
```
Scatterplots of the intensities and compare samples
```{r}
input.wide <- dcast(input, ProteinName+PeptideSequence+PrecursorCharge~Condition+BioReplicate, value.var = 'Intensity')

ggplot(input.wide, aes(x=log2(AT15_NP00_7), y=log2(AT15_NP00_9), color=c('blue'))) + 
  geom_point()

```
Identify if there any multiple-feature peptide ions in the runs. We expect one peptide feature per row per run, so identify duplicates.
No multi-feature peptides detected

```{r}
any(input[, .N, by = .(Run, PeptideSequence, PrecursorCharge)]$N) > 1
```
Plot of number of peptide ion measurements per run. Counts seem pretty consistent across replicates & conditions

```{r}
### plot of peptide ions per run
p <- input %>%  
      ggplot(mapping=aes(x=reorder(interaction(Condition,BioReplicate)), fill=Condition)) +
      geom_bar() + 
      theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

BackupAsPDF(p, "BarPlot_NPeptides_")
```
Boxplots of distributions of raw intensity values per condition/replicate. Again abundances are pretty consistent and no real concerns with raw data quality. 
```{r create contrasts}
p <- ggplot(data = input, mapping=aes(x=reorder(interaction(Condition, BioReplicate)), y=log2(Intensity), fill=Condition)) +
      geom_boxplot() +
      theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

BackupAsPDF(p, "BoxPlot_RawIntensities_")

```


Data Processing
----
Convert spectronaut data to msstats format
```{r msstats-formatting}
ab.mss <- specFileToCompleteMSstats(input)
ab.mss[, IsotopeLabelType := 'L'] #label free
setDF(ab.mss)  # convert to DF for msstats processing

```
MSStats DataProcess: Normalise (equilizemedians), filter & summarize peptides to protein intensities

```{r mss-dataproc}

dp.out <- MSstats::dataProcess(ab.mss,
                               normalization = 'equalizeMedians',
                               summaryMethod="TMP",
                               featureSubset = "highQuality",
                               remove_uninformative_feature_outlier = TRUE,
                               MBimpute = FALSE )
```
Write out protein level data to file (provide feature level data also)
```{r writeout-data}
# the main data of interest is this one:
fwrite(dp.out$ProteinLevelData, ScriptAndDatedFileName("ProteinLevelData.csv.gz"))

#this table is mostly a copy of the input with some additional columns for hte post-normalization intensity (labeled ABUNDANCE), and columns for is_outlier and is_informative
fwrite(dp.out$FeatureLevelData, ScriptAndDatedFileName("FeatureLevelData.csv.gz"))
```
```{r}
prot.quant <- setDT(dp.out$ProteinLevelData)

p <- prot.quant %>% 
      ggplot(mapping=aes(x=interaction(GROUP,SUBJECT), y=LogIntensities, fill=GROUP)) +
      geom_boxplot() +
      theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))
p
BackupAsPDF(p, "BoxPlot_NormalizedProteinIntensities_")
```

Possibly might need another round of normalisation. Run some EDA before deciding

Looking at log2 intensities can't see a lot of smaple clustering

```{r}

  rowClusterWithNA <- function(mat, na.value = 0, ...){
    mat[is.na(mat)] <- na.value
    hclust(dist(mat), ...)
  }



intensity.mat <- as.matrix(dcast(prot.quant, Protein ~GROUP+SUBJECT, value.var = "LogIntensities"),
                           rownames = "Protein")
  
subMat <- intensity.mat[sample.int(nrow(intensity.mat), 2000),]


colnames(subMat)
unique(sub("_[0-9]{1,2}$", "", colnames(subMat)))
#create heatmap annotation col
type <-sub("_[0-9]{1,2}$", "", colnames(subMat))
ha <- HeatmapAnnotation(type = type, annotation_name_side = "right", col= list(type = c('AT00_NP00'='red', 'AT00_NP30' = 'blue', 'AT15_NP00'='green', 'AT15_NP30'='orange', 'AT22_NP00'='cyan', 'AT22_NP30'='purple', 'AT56_NP00'='yellow',  "AT56_NP30"='pink')))

  #generate heatmap
  hm <- Heatmap(subMat,
        #clustering_distance_rows = function(m) dist(m, method = 'euclidean'),
        #clustering_method_rows= "complete",
        top_annotation = ha,
        cluster_rows = rowClusterWithNA(subMat),
        show_row_names = FALSE,
        row_title = sprintf("%d subsampled proteins", nrow(subMat)),
        na_col = 'grey',
       # column_title = paste0(' heatmap'),
        name = "Log2 Int.")
  hm
  BackupAsPDF(hm,"heatmap_subsample_log2Intensities_")

```
Subtract median intensies per run and re-plot

```{r}
intensity.mat <- sweep(intensity.mat, 1, apply(intensity.mat, 1, median, na.rm=T)) 

  
subMat <- intensity.mat[sample.int(nrow(intensity.mat), 2000),]

#generate heatmap
hm <- Heatmap(subMat,
        #clustering_distance_rows = function(m) dist(m, method = 'euclidean'),
        #clustering_method_rows= "complete",
        top_annotation = ha,
        cluster_rows = rowClusterWithNA(subMat),
        show_row_names = FALSE,
        row_title = sprintf("%d subsampled proteins", nrow(subMat)),
        na_col = 'grey',
       # column_title = paste0( heatmap'),
        name = "Log2 Int.")


BackupAsPDF(hm, "heatmap_subsample_log2Intensities_scaled_")

```

Apply second round of normalisation using TMP and rerun the plots

```{r}
prot.quant[, group_subject := paste(GROUP, SUBJECT, sep = "_")]


 #generate matrix
prot.mat <- dcast(prot.quant, Protein~GROUP + SUBJECT, value.var = "LogIntensities") %>%  #long to wide format dt
  as.matrix (rownames = "Protein")
  
#perform tmp normalisation
tmp <- medpolish(prot.mat, na.rm = TRUE) #iteratively subtract row and col median values 
tmp.dt <- data.table(offset = tmp$col, group_subject = names(tmp$col))

prot.quant[tmp.dt, normalize.offset := i.offset, on = "group_subject"]
prot.quant[, LogIntensities := LogIntensities - normalize.offset]


```

Looks better after second round of normalisation. Write this data to file

```{r}
p <- ggplot(prot.quant, aes(x=reorder(interaction(GROUP,SUBJECT)), y=LogIntensities, fill = GROUP)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=90)) 

p
BackupAsPDF(p, "BoxPlot_TMP_NormalizedProteinIntensities_")
```
Write out protein level data to file (provide feature level data also)
```{r writeout-data}
# the main data of interest is this one:
fwrite(prot.quant, ScriptAndDatedFileName("TMPnorm_ProteinLevelData.csv.gz"))
```

Lets regenerate the clustering and plot PCAs

Seems to be some clustering by NP00 and NP30 group, more visible than agonist clustering anyway

```{r}
intensity.mat <- as.matrix(dcast(prot.quant, Protein ~GROUP+SUBJECT, value.var = "LogIntensities"),
                           rownames = "Protein")
  
subMat <- intensity.mat[sample.int(nrow(intensity.mat), 2000),]


unique(sub("_[0-9]{1,2}$", "", colnames(subMat)))
#create heatmap annotation col
type <-sub("_[0-9]{1,2}$", "", colnames(subMat))

ha <- HeatmapAnnotation(type = type, annotation_name_side = "right", col= list(type = c('AT00_NP00'='red', 'AT00_NP30' = 'pink', 'AT15_NP00'='green', 'AT15_NP30'='purple', 'AT22_NP00'='cyan', 'AT22_NP30'='blue', 'AT56_NP00'='yellow',  "AT56_NP30"='orange')))

  #generate heatmap
  hm <- Heatmap(subMat,
        #clustering_distance_rows = function(m) dist(m, method = 'euclidean'),
        #clustering_method_rows= "complete",
        top_annotation = ha,
        cluster_rows = rowClusterWithNA(subMat),
        show_row_names = FALSE,
        row_title = sprintf("%d subsampled proteins", nrow(subMat)),
        na_col = 'grey',
       # column_title = paste0(' heatmap'),
        name = "Log2 Int.")
  hm
 BackupAsPDF(hm,"tmpnorm_heatmap_subsample_")
```

Scaled heatmap

```{r}
intensity.mat <- sweep(intensity.mat, 1, apply(intensity.mat, 1, median, na.rm=T)) 

  
subMat <- intensity.mat[sample.int(nrow(intensity.mat), 2000),]

#generate heatmap
hm <- Heatmap(subMat,
        #clustering_distance_rows = function(m) dist(m, method = 'euclidean'),
        #clustering_method_rows= "complete",
        top_annotation = ha,
        cluster_rows = rowClusterWithNA(subMat),
        show_row_names = FALSE,
        row_title = sprintf("%d subsampled proteins", nrow(subMat)),
        na_col = 'grey',
       # column_title = paste0( heatmap'),
        name = "Log2 Int.")


BackupAsPDF(hm, "tmpnorm_heatmap_subsample_scaled_")

```
PCA plot
-----
Based on the sample clustering and the above heatmaps, I think its best to drop the following 4 samples: AT15_NP00_9, AT00_NP30_4, AT56_NP30_24, AT15_NP30_12
However we only have 3 replicates per condition so just proceed as is 
```{r pca}
intensity.mat <- as.matrix(dcast(prot.quant, Protein ~GROUP+SUBJECT, value.var = "LogIntensities"),
                           rownames = "Protein")

complete.mat <- intensity.mat[complete.cases(intensity.mat),]

pcaOut <- prcomp(t(complete.mat))
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
#pcaDT[, mainGroup := tstrsplit(rn, "_")[[1]]] #transpose & split
pcaDT[, mainGroup := sub("_[0-9]{1,2}$", "", rn)]

pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)

p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color = mainGroup )) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA using %d proteins (log intensity)", nrow(complete.mat))) 
p

BackupAsPDF(p, "PCA_Complete_Proteins")
```
With the 4 samples dropped we can see clearer clustering of NP00 and NP30 groups along the x axis
```{r}
out.samples <- c('AT15_NP00_9', 'AT00_NP30_4', 'AT56_NP30_24', 'AT15_NP30_12')
complete.mat <- intensity.mat[complete.cases(intensity.mat), !colnames(intensity.mat) %in% out.samples ]

pcaOut <- prcomp(t(complete.mat))
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
#pcaDT[, mainGroup := tstrsplit(rn, "_")[[1]]] #transpose & split
pcaDT[, mainGroup := sub("_[0-9]{1,2}$", "", rn)]

pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)

p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color = mainGroup )) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA using %d proteins (log intensity)", nrow(complete.mat))) 
p

BackupAsPDF(p, "PCA_RmOutliers_Proteins")
```
Drop outliers
------
Drop the outlier samples and visualise

```{r}
prot.quant <- prot.quant[!group_subject %in% out.samples,]
fwrite(prot.quant,ScriptAndDatedFileName("ProteinLevelData_rmoutliers.csv.gz"))
```



Differential Expression Analysis
------

Perform DEA with provided contrasts
Generate a contrast matrix for requested pairwise comparisons of all groups:

```{r make contrasts}

#all pairwise; need to customise
all.dp.contrasts <- MSstatsContrastMatrix("pairwise", unique(dp.out$ProteinLevelData$GROUP))
all.dp.contrasts
#np0 agonists
#all.dp.contrasts[c(1,15,17,24),]
#NP30 vs NP30
#all.dp.contrasts[c(20,22,27),]

dp.contrasts <- all.dp.contrasts[c(1,2,4,6,14,15,17,24,20,22,23,27,28),]
dp.contrasts <- rbind(dp.contrasts, c(-1, 1,0,0,0,0,0,0))
rownames(dp.contrasts)[14] <- 'AT00_NP30 vs AT00_NP00'
dp.contrasts <- rbind(dp.contrasts, c(0, 0,0,0,0,0,-1,1))
rownames(dp.contrasts)[15] <- 'AT56_NP30 vs AT56_NP00'
dp.contrasts <- rbind(dp.contrasts, c(0, 0,0,0,-1,1,0,0))
rownames(dp.contrasts)[16] <- 'AT22_NP30 vs AT22_NP00'
dp.contrasts <- rbind(dp.contrasts, c(0, 0,-1,1,0,0,0,0))
rownames(dp.contrasts)[17] <- 'AT15_NP30 vs AT15_NP00'

dp.contrasts <- dp.contrasts[c(2,3,4,6,7,8,9,10,12,14,15,16,17),]

dp.contrasts <- rbind(dp.contrasts, c(0, -1,0,1,0,0,0,0))
rownames(dp.contrasts)[14] <- 'AT15_NP30 vs AT00_NP30'
dp.contrasts <- rbind(dp.contrasts, c(0, -1,0,0,0,1,0,0))
rownames(dp.contrasts)[15] <- 'AT22_NP30 vs AT00_NP30'
dp.contrasts <- rbind(dp.contrasts, c(0, -1,0,0,0,0,0,1))
rownames(dp.contrasts)[16] <- 'AT56_NP30 vs AT00_NP30'

dp.contrasts
```

```{r mss-pairwise}

# use the tpm normlaised data
dp.out$ProteinLevelData <- prot.quant

test.mss <- groupComparison(contrast.matrix=dp.contrasts, data=dp.out)
test.mss.all <- setDT(test.mss$ComparisonResult)

# adjust p values using the Benjamini Hochberg method, which estimates False Discovery Rate
test.mss.all[, fdr := p.adjust(pvalue, method = "BH")] 
#get gene names
test.mss.all[, gene := multiUniprots2multiGenes(as.character(Protein), simplify = TRUE, allowDups = FALSE)]


fwrite(test.mss.all, ScriptAndDatedFileName("msstats_groupcomparisons_rmoutliers.csv.gz"))
```

Generate Volcano plots for each of the pairwise comparisons

```{r multi-volcanoplot}
## volcano plots

# define significant proteins
test.mss.all[, sig := "Not"]
test.mss.all[fdr < 0.05 & abs(log2FC) > log2(1.5),sig := ifelse (log2FC  > 0, "Up", "Down")]

p <- ggplot (test.mss.all, aes (x = log2FC, y = -log10(fdr), color = sig, label = gene)) +
  geom_point() +
  scale_color_manual(values = c(Not = "gray", Down = "blue", Up = "red")) +
  ggrepel::geom_text_repel(data = test.mss.all[sig != "Not"]  ,  size = 2, max.overlaps = 20) +
  theme_bw() +
  facet_wrap(~ Label,)
p <- p + theme(strip.text.x.top = element_text(size=6))
BackupAsPDF(p, "norm_tmp-volcanoplot_multipanel")

test.mss.all[, gene %in% "HGS"]
prot.quant[Protein == 'O14964', .(LogIntensities,GROUP,SUBJECT)]
prot.quant
```


Over plotting. Generate each volcano plot separately.
```{r split-dt}
#split datatable by comparison

test.mss.list <- split(test.mss.all, test.mss.all$Label)
```

```{r cal-geommean}
#calculate geomemtric mean
test.mss.list <- test.mss.list %>% 
  map( ~{
    dt <- .x
    dt$gm.pval.log2fc <- sqrt(dt[,'fdr']*abs(dt[,'log2FC']))
    
    return(dt)
  })

```

```{r volcanoplots}
# use purr to plot each list element 


  imap(test.mss.list, ~ {
    dt <- .x
    dt.name <- .y
    
    p <- ggplot (dt, aes (x = log2FC, y = -log2(fdr), color = sig, label = gene)) +
      geom_point() +
      scale_color_manual(values = c(Not = "gray", Down = "blue", Up = "red")) +
      ggrepel::geom_text_repel(data = dt[sig != "Not"] ,  size = 2, max.overlaps = 20) +
      ggtitle(paste0(dt.name, ' Volcano plot')) +
      theme_minimal()
    p
   BackupAsPDF(p, paste0(dt.name,"_rmoutliers_volcanoplot"))
  })
```


Enrichment Heatmaps for the different comparisons

Perform enrichment analysis for each of the comparisons
```{r}
#creat new factor col of combined label sig info
test.mss.all[, enrich_grp:= interaction(Label,sig)]

# all go ontology genes (biological process, cellular componetn & metabolic function) for enrichment analysis
gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")

# all identified genes as the background set
universe <- unique(test.mss.all$gene)

# perform enrichment on significant genes for all groups
enrich.dt <- enricherOnGroups(test.mss.all[test.mss.all$sig != 'Not'], groupColumns = 'enrich_grp', term2gene.gmt = gmt.go, universe = universe)

#simplify enrich results to remove redundant GO terms
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')


# failed.. troubleshoot
#enrichmentOnGroupsPL(test.mss.all[test.mss.all$sig != 'Not'], geneColumn = 'gene', groupColumns = 'enrich_grp', gmt = gmt.go, universe = universe)

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, NULL, groupColumn = 'enrich_grp', topN = 8, negCols =unique(simp.enrich$simplified$enrich_grp[grep('Down', simp.enrich$simplified$enrich_grp)]), row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 4), upperThreshold = 8) 
ht# compare enrichment in the top 10 terms per group
BackupAsPDF(ht, "heatmap_GO_rmoutliers")
```
Barplots of interesting genes

```{r}
genes.IO <- c('SKP1', 'FBXW11', 'CUL1')
contrasts.IO <- c('AT00_NP30 vs AT00_NP00','AT15_NP30 vs AT15_NP00','AT22_NP30 vs AT22_NP00','AT56_NP30 vs AT56_NP00')

p <- ggplot(test.mss.all[gene %in% genes.IO & Label %in% contrasts.IO, ], aes(x=Label, y=log2FC, fill=Label)) +
         geom_col() +
         geom_errorbar( aes(x=Label, ymin=log2FC-SE, ymax=log2FC+SE), width=0.2, colour="black", alpha=0.5, size=0.5) +
         theme_classic() +
         theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 6 )) +
         facet_wrap(~gene, nrow = 3)
BackupAsPDF(p, '_barplot-NP30-vsNP00')

test.mss.all[gene %in% genes.IO & Label %in% contrasts.IO,]

test.mss.all
```
Plot comparisons for all conditions

```{r}

p <- ggplot(test.mss.all[gene %in% genes.IO , ], aes(x=Label, y=log2FC, fill=adj.pvalue < 0.05)) +
         geom_col() +
         theme_classic() +
         theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5)) +
         facet_wrap(~gene, nrow = 3)
p
BackupAsPDF(p, 'barplot-all-contrasts')

```

Also generate heatmap of significant genes (use log2fc values)
sub

```{r}
test.mss.all

#take isg genes found in the four contrasts
sig.genes <- unique(test.mss.all[adj.pvalue < 0.05 & Label %in% contrasts.IO & is.finite(log2FC), gene ])

log2fc.mat <- as.matrix(dcast(test.mss.all[gene %in% sig.genes & Label %in% contrasts.IO, ], gene~Label, value.var = c('log2FC')), rownames = 'gene')

# remove infinite vlaues
log2fc.mat[is.infinite(log2fc.mat)] <- NA


#generate heatmap
hm <- Heatmap(log2fc.mat,
        #clustering_distance_rows = function(m) dist(m, method = 'euclidean'),
        #clustering_method_rows= "complete",
       # top_annotation = ha,
        cluster_rows = rowClusterWithNA(log2fc.mat),
        cluster_columns = FALSE,
        show_row_names = TRUE,
        row_title = sprintf("%d differentialy expressed proteins", nrow(log2fc.mat)),
        na_col = 'grey',
        row_names_gp = gpar(fontsize = 3), 
       column_names_gp= gpar(fontsize = 5),
       # title = 'differentially expressed proteins  NP30 vs NP00 (p.adj < 0.05)',
       # column_title = paste0( heatmap'),
        name = "Log2FC.")
hm
BackupAsPDF(hm, 'heatmap-sig-genes', dimensions = c(8, 13))

```
Raw counts of the BTRC in prot.quant data

```{r}
prot.quant[,  gene := multiUniprots2multiGenes(as.character(Protein), simplify = TRUE, allowDups = FALSE)]


ggplot(prot.quant[gene == 'BTRC'], aes(x=GROUP, y=LogIntensities)) +
  geom_point()

prot.quant[gene == 'BTRC']

rownames(dp.contrasts)
```


07-05-23
Generate a dot plot of the intensities for proteins of interest in each condition. Add a mean bar
Generate heatmaps of log2FC changes for NPFF vs Ctrl and Anatagonists + NPFF vs Antagonists


Dotplot
Read in the data
```{r}
#prot.quant <- fread("./232306_QLi_DARPA_data/2023_06_27_ProteinLevelData_rmoutliers.csv.gz")
prot.quant <- fread("./232306_QLi_DARPA_data/2023_06_26_TMPnorm_ProteinLevelData.csv.gz")


mss <- fread("./232306_QLi_DARPA_data/2023_06_27_msstats_groupcomparisons_rmoutliers.csv.gz")
```

dotplot of genes of interest for the comparisons
```{r}
# check for others of interest
genesOI <- c('BTRC', 'SKP1', 'ARRB2', 'FBXW11', 'STAM2', 'HGS')


prot.quant[, gene := multiUniprots2multiGenes(Protein , species='HUMAN', simplify = T)]

p <- ggplot(prot.quant[gene %in% genesOI,], aes(x=GROUP, y=LogIntensities, color=GROUP)) + 
  geom_point(size=1.5) +
  facet_wrap(~gene, scales = 'free') + 
  stat_summary(fun.data = "mean_cl_normal",
               geom = "errorbar",
               color = 'black',
               width = .4, 
               alpha=0.4) +
  stat_summary(fun = "mean", geom = "point", color='black', alpha=0.4) +
  #stat_summary(fun.y = mean, geom='point') +#  fun.args = list(mult = 1), geom = "pointrange", colour = "red")
  theme_bw() +
  theme(
    axis.text.x = element_text(angle=90, hjust=1, size= 6)
  )

p

#BackupAsPDF(p, 'geneIO.intensities.dotplot.pdf', dimensions = c(10,8))
```

Plot each of these individually

```{r}
for (g in genesOI){
  
 p <- ggplot(prot.quant[gene %in% g,], aes(x=GROUP, y=LogIntensities, color=GROUP, alpha=0.2)) + 
  geom_point() +
  geom_hline(aes(yintercept = mean.Ints,col=GROUP))
 # facet_wrap(~gene, scales = 'free') + 
 # stat_summary(fun.data = 'mean_cl_normal', geom='errorbar') +
  ggtitle(paste0(g, ' Log2Intensity'))
  theme_classic() +
  theme(
    axis.text.x = element_text(angle=90, hjust=1, size= 6)
  )
 
 print(p)
}

```
Generate heatmaps of the LOg2FCs for the same genes

NPFF vs Ctrl

```{r}
ms.mat <- dcast(mss[gene %in% genesOI,],  gene~Label, value.var = c('log2FC'), rownames = 'gene') %>% 
  as.matrix(rownames='gene')

#define order here
contrasts.OI <- c("AT00_NP30 vs AT00_NP00", "AT15_NP30 vs AT15_NP00", "AT22_NP30 vs AT22_NP00",  "AT56_NP30 vs AT56_NP00", "AT15_NP00 vs AT22_NP00", "AT15_NP00 vs AT56_NP00", "AT22_NP00 vs AT56_NP00")

submat <- ms.mat[, colnames(ms.mat) %in% contrasts.OI]

# remove infinite vlaues
ms.mat[is.infinite(ms.mat)] <- NA


#generate heatmap
hm <- Heatmap(submat,
        cluster_rows = hclust(dist(ms.mat)),
        cluster_columns = FALSE,
        column_order = contrasts.OI,
        show_row_names = TRUE,
        col = colorRamp2(c(-4,0, 4), c("blue", "white", "red")),
        na_col = 'grey',
        name = "Log2FC.")
hm
BackupAsPDF(hm, 'heatmap-genesOI')


colnames(ms.mat)

```
```{r}
ms.mat[, colnames(ms.mat) %in% contrasts.OI]
```

08-28-23
----
Read in the list of differentially expressed


```{r}

```

