---
title: "071824_NMF_SAINT_WF"
author: "Martin Gordon"
date: "2024-07-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NMF-SAINT workflow

```{r packages}
library(nnls)
library(ggplot2)
library(data.table)
library(NMF)
library(ggplot2)
library(magrittr)
library(stringr)
library(ComplexHeatmap)
```

read in and tidy data
```{r}
spec.mat <- fread("./data/2024_06_25_PCMI_spectral_count.csv", header=T) %>% 
  as.matrix(rownames='V1')


colnames(spec.mat)

# batch14 contrast mat
batch14.int <- fread('./data/2021Nov_PCMI_emptyGFPbatch14-saint-baits-spc.txt', header = F)
setnames(batch14.int, new=c('IP', 'Bait', 'Condition'))

#batch14 interaxctions
batch14.spc <- fread('./data/2021Nov_PCMI_emptyGFPbatch14-saint-interactions-spc.txt', header = F)
setnames(batch14.spc, new=c('IP', 'Bait', 'Prey', 'Spc'))

# match mat names
batch14.dt[, IP := sub('[.]', '-', IP)]

# remove batch 14
f.mat <- spec.mat[, !colnames(spec.mat) %in% batch14.dt$IP]
```

Filtering out rows with no measurements (why are these in the table?)
```{r}
# filter rows with no values
f.mat <- f.mat[rowSums(f.mat) != 0,] 
# convert 0 to NA
f.mat[f.mat == 0] <- NA
```

Normalization using tmp 
```{r}
# calc scaling factors using prey/rows with < 50% missingness
NormalizationVec <- apply(f.mat, 1, function(x){sum(is.na(x)) < (ncol(f.mat)/2) })

#calculate offset and convert back to linear space
offsetCol <- 2^(medpolish(log2(f.mat[NormalizationVec,]), na.rm=T)$col)

#normalize IPs
f.mat <- sweep(f.mat, 2, offsetCol, "/")

ggplot(reshape2::melt(f.mat), aes(x=Var2, y=log2(value))) +
  geom_boxplot()
```

```{r}
f.mat[is.na(f.mat)] <- 0

# sqrt('completeness')/rowmaxCounts
# weighting 'fullness' of protein detection across grps in conversion to linear scale
SAINTScalingFactor <- sqrt(apply(f.mat, 1, function(x)sum(x!=0)) /
      ncol(f.mat)) / 
      apply(f.mat, 1, max)

scaled.f.mat <- sweep(f.mat, 1, SAINTScalingFactor, "*")

# scaled 0-1
apply(scaled.f.mat,1, max)
```

```{r}
SAINTnmf.out <- NMF::nmf(scaled.f.mat, rank = 6)
```

```{r}
SAINT.BV <- SAINTnmf.out@fit@W

# coeff matrix n * r;
Heatmap(coef(SAINTnmf.out))
# basis vec r * n
Heatmap((SAINT.BV))
```

Now have our NMF basis vectors; assume these capture binding background patterns 
Want to use this as bg for our test set; scale batch 14 to match our matrix

Need to rescale this data

```{r}
batch14.mat <- dcast(batch14.spc, Prey~IP, value.var='Spc')

merged.Batch14 <- merge(batch14.mat, 
                        SAINTScalingFactor |> as.data.table(keep.rownames = TRUE), by.x= "Prey", by.y = "rn") 

rownames(merged.Batch14) <- merged.Batch14$Prey
merged.Batch14$Prey <- NULL

scaled.Batch14 <- sweep(merged.Batch14, 1, merged.Batch14$SAINTScalingFactor, '*')
```

filtering out stuff not found in batch 14

Worry with incorrect scaling is that when identifying the contribution of 
```{r}
# also not sure about this filtering.. d
SAINT.BV <- SAINTnmf.out@fit@W[row.names(scaled.Batch14),]

scaled.Batch14[is.na(scaled.Batch14)] <- 0

SAINT.nnls <- apply(scaled.Batch14, 2, function(coln) nnls(SAINT.BV, coln)$x)


SAINT.nnls


SAINT.BV

scaled.Batch14[,2]


help(nnls)
```

```{r}
dim(scaled.Batch14)


23.2998058 * 

dim(SAINT.Pseudo.Control)
SAINT.Pseudo.Control <- SAINT.BV %*% SAINT.nnls

SAINT.Pseudo.Control %>% View()
SAINT.PC.long <- melt(SAINT.Pseudo.Control, variable.name = "run", id.vars = "rn", value.name = "PC spc") |> setDT()


SAINT.Pseudo.Control[rownames(SAINT.Pseudo.Control) =='Q8NFD5',]

SAINT.PC.long[, c("bait", "replicate") := tstrsplit(Var2, "\\-")]

setnames(SAINT.PC.long, old = "Var1", new = "prey")

setnames(SAINT.PC.long, old = "Var2", new = "run")

SAINT.PC.long$replicate <- NULL
```
 first test 

filter baits used in the SAINT test 
```{r}
preys.txt <- fread('./data/SAINTinput_batches/2021Nov_PCMI_emptyGFPbatch14-saint-preys-spc.txt')

out.int <- fread('./data/merged.Batch14.interactions.txt')
out.bait <- fread('./data/merged.Batch14.Baits.txt')

fwrite(out.int[,.(V2,V3,V1,V4)], './data/reordered.merged.Batch14.interactions.txt', sep='\t', col.names = F)

# writes to output, but need to filter to a smaller set
fwrite(out.int[,.(V2,V3,V1,V4)], './data/minimal.Batch14.interactions.txt', sep='\t')

min.ints <- rbind(out.int[grep('[-]14', V2)],out.int[grep('[-]14', V2, invert=T)][grep('GFP|EmptyVector', V2)])
fwrite(min.ints[,.(V2,V3,V1,V4)], './data/reduced.Batch14.interactions.txt', sep='\t', col.names = F)
```
 Interactions with 0 counts must be removed from file in manual
 
```{r}
out.int[V1 == 'Q14192']
preys.txt[V1 == 'Q14192',]



fwrite(out.int[V4 != 0,.(V2,V3,V1,V4)], './data/reordered.merged.Batch14.interactions.txt', sep='\t', col.names = F)

 
fread('./data/reordered.merged.Batch14.interactions.txt')

# what about if we try this with the other files
fread('./data/SAINTinput_batches/2021Nov_PCMI_emptyGFPbatch14-saint-interactions-spc.txt')
```
 




Considerations: 
need to normalize the batch14 input for comparisons (can reuse median polish proceedure)
Also, I think you have a problem with your reuse of scaling factors for the test batch;these values dont range between 0-1
Rounding: do we want our values to be 0??

Why in 0-1 scale? Beacuse we don't want proteins with different ranges of ints to have similar weight; interested in recovering patterns, but we dont want to test in this space
