---
title: "01-12-23.share.Run1.results"
author: "Martin Gordon"
date: "2023-12-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Share results from the first DEG analysis

share the output from the first DGE analysis


what to share
: allTimepoints.wide.csv ~6pm


```{r}
library(magrittr)
library(data.table)
library(stringr)
library(ggplot2)
library(tximport)
library(ggrepel)
library(DESeq2)
# visualisation
library(ComplexHeatmap)
library(EnhancedVolcano)
# annotation & filtering 
library(AnnotationHub)
library(genefilter) # not available 

#Ben utils
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/enrichmentTestFunctions.R")

source("../../utils/mg_utils/HelperFunctions.R")
```

First pass just share the results from the analysis
```{r}
path <- 'output/JyotiB_results/tables'

# create dt
input <- dir(path=path, pattern = "*_(genes|output)\\.csv$", recursive = T, full.names = TRUE) %>% 
  as.data.table() %>% 
  .[, fname := gsub("output/JyotiB_results/tables/|_(genes|output)\\.csv", '', .)] %>% 
  setnames(new=c('path', 'fname'))
  
in.list <- apply(input[,1],1, fread)
names(in.list) <- input$fname
```
# quick look at the results
```{r}
lapply(in.list, head)
```


need to rerun the filtering and extract everyhting that is significant
filtered out pval = 0, keep these as just means the number is so small it isnt included..

get the full list of significant things
```{r}
full.list <- in.list[grep('_dds', names(in.list))]

full.df <-  rbindlist(full.list, idcol = 'group')
full.df %>%  head()

# filter by geommetric mean of pval and log2FC to get most significant hits
full.df[, sigScore := magSigScore(log2FoldChange, pvalue)]

full.df[, `:=`(timepoint=str_extract(group, '[0-9]{1,2}'),
               comparison=gsub("_[0-9]{1,2}|_[0-9]{1,2}_dds", "", group)
               )]

full.df %>%  dim()

full.list <- split(full.df, full.df$comparison)

full.df[comparison == 'Mock-Brisbane' & symbol == 'TMSB15B']
```
share the complete list of files
identify the duplicate genes, take the largest value and share

```{r}
# recalcualte geom_mean and use use for filtering

lapply(full.list, function(x){


  rm.dups <- x[, .N, by=.(symbol, group)] %>% 
    .[N > 1,]
  
  rm.rows <- x[symbol %in% rm.dups$symbol, .SD[which.min(sigScore)], by=.(symbol, group)]
  rm.rows[, filter.col := paste0(symbol,'.',group,'.', gm.pval.log2fc)]

  x[, filter.col := paste0(symbol,'.',group,'.', gm.pval.log2fc)]
  filtered.x <- x[!filter.col %in% rm.rows$filter.col, ]
  filtered.x <- filtered.x[, filter.col := NULL]

  # now convert to wide format
  wide.dt <- dcast(filtered.x, symbol+entrez+comparison~timepoint, value.var=c('log2FoldChange', 'pvalue', 'padj'))
  
  col.order <- c('symbol', 'entrez', 'comparison',  
                 "log2FoldChange_0",  "log2FoldChange_3",  "log2FoldChange_6",  "log2FoldChange_12",  "log2FoldChange_24",
                 "pvalue_0", "pvalue_3" ,"pvalue_6","pvalue_12",  "pvalue_24",
                 "padj_0","padj_3", "padj_6","padj_12","padj_24" )
  
  
  setcolorder(wide.dt, col.order)

  fname <-  unique(x$comparison)

 fwrite(wide.dt, ScriptAndDatedFileName(paste0(fname, '.allDESeqresults.wide.csv')))
})
```

```{r}

# sanity check 
full.df[comparison == 'Mock-Brisbane' & symbol =='A1BG']
```



go through the significant list of things
create the full list of sig sites

```{r}
sig.list <- lapply(full.list, function(x){
  sig.hits <- x[sig != 'Not',]
  return(sig.hits)
})

lapply(sig.list, function(x){
  
  x[symbol == 'CACNA1S']
})
```

summarise the list of sites and take the max pval

```{r}
# load in functions 
source("../../utils/mg_utils/HelperFunctions.R")

comb.dt <- rbindlist(sig.list)[, sigScore := magSigScore(log2FoldChange,pvalue)]# 

comb.dt[, .N, by=.(symbol, comparison)] %>% 
  .[order(-N)]

```

write out the unfiltered results; will need these for plotting 

```{r}
unfiltered.list <-split(comb.dt, comb.dt$comparison)

lapply(seq_along(unfiltered.list), function(x,n,i){
  fwrite(x[[i]], ScriptAndDatedFileName(paste0(n[[i]], 'sig.genes.unfiltered.csv')))
},x=unfiltered.list, n=names(unfiltered.list))
```


```{r}
filter.dt <- comb.dt[, .SD[which.max(sigScore)], by=.(symbol,comparison)]

# sanity check 
#CACNA1S	Mock-Brisbane	5
comb.dt[symbol == 'CACNA1S' & comparison == 'Mock-Brisbane', ]
filter.dt[symbol == 'CACNA1S' & comparison == 'Mock-Brisbane', ]

# also want to add a column of n sig timepoints
# create a dt of N sig for each gene/contrast, and add this col to the dt
sig.counts <- comb.dt[, .N, by=.(symbol, comparison)]
setnames(sig.counts, new=c('symbol', 'comparison', 'N.SigTimePoints'))

# combine the datatables; does the DT 
# sanity check
merge.dt <- merge(x=filter.dt, y=sig.counts, all.x=T, all.y=T, by=c('symbol', 'comparison'))

# filter the dt, split and merge
filtered.list <- merge.dt[, .(symbol,entrez, comparison,timepoint,log2FoldChange, padj, pvalue, N.SigTimePoints)] %>% 
  split(., .$comparison)

# sanity check 
lapply(filtered.list, function(x){
  
  x[symbol ==  'CACNA1S']
})
```

write out this filtered dt in long format

```{r}
lapply(seq_along(filtered.list), function(x,n,i){
  fwrite(x[[i]], ScriptAndDatedFileName(paste0(n[[i]], 'sig.genes.filtered.csv')))
},x=filtered.list, n=names(filtered.list))
```

convert to wide and share
```{r}
wide.filtered <- dcast(rbindlist(filtered.list), symbol+entrez~comparison, value.var = c('timepoint', 'log2FoldChange', 'padj', 'pvalue', 'N.SigTimePoints'))

fwrite(wide.filtered, ScriptAndDatedFileName('combined.filtered.sig.results.wide.csv'))


wide.list <-  lapply(seq_along(filtered.list), function(x,n,i){
  wide.x <- dcast(x[[i]], symbol+entrez+timepoint+N.SigTimePoints~comparison, value.var = c('log2FoldChange', 'padj', 'pvalue'))
  
  return(wide.x)
 fwrite(x[[i]], ScriptAndDatedFileName(paste0(n[[i]], 'sig.genes.filtered.wide.csv')))
},x=filtered.list, n=names(filtered.list))


#fread(ScriptAndDatedFileName('Mock-Victoriasig.genes.filtered.wide.csv'))

fread(ScriptAndDatedFileName('sig.genes.allTimepoints.csv'))
```
use the list of sig genes to filter to the full list

```{r}

full.sig.dt <- lapply(seq_along(full.list), function(x,y,n,i){
  
  compar.oi <- n[[i]]
  
  sig.genes <- y[[compar.oi]][, unique(symbol)]

  return(x[[i]][symbol %in% sig.genes,])
  
},x=full.list, n=names(full.list), y=sig.list) %>% 
  rbindlist()

fwrite(full.sig.dt, ScriptAndDatedFileName('sig.genes.allTimepoints.csv'))

full.sig.dt[, .N, by=.(symbol, comparison, timepoint)] %>% 
  .[order(-N)]


full.sig.dt[symbol =='PDE11A' & comparison == 'Mock-Brisbane',]


full.df[symbol =='PDE11A' & comparison == 'Mock-Brisbane',]


full.list <- in.list[grep('_dds', names(in.list))]
full.df <-  rbindlist(full.list, idcol = 'group')

full.df$group %>%  unique(0)
full.df[symbol =='PDE11A' & group == 'Mock_0-Brisbane_0_dds',]

sig.df[symbol =='PDE11A' & group == 'Mock_0-Brisbane_0_dds',]
```

now, split to list and convert to wide format before sharing
a couple of gene IDs  multimapped to same gene symbol rm these
```{r}
# case here of mapping a couple of genes symbols to same gene; just take the max score
full.sig.dt[, .SD[which.min(gm.pval.log2fc)], by=.(symbol,group)]

rm.dups <- full.sig.dt[, .N, by=.(symbol, group)] %>% 
  .[N > 1,]
rm.dups

rm.rows <- full.sig.dt[symbol %in% rm.dups$symbol & group %in% rm.dups$group, .SD[which.min(gm.pval.log2fc)], by=.(symbol,group)]
rm.rows[, filter.col := paste0(symbol,'.', group,'.', gm.pval.log2fc)]
full.sig.dt[, filter.col := paste0(symbol,'.', group,'.', gm.pval.log2fc)]

full.sig.dt <- full.sig.dt[!filter.col %in% rm.rows$filter.col, ]
full.sig.dt[, filter.col := NULL]


sig.genes.all <- split(full.sig.dt, full.sig.dt$comparison)
```

```{r}
col.order <- c("symbol","entrez","log2FoldChange_0","log2FoldChange_3","log2FoldChange_6","log2FoldChange_12","log2FoldChange_24",
               "pvalue_0","pvalue_3","pvalue_6","pvalue_12","pvalue_24",
               "padj_0", "padj_3","padj_6","padj_12","padj_24",
               "sig_0", "sig_3", "sig_6", "sig_12", "sig_24")


new.wide.list <- lapply(wide.list, function(x){
  colnames(x) <- gsub("_(Mock|Brisbane|Victoria)-(Mock|Brisbane|Victoria)", "", colnames(x))
  setcolorder(x, col.order)
  return(x)
})

# convert to wide and share
lapply(seq_along(sig.genes.all), function(x,n,i){
 x.wide <-  dcast(x[[i]], symbol+entrez~comparison+timepoint, value.var=c('log2FoldChange', 'sig', 'pvalue', 'padj'))
 print(colnames(x.wide))
 colnames(x.wide) <- gsub("_(Mock|Brisbane|Victoria)-(Mock|Brisbane|Victoria)", "", colnames(x.wide))
 setcolorder(x.wide, col.order)
 

 fwrite(x.wide, ScriptAndDatedFileName(paste0(n[[i]], '.sigGenes.allTimepoints.wide.csv')))
},x=sig.genes.all, n=names(sig.genes.all))
```


Not used
-----



how can we tidy this further? 
Look for repeated genes that are sig across many time points, find the gene with the largest, magSig score (or just log2FC?) and keep this


many hits significant across timepoints
filter out to per gene score taking the max sig hit
```{r}
lapply(sig.list, function(x){
  x[, .N, by=symbol] %>% 
    .[order(-N)]
})
```

calculate magsig score and use this to select top hit
add a col indicating number of other timepoints the gene was significant in 

```{r}
# load in functions 
source("../../utils/mg_utils/HelperFunctions.R")

comb.dt <- rbindlist(sig.list)[, sigScore := magSigScore(log2FoldChange,pvalue)]# 
filter.dt <- comb.dt[, .SD[which.max(sigScore)], by=.(symbol,comparison)]

comb.dt[, .N, by=.(symbol, comparison)] %>% 
  .[order(-N)]

# sanity check 
#CACNA1S	Mock-Brisbane	5
comb.dt[symbol == 'CACNA1S' & comparison == 'Mock-Brisbane', ]
filter.dt[symbol == 'CACNA1S' & comparison == 'Mock-Brisbane', ]




# also want to add a column of n sig timepoints
# create a dt of N sig for each gene/contrast, and add this col to the dt
sig.counts <- comb.dt[, .N, by=.(symbol, comparison)]
setnames(sig.counts, new=c('symbol', 'comparison', 'N.SigTimePoints'))


# combine the datatables; does the DT 
# sanity check
merge.dt <- merge(x=filter.dt, y=sig.counts, all.x=T, all.y=T, by=c('symbol', 'comparison'))

# filter the dt, split and merge
filtered.list <- merge.dt[, .(symbol,entrez, comparison,timepoint,log2FoldChange, padj, pvalue, N.SigTimePoints)] %>% 
  split(., .$comparison)
```


```{r}
lapply(seq_along(filtered.list), function(x,n,i){
  fwrite(x[[i]], ScriptAndDatedFileName(paste0(n[[i]], 'sig.genes.filtered.csv')))
},x=filtered.list, n=names(filtered.list))
```
# collapse list and then share

```{r}
wide.filtered <- dcast(rbindlist(filtered.list), symbol+entrez~comparison, value.var = c('timepoint', 'log2FoldChange', 'padj', 'pvalue', 'N.SigTimePoints'))

fwrite(wide.filtered, ScriptAndDatedFileName('combined.filtered.sig.results.wide.csv'))


wide.list <-  lapply(seq_along(filtered.list), function(x,n,i){
  wide.x <- dcast(x[[i]], symbol+entrez+timepoint+N.SigTimePoints~comparison, value.var = c('log2FoldChange', 'padj', 'pvalue'))
  
  return(wide.x)
 fwrite(x[[i]], ScriptAndDatedFileName(paste0(n[[i]], 'sig.genes.filtered.wide.csv')))
},x=filtered.list, n=names(filtered.list))


colnames(wide.filtered)
```

# first things first, extract the sig list of genes
```{r}
sig.list <- in.list[grep('_sig', names(in.list))]

sig.df <-  rbindlist(sig.list, idcol = 'group')
sig.df %>%  head()

sig.df[, `:=`(timepoint=str_extract(group, '[0-9]{1,2}'),
               comparison=gsub("_[0-9]{1,2}|_[0-9]{1,2}_sig", "", group)
               )]

sig.list <- split(sig.df, sig.df$comparison)

# many more hits comparing Mock-virus vs virus-virus
lapply(sig.list, dim)


# this is now the list of sig genes for each comparison
lapply(seq_along(sig.list), function(x,n,i){
  fwrite(x[[i]], ScriptAndDatedFileName(paste0(n[[i]], 'sig.genes.unfiltered.csv')))
},x=sig.list, n=names(sig.list))


# write out the sig list in wide 
wide.list <- lapply(seq_along(sig.list), function(x,n,i){
  # conver to wide for each tp
  dt <- x[[i]]
 wide.dt <- dcast(dt, symbol+entrez~comparison+timepoint, value.var = c('log2FoldChange','pvalue', 'padj'))
 #fwrite(x[[i]], ScriptAndDatedFileName(paste0(n[[i]], 'sig.genes.unfiltered.csv')))
},x=sig.list, n=names(sig.list))

lapply(wide.list, colnames)


# reorder col names
col.order <- c("symbol","entrez","log2FoldChange_0","log2FoldChange_3","log2FoldChange_6","log2FoldChange_12","log2FoldChange_24",
               "pvalue_0","pvalue_3","pvalue_6","pvalue_12","pvalue_24",
               "padj_0", "padj_3","padj_6","padj_12","padj_24")


new.wide.list <- lapply(wide.list, function(x){
  colnames(x) <- gsub("_(Mock|Brisbane|Victoria)-(Mock|Brisbane|Victoria)", "", colnames(x))
  setcolorder(x, col.order)
  return(x)
})

names(new.wide.list) <- c("Mock-Brisbane", "Mock-Victoria", "Victoria-Brisbane")


lapply(seq_along(new.wide.list), function(x,n,i){
  
  fwrite(x[[i]], ScriptAndDatedFileName(paste0(n[[i]], '.siggenes.wide.csv')))
  
},x=new.wide.list, n=names(new.wide.list))
```


