---
title: "011223.regeneratePlots"
author: "Martin Gordon"
date: "2023-12-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## regenerate plots for Jyoti

First create a plot of the number of sig genes in each group and compare
Take the filtered list and plot

Normalise counts look good; counts align with DESeq2 res
Important to note with naming Denominator-Numerator (so Down means reduced in Numerator relativr to Denominator)
Important note:



```{r}
library(randomcoloR)
library(magrittr)
library(data.table)
library(stringr)
library(ggplot2)
library(tximport)
library(ggrepel)
library(DESeq2)
# visualisation
library(ComplexHeatmap)
library(EnhancedVolcano)
# annotation & filtering 
library(AnnotationHub)
library(genefilter) # not available 

#Ben utils
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/enrichmentTestFunctions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping

```


load in the unfiltered lists
```{r}
vic_bris <- fread('011223.share.1st.run.results_data/2023_12_01_Victoria-Brisbanesig.genes.unfiltered.csv')
mock_vic <-  fread('011223.share.1st.run.results_data/2023_12_01_Mock-Victoriasig.genes.unfiltered.csv')
mock_bris <-  fread('011223.share.1st.run.results_data/2023_12_01_Mock-Brisbanesig.genes.unfiltered.csv')

degs.dt <- rbind(vic_bris, mock_vic, mock_bris)
```

wont do as I will need the 

barplot plot by timepoint
```{r}
degs.dt[sig != 'Not', .N, by=.(timepoint, comparison,sig)]
degs.dt[sig != 'Not' & timepoint == 24 & comparison == 'Mock-Victoria', unique(symbol)] %>%  length()

degs.dt[, c("denom", "num") := tstrsplit(comparison, '-', fixed=T)]
degs.dt[, new_Label := paste0(num, '-', denom)]

drop.cols <-  c('denom', 'num')
degs.dt[, (drop.cols) := NULL]


col.pal<- randomcoloR::distinctColorPalette(k=3)

g <- ggplot(degs.dt[sig != 'Not', .N, by=.(timepoint, new_Label,sig)], aes(x=factor(timepoint), y=N, fill = forcats::fct_rev(sig))) +
  geom_bar(stat = 'Identity') +
  scale_fill_manual('sig', values=c('Up'='darkred', 'Down'='darkblue')) +
  ylab('Counts (log2 FC > 1 & adj.p < 0.05)') +
  xlab('Timepoint') +
  facet_wrap(~new_Label, nrow=3, strip.position="right") +
  theme_classic()

BackupAsPDF(g, 'Nsig.3comparisons')

g <- ggplot(degs.dt[sig != 'Not' & comparison != 'Victoria-Brisbane', .N, by=.(timepoint, new_Label,sig)], aes(x=factor(timepoint), y=N, fill = forcats::fct_rev(sig))) +
  geom_bar(stat = 'Identity') +
  scale_fill_manual('sig', values=c('Up'='darkred', 'Down'='darkblue')) +
  ylab('Counts (log2 FC > 1 & adj.p < 0.05)') +
  xlab('Timepoint') +
  facet_wrap(~new_Label, nrow=3, strip.position="right") +
  theme_classic()
g
BackupAsPDF(g, 'Nsig.3comparisons')
```
To generate the requested heatmap, just rerun the enrichment analysis, simplify the heatmap  

get mapping for ensembl to gene symbol
```{r}
gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")

ah <- AnnotationHub()
#download dbs matching the follwoing query (only one element matching)
hs.embl.101 <- AnnotationHub::query(ah, c("Hsapiens", "EnsDb", "101"))[[1]]

dds <- readRDS('/Users/martingordon/Documents/projects/061223_JyotiB_RNAseq/output/011223.dds.obj.rds')

# map the names 
universe <- AnnotationDbi::mapIds(hs.embl.101, keys = rownames(assay(dds)),column = "SYMBOL", keytype = "GENEID", multiVals = "first")
```

create sig_grps and run the enrichment analysis

041223
---
Edit adjusting the min threshold for gene set size from 10 to 6 to include one of the samples

```{r}
subDT <- degs.dt[new_Label != 'Brisbane-Victoria',]

#subDT[, new_Label := sub('-','.', new_Label)]
subDT[, enrich_grp := paste0(new_Label,'_',timepoint, '_', sig)]
subDT$enrich_grp %>% unique() %>%  length() # 20 go in

enrich.dt <- enricherOnGroups(subDT[sig != 'Not',], geneColumn = "symbol", groupColumns = 'enrich_grp', universe=universe, term2gene.gmt = gmt.go, minGSSize=10, numProcessors = 6)
# best per group enrichment heatmap

simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')
simp.enrich$simplified$enrich_grp %>%  unique()

col.template <- c("Brisbane-Mock_0_Down","Brisbane-Mock_3_Down","Brisbane-Mock_6_Down","Brisbane-Mock_12_Down","Brisbane-Mock_24_Down")
col.order <- c(col.template, sub('Brisbane', 'Victoria', col.template))
col.order <- c(col.order, sub('Down', 'Up', col.template))
col.order <- c(col.order,c("Victoria-Mock_0_Up","Victoria-Mock_3_Up","Victoria-Mock_6_Up","Victoria-Mock_12_Up","Victoria-Mock_24_Up"))

col.order[!col.order %in% unique(simp.enrich$simplified$enrich_grp)]
col.order <- col.order[!col.order %in% ("Victoria-Mock_0_Down")]

unique(simp.enrich$simplified$enrich_grp) %in% col.order
col.order %in% unique(simp.enrich$simplified$enrich_grp) 

#set the column ordering before running
simp.enrich$simplified[, enrich_grp := factor(enrich_grp, levels=col.order)]
simp.enrich$clusterInfo$enrich_grp %>%  unique()

hm <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, cols=col.order, groupColumn = 'enrich_grp', topN=8, upperThreshold=10, 
                                row_names_gp = gpar(fontsize = 8), column_names_gp = gpar(fontsize = 6), negCols=unique(simp.enrich$simplified$enrich_grp[grep('Down', simp.enrich$simplified$enrich_grp)]),
                                cluster_columns=FALSE, column_title_gp=gpar(fontsize=7.5),
                                column_split=list(rep(c('Down', 'Up'), c(9,10)),
                                                 rep(c('Brisbane', 'Victoria', 'Brisbane', 'Victoria'),c(5,4,5,5)) ) 
                                )

BackupAsPDF(hm, 'ordered.direction.nocolor.heatmap.',dimensions=c(8,8))                                            
```
Extract the victoria mock_3_Down rows and rbind to the dts for the original

```{r}
dummy.row <- simp.enrich$simplified[enrich_grp == "Victoria-Mock_3_Down",]
dummy.row$enrich_grp <- "Victoria-Mock_0_Down"

dummy.clusterinfo <-  simp.enrich$clusterInfo[enrich_grp == "Victoria-Mock_3_Down",]
dummy.clusterinfo$enrich_grp <- "Victoria-Mock_0_Down"

simp.enrich$simplified <- rbind(dummy.row, simp.enrich$simplified)
simp.enrich$clusterInf <- rbind(dummy.clusterinfo,  simp.enrich$clusterInfo)


col.template <- c("Brisbane-Mock_0_Down","Brisbane-Mock_3_Down","Brisbane-Mock_6_Down","Brisbane-Mock_12_Down","Brisbane-Mock_24_Down")
col.order <- c(col.template, sub('Brisbane', 'Victoria', col.template))
col.order <- c(col.order, sub('Down', 'Up', col.template))
col.order <- c(col.order,c("Victoria-Mock_0_Up","Victoria-Mock_3_Up","Victoria-Mock_6_Up","Victoria-Mock_12_Up","Victoria-Mock_24_Up"))

hm <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, cols = col.order,
                                groupColumn = 'enrich_grp', topN=8, upperThreshold=10, 
                                row_names_gp = gpar(fontsize = 8), column_names_gp = gpar(fontsize = 6), negCols=unique(simp.enrich$simplified$enrich_grp[grep('Down', simp.enrich$simplified$enrich_grp)]),
                                cluster_columns=FALSE, column_title_gp=gpar(fontsize=7.5),
                                column_split=list(rep(c('Down', 'Up'), c(10,10)),
                                                 rep(c('Brisbane', 'Victoria', 'Brisbane', 'Victoria'),c(5,5,5,5)) ) 
                                )

BackupAsPDF(hm, 'ordered.direction.nocolor.heatmap.allSamples.',dimensions=c(8,8))    
```
clear that the heatmap is dominated by upregulated sites which are mainly response to infection...
Not a lot of difference in the virus states.. lets just plot what is downregulated in the host


only 4-5 weakly significant negatively enriched terms
```{r}

hm <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified[enrich_grp %like% 'Down'], 
                                groupColumn = 'enrich_grp', topN=8, upperThreshold=10, 
                                row_names_gp = gpar(fontsize = 8), column_names_gp = gpar(fontsize = 6), negCols=unique(simp.enrich$simplified$enrich_grp[grep('Down', simp.enrich$simplified$enrich_grp)]),
                                cluster_columns=FALSE, column_title_gp=gpar(fontsize=7.5))
```
Not much happening.. lets try Victoria vs Brisbane

```{r}
subDT <- degs.dt[group %in% c("Victoria_0-Brisbane_0_dds", "Victoria_3-Brisbane_3_dds",  "Victoria_6-Brisbane_6_dds", "Victoria_12-Brisbane_12_dds", "Victoria_24-Brisbane_24_dds"),]
subDT[, enrich_grp := paste0(new_Label,'_',timepoint, '_', sig)]


enrich.dt <- enricherOnGroups(subDT[sig != 'Not',], geneColumn = "symbol", groupColumns = 'enrich_grp', universe=universe, term2gene.gmt = gmt.go, minGSSize=10, numProcessors = 6)
# best per group enrichment heatmap

simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')
simp.enrich$simplified$enrich_grp %>%  unique()

col.order <- c("Brisbane-Victoria_0_Down","Brisbane-Victoria_3_Down","Brisbane-Victoria_6_Down","Brisbane-Victoria_12_Down","Brisbane-Victoria_24_Down",
                  "Brisbane-Victoria_0_Up","Brisbane-Victoria_3_Up","Brisbane-Victoria_6_Up","Brisbane-Victoria_12_Up","Brisbane-Victoria_24_Up")

col.order[!col.order %in% unique(simp.enrich$simplified$enrich_grp)]
col.order <- col.order[!col.order %in% ("Victoria-Mock_0_Down")]

unique(simp.enrich$simplified$enrich_grp) %in% col.order
col.order %in% unique(simp.enrich$simplified$enrich_grp) 

#set the column ordering before running
simp.enrich$simplified[, enrich_grp := factor(enrich_grp, levels=col.order)]
simp.enrich$clusterInfo$enrich_grp %>%  unique()

hm <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, cols=col.order, groupColumn = 'enrich_grp', topN=16, upperThreshold=10, 
                                row_names_gp = gpar(fontsize = 8), column_names_gp = gpar(fontsize = 6), negCols=unique(simp.enrich$simplified$enrich_grp[grep('Down', simp.enrich$simplified$enrich_grp)]),
                                cluster_columns=FALSE, column_title_gp=gpar(fontsize=7.5),
                                column_split=list(rep(c('Down', 'Up'), c(5,5))) )
                             
BackupAsPDF(hm, 'vic.brisbane.heatmap.',dimensions=c(8,8))    


simp.enrich$simplified[, .N, by=.(enrich_grp)]
fwrite(enrich.dt, '
```
write out the simplified enrichment table for virus vs virus

```{r}
fwrite(enrich.dt, ScriptAndDatedFileName('vic.brisbane.enrichment.csv'))

enrich.dt
```


retry with this

```{r}
col.template <- c("Brisbane-Mock_0_Down","Brisbane-Mock_3_Down","Brisbane-Mock_6_Down","Brisbane-Mock_12_Down","Brisbane-Mock_24_Down")
col.order <- c(col.template, sub('Brisbane', 'Victoria', col.template))
col.order <- c(col.order, sub('Down', 'Up', col.template))
col.order <- c(col.order,c("Victoria-Mock_0_Up","Victoria-Mock_3_Up","Victoria-Mock_6_Up","Victoria-Mock_12_Up","Victoria-Mock_24_Up"))


#col.order[!col.order %in% unique(simp.enrich$simplified$enrich_grp)]
#col.order <- col.order[!col.order %in% ("Victoria-Mock_0_Down")]
simp.enrich$simplified$enrich_grp %>%  unique()

unique(simp.enrich$simplified$enrich_grp) %in% col.order
col.order %in% unique(simp.enrich$simplified$enrich_grp) 

#set the column ordering before running
simp.enrich$simplified[, enrich_grp := factor(enrich_grp, levels=col.order)]
simp.enrich$clusterInfo$enrich_grp %>%  unique()

hm <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, cols=col.order, groupColumn = 'enrich_grp', topN=8, upperThreshold=10, 
                                row_names_gp = gpar(fontsize = 8), column_names_gp = gpar(fontsize = 6), negCols=unique(simp.enrich$simplified$enrich_grp[grep('Down', simp.enrich$simplified$enrich_grp)]),
                                cluster_columns=FALSE, column_title_gp=gpar(fontsize=7.5),
                                column_split=list(rep(c('Down', 'Up'), c(9,10)),
                                                 rep(c('Brisbane', 'Victoria', 'Brisbane', 'Victoria'),c(5,4,5,5)) ) 
                                )

```


write output

```{r}

# want this but also want the time ordering...
split.list <- data.table(virus = str_extract(col.order, '(Brisbane|Victoria)'),
                         direction = str_extract(col.order, '(Up|Down)'))

ifelse(grepl('Down', unique(simp.enrich$simplified$enrich_grp)), 'Down',"up")

fwrite(enrichment.dt, ScriptAndDatedFileName('simplifiedenrichmentResults.csv'))

enrichment.dt
```


```{r}
gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")

# our universe will be all genes used for the DGE testing
# read in the deseq2 object
dds <- readRDS('/Users/martingordon/Documents/projects/061223_JyotiB_RNAseq/output/011223.dds.obj.rds')
rownames(assay(dds))

# sanity checks look good
#plotCounts(dds, gene='ENSG00000081248', intgroup = 'grp')

degs.dt[symbol == 'GTF3C2'& comparison == 'Mock-Victoria',]
count.mat['ENSG00000115207',grep('(Victoria|Mock)_24', colnames(count.mat))]



lapply(seq_along(comb.res.dt.list), function(x,n,i){
  #perform enrihcment analysis on each of the groups
  enrich.comb <- enricherOnGroups(x[[i]][sig != 'Not',], geneColumn = "symbol", groupColumns = c("time","sig"), universe=universe, term2gene.gmt = gmt.go, numProcessors = 6)
  # best per group enrichment heatmap
  hm <- enrichHeatmapBestPerGroup(enrich.comb, NULL, groupColumn = 'time.sig', topN=15, upperThreshold=8, title=n[[i]],
                                 row_names_gp = gpar(fontsize = 6), column_names_gp = gpar(fontsize = 8), negCols = enrich.comb$time.sig[grep("Down", enrich.comb$time.sig)])
  
  
  
  BackupAsPDF(hm, paste0(n[[i]],"-GO.enrichment_"), dimensions=c(13,11.5))
}, x=comb.res.dt.list, n=names(comb.res.dt.list))



read in the DESeq2 obj from the analysis
clear from the output that the 

```{r}
dds <- readRDS('/Users/martingordon/Documents/projects/061223_JyotiB_RNAseq/output/011223.dds.obj.rds')

# sanity checks look good
#plotCounts(dds, gene='ENSG00000081248', intgroup = 'grp')

degs.dt[symbol == 'GTF3C2'& comparison == 'Mock-Victoria',]
count.mat['ENSG00000115207',grep('(Victoria|Mock)_24', colnames(count.mat))]
```

