---
title: "ERGS_metadata_quicklook"
author: "Martin Gordon"
date: "2024-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## notes 
Dropped sample "Human.PRKAG2.Propranolol.#367.106" from the analysis
Features I included in my PCA:
"Age_Years"         "Weight_KG"         "BSA"               "AVNN"              "SDNN"              "RMSSD"             "pNN50"             "SEM"              
 [9] "PIP"               "IALS"              "PSS"               "PAS"               "BETA_AR"           "HF_NORM_AR"        "HF_PEAK_AR"        "HF_POWER_AR"      
[17] "LF_NORM_AR"        "LF_POWER_AR"       "LF_TO_HF_AR"       "TOTAL_POWER_AR"    "VLF_NORM_AR"       "VLF_POWER_AR"      "VLF_TO_LF_AR"      "BETA_WELCH"       
[25] "HF_NORM_WELCH"     "HF_PEAK_WELCH"     "HF_POWER_WELCH"    "LF_NORM_WELCH"     "LF_PEAK_WELCH"     "LF_POWER_WELCH"    "LF_TO_HF_WELCH"    "TOTAL_POWER_WELCH"
[33] "VLF_NORM_WELCH"    "VLF_POWER_WELCH"   "VLF_TO_LF_WELCH"   "SD1"               "SD2"               "alpha1"            "alpha2"            "SampEn"           
[41] "SD1-SD2" 


Match levels? Drop samples collected not in one set of the data, or else we need to control for them eg Gender..

Not possible to normalise across all these things... for now, lets look at each species individually, perform linear models to identify which of these variables change in response to treatment
... maybe we want to look at the drugs separately. What variables are significantly different in response to genotype, per species
... do we control for drug, look at interaction between drug and genotype

## Quick look at the ERG data

```{r}
library(data.table)
library(ggplot2)
library(magrittr)
library(GGally)
library(ggrepel)
library(ComplexHeatmap)
library(circlize)
library(broom) # tidymodel outputs
library(eulerr) # venn diagrams



source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/LinearModels.R")
```

```{r}
anova.dt <- fread('./data/2-way_ANOVA_results.csv')
EKGs.dt <- fread('./data/EKGs_Clean.csv')
```
inpsect the original data
250 individuals measured with 50 variables
```{r}
dim(EKGs.dt)
EKGs.dt %>%  str()
```

First create a an identifier variable
Just give 1:nrow as no easy way to combine atm
```{r}
EKGs.dt$sample %>%  unique()
```

sample breakdown
```{r}

g <- ggplot(EKGs.dt[, .N, by=.(Species,Drug,Genotype)], aes(x=interaction(Species, Genotype), y=N, fill=Drug)) +
  geom_bar(stat='Identity') +
  scale_fill_manual(values=col.pal) +
  labs(title='Sample breakdown') +
  theme_classic()

BackupAsPDF(g, 'samplebreakdown.barplot')
```
Weirdly one of the rows is identical based on the identifiers below.. looks like repeated measures
```{r}
EKGs.dt[, rowID := .I]
EKGs.dt[, sample := paste0(Species, '.', Genotype, '.', Drug, '.', ID, '.', rowID)]
```

Create a matrix to run PCA; we want sample names as columns, which will be this new ID value, and to remove rowIDs and other variables

```{r}
col.pal <- randomcoloR::distinctColorPalette(k=8)

cols.noi <- c('ID', 'Genotype', 'Species', 'Height', 'BMI', 'Age_Normalized','Gender', 'Drug','Pacemaker','rowID')

# create an actual copy of the datatable
EKGs.sub.dt <- copy(EKGs.dt)
EKGs.sub.dt[, (cols.noi) := NULL]
```

First, plot the distribution of the variables per species
Do these overlap? If they are completely different is there a point of comparison? We need to think of a normalization
Need to plot distributions of the variables per sample, we need to see if the distributions overlap, otherwise we probably need to normalize

```{r}
tidy.dt <- melt(EKGs.sub.dt, id.vars=c("sample"))
tidy.dt[, c("Species","Genotype", "Drug") := tstrsplit(sample, "[.]", keep = c(1,2,3)) ]

g <- ggplot(tidy.dt, aes(x=value, fill=interaction(Genotype,Species))) + 
  geom_histogram() +
  facet_wrap(~variable, scales='free') +
  scale_fill_manual(values=col.pal) +
  theme_bw()
g
```
many variables...lets find strongly correlated variables first

```{r}
ekg.mat <- melt(EKGs.sub.dt, id.vars=c("sample")) %>% 
  dcast(., sample~variable, value.var = 'value') %>% 
  as.matrix(rownames='sample')

# also need to drop these as many missing values 
cols.noi <- c('Height', 'BMI')

ekg.mat <- ekg.mat[, !colnames(ekg.mat) %in% cols.noi]

dim(ekg.mat)
ekg.mat[complete.cases(ekg.mat),]  %>%  dim()
```

```{r}
hm <- Heatmap(cor(ekg.mat, use='all.obs', method='pearson'), col=colorRamp2(breaks=c(-1, 0, .8), colors = c('blue', 'white', 'red')), 
                                                                            column_names_gp = gpar(fontsize=6), row_names_gp = gpar(fontsize=6))
hm
BackupAsPDF(hm,'variables.pearsonCorrelation.heatmaps')
```


```{r}
scaled.mat <- scale(ekg.mat)

ekg.clean.dt <- melt(data.table(scaled.mat, keep.rownames = T), id.vars = 'rn')
ekg.clean.dt[,c("Species","Genotype", "Drug") := tstrsplit(rn, "[.]", keep = c(1,2,3)) ]


g <- ggplot(ekg.clean.dt, aes(x=variable, y=value, fill=interaction(Species,Genotype), label=rn)) +
  geom_boxplot() +
  geom_text_repel(data=ekg.clean.dt[value > 2.5,], aes(label=rn)) +
  theme_bw() +
  scale_fill_manual(values=col.pal) +
  facet_grid(Genotype~Species, scales = 'free_x') +
  theme(axis.text.x = element_text(angle=90))
g
BackupAsPDF(g, 'variableDistributionScaled.boxplot', dimensions=c(12,12))
```
plot distributions of the raw scores

```{r}
g <- ggplot(ekg.clean.dt, aes(x=interaction(Genotype, Species), y=value, fill=interaction(Genotype, Species), label=rn)) + 
  geom_violin() +
  geom_point() +
  geom_hline(yintercept = 0, alpha=0.4, linetype=2) +
  geom_text_repel(size=1.5) +
  facet_wrap(~variable, ncol=4, scales='free_y') +
  scale_fill_manual(values = col.pal) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90))
g
BackupAsPDF(g, 'variableDistribution.groupFaceted.violinplot', dimensions = c(14,30))
```

Lets look at how scaling works for this data; driven by species
Difference will be clearly driven by species along PC1

```{r}
#incorrect as scaling at species level will affect the output..
scale.dt <- scale(ekg.mat, center=T, scale=T) %>% 
  as.data.table(keep.rownames=T) %>% 
  melt(., id.vars='rn')

scale.dt[, c("Species","Genotype", "Drug") := tstrsplit(rn, "[.]", keep = c(1,2,3)) ]

ggplot(scale.dt, aes(x=value, fill=(Species))) +
  geom_histogram() +
  facet_wrap(~variable, scales='free')
```

We know Species is the main driver of effect here

```{r}

pcaOut <- prcomp(ekg.mat, center = T, scale. = T)
colInfo <- data.table(colname = rownames(ekg.mat))

colInfo
colInfo[,c("Species","Genotype", "Drug") := tstrsplit(colname, "[.]", keep = c(1,2,3)) ]

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
```

plot a scree plot of the PCs
Beyond the first 4 PCs not a lot of variance explained..

```{r}
# scree plot of the PCAs
q <- qplot(1:10,  pcaPercentVar[1:10]) +
  geom_line() +
  xlab('Princpal Component 1:10') +
  scale_x_continuous(breaks=seq(1,10)) +
  ylab(('% Variance explained')) +
  ggtitle('PCA scree plot') +
  theme_bw()

q
BackupAsPDF(q, 'pca.screeplot.')
```

```{r}
#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = Drug, shape = Drug)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  geom_vline(xintercept=0, linetype=2, alpha=0.4) +
  geom_hline(yintercept = 0, linetype=2, alpha=0.4) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  scale_fill_manual(values = col.pal) +
  
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, "PCA_PC12")
```
Issue with the PCA is signal is dominated by species. We need to subset this to species seperately and run

```{r}
# create a human and mouse matrix seperately
hs.mat <- ekg.mat[grepl('Human', rownames(ekg.mat)), ]
mm.mat <- ekg.mat[!rownames(ekg.mat) %in% rownames(hs.mat),]

mat.list <- list(human = hs.mat, 
                 mouse = mm.mat)

```

run prcomp indiviudally

```{r}
lapply(seq_along(mat.list), function(x,n,i){
  
  pcaOut <- prcomp(x[[i]], center = T, scale. = T)
  colInfo <- data.table(colname = rownames(x[[i]]))


  colInfo[,c("Genotype", "Drug") := tstrsplit(colname, "[.]", keep = c(2,3)) ]

  #PCA
  pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
  pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
  pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
  
  
  #
  p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = Genotype, shape = Genotype)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 2) +
  geom_vline(xintercept=0, linetype=2, alpha=0.4) +
  geom_hline(yintercept = 0, linetype=2, alpha=0.4) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  scale_fill_manual(values = col.pal) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
  
},x=mat.list, n=names(mat.list))

```
Drop the outlier human samples and rerun
The second one is not as bad.. may yet include

```{r}
mat.list$human <-  mat.list$human[rownames(mat.list$human) != "Human.PRKAG2.Propranolol.#367.106",]
# dropping. two...
mat.list$human <-  mat.list$human[!rownames(mat.list$human) %in% c("Human.PRKAG2.Propranolol.#367.106", "Human.PRKAG2.Atropine.#677.66"),]
rownames(mat.list$human)
```

Looked at the other PCs and couldnt seem to recover genotype as a key driver distinguishing the samples
```{r}
lapply(seq_along(mat.list), function(x,n,i){
  
  pcaOut <- prcomp(x[[i]], center = T, scale. = T)
  colInfo <- data.table(colname = rownames(x[[i]]))


  colInfo[,c("Genotype", "Drug") := tstrsplit(colname, "[.]", keep = c(2,3)) ]

  #PCA
  pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
  pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
  pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
  
  
  #
  p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = Genotype, shape = Drug)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 2) +
  geom_vline(xintercept=0, linetype=2, alpha=0.4) +
  geom_hline(yintercept = 0, linetype=2, alpha=0.4) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  scale_fill_manual(values = col.pal) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

# BackupAsPDF(p, paste0(n[[i]],'.pc1pc2.genotype.pca'))
  
},x=mat.list, n=names(mat.list))
```

More metadata information? combine the sample metadata table with the rownames. Color the variables by other metadata characteristics to see what is driving variance

PC4 some speeration with pacemaker and BMI for human, for mouse on PC4 clear seperation on Age

```{r}

meta.dt <- EKGs.dt[, .(sample, Age_Years, Weight_KG, BMI, Gender, Pacemaker)]

lapply(seq_along(mat.list), function(x,n,i){
  
  pcaOut <- prcomp(x[[i]], center = T, scale. = T)
  colInfo <- data.table(colname = rownames(x[[i]]))


  colInfo[,c("Genotype", "Drug") := tstrsplit(colname, "[.]", keep = c(2,3)) ]

  #PCA
  pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
  pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
  pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
  
  # add in extra metadata
  pcaDT <- merge(pcaDT, meta.dt, by.x="rn", by.y="sample", all.x=T)
  
    #
  p <- ggplot (pcaDT, aes(x=PC3, y=PC4,  fill = Age_Years, shape = Drug)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 2) +
  geom_vline(xintercept=0, linetype=2, alpha=0.4) +
  geom_hline(yintercept = 0, linetype=2, alpha=0.4) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
 # scale_fill_manual(values = col.pal) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

  
},x=mat.list, n=names(mat.list))
```
What variables have a significantly different response to genotype?
------

Split the dataset into human and mouse and tidy

```{r}
hs.ekg.dt <- EKGs.dt[Species == 'Human']
mm.ekg.dt <- EKGs.dt[Species == 'Mouse']


#get all the numeric columns
num.cols <- colnames(hs.ekg.dt)[unlist(lapply(hs.ekg.dt, is.numeric))]
cols.oi <- c(num.cols, 'sample')
hs.ekg.dt <- hs.ekg.dt[, ..cols.oi]


num.cols <- colnames(mm.ekg.dt)[unlist(lapply(mm.ekg.dt, is.numeric))]
cols.oi <- c(num.cols, 'sample')
mm.ekg.dt <-  mm.ekg.dt[, ..cols.oi]

# remove variables not of interest to use 
cols.noi <- c('ID','Species','BMI', 'Age_Normalized', 'rowID', 'Height')

mm.ekg.dt <- mm.ekg.dt[, (cols.noi) := NULL]
hs.ekg.dt <- hs.ekg.dt[, (cols.noi) := NULL]


ekg.dts <- list(human = hs.ekg.dt,
                mouse =mm.ekg.dt)

```


```{r}
lapply(seq_along(ekg.dts), function(x,n,i){
  
  dt <- melt(x[[i]], id.vars='sample')
  dt[, c("Genotype", "Drug") := tstrsplit(sample, "[.]", keep = c(2,3)) ]
  
  g <- ggplot(dt, aes(y=value, x=Genotype, fill=Genotype)) +
    geom_boxplot() +
    geom_jitter(color='black', alpha=0.4, width=0.3, height=0.1) +
    labs(title=n[[i]]) +
    scale_fill_manual(values=col.pal) +
    facet_wrap(~variable, scales='free_y')  +
    theme_classic()
  
  BackupAsPDF(g, paste0(n[[i]],'variables.boxplot'), dimensions = c(14,18))
  
},x=ekg.dts, n=names(ekg.dts))


lapply(seq_along(ekg.dts), function(x,n,i){
  
  dt <- melt(x[[i]], id.vars='sample')
  dt[, c("Genotype", "Drug") := tstrsplit(sample, "[.]", keep = c(2,3)) ]
  
  g <- ggplot(dt, aes(y=value, x=interaction(Genotype,Drug), fill=interaction(Genotype,Drug))) +
    geom_boxplot() +
    geom_jitter(color='black', alpha=0.4, width=0.3, height=0.1) +
    labs(title=paste0(n[[i]], ' treatment breakdown')) +
    scale_fill_manual(values=col.pal) +
    facet_wrap(~variable, scales='free_y')  +
    theme_classic() +
    theme(axis.text.x = element_text(angle=90))
  
  BackupAsPDF(g, paste0(n[[i]],'variables.drugbreakdown.boxplot'), dimensions = c(18,18))
  
},x=ekg.dts, n=names(ekg.dts))

```

convert some of the variables to log scale and drop BSA variable as covered by Weight

```{r}

ekgs.long$human <- ekgs.long$human[ variable != 'BSA',]
ekgs.long$mouse<- ekgs.long$mouse[ variable != 'BSA',]

##ekg.dts[[1]][, c("Genotype", "Drug") := NULL]
ekg.dts[[2]] <- ekg.dts[[2]][, c("ID","Species","Genotype", "Drug","Gender", "Pacemaker") := NULL]


# convert the datatable to long format
ekgs.long <- lapply(ekg.dts, function(x){
 melt(x, id.vars='sample')
})
names(ekgs.long) <- names(ekg.dts)

# create a bkup copy
ekgs.long.bak <- copy(ekgs.long)

#hu8man
# remove SEM and SD1-SD2
# add VLF to FL AR

vars.logtransform <- c( "pNN50","RMSSD", "SDNN", "HF_POWER_AR", "LF_POWER_AR", "TOTAL_POWER_AR", "VLF_POWER_AR", "HF_POWER_WELCH", "LF_POWER_WELCH", "TOTAL_POWER_WELCH", "VLF_POWER_WELCH", "VLF_TO_LF_WELCH", "VLF_TO_LF_AR", "SD1", "SD2") 

ekgs.long$human[variable %in% vars.logtransform, value := log2(value)]

#mosue
vars.logtransform <- c("Age_Years", "Weight_KG", "pNN50","RMSSD", "SDNN", "PIP", "LF_POWER_AR", "TOTAL_POWER_AR", "LF_POWER_AR", "TOTAL_POWER_AR", "VLF_TO_LF_AR", "LF_PEAK_WELCH", "LF_POWER_WELCH", "VLF_POWER_AR", "LF_TO_HF_WELCH", "LF_POWER_WELCH", "TOTAL_POWER_WELCH", "VLF_POWER_WELCH", "VLF_TO_LF_WELCH") 

ekgs.long$mouse[variable %in% vars.logtransform, value := log2(value)]
```
replot 

```{r}
lapply(seq_along(ekgs.long), function(x,n,i){
  dt <-  x[[i]]
  dt[, c("Genotype", "Drug") := tstrsplit(sample, "[.]", keep = c(2,3)) ]
  
  print(dt)
  
  g <- ggplot(dt, aes(y=value, x=interaction(Genotype,Drug), fill=interaction(Genotype,Drug))) +
    geom_boxplot() +
    geom_jitter(color='black', alpha=0.4, width=0.3, height=0.1) +
    labs(title=paste0(n[[i]], ' treatment breakdown')) +
    scale_fill_manual(values=col.pal) +
    facet_wrap(~variable, scales='free_y')  +
    theme_classic() +
    theme(axis.text.x = element_text(angle=90))
  
  #BackupAsPDF(g, paste0(n[[i]],'variables.drugbreakdown.logtransformed.boxplot'), dimensions = c(18,18))
  
},x=ekgs.long, n=names(ekgs.long))


g <- ggplot(ekgs.long$mouse, aes(y=value, x=interaction(Genotype,Drug), fill=interaction(Genotype,Drug))) +
    geom_boxplot() +
    geom_jitter(color='black', alpha=0.4, width=0.3, height=0.1) +
   # labs(title=paste0(n[[i]], ' treatment breakdown')) +
    scale_fill_manual(values=col.pal) +
    facet_wrap(~variable, scales='free_y')  +
    theme_classic() +
    theme(axis.text.x = element_text(angle=90))


g
```
ANOVA models
-------
Run modelling
List of models including terms to control for age, gender etc

```{r}
# just take entire dts
hs.ekg.dt <- EKGs.dt[Species == 'Human']
mm.ekg.dt <- EKGs.dt[Species == 'Mouse']

ekg.dts <- list(human = hs.ekg.dt,
                mouse =mm.ekg.dt)

# set as factors the things we wish to control for (set control base levels for genotype and treatment)
#human
ekg.dts[[1]][, Gender := factor(Gender, levels=c('M','F'))]
ekg.dts[[1]][, Drug := factor(Drug, levels=c("Basal","Atropine", "Propranolol"))]
ekg.dts[[1]][, Genotype := factor(Genotype, levels=c("Control","PRKAG2"))]

# mouse
ekg.dts[[2]][, Drug := factor(Drug, levels=c("Basal","Atropine", "Propranolol", "Double"))]
ekg.dts[[2]][, Genotype := factor(Genotype, levels=c("Control","PRKAG2"))]
```

Select the variables we wish to measure 
```{r}
# list of variables to look at
varToMeasure <- c("AVNN","SDNN","RMSSD","pNN50","SEM","PIP","IALS","PSS","PAS","BETA_AR","HF_NORM_AR",
                  "HF_PEAK_AR","HF_POWER_AR","LF_NORM_AR","LF_POWER_AR","LF_TO_HF_AR","TOTAL_POWER_AR","VLF_NORM_AR","VLF_POWER_AR","VLF_TO_LF_AR","BETA_WELCH",
                  "HF_NORM_WELCH","HF_PEAK_WELCH","HF_POWER_WELCH","LF_NORM_WELCH","LF_PEAK_WELCH","LF_POWER_WELCH","LF_TO_HF_WELCH","TOTAL_POWER_WELCH","VLF_NORM_WELCH",
                  "VLF_POWER_WELCH","VLF_TO_LF_WELCH","SD1","SD2","alpha1","alpha2","SampEn","SD1-SD2") 

ekg.dts[[1]][, c("Species", "Genotype", "Drug") := tstrsplit(sample, "[.]", keep = c(1, 2,3)) ]
ekg.dts[[2]][, c("Species", "Genotype", "Drug") := tstrsplit(sample, "[.]", keep = c(1, 2,3)) ]

#ekg.dts$human <- merge(x=ekg.dts$human, y=meta.dt$human[, .(Gender,sample)], by.x='sample', by.y='sample')

# we only want to model gender for humams, no point controlling for mice if only 1 level and waste a DF
human.lm.out <- lapply(varToMeasure, function(singleVar) {
    
    print(paste0('Running linear model on ', singleVar))

    # define a new formula list
    formulas = list (full = as.formula(paste(singleVar , "~Gender+Age_Years+Genotype*Drug")),
                 noInteraction = as.formula(paste(singleVar ,'~Gender+Age_Years+Genotype+Drug')),
                 basic = as.formula(paste(singleVar ,'~Genotype+Drug'))
                 )

    # run the linearModels on each response var  
    lms <- lapply(formulas, function(x){
      lm(formula = x, data=ekg.dts$human[, !"sample"]) #drop non-numeric variables
    })
    
    # extract co-efficients as data.table
    coef.table <- lapply (lms, function(lm.out)as.data.table(coefficients(summary(lm.out)), keep.rownames = TRUE)) %>% 
      rbindlist(., idcol = "model")
    setnames (coef.table, old = c("rn", "Pr(>|t|)"), new = c("term", "p.value"), skip_absent = TRUE)
  
    #f.tests from anova function; visualise pval histograms to assess model performance
    anova.list <- lapply (lms, function(l)as.data.table(anova(l), keep.rownames=TRUE))
    anovaTables <- rbindlist(anova.list, idcol = "model")
    
    anovaTables[, sigCode := dplyr::case_when(`Pr(>F)`< 0.001 ~ "***",
                                              `Pr(>F)`< 0.01  ~ "**",
                                              `Pr(>F)`< 0.05  ~ "*",
                                              `Pr(>F)`< 0.1  ~ ".",
                                              TRUE ~ "")]
    setnames(anovaTables, old = c("rn", "Pr(>F)"), new = c("term", "p.value"), skip_absent = TRUE)

    
     return(list(anova = anovaTables, 
                 coef = coef.table))
})

names(human.lm.out) <-  varToMeasure

hu.anova <- lapply(human.lm.out, "[[", "anova") %>%  rbindlist(idcol=T)
hu.coeff <- lapply(human.lm.out, "[[", "coef") %>%  rbindlist(idcol=T)

# write out the results of the anova
fwrite(hu.anova, ScriptAndDatedFileName('human.anova.txt'))
fwrite(hu.coeff, ScriptAndDatedFileName('human.coefficients.txt'))
```

```{r}
mouse.lm.out <- lapply(varToMeasure, function(singleVar) {
    
    print(paste0('Running linear model on ', singleVar))

    # define a new formula list
    formulas = list (full = as.formula(paste(singleVar , "~Age_Years+Weight_KG+Genotype*Drug")),
                 noInteraction = as.formula(paste(singleVar ,'~Age_Years+Weight_KG+Genotype+Drug')),
                 basic = as.formula(paste(singleVar ,'~Genotype+Drug'))
                 )

    # run the linearModels on each response var  
    lms <- lapply(formulas, function(x){
      lm(formula = x, data=ekg.dts$mouse[, !"sample"]) #drop non-numeric variables
    })
    
    # extract co-efficients as data.table
    coef.table <- lapply (lms, function(lm.out)as.data.table(coefficients(summary(lm.out)), keep.rownames = TRUE)) %>% 
      rbindlist(., idcol = "model")
    setnames (coef.table, old = c("rn", "Pr(>|t|)"), new = c("term", "p.value"), skip_absent = TRUE)
  
    #f.tests from anova function; visualise pval histograms to assess model performance
    # only compares fit of full model to null model (just intercept)
    anova.list <- lapply (lms, function(l)as.data.table(anova(l), keep.rownames=TRUE))
    anovaTables <- rbindlist(anova.list, idcol = "model")
    
    anovaTables[, sigCode := dplyr::case_when(`Pr(>F)`< 0.001 ~ "***",
                                              `Pr(>F)`< 0.01  ~ "**",
                                              `Pr(>F)`< 0.05  ~ "*",
                                              `Pr(>F)`< 0.1  ~ ".",
                                              TRUE ~ "")]
    
     setnames(anovaTables, old = c("rn", "Pr(>F)"), new = c("term", "p.value"), skip_absent = TRUE)
    
    
     return(list(anova = anovaTables, 
                 coef = coef.table))
})

names(mouse.lm.out) <-  varToMeasure

mm.anova <- lapply(mouse.lm.out, "[[", "anova") %>%  rbindlist(idcol=T)
mm.coeff <- lapply(mouse.lm.out, "[[", "coef") %>%  rbindlist(idcol=T)

```

# losing about 20% of hits when you don't include Gender in model, how much interaction includes the model is debatable...

```{r}
g <- ggplot(hu.anova, aes(x=`Pr(>F)`, fill=model)) +
  geom_histogram(boundary=0.0) +
  facet_wrap(~model) +
  scale_fill_manual(values=col.pal)

BackupAsPDF(g, 'human.pvalhistogram.anovaModelComparison')

g <- ggplot(mm.anova, aes(x=p.value, fill=model)) +
  geom_histogram(boundary=0.0) +
  facet_wrap(~model) +
  scale_fill_manual(values=col.pal)
g
BackupAsPDF(g, 'mouse.pvalhistogram.anovaModelComparison')
```
Write out the tables used to produce the ANOVA

```{r}
lapply(seq_along(ekg.dts), function(x,n,i){
  
  fwrite(x[[i]], ScriptAndDatedFileName(paste0(n[[i]], '.anova.input.tsv')))
},x=ekg.dts, n=names(ekg.dts))
```
First place to assess? Lets just pull out the list of terms that are significant to genotype in both human and mouse data
Which of these terms are significant (after correction) in mouse but not in human and vice versa

```{r}
# list of the anova results
anova.list <- list(human = hu.anova,
                   mouse = mm.anova)

lapply(anova.list, function(x){
  # correct at level of each model
  x[, p.adj := p.adjust(p.value, method='BH'), by=.(model,term)]
})
```
For genotype alone, identify how many variables this is significant for in both datasets

Lets just look at raw pvalue alone for now

```{r}
lapply(anova.list, function(x){
  # correct at level of each model
  x[term == 'Genotype' & p.value < 0.05, unique(.id), by=model]
})

```


A couple of terms for genotype weakly significant for mouse, but much stronger evidence for human
Just use a simple model for mouse and drop the other terms?

```{r}
sig.genotype <- lapply(anova.list, function(x){
  # correct at level of each model
  x[term == 'Genotype' & p.value < 0.05, unique(.id)]
})

names(sig.genotype) <- names(anova.list)

p <- plot(euler(sig.genotype),
    quantities = TRUE,
    main = 'Sig (p.val < 0.05) Genotype hits',
    fills = list(fill = c("red", "steelblue4"), alpha = 0.5)
     )

BackupAsPDF(p, 'genotype.pval.sig.eulerplot')

```
What about looking at drugs? How many of these terms respond differently to rtreatment?

```{r}
sig.drug <- lapply(anova.list, function(x){
  # correct at level of each model
  x[term == 'Drug' & p.value < 0.05, unique(.id)]
})

names(sig.drug) <- names(anova.list)

p <- plot(euler(sig.drug),
    quantities = TRUE,
    main = 'Sig (p.val < 0.05) Treatment hits'#,
    #fills = list(fill = c("", "steelblue4"), alpha = 0.5)
     )
p
BackupAsPDF(p, 'treatment.pval.sig.eulerplot')
```
Interaction of drug and genotype

```{r}
sig.interactions <- lapply(anova.list, function(x){
  # correct at level of each model
  x[term == 'Genotype:Drug' & p.value < 0.05, unique(.id)]
})

names(sig.interactions) <- names(anova.list)

p <- plot(euler(sig.interactions),
    quantities = TRUE,
    main = 'Sig (p.val < 0.05) Genotype Treatment Interactions',
    fills = list(fill = c("#ffcf20FF", "#2f9aa0FF"), alpha = 0.5)
     )
p
BackupAsPDF(p, 'treatment.genotype.pval.sig.eulerplot')
```
Interesting set of results from the analysis. Looks like a lot of things 


Also just produce a simple interaction plot to illustrate to Jack what the interaction highlights


test an anova for this
basic model performs best for mouse, at least for this variable

```{r}

lm1 <- lm(AVNN~Weight_KG+Genotype+Drug, data=ekg.dts$mouse)
lm2 <- lm(AVNN~WGenotype+Drug, data=ekg.dts$mouse)

anova(lm1, lm2)
# need to model drug and genotype
# again, interaction doesnt help

# tests with human
# only thing not worth including for this is the interaction term..

lm1 <- lm(AVNN~Gender+Age_Years+Weight_KG+Genotype*Drug, data=ekg.dts$human)
lm2 <- lm(AVNN~Gender+Age_Years+Genotype*Drug, data=ekg.dts$human)

anova(lm1, lm2)
```



03-22-24
----
I think we need to take a simplified approach. Look for things significantly different given genotype, per drug type within each species

```{r}

em.contrastOfContrasts <- function (l, factorFormula = ~status|timepoint){  # how does the drug effect change per tissue
  emm <- emmeans(l, factorFormula)
  contrast1 <- pairs(emm)
  contrast2 <- pairs(contrast1, by = NULL, adjust = "none")
  return (as.data.table(contrast2))
}

formulas = list (full = LogIntensities~timepoint*status)


## testing 
lm.out <- linearModelsAllProteins(ekg.dts$mouse, 
                                  formulaList = formulas, 
                                  splitColumn = "Protein", postProcessFunction = em.interactionContrast, cl = 6) |> suppressWarnings()



mouse.lm.out.bp <- lapply(varToMeasure, function(singleVar) {
    
    print(paste0('Running linear model on ', singleVar))

    # define a new formula list
    formulas = list (full = as.formula(paste(singleVar , "~Age_Years+Genotype")),
                 basic = as.formula(paste(singleVar ,'~Genotype'))
                 )
    
    lm.out <- linearModelsAllProteins(ekg.dts$mouse, 
                                  formulaList = as.formula(paste(singleVar , "~Age_Years+Genotype")), 
                                  splitColumn = "Species") 
})

mouse.lm.out.bp

mouse.lm.out.bp


ekg.dts$mouse


split (ekg.dts$mouse, ekg.dts$mouse[['Drug']])
```




Find the human terms significant for an interaction

```{r}
hu.anova[rn == 'Genotype:Drug', .N, .(.id, model)]

hu.sig <- hu.anova[rn == 'Genotype' & `Pr(>F)` < 0.05, .N, .(.id)]$.id
hu.anova[rn == 'Genotype:Drug' & `Pr(>F)` < 0.05,]

hu.coeff
hu.coeff[rn == 'Genotype:Drug' & `Pr(>F)` < 0.05,]
# 


hist(hu.anova$`Pr(>F)`)


g <- ggplot(hu.anova, aes(x=`Pr(>F)`)) +
  geom_histogram() +
  facet_wrap(~model)

BackupAsPDF(g, 'human.anova.histogram')
```




```{r}




ekg.dts$human$Gender
lm.out
  
ekgs.long


## testing 
lm.out <- linearModelsAllProteins(p.quant, 
                                  formulaList = formulas, 
                                  splitColumn = "Protein", postProcessFunction = em.interactionContrast, cl = 6) |> suppressWarnings()

varToMeasure <- c("Weight_KG","BSA","AVNN","SDNN","RMSSD","pNN50","SEM","PIP","IALS","PSS","PAS","BETA_AR","HF_NORM_AR",
                  "HF_PEAK_AR","HF_POWER_AR","LF_NORM_AR","LF_POWER_AR","LF_TO_HF_AR","TOTAL_POWER_AR","VLF_NORM_AR","VLF_POWER_AR","VLF_TO_LF_AR","BETA_WELCH",
                  "HF_NORM_WELCH","HF_PEAK_WELCH","HF_POWER_WELCH","LF_NORM_WELCH","LF_PEAK_WELCH","LF_POWER_WELCH","LF_TO_HF_WELCH","TOTAL_POWER_WELCH","VLF_NORM_WELCH",
                  "VLF_POWER_WELCH","VLF_TO_LF_WELCH","SD1","SD2","alpha1","alpha2","SampEn","SD1-SD2" ,"sample")  


View(linearModelsAllProteins)
```


I think lets try the 


I think best thing to do is process seperately
For now lets look individually and find things that response differently to Genotype

We also need to control for variables in our analysis also... need to control for Sex, Age etc (uninteresting..)

```{r}

```

Compare the models with an F-test to see if adding these co-variates explains significantly more variance

```{r}

```


Q. how do we handle these unmatched levels?

Lets show the 

Run Jacks anova first and extract the hits with significant p-value
Maybe first normalise the prots of interest 

```{r}
lapply((mat.list), function(x,n,i){
  colnames(x)})
```



Thinks to look at:
i) multicolinearilty; need to simplify number of variables, can we find ones that strongly correlate with each other?
ii) principal component regression:

