---
title: "100824_makePCProphetInput"
author: "Martin Gordon"
date: "2024-10-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## packages

```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)


source('~/Documents/utils/bp_utils/ManageScriptData.R')
source('~/Documents/utils/mg_utils/r_utils/SEC_MS_utils.mgupdate.R')
source('~/Documents/utils/bp_utils/UniprotIDMapping.R')
```

# read in the 3 datasets 
```{r}
secms.list <- list(ctrl = fread('./data/Biosep_MDAMB468_CTRL_2_Report.tsv'),
                   ptxl = fread('./data/Biosep_MDAMB468_PTXL_2_Report.tsv'),
                   vrst = fread('./data/Biosep_MDAMB468_VRST_1_Report.tsv'))

 
col.order <- c("PG.ProteinGroups","PG.ProteinAccessions","PG.Genes", "PG.UniProtIds", "PG.ProteinNames", 
               "[5] BiosepMDA_MB468_Ptxl2_072724_F11_A11_1_6617.d.PG.Quantity", "[76] BiosepMDA_MB468_Ptxl_072724_F12_A12_1_6618.d.PG.Quantity", "[6] BiosepMDA_MB468_Ptxl2_072724_F13_B12_1_6620.d.PG.Quantity", 
               "[7] BiosepMDA_MB468_Ptxl2_072724_F14_B11_1_6621.d.PG.Quantity",  "[8] BiosepMDA_MB468_Ptxl2_072724_F15_B10_1_6622.d.PG.Quantity" ,"[9] BiosepMDA_MB468_Ptxl2_072724_F16_B9_1_6623.d.PG.Quantity",   
               "[10] BiosepMDA_MB468_Ptxl2_072724_F17_B8_1_6624.d.PG.Quantity",  "[11] BiosepMDA_MB468_Ptxl2_072724_F18_B7_1_6625.d.PG.Quantity", "[12] BiosepMDA_MB468_Ptxl2_072724_F19_B6_1_6629.d.PG.Quantity" ,
               "[13] BiosepMDA_MB468_Ptxl2_072724_F20_B5_1_6630.d.PG.Quantity" , "[14] BiosepMDA_MB468_Ptxl2_072724_F21_B4_1_6631.d.PG.Quantity" , "[15] BiosepMDA_MB468_Ptxl2_072724_F22_B3_1_6632.d.PG.Quantity",
               "[16] BiosepMDA_MB468_Ptxl2_072724_F23_B2_1_6633.d.PG.Quantity",  "[17] BiosepMDA_MB468_Ptxl2_072724_F24_B1_1_6634.d.PG.Quantity", "[18] BiosepMDA_MB468_Ptxl2_072724_F25_C1_1_6636.d.PG.Quantity",
               "[19] BiosepMDA_MB468_Ptxl2_072724_F26_C2_1_6637.d.PG.Quantity",  "[20] BiosepMDA_MB468_Ptxl2_072724_F27_C3_1_6638.d.PG.Quantity", "[21] BiosepMDA_MB468_Ptxl2_072724_F28_C4_1_6639.d.PG.Quantity",
               "[22] BiosepMDA_MB468_Ptxl2_072724_F29_C5_1_6640.d.PG.Quantity",  "[23] BiosepMDA_MB468_Ptxl2_072724_F30_C6_1_6641.d.PG.Quantity", "[24] BiosepMDA_MB468_Ptxl2_072724_F31_C7_1_6645.d.PG.Quantity",
               "[25] BiosepMDA_MB468_Ptxl2_072724_F32_C8_1_6646.d.PG.Quantity",  "[26] BiosepMDA_MB468_Ptxl2_072724_F33_C9_1_6647.d.PG.Quantity", "[27] BiosepMDA_MB468_Ptxl2_072724_F34_C10_1_6648.d.PG.Quantity",
               "[28] BiosepMDA_MB468_Ptxl2_072724_F35_C11_1_6649.d.PG.Quantity", "[29] BiosepMDA_MB468_Ptxl2_072724_F36_C12_1_6650.d.PG.Quantity","[30] BiosepMDA_MB468_Ptxl2_072724_F37_D12_1_6652.d.PG.Quantity",
               "[31] BiosepMDA_MB468_Ptxl2_072724_F38_D11_1_6653.d.PG.Quantity", "[32] BiosepMDA_MB468_Ptxl2_072724_F39_D10_1_6654.d.PG.Quantity","[33] BiosepMDA_MB468_Ptxl2_072724_F40_D9_1_6655.d.PG.Quantity",
               "[34] BiosepMDA_MB468_Ptxl2_072724_F41_D8_1_6656.d.PG.Quantity",  "[35] BiosepMDA_MB468_Ptxl2_072724_F42_D7_1_6657.d.PG.Quantity", "[36] BiosepMDA_MB468_Ptxl2_072724_F43_D6_1_6661.d.PG.Quantity",
               "[37] BiosepMDA_MB468_Ptxl2_072724_F44_D5_1_6662.d.PG.Quantity",  "[38] BiosepMDA_MB468_Ptxl2_072724_F45_D4_1_6663.d.PG.Quantity", "[39] BiosepMDA_MB468_Ptxl2_072724_F46_D3_1_6664.d.PG.Quantity",
               "[40] BiosepMDA_MB468_Ptxl2_072724_F47_D2_1_6665.d.PG.Quantity",  "[41] BiosepMDA_MB468_Ptxl2_072724_F48_D1_1_6666.d.PG.Quantity" ,"[42] BiosepMDA_MB468_Ptxl2_072724_F49_E1_1_6668.d.PG.Quantity",
               "[43] BiosepMDA_MB468_Ptxl2_072724_F50_E2_1_6669.d.PG.Quantity",  "[44] BiosepMDA_MB468_Ptxl2_072724_F51_E3_1_6670.d.PG.Quantity","[45] BiosepMDA_MB468_Ptxl2_072724_F52_E4_1_6671.d.PG.Quantity",
               "[46] BiosepMDA_MB468_Ptxl2_072724_F53_E5_1_6672.d.PG.Quantity",  "[47] BiosepMDA_MB468_Ptxl2_072724_F54_E6_1_6673.d.PG.Quantity", "[48] BiosepMDA_MB468_Ptxl2_072724_F55_E7_1_6677.d.PG.Quantity" ,
               "[49] BiosepMDA_MB468_Ptxl2_072724_F56_E8_1_6678.d.PG.Quantity" , "[50] BiosepMDA_MB468_Ptxl2_072724_F57_E9_1_6679.d.PG.Quantity" ,"[51] BiosepMDA_MB468_Ptxl2_072724_F58_E10_1_6680.d.PG.Quantity",
               "[52] BiosepMDA_MB468_Ptxl2_072724_F59_E11_1_6681.d.PG.Quantity", "[53] BiosepMDA_MB468_Ptxl2_072724_F60_E12_1_6682.d.PG.Quantity","[54] BiosepMDA_MB468_Ptxl2_072724_F61_F12_1_6684.d.PG.Quantity",
               "[55] BiosepMDA_MB468_Ptxl2_072724_F62_F11_1_6685.d.PG.Quantity", "[56] BiosepMDA_MB468_Ptxl2_072724_F63_F10_1_6686.d.PG.Quantity","[57] BiosepMDA_MB468_Ptxl2_072724_F64_F9_1_6687.d.PG.Quantity" ,
               "[58] BiosepMDA_MB468_Ptxl2_072724_F65_F8_1_6688.d.PG.Quantity",  "[59] BiosepMDA_MB468_Ptxl2_072724_F66_F7_1_6689.d.PG.Quantity" ,"[60] BiosepMDA_MB468_Ptxl2_072724_F67_F6_1_6693.d.PG.Quantity" ,
               "[61] BiosepMDA_MB468_Ptxl2_072724_F68_F5_1_6694.d.PG.Quantity" , "[62] BiosepMDA_MB468_Ptxl2_072724_F69_F4_1_6695.d.PG.Quantity" ,"[63] BiosepMDA_MB468_Ptxl2_072724_F70_F3_1_6696.d.PG.Quantity",
               "[64] BiosepMDA_MB468_Ptxl2_072724_F71_F2_1_6697.d.PG.Quantity",  "[65] BiosepMDA_MB468_Ptxl2_072724_F72_F1_1_6698.d.PG.Quantity", "[66] BiosepMDA_MB468_Ptxl2_072724_F73_G1_1_6700.d.PG.Quantity",
               "[67] BiosepMDA_MB468_Ptxl2_072724_F74_G2_1_6701.d.PG.Quantity",  "[68] BiosepMDA_MB468_Ptxl2_072724_F75_G3_1_6702.d.PG.Quantity", "[69] BiosepMDA_MB468_Ptxl2_072724_F76_G4_1_6703.d.PG.Quantity" ,
               "[70] BiosepMDA_MB468_Ptxl2_072724_F77_G5_1_6704.d.PG.Quantity" , "[71] BiosepMDA_MB468_Ptxl2_072724_F78_G6_1_6705.d.PG.Quantity","[72] BiosepMDA_MB468_Ptxl2_072724_F79_G7_1_6707.d.PG.Quantity",
               "[73] BiosepMDA_MB468_Ptxl2_072724_F80_G8_1_6708.d.PG.Quantity",  "[74] BiosepMDA_MB468_Ptxl2_072724_F81_G9_1_6709.d.PG.Quantity" ,"[75] BiosepMDA_MB468_Ptxl2_072724_F82_G10_1_6710.d.PG.Quantity")

# adjust the column order
secms.list[['ptxl']] <- secms.list[['ptxl']][, ..col.order]


secms.mat <- lapply(secms.list, function(x) generatePCProphetMatrix(x, gene="PG.Genes", uniprot="PG.UniProtIds", fracRegx = "[0-9]{1,2}"))

# filter out the incharacterised protein; wont be mapped to PCP database anyway

secms.mat <- lapply(secms.mat, function(x){
  return(x[GN != '',])
})
```

Seems the search was performed with isoforms; not sure how complex detection will be handled with this. For now, just extract the canonical and drop the others
Handle the issues with peptide IDs; for now iterate through the rows, order/sort, extract the first peptide/uniprot ID

```{r}
# how to handle this; take the first ID
 secms.f.mat <- lapply(secms.mat, function(x){
  
  # for sanity check, count the number of dupliate rows in eahc
  dup.genes <- x[,.N, by=GN] %>% 
    .[N> 1, GN]
  
  message(paste0('Found ', length(dup.genes), ' genes with at least 1 duplicated geneID'))
  
  # first thing; extract the first uniprot ID in each of these instances
  x[, firstUniprot := tstrsplit(ID, ';', keep=1)]
  message(paste0('Extracting canonical from this set..'))
  
  canonical.set <- x[GN %in% dup.genes,] %>% 
    .[grep('[-]', firstUniprot, invert=T),]
  
  message(paste0('Found ', length(canonical.set[, unique(GN)]), ' canonical uniprot IDs'))
  message(paste0('Taking the first record for the remaining geneIDs: ',  
                 paste(unique(x[(GN %in% dup.genes & !GN %in% canonical.set$GN),]$GN), collapse=';')))
  
  
  non.canonical.set <- x[(GN %in% dup.genes & !GN %in% canonical.set$GN),] %>% 
    .[, head(.SD, 1), by=GN]
  
  clean.dt <- rbind(x[!GN %in% dup.genes,],
                    canonical.set,
                    non.canonical.set)
  
  # now tidy the output to match expected genes
  frac.cols <- grep('frac_', colnames(x), value=T)
  cols.oi <- c('GN','firstUniprot', frac.cols)
  
  clean.dt <- clean.dt[, ..cols.oi]
  setnames(clean.dt, old='firstUniprot', new="ID")
  
  message('Checking output geneIDs...')
  stopifnot(all(unique(x$GN) %in% unique(clean.dt$GN)))
  
  return(clean.dt)
})


dropNonCanonicalIsoforms <- function(dt, canonical.only=F){

  dup.genes <- dt[,.N, by=GN] %>% 
    .[N> 1, GN]
  
  message(paste0('Found ', length(dup.genes), ' genes with at least 1 duplicated geneID'))
  
  # first thing; extract the first uniprot ID in each of these instances
  dt[, firstUniprot := tstrsplit(ID, ';', keep=1)]
  message(paste0('Extracting canonical from this set..'))
  
  canonical.set <- dt[GN %in% dup.genes,] %>% 
    .[grep('[-]', firstUniprot, invert=T),]
  
  message(paste0('Found ', length(canonical.set[, unique(GN)]), ' canonical uniprot IDs'))
  message(paste0('Taking the first record for the remaining geneIDs: ',  
                 paste(unique(dt[(GN %in% dup.genes & !GN %in% canonical.set$GN),]$GN), collapse=';')))
  
  
  non.canonical.set <- dt[(GN %in% dup.genes & !GN %in% canonical.set$GN),] %>% 
    .[, head(.SD, 1), by=GN]
  
  if (!canonical.only){
    
    message('Including non-canonical uniprots...')
    clean.dt <- rbind(dt[!GN %in% dup.genes,],
                    canonical.set,
                    non.canonical.set)
    
  } else {
    
    message('Removing non-canonical uniprots...')
    clean.dt <- rbind(dt[!GN %in% dup.genes,],
                    canonical.set)
    
  }
  
  # nset output columns
  frac.cols <- grep('frac_', colnames(dt), value=T)
  cols.oi <- c('GN','firstUniprot', frac.cols)
 
  clean.dt <- clean.dt[, ..cols.oi]
  setnames(clean.dt, old='firstUniprot', new="ID") 
  
  message('Checking output geneIDs...')
  
  if (!canonical.only){
    
    stopifnot(all(unique(dt$GN) %in% unique(clean.dt$GN)))
    
  } else {
    
    stopifnot(all(unique(dt[!GN %in% non.canonical.set$GN, GN]) %in% unique(clean.dt$GN)))
    
  }
  
   message('Returning clean output')
  return(clean.dt)
}
```
```{r}

lapply(secms.mat, function(x) dropNonCanonicalIsoforms(dt=x, canonical.only = F))
```


```{r}
lapply(seq_along(secms.f.mat), function(x,n,i){

  fwrite(x[[i]], sep='\t', ScriptAndDatedFileName(paste0(n[[i]], '.pcprophet.canonical.mat.tsv')))

},x=secms.f.mat, n=names(secms.f.mat))
```




#check that each gene ID is unique
# looks like there is ~100 genes in each
genes.oi <- secms.mat[[2]][,.N, by=GN] %>% 
  .[N > 1, GN]


secms.mat[[2]][GN %in% genes.oi,]


secms.mat[[2]][GN == 'ESYT2',]
#generatePCProphetMatrix(secms.list[[1]], gene="PG.Genes", uniprot="PG.UniProtIds", fracRegx = "[0-9]{1,2}")

lapply(seq_along(secms.mat), function(x,n,i){

  fwrite(x[[i]], sep='\t', ScriptAndDatedFileName(paste0(n[[i]], '.pcprophet.mat.tsv')))

},x=secms.mat, n=names(secms.mat))


secm
```

```{r}
lapply(secms.mat, function(x){
  
  genes.dup <- x[,.N, by=GN] %>% 
    .[N> 1, GN]
  x[GN %in% genes.dup,]
  x[GN == 'ARMC10',]
  
})


lapply(secms.list, colnames)
lapply(secms.list, function(x){
  
  x[PG.Genes == 'ARMC10',]
})
```

```{r}

```

```{r}
# read in PCprophet  input matrices
old.secms.list <- list(ctrl_1=fread('/Users/martingordon/Documents/projects/100324_AForget_PCProphetPipeline/100324.makePCPFOrmatFiles_data/2024_10_04_Parental_052324_Report.mat.txt'),
                  ctrl_2=fread('/Users/martingordon/Documents/projects/100324_AForget_PCProphetPipeline/100324.makePCPFOrmatFiles_data/2024_10_04_Parental_091024_Report.mat.txt'),
                  ctrl_3=fread('/Users/martingordon/Documents/projects/100324_AForget_PCProphetPipeline/100824.PCP.perturbSeq_data/2024_10_08_Parental_100724_Report.mat.txt'))



lapply(old.secms.list, function(x){
  
  genes.dup <- x[,.N, by=GN] %>% 
    .[N> 1, GN]
  
  x[GN %in% genes.dup,]
})


lapply(old.secms.list, function(x){
  
  x[, gene := multiUniprots2multiGenes(ID, species='HUMAN')]
})


lapply(old.secms.list, function(x){
  
    genes.dup <- x[,.N, by=GN] %>% 
    .[N> 1, GN]
  
  x[GN %in% genes.dup, .(GN,gene,ID)]
})


multiUniprots2multiGenes()
```

ˆ