---
title: "050625_makePlots"
author: "Martin Gordon"
date: "2025-05-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Plots to generate to visualize as part of the HARC meeting thursday
```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)
library(stringr)
library(viridis)
library(pracma) # needed for the peak finding algorithm
library(magick)
library(RColorBrewer)
#library(UniProt.ws) #cant load this packge; get naming conventions another way

# source Bens scripts so we can do a quick QC assessment of the input data before running
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/UniprotIDMapping.R") # map hu to mm 
source("~/Documents/utils/bp_utils/enrichmentTestFunctions.R")

# flag consec proteins
source("~/Documents/utils/bp_utils/SEC_MS_utils.localMods.R")
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")


source('~/Documents/utils/mg_utils/r_utils/plottingHelperFunctions.R')


de.pal <- c("dodgerblue2", "white","#E31A1C")
```

Read in the normalized SEC_MS data 
```{r}
sec.long <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_HIV_CD4T_TwoDonors_data/2025_05_06_sec.long.donor1donor2.normalized.interpolated.scaled.csv.gz')
apms.int <-  fread('~/Documents/projects/040825_MMuralidharan_PCProphet_HIV_CD4Tcells/040825_preparePCProphetInput_data/2025_04_11_yager.ints.toCompare.csv.gz') # Yager paper data

yager.hits <- apms.int[bestMistScore > .75, .(protein1, protein2, gene1, gene2)]

viral.prots <- sec.long[org != 'Homo sapiens', unique(protein)]

robyn.complexes <- setDT(readxl::read_xlsx('./docs/Complexes_Database_RMK_20250506.xlsx'))
#filter to just described complexes 
robyn.complexes <- robyn.complexes[, (.SD), .SDcols = !patterns('[.]PG[.]Quantity')]
ro
# differential peaks
peakClusters <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_donor1donor2.de.anova.results.csv.gz')

peakClusters[treatmentDiff.p < 0.05 & protein %in% viral]



getQualitati
```
color palette 
```{r, col}
donor.col <- brewer.pal(n=2,name='Pastel2')

 
status.col <- brewer.pal(n=3,name='Pastel1')
names(status.col) <- c('infected', 'uninfected', 'challenged')

 
status.col <-  c("#E41A1C","#377EB8","#4DAF4A")
names(status.col) <-  c('infected', 'uninfected', 'challenged')
```

tidy the complex info

```{r}
robyn.complexes <- robyn.complexes[, .(protein=PG.ProteinAccessions, gene=PG.Genes, org=PG.Organisms, 
                    complexID=`Complex ID...84`, subcomplexID=`SubComplex ID...85`)]

complexes.oi <- unique(robyn.complexes$complexID)
robyn.complexes <- robyn.complexes[complexID %in% complexes.oi & !is.na(complexID)]
```

create the plot data 
```{r}
plot.dt <- merge(sec.long[, .(sample, protein, gene, org, mw, fraction, status, donor,intensity_totalScaled)], robyn.complexes, by=c('protein', 'gene', 'org'), all.x=T)
plot.dt[, donor := as.character(donor)]
plot.dt[, status := factor(status, levels=c('uninfected', 'challenged', 'infected'))]
```

Get our functions organised to plot these different complexes
```{r}
customintensityHeatmaps <- function(intMats, 
                                    intensityName = "Relative Protein Intensity", 
                                    sample='norm_CD4_infected_1', 
                                    proteinsOI, 
                                    split_rows=FALSE,
                                    topOfColorRange = 0.45,...){
  denom = 50 / topOfColorRange
  colorFun <- circlize::colorRamp2(breaks = (0:50)/denom, colors = viridis(51,direction = 1))
  
  subMat <- intMats[rownames(intMats) %in% proteinsOI$protein,]
  print(rownames(subMat))
  
  rowgrp <- proteinsOI[match(rownames(subMat), protein), info]
  # convert names to subset we want
  #print(rownames(subMat))
  rownames(subMat) <- comb.mapper[match(gsub('[;].+', '', rownames(subMat)), comb.mapper$protein), simplifiedGene]
  subMat <- subMat[!is.na(rownames(subMat)),]
  
  hml <- Heatmap (subMat,
                  name = intensityName,
                  cluster_rows = TRUE,
                  #row_split = rowgrp,
                  border=T,
                  row_dend_reorder = FALSE,
                  cluster_columns = FALSE,
                  col = colorFun,
                  show_row_names = ifelse(nrow(subMat) > 100, FALSE, TRUE),
                  row_names_side = "left",
                  column_names_gp = gpar(fontsize = 10),
                  column_labels = ifelse(as.integer(colnames(subMat)) %% 5 == 0, colnames(subMat), ""),
                  column_title = sample,
                  #first only:
                  show_heatmap_legend = TRUE, 
                 # row_title = sprintf ("%d Proteins", nrow(subMat)),
                  ...
  )
  return(hml)
}
```

ridgeline plots

```{r, fig.width=10,fig.height=6}

# plot the donors on top of each other... color by group maybe? use lines to distinguish groups
p <- ggplot(plot.dt[complexID == 'CCC/WASH'], aes(x = fraction, y=gene, color=status)) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled), alpha=0.3, fill='white', scale = 8) +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=3, aes(color='maroon')) +
    #scale_fill(low = "navy", mid = "gray98", high = "firebrick", midpoint = log2(1), name = "log2 mass ratio") +
    scale_color_manual(values=status.col) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) +
    facet_grid(subcomplexID~donor, scales='free_y') + 
    theme_bw() +
    theme(strip.text.y=element_text(size=10),
          strip.text=element_text(color='white'),
          strip.background = element_rect(colour = "white", fill = "grey80")) +
  guides(linetype=guide_legend('donor'))
p
BackupAsPDF(p, 'test.ridgeplot')
```
Now try the same plot with fill
```{r, fig.width=10,fig.height=6}

# plot the donors on top of each other... color by group maybe? use lines to distinguish groups
p <- ggplot(plot.dt[complexID == 'CCC/WASH'], aes(x = fraction, y=gene, fill=status)) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled), alpha=0.4,  scale = 8) +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=3, aes(color='maroon')) +
    #scale_fill(low = "navy", mid = "gray98", high = "firebrick", midpoint = log2(1), name = "log2 mass ratio") +
    scale_fill_manual(values=status.col) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) +
    facet_grid(subcomplexID~donor, scales='free_y') + 
    theme_bw() +
    theme(strip.text.y=element_text(size=10),
          strip.text=element_text(color='white'),
          strip.background = element_rect(colour = "white", fill = "grey80")) +
  guides(linetype=guide_legend('donor'))
p
BackupAsPDF(p, 'test.ridgeplot.fill')
```
Plot the viral complex for all things

```{r, fig.width=10,fig.height=6}
# plot the donors on top of each other... color by group maybe? use lines to distinguish groups
p <- ggplot(plot.dt[protein %in% viral.prots,], aes(x = fraction, y=gene, fill=status)) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled), alpha=0.4,  scale = 8) +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=3, aes(color='maroon')) +
    #scale_fill(low = "navy", mid = "gray98", high = "firebrick", midpoint = log2(1), name = "log2 mass ratio") +
    scale_fill_manual(values=status.col) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) +
    facet_grid(subcomplexID~donor, scales='free_y') + 
    theme_bw() +
    theme(strip.text.y=element_text(size=10),
          strip.text=element_text(color='white'),
          strip.background = element_rect(colour = "white", fill = "grey80")) +
  guides(linetype=guide_legend('donor'))
p

BackupAsPDF(p, 'viralprots.ridgeplot.fill')
```
plot the linechart for the 

```{r, fig.width=10,fig.height=6}
# plot the donors on top of each other... color by group maybe? use lines to distinguish groups
p <- ggplot(plot.dt[protein %in% viral.prots,], aes(x = fraction, y=gene, color=status)) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled), alpha=0.4, fill='white', scale = 8) +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=3, aes(color='maroon')) +
    #scale_fill(low = "navy", mid = "gray98", high = "firebrick", midpoint = log2(1), name = "log2 mass ratio") +
    scale_color_manual(values=status.col) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) +
    facet_grid(subcomplexID~donor, scales='free_y') + 
    theme_bw() +
    theme(strip.text.y=element_text(size=10),
          strip.text=element_text(color='white'),
          strip.background = element_rect(colour = "white", fill = "grey80")) +
  guides(linetype=guide_legend('donor'))
p
BackupAsPDF(p, 'viralprots.ridgeplot.line')
```

Read in the new complex info; iterate over the complexes to plot

```{r}
robyn.complexes <- setDT(readxl::read_xlsx('./docs/Complexes_Database_RMK_20250506.xlsx'))

robyn.complexes <- robyn.complexes[, .(protein=PG.ProteinAccessions, gene=PG.Genes, org=PG.Organisms, 
                    complexID=`Complex ID...84`, subcomplexID=`SubComplex ID...85`)]

complexes.oi <- unique(robyn.complexes$complexID)
robyn.complexes <- robyn.complexes[complexID %in% complexes.oi & !is.na(complexID)]
```

create the plot data 
```{r}
plot.dt <- merge(sec.long[, .(sample, protein, gene, org, mw, fraction, status, donor,intensity_totalScaled)], robyn.complexes, by=c('protein', 'gene', 'org'), all.x=T)
plot.dt[, donor := as.character(donor)]
plot.dt[, status := factor(status, levels=c('uninfected', 'challenged', 'infected'))]
```
## 05-06-25 
New set of complexes  and use BPs sec long set so we can look at visualizations
```{r}
sec.long <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/docs/2025_05_07_OutlierFractionsRemoved_interpNorm_sec.csv.gz')
complexes.oi <- unique(robyn.complexes$complexID)
robyn.complexes <- robyn.complexes[complexID %in% complexes.oi & !is.na(complexID)]

plot.dt <- merge(sec.long[, .(sample, protein, gene, org, mw, fraction, status, donor, normInts, intensity_totalScaled)], robyn.complexes, by=c('protein', 'gene', 'org'), all.x=T)
plot.dt[, donor := as.character(donor)]
plot.dt[, status := factor(status, levels=c('uninfected', 'challenged', 'infected'))]
```

iterate over the list of protein complexes; split by subcomplex id and plot both

```{r, elution-plots, fig.width=10,fig.height=6}
plot.dt[is.na(subcomplexID), subcomplexID := '']

lapply(complexes.oi, function(comp){
  
  fname <- gsub(' |[/]', '_', comp)
  # plot the donors on top of each other... color by group maybe? use lines to distinguish groups
  p <- ggplot(plot.dt[complexID == comp,], aes(x = fraction, y=gene, fill=status)) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled), alpha=0.4,  scale = 8) +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=2, aes(color='maroon'), alpha=0.4) +
    scale_fill_manual(values=status.col) +
    labs(title=paste0(comp)) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) + # to reorder colnames
    facet_grid(subcomplexID~donor, scales='free_y') + 
    theme_bw() +
    theme(strip.text.y=element_text(size=10),
          strip.text=element_text(color='white'),
          strip.background = element_rect(colour = "white", fill = "grey80")) +
  guides(linetype=guide_legend('donor'))
  p
  print(paste0(fname, '.complexSubunits.ridgeline'))
  BackupAsPDF(p, paste0(fname, '.complexSubunits.ridgeline'))
})
```

```{r, elution-plots, fig.width=10,fig.height=12}
replotComplexes <- c('CRL', 'Proteasome', 'NPC', 'Dynactin', 'RNAPII')

lapply(complexes.oi, function(comp){
  
  fname <- gsub(' |[/]', '_', comp)
  # plot the donors on top of each other... color by group maybe? use lines to distinguish groups
  p <- ggplot(plot.dt[complexID == comp,], aes(x = fraction, y=gene, fill=status)) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled), alpha=0.4,  scale = 8) +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=2, aes(color='maroon'), alpha=0.4) +
    scale_fill_manual(values=status.col) +
    labs(title=paste0(comp)) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) + # to reorder colnames
    facet_grid(subcomplexID~donor, scales='free_y') + 
    theme_bw() +
    theme(strip.text.y=element_text(size=10),
          strip.text=element_text(color='white'),
          strip.background = element_rect(colour = "white", fill = "grey80")) +
  guides(linetype=guide_legend('donor'))
  print(p)
  #print(paste0(fname, '.complexSubunits.ridgeline'))
  #BackupAsPDF(p, paste0(fname, '.complexSubunits.ridgeline'))
})
```

To tidy  (just make longer for now )
- RNAp11 complex
- Proteasome
- NPC complex
- Dynactin complex
- CRL complex

heatmaps of absolute intensity
color the complexes by intensity and replot; wont be able to ovelay; for now use donor 1
elution profiles of the Yager interactors with our data
```{r, ridgeline-ints, fig.width=11,fig.height=6}
#
plot.dt[, totalInts := sum(normInts, na.rm=T), by=.(protein, sample)]
plot.dt[is.na(subcomplexID), subcomplexID := complexID]

lapply('BLOC1', function(comp){
  
  fname <- gsub(' |[/]', '_', comp)
  # plot the donors on top of each other... color by group maybe? use lines to distinguish groups
  p <- ggplot(plot.dt[complexID == comp & donor == '1',], aes(x = fraction, y=gene, fill=intensity)) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled), scale = 8) +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=2, aes(color='maroon'), alpha=0.4) +
    scale_fill_viridis_c(name = "log2 total normalized intensity", option = "C") +
    labs(title=paste0(comp)) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) + # to reorder colnames
    facet_grid(subcomplexID~status, scales='free_y') + 
    theme_bw() +
    theme(strip.text=element_text(color='white', size=9),
          strip.background = element_rect(colour = "white", fill = "grey80"))
 
  print(p)
  #print(paste0(fname, '.complexSubunits.ridgeline'))
  #BackupAsPDF(p, paste0(fname, '.complexSubunits.ridgeline.totalInts'))
})
```

## 05-07-25 Evening 
Plot the remaining HIV complexes reported by the group
Read in Robyns latest data file and plot

```{r}

```







add Yager interactors group to our data
plot both the overlap per donor and also the total intensity color
```{r, ridgeline-ints, fig.width=11,fig.height=6}
# only 7 viral proteins?? REV is filtered out at some step of the filtering... check if it only had one REV
# REV seems to have been filtered out at the peptide filtering s
sec.long[org != 'Homo sapiens', unique(gene)]
sec.ori[org != 'Homo sapiens', unique(gene)]


robyn.complexes[gene == 'rev']

sec.ori <- fread('050625_HIV_CD4T_TwoDonors_data/2025_05_06_msstats.input.singleProtein.filtered.csv.gz')
sec.ori[ProteinName %in% viral.prots,]

sec.ori
sec.ori[ProteinName == 'Q8ADQ7',]

for (i in unique(yager.hits$gene1)){
  
  sig.interactors <- yager.hits[gene1 == i, protein2]
  prot.oi <- rbind(plot.dt[protein %in% sig.interactors,],
                   plot.dt[gene == tolower(i),])
  
  p <- ggplot (plot.dt[protein %in% prot.oi$protein,], aes(x = fraction, y = gene)) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled, fill=status),alpha=0.4, scale = 8) +
    labs(title=paste0(i, ' interactors (Jäger et al. 2011)')) +
    facet_grid(.~donor) +
    scale_fill_manual(values=status.col) +
    theme_bw() +
    theme(strip.text.y = element_text(size = 5.5))
  
   print(p)
   BackupAsPDF(p, paste0(i, '.jagerinteractors.ridgeline.'))
}
```

Investigate why some complexes dont plot
```{r}
# dont know why RNAp is changing; like a protein has two summarized 
subdt <- plot.dt[donor == 1 & complexID == 'RNAPII',]
subdt[is.na(intensity_totalScaled), intensity_totalScaled := 0]
subdt %>% str()

 ggplot(subdt, aes(x = fraction, y=gene, fill=log2(totalInts))) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled)) +
    #geom_vline(xintercept=c(10,20,30,40,50,60), linetype=2, aes(color='maroon'), alpha=0.4) +
    #scale_fill_viridis_c(name = "log2 total normalized intensity", option = "C") +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) + # to reorder colnames
    #facet_grid(subcomplexID~status, scales='free_y') + 
    theme_bw() +
    theme(strip.text.y=element_text(size=10),
          strip.text=element_text(color='white'),
          strip.background = element_rect(colour = "white", fill = "grey80"))
```

```{r}

```


```{r}


p <- ggplot (plot.dt[complexID == 'CRL' & donor == 1,], aes(x = fraction, y = gene)) +
  ggridges::geom_ridgeline_gradient(aes(height = intensity_totalScaled, fill =  log2(totalInts+1)), scale = 10) +
  #scale_fill_gradient2(low = "navy", mid = "gray98", high = "firebrick", name = "norm Ints")+
  scale_fill_viridis_c(name = "log2 normalized intensity", option = "C") +
  scale_y_discrete(name = "protein") +
  facet_grid(.~status) +
  theme_bw()
p
```

Work on a heatmap function to plot the heatmap list of the different complexes; use the scaledheatmap 
```{r}
sec.long$intensity_totalScaled <- NULL
# this is total scaled
intMats <- scaledIntensityMatrices(sec.long, scaleDenom = 'max')

sec.long$intensity_totalScaled <- NULL
donor1_intMats <- scaledIntensityMatrices(secLong.dt = sec.long[donor == 1])
```

adjust these heatmap plots 
```{r}
# denom to make 50 different color scales
complexIntensityHeatmaps <- function(intMats, complexDT, complexOI,  intensityName = "Relative Protein Intensity", topOfColorRange = 0.45,...){
  denom = 50 / topOfColorRange
  colorFun <- circlize::colorRamp2(breaks = (0:50)/denom, colors = viridis(51,direction = 1))
  # get the individual sample lists
  samples <- names(intMats)
  
  sample <- samples[1]
  
  mapping.dt <- complexDT[complexID == complexOI,]
  mapping.dt[is.na(subcomplexID), subcomplexID:= complexID]
  subMat <- intMats[[sample]][rownames(intMats[[sample]]) %in% mapping.dt$protein,]
  
  rowgrp <- mapping.dt[match(rownames(subMat), protein), subcomplexID]
  
  rowOrder <-  mapping.dt[, gene]
  
  # convert names to subset we want
  #print(rownames(subMat))
  rownames(subMat) <- mapping.dt[match(gsub('[;].+', '', rownames(subMat)), protein), gene]
  
  hml <- Heatmap (subMat,
                  name = intensityName,
                  cluster_rows = FALSE,
                  row_order = rowOrder, # enforce roworder here, later heatmaps will follow this clustering
                  show_row_dend = F,
                  row_split = rowgrp, 
                  cluster_columns = FALSE,
                  col = colorFun ,
                  show_row_names = TRUE,
                  row_names_side = "left",
                  row_names_gp = gpar(fontsize = 8),
                  row_title_rot = 0,
                  column_names_gp = gpar(fontsize = 5),
                 # column_labels = ifelse(as.integer(colnames(intMats[[sample]])) %% 5 == 0, colnames(intMats[[sample]]), ""),
                  column_title = sample,
                  column_title_gp = gpar(fontsize=12, fontface='bold'),
                  #first only:
                  show_heatmap_legend = TRUE, 
                  #row_title = sprintf ("%d Proteins", nrow(subMat)),
                  ...
  )
  
  # first one gets row title, also gets a legend
  
  if (length(intMats) > 1){
    for (sample in samples[2:length(samples)]){
      hml <- hml + Heatmap (intMats[[sample]][rownames(intMats[[sample]]) %in% mapping.dt$protein,],
                            name = intensityName,
                            cluster_rows = FALSE,
                            row_dend_reorder = FALSE,
                            cluster_columns = FALSE,
                            col = colorFun ,
                            show_row_names = FALSE, 
                            column_names_gp = gpar(fontsize = 5),
                            #column_labels = ifelse(as.integer(colnames(intMats[[sample]])) %% 5 == 0, colnames(intMats[[sample]]), ""),
                            column_title = sample,
                            column_title_gp = gpar(fontsize=12, fontface='bold'),
                            # 
                            show_heatmap_legend = FALSE,
                            ...)      
    }
  }
  return (hml)
}
```

```{r}
fwrite(robyn.complexes,  ScriptAndDatedFileName('robynComplexes.formattd.csv.gz'))
```
Heatmap of all samples
```{r, test-heatmap, fig.height=6, fig.width=12}
robyn.complexes[complexID == 'AP1 complex']

hm <- complexIntensityHeatmaps(intMats = donor1_intMats, 
                         complexDT = robyn.complexes,
                         complexOI = 'AP1 complex', 
                         border=T)

BackupAsPDF(hm, 'testheatmap')
```

Example ridgeline plot
```{r}
# overlay the infected vs uninfected 
p <- ggplot(plot.dt[!grepl('challenged', sample),], aes(x = fraction, y=gene, fill=sample)) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled), alpha=0.8, scale = 8) +
    #scale_fill(low = "navy", mid = "gray98", high = "firebrick", midpoint = log2(1), name = "log2 mass ratio") +
    scale_fill_brewer(type='qual', palette=4) +
    scale_y_discrete(name = "protein", limits=rev(new.complexes$gene)) +
    #facet_grid(.~sample, scales='free_y') + 
    theme_bw() +
    theme(strip.text.y=element_text(size=10))
p
BackupAsPDF(p, 'geneSet.infectedVsuninfected.overlap.elution')
```
**todo**
Create plots of normalized intensity matrices (log2 transform)

```{r}
View(scaledIntensityMatrices)
```


I think for the matrices, we can plan to plot either the CD4 infected, or plot all the heatmaps beside each other I think the best way is to look at the IntensityHeatmap function and ape that..
```{r, fig.height=8, fig.width=10}
plot.dt <- merge(sec.long[, .(sample, protein, gene, org, mw, fraction, status, donor,intensity_totalScaled)], robyn.complexes, by=c('protein', 'gene', 'org'), all.x=T)
plot.dt[, donor := as.character(donor)]
plot.dt[, status := factor(status, levels=c('uninfected', 'challenged', 'infected'))]


# for first pass, lets plot the 
plot.dt$complexID %>% unique()
plot.dt[complexID == 'proteomsome']

library(RColorBrewer)

# plot the donors on top of each other... color by group maybe? use lines to distinguish groups
p <- ggplot(plot.dt[complexID == 'CCC/WASH'], aes(x = fraction, y=gene, fill=donor, linetype=as.character(donor))) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled), alpha=0.6, scale = 8) +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=3, aes(color='maroon'), alpha=0.6) +
    #scale_fill(low = "navy", mid = "gray98", high = "firebrick", midpoint = log2(1), name = "log2 mass ratio") +
    scale_fill_manual(values=donor.col) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) +
    facet_grid(subcomplexID~status, scales='free_y') + 
    theme_bw() +
    theme(strip.text.y=element_text(size=10),
          strip.background=element_rect()) +
  guides(linetype=guide_legend('donor'))
p
BackupAsPDF(p ,'ccc.wash.treatmentOverlap.ridgeline.test')


# plot the donors on top of each other... color by group maybe? use lines to distinguish groups
p <- ggplot(plot.dt[complexID == 'CCC/WASH'], aes(x = fraction, y=gene, fill=status, linetype=as.character(donor))) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled), alpha=0.6, scale = 8) +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=3, aes(color='maroon'), alpha=0.6) +
    #scale_fill(low = "navy", mid = "gray98", high = "firebrick", midpoint = log2(1), name = "log2 mass ratio") +
    scale_fill_manual(values=status.col) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) +
    facet_grid(subcomplexID~donor, scales='free_y') + 
    theme_bw() +
    theme(strip.text.y=element_text(size=10),
          strip.background = element_rect(colour = "black", fill = "white")) +
  guides(linetype=guide_legend('donor'))
p
BackupAsPDF(p ,'ccc.wash.donorOverlap.ridgeline.test')
```

```{r}
p <- ggplot(plot.dt[complexID == 'CCC/WASH'], aes(x = fraction, y=gene, color=donor, linetype=as.character(donor))) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled), alpha=0.6, scale = 8) +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=3, aes(color='maroon'), alpha=0.6) +
    #scale_fill(low = "navy", mid = "gray98", high = "firebrick", midpoint = log2(1), name = "log2 mass ratio") +
    scale_fill_manual(values=donor.col) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) +
    facet_grid(subcomplexID~status, scales='free_y') + 
    theme_bw() +
    theme(strip.text.y=element_text(size=10),
          strip.background=element_rect()) +
  guides(linetype=guide_legend('donor'))
p
```


