---
title: "050625_peakDetectionAndDifferentialExpression"
author: "Martin Gordon"
date: "2025-05-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Now happy with a set of input proteins for each of the samples, we need to perform peak detection and differential testing on each of the samples
My first inking is that we will need to try the peak alignment for the samples, then proceed with differential testing... I don't have a good handle on the sample clustering/intra-sample variation atm, but it may be the case that samples need to be processed donor-wise and then search for overlap in ids/differential proteins/complexes

```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)
library(stringr)
library(viridis)
library(pracma) # needed for the peak finding algorithm
library(magick)
#library(UniProt.ws) #cant load this packge; get naming conventions another way

# source Bens scripts so we can do a quick QC assessment of the input data before running
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/UniprotIDMapping.R") # map hu to mm 
source("~/Documents/utils/bp_utils/enrichmentTestFunctions.R")

# flag consec proteins
source("~/Documents/utils/bp_utils/SEC_MS_utils.localMods.R")
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")
```

read in the processed SEC_MS data and lets look to see if we can identify peaks, and finally, differential peaks between the conditions

```{r, readFiles}
sec.long <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_HIV_CD4T_TwoDonors_data/2025_05_06_sec.long.donor1donor2.normalized.interpolated.scaled.csv.gz')
```

```{r, peakDetection}
# we wan
normInt.mats <- scaledIntensityMatrices(sec.long)

peakTables <- lapply(normInt.mats, goodPeaksTableFromIntensityMatrix, minCV = -Inf)
allPeaks <- rbindlist(peakTables, idcol = "sample")
#fwrite(allPeaks, ScriptAndDatedFileName('donor1donor2.detectedPeaks.csv.gz'))

allPeaks <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_06_donor1donor2.detectedPeaks.csv.gz')
```

This looks really good; shows the peak detection algorithm is doing its job to seperate good and bad peaks
It seems that fract 60 in cd4 infected2 just has globally more proteins than the surrounding 
Speak to BP tomorrow about the above fraction; may want to drop and just interpolate to remove this inflation

```{r, fig.width = 5, fig.height = 7}
# udpate height for more samples
p <- ggplot(allPeaks, aes(x= peakLocation, fill = goodPeak)) + 
  geom_bar() + 
  facet_grid(sample~.) +
  theme_bw()
p
BackupAsPDF(p, "peakCountsPerFraction.barchart")
```
Dont really think aligning peaks across these fractions is going to do a good job...  maybe for now we can just 
```{r, peak-fraction-adjustments}

standardizeAllPeakTablesToStandard(peakTables, sec.long, standardIdx = "cd4_uninfected_2", fitPortion = 0.65, minPeaksPerFraction = 45, startFitAtPeak = 20)

#fwrite(sec.long, ScriptAndDatedFileName('hiv.donor1donor2.sec.ms.wStandardisedFractions.csv.gz'))
```
To me, the infected and the challenged look good and similiar to the uninfected for donor2
Undo the standardisation for both of these samples as they look good.. not sure what to do for donor1 for now.... proceed as is and speak to BP tomorrow morning.

```{r, fix-peak-alignments}
peakTables$cd4_infected_2[, cofmN.standardized := cofmN]
sec.long[sample == "cd4_infected_2", standardFraction := fraction]

peakTables$cd4_challenged_2[, cofmN.standardized := cofmN]
sec.long[sample == "cd4_challenged_2", standardFraction := fraction]
```

We dont have the molecular weight peak standards for the second file; considering these files are run on very different dates
I dont think we want to use the molecular weight stadnards for filtering and should just proceed with everything
Label with their mass, but dont use this to define good/bad peaks
```{r, mw-standards}
mw.standards <- fread('../040825_MMuralidharan_PCProphet_HIV_CD4Tcells/data/HIV_Infection_CD4T_cells/Donor-1-032625/PCprophet/cal_SRT.txt')
setnames(mw.standards, new=c('fraction', 'mw'))

# want the standards in daltons (right?)
mw.standards[, mw := mw*1000]
mc <- calculateFractionMassConverters(mw.standards)
```

```{r mass-ratio-labelling}
mw.dt <- sec.long[,.(protein, mw=as.numeric(mw))] %>% 
  unique()
# for joining and calculating mass below
setkey(mw.dt, protein)
mw.dt <- mw.dt[!is.na(mw)] # not including mass for protein groups
proteinMW <- mw.dt

# label fractions by mass, and fraction/proteins by mass ratio, aka log2MassNumber
sec.long[, fractionMass := mc$fraction2Mass(standardFraction)]
# this is kind 
sec.long[, log2MassNumber := log2(fractionMass/mw.dt[protein, .(mw), on = "protein"])]


# add a label for the filtered good peak and decide later if we want to keep
scorePeakByMassShift <- function(peak.dt, mw){
  peak.dt[proteinMW, fiftyPCFraction := mc$mass2Fraction(1.2 * mw), on = "protein"]
  peak.dt[!is.na(fiftyPCFraction), filtteredGoodPeak := goodPeak & cofmN.standardized < fiftyPCFraction]
  peak.dt
}

# need protein MW to use this as hard coded in the function
purrr::walk(peakTables, scorePeakByMassShift, mw)
```
Save the output, plot the number of good peaks before and after filtering 
```{r, fig.width = 5, fig.height = 7}
# udpate height for more samples
p <- ggplot(allPeaks, aes(x= peakLocation, fill = filtteredGoodPeak)) + 
  geom_bar() + 
  facet_grid(sample~.) +
  theme_bw()
p
BackupAsPDF(p, "filteredPeakCountsPerFraction.barchart")
```

```{r}
#fwrite(allPeaks, ScriptAndDatedFileName("allPeaks.donor1donor2.monomerMassLabelled.csv.gz"))
allPeaks <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_allPeaks.donor1donor2.monomerMassLabelled.csv.gz')
```

plot the MW standards for both both donors
```{r, proteinPeaks, fig.height=8, fig.width=14}
# merge the mw info from the spectronaut search to the table
allPeaks[mw.dt, proteinMW := i.mw, on = "protein"]
# no obvious impacts here... cant really say atm why these peaks behave as they do
allPeaks[, proteinMW := as.numeric(proteinMW)]
allPeaks[!is.na(proteinMW),.N, by=sample]


g <- ggplot(allPeaks[grepl('_1$', sample), .SD[which.max(peakHeight)], by = .(protein, sample)  ], aes( x  = as.factor(sprintf("%02d", as.integer(peakLocation))), y = proteinMW, color = log10(proteinMW))) + 
  #geom_density_2d() + 
  scale_y_log10() +
  geom_boxplot(outlier.shape = NA) + 
  ggforce::geom_sina(shape = ".",scale = "area")+
  geom_point(data = mw.standards[, .(peakLocation = fraction, proteinMW = mw)], color = "red") + 
  rotate.x.axis.text +
  scale_color_viridis_c() + 
  facet_wrap(~sample, ncol = 2) + 
  ggtitle("Protein Peaks")
g
BackupAsPDF(g, 'proteinPeaksVsMWstandards.donor1.boxplots')

g <- ggplot(allPeaks[grepl('_2$', sample), .SD[which.max(peakHeight)], by = .(protein, sample)  ], aes( x  = as.factor(sprintf("%02d", as.integer(peakLocation))), y = proteinMW, color = log10(proteinMW))) + 
  #geom_density_2d() + 
  scale_y_log10() +
  geom_boxplot(outlier.shape = NA) + 
  ggforce::geom_sina(shape = ".",scale = "area")+
  geom_point(data = mw.standards[, .(peakLocation = fraction, proteinMW = mw)], color = "red") + 
  rotate.x.axis.text +
  scale_color_viridis_c() + 
  facet_wrap(~sample, ncol = 2) + 
  ggtitle("Protein Peaks")
g
BackupAsPDF(g, 'proteinPeaksVsMWstandards.donor2.boxplots')
```


**Question**
Why do we seem to see a difference in plotting density??
We actually have less proteins...what are we actually plotting here?
Wont figure this out tonight, try just generate some plots of the different complexes and share tomorrow...
```{r}
allPeaks[goodPeak == TRUE, .N, by=sample]
sec.long[,length(unique(protein)), by=sample]
allPeaks[ , .SD[which.max(peakHeight)], by = .(protein, sample)][!is.na(peakLocation),.N, by=sample]

allPeaks[goodPeak == TRUE, .N, by=sample]
allPeaks[ , .SD[which.max(peakHeight)], by = .(protein, sample)  ]
```

Leave as is for now, not using the MW for filtering, so lets push forward
Run analysis both ways and if we get a better TP/FP ratio, can filter out these masses

## GS and decoy set

First read in our ID mapper
```{r}
hu.string.dt <- fread('~/Documents/utils/mg_utils/data/stringDB/9606.protein.aliases.v12.0.txt.gz')
huString.mapper <- hu.string.dt[source == 'Ensembl_UniProt', .(protein=alias, stringID=`#string_protein_id`)]
huString.mapper[protein %in% sec.long$protein, unique(stringID)]

# save this file for future use
#fwrite(huString.mapper, ScriptAndDatedFileName('protein.string.idMapper.csv.gz'))
huString.mapper <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_protein.string.idMapper.csv.gz')

# also, read int he 
hiv.nodes <- fread('~/Documents/utils/mg_utils/data/stringDB/11676.protein.aliases.v10.5.txt.gz')
setnames(hiv.nodes, new=c('stringID', 'gene', 'db'))
hiv.nodes[, gene := toupper(gene)]
hiv.nodes$db <- NULL # not used for now
```

```{r, gs-decoy-sets eval = FALSE}
#need to merge our protein ids with 
gs.interactors <- goldStandardPairs(huString.mapper[protein %in% sec.long$protein, unique(stringID)], 
                                    corum.path = NULL,
                                    string.links.path = "~/Documents/utils/mg_utils/data/stringDB/9606.protein.physical.links.detailed.v12.0.txt.gz",
                                    string.info.path =  NULL,
                                    stringCombinedScoreThreshold = 250)
gs.decoys <- decoysFromString(huString.mapper[protein %in% sec.long$protein, unique(stringID)], 
                              links.path = "~/Documents/utils/mg_utils/data/stringDB/9606.protein.physical.links.detailed.v12.0.txt.gz",
                              info.path =  NULL)
gs.decoys[, source := 'string']

gs.comb.interactors <- rbind(gs.interactors, gs.decoys)
```

## peak-peak correlations

```{r, peak-correlations}
# just ensure a refresh
sec.long$intensity_totalScaled <- NULL
intMats <- scaledIntensityMatrices(sec.long)
intensityHeatmaps(intMats, border=T)

#saveRDS(intMats, ScriptAndDatedFileName('hiv.donor1donor2.norm.totalScaled.intsMats.rds'))
intMats <- readRDS('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_hiv.donor1donor2.norm.totalScaled.intsMats.rds')

normMatrices <- intMats

peakMatrices <- purrr::map2(normMatrices, peakTables, goodPeaksMatFromPeaksTable)
#corPeaks6 <- windowedCorrelation(npc1Mat, peaksMatrix, outerRadius = 6)
corPeaks.ls <- purrr::map2(normMatrices, peakMatrices, windowedCorrelation, outerRadius = 6)
allPeakCorrs <- rbindlist(purrr::map(corPeaks.ls, "cor"), idcol = "sample")
```

```{r annotate-peakDT}
anno.dt <- sec.long[, .(protein, gene)] %>% 
  unique()

allPeakCorrs[huString.mapper, gene1 := i.stringID, on = c(protein1 = "protein")]
allPeakCorrs[huString.mapper, gene2 := i.stringID, on = c(protein2 = "protein")]
allPeakCorrs[anno.dt, alias1 := i.gene, on = c(protein1 = "protein")]
allPeakCorrs[anno.dt, alias2 := i.gene, on = c(protein2 = "protein")]

# for now use protein as gene for the NA gene
#allPeakCorrs[is.na(gene1), gene1 := protein1]
#allPeakCorrs[is.na(gene2), gene2 := protein2]

#fwrite(allPeakCorrs, ScriptAndDatedFileName('donor1donor2.allByallPeakCorr.csv.gz'))
allPeakCorrs <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_donor1donor2.allByallPeakCorr.csv.gz')
```

Rerun with the peak filtering included and see how this looks with regards to TP/FP recovery

```{r}
temp <- names(peakTables)


lapply(peakTables, function(x){
  
  x[, originalGoodPeak := goodPeak]
  x[, goodPeak := filtteredGoodPeak]
})

f.peakMatrices <- purrr::map2(normMatrices, peakTables, goodPeaksMatFromPeaksTable)
#corPeaks6 <- windowedCorrelation(npc1Mat, peaksMatrix, outerRadius = 6)
f.corPeaks.ls <- purrr::map2(normMatrices, f.peakMatrices, windowedCorrelation, outerRadius = 6)
f.allPeakCorrs <- rbindlist(purrr::map(f.corPeaks.ls, "cor"), idcol = "sample")
```
```{r, annotate-filtered}
f.allPeakCorrs[huString.mapper, gene1 := i.stringID, on = c(protein1 = "protein")]
f.allPeakCorrs[huString.mapper, gene2 := i.stringID, on = c(protein2 = "protein")]
f.allPeakCorrs[anno.dt, alias1 := i.gene, on = c(protein1 = "protein")]
f.allPeakCorrs[anno.dt, alias2 := i.gene, on = c(protein2 = "protein")]

# for now use protein as gene for the NA gene
f.allPeakCorrs[is.na(gene1), gene1 := protein1]
f.allPeakCorrs[is.na(gene2), gene2 := protein2]

#fwrite(f.allPeakCorrs, ScriptAndDatedFileName('donor1donor2.allByallPeakCorr.filteredPeaks.csv.gz'))
```
Read in the original and filtered peak correlations

```{r}
allPeakCorrs <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_donor1donor2.allByallPeakCorr.csv.gz')
filteredPeakCorrs <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_donor1donor2.allByallPeakCorr.filteredPeaks.csv.gz')
```

Run the rest of the analysis on both
```{r}
peakCor.ls <- list(filtered=filteredPeakCorrs,
                   all=allPeakCorrs)
```
Label the genesets in each
```{r}
lapply(peakCor.ls, function(x){
  
  x[, gs := "unknown"]
  x[gs.interactors ,gs := "interactor" , on = c("gene1", "gene2")]
  x[gs.decoys ,gs := "decoy" , on = c("gene1", "gene2")]
})
```
plot the recovery of peaks in each
```{r,plot-correlations, fig.width=9,fig.height=6}

lapply(names(peakCor.ls), function(x){
  
  g <- ggplot(peakCor.ls[[x]][gene1 < gene2 & gs != 'unknown'], aes(x = sample, fill = gs)) +
    geom_bar() + 
    labs(title=paste0(x, ' Breakdown of detected interactors')) +
    theme_bw() + 
    scale_fill_manual(values=c('decoy'='grey70', 'interactor'='gold'))

  BackupAsPDF(g, paste0(x,'.interactorSet.breakdown.barplot'))
})
```
## scoring PPI based on peak-peak correlation
Setup
```{r}
allGenes <- huString.mapper$stringID %>% 
  unique()
denomInteractor <- nrow(gs.interactors[gene1 %in% allGenes & gene2 %in% allGenes])
denomDecoy <- nrow(gs.decoys[gene1 %in% allGenes & gene2 %in% allGenes])
```
Using log-likihood ratio approach; how more likey is the PPI derived from the truth set than the random set
```{r ppi-scoring}
lapply(names(peakCor.ls), function(x){
  scoreByGS(peakCor.ls[[x]], denomDecoy, denomInteractor, column = "corScore", groupByVariable = c("sample", "start"))
})
```
plotting the LLR; distance between the line is what we are scoring 
```{r}


# an example LL plot
startOI = 31

peakCor.ls[['filtered']]

sampleOI <- 'cd4_uninfected_2'

corScoreLabels <- function(corScore){
  1- signif(10^(-corScore), 2)
}

allwindows <- peakCor.ls[['filtered']]$start %>%
  unique() %>% 
  sort()

outDir <- './050625_peakDetectionAndDifferentialExpression_data/pdfs/logliklihood_plots/'

for (i in allwindows){
  
  p <- ggplot(peakCor.ls[['filtered']][start == i & sample == sampleOI], aes(x = corScore)) +
    geom_line(aes(y = log10IntRate, color = "interactor")) +
    #geom_segment(aes(y =   log10IntRate-log10RateRatioID, yend = log10IntRate,color = "ratio", xend = corScore)) + 
    geom_line(aes(y = log10DecoyRate, color = "decoy")) +
    scale_color_manual(values = c(decoy = "grey", interactor = "red", ratio = "purple")) +
    ggtitle(label= paste0(sampleOI, " fraction ", i)) + 
    scale_x_continuous(labels = corScoreLabels, name = "Pearson's R") + 
    scale_y_continuous(name = "Log Likelihood (i.e. what portion passes threshold)") + 
    theme_bw()
  BackupAsPDF(p, paste0('loglikelihood_plots/',i,'.decoyTruthRatio.linechart'))
}
```

LLR plots; see where we have the best descriminatory power
```{r, llr-plots, fig.width=11, fig.height=7}
lapply(names(peakCor.ls), function(x){
  
  g <- ggplot(peakCor.ls[[x]], aes(x = corScore, y = log10RateRatioID, color = start)) + geom_hline(yintercept = 0.0, lty = "dotted") + 
    geom_line(aes(group = start)) + 
    labs(title=x) +
    facet_wrap(~sample, ncol=3) + 
    scale_color_viridis_c(option = "H", name = "fraction") + 
    theme_bw()
  g
  BackupAsPDF(g, paste0(x,'.llratios.linechart'))
})
```

peak summarization to proteins 
focus on. the filtereSet
```{r, protein-summarization}
allPeakCorrs <- peakCor.ls[['filtered']]

stoichDevThreshold <- log2(1 + 0.2)
setorder(allPeakCorrs, sample, -log10RateRatioID )

summary <- allPeakCorrs[, .SD[1], by= .(sample, protein1, protein2, gene1, gene2, alias1=gsub('[.].+','',alias1), alias2=sub('[.].+','',alias2), prot1Peak, prot2Peak, gs) # best correlation per pair of peaks
                          ][, .(sumLLRatio = sum(log10RateRatioID[abs(log2Ratio - log2Ratio[1]) < stoichDevThreshold]), # only sum peaks that are within 20% of the first
                                numPeaks = .N,
                                numSameStoichPeaks  = sum (abs(log2Ratio - log2Ratio[1]) < stoichDevThreshold)),
                            by = .(sample, protein1, protein2, gene1, gene2, alias1, alias2, gs)
                            ]
setorder(summary, sample, -sumLLRatio)
fwrite(summary[gene1 < gene2], ScriptAndDatedFileName("filteredset.peakSimilarityScores.csv.gz"))

allPeakCorrs <- peakCor.ls[['all']]
setorder(allPeakCorrs, sample, -log10RateRatioID )

summary <- allPeakCorrs[, .SD[1], by= .(sample, protein1, protein2, gene1, gene2, alias1=gsub('[.].+','',alias1), alias2=sub('[.].+','',alias2), prot1Peak, prot2Peak, gs) # best correlation per pair of peaks
                          ][, .(sumLLRatio = sum(log10RateRatioID[abs(log2Ratio - log2Ratio[1]) < stoichDevThreshold]), # only sum peaks that are within 20% of the first
                                numPeaks = .N,
                                numSameStoichPeaks  = sum (abs(log2Ratio - log2Ratio[1]) < stoichDevThreshold)),
                            by = .(sample, protein1, protein2, gene1, gene2, alias1, alias2, gs)
                            ]
setorder(summary, sample, -sumLLRatio)
fwrite(summary[gene1 < gene2], ScriptAndDatedFileName("fullSet.peakSimilarityScores.csv.gz"))
```

```{r}
sec.long <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_HIV_CD4T_TwoDonors_data/2025_05_06_sec.long.donor1donor2.normalized.interpolated.scaled.csv.gz')

anno.dt <- sec.long[, .(protein, gene)] %>% 
  unique()

summary <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_filteredset.peakSimilarityScores.csv.gz')
allPeakCorrs <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_donor1donor2.allByallPeakCorr.csv.gz')
normMatrices <- readRDS('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_hiv.donor1donor2.norm.totalScaled.intsMats.rds')
```


```{r}
cosMats <- pbapply::pblapply(normMatrices, cosineMatrix)
```
```{r}

cosineMat2Table <- function(mat, threshold = 0.9){
  idx <- which(mat > threshold , arr.ind = TRUE)
  
  dt <- as.data.table(idx,
                    keep.rownames = TRUE)
  setnames(dt, "rn", "protein1")
  dt[, protein2 := colnames(mat)[col]]
  dt[, cosSim := mat[idx]]
  dt[, c("row", "col") := NULL ]
  return (dt[protein1 != protein2,])
}

cosSim.dt <- pbapply::pblapply(cosMats, cosineMat2Table, threshold = -1) %>% 
  rbindlist(idcol = "sample")
cosSim.dt[, gene1 := protein1]
cosSim.dt[, gene2 := protein2]

# get stringIDs where I have them...
cosSim.dt[huString.mapper, gene1 := stringID, on = c(protein1 = "protein")]
cosSim.dt[huString.mapper, gene2 := stringID, on = c(protein2 = "protein")]
cosSim.dt[anno.dt, alias1 := i.gene, on = c(protein1 = "protein")]
cosSim.dt[anno.dt, alias2 := i.gene, on = c(protein2 = "protein")]

cosSim.dt[, gs := "unknown"]
cosSim.dt[gene1 > gene2, gs := "reverse"]
cosSim.dt[gs.interactors ,gs := "interactor" , on = c("gene1", "gene2")]
cosSim.dt[gs.decoys, gs := "decoy" , on = c("gene1", "gene2")]
```
Distribution of the cosine similarity of the data; doesnt look good, not freat seperation of the two sets
Weird plot... why arewe biased towards 0 in the donor1 samples? For now just use the peak differential analysis
```{r, cosSim-histogram, fig.width=10, fig.height=7}
g <- ggplot(cosSim.dt[gs %in% c("decoy", "interactor")], aes (x = cosSim, fill=gs)) + 
  geom_histogram(boundary = 1.0, binwidth = 0.003) +
  scale_fill_manual(values=c('interactor'='gold', 'decoy'='grey70')) +
  facet_grid(gs~sample, scales = "free_y") +
  theme_bw()
g
BackupAsPDF(g, 'cosineSimilarity.decoyVsinteractor.histogram')
```

Save the cosine similarity results
```{r}
cosSim.dt <- cosSim.dt[cosSim > 0.8][gene1 < gene2]
fwrite(cosSim.dt, ScriptAndDatedFileName('donor1donor2.hiv.cosineSimilarity.highScores.csv.gz'))

cosSim.dt <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_donor1donor2.hiv.cosineSimilarity.highScores.csv.gz')
```

## Differential peak analysis

add A standard fractions to the allPeaksTable
```{r}
# all filtered peaks 
allPeaks <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_allPeaks.donor1donor2.monomerMassLabelled.csv.gz')
allPeaks[, unfilteredPeaks := goodPeak]
allPeaks[, goodPeak := filtteredGoodPeak]

# with standardised fractions
sec.long <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_hiv.donor1donor2.sec.ms.wStandardisedFractions.csv.gz') 
sec.long[, treatment := status] # I think we use status?
sec.long[, replicate := donor]
```

```{r}
# per protein, join peaks across runs into clusters
allPeaks[goodPeak == TRUE, proteinPeakCluster := clusterPeaks(cofmN.standardized, maxDistance = 2.5), by = protein]

# per peak cluster (per protein), summarize to a center (and height etc, but most important is center)
peakClusters <- allPeaks[!is.na(proteinPeakCluster), 
                         .(.N, meanPeakHeight  = mean(peakHeight), center = mean(cofmN.standardized)),
                         keyby = .(protein, proteinPeakCluster)]


# split data into single-protein chunks
# peaks in the protein
peakClusters.subTables <- split(peakClusters, by = "protein")

# sec data per protein (only those in above table)
allProteins <- names(peakClusters.subTables)
sec.subTables <- split(sec.long[protein %in% allProteins], by = "protein" )

names(allProteins) <- allProteins

anova.ls <- pbapply::pblapply(allProteins, function(proteinOI) anovaPeaksInOneProtein(sec.subTables[[proteinOI]], peakClusters.subTables[[proteinOI]], radius = 5))
# format output
anova.dt <- rbindlist(anova.ls, use.names = TRUE, fill = TRUE, idcol = "protein")

setnames(anova.dt,
         old = c( "rn",        "Sum.Sq",     "Mean.Sq",    "F.value", "Pr..F."), # anova col names to change
         new = c( "modelTerm", "SumSquares", "MeanSquare", "F",       "p.F"))

peakClusters[anova.dt[modelTerm == "poly(standardFraction, 4):treatment"], treatmentDiff.p := i.p.F]
peakClusters[anova.dt[modelTerm == "treatment"], treatmentIntensity.p := i.p.F]

# term we are interested in: what peak profiles change in response to treatment?
setorder(peakClusters, treatmentDiff.p, na.last = T)
#fwrite(peakClusters, ScriptAndDatedFileName('donor1donor2.de.anova.results.csv.gz'))

peakClusters <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_donor1donor2.de.anova.results.csv.gz')
```


Save the peaks and combine later...
Combine the peak and elution correlation scores (maybe for now just use the peak correlation score?)
```{r}
allGenes <- protein.info$stringID |> unique()
denomInteractor <- nrow(gs.interactors[gene1 %in% allGenes & gene2 %in% allGenes])
denomDecoy <- nrow(gs.decoys[gene1 %in% allGenes & gene2 %in% allGenes])

```


```{r}

scoreByGS(allPeakCorrs, denomDecoy, denomInteractor, column = "corScore", groupByVariable = c("sample", "start"))
```


**Not used **
apms data integration.. ignore for now
```{r}
#take our interactors passing mist and see if these overlap 
hiv.hu.mist <- fread('/Users/martingordon/Documents/projects/112024_RKakke_AF2_HIV_hitsInvestigation/201124_AFInteractiorsIntegration_data/2024_11_20_af_interactors_mistScores.csv')
# REV, VPR,  
# GAG-POL are there but seperate entities
# ENV is either GP120 or GP160

# just take those with score > 0.75 in either set
# add this to the gold standard set
highConf.interactors <- hiv.hu.mist[Bait %in% c('REV', 'VPR') & (HEK_MiST_score > 0.75 | Jurkat_MiST_score > 0.75), .(protein1=Bait, protein2=Protein_1, source='HIV-HU_MIST')]

# a datatable merge ; the on argument  in quotes is the col in i..
highConf.interactors[huString.mapper, stringID2 := i.stringID, on=c(protein2='protein')]
# now merge the viral prot IDs
highConf.interactors[hiv.nodes, stringID1 := i.stringID, on=c(protein1='gene')]
hu.hiv.gs <- highConf.interactors[, .(gene1=stringID1, gene2=stringID2, source='Jager et al')]
```


```{r}
#sec.dt[up2gene.dt, gene := i.geneLabel, on = c(protein = "uniprot")]
gs.interactors <- goldStandardPairs(unique(protein.info[!is.na(stringID)]$stringID), corum.path = NULL,
                  string.links.path = "/Users/ben/Downloads/208964.protein.physical.links.detailed.v12.0.txt.gz",
                  string.info.path =  NULL,
                  stringCombinedScoreThreshold = 250)

gs.decoys <- decoysFromString(unique(protein.info[!is.na(stringID)]$stringID),
                  links.path = "/Users/ben/Downloads/208964.protein.physical.links.detailed.v12.0.txt.gz",
                  info.path =  NULL)
```


