---
title: "050725_newplots"
author: "Martin Gordon"
date: "2025-05-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview 

Generate plots for HARC meeting tomorrow morningn

```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)
library(stringr)
library(viridis)
library(pracma) # needed for the peak finding algorithm
library(magick)
library(RColorBrewer)
#library(UniProt.ws) #cant load this packge; get naming conventions another way

# source Bens scripts so we can do a quick QC assessment of the input data before running
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/UniprotIDMapping.R") # map hu to mm 
source("~/Documents/utils/bp_utils/enrichmentTestFunctions.R")

# flag consec proteins
source("~/Documents/utils/bp_utils/SEC_MS_utils.localMods.R")
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")
```

color palette 
```{r, col}
donor.col <- brewer.pal(n=2,name='Pastel2')

status.col <-  c("#E41A1C","#377EB8","#4DAF4A")
names(status.col) <-  c('infected', 'uninfected', 'challenged')
```
functions

```{r}
complexIntensityHeatmaps <- function(intMats, complexDT, complexOI, dataType='relative', topOfColorRange = 0.5,...){
  
  
  if (dataType == 'relative'){
    denom = 50 / topOfColorRange
    colorFun <- circlize::colorRamp2(breaks = (0:50)/denom, colors = viridis(51,direction = 1))
    intensityName = "Relative Protein Intensity"
  } else if(dataType == 'absolute') {
    colorFun <- viridis(30, direction = 1)
    intensityName = "Log2 Intensity"
  }
  # get the individual sample lists
  samples <- names(intMats)
  
  sample <- samples[1]
  
  
  cols.oi <- c("protein","gene" ,"org", "complexID","subcomplexID")
  
  mapping.dt <- complexDT[complexID == complexOI, ..cols.oi] %>% 
    unique()
  mapping.dt[is.na(subcomplexID), subcomplexID:= complexID]
  print(mapping.dt)
  
  subMat <- intMats[[sample]][rownames(intMats[[sample]]) %in% mapping.dt$protein,]
  rowgrp <- mapping.dt[match(rownames(subMat), protein), subcomplexID]
  
  rowOrder <-  mapping.dt[, gene]
  
  # convert names to subset we want
  #print(rownames(subMat))
  rownames(subMat) <- mapping.dt[match(gsub('[;].+', '', rownames(subMat)), protein), gene]
  
  hml <- Heatmap (subMat,
                  name = intensityName,
                  cluster_rows = TRUE,
                  #row_order = rowOrder, # enforce roworder here, later heatmaps will follow this clustering
                  show_row_dend = F,
                  row_split = rowgrp, 
                  cluster_columns = FALSE,
                  col = colorFun ,
                  show_row_names = TRUE,
                  row_names_side = "left",
                  row_names_gp = gpar(fontsize = 10, col=ifelse(rownames(subMat) %in% viral.genes, 'red', 'black')),
                  row_title_rot = 0,
                  column_names_gp = gpar(fontsize = 5),
                 # column_labels = ifelse(as.integer(colnames(intMats[[sample]])) %% 5 == 0, colnames(intMats[[sample]]), ""),
                  column_title = sample,
                  column_title_gp = gpar(fontsize=12, fontface='bold'),
                  #first only:
                  show_heatmap_legend = TRUE, 
                  #row_title = sprintf ("%d Proteins", nrow(subMat)),
                  ...
  )
  
  # first one gets row title, also gets a legend
  
  if (length(intMats) > 1){
    for (sample in samples[2:length(samples)]){
      hml <- hml + Heatmap (intMats[[sample]][rownames(intMats[[sample]]) %in% mapping.dt$protein,],
                            name = intensityName,
                            cluster_rows = FALSE,
                            row_dend_reorder = FALSE,
                            cluster_columns = FALSE,
                            col = colorFun ,
                            show_row_names = FALSE, 
                            column_names_gp = gpar(fontsize = 5),
                            #column_labels = ifelse(as.integer(colnames(intMats[[sample]])) %% 5 == 0, colnames(intMats[[sample]]), ""),
                            column_title = sample,
                            column_title_gp = gpar(fontsize=12, fontface='bold'),
                            # 
                            show_heatmap_legend = FALSE,
                            ...)      
    }
  }
  return (hml)
}
```

data

Just read this in to get relative intensity
```{r}
sec.long <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/docs/2025_05_07_OutlierFractionsRemoved_interpNorm_sec.csv.gz')
relIntsMat <- scaledIntensityMatrices(secLong.dt = sec.long, scaleDenom = 'max')
```


```{r}
# using BPs normalized interpolated data
sec.long <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/docs/2025_05_07_OutlierFractionsRemoved_interpNorm_sec.csv.gz')
# log transform the intensity for visualization
sec.long[, intensity := log2(intensity+1)]

apms.int <-  fread('~/Documents/projects/040825_MMuralidharan_PCProphet_HIV_CD4Tcells/040825_preparePCProphetInput_data/2025_04_11_yager.ints.toCompare.csv.gz') # Yager paper data
yager.hits <- apms.int[bestMistScore > .75, .(protein1, protein2, gene1, gene2)]

viral.prots <- sec.long[org != 'Homo sapiens', unique(protein)]
viral.genes <- sec.long[org != 'Homo sapiens', unique(gene)]

robyn.complexes <- setDT(readxl::read_xlsx('./docs/Complexes_Database_RMK_20250506.xlsx', sheet = 1))
#filter to just described complexes 
robyn.complexes <- robyn.complexes[, (.SD), .SDcols = !patterns('[.]PG[.]Quantity')]
robyn.complexes <- robyn.complexes[, .(protein=PG.ProteinAccessions, gene=PG.Genes, org=PG.Organisms, 
                    complexID=`Complex ID...84`, subcomplexID=`SubComplex ID...85`)]
complexes.oi <- unique(robyn.complexes$complexID)
robyn.complexes <- robyn.complexes[complexID %in% complexes.oi & !is.na(complexID)]
robyn.complexes[is.na(subcomplexID), subcomplexID := complexID]

# differential peaks
peakClusters <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_07_donor1donor2.de.anova.results.csv.gz')
peakClusters[treatmentDiff.p < 0.05 & protein %in% viral.prots]
```

```{r}
hiv.set <- setDT(readxl::read_xlsx('./docs/Complexes_Database_RMK_20250506.xlsx', sheet = 2))
setnames(hiv.set, new=c('mw', 'protein', 'gene', 'org'))
hiv.set[, complexID := 'HIV related']

hiv.complexes.set <-  setDT(readxl::read_xlsx('./docs/Complexes_Database_RMK_20250506.xlsx', sheet = 3))
setnames(hiv.complexes.set, new=c('mw', 'protein', 'gene', 'org', 'descriptions', 'complexID', 'subcomplexID'))
```

Make a plotting data table that contains the  relevant col fields
```{r}
# first merge with the complexes
plot.dt <- merge(sec.long[, .(sample, protein, gene, org, mw, fraction, status, donor, normInts, intensity, intensity_totalScaled)], robyn.complexes, by=c('protein', 'gene', 'org'), all.x=T)
plot.dt[, donor := as.character(donor)]
plot.dt[, status := factor(status, levels=c('uninfected', 'challenged', 'infected'))]

plot.dt <- rbind(plot.dt, hiv.set, hiv.complexes.set, fill=T)
# dont think we need these merges now 
# then merge with the hiv associated
#plot.dt <- merge(plot.dt, hiv.set[,.(protein, gene, group)], by=c('protein', 'gene'), all.x=T)

# now merge with the HIV set
#plot.dt <- merge(plot.dt, hiv.complexes.set[, .(protein, gene, hiv_complexID, hiv_subcomplexID)], by=c('protein', 'gene'), all.x=T)
#plot.dt[, c('status', 'donor') := tstrsplit(sample, '_', keep=c(1,2))]

```

For now, zero out the uninfected intensities (scaled and normalized) as we think these are faulty IDs
```{r}
plot.dt[status == 'uninfected' & protein %in% viral.prots, intensity_totalScaled := NA]
plot.dt[status == 'uninfected' & protein %in% viral.prots, normInts := NA]
```
Heatmap of the scaled and unscaled (just log2 normalized) of the whole set
First plot all side-by-side, then donor-by-donor... not really sure how best to show these six samples
```{r}
intsMat <- scaledIntensityMatrices(secLong.dt = sec.long, scaleDenom = 'none')
```

Seems to work OK for now.... takes the input vector as a matrix
```{r, hm-hiv-related, fig.height=7, fig.width=10}
hm <- complexIntensityHeatmaps(intsMat, 
                         complexDT = plot.dt, 
                         dataType = 'absolute',
                         complexOI='HIV related')

BackupAsPDF(hm, 'hiv.associated.absInts.heatmap')


hm <- complexIntensityHeatmaps(relIntsMat, 
                         complexDT = plot.dt, 
                         dataType = 'relative',
                         complexOI='HIV related')
hm
BackupAsPDF(hm, 'hiv.associated.relativeInts.heatmap')
```
Run the same analysis for the second set
```{r, hm-hiv-related, fig.height=7, fig.width=12}
hiv.complexes.set[complexID == "ER membrane protein complex (EMC8 variant)", complexID := "ER membrane protein complex"]
hiv.complexes.set[complexID == "ER membrane protein complex (EMC9 variant)", complexID := "ER membrane protein complex"]

complexes.oi <- hiv.complexes.set$complexID %>% 
  unique()
# cant plot a heatmap of one gene

lapply(complexes.oi, function(comp){
  
  fname <- gsub('[/]| ', '', comp)
  
  hm <- complexIntensityHeatmaps(relIntsMat, 
                         complexDT = plot.dt, 
                         dataType = 'relative',
                         complexOI=comp)
 BackupAsPDF(hm, paste0(fname, '.hiv.associated.relativeInts.heatmap'))
})
```
plot the same complexes absolute intensities

```{r, hiv-abs, fig.height=7, fig.width=12}
lapply(complexes.oi, function(comp){
  
  fname <- gsub('[/]| ', '', comp)
  
  hm <- complexIntensityHeatmaps(intsMat, 
                         complexDT = plot.dt, 
                         dataType = 'absolute',
                         complexOI=comp)
 BackupAsPDF(hm, paste0(fname, '.hiv.associated.absInts.heatmap'))
})
```
plot the absolute and relative intensity heatmaps for everything

```{r, hm-hiv-related, fig.height=7, fig.width=12}
lapply(complexes.oi, function(comp){
  
  fname <- gsub('[/]| ', '', comp)
  
  hm.rel <- complexIntensityHeatmaps(relIntsMat, 
                         complexDT = plot.dt, 
                         dataType = 'relative',
                         complexOI=comp)
  
  BackupAsPDF(hm.rel, paste0(fname, '.hiv.associated.relativeInts.heatmap'))
  
  hm.abs <- complexIntensityHeatmaps(intsMat, 
                         complexDT = plot.dt, 
                         dataType = 'absolute',
                         complexOI=comp)
 BackupAsPDF(hm.abs, paste0(fname, '.hiv.associated.absInts.heatmap'))
})
```



 Another plot; collapsed linecharts; take the average of the normalized of protein intensities of each oft he groups; loop over the plot and highlight the different complexes
 
```{r}
plot.dt
summary.dt <- plot.dt[!is.na(complexID), 
                     .(mean_totalInts=mean(intensity_totalScaled, na.rm=T),mean_normInts=mean(intensity, na.rm=T), sd_totalInts=sd(intensity_totalScaled, na.rm=T),sd_normInts=sd(intensity, na.rm=T),
                       median_totalInts=median(intensity_totalScaled, na.rm=T),median_normInts=median(intensity, na.rm=T)),
                      by=.(complexID, status, donor, fraction)]

# global summary per donor
global.summary.dt <- plot.dt[!is.na(complexID), 
                     .(mean_totalInts=mean(intensity_totalScaled, na.rm=T),mean_normInts=mean(intensity, na.rm=T), sd_totalInts=sd(intensity_totalScaled, na.rm=T),sd_normInts=sd(intensity, na.rm=T),
                      median_totalInts=median(intensity_totalScaled, na.rm=T),median_normInts=median(intensity, na.rm=T)),
                      by=.(complexID, donor, fraction)]


# practise plot 
g <- ggplot(global.summary.dt, aes(x=fraction, y=mean_normInts, group=interaction(donor,complexID))) +
  geom_line(alpha=0.2) +
  geom_line(data=summary.dt[complexID == 'RNAPII',], aes(x=fraction, y=mean_normInts, group=interaction(status, donor,complexID), linetype=donor, color=status), linewidth=1) +
  scale_x_continuous(breaks =seq(1,72, 10),  labels = seq(1,72, 10)) +
  scale_color_manual(values=status.col) +
  labs(title='complexID', subtitle='',x='fraction', y='group intensiy (scaled to total)') +
  theme_bw() +
  guides(linetype=guide_legend('donor'))
g
```
For each complex, take the average and plot he 

```{r, fig.width=8,fig.height=5}
complexes.oi <- plot.dt$complexID %>%
  unique()
complexes.oi <- complexes.oi[-1]

lapply(complexes.oi, function(comp){
  
  fname <- gsub('[/]| ', '', comp)
  
  g <- ggplot(global.summary.dt[!is.na(donor) & fraction %in% seq(10, 62,1)], aes(x=fraction, y=median_totalInts, group=interaction(donor,complexID))) +
  geom_line(alpha=0.1) +
  geom_line(data=summary.dt[complexID == comp & !is.na(donor) & fraction %in% seq(10, 62,1),], aes(x=fraction, y=median_totalInts, group=interaction(status, donor,complexID), linetype=donor, color=status), linewidth=0.8) +
  scale_x_continuous(breaks =seq(1,72, 10),  labels = seq(1,72, 10)) +
  scale_color_manual(values=status.col) +
  labs(title=comp,x='fraction', y='protein group median intensity (scaled to total)') +
  theme_bw() +
  theme(strip.text.y=element_text(size=10),
          strip.text=element_text(color='white'),
          strip.background = element_rect(colour = "white", fill = "grey80")) + 
  guides(linetype=guide_legend('donor'))
  
  g
  BackupAsPDF(g, paste0(fname,'.complexes.medianRelInts.linechart'))
})
```

Plot the absolute intensity of donor1 and donor 2 for the ridgelineplots
```{r}
plot.dt[, totalInts := sum(intensity, na.rm=T), by=.(protein, sample)]
plot.dt[is.na(subcomplexID), subcomplexID := complexID]

# test... looks goos
lapply('BLOC1', function(comp){
  
  fname <- gsub(' |[/]', '_', comp)
  # plot the donors on top of each other... color by group maybe? use lines to distinguish groups
  p <- ggplot(plot.dt[complexID == comp & donor == '1',], aes(x = fraction, y=gene, height = intensity_totalScaled, fill=totalInts)) +
    ggridges::geom_ridgeline(scale = 8, stat = "identity") +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=2, aes(color='maroon'), alpha=0.4) +
    scale_fill_viridis_c(name = "log2 total normalized intensity", option = "C") +
    labs(title=paste0(comp)) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) + # to reorder colnames
    facet_grid(subcomplexID~status, scales='free_y') + 
    theme_bw() +
    theme(strip.text=element_text(color='white', size=9),
          strip.background = element_rect(colour = "white", fill = "grey80"))
 
  print(p)
  #print(paste0(fname, '.complexSubunits.ridgeline'))
  #BackupAsPDF(p, paste0(fname, '.complexSubunits.ridgeline.totalInts'))
})
```
```{r, ridgeline-ints, fig.width=11,fig.height=6}
# sum normalized intensitt and then taker the log2 val
plot.dt[, totalInts := log2(sum(normInts, na.rm=T)+1), by=.(protein, sample)]

lapply(complexes.oi[-1], function(comp){
  
  fname <- gsub(' |[/]', '_', comp)
  print(fname)
  # plot the donors on top of each other... color by group maybe? use lines to distinguish groups
  p <- ggplot(plot.dt[complexID == comp & donor == '1',], aes(x = fraction, y=gene, height = intensity_totalScaled, fill=totalInts)) +
    ggridges::geom_ridgeline(scale = 8, stat = "identity") +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=2, aes(color='maroon'), alpha=0.4) +
    scale_fill_viridis_c(name = "log2 total normalized intensity", option = "C") +
    labs(title=paste0(comp)) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) + # to reorder colnames
    facet_grid(subcomplexID~status, scales='free_y') + 
    theme_bw() +
    theme(strip.text=element_text(color='white', size=9),
          strip.background = element_rect(colour = "white", fill = "grey80"))
 
  BackupAsPDF(p, paste0(fname, '.complexSubunits.ridgeline.totalInts'))
})
```


## 05-08-25
Plot ridge line plots of the two listed complexes
```{r, fig.width=10,fig.height=7}

paf <- fread('./docs/PAF1C_Associated.txt')$PG.ProteinAccessions
hiv <- fread('./docs/HIV_Set_Short.txt')$PG.ProteinAccessions
crl.apobec <- fread('./docs/CRL_APOBEC.txt')$PG.ProteinAccessions

plot.dt

plot.dt[protein %in% viral.prots & status == 'uninfected', intensity_totalScaled := NA]


p <- ggplot(plot.dt[protein %in% crl.apobec & !is.na(status),], aes(x = fraction, y=gene, height = intensity_totalScaled, linetype=donor, fill=totalInts)) +
    ggridges::geom_ridgeline(scale = 8, stat = "identity") +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=2, aes(color='maroon'), alpha=0.4) +
    scale_fill_viridis_c(name = "log2 total normalized intensity", option = "C") +
    labs(title=paste0('CRL-APOBEC complex')) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) + # to reorder colnames
    facet_grid(.~status, scales='free_y') + 
    theme_bw() +
    theme(strip.text=element_text(color='white', size=9),
          strip.background = element_rect(colour = "white", fill = "grey80"))

p 
BackupAsPDF(p, 'crl.apobec.associated.bothdonors.ridgeline')



p <- ggplot(plot.dt[protein %in% crl.apobec & donor== '1' & !is.na(status),], aes(x = fraction, y=gene, height = intensity_totalScaled,fill=totalInts)) +
    ggridges::geom_ridgeline(scale = 8, stat = "identity") +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=2, aes(color='maroon'), alpha=0.4) +
    scale_fill_viridis_c(name = "log2 total normalized intensity", option = "C") +
    labs(title=paste0('CRL-APOBEC complex')) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) + # to reorder colnames
    facet_grid(.~status, scales='free_y') + 
    theme_bw() +
    theme(strip.text=element_text(color='white', size=9),
          strip.background = element_rect(colour = "white", fill = "grey80"))

p 
BackupAsPDF(p, 'crl.apobec.associated.donor1.ridgeline')


p <- ggplot(plot.dt[protein %in% crl.apobec & donor== '2' & !is.na(status),], aes(x = fraction, y=gene, height = intensity_totalScaled,fill=totalInts)) +
    ggridges::geom_ridgeline(scale = 8, stat = "identity") +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=2, aes(color='maroon'), alpha=0.4) +
    scale_fill_viridis_c(name = "log2 total normalized intensity", option = "C") +
    labs(title=paste0('CRL-APOBEC complex')) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) + # to reorder colnames
    facet_grid(.~status, scales='free_y') + 
    theme_bw() +
    theme(strip.text=element_text(color='white', size=9),
          strip.background = element_rect(colour = "white", fill = "grey80"))

p 
BackupAsPDF(p, 'crl.apobec.associated.donor2.ridgeline')


# plot the donors on top of each other... color by group maybe? use lines to distinguish groups
p <- ggplot(plot.dt[protein %in% crl.apobec & !is.na(status)], aes(x = fraction, y=gene, fill=status)) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled), alpha=0.4,  scale = 8) +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=3, aes(color='maroon')) +
    #scale_fill(low = "navy", mid = "gray98", high = "firebrick", midpoint = log2(1), name = "log2 mass ratio") +
    scale_fill_manual(values=status.col) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) +
    facet_grid(.~donor, scales='free_y') + 
    labs(title='CRL-APOBEC complex') +
    theme_bw() +
    theme(strip.text.y=element_text(size=10),
          strip.text=element_text(color='white'),
          strip.background = element_rect(colour = "white", fill = "grey80")) +
  guides(linetype=guide_legend('donor'))
p
BackupAsPDF(p, 'crl.apobec.statuscol.ridgeplot')
```

```{r, fig.width=10,fig.height=7}

# plot the donors on top of each other... color by group maybe? use lines to distinguish groups
p <- ggplot(plot.dt[protein %in% paf], aes(x = fraction, y=gene, fill=status)) +
    ggridges::geom_ridgeline(aes(height = intensity_totalScaled), alpha=0.4,  scale = 8) +
    geom_vline(xintercept=c(10,20,30,40,50,60), linetype=3, aes(color='maroon')) +
    #scale_fill(low = "navy", mid = "gray98", high = "firebrick", midpoint = log2(1), name = "log2 mass ratio") +
    scale_fill_manual(values=status.col) +
    #scale_y_discrete(name = "protein", limits=rev(robyn.complexes[complexID == 'CCC/WASH', gene])) +
    facet_grid(.~donor, scales='free_y') + 
    labs(title='PAF-associated complex') +
    theme_bw() +
    theme(strip.text.y=element_text(size=10),
          strip.text=element_text(color='white'),
          strip.background = element_rect(colour = "white", fill = "grey80")) +
  guides(linetype=guide_legend('donor'))
p
BackupAsPDF(p, 'paf.statuscol.ridgeplot')
```




 
 Generate linecharts of the different complexes (linetype can be donor, color can treatment; the bg lines can be average over all other complexes)




