---
title: "redo_pwcomparions_wMatchProts"
author: "Martin Gordon"
date: "2024-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Redo of the pairwise comparisons with the matching proteins included

Todo 
--
Find out why the unique merging has N number of different rows

```{r}
library(data.table)
library (ComplexHeatmap)
library (circlize)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(magrittr)
library(MSstats)
library(ggrepel)
library(circlize)
library(patchwork)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```


```{r}
spec.proc <- prepareDataForMSStats('/Users/martingordon/Documents/projects/020124_R_MCCall_Cox_APEX/data/RitaMC_MQOut/combined/txt/evidence.txt',
                                   './docs/keys.txt')

spec.rmc.proc <- prepareDataForMSStats('~/Library/CloudStorage/Box-Box/APEX\ Data\ to\ Share/code/evidenceApex12.txt',
                                       '~/Library/CloudStorage/Box-Box/APEX\ Data\ to\ Share/keys.tidy.txt')
```

lets process this data ourselves..
```{r}
spec <- fread('/Users/martingordon/Documents/projects/020124_R_MCCall_Cox_APEX/data/RitaMC_MQOut/combined/txt/evidence.txt')
keys <- fread('./docs/keys.txt')

spec <- merge(x=spec, y=keys, by.x='Raw file', by.y='RawFile')

#remove contaminants
spec <- spec[!grepl("^$|CON__|REV__", Proteins),]

# just map to sequences for the matches and append modifications... better approach wou
#spec[Type ==  "TIMS-MULTI-MATCH", `Modified sequence` := paste0(Modifications,Sequence)]
#spec[Type ==  "TIMS-MULTI-MATCH", `Modified sequence` := Sequence]

# for this assuming the mod id is a combo of the modifications and sequence cols only...##

mod.seqs <- spec[Type !=  "TIMS-MULTI-MATCH", .(`Modified sequence`), by=.(Sequence,Modifications)] %>% unique()

# merge this info to the match prots
nrow(spec[Type == "TIMS-MULTI-MATCH",])
nrow(spec[Type != "TIMS-MULTI-MATCH",])
spec %>% nrow()

nrow(spec)

# wht are we gaining nearly 1k extra rows?....
spec.match <- merge(x=spec[Type == "TIMS-MULTI-MATCH", -c("Modified sequence")], y=mod.seqs, all.x=TRUE, all.y=FALSE, by=c('Sequence', 'Modifications'))
spec <- rbind(x=spec.match, y=spec[Type !=  "TIMS-MULTI-MATCH",])


# sum the intensity values for multiple peptides in each run
spec.noDups <- spec[, .(Intensity = sum(Intensity), count.features = .N),
                                by = .(Proteins, 
                                       `Modified sequence`,
                                        Charge,
                                        IsotopeLabelType,
                                        Condition,
                                        BioReplicate,
                                        Run)]

spec.mod.proc <- spec.noDups[,.(ProteinName = Proteins,
                                               PeptideSequence = `Modified sequence`,
                                                PrecursorCharge = Charge,
                                                FragmentIon = NA,
                                                ProductCharge  = NA,
                                                IsotopeLabelType,
                                                Condition,
                                                BioReplicate,
                                                Run,
                                                Intensity)]

nrow(spec.mod.proc[!is.na(Intensity)])
nrow(spec.proc[!is.na(Intensity)])
```


```{r}
# no multiple feature peptide ions detected
spec.proc[!is.na(Intensity),.N, by=.(PeptideSequence,PrecursorCharge,Run)][N >1]
spec.mod.proc[!is.na(Intensity),.N, by=.(PeptideSequence,PrecursorCharge,Run)][N >1]
```

```{r}
col.pal <-  randomcoloR::distinctColorPalette(length(unique(spec.proc$Condition)))
col.pal

unique(spec.mod.proc$Condition)
spec.list <-  list('ori'=spec.proc,
                   'cox'=spec.rmc.proc,
                   'ori-w-matches' = spec.mod.proc)


# change labelling to match others
spec.list[[2]][,Condition := ifelse(Condition == 'EccC', 'EccCmut_inf_BR1', 'WT_inf_BR1')]
# assign biorep 1-3 based on Run ID
spec.list[[2]][, BioReplicate := ifelse(Run == 1 | Run == 4, 1,
                                        ifelse(Run == 2 | Run == 5, 2, 3)) ] 

lapply(seq_along(spec.list), function(x, n, i){

g <- ggplot(x[[i]][!is.na(Intensity),.N, by=.(Condition,BioReplicate,Run)], aes(x=reorder(interaction(Condition,BioReplicate,Run)), y=N, fill=Condition)) +
  geom_bar(stat='Identity') +
  ggtitle('N features per sample') +
  ggtitle(n[[i]]) +
  scale_fill_manual(values= col.pal) +
  theme(axis.text.x = element_text(angle=90))
g

 BackupAsPDF(g, paste0(n[[i]],'.rawFeatureCounts.barplot'))

}, x=spec.list, n=names(spec.list))
```

How do the intensities look?

```{r}

lapply(seq_along(spec.list), function(x, n, i){

g <- ggplot(x[[i]], aes(x=reorder(interaction(Condition,BioReplicate,Run)), y=log2(Intensity), fill=Condition)) +
  geom_boxplot() +
  ggtitle('N features per sample') +
  ggtitle(n[[i]]) +
  scale_fill_manual(values= col.pal) +
  theme(axis.text.x = element_text(angle=90))
g

 BackupAsPDF(g, paste0(n[[i]],'.rawIntensities.boxplots'))

}, x=spec.list, n=names(spec.list))
```
PCA of features
No real changes 

```{r}
                                 
lapply(seq_along(spec.list), function(x,n,i){

  featureMat <- dcast(x[[i]],
                     paste0(PeptideSequence, PrecursorCharge)~interaction(Condition,BioReplicate), value.var = "Intensity") |> as.matrix(rownames = "PeptideSequence")

  featureMat <- featureMat[complete.cases(featureMat), ]

  pcaOut <- prcomp(t(log2(featureMat)))

  colInfo <- data.table(colname = colnames(featureMat))
  colInfo[,c("status","rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]
  colInfo[, biorep := str_extract(status,"[123]$") ]

  #PCA  
  pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
  pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
  pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
  #plot first two components
  p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = status, shape = rep)) + 
    geom_point(alpha=1.0, size=4) + 
    ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
    theme_bw() + 
    ggtitle(n[[i]]) +
    xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
    ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
    ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_shape_manual(values = 21:25) +
    #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
    guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

  BackupAsPDF(p, paste0(n[[i]],".PCA_Complete_Features"))
  
}, x=spec.list, n=names(spec.list))

```
Heatmaps of the features in each; can we see v obvious intensity inflation for the summed 

```{r}

lapply(seq_along(spec.list), function(x,n,i){

  featureMat <- dcast(x[[i]],
                     paste0(PeptideSequence, PrecursorCharge)~interaction(Condition,BioReplicate), value.var = "Intensity") |> as.matrix(rownames = "PeptideSequence")
  
  #  log transform
  featureMat <-  log2(featureMat)
  
  # create annotation
  ann <- data.table(replicate=str_extract(colnames(featureMat),'[1-3]$'),
                  biogroup=str_extract(colnames(featureMat),'BR[1-3]'),
                  status=ifelse(grepl("EccC", colnames(featureMat)), 'EccCmut', 'WT'))


  colours <- list('biogroup' = c('BR1' = '#4477AA', 'BR2'="#228833", 'BR3'="#CCBB44"),
                'replicate' = c('1' = '#4477AA', '2'='#66CCEE', '3'="#228833"),
                'status' =c('WT'="#CCBB44", 'EccCmut'='#EE6677')
                )

  colAnn <- HeatmapAnnotation(df = ann, col = colours)

  # subset the data
  submat <- featureMat[sample(rownames(featureMat), 2000), ]
  
  print(dim(featureMat))
  
  hm <- Heatmap(submat, 
                cluster_rows=clusterWNA(submat),
                show_row_names = F, 
                cluster_columns = T, 
                top_annotation = colAnn, 
                row_title = paste0(nrow(submat), ' N features'), 
                na_col = 'darkgrey', 
                name='LogIntensities', 
                column_names_gp = gpar(fontsize=6))
  hm
  BackupAsPDF(hm, paste0(n[[i]],'.features.clustered.heatmap'))


  submat <-  featureMat[complete.cases(featureMat),]

  hm <- Heatmap(submat, cluster_rows=clusterWNA(submat), show_row_names = F, row_title = paste0(nrow(submat), ' N features'), cluster_columns = T, top_annotation = colAnn, na_col = 'darkgrey', name='LogIntensities', column_names_gp = gpar(fontsize=6))
  hm
  BackupAsPDF(hm, paste0(n[[i]],'.features.completeCases.clustered.heatmap'))


  submat <- sweep(submat,1, STATS=apply(submat, 1, median, na.rm=T))
  hm <- Heatmap(submat, cluster_rows=clusterWNA(submat), show_row_names = F, cluster_columns = T, top_annotation = colAnn, row_title = paste0(nrow(submat), ' N features'),  na_col = 'darkgrey',name='LogIntensities/Median',  column_names_gp = gpar(fontsize=6))
  BackupAsPDF(hm, paste0(n[[i]],'.features.medianScaled.clustered.heatmap'))

}, x=spec.list, n=names(spec.list))



lapply(seq_along(spec.list), function(x,n,i){
  
  featureMat <- dcast (x[[i]],
                     paste0(PeptideSequence, PrecursorCharge)~interaction(Condition,BioReplicate), value.var = "Intensity") |> as.matrix(rownames = "PeptideSequence")

#  log transform
featureMat <-  log2(featureMat)

dim(featureMat)
dim(featureMat[complete.cases(featureMat),]) #only 234 cases...
colnames(featureMat)

# create annotation
ann <- data.table(replicate=str_extract(colnames(featureMat),'[1-3]$'),
                  biogroup=str_extract(colnames(featureMat),'BR[1-3]'),
                  status=ifelse(grepl("EccCmut", colnames(featureMat)), 'EccCmut', 'WT'))


colours <- list('biogroup' = c('BR1' = '#4477AA', 'BR2'="#228833", 'BR3'="#CCBB44"),
                'replicate' = c('1' = '#4477AA', '2'='#66CCEE', '3'="#228833"),
                'status' =c('WT'="#CCBB44", 'EccCmut'='#EE6677')
                )

colAnn <- HeatmapAnnotation(df = ann, col = colours)



# subset the data
submat <- featureMat[sample(rownames(featureMat), 2000), ]
  

hm <- Heatmap(submat, cluster_rows=clusterWNA(submat), show_row_names = F, cluster_columns = T, top_annotation = colAnn, name='LogIntensities', column_names_gp = gpar(fontsize=6))
hm
BackupAsPDF(hm, 'features.clustered.heatmap')


submat <-  featureMat[complete.cases(featureMat),]

hm <- Heatmap(submat, cluster_rows=clusterWNA(submat), show_row_names = F, cluster_columns = T, top_annotation = colAnn, name='LogIntensities', column_names_gp = gpar(fontsize=6))
hm
BackupAsPDF(hm, 'features.completeCases.clustered.heatmap')


submat <- sweep(submat,1, STATS=apply(submat, 1, median, na.rm=T))
hm <- Heatmap(submat, cluster_rows=clusterWNA(submat), show_row_names = F, cluster_columns = T, top_annotation = colAnn, name='LogIntensities/Median',  column_names_gp = gpar(fontsize=6))
hm

}, x=spec.list, n=names(spec.list))

```
Matching between Runs?
I dont think we will gain much from these comparisons

prepare the spec files.. not working for the matchbetweenRuns set

```{r}
lapply(seq_along(spec.list), function(x,n,i){
  
  x[[i]][, oldName := paste0(Condition, '.',BioReplicate)]
  x[[i]][, TechReplicate := BioReplicate]
  x[[i]][, BioReplicate := str_extract(Condition, '[123]$')]
  x[[i]][, Condition := str_extract(Condition, "WT|EccCmut")]

},x=spec.list, n=names(spec.list))

# now process the data 
dp.list <- lapply(seq_along(spec.list), function(x,n,i){
  
  dp.out <- MSstats::dataProcess(x[[i]], 
                              MBimpute =  FALSE, 
                              featureSubset = "highQuality", 
                              remove_uninformative_feature_outlier = TRUE)
  
  return(dp.out)

},x=spec.list, n=names(spec.list))

names(dp.list) <-  names(spec.list)
```
save mss processed output
```{r}
lapply(seq_along(dp.list), function(x,n,i){
  
  saveRDS(x[[i]], paste0("./output/", n[[i]], ".150224_dp.out.rds"))
  fwrite (x[[i]]$ProteinLevelData, paste0("./output/", n[[i]],".AB.ProteinLevelData.csv"))
  fwrite (x[[i]]$FeatureLevelData,  paste0("./output/", n[[i]],".AB.FeatureLevelData.csv.gz"))

}, x=dp.list, n=names(dp.list))
```
Read in output of the dataProcess function 

```{r}
dp.list <- list("ori" = readRDS("./output/ori.150224_dp.out.rds"),
                "cox" = readRDS("./output/cox.150224_dp.out.rds"),
                "ori-w-matches" = readRDS("./output/ori-w-matches.150224_dp.out.rds")
                )
```


Read in the three p.quants and do boxplots and counts of proteins in each of the samples

```{r}
p.quant.list <- list("ori" = fread('./output/ori.AB.ProteinLevelData.csv'),
                     "cox" = fread('./output/cox.AB.ProteinLevelData.csv'),
                     "ori-w-matches" =  fread('./output/ori-w-matches.AB.ProteinLevelData.csv')
                     )

names(p.quant.list)
```

boxplots of normalised intensities

```{r}
lapply(seq_along(p.quant.list), function(x,n,i){
  
  g <- ggplot(x[[i]], aes (x=interaction(SUBJECT, GROUP,RUN), y = LogIntensities, fill = GROUP)) +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values=col.pal) +
  theme(axis.text.x = element_text(angle=90))
g
BackupAsPDF(g, paste0(n[[i]], 'proteinIntensities.boxplots'))
  
}, x=p.quant.list, n=names(p.quant.list))

```

Protein Counts per sample

```{r}

lapply(seq_along(p.quant.list), function(x,n,i){
  
  g <- ggplot(x[[i]][,.N, by=.(GROUP,SUBJECT,RUN)], aes (x=interaction(SUBJECT, GROUP,RUN), y = N, fill = GROUP)) +
  geom_bar(stat='Identity') +
  theme_classic() +
  scale_fill_manual(values=col.pal) +
  theme(axis.text.x = element_text(angle=90))
g
BackupAsPDF(g, paste0(n[[i]], 'nProteins.barplot'))
  
}, x=p.quant.list, n=names(p.quant.list))
```
Lets just quickly check the PCA with all three samples, then proceed with ours with matched runs enabled.

```{r}
p.quant.list
lapply(seq_along(p.quant.list), function(x,n,i){

  featureMat <- dcast(x[[i]],
                     Protein~interaction(GROUP,SUBJECT,originalRUN), value.var = "LogIntensities") |> as.matrix(rownames = "Protein")

  featureMat <- featureMat[complete.cases(featureMat), ]

  pcaOut <- prcomp(t(log2(featureMat)))

  colInfo <- data.table(colname = colnames(featureMat))
  colInfo[,c("status","rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]
  colInfo[, biorep := str_extract(status,"[123]$") ]

  #PCA  
  pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
  pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
  pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
  #plot first two components
  p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = status, shape = rep)) + 
    geom_point(alpha=1.0, size=4) + 
    ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
    theme_bw() + 
    ggtitle(n[[i]]) +
    xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
    ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
    ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_shape_manual(values = 21:25) +
    #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
    guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

  BackupAsPDF(p, paste0(n[[i]],".PCA_Complete_Proteins"))
  
}, x=p.quant.list, n=names(p.quant.list))
```
I dont know if we can proceed with analysing the BR1 group for the mutant.. may need to proceed with just one per group
None of these proteins recovered....


```{r}
biotin.carboxylases.up <- c("O00763","P05165","P11498","Q13085","Q96RQ3")

lapply(p.quant.list, function(x){
  
  x[,ProteinName := tidyUniProtNames(Protein)]
  
})

p <- ggplot(p.quant.list[[1]][ProteinName %in% biotin.carboxylases.up], 
            aes (x = interaction ( SUBJECT, GROUP,RUN ), y = LogIntensities, color = Protein)) + 
  geom_line(aes(group = Protein)) + 
  geom_point(aes(shape = as.factor(SUBJECT))) +
  theme(axis.text.x = element_text(angle=90))
p

```

Looking at the PCA, doesnt seem clear and obvious 'pairing' between the samples..
Regardless, handle this pairing design by adding batch term to model to control for random effect

```{r}
dp.list <- lapply(seq_along(dp.list), function(x,n,i){
  
  p.q <- setDT(x[[i]]$ProteinLevelData)
  f.q <- setDT(x[[i]]$FeatureLevelData)
  
  # subject is 
  p.q[, SUBJECT := paste0('batch.', SUBJECT)]
  f.q[, SUBJECT := paste0('batch.', SUBJECT)]
  
  x[[i]]$ProteinLevelData <- p.q
  x[[i]]$FeatureLevelData <- f.q
  
  return(x[[i]])
  
},x=dp.list, n=names(dp.list))
```
Run MSstats and compare number of hits
====
If not many, proceed as is 

```{r}
contrasts.mat <- matrix(c(-1,1), nrow=1, ncol=2)
colnames(contrasts.mat) <-  c('WT', 'EccCmut')
rownames(contrasts.mat) <- 'EccCmut-WT'

mss.list <- lapply(seq_along(dp.list), function(x,n,i){
  
  mss <- groupComparison(contrast.matrix=contrasts.mat, data=x[[i]])
  mss.dt <- setDT(mss$ComparisonResult)
  return(mss.dt)
  
},x=dp.list, n=names(dp.list))

names(dp.list) <- names(p.quant.list)
names(mss.list) <- names(p.quant.list)
```

tidy the protein names, get gene names and extract the sig hits

```{r}
# function to extract UniProtID substr from longer string and collapse
tidyUniProtNames <-  function(Protein, sep=';'){
 return(sapply(strsplit(Protein, split = ';', fixed=T), function(x){ 
   # assuming the UNIPROT are surrounded by '|'
  str_extract(x,  "(?<=[|])[A-Z0-9]+(?=[|])") %>% 
    paste(., collapse=sep)
  }))
}

mss.comb <- rbindlist(mss.list, use.names = T, idcol=T)
mss.comb[, ProteinName := tidyUniProtNames(as.character(Protein))]

mss.comb[, gene := multiUniprots2multiGenes(ProteinName)]
mss.comb[, sig := 'not']

#lets perfrom the multiple testing ourselves after removing things absent in both conditions
mss.comb[abs(log2FC) > 0.58 & adj.pvalue < 0.05, sig := ifelse(log2FC > 0, 'up' , 'down')]
```

N sig hits per comparison:
Low number of hits in our comparisons a result of the multiple testing correction..

```{r}
mss.comb[sig != 'not',.N, by=.(.id,sig)]

g <-  ggplot(mss.comb[sig != 'not',.N, by=.(.id,sig)], aes(x=.id, y=N, fill=sig)) +
  geom_bar(stat='Identity') +
  ggtitle('N sig hits') +
  scale_fill_manual(values=c('up'='red', 'down'='blue'))
g
BackupAsPDF(g, 'N.sighits.byMatches')

mss.comb[, sig := 'not']
#lets perfrom the multiple testing ourselves after removing things absent in both conditins
mss.comb[abs(log2FC) > 0.58 & adj.pvalue < 0.05, sig := ifelse(log2FC > 0, 'up' , 'down')]

g <-  ggplot(mss.comb[sig != 'not' & issue != 'oneConditionMissing', .N, by=.(.id,sig)], aes(x=.id, y=N, fill=sig)) +
  geom_bar(stat='Identity') +
  ggtitle('N sig hits (raw pval < 0.05') +
  scale_fill_manual(values=c('up'='red', 'down'='blue'))
g
BackupAsPDF(g, 'N.sighits.rawPval.byMatches')
```

How compariable are thelog2FC  estimates?
Look at scatterplots of the values

```{r}
mss.comb[,.id := gsub('-','_', .id)]

mss.comb.wide <-dcast(mss.comb, Protein+gene~.id, value.var='log2FC')

g <-  ggplot(mss.comb.wide, aes(x=ori, y=ori_w_matches)) +
  geom_point() +
  geom_abline(intercept = 0, slope=1, col='red') +
  geom_vline(xintercept=0, linetype=2, alpha=0.6) +
  geom_hline(yintercept = 0, linetype=2, alpha=0.6)

g


gg <-  ggplot(mss.comb.wide, aes(x=cox, y=ori_w_matches)) +
  geom_point() +
  geom_abline(intercept = 0, slope=1, col='red') +
  geom_vline(xintercept=0, linetype=2, alpha=0.6) +
  geom_hline(yintercept = 0, linetype=2, alpha=0.6)

gg

ggg <-  ggplot(mss.comb.wide, aes(x=ori, y=cox)) +
  geom_point() +
  geom_abline(intercept = 0, slope=1, col='red') +
  geom_vline(xintercept=0, linetype=2, alpha=0.6) +
  geom_hline(yintercept = 0, linetype=2, alpha=0.6)

ggg
#  ggtitle('N sig hits (raw pval < 0.05') +
 # scale_fill_manual(values=c('up'='red', 'down'='blue'))

BackupAsPDF(g + gg + ggg, 'log2FC.acatterplots', dimensions = c(14,8))

mss.comb[, unique(.id), ]


mss.comb[.id == 'ori_w_matches' & issue == 'oneConditionMissing', ]
```

Look at the heatmap of sig results for each 

```{r}
sigProts <- mss.comb[sig != 'not', unique(ProteinName), by=.id]

sigProts[,.N, by=.id]

p.quant.comb <- rbindlist(p.quant.list, idcol = T)
p.quant.comb[,ProteinName := tidyUniProtNames(Protein)]

ori.mat <- dcast(p.quant.comb[.id == 'ori',], ProteinName~interaction(GROUP,SUBJECT,originalRUN), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='ProteinName')

ori.wm.mat <- dcast(p.quant.comb[.id == 'ori-w-matches',], ProteinName~interaction(GROUP,SUBJECT,originalRUN), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='ProteinName')

cox.mat <- dcast(p.quant.comb[.id == 'cox',], ProteinName~interaction(GROUP,SUBJECT,originalRUN), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='ProteinName')
```

plot the significant hits in each of these comparisons

```{r}
mat <-  cox.mat[rownames(cox.mat) %in% sigProts[.id == 'cox',V1],]

Heatmap(cox.mat, cluster_rows = clusterWNA(cox.mat))
Heatmap(ori.mat, cluster_rows = clusterWNA(ori.mat))

submat <- ori.wm.mat[rownames(ori.wm.mat) %in% sigProts[.id == 'ori_w_matches', V1],]
rownames(submat) <- multiUniprots2multiGenes(rownames(submat))

hm <- Heatmap(submat, 
        cluster_rows = clusterWNA(submat),
        column_split = ifelse(grepl('Ecc', colnames(submat)), 'Ecc', 'WT')
        )
BackupAsPDF(hm, 'searchWMatches.sighits.heatmap', dimensions=c(8,8))
```
```{r}
submat <- ori.mat[rownames(ori.mat) %in% sigProts[.id == 'ori', V1],]
rownames(submat) <- multiUniprots2multiGenes(rownames(submat))

hm <- Heatmap(submat, 
        cluster_rows = clusterWNA(submat),
        column_split = ifelse(grepl('Ecc', colnames(submat)), 'Ecc', 'WT')
        )

BackupAsPDF(hm, 'ori.sighits.heatmap', dimensions=c(8,8))

```
cox heatmaps

```{r}
submat <- cox.mat[rownames(cox.mat) %in% sigProts[.id == 'ori', V1],]
rownames(submat) <- multiUniprots2multiGenes(rownames(submat))

hm <- Heatmap(submat, 
        cluster_rows = clusterWNA(submat),
        column_split = ifelse(grepl('Ecc', colnames(submat)), 'Ecc', 'WT')
        )

BackupAsPDF(hm, 'cox.sighits.heatmap', dimensions=c(8,12))


# what aboutt he entire cox matrix
hm <- Heatmap(cox.mat, 
        cluster_rows = clusterWNA(cox.mat),
        show_row_names = F,
        column_split = ifelse(grepl('Ecc', colnames(submat)), 'Ecc', 'WT')
        )
hm
BackupAsPDF(hm, 'cox.fullDT.heatmap', dimensions=c(8,12))
```

THis is not a good comparison between the groups.. too many missing values, 
Maybe want to use SAINT instead?? 
I think first we rerun with the correct parameters enabled... allow a fair comparison between the groups..




tidy the datatables, get gene names, highlight sig genes


When we remove missing value hits, there is no significantly different things between the conditions...

Looking the matrices, out of the 2k proteins detected, 

 # our first run
[1] 3120   19
[1] 287  19 # complete cases

# cox run
[1] 2527    7
[1] 848   7 # complete cases

# run with peptide matches
[1] 3699   19
[1] 660  19 # complete cases


```{r}
lapply(p.quant.list, function(x){
  x <- dcast(x, Protein~interaction(GROUP,SUBJECT,originalRUN), value.var = 'LogIntensities')
  dim(x) %>% print()
  
  x <- x[complete.cases(x),]
  dim(x) %>%  print()
})
```



```{r}
lapply(mss.list, function(x){
  
  x[, ProteinName := lapply(Protein, function(x){ tidyUniProtNames(x)})]
  
})
  
  
  grep("(?<=[|])[A-Z1-9]+")
  
  q(?=u) #matches a q that is followed by a u, without making the u part of the match
  
  (?<=a) # only matches b preceded by an a
  
}

test <- as.character(mss.list[[1]]$Protein)


sapply(strsplit(test, split = ';', fixed=T), function(x){ 
  str_extract(x,  "(?<=[|])[A-Z0-9]+(?=[|])") %>% 
    paste(., collapse=';')
  })




unlist()
strsplit(test[1],

help(strsplit)

strsplit(test, split=';', fixed = TRUE)


strsplit()

grep("(?<=[|])[A-Z0-9]+(?=[|])", perl=TRUE, mss.list[[1]]$Protein, ignore.case = F, value = TRUE)

str_extract(mss.list[[1]]$Protein, "(?<=[|])[A-Z0-9]+(?=[|])")

grep("[A-Z0-9]+(?=\\|)", mss.list[[1]]$Protein, value = T)

mss.list[[1]]$Protein

lapply(seq_along(mss.list), function(x,n,i){
  
  x[[i]][,ProteinName := gsub('sp[|]', '', Protein)]
  x[[i]][,ProteinName := gsub('[|].+', '', ProteinName)]
  
  print(x[[i]][, .(Protein,ProteinName)])
  
}, x=mss.list, n=names(mss.list))


lapply(mss.list, function(x){
  print(x[abs(log2FC) > 0.58 & adj.pvalue < 0.05,])
  print(x[abs(log2FC) > 0.58 & pvalue < 0.05,])
})
```



```{r}
mss.list %>%  names()
```



not used 
-----
 
```{r}
spec <- fread('/Users/martingordon/Documents/projects/020124_R_MCCall_Cox_APEX/data/RitaMC_MQOut/combined/txt/evidence.txt')
keys <- fread('./docs/keys.txt')

spec <- merge(x=spec, y=keys, by.x='Raw file', by.y='RawFile')

#remove contaminants
spec <- spec[!grepl("^$|CON__|REV__", Proteins),]

# for this assuming the mod id is a combo of the modifications and sequence cols only...
mod.seqs <- spec[Type !=  "TIMS-MULTI-MATCH", .(`Modified sequence`), by=.(
                                        Sequence,
                                        Modifications)] %>% unique()

# merge this info to the match prots
spec.match <- merge(x=spec[Type == "TIMS-MULTI-MATCH", -c("Modified sequence")], y=mod.seqs, by=c('Sequence', 'Modifications'))
```
 


try w/o the new sequences just drop the matched prots for testing...
works... need to find a better way to handle the naming of the matched proteins....
```{r}
spec <- fread('/Users/martingordon/Documents/projects/020124_R_MCCall_Cox_APEX/data/RitaMC_MQOut/combined/txt/evidence.txt')
keys <- fread('./docs/keys.txt')

spec <- merge(x=spec, y=keys, by.x='Raw file', by.y='RawFile')

#remove contaminants
spec <- spec[!grepl("^$|CON__|REV__", Proteins),]

#spec <- spec[Type !=  "TIMS-MULTI-MATCH",]

# just map to sequences for the matches... adding in the mod info causes the pipeline to fail
# dfix this first and reun
spec[Type ==  "TIMS-MULTI-MATCH", `Modified sequence` := paste0(Sequence,Modifications)]

# sum the intensity values for multiple peptides in each run
spec.noDups <- spec[, .(Intensity = sum(Intensity), count.features = .N),
                                by = .(Proteins, 
                                       `Modified sequence`,
                                        Charge,
                                        IsotopeLabelType,
                                        Condition,
                                        BioReplicate,
                                        Run)]


spec.mod.proc <- spec.noDups[,.(ProteinName = Proteins,
                                               PeptideSequence = `Modified sequence`,
                                                PrecursorCharge = Charge,
                                                FragmentIon = NA,
                                                ProductCharge  = NA,
                                                IsotopeLabelType,
                                                Condition,
                                                BioReplicate,
                                                Run,
                                                Intensity)]





spec.mod.proc[, oldName := paste0(Condition, '.',BioReplicate)]
spec.mod.proc[, TechReplicate := BioReplicate]
spec.mod.proc[, BioReplicate := str_extract(Condition, '[123]$')]
spec.mod.proc[, Condition := str_extract(Condition, "WT|EccCmut")]



dp.out <- MSstats::dataProcess(spec.mod.proc, 
                              MBimpute =  FALSE, 
                              featureSubset = "highQuality", 
                              remove_uninformative_feature_outlier = TRUE)

```


