---
title: "APEX_FirstLook"
author: "Martin Gordon"
date: "2024-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R McCall First Look

issue with number of features in each runs... compared this with the Berkley group and they are detecting more, but still issue with the 

packages

```{r}
library(data.table)
library (ComplexHeatmap)
library (circlize)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(magrittr)
library(MSstats)
library(ggrepel)
library(circlize)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

source ("../../utils/bp_utils/EvidenceFile2MSstatsInput.R")
# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```

create keys file
```{r}
keys <- fread('./data/RitaMC_MQOut/experimentalDesignTemplate.txt')
keys[, IsotopeLabelType := 'L']
keys[, Condition := sub("_[123]$","", Experiment)]
keys[, BioReplicate := str_extract(Experiment, "[123]$")]
keys[, RawFile := Name]
keys[, Run := .I]


fwrite(keys[, .(RawFile, IsotopeLabelType, Condition, BioReplicate,Run)], './docs/keys.txt')
```

read in data and create the evidence file

```{r}

anno <- fread('./data//RitaMC_MQOut/combined/txt/summary.txt')
protGroups <-  fread('./data//RitaMC_MQOut/combined/txt/proteinGroups.txt')
keys <-  fread('./docs/keys.txt')

keys
```

Read in the data and do a quick. QC

```{r}
spec <- fread('./data/RitaMC_MQOut/combined/txt/evidence.txt')

# convert MQ output to suitable format (drop multi-mapping peptides etc..)
spec.dt <- MaxQtoMSstatsFormat(evidence=spec, proteinGroups=protGroups, annotation=keys)
spec.dt <- as.data.frame(spec.dt) %>% 
  setDT()
spec.dt <- merge(x=keys, y=spec.dt[, .(ProteinName, PeptideSequence, PrecursorCharge, FragmentIon, ProductCharge,Run,Fraction, Intensity)], by.x='RawFile', by.y='Run')
```
Use BPs evidenceFile to MSstats input..

```{r}
spec.proc <- prepareDataForMSStats('./data/RitaMC_MQOut/combined/txt/evidence.txt', './docs/keys.txt')

any(spec.proc[, .N, by = .(Run, PeptideSequence, PrecursorCharge)]$N) > 1

col.pal <-  randomcoloR::distinctColorPalette(length(unique(spec.proc$Condition)))

g <- ggplot(spec.proc[!is.na(Intensity),.N, by=.(Condition,BioReplicate)], aes(x=reorder(interaction(Condition,BioReplicate)), y=N, fill=Condition)) +
  geom_bar(stat='Identity') +
  ggtitle('N detected features per sample') +
  scale_fill_manual(values= col.pal) +
  theme(axis.text.x = element_text(angle=90))

g
#BackupAsPDF(g, 'rawFeatureCounts.barplot')

```

Read in BPs evidence file and lets look at the number of features detected

```{r}
spec <- fread('/Users/martingordon/Documents/projects/020124_R_MCCall_Cox_APEX/data/RitaMC_MQOut/bp_res/evidence.txt')
keys.subset <- keys[RawFile %in% spec$`Raw file`,]
fwrite(keys.subset, './docs/keys.subset.txt')

spec.proc <- prepareDataForMSStats('/Users/martingordon/Documents/projects/020124_R_MCCall_Cox_APEX/data/RitaMC_MQOut/bp_res/evidence.txt',
                                   './docs/keys.subset.txt')


g <- ggplot(spec.proc[!is.na(Intensity),.N, by=.(Condition,BioReplicate)], aes(x=reorder(interaction(Condition,BioReplicate)), y=N, fill=Condition)) +
  geom_bar(stat='Identity') +
  ggtitle('N detected features per sample') +
  scale_fill_manual(values= col.pal) +
  theme(axis.text.x = element_text(angle=90))

g
```



check for and drop contaminants
```{r}
# remove contaminants
spec.dt <- spec.dt[grep('CON_|REV_', ProteinName, invert = T)]

#fwrite(spec.dt, './data/MQ_out_clean.txt')
spec.dt <- fread('./data/MQ_out_clean.txt')

```
intensities look fine, no spec junk on lower shoulder..
No multi peptides detected per condition
```{r}
any(spec.dt[, .N, by = .(Run, PeptideSequence, PrecursorCharge)]$N) > 1
```

```{r}
col.pal <-  randomcoloR::distinctColorPalette(length(unique(spec.dt$Condition)))


g <- ggplot(spec.dt[!is.na(Intensity),.N, by=.(Condition,BioReplicate)], aes(x=reorder(interaction(Condition,BioReplicate)), y=N, fill=Condition)) +
  geom_bar(stat='Identity') +
  ggtitle('N detected features per sample') +
  scale_fill_manual(values= col.pal) +
  theme(axis.text.x = element_text(angle=90))

g

BackupAsPDF(g, 'rawFeatureCounts.barplot')
```
```{r}

g <- ggplot(spec.dt, aes(x=reorder(interaction(Condition,BioReplicate)), y=log2(Intensity), fill=Condition)) +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values=col.pal) +
  theme(axis.text.x = element_text(angle=90))

g

BackupAsPDF(g, 'raw.intensities.boxplot')
```

```{r}
wide <- dcast(spec.dt, ProteinName+PeptideSequence+PrecursorCharge~Condition+BioReplicate, value.var = "Intensity")

# not v similiar and lots of missing values..
ggplot (wide, aes (x =log2(EccCmut_inf_BR1_1), y = log2(EccCmut_inf_BR2_1))) + geom_point(shape = ".") + geom_density_2d() + coord_fixed()
ggplot (wide, aes (x =log2(EccCmut_inf_BR1_1), y = log2(EccCmut_inf_BR1_2))) + geom_point(shape = ".") + geom_density_2d() + coord_fixed()

ggplot (wide, aes (x =log2(WT_inf_BR1_1), y = log2(EccCmut_inf_BR1_1))) + geom_point(shape = ".") + geom_density_2d() + coord_fixed()
```
```{r}
featureMat <- dcast(spec.dt,
                     paste0(PeptideSequence, PrecursorCharge)~interaction(Condition,BioReplicate), value.var = "Intensity") |> as.matrix(rownames = "PeptideSequence")

featureMat <- featureMat[complete.cases(featureMat), ]

```

As i thought.. one of the runs is problematic... maybe drop?
Look at the other MQ run to see how the other samples look...
May need to drop these..
Also, why is there a large amt of variance in the different samples?
```{r}

pcaOut <- prcomp(t(log2(featureMat)))

colInfo <- data.table(colname = colnames(featureMat))

colInfo
colInfo[,c("status","rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]
colInfo[, biorep := str_extract(status,"[123]$") ]

title <- "PCA"

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = status, shape = rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, "PCA_Complete_Features")
```
Regenerate and remove the other reps 
```{r}

dim(featureMat) # only 234 features w/o NA values??? This seeems v low... continue QC but check with Ben tomorrow
featureMat[,!colnames(featureMat) %like% "EccCmut_inf_BR1" ]

colnames(featureMat)
pcaOut <- prcomp(t(log2(featureMat[,!colnames(featureMat) %like% "EccCmut_inf_BR1" ])))

colInfo <- data.table(colname = colnames(featureMat))

colInfo
colInfo[,c("status","rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]
colInfo[, biorep := str_extract(status,"[123]$") ]

title <- "PCA"

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = status, shape = rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, "PCA_Complete_Features_noMut1Grp")
```
```{r}
featureMat <- dcast (spec.dt,
                     paste0(PeptideSequence, PrecursorCharge)~interaction(Condition,BioReplicate), value.var = "Intensity") |> as.matrix(rownames = "PeptideSequence")

#  log transform
featureMat <-  log2(featureMat)

dim(featureMat)
dim(featureMat[complete.cases(featureMat),]) #only 234 cases...
colnames(featureMat)

# create annotation
ann <- data.table(replicate=str_extract(colnames(featureMat),'[1-3]$'),
                  biogroup=str_extract(colnames(featureMat),'BR[1-3]'),
                  status=ifelse(grepl("EccCmut", colnames(featureMat)), 'EccCmut', 'WT'))


colours <- list('biogroup' = c('BR1' = '#4477AA', 'BR2'="#228833", 'BR3'="#CCBB44"),
                'replicate' = c('1' = '#4477AA', '2'='#66CCEE', '3'="#228833"),
                'status' =c('WT'="#CCBB44", 'EccCmut'='#EE6677')
                )

colAnn <- HeatmapAnnotation(df = ann, col = colours)



# subset the data
submat <- featureMat[sample(rownames(featureMat), 2000), ]
  

hm <- Heatmap(submat, cluster_rows=clusterWNA(submat), show_row_names = F, cluster_columns = T, top_annotation = colAnn, name='LogIntensities', column_names_gp = gpar(fontsize=6))
hm
BackupAsPDF(hm, 'features.clustered.heatmap')


submat <-  featureMat[complete.cases(featureMat),]

hm <- Heatmap(submat, cluster_rows=clusterWNA(submat), show_row_names = F, cluster_columns = T, top_annotation = colAnn, name='LogIntensities', column_names_gp = gpar(fontsize=6))
hm
BackupAsPDF(hm, 'features.completeCases.clustered.heatmap')


submat <- sweep(submat,1, STATS=apply(submat, 1, median, na.rm=T))
hm <- Heatmap(submat, cluster_rows=clusterWNA(submat), show_row_names = F, cluster_columns = T, top_annotation = colAnn, name='LogIntensities/Median',  column_names_gp = gpar(fontsize=6))
hm
BackupAsPDF(hm, 'features.completeCases.clustered.medianscaled.heatmap')
```
I need to check this with Ben tomorrow..

## Summarisation & Normalisation

Running MSstats processing

# protein summarization
Takes median intensity from the samples.. add a tech replicate column
see https://groups.google.com/g/msstats/c/6IMm-AMMOGA
```{r}

spec.dt[, Run := RawFile]
spec.dt[, oldName := paste0(Condition, '.',BioReplicate)]
spec.dt[, TechReplicate := BioReplicate]
spec.dt[, BioReplicate := str_extract(Condition, '[123]$')]
spec.dt[, Condition := str_extract(Condition, "WT|EccCmut")]

# check formatting
spec.dt[,.N, by=.(Condition, BioReplicate, oldName,TechReplicate,RawFile,Run)]


dp.out <- MSstats::dataProcess(spec.dt, 
                              MBimpute =  FALSE, 
                              featureSubset = "highQuality", 
                              remove_uninformative_feature_outlier = TRUE)
```


```{r}
saveRDS(dp.out,  "./output/120224_dp.out.rds")

fwrite (dp.out$ProteinLevelData, "./output/AB.ProteinLevelData.csv")
fwrite (dp.out$FeatureLevelData, "./output/AB.FeatureLevelData.csv.gz")
```



```{r}
p.quant <- fread("./output/AB.ProteinLevelData.csv")
```

boxplots of normalised intensities

```{r}
g <- ggplot(p.quant, aes (x= interaction(SUBJECT, GROUP,RUN), y = LogIntensities, fill = GROUP)) +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values=col.pal) +
  theme(axis.text.x = element_text(angle=90))
g
BackupAsPDF(g, 'protein.intensities.boxplot')
```

Protein Counts per sample

```{r}
g <- ggplot(p.quant[,.N,by=.(GROUP,SUBJECT,RUN)], aes(x=reorder(interaction(GROUP,SUBJECT,RUN)), y = N, fill = GROUP)) +
  geom_bar(stat='Identity') +
  theme_classic() +
  scale_fill_manual(values=col.pal) +
  theme(axis.text.x = element_text(angle=90))
g
BackupAsPDF(g, 'nProteins.barplot')
```
# pca of the protein summarised data
 
```{r}
prot.mat <- dcast (p.quant, Protein~interaction(GROUP,SUBJECT,RUN), value.var = "LogIntensities") |> as.matrix (rownames = "Protein")
colnames(prot.mat)
prot.mat %>% View()

p.quant[Protein == 'sp|A0MZ66|SHOT1_HUMAN' & GROUP == 'WT' & SUBJECT == '1',]

prot.mat <- prot.mat[complete.cases(prot.mat),]
```


```{r}

pcaOut <- prcomp(t(prot.mat))

colInfo <- data.table(colname = colnames(prot.mat))
colInfo[,c("status","rep") := tstrsplit(colnames(prot.mat), "[.]", keep = c(1,2)) ]
colInfo[, run := str_extract(status,"[123]$") ]

colInfo
title <- "PCA"

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)


pcaDT %>%  head()
#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = interaction(status,rep), shape = rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, "PCA_Complete_Proteins.")


#and with BR1 dropped...

pcaOut <- prcomp(t(prot.mat[, !colnames(prot.mat) %like% "EccCmut_inf_BR1"]))

colInfo
colInfo[,c("status","rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]
colInfo[, biorep := str_extract(status,"[123]$") ]

title <- "PCA"

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
pcaDT[, `:=`(timepoint = ifelse(grepl("[47]D", rn), str_extract(rn, "[47]D"), 'na'),
             treat = gsub("[_.].+", "", rn)
             )]

#plot first two components
p <- ggplot (pcaDT, aes(x=PC2, y=PC3,  fill = status, shape = rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, "PCA_Complete_Proteins_nomut1grp.")
```

variance explained:
Huge amt explained due to first PC which corresponds to condition, which is good, but why is there so few proteins and features identified....
I dont know anything else about model ro conditions, but 

```{r}
q <- qplot(1:10,  pcaPercentVar[1:10]) +
  geom_line() +
  xlab('Princpal Component 1:10') +
  ylab(('% Variance explained')) +
  ggtitle('All samples scree plot')

q
BackupAsPDF(q, 'pca.all.screeplot.')
```
biotinylated carboxylases
-----
Plot the biotinylated carboxylases
None of the endogenous proteins detected... have we used the correct uniprot file?
```{r}


biotin.carboxylases.up <- c("O00763","P05165","P11498","Q13085","Q96RQ3")
p <- ggplot(p.quant[Protein %in% biotin.carboxylases.up], aes (x = interaction ( SUBJECT, GROUP,RUN ), y = LogIntensities, color = Protein)) + geom_line(aes(group = Protein)) + geom_point(aes(shape = as.factor(SUBJECT)))
#ackupAsPDF(p, "BiotinCarboxylaseLevels_")
```



read back in the R object and lets run DEA
We want to model using lmm; add a 'batch' term to the subject

```{r}
dp.out <- readRDS("./output/120224_dp.out.rds")

# we want to model 'batch' random effects
# two options: pseudobulk (just take the mean of each group to summarise or lmm)
# we want to 'batch' the groupd
p.quant <-  setDT(dp.out$ProteinLevelData)
#p.quant[, SUBJECT := paste0(GROUP,SUBJECT)]
p.quant[, SUBJECT := paste0('batch.', SUBJECT)]

dp.out$ProteinLevelData <- p.quant

f.quant <- setDT(dp.out$FeatureLevelData)
#f.quant[,SUBJECT := paste0(GROUP,SUBJECT)]
f.quant[,SUBJECT := paste0('batch.', SUBJECT)]
        
        
dp.out$ProteinLevelData <- p.quant
dp.out$FeatureLevelData <- f.quant

dp.out$ProteinLevelData
```



```{r}
contrasts.mat <- matrix(c(-1,1), nrow=1, ncol=2)
colnames(contrasts.mat) <-  c('WT', 'EccCmut')
rownames(contrasts.mat) <- 'EccCmut-WT'

# run msstats correcting for batch 
mss <- groupComparison(contrast.matrix=contrasts.mat, data=dp.out)
mss.dt <- setDT(mss$ComparisonResult)

mss.dt[abs(log2FC) > 0.58 & pvalue < 0.05 & is.finite(log2FC), ]

# larger SE when. modelling with lmm (batch)
#7 and 12 DF: SE val with lmm 0.1879289	 0.1022959	
```

Contact Rita about the analysis and request access to the ArtMs sata



not needed...
---------

```{r}
spec.dt <-  spec
spec.dt[, ProteinName := `Leading proteins`]

spec.dt <- spec.dt[grep('CON_', ProteinName, invert = T)]
spec.dt <- spec.dt[grep('REV_', ProteinName, invert = T)]

spec.dt %>%  names()

spec

sp
```

```{r}

col.pal <-  randomcoloR::distinctColorPalette(length(unique(spec.dt$Condition)))

g <- ggplot(spec.dt, aes(x=reorder(interaction(Condition,BioReplicate)), y=log2(Intensity), fill=Condition)) +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values=col.pal) +
  theme(axis.text.x = element_text(angle=90))

g

BackupAsPDF(g, 'raw.intensities.boxplot')
```

Missed clevages

```{r}
spec.dt$`Potential contaminant` %>%  unique()
```

Issue here, the first mutant biorep has far less identified peptides than the others,
Only 30k features identified also??? consistent across all tech reps..

```{r}

Portein.
spec.dt <- spec.dt[Intensity > 2^5,]

g <- ggplot(spec.dt[,.N, by=.(Condition,BioReplicate)], aes(x=reorder(interaction(Condition,BioReplicate)), y=N, fill=Condition)) +
  geom_bar(stat='Identity') +
  ggtitle('N features per sample') +
  scale_fill_manual(values= col.pal) +
  theme(axis.text.x = element_text(angle=90))

g

BackupAsPDF(g, 'rawFeatureCounts.barplot')

spec.dt$Condition %>%  unique()

```
```{r}
wide <- dcast(spec.dt, Proteins+PeptideSequence+PrecursorCharge+Type~Condition+BioReplicate, value.var = "Intensity")

(wide)


spec.dt[PeptideSequence == 'GFAFVEYESHR' & Condition == 'EccCmut_inf_BR2',]
# good agreement between replicates it looks like 

# one of the mocks may be an issue but generally v simialr 
ggplot (wide, aes (x =log2(EccCmut_inf_BR1), y = log2(Vehicle_7D_2))) + geom_point(shape = ".") + geom_density_2d() + coord_fixed()
ggplot (wide, aes (x =log2(x4206_7D_2), y = log2(x4206_7D_4))) + geom_point(shape = ".") + geom_density_2d() + coord_fixed()

spec$Type %>% unique()
```
 PCA features 
 ----
```{r}
featureMat <- dcast(spec.dt,
                     paste0(PeptideSequence,'.', PrecursorCharge)~interaction(Condition,BioReplicate), value.var = "Intensity") |> as.matrix(rownames = "PeptideSequence")


featureMat
featureMat <- featureMat[complete.cases(featureMat), ]

spec.dt[PeptideSequence == 'AAAAAAALQAK' & Condition ==  "EccCmut_inf_BR1" & BioReplicate == 1,]


spec.dt$Condition %>%  unique()
```


```{r}
pcaOut <- prcomp(t(log2(featureMat)))

colInfo <- data.table(colname = colnames(featureMat))

colInfo
colInfo[,c("status","rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]

title <- "PCA"

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = status, shape = rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, "PCA_Complete_Features")
```

```{r}
featureMat <- dcast (spec.dt,
                     paste0(PeptideSequence, PrecursorCharge)~interaction(Condition,BioReplicate), value.var = "Intensity") |> as.matrix(rownames = "PeptideSequence")

#  log transform
featureMat <-  log2(featureMat)

dim(featureMat)
dim(featureMat[complete.cases(featureMat),])
colnames(featureMat)

# create annotation
ann <- data.table(replicate=str_extract(colnames(featureMat),'[1-5]$'),
                  treatment=sub("[_.].+", "", colnames(featureMat)),
                  timepoint=ifelse(grepl("[47]D", colnames(featureMat)), str_extract(colnames(featureMat), "[47]D"), 'na'))


colours <- list('timepoint' = c('na' = '#4477AA', '4D'="#228833", '7D'="#CCBB44"),
                'replicate' = c('1' = '#4477AA', '2'='#66CCEE', '3'="#228833", '4'="#CCBB44", '5'='#EE6677'),
                'treatment' =c('Mock' = col.pal[1], 'Vehicle'=col.pal[2], 'N1062D'=col.pal[3], 'x3769'=col.pal[4], 'x4052'=col.pal[5], 'x4206'=col.pal[6])
                )

colAnn <- HeatmapAnnotation(df = ann, col = colours)



# subset the data
submat <- featureMat[sample(rownames(featureMat), 2000), ]
  


hm <- Heatmap(submat, cluster_rows=clusterWNA(submat), show_row_names = F, cluster_columns = T, top_annotation = colAnn, name='LogIntensities', column_names_gp = gpar(fontsize=6))
hm
BackupAsPDF(hm, 'features.clustered.heatmap')

submat <- sweep(submat,1, STATS=apply(submat, 1, median, na.rm=T))
hm <- Heatmap(submat, cluster_rows=clusterWNA(submat), show_row_names = F, cluster_columns = T, top_annotation = colAnn, name='LogIntensities/Median',  column_names_gp = gpar(fontsize=6))
hm
BackupAsPDF(hm, 'features.clustered.medianscaled.heatmap')

```