---
title: "DDA_BPdata_firstLook"
author: "Martin Gordon"
date: "2024-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

packages
```{r}
library(data.table)
library (ComplexHeatmap)
library (circlize)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(magrittr)
library(MSstats)
library(ggrepel)
library(circlize)
library(ggvenn)
library(readxl)
library(MSstats)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

source ("../../utils/bp_utils/EvidenceFile2MSstatsInput.R")
# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```


## BP data check

Due to issues of low counts in my maxquant search output, perform quick QC of BPs run and compare N detected features etc.
Low count issues still an issue so not related to my MQ params ( I think)


Next steps:
----
Look at the fasta; should we blast it to identify the spec it maps to?

```{r}
#spec <- fread('/Users/martingordon/Documents/projects/020124_R_MCCall_Cox_APEX/data/RitaMC_MQOut/combined/txt/evidence.txt')
#keys.subset <- keys[RawFile %in% spec$`Raw file`,]
#fwrite(keys.subset, './docs/keys.subset.txt')


spec.proc <- prepareDataForMSStats('/Users/martingordon/Documents/projects/020124_R_MCCall_Cox_APEX/data/RitaMC_MQOut/combined/txt/evidence.txt',
                                   './docs/keys.txt')
```

```{r}
g <- ggplot(spec.proc[!is.na(Intensity),.N, by=.(Condition,BioReplicate)], aes(x=reorder(interaction(Condition,BioReplicate)), y=N, fill=Condition)) +
  geom_bar(stat='Identity') +
  ggtitle('N detected features per sample') +
  scale_fill_manual(values= col.pal) +
  theme(axis.text.x = element_text(angle=90))

BackupAsPDF(g, 'Nfeaturesdetected.UCSF.barplot')
```
Compare this to Ritas evidence.txt

```{r}
spec <- fread('~/Library/CloudStorage/Box-Box/APEX\ Data\ to\ Share/code/evidenceApex12.txt')
spec$`Raw file` %>%  unique()
spec$Experiment %>%  unique()

#keys <- fread('~/Library/CloudStorage/Box-Box/APEX\ Data\ to\ Share/keys.txt')
#keys[,RawFile := sub('S','', RawFile)]
#fwrite(keys, '~/Library/CloudStorage/Box-Box/APEX\ Data\ to\ Share/keys.tidy.txt')
keys <- fread('~/Library/CloudStorage/Box-Box/APEX\ Data\ to\ Share/keys.tidy.txt')

spec.proc <- prepareDataForMSStats('~/Library/CloudStorage/Box-Box/APEX\ Data\ to\ Share/code/evidenceApex12.txt',
                                   '~/Library/CloudStorage/Box-Box/APEX\ Data\ to\ Share/keys.tidy.txt')
```

```{r}

g <- ggplot(spec.proc[!is.na(Intensity),.N, by=.(BioReplicate,Run)], aes(x=reorder(interaction(BioReplicate,Run)), y=N, fill=BioReplicate)) +
  geom_bar(stat='Identity') +
  ggtitle('N detected features per sample') +
  scale_fill_manual(values= col.pal) +
  theme(axis.text.x = element_text(angle=90))

g

BackupAsPDF(g, 'Nfeaturesdetected.berkely.barplot')

```

Lets look at the output from MaxQuant; view the txt.summary file
Seems that their search returned ~ 30% more identifications on average... how can we correct this?

```{r}
# compare to our summary DT
sum.dt <- fread ('/Users/martingordon/Documents/projects/020124_R_MCCall_Cox_APEX/data/RitaMC_MQOut/combined/txt/summary.txt')
summary.dt <- fread('~/Library/CloudStorage/Box-Box/APEX\ Data\ to\ Share/summary.txt')

summary.dt

wide.dt <- merge(x=sum.dt, y=summary.dt, by.x='Raw file', by.y='Raw file', suffixes = c(x=".ucsf", y=".berkely"))
wide.dt <- merge(x=wide.dt, y=keys, by.x='Raw file', by.y='RawFile')

# peptides seem the same
g <-  ggplot(wide.dt, aes(x=`Peptide sequences identified.ucsf`, y=`Peptide sequences identified.berkely`, color=interaction(Condition,BioReplicate))) +
  geom_point(size=2.4) +
  geom_abline(intercept=0, slope = 1) +
    xlim(0,20000) +
  ylim(0,20000) 
g
#BackupAsPDF(g, 'npeptidesMQout.scatterplot')

# lets plot more than one variable.... peptides and MS/MS scans?
g <-  ggplot(wide.dt, aes(x=`MS/MS identified.ucsf`, y=`MS/MS identified.berkely`, color=interaction(Condition,BioReplicate))) +
  geom_point(size=2.4) +
  geom_abline(intercept=0, slope = 1) +
  xlim(5000,22000) +
  ylim(5000,22000) 
g

#BackupAsPDF(g, 'nscansIdentified.scatterplot')


g <- ggplot(comb.dt[`Raw file` != 'Total'], aes(x=id, y=`Peptide sequences identified`, fill=id)) +
  geom_bar(stat='Identity') + 
  theme(axis.text.x = element_text(angle=90))

#BackupAsPDF(g, 'oldVsNewMQsearch.peptides identified')



# get amt
comb.dt[, `Peptide sequences identified` := as.double(`Peptide sequences identified`)]
comb.dt[, N.diff := (`Peptide sequences identified`[id == 'berkley'] - `Peptide sequences identified`[id == 'ucsf']), by=.(`Raw file`)]

ggplot(comb.dt, aes(x=`Raw file`, y=N.diff)) +
  geom_bar(stat='Identity') +
   theme(axis.text.x = element_text(angle=90))


head(comb.dt)

comb.dt[,.N,by=.( `Peptide sequences identified`, `Raw file`, `id`, N.diff )]
```

What about PEP scores in both files
Berkley has a greater spread of PEP values... less trustworhty?
PEP scores are NA for peptides identified through matching

```{r}
# compare to our summary DT
spec.ucsf <- fread('/Users/martingordon/Documents/projects/020124_R_MCCall_Cox_APEX/data/RitaMC_MQOut/combined/txt/evidence.txt')  
spec.ucsf[, source := 'ucsf']
spec.berkley <- fread('~/Library/CloudStorage/Box-Box/APEX\ Data\ to\ Share/code/evidenceApex12.txt')
spec.berkley[, source := 'berkley']

files.oi <- spec.berkley$`Raw file` %>% unique()

spec.merge <- rbind(spec.ucsf, spec.berkley, fill=T)


# boxplots of pep; more variance in berkley scores
ggplot(spec.merge[`Raw file` %in% files.oi & !is.na(PEP), ], aes(y=PEP, color=source)) +
  geom_boxplot() +
  facet_wrap(~`Raw file`) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90))


# plot this 
# proportion of NA PEP scores
g <- ggplot(spec.merge[`Raw file` %in% files.oi, ], aes(x=source, fill=is.na(PEP))) +
  geom_bar() +
  facet_wrap(~`Raw file`)
g

# what are the PEP  NA vals? these are all from multi-matches
g <- ggplot(spec.merge[`Raw file` %in% files.oi, ], aes(x=source, fill=Type)) +
  geom_bar() +
  facet_wrap(~`Raw file`)
g
```
Look at breakdown of mods in each set; could the unexplained be related to the deamination proportion?

```{r}
# look at the mods; the top ones are the ones we have included...
#< Modifications included in protein quantification      Oxidation (M);Acetyl (Protein N-term)
#> Modifications included in protein quantification      Oxidation (M);Acetyl (Protein N-term);Deamidation (NQ)

spec.merge[, Mod := ifelse(grepl('Acetyl', Modifications), 'Acetyl',
                                       ifelse(grepl('Deamidation', Modifications), 'Deamidation',
                                                          ifelse(grepl('Oxidation', Modifications), 'Oxidation',
                                                                 ifelse(grepl('Glu->pyro-Glu', Modifications), 'Glu-Pyro-Glue',
                                                                             'Unmodified' ))) )]

# lets simplify even further; just seperate into modifiied and unmodified
spec.merge[, Mod.simplified := ifelse(Modifications == 'Unmodified', 'Unmodified', 'Modified')]

col.vec <- randomcoloR::distinctColorPalette(k=length(unique(spec.merge$Mod)))

g <- ggplot(spec.merge[`Raw file` %in% files.oi, .N, by=.(source, Mod, `Raw file`)], aes(x=interaction(source), y=N, fill=Mod)) +
  geom_bar(stat='Identity') +
  scale_fill_manual(values=col.vec) +
  facet_wrap(~`Raw file`)
g
#BackupAsPDF(g, 'ModificationCounts.barplots')

```
Lets look again at the Venn Diagrams
Exclude the TIMS=MULTI-MSMS category and lets look within those detected directly in the run only

No still doesnt seem to resolve the issue....

```{r}
spec.merge[`Raw file` %in% files.oi,  .N, .(source, Sequence, Mod)] %>% 
  .[order(Sequence)]


# convert to a list of sequences assigned to each of these groups and inspect overlap in both
# I think we only want to look at one run at a time for this...

spec.merge[`Raw file` == "14Apr2023_DDH_RM1_2_S3-A2_1_5548" & !is.na(PEP), .N, by=.(source, Mod.simplified)]

venn.list <- list(ucsf.unmodified = spec.merge[`Raw file` == "14Apr2023_DDH_RM1_2_S3-A2_1_5548" & !is.na(PEP) & source == 'ucsf' & Modifications =='Unmodified', Sequence],
                  ucsf.modified = spec.merge[`Raw file` == "14Apr2023_DDH_RM1_2_S3-A2_1_5548"  & !is.na(PEP)   & source == 'ucsf' & Modifications !='Unmodified', Sequence],
                  berkley.unmodified = spec.merge[`Raw file` == "14Apr2023_DDH_RM1_2_S3-A2_1_5548" & !is.na(PEP)  & source == 'berkley' & Modifications =='Unmodified', Sequence],
                  berkley.modified = spec.merge[`Raw file` == "14Apr2023_DDH_RM1_2_S3-A2_1_5548"  & !is.na(PEP)   & source == 'berkley' & Modifications !='Unmodified', Sequence]
                  )

g <- ggvenn(venn.list)
g
#BackupAsPDF(g, 'run5548.peptideBreakdown.vennDiagram')

```

What about the subset of sequences not in ours, what do their scores look like
more unique. sequecnes identified in our data than theres, but only when NA peptides with PEP scores included...

```{r}

pep.berkley <- spec.merge[source == 'berkley', unique(Sequence)]
pep.ucsf<- spec.merge[source == 'ucsf', unique(Sequence)]

# looks 
pdf('pepCompareSource.histograms.pdf')
par(mfrow=c(1,2))
hist(spec.merge[`Raw file` %in% files.oi & source == 'berkley' & !Sequence %in% pep.ucsf, ]$PEP) # this is saying many of the sequences in the dataset
hist(spec.merge[`Raw file` %in% files.oi & source == 'ucsf' & !is.na(PEP) & !Sequence %in% pep.berkley, ]$PEP)
dev.off()

seq.counts.dt <- spec.merge[, unique(Sequence), by=.(`Raw file`,source)] %>% 
  .[, .N, by=.(source, `Raw file`)]

ggplot(seq.counts.dt[`Raw file` %in% files.oi,], aes(x=source, y=N, fill=source)) +
  geom_bar(stat='Identity') +
  facet_wrap(~`Raw file`)

```
At this stage, maybe best to summarise both to proteins and see what is different between the two
-----------------
Straight off, we are not importing the matched peptides from the MQ run


```{r}
nrow(spec.proc[PeptideSequence %in% no.match.peps,]) == nrow(spec.proc)

# lets get list of the na PEP scores and see if they are included in the output after processing 
match.peps <- spec.merge[source == 'ucsf' & is.na(PEP), unique(`Modified sequence`)]
no.match.peps <- spec.merge[source == 'ucsf' & !is.na(PEP), unique(`Modified sequence`)]

spec.merge[is.na(PEP), .N, .(Type, Sequence, `Modified sequence`)]$`Modified sequence` %>%  unique()

spec.proc <- prepareDataForMSStats('/Users/martingordon/Documents/projects/020124_R_MCCall_Cox_APEX/data/RitaMC_MQOut/combined/txt/evidence.txt',
                                   './docs/keys.txt')

spec.rmc.proc <- prepareDataForMSStats('~/Library/CloudStorage/Box-Box/APEX\ Data\ to\ Share/code/evidenceApex12.txt',
                                       '~/Library/CloudStorage/Box-Box/APEX\ Data\ to\ Share/keys.tidy.txt')
```
lets process this data oursleves..
```{r}
spec <- fread('/Users/martingordon/Documents/projects/020124_R_MCCall_Cox_APEX/data/RitaMC_MQOut/combined/txt/evidence.txt')
keys <- fread('./docs/keys.txt')

spec <- merge(x=spec, y=keys, by.x='Raw file', by.y='RawFile')

#remove contaminants
spec <- spec[!grepl("^$|CON__|REV__", Proteins),]

# just map to sequences for the matches...
spec[Type ==  "TIMS-MULTI-MATCH", `Modified sequence` := Sequence]

# sum the intensity values for multiple peptides in each run
spec.noDups <- spec[, .(Intensity = sum(Intensity), count.features = .N),
                                by = .(Proteins, 
                                       `Modified sequence`,
                                        Charge,
                                        IsotopeLabelType,
                                        Condition,
                                        BioReplicate,
                                        Run)]


spec.mod.proc <- spec.noDups[,.(ProteinName = Proteins,
                                               PeptideSequence = `Modified sequence`,
                                                PrecursorCharge = Charge,
                                                FragmentIon = NA,
                                                ProductCharge  = NA,
                                                IsotopeLabelType,
                                                Condition,
                                                BioReplicate,
                                                Run,
                                                Intensity)]


```

no multifeatures detected 

```{r}
# no multiple feature peptide ions detected
spec.proc[!is.na(Intensity),.N, by=.(PeptideSequence,PrecursorCharge,Run)][N >1]
spec.mod.proc[!is.na(Intensity),.N, by=.(PeptideSequence,PrecursorCharge,Run)][N >1]
```

```{r}
spec.list <-  list('ori'=spec.proc,
                   'cox'=spec.rmc.proc,
                   'ori-w-matches' = spec.mod.proc)



col.pal <-  randomcoloR::distinctColorPalette(length(unique(spec.proc$Condition)))


lapply(spec.list, function(x){

g <- ggplot(x[!is.na(Intensity),.N, by=.(Condition,BioReplicate,Run)], aes(x=reorder(interaction(Condition,BioReplicate,Run)), y=N, fill=Condition)) +
  geom_bar(stat='Identity') +
  ggtitle('N features per sample') +
  scale_fill_manual(values= col.pal) +
  theme(axis.text.x = element_text(angle=90))
g

#BackupAsPDF(g, 'rawFeatureCounts.barplot')

})
#BackupAsPDF(g, 'rawFeatureCounts.barplot')
spec.list
```
try this modified mapping... definitely better than just pasting sequences together...
```{r}


# identify the modified sequences in each group
mod.seq <- spec[Type !=  "TIMS-MULTI-MATCH", .(`Modified sequence`, count.mods = .N), by=.(
                                                                                Sequence,
                                                                                Modifications
                                                                                )]


mod.seq
# now merge these identifications to the mod table 
# look up this... 
test <- merge(x=spec, y=mod.seqs, by.x=c('Sequence', 'Modifications'), by.y=c('Sequence', 'Modifications'))

spec.mod.proc <- spec.noDups[,.(ProteinName = Proteins,
                                              PeptideSequence = `Modified sequence`,
                                                PrecursorCharge = Charge,
                                                FragmentIon = NA,
                                                ProductCharge  = NA,
                                                IsotopeLabelType,
                                                Condition,
                                                BioReplicate,
                                                Run,
                                                Intensity)]


# no; for this match, you want to get the IDs of the other 
mod.seqs <- spec[Type !=  "TIMS-MULTI-MATCH", .(`Modified sequence`), by=.(
                                        Sequence,
                                        Modifications)]

mod.seqs[.SD, on=.(Sequence, Modifications)]

# looking at htis but look later...
table1[is.na(C), C := table2[.SD, on=.(A, B), x.C]]


# 



spec[Type ==  "TIMS-MULTI-MATCH", `Modified sequence` := mod.seqs[.SD, on=.(Sequence, Modifications), x.`Modified sequence`] ]



mod.seqs[.SD, on=.(Sequence, Modifications)]
spec[Type ==  "TIMS-MULTI-MATCH", `Modified sequence` := Sequence]

```

lets compare the peptides in the new processed vs old

```{r}
spec.mod.proc[PeptideSequence %in% match.peps,]

nrow(spec.proc[PeptideSequence %in% no.match.peps,]) == nrow(spec.proc)


nrow(spec.mod.proc)
nrow(spec.proc)
nrow(spec.rmc.proc)

spec.proc[is.na(Intensity),]

```





g <- ggplot(spec.merge[`Raw file` %in% files.oi, ], aes(x=PEP, fill=source)) +
  geom_histogram(binwidth = 0.01) +
  facet_wrap(~`Raw file`)
g
#BackupAsPDF(g, 'PEPscores.histogram')

# plot this 
# proportion of NA PEP scores
g <- ggplot(spec.merge[`Raw file` %in% files.oi, ], aes(x=source, fill=is.na(PEP))) +
  geom_bar() +
  facet_wrap(~`Raw file`)
g

#BackupAsPDF(g, 'PEPscores.propNA.barplots')


# peptides seem the same
g <-  ggplot(wide.dt, aes(x=`Peptide sequences identified.ucsf`, y=`Peptide sequences identified.berkely`, color=interaction(Condition))) +
  geom_point(size=2.4) +
  geom_abline(intercept=0, slope = 1) +
    xlim(0,20000) +
  ylim(0,20000) 

g

#BackupAsPDF(g, 'npeptidesMQout.scatterplot')

PEP score... why so many NA values in our runs??
PEP NA scores are found across all modification groups.... ! reason we have these PEP NA scores is match between Runs!!!
Still doesnt explain why we find less hits...
```{r}
ggplot(spec.merge[`Raw file` %in% files.oi, .N, by=.(source, `Raw file`, is.na(PEP))], 
       aes(x=source, y=N, fill=is.na)) +
  geom_bar(stat='Identity') +
  facet_wrap(~`Raw file`)

spec.merge[`Raw file` %in% files.oi & is.na(PEP), .N, by=.(source, Mod)]


spec.merge[is.na(PEP),]
```
Seems that UCSF has more detected, but more of our PEP values are NA.. why??
```{r}
# look at the mods; the top ones are the ones we have included...
#< Modifications included in protein quantification      Oxidation (M);Acetyl (Protein N-term)
#> Modifications included in protein quantification      Oxidation (M);Acetyl (Protein N-term);Deamidation (NQ)


spec.merge[, Mod := ifelse(grepl('Acetyl', Modifications), 'Acetyl',
                                       ifelse(grepl('Deamidation', Modifications), 'Deamidation',
                                                          ifelse(grepl('Oxidation', Modifications), 'Oxidation',
                                                                 ifelse(grepl('Glu->pyro-Glu', Modifications), 'Glu-Pyro-Glue',
                                                                             'Unmodified' ))) )]


spec.merge[Mod == 'Unmodified', unique(Modifications)]

col.vec <- randomcoloR::distinctColorPalette(k=length(unique(spec.merge$Mod)))

g <- ggplot(spec.merge[`Raw file` %in% files.oi, .N, by=.(source, Mod, `Raw file`)], aes(x=interaction(source), y=N, fill=Mod)) +
  geom_bar(stat='Identity') +
  scale_fill_manual(values=col.vec) +
  facet_wrap(~`Raw file`)
g
BackupAsPDF(g, 'ModificationCounts.barplots')
```
peptide IDs; 
Look for how many times peptide IDs may have cropped up; are the same IDs in the unmodified and other modified group

venn diagram
Problem is, even though Berkley has many more sequences, the

```{r}
spec.merge[`Raw file` %in% files.oi,  .N, .(source, Sequence, Mod)] %>% 
  .[order(Sequence)]


# convert to a list of sequences assigned to each of these groups and inspect overlap in both
# I think we only want to look at one run at a time for this...

(spec.merge[`Raw file` == "14Apr2023_DDH_RM1_2_S3-A2_1_5548", .N, by=.(source, Mod)])


venn.list <- list(ucsf.unmodified = spec.merge[`Raw file` == "14Apr2023_DDH_RM1_2_S3-A2_1_5548"  & source == 'ucsf' & Modifications =='Unmodified', Sequence],
                  ucsf.modified = spec.merge[`Raw file` == "14Apr2023_DDH_RM1_2_S3-A2_1_5548"  & source == 'ucsf' & Modifications !='Unmodified', Sequence],
                  berkley.unmodified = spec.merge[`Raw file` == "14Apr2023_DDH_RM1_2_S3-A2_1_5548"  & source == 'berkley' & Modifications =='Unmodified', Sequence],
                  berkley.modified = spec.merge[`Raw file` == "14Apr2023_DDH_RM1_2_S3-A2_1_5548"  & source == 'berkley' & Modifications !='Unmodified', Sequence]
                  )


g <- ggvenn(venn.list)
g
BackupAsPDF(g, 'run5548.peptideBreakdown.vennDiagram')


files.oi

spec.merge
```

