---
title: "R Notebook"
output: html_notebook
---

Script to loop through PTM analysis

```{r}
library (data.table)
library (ComplexHeatmap)
library (ggplot2)
library(tidyverse)
rotate.x.axis.text <- theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
#source("./scripts/LocatePeptidePtmInProteins.R") #jacks version of script

source("../../utils/bp_utils/LocatePeptidePtmInProteins.R")

WriteSessionInfo()
WriteInstalledPackages()
```

Configure paths to files

```{r setup}
#path to files on your machine
path <- "/Users/martingordon/Documents/projects/071723_JMoen_SquidPTM"

fasta <- paste0(path,'/docs/squid_Annotated_v5_tryptic_edits_fixed.fasta')  #use this as dups removed

f_input <- dir(path = paste0(path, '/data/'), pattern = "*.All_PTMs_fixed.txt", recursive = FALSE, full.names = TRUE) 
k_input <- dir(path = paste0(path, '/data/'), pattern = "*_keys.txt", recursive = FALSE, full.names = TRUE)


input <-  expand.grid(f_input,k_input)


input <- input %>% 
  mutate(ptm_type=as.factor(str_extract(k_input, pattern="AcK|diMeK|monoMeK|triMeK|PH|UB"))) %>% 
  arrange(ptm_type)
```

```{r functions}

#file preprocessing
preprocessSpec <- function(spec_f,keys_f){

  #fread stores as data.table by default
  spec <- fread(spec_f)
  keys <- fread(keys_f)

  #add semicolon if required 
  spec[, ProteinName := gsub ("(?<![_>])PAC", ";PAC", ProteinName, perl = TRUE)] #uses perl regex v with negative lookbehind to negate -/> characters 


  # for some reason the '>' was kep in the protein names here.  Remove those:
  spec[, ProteinName := gsub (">", "", ProteinName)]

  # apply Condition, BioReplicate from keys to spec
  spec[keys , Condition := i.Condition , on = c(Run = "RawFile")] #overwrite condition col w values from keys file, match run col in spec to rawfile col in keys
  spec[keys , BioReplicate := i.BioReplicate , on = c(Run = "RawFile")]
  
  #only select defined runs
  spec <- spec[Condition != 'Not Defined',]
  
  return(spec)
  setnames(spec) = dat_name
}

#run msstats

runMSStats <- function(spec_in){

  # required column for MSstats
  spec_in[, IsotopeLabelType := "L"]

  dp.out <- MSstats::dataProcess(spec_in, MBimpute = FALSE, featureSubset = "highQuality", remove_uninformative_feature_outlier = TRUE)
  return(dp.out)
}

# create volcano plot
```

```{r make-contrast-matrix}
contrasts <- scan(paste0(path,"/data/contrasts.txt"), what = "character")
```

```{r run-preprocesing}
# not working.. read in the files and see 
test.keys <- fread('/Users/martingordon/Documents/projects/071723_JMoen_SquidPTM/data//diMeK_keys.txt')
test.f <- fread('/Users/martingordon/Documents/projects/071723_JMoen_SquidPTM/data/MSstats_All_PTMs_fixed.txt')

unique(test.keys$RawFile)

test.f[Run %in% unique(test.keys$RawFile),] %>% 
  .[, unique(Run)]

head(test.keys)
head(test.f)

test.f[test.keys, Condition := i.Condition, on = c(Run = 'RawFile')]

#apply to each row of df and name after val in third column
results_list <- setNames(apply(input, 1,function(row) {
  spec=row[1]
  keys=row[2]
  preprocessSpec(spec,keys) }), input$ptm_type)

```

```{r}
lapply(results_list, function(x){
  x[, .N, by=.(Condition, BioReplicate, Run)]
})
```

Add run info for modelling
```{r}
# check out the nbioconditions
lapply(results_list, function(x){
  x[, ID:=.GRP, by=.(Condition,BioReplicate)]
  x[, BioReplicate := ID]
  x[, .N, by=.(Condition, BioReplicate, Run)]
  
})
```


Spectronaut data trim left tail which is obviously garbage,
Left shoulder peak: seems imputiation may have been used

```{r plot-hist}

#decide on cutoffs

lapply(results_list, function(dt){
  
  ggplot(data=dt, aes(x=log2(Intensity))) +
    geom_histogram(stat='bin', bins=100)

} )


# specify threshold (may need to do this individually for each list element)
results_list <- lapply(results_list, function(x){
  x <- x[Intensity >= 2^4]
} )


lapply(results_list, function(dt){
  
  ggplot(data=dt, aes(x=log2(Intensity))) +
    geom_histogram(stat='bin', bins=100)

} )

```
Run the PH

```{r}

ptm.list <- lapply(seq_along(results_list), function(x, n, i){

  print(n[[i]])
  
  if (n[[i]] == 'AcK') {
    sitifyProteins_SpectronautFile(x[[i]], fastaFile = fasta, site = "NAC") # for n-acetyl
  } else if (n[[i]] == 'monoMeK') {
    sitifyProteins_SpectronautFile(x[[i]], fastaFile = fasta, site = "MMET") 
  } else if (n[[i]] == 'diMeK') {
    sitifyProteins_SpectronautFile(x[[i]], fastaFile = fasta, site = "DMET") 
  } else if (n[[i]] == 'triMeK') {
    sitifyProteins_SpectronautFile(x[[i]], fastaFile = fasta, site = "TMET") 
  } else if (n[[i]] == 'UB' | n[[i]] == 'PH') {
    sitifyProteins_SpectronautFile(x[[i]], fastaFile = fasta, site = n[[i]]) 
  }
  
}, x=results_list, n=names(results_list))

names(ptm.list) <- names(results_list)
```

```{r filter dts}

ptm.list <- lapply(ptm.list, function(dt){
  dt <- dt[!is.na(dt$oldProteinName)]
  return(dt)
} )

lapply(ptm.list, function(dt){
  
  ggplot(data=dt, aes(x=log2(Intensity))) +
    geom_histogram(stat='bin', bins=100)

} )

```

```{r run-msstats}
# make contrasts
#put in function

# run msstats on each list element
msstats_proc <- lapply(ptm.list, function(dt) {
  runMSStats(dt)
})

# write out files to output directory
imap(msstats_proc, function(dt, idt) {
  fwrite(dt$ProteinLevelData, ScriptAndDatedFileName(paste0(idt,"-ProteinLevelData.csv")))
  fwrite(dt$FeatureLevelData, ScriptAndDatedFileName(paste0(idt,"-FeatureLevelData.csv")))
})

# MSStats Differential Expression
msdiff_list <- lapply(msstats_proc, function(dt) {
  contrast.mat <- makeContrast.AllByAll(mssQ = dt)
  gc.out <- MSstats::groupComparison(contrast.mat, dt)
  return(gc.out)
})

# write out differential expression results to file
imap(msdiff_list, function(dt, idt) {
  fwrite(dt$ComparisonResult, ScriptAndDatedFileName(paste0(idt,"-diff-results.txt")))
})

```

Generate volcano plots
```{r}
lapply(seq_along(msdiff_list), function(x, n, i) {
  
  
  dt <- data.table(x[[i]]$ComparisonResult)
  
  pvalueThreshold <- 0.05
  log2FCThreshold <- log2(1.5) # modify as needed
  pvalueVariable <-  "adj.pvalue" # or "pvalue"
  dt[, sig := "not"]
  dt[dt[[pvalueVariable]] < pvalueThreshold & abs(log2FC) > log2FCThreshold, sig := ifelse(log2FC  < 0, "down", "up")]
  dt[, yVariable := -log10(dt[[pvalueVariable]])]
  
  p <- ggplot (dt, aes(x=log2FC, y = yVariable, color = sig)) + 
  geom_point(show.legend = FALSE, alpha = 0.5, size = 1, shape = 16) + 
  facet_wrap (~Label) + 
  scale_color_manual(values = c(down= "blue", not = "gray", up = "red")) +
  #scale_x_continuous(limits = c(-4,4)) +
  scale_y_continuous(name = paste(pvalueVariable, "(-log10)", collapse = "")) +
  geom_hline(yintercept = -log10(pvalueThreshold), lty="dotted") +
  geom_vline(xintercept = c(-1, 1) * log2FCThreshold, lty="dotted") + 
  theme_bw() 

  pdf(paste0(n[[i]],"-Volcano_plot.pdf"))
  print(p)
  dev.off()
  
  
}, x=msdiff_list, n=names(msdiff_list))

```


# don't run: just examples of command for each. PTM
sitifyProteins_SpectronautFile(results_list[['PH']], fastaFile = fasta, site = "PH") 
sitifyProteins_SpectronautFile(results_list[['AcK']], fastaFile = fasta, site = "NAC") 
sitifyProteins_SpectronautFile(results_list[['monoMeK']], fastaFile = fasta, site = "MMET")
sitifyProteins_SpectronautFile(results_list[['diMeK']], fastaFile = fasta, site = "DMET") 
sitifyProteins_SpectronautFile(results_list[['triMeK']], fastaFile = fasta, site = "TMET") 
sitifyProteins_SpectronautFile(results_list[['UB']], fastaFile = fasta, site = "UB")

