---
title: "043025_QCpFunctions"
author: "Martin Gordon"
date: "2025-04-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overivew 

Write a number of functions to visualize the QC quality of different replicates; these need to be implemented as part of the standard SEC-MS workflow
- first; start from peptide data; create a clean long format file and then write a funciton to summarize to Proteins (will use mstats)
- at protein level; look at correlations of sister peptides across the replicates (with/wo phagoSome and riboSome proteins)


Look at Bens complexes and see how they appear

```{r cars}
library(MSstats)
library(magrittr)
library(data.table)
library(ggplot2)
library(ComplexHeatmap)
library(stringr)
library(eulerr)

source("~/Documents/utils/bp_utils/ManageScriptData.R") # read the SECms utils library
source("~/Documents/utils/bp_utils/SEC_MS_utils.R") # read the SECms utils library
source("~/Documents/utils/bp_utils/SEC_MS_utils.localMods.R")
```
Read in the peptide level data
```{r}
lysisPrep.list[[1]]
lysisPrep.list <- list('sonic'=fread("/Users/martingordon/Documents/projects/043025_MGordon_SECMS_QCplotsWork/data/2025_02_SEC_Lysis_Test/Spectronaut/2. NP40-Lysis buffer + Sonication/20250115_092735_MM_Sonic_Lysis_122724_MSStats_Report.txt"),
                       'FT'=fread("/Users/martingordon/Documents/projects/043025_MGordon_SECMS_QCplotsWork/data/2025_02_SEC_Lysis_Test/Spectronaut/3. NP40-Lysis buffer + Freeze-Thaw/20250114_172438_MM_FT_Lysis_122724_MSStats_Report.txt"),
                       'CL'=fread("/Users/martingordon/Documents/projects/043025_MGordon_SECMS_QCplotsWork/data/2025_02_SEC_Lysis_Test/Spectronaut/1. CellLytic M Lysis buffer/20250114_135820_MM_CL_Lysis_122724_MSStat_Report.txt"))
```
Need to tidy up this list; get the fraction ID

```{r}
lapply(lysisPrep.list, function(x){
  
  x[, Fraction := str_extract(Run, '(?<=Frac-)[0-9]{1,2}')]
  x[, Condition :=  paste0(Condition, '.', Fraction)] 
  x[, IsotopeLabelType := 'L']
})
```

```{r}
# looking at some missing fractions in the two groups.. lets summarize first, then remove
lapply(lysisPrep.list, function(x){
  x[,.N, by=.(BioReplicate,Condition)][,.N, by=Condition][N < 2,]
})

# for a quick QC check, lets look at the data
lysisPrep.dt <- lysisPrep.list %>% 
  rbindlist(., idcol='sample')

# tidy colnames to conform with SEDC_MS standards
lysisPrep.dt[, treatment := sample]
lysisPrep.dt[, sample := paste0(sample, '_', BioReplicate)] #unique ID
lysisPrep.dt[, fraction := as.integer(str_extract(Run, '(?<=Frac-)[0-9]{1,2}'))]
#lysisPrep.dt[, fraction := factor(fraction, levels=as.character(seq(1,72,1)))]
lysisPrep.dt[, protein := ProteinName]
lysisPrep.dt[, peptide := paste0(PeptideSequence, '.', PrecursorCharge)]
lysisPrep.dt[, replicate := BioReplicate]
lysisPrep.dt[, intensity := Intensity]

cols.oi <- c('sample', 'treatment', 'replicate', 'fraction', 'protein', 'peptide', 'intensity')

# work with this for now
sec.dt <- lysisPrep.dt[, ..cols.oi]
sec.dt[, peptideSequence := gsub('[.][0-9]+$','', peptide)]

# how many peptides per protein and how often are they observed
sec.dt[, nfeatures := .N, by=.(protein,sample, fraction)] # n features per protein/fraction
```

save the file to output
```{r}
#fwrite(sec.dt, ScriptAndDatedFileName('lysis.test.raw.peptide.longformat.csv.gz'))
sec.dt <- fread('~/Documents/projects/043025_MGordon_SECMS_QCplotsWork/043025_SECMS_QCplots_data/2025_04_30_lysis.test.raw.peptide.longformat.csv.gz')
```

collapse features to peptides (CCprofiler does this to help with completeness; just sum precursor ints)
Look at the distributions of the nFractions observed

```{r}
sec.dt <- sumFeatureIntensities(sec.dt)

sec.dt[, .(nTimesObserved=.N), by=.(treatment,replicate,peptideSequence)][replicate == 2]
g <- ggplot(sec.dt[, .(nTimesObserved=.N), by=.(treatment,replicate,peptideSequence)], aes(x=nTimesObserved, color=treatment, linetype=as.factor(replicate))) +
  geom_freqpoly(bins=72) +
  labs(title='Distribution of peptide detections', x='n fractions observed in') +
  scale_color_brewer(palette = 1, type='qual') +
  theme_bw()
g
BackupAsPDF(g, 'npeptideObsAcrossfractions.frequencypoly')

g <- ggplot(sec.dt[, .(nTimesObserved=.N), by=.(treatment,replicate,peptideSequence)], aes(x=nTimesObserved, color = treatment, linetype=as.factor(replicate))) +
  geom_density() +
  labs(title='Dpeptide detections', x='n fractions observed in') +
  scale_color_brewer(palette = 1, type='qual') +
  theme_bw() +
  guides(linetype=guide_legend('replicate'))
g
BackupAsPDF(g, 'npeptideObsAcrossfractions.density')
```
N observed peptides per fraction

```{r}
g <- ggplot(sec.dt[, .(nObserved=length(unique(protein))), by=.(treatment,replicate,fraction)], aes(x=fraction, y=nObserved, color = treatment, shape=as.factor(replicate))) +
  geom_line(aes(group=interaction(treatment, replicate), linetype = as.factor(replicate))) +
  labs(title='Number of peptides detected', x='fraction number') +
  scale_fill_brewer(palette = 1, type='qual') +
  scale_color_brewer(palette = 1, type='qual') +
  theme_bw() +
  guides(linetype=guide_legend('replicate'))
g
BackupAsPDF(g, 'numPeptidesDetected.linechart')

# another way to look at it; plot
g <- ggplot(sec.dt[, .(nObserved=length(unique(protein))), by=.(treatment,replicate,fraction)], aes(x=fraction, y=nObserved, fill = treatment)) +
  geom_bar(stat='identity') +
  labs(title='Number of peptides detected', x='fraction number') +
  scale_fill_brewer(palette = 1, type='qual') +
  scale_alpha_manual(values=c('1'=1, '2'=0.5)) +
  theme_bw() +
  facet_grid(replicate~treatment) +
  guides(linetype=guide_legend('replicate'))
g
BackupAsPDF(g, 'numPeptidesDetected.linechart')
```
Look at distributions of the consecutive detections per run;
What does a good protein look like?

```{r}
#' Flag low quality peptides based on missingness proportion, lack of consecutive detections <5 for now as required for peak detection
FlagLowQualityPeptides <- function(secLong.dt, idcol='peptideSequence', intsCol='intensity', ConsecFractionThreshold=5, detectionCutoff=0){
  
  .oneMatrix <- function(sub.dt){
    mat <- dcast(sub.dt, sub.dt[[idcol]]~fraction, value.var = intsCol) %>% 
      as.matrix(rownames = 1)
    mat[is.na(mat)] <- 0.0
    mat[order(rownames(mat)),]
  }
  
    #' Identify the max number of consecitive measurmeents
  .MaxConsecFractionMeasurments <- function(row, cutoff=detectionCutoff){
    # get unique values in a vector and the maximum run of consec vals
    detectVec <- rle(row > cutoff)
    return(max(detectVec$lengths[detectVec$values == TRUE]))
  }
  
  # one matrix per sample 
  mats <- lapply(split(secLong.dt, list(secLong.dt$sample)), .oneMatrix)
  
  consec.dt <- lapply(mats, function(x) {
    apply(x, 1, function(pep){ MaxRunOfConsecFractionMeasurments(row = pep, cutoff = 0) }) %>% 
    as.data.table(keep.rownames=T)
  }) %>% 
  rbindlist(idcol='sample')
  
  setnames(consec.dt, new=c('sample', 'peptideSequence', 'consecutiveDetections'))
  consec.dt[, passThreshold := ifelse(consecutiveDetections >= ConsecFractionThreshold, TRUE, FALSE)]
  return(consec.dt)
}
```

```{r}
consec.dt <- FlagLowQualityPeptides(sec.dt, idcol = 'peptide')
consec.dt[, c('treatment', 'replicate') := tstrsplit(sample, '_', keep=c(1,2))]
```

```{r}
g <- ggplot(consec.dt, aes(x=consecutiveDetections, fill=interaction(treatment,replicate), linetype=replicate)) +
  geom_freqpolygon(bins=72) +
  labs(title='Max number of consecutive detections per peptide', x='N fractions', y='N peptides') +
  scale_color_brewer(palette = 1, type='qual') +
  scale_alpha_manual(values=c('1'=1, '2'=0.5)) +
  theme_bw() +
  guides(linetype=guide_legend('replicate'))
g
#BackupAsPDF(g, 'npeptideConsecutiveMeasurements.linechart')

g <- ggplot(consec.dt, aes(x=consecutiveDetections, fill=interaction(treatment,replicate), linetype=replicate)) +
  geom_histogram(bins=72) +
  labs(title='Max number of consecutive detections per peptide', x='N fractions', y='N peptides') +
  scale_color_brewer(palette = 1, type='qual') +
  scale_alpha_manual(values=c('1'=1, '2'=0.5)) +
  theme_bw() +
  guides(linetype=guide_legend('replicate'))
g

g <- ggplot(consec.dt, aes(x=paste0(treatment,'_',replicate), fill=passThreshold)) +
  geom_bar(position = 'stack') +
  labs(title='Number of passing peptides per sample', x='sample', y='N peptides') +
  scale_fill_brewer(palette = 3, type='qual') +
  scale_alpha_manual(values=c('1'=1, '2'=0.5)) +
  theme_bw() +
  guides(linetype=guide_legend('replicate'))
g
#BackupAsPDF(g, 'npeptidePassingThresholds.barplot')
```
What is a good peptide? 
Get an indication of peptide ids
```{r}
# adds col for filtering and n consecutive runs
filterPeptides(sec.dt, consec.dt)
```
Now of the good peptides, see how many overlap between the different conditions
```{r}
g <- ggplot(sec.dt[passThreshold==TRUE,unique(peptideSequence), by=.(treatment, replicate)][,.N, by=.(treatment,replicate)], aes(x=paste0(treatment, replicate), y=N, fill=treatment)) +
  geom_bar(stat='identity') +
  labs(title='Good peptides per sample', x='sample', y='counts') +
  scale_fill_brewer(palette = 1, type='qual') +
  scale_alpha_manual(values=c('1'=1, '2'=0.5)) +
  theme_bw() +
  guides(linetype=guide_legend('replicate'))
g
BackupAsPDF(g, 'nGoodpeptidesPersample.barplot')
```

Overlapping good peptide ids... what lysis prep method looks best?
```{r}

protTraces
# need to get htis as character vector
ids <- sec.dt[passThreshold == TRUE, unique(peptideSequence), by=.(sample)] %>% 
  split(., by='sample', keep.by=F) 
ids <- lapply(ids, function(x) unname(unlist(x)))

g <- plot(euler(ids),
     labels = list(font = 4),
     alpha=0.7,
     fill=c(values=c("#3a5e8cFF","#3a5e8cFF", "#ffcf20FF","#ffcf20FF","#541352FF","#541352FF")),
    quantities = TRUE)
g
BackupAsPDF(g, 'sig.ppiOverlap.euler')
```
sec.ms imputation; question is if imputation can clean up the results protein summarization
Question is whether imputation cleans the intensity summarization? We dont know; need to run  and assess
MStstats imputation looks to impute censored data (true 0s under limit of detection),  looks across runs, find the lowest val (LOD) and imputes a value lower than this
```{r}
#' Read in the peptide DT, filter low qual proteinsand then run
MSstatsFormatting <-  function(secLong.dt){
 
  stopifnot(c('peptide', 'protein', 'intensity', 'sample', 'treatment', 'fraction', 'replicate') %in% colnames(secLong.dt))
  stopifnot(is.numeric(secLong.dt$fraction) & is.numeric(secLong.dt$replicate))
   
  message('Converting input DT to format expected by MSstats...')
  .msstatsifyDTformat <- function(secLong.dt){
    # 
    ms.dt <- secLong.dt[, .(
      Condition = paste(treatment, fraction, sep='_'),
      BioReplicate = replicate,
      ProteinName=protein,
      Intensity=intensity,
      PeptideSequence=peptideSequence,
      PrecursorCharge = as.numeric(gsub('.+[_][.]', '', peptide)),
      FragmentIon = NA,
      ProductCharge = NA,
      IsotopeLabelType = 'L')
    ]
    ms.dt[, Run := paste(Condition, BioReplicate, sep='_')]
    return(ms.dt)
  }
  mss.dt <- .msstatsifyDTformat(secLong.dt)
  return(mss.dt)
}

```

## Run MSSTats summarization w and without imputation; save both files seperately and investigate their sister peptide correlations tomorrow
Run failed; try facet by sample or at least condition
Remove peptides with 0 across all rows
```{r}
mss.dt <- MSstatsFormatting(secLong.dt = sec.dt[passThreshold == TRUE,])
#fwrite(mss.dt, ScriptAndDatedFileName('mssInput.csv.gz'))
mss.dt <- fread('~/Documents/projects/043025_MGordon_SECMS_QCplotsWork/043025_SECMS_QCplots_data/2025_05_02_mssInput.csv.gz')

```
msstats summarize
```{r}
msStatsSummarization <- function(mss.dt, imputeVals=FALSE){
  
  stopifnot(c('Condition', 'BioReplicate', 'ProteinName', 'Intensity', 'PeptideSequence', 'PrecursorCharge', 'FragmentIon','ProductCharge', 'IsotopeLabelType', 'Run') %in% colnames(mss.dt))
  
  message('Converting 0 values to NA...')
  mss.dt[Intensity == 0, Intensity := NA]
  message('Running msstats summarization...')
  if (imputeVals){
    message('Warning: imputation is enabled')
  }
  dp.out <- MSstats::dataProcess(mss.dt, 
                                 MBimpute =  imputeVals, 
                                 normalization = FALSE, # turn off for SECMS
                                 summaryMethod = "TMP",
                                 featureSubset = 'highQuality',
                                 censoredInt = "NA", # consider NA not missing at random and impute
                                 remove_uninformative_feature_outlier=T)

  return(dp.out)
}

View(ScriptNamedDir)
```

Run per sample, lets see if its just an issue with one sample and we can troubleshoot
```{r}
#lets just run the imputation per sample as a test and save the outpu
mss.dt[, treatment := gsub('_.+','', Condition)]
mss.dt[, conditionOI := paste0(treatment, '.',BioReplicate)]

# all unique set
mss.dt[,.N, by=.(Run,ProteinName, PeptideSequence, PrecursorCharge)]$N %>% unique() 

condition.oi <- names(split(mss.dt, list(mss.dt$treatment, mss.dt$BioReplicate)))
#dont know why this failed
lapply(condition.oi, function(cond){
  
  dp.out <- msStatsSummarization(mss.dt = mss.dt[conditionOI == cond,], imputeVals=FALSE)
  saveRDS(dp.out, ScriptAndDatedFileName(paste0(cond,'.mss.dp.no.impute.rds')))
})
```
So individual samples run fine; lets combine two too see if this still throws an issue, then test by modifying birep column
```{r}
lapply(c("sonic.1", "CL.2","FT.2","sonic.2"), function(cond){
  
  dp.out <- msStatsSummarization(mss.dt = mss.dt[conditionOI == cond,], imputeVals=FALSE)
  saveRDS(dp.out, ScriptAndDatedFileName(paste0(cond,'.mss.dp.no.impute.rds')))
})
```

Run summarisation with imputation turned on
```{r}
lapply(condition.oi, function(cond){
  
  dp.out <- msStatsSummarization(mss.dt = mss.dt[conditionOI == cond,], imputeVals=TRUE)
  saveRDS(dp.out, ScriptAndDatedFileName(paste0(cond,'.mss.dp.impute.rds')))
})
```

Read in and combine all the protein level quants, and lets see if we see an improvement on sister protein quantification/correlation
First, try run the quantification on all files at onece and savew

```{r}
dp.out <- msStatsSummarization(mss.dt = mss.dt,imputeVals = TRUE)
saveRDS(dp.out, ScriptAndDatedFileName('mss.dp.impute.combined.rds'))
```

plots; prot-prot correlation in specific complexes and correlate replicates with each other

I think we need to first test this on one run and see where it falls down, or is it one sample problematic?
```{r}
oneSample <- mss.dt[Condition %like% 'sonic' & BioReplicate == 1,]

testRun <- msStatsSummarization(oneSample, imputeVals = F)
```


# lets try summarizing the results
```{r}
ms.noimpute.dt <- msStatsSummarization(mss.dt=mss.dt, minNumberFeatures = 2, imputeVals = TRUE)
saveRDS(ms.noimpute.dt, ScriptAndDatedFileName('mss.dp.no.impute.rds'))

ms.impute.dt <- msStatsSummarization(mss.dt=mss.dt, minNumberFeatures = 2, imputeVals = FALSE)
saveRDS(ms.impute.dt, ScriptAndDatedFileName('mss.dp.impute.rds'))
```
Keep an et

```{r}

mss.dt
```


```{r}


sec.dt
View(data)

mss.impiute <- MSstatsSummarize(sec.dt, minNumberFeatures = 2, imputeVals=TRUE)
saveRDS(mss.impiute, ScriptAndDatedFileName('mss.dp.impute.rds'))
```

Go look at the quantifications that go up or down and investigat

```{r}
sec.dt
mss.dt <- MSstatsSummarize(secLong.dt = sec.dt,imputeZero = TRUE)
mss.d

t
```
```{r}
lysisPrep.dt
```


```{r}

print('heel')
# filters out proteins missing in at least 5 fractions
FlagLowQualityPeptides <- function(secLong.dt, idcol='peptideSequence', intsCol='intensity', ConsecFractionThreshold=5, detectionCutoff=0){
  
  .oneMatrix <- function(sub.dt){
    mat <- dcast(sub.dt, sub.dt[[idcol]]~fraction, value.var = intsCol) %>% 
      as.matrix(rownames = 1)
    mat[is.na(mat)] <- 0.0
    mat[order(rownames(mat)),]
  }
  
    #' Identify the max number of consecitive measurmeents
  .MaxConsecFractionMeasurments <- function(row, cutoff=detectionCutoff){
    # get unique values in a vector and the maximum run of consec vals
    detectVec <- rle(row > cutoff)
    return(max(detectVec$lengths[detectVec$values == TRUE]))
  }
  
  # one matrix per sample 
  mats <- lapply(split(secLong.dt, list(secLong.dt$sample)), .oneMatrix)
  
  consec.dt <- lapply(mats, function(x) {
    apply(x, 1, function(pep){ MaxRunOfConsecFractionMeasurments(row = pep, cutoff = 0) }) %>% 
    as.data.table(keep.rownames=T)
  }) %>% 
  rbindlist(idcol='sample')
  
  setnames(consec.dt, new=c('sample', 'peptideSequence', 'consecutiveDetections'))
  consec.dt[, passThreshold := ifelse(consecutiveDetections >= ConsecFractionThreshold, TRUE, FALSE)]
  return(consec.dt)
}
```



```{r}
testmat <- dcast(sec.dt[sample == 'FT_1'], peptideSequence~fraction, value.var='intensity') %>% 
  as.matrix(rownames='peptideSequence')

testmat[is.na(testmat)] <- 0


sec.dt


cor.dt <- apply(testmat, 1, function(row) MaxRunOfConsecFractionMeasurments(row, 0))
cor.dt <- data.table(peptide=names(cor.dt), nConsec=cor.dt)


ggplot(cor.dt, aes(x=nConsec)) +
  geom_histogram(bins=72)



badpeptides <- cor.dt[nConsec < 5]$peptide 
```

```{r}
hm <- Heatmap(testmat[rownames(testmat) %in% badpeptides,], cluster_columns = F, show_row_names = F)
hm
BackupAsPDF(hm, ScriptAndDatedFileName('badproteins.heatmap'))
```


```{r}
# there are some genuine waringingas alos
sec.dt[sample == 'CL_1' & peptideSequence == '_[Acetyl (Protein N-term)]AQPGTLNLNNEVVK_']
lysisPrep.dt[sample == 'CL' & BioReplicate == 1 & grepl('AQPGTLNLNNEVVK_', PeptideSequence)]



```


Tag good peptides per protein
A good peptide should i) have a run of at least 5 fractions (peak picking algorithm needs this)

```{r}
sec.dt
test.dt[]
DetectGoodQualityPeptides(test.dt[sample == 'sonic_1'])
```


Define some QC summary metrics

First off; define a set of proteins we may want to exclude from the analysis
Pull out phagosome and ribosome proteins
```{r}
source("~/Documents/utils/bp_utils/enrichmentTestFunctions.R")

gmt.cc <- loadGmtFromBioconductor(ontology = "CC")

bigComplexes <- rbind(data.table(gmt.cc[grepl('ribosome', ont), .(GO_CC='ribosome', gene)]),
                      data.table(gmt.cc[grepl('phagosome', ont), .(GO_CC='phagosome', gene)]))
```
Ok here is the plots of the sister peptide correlations in each sample
I think we can filter this to remove noisy peptides and replot
```{r}
rawpeptideCor <- correlateSisterPeptideElutionProfiles(sec.dt)

g <- ggplot(rawpeptideCor, aes(x=r, color=treatment, linetype = replicate)) +
  geom_density() +
  labs(title='sister peptide elution profiles') +
  scale_color_brewer(palette = 1, type='qual') +
  theme_bw()

BackupAsPDF(g, 'sisterPeptideElutions.density')
# rerun w/o all the ribosomal stuff and see how things look

# for some reason we are not filtering out bg complexes; need to fix this...
filteredCor.dt <- correlateSisterPeptideElutionProfiles(secLong.dt = sec.dt[!protein %in% bigComplexes$gene,])

g <- ggplot(filteredCor.dt, aes(x=r, color=treatment, linetype = replicate)) +
  geom_density() +
  labs(title='sister peptide elution profiles') +
  scale_color_brewer(palette = 1, type='qual') +
  theme_bw()
g
BackupAsPDF(g, 'sisterPeptideElutions.density')

```

sum feature intensities to the peptide level and replot; see if this improves the agreement
```{r}
#' ccprofiler setup; summed peptide intensities
#' Collapse to the peptide level 
#' 
sec.dt
sumFeatureIntensities <- function(secLong.dt){
  
  return(secLong.dt[, .(intensity=sum(intensity, na.rm=TRUE), npeptides=npeptides, ntimesObserved), by=.(sample, treatment, replicate, fraction, protein, peptideSequence)])
  
}

collapse.sec.dt <- sumFeatureIntensities(sec.dt)

collapse.sec.dt[npeptides == 32,]
```


*not used*

sister peptide correlations
```{r}
#' within each sample, calculate sister peptide correlations basically do an all by all correlation within each protein
#' remove multi-proteins; just want a quick overview for simplicity
#' excludedProteins: vecotr of proteins to remove from the datasrt

test.dt <- sec.dt[treatment == 'sonic' & protein %in% c('P02538', 'P04264')]

test.dt[, if(.N > 30) .(meanints=mean(intensity, na.rm=T)), by=.(sample, protein)]

correlateElutionProfiles <- function(secLong.dt, idcol='peptide', intsCol='intensity', excludedSet=c()){
  
  if (is.null(excludedSet)){ 
    allProteins <- unique(secLong.dt$protein)
  } else {
    rmSet <- sum(unique(secLong.dt$protein) %in% unique(excludedSet))
    message('Removing ', rmSet, ' proteins from calculations..')
    allProteins <- unique(secLong.dt$protein[!secLong.dt$protein %in% excludedSet])
  }
  .oneMatrix <- function(sub.dt){
    mat <- dcast(sub.dt, sub.dt[[idcol]]~fraction, value.var = intsCol) %>% 
      as.matrix(rownames = 1)
    mat[is.na(mat)] <- 0.0
    mat[order(rownames(mat)),]
  }
  
  .correlateSisterPeptides <- function(mat){
    cor.mat <- cor(t(mat), use='pairwise.complete.obs')
    r.dt <- setDT(reshape2::melt(cor.mat))
    setnames(r.dt, new=c('feature1', 'feature2', 'r'))
    return(r.dt)
  }
  
  multipep.dt <-  secLong.dt[npeptides > 2,]
  mats <- lapply(split(multipep.dt, list(multipep.dt$sample, multipep.dt$protein)), .oneMatrix)
  print(mats[1])
  
  cor.dt <- lapply(mats, .correlateSisterPeptides) %>% 
    rbindlist(idcol = 'subset')
  cor.dt[, c('sample', 'protein') := tstrsplit(subset, '[.]', keep=c(1,2))]
  
  return(cor.dt[feature1 != feature2, .(sample, protein, feature1, feature2, r)])
}
```
Look at distributions of sister peptide correlations in the same sample; do the distirbutions look good?
```{r}
# sanity check that this runs for more than none treatment
test.dt <- sec.dt[treatment == 'sonic']

rawpeptideCor <- correlateSisterPeptideElutionProfiles(sec.dt)
rawpeptideCor
g <- ggplot(rawpeptideCor, aes(x=r, color=treatment, linetype = replicate)) +
  geom_density() +
  labs(title='sister peptide elution profiles') +
  scale_color_brewer(palette = 1, type='qual') +
  theme_bw()
g
BackupAsPDF(g, 'sisterPeptideElutions.density')
# rerun w/o all the ribosomal stuff and see how things look

# for some reason we are not filtering out bg complexes; need to fix this...
filteredCor.dt <- correlateSisterPeptideElutionProfiles(secLong.dt = sec.dt[!protein %in% bigComplexes$gene,])

g <- ggplot(filteredCor.dt, aes(x=r, color=treatment, linetype = replicate)) +
  geom_density() +
  labs(title='sister peptide elution profiles') +
  scale_color_brewer(palette = 1, type='qual') +
  theme_bw()
g
BackupAsPDF(g, 'sisterPeptideElutions.density')

features.oi <- c('_NAWGNLSYADLITR_.2', '_AGSAMAIGGGGGSR_.2')

# a lot of missingness for these 'crap' scores; remove 
mat <- dcast(sec.dt[treatment == 'CL' & replicate == '1' & peptide %in% features.oi,], peptide~fraction, value.var='intensity') %>% 
 as.matrix(rownames=1)
mat[is.na(mat)] <- 0
Heatmap(mat)
```

# peptide max number of detected consec runs, and summarize this at protein level (avg consec runs, sd consec runs)

```{r}

sec.dt
#' ccprofiler setup; summed peptide intensities
#' Collapse to the peptide level 
sumFeatureIntensities <- function(secLong.dt){
  
  return(secLong.dt[, .(peptide=peptideSequence, intensity=sum(intensity, na.rm=TRUE), npeptides=npeptides), by=.(sample, treatment, replicate, fraction, protein)])
  
}

collapse.sec.dt <- sumFeatureIntensities(sec.dt)



sec.dt[peptideSequence == '_QQVSPIR_']
collapse.sec.dt[peptide == '_QQVSPIR_']
```


Other summary stats; n consecutive measurements per feature, missingness per peptide/protein
Write generally, so that we can modify to run for everyhtin


```{r}
#' Flag low quality peptides based on missingness proportion, lack of consecutive detections <5 for now as required for peak detection
FlagLowQualityPeptides <- function(secLong.dt, idcol='peptideSequence', intsCol='intensity', ConsecFractionThreshold=5, detectionCutoff=0){
  
  .oneMatrix <- function(sub.dt){
    mat <- dcast(sub.dt, sub.dt[[idcol]]~fraction, value.var = intsCol) %>% 
      as.matrix(rownames = 1)
    mat[is.na(mat)] <- 0.0
    mat[order(rownames(mat)),]
  }
  
    #' Identify the max number of consecitive measurmeents
  .MaxConsecFractionMeasurments <- function(row, cutoff=detectionCutoff){
    # get unique values in a vector and the maximum run of consec vals
    detectVec <- rle(row > cutoff)
    return(max(detectVec$lengths[detectVec$values == TRUE]))
  }
  
  # one matrix per sample 
  mats <- lapply(split(secLong.dt, list(secLong.dt$sample)), .oneMatrix)
  
  consec.dt <- lapply(mats, function(x) {
    apply(x, 1, function(pep){ MaxRunOfConsecFractionMeasurments(row = pep, cutoff = 0) }) %>% 
    as.data.table(keep.rownames=T)
  }) %>% 
  rbindlist(idcol='sample')
  
  setnames(consec.dt, new=c('sample', 'peptideSequence', 'consecutiveDetections'))
  consec.dt[, passThreshold := ifelse(consecutiveDetections >= ConsecFractionThreshold, TRUE, FALSE)]
  return(consec.dt)
}

View(MaxRunOfConsecFractionMeasurments)
consec.dt <- FlagLowQualityPeptides(sec.dt)
consec.dt[, c('treatment', 'replicate') := tstrsplit(sample, '_', keep=c(1,2))]

g <- ggplot(consec.dt, aes(x=consecutiveDetections, fill=interaction(treatment,replicate), linetype=replicate)) +
  geom_freqpolygon(bins=72) +
  labs(title='Max number of consecutive detections per peptide', x='N fractions', y='N peptides') +
  scale_color_brewer(palette = 1, type='qual') +
  scale_alpha_manual(values=c('1'=1, '2'=0.5)) +
  theme_bw() +
  guides(linetype=guide_legend('replicate'))
g
BackupAsPDF(g, 'npeptideConsecutiveMeasurements.linechart')

g <- ggplot(consec.dt, aes(x=paste0(treatment,'_',replicate), fill=passThreshold)) +
  geom_bar(position = 'stack') +
  labs(title='Number of passing peptides per sample', x='sample', y='N peptides') +
  scale_fill_brewer(palette = 3, type='qual') +
  scale_alpha_manual(values=c('1'=1, '2'=0.5)) +
  theme_bw() +
  guides(linetype=guide_legend('replicate'))
g
BackupAsPDF(g, 'npeptidePassingThresholds.barplot')
```

What does a good peptide look like?
A good number of consecutive non-zero values (what is good? Again shared with sibling peptides)
```{r}
#' Identify the max number of consecitive measurmeents
MaxRunOfConsecFractionMeasurments <- function(row, cutoff=0){
  # rle finds the unique values in a vector and the maximum run of consec vals
  binVec <- rle(row > cutoff)
  return(max(binVec$lengths[binVec$values == TRUE]))
} 


MaxRunOfConsecFractionMeasurments(c(0,1,0,2,2,3,3,4,4,0,0,1), cutoff=3)



correlateElutionProfiles <- function(secLong.dt, idcol='peptide', intsCol='intensity', excludedSet=c()){
  
  if (is.null(excludedSet)){ 
    allProteins <- unique(secLong.dt$protein)
  } else {
    rmSet <- sum(unique(secLong.dt$protein) %in% unique(excludedSet))
    message('Removing ', rmSet, ' proteins from calculations..')
    allProteins <- unique(secLong.dt$protein[!secLong.dt$protein %in% excludedSet])
  }
  .oneMatrix <- function(sub.dt){
    mat <- dcast(sub.dt, sub.dt[[idcol]]~fraction, value.var = intsCol) %>% 
      as.matrix(rownames = 1)
    mat[is.na(mat)] <- 0.0
    mat[order(rownames(mat)),]
  }
  
  .correlateSisterPeptides <- function(mat){
    cor.mat <- cor(t(mat), use='pairwise.complete.obs')
    r.dt <- setDT(reshape2::melt(cor.mat))
    setnames(r.dt, new=c('feature1', 'feature2', 'r'))
    return(r.dt)
  }
  
  multipep.dt <-  secLong.dt[npeptides > 2,]
  mats <- lapply(split(multipep.dt, list(multipep.dt$sample, multipep.dt$protein)), .oneMatrix)
  print(mats[1])
  
  cor.dt <- lapply(mats, .correlateSisterPeptides) %>% 
    rbindlist(idcol = 'subset')
  cor.dt[, c('sample', 'protein') := tstrsplit(subset, '[.]', keep=c(1,2))]
  
  return(cor.dt[feature1 != feature2, .(sample, protein, feature1, feature2, r)])
}
```



```{r}
help(cor)

test.dt
correlateProfiles(test.dt, idcol='peptide')

sisterPepetideCorrelations <- function(dt, excludedSet=c()){
  
  proteins <- dt[!grepl(';', protein), unique(protein)]
  
  if (!is.null(excludedSet)){ 
    rmSet <- sum(unique(proteins) %in% unique(excludedSet))
    message('Removing ', rmSet, ' proteins from calculations..')
    proteins <- proteins[!proteins %in% excludedSet]
  }
}
  
  # subset the data.table to each protein per 
  lapply(proteins, function(prot){
    
    subdt <- dt[protein ]
    
  }
}



test.dt <- sec.dt[treatment == 'sonic']

sec.dt[,.SD(), .SDcols='protein']



# use the .N > 20 filter to exclude teams with few observations
test.dt[ , if (.N > 1) .(peptideCor = cor(x=intensity[sample]), by = .(sample,protein, peptide)]



          ][ , hist(w_coef, 20L, las = 1L,
                    xlab = 'Fitted Coefficient on W',
                    ylab = 'Number of Teams', col = 'darkgreen',
                    main = 'Team-Level Distribution\nWin Coefficients on ERA')]
abline(v = overall_coef, lty = 2L, col = 'red')

sisterPepetideCorrelations(sec.dt, excludedSet = bigComplexes$gene) %>% length()
```


