---
title: "060325_coelutionDetection"
author: "Martin Gordon"
date: "2025-06-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Coelution detection of the SEC-MS datasets
After qucik overview of QC and tidy of missing fractions, run BP pipeline to detect co-eluting proteins in the data
Our focus here should be to detect co-eluting viral and host proteins, and then see if any of these proteins shift in one condition vs another

```{r cars}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)
library(stringr)
library(viridis)
library(pracma) # needed for the peak finding algorithm
library(magick)
#library(UniProt.ws) #cant load this packge; get naming conventions another way

# source Bens scripts so we can do a quick QC assessment of the input data before running
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/UniprotIDMapping.R") # map hu to mm 
source("~/Documents/utils/bp_utils/enrichmentTestFunctions.R")

# flag consec proteins
source("~/Documents/utils/bp_utils/SEC_MS_utils.localMods.R")
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")
```

Read in the fitlered normalized and interpolated protein set
```{r}
sec.long <- fread('~/Documents/projects/060325_APelin_POX_SECMS/060325_poxSECMS_firstLook_data/2025_06_03_sec.long.normalized.interpolated.filtered.csv.gz')
sec.long[, c('treatment', 'replicate') := tstrsplit(sample, '_', keep=c(1,2))]
```

```{r, peakDetection}
# to be safe drop the empty rows
normInt.mats <- scaledIntensityMatrices(sec.long[protein != 'MCHERRY'])

peakTables <- lapply(normInt.mats, goodPeaksTableFromIntensityMatrix, minCV = -Inf)
allPeaks <- rbindlist(peakTables, idcol = "sample")
fwrite(allPeaks, ScriptAndDatedFileName('poxInf.detectedPeaks.csv.gz'))

allPeaks <- fread('~/Documents/projects/043025_MGordon_SECMS_QCplotsWork/060325_coelutionDetection_data/2025_06_03_poxInf.detectedPeaks.csv.gz')
```

```{r}
```

