---
title: "060325_POX_SECMS_firstLook"
author: "Martin Gordon"
date: "2025-06-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 ## Preporcessing of the POX virus SEC-MS data
Inlcude QC assessment of the replicates
Loook at the viral protein profiles and see how they look
```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)
library(stringr)
library(viridis)
library(pracma) # needed for the peak finding algorithm
library(magick)
#library(UniProt.ws) #cant load this packge; get naming conventions another way

# source Bens scripts so we can do a quick QC assessment of the input data before running
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/UniprotIDMapping.R") # map hu to mm 
source("~/Documents/utils/bp_utils/enrichmentTestFunctions.R")

# flag consec proteins
source("~/Documents/utils/bp_utils/SEC_MS_utils.localMods.R")
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")
```
Read in hte pox SECMS data and visualize
So what are these samples?human cell-lin einfected pox samples?
```{r}
sec.long <- fread('./data/2024_10_25_SEC_MS_Input_Long.csv.gz')
sec.long$sample %>% unique()


# viral proteins; quick adn dirty way to find them... does thisd work?
viral.prots <- fread('./data/2024_10_25_SEC_MS_Input_Long.csv.gz') %>% 
  .[!grepl('HUMAN|mCherry', Protein.Names), unique(Protein.Names)]

```

```{r}
#fwrite(sec.long[,.(protein=Protein.Group, gene=Genes, sample, treatment, replicate, fraction, intensity)], ScriptAndDatedFileName('seclong.input.csv.gz'))
sec.long <- fread('~/Documents/projects/060325_APelin_POX_SECMS/060325_poxSECMS_firstLook_data/2025_06_03_seclong.input.csv.gz')
```
Run standard QC on the data to see how the samples compare
Maybe also a all-by-all correlation? Look at Bens work on this

*QC*
Looking at N proteins, seems there are a number of troublesome fractions. Can we try tidy these, impute missing values etc?
Looks like Vaccinia rep1 just failed, need to remove this sample from the analysis, it's not going to be of much use
```{r nproteins, fig.width=8, fig.height=6}
qc.dt <- qcSummaryTable(sec.long)
qcPlotProteinCount(qc.dt)
# number of protiens eluting looks good

BackupAsPDF(qcPlotProteinCount(qc.dt), 'Nproteins.linechart') +
  theme_classic()
```
```{r medpolish, fig.width=8, fig.height=6}
g <- qcPlotMedianInt(qc.dt) +
  theme_bw()
g
BackupAsPDF(g, 'medianPolish.linechart')
```
plot intensity heatmap with all proteins;

```{r, heatmaps, fig.width=8, fig.height=6}
intMats <- scaledIntensityMatrices(sec.long)
intensityHeatmaps(intMats)
BackupAsPDF(intensityHeatmaps(intMats), 'filtered.prots.heatmap')
```
 remove rep1 for VacV and then focus on tidying up the complexes
 
```{r}
sec.long <- sec.long[sample != 'VacV_R1']
#fwrite(sec.long, ScriptAndDatedFileName('seclong.input.dropVacV1.csv.gz'))
sec.long <- fread('~/Documents/projects/060325_APelin_POX_SECMS/060325_poxSECMS_firstLook_data/2025_06_03_seclong.input.dropVacV1.csv.gz')
qc.dt <- qcSummaryTable(sec.long)
```

We need to tidy the problematic fractions and impute missing fractions
```{r}
intMats <- scaledIntensityMatrices(sec.long)
intensityHeatmaps(intMats)
```
Missing value imputation; first identfiy the problematic fractions, zero out those measured values and impute the missing
then we can run the co-elution detection and the differential analysis

**Normalization**
Fit loess to each sample (fit an additive model;plot outlier samples

```{r}
allFits <- fitLocalCubics(qc.dt, window = 15, sampleTerm = "additive")
labelOutliers(qc.dt, allFits, threshold = 2) 

p <-plotNormAndOutlierFits(qc.dt , allFits)
p
BackupAsPDF(p, 'localRegression.window15.linechart')
```
Normalize the intensities 

```{r}
normalizeByResiduals(secLong.dt = sec.long, qc.dt)
#fwrite(sec.long, ScriptAndDatedFileName('sec.long.normalized.csv.gz'))
sec.long <- fread('~/Documents/projects/060325_APelin_POX_SECMS/060325_poxSECMS_firstLook_data/2025_06_03_sec.long.normalized.csv.gz')
```

Visualize output;
```{r}
new.qc.dt <- qcSummaryTable(sec.long)

g <- qcPlotMedianInt(new.qc.dt) +
  theme_bw()
g
BackupAsPDF(g, 'normalized.qc')
```

Clear issues with normalization of some outlier samples; interpolate missing fractions and proceed
interpolate the normalized values and proceed
```{r}
interp.dt <- handleMissingandOutlierFractions(sec.long, qc.dt, fractionOrder = seq(1,72,1))
#fwrite(interp.dt, ScriptAndDatedFileName('sec.long.normalized.interpolated.csv.gz'))
sec.long <- fread('~/Documents/projects/060325_APelin_POX_SECMS/060325_poxSECMS_firstLook_data/2025_06_03_sec.long.normalized.interpolated.csv.gz')
```

Results with the interpolated fractions look better. Lets run the differential analysis and proceed
```{r}
new.qc.dt <- qcSummaryTable(interp.dt)

g <- qcPlotMedianInt(new.qc.dt) +
  theme_bw()
g
BackupAsPDF(g, 'normalizedInterpolated.qc')
```

```{r, heatmaps, fig.width=10, fig.height=6}
intMats <- scaledIntensityMatrices(sec.long)
intensityHeatmaps(intMats)
BackupAsPDF(intensityHeatmaps(intMats), 'norm.interpolated.prots.heatmap')
```
Remove proteins with < 5 consecutvie measurements
Do this after interpolation on expanded but non-interpolated data to ensure the consecutvie meaurements are consecutive functions...
```{r}
# setNA ori.intensity to 0 for detection; these are the normalized values
sec.long[is.na(ori.intensity), ori.intensity := 0]
consec.dt  <- flagLowQualityFeatures(secLong.dt = sec.long, idcol = 'protein', intsCol = 'ori.intensity', ConsecFractionThreshold=5, detectionCutoff=0)

g <- ggplot(consec.dt[passThreshold == FALSE,.N, by=sample], aes(x=sample, y=N)) +
  geom_bar(stat='identity') +
  theme_bw()
g
BackupAsPDF(g, 'nProteinsFailConsecDetections.barplot')
```
Filter out the proteins without sufficient number of consecutive detections
```{r}
# keep passing proteins in sec.long 
fwrite(sec.long[consec.dt[passThreshold == TRUE, .(sample, protein=feature)],  on=.(protein, sample)], ScriptAndDatedFileName('sec.long.normalized.interpolated.filtered.csv.gz'))
```



**Functions**
detect features with <N consecutvie measurements
And fix the interpolation

```{r}
#' Flag low quality peptides based on missingness proportion, lack of consecutive detections <5 for now as required for peak detection
flagLowQualityFeatures <- function(secLong.dt, idcol='peptideSequence', intsCol='intensity', ConsecFractionThreshold=5, detectionCutoff=0){
  
  .oneMatrix <- function(sub.dt){
    mat <- dcast(sub.dt, sub.dt[[idcol]]~fraction, value.var = intsCol) %>% 
      as.matrix(rownames = 1)
    mat[is.na(mat)] <- 0.0
    mat[order(rownames(mat)),]
  }
  
    #' Identify the max number of consecitive measurmeents
  .MaxConsecFractionMeasurments <- function(row, cutoff=detectionCutoff){
    # get unique values in a vector and the maximum run of consec vals
    detectVec <- rle(row > cutoff)
    return(max(detectVec$lengths[detectVec$values == TRUE]))
  }
  
  # one matrix per sample 
  mats <- lapply(split(secLong.dt, list(secLong.dt$sample)), .oneMatrix)
  
  consec.dt <- lapply(mats, function(x) {
    apply(x, 1, function(pep){ .MaxConsecFractionMeasurments(row = pep, cutoff = 0) }) %>% 
    as.data.table(keep.rownames=T)
  }) %>% 
  rbindlist(idcol='sample')
  
  setnames(consec.dt, new=c('sample', 'feature', 'consecutiveDetections'))
  consec.dt[, passThreshold := ifelse(consecutiveDetections >= ConsecFractionThreshold, TRUE, FALSE)]
  return(consec.dt)
}

#interpolation function
#' functin that when given sec.dt and the qc output, fills in missing and outlier fractions with NA
#' datatable join to replace values in problematic fractions with NA
handleMissingandOutlierFractions <- function(sec.dt, qc.dt, fractionOrder){
  
  # where the sample fraction is flagged an outlier, convert all the intensities to NA
  sec.dt[, ori.intensity := intensity]
  sec.dt[qc.dt[isOutlier == TRUE], on=.(sample, fraction), intensity := NA]
  
  # add NA for missing fractions
  .addMissingFractions <- function(subDT){
    subMat <- dcast(subDT, protein~fraction, value.var='intensity') %>% 
      as.matrix(rownames='protein')
    
    if (!all(colnames(subMat) == fractionOrder)){
      message('Some fractions are missing and/or not in correct order\nFilling in missing fractions with NA...')
      message('Missing fractions:\n', setdiff(fractionOrder, colnames(subMat)))
      subMat <- subMat[, match(fractionOrder, colnames(subMat)), drop=FALSE] #missing fractions assigned an NA col. drop=FALSE to avoid collapsing to vector
      colnames(subMat) <- fractionOrder
    }
    return(subMat)
  }
  
  .interpolateOutlierFractions <- function(subMat, sampleOI){
    # interpolate all values rowwise
    interpMat <- apply(subMat, 1, function(x) zoo::na.approx(x, na.rm=F, maxgap=1)) %>% 
      t()
    # fractions to update
    fractionsOI <- c(qc.dt[sample == sampleOI & isOutlier == TRUE, unique(fraction)],
                     setdiff(fractionOrder, sec.dt[sample == sampleOI, unique(fraction)])
                     )
    
    print(fractionsOI)
    colsToupdate <- colnames(subMat) %in% fractionsOI
    
    # now apply the values from the interpolated matrix to the original
    subMat[, colsToupdate] <- interpMat[, colsToupdate]
    subdt <- reshape2::melt(subMat)
    setnames(subdt, new=c('protein', 'fraction', 'intensity'))
    return(subdt)
  }

  # handling missing/outlir fractions
  sec.list <- lapply(split(sec.dt, sec.dt$sample), .addMissingFractions)
  
  # interpolate missing values
  message('interpolating missing values...')
  interp.list <- lapply(names(sec.list), function(n){.interpolateOutlierFractions(subMat = sec.list[[n]], sampleOI = n)}) 
  names(interp.list) <- names(sec.list)
  interp.dt <- rbindlist(interp.list, idcol='sample')
  
  # now combine the dt with the treatment, replicate and ori intensity and return
  print(nrow(interp.dt))
  interp.dt <- merge(x=interp.dt, y=sec.dt[,.(protein, sample,fraction, ori.intensity)], by=c('protein', 'sample', 'fraction'), all.x=T, all.y=F) 
  print(nrow(interp.dt))

  return(interp.dt)
}
```
