---
title: "060925_PPIenrichment"
author: "Martin Gordon"
date: "2025-06-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Have a list of coeluting PPI from the SEC-MS experiments. Need to get a handle on how info-rich this data is. 
Two things to do first i) Check if the set is enrihced for AP-MS PPI and ii) check if the list is enriched for higher AF3/RosettaFold scoring 

Look at differential vs Mock; is there enrichment for the Prey genes in the PPI set

```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)
library(stringr)
library(viridis)
library(pracma) # needed for the peak finding algorithm
library(magick)
library(eulerr)
library(pROC)
#library(UniProt.ws) #cant load this packge; get naming conventions another way

# source Bens scripts so we can do a quick QC assessment of the input data before running
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/UniprotIDMapping.R") # map hu to mm 
source("~/Documents/utils/bp_utils/enrichmentTestFunctions.R")

# flag consec proteins
source("~/Documents/utils/bp_utils/SEC_MS_utils.localMods.R")
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")
```

First to look for enrichment of RosettaFold 

```{r}
#dca.scores  <- fread('./data/DCA_scores.zip') Dont read DCA scores tet
af.scores <- fread('./data/AF_scores.zip')
rf2.scores <- fread('./data/RF2-PPI_scores.zip')


#46 million PPI
dim(rf2.scores)

# 3 million PPI
dim(af.scores)


af.scores[, ppi_uniprot_ordered := {
  strsplit(V1, '_')[[1]] %>% 
    sort() %>% 
    paste(., collapse='_')
}, by=.I]

rf2.scores[, ppi_uniprot_ordered := {
  strsplit(V1, '_')[[1]] %>% 
    sort() %>% 
    paste(., collapse='_')
}, by=.I]


fwrite(af.scores, ScriptAndDatedFileName('af.scores.csv.gz'))
fwrite(rf2.scores, ScriptAndDatedFileName('rf2.ppi.scores.csv.gz'))

rm(af.scores)
rm(rf2.scores)

af.scores <- fread('~/Documents/projects/060325_APelin_POX_SECMS/060925_PPIenrichment_data/2025_06_09_af.scores.csv.gz')
rf.scores <- fread('~/Documents/projects/060325_APelin_POX_SECMS/060925_PPIenrichment_data/2025_06_09_rf2.ppi.scores.csv.gz')
```
Cosine and peak similarity scores
```{r}
allScores <- fread('~/Documents/projects/060325_APelin_POX_SECMS/060325_coelutionDetection_data/2025_06_05_pox.combinedCosinePeakSimilarity.scores.csv.gz')
```
A good RF2/AF score is >5; how do the cosine similarity scores compare between the two?
Scores dont look great, but just so much noise is the issue...

```{r}
allScores[af.scores, af.scores := i.V2, on=.(ppi_uniprot_ordered)]
allScores[rf.scores, rf.ppi.scores := i.V2, on=.(ppi_uniprot_ordered)]

allScores[, AF_pair := 'no']
allScores[af.scores >= 0.5, AF_pair := 'yes']
allScores[bestLLRatio > 3]


allScores[, RFppi_pair := 'no']
allScores[rf.ppi.scores >= 0.5, RFppi_pair := 'yes']
allScores[bestLLRatio > 3]


ggplot(allScores, aes(x=peakLLRatio, fill=AF_pair)) +
  geom_density() +
  facet_wrap(~AF_pair) +
  theme_bw()
```

plot the ecdf of the two distributions and maybe a statistical test to see if they are different
This is the combined coSine and peakLLR... seperation v poor
```{r}
g <- ggplot(allScores, aes(x=peakLLRatio, color=AF_pair)) +
  stat_ecdf(geom='line') +
  labs(title='ECDFs of peak coelution scores', subtitle='AF iPTM >= .5') +
  theme_bw()

BackupAsPDF(g, 'peakCor.AF.iPTM.ecdf')
```
How many are sig in both R2 anf AF2

```{r}
# much more of the RFppi pass the threshold
allScores[,.N, by=RFppi_pair]
allScores[,.N, by=AF_pair]
```


Generate a mosaic plot; are the coelutions enriched for 
```{r}
g <- ggplot(allScores[peakLLRatio >= 0,], aes(x=peakLLRatio, color=AF_pair)) +
  stat_ecdf(geom='line') +
  labs(title='ECDFs of peak coelution scores', subtitle='AF iPTM >= .5') +
  theme_bw()
g


g <- ggplot(allScores[peakLLRatio > 0,], aes(x=peakLLRatio, color=AF_pair)) +
  stat_ecdf(geom='line') +
  labs(title='ECDFs of peak coelution scores', subtitle='AF iPTM >= .5') +
  theme_bw()

g
BackupAsPDF(g, 'peakCor.LLRgreaterZero.AF.iPTM.ecdf')
```
Look at the RF scores also and see if we get good seperation
Not as good as AF, which seems a higher quality set, but its not nothing!
```{r}
g <- ggplot(allScores[peakLLRatio>0,], aes(x=peakLLRatio, color=RFppi_pair)) +
  stat_ecdf(geom='line') +
  labs(title='ECDFs of peak coelution scores', subtitle='AF iPTM >= .5') +
  theme_bw()

g
BackupAsPDF(g, 'peakCor.LLRgreaterZero.RF2.iPTM.ecdf')
```
Cosine is not as good a seperator
```{r}
g <- ggplot(allScores[cosLLRatio > 0,], aes(x=cosLLRatio, color=RFppi_pair)) +
  stat_ecdf(geom='line') +
  labs(title='ECDFs of peak coelution scores', subtitle='AF iPTM >= .5') +
  theme_bw()

g
BackupAsPDF(g, 'cosCor.LLRgreaterZero.RF2.iPTM.ecdf')
```

So there is clearly some seperation between the groups based on peak based similarity
Huge number o fPPI in the dataset, to evaluate just subset to the string set and see how the scores compare

```{r}
rm(allScores)

allScores <- fread('~/Documents/projects/060325_APelin_POX_SECMS/060325_coelutionDetection_data/2025_06_05_pox.combinedCosinePeakSimilarity.scores.csv.gz')
allScores <- allScores[gs %in% c('interactor', 'decoy')]
```

merge this with the AF and Rosettafold scores
```{r}
allScores <- merge(x=allScores, y=rf.scores[, .(ppi_uniprot_ordered, rf2_ppi=V2)], by='ppi_uniprot_ordered', all.x = T)
allScores <- merge(x=allScores, y=af.scores[, .(ppi_uniprot_ordered, af2_ppi=V2)], by='ppi_uniprot_ordered', all.x = T)
```

scatterplot of peak cor scores and best PPI score. color by i) sig in AF, ii) sig in RF-2 and iii) sig in both
```{r}
allScores[, AF_pair := 'no']
allScores[af2_ppi >= 0.5, AF_pair := 'yes']
allScores[, RFppi_pair := 'no']
allScores[rf2_ppi >= 0.5, RFppi_pair := 'yes']

allScores[, grp := ifelse(AF_pair == 'yes' & RFppi_pair == 'yes', 'RF2-PPI & AF2 hit',
                          ifelse(AF_pair == 'yes' & RFppi_pair != 'yes', 'AF2 hit',
                                 ifelse(AF_pair != 'yes' & RFppi_pair == 'yes', 'RF2-PPI hit',
                                        'no hit')))]


```
This is problematic.. looks like the scores for decoy seem to be higher!!
```{r}
col.pal <- c('gold', 'grey')
names(col.pal) <- c('interactor', 'decoy')

g <- ggplot(allScores, aes(x=rf2_ppi, fill=gs)) +
  geom_density(alpha=0.5) +
  geom_vline(xintercept=0.5) +
  labs(title='RF2-PPI iPTM scores') +
  scale_fill_manual(values=col.pal) + 
  theme_bw() 

BackupAsPDF(g, 'RF2.iptm.distributions.density')

g <- ggplot(allScores, aes(x=af2_ppi, fill=gs)) +
  geom_density(alpha=0.5) +
  geom_vline(xintercept=0.5) +
  labs(title='AF2 iPTM scores') +
  scale_fill_manual(values=col.pal) + 
  theme_bw() 
g
BackupAsPDF(g, 'AF2.iptm.distributions.density')


g <- ggplot(allScores[rf2_ppi > .5], aes(x=rf2_ppi, fill=gs)) +
  geom_density(alpha=0.5) +
  geom_vline(xintercept=0.5) +
  labs(title='AF2 iPTM scores') +
  scale_fill_manual(values=col.pal) + 
  theme_bw() 
g
```
Clearly some enrichment here for high scores... if we look at differential hits,are they enrihced in the AP-MS and/or structural prediction scores?

```{r}
# take the best score and plot that
allScores[, best_iptm := ifelse(af2_ppi > rf2_ppi, af2_ppi, rf2_ppi)]

ggplot(allScores, aes(x=peakLLRatio)) +
  geom_density()


g <- ggplot(allScores[peakLLRatio >0,], aes(x=af2_ppi, y=peakLLRatio, color=gs)) +
  geom_point(alpha=0.2) +
  geom_density2d(color='black') +
  scale_color_manual(values=col.pal) +
  theme_bw() +
  facet_wrap(~gs)

BackupAsPDF(g, 'STRINGgs.LLRvsAF2ppi.dotplot')


g <- ggplot(allScores[peakLLRatio >0,], aes(x=rf2_ppi, y=peakLLRatio, color=gs)) +
  geom_point(alpha=0.2) +
  geom_density2d(color='black') +
  scale_color_manual(values=col.pal) +
  theme_bw() +
  facet_wrap(~gs)
g
BackupAsPDF(g, 'STRINGgs.LLRvsRF2ppi.dotplot')
```
gsset iPTM; are they useful to discriminate?
```{r}
g <- ggplot(allScores[peakLLRatio>0,], aes(x=af2_ppi, color=gs)) +
  stat_ecdf(geom='line') +
  geom_vline(xintercept = 0.5) +
  scale_color_manual(values=col.pal) +
  labs(title='ECDFs of peak coelution scores', subtitle='AF iPTM >= .5') +
  theme_bw()

g
BackupAsPDF(g, 'afiPTM.LLRgreaterZero.gsString.ecdf')
```
Try the same with the RF2 iPTM

```{r}
g <- ggplot(allScores[peakLLRatio>0,], aes(x=rf2_ppi, color=gs)) +
  stat_ecdf(geom='line') +
  geom_vline(xintercept = 0.5) +
  scale_color_manual(values=col.pal) +
  labs(title='ECDFs of peak coelution scores', subtitle='RF2-PPI iPTM >= .5') +
  theme_bw()

g
BackupAsPDF(g, 'rf2iPTM.LLRgreaterZero.gsString.ecdf')
```
Mosaic plot; are the LLRatio > 3 enriched for AF/RF high scoring PPI?
```{r}
allScores
allScores[peakLLRatio >=3, LLRsig := 'yes']
allScores[peakLLRatio >=3,.N, by=AF_pair]

contTab <- xtabs(~LLRsig + AF_pair, data = allScores)

pdf('/Users/martingordon/Documents/projects/060325_APelin_POX_SECMS/gs.mosaic.afIPTM.mosaic.pdf')
g <- mosaicplot(contTab,
           sub = NULL, 
            main = 'STRING geneset',
           xlab = 'LLRratio > 3', ylab = 'AF iPTM > 0.5',
           col=c('grey', 'gold'),
           border=T)
g 
dev.off()

contTab <- xtabs(~LLRsig + RFppi_pair, data = allScores[peakLLRatio > 0])

pdf('/Users/martingordon/Documents/projects/060325_APelin_POX_SECMS/gs.mosaic.rf2ppiIPTM.mosaic.pdf')
g <- mosaicplot(contTab,
           sub = NULL, 
            main = 'STRING geneset',
           xlab = 'LLRratio > 3', ylab = 'RF2-PPI iPTM > 0.5',
           col=c('grey', 'gold'),
           border=T)
g 
dev.off()

# incredibly significant for the Genseset
fisher.test(contTab, alternative = 'greater')

# both sets are hugely significant when we look at STRING HC
contTab <- xtabs(~LLRsig + AF_pair, data = allScores)
```
This looks good! Read in all the scores and repeat and check the significance
```{r}
allScores <- fread('~/Documents/projects/060325_APelin_POX_SECMS/060325_coelutionDetection_data/2025_06_05_pox.combinedCosinePeakSimilarity.scores.csv.gz')
allScores <- merge(x=allScores, y=rf.scores[, .(ppi_uniprot_ordered, rf2_ppi=V2)], by='ppi_uniprot_ordered', all.x = T)
allScores <- merge(x=allScores, y=af.scores[, .(ppi_uniprot_ordered, af2_ppi=V2)], by='ppi_uniprot_ordered', all.x = T)


allScores[, AF_pair := 'no']
allScores[af2_ppi >= 0.5, AF_pair := 'yes']
allScores[, RFppi_pair := 'no']
allScores[rf2_ppi >= 0.5, RFppi_pair := 'yes']
```

```{r}
allScores[, LLRsig := 'no']
allScores[peakLLRatio >=2, LLRsig := 'yes']
contTab <- xtabs(~LLRsig + AF_pair, data = allScores)


dt <- data.table(contTab)
dt[, total := sum(N), by=LLRsig]
dt[, labOffset := 1- N/total]
dt[, AF_pair := factor(AF_pair, levels=c('yes', 'no'))]
dt[, LLRsig := factor(LLRsig, levels=c('no', 'yes'))]

g <- ggplot(dt, aes(x=LLRsig, y=N, fill=AF_pair, label=as.character(N))) +
  geom_bar(stat='identity', position='fill') +
  geom_col(color='black', position='fill') +
  labs(x='LLRatio > 2') +
 # facet_grid(~model, scales = "free_x", space = "free_x") +
  geom_label(data=dt, aes(y=labOffset),  color='black', fill='white', size=4) +
  scale_fill_brewer(type='qual', direction=-1, palette=6) +
  theme_bw() +
  guides(fill=guide_legend(title='AF iPTM > .5'))
g

BackupAsPDF(g, 'af2iptm.llr2threshold.mosaic')

allScores[, LLRsig := 'no']
allScores[peakLLRatio >=3, LLRsig := 'yes']

contTab <- xtabs(~LLRsig + AF_pair, data = allScores)

dt <- data.table(contTab)
dt[, total := sum(N), by=LLRsig]
dt[, labOffset := 1- N/total]
dt[, AF_pair := factor(AF_pair, levels=c('yes', 'no'))]
dt[, LLRsig := factor(LLRsig, levels=c('no', 'yes'))]

g <- ggplot(dt, aes(x=LLRsig, y=N, fill=AF_pair, label=as.character(N))) +
  geom_bar(stat='identity', position='fill') +
  geom_col(color='black', position='fill') +
  labs(x='LLRatio > 2') +
 # facet_grid(~model, scales = "free_x", space = "free_x") +
  geom_label(data=dt, aes(y=labOffset),  color='black', fill='white', size=4) +
  scale_fill_brewer(type='qual', direction=-1, palette=6) +
  theme_bw() +
  guides(fill=guide_legend(title='AF iPTM > .5'))
g
BackupAsPDF(g, 'af2iptm.llr2threshold.mosaic')

```
Look at rosettafold2 scores 

```{r}
allScores[, LLRsig := 'no']
allScores[peakLLRatio >=3, LLRsig := 'yes']

contTab <- xtabs(~LLRsig + RFppi_pair, data = allScores)

dt <- data.table(contTab)
dt[, total := sum(N), by=LLRsig]
dt[, labOffset := 1- N/total]
dt[, RFppi_pair := factor(RFppi_pair, levels=c('yes', 'no'))]
dt[, LLRsig := factor(LLRsig, levels=c('no', 'yes'))]

g <- ggplot(dt, aes(x=LLRsig, y=N, fill=RFppi_pair, label=as.character(N))) +
  geom_bar(stat='identity', position='fill') +
  geom_col(color='black', position='fill') +
  labs(x='LLRatio > 3') +
 # facet_grid(~model, scales = "free_x", space = "free_x") +
  geom_label(data=dt, aes(y=labOffset),  color='black', fill='white', size=4) +
  scale_fill_brewer(type='qual', direction=-1, palette=6) +
  theme_bw() +
  guides(fill=guide_legend(title='RF2-PPI iPTM > .5'))
g
BackupAsPDF(g, 'rf2iptm.llr3threshold.mosaic')


allScores[, LLRsig := 'no']
allScores[peakLLRatio >=2, LLRsig := 'yes']

contTab <- xtabs(~LLRsig + RFppi_pair, data = allScores)

dt <- data.table(contTab)
dt[, total := sum(N), by=LLRsig]
dt[, labOffset := 1- N/total]
dt[, RFppi_pair := factor(RFppi_pair, levels=c('yes', 'no'))]
dt[, LLRsig := factor(LLRsig, levels=c('no', 'yes'))]

g <- ggplot(dt, aes(x=LLRsig, y=N, fill=RFppi_pair, label=as.character(N))) +
  geom_bar(stat='identity', position='fill') +
  geom_col(color='black', position='fill') +
  labs(x='LLRatio > 2') +
 # facet_grid(~model, scales = "free_x", space = "free_x") +
  geom_label(data=dt, aes(y=labOffset),  color='black', fill='white', size=4) +
  scale_fill_brewer(type='qual', direction=-1, palette=6) +
  theme_bw() +
  guides(fill=guide_legend(title='RF2-PPI iPTM > .5'))
g
BackupAsPDF(g, 'rf2iptm.llr2threshold.mosaic')
```
Now check the DE results and see if these are eenriched for the AP-MS hits

```{r}
peakClusters <- fread('~/Documents/projects/060325_APelin_POX_SECMS/060325_differentialAnalysis_data/2025_06_05_proteinPeaks.anova.pval.anno.csv.gz')
viral.apms <- fread('~/Documents/projects/060325_APelin_POX_SECMS/060325_coelutionDetection_data/2025_06_05_viralAPMS.sigHits.anno.csv.gz')
```
What is interesting is that our co-elution is clearly enriched for sig hits, but differential expression set is not.... process of my analysis? averaging over batches? Worht considering how we handle these different batches

```{r}
peakClusters[, inPreySet := 'no']
peakClusters[protein %in% preyPullDowns, inPreySet:= 'yes']

peakClusters[,.N, by=inPreySet]

peakClusters[, sigSet := 'no']
peakClusters[treatmentDiff.p < 0.05, sigSet := 'yes']

contTab <- xtabs(~sigSet+inPreySet, data=peakClusters)

dt <- data.table(contTab)
dt[, total := sum(N), by=inPreySet]
dt[, labOffset := 1- N/total]
dt[, sigSet := factor(sigSet, levels=c('yes', 'no'))]
dt[, inPreySet := factor(inPreySet, levels=c('no', 'yes'))]


g <- ggplot(dt, aes(x=inPreySet, y=N, fill=sigSet, label=as.character(N))) +
  geom_bar(stat='identity', position='fill') +
  geom_col(color='black', position='fill') +
  labs(x='') +
 # facet_grid(~model, scales = "free_x", space = "free_x") +
  geom_label(data=dt, aes(y=labOffset),  color='black', fill='white', size=4) +
  scale_fill_brewer(type='qual', direction=-1, palette=6) +
  theme_bw() +
  guides(fill=guide_legend(title='gene in HC AP-MS set'))
g
BackupAsPDF(g, 'rf2iptm.llr2threshold.mosaic')
```
Clearly no enrichment for the AP-MS set in the DE results

```{r}
sec.long <- fread('~/Documents/projects/060325_APelin_POX_SECMS/060325_poxSECMS_firstLook_data/2025_06_03_sec.long.normalized.interpolated.filtered.csv.gz')
sec.long[, c('treatment', 'replicate') := tstrsplit(sample, '_', keep=c(1,2))]


# summarize the intensity
sec.long[!is.na(intensity), range(intensity)]
sec.long[, sumIntensity := log2(sum(intensity, na.rm=T)), by=.(protein,sample)]
```

divide the set into prey and non-prey proteins; do the intensity values in the two groups look different?
```{r}
sec.long[, isPrey := 'no']
sec.long[protein %in% preyPullDowns, isPrey := 'yes']

ggplot(sec.long, aes(x=sumIntensity, fill=isPrey)) +
  geom_density(alpha=0.5) +
  facet_wrap(~sample)
```

```{r}

```

