---
title: "020325_msstatsDEanalysis"
author: "Martin Gordon"
date: "2025-02-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}

library(data.table)
library(ggplot2)
library(magrittr)
library(ggrepel)
library(stringr)
library(ComplexHeatmap)
library(circlize)
library(MSstats)
library(viridis)
library(RColorBrewer)
library(hrbrthemes) # visually nice set of themes
library(showtext)
library(readxl)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")
source ("../../utils/bp_utils/NMF_Helpers.R")

source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("../../utils/mg_utils/r_utils/msstats_helperFunctions.R")


# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}

customTheme <- theme_ipsum_rc() +
  theme(panel.border = element_rect(color = "lightgrey", 
                                    fill = NA, 
                                    size = 1),
        axis.text.x = element_text(angle=90)
        )
#set one
col.pal <- getQualitativePalette(n=13)
#col.pal <- randomcoloR::distinctColorPalette(k=13)

# fonts needed for hrbrthemes
library(extrafont)
font_import()
loadfonts()

# directly from google font
sysfonts::font_add_google("Roboto Condensed")
showtext_auto()
```
```{r}
spec <- fread('./data/CR2/20250115_CR2_SMO_APEX_Report.tsv')
keys <- fread('./data/CR2//CR2_SMO_APEX_01152025_ConditionSetup.tsv')
```
Look at intensity distributions 

```{r}
ggplot(spec, aes(x=log2(Intensity))) +
  geom_histogram()
# cut off the poor left hand side of the plot
ggplot(spec[log2(Intensity) > 5,], aes(x=log2(Intensity))) +
  geom_histogram()

g <- ggplot(spec[log2(Intensity) > 4,], aes(x=log2(Intensity))) +
  geom_histogram() + 
  customTheme

BackupAsPDF(g, 'intensityDistribution.histogram')

spec <- spec[log2(Intensity) > 4,]
```
process the 3 datqsets seperately and look at boxplots of the distributions before running the differential analysis comparisons
```{r}
spec[, probe := str_extract(Condition, 'GPR135|SMO|PTH2R')]

spec.list <- split(spec, spec$probe)

dp.out.list <- lapply(names(spec.list), function(x){

  spec.list[[x]][, IsotopeLabelType := 'L']
  
  dp.out <- MSstats::dataProcess(spec.list[[x]], 
                                 MBimpute =  FALSE, 
                                 normalization = 'equalizeMedians',
                                 summaryMethod = "TMP",
                                 featureSubset = 'highQuality',
                                 remove_uninformative_feature_outlier=T)
  
  saveRDS(dp.out, ScriptAndDatedFileName(paste0(x, 'dp.out.rds')))
})
```
read in the dataprocess objects

```{r}
dp.list <- list(SMO = readRDS('./020325_msstatsDEanalysis_data/2025_02_03_SMOdp.out.rds'),
                PTH2R = readRDS('./020325_msstatsDEanalysis_data/2025_02_03_PTH2Rdp.out.rds'),
                GPR135 = readRDS('./020325_msstatsDEanalysis_data/2025_02_03_GPR135dp.out.rds'))
```

fix SUBJECT field for group comparison

```{r}
lapply(dp.list, function(x){
  
  p.quant <- setDT(x$ProteinLevelData)
  p.quant[, SUBJECT := paste0(GROUP,'.',SUBJECT)]
  p.quant[, GROUP := factor(GROUP)]
  
  x$ProteinLevelData <- p.quant
  return(x)
})

```


```{r}
lapply(dp.list, function(x){
  
  p.quant <- x$ProteinLevelData
  probe <- unique(str_extract(p.quant$GROUP, 'SMO|GPR135|PTH2R'))
  
  g <- ggplot(p.quant, aes(x=paste0(GROUP,SUBJECT), y=LogIntensities, fill=GROUP)) +
    geom_boxplot() +
    ggtitle(probe) +
    scale_fill_ipsum() +
    customTheme
  g
  BackupAsPDF(g, paste0(probe,'.protein.ints.boxplots'))
})
```
PW comparisons between groups

```{r}
contrasts.list <- list('+cilia_SMO_ag vs +cilia_SMO' = data.table(V1='SMO_+Cilia_+Agonist', V2='SMO_+Cilia_-Agonist'),
                       '+cilia_SMO vs -cilia_SMO' = data.table(V1='SMO_+Cilia_-Agonist', V2='SMO_-Cilia_-Agonist'),
                       '+cilia_PTH2R_ag vs +cilia_PTH2R' = data.table(V1='PTH2R_+Cilia_+Agonist', V2='PTH2R_+Cilia_-Agonist'),
                       '+cilia_PTH2R vs -cilia_PTH2R' = data.table(V1='PTH2R_+Cilia_-Agonist', V2='PTH2R_-Cilia_-Agonist'),
                       '+cilia_GPR135 vs -cilia_GPR135' = data.table(V1='GPR135_+Cilia', V2='GPR135_-Cilia'))



rownames(contrasts.mat)
contrasts.mat <- MSstats::MSstatsContrastMatrix(contrasts.list, 
                               conditions = unique(p.quant$GROUP), 
                               labels = names(contrasts.list))

```

Run lm on each of the groups, just want a standard lm for group comparison
```{r}
contrasts.mat[grepl('PTH2R', rownames(contrasts.mat)), grepl('PTH2R', colnames(contrasts.mat))]
contrasts.mat[grepl('SMO', rownames(contrasts.mat)), grepl('SMO', colnames(contrasts.mat))]
contrasts.mat[grepl('GPR135', rownames(contrasts.mat)), grepl('GPR135', colnames(contrasts.mat))]

# rbind the rows from the different groups
mss <- groupComparison(contrast.matrix=contrasts.mat[grepl('PTH2R', rownames(contrasts.mat)), grepl('PTH2R', colnames(contrasts.mat))],
                       verbose=F,
                       data=dp.list[['PTH2R']])
pth2r.mss <- setDT(mss$ComparisonResult)

mss <- groupComparison(contrast.matrix=contrasts.mat[grepl('SMO', rownames(contrasts.mat)), grepl('SMO', colnames(contrasts.mat))],
                       verbose=F,
                       data=dp.list[['SMO']])
smo.mss <- setDT(mss$ComparisonResult)


gpcr.mat <- matrix(contrasts.mat[grepl('GPR135', rownames(contrasts.mat)), grepl('GPR135', colnames(contrasts.mat))], nrow=1)
rownames(gpcr.mat) <- '+cilia_GPR135 vs -cilia_GPR135'
colnames(gpcr.mat) <- c('GPR135_+Cilia','GPR135_-Cilia')

# rbind the rows from the different groups
mss <- groupComparison(contrast.matrix=gpcr.mat,
                       verbose=F,
                       data=dp.list[['GPR135']])
gpr.mss <- setDT(mss$ComparisonResult)

```
adjust the pvalues globally across all levels of the contrasts; also try the 

```{r}
mss.dt <- rbind(gpr.mss, smo.mss, pth2r.mss)
# recal padj to account for the global MTC
mss.dt[, padj := p.adjust(pvalue, method='BH'), by=(Label)]
mss.dt[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]
mss.dt[, c('numerator','denominator') := tstrsplit(Label, ' vs ', keep=c(1,2))]
mss.dt[, sig := 'not']
mss.dt[abs(log2FC) > 0.58 & pvalue < 0.005, sig := ifelse(log2FC > 0, 'up', 'down')]


# save the results to file 
fwrite(mss.dt, ScriptAndDatedFileName('pwcomparisons.probelvlgrouping.csv.gz'))
```

set significance at LFC 50% and pval < 0.005
```{r}
mss.dt <- fread('~/Documents/projects/012425_AMarley_CiliaAPEX_Rerun/020325_msstatsDEanalysis_data/2025_02_03_pwcomparisons.probelvlgrouping.csv.gz')
```
Ok, we can also run the analysis on the combined dataset and see if that recovers a greater number of hits
Whatever works best, we will then try to use fishers method to combine p-values from the two seperate experiments

First lets test with first round of normalization

```{r}
dp.out <- readRDS('./012425_ciliaAPEX_firstLook_data/2025_01_24_mss.dataProc.rds')
p.quant <- fread('012425_ciliaAPEX_firstLook_data/2025_01_24_ProteinlvlQuant.2ndRoundNorm.csv')


p.quant[, GROUP := factor(GROUP)]
p.quant[, SUBJECT := paste0(GROUP,'.', SUBJECT)] # to enforce lm for group comparisons

#p.quant[, LogIntensities := tmpAdjLogIntensities]

dp.out$ProteinLevelData  <- p.quant
```


```{r}
mss <- groupComparison(contrast.matrix=contrasts.mat,
                       verbose=F,
                       data=dp.out)

mss.out <- setDT(mss$ComparisonResult)

mss.out[, padj := p.adjust(pvalue, method='BH'), by=(Label)]
mss.out[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]
mss.out[, c('numerator','denominator') := tstrsplit(Label, ' vs ', keep=c(1,2))]
mss.out[, sig := 'not']
mss.out[abs(log2FC) > 0.58 & padj < 0.05, sig := ifelse(log2FC > 0, 'up', 'down')]

# this is with one round normalization
firstPass.dt <- mss.out[sig != 'not', .N, by=Label]
#tmpPass.dt <- mss.out[sig != 'not', .N, by=Label]
#lmPass.dt <- mss.out[sig != 'not', .N, by=Label]

firstPass.dt
tmpPass.dt
```
Looking at this output, I think we want to run another round of normalization, but lets try the global median approach

```{r}
p.mat <- dcast(p.quant, Protein~paste0(GROUP,'.',SUBJECT), value.var='LogIntensities') %>% 
  as.matrix(rownames='Protein')



# like 1147 proteins shared
sharedProts <-  p.mat[complete.cases(p.mat),] %>% 
  rownames()


p.quant %>% colnames()
ggplot(p.quant[Protein %in% sharedProts,], aes(x=interaction(GROUP,SUBJECT), y=tmpAdjLogIntensities, fill=GROUP)) +
  geom_boxplot() +
  customTheme

```

*not used*
Just proceed with the tmp normalization data

```{r}
dp.out <- readRDS('./012425_ciliaAPEX_firstLook_data/2025_01_24_mss.dataProc.rds')
p.quant <- fread('012425_ciliaAPEX_firstLook_data/2025_01_24_ProteinlvlQuant.2ndRoundNorm.csv')

p.quant[, GROUP := factor(GROUP)]
p.quant[, SUBJECT := paste0(GROUP,'.', SUBJECT)] # to enforce lm for group comparisons

p.quant[, LogIntensities := tmpAdjLogIntensities]

dp.out$ProteinLevelData  <- p.quant


# try with global median norm
p.quant <- fread('020325_msstatsDEanalysis_data/2025_02_04_proteinlevelquant.renorm.csv')

p.quant[, GROUP := factor(GROUP)]
p.quant[, LogIntensities := medianNormInts]

dp.out$ProteinLevelData  <- p.quant
```

```{r}
mss <- groupComparison(contrast.matrix=contrasts.mat,
                       verbose=F,
                       data=dp.out)

mss.out <- setDT(mss$ComparisonResult)

mss.out[, padj := p.adjust(pvalue, method='BH'), by=(Label)]
mss.out[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]
mss.out[, c('numerator','denominator') := tstrsplit(Label, ' vs ', keep=c(1,2))]
mss.out[, sig := 'not']
mss.out[abs(log2FC) > 0.58 & padj < 0.05, sig := ifelse(log2FC > 0, 'up', 'down')]
mss.out[, probe := str_extract(Label, 'SMO|PTH2R|GPR135')]
mss.out[, label := sig]
mss.out[Protein %in% c(cilia.AM.prots, cilia.Consensus.prots), label := 'cilia assoc.']


mss.out[sig != 'not',.N, by=Label]
#fwrite(mss.out, ScriptAndDatedFileName('mss.pwcomparisons.allGroups.TMPrenormData.csv.gz'))

mss.out <- fread('~/Documents/projects/012425_AMarley_CiliaAPEX_Rerun/020325_msstatsDEanalysis_data/2025_02_04_mss.pwcomparisons.allGroups.csv.gz')
#mss.out <- fread('~/Documents/projects/012425_AMarley_CiliaAPEX_Rerun/020325_msstatsDEanalysis_data/2025_02_04_mss.pwcomparisons.allGroups.TMPrenormData.csv.gz')
```

barplot of number of significant hits in each group
produce all the standard plots (volcanoplot, de heatmap, enrichment etc) and tables and return results by end of day
Try fishers method to combine p-values from the two datasets

Volcanoplots of the different tests
```{r}
mss.out[, label := factor(label,levels=c('not', 'down', 'up', 'cilia assoc.'))]
mss.out <- mss.out[order(label)]

g <- ggplot(mss.out, aes(y=-log10(padj), x=log2FC, color=label, label=gene)) +
  geom_point(size=1.5) +
  geom_text_repel(data=mss.out[label != 'not',]) +
  geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
  geom_hline(yintercept = -log10(0.05), linetype=2, alpha=0.4) +
  labs(x=expression(log[2]~'FoldChange'), y=expression(-log[10]~'adj.pvalue')) +
  scale_color_manual(values=c('up'=col.pal[2], 'down'=col.pal[1], 'cilia assoc.'='green4', 'not'='grey')) +
  facet_wrap(~Label, scales='free', ncol=2) +
  customTheme +
  theme(strip.text.x = element_text(face='bold', size=12))

g
#BackupAsPDF(g, 'allContrasts.ajdpval.volcanoplots', dimensions = c(16,14))
```
pull out the agonist comparisons and plot them seperately

```{r}
contrast.oi <- c("+cilia_SMO_ag vs +cilia_SMO", "+cilia_PTH2R_ag vs +cilia_PTH2R")

lapply(unique(mss.dt$Label), function(x){
  
  g <- ggplot(mss.out[Label == x,], aes(y=-log10(padj), x=log2FC, color=label, label=gene)) +
  geom_point(size=1.5) +
  geom_text_repel(data=mss.out[Label == x & label != 'not',]) +
  geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
  geom_hline(yintercept = -log10(0.05), linetype=2, alpha=0.4) +
  ggtitle(paste0(x, ' comparison')) +
  labs(x=expression(log[2]~'FoldChange'), y=expression(-log[10]~'adj.pvalue')) +
  scale_color_manual(values=c('up'=col.pal[2], 'down'=col.pal[1], 'cilia assoc.'='green4', 'not'='grey')) +
  customTheme 
  g
  #BackupAsPDF(g, paste0(x,'.ajdpval.volcanoplots'), dimensions = c(10,8))
})
```
for the agonist comparisons, the pvalues are again quite high...
reduce the significance threshold and try those labels

```{r}
mss.out[, reduced.sig := 'not']
mss.out[abs(log2FC) >= 1 & pvalue < 0.005, sig := ifelse(log2FC > 0, 'up', 'down')]
mss.out[, reduced.label := reduced.sig]
mss.out[Protein %in% c(cilia.AM.prots, cilia.Consensus.prots), reduced.label := 'cilia assoc.']
mss.out[, reduced.label := factor(label,levels=c('not', 'down', 'up', 'cilia assoc.'))]
mss.out <- mss.out[order(reduced.label)]
```


```{r}
lapply(unique(mss.dt$Label), function(x){
  
  g <- ggplot(mss.out[Label == x,], aes(y=-log10(pvalue), x=log2FC, color=reduced.label, label=gene)) +
  geom_point(size=1.5) +
  geom_text_repel(data=mss.out[Label == x & label != 'not',]) +
  geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
  geom_hline(yintercept = -log10(0.005), linetype=2, alpha=0.4) +
  ggtitle(paste0(x, ' comparison')) +
  labs(x=expression(log[2]~'FoldChange'), y=expression(-log[10]~'raw pvalue')) +
  scale_color_manual(values=c('up'=col.pal[2], 'down'=col.pal[1], 'cilia assoc.'='green4', 'not'='grey')) +
  customTheme 
  g
  BackupAsPDF(g, paste0(x,'.rawpval.volcanoplots'), dimensions = c(10,8))
})
```
Enrichment analysis 
```{r}
g <- ggplot(mss.out[sig != 'not',.N, by=.(sig,Label)], aes(x=reorder(Label,-N), y=N, fill=sig)) +
  geom_bar(stat='identity') +
  scale_fill_manual(values=c('up'=col.pal[2], 'down'=col.pal[1])) +
  ggtitle('Overview of comparisons') +
  labs(x='Contrast', y=expression('N sig hits ('~log[2]~'FC & padj < 0.05)')) +
  customTheme + 
  theme(axis.text.x = element_text(size=8))

BackupAsPDF(g, 'nSighits.barplot', dimensions = c(5,5))
```
run enrichment analysis of the results

Run enrichment of these proteins 

GO CC enrichment first
```{r}
gmt.go <- loadGmtFromBioconductor(dbName='org.Mm.eg.db', ontology='CC', keyType='UNIPROT')
```

```{r}
# define the universe, the total set of identified genes in our study
universe <- as.character(unique(p.quant$Protein))

mss.out[, enrich.grp := interaction(Label,sig)]

enrich.dt <- enricherOnGroups(mss.out[sig != 'not'], 
                              groupColumns = 'enrich.grp', 
                              geneColumn = "Protein", 
                              term2gene.gmt = gmt.go, 
                              universe = universe)

fwrite(enrich.dt, ScriptAndDatedFileName('pwContrasts.GO.CC.enrichments.csv'))

simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich.grp')
fwrite(simp.enrich$simplified, ScriptAndDatedFileName('pwContrasts.GO.CC.simplified.enrichments.csv'))

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, 
                                groupColumn = 'enrich.grp', 
                                topN = 10,
                                title='GO Cellular Component', 
                                row_names_gp = gpar(fontsize = 7), 
                                negCols=unique(grep('down', simp.enrich$simplified$enrich.grp, value=T)),
                                column_names_gp= gpar(fontsize = 6), 
                                upperThreshold = 8)
ht
BackupAsPDF(ht, 'pwcontrast.GOCC.enrichments.heatmap', dimensions=c(6,7))
```
Run GO BP enrichment

```{r}
gmt.go <- loadGmtFromBioconductor(dbName='org.Mm.eg.db', ontology='BP', keyType='UNIPROT')
```

```{r}
# define the universe, the total set of identified genes in our study

enrich.dt <- enricherOnGroups(mss.out[sig != 'not'], 
                              groupColumns = 'enrich.grp', 
                              geneColumn = "Protein", 
                              term2gene.gmt = gmt.go, 
                              universe = universe)

# this is a very small set, just save the output
# cant simplify further...
fwrite(enrich.dt, ScriptAndDatedFileName('pwcontrast.GO.BP.enrichments.csv'))
```

```{r}
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich.grp')
fwrite(simp.enrich$simplified, ScriptAndDatedFileName('pwcontrasts.GO.BP.simplified.enrichments.csv'))

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, 
                                groupColumn = 'enrich.grp', 
                                topN = 12,
                                title='GO Biological Process', 
                                row_names_gp = gpar(fontsize = 7), 
                                negCols=unique(grep('down', simp.enrich$simplified$enrich.grp, value=T)),
                                column_names_gp= gpar(fontsize = 6), 
                                upperThreshold = 8)
ht
BackupAsPDF(ht, 'agonistContrast.GOCC.enrichments.heatmap', dimensions=c(10,10))
```
I think we want to take the normalized counts (tmp) and add these to the msstats data table
Do we want to look at heatmaps of the differentially expressed proteins? Possibly, just add in anyways so we can visualize.
```{r}
p.mat <- dcast(p.quant, Protein~SUBJECT, value.var='tmpAdjLogIntensities') %>% 
  as.matrix(rownames='Protein')

sigProts <- mss.out[label != 'not', unique(Protein)]

submat <- p.mat[rownames(p.mat) %in% sigProts, !grepl('[-]Cilia_[+]Agonist', colnames(p.mat))]
rownames(submat) <- multiUniprots2multiGenes(rownames(submat), species='MOUSE')

# 260 proteins in total that move, but larger number of contrasts
submat <- sweep(submat, 1, apply(submat,1, median, na.rm=T))


# get the sig genes and contrasts they are sig in 
sigGenes.dt <- unique(mss.out[label != 'not',.(gene, Label, label)])
sigGenes.dt <- sigGenes.dt[, lapply(.SD, function(x) paste(x, collapse='\n')), .SDcols=c('Label', 'label'), by=gene]

# tidy names to remove duplicates in collapse (how doe we get them? different protein same genename)
sigGenes.dt[Label == '+cilia_SMO_ag vs +cilia_SMO']
sigGenes.dt


mss.out[gene %in% c('Wasl','Tnks1bp1') & sig != 'not',]
# its a merge; merging names but just want the Label col in the order of the rownames
splitGenes <- sigGenes.dt[rownames(submat), Label, on='gene']


column_ha = HeatmapAnnotation(
                              ciliaStatus = ifelse(grepl('[+]Cilia', colnames(submat)), 'cilia', 'na'), 
                              agonistStatus= ifelse(grepl('[+]Agonist', colnames(submat)), 'agonist', 'na'), 
                              col=list(ciliaStatus=c('cilia'='darkgreen', 'na'='white'), agonistStatus=c('agonist'='darkred', na='white')),
                              border=T
                              )

hm <- Heatmap(submat, 
        name='Ints/\nmedian',
        top_annotation = column_ha,
        row_split = splitGenes,
        cluster_columns = F,
        column_split = list(str_extract(colnames(submat), 'SMO|PTH2R|GPR135')),
        col=colorRamp2(breaks=c(-3,0,3), colors = c(col.pal[1], 'white', col.pal[2])),
        column_names_gp = gpar(fontsize=6),
        row_names_gp = gpar(fontsize=5),
        row_title_gp = gpar(fontsize=8),
        border=T,
        row_title_rot = 0
        )

hm
BackupAsPDF(hm, 'sigGenes.onlyTestedConditions.heatmap', dimensions=c(13,17))
```
Check out the gene profiles we see; run enrichment andfigure out what they are enriched for 

```{r}
genes.oi <- c('Arpc3', 'Arpc4', 'Arpc1b', 'Actr3', 'Actb', 'Dbn1', 'Tpm3', 'Actg1', 'Tip1', 'Ocin', 'Tip2', 'Capzb', 'Myh14', 'Tpm4')
fwrite(data.table(genes.oi), ScriptAndDatedFileName('geneslist.csv'))

test.dt <- data.table(gene=genes.oi, 
                      enrich.grp='genes.oi'
                      )

universe <- mss.out$gene

gmt.go <- loadGmtFromBioconductor(dbName='org.Mm.eg.db', ontology='CC', keyType='SYMBOL')
enrich.dt <- enricherOnGroups(test.dt, groupColumns = 'enrich.grp', geneColumn='gene',universe = universe, term2gene.gmt=gmt.go)
fwrite(enrich.dt, ScriptAndDatedFileName('proteinsOI.enrich.GO.BP.csv'))
```

Rerun the analysis with a term for batch/subject
```{r}
mss.out[, c('numerator', 'denominator') := tstrsplit(Label, ' vs ', keep=c(1,2))]
mss.out


p.quant.wide <- dcast(p.quant, Protein~GROUP, value.var='tmpAdjLogIntensities', fun.aggregate = function(x){ paste0(x, collapse='|')})
colnames(p.quant.wide) <- paste0(colnames(p.quant.wide), '__normalizedLog2Intensities')
setnames(p.quant.wide, old='Protein__normalizedLog2Intensities', new='Protein')

p.quant.wide[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]
p.quant.wide %>% colnames()

p.quant.wide <- p.quant.wide[,.(Protein, gene, `GPR135_-Cilia__normalizedLog2Intensities`, `GPR135_+Cilia__normalizedLog2Intensities`, 
                `PTH2R_-Cilia_-Agonist__normalizedLog2Intensities`,`PTH2R_-Cilia_+Agonist__normalizedLog2Intensities`, `PTH2R_+Cilia_-Agonist__normalizedLog2Intensities`,`PTH2R_+Cilia_+Agonist__normalizedLog2Intensities`,
                `SMO_-Cilia_-Agonist__normalizedLog2Intensities`, `SMO_-Cilia_+Agonist__normalizedLog2Intensities`,`SMO_+Cilia_-Agonist__normalizedLog2Intensities`, `SMO_+Cilia_+Agonist__normalizedLog2Intensities`)]


mss.out <- merge(mss.out[,.(Protein,gene, Label, log2FC, SE, pvalue, padj)], p.quant.wide, by=c('Protein', 'gene'))
fwrite(mss.out, ScriptAndDatedFileName('mss.pwcomparisons.firstPass.wIntensities.csv.gz'))
```

Look at the counts matrix
```{r}
mss.out
submat <- sweep(p.mat, 1, apply(p.mat,1, median, na.rm=T))

Heatmap(submat,
        border=T,
        cluster_rows = clusterWNA(submat),
        row_title = sprint()
        column_split = list(str_extract(colnames(submat), 'SMO|PTH2R|GPR135')),
        col=colorRamp2(breaks=c(-3,0,3), colors = c(col.pal[1], 'white', col.pal[2])),
        show_row_names = F
        )
```






```{r}

 "green4"      "#6A3D9A"     "#FF7F00"     "black"       "gold1"       "skyblue2"    "#FB9A99"     "palegreen2"  "#CAB2D6"     "#121111"     "gray70" 
```




run another round of median normalzation on the data

Calculate the median value of each sample
Determine a refence value( dan be the median of all samples or user-defined) 
Divide each datapoint by the samples median and multiply it by reference value
```{r}
calMedianNormalization <- function(protDT){
  
  p.mat <- dcast(protDT, Protein~SUBJECT, value.var='LogIntensities') %>% 
  as.matrix(rownames='Protein')
  
  sample.median <- apply(p.mat, 2, median, na.rm=T)
  global.ref <- median(sample.median, na.rm=T)

  norm.mat <- apply(p.mat, MARGIN = 1, function(x){ (x/sample.median)*global.ref })
  
  norm.dt <- setDT(reshape2::melt(norm.mat))
  setnames(norm.dt, c('sample', 'Protein', 'medianNormInts'))

  
  #norm.dt[, c('GROUP','SUBJECT') := tstrsplit(sample, '[.]', keep=c(1,2))]
  message('returning normalized values...')
  print(head(norm.dt))
  
  out.dt <- merge(x=protDT, y=norm.dt, by.x=c('Protein','SUBJECT'), by.y=c('Protein','sample'))
  return(out.dt)
}
```

compare the output of this comparison withe tmp adjusted

```{r}
dp.out <- readRDS('./012425_ciliaAPEX_firstLook_data/2025_01_24_mss.dataProc.rds')
p.quant <- fread('012425_ciliaAPEX_firstLook_data/2025_01_24_ProteinlvlQuant.2ndRoundNorm.csv')

p.quant[,GROUP := factor(GROUP)]
p.quant[,SUBJECT := paste0(GROUP,'.', SUBJECT)]

p.quant <- calMedianNormalization(p.quant)
#fwrite(p.quant, ScriptAndDatedFileName('proteinlevelquant.renorm.csv'))
p.quant <- fread('~/Documents/projects/012425_AMarley_CiliaAPEX_Rerun/020325_msstatsDEanalysis_data/2025_02_04_proteinlevelquant.renorm.csv')
```

```{r}

g <- ggplot(p.quant, aes(x=SUBJECT, y=medianNormInts, fill=GROUP)) +
  geom_boxplot() +
  ggtitle('Normalized protein intensity distributions') +
  scale_fill_manual(values=unname(col.pal)) +
  customTheme
g

p.quant[, LogIntensities := medianNormInts]

dp.out$ProteinLevelData  <- p.quant
```


```{r}
mss <- groupComparison(contrast.matrix=contrasts.mat,
                       verbose=F,
                       data=dp.out)

mss.out <- setDT(mss$ComparisonResult)

mss.out[, padj := p.adjust(pvalue, method='BH'), by=(Label)]
mss.out[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]
mss.out[, c('numerator','denominator') := tstrsplit(Label, ' vs ', keep=c(1,2))]
mss.out[, sig := 'not']
mss.out[abs(log2FC) > 0.58 & padj < 0.05, sig := ifelse(log2FC > 0, 'up', 'down')]


mss.out[sig != 'not', .N, by=Label]
# this is with one round normalization
medianNormPass.dt <- mss.out[sig != 'not', .N, by=Label]
medianNormPass.dt[order(-N)]

g <- ggplot(p.quant[Protein %in% sharedProts,], aes(x=SUBJECT, y=tmpAdjLogIntensities, fill=GROUP)) +
  geom_boxplot() +
  ggtitle('Normalized protein intensity distributions') +
  scale_fill_manual(values=unname(col.pal)) +
  customTheme
g
```
I think we take the renormalized set and got for

```{r}
mss.out
```

