---
title: "021025_agonistContrastofContrast"
author: "Martin Gordon"
date: "2025-02-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}

library(data.table)
library(ggplot2)
library(magrittr)
library(ggrepel)
library(stringr)
library(ComplexHeatmap)
library(circlize)
library(MSstats)
library(viridis)
library(RColorBrewer)
library(hrbrthemes) # visually nice set of themes
library(showtext)
library(readxl)
library(emmeans)


source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")
source ("../../utils/bp_utils/NMF_Helpers.R")


source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("../../utils/mg_utils/r_utils/msstats_helperFunctions.R")

source("../../utils/bp_utils/LinearModels.R")

# for converting output to artMS format
source("../../utils/bp_utils/spectronautFile2ArtMS.R")
library(artMS)

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}

customTheme <- theme_ipsum_rc() +
  theme(panel.border = element_rect(color = "lightgrey", 
                                    fill = NA, 
                                    size = 1),
        axis.text.x = element_text(angle=90)
        )
#set one
col.pal <- getQualitativePalette(n=13)
#col.pal <- randomcoloR::distinctColorPalette(k=13)

# fonts needed for hrbrthemes
library(extrafont)
font_import()
loadfonts()

# directly from google font
sysfonts::font_add_google("Roboto Condensed")
showtext_auto()
```

read in the pquant and the other annotation data

```{r}
p.quant <- fread('012425_ciliaAPEX_firstLook_data/2025_01_24_ProteinlvlQuant.2ndRoundNorm.csv')
p.quant[, LogIntensities := tmpAdjLogIntensities]
```

read in the annotation files
```{r}
mh_mappings <- fread('~/Documents/utils/mg_utils/data/mouseHumanIDConversion.txt')

cilia.AM <- c('INPP5E', 'LCA5', 'PRKAR1A', 'ARL13B', 'SMO', 'EVC2', 'EVC', 'PKD2', 'TTC21B', 'PKHD1', 'IFT88', 'NPHP4', 'EFCAB7', 'CYS1')

cilia.Consensus <- c('Pdap1','Cep131','Cep41','Ift122','Ift172','Ift52','Ift57','Ift74','Ift81','Ift88','Cfap97','Lca5','Cemip2','Wdr35',
                     'Arl13b','Arl3','Inpp5e','Nme7','Kif3a','Kif3b','Kifap3','Ywhaz','Cys1','Evc2','Ttc21b','Ift56')

cilia.Consensus[!cilia.Consensus %in% unique(mh_mappings$Symbol.mouse)]

# all found
cilia.AM.prots <- mh_mappings[Symbol.human %in% cilia.AM, SWISS_PROT_IDs.mouse]
cilia.Consensus.prots <- mh_mappings[Symbol.mouse %in% cilia.Consensus, SWISS_PROT_IDs.mouse]
```

read up contrasts of contrast methodolgy and see how I can speciy a subset of contrasts.
*lm all proteins*

tidy the tables 

```{r}
p.quant[, cilia_present := factor(ifelse(grepl('[+]Cilia', GROUP), 'Cilia', 'noCilia'), levels=c('Cilia', 'noCilia'))]
p.quant[, ligand_present := factor(ifelse(grepl('[+]Agonist', GROUP), 'Ligand', 'noLigand'), levels=c('Ligand', 'noLigand'))]
p.quant[, batch := factor(SUBJECT, levels=c('1','2','3'))]
p.quant[, probe := factor(gsub('[_]','', str_extract(GROUP, "GPR135_|PTH2R_|SMO_")))]

p.quant[,.N, by=.(cilia_present, ligand_present, probe, batch, LogIntensities, tmpAdjLogIntensities)]
```

```{r}
# per probe, how does impact of ligand change given cilia status of cell. IE is the effect independent of cilia presence?
# need to fix these contrasts tomorrow
emmeans.contrastOfContrasts <- function (l, factorFormula = ~ probe*ligand_present*cilia_present){  
  emm <- emmeans::emmeans(l, factorFormula)
  contrast1 <- pairs(emm)
  contrast2 <- pairs(contrast1)
  return (as.data.table(contrast2))
}

# postProcessFunction = emmeans.contrastOfContrasts tells it to use the function above after doing anova.
# it is optional, adds significant time to the computation (~5x), but is nice to
# put a magnitude and p value on specific differences of differences
#
# cl = 6 parameter says to use 6 processes.  Set to NULL if multiprocess is failing, or higher or lower as you wish.
# Numbers near the number of processors available in your machine is probably counterproductive.

lm.out <- linearModelsAllProteins(p.quant, 
                                  formulaList = list(Interaction =  LogIntensities ~ batch + probe + ligand_present*cilia_present),
                                  splitColumn = "Protein",
                                  postProcessFunction = emmeans.contrastOfContrasts, cl = 6) %>% 
  suppressWarnings()

```
```{r}
contrasts.oi <- c("(SMO Ligand Cilia - SMO noLigand Cilia) - (SMO Ligand noCilia - SMO noLigand noCilia)",
                  "(PTH2R Ligand Cilia - PTH2R noLigand Cilia) - (PTH2R Ligand noCilia - PTH2R noLigand noCilia)")


lm.out$postProcess[, gene := multiUniprots2multiGenes(Protein, species = 'MOUSE')]
lm.out$postProcess[contrast %in% contrasts.oi & gene == 'Cep41', ] #pvalue the same..need to adjust..
```

```{r}
g <- ggplot(lm.out$anova[term != 'Residuals'], aes(x=p.value)) +
  geom_histogram() +
  facet_wrap(~term) +
  customTheme
g
BackupAsPDF(g, 'pvalue.histogram', dimensions = c(12,8))
```
```{r}
prots.oi <- lm.out$anova[term == 'ligand_present:cilia_present' & p.value < 0.05, unique(Protein)]
length(prots.oi)
```


```{r}
p.mat <- dcast(p.quant, Protein~paste0(GROUP,'.',SUBJECT), value.var='tmpAdjLogIntensities') %>% 
  as.matrix(rownames='Protein')

submat <- p.mat[rownames(p.mat) %in% prots.oi, !grepl('GPR135', colnames(submat))]
rownames(submat) <- multiUniprots2multiGenes(rownames(submat), species='MOUSE')

# 260 proteins in total that move, but larger number of contrasts
submat <- sweep(submat, 1, apply(submat,1, median, na.rm=T))

column_ha = HeatmapAnnotation(
                              agonistStatus= ifelse(grepl('[+]Agonist', colnames(submat)), 'agonist', 'na'), 
                              col=list(ciliaStatus=c('cilia'='darkgreen', 'na'='white'), agonistStatus=c('agonist'='darkred', na='white')),
                              border=T
                              )

row_ha = rowAnnotation(ciliaMarker = ifelse(rownames(submat) %in% multiUniprots2multiGenes(cilia.Consensus.prots, species='MOUSE'), 'yes', 'no'),
                       col=list(ciliaMarker=c('yes'=col.pal[4], 'no'='white')),
                       border=T)

hm <- Heatmap(submat, 
        name='Ints/\nmedian',
        top_annotation = column_ha,
        right_annotation = row_ha,
        cluster_columns = F,
        row_title = sprintf('%s Proteins pval < 0.05 interaction effects', nrow(submat)),
        column_split = list( str_extract(colnames(submat), 'SMO|PTH2R'),
                             str_extract(colnames(submat), '[+-]Cilia')),
        col=colorRamp2(breaks=c(-2,0,2), colors = c(col.pal[1], 'white', col.pal[2])),
        column_names_gp = gpar(fontsize=7),
        row_names_gp = gpar(fontsize=8),
        row_title_gp = gpar(fontsize=10),
        border=T
        )
hm
BackupAsPDF(draw(hm, column_title='differential (pval < 0.05) response to agonist based on cilia background'), 'sigGenes.contrastOfcontrasts.heatmap', dimensions=c(12,9))
```



```{r}
contrasts.oi <- c("(SMO Ligand Cilia - SMO noLigand Cilia) - (SMO Ligand noCilia- SMO noLigand noCilia)",
                  "(PTH2R Ligand Cilia - PTH2R noLigand Cilia) - (PTH2R Ligand noCilia - PTH2R noLigand noCilia)")


lm.out$postProcess$contrast %>% unique()
# get the subset
emm.dt <- lm.out$postProcess[contrast %in% contrasts.oi,]
emm.dt
# nothing passes threshold
emm.dt[model == 'interaction', padj := p.adjust(p.value, method='BH'), by=.(model, contrast)]
emm.dt[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]
```

```{r}
emm.dt[!is.na(estimate),.N, by=.(model,contrast)]

prots.oi <- emm.dt[p.value < 0.05, unique(Protein)]
```

plot heatmap of the proteins that are de using the contrast of contrasts package

```{r}
p.mat <- dcast(p.quant, Protein~paste0(GROUP,'.',SUBJECT), value.var='tmpAdjLogIntensities') %>% 
  as.matrix(rownames='Protein')

submat <- p.mat[rownames(p.mat) %in% prots.oi, !grepl('GPR135', colnames(submat))]
rownames(submat) <- multiUniprots2multiGenes(rownames(submat), species='MOUSE')

# 260 proteins in total that move, but larger number of contrasts
submat <- sweep(submat, 1, apply(submat,1, median, na.rm=T))

# get the sig genes and contrasts they are sig in 
sigGenes.dt <- unique(emm.dt[p.value < 0.05, .(gene, contrast)])
sigGenes.dt <- sigGenes.dt[, lapply(.SD, function(x) paste(x, collapse='\n')), .SDcols=c('contrast'), by=gene]
sigGenes.dt$contrast %>% unique()

# its a merge; merging names but just want the Label col in the order of the rownames
splitGenes <- sigGenes.dt[rownames(submat), contrast, on='gene']


column_ha = HeatmapAnnotation(
                              agonistStatus= ifelse(grepl('[+]Agonist', colnames(submat)), 'agonist', 'na'), 
                              col=list(ciliaStatus=c('cilia'='darkgreen', 'na'='white'), agonistStatus=c('agonist'='darkred', na='white')),
                              border=T
                              )


rownames(submat) 
row_ha = rowAnnotation(ciliaMarker = ifelse(rownames(submat) %in% multiUniprots2multiGenes(cilia.Consensus.prots, species='MOUSE'), 'yes', 'no'),
                       col=list(ciliaMarker=c('yes'=col.pal[4], 'no'='white')),
                       border=T)

hm <- Heatmap(submat, 
        name='Ints/\nmedian',
        top_annotation = column_ha,
        right_annotation = row_ha,
        row_split = splitGenes,
        cluster_columns = F,
        
        column_split = list( str_extract(colnames(submat), 'SMO|PTH2R'),
                             str_extract(colnames(submat), '[+-]Cilia')),
        col=colorRamp2(breaks=c(-2,0,2), colors = c(col.pal[1], 'white', col.pal[2])),
        column_names_gp = gpar(fontsize=7),
        row_names_gp = gpar(fontsize=8),
        row_title_gp = gpar(fontsize=6),
        border=T,
        row_title_rot = 0
        )
hm
BackupAsPDF(hm, 'sigGenes.contrastOfcontrasts.heatmap', dimensions=c(12,9))
```
pvalues all wrong... need to check the model...
Need to take one or two example proteins and play with the package and inspect the output

**Not used**
Play with the emmeans package to get used to the contrast method. Take a subset of our data and use that..
Worst comes to worst, subset and run on each probe seperately with two variables with two levels each

```{r}
p.quant[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]
test.quant <- p.quant[gene == 'Timm50',]

# model investigation
# go with the interaction term
mod <- lm(LogIntensities~batch+probe+cilia_present, data=test.quant)
mod1 <- lm(LogIntensities~batch+probe+cilia_present+ligand_present, data=test.quant)
mod2 <- lm(LogIntensities~batch+probe+cilia_present*ligand_present, data=test.quant) # w interaction


mod3 <- lm(LogIntensities~batch+probe*cilia_present*ligand_present, data=test.quant) # w interaction between all terms.. # we assume the difference is impacted by the probe, the cilia and ligand impact


# adding ligand adds no significant difference to the model, but  interaction of ligand and cilia does
# ligand only has interesting effect dependent on cilia level (ie in cilicated cells as expected)
# basically, I do think we want the interaction of these terms
anova(mod1, mod2) 
anova(mod1, mod3)# for this protien, no material difference, but we do want to check if the term changes for each 
```

The problem with my lm input, I think probe is averaged over... need to include the interaction term in the input model (sacrifice df... better to run each seperately with a simplier model?)
ok, now we have a good model create an emmeans object

```{r}
# dont want to look for probe interaction; wqnt to 
EMM <- emmeans(mod3, ~ probe*ligand_present*cilia_present)
c1 <- pairs(EMM, by='probe') # run pw contrasts within each level of probe
c2 <- pairs(c1, by='probe') # this is the contrast of contrast; from the first level above- pw comparison


c2.dt[contrast == '(Ligand Cilia - noLigand Cilia) - (Ligand noCilia - noLigand noCilia)']
# same issue; think we are averaging over the same levels
c2.dt[contrast =='(PTH2R Ligand Cilia - PTH2R noLigand Cilia) - (PTH2R Ligand noCilia - PTH2R noLigand noCilia)']
c2.dt[contrast =='(SMO Ligand Cilia - SMO noLigand Cilia) - (SMO Ligand noCilia - SMO noLigand noCilia)']
```

Rerun the emm with the fixed format and collapse to 

```{r}
# per probe, how does impact of ligand change given cilia status of cell. IE is the effect independent of cilia presence?
emmeans.contrastOfContrasts <- function (l, factorFormula = ~ probe*ligand_present*cilia_present){  
  emm <- emmeans::emmeans(l, factorFormula)
  contrast1 <- pairs(emm, by='probe')
  contrast2 <- pairs(contrast1, by='probe')
  return (as.data.table(contrast2))
}

# postProcessFunction = emmeans.contrastOfContrasts tells it to use the function above after doing anova.
# it is optional, adds significant time to the computation (~5x), but is nice to
# put a magnitude and p value on specific differences of differences
#
# cl = 6 parameter says to use 6 processes.  Set to NULL if multiprocess is failing, or higher or lower as you wish.
# Numbers near the number of processors available in your machine is probably counterproductive.

lm.out <- linearModelsAllProteins(p.quant, 
                                  formulaList = list(Interaction =  LogIntensities ~ batch + probe*ligand_present*cilia_present), # control for batch and look for interactions in the other variables
                                  splitColumn = "Protein",
                                  postProcessFunction = emmeans.contrastOfContrasts, cl = 6)
```

First look at the lm/anova and see what terms are significant

```{r}
g <- ggplot(lm.out$anova, aes(x=p.value)) +
  geom_histogram() +
  facet_wrap(~term) +
  ggtitle('pvalue distributions of lm predictors', subtitle = "n=1 model per protein" ) +
  customTheme
g
BackupAsPDF(g, 'lmterms.pvalDistributions.histogram', dimensions = c(8,6))
```
pull out the contrast of contrast results and lets see which term is significant
```{r}
lm.out$anova[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]
lm.out$anova[, padj := p.adjust(p.value, method='BH'), by=.(term, model)]
# this is a different set of genes. lets plot them!

intGenes <- lm.out$anova[term == 'probe:ligand_present:cilia_present' & p.value < 0.005, unique(Protein)]
```

```{r}
p.mat <- dcast(p.quant, Protein~paste0(GROUP,'.',SUBJECT), value.var='tmpAdjLogIntensities') %>% 
  as.matrix(rownames='Protein')

submat <- p.mat[rownames(p.mat) %in% intGenes, !grepl('GPR135', colnames(submat))]
rownames(submat) <- multiUniprots2multiGenes(rownames(submat), species='MOUSE')

# 260 proteins in total that move, but larger number of contrasts
submat <- sweep(submat, 1, apply(submat,1, median, na.rm=T))


column_ha = HeatmapAnnotation(
                              agonistStatus= ifelse(grepl('[+]Agonist', colnames(submat)), 'agonist', 'na'), 
                              col=list(ciliaStatus=c('cilia'='darkgreen', 'na'='white'), agonistStatus=c('agonist'='darkred', na='white')),
                              border=T
                              )


rownames(submat) 
row_ha = rowAnnotation(ciliaMarker = ifelse(rownames(submat) %in% multiUniprots2multiGenes(cilia.Consensus.prots, species='MOUSE'), 'yes', 'no'),
                       col=list(ciliaMarker=c('yes'=col.pal[4], 'no'='white')),
                       border=T)

hm <- Heatmap(submat, 
        name='Ints/\nmedian',
        top_annotation = column_ha,
        right_annotation = row_ha,
        cluster_columns = F,
        
        column_split = list( str_extract(colnames(submat), 'SMO|PTH2R'),
                             str_extract(colnames(submat), '[+-]Cilia')),
        col=colorRamp2(breaks=c(-2,0,2), colors = c(col.pal[1], 'white', col.pal[2])),
        column_names_gp = gpar(fontsize=7),
        row_names_gp = gpar(fontsize=8),
        row_title_gp = gpar(fontsize=6),
        border=T,
        row_title_rot = 0
        )
hm
BackupAsPDF(draw(hm, column_title='genes with differential response (pval < 0.05) to agonist dependent on cilia background'), 'sigGenes.probeInteractions.heatmap', dimensions=c(12,12))
```
Subset to the high confidence set of interactions.. not too confident in any of these 

```{r}
intGenes <- lm.out$anova[term == 'probe:ligand_present:cilia_present' & p.value < 0.005, unique(Protein)]

submat <- p.mat[rownames(p.mat) %in% intGenes, !grepl('GPR135', colnames(submat))]
rownames(submat) <- multiUniprots2multiGenes(rownames(submat), species='MOUSE')

# 260 proteins in total that move, but larger number of contrasts
submat <- sweep(submat, 1, apply(submat,1, median, na.rm=T))


column_ha = HeatmapAnnotation(
                              agonistStatus= ifelse(grepl('[+]Agonist', colnames(submat)), 'agonist', 'na'), 
                              col=list(ciliaStatus=c('cilia'='darkgreen', 'na'='white'), agonistStatus=c('agonist'='darkred', na='white')),
                              border=T
                              )


rownames(submat) 
row_ha = rowAnnotation(ciliaMarker = ifelse(rownames(submat) %in% multiUniprots2multiGenes(cilia.Consensus.prots, species='MOUSE'), 'yes', 'no'),
                       col=list(ciliaMarker=c('yes'=col.pal[4], 'no'='white')),
                       border=T)

hm <- Heatmap(submat, 
        name='Ints/\nmedian',
        top_annotation = column_ha,
       # right_annotation = row_ha,
        cluster_columns = F,
        
        column_split = list( str_extract(colnames(submat), 'SMO|PTH2R'),
                             str_extract(colnames(submat), '[+-]Cilia')),
        col=colorRamp2(breaks=c(-2,0,2), colors = c(col.pal[1], 'white', col.pal[2])),
        column_names_gp = gpar(fontsize=7),
        row_names_gp = gpar(fontsize=8),
        row_title_gp = gpar(fontsize=6),
        border=T,
        row_title_rot = 0
        )
hm
BackupAsPDF(draw(hm, column_title='genes with differential response (pval < 0.005) to agonist dependent on cilia background'), 'sigGenes.pval0.005.probeInteractions.heatmap', dimensions=c(12,8))
```


```{r}
# these are the terms that passed. none will pass correction I
lm.out$anova[term == "probe:ligand_present:cilia_present" & p.value < 0.05, unique(gene)]
lm.out$anova[, padj := p.adjust(p.value, method='BH'), by=.(term, model)]

# no interaction term passess correction
lm.out$anova[p.value < 0.005 & term == "ligand_present:cilia_present"]

# write out all the results for me..
fwrite(lm.out$anova, ScriptAndDatedFileName('probes.agonistLigand.interactions.anova.csv'))

lm.anova.tidy <- dcast(lm.out$anova[term %in% c('probe:ligand_present:cilia_present'), .(Protein, gene, term, p.value, padj)], gene+Protein~term, value.var=c('p.value', 'padj'))
setnames(lm.anova.tidy, old=c("p.value_probe:ligand_present:cilia_present","padj_probe:ligand_present:cilia_present"), new=c('pvalue_cilia:ligand_interaction', 'adj.pvalue_cilia:ligand_interaction'))

# use this set
fwrite(lm.anova.tidy, ScriptAndDatedFileName('probes.agonistLigand.interactions.anova.wide.csv'))
```

ok look at the contrast of contrast results
```{r}
# pull out the contrasts that we weant to use which is comparing FC in the two cell backgrounds
contrast.oi <- c('(Ligand Cilia - noLigand Cilia) - (Ligand noCilia - noLigand noCilia)')

emm.dt <- lm.out$postProcess[contrast %in% contrast.oi,]
emm.dt[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]
emm.dt[, padj := p.adjust(p.value, method='BH'), by=.(contrast, probe)]
```
now lets look at the heatmap of the contrast proteins
```{r}
#plot heatmap of the contrast of contrast proteins
intGenes <- emm.dt[p.value < 0.05, unique(Protein)] # 13 proteins found..
intGenes
# all 53 are in the SMO probe
emm.dt[p.value < 0.05, .N, by=probe]

submat <- p.mat[rownames(p.mat) %in% intGenes, grepl('[SMO]', colnames(p.mat))]
rownames(submat) <- multiUniprots2multiGenes(rownames(submat), species='MOUSE')

# 260 proteins in total that move, but larger number of contrasts
submat <- sweep(submat, 1, apply(submat,1, median, na.rm=T))


column_ha = HeatmapAnnotation(
                              agonistStatus= ifelse(grepl('[+]Agonist', colnames(submat)), 'agonist', 'na'), 
                              col=list(ciliaStatus=c('cilia'='darkgreen', 'na'='white'), agonistStatus=c('agonist'='darkred', na='white')),
                              border=T
                              )

row_ha = rowAnnotation(ciliaMarker = ifelse(rownames(submat) %in% multiUniprots2multiGenes(cilia.Consensus.prots, species='MOUSE'), 'yes', 'no'),
                       col=list(ciliaMarker=c('yes'=col.pal[4], 'no'='white')),
                       border=T)

hm <- Heatmap(submat, 
        name='Ints/\nmedian',
        top_annotation = column_ha,
       # right_annotation = row_ha,
        cluster_columns = F,
        column_split = list( str_extract(colnames(submat), 'SMO|PTH2R'),
                             str_extract(colnames(submat), '[+-]Cilia')),
        col=colorRamp2(breaks=c(-2,0,2), colors = c(col.pal[1], 'white', col.pal[2])),
        column_names_gp = gpar(fontsize=7),
        row_names_gp = gpar(fontsize=8),
        row_title_gp = gpar(fontsize=6),
        border=T,
        row_title_rot = 0
        )
hm
BackupAsPDF(draw(hm, column_title='significant (raw pval < 0.05) agonist response in SMO cilated vs non-cilated cell-lines'), 'sigGenes.pval0.05.SMO.interactions.heatmap', dimensions=c(12,8))
```
write out the interaction analysis files 

```{r}
emm.subset.dt <- emm.dt[probe != 'GPR135',]
emm.subset.dt[, padj := p.adjust(p.value, method='BH'), by=.(probe)]
#fwrite(emm.subset.dt, ScriptAndDatedFileName('probes.agonistLigand.interaction.contrastOfcontrast.csv'))

emm.subset.dt <- dcast(emm.subset.dt, Protein+gene~probe, value.var=c('estimate', 'p.value', 'padj'))
setnames(emm.subset.dt, old=c('estimate_PTH2R', 'estimate_SMO'), new=c('LFC_PTH2R', 'LFC_SMO'))

#fwrite(emm.subset.dt, ScriptAndDatedFileName('probes.agonistLigand.interaction.contrastOfcontrast.wide.csv'))
```
```{r}
test <- fread('./020425_pwcomparisons/DEA/tables/2025_02_04_mss.pwcomparisons.firstPass.wIntensities.csv.gz')
test$Label %>% unique()


test[Label == "+cilia_PTH2R_ag vs +cilia_PTH2R", .(pvalue, padj)]
```


