---
title: "021025_additionalContrastsanQCplots"
author: "Martin Gordon"
date: "2025-02-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}

library(data.table)
library(ggplot2)
library(magrittr)
library(ggrepel)
library(stringr)
library(ComplexHeatmap)
library(circlize)
library(MSstats)
library(viridis)
library(RColorBrewer)
library(hrbrthemes) # visually nice set of themes
library(showtext)
library(readxl)
library(emmeans)


source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")
source ("../../utils/bp_utils/NMF_Helpers.R")

source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("../../utils/mg_utils/r_utils/msstats_helperFunctions.R")


# for converting output to artMS format
source("../../utils/bp_utils/spectronautFile2ArtMS.R")
library(artMS)

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}

customTheme <- theme_ipsum_rc() +
  theme(panel.border = element_rect(color = "lightgrey", 
                                    fill = NA, 
                                    size = 1),
        axis.text.x = element_text(angle=90)
        )
#set one
col.pal <- getQualitativePalette(n=13)
#col.pal <- randomcoloR::distinctColorPalette(k=13)

# fonts needed for hrbrthemes
library(extrafont)
font_import()
loadfonts()

# directly from google font
sysfonts::font_add_google("Roboto Condensed")
showtext_auto()
```

read in the output of dataProcess that was used for the differential comparisons (compare to old data).
Plot Justins recommendations and also the 

Read in data
```{r}
spec <- fread('./data/CR2/20250115_CR2_SMO_APEX_Report.tsv')

dp.out <- readRDS('./012425_ciliaAPEX_firstLook_data/2025_01_24_mss.dataProc.rds')
p.quant <- fread('012425_ciliaAPEX_firstLook_data/2025_01_24_ProteinlvlQuant.2ndRoundNorm.csv')

p.quant[, GROUP := factor(GROUP)]
p.quant[, SUBJECT := paste0(GROUP,'.', SUBJECT)] # to enforce lm for group comparisons

p.quant[, LogIntensities := tmpAdjLogIntensities]
dp.out$ProteinLevelData  <- p.quant
```
cilia protein markers 

```{r}
mh_mappings <- fread('~/Documents/utils/mg_utils/data/mouseHumanIDConversion.txt')

cilia.AM <- c('INPP5E', 'LCA5', 'PRKAR1A', 'ARL13B', 'SMO', 'EVC2', 'EVC', 'PKD2', 'TTC21B', 'PKHD1', 'IFT88', 'NPHP4', 'EFCAB7', 'CYS1')

cilia.Consensus <- c('Pdap1','Cep131','Cep41','Ift122','Ift172','Ift52','Ift57','Ift74','Ift81','Ift88','Cfap97','Lca5','Cemip2','Wdr35',
                     'Arl13b','Arl3','Inpp5e','Nme7','Kif3a','Kif3b','Kifap3','Ywhaz','Cys1','Evc2','Ttc21b','Ift56')

cilia.Consensus[!cilia.Consensus %in% unique(mh_mappings$Symbol.mouse)]

# all found
cilia.AM.prots <- mh_mappings[Symbol.human %in% cilia.AM, SWISS_PROT_IDs.mouse]
cilia.Consensus.prots <- mh_mappings[Symbol.mouse %in% cilia.Consensus, SWISS_PROT_IDs.mouse]
```
Lets look at the sum of intensities in each run

```{r}
g <- ggplot(spec[, .(sumInts=sum(Intensity, na.rm=T)), by=.(Condition,BioReplicate)], aes(x=paste0(Condition), y=(sumInts), fill=Condition, group=as.factor(BioReplicate))) +
  geom_bar(stat='identity', position='dodge', color='black') +
  ggtitle('Sum of intensities per sample') +
  scale_color_manual(values=col.pal) +
  scale_fill_manual(values=col.pal) +
  customTheme

g
BackupAsPDF(g, 'sumIntensities.barplot')


g <- ggplot(spec[, .(sumInts=sum(Intensity, na.rm=T)), by=.(RunID=gsub('exD00|[.]raw','', Run), Condition,BioReplicate)], aes(x=RunID, y=sumInts, fill=Condition)) +
  geom_bar(stat='identity',color='black') +
  ggtitle('Sum of intensities per sample') +
  scale_color_manual(values=col.pal) +
  scale_fill_manual(values=col.pal) +
  customTheme

g
BackupAsPDF(g, 'sumIntensities.runOrder.barplot')
```
find the median intensity peptides in each run and plot them; how do they look across samples and hwo do they look pre- and post- normalization

```{r}
g <- ggplot(spec[, .(sumInts=sum(Intensity, na.rm=T)), by=.(RunID=gsub('exD00|[.]raw','', Run), Condition,BioReplicate)], aes(x=RunID, y=sumInts, fill=Condition)) +
  geom_bar(stat='identity',color='black') +
  ggtitle('Sum of intensities per sample') +
  scale_color_manual(values=col.pal) +
  scale_fill_manual(values=col.pal) +
  customTheme

g
```
First lets look at levels of biotin carboxylases before and after normalization 
get sample median, subtract each value from the median and find the peptide closest

Look at peptide distributions across groups
```{r}
spec[, sampleMedian := median(log2(Intensity), na.rm=T), by=Run]
spec[, intsDiff := log2(Intensity) - sampleMedian]

# get peptides closest to median value; how do these peptide values look across groups?
medianPeptides <- spec[, .SD[which.min(abs(intsDiff)),], by=Run] %>% 
  .[,.(Condition, BioReplicate, ProteinName, PeptideSequence, PrecursorCharge)]
medianPeptides[, protGrp := paste0(ProteinName, '__', PeptideSequence, '__', PrecursorCharge)]
```


```{r}
spec[, protGrp :=  paste0(ProteinName, '__', PeptideSequence, '__', PrecursorCharge)]

f.mat <- dcast(spec, protGrp~paste0(Condition, '.', BioReplicate), value.var='Intensity') %>% 
  as.matrix(rownames='protGrp')

# basically pull out median peptides with measuments across majority samples (n = smallest probe group size = 30 - 6 =24)
subset.oi <- apply(f.mat, 1, function(x){sum(!is.na(x))})[medianPeptides$protGrp][apply(f.mat, 1, function(x){sum(!is.na(x))})[medianPeptides$protGrp] >= 24]


g <- ggplot(spec[protGrp %in% names(subset.oi),], aes(x=paste0(Condition, BioReplicate), y=log2(Intensity), group=protGrp, color=protGrp)) +
  geom_line() +
  geom_point() +
  ggtitle('median peptide log2 ints.') +
  customTheme
g
BackupAsPDF(g, 'medianpetide.ints.linechart', dimensions = c(10,8))
```
Look at the endogenous biotinylases across each of the groups. cant find our standard set in the data, look for more for potential cilia markers
No endo biotinylases detected in this set
```{r}
p.quant <- fread('012425_ciliaAPEX_firstLook_data/2025_01_24_ProteinlvlQuant.2ndRoundNorm.csv')

biotin.carboxylases.up <- c("O00763","P05165","P11498","Q13085","Q96RQ3")
p.quant[, gene := multiUniprots2multiGenes(Protein)]


g <- ggplot(p.quant[Protein %in% biotin.carboxylases.up,], aes(x=paste0(GROUP, SUBJECT), y=LogIntensities)) +
  geom_line() +
  geom_point() +
  ggtitle('median peptide log2 ints.') +
  customTheme
g
```
Run artMS on the output so we can see the differences 
```{r}
artms.in <- spectronautFile2ArtMS('./data/CR2/20250115_CR2_SMO_APEX_Report.tsv')

```


*to do ahead of data return*
- look at endo biotinylases and profile of median peptides across runs - not found in the dataset, so have to ignore for now
- rerun the contrasts controlling for biorep1; include the other contrasts, then try a contrast of contrasts approach

Run the QC analysis of the data using artSM to
```{r}
artmsQualityControlEvidenceBasic(evidence_file = artms.in$evidence_file,
                                 keys_file = artms.in$keys_file, 
                                 prot_exp = 'AB')

artmsQualityControlEvidenceExtended(evidence_file = artms.in$evidence_file,
                                  keys_file = artms.in$keys_file)
```
**redo contrasts**

lets redo the PW comparisons controlling for the outlier impact from biorep 1
Take the list of 
```{r}
contrasts.list <- list('+cilia_SMO_ag vs +cilia_SMO' = data.table(V1='SMO_+Cilia_+Agonist', V2='SMO_+Cilia_-Agonist'),
                       '+cilia_SMO vs -cilia_SMO' = data.table(V1='SMO_+Cilia_-Agonist', V2='SMO_-Cilia_-Agonist'),
                       '+cilia_PTH2R_ag vs +cilia_PTH2R' = data.table(V1='PTH2R_+Cilia_+Agonist', V2='PTH2R_+Cilia_-Agonist'),
                       '+cilia_PTH2R vs -cilia_PTH2R' = data.table(V1='PTH2R_+Cilia_-Agonist', V2='PTH2R_-Cilia_-Agonist'),
                       '+cilia_GPR135 vs -cilia_GPR135' = data.table(V1='GPR135_+Cilia', V2='GPR135_-Cilia'),
                       # look at the agonist in cilia vs agonist no-cilia comparison
                       '+cilia_SMO_ag vs -cilia_SMO_ag' = data.table(V1='SMO_+Cilia_+Agonist', V2='SMO_-Cilia_+Agonist'),
                       '+cilia_PTH2R_ag vs -cilia_PTH2R_ag' = data.table(V1='PTH2R_+Cilia_+Agonist', V2='PTH2R_-Cilia_-Agonist'))



contrasts.mat <- MSstats::MSstatsContrastMatrix(contrasts.list, 
                               conditions = unique(p.quant$GROUP), 
                               labels = names(contrasts.list))
```
use the TMP normalized data and try lmm controlling for biorep
```{r}
# I think we want to control at the level of probe?
p.quant[, oldInts := LogIntensities]
p.quant[, LogIntensities := tmpAdjLogIntensities]

dp.out$ProteinLevelData <- p.quant
```

We want to run a mixed effects model to control for impact in the biorep1 and compare this to our previous hit-list
Then run a contrast of contrasts for the agonist activation in both backgrounds

```{r}
mss <- groupComparison(contrast.matrix=contrasts.mat,
                       verbose=T,
                       data=dp.out)
mss.dt <- setDT(mss$ComparisonResult)

```

save the results of the lmm test and compare to our original data; important to model for this batch effect

```{r}
mss.dt[, padj := p.adjust(pvalue, method='BH'), by=(Label)]
mss.dt[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]
mss.dt[, c('numerator','denominator') := tstrsplit(Label, ' vs ', keep=c(1,2))]
mss.dt[, sig := 'not']
mss.dt[abs(log2FC) > 0.58 & padj < 0.05, sig := ifelse(log2FC > 0, 'up', 'down')]
mss.dt[, probe := str_extract(Label, 'SMO|PTH2R|GPR135')]
mss.dt[, label := sig]
mss.dt[Protein %in% c(cilia.AM.prots, cilia.Consensus.prots), label := 'cilia assoc.']

# save the results to file 
fwrite(mss.dt, ScriptAndDatedFileName('pwcomparisons.allContrasts.batchControlled.csv.gz'))
```

So modelling for biorep1 recovered more hits, so proceed with the new analysis set
```{r}
mss.dt <- fread('~/Documents/projects/012425_AMarley_CiliaAPEX_Rerun/021025_additionalContrastsandQCplots_data/2025_02_10_pwcomparisons.allContrasts.batchControlled.csv.gz')
old.mss.dt <- fread('~/Documents/projects/012425_AMarley_CiliaAPEX_Rerun/020325_msstatsDEanalysis_data/2025_02_04_mss.pwcomparisons.allGroups.csv.gz')

old.mss.dt[, padj := p.adjust(pvalue, method='BH'), by=(Label)]
old.mss.dt[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]
old.mss.dt[, c('numerator','denominator') := tstrsplit(Label, ' vs ', keep=c(1,2))]
old.mss.dt[, sig := 'not']
old.mss.dt[abs(log2FC) > 0.58 & padj < 0.05, sig := ifelse(log2FC > 0, 'up', 'down')]
rm(old.mss.dt)
```

Controlling for batch (biorep 1), improves results. Regenerate the plots/tables used previously

reorder the points plotted so 
```{r}
mss.dt[, label := factor(label, levels = c('not', 'down', 'up', 'cilia assoc.'))]
```


```{r}
lapply(unique(mss.dt$Label), function(x){
  
  g <- ggplot(mss.dt[Label == x,][order(label)], aes(y=-log10(padj), x=log2FC, color=label, label=gene)) +
  geom_point(size=1.5) +
  geom_text_repel(data=mss.dt[Label == x & label != 'not',], show.legend=FALSE) +
  geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
  geom_hline(yintercept = -log10(0.05), linetype=2, alpha=0.4) +
  ggtitle(paste0(x, ' comparison')) +
  labs(x=expression(log[2]~'FoldChange'), y=expression(-log[10]~'adjusted pvalue')) +
  scale_color_manual(values=c('up'=col.pal[2], 'down'=col.pal[1], 'cilia assoc.'='green4', 'not'='grey')) +
  customTheme 
  
  g
  BackupAsPDF(g, paste0(x,'.volcanoplots'), dimensions = c(10,8))
})
```
```{r}
g <- ggplot(mss.dt[sig != 'not',.N, by=.(sig,Label)], aes(x=reorder(Label,-N), y=N, fill=sig)) +
  geom_bar(stat='identity') +
  scale_fill_manual(values=c('up'=col.pal[2], 'down'=col.pal[1])) +
  ggtitle('Overview of comparisons') +
  labs(x='Contrast', y=expression('N sig hits ('~log[2]~'FC & padj < 0.05)')) +
  customTheme + 
  theme(axis.text.x = element_text(size=8))
g
BackupAsPDF(g, 'nSighits.barplot', dimensions = c(5,5))
```
GO CC enrichment first
```{r}
gmt.go <- loadGmtFromBioconductor(dbName='org.Mm.eg.db', ontology='CC', keyType='UNIPROT')
```

```{r}
# define the universe, the total set of identified genes in our study
universe <- as.character(unique(p.quant$Protein))

mss.dt[, enrich.grp := interaction(Label,sig)]

enrich.dt <- enricherOnGroups(mss.dt[sig != 'not'], 
                              groupColumns = 'enrich.grp', 
                              geneColumn = "Protein", 
                              term2gene.gmt = gmt.go, 
                              universe = universe)

fwrite(enrich.dt, ScriptAndDatedFileName('pwContrasts.GO.CC.enrichments.csv'))

simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich.grp')
fwrite(simp.enrich$simplified, ScriptAndDatedFileName('pwContrasts.GO.CC.simplified.enrichments.csv'))

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, 
                                groupColumn = 'enrich.grp', 
                                topN = 10,
                                title='GO Cellular Component', 
                                row_names_gp = gpar(fontsize = 7), 
                                negCols=unique(grep('down', simp.enrich$simplified$enrich.grp, value=T)),
                                column_names_gp= gpar(fontsize = 6), 
                                upperThreshold = 8)
ht
BackupAsPDF(ht, 'pwcontrast.GOCC.enrichments.heatmap', dimensions=c(6,7))
```
```{r}
gmt.go <- loadGmtFromBioconductor(dbName='org.Mm.eg.db', ontology='BP', keyType='UNIPROT')
```

```{r}
# define the universe, the total set of identified genes in our study

enrich.dt <- enricherOnGroups(mss.dt[sig != 'not'], 
                              groupColumns = 'enrich.grp', 
                              geneColumn = "Protein", 
                              term2gene.gmt = gmt.go, 
                              universe = universe)

# this is a very small set, just save the output
# cant simplify further...
fwrite(enrich.dt, ScriptAndDatedFileName('pwcontrast.GO.BP.enrichments.csv'))
```

```{r}
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich.grp')
fwrite(simp.enrich$simplified, ScriptAndDatedFileName('pwcontrasts.GO.BP.simplified.enrichments.csv'))

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, 
                                groupColumn = 'enrich.grp', 
                                topN = 12,
                                title='GO Biological Process', 
                                row_names_gp = gpar(fontsize = 7), 
                                negCols=unique(grep('down', simp.enrich$simplified$enrich.grp, value=T)),
                                column_names_gp= gpar(fontsize = 6), 
                                upperThreshold = 8)
ht
BackupAsPDF(ht, 'agonistContrast.GOCC.enrichments.heatmap', dimensions=c(10,10))
```

```{r}
p.mat <- dcast(p.quant, Protein~paste0(GROUP,'.',SUBJECT), value.var='tmpAdjLogIntensities') %>% 
  as.matrix(rownames='Protein')

sigProts <- mss.dt[label != 'not', unique(Protein)]

submat <- p.mat[rownames(p.mat) %in% sigProts,]
rownames(submat) <- multiUniprots2multiGenes(rownames(submat), species='MOUSE')

# 260 proteins in total that move, but larger number of contrasts
submat <- sweep(submat, 1, apply(submat,1, median, na.rm=T))


# get the sig genes and contrasts they are sig in 
sigGenes.dt <- unique(mss.dt[label != 'not',.(gene, Label, label)])
sigGenes.dt <- sigGenes.dt[, lapply(.SD, function(x) paste(x, collapse='\n')), .SDcols=c('Label', 'label'), by=gene]


# its a merge; merging names but just want the Label col in the order of the rownames
splitGenes <- sigGenes.dt[rownames(submat), Label, on='gene']


column_ha = HeatmapAnnotation(
                              ciliaStatus = ifelse(grepl('[+]Cilia', colnames(submat)), 'cilia', 'na'), 
                              agonistStatus= ifelse(grepl('[+]Agonist', colnames(submat)), 'agonist', 'na'), 
                              col=list(ciliaStatus=c('cilia'='darkgreen', 'na'='white'), agonistStatus=c('agonist'='darkred', na='white')),
                              border=T
                              )

row_ha = rowAnnotation(ciliaMarker = ifelse(rownames(submat) %in% multiUniprots2multiGenes(cilia.Consensus.prots, species='MOUSE'), 'yes', ''),
                       col=list(ciliaMarker=c('yes'=col.pal[4], ''='white')),
                       border=T)

hm <- Heatmap(submat, 
        name='Ints/\nmedian',
        top_annotation = column_ha,
        right_annotation = row_ha,
        row_split = splitGenes,
        cluster_columns = F,
        column_split = list(str_extract(colnames(submat), 'SMO|PTH2R|GPR135')),
        col=colorRamp2(breaks=c(-3,0,3), colors = c(col.pal[1], 'white', col.pal[2])),
        column_names_gp = gpar(fontsize=6),
        row_names_gp = gpar(fontsize=3.5),
        row_title_gp = gpar(fontsize=6),
        border=T,
        row_title_rot = 0
        )
hm
BackupAsPDF(hm, 'sigGenes.heatmap', dimensions=c(17,29))
```
