---
title: "121223.APMS.FragPipeIn"
author: "Martin Gordon"
date: "2023-12-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 
## previous analysis didnt proceed with the correct format
Fragpipe input - reprocess using bp_utils scripts


## questions
What are our controls?
What are the backgrounds? tools we need to use?
```{r}
library (data.table)
library (ComplexHeatmap)
library (ggplot2)
library(readxl)
library(MSstats)
library(circlize)
library(ggvenn)

source ("../../utils/bp_utils/spectronautFile2ArtMS.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")

getwd()
```

path to spec file from FragPipe

```{r}
spectronautPeptideFile <- './data/MSstats.csv'
```


```{r}
a <- fread (spectronautPeptideFile)

# histogram for log2 intensity; alredy filtered
hist(log2(a$Intensity), breaks = 100)
abline(v = 5)


boxplot(log2(Intensity)~interaction(Condition,BioReplicate), data=a) #doesnt look normalised anyway....
```

```{r}
cf<- list()
# normalization method FALSE = no normalization; default is global medians which you can get my removing/commenting out all normalization lines
# cf$msstats$normalization_method = FALSE

#cf$msstats$normalization_method = "globalStandards"
#cf$msstats$normalization_reference <-  "P38398"

# should artms attempt to annotate proteins 1 = yes; 0 = no
cf$output_extras$annotate$enabled <- as.integer(1)
# should artms do extended QC 1 = yes; 0= no
cf$qc$extended <- as.integer(1)
cf$qc$basic <- as.integer(1)

cf$output_extras$annotate$species <- "HUMAN"


# make files in artMS format
# difference with prev attempt is we have included dummy variables 
globalInput <- spectronautFile2ArtMS(spectronautPeptideFile, 
                                     outFilePrefix = "./121223.APMS.FragPipeIn_data", 
                                     artmsConfig = cf)

# should we change to FragPipe??
globalInput$evidence_file$type %>%  unique()
```


```{r}
evidence <- read.table(file = "121223.APMS.FragPipeIn_data/evidence.txt", header = T, sep = "\t", stringsAsFactors = F, check.names = F)
evidence_sub <- evidence[-which(is.na(evidence$Intensity)), ] # remove rows with NA intensities


evidence_sub
# check contaminants
contaminate <- c("O77727", "P00698", "P00761", "P00883", "P02662", "P02663", "P02666", "P02668", "P02769")
# check Leading proteins formate

evidence.dt <-  data.table(evidence_sub)

# isloate the leading proteins(|*|) and append CON__ for contaminants
evidence.dt[, `Leading proteins` :=  str_extract(`Leading proteins`, '\\|[A-Z0-9]+\\|')]
evidence.dt[, `Leading proteins` := gsub('\\|', '', `Leading proteins`)]
evidence.dt[, `Leading proteins` := ifelse(grepl("contam_sp\\|", Proteins), paste0("CON__", `Leading proteins`), `Leading proteins`)]

#looks good
evidence.dt
write.table(evidence.dt, "121223.APMS.FragPipeIn_data/evidence_sub.txt", sep = "\t", row.names = F, quote = F)

```

```{r}
# QC
artmsQualityControlEvidenceBasic(evidence_file = "./121223.APMS.FragPipeIn_data/evidence_sub.txt", 
                                 keys_file = "./121223.APMS.FragPipeIn_data/keys.txt", 
                                 prot_exp = "APMS")
artmsQualityControlEvidenceExtended(evidence_file = "./121223.APMS.FragPipeIn_data/evidence_sub.txt", 
                                    keys_file = "./121223.APMS.FragPipeIn_data/keys.txt", 
                                    plotPCA = FALSE)

```

QC plots produced. just share these with Danielle for now in addition to the PCA plot

```{r}
# PCA
# MSstats
peptideDF <- fread(spectronautPeptideFile)
peptideDF$PeptideModifiedSequence <- peptideDF$PeptideSequence
peptideDF$IsotopeLabelType <- "L"

library(MSstats)


mssquant <- dataProcess(raw = as.data.frame(peptideDF),
                        MBimpute=F)


write.table(
  mssquant$FeatureLevelData,
  file = 'output/mss-FeatureLevelData.txt',
  eol = "\n",
  sep = "\t",
  quote = FALSE,
  row.names = FALSE,
  col.names = TRUE
)


write.table(
  mssquant$ProteinLevelData,
  file = 'output/mss-ProteinLevelData.txt',
  eol = "\n",
  sep = "\t",
  quote = FALSE,
  row.names = FALSE,
  col.names = TRUE
)
```

read in data and generate PCA from peptide IONs

```{r}
f.quant <- fread('./output/mss-FeatureLevelData.txt')

#normPepInt[, logCenteredIntensity := log2(INTENSITY/(median(INTENSITY, na.rm=TRUE))), by = PEPTIDE]

f.mat <- dcast(f.quant,  PEPTIDE~GROUP+SUBJECT, value.var = "INTENSITY") %>% 
  as.matrix(rownames='PEPTIDE')

f.mat <-  log2(f.mat)

f.mat <- f.mat[complete.cases(f.mat),]

pcaOut <- prcomp(t(f.mat))

colInfo <-  data.table(
  colname= colnames(f.mat)
)
colInfo[,c("bait","rep") := tstrsplit(colname, "[._]", keep = c(1,2)) ] # v useful fucntion]


#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)

#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = bait, shape=rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  theme(legend.title = element_text(size = 8), 
        legend.text = element_text(size = 6),
        legend.key.size =  unit(0.4, 'cm')) +
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  #scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_fill_manual(values=c(col.pal)) +
  scale_shape_manual(values=c(21:24)) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,#legend settings
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, 'PCA.complete.features.')
```
generate a PCA plot from the proteins

```{r}
p.quant <-  fread('./output/mss-ProteinLevelData.txt')

p.quant
#normPepInt[, logCenteredIntensity := log2(INTENSITY/(median(INTENSITY, na.rm=TRUE))), by = PEPTIDE]

p.mat <- dcast(p.quant,  Protein~GROUP+SUBJECT, value.var = "LogIntensities") %>% 
  as.matrix(rownames='Protein')


p.mat <- p.mat[complete.cases(p.mat),]

pcaOut <- prcomp(t(p.mat))

colInfo <-  data.table(
  colname= colnames(p.mat)
)
colInfo[,c("bait","rep") := tstrsplit(colname, "[._]", keep = c(1,2)) ] # v useful fucntion]


#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)

#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = bait, shape=rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  theme(legend.title = element_text(size = 8), 
        legend.text = element_text(size = 6),
        legend.key.size =  unit(0.4, 'cm')) +
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  #scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_fill_manual(values=c(col.pal)) +
  scale_shape_manual(values=c(21:24)) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,#legend settings
         color = guide_legend(override.aes = list(shape =21) ) )
p

BackupAsPDF(p, 'PCA.complete.proteins.')
```

13/12/23
---
Now that the QC run is complete, process the spec table 

```{r}
# timsTOF to SAINTexpress
spc <- data.table::fread(file = "./data/reprint.spc.tsv", stringsAsFactors = F)

spc.dt <- reshape2::melt(spc, id.vars= c("PROTID", "GENEID", "PROTLEN")) %>% 
  .[-1,] %>% 
  data.table()# rm first row 

spc.dt %>%  str()

spc.dt[, `:=`(Condition = str_extract(variable, '[A-Za-z1-9]+(?=_[123]{1}_SPC)'), # lookahead for _SPC
              Spectral.Count = as.integer(value),
              BioReplicate = str_extract(variable, '[123]{1}(?=_SPC)')  )]


#add biorep identifier to condtiion
spc.dt[, BioReplicate := paste0(Condition, '-', BioReplicate)]

#remove any contaminants
# use the evidence file to identify contaminations and screen spec.counts for this vector to remove
evidence <- read.table(file = "121223.APMS.FragPipeIn_data/evidence.txt", header = T, sep = "\t", stringsAsFactors = F, check.names = F) %>% 
  data.table()

# identify contaminants in the ms
contam.sp <- evidence %>%
  .[Proteins %like% 'contam', str_extract(Proteins,'(?<=\\|).+(?=\\|)')] %>% #lookahead and lookbehind to capture between | |
  unique()

# sanity check for remaining contam...
evidence[, uniprot := str_extract(Proteins,'(?<=\\|).+(?=\\|)')]
evidence[!uniprot %in%  contam.sp, grep('contam', Proteins)]

# no contaminant sp found.. cleaned??
spec.dt[PROTID %in% contam.sp,]

spec.dt
# dropping runs? Will wait for Danielle to make call.. include all for now
```
No runs dropped atm. proceed...


19-12-23
-----

preparing APMS controls for SAINT
Using samplesheet provided by Danielle; process the samples in two batches

```{r}
orf3a.grps <- c('Empty', 'Lempty', 'Lorf3aBeta', 'Lorf3aQ57H', 'Lorf3aS171L', 'Lorf3aWT', 
               'orf3aBeta', 'orf3aQ57H', 'orf3aS171L', 'orf3aWT', 'UT')

nsp6.gprs <- c('DemptyVector', 'Dnsp6BA1', 'Dnsp6BA2', 'Dnsp6WT', 'Dnsp6XBB', 'Empty',
               'nsp6BA1', 'nsp6BA2', 'nsp6WT', 'nsp6XBB', 'UT')

orf3a.spc.dt <-  spc.dt[Condition %in% orf3a.grps, ]
# sanity check
orf3a.spc.dt[, .N, by=.(Condition,BioReplicate)]


nsp6.spc.dt <-spc.dt[Condition %in% nsp6.gprs, ]
nsp6.spc.dt[, .N, by=.(Condition,BioReplicate)]


# add these to a list and process
saint.in.dt <- list(orf3a=orf3a.spc.dt,
                    nsp6=nsp6.spc.dt)


# control conditions
ctrl.grps <- c('DemptyVector', 'Empty', 'Lempty', 'UT')

makeBaitCtrl <- function(dt,ctrls){
  
  # Bait-CTRL table
  Bait <- unique(dt$BioReplicate)
  Bait <- gsub("-", "_", Bait)
  Bait <- sub("_([^_]*)$", "-\\1", Bait)
  Bait <- data.table(BioReplicate = Bait, Condition = gsub("-.*", "", Bait))
  Bait[, SAINT := "T"]
  Bait[Condition %in% ctrls, SAINT := 'C']
  #Bait$SAINT[Bait$Condition %in% ctrls] <- "C"
  
  return(Bait)
}


bait.tables <- lapply(saint.in.dt, function(x){
  x.dt <- makeBaitCtrl(x, ctrl.grps)
})

# wirte out the data
lapply(seq_along(bait.tables), function(x,n,i){
  
  fwrite(x[[i]], paste0('./output/', n[[i]], '.speccounts.saint.baits.tsv'), sep = "\t", quote = F, row.names = F, col.names = F)
  
}, x=bait.tables, n=names(bait.tables))

#looks good
#fread('./output/nsp6.speccounts.saint.baits.tsv', header=F)
#fread('./output/orf3a.speccounts.saint.baits.tsv', header=F)
```
create the prey table from the spec files

```{r}
makeInteractionTable <- function(dt){
  
  # interactions
  interactions <- data.table(BioReplicate = dt$BioReplicate, 
                           Condition = gsub("-.*", "", dt$BioReplicate), 
                           PROTID = dt$PROTID, 
                           spc = dt$value)
  interactions[, BioReplicate := gsub("-", "_", BioReplicate)]
  interactions[, BioReplicate := sub("_([^_]*)$", "-\\1", BioReplicate)]
  interactions[, Condition := gsub("-.*", "", BioReplicate)]
  # remove 0 counts
  interactions <- interactions[spc != 0 & PROTID != 'na', ]
  #write.table(interactions, "spectral_counts-saint-interactions.txt", sep = "\t", quote = F, row.names = F, col.names = F)
}


int.tables <- lapply(saint.in.dt, function(x){
  x.dt <- makeInteractionTable(x)
})

# looks good write to file
int.tables

# wirte out the data
lapply(seq_along(int.tables), function(x,n,i){
  
  fwrite(x[[i]], paste0('./output/', n[[i]], '.speccounts.saint.interactions.tsv'), sep = "\t", quote = F, row.names = F, col.names = F)
  
}, x=int.tables, n=names(int.tables))

# now create the prey table 

makePreyTable <-  function(dt){
  
  preys <- data.table(PROTID = dt$PROTID, 
                    PROTLEN = dt$PROTLEN, 
                    GENEID = dt$GENEID)
  
  preys <-  preys[PROTID != 'na',]
  preys <- unique(preys)
}


prey.tables <- lapply(saint.in.dt, function(x){
  x.dt <- makePreyTable(x)
})


# write out the data
lapply(seq_along(prey.tables), function(x,n,i){
  
  fwrite(x[[i]], paste0('./output/', n[[i]], '.speccounts.saint.preys.tsv'), sep = "\t", quote = F, row.names = F, col.names = F)
  
}, x=prey.tables, n=names(prey.tables))
```

create the same tables with the intensity data
-----
read in the intensity tables
```{r}
# intensity prep to SAINTexpress
spc <- fread(file = "./data/reprint.int.tsv", stringsAsFactors = F)

spc.dt <- reshape2::melt(spc, id.vars= c("PROTID", "GENEID")) %>% 
  .[-1,] %>% 
  data.table()# rm first row 

spc.dt %>%  str()

spc.dt[, `:=`(Condition = str_extract(variable, '[A-Za-z1-9]+(?=_[123]{1}_INT)'), # lookahead for _INT
              Intensity = as.integer(value),
              BioReplicate = str_extract(variable, '[123]{1}(?=_INT)')  )]


#add biorep identifier to condtiion
spc.dt[, BioReplicate := paste0(Condition, '-', BioReplicate)]

#remove any contaminants
# use the evidence file from above to identify contaminations and screen spec.counts for this vector to remove

# no contaminant sp found.
spc.dt[PROTID %in% contaminate,]
spc.dt[PROTID %in% contam.sp,]

```

probably actually dont need to recreate the condititon table as can reuse above..

```{r}

orf3a.spc.dt <-  spc.dt[Condition %in% orf3a.grps, ]
# sanity check
orf3a.spc.dt[, .N, by=.(Condition,BioReplicate)]


nsp6.spc.dt <-spc.dt[Condition %in% nsp6.gprs, ]
nsp6.spc.dt[, .N, by=.(Condition,BioReplicate)]


# add these to a list and process
saint.in.dt <- list(orf3a=orf3a.spc.dt,
                    nsp6=nsp6.spc.dt)


# control conditions
ctrl.grps <- c('DemptyVector', 'Empty', 'Lempty', 'UT')

bait.tables <- lapply(saint.in.dt, function(x){
  x.dt <- makeBaitCtrl(x, ctrl.grps)
})

# sanity check
bait.tables[[1]][, .N, .(BioReplicate,SAINT)] %>%  .[order(SAINT)]

# wirte out the data
lapply(seq_along(bait.tables), function(x,n,i){
  
  fwrite(x[[i]], paste0('./output/', n[[i]], '.ints.saint.baits.tsv'), sep = "\t", quote = F, row.names = F, col.names = F)
  
}, x=bait.tables, n=names(bait.tables))

#looks good
#fread('./output/nsp6.ints.saint.baits.tsv', header=F)
#fread('./output/orf3a.ints.saint.baits.tsv', header=F)
```
use same functions as above and just rename cols as necessary...
```{r}

int.tables <- lapply(saint.in.dt, function(x){
  x.dt <- makeInteractionTable(x)
})

# just rename the spc col to int to match Yuan processing
lapply(int.tables, function(x){
  setnames(x, old='spc', new='int')
  return(x)
})

# wirte out the data
lapply(seq_along(int.tables), function(x,n,i){
  
  fwrite(x[[i]], paste0('./output/', n[[i]], '.ints.saint.interactions.tsv'), sep = "\t", quote = F, row.names = F, col.names = F)
  
}, x=int.tables, n=names(int.tables))

# now create the prey table 
# wont cretae PROTLEN col as doesnt exist in intensity input dt
prey.tables <- lapply(saint.in.dt, function(x){
  x.dt <- makePreyTable(x)
})

# write out the data
lapply(seq_along(prey.tables), function(x,n,i){
  
  fwrite(x[[i]], paste0('./output/', n[[i]], '.ints.saint.preys.tsv'), sep = "\t", quote = F, row.names = F, col.names = F)
  
}, x=prey.tables, n=names(prey.tables))
```
Ok, have now created the necessary input files for SAINT processing.
create two subdirectories; one for ints data and one for spec counts and run saint for each
Can also try running the other tools afterwards if needed 

```{bash, eval=FALSE}
#mkdir -p ./output/{spc,ints}/{orf3a,nsp6}
#find ./output -type f -name "nsp6.ints.*" -exec mv {} ./output/ints/nsp6 \;
#find ./output -type f -name "orf3a.ints.*" -exec mv {} ./output/ints/orf3a \;
#find ./output -type f -name "nsp6.speccounts.*" -exec mv {} ./output/spc/nsp6 \;
#find ./output -type f -name "orf3a.speccounts.*" -exec mv {} ./output/spc/orf3a \;
```
check output before processing with SAINT
---
tables seem ok; lets run 


clone the docker image locally and run
```{bash}
#mkdir -p output/saintdocker
#wget https://raw.githubusercontent.com/kroganlab/BenPolacco/master/saintDocker/Dockerfile -O output/saintdocker/Dockerfile

#pull in saintExpress img
#docker pull brianyee/saintexpress:3.6.3
```

Run SAINT; 
Just share exact commands to run
```{bash runSAINT eval=F}
#cd ./output/ints/orf3a
#../../../../../utils/saintDocker/saintDocker.sh int orf3a.ints.saint.interactions.tsv orf3a.ints.saint.preys.tsv orf3a.ints.saint.baits.tsv
#cd -
#cd ./output/ints/nsp6
#../../../../../utils/saintDocker/saintDocker.sh int nsp6.ints.saint.interactions.tsv nsp6.ints.saint.preys.tsv nsp6.ints.saint.baits.tsv
#cd - 
#cd ./output/spc/orf3a
#../../../../../utils/saintDocker/saintDocker.sh spc orf3a.speccounts.saint.interactions.tsv orf3a.speccounts.saint.preys.tsv orf3a.speccounts.saint.baits.tsv
#cd ./output/spc/nsp6
#../../../../../utils/saintDocker/saintDocker.sh spc nsp6.speccounts.saint.interactions.tsv nsp6.speccounts.saint.preys.tsv nsp6.speccounts.saint.baits.tsv
```


now we have the `list.txt` file produced from each run

```{r}
#ints data
nsp6.ints <- fread('./output/ints/nsp6/list.txt')
orf3a.ints <- fread('./output/ints/orf3a/list.txt')

# spec data
nsp6.spc <- fread('./output/spc/nsp6/list.txt')
orf3a.spc <- fread('./output/spc/orf3a/list.txt')



saint.list <-  list(nsp6.ints = nsp6.ints,
                    orf3a.ints = orf3a.ints,
                    nsp6.spc = nsp6.spc,
                    orf3a.spc = orf3a.spc
                    )

saint.list
```
Ok, now that we have the SAINT output, look. at
histogram of the SAINT scores; looks like majority are low and some high scorers 

```{r}
lapply(saint.list, function(x){
  hist(x$TopoAvgP)
  })
```
```{r}
lapply(saint.list, function(x){
  summary(x)
  })
```


```{r}
filtered.saint.list <-  lapply(seq_along(saint.list), function(x,n,i){
  fwrite(x[[i]], paste0('./output/',n[[i]], '.saint.list.txt'))
  
  x.dt <-  x[[i]][BFDR < 0.05] 
  
  #fwrite(x.dt, paste0('./output/',n[[i]], '.saint.list.filtered.txt'))
  return(x.dt)
},x=saint.list, n=names(saint.list))

filtered.saint.list

# assign names
names(filtered.saint.list) <- names(saint.list)
```

Todo
---
bait protein intensity and spectral counts across samples. This would be for these 2 proteins:  nsp6 and orf3a
also, I would like to see a heatmap of the bait protein intensity and spectral counts across samples. This would be for these 2 proteins:  nsp6 and orf3a

need intensity and spectral count matrices and generate for the two proteins

```{r}
int.spec <- fread('./data/reprint.int.tsv')
spec.counts <- fread('./data/reprint.spc.tsv')


seqinr::read

int.spec[!GENEID %like% 'NSP',]

int.spec


fread('./data/-manifest')

int.spec[GENEID %like% c('ORF'),]

int.spec%>%  View()

spec.counts[PROTID %like% c('ORF3A'),]


mss <- fread('./data/MSstats.csv')

mss[ProteinName %like% 'P',]

mss


spec.counts[PROTID %like% 'P0DTF1']


# only 6 found
spec.counts[PROTID %like% 'P0D',]
int.spec[PROTID %like% 'P0D',]

# NA..
spec.counts[PROTID == 'A0A663DJA2']

prots.oi <- spec.counts[PROTID %in% c('P0DTC1', 'P0DTC3'),]$PROTID
prots.oi 
#plot for PODTC3 (3a) and  P0DTD1
# generate plots of all

int.mat <-  int.spec[-1,-2] %>% 
  as.matrix(rownames='PROTID')

spec.mat <- spec.counts[-1,-c(2,3)] %>% 
  as.matrix(rownames='PROTID') 


class(spec.mat) <- 'numeric'
class(int.mat) <- 'numeric'


subMat <- spec.mat[rownames(spec.mat) %in% prots.oi,]

# spec heatmap 
subMat <- spec.mat[rownames(spec.mat) %in% prots.oi,]
subMat


hm <- Heatmap(subMat, 
        na_col = 'grey',
        name='spectral counts',
        column_names_gp = gpar(fontsize=6),
        col = colorRamp2(breaks=c(0,10,20,30), colors = c('grey', 'pink', 'red', 'darkred')))

BackupAsPDF(draw(hm, column_title='Spectral counts of ORF3a (P0DTC3) & Replicase (P0DTC1)'), prefix='ORF3a.Replicase.spc.heatmap', dimensions=c(10,8))


#intensity heatmap

subMat <- int.mat[rownames(int.mat) %in% prots.oi,]
subMat <- log2(subMat+1)

subMat
hm <- Heatmap(subMat, 
        na_col = 'grey',
        name='log2(Ints + psuedocount1)',
        column_names_gp = gpar(fontsize=6),
        col = colorRamp2(breaks=c(0,25,28,32), c('grey', 'pink', 'red', 'darkred')) )

hm
BackupAsPDF(draw(hm, column_title='Intensity values of ORF3a (P0DTC3) & Replicase (P0DTC1)'), prefix='ORF3a.Replicase.ints.heatmap', dimensions=c(10,8))



hm <- Heatmap(subMat, 
        na_col = 'grey',
        name='log2(Ints + psuedocount1)',
        column_names_gp = gpar(fontsize=6),
        col = colorRamp2(breaks=c(0,5), c('grey','darkred')) )

hm

```

21-12-23
-----
Produce a venn diagram of the filtered hits from each condition to view the spec and intensitity data overlap

```{r}
filtered.saint.list

nsp6.f.list <- filtered.saint.list[c(1,3)]
orf3a.f.list <- filtered.saint.list[c(2,4)]

# create a venn diagram of these and also return the list of Venn Diagram hits
nsp6.f.list %>%  names()
orf3a.f.list

# format correclty; extract the bait and prey cols, add a ID col for the set and rbin
nsp6.v.dt <- lapply(seq_along(nsp6.f.list), function(x,n,i){
  
  x.dt  <- x[[i]][, .(Bait,Prey)]
  x.dt <-  x.dt[, ID := n[[i]]]
  return(x.dt)
  
}, x=nsp6.f.list, n=names(nsp6.f.list)) %>% rbindlist()


nsp6.v.dt

# format correclty; extract the bait and prey cols, add a ID col for the set and rbin
orf3a.v.dt <- lapply(seq_along(orf3a.f.list), function(x,n,i){
  
  x.dt  <- x[[i]][, .(Bait,Prey)]
  x.dt <-  x.dt[, ID := n[[i]]]
  return(x.dt)
  
}, x=orf3a.f.list, n=names(orf3a.f.list)) %>% rbindlist()



orf3a.v.dt
```

generate the venn diagrams of the data 

```{r}
g <- ggvenn(list('NSP6 Intensity' = nsp6.v.dt[ID == 'nsp6.ints', unique(Prey)],
                 'NSP6 Spec Counts' = nsp6.v.dt[ID == 'nsp6.spc', unique(Prey)]),
       fill_color = c("#0073C2FF", "#EFC000FF"),
       stroke_size = 0.5, set_name_size = 2) +# + position = position_nudge(x=-0.1)
       ggtitle('NSP6 Unique Prey overlaps')

g

BackupAsPDF(g, 'NSp6.allSigOverlaps.')  


g <- ggvenn(list('ORF3a Intensity'   = orf3a.v.dt[ID == 'orf3a.ints', unique(Prey)],
                 'ORF3a Spec Counts' = orf3a.v.dt[ID == 'orf3a.spc', unique(Prey)]),
       fill_color = c("#0073C2FF", "#EFC000FF"),
       stroke_size = 0.5, set_name_size = 2) +# + position = position_nudge(x=-0.1)
       ggtitle('ORF3a Unique Prey overlaps')
g

BackupAsPDF(g, 'ORF3a.allSigOverlaps.')  

```
look at overlaps within the individual sets
Have 8 sets in each; look at overlap for each subset

```{r}
nsp6.bait <- filtered.saint.list[[1]][, unique(Bait)]
orf3a.bait <- filtered.saint.list[[2]][, unique(Bait)]

# now have our list of baits for each; we want to loop through chr vector and plot a venn diagram for each subset to look at the overlap 
col.pal <- randomcoloR::distinctColorPalette(k=4)
  
  
for (i in orf3a.bait){

  # set names
  ints.set <- paste0(i,' Intensity')
  spec.set <-  paste0(i,' Spec Counts')
  
  #create list
  set.list <-  list(ints.set = orf3a.v.dt[ID == 'orf3a.ints' & Bait == i, Prey],
                    spec.set = orf3a.v.dt[ID == 'orf3a.spc' & Bait == i, Prey]
                    )
  
  print(set.list)
  
  g <- ggvenn(list(ints.set  = orf3a.v.dt[ID == 'orf3a.ints' & Bait == i, Prey],
                   spec.set = orf3a.v.dt[ID == 'orf3a.spc' & Bait == i, Prey]),
       fill_color = c(col.pal[1], col.pal[2]),
       stroke_size = 0.5, set_name_size = 2) +# + position = position_nudge(x=-0.1)
       ggtitle(paste0(i, ' Prey overlaps'))
  print(g)
  
  BackupAsPDF(g, paste0('orf3a.', i, '.SigOverlaps.'))
  
}


for (i in nsp6.bait){
  # set names
  ints.set <- paste0(i,' Intensity')
  spec.set <-  paste0(i,' Spec Counts')
  
  #create list
  set.list <-  list(ints.set = nsp6.v.dt[ID == 'nsp6.ints' & Bait == i, Prey],
                    spec.set = nsp6.v.dt[ID == 'nsp6.spc' & Bait == i, Prey]
                    )
  
  print(set.list)
  
  g <- ggvenn(list(ints.set  = nsp6.v.dt[ID == 'nsp6.ints' & Bait == i, Prey],
                   spec.set =  nsp6.v.dt[ID == 'nsp6.spc' & Bait == i, Prey]),
       fill_color = c(col.pal[3], col.pal[4]),
       stroke_size = 0.5, set_name_size = 2) +# + position = position_nudge(x=-0.1)
       ggtitle(paste0(i, ' Prey overlaps'))
  print(g)
  
  BackupAsPDF(g, paste0('nsp6.', i, '.SigOverlaps.'))
  
}

orf3a.v.dt
nsp6.v.dt
```
extract the overlapping genes from each of the sets

```{r}
# all sig overlapping
nsp6.all.overlaps <- intersect(nsp6.v.dt[ID == 'nsp6.ints', unique(Prey)],
          nsp6.v.dt[ID == 'nsp6.spc', unique(Prey)]
          ) %>% 
  data.table() %>% 
  .[, ID := 'NSP6.all.overlaps'] %>% 
  setnames(old='.', new='Protein')

# all sig overlapping
orf3a.all.overlaps <- intersect(orf3a.v.dt[ID == 'orf3a.ints', unique(Prey)],
          orf3a.v.dt[ID == 'orf3a.spc', unique(Prey)]
          ) %>% 
  data.table() %>% 
  .[, ID := 'ORF3A.all.overlaps'] %>% 
  setnames(old='.', new='Protein')


all.overlaps <- rbind(orf3a.all.overlaps, nsp6.all.overlaps)

all.overlaps[, gene := multiUniprots2multiGenes(Protein)]

fwrite(all.overlaps[,.(Protein, gene, ID)], ScriptAndDatedFileName('orf3a.nsp6.all.overlaps'))
```
extract overlaps from the subsets

```{r}
orfa.list <- for (i in orf3a.bait){

  # set names
  ints.set <- paste0(i,' Intensity')
  spec.set <-  paste0(i,' Spec Counts')
  
  
  dt.overlaps <- intersect(orf3a.v.dt[ID == 'orf3a.ints' & Bait == i,Prey],
                         orf3a.v.dt[ID == 'orf3a.spc' & Bait == i,Prey])

  dt.overlaps <- data.table(Protein = dt.overlaps,
                            Gene = multiUniprots2multiGenes(dt.overlaps),
                            ID = i)
  
}


# try lappy 

orf3a.overlaps <- lapply(orf3a.bait, function(i){
  
  dt.overlaps <- intersect(orf3a.v.dt[ID == 'orf3a.ints' & Bait == i,Prey],
                         orf3a.v.dt[ID == 'orf3a.spc' & Bait == i,Prey])

  dt.overlaps <- data.table(Protein = dt.overlaps,
                            ID = i)
  
  dt.overlaps[,Gene := multiUniprots2multiGenes(Protein)]
  
  return(dt.overlaps[, .(Protein,Gene, ID)])
  
}) %>%  rbindlist()


orf3a.overlaps


nsp6.overlaps <- lapply(nsp6.bait, function(i){
  
  dt.overlaps <- intersect(nsp6.v.dt[ID == 'nsp6.ints' & Bait == i, Prey],
                           nsp6.v.dt[ID == 'nsp6.spc' & Bait == i, Prey])
  
  print(dt.overlaps)

  dt.overlaps <- data.table(Protein = dt.overlaps,
                            ID = i)
  
  return(dt.overlaps[, .(Protein, ID)])
  
}) %>%  rbindlist()

nsp6.overlaps[, Gene := multiUniprots2multiGenes(Protein)]



fwrite(nsp6.overlaps[, .(Protein,Gene,ID)],  ScriptAndDatedFileName('nsp6.bait.overlaps'))
fwrite(orf3a.overlaps[, .(Protein,Gene,ID)],  ScriptAndDatedFileName('orf3a.bait.overlaps'))
```


