---
title: "240124.APMS.ContrastsofContrasts"
author: "Martin Gordon"
date: "2024-01-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Redo of contrasts of contrasts heatmap plots

Essentially, we want to collapse/summarise the columns (conditions/replicates), but still keep the plots the same, so no need to rerun the analysis


read in the processed p.quant data 
```{r}
dataDir <-'/Users/martingordon/Documents/projects/121123_DSwaney_APMSFirstRun/090124.APMS.contrastofcontrast_data/'

p.quant.nsp6 <- fread(paste0(dataDir, '2024_01_16_nsp6_ProteinLevelData.csv'))
p.quant.orf3a <- fread(paste0(dataDir, '2024_01_16_orf3a_ProteinLevelData.csv'))

# combine in list
p.quant.list <-  list(orf3a = p.quant.orf3a,
                      nsp6 = p.quant.nsp6)
```


read in the contrasts of contrasts data:
```{r}
dataDir <-'/Users/martingordon/Documents/projects/121123_DSwaney_APMSFirstRun/090124.APMS.contrastofcontrast_data/interaction.contrasts/'


nsp6.emm  <- fread(paste(dataDir,'2024_01_17_nsp6.emm.interactionContrasts.csv', sep='/'))
orf3a.emm <- fread(paste(dataDir,'2024_01_17_orf3a.emm.interactionContrasts.csv', sep='/'))

# read in the interaction data and convert to wide format
emm.interactions <- list(orf3a = orf3a.emm,
                         nsp6 = nsp6.emm)
```

Read in SAINT scoring output, we will take all prots passing APMS thresholds and plot

```{r}
nsp6.spc.saint <-  fread('./output/SAINT.scoring/nsp6.spc.saint.list.txt')
nsp6.int.saint <-  fread('./output/SAINT.scoring/nsp6.ints.saint.list.txt')
orf3a.spc.saint <-fread('./output/SAINT.scoring/orf3a.spc.saint.list.txt')
orf3a.int.saint <-fread('./output/SAINT.scoring/orf3a.ints.saint.list.txt')

saint.list <- list(nsp6.spc = nsp6.spc.saint,
                   nsp6.int = nsp6.int.saint,
                   orf3a.spc = orf3a.spc.saint,
                   orf3a.int = orf3a.int.saint)


lapply(saint.list, str)

f.saint.list <- lapply(saint.list, function(x){
  
  x.dt <- x[BFDR < 0.1, unique(Prey)]
  return(x.dt)
})

```

Just combine the ints and spec count edges and lets look at comparisons of the entire set
```{r}
nsp6.saint.fdr.01 <- unique(c(f.saint.list[[1]],f.saint.list[[2]]))
orf3a.saint.fdr.01 <- unique(c(f.saint.list[[3]],f.saint.list[[4]]))
```


read the emm.interactions data

```{r}
emm.interactions$orf3a <- emm.interactions$orf3a[Protein %in% orf3a.saint.fdr.01,]
emm.interactions$nsp6 <- emm.interactions$nsp6[Protein %in% nsp6.saint.fdr.01,]

lapply(emm.interactions, function(x){
  x[, fdr := p.adjust(p.value, method='BH'), by=Label]
})

```
# replot the data, this time using the z-score 
I think best if we calculate once per sample_group, collapse to condition (take mean or z-scores among replicates) then plot the sig hits identified in each interaction contrast


now that we have the matrix, calculate the z score and get the mean of each group. We will use this to plot all groups together
# finish tonight 
```{r}
lapply(p.quant.list, function(x){
  # calculate z-score per sample
  x[, z.score := (LogIntensities - mean(LogIntensities))/sd(LogIntensities), by=.(GROUP,SUBJECT) ]
  #now calculate avg z-score per condition
  x[, avg.z.score := mean(z.score), by=.(GROUP,Protein)]
})

lapply(p.quant.list, head)

p.mat.list <- lapply(p.quant.list, function(x){
  p.mat <- dcast(x[SUBJECT == 1,], Protein~GROUP, value.var = 'avg.z.score') %>% 
    as.matrix(rownames='Protein')
})
```

ok now that we have the mat, plot and filter the results

```{r}
rowClusterWithNA <- function(mat, na.value = 0, ...){
    mat[is.na(mat)] <- na.value
    hclust(dist(mat), ...)
  }

#play with thresholds here
lapply(seq_along(emm.interactions), function(x,n,y,i){
  
  bait <- n[[i]] 
  print(paste0('Working on: ', bait))
  
  # extract sig proteins in the interaction list
  sigProtein <- x[[bait]][fdr < 0.05 & Protein %in% y[[bait]] & !is.infinite(p.value),] %>% 
    .[,unique(Protein)]

  # filter just to our data
   prot.mat <- p.mat.list[[bait]][rownames(p.mat.list[[bait]]) %in% sigProtein,]

  # convert to gene names
  rownames(prot.mat) <- multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
  # median sweep
  #prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))
   #print(prot.mat[1:8,1:8])
  
  hm <- Heatmap(prot.mat, 
                cluster_columns = FALSE, 
                cluster_rows = rowClusterWithNA(prot.mat),
                name = 'avg z-score',
                show_row_names = T,
                #column_title = paste0(n[[i]]),
                row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(prot.mat)),
                row_names_gp = gpar(fontsize = 8),
                column_names_gp = gpar(fontsize = 8),#,
                column_split = ifelse(grepl("^[DL]{1}", colnames(prot.mat)), 'pre-treatment', "no pre-treatment") )
  
 #BackupAsPDF(draw(hm, column_title=paste0(bait, ' treatment interactions')), paste0(bait, 'interaction.contrasts.fdr<0.05.avg.zscores.heatmap'), dimensions = c(8,7))
  
},x=emm.interactions, y=saint.fdr1prc.list, n=names(emm.interactions))
```
plot the fdr < 0.1

```{r}
#play with thresholds here
lapply(seq_along(emm.interactions), function(x,n,y,i){
  
  bait <- n[[i]] 
  print(paste0('Working on: ', bait))
  
  # extract sig proteins in the interaction list
  sigProtein <- x[[bait]][fdr < 0.1 & Protein %in% y[[bait]] & !is.infinite(p.value),] %>% 
    .[,unique(Protein)]

  # filter just to our data
   prot.mat <- p.mat.list[[bait]][rownames(p.mat.list[[bait]]) %in% sigProtein,]

  # convert to gene names
  rownames(prot.mat) <- multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
  # median sweep
  #prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))
   #print(prot.mat[1:8,1:8])
  
  hm <- Heatmap(prot.mat, 
                cluster_columns = FALSE, 
                cluster_rows = rowClusterWithNA(prot.mat),
                name = 'avg z-score',
                show_row_names = T,
                #column_title = paste0(n[[i]]),
                row_title = sprintf('%s Proteins (fdr < 0.1)', nrow(prot.mat)),
                row_names_gp = gpar(fontsize = 8),
                column_names_gp = gpar(fontsize = 8),#,
                column_split = ifelse(grepl("^[DL]{1}", colnames(prot.mat)), 'pre-treatment', "no pre-treatment") )
  
 BackupAsPDF(draw(hm, column_title=paste0(bait, ' treatment interactions')), paste0(bait, 'interaction.contrasts.fdr<0.1.avg.zscores.heatmap'), dimensions = c(8,7))
  
},x=emm.interactions, y=saint.fdr1prc.list, n=names(emm.interactions))
```
make new pmat for use to match colnames of testing

```{r}
# create a pretreatment and condition col 
lapply(p.quant.list, function(x){
  x[, condition   :=  sub("^[LD]{1}", '', GROUP)]
  # tidy
  x[ condition == 'Empty' | condition == "emptyVector", condition := 'empty']
  x[, pretreat := ifelse(grepl("^[LD]{1}.*", GROUP), str_extract(GROUP, "^[LD]{1}"), 'Ctrl')] 
})


# create new matrix
p.mat.list <- lapply(p.quant.list, function(x){
  x.mat <- dcast (x[SUBJECT == 1,], Protein~condition+pretreat, value.var = "avg.z.score") %>% 
        as.matrix(rownames = "Protein")
})
```
plot each list of sig hits @ 0.1
```{r}
#play with thresholds here
lapply(seq_along(emm.interactions), function(x,y,n,i){
  
  bait <- n[[i]]

  # identify each contrast and iterate 
  contrasts.oi <-  x[[bait]][, unique(Label)]
  
  lapply(contrasts.oi, function(lab){
    
   
   # extract sig proteins in the interaction list
   sigProtein <- x[[bait]][Label == lab & fdr < 0.1 & Protein %in% y[[bait]] & !is.infinite(p.value),] %>% 
    .[,unique(Protein)]

   # only create heatmap if more than 1 
   if (length(sigProtein) > 1){
     
    # want to reduce the number of columns in each: try strsplit the label and extract the conditions, then look for something similiar in the data
    # convert the label to a regex pattern  to match matrix cols and and use to grep
    regex.pattern <- strsplit(lab, ' ') %>% 
      lapply(., function(x){
      paste( x[2], '_', x[5], '|', 
           x[4], '_', x[5], '|',
           x[9], '_', x[12],'|',
           x[11], '_', x[12],  sep='')
    })
       # filter just to our data
      prot.mat <- p.mat.list[[bait]][rownames(p.mat.list[[bait]]) %in% sigProtein,]
     
     
      # extract specific cols to the contrast
      prot.mat <- prot.mat[, grep(regex.pattern, colnames(prot.mat))]
    

      # convert to gene names
      rownames(prot.mat) <- multiUniprots2multiGenes(rownames(prot.mat), species = "HUMAN")
      # median sweep
      #prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))
  
      hm <- Heatmap(prot.mat, 
                cluster_columns = FALSE, 
                cluster_rows = rowClusterWithNA(prot.mat),
                name = 'zscore',
                show_row_names = T,
               #column_title = paste0(n[[i]], ' ', lab),
                row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(prot.mat)),
                row_names_gp = gpar(fontsize = 4),
                column_names_gp = gpar(fontsize = 8),#,
                column_split = ifelse(grepl("_[DL]{1}", colnames(prot.mat)), 'pre-treatment', "no pre-treatment") )
  

    BackupAsPDF(draw(hm, column_title=paste0(n[[i]], ' ', lab)), paste0(n[[i]], '.', lab, '.interaction.contrasts.fdr<0.1.heatmap'), dimensions = c(8,7))
   }
 })
},x=emm.interactions, y=saint.fdr1prc.list, n=names(emm.interactions))

regex.pattern
```

return to this tomorrow nd finish
----
need to consider a way of doing the z-score; does this make sense over all the data?
