---
title: "230124.APMS.pwContrasts"
author: "Martin Gordon"
date: "2024-01-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Rerun the pw comparisons for the APMS data using exclusively any genes that passed the SAINT scoring threshold

- Read in the SAINT output and filter to this list of genes
- Redo the PW comparisons 
- Also color the bait protein in each of the comparisons so we can see this is not moving

Reran with bait normalisation; unfortunately doesnt do a good job of normalising the prey proteins

25-01-24
----
Other normalisation techiques show its not too comparable 


```{r}
library (data.table)
library (ComplexHeatmap)
library (ggplot2)
library(readxl)
library(MSstats)
library(circlize)
library(ggvenn)
library (emmeans)
library(stringr)

source ("../../utils/bp_utils/spectronautFile2ArtMS.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source("/Users/martingordon/Documents/utils/bp_utils/LinearModels.R")


rotate.x.axis.text <- theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

clusterwNA <- function(x, na.val=0){
  
  x[is.na(x)] <- na.val
  dist(x) %>%  hclust(method='complete')
}
```

Read in SAINT scoring output, we will take all prots passing APMS thresholds and plot

```{r}
nsp6.spc.saint <-  fread('./output/SAINT.scoring/nsp6.spc.saint.list.txt')
nsp6.int.saint <-  fread('./output/SAINT.scoring/nsp6.ints.saint.list.txt')
orf3a.spc.saint <-fread('./output/SAINT.scoring/orf3a.spc.saint.list.txt')
orf3a.int.saint <-fread('./output/SAINT.scoring/orf3a.ints.saint.list.txt')

saint.list <- list(nsp6.spc = nsp6.spc.saint,
                   nsp6.int = nsp6.int.saint,
                   orf3a.spc = orf3a.spc.saint,
                   orf3a.int = orf3a.int.saint)


lapply(saint.list, str)

f.saint.list <- lapply(saint.list, function(x){
  
  x.dt <- x[BFDR < 0.1, unique(Prey)]
  return(x.dt)
})
```

Just combine the ints and spec count edges and lets look at comparisons of the entire set
```{r}
nsp6.saint.fdr.01 <- unique(c(f.saint.list[[1]],f.saint.list[[2]]))
orf3a.saint.fdr.01 <- unique(c(f.saint.list[[3]],f.saint.list[[4]]))

saint.fdr1prc.list <- list(orf3a = orf3a.saint.fdr.01,
                           nsp6 = nsp6.saint.fdr.01)
```

Read in the R object and perform the pw contrasts in msstats

```{r}
orf3a.mss <- readRDS("/Users/martingordon/Documents/projects/121123_DSwaney_APMSFirstRun/090124.APMS.contrastofcontrast_data/orf3a_mss.rds")
nsp6.mss <- readRDS("/Users/martingordon/Documents/projects/121123_DSwaney_APMSFirstRun/090124.APMS.contrastofcontrast_data/nsp6_mss.rds")

dp.out.list <-  list(orf3a=orf3a.mss,
                      nsp6=nsp6.mss)

lapply(dp.out.list, function(x){
  
  setDT(x$ProteinLevelData)
  setDT(x$FeatureLevelData)
  x$ProteinLevelData[,SUBJECT := interaction(GROUP,SUBJECT)] #for PW comparisonsgive each subject a unique label
  x$FeatureLevelData[,SUBJECT := interaction(GROUP,SUBJECT)]
  return(x)
})

lapply(dp.out.list,str)



lapply(seq_along(dp.out.list),  function(x,y,n,i){
  
  bait <- n[[i]]
  
  dt <- setDT(x[[bait]]$ProteinLevelData)
  
  #make new tables
  bait.dt <- dt[Protein == y[[bait]], .(GROUP,SUBJECT,LogIntensities)] %>% 
    .[,NAME := 'bait']
  gl.mean.dt <- dt[Protein != y[[bait]], mean(LogIntensities, na.rm=T), by=.(GROUP,SUBJECT)] %>% 
    .[,NAME := 'mean']
  gl.med.dt <-dt[Protein != y[[bait]], median(LogIntensities, na.rm=T), by=.(GROUP,SUBJECT)] %>% 
    .[,NAME := 'median']
  
  setnames(gl.mean.dt, old='V1', new='LogIntensities')
  setnames(gl.med.dt, old='V1', new='LogIntensities')
  
  p.dt <- rbind(bait.dt, gl.mean.dt, gl.med.dt)
  
  g <-  ggplot(p.dt, aes(x=reorder(interaction(GROUP,SUBJECT)), y=LogIntensities, color=NAME, group=NAME)) +
    geom_point() +
    geom_line() +
    ggtitle(bait) +
    theme_classic() +
    theme(axis.text.x = (element_text(angle=90)))

  BackupAsPDF(g, paste0(bait,'equilizeMedian.norm.linechart' ))
},x=dp.out.list, y=mssGS.list, n=names(dp.out.list))
```

subset both DTs to the list of sites significant in the SAINT scoring
Dont need! just subset afterwards...
```{r}
dp.out.list[['orf3a']]$ProteinLevelData
dp.out.list[['orf3a']]$FeatureLevelData


dp.out.list[['orf3a']]$ProteinLevelData <-  dp.out.list[['orf3a']]$ProteinLevelData[Protein %in% orf3a.saint.fdr.01,]
#dp.out.list[['orf3a']]$FeatureLevelData <-  dp.out.list[['orf3a']]$FeatureLevelData[PROTEIN %in% orf3a.saint.fdr.01,]
dp.out.list[['nsp6']]$ProteinLevelData <-  dp.out.list[['nsp6']]$ProteinLevelData[Protein %in% nsp6.saint.fdr.01,]
#dp.out.list[['nsp6']]$FeatureLevelData <-  dp.out.list[['nsp6']]$FeatureLevelData[PROTEIN %in% nsp6.saint.fdr.01,]
```

create the contrast matrices for both groups

```{r}
orf3a.contrasts <- makeContrast.AllByAll(dp.out.list$orf3a)
nsp6.contrasts  <- makeContrast.AllByAll(dp.out.list$nsp6)
```

'fix' the contrast lists by converting controls to denominators for relevant contrasts

```{r}
# orf3a
submat <- orf3a.contrasts[grep('^Lempty|^Empty', rownames(orf3a.contrasts)), ]
submat <-  submat*-1

old.names <- rownames(submat)
rev.names <- tstrsplit(rownames(submat), '-')

new.names <- sapply(seq_along(rev.names[[1]]), function(i){
  paste0(rev.names[[2]][i], '-', rev.names[[1]][i])
})
rownames(submat) <- new.names

orf3a.contrasts <- rbind(submat, orf3a.contrasts[!rownames(orf3a.contrasts) %in% old.names,])
orf3a.contrasts

#sanity check - should all = 0
apply(orf3a.contrasts, 1, sum)
```

## TODO convert this to a function to get contrast matrix inverse so can run easily

```{r}
# nsp6
submat <- nsp6.contrasts[grep('^DemptyVector|^Empty', rownames(nsp6.contrasts)), ]
submat <-  submat*-1

old.names <- rownames(submat)
rev.names <- tstrsplit(rownames(submat), '-')

new.names <- sapply(seq_along(rev.names[[1]]), function(i){
  paste0(rev.names[[2]][i], '-', rev.names[[1]][i])
})
rownames(submat) <- new.names

nsp6.contrasts <- rbind(submat, nsp6.contrasts[!rownames(nsp6.contrasts) %in% old.names,])
apply(nsp6.contrasts, 1, sum)

# sanity check 
nsp6.contrasts
```


```{r}
contrast.list <-  list(orf3a = orf3a.contrasts,
                       nsp6 = nsp6.contrasts)
```

run the PW comparisons on the reduced dataset


edit
---- 
For APMS, we are interested in perhaps sites with oneConditionMissing, but we want to ensure these are

```{r}

mss.out.list <- lapply(seq_along(dp.out.list), function(x,n,i){

  bait <- n[[i]]
  print(paste0('running comparison on ', bait))
  
  # run msstats correcting for batch 
  mss <- groupComparison(contrast.matrix=contrast.list[[bait]], data=x[[bait]])
  mss.dt <- setDT(mss$ComparisonResult)
  
  return(mss.dt)
  
},x=dp.out.list, n=names(dp.out.list))

names(mss.out.list) <- names(dp.out.list)

```
filter out the non-sig sites and count the number of DEGs within the SAINT sig hits

```{r}
# 
lapply(seq_along(mss.out.list), function(x,n,i){
  dt <- x[[i]]
  
   dt[,gene := multiUniprots2multiGenes(as.character(Protein))] %>% #needed as in factor format
  .[, p.adj := p.adjust(pvalue, method='BH'), by=.(Label)] %>% 
  .[, sig := 'not'] %>% 
  .[!issue %in% c('completeMissing', 'oneConditionMissing') & abs(log2FC) > 0.58 & p.adj < 0.05, sig := ifelse(log2FC > 0, 'up', 'down')] 
   fwrite(dt, paste0('./output/24.01.24.', n[[i]], '.mss.pwcomparisons.proc.csv'))
  }, x=mss.out.list, n=names(mss.out.list))

mss.out.list

lapply(mss.out.list, print)

# create another table to filter out to only sig hits in the 
f.mss.out.list <- lapply(seq_along(mss.out.list), function(x,y,n,i){
  
  bait <-  n[[i]]
  dt <- x[[bait]] %>% 
    .[Protein %in% y[[bait]], ] %>% 
    .[, p.adj := p.adjust(pvalue, method='BH'), by=.(Label)] %>%
    .[, sig := 'not'] %>% 
    .[!issue %in% c('completeMissing', 'oneConditionMissing') & abs(log2FC) > 0.58 & p.adj < 0.05, sig := ifelse(log2FC > 0, 'up', 'down')] 
   # fwrite(dt, paste0('./output/24.01.24.SAINTsigprots.',bait, '.mss.pwcomparisons.proc.csv'))
  
  return(dt)
  }, x=mss.out.list, y=saint.fdr1prc.list, n=names(mss.out.list))

```
Heatmap of one condition missing proteins
```{r}
lapply(seq_along(f.mss.out.list), function(x,n,i){
  
  bait <- n[[i]]  
  missing.prots <- (x[[bait]][issue == 'oneConditionMissing', unique(Protein)])
  
  print(head(dp.out.list[[bait]]$ProteinLevelData))
  
  p.mat <-  dcast(dp.out.list[[bait]]$ProteinLevelData, Protein~GROUP+SUBJECT, value.var = 'LogIntensities') %>% 
    as.matrix(rownames='Protein')


  submat <-  p.mat[rownames(p.mat) %in% missing.prots,]
  
  rownames(submat) <- multiUniprots2multiGenes(rownames(submat))
  
  # just convert to 0 now
  submat[is.na(submat)] <- 0
  
  
  print(submat)
  
  Heatmap(submat, cluster_rows = T, na_col='grey' )
  
},x=f.mss.out.list, n=names(f.mss.out.list))

names(f.mss.out.list) <-names(mss.out.list)
```
generateVolcano plots of all the comparisons
Over 100 in total so dont rerun....
also highlight the bait 

```{r}
dp.out.list

prots.oi <- c("P0DTC1","P0DTC3")

lapply(seq_along(f.mss.out.list), function(x,n,i){
  dt <- x[[i]]
  
  # vecotr to iterate through
  contrasts.oi <- dt$Label %>%  unique()
  
  # nested loop w 
  lapply(contrasts.oi, function(cont){
    
      #print(dt[Label == cont & issue %in% c('oneConditionMissing')])
      g <- ggplot(dt[Label == cont & !issue %in% c('completeMissing'),], aes(x=log2FC, y=-log10(p.adj), color=sig, label=gene)) + 
      geom_point() +
      geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
      # add bait to plot
      geom_point(data=dt[Label == cont & Protein %in% prots.oi,], aes(x=log2FC, y=-log10(p.adj), color='bait')) +
      ggrepel::geom_text_repel(data=dt[Label == cont & Protein %in% prots.oi,],  size = 2, max.overlaps = 20, segment.color = 'grey80', colour = "green") +
      geom_hline(yintercept = -log10(0.05), linetype=2, alpha=0.4) +
      ggrepel::geom_text_repel(data=dt[Label == cont & sig != 'not' & !Protein %in% prots.oi,],  size = 2, max.overlaps = 20) +
      scale_color_manual(values=c('up'='red', 'down'='blue', 'not'='grey', 'bait'='green')) +
      ggtitle(paste0(n[[i]], ' ', cont, ' (FC +/ 50% & p.adj < 0.05)')) +
      # facet_wrap(~Label, scales = 'free') +
      theme_bw() +
      theme(strip.text.x = element_text(size = 5))
      
      BackupAsPDF(g, paste0(n[[i]],'.',cont,'.SAINT.bfdr0.1.volcanoplot.'))
    
  })
  
  }, x=f.mss.out.list, n=names(f.mss.out.list))

names(f.mss.out.list)
```

