---
title: "26.01.24"
author: "Martin Gordon"
date: "2024-01-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

First lets see how data looks when combined

```{r}
library (data.table)
library (ComplexHeatmap)
library (ggplot2)
library(readxl)
library(MSstats)
library(circlize)
library(ggvenn)
library (emmeans)
library(stringr)
library(patchwork) # combine plots

source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/spectronautFile2ArtMS.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source("/Users/martingordon/Documents/utils/bp_utils/LinearModels.R")
source ("../../utils/bp_utils/MSstats_Helper_Functions.R")


rotate.x.axis.text <- theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))
```


## em normalisation first, then 
-----
```{r}
dataDir <- '/Users/martingordon/Documents/projects/121123_DSwaney_APMSFirstRun/data/'

```

run tmp, subtract the column offset for each sample to 'normalise',  then run TMP normalising by global standards
```{r}
spec <- fread(paste(dataDir, 'MSstats.csv', sep=''))
#remove contaminant proteins
spec <- spec[!ProteinName %like% 'contam_sp',]

spec[, group_subject := paste(Condition, BioReplicate,sep='_')]

# lets create a flag for pre-treatment; will use this to split the datasets
spec %>%  str()
spec[, treatment := ifelse(grepl("^[DL]{1}", Condition), 'treatment', 'control') ]
spec[, condition := sub("^[DL]{1}","",Condition)]

#sanity check
spec[, .N, by=.(Condition,treatment, condition)]
```
seperate the two bait groups into two treatment groups

```{r}
orf3a.grps <- c('Lorf3aBeta', 'Lorf3aQ57H', 'Lorf3aS171L', 'Lorf3aWT', 
               'orf3aBeta', 'orf3aQ57H', 'orf3aS171L', 'orf3aWT')

nsp6.gprs <- c('Dnsp6BA1', 'Dnsp6BA2', 'Dnsp6WT', 'Dnsp6XBB',
               'nsp6BA1', 'nsp6BA2', 'nsp6WT', 'nsp6XBB')


spec.orf3a <- spec[Condition %in% orf3a.grps,]
spec.nsp6 <- spec[Condition %in% nsp6.gprs,]


spec.list <-  list(orf3a = spec.orf3a,
                   nsp6 = spec.nsp6)

```

```{r}
mssGS.list <-  list(orf3a ='P0DTC3',
                    nsp6 ='P0DTC1')


names(mssGS.list) == names(spec.list)
```

```{r}
mssInput.list <- lapply(seq_along(spec.list), function(x,n,i){
  
  # iterate through list grp
  group <- n[[i]]
  
  # tidy the protein names
  dt <- x[[group]][, uniprot := str_extract(ProteinName,'(?<=\\|).+(?=\\|)')]
  dt[, ProteinName := uniprot]

  # configure for MSstats input
  dt[, PeptideModifiedSequence := PeptideSequence]
  mssInput <- specFileToCompleteMSstats(dt)
  mssInput[, IsotopeLabelType := 'L']
  
  return(mssInput)
},x=spec.list, n=names(spec.list))

names(mssInput.list) <-  names(spec.list)

mssInput.list
```

```{r}
lapply(seq_along(mssInput.list),  function(x,n,i){
g <-  ggplot(x[[i]], aes(x=group_subject, y=log2(Intensity), fill=Condition)) +
    geom_boxplot() +
    ggtitle(n[[i]]) +
    theme(axis.text.x = (element_text(angle=90)))
    BackupAsPDF(g, paste0(n[[i]], '.raw.intensities.boxplots.'))
},x=mssInput.list, n=names(mssInput.list))
```

bait peptide vs global raw intensities
```{r}
lapply(seq_along(mssInput.list),  function(x,y,n,i){
  
  bait <- n[[i]]
  dt <- x[[bait]]
  print(dt)
  
  #make new tables
  bait.dt <- dt[ProteinName == y[[bait]], .(Condition,BioReplicate,Intensity)] %>% 
  .[, mean(Intensity, na.rm=T), by=.(Condition,BioReplicate)] %>% 
    .[,NAME := 'mean_bait']
   med.bait.dt <- dt[ProteinName == y[[bait]], .(Condition,BioReplicate,Intensity)] %>% 
  .[, median(Intensity, na.rm=T), by=.(Condition,BioReplicate)] %>% 
    .[,NAME := 'median_bait']
  
  gl.mean.dt <- dt[ProteinName != y[[bait]], mean(Intensity, na.rm=T), by=.(Condition,BioReplicate)] %>% 
    .[,NAME := 'mean']
  gl.med.dt <-dt[ProteinName != y[[bait]], median(Intensity, na.rm=T), by=.(Condition,BioReplicate)] %>% 
    .[,NAME := 'median']
  
  setnames(med.bait.dt, old='V1', new='Intensity')
  setnames(bait.dt, old='V1', new='Intensity')
  setnames(gl.mean.dt, old='V1', new='Intensity')
  setnames(gl.med.dt, old='V1', new='Intensity')
  
  p.dt <- rbind(bait.dt,med.bait.dt, gl.mean.dt, gl.med.dt)
  
  g <-  ggplot(p.dt, aes(x=reorder(interaction(Condition,BioReplicate)), y=log2(Intensity), color=NAME, group=NAME)) +
    geom_point() +
    geom_line() +
    ggtitle(paste0(bait, ' raw intensities')) +
    theme_classic() +
    theme(axis.text.x = (element_text(angle=90)))

  BackupAsPDF(g, paste0(bait,'raw.intensities.linechart'))
},x=mssInput.list, y=mssGS.list, n=names(mssInput.list))
```

```{r}
mss.em.list <-  lapply(seq_along(mssInput.list), function(x,i){
  
  mss.x <-  dataProcess(x[[i]], 
                        normalization = 'EQUALIZEMEDIANS', #preran with globalStandards and didnt work well...
                        summaryMethod="TMP",
                        remove_uninformative_feature_outlier=TRUE,
                        featureSubset="highQuality",
                        MBimpute =  FALSE)
 return(mss.x)
  
}, x=mssInput.list)

names(mss.em.list) <- names(mssInput.list)
```

```{r}

lapply(seq_along(mss.em.list),  function(x,y,n,i){
  
  group <- n[[i]]
  dt <- setDT(x[[group]]$ProteinLevelData)
  
  #make new tables
  bait.dt <- dt[Protein == y[[group]], .(GROUP,SUBJECT,LogIntensities)] %>% 
    .[,NAME := 'bait']
  
  gl.mean.dt <- dt[Protein != y[[group]], mean(LogIntensities, na.rm=T), by=.(GROUP,SUBJECT)] %>% 
    .[,NAME := 'global_mean']
  gl.med.dt <-dt[Protein != y[[group]], median(LogIntensities, na.rm=T), by=.(GROUP,SUBJECT)] %>% 
    .[,NAME := 'global_median']
  
  setnames(gl.mean.dt, old='V1', new='LogIntensities')
  setnames(gl.med.dt, old='V1', new='LogIntensities')
  
  p.dt <- rbind(bait.dt,gl.mean.dt, gl.med.dt)
  
  g <-  ggplot(p.dt, aes(x=reorder(interaction(GROUP,SUBJECT)), y=LogIntensities, color=NAME, group=NAME)) +
    geom_point() +
    geom_line() +
    ggtitle(paste0(group, ' equilize median normalization')) +
    theme_classic() +
    theme(axis.text.x = (element_text(angle=90)))

  BackupAsPDF(g, paste0(group,'equilizeMedian.intensities.linechart'))
},x=mss.em.list, y=mssGS.list, n=names(mss.em.list))
```

now return to the mssInput data.table and run with global standard normalisation

```{r}
mss.gs.list <-  lapply(seq_along(mssInput.list), function(x,y,i){
  
  bait <- y[[i]]
  
  mss.x <-  dataProcess(x[[i]], 
                        normalization = 'globalStandards', #preran with globalStandards and didnt work well...
                        nameStandards = bait, # name of protein to use for standards
                        summaryMethod="TMP",
                        remove_uninformative_feature_outlier=TRUE,
                        featureSubset="highQuality",
                        MBimpute =  FALSE)
 return(mss.x)
  
}, x=mssInput.list, y=mssGS.list)

names(mss.gs.list) <-  names(mssInput.list)
```
      
```{r}
lapply(seq_along(mss.gs.list),  function(x,y,n,i){
  
  group <- n[[i]]
  dt <- setDT(x[[group]]$ProteinLevelData)
  
  #make new tables
  bait.dt <- dt[Protein == y[[group]], .(GROUP,SUBJECT,LogIntensities)] %>% 
  #.[, LogIntensities, na.rm=T), by=.(GROUP,SUBJECT)] %>% 
    .[,NAME := 'bait']
  
  gl.mean.dt <- dt[Protein != y[[group]], mean(LogIntensities, na.rm=T), by=.(GROUP,SUBJECT)] %>% 
    .[,NAME := 'global_mean']
  gl.med.dt <-dt[Protein != y[[group]], median(LogIntensities, na.rm=T), by=.(GROUP,SUBJECT)] %>% 
    .[,NAME := 'global_median']
  
  setnames(gl.mean.dt, old='V1', new='LogIntensities')
  setnames(gl.med.dt, old='V1', new='LogIntensities')
  
  p.dt <- rbind(bait.dt,gl.mean.dt, gl.med.dt)
  
  g <-  ggplot(p.dt, aes(x=reorder(interaction(GROUP,SUBJECT)), y=LogIntensities, color=NAME, group=NAME)) +
    geom_point() +
    geom_line() +
    ggtitle(paste0(group, ' global standards (bait) normalization')) +
    theme_classic() +
    theme(axis.text.x = (element_text(angle=90)))

  BackupAsPDF(g, paste0(group,'globalStandards.intensities.linechart'))
},x=mss.gs.list, y=mssGS.list, n=names(mss.gs.list))
```


T-test to assess if FC estimates between each of the conditions

First need to check if the data meets assumptions
- Are the values (log2FC) normally distributed? Will need to look within each WT-* comparison and SAINT subset
- Are the variances uniform?

lets just try first with equilizeMedians...


read in all data

```{r}
outdir <-  paste0(getwd(), '/output/')

# equilizemedians
em.mss.list <- list(orf3a_control = fread(paste0(outdir, '26.01.24.orf3a_control.batch.mss.pwcomparisons.csv')),
                       orf3a_treatment = fread(paste0(outdir, '26.01.24.orf3a_treatment.batch.mss.pwcomparisons.csv')),
                       nsp6_control = fread(paste0(outdir, '26.01.24.nsp6_control.batch.mss.pwcomparisons.csv')),
                       nsp6_treatment =fread(paste0(outdir, '26.01.24.nsp6_treatment.batch.mss.pwcomparisons.csv')) )

# equilizemedians
gs.mss.list <- list(orf3a_control = fread(paste0(outdir, '26.01.24.orf3a_control.batch.gsnorm.mss.pwcomparisons.csv')),
                       orf3a_treatment = fread(paste0(outdir, '26.01.24.orf3a_treatment.batch.gsnorm.mss.pwcomparisons.csv')),
                       nsp6_control = fread(paste0(outdir, '26.01.24.nsp6_control.batch.gsnorm.mss.pwcomparisons.csv')),
                       nsp6_treatment =fread(paste0(outdir, '26.01.24.nsp6_treatment.batch.gsnorm.mss.pwcomparisons.csv')) )


# filter these just to the comparisons we want 
```

Q. are the log2FC values distributed normally?
The superset is 

```{r}
lapply(em.mss.list, function(x){
  
  for(i in unique(x$Label)){
    hist(x[Label == i,log2FC])
  }
})
```
what about the subset of SAINT interactors?
-----

Read in SAINT scoring output, we will take all prots passing APMS thresholds and plot

```{r}
nsp6.spc.saint <-  fread('./output/SAINT.scoring/nsp6.spc.saint.list.txt')
nsp6.int.saint <-  fread('./output/SAINT.scoring/nsp6.ints.saint.list.txt')
orf3a.spc.saint <-fread('./output/SAINT.scoring/orf3a.spc.saint.list.txt')
orf3a.int.saint <-fread('./output/SAINT.scoring/orf3a.ints.saint.list.txt')

saint.list <- list(nsp6.spc = nsp6.spc.saint,
                   nsp6.int = nsp6.int.saint,
                   orf3a.spc = orf3a.spc.saint,
                   orf3a.int = orf3a.int.saint)


lapply(saint.list, str)

f.saint.list <- lapply(saint.list, function(x){
  
  x.dt <- x[BFDR < 0.1, unique(Prey)]
  return(x.dt)
})

nsp6.saint.fdr.01 <- unique(c(f.saint.list[[1]],f.saint.list[[2]]))
orf3a.saint.fdr.01 <- unique(c(f.saint.list[[3]],f.saint.list[[4]]))

saint.fdr1prc.list <- list(orf3a_control = orf3a.saint.fdr.01,
                           orf3a_treatment = orf3a.saint.fdr.01,
                           nsp6_control= nsp6.saint.fdr.01,
                           nsp6_treatment= nsp6.saint.fdr.01)
```

subset of label log2FC not normally distributed.. makes sense also with normalisation issues

```{r}
lapply(seq_along(em.mss.list), function(i, x, y) {
  for (l in unique(x[[i]]$Label)) {
      dt <- x[[i]][x[[i]]$Label == l & x[[i]]$Protein %in% y[[i]], ]
      print(ggplot(dt, aes(log2FC)) +
        geom_histogram())
  }
}, x = em.mss.list, y = saint.fdr1prc.list)
```

other assumption; need to assess the variance; what does the standard deviation look like for each comparison?
esitmate SD: SE * sqrt(N)
add N coefficient estimates to DF

```{r}
summary(em.mss.list[[1]])

lapply(seq_along(em.mss.list), function(i, x, y) {
    x[[i]][, N := DF + 4] # N mean estimates
    x[[i]][, SD :=  SE * sqrt(N)] # N mean estimates
}, x = em.mss.list, y = saint.fdr1prc.list)
```

Stack the dts and look at the SDs side by side

```{r}
em.mss.list[[1]][, status := 'control']
em.mss.list[[2]][, status := 'treatment']
em.mss.list[[3]][, status := 'control']
em.mss.list[[4]][, status := 'treatment']

em.mss.merge <- list(orf3a=rbind(em.mss.list[[1]], em.mss.list[[2]]),
                     nsp6=rbind(em.mss.list[[3]], em.mss.list[[4]])
                     )


lapply(seq_along(em.mss.list), function(i, x, y) {
  for (l in unique(x[[i]]$Label)) {
      dt <- x[[i]][x[[i]]$Label == l & x[[i]]$Protein %in% y[[i]], ]
      print(ggplot(dt, aes(SD)) +
        geom_histogram())
  }
}, x = em.mss.list, y = saint.fdr1prc.list)
```

```{r}
# create a pretreatment col 
lapply(em.mss.list, function(x){
  x[, condition   :=  ifelse(grepl("^[LD]{1}", Label), 'treated', 'control')]
  x[, contrast := sub("^[LD]{1}", "", Label)]
  x[, contrast := str_replace(contrast, "(?<=-)[LD]", "")]
})

# view 
lapply(em.mss.list, function(x){
  x[,.N,  by=.(Label,condition,contrast)]
})

# merge the two datatables
orf3a.dt <- merge(x=em.mss.list[[1]], y=em.mss.list[[2]], by=c('Protein','contrast'), suffixes = c(".ctrl", ".treat"))
nsp6.dt <-  merge(x=em.mss.list[[3]], y=em.mss.list[[4]], by=c('Protein','contrast'), suffixes = c(".ctrl", ".treat"))

em.list <- list(orf3a = orf3a.dt,
                nsp6 = nsp6.dt)
```


Following msstatsPTM methodology:
Now we want to 
i) subtract the log2FC estimates from each other
ii) recalculate pooled SE and DF
iii) delta FC/pooled SE
iv) compare to Student dist with DF from ii)
v) multtiple testing correction

```{r}
lapply(em.list, function(x){

  x[, delta_log2FC := log2FC.treat - log2FC.ctrl]
  # Assuming that the sources of variation independent..
  x[, delta_SE := sqrt(SE.treat^2 + SE.ctrl^2)]
  
  # satherwaite approximation for DF estimate see msstatsPTM...
  x[ ,delta_df := (((SE.treat^2) + (SE.ctrl^2))^2) / (((SE.treat^4)/DF.treat) + ((SE.ctrl^4)/DF.ctrl)) ] 
  
  x[, delta_tstat := delta_log2FC/delta_SE]
  x[, delta_pval :=2 * pt(-abs(delta_tstat), delta_df)]
  x[, delta_fdr := p.adjust(delta_pval, method='BH')]
})


# looks to be many hits, lets see the 
lapply(em.list, function(x){
  hist(x$delta_log2FC)
  hist(x$delta_pval)
})
```
Lets calculate it the other way and compare the results

```{r}
# dont use; think this m
lapply(em.list, function(x){

  x[, delta_log2FC := log2FC.treat - log2FC.ctrl]
  x[, delta2_SE := sqrt(SE.treat^2 / 6 + SE.ctrl^2 / 6)]
  x[, delta2_DF := 6 + 6 - 2 ]
  x[, delta2_tstat := delta_log2FC/delta2_SE ]
  x[, delta2_pval := 2 * pt(-abs(delta2_tstat), delta2_DF) ]
  x[, delta2_fdr := p.adjust(delta2_pval, method='BH')]
})


lapply(em.list,function(x){
  
  ggplot(x, aes(x=delta_pval, y=delta2_pval)) +
    geom_point()
})


lapply(em.list,function(x){
  

  x[, sig := 'not']
  x[, sig2 := 'not']
  x[abs(delta_log2FC) > 0.58 & delta_pval < 0.005, sig := ifelse(delta_log2FC > 0, 'up', 'down')]
   x[abs(delta_log2FC) > 0.58 & delta2_fdr < 0.05, sig2 := ifelse(delta_log2FC > 0, 'up', 'down')]
})


lapply(em.list,function(x){
 x[,.N, by=.(sig)]
 #x[sig != 'not',.N, by=.(sig,contrast)] 
 #x[,.N, by=.(sig2)] 
})
```
extract the list of sig hits and lets plot for each heatmap
Abit difficult to merge dts as processed seperately so scaled differently...  
only sensible way to visualise is z-scores due to batch issues

```{r}
p.quant.list <- list(orf3a = rbind(mss.em.list[[1]]$ProteinLevelData),
                     nsp6 =  rbind(mss.em.list[[2]]$ProteinLevelData))

clusterwNA <- function(x, na.val=0){
  
  x[is.na(x)] <- na.val
  dist(x) %>%  hclust(method='complete')
}

lapply(seq_along(em.list), function(x,y, n,i){
  
   grp <- n[[i]]
   print(grp)
   
   sigProts <- x[[i]][sig != 'not', unique(Protein)]
   
   p.mat <-  data.table(y[[i]]$ProteinLevelData) %>% 
    dcast(., Protein~paste0(GROUP,'.',SUBJECT), value.var = 'LogIntensities') %>% 
    as.matrix(rownames='Protein')
   
   # z scores of heatmap
   p.mat <-  scale(p.mat)
   #p.mat <- sweep(p.mat, 1, STATS=apply(p.mat,1,median, na.rm=T))
   
   subMat <- p.mat[rownames(p.mat) %in% sigProts,]
   
  rownames(subMat) <-  multiUniprots2multiGenes(rownames(subMat))
   
  #heatmap annotation 
  grp <- sub(".[1-6]$", "", colnames(subMat))
  grp <- sub("^[LD]{1}", "", grp)
  grp <-  str_replace(grp, "(?<=-)[LD]", "")
  treat <- ifelse(grepl("^[D|L]{1}", colnames(subMat)), 'treatment', 'control')
  rep <- sub(".","",str_extract(colnames(subMat), ".[1-3]$"))

  col <-  colorRamp2(c(-3,0,3), c("blue","white","red"))
  
  ha <- HeatmapAnnotation(group = grp, rep=rep, col=list(
                                                                          group=c('orf3aBeta'='red', 'orf3aQ57H'='blue', 'orf3aS171L'='green', 'orf3aWT'='orange',
                                                                           'nsp6BA1'='yellow', 'nsp6BA2'='purple', 'nsp6XBB'='pink', 'nsp6WT'='darkred'),
                                                                              rep=c('1'='pink','2'='cyan', '3'='palegreen')))
  
  hm <-  Heatmap(subMat, 
          cluster_rows = clusterwNA(subMat),
          show_row_names = T,
          cluster_columns = F,
          row_names_gp = gpar(fontsize=3),
          col=col,
          name = 'z-score (Ints.)',
          row_title = paste0(nrow(subMat), ' proteins +/- 50% FC & p.val < 0.005'),
          column_split = ifelse(grepl('^[LD]{1}', colnames(subMat)), 'treatment', 'control'),
         top_annotation = ha)


  BackupAsPDF(draw(hm, column_title=paste0(n[[i]], ' DE between treatment groups')), paste0(n[[i]], '.contrastofcontrast.heatmap.'),  dimensions=c(10,11))
  
}, x=em.list, y=mss.em.list, n=names(em.list))
```
collapse to group mean
Some issue here with accessing address for the DT?

```{r}
lapply(p.quant.list,str)

lapply(p.quant.list, function(x){
  return(setDT(x))
})

p.quant.list <- lapply(p.quant.list, function(x,i){
  # calculate z-score per sample
  x[, z.score := (LogIntensities - mean(LogIntensities))/sd(LogIntensities), by=.(GROUP,SUBJECT) ]
  #now calculate avg z-score per condition
  x[, avg.z.score := mean(z.score), by=.(GROUP,Protein)]

})

p.mat.list <- lapply(seq_along(p.quant.list), function(x,i){
  p.mat <- dcast(x[[i]][SUBJECT == 1,], Protein~GROUP, value.var = 'avg.z.score') %>% 
    as.matrix(rownames='Protein')
}, x=p.quant.list)

names(p.mat.list) <- names(p.quant.list)
```
plot the average z score

```{r}
lapply(seq_along(em.list), function(x,y, n,i){
  
   grp <- n[[i]]
   print(grp)
   
   sigProts <- x[[i]][sig != 'not', unique(Protein)]
   
   p.mat <-  y[[i]]
   
   # z scores of heatmap
   p.mat <-  scale(p.mat)
   #p.mat <- sweep(p.mat, 1, STATS=apply(p.mat,1,median, na.rm=T))
   
  subMat <- p.mat[rownames(p.mat) %in% sigProts,]
  rownames(subMat) <- multiUniprots2multiGenes(rownames(subMat))
   
  #heatmap annotation 
  #grp <- sub(".[1-6]$", "", colnames(subMat))
  #grp <- sub("^[LD]{1}", "", grp)
  #grp <-  str_replace(grp, "(?<=-)[LD]", "")
  treat <- ifelse(grepl("^[D|L]{1}", colnames(subMat)), 'treatment', 'control')
  rep <- sub(".","",str_extract(colnames(subMat), ".[1-3]$"))

  col <-  colorRamp2(c(-3,0,3), c("blue","white","red"))
  
  ha <- HeatmapAnnotation(treatment=treat,  col=list(treatment=c('treatment'='red', 'control'='blue'),
                                                                          group=c('orf3aBeta'='red', 'orf3aQ57H'='blue', 'orf3aS171L'='green', 'orf3aWT'='orange',
                                                                           'nsp6BA1'='yellow', 'nsp6BA2'='purple', 'nsp6XBB'='pink', 'nsp6WT'='darkred', 'nsp6B'='lightblue'
                                                                             )) )
  hm <-  Heatmap(subMat, 
          cluster_rows = clusterwNA(subMat),
          show_row_names = T,
          row_names_gp = gpar(fontsize=3),
          col=col,
          name = 'mean z-score (Ints.)',
          row_title = sprintf('%s proteins', nrow(subMat)),
          column_split = ifelse(grepl('^[LD]{1}', colnames(subMat)), 'treatment', 'control'),
         top_annotation = ha)


  #BackupAsPDF(draw(hm, column_title=paste0(n[[i]], ' DE between treatment groups')), paste0(n[[i]], '.contrastofcontrast.collapsed.heatmap.'),  dimensions=c(8,13))
  
}, x=em.list, y=p.mat.list, n=names(em.list))


p.mat.list
```
now loop through each of the contrasts

```{r}
lapply(seq_along(em.list), function(x,y, n,i){
  
   grp <- n[[i]]

   
   for (lab in unique(x[[grp]]$contrast)){
     print(lab)
    
     sigProts <- x[[i]][sig != 'not' & contrast == lab, unique(Protein)]
     #print(sigProts)
   
      p.mat <-  data.table(y[[i]]$ProteinLevelData) %>% 
        dcast(., Protein~paste0(GROUP,'.',SUBJECT), value.var = 'LogIntensities') %>% 
        as.matrix(rownames='Protein')
   
      # z scores of heatmap
      p.mat <-  scale(p.mat)
      #p.mat <- sweep(p.mat, 1, STATS=apply(p.mat,1,median, na.rm=T))
   
    subMat <- p.mat[rownames(p.mat) %in% sigProts,]
    
    rownames(subMat) <-  multiUniprots2multiGenes(rownames(subMat))
   
  #heatmap annotation 
  grp <- sub(".[1-6]$", "", colnames(subMat))
  grp <- sub("^[LD]{1}", "", grp)
  grp <-  str_replace(grp, "(?<=-)[LD]", "")
  treat <- ifelse(grepl("^[D|L]{1}", colnames(subMat)), 'treatment', 'control')
  rep <- sub(".","",str_extract(colnames(subMat), ".[1-3]$"))

  col <-  colorRamp2(c(-3,0,3), c("blue","white","red"))
  
  ha <- HeatmapAnnotation(group = grp, col=list(
                                                                          group=c('orf3aBeta'='red', 'orf3aQ57H'='blue', 'orf3aS171L'='green', 'orf3aWT'='orange',
                                                                           'nsp6BA1'='yellow', 'nsp6BA2'='purple', 'nsp6XBB'='pink', 'nsp6WT'='darkred')))
  
  hm <-  Heatmap(subMat, 
          cluster_rows = clusterwNA(subMat),
          cluster_columns = F,
          show_row_names = T,
          row_names_gp = gpar(fontsize=3),
          col=col,
          name = 'z-score (Ints.)',
          row_title = sprintf('%s proteins', nrow(subMat)),
          column_split = ifelse(grepl('^[LD]{1}', colnames(subMat)), 'treatment', 'control'),
         top_annotation = ha)
  
 draw(hm, column_title=paste0(lab))
BackupAsPDF(draw(hm, column_title=paste0(lab)), paste0(n[[i]], '.', lab, '.contrastofcontrast.collapsed.heatmap.'),  dimensions=c(9,11))

}
  
}, x=em.list, y=mss.em.list, n=names(em.list))
```

now collapse each of these to the mean z-score per group

```{r}
lapply(seq_along(em.list), function(x,y, z,n,i){
  
   grp <- n[[i]]

   
   for (lab in unique(x[[grp]]$contrast)){
     print(lab)
    
     sigProts <- x[[grp]][sig != 'not' & contrast == lab, unique(Protein)]
     #print(sigProts)
      p.mat <- p.mat.list[[grp]]
   
      # z scores of heatmap
      p.mat <-  scale(p.mat)
      #p.mat <- sweep(p.mat, 1, STATS=apply(p.mat,1,median, na.rm=T))
   
    subMat <- p.mat[rownames(p.mat) %in% sigProts,]
    
    rownames(subMat) <-  multiUniprots2multiGenes(rownames(subMat))
   
  #heatmap annotation 
 # grp <- sub(".[1-6]$", "", colnames(subMat))
#  grp <- sub("^[LD]{1}", "", grp)
 # grp <-  str_replace(grp, "(?<=-)[LD]", "")
#  treat <- ifelse(grepl("^[D|L]{1}", colnames(subMat)), 'treatment', 'control')
#  rep <- sub(".","",str_extract(colnames(subMat), ".[1-3]$"))

  col <-  colorRamp2(c(-3,0,3), c("blue","white","red"))
  
  ha <- HeatmapAnnotation(group = grp, col=list(
                                                                          group=c('orf3aBeta'='red', 'orf3aQ57H'='blue', 'orf3aS171L'='green', 'orf3aWT'='orange',
                                                                           'nsp6BA1'='yellow', 'nsp6BA2'='purple', 'nsp6XBB'='pink', 'nsp6WT'='darkred')))
  
  hm <-  Heatmap(subMat, 
          cluster_rows = clusterwNA(subMat),
          cluster_columns = F,
          show_row_names = T,
          row_names_gp = gpar(fontsize=3),
          col=col,
          name = 'z-score (Ints.)',
          row_title = paste0(nrow(subMat), ' proteins +/- 50% FC & p.val < 0.005'),
          column_split = ifelse(grepl('^[LD]{1}', colnames(subMat)), 'treatment', 'control')) 
       #  top_annotation = ha)
  
 draw(hm, column_title=paste0(lab))
 BackupAsPDF(draw(hm, column_title=paste0(lab)), paste0(n[[i]], '.', lab, '.contrastofcontrast.collapsed.heatmap.'),  dimensions=c(9,11))

}
  
}, x=em.list, y=mss.em.list, z=p.mat.list, n=names(em.list))
```
ok now tidy output for both and write out table
```{r}
em.list %>%  names()

lapply(seq_along(em.list), function(x,n,i){
  print(head(x[[i]]))
  x.dt <-  dcast(x[[i]], Protein+gene.ctrl~contrast, value.var = c('delta_log2FC','delta_pval', 'delta_fdr'))
  print(head(x.dt))
  setnames(x.dt, old=c('gene.ctrl'), new=c('gene'))
  
  fwrite(x.dt, paste0('./output/29.01.24.', n[[i]], '.treatmentstatus.contrasts.wide.csv'))
},x=em.list, n=names(em.list))
```


```{r}
lapply(seq_along(em.list), function(x,y, z,n,i){
  
   grp <- n[[i]]

   
   for (lab in unique(x[[grp]]$contrast)){
     print(lab)
    
     sigProts <- x[[grp]][sig != 'not' & contrast == lab, unique(Protein)]
     #print(sigProts)
      p.mat <- p.mat.list[[grp]]
   
      # z scores of heatmap
      p.mat <-  scale(p.mat)
      #p.mat <- sweep(p.mat, 1, STATS=apply(p.mat,1,median, na.rm=T))
   
    subMat <- p.mat[rownames(p.mat) %in% sigProts,]
    
    rownames(subMat) <-  multiUniprots2multiGenes(rownames(subMat))
   
  #heatmap annotation 
 # grp <- sub(".[1-6]$", "", colnames(subMat))
#  grp <- sub("^[LD]{1}", "", grp)
 # grp <-  str_replace(grp, "(?<=-)[LD]", "")
#  treat <- ifelse(grepl("^[D|L]{1}", colnames(subMat)), 'treatment', 'control')
#  rep <- sub(".","",str_extract(colnames(subMat), ".[1-3]$"))

  col <-  colorRamp2(c(-3,0,3), c("blue","white","red"))
  
  ha <- HeatmapAnnotation(group = grp, col=list(
                                                                          group=c('orf3aBeta'='red', 'orf3aQ57H'='blue', 'orf3aS171L'='green', 'orf3aWT'='orange',
                                                                           'nsp6BA1'='yellow', 'nsp6BA2'='purple', 'nsp6XBB'='pink', 'nsp6WT'='darkred')))
  
 # hm <-  Heatmap(subMat, 
 #         cluster_rows = clusterwNA(subMat),
#          cluster_columns = F,
#          show_row_names = T,
#          row_names_gp = gpar(fontsize=3),
#          col=col,
##          name = 'z-score (Ints.)',
#          row_title = paste0(nrow(subMat), ' proteins +/- 50% FC & p.val < 0.005'),
#          column_split = ifelse(grepl('^[LD]{1}', colnames(subMat)), 'treatment', 'control')) 
       #  top_annotation = ha)
  
# draw(hm, column_title=paste0(lab))
# BackupAsPDF(draw(hm, column_title=paste0(lab)), paste0(n[[i]], '.', lab, '.contrastofcontrast.collapsed.heatmap.'),  dimensions=c(9,11))

#}
  
}, x=em.list, y=mss.em.list, z=p.mat.list, n=names(em.list))
```