---
title: "Roche.ABUB.scatterplot"
author: "Martin Gordon"
date: "2023-11-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Generate AB UB scatteplot for the Roche report

```{r}
library(ggplot2)
library(ggrepel)
library(data.table)
library(magrittr)
library(ggpubr) #publication ready plots
library(ComplexHeatmap)
library(circlize)

source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping

```


read in the files for the plot
```{r}
ub.dt <- fread('data/A549_UB_results.txt')
ab.dt <-  fread('data/A549_AB_results.txt')

ub.dt[abs(log2FC) > 1 & adj.pvalue < 0.05, .N] # 2665 in total


str(ub.dt) # no sites; all the info collapsed per gene?
ab.dt$issue %>%  unique() # some conditions mising here, lets plot as is for now but remove afterwards if necissary

ub.dt[Label == 'PIV_6h-Mock_6h' & Protein %like% c('P07873'),]


ub.dt[,Label] %>%  unique()


ub.dt[log2FC > 1 & adj.pvalue < 0.05, .N, by=Label]
```
subset to the desired labels 
```{r}
labels.oi <-c("PIV_6h-Mock_6h",  "PIV_24h-Mock_24h", "RSV_6h-Mock_6h", "RSV_24h-Mock_24h")


omics.list <-  list(ab = ab.dt[Label %in% labels.oi,],
                    ub = ub.dt[Label %in% labels.oi,])



ub.dt[Label %in% labels.oi & adj.pvalue < 0.05 & abs(log2FC) > 1, ]
omics.list[['ub']][adj.pvalue < 0.05 & abs(log2FC) > 1, ] # shoud be 600 significant things
```

add a omics identifier to each dt and translate prots to genes
clean the data; add labels etc.
```{r}
lapply(seq_along(omics.list), function(x,n,i){
  
  x[[i]][, omics := n[[i]]]
  x[[i]][, gene := ifelse(omics == 'ub', multiUniprotSites2multiGeneSites(Protein), multiUniprots2multiGenes(Protein)) ]
  
  x[[i]][, sig := 'not']
  x[[i]][!issue %in% c("oneConditionMissing", "completeMissing") & abs(log2FC) > 1 & adj.pvalue < 0.05, sig := 'log2FC > 1 & adj.pval < 0.05']
  
},x=omics.list, n=names(omics.list))
```


Clean UB sites with duplicate entries
```{r}
#omics.list[['ub']][sig != 'not', ] # shoud be 600 significant things

# scrub the ub site from the data and rename
omics.list[['ub']][, ProteinSite := Protein]

omics.list[['ub']][, Protein := gsub("_K[0-9]+", "", ProteinSite) %>% #strip UB info
          strsplit (";") %>%  #split multiprots; creates a list of vectors of each element
          lapply (unique) %>% #removes any duplicate elements in list
          lapply (paste, collapse = ";") %>% #colapses back to unique prot names
          unlist]

# clean the gene names col
omics.list[['ub']][, geneSite := gene]
omics.list[['ub']][, gene := multiUniprots2multiGenes(Protein)]

```
identify best scoring site per protein
Use geommetric mean 

```{r}
#similiar to above
magSigScore <- function(log2FC, pvalue, magnitudeScale = 2){
  magnitude <- magnitudeScale * abs(log2FC)
  significance <- -log10(pvalue)
  significance <- ifelse (significance > magnitude, magnitude, significance)
  sqrt(magnitude * significance)
}

omics.list[['ub']][, geom.mean := magSigScore(log2FC = log2FC, pvalue = pvalue, magnitudeScale = 2)]

#only take site with max sig score wehn mulitple measurements
omics.list[['ub']] <- omics.list[['ub']][!is.na(geom.mean) & !issue %in% c("oneConditionMissing", "completeMissing"), .SD[which.max(geom.mean)], by=.(Protein,Label)]

omics.list[['ub']][sig != 'not',.N, by=Label] 
omics.list[['ub']][gene=='MX1' & Label == 'RSV_24h-Mock_24h',]
omics.list[['ab']][gene=='MX1' & Label == 'RSV_24h-Mock_24h',]
```

now have one value per protein/UB site
combine the DTs 
```{r}
omics.list[['ab']][omics.list[['ub']], `:=`(UB.log2FC = i.log2FC, UB.sig = i.sig), on=c('Protein', 'Label')]
```


```{r}
omics.list[['ab']][, Label := factor(Label, levels = c(labels.oi))]

g <-  ggplot(omics.list[['ab']][!issue %in% c("oneConditionMissing", "completeMissing"),], aes(x=UB.log2FC, y=log2FC, col=UB.sig, label=gene)) +
  geom_point() +
  geom_abline(intercept=0, slope=1, linetype=2, alpha=0.4) +
  geom_vline(xintercept=0) +
  geom_hline(yintercept=0) +
  ylab('Abundance (Log2FC)') +
  xlab('Ubiquitination (Log2FC)') +

  scale_color_manual(values=c('not'='black', 'log2FC > 1 & adj.pval < 0.05'='red'), name='Significant in UB') +
  geom_text_repel(data=omics.list[['ab']][UB.sig != 'not',], aes(label=gene), size = 2, max.overlaps = 20, color='grey', segment.linetype=3, fontface='bold') +
  facet_wrap(~Label) +
  theme_bw()
g

BackupAsPDF(g, 'roche.scatterplot.ABvsUB.', dimensions = c(12,8))
```

results look good, dropped sites are those not present in both sets

```{r}
omics.list[['ab']][, Label := factor(Label, levels = c(labels.oi))]

g <-  ggplot(omics.list[['ab']][!issue %in% c("oneConditionMissing", "completeMissing"),], aes(x=UB.log2FC, y=log2FC, col=UB.sig, label=gene)) +
  geom_point() +
  geom_abline(intercept=0, slope=1, linetype=2, alpha=0.4) +
  geom_vline(xintercept=0) +
  geom_hline(yintercept=0) +
  ylab('Abundance (Log2FC)') +
  xlab('Ubiquitination (Log2FC)') +
  ylim(-4,4) +
  scale_color_manual(values=c('not'='black', 'log2FC > 1 & adj.pval < 0.05'='red'), name='Significant in UB') +
  geom_text_repel(data=omics.list[['ab']][UB.sig != 'not',], aes(label=gene), size = 2, max.overlaps = 20, color='grey', segment.linetype=3, fontface='bold') +
  facet_wrap(~Label) +
  theme_bw()
g

BackupAsPDF(g, 'roche.scatterplot.ABvsUB.adjused.y.limits.', dimensions = c(12,8))
```


How many UB sites are sig in the different datasets?
```{r}
omics.list[['ub']][sig != 'not' & Label == 'PIV_6h-Mock_6h']
omics.list[['ab']][sig != 'not' & Label == 'PIV_6h-Mock_6h' & Protein %in% c('P07873', 'Q8IVM0')]
```


Regenerate the barplot and the Go enrihcment heatmap

```{r}

ub.dt <- fread('data/A549_UB_results.txt')
labels.oi <-c("PIV_6h-Mock_6h",  "PIV_24h-Mock_24h", "RSV_6h-Mock_6h", "RSV_24h-Mock_24h")


str(ub.dt) # no sites; all the info collapsed per gene?
ub.dt[,Label] %>%  unique()

ub.dt[, p.adj := p.adjust(pvalue, method='BH'), by=Label]
ub.dt[, sig := 'not']
ub.dt[abs(log2FC) > 1 & p.adj < 0.05, sig := ifelse(log2FC > 0, 'up', 'down')] 

ub.dt[sig != 'not' & Label %in% labels.oi, .N, by=.(Label,sig)] 


lab.order <-c("RSV_6h-Mock_6h", "RSV_24h-Mock_24h", "PIV_6h-Mock_6h",  "PIV_24h-Mock_24h")

plot.dt <-  ub.dt[sig != 'not' & Label %in% labels.oi, .N, by=.(Label,sig)] 
plot.dt[, Label := factor(Label, levels=c(lab.order))]


g <-   ggplot(plot.dt, aes(x=Label, y=N, fill=sig)) + 
  geom_bar(stat='Identity') +
  scale_fill_manual(values=c('up'='red', 'down'='blue'), name='significance') +
  xlab('Contrasts') +
  ylab('Significant at abs(Log2FC) > 1 & adj.pval < 0.05' ) +
  theme_classic()
g

BackupAsPDF(g, 'N.sig.sites.barplot', dimensions = c(8,6))


g <-   ggplot(plot.dt, aes(x=Label, y=N, fill=sig)) + 
  geom_bar(stat='Identity') +
  scale_fill_manual(values=c('up'='red', 'down'='blue'), name='significance') +
  ylab('Count UB Sites' ) +
  xlab(NULL) +
  theme_classic() +
  theme(axis.text.x = element_text(angle=90))
g

BackupAsPDF(g, 'N.sig.sites.barplot', dimensions = c(5,5))

```
run enrichment analysis of the significant genes  in each set
```{r}
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")

# universe will be the proteins of all the detected sig sites
ub.dt[, uniprot := gsub("_K[0-9]+", "", Protein) %>% #strip UB info
          strsplit (";") %>%  #split multiprots; creates a list of vectors of each element
          lapply (unique) %>% #removes any duplicate elements in list
          lapply (paste, collapse = ";") %>% #colapses back to unique prot names
          unlist]

ub.dt[, gene := multiUniprots2multiGenes(uniprot)]

universe <- unique(ub.dt$gene)
universe %>%  length()


ub.dt[, enrich.grp := interaction(Label, sig)]
  
enrich.dt <- enricherOnGroups(ub.dt[Label %in% labels.oi & sig != 'not',], groupColumns = 'enrich.grp', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)
#simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich.grp')

hm <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt, NULL, groupColumn = 'enrich.grp', topN = 10,
                                  negCols=unique(enrich.dt$enrich_grp[grep('down', enrich.dt$enrich_grp)]), 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 5), upperThreshold = 8) 

BackupAsPDF(hm, 'UBsite.GOenrichment.enrichgrp.full.heatmap')
```

Changes; leave out the 6hr and 

```{r}
enrich.dt <- enricherOnGroups(ub.dt[Label %in% labels.oi & !Label %like% '6h' & sig != 'not',], groupColumns = 'enrich.grp', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)

# nothing to simplify; just plot the original heatmap
#simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go)
simplifyEnrichBySimilarUniverseMembership.general(enrichResultsTable = enrich.dt, gmt=gmt.go, group='Label')


hm <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt, NULL, subset.enrich.dt, topN = 10,
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 7),
                                  cluster_columns=F, upperThreshold = 8) 


BackupAsPDF(hm, 'go.enrichment.24hr.heatmap', dimensions=c(6,5))


simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go)


View(enrichHeatmapBestPerGroup)


ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt, groupColumn = 'enrich_grp', topN = 10, title='GO term enrichment', 
                                  negCols=unique(simp.enrich$simplified$enrich_grp[grep('down', simp.enrich$simplified$enrich_grp)]), 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 4), upperThreshold = 10)


```

```{r}
# try combining the up and down genes
ub.dt[, enrich.grp := Label]


enrich.dt <- enricherOnGroups(ub.dt[Label %in% labels.oi & !Label %like% '6h' & sig != 'not',], groupColumns = 'enrich.grp', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)

# nothing to simplify; just plot the original heatmap
#simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go)
simplifyEnrichBySimilarUniverseMembership.general(enrichResultsTable = enrich.dt, gmt=gmt.go, group='Label')


hm <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt, NULL, topN = 10, groupColumn = 'enrich.grp',
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 7),
                                  cluster_columns=F, upperThreshold = 8) 

Heatmap()



enrich.dt
enrich.dt[p.adjust < 0.1,]

enrichHeatmapBestPerGroup()


BackupAsPDF(hm, 'go.enrichment.24hr.heatmap', dimensions=c(6,5))
enrichHeatmapBestPerGroup()

simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go)


View(enrichHeatmapBestPerGroup)


ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt, groupColumn = 'enrich_grp', topN = 10, title='GO term enrichment', 
                                  negCols=unique(simp.enrich$simplified$enrich_grp[grep('down', simp.enrich$simplified$enrich_grp)]), 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 4), upperThreshold = 10)

```

Try generate your own heatmap using just the 

```{r}

main.wide <- dcast (enrich.dt[ID %in% ID[p.adjust < 0.1]], as.formula(paste("Description", 'enrich.grp', sep="~")), value.var="p.adjust")
main.mat <- -log10(as.matrix(main.wide, rownames = "Description"))


counts.wide <- dcast (enrich.dt[ID %in% ID[p.adjust < 0.1]], as.formula(paste("Description", 'enrich.grp', sep="~")), value.var="Count")
counts.mat <- as.matrix(counts.wide, rownames="Description")


grp.sizes <-  ub.dt[Label %in% labels.oi & !Label %like% '6h' & sig != 'not', .N, by=Label]
(grp.sizes$N)


topBars <- HeatmapAnnotation(`Group Sizes` = anno_barplot(rev(grp.sizes$N)))

col.palette = circlize::colorRamp2(breaks=c(0,2), c("#EEEEEE", "blue"))

colorRamp2()

hm <- heatmapNumbered(main.mat, counts.mat, max_pAdjust=0.1, bottom_annotation=topBars, column_names_gp = gpar(fontsize=7), row_names_gp = gpar(fontsize=7), upperThreshold = 2.5)
BackupAsPDF(hm, 'go.enrichment.24hr.labels.heatmap', dimensions=c(6,5))

```
Rerun the plot w/o using the max vals per UB site

read in the files for the plot
```{r}
ub.dt <- fread('data/A549_UB_results.txt')
ab.dt <-  fread('data/A549_AB_results.txt')

str(ub.dt) # no sites; all the info collapsed per gene?
ab.dt$issue %>%  unique() # some conditions mising here, lets plot as is for now but remove afterwards if necissary

ub.dt[Label == 'PIV_6h-Mock_6h' & Protein %like% c('P07873'),]


ub.dt[,Label] %>%  unique()


ub.dt[log2FC > 1 & adj.pvalue < 0.05, .N, by=Label]
```
subset to the desired labels 
```{r}
labels.oi <-c("PIV_6h-Mock_6h",  "PIV_24h-Mock_24h", "RSV_6h-Mock_6h", "RSV_24h-Mock_24h")


omics.list <-  list(ab = ab.dt[Label %in% labels.oi,],
                    ub = ub.dt[Label %in% labels.oi,])



ub.dt[Label %in% labels.oi & adj.pvalue < 0.05 & abs(log2FC) > 1, ]
omics.list[['ub']][adj.pvalue < 0.05 & abs(log2FC) > 1, ] # shoud be 600 significant things
```

add a omics identifier to each dt and translate prots to genes
clean the data; add labels etc.
```{r}
lapply(seq_along(omics.list), function(x,n,i){
  
  x[[i]][, omics := n[[i]]]
  x[[i]][, gene := ifelse(omics == 'ub', multiUniprotSites2multiGeneSites(Protein), multiUniprots2multiGenes(Protein)) ]
  
  x[[i]][, sig := 'not']
  x[[i]][!issue %in% c("oneConditionMissing", "completeMissing") & abs(log2FC) > 1 & adj.pvalue < 0.05, sig := 'log2FC > 1 & adj.pval < 0.05']
  
},x=omics.list, n=names(omics.list))

```

```{r}
#omics.list[['ub']][sig != 'not', ] # shoud be 600 significant things

# scrub the ub site from the data and rename
omics.list[['ub']][, ProteinSite := Protein]

omics.list[['ub']][, Protein := gsub("_K[0-9]+", "", ProteinSite) %>% #strip UB info
          strsplit (";") %>%  #split multiprots; creates a list of vectors of each element
          lapply (unique) %>% #removes any duplicate elements in list
          lapply (paste, collapse = ";") %>% #colapses back to unique prot names
          unlist]

# clean the gene names col
omics.list[['ub']][, geneSite := gene]
omics.list[['ub']][, gene := multiUniprots2multiGenes(Protein)]


omics.list[['ub']][sig != 'not', .N, by=Label]


#only take site with max sig score wehn mulitple measurements
test <-  omics.list[['ub']]
test[, geom.mean := magSigScore(log2FC = log2FC, pvalue = pvalue, magnitudeScale = 2)]

test <- test[!is.na(geom.mean) & !issue %in% c("oneConditionMissing", "completeMissing"), .SD[which.max(geom.mean)], by=.(Protein,Label)]

# different number of hits in both
omics.list[['ub']][sig != 'not',.N, by=Label] 
test[sig != 'not', .N, by=Label]
```

combine the DTs and plot all sites vs 

now have one value per protein/UB site
combine the DTs 
```{r}
omics.list[['ab']] <- omics.list[['ab']][omics.list[['ub']], `:=`(UB.log2FC = i.log2FC, UB.sig = i.sig), on=c('Protein', 'Label')]
omics.list[['ab']][Label %in% labels.oi & sig != 'not',.N, by=Label]
omics.list[['ub']][Label %in% labels.oi & sig != 'not',.N, by=Label]


test.dt <-  merge(x=omics.list[['ab']], y=omics.list[['ub']], by.x=c('Protein', 'Label'), by.y=c('Protein', 'Label'), all.y=T)
test.dt
# convert NAs to 0 for plotting
test.dt <- test.dt[is.na(log2FC.x), log2FC.x := 0]
test.dt <- test.dt[is.na(log2FC.y), log2FC.y := 0]

test.dt[, Label:= factor(Label, levels=labels.oi)]

g <-  ggplot(test.dt, aes(x=log2FC.y, y=log2FC.x, col=sig.y, label=geneSite)) +
  geom_point() +
  geom_abline(intercept=0, slope=1, linetype=2, alpha=0.4) +
  geom_vline(xintercept=0) +
  geom_hline(yintercept=0) +
  ylab('Abundance (Log2FC)') +
  xlab('Ubiquitination (Log2FC)') +
  scale_color_manual(values=c('not'='black', 'log2FC > 1 & adj.pval < 0.05'='red'), name='Significant in UB') +
  geom_text_repel(data=test.dt[sig.y != 'not',], aes(label=geneSite), size = 2, max.overlaps = 20, color='grey', segment.linetype=3, fontface='bold') +
  facet_wrap(~Label, scales='free_y') +
  theme_bw()

g

BackupAsPDF(g, 'roche.scatterplot.ABvsUB.allUB.labelSites', dimensions = c(12,8))
```
Todo
---
Compare the output of the two join methods

```{r}
omics.list[['ab']][, Label := factor(Label, levels = c(labels.oi))]


omics.list[['ab']][UB.sig != 'not', .N, by=Label]

g <-  ggplot(omics.list[['ab']][!issue %in% c("oneConditionMissing", "completeMissing"),], aes(x=UB.log2FC, y=log2FC, col=UB.sig, label=gene)) +
  geom_point() +
  geom_abline(intercept=0, slope=1, linetype=2, alpha=0.4) +
  geom_vline(xintercept=0) +
  geom_hline(yintercept=0) +
  ylab('Abundance (Log2FC)') +
  xlab('Ubiquitination (Log2FC)') +

  scale_color_manual(values=c('not'='black', 'log2FC > 1 & adj.pval < 0.05'='red'), name='Significant in UB') +
  geom_text_repel(data=omics.list[['ab']][UB.sig != 'not',], aes(label=gene), size = 2, max.overlaps = 20, color='grey', segment.linetype=3, fontface='bold') +
  facet_wrap(~Label,scales = 'free_y') +
  theme_bw()
g

BackupAsPDF(g, 'roche.scatterplot.ABvsUB.allUB', dimensions = c(12,8))
```



Not needed
----

for UB, we have multiple UBylated sites per protein. 
For now we need a representative value for per protein for plotting 
Take the geometric mean of the log2FC and pval
For now, extract the site per Protein with the highest log2FC pval geometric mean

```{r}
# scrub the ub site from the data
omics.list[['ub']][, newProtein := gsub('_K[0-9]+','', Protein)]
# calculate geommetric mean for filtering rows
#omics.list[['ub']][, geom.mean := sqrt( (-log10(pvalue))*(abs(log2FC))  )]


# threshold pval is for RNAseq data as tend to have inflated pvalues
# work on this...
calculateGeoMean <- function(x='log2FC', y='pvalue', thresholdpval=T, threshold=2){
  
  y.log <- -log10(y)
  
  if (thresholdpval %in% c('T', 'TRUE')){
    # if pval > 2* abs, just take the log2FC
    if (y.log > (threshold * abs(x)) ) {
      y.log == abs(x)
    }
  }
  
  return(sqrt(abs(x) * y.log))
}

calculateGeoMean(x=2.5, y=0.0001, threshold=2)
# ge thtis working for data.tables.. could probably use apply to go through the dt and calculate per line.
#test[, geom_m := calculateGeoMean(x=log2FC, y=pvalue, thresholdpval = T, threshold = 2)]
```

per protein, extract the site with the max geom.mean value
```{r}
test <- omics.list[['ub']]
test <-  test[, .SD[which.max(geom.mean)], by=.(Protein,Label)]
test
test[newProtein == 'A0AVT1']
omics.list[['ub']][newProtein == 'A0AVT1']

```

