---
title: "130923_DAR31_contrastofcontrasts"
author: "Martin Gordon"
date: "2023-09-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DAR31 looking at contrast of the different ligand treatments per receptor group
Issue with this is normalisation factors are different; have now rerun the analysis to process each group together

```{r}
library(data.table)
library(ComplexHeatmap)
library(circlize)
library(ggplot2)
library(magrittr)
library(circlize)
library(stringr)
library(emmeans) #EMM package for post-hoc analysis on models
library(parallelDist) #needed for some of Bens custom enrichment functions


rotate.x.axis.text <- theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

#other R scripts with heler functions for the analysis
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R") # lm modesl with emmeans functionality
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R") #custom enrichment functions using complex heatmap

WriteSessionInfo()
WriteInstalledPackages()
```

For DAR31, we have 3 conditions followed over 5 timepoints
Canabonid receptor 1 (CB1) with 3 treatments
We want to see if there is any difference in the response over time; check for an interaction effect between the receptor 
Similar to the analysis for DAR34; run lm model per prot, look for significant interactions first, plot and run GO enrichment

150923
--- 
reprocess the expression data, as each of the DARPA datasets were processed individually

Data generated from Spectronaut DIA 
Assuming is CBXXX is receptor, _[0=9] refers to timepoint, there are 3 receptors (CB3234,CB1066,CBCP55) measured across 4 timepoints (00,01,05,10,30) and a control (Cyto)
Splitting dataset into each individual receptor

```{r}
spec <- fread('./data/DAR031/EX/All/DIA/Results/DirDIA/20230524_185013_DAR31_EX_dirDIA_all_MSStatsFormatReport.xls')
unique(spec$Condition) #3 unique drugs conditions : CB1066, CB3234, CBCP55

spec <- spec[!Condition %like% 'Cyto',]

spec$Condition %>%  unique()
```
remove the low intensity 'trash' on the 

```{r}
spec <- spec[Intensity > 2^5,]
ggplot(spec, aes(x=log2(Intensity))) + geom_histogram(bins=100)
```
msstats data process

```{r}

spec.mss <- specFileToCompleteMSstats(spec)
spec.mss[, IsotopeLabelType := "L"]

setDF(spec.mss)

dp.out <- MSstats::dataProcess(spec.mss,
                               normalization = 'equalizeMedians',
                               summaryMethod="TMP",
                               featureSubset = "highQuality",
                               remove_uninformative_feature_outlier = TRUE,
                               MBimpute = FALSE )
```
#save the output as RData file
```{r}
#saveRDS(dp.out, file = './output/DAR31.msstats.proc.out.Rdata')

fwrite(dp.out$ProteinLevelData, './output/mss.dataProc.Protein.csv.gz')
fwrite(dp.out$FeatureLevelData, './output/mss.dataProc.Feature.csv.gz')
```


plot the normalised values and assess if we need to run a second round of normalisation
proceed as is for now; dont want to over-correct for true biological differences

```{r}
mss <- as.data.table(dp.out$ProteinLevelData)
mss
```


```{r}
# including the dropped sample for now
protQuant <- fread ("./output/mss.dataProc.Protein.csv.gz")

protQuant %>%  str()
protQuant$GROUP %>%  unique()

# parse GROUP names to define the actual columns that the linear models will work on
# this work depends on how your columsn are defined
protQuant[, ligand   := gsub("_[0-9]{2}$", "",GROUP)]

# fix subject info to account for treatment batch 
protQuant[,  SUBJECT := as.factor(interaction(ligand, SUBJECT))]

protQuant[, gene := multiUniprots2multiGenes(Protein)]
protQuant[, rep := str_extract(SUBJECT, '[0-9]{1}$')]

protQuant[, timeStr := str_extract(GROUP, "[0-9]{2}$")]
protQuant[, rankTime := as.integer(as.factor(timeStr))] #timeranked to 0-5; we do this to allow even intervals between the timepoints
protQuant[, .(timeStr, rankTime)] %>% unique() #timepoint to time rank 

protQuant[, .N, by=.(ligand, timeStr,SUBJECT)] %>% 
  .[order(timeStr,ligand)]



#removing troublesome samples
protQuant <- protQuant[ !(ligand == "CBCP55" & SUBJECT == 'CBCP55.3')]
protQuant <- protQuant[ !(ligand == "CBCP55" & SUBJECT =='CBCP55.1' & timeStr %in% c("00", "01"))]
protQuant <- protQuant[ !(ligand == "CB1066" & SUBJECT =='CB1066.4' & timeStr %in% c("05"))]


# calculate vsT0 expression
# check if this is correct? does ligand need to be changed to GROUP?
protQuant[, intVsMeanTime0 := LogIntensities - mean(LogIntensities[timeStr == '00']), by = .(ligand, Protein)]
protQuant[, vsTimeZero := LogIntensities - mean(LogIntensities[timeStr == timeStr[!is.na(LogIntensities)][1] ]), by = .(Protein, ligand)]


ggplot(data=protQuant, aes(x=paste0(GROUP,'.', SUBJECT), y=LogIntensities, fill=GROUP)) + geom_boxplot() + rotate.x.axis.text

#write out the cleaned data used in the downstream analysis
#fwrite(protQuant,"./output/mss.dataProc.Protein.clean.csv.gz")

protQuant <- fread("./output/mss.dataProc.Protein.clean.csv.gz")
protQuant[, timeStr := as.factor(timeStr)]
```

model to look at intensity over time
emtrends: contrast trendlines over ligand factor levels
```{r}
model <- lm(LogIntensities ~ SUBJECT + poly(rankTime,degree=3)*ligand, data=protQuant)
anova(model) # model suggests there is good distinction between the different ligand effects
emtrends(model, pairwise ~ ligand, var='rankTime')$contrasts
emmip(model,  ligand ~ rankTime, cov.reduce= range)
```
CP55 vs 1066
----
Lets just look at pairwise comparisons for now
Drop CP3234 for the first comparison - also calculate vs T0 intensities

```{r}
ligQuant <- protQuant[ligand != 'CB3234']
ligQuant <- ligQuant[NumMeasuredFeature != 1,] # remove proteins with only 1 measured feature

# remove proteins with < 15 measured per group
ligQuant[, prots.per.condition := .N, by=.(Protein,ligand)] 
rm.prots <- ligQuant[prots.per.condition < 15, unique(Protein)]

ligQuant <- ligQuant[!Protein %in% rm.prots]
```

Run the model
```{r}
#using lm code, look for interactions between ligand response
lm.out <- linearModelsAllProteins(ligQuant, 
                                  formulaList = list(full  = LogIntensities ~ SUBJECT + ligand + rankTime + I(rankTime^2) + I(rankTime^3) + ligand:I(rankTime^2)  + ligand:I(rankTime^3)),
                                  splitColumn = "Protein") 
```


plot results histogram
nothing significant found here
```{r}
g <- ggplot(lm.out$anova, aes(x=p.value)) +
  geom_histogram(boundary=0.0, binwidth = 0.005) + 
  facet_wrap(~term)

g

interaction(protQuant$GROUP, protQuant$SUBJECT)
BackupAsPDF(g, 'cp55.cp1066.pva.histogram')
```


```{r}
lm.out$anova[, fdr := p.adjust(p.value, method = 'BH')]
lm.out$anova[, gene := multiUniprots2multiGenes(Protein)]

# write out the results
fwrite(lm.out$anova, './output/cp55.cp1066.anova.csv.gz')
```

plot the CP55:CP1066 data

```{r}
cb55.cb1066 <- fread('./output/cp55.cp1066.anova.csv.gz')

sigProteins <- cb55.cb1066[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & fdr < 0.05, unique(gene)]

p.mat <- dcast(ligQuant, Protein~paste0(GROUP,'.', rep), value.var = 'intVsMeanTime0') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))

p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]


hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints. vs T0',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("CP55", colnames(p.mat)), "CP55", "CP1066")))
hm
BackupAsPDF(hm, '1066.CP55.interaction.fdr<0.05.heatmap')
```
Collapse replicates and plot

```{r, fig.width=4, fig.height=5}

# calculate mean expression per ligand/tp/protein
ligQuant.sum <- ligQuant[, .(gene, GROUP, mean.Ints.vsT0 = mean(intVsMeanTime0, na.rm=T), mean.Intensity = mean(LogIntensities, na.rm=T)), by=.(ligand, timeStr, Protein)] %>%  unique() %>% 
  .[order(timeStr, ligand)]

# now to get vsT0 subtract from 
ligQuant.sum[, mean.vsT0 := mean.Intensity - mean.Intensity[timeStr == '0'],  by=.(Protein,ligand)]


cb55.cb1066 <- fread('./output/cp55.cp1066.anova.csv.gz')

sigProteins <- cb55.cb1066[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & fdr < 0.05, unique(gene)]

p.mat <- dcast(ligQuant.sum[!timeStr %in% '0',], Protein~GROUP, value.var = 'mean.Ints.vsT0') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))


p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]
p.mat

hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints. vs T0',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("CP55", colnames(p.mat)), "CBCP55", "CB1066")))

BackupAsPDF(hm, 'cb1066.cbcp55.sigProtes.collapsed.heatmap')
```



Plot raw intensity values 
No real evidence of consistent change
```{r}

sigProteins <- cb55.cb1066[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & fdr < 0.05, unique(gene)]


p.mat <- dcast(ligQuant, Protein~paste0(GROUP,'.', rep), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))


p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]

col_fun = colorRamp2(c(-2, 0, 2), c("blue", "white", "red"))


hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints.',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("CP55", colnames(p.mat)), "CP55", "CP1066")))
hm
BackupAsPDF(hm, '1066.CP55.interaction.intensities.fdr<0.05.heatmap')

```

Go enrichment of all significant results
No significant hits
```{r}
gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")
universe <- protQuant$gene %>%  unique()

lm.out$anova[, enrich_grp := 'no']
lm.out$anova[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & p.value < 0.01, enrich_grp := 'yes']

enrich.dt <- enricherOnGroups(lm.out$anova[enrich_grp == 'yes'], groupColumns = 'enrich_grp', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt,  gmt=gmt.go, groupColumn = 'enrich_grp', )

#how the heatmap is constructed
#dcast(enrich.dt[pvalue < 0.001], as.formula(paste("Description", 'enrich_grp', sep="~")), value.var="p.adjust") %>% as.matrix(rownames='Description') %>%  Heatmap() 

ggplot(enrich.dt[p.adjust < 0.05], aes(x=Description, y=-log10(pvalue))) + geom_col(fill='darkblue') + ggtitle('CP1066:CBCP55 sig GO terms') + coord_flip() + theme_minimal() %>% 
  BackupAsPDF('cp1066:cbcp55.sigGOterms.pdf')

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable =  enrich.dt, groupColumn = NULL, topN = 1, title='GO term enrichment (proteins FDR < 0.05)', 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 4), reduceRedundantsAcrossGroups=FALSE, upperThreshold = 10)
```

re-run lm and emmeans using emtrends to extract estimates of covariate slope for each drug level
```{r}

prot.mat <- dcast(ligQuant, Protein~paste0(ligand,'.',timeStr, rep), value.var = 'vsTimeZero') %>% 
  as.matrix(rownames='Protein')


#just filter out proteins with more than 25% of values misssing
na.vals <- apply(prot.mat, 1, function(x){
  # use the any function to detect 
  any(sum(is.na(x)) > 11) })

sum(na.vals) # removing over 1k proteins
names(na.vals[na.vals == TRUE])



#run our own analysis;anova per protein
# try again with this filter, will also rm prots with low numbers of reps
test.dt <- ligQuant[!Protein %in% names(na.vals[na.vals == TRUE]),] %>% # filter out missing vals
split(., .$Protein)


# run lm on each. of the proteins, 
test.out <- lapply(test.dt, function(x){
     lm.x <- lm(vsTimeZero ~ SUBJECT + ligand*poly(rankTime, degree=3), data=x) # different response to drug over time
     #emm1 = emmeans(lm.x, specs = pairwise ~ drug|poly(rankTime, degree=3)) #compare the CZ SP drug response over time 
     emm1 = emtrends(lm.x, pairwise ~ ligand, var='rankTime') #asking is the slope different for each level of the factor
     emm1$contrasts %>%  
     as.data.table()
})


test.out <- rbindlist(test.out, idcol = 'Protein')
test.out[, fdr := p.adjust(p.value, method='BH')]
```

Visualise the proteins with different trends per ligand
None found at a significance < 0.05

```{r}

sigProteins <- test.out[p.value < 0.01, unique(Protein)]

p.mat <- dcast(ligQuant, Protein~paste0(GROUP,'.', rep), value.var = 'intVsMeanTime0') %>% 
  as.matrix(rownames='Protein')

p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]

dim(p.mat)

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))



hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints. vs T0',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (p.value < 0.01)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("CP55", colnames(p.mat)), "CP55", "CP1066")))


hm
BackupAsPDF(hm, '1066.CP55.interaction.p.value<0.01.emmeans.trends.heatmap')
```

Compare CP55 to CP3234

```{r}
ligQuant <- protQuant[ligand != 'CB1066']
ligQuant <- ligQuant[NumMeasuredFeature != 1,] # remove proteins with only 1 measured feature

# remove proteins with < 15 measured per group
ligQuant[, prots.per.condition := .N, by=.(Protein,ligand)] 
rm.prots <- ligQuant[prots.per.condition < 15, unique(Protein)]

ligQuant <- ligQuant[!Protein %in% rm.prots]
```

Run the model per prot
```{r}
#using lm code, look for interactions between ligand response
lm.out <- linearModelsAllProteins(ligQuant, 
                                  formulaList = list(full  = LogIntensities ~ SUBJECT + ligand + rankTime + I(rankTime^2) + I(rankTime^3) + ligand:I(rankTime^2)  + ligand:I(rankTime^3)),
                                  splitColumn = "Protein") 
```


plot results histogram
again not much evidence of interaction between treatment type and time
```{r}
g <- ggplot(lm.out$anova, aes(x=p.value)) +
  geom_histogram(boundary=0.0, binwidth = 0.005) + 
  facet_wrap(~term)

g

BackupAsPDF(g, 'cb3234.cbcp55.pva.histogram')
```

```{r}
lm.out$anova[, fdr := p.adjust(p.value, method = 'BH')]
lm.out$anova[, gene := multiUniprots2multiGenes(Protein)]

# write out the results
fwrite(lm.out$anova, './output/cb3234.cbcp55.anova.csv.gz')
```


plot the cb3234.cbcp55 data

```{r}
cbcp3234.cbcp55 <- fread('./output/cb3234.cbcp55.anova.csv.gz')

sigProteins <- cbcp3234.cbcp55[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & fdr < 0.05, unique(gene)]

p.mat <- dcast(ligQuant, Protein~paste0(GROUP,'.', rep), value.var = 'intVsMeanTime0') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))


p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]


hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints. vs T0',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("CP55", colnames(p.mat)), "CBCP55", "CP3234")))
hm
BackupAsPDF(hm, 'CP3234.CBCP55.interaction.fdr<0.05.heatmap')
```

collapse replicates and plot 

```{r, fig.width=4, fig.height=6}

# calculate mean expression per ligand/tp/protein
ligQuant.sum <- ligQuant[, .(gene, GROUP, mean.Ints.vsT0 = mean(intVsMeanTime0, na.rm=T), mean.Intensity = mean(LogIntensities, na.rm=T)), by=.(ligand, timeStr, Protein)] %>%  unique() %>% 
  .[order(timeStr, ligand)]


ligQuant

# now to get vsT0 subtract from 
ligQuant.sum[, mean.vsT0 := mean.Intensity - mean.Intensity[timeStr == '0'],  by=.(Protein,ligand)]


cbcp3234.cbcp55 <- fread('./output/cb3234.cbcp55.anova.csv.gz')

sigProteins <- cbcp3234.cbcp55[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & fdr < 0.05, unique(gene)]

p.mat <- dcast(ligQuant.sum[!timeStr %in% '0',], Protein~GROUP, value.var = 'mean.Ints.vsT0') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))

p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]
p.mat

hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints. vs T0',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("CP55", colnames(p.mat)), "CBCP55", "CB3234")))

hm
BackupAsPDF(hm, 'cb3234.cbcp55.sigProtes.collapsed.heatmap')
```

Plot the raw intensities

```{r}
p.mat <- dcast(ligQuant, Protein~paste0(GROUP,'.', rep), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))


p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]


hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints.',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("CP55", colnames(p.mat)), "CBCP55", "CP3234")))
hm
BackupAsPDF(hm, 'CP3234.CBCP55.interaction.intensities.fdr<0.05.heatmap')
```
Go enrichment 
No significant results
```{r}
lm.out$anova[, enrich_grp := 'no']
lm.out$anova[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & fdr < 0.05, enrich_grp := 'yes']

enrich.dt <- enricherOnGroups(lm.out$anova[enrich_grp == 'yes'], groupColumns = 'enrich_grp', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt, groupColumn = 'enrich_grp', topN = 10, title='GO term enrichment (proteins FDR < 0.05)', 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 4), upperThreshold = 10)
```

Compare CP1066 vs CP3234

```{r}
ligQuant <- protQuant[ligand != 'CBCP55']
ligQuant <- ligQuant[NumMeasuredFeature != 1,] # remove proteins with only 1 measured feature as likely ID mismatch

# remove proteins with < 15 measured per group
ligQuant[, prots.per.condition := .N, by=.(Protein,ligand)] 
rm.prots <- ligQuant[prots.per.condition < 15, unique(Protein)]

ligQuant <- ligQuant[!Protein %in% rm.prots]
```

Run the model per prot
```{r}
#using lm code, look for interactions between ligand response
lm.out <- linearModelsAllProteins(ligQuant, 
                                  formulaList = list(full  = LogIntensities ~ SUBJECT + ligand + rankTime + I(rankTime^2) + I(rankTime^3) + ligand:I(rankTime^2)  + ligand:I(rankTime^3)),
                                  splitColumn = "Protein") 
```
plot results histogram
again not much evidence of interaction between treatment type and time
```{r}
g <- ggplot(lm.out$anova, aes(x=p.value)) +
  geom_histogram(boundary=0.0, binwidth = 0.005) + 
  facet_wrap(~term)

g

BackupAsPDF(g, 'cp1066.cp3234.pval.histogram')
```

```{r}
lm.out$anova[, fdr := p.adjust(p.value, method = 'BH')]
lm.out$anova[, gene := multiUniprots2multiGenes(Protein)]

# write out the results
fwrite(lm.out$anova, './output/cb1066.cb3234.anova.csv.gz')
```


plot the cb1066.cb3234 data

```{r}
cb1066.cb3234 <- fread('./output/cb1066.cb3234.anova.csv.gz')

sigProteins <- cb1066.cb3234[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & fdr < 0.05, unique(gene)]

p.mat <- dcast(ligQuant, Protein~paste0(GROUP,'.', rep), value.var = 'intVsMeanTime0') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))


p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]


hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints. vs T0',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("CB3234", colnames(p.mat)), "CB1066", "CB3234")))
hm
BackupAsPDF(hm, 'CP3234.CB1066.interaction.fdr<0.05.heatmap')
```

collapse replicates and plot 

```{r, fig.width=4, fig.height=5}

# calculate mean expression per ligand/tp/protein
ligQuant.sum <- ligQuant[, .(gene, GROUP, mean.Ints.vsT0 = mean(intVsMeanTime0, na.rm=T), mean.Intensity = mean(LogIntensities, na.rm=T)), by=.(ligand, timeStr, Protein)] %>%  unique() %>% 
  .[order(timeStr, ligand)]


# now to get vsT0 subtract from 
ligQuant.sum[, mean.vsT0 := mean.Intensity - mean.Intensity[timeStr == '0'],  by=.(Protein,ligand)]


cb1066.cb3234 <- fread('./output/cb1066.cb3234.anova.csv.gz')

sigProteins <- cb1066.cb3234[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & fdr < 0.05, unique(gene)]

p.mat <- dcast(ligQuant.sum[!timeStr %in% '0',], Protein~GROUP, value.var = 'mean.Ints.vsT0') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))

p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]
p.mat

hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints. vs T0',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("1066", colnames(p.mat)), "CB1066", "CB3234")))


BackupAsPDF(hm, 'cb3234.cbcp55.sigProtes.collapsed.heatmap')
```


Plot raw intensities 
```{r}
p.mat <- dcast(ligQuant, Protein~paste0(GROUP,'.', rep), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))


p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]


hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints.',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("CB3234", colnames(p.mat)), "CB1066", "CB3234")))
hm
BackupAsPDF(hm, 'CP3234.CB1066.interaction.intensities.fdr<0.05.heatmap')
```
Look at GO enrichment 
No significant results

```{r}
lm.out$anova[, enrich_grp := 'no']
lm.out$anova[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & fdr < 0.05, enrich_grp := 'yes']

enrich.dt <- enricherOnGroups(lm.out$anova[enrich_grp == 'yes'], groupColumns = 'enrich_grp', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt, groupColumn = 'enrich_grp', topN = 10, title='GO term enrichment (proteins FDR < 0.05)', 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 4), upperThreshold = 10)
```





```{r}
# including the dropped sample for now
protQuant <- fread ("/Users/martingordon/Documents/projects/052623_BPolacco_DARPA/052623_DARPa31_data/DAR31-ProteinLevelData.csv.gz")

# parse GROUP names to define the actual columns that the linear models will work on
# this work depends on how your columsn are defined
protQuant[, ligand   := gsub("CB|_[0-9]{2}", "",GROUP)]

# fix subject info to account for treatment batch 
protQuant[,  SUBJECT := as.factor(interaction(ligand, SUBJECT))]

protQuant[, timeStr := str_extract(GROUP, "[0-9]{2}$")]
protQuant[, rankTime := as.integer(as.factor(timeStr))] #timeranked to 0-5; we do this to allow even intervals between the timepoints
protQuant[, .(timeStr, rankTime)] %>% unique() #timepoint to time rank 

protQuant[, .N, by=.(ligand, timeStr,SUBJECT)] %>% 
  .[order(timeStr,ligand)]


# this doesnt account for missing values in the data
protQuant[, intVsMeanTime0 := LogIntensities - mean(LogIntensities[timeStr == '00']), by = .(ligand, Protein)]
protQuant[, vsTimeZero := LogIntensities - mean(LogIntensities[timeStr == timeStr[!is.na(LogIntensities)][1] ]), by = .(Protein, ligand)]
protQuant <- protQuant[ligand != '1066']
protQuant[, gene := multiUniprots2multiGenes(Protein)]

protQuant[, rep := str_extract(SUBJECT, '.[0-9]{1}$')]
```


```{r}
lm.out <- linearModelsAllProteins(protQuant, 
                                  formulaList = list(full  = LogIntensities ~ SUBJECT + ligand + rankTime + I(rankTime^2) + I(rankTime^3) + ligand:I(rankTime^2)  + ligand:I(rankTime^3)),
                                  splitColumn = "Protein", cl=6) 
```

Save results, visualize and 

```{r}
ggplot(lm.out$anova, aes(x=p.value)) + geom_histogram(boundary=0) + facet_wrap(~term) %>% 
  BackupAsPDF('cp55.cp3234.histogram')
```

```{r}
lm.out$anova[, fdr := p.adjust(p.value, method = 'BH')]
lm.out$anova[, gene := multiUniprots2multiGenes(Protein)]

# write out the results
fwrite(lm.out$anova, './output/cb55.cb3234.anova.csv.gz')
```

```{r}

cb55.cb3234 <- fread('./output/cb55.cb3234.anova.csv.gz')

sigProteins <- cb55.cb3234[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & fdr < 0.05, unique(gene)]

p.mat <- dcast(protQuant, Protein~paste0(ligand,'.',timeStr, rep), value.var = 'intVsMeanTime0') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))


p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]


hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints. vs T0',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("CP55", colnames(p.mat)), "CP55", "CP3234")))


hm

BackupAsPDF(hm, 'CP3234.CP55.interaction.fdr<0.05.heatmap')
```
LogIntensities plot
```{r}
p.mat <- dcast(protQuant, Protein~paste0(ligand,'.',timeStr, rep), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))


p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]


hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints.',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("CP55", colnames(p.mat)), "CP55", "CP3234")))


hm
BackupAsPDF(hm, 'CP3234.CP55.interaction.intensities.fdr<0.05.heatmap')
```
No significany Go enrichment terms
```{r}
lm.out$anova[, enrich_grp := 'no']
lm.out$anova[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & fdr < 0.1, enrich_grp := 'yes']

enrich.dt <- enricherOnGroups(lm.out$anova[enrich_grp == 'yes'], groupColumns = 'enrich_grp', geneColumn = "gene", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt, groupColumn = 'enrich_grp', topN = 10, title='GO term enrichment (proteins FDR < 0.05)', 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 4), upperThreshold = 10)
```

Compare CP1066 to CP3234
```{r}
# including the dropped sample for now
protQuant <- fread ("/Users/martingordon/Documents/projects/052623_BPolacco_DARPA/052623_DARPa31_data/DAR31-ProteinLevelData.csv.gz")

# parse GROUP names to define the actual columns that the linear models will work on
# this work depends on how your columsn are defined
protQuant[, ligand   := gsub("CB|_[0-9]{2}", "",GROUP)]

# fix subject info to account for treatment batch 
protQuant[,  SUBJECT := as.factor(interaction(ligand, SUBJECT))]

protQuant[, timeStr := str_extract(GROUP, "[0-9]{2}$")]
protQuant[, rankTime := as.integer(as.factor(timeStr))] #timeranked to 0-5; we do this to allow even intervals between the timepoints
protQuant[, .(timeStr, rankTime)] %>% unique() #timepoint to time rank 

protQuant[, .N, by=.(ligand, timeStr,SUBJECT)] %>% 
  .[order(timeStr,ligand)]


# this doesnt account for missing values in the data
protQuant[, intVsMeanTime0 := LogIntensities - mean(LogIntensities[timeStr == '00']), by = .(ligand, Protein)]
protQuant[, vsTimeZero := LogIntensities - mean(LogIntensities[timeStr == timeStr[!is.na(LogIntensities)][1] ]), by = .(Protein, ligand)]
protQuant <- protQuant[ligand != 'CP55']
protQuant[, gene := multiUniprots2multiGenes(Protein)]

protQuant[, rep := str_extract(SUBJECT, '.[0-9]{1}$')]

protQuant$ligand %>%  unique()
```

```{r}
lm.out <- linearModelsAllProteins(protQuant, 
                                  formulaList = list(full  = LogIntensities ~ SUBJECT + ligand + rankTime + I(rankTime^2) + I(rankTime^3) + ligand:I(rankTime^2)  + ligand:I(rankTime^3)),
                                  splitColumn = "Protein", cl=6) 
```

Save results, visualize and 

```{r}
ggplot(lm.out$anova, aes(x=p.value)) + geom_histogram(boundary=0) + facet_wrap(~term) %>% 
  BackupAsPDF('cp1066.cp3234.histogram')
```

```{r}
lm.out$anova[, fdr := p.adjust(p.value, method = 'BH')]
lm.out$anova[, gene := multiUniprots2multiGenes(Protein)]

# write out the results
fwrite(lm.out$anova, './output/cb3234.cb1066.anova.csv.gz')
```

```{r}
sigProteins <- lm.out$anova[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & fdr < 0.05, unique(gene)]

p.mat <- dcast(protQuant, Protein~paste0(ligand,'.',timeStr, rep), value.var = 'intVsMeanTime0') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))


p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]


hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints. vs T0',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("1066", colnames(p.mat)), "CP1066", "CP3234")))


hm

BackupAsPDF(hm, 'CP3234.CP1066.interaction.fdr<0.05.heatmap')
```
```{r}
sigProteins <- lm.out$anova[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & fdr < 0.05, unique(gene)]

p.mat <- dcast(protQuant, Protein~paste0(ligand,'.',timeStr, rep), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))


p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]


hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints. vs T0',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("1066", colnames(p.mat)), "CP1066", "CP3234")))

hm
```
150923
----
Redo the plots; remove the noisy proteins and rm any replicates that are proble

CB1066 vs CB3234
filter dataset
```{r}
protQuant$ligand %>% unique()

ligQuant <- protQuant[ligand != 'CBCP55']
ligQuant <- ligQuant[NumMeasuredFeature != 1,] # remove proteins with only 1 measured feature

# remove proteins with < 15 measured per group
ligQuant[, prots.per.condition := .N, by=.(Protein,ligand)] 
rm.prots <- ligQuant[prots.per.condition < 15, unique(Protein)]

ligQuant <- ligQuant[!Protein %in% rm.prots]

```


```{r}
lm.out <- linearModelsAllProteins(ligQuant, 
                                  formulaList = list(full  = LogIntensities ~ SUBJECT + ligand + rankTime + I(rankTime^2) + I(rankTime^3) + ligand:I(rankTime^2)  + ligand:I(rankTime^3)),
                                  splitColumn = "Protein", cl=6) 
```

Save results, visualize and 

```{r}
ggplot(lm.out$anova, aes(x=p.value)) + geom_histogram(boundary=0) + facet_wrap(~term) %>% 
  BackupAsPDF('cp1066.cp3234.histogram')
```

```{r}
lm.out$anova[, fdr := p.adjust(p.value, method = 'BH')]
lm.out$anova[, gene := multiUniprots2multiGenes(Protein)]

# write out the results
fwrite(lm.out$anova, './output/cb1066.cb3234.anova.csv.gz')
```

```{r}

cb1066.cb3234 <- fread('./output/cb1066.cb3234.anova.csv.gz')

sigProteins <- cb1066.cb3234[term %in% c("ligand:I(rankTime^2)","ligand:I(rankTime^3)") & fdr < 0.05, unique(gene)]

p.mat <- dcast(ligQuant, Protein~paste0(ligand,'.',timeStr, rep), value.var = 'intVsMeanTime0') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))


p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]


hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints. vs T0',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("CB1066", colnames(p.mat)), "CB1066", "CP3234")))


hm

BackupAsPDF(hm, 'CP3234.CP55.interaction.fdr<0.05.heatmap')


ggplot(protQuant, aes(x=paste0(GROUP,SUBJECT), y=LogIntensities)) + geom_boxplot()
```
LogIntensities plot
```{r}
p.mat <- dcast(protQuant, Protein~paste0(ligand,'.',timeStr, rep), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat))


p.mat <- p.mat[rownames(p.mat) %in% sigProteins,]


hm <- Heatmap(p.mat, cluster_columns = FALSE, 
              name = 'Log2 Ints.',
              #column_title = 'CP55 vs CP1066',
              row_title = sprintf('%s Proteins (fdr < 0.05)', nrow(p.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("CP55", colnames(p.mat)), "CP55", "CP3234")))


hm
BackupAsPDF(hm, 'CP3234.CP55.interaction.intensities.fdr<0.05.heatmap')
```



