---
title: "061225_AF3set_dataInspection"
author: "Martin Gordon"
date: "2025-06-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Inspect the files shared by Jyoti and clean if neccessary prior to the AF3 run
Two cell-lines, ~400 ppi in each 
```{r packages}
library(data.table)
library(seqinr)
library(magrittr)
library(ggplot2)
library(seqinr)
library(readxl)

source("../../utils/bp_utils/ManageScriptData.R")
```

```{r read-data}
hek.set <- setDT(readxl::read_xlsx('./data/FluB_PPI.xlsx', sheet=1))
a549.set <- setDT(readxl::read_xlsx('./data/FluB_PPI.xlsx', sheet=2))

hek.set$Bait.x %>% unique()
a549.set$Bait.x %>% unique()
fa.obj <- seqinr::read.fasta(file='./data/ibv.bait.seqs.fa', seqtype = 'AA', as.string = T)
names(fa.obj)
```

we want to create two seperate tables and concatenate them
```{r get-unique-baits}
hek.baits <- unique(hek.set$Bait.x)
a549.baits <- unique(a549.set$Bait.x)
```
Names dont match, I think easiest way to fix is to create a keys DT and manually match the IDs, then merge this to our bait-prey table 
```{r merge-dts}
keys.dt <- data.table(fa.header=c('up|I0B7A3|Hemagglutinin_IBV', 'up|Q596H2|Hemagglutinin_IBV', 'up|C0LT34|Neuraminidase_IBV', 'up|Q596G9|Neuraminidase_IBV','up|I0B7B2|PB1_IBV','up|Q596H4|PB1_IBV',
                                  'up|I0B7B1|PA_IBV', 'up|Q596H3|PA_IBV', 'up|I0B7A8|Nucleoprotein_IBV', 'up|Q596H1|Nucleoprotein_IBV','up|I0B7A4|Matrix_protein','up|Q596G8|Matrix_protein',
                                  'up|I0B7B0|Nuclear_export_protein_IBV', 'up|Q596G6|Nuclear_export_protein_IBV','up|I0B7A9|NS1_IBV','up|Q596G5|NS1_IBV','up|C0LT33|Glycoprotein_NB_IBV','up|Q596H0|Glycoprotein_NB_IBV',
                                  'up|I0B7A5|BM2_IBV', 'up|Q596G7|BM2_IBV', "up|I0B7B0|Nuclear_export_protein_IBV"),
                      bait.name =c('Brisbane_HA', 'Victoria_HA', 'Brisbane_NA', 'Victoria_NA','Brisbane_PB1', 'Victoria_PB1',
                                   'Brisbane_PA', 'Victoria_PA','Brisbane_NP', 'Victoria_NP', 'Brisbane_M1', 'Victoria_M1',
                                   'Bribane_NEP', 'Victoria_NS2','Brisbane_NS1','Victoria_NS1','Brisbane_NB', 'Victoria_NB',
                                   'Brisbane_BM2', 'Victoria_BM2', 'Brisbane_NS2'
                                   ))


# same set now!
setdiff(unique(apms.dt$bait_id), keys.dt$bait.name)
setdiff(keys.dt$bait.name, unique(apms.dt$bait_id))

keys.dt[, uniprot := tstrsplit(fa.header, '[|]', keep=c(2))]
setkey(keys.dt, bait.name)

apms.dt <- rbind(hek.set[, .(bait_id=Bait.x, prey=Prey.x)], a549.set[,.(bait_id=Bait.x, prey=Prey.x)]) %>% 
  unique()
setkey(apms.dt, bait_id)


keys.dt[bait.name == 'Brisbane_NS2']
apms.dt[bait_id == 'Brisbane_NS2',]
apms.dt[bait_id == 'Brisbane_NS2',] %>% 
  .[keys.dt, .(.I, uniprot, prey)]

# merge and write
apms.dt[bait_id %in% c('Brisbane_NS2', 'Victoria_PA')]
keys.dt


missing.set <- merge(apms.dt[bait_id %in% c("Brisbane_NS2","Victoria_PA")], keys.dt, by.x='bait_id', by.y='bait.name')[, .(.I, uniprot, prey)]
fwrite(missing.set, sep=',', col.names = F, ScriptAndDatedFileName('hiv.missingSet.csv'))
fwrite(apms.dt[keys.dt, .(.I, uniprot, prey)], sep=',', col.names = F, ScriptAndDatedFileName('hiv.fullbaitSet.csv'))


```

get the list of unique bait ids, wirte to file and download these fa records from uniprot website

```{r get-prey-ids}
#fwrite(data.table(unique(apms.dt$prey)), col.names = F, ScriptAndDatedFileName('hiv.apms.preyuniprots.txt')
```
```{bash get-prey-fasta, eval=FALSE}
python ~/Documents/utils/mg_utils/py_utils/fetchFastafromUniProt.py \
--input '~/Documents/projects/061225_JBatra_HostViralSecondSet_AF3/061225_AF3set_dataInspection_data/2025_06_12_hiv.apms.preyuniprots.txt' \
--output  '~/Documents/projects/061225_JBatra_HostViralSecondSet_AF3/061225_AF3set_dataInspection_data/2025_06_12_hiv.apms.preyuniprots.fa'
```


missing baits in fasta file... ask Jyoti to share remainder and run this set for now 
```{r missing-baits}
setdiff(union(x=a549.baits, y=hek.baits), detected.set)  
c('up|N0BQ34|PB2_IBV','up|Q596H5|PB2_IBV')
```

## 061425

Review the set of ppi and see how it scores

```{r}
ppi.set <- fread('~/Documents/projects/061225_JBatra_HostViralSecondSet_AF3/061225_AF3set_dataInspection_data/2025_06_12_hiv.baitSet2.csv')

apms.dt <- rbind(hek.set[, .(bait_prey=Bait_Prey, bait_id=Bait.x, prey=Prey.x, prey_gene=PreyGene.x)], a549.set[,.(bait_prey=Bait_Prey, bait_id=Bait.x, prey=Prey.x, prey_gene=PreyGene.x)]) %>% 
  unique()
apms.dt
apms.dt[grepl('Brisbane_NS2', bait_prey)]


names(fa.obj)
```

