---
title: "070725_NP_InteractorsInspection"
author: "Martin Gordon"
date: "2025-07-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Invesitgate the results for the NP interactor set from AF3; only one of the proteins passed the expected threshold
What is the proportion of the different proteins that passed the AF3 score

```{r packages}
library(data.table)
library(seqinr)
library(magrittr)
library(ggplot2)
library(seqinr)
library(readxl)
library(ggforce)
library(ggrepel)

source("../../utils/bp_utils/ManageScriptData.R")
```



```{r}
af3.scores <- fread('/Users/martingordon/Documents/projects/061225_JBatra_HostViralSecondSet_AF3/062325_AFscoresInspection_data/2025_06_23_allAF3Scores.tidy.csv')
af3.scores[, virus := tstrsplit(ppi_gene, '_', keep=1)]

paf1 <- toupper(c('Paf1', 'Ctr9', 'Cdc73', 'Rtf1', 'Leo1'))
paf1

af3.scores[, paf1c := 'no']
af3.scores[gene2 %in% paf1, paf1c := 'yes']

g <- ggplot(af3.scores[grepl('NP$', gene1),],  aes(x=virus, y=iptm, label=ppi_gene, color=paf1c)) + 
  geom_hline(yintercept=0.5) +
  geom_violin() +
  geom_sina() +
  geom_text_repel(data=af3.scores[grepl('_NP$', gene1) & gene2 %in% paf1,])
g
```
Some of the PPI score well, but most score very poorly likely a problem with modelling the NP protein
```{r}
g <- ggplot(af3.scores[grepl('NP$', gene1),],  aes(x=virus, y=iptm, label=ppi_gene, color=paf1c)) + 
  geom_hline(yintercept=0.5) +
  geom_violin() +
  geom_sina() +
  geom_text_repel(data=af3.scores[grepl('_NP$', gene1) & gene2 %in% paf1,])
g
```

Remove LEO1; nothing scores above 0.2. no indication fo high quality interactors
```{r}
g <- ggplot(af3.scores[grepl('NP$', gene1) & gene2 != 'LEO1',],  aes(x=virus, y=iptm, label=ppi_gene, color=paf1c)) + 
  geom_hline(yintercept=0.5) +
  geom_violin() +
  geom_sina() 
g
```

