---
title: "030325_PHSanityCheck"
author: "Martin Gordon"
date: "2025-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Considering the large number of DE sites, I want to process this w/o artMS, run the comparisons to ensure the n sig hits are comparable before sharing hte results
```{r packages}
library(data.table)
library(ggplot2)
library(magrittr)
library(ggrepel)
library(stringr)
library(ComplexHeatmap)
library(circlize)
library(MSstats)
library(viridis)
library(ggbeeswarm)
library(RColorBrewer)
library(hrbrthemes) # visually nice set of themes
library(patchwork)
library(showtext)
library(artMS)
library(readxl)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/spectronautFile2ArtMS.R")

source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")

# kinase enrichment analysis
source("/Users/martingordon/Documents/utils/bp_utils/KinaseActivityScores.R")


# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}

#set one
col.pal <- getQualitativePalette(n=13)
col.pal <- randomcoloR::distinctColorPalette(k=13)

# fonts needed for hrbrthemes
library(extrafont)
font_import()
loadfonts()

# directly from google font
sysfonts::font_add_google("Roboto Condensed")
showtext_auto()

customTheme <- theme_ipsum_rc(base_family = "serif") +
  theme(panel.border = element_rect(color = "lightgrey", 
                                    fill = NA, 
                                    size = 1),
        axis.text.x = element_text(angle=90,size=6)
        )
#set one
col.pal <- getQualitativePalette(n=13)
```

Read in the files with the modified sequences and protein sites named

```{r}
#these just include the modified sites 
spec.list <- list(MRC5=fread('./022525_PH_pwComparisons_data/MRC5/evidence.PH.txt'),
                  RFe=fread('./022525_PH_pwComparisons_data/RFe/evidence.PH.txt')
                  )

keys.list <- list(MRC5=fread('./022525_PH_pwComparisons_data/MRC5/keys.txt'),
                  RFe=fread('./022525_PH_pwComparisons_data/RFe/keys.txt')
                  )

# combine the two lists element wise
comb.list <- mapply(function(spec, key) merge(x=spec, y=key, by='RawFile'), spec.list, keys.list, SIMPLIFY = F)
```
Actually, I think the safest method  may be to just merge the sequences on Protein
```{r}

input.list <- list(MRC5=fread('~/Documents/projects/121924_JBatra_SARS2_HuBatComparison_ABPHProteomics/123024_PH_InitialQC_data/2025_02_25_MRC5.mssInput.csv'),
                  RFe=fread('~/Documents/projects/121924_JBatra_SARS2_HuBatComparison_ABPHProteomics/123024_PH_InitialQC_data/2025_02_25_Rfe.mssInput.csv'))


# create necessary cols and simplify 
comb.list <- lapply(comb.list, function(x){
  # reverse engineering to get MSS input
  x[, `:=`(ProteinName = Proteins,
           PeptideSequence = sequence,
           #PeptideSequence =gsub('[(][a-z]+[)]' ,'', `Modified sequence`),
           PeptideModifiedSequence = `Modified sequence`,
           PrecursorCharge = Charge,
           FragmentIon = NA,
           ProductCharge = NA
           )]
  
  
  x <- x[, .(Condition, Run=RawFile, BioReplicate=as.numeric(gsub('.+[.]' ,'', BioReplicate)), ProteinName, PeptideSequence=PeptideModifiedSequence, FragmentIon,
             PrecursorCharge, Intensity, ProductCharge, IsotopeLabelType)]
  return(x)
})


dp.list <- lapply(comb.list, function(x){
  
  dp.out <- MSstats::dataProcess(x, 
                                 MBimpute =  FALSE, 
                                 normalization = 'equalizeMedians',
                                 summaryMethod = "TMP")

  return(dp.out)
})

```

```{r}
lapply(names(dp.list), function(x){
  
  saveRDS(dp.list[[x]], ScriptAndDatedFileName(paste0(x, 'dataProc.rds')))
})
```

```{r}
dp.list <- list(MRC5=readRDS('~/Documents/projects/121924_JBatra_SARS2_HuBatComparison_ABPHProteomics/030325_PHpwComparisonsSanityCheck_data/2025_03_03_MRC5dataProc.rds'),
                RFe=readRDS('~/Documents/projects/121924_JBatra_SARS2_HuBatComparison_ABPHProteomics/030325_PHpwComparisonsSanityCheck_data/2025_03_03_RFedataProc.rds'))


# look fine 
lapply(dp.list, function(x){
  
  ggplot(x$ProteinLevelData, aes(x=paste0(GROUP,SUBJECT), y=LogIntensities, fill=GROUP)) +
    geom_boxplot()
})
```
Quickly want to run PW comparisons to ensure results are not v far off what we have recovered before sharing

```{r}
contrasts.oi <- fread('~/Documents/projects/121924_JBatra_SARS2_HuBatComparison_ABPHProteomics/022525_PH_pwComparisons_data/RFe/contrast.txt', header = F)

input <- split(contrasts.oi, contrasts.oi$V1)

input <- lapply(input, function(x){
  
  cond <- unlist(strsplit(x$V1, '[-]'))
  num <- cond[1]; denom <- cond[2]
  return(c(num,denom))
})

contrast.mat <- MSstatsContrastMatrix(contrasts=input, conditions = unique(dp.list[[2]]$ProteinLevelData$GROUP))
```

Run the pw comparisons
```{r}
ms.list <- lapply(names(dp.list), function(x){
  
  obj <- dp.list[[x]]

  if (x == 'MRC5'){
    cmat <- contrast.mat[!grepl('06hpi', rownames(contrast.mat)),!grepl('06hpi', colnames(contrast.mat))]
  } else {
    cmat <-  contrast.mat
  }

  print(obj$ProteinLevelData)
  p.quant <- setDT(obj$ProteinLevelData)
  #p.quant[, SUBJECT := factor(paste0(GROUP, '.', SUBJECT))]
  #p.quant[, GROUP := factor(GROUP)]

  obj$ProteinLevelData <-  p.quant
  message('runnning MS contrasts on ', x)
  mss <- groupComparison(contrast.matrix=cmat, data=obj)
  mss.dt <- setDT(mss$ComparisonResult)
  return(mss.dt)
})
names(ms.list) <- c('MRC5', 'RFe')
```
save to file, create a sig column and compare the number of hits across conditions 
```{r}
#fwrite(rbindlist(ms.list, idcol='cell_line'), ScriptAndDatedFileName('mss.pwcomparisons.customRun.csv.gz'))

mss.dt <- fread('~/Documents/projects/121924_JBatra_SARS2_HuBatComparison_ABPHProteomics/030325_PHpwComparisonsSanityCheck_data/2025_03_03_mss.pwcomparisons.customRun.csv.gz')
mss.dt[, gene := multiUniprotSites2multiGeneSites(Protein, species = 'HUMAN')]
mss.dt[, p.adj := p.adjust(pvalue, method='BH'), by=.(Label)]
mss.dt[, sig := 'not']
mss.dt[abs(log2FC) > 0.58 & p.adj < 0.05 & !issue %in% c("oneConditionMissing","completeMissing"), sig := ifelse(log2FC > 0, 'up', 'down')]
# remove modifications
mss.dt[, simplifiedProtein := gsub('[_][STY][0-9]+', '', Protein)]
mss.dt[, sites := lapply(.SD, function(x) paste0(unlist(str_extract_all(x,  '[_][STY][0-9]+')), collapse=';')), .SDcols = 'Protein', by=.I]
```
just get the n hits in eahc 
Results look identical, which is great! Why so many differences in comparison to the abundance data? Sure we have enriched, but no evidence of this DE in abundance?
```{r}
g <- ggplot(mss.dt[sig != 'not',.N, by=.(cell_line, sig, Label)], aes(x=reorder(Label,-N), y=N, fill=sig)) +
  geom_bar(stat='Identity') +
  ggtitle('Number of significant hits per contrast', "abs LFC > 0.58 & adj.pval < 0.05") +
  scale_fill_manual(values=col.pal) +
  facet_wrap(cell_line~.) +
  customTheme
g
```

```{r}

```

