---

title: "PWcomparisons_LM"
author: "Martin Gordon"
date: "2024-05-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As an alternative to the MAGECK pipeline, try the approach favoured by the other Genome-wide screen assays
Read in the raw count tables, drop  log normalize

Not the output I expected at all... still seems that most of the contrasts comparing the group to the treatment groups to baseline at 

## packages
```{r}
library(magrittr)
library(data.table)
library(circlize)
library(ComplexHeatmap)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(ggrepel)
library(ggridges) # ridgeline plots
library(eulerr) # for venn diagram plots
library(ggvenn)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/enrichmentTestFunctions.R")
source ("../../utils/mg_utils/r_utils/IDmapping.R")
source ("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("/Users/martingordon/Documents/utils/bp_utils/LinearModels.R")


# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0,...){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```
read in the count matrices, log normalize and perform the PW comparisons

```{r}
counts.dt <- fread('/Users/martingordon/Documents/projects/032124_RBabu_CRISPRMutantScreen/output/mageck_count_comb_sgRNA/240424_RBabu_CRISPRe.count.txt')
norm.dt <- fread('/Users/martingordon/Documents/projects/032124_RBabu_CRISPRMutantScreen/output/mageck_count_comb_sgRNA/240424_RBabu_CRISPRe.count_normalized.txt')
```
Perform lognormalization of the raw counts
```{r}
# convert to long format 
counts.long <- melt(counts.dt, id.vars = c('sgRNA', 'Gene'), value.name = 'counts', variable.name = 'sample')
counts.long[, logNorm_counts := round(log2(counts/sum(counts)*1e+06 + 1),2), by=sample]
```
Add additional annotation info to the dt
```{r}
counts.long[, .N, .(cas9, sample, group)]

counts.long[, group := sub("_[0-9]{1,2}$","", sample)]
counts.long[, cas9 := ifelse(grepl('Abe8e|Bemax', group), str_extract(sample, 'Abe8e|Bemax'), 'Ctrl')]

counts.long[, .N, .(cas9, sample, group)]

# fix some of the labels
counts.long[group %like% 'Bemax_Lib', group := 'Bemax_Lib']
counts.long[group %like% 'Abe8e_Lib', group := 'Abe8e_Lib']

counts.long[, sample_no := str_extract(sample, "[0-9]{1,2}$")]
counts.long[, treatment := sub("Abe8e_|Bemax_|Abe8eDMSO_|Abe8eSerum_|BemaxDMSO_|BemaxSerum_", "", group)]
```
Get how many sd each sgRNA expression is from the sample mean
Filter out groups +/- 3SD from mean in the plasmid condition
```{r}
counts.long[, logNorm_zscore := scale(counts), by=.(sample)]

# these are the 
drop.sgRNA %>%  length()
drop.sgRNA <- counts.long[sample == 'plasmid_BE_38' & abs(logNorm_zscore) > 3, sgRNA]

counts.long[sample == 'plasmid_BE_38' & abs(logNorm_zscore) > 3,]

counts.mat <- dcast(counts.long, interaction(sgRNA)~sample, value.var = 'counts') %>% 
  as.matrix(rownames='sgRNA')
```

I guess for our PW comparisons, we want to compare all conditions to plasmid, then do a contrast of contrasts to compare the difference in differences and I guess this is what we want to work with

filter out noisy sgRNA by i) removing sgRNA with extreme counts (+/- 3SD) in the plasmid condition 
```{r}
counts.f <- counts.long[!sgRNA %in% drop.sgRNA,]

zeros.sgRNA <- counts.f[counts == 0, .N, by=.(sgRNA, group)] %>%
  .[group =='plasmid_BE', sgRNA]

counts.f <- counts.f[!sgRNA %in% zeros.sgRNA,]
```

# now that we have the normalised counts, lets run LM of each of the groups

Setup a custom contrast comparing each group against plasmid

```{r}
counts.f$treatment %>%  unique()
counts.f[, treatment := factor(treatment, levels=c("plasmid_BE","Lib","gRNA_noBE","CTRL1","CTRL2","CTRL3","GPN","TORIN","BAFA1","STARVE","DOXO"))]
counts.f[, cas9 := factor(group,levels=c('Ctrl',"Bemax" ,"Abe8e"))]
counts.f[, group := factor(group, levels=c("plasmid_BE","Bemax_Lib","Abe8e_Lib","gRNA_noBE","BemaxDMSO_CTRL1","BemaxDMSO_CTRL2","Bemax_GPN","Bemax_TORIN","Bemax_BAFA1",
                                              "BemaxSerum_STARVE","BemaxDMSO_CTRL3","Bemax_DOXO","Abe8eDMSO_CTRL1","Abe8eDMSO_CTRL2","Abe8e_GPN","Abe8e_TORIN","Abe8eSerum_STARVE",
                                              "Abe8e_BAFA1","Abe8eDMSO_CTRL3","Abe8e_DOXO"))]
```

Create a custom contrast matrix of the contrasts we want
```{r}
contrast.mat <- sapply(levels(counts.f$group), function(x){
  sample.vec <- rep(0, length(levels(counts.f$group)))
  sample.vec[1] <-  -1 
  sample.idx <- which(levels(counts.f$group) == x)
  sample.vec[sample.idx] <-  1
  return(sample.vec)
}) %>%  cbind()

contrast.mat

colnames(contrast.mat) <- paste0(colnames(contrast.mat), '-plasmid_BE')
contrast.mat <- contrast.mat[,-1] %>% 
  as.matrix()

rownames(contrast.mat)  <- levels(counts.f$group)
```

Run the PW comparisons


```{r}

list(contrast.mat)

# interaction contrasts functions
em.interactionContrast <-  function (l, factorFormula = ~group){  # assuming an interaction of timepoint and status
  emm <- emmeans::emmeans(l, factorFormula) # obtain the estimated marginal means for each timepoint per group (WT, mut, treatment) from the model
 # contrast1 <- contrast(emm, interaction = c("pairwise"), by = NULL) #interaction contrast, or contrasts of contrasts; compare response between groups
  contrast1 <- contrast(emm, method=as.list(as.data.table(contrast.mat)))
  return (as.data.table(contrast1))
}

em.contrastOfContrasts <- function (l, factorFormula = ~status|timepoint){  # how does the drug effect change per tissue
  emm <- emmeans(l, factorFormula)
  contrast1 <- pairs(emm)
  contrast2 <- pairs(contrast1, by = NULL, adjust = "none")
  return (as.data.table(contrast2))
}

formulas = list(full = logNorm_counts~group)


## testing 
lm.out <- linearModelsAllProteins(counts.f, 
                                  formulaList = formulas, 
                                  splitColumn = "sgRNA", postProcessFunction = em.interactionContrast, cl = 12) |> suppressWarnings()


lm.out$postProcess[, p.adj := p.adjust(p.value, method='BH'), by=contrast]
lm.out$postProcess[abs(estimate) > 0.58 & p.value < 0.001,] %>% 
  .[order(p.value)]

```

22-05-24
----
Recovering very few hits; possibly due to dynamic range of log2 normalized scores??
Need to use the normalized data to run the PW comparisons
Use the control normalized data for the comparisons


```{r}
norm.dt <- fread('/Users/martingordon/Documents/projects/032124_RBabu_CRISPRMutantScreen/output/mageck_count_comb_sgRNA/240424_RBabu_CRISPRe.count_normalized.txt')
```


```{r}
# convert to long format 
counts.long <- melt(norm.dt, id.vars = c('sgRNA', 'Gene'), value.name = 'counts', variable.name = 'sample')

ggplot(counts.long, aes(x=sample, y=log2(counts+1))) +
  geom_boxplot()
```
Add additional annotation info to the dt
```{r}
counts.long[, cas9 := ifelse(grepl('Abe8e|Bemax', sample), str_extract(sample, 'Abe8e|Bemax'), 'Ctrl')]
counts.long[, group := sub("_[0-9]{1,2}$","", sample)]

counts.long[, .N, .(cas9, sample, group)]

# fix some of the labels
counts.long[group %like% 'Bemax_Lib', group := 'Bemax_Lib']
counts.long[group %like% 'Abe8e_Lib', group := 'Abe8e_Lib']

counts.long[, sample_no := str_extract(sample, "[0-9]{1,2}$")]
counts.long[, treatment := sub("Abe8e_|Bemax_|Abe8eDMSO_|Abe8eSerum_|BemaxDMSO_|BemaxSerum_", "", group)]
```
Get how many sd each sgRNA expression is from the sample mean
Filter out groups +/- 3SD from mean in the plasmid condition
```{r}
counts.long[, NormCounts_zscore := scale(counts), by=.(sample)]

# these are the 
drop.sgRNA <- counts.long[sample == 'plasmid_BE_38' & abs(NormCounts_zscore) > 3, sgRNA]
drop.sgRNA %>%  length()

counts.long[sample == 'plasmid_BE_38' & abs(NormCounts_zscore) > 3,]

counts.mat <- dcast(counts.long, interaction(sgRNA)~sample, value.var = 'counts') %>% 
  as.matrix(rownames='sgRNA')
```

I guess for our PW comparisons, we want to compare all conditions to plasmid, then do a contrast of contrasts to compare the difference in differences and I guess this is what we want to work with

filter out noisy sgRNA by i) removing sgRNA with extreme counts (+/- 3SD) in the plasmid condition 
```{r}
counts.f <- counts.long[!sgRNA %in% drop.sgRNA,]

zeros.sgRNA <- counts.f[counts == 0, .N, by=.(sgRNA, group)] %>%
  .[group =='plasmid_BE', sgRNA]

counts.f <- counts.f[!sgRNA %in% zeros.sgRNA,]
```

# now that we have the normalised counts, lets run LM of each of the groups

Setup a custom contrast comparing each group against plasmid

```{r}
counts.f[, treatment := factor(treatment, levels=c("plasmid_BE","Lib","gRNA_noBE","CTRL1","CTRL2","CTRL3","GPN","TORIN","BAFA1","STARVE","DOXO"))]
counts.f[, cas9 := factor(group,levels=c('Ctrl',"Bemax" ,"Abe8e"))]
counts.f[, group := factor(group, levels=c("plasmid_BE","Bemax_Lib","Abe8e_Lib","gRNA_noBE","BemaxDMSO_CTRL1","BemaxDMSO_CTRL2","Bemax_GPN","Bemax_TORIN","Bemax_BAFA1",
                                              "BemaxSerum_STARVE","BemaxDMSO_CTRL3","Bemax_DOXO","Abe8eDMSO_CTRL1","Abe8eDMSO_CTRL2","Abe8e_GPN","Abe8e_TORIN","Abe8eSerum_STARVE",
                                              "Abe8e_BAFA1","Abe8eDMSO_CTRL3","Abe8e_DOXO"))]
```

Run the PW comparisons on just the normalized data
```{r}
# interaction contrasts functions
em.interactionContrast <-  function (l, factorFormula = ~group){  # assuming an interaction of timepoint and status
  emm <- emmeans::emmeans(l, factorFormula) # obtain the estimated marginal means for each timepoint per group (WT, mut, treatment) from the model
 # contrast1 <- contrast(emm, interaction = c("pairwise"), by = NULL) #interaction contrast, or contrasts of contrasts; compare response between groups
  contrast1 <- contrast(emm, method=as.list(as.data.table(contrast.mat)))
  return (as.data.table(contrast1))
}


em.contrastOfContrasts <- function (l, factorFormula = ~status|timepoint){  # how does the drug effect change per tissue
  emm <- emmeans(l, factorFormula)
  contrast1 <- pairs(emm)
  contrast2 <- pairs(contrast1, by = NULL, adjust = "none")
  return (as.data.table(contrast2))
}

formulas = list(full = counts~group)


## testing 
lm.out <- linearModelsAllProteins(counts.f, 
                                  formulaList = formulas, 
                                  splitColumn = "sgRNA", postProcessFunction = em.interactionContrast, cl = 12)

lm.out$postProcess

lm.out$postProcess[, p.adj := p.adjust(p.value, method='BH'), by=contrast]
lm.out$postProcess[abs(estimate) > 1 & p.value < 0.01,] %>% 
  .[order(p.value)]

lm.out$postProcess[, sig := 'not']
lm.out$postProcess[abs(estimate) > 0.58 & p.value < 0.01, sig := ifelse(abs(estimate) > 0, 'up', 'down')]
```

```{r}
lapply(unique(lm.out$postProcess$contrast), function(x){
  
  ggplot(lm.out$postProcess[contrast == x,], aes(x=estimate, y=-log10(p.value), color=sig, label=sgRNA)) +
    geom_point() +
    geom_text_repel(data=lm.out$postProcess[contrast == x & sig != 'not',]) +
    ggtitle(x)
  
})
```
```{r}
counts.f[sgRNA == 'BE_5341' & group %in% c('plasmid_BE', 'BemaxDMSO_CTRL3'),]
```
Maybe a single guide RNA condition is not a good baseline comparison... 
Lets try compare the different groups


Lets try BeMax Lib and Abe8e Lib as the baseline groups
Split the datatables into two and create two contrast matricies

```{r}
counts.f[, cas9 := ifelse(grepl('Abe8e|Bemax', group), str_extract(sample, 'Abe8e|Bemax'), 'Ctrl')]
counts.f[,.N, by=.(group,cas9)]

counts.list <- split(counts.f, counts.f$cas9)

counts.list[['Abe8e']] <-  rbind(counts.list[['Abe8e']], counts.list[['Ctrl']])
counts.list[['Bemax']] <-  rbind(counts.list[['Bemax']], counts.list[['Ctrl']])
counts.list[['Ctrl']] <-  NULL

counts.list[['Abe8e']][, group := factor(group, levels=c('Abe8e_Lib', 'gRNA_noBE', 'plasmid_BE', 'Abe8eDMSO_CTRL1', 'Abe8eDMSO_CTRL2', 'Abe8eDMSO_CTRL3',
                                          'Abe8e_GPN', 'Abe8e_TORIN', 'Abe8eSerum_STARVE', 'Abe8e_BAFA1', 'Abe8e_DOXO'))]

counts.list[['Bemax']][, group := factor(group, levels=c('Bemax_Lib', 'gRNA_noBE', 'plasmid_BE', 'BemaxDMSO_CTRL1', 'BemaxDMSO_CTRL2', 'BemaxDMSO_CTRL3',
                                          'Bemax_GPN', 'Bemax_TORIN', 'BemaxSerum_STARVE', 'Bemax_BAFA1', 'Bemax_DOXO'))]


# contrast matrices
contrast.list <- lapply(counts.list, function(x){
  
  contrast.mat <- sapply(levels(x$group), function(i){
  sample.vec <- rep(0, length(levels(x$group)))
  sample.vec[1] <-  -1 
  sample.idx <- which(levels(x$group) == i)
  sample.vec[sample.idx] <-  1
  return(sample.vec)
  
  }) %>%  cbind()
  
  rownames(contrast.mat) <- levels(x$group)
  colnames(contrast.mat) <- paste0(levels(x$group), '-', levels(x$group)[1])
                             
  return(contrast.mat[,-1])
})
  
# check names are the same 
names(counts.list) == names(contrast.list)
```
Run the comparisons vs the Library at D0 (with BE added)
Want to contrast all groups at D0
Lets see how the output looks, then we can try a contrast of contrasts

```{r}
library(scales)
# interaction contrasts functions
# return both results from individual contrasts and also contrast of contrast
formulas = list(full = counts~group)

lm.list <- lapply(seq_along(counts.list), function(x,n,i){
  
  editor <- n[[i]]
  print(paste0('Working on editor ', editor, '....'))
  
  # set up the interaciton contrast function
  em.interactionContrast <-  function (l, factorFormula = ~group){  # assuming an interaction of timepoint and status
    emm <- emmeans::emmeans(l, factorFormula) # obtain the estimated marginal means for each timepoint per group (WT, mut, treatment) from the model
    contrast1 <- as.data.table(contrast(emm, method=as.list(as.data.table(contrast.list[[editor]]))))
    return (contrast1)
  }
  
  lm.out <- linearModelsAllProteins(x[[editor]], 
                                  formulaList = formulas, 
                                  splitColumn = "sgRNA", postProcessFunction = em.interactionContrast, cl = 12)
  
  print(paste0('Finshed working on editor ', editor, '....'))
  return(lm.out)
  
},x=counts.list,n=names(counts.list))

names(lm.list) <- names(counts.list)


lapply(lm.list, function(x){
  
  dt <- copy(x$postProces)
  dt[, sig := 'not']
  dt[p.value < 0.01 & abs(estimate) > 0.58, sig := ifelse((estimate) > 0, 'up', 'down')]
  
  lapply(unique(dt$contrast), function(i){
   g <-   ggplot(dt[contrast==i,], aes(x=estimate, y=-log10(p.value), color=sig, label=sgRNA)) +
    geom_vline(xintercept=c(-0.58,0.58), linetype=2, alpha=0.4) +
    geom_hline(yintercept=c(-log10(0.01)), linetype=2, alpha=0.4) +
    geom_text_repel(data=dt[contrast==i & sig != 'not']) +
    geom_point() +
    xlab('Fold Change') +
    scale_color_manual(values=c('down'=muted('blue'), 'up'=muted('red'), 'not'='grey')) +
    ggtitle(i) +
    theme_bw()
   
   BackupAsPDF(g, paste0(i, 'lm.Lib_baseline.volcanoplot'))
  })
})
```

Save these output tables

```{r}
lm.list[['Abe8e']]$postProcess[, fdr := p.adjust(p.value, method='BH'), by=contrast]
lm.list[['Bemax']]$postProcess[, fdr := p.adjust(p.value, method='BH'), by=contrast]

fwrite(lm.list[['Abe8e']]$anova, ScriptAndDatedFileName('abe8e.Lib_baseline_anova.csv'))
fwrite(lm.list[['Bemax']]$anova, ScriptAndDatedFileName('bemax.Lib_baseline_anova.csv'))
fwrite(lm.list[['Abe8e']]$postProcess, ScriptAndDatedFileName('abe8e.Lib_baseline_LMContrasts.csv'))
fwrite(lm.list[['Bemax']]$postProcess, ScriptAndDatedFileName('bemax.Lib_baseline_LMContrasts.csv'))
```

