---
title: "082624_plotRequests"
author: "Martin Gordon"
date: "2024-08-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load in the packages and prepare the files requested by Jyoti

```{r packages}

library(DESeq2)
library(data.table)
library(magrittr)
library(ggplot2)
library(stringr)
library(circlize)
library(ComplexHeatmap)
library(ggrepel)
library(usedist) #package for working with distances
library(ggh4x) # additional functionality for ggplot2 obj; eg facet_grid allow x/y axis to vary
library(scales)
library(patchwork)
library(RColorBrewer)
library(cluster) # pam clustering of genes
library(eulerr) # eulerr plot 
library(ggvenn)
library(viridis)
library(readxl)
library(viridis)

library(biomaRt)

# RNAseq DE functions
library(tximport)
library(DESeq2)
library(IHW)

source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/enrichmentTestFunctions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping

source("../../utils/mg_utils/r_utils/IDmapping.R")
source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")
source("../../utils/mg_utils/r_utils/HelperFunctions.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0,...){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}

redbluColpal <- c('#D01B1B', '#FF4242', '#FFFFFF', '#95D2EC','#47abd8')
```

Load relevant datasets
```{r}
id.mapping.dt <- read_xlsx('./docs/human_to_RFE_gene_conversion_table_FINAL.xlsx')

isGenes <- fread('/Users/martingordon/Documents/projects/022624_AViDD_AB_PH_data/docs/ISGs.txt', header=F) %>% 
  .[,V1]

viralGenes <- c('orf1ab', 'S', 'orf3a', 'E', 'M',  'orf6', 'orf7a', 'orf8', 'N', 'orf10')

# need to highlight these 3 genes in the heatmap
proinflam <- readxl::read_xlsx('./docs/List of ISGs and cytokines.xlsx', sheet=1, col_names = T)$Proinflammatory
isg <- readxl::read_xlsx('./docs/List of ISGs and cytokines.xlsx', sheet=2, col_names = T)$ISG
ckines <- readxl::read_xlsx('./docs/List of ISGs and cytokines.xlsx', sheet=3, col_names = T)$GENES


# id mapper
idMap.txt <- readxl::read_xlsx('./docs/human_to_RFE_gene_conversion_table_FINAL.xlsx', sheet=2) %>% 
  setDT()

genes.oi <- c(isg, proinflam, ckines)

# some overlap in the sets
length(genes.oi)
length(unique(genes.oi))
```
metadata

```{r}
meta.data <- data.table(sample=dir('./output/salmonOut', full.names = F, pattern='quant.sf'))
meta.data[, `:=`(host=ifelse(grepl('MRC5', sample), 'human', 'bat'),
                 timepoint=ifelse(grepl('Mock', sample), '', str_extract(sample, '[0-9]{1,2}h')),
                 virus=ifelse(grepl('Mock', sample), 'Mock', gsub('MRC5_|RFe_|_[0-9]{1,2}h.+', '', sample))
                 )]

# now want to set T0 and Mock as baseline for factor levels
meta.data[, `:=`(virus=factor(virus, levels=c('Mock','WA', 'N_P80T', '9bI_N_P80T')),
                 timepoint=factor(timepoint, levels=c('','6h', '12h', '24h', '48h'))
                 )]


# another fix; tp and virus are confounded for Mock, so we need to create a 'group' variable 
meta.data[, condition := factor(ifelse(virus=='Mock', 'Mock', paste(virus, timepoint, sep='.')))]
meta.data[, condition := relevel(condition, ref='Mock')]

bat.meta <- data.frame(meta.data[host=='bat',], row.names = 'sample')
hu.meta <- data.frame(meta.data[host=='human',], row.names = 'sample')

# reset the tp levels for human as missing 6hr
hu.meta$timepoint <-  factor(hu.meta$timepoint, levels=c('12h', '24h', '48h'))
bat.meta$timepoint <-  factor(bat.meta$timepoint, levels=c('6h', '12h', '24h', '48h'))
```

Count matrices 
Also include the output with the viral counts

```{r}
dds.bat <- readRDS('~/Documents/projects/080624_JBatra_SARS2RNAseqHuBat/080824_PWcomparisons_data/2024_08_13_dds.bat.vsMock.rds')
dds.hu <- readRDS('~/Documents/projects/080624_JBatra_SARS2RNAseqHuBat/080824_PWcomparisons_data/2024_08_13_dds.human.vsMock.rds')

# get count matrices with viral counts
dds.viral.bat <- readRDS('~/Documents/projects/080624_JBatra_SARS2RNAseqHuBat/viralProteinPlots_data/2024_08_26_dds.bat.vsMock.featureCountswVirus.rds')
dds.viral.hu <- readRDS('~/Documents/projects/080624_JBatra_SARS2RNAseqHuBat/viralProteinPlots_data/2024_08_26_dds.human.vsMock.featureCountswVirus.rds')
```

**Creating output tables**
PWcontrasts
(write out the contrasts in wide format)
```{r}
tp.dt <- fread('~/Documents/projects/080624_JBatra_SARS2RNAseqHuBat/PWComparisonsTimepoint_data/2024_08_13_PWcomparisons.viralTimepoint.csv')
vsMock.dt <- fread('~/Documents/projects/080624_JBatra_SARS2RNAseqHuBat/080824_PWcomparisons_data/2024_08_08_PWcomparisonsvsMock.PostProc.csv')

# combine the two datasets to make it easier to analyse
tp.dt[, Label := 'SARS2 strain contrasts']
vsMock.dt[, Label := 'Vs Mock contrasts']

comb.dt <- rbind(tp.dt, vsMock.dt)

# tp 
fwrite(dcast(tp.dt[species=='human',], gene~contrast, value.var = c('log2FoldChange', 'lfcSE', 'pvalue', 'padj')), ScriptAndDatedFileName('_humanPWcomparisons.viralTimepoint.wide.csv'))
fwrite(dcast(tp.dt[species=='bat',], gene~contrast, value.var = c('log2FoldChange','lfcSE', 'pvalue', 'padj')), ScriptAndDatedFileName('_batPWcomparisons.viralTimepoint.wide.csv'))

# vs mock
fwrite(dcast(vsMock.dt[species=='human',], gene~contrast, value.var = c('log2FoldChange','lfcSE', 'pvalue', 'padj')), ScriptAndDatedFileName('_humanPWcomparisons.vsMock.wide.csv'))
fwrite(dcast(vsMock.dt[species=='bat',], gene~contrast, value.var = c('log2FoldChange','lfcSE', 'pvalue', 'padj')), ScriptAndDatedFileName('_batPWcomparisons.vsMock.wide.csv'))

```

```{r}
hu.counts <- assay(vst(dds.hu, blind=F))
bat.counts <- assay(vst(dds.bat, blind=F))
```

get the viral protein counts for use with the heatmaps

```{r}
#CoV2
sars2.gtf <- fread(cmd = "grep -v '#' ./output/annoFiles/SARS2_WA.gtf") %>% 
  .[V3 == 'gene', .(ref=V1,start=V4, stop=V5, info=V9)]
sars2.gtf[, c('gene_id', 'gene') := tstrsplit(gsub('\\"|gene_id\\s|gene\\s', '', info), ';', keep=c(1,5))]

viral.genes <- sars2.gtf[, gene_id]

# assay and counts are the same, provided normalize is set to false for counts,
assay(dds.viral.hu)[rownames(assay(dds.viral.hu)) %in% viral.genes,][, ] %>% head()
counts(dds.viral.hu, normalize=T)[rownames(assay(dds.viral.hu)) %in% viral.genes,][, ] %>%  head()

# the rlog and vst transformations are strange with thiese lo counts..seem to impute values... just use log2 + pseudocount
# look into why or how it does this.. definitely dont trust the rlog values, but I gues

# look into why... for now 
# for now take the normalized coutns so we can compare across groups
hu.viral.mat <-  counts(dds.viral.hu, normalized=T)[rownames(counts(dds.viral.hu, normalized=T)) %in% viral.genes,]
bat.viral.mat <- counts(dds.viral.bat, normalized=T)[rownames(counts(dds.viral.bat, normalized=T)) %in% viral.genes,]

# convert the names 
rownames(hu.viral.mat)[rownames(hu.viral.mat) %in% viral.genes] <- sars2.gtf[, .(gsub('\\s','',gene))]$V1
rownames(bat.viral.mat)[rownames(bat.viral.mat) %in% viral.genes] <- sars2.gtf[, .(gsub('\\s','',gene))]$V1

```

Initally finding only ~50% of genes in the dataset; could be due to non-stranded? low coverage etc??
Update: looked at aliases not the issue
Query the org.db structure, filter on the names of the genes in the list
```{r}
library(org.Hs.eg.db)

id.map <- mapIds(org.Hs.eg.db, 
       keys=genes.oi,
       column='ALIAS',
       keytype = 'SYMBOL',multiVals = 'list')


# lets make a datatable of the output
ids.dt <- lapply(id.map, function(x) paste(x, collapse=';')) %>% 
  do.call(rbind, .) %>% 
  as.data.table(., keep.rownames=T)

setnames(ids.dt, new=c('symbol', 'alias'))

# doesnt work as rbindlist expects list of dt or list of list, doesnt know how to handle list of char vectors
##lapply(id.map, function(x) paste(x, collapse=';')) %>% 
#  rbindlist()
```
```{r}
rownames(hu.counts) %in% genes.oi %>%  sum()
# just recovered 10 more genes... looks like we just arent picking up most of these
rownames(hu.counts) %in% do.call(c, strsplit(ids.dt$alias, ';')) %>%  sum()

genes.oi <- unique(c(genes.oi, do.call(c, strsplit(ids.dt$alias, ';'))))
bat.genes.oi <- unique(c(genes.oi, idMap.txt[Human_gene_name %in% genes.oi, RFE_gene_name]))
```

add identifier so can plot these
```{r}
# subset to the genes of interest and write these out as a table
tp.dt[, info := ifelse(gene %in% genes.oi, 'proinflammatory/ISG/chemokine/cytokine', NA)]
vsMock.dt[, info := ifelse(gene %in% genes.oi, 'proinflammatory/ISG/chemokine/cytokine', NA)]
```
plot the requested ISGs and cytokines, I’m particularly interested in looking at specific genes in the attached file. Could you make a heatmap and a wide-format file for these?

```{r}
# all tables with the different gene sets
fwrite(dcast(vsMock.dt[species=='bat' & !is.na(info),], gene~contrast, value.var = c('log2FoldChange', 'lfcSE', 'pvalue', 'padj')), ScriptAndDatedFileName('_batPWcomparisons.vsMock.genesOI.wide.csv'))
fwrite(dcast(vsMock.dt[species=='human' & !is.na(info),], gene~contrast, value.var = c('log2FoldChange', 'lfcSE', 'pvalue', 'padj')), ScriptAndDatedFileName('_humanPWcomparisons.vsMock.genesOI.wide.csv'))
fwrite(dcast(tp.dt[species=='bat' & !is.na(info),], gene~contrast, value.var = c('log2FoldChange', 'lfcSE', 'pvalue', 'padj')), ScriptAndDatedFileName('_batPWcomparisons.viralTimepoints.genesOI.wide.csv'))
fwrite(dcast(tp.dt[species=='human' & !is.na(info),], gene~contrast, value.var = c('log2FoldChange', 'lfcSE', 'pvalue', 'padj')), ScriptAndDatedFileName('_humanPWcomparisons.viralTimepoints.genesOI.wide.csv'))
```
ok, now to produce the different plots of the data
combine all the datatables and process this as unique combos of species/comparison

```{r}
# create heatmap annotation 
col.pal <- randomcoloR::distinctColorPalette(k=4)
```

```{r}
submat <- hu.counts[rownames(hu.counts) %in% genes.oi,]

submat <- sweep(submat, 1, apply(submat, 1, median, na.rm=T))
colnames(submat) <- gsub('[.]quant[.]sf','', colnames(submat)) # tidy colnames

# column annotation with counts of a SARS protein
colours <- list('timepoint' = c('12h'="#228833", '24h'="#CCBB44", '48h'='#EE6677'),
                'virus' = c('WA'=col.pal[1], 'N_P80T'=col.pal[2], '9bI_N_P80T'=col.pal[3], 'Mock'=col.pal[4]))

colAnn <- HeatmapAnnotation(df = meta.data[host == 'human',.(virus)],
                            'SARS2 Spike protein counts' = anno_barplot(log2(hu.viral.mat[rownames(hu.viral.mat) == 'S',] + 1)),  # axis=T, axis_param = list(at=seq(0, 20, by=5), labels=seq(0, 20, by=5))),
                                                        col = colours)

# row annotation; identify proteins belonging to the different groups
proinflam_bar <- as.numeric(rownames(submat) %in% proinflam) #logical vec
isg_bar <- as.numeric(rownames(submat) %in% isg)
ckine_bar <-  as.numeric(rownames(submat) %in% ckines)

#matrix rownames
names_bar <-  rownames(submat) %in% c(proinflam, isg, ckines)


#sanity check
colnames(submat) ==colnames(hu.viral.mat)

ht_list <- Heatmap(submat,
        top_annotation = colAnn,
        cluster_columns = F,
        show_column_names = T,
        column_names_gp = gpar(fontsize=7),
        name = 'norm counts/median',
        border=T,
        show_row_names = T,
        column_split =  list(#meta.data[host == 'human',.(virus)],
                             meta.data[host == 'human',.(timepoint)]) ) +
  Heatmap(proinflam_bar, name = "proinflammatory", col = c("0" = "white", "1" = '#fde725'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"), column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  Heatmap(isg_bar, name = "interferon stimulated genes (ISGs)", col = c("0" = "white", "1" = '#440154FF'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"),  column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  Heatmap(ckine_bar, name = "cytokines/chemokines", col = c("0" = "white", "1" = '#5ec962'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"),  column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  rowAnnotation(link = anno_mark(at = which(names_bar), 
        labels = rownames(submat)[names_bar], 
        labels_gp = gpar(fontsize = 4), padding = unit(0.2, "mm"))) 

hm <- draw(ht_list, ht_gap = unit(0.2, "mm"), heatmap_legend_side="right")
BackupAsPDF(hm, 'human.genes.oi.medianScaled.heatmap', dimensions = c(12,9))
```
Now plot vs Mock and remove that group from the contrasts

```{r}
submat <- sweep(hu.counts, 1, apply(hu.counts[, grepl('Mock', colnames(hu.counts))], 1, mean, na.rm=T), FUN='-')
submat <- submat[rownames(submat) %in% genes.oi, grep('Mock', colnames(submat), invert=T)]
colnames(submat) <- gsub('[.]quant[.]sf','', colnames(submat)) # tidy colnames

# row annotation; identify proteins belonging to the different groups
proinflam_bar <- as.numeric(rownames(submat) %in% proinflam) #logical vec
isg_bar <- as.numeric(rownames(submat) %in% isg)
ckine_bar <-  as.numeric(rownames(submat) %in% ckines)

#matrix rownames
names_bar <-  rownames(submat) %in% c(proinflam, isg, ckines)

#sanity check
colnames(submat) ==colnames(hu.viral.mat[, grep('Mock', colnames(hu.viral.mat), invert=T)])

# lets make a color vectoe
colours <- list('timepoint' = c('12h'="#228833", '24h'="#CCBB44", '48h'='#EE6677'),
                'virus' = c('WA'=col.pal[1], 'N_P80T'=col.pal[2], '9bI_N_P80T'=col.pal[3]))

colAnn <- HeatmapAnnotation(df = meta.data[host == 'human' & virus != 'Mock',.(virus)],
                            'SARS2 Spike protein counts' = anno_barplot(log2(hu.viral.mat[rownames(hu.viral.mat) == 'S', grep('Mock', colnames(hu.viral.mat), invert=T)] + 1)), 
                                                        col = colours)

ht_list <- Heatmap(submat,
        top_annotation = colAnn,
        cluster_columns =F,
        column_names_gp = gpar(fontsize=6),
        show_column_names = F,
        column_title_gp = gpar(fontsize=8, fontface='bold'),
        row_names_gp = gpar(fontsize=6), #to reordr , treat the extracted vector as factor and set levls
        column_split = factor(str_extract(colnames(submat), '[0-9]{1,2}h'), levels=c('Na', '6h','12h', '24h','48h')),
        col = colorRamp2(breaks=c(-2,0,2), colors=c('blue','white','red')),
        name=('Counts vs Mock'),
        border=T,
        show_row_names = T) +
  Heatmap(proinflam_bar, name = "proinflammatory", col = c("0" = "white", "1" = '#fde725'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"), column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  Heatmap(isg_bar, name = "interferon stimulated genes (ISGs)", col = c("0" = "white", "1" = '#440154FF'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"),  column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  Heatmap(ckine_bar, name = "cytokines/chemokines", col = c("0" = "white", "1" = '#5ec962'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"),  column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  rowAnnotation(link = anno_mark(at = which(names_bar), 
        labels = rownames(submat)[names_bar], 
        labels_gp = gpar(fontsize = 4), padding = unit(0.2, "mm"))) 

hm <- draw(ht_list, ht_gap = unit(0.2, "mm"), heatmap_legend_side="right")
BackupAsPDF(hm, 'human.genes.oi.vsMock.heatmap', dimensions = c(12,9))
```





Now produce the plot with the bat samples

```{r}
submat <- bat.counts[rownames(bat.counts) %in% genes.oi,]

submat <- sweep(submat, 1, apply(submat, 1, median, na.rm=T))
colnames(submat) <- gsub('[.]quant[.]sf','', colnames(submat)) # tidy colnames

# create heatmap annotation 
#col.pal <- randomcoloR::distinctColorPalette(k=4)

# column annotation with counts of a SARS protein
colours <- list('timepoint' = c('6h'="#66CCEE", '12h'="#228833", '24h'="#CCBB44", '48h'='#EE6677'),
                'virus' = c('WA'=col.pal[1], 'N_P80T'=col.pal[2], '9bI_N_P80T'=col.pal[3], 'Mock'=col.pal[4]))

colAnn <- HeatmapAnnotation(df = meta.data[host == 'bat',.(virus)],
                            'SARS2 Spike protein counts' = anno_barplot(log2(bat.viral.mat[rownames(bat.viral.mat) == 'S',] + 1)),  # axis=T, axis_param = list(at=seq(0, 20, by=5), labels=seq(0, 20, by=5))),
                                                        col = colours)

# row annotation; identify proteins belonging to the different groups
proinflam_bar <- as.numeric(rownames(submat) %in% proinflam) #logical vec
isg_bar <- as.numeric(rownames(submat) %in% isg)
ckine_bar <-  as.numeric(rownames(submat) %in% ckines)

#matrix rownames
names_bar <-  rownames(submat) %in% c(proinflam, isg, ckines)

#sanity check
colnames(submat) ==colnames(hu.viral.mat)

ht_list <- Heatmap(submat,
        top_annotation = colAnn,
        cluster_columns = F,
        show_column_names = T,
        column_names_gp = gpar(fontsize=7),
        name = 'norm counts/median',
        border=T,
        show_row_names = T,
        column_split =  list(#meta.data[host == 'human',.(virus)],
                             meta.data[host == 'bat',.(timepoint)]) ) +
    Heatmap(proinflam_bar, name = "proinflammatory", col = c("0" = "white", "1" = '#fde725'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"), column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  Heatmap(isg_bar, name = "interferon stimulated genes (ISGs)", col = c("0" = "white", "1" = '#440154FF'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"),  column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  Heatmap(ckine_bar, name = "cytokines/chemokines", col = c("0" = "white", "1" = '#5ec962'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"),  column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  rowAnnotation(link = anno_mark(at = which(names_bar), 
        labels = rownames(submat)[names_bar], 
        labels_gp = gpar(fontsize = 4), padding = unit(0.2, "mm"))) 

hm <- draw(ht_list, ht_gap = unit(0.2, "mm"), heatmap_legend_side="right")
BackupAsPDF(hm, 'bat.genes.oi.medianScaled.heatmap', dimensions = c(12,9))
```
```{r}
submat <- sweep(bat.counts, 1, apply(bat.counts[, grepl('Mock', colnames(bat.counts))], 1, mean, na.rm=T), FUN='-')
submat <- submat[rownames(submat) %in% genes.oi, grep('Mock', colnames(submat), invert=T)]
colnames(submat) <- gsub('[.]quant[.]sf','', colnames(submat)) # tidy colnames


# row annotation; identify proteins belonging to the different groups
proinflam_bar <- as.numeric(rownames(submat) %in% proinflam) #logical vec
isg_bar <- as.numeric(rownames(submat) %in% isg)
ckine_bar <-  as.numeric(rownames(submat) %in% ckines)

#matrix rownames
names_bar <-  rownames(submat) %in% c(proinflam, isg, ckines)

#sanity check
colnames(submat) ==colnames(bat.viral.mat[, grep('Mock', colnames(bat.viral.mat), invert=T)])

# lets make a color vectoe
colours <- list('timepoint' = c('12h'="#228833", '24h'="#CCBB44", '48h'='#EE6677'),
                'virus' = c('WA'=col.pal[1], 'N_P80T'=col.pal[2], '9bI_N_P80T'=col.pal[3]))

colAnn <- HeatmapAnnotation(df = meta.data[host == 'bat' & virus != 'Mock',.(virus)],
                            'SARS2 Spike protein counts' = anno_barplot(log2(bat.viral.mat[rownames(hu.viral.mat) == 'S', grep('Mock', colnames(bat.viral.mat), invert=T)] + 1)), 
                                                        col = colours)

ht_list <- Heatmap(submat,
        top_annotation = colAnn,
        cluster_columns =F,
        column_names_gp = gpar(fontsize=6),
        show_column_names = F,
        column_title_gp = gpar(fontsize=8, fontface='bold'),
        row_names_gp = gpar(fontsize=6), #to reordr , treat the extracted vector as factor and set levls
        column_split = factor(str_extract(colnames(submat), '[0-9]{1,2}h'), levels=c('Na', '6h','12h', '24h','48h')),
        col = colorRamp2(breaks=c(-2,0,2), colors=c('blue','white','red')),
        name=('Counts vs Mock'),
        border=T,
        show_row_names = T) +
  Heatmap(proinflam_bar, name = "proinflammatory", col = c("0" = "white", "1" = '#fde725'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"), column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  Heatmap(isg_bar, name = "interferon stimulated genes (ISGs)", col = c("0" = "white", "1" = '#440154FF'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"),  column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  Heatmap(ckine_bar, name = "cytokines/chemokines", col = c("0" = "white", "1" = '#5ec962'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"),  column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  rowAnnotation(link = anno_mark(at = which(names_bar), 
        labels = rownames(submat)[names_bar], 
        labels_gp = gpar(fontsize = 4), padding = unit(0.2, "mm"))) 

hm <- draw(ht_list, ht_gap = unit(0.2, "mm"), heatmap_legend_side="right")
hm
BackupAsPDF(hm, 'bat.genes.oi.vsMock.heatmap', dimensions = c(12,9))
```
Lets collapse the counts for each replicate and plot vs Mock for both, but warn Jyoti that this hides differences between the group

```{r}

hu.counts %>%  colnames()

submat <- sweep(hu.counts, 1, apply(hu.counts[, grepl('Mock', colnames(hu.counts))], 1, mean, na.rm=T), FUN='-')
submat <- submat[rownames(submat) %in% genes.oi, grep('Mock', colnames(submat), invert=T)]
colnames(submat) <- gsub('[.]quant[.]sf','', colnames(submat)) # tidy colnames

# row annotation; identify proteins belonging to the different groups
proinflam_bar <- as.numeric(rownames(submat) %in% proinflam) #logical vec
isg_bar <- as.numeric(rownames(submat) %in% isg)
ckine_bar <-  as.numeric(rownames(submat) %in% ckines)

#matrix rownames
names_bar <-  rownames(submat) %in% c(proinflam, isg, ckines)

#sanity check
colnames(submat) ==colnames(hu.viral.mat[, grep('Mock', colnames(hu.viral.mat), invert=T)])

# lets make a color vectoe
colours <- list('timepoint' = c('12h'="#228833", '24h'="#CCBB44", '48h'='#EE6677'),
                'virus' = c('WA'=col.pal[1], 'N_P80T'=col.pal[2], '9bI_N_P80T'=col.pal[3]))

colAnn <- HeatmapAnnotation(df = meta.data[host == 'human' & virus != 'Mock',.(virus)],
                            'SARS2 Spike protein counts' = anno_barplot(log2(hu.viral.mat[rownames(hu.viral.mat) == 'S', grep('Mock', colnames(hu.viral.mat), invert=T)] + 1)), 
                                                        col = colours)

ht_list <- Heatmap(submat,
        top_annotation = colAnn,
        cluster_columns =F,
        column_names_gp = gpar(fontsize=6),
        show_column_names = F,
        column_title_gp = gpar(fontsize=8, fontface='bold'),
        row_names_gp = gpar(fontsize=6), #to reordr , treat the extracted vector as factor and set levls
        column_split = factor(str_extract(colnames(submat), '[0-9]{1,2}h'), levels=c('Na', '6h','12h', '24h','48h')),
        col = colorRamp2(breaks=c(-2,0,2), colors=c('blue','white','red')),
        name=('Counts vs Mock'),
        border=T,
        show_row_names = T) +
  Heatmap(proinflam_bar, name = "proinflammatory", col = c("0" = "white", "1" = '#fde725'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"), column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  Heatmap(isg_bar, name = "interferon stimulated genes (ISGs)", col = c("0" = "white", "1" = '#440154FF'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"),  column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  Heatmap(ckine_bar, name = "cytokines/chemokines", col = c("0" = "white", "1" = '#5ec962'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"),  column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  rowAnnotation(link = anno_mark(at = which(names_bar), 
        labels = rownames(submat)[names_bar], 
        labels_gp = gpar(fontsize = 4), padding = unit(0.2, "mm"))) 

hm <- draw(ht_list, ht_gap = unit(0.2, "mm"), heatmap_legend_side="right")
BackupAsPDF(hm, 'human.genes.oi.vsMock.heatmap', dimensions = c(12,9))
```
Get average of each group

```{r}
hu.dt <- setDT(reshape2::melt(hu.counts))
hu.dt$Var2 %>%  unique(0)
hu.dt[, Var2 := gsub('_T1[.]quant[.]sf|[.]quant[.]sf','', Var2)]
hu.dt[, `:=`(timepoint=ifelse(grepl('Mock', Var2), '', str_extract(Var2, '[0-9]{1,2}h')),
             virus=ifelse(grepl('Mock', Var2), 'Mock', gsub('MRC5_|RFe_|_[0-9]{1,2}h.+', '', Var2)))]
hu.dt[, replicate := str_extract(Var2,'[123]$'),]

# convert to long and summarise by group
summary.mat <- dcast(hu.dt, Var1~interaction(virus, timepoint), value.var = 'value', fun.aggregate = mean) %>% 
  as.matrix(rownames='Var1')

submat <- summary.mat[rownames(summary.mat) %in% genes.oi,]
submat <- sweep(submat, 1, submat[, grep('Mock', colnames(submat))])
submat <-  submat[, grep('Mock', colnames(submat), invert=T)]


# row annotation; identify proteins belonging to the different groups
proinflam_bar <- as.numeric(rownames(submat) %in% proinflam) #logical vec
isg_bar <- as.numeric(rownames(submat) %in% isg)
ckine_bar <-  as.numeric(rownames(submat) %in% ckines)

#matrix rownames
names_bar <-  rownames(submat) %in% c(proinflam, isg, ckines)

ht_list <- Heatmap(submat, 
        cluster_column_slices = F,
        name = 'Counts vs Mock (avg)',
       top_annotation = HeatmapAnnotation(virus=gsub('[.][0-9]{1,2}h','', colnames(submat)),
                                           col=list(virus=c('WA'=col.pal[1], 'N_P80T'=col.pal[2], '9bI_N_P80T'=col.pal[3]))
                                          ),
        column_split=str_extract(colnames(submat), '[0-9]{1,2}h')) +
  Heatmap(proinflam_bar, name = "proinflammatory", col = c("0" = "white", "1" = '#fde725'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"), column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  Heatmap(isg_bar, name = "interferon stimulated genes (ISGs)", col = c("0" = "white", "1" = '#440154FF'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"),  column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  Heatmap(ckine_bar, name = "cytokines/chemokines", col = c("0" = "white", "1" = '#5ec962'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"),  column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  rowAnnotation(link = anno_mark(at = which(names_bar), 
        labels = rownames(submat)[names_bar], 
        labels_gp = gpar(fontsize = 4), padding = unit(0.2, "mm"))) 

hm <- draw(ht_list, ht_gap = unit(0.2, "mm"), heatmap_legend_side="right")
BackupAsPDF(hm, 'human.genes.oi.vsMock.meanCollapsed.heatmap', dimensions = c(9,9))
```
replot the viral samples now

```{r}
bat.dt <- setDT(reshape2::melt(bat.counts))
bat.dt$Var2 %>%  unique()

bat.dt[, Var2 := gsub('_T1[.]quant[.]sf|[.]quant[.]sf','', Var2)]
bat.dt[, `:=`(timepoint=ifelse(grepl('Mock', Var2), '', str_extract(Var2, '[0-9]{1,2}h')),
             virus=ifelse(grepl('Mock', Var2), 'Mock', gsub('MRC5_|RFe_|_[0-9]{1,2}h.+', '', Var2)))]
bat.dt[, replicate := str_extract(Var2,'[123]$'),]

# convert to long and summarise by group
summary.mat <- dcast(bat.dt, Var1~interaction(virus, timepoint), value.var = 'value', fun.aggregate = mean) %>% 
  as.matrix(rownames='Var1')

submat <- summary.mat[rownames(summary.mat) %in% genes.oi,]
submat <- sweep(submat, 1, submat[, grep('Mock', colnames(submat))])
submat <-  submat[, grep('Mock', colnames(submat), invert=T)]


# row annotation; identify proteins belonging to the different groups
proinflam_bar <- as.numeric(rownames(submat) %in% proinflam) #logical vec
isg_bar <- as.numeric(rownames(submat) %in% isg)
ckine_bar <-  as.numeric(rownames(submat) %in% ckines)

#matrix rownames
names_bar <-  rownames(submat) %in% c(proinflam, isg, ckines)

submat <- submat[,c( "9bI_N_P80T.6h","N_P80T.6h","WA.6h", "9bI_N_P80T.12h", "N_P80T.12h",     "WA.12h" ,  "9bI_N_P80T.24h" ,"N_P80T.24h" ,    "WA.24h", "9bI_N_P80T.48h" ,"N_P80T.48h" ,"WA.48h")]         

Heatmap(submat,
        cluster_columns = F)

ht_list <- Heatmap(submat, 
        cluster_column_slices = F,
        column_order = colnames(submat),
        name = 'Counts vs Mock (avg)',
       top_annotation = HeatmapAnnotation(virus=gsub('[.][0-9]{1,2}h','', colnames(submat)),
                                           col=list(virus=c('WA'=col.pal[1], 'N_P80T'=col.pal[2], '9bI_N_P80T'=col.pal[3]))),
        column_split = factor(ifelse(grepl('Mock', colnames(submat)), 'Na', str_extract(colnames(submat), '[0-9]{1,2}h')), levels=c('Na', '6h','12h', '24h','48h'))) +
  Heatmap(proinflam_bar, name = "proinflammatory", col = c("0" = "white", "1" = '#fde725'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"), column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  Heatmap(isg_bar, name = "interferon stimulated genes (ISGs)", col = c("0" = "white", "1" = '#440154FF'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"),  column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  Heatmap(ckine_bar, name = "cytokines/chemokines", col = c("0" = "white", "1" = '#5ec962'), 
        show_heatmap_legend = FALSE, width = unit(3, "mm"),  column_names_gp = gpar(fontsize=7, fontface='bold'), column_gap = unit(0.2, "cm"), gap = unit(3, "cm")) +
  rowAnnotation(link = anno_mark(at = which(names_bar), 
        labels = rownames(submat)[names_bar], 
        labels_gp = gpar(fontsize = 4), padding = unit(0.2, "mm"))) 

hm <- draw(ht_list, ht_gap = unit(0.2, "mm"), heatmap_legend_side="right")


BackupAsPDF(hm, 'bat.genes.oi.vsMock.meanCollapsed.heatmap', dimensions = c(9,9))
```


More important plot; plot proportion of viral reads in each of the matrices; is this increasing overtime?
Strange; no obvious increase in viral proteins over time it seems?
```{r}
test <- log2(hu.viral.mat+1)

subtest <- sweep(test, 1, apply(test, 1, median, na.rm=T))

hm <- Heatmap(subtest,
              top_annotation = HeatmapAnnotation(virus=gsub('_[123]','', gsub('MRC5_|RFe_|[_][0-9].+','', colnames(subtest))),
                                           col=list(virus=c('WA'=col.pal[1], 'N_P80T'=col.pal[2], '9bI_N_P80T'=col.pal[3], 'Mock'=col.pal[4]))),
              column_split = meta.data[host == 'human',.(timepoint)],
              name='log2 Counts/Median',
              border=T,
              column_names_gp = gpar(fontsize=6),
              col=colorRamp2(breaks=c(-2,0,2), col=c('blue','white','red')))

hm
BackupAsPDF(hm, 'human.viralCounts.medianScale.heatmap')


# do the bat 
test <- log2(bat.viral.mat+1)
subtest <- sweep(test, 1, apply(test, 1, median, na.rm=T))


hm <- Heatmap(subtest,
              cluster_column_slices = F,
              top_annotation = HeatmapAnnotation(virus=gsub('_[123]','', gsub('MRC5_|RFe_|[_][0-9].+','', colnames(subtest))),
                                           col=list(virus=c('WA'=col.pal[1], 'N_P80T'=col.pal[2], '9bI_N_P80T'=col.pal[3], 'Mock'=col.pal[4]))),
             # column_split = meta.data[host == 'bat',.(timepoint)],
              column_split = factor(ifelse(grepl('Mock', colnames(subtest)), '', str_extract(colnames(subtest), '[0-9]{1,2}h')), levels=c('', '6h','12h', '24h','48h')),
              name='log2 Counts/Median',
              border=T,
              column_names_gp = gpar(fontsize=6),
              col=colorRamp2(breaks=c(-4,0,4), col=c('blue','white','red')))

hm
BackupAsPDF(hm, 'bat.viralCounts.medianScale.heatmap')


 sort(meta.data[host == 'bat',.(timepoint)]

colnames(subtest)
gsub('_[123]','', gsub('MRC5_|RFe_|[_][0-9].+','', colnames(subtest)))



2^10
```
Lets just try the unnormalized to inspect

```{r}
counts(dds.viral.hu, normalized=F)


# these are raw counts no obvious trend related to timepoint
test <- log2(counts(dds.viral.hu, normalized=F)+1)

subtest <- sweep(test, 1, apply(test, 1, median, na.rm=T))

Heatmap(subtest[rownames(subtest) %in% viral.genes,],
        column_split = meta.data[host == 'human',.(timepoint)],
        name='log2 Counts/Median',
        col=colorRamp2(breaks=c(-4,0,4), col=c('blue','white','red')))





tp.dt
2^4
```


