---
title: "010725_plotsForKroganWeekly"
author: "Martin Gordon"
date: "2025-01-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(stringr)
library(ggbeeswarm)
library(ggrepel)
library(patchwork)
library(readxl)
library(circlize)
library(ComplexHeatmap)
library(randomcoloR)
library(scales) #for muted colors
library(eulerr)

# this is for the Venn diagram plots
library(ggvenn)
library(ggVennDiagram) # main diff to above is you can quantify overlapping regions

#Ben utils
source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")


# helper scripts for module/network export etc
source ("/Users/martingordon/Documents/utils/mg_utils/r_utils/HelperFunctions.R")


# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}

```

Read in the WGCNA data: module assignment from the WGCNA run
```{r}
module.assignment <- readRDS('./output/gpcr.11.receptors.moduleAssignment.Rdata')
exp.mat <- readRDS('./output/gpcr.11.receptors.exp.mat.Rdata') # the expression matrix used as input for each of these

clust.genes <- fread('./output/wgcna.clustergenes.adjmat.csv')
```

Quick look at the data; seems like the same module info
```{r}
Heatmap(exp.mat,
        cluster_columns = F,
        cluster_column_slices = T,
        show_row_names = F,
        border=T,
        column_split = tstrsplit(colnames(exp.mat), "[_.]")[[1]],
        )   
```

```{r}
protQuant <- fread('data/2022_08_17_ProteinLevelData.lib.csv.gz')

group.noi <- protQuant[!GROUP %like% c("DOR|KOR|MOR|MRX2|MRX4|NK1R|NPFFR1|NPY1R|PTGER1|PTGER4|QRFPR|DOR"),unique(GROUP)]

protQuant <- protQuant[!GROUP %in% group.noi, ]
protQuant <- protQuant[!GROUP %like% "MOR_LG94",]

protQuant[, `:=`(timeStr = str_extract(GROUP,'[0-9]{2}$'),
                 receptor = gsub("_[0-9]{2}$", '', protQuant$GROUP)
               )]

setorder(protQuant, receptor, Protein, timeStr)

protQuant[, batchControlled := LogIntensities - median(LogIntensities, na.rm = TRUE), by = .(Protein, receptor, SUBJECT)]
protQuant[, vsTimeZero := batchControlled - mean(batchControlled[timeStr == timeStr[!is.na(batchControlled)][1] ]), by = .(Protein, receptor)]
protQuant[, gene := multiUniprots2multiGenes(Protein)]
```

```{r}
# max shift across timepoints per protein/receptor
protQuant[, maxshiftVsT0 :=  ifelse(max(vsTimeZero) > abs(min(vsTimeZero)), max(vsTimeZero), min(vsTimeZero)), by=.(Protein,receptor)]
# find the timepoint wihtt he maximum shift
protQuant[, maxshiftVsT0_timepoint :=  ifelse(maxshiftVsT0 == vsTimeZero, 'yes', 'no'), by=.(Protein,receptor)]
```

```{r}
names(clust.genes) <- c('gene', 'Cluster')


# now add the cluster info to the protQuant; subset to this 
protQuant <- merge(protQuant, clust.genes, by='gene', all.x=F)
protQuant[receptor == 'MOR_DAMGO', receptor := 'MOR']

# use this to filter and calculate the median value per module/receptor
protQuant[, medianClustervsT0 := median(maxshiftVsT0, na.rm=T), by=.(Cluster,receptor)]
```

```{r}
# add this for cluster coloring 
col.pal <- c("#D57AC7","#8DDDE1","#A1E4AC","#6EE766","#CEDF5F","#DEC5D2","#85A693","#DBBB7C","#7D9AD6","#D96A6E","#A94FDD")  #"#999933" '#DDCC77' "grey"


# heatmap annotation\
row_ha <- HeatmapAnnotation(receptor=protQuant$receptor, 
                        col=list(receptor = c("DOR"=col.pal[1], 'KOR'=col.pal[2], 'MOR'=col.pal[3], 'MRX2'=col.pal[4], 'MRX4'=col.pal[5], 'NK1R'=col.pal[6], 'NPFFR1'=col.pal[7], 'NPY1R'=col.pal[8], 'PTGER1'=col.pal[9], 'PTGER4'=col.pal[10], 'QRFPR'=col.pal[11])) )

p.mat <- dcast(protQuant, Cluster~receptor, value.var = 'medianClustervsT0', fun.aggregate = unique) %>% 
  as.matrix(rownames='Cluster') 


ha <- Heatmap(p.mat, top_annotation =row_ha,
          name = "max log2(intensity/\nmean(timeZero))")

rownames(p.mat) <- c('C1 Endosome trafficking', 'C2 PM structure/adhesion', 'C3 Actin cytoskeleton', 'C4 Golgi trafficking', 'C5 G-Protein coupled receptor signalling', 
                     'C6 Protein Kinase C', 'C7 Clathrin', 'C8 Cytoplasmic stress granule', 'C9 Actin polymerization', 'C10 Cell morphogenesis', 'C11 RNA processing')

ha <- Heatmap(p.mat, 
              top_annotation =row_ha,
              border=T,
              name = "cluster median/\ntimeZero",
              row_names_gp = gpar(fontsize=8)
          )
BackupAsPDF(ha, 'clusterMedian.heatmap', format='png')
```
Lets plot an example of what this looks like

```{r}
protQuant[, cluster_label := fcase(Cluster == 'cluster1', 'C1 Endosome trafficking',
                              Cluster == 'cluster2', 'C2 PM structure/adhesion',
                              Cluster == 'cluster3', 'C3 Actin cytoskeleton',
                              Cluster == 'cluster4', 'C4 Golgi trafficking',
                              Cluster == 'cluster5', 'C5 G-Protein coupled receptor signalling',
                              Cluster == 'cluster6', 'C6 Protein Kinase C',
                              Cluster == 'cluster7', 'C7 Clathrin',
                              Cluster == 'cluster8', 'C8 Cytoplasmic stress granule',
                              Cluster == 'cluster9', 'C9 Actin polymerization',
                              Cluster == 'cluster10', 'C10 Cell morphogenesis',
                              Cluster == 'cluster11', 'C11 RNA processing'
                              )]
```


```{r}

# calculate the median shift vs T0 per timepoint/cluster/receptor
protQuant[, medianVsTimeZero := median(vsTimeZero, na.rm=T), by=.(Cluster, receptor, timeStr)]
protQuant[, GO.terms := cluster_label]

protQuant$GO.terms <- factor(protQuant$GO.terms, levels=c("C1 Endosome trafficking", "C2 PM structure/adhesion","C3 Actin cytoskeleton","C4 Golgi trafficking","C5 G-Protein coupled receptor signalling","C6 Protein Kinase C",
                                                          "C7 Clathrin", "C8 Cytoplasmic stress granule", "C9 Actin polymerization", "C10 Cell morphogenesis","C11 RNA processing"))

# save the protQuant table to file 
fwrite(protQuant, ScriptAndDatedFileName('proteinQuant.anno.csv.gz'))
```

```{r}

g <- ggplot(protQuant[!is.na(Cluster)], aes(x=timeStr, y=medianVsTimeZero, group=receptor, color=receptor)) + 
  stat_summary(fun = "median", geom = "line") +
  scale_color_manual(values = col.pal) + 
  ggtitle('Median expression vs Time 0') +
  geom_hline(yintercept = 0, linetype=2, alpha=0.6) +
  labs(x='timepoint (mins)', y="Intensity (log2)") +
  facet_wrap(~GO.terms, scales='free_y', ncol=3) +
  theme_classic() +
  scale_x_discrete(expand=c(0,0.03)) +
  theme(
   strip.text = element_text(size = 6))

g
BackupAsPDF(g, 'medianVsT0.modulenamepanel.with0line.linechart.', dimensions = c(8,8))

g <- ggplot(protQuant[!is.na(Cluster)], aes(x=timeStr, y=medianVsTimeZero, group=GO.terms, color=GO.terms)) + 
  stat_summary(fun = "median", geom = "line") +
  scale_color_manual(values = col.pal) + 
  ggtitle('Median expression vs Time 0') +
  geom_hline(yintercept = 0, linetype=2, alpha=0.6) +
  labs(x='timepoint (mins)', y="Intensity (log2)") +
  facet_wrap(~receptor, scales='free_y', ncol=3) +
  theme_classic() +
  scale_x_discrete(expand=c(0,0.03)) +
  theme(
   strip.text = element_text(size = 6))

g

BackupAsPDF(g, 'medianVsT0.receptornamepanel.with0line.linechart.', dimensions = c(8,8))
```
Subset to cluster 1, cluster 6

```{r}
cluster.oi <- c('cluster1', 'cluster6')

g <- ggplot(protQuant[Cluster %in% cluster.oi,], aes(x=timeStr, y=vsTimeZero, group=receptor, color=receptor)) + 
  geom_smooth(stat = 'summary', fun.y = median, se =TRUE, show.legend = T, aes(fill=receptor), alpha=0.4) +
  scale_color_manual(values = col.pal) + 
  scale_fill_manual(values = col.pal) + 
  ggtitle('Median expression vs Time 0') +
  geom_hline(yintercept = 0, linetype=2, alpha=0.6) +
  labs(x='timepoint (mins)', y="Intensity (log2)") +
  facet_wrap(~cluster_label, ncol=3) +
  theme_classic() +
  scale_x_discrete(expand=c(0,0.03)) +
  theme(
   strip.text = element_text(size = 6)) +
  guides(color=guide_legend(override.aes=list(fill=NA))) # allows you to modify sections of the legend

g 
BackupAsPDF(g, 'clusters.oi.wSE.scales.fixed.linechart')


g <- ggplot(protQuant[Cluster %in% cluster.oi,], aes(x=timeStr, y=vsTimeZero, group=receptor, color=receptor)) + 
  geom_smooth(stat = 'summary', fun.y = median, se =TRUE, show.legend = T, aes(fill=receptor, alpha=0.2)) +
  scale_color_manual(values = col.pal) + 
  scale_fill_manual(values = col.pal) + 
  ggtitle('Median expression vs Time 0') +
  geom_hline(yintercept = 0, linetype=2, alpha=0.6) +
  labs(x='timepoint (mins)', y="Intensity (log2)") +
  facet_wrap(~cluster_label, scales='free_y', ncol=3) +
  theme_classic() +
  scale_x_discrete(expand=c(0,0.03)) +
  theme(
   strip.text = element_text(size = 6)) +
  guides(color=guide_legend(override.aes=list(fill=NA))) # allows you to modify sections of the legend

g 
BackupAsPDF(g, 'clusters.oi.wSE.scales.free.linechart')

g <- ggplot(protQuant[Cluster %in% cluster.oi,], aes(x=timeStr, y=vsTimeZero, group=receptor, color=receptor)) + 
  #geom_smooth(stat = 'summary', fun.y = median, se =TRUE) +
  stat_summary(fun = median, geom="line", se=T) +
  #stat_summary(fun = "median", geom = "line", se=TRUE) +
  scale_color_manual(values = col.pal) + 
  ggtitle('Median expression vs Time 0') +
  geom_hline(yintercept = 0, linetype=2, alpha=0.6) +
  labs(x='timepoint (mins)', y="Intensity (log2)") +
  facet_wrap(~cluster_label, ncol=3) +
  theme_classic() +
  scale_x_discrete(expand=c(0,0.03)) +
  theme(
   strip.text = element_text(size = 6))
g
BackupAsPDF(g, 'clusters.oi.scales.fixed.linechart')
```
plot the max shift vs T0; include a red dot if this tp is the max shift vs t0
Can try add red points for the max shift; easeir to share table to annotate

```{r}
protQuant[,.N, by=maxshiftVsT0_timepoint]


protQuant[Cluster %in% cluster.oi & maxshiftVsT0_timepoint == 'yes',.(medianVsTimeZero), by=.(cluster_label,receptor, timeStr)]

g <- ggplot(protQuant[Cluster %in% cluster.oi,], aes(x=timeStr, y=vsTimeZero, group=receptor, color=receptor, shape=maxshiftVsT0_timepoint)) + 
  stat_summary(fun = median, geom="line") +
  scale_color_manual(values = col.pal) + 
  ggtitle('Median expression vs Time 0') +
  geom_hline(yintercept = 0, linetype=2, alpha=0.6) +
  labs(x='timepoint (mins)', y="Intensity (log2)") +
  facet_wrap(~cluster_label, ncol=3) +
  theme_classic() +
  scale_x_discrete(expand=c(0,0.03)) +
  theme(
   strip.text = element_text(size = 6))
g

protQuant[Cluster %in% cluster.oi & maxshiftVsT0_timepoint == 'yes',]

#fwrite(protQuant[, .N, by=.(maxshiftVsT0_timepoint, receptor, cluster_label, timeStr)][maxshiftVsT0_timepoint == 'yes', .(receptor, cluster_label, timepoint_maxShiftvsT0=timeStr)], ScriptAndDatedFileName('maxShiftTPvsT0.csv'))

```

Try a different visualization of the heatmap.. use this vs T0, then take the max shift

```{r}
# calculate the median shift vs T0 per cluster, per receptor, per timepoint
protQuant[, medianVsTimeZero := median(vsTimeZero, na.rm=T), by=.(Cluster, receptor, timeStr)]

protQuant[,medianVsTimeZero, by=.(Cluster, receptor, timeStr)] %>% 
  unique() %>% 
  .[Cluster == 'cluster8']


# get the maximum shift vs t0 basically taking the max timepoint value
maxShift.mat <- dcast(protQuant, receptor~cluster_label, value.var = 'medianVsTimeZero', fun.aggregate = function(x){x[which.max(abs(x))]}) %>% 
  as.matrix(rownames='receptor') %>% 
  t()

# sanity check; checking in both directions
protQuant[,.N, by=.(Cluster, receptor, timeStr, medianVsTimeZero)][Cluster == 'cluster10' & receptor == 'DOR']
protQuant[,.N, by=.(Cluster, receptor, timeStr, medianVsTimeZero)][Cluster == 'cluster1' & receptor == 'DOR']
```
plot the max shift per timepoint

```{r}
# reorder so it matches the rownames of p.mat if not clustering..
maxShift.mat <- maxShift.mat[rownames(p.mat),]


rownames(p.mat)
maxShift.mat

ha <- Heatmap(maxShift.mat, 
              top_annotation =row_ha,
              border=T,
              name = "cluster median vs T0\n(max shift)",
              row_names_gp = gpar(fontsize=8)
          )

ha
BackupAsPDF(ha, 'maxShift.clusterMedian.vsT0.heatmap', format='pdf')
```

01-21-25
---
Plot requests for Qiongyu, lets
First the data driven heatmap with clustering enforce in the columns
Also need to split the row groups based on wgcna order
```{r}
# order here is DOR, KOR, MOR, MRX2, MRX4, NK1R, NPFFR1, NPY1R, PTGR1 PTGR4, QRFPR
col.pal <- c("#D57AC7","#8DDDE1","#A1E4AC","#6EE766","#CEDF5F","#DEC5D2","#85A693","#DBBB7C","#7D9AD6","#D96A6E","#A94FDD")  #"#999933" '#DDCC77' "grey"
# this is Qiopngyus new palette
receptorColors <- c(
DOR = "#99CCFF",
KOR = "#66CCFF",
MOR = "#3399FF",
MRX2 = "#FF9900",
MRX4 = "#FF6666",
NK1R = "#FF6633",
NPFFR1 = "#9EBCDA",
NPY1R = "#0099CC",
PTGER1 = "#FFCCCC",
PTGER4 = "#85CB9C",
QRFPR = "#FF9999")


split <- factor(module.assignment, levels=c('1','2','3','4','5','6','7','8','9','10','11'))

# add wgcna module annotation
row_ha <- rowAnnotation(modules=modules.adj, 
                        #set any annotaiton settings for CH 
                        annotation_legend_param = list(modules = list(
                       	ncol = 2, 
                       	at = split_level,
                       	title = "wgcna modules")),
                        #setting annotation cols
                        col=list(modules = c("1"=col.pal[1], '2'=col.pal[2], '3'=col.pal[3],
                                                       '4'=col.pal[4], '5'=col.pal[5], '6'=col.pal[6], 
                                                       '7'=col.pal[7], '8'=col.pal[8],'9'=col.pal[9], 
                                                       '10'=col.pal[10], '11'=col.pal[11])))


Heatmap(exp.mat,
        cluster_column_slices = T,
        cluster_columns=F,
        show_row_names = F,
        border=T,
        column_split = factor(tstrsplit(colnames(exp.mat), "[_.]")[[1]]),
        )   

```
tidier version of the plot
```{r}
library(RColorBrewer)


# add split order
split <- factor(modules.adj, levels=c('1','2','3','4','5','6','7','8','9','10','11'))
#split <- factor(modules.tom, levels=c('1','2','3','4','5','6'))

split_level = intersect(1:11, split)

# add wgcna module annotation
# not using right now.. maybe try set names?
row_ha <- rowAnnotation(modules=module.assignment, 
                        #set any annotaiton settings for CH 
                        annotation_legend_param = list(modules = list(
                       	ncol = 2, 
                       	at = split_level,
                       	title = "wgcna modules")),
                        #setting annotation cols
                        col=list(modules = c("1"=col.pal[1], '2'=col.pal[2], '3'=col.pal[3],
                                                       '4'=col.pal[4], '5'=col.pal[5], '6'=col.pal[6], 
                                                       '7'=col.pal[7], '8'=col.pal[8],'9'=col.pal[9], 
                                                       '10'=col.pal[10], '11'=col.pal[11])))

tp.pal <- brewer.pal(5,'Blues')
names(tp.pal) <- unique(tstrsplit(colnames(exp.mat), "[_.]")[[2]])

col_ha <- HeatmapAnnotation(receptor=tstrsplit(colnames(exp.mat), "[_.]")[[1]],
                            timepoint=tstrsplit(colnames(exp.mat), "[_.]")[[2]],
                            col=list(receptor=receptorColors, timepoint=tp.pal))


hm <- Heatmap(exp.mat, cluster_columns = FALSE, 
         cluster_column_slices = T,
         show_row_names = F,
         show_column_names = F,
         top_annotation = col_ha,
         #right_annotation = row_ha,
         split = split,
         col=colorRamp2(colors=c("dodgerblue2", "white", "#E31A1C"),  breaks=c(-2,0,2)), 
         border=T,
         name = "log2(intensity/\nmean(timeZero))",
         column_split = tstrsplit(colnames(exp.mat), "[_.]")[[1]],
         na_col = 'darkgrey',
         row_title =("M %s"),
         row_title_side = 'right',
         #row_title =sprintf("%d Proteins",
         #                   nrow(exp.mat)),
         column_names_gp = gpar (fontsize= 3),
         column_title_gp = gpar (fontsize= 9))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=12))

hm
BackupAsPDF(hm, 'wgcna.modules.firstPass.heatmap', dimensions = c(9,6))



# want to cluster between groups, not within
column_split = tstrsplit(colnames(exp.mat), "[_.]")[[1]]

# here we are clustering means between different groups but not chaning the cluster order
dend = cluster_between_groups(exp.mat, column_split)

# try to reproduce but cluster the 
hm <- Heatmap(exp.mat, 
         cluster_columns = dend,
         cluster_column_slices = T,
         cluster_row_slices = T,
         show_row_names = F,
         show_column_names = F,
         top_annotation = col_ha,
         split = split,
         col=colorRamp2(colors=c("dodgerblue2", "white", "#E31A1C"),  breaks=c(-2,0,2)), 
         border=T,
         name = "log2(intensity/\nmean(timeZero))",
         column_split = 11, # need to specify the number of groups to split into 
         na_col = 'darkgrey',
         #row_title =sprintf("%d Proteins",
         #                   nrow(exp.mat)),
         row_title =("M %s"),
         row_title_side = 'right',
         row_title_gp = gpar(fontsize=8),
         column_names_gp = gpar (fontsize= 3),
         column_title_gp = gpar (fontsize= 9))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=12))

BackupAsPDF(hm,  'wgcna.modules.colSlicesClustered.heatmap', dimensions = c(9,8))
#column_split = tstrsplit(colnames(exp.mat), "[_.]")[[1]],
```
The col clustering just doesnt look so great, instead apply the ordering from the summary plot
Looks good, but now we want to collapse and just take the average of each group and plot that

```{r}
exp.dt <- reshape2::melt(exp.mat) %>% as.data.table()
setnames(exp.dt, new=c('Protein', 'sample', 'value'))
         
exp.dt[, c('receptor', 'timepoint', 'batchID') := tstrsplit(sample, "[_.]", keep = c(1,2,4))] 

exp.dt[, .N, by=.(sample, receptor, timepoint, batchID)]
exp.dt[, GROUP := paste0(receptor, '_', timepoint)]

# convert back to wide, first w/o agg fucntion to ensure it is converted correclty
collapse.mat <- dcast(exp.dt, Protein~GROUP, value.var = 'value', fun.aggregate = mean, na.rm=T) %>% 
  as.matrix(rownames='Protein')

fwrite(exp.dt, ScriptAndDatedFileName('11.receptor.vsT0.csv'))


collapse.dt <- reshape2::melt(collapse.mat)
setnames(collapse.dt, old=c('Protein', 'GROUP', 'vsTimeZero'))


fwrite(collapse.dt, ScriptAndDatedFileName('11.receptor.collapsed.vsT0.csv'))
```

adjust column clustering and order
```{r}
# need to adjust the column annotations
names(tp.pal) <- unique(tstrsplit(colnames(collapse.mat), "[_.]")[[2]])

col_ha <- HeatmapAnnotation(receptor=tstrsplit(colnames(collapse.mat), "[_.]")[[1]],
                            timepoint=tstrsplit(colnames(collapse.mat), "[_.]")[[2]],
                            col=list(receptor=receptorColors, timepoint=tp.pal))


# want to cluster between groups, not within
column_split = tstrsplit(colnames(collapse.mat), "[_.]")[[1]]
# here we are clustering means between different groups but not chaning the cluster order
dend = cluster_between_groups(collapse.mat, column_split)


 tstrsplit(colnames(collapse.mat), "[_.]")[[1]]

View(cluster_between_groups)

# try to reproduce but cluster the 
hm <- Heatmap(collapse.mat, 
         cluster_columns = dend,
         cluster_column_slices = T,
         cluster_row_slices = F,
         show_row_names = F,
         show_column_names = F,
         top_annotation = col_ha,
         split = split,
         col=colorRamp2(colors=c("dodgerblue2", "white", "#E31A1C"),  breaks=c(-2,0,2)), 
         border=T,
         name = "log2(intensity/\nmean(timeZero))",
         column_split = 11, # need to specify the number of groups to split into 
         na_col = 'darkgrey',
         row_title ="M %s",
         row_title_side = 'right',
         row_title_gp = gpar(fontsize=8),
         column_names_gp = gpar (fontsize= 3),
         column_title_gp = gpar (fontsize= 9))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=12))

BackupAsPDF(hm,  'wgcna.modules.collapsed.colSlicesClustered.heatmap', dimensions = c(9,7))
```

recluster trying pca order and/or heatmap summary

```{r}
summary.order <- c('QRFPR', 'NK1R', 'MRX4', 'PTGER1', 'NPY1R', 'PTGER4', 'KOR', 'NPFFR1', 'MRX2', 'MOR', 'DOR')


tstrsplit(colnames(collapse.mat), "[_.]")[[1]]

# try to reproduce but cluster the 
hm <- Heatmap(collapse.mat, 
         cluster_columns = F,
         cluster_column_slices = F,
         cluster_row_slices = T,
         show_row_names = F,
         show_column_names = F,
         top_annotation = col_ha,
         split = split,
         column_split = factor(tstrsplit(colnames(collapse.mat), "[_.]")[[1]], levels=summary.order),
         col=colorRamp2(colors=c("dodgerblue2", "white", "#E31A1C"),  breaks=c(-2,0,2)), 
         border=T,
         name = "log2(intensity/\nmean(timeZero))",
         na_col = 'darkgrey',
         row_title ="M %s",
         row_title_side = 'right',
         row_title_gp = gpar(fontsize=8),
         column_names_gp = gpar (fontsize= 3),
         column_title_gp = gpar (fontsize= 9))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=12))

BackupAsPDF(hm,  'wgcna.modules.collapsed.summaryMatOrdered.heatmap', dimensions = c(9,7))
```
Try an order that just makes sense to us to view

```{r}
pca.order <- c('DOR', 'MOR', 'KOR', 'NPY1R','MRX2', 'NK1R', 'PTGER4', 'PTGER1', 'MRX4', 'QRFPR', 'NPFFR1')

# try to reproduce but cluster the 
hm <- Heatmap(collapse.mat, 
         cluster_columns = F,
         cluster_column_slices = F,
         cluster_row_slices = T,
         show_row_names = F,
         show_column_names = F,
         top_annotation = col_ha,
         split = split,
         column_split = factor(tstrsplit(colnames(collapse.mat), "[_.]")[[1]], levels=pca.order),
         col=colorRamp2(colors=c("dodgerblue2", "white", "#E31A1C"),  breaks=c(-2,0,2)), 
         border=T,
         name = "log2(intensity/\nmean(timeZero))",
         na_col = 'darkgrey',
         row_title ="M %s",
         row_title_side = 'right',
         row_title_gp = gpar(fontsize=8),
         column_names_gp = gpar (fontsize= 3),
         column_title_gp = gpar (fontsize= 9))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=12))
hm
BackupAsPDF(hm,  'wgcna.modules.collapsed.pca1Ordered.heatmap', dimensions = c(9,7))
```
alphabetical order

```{r}
# try to reproduce but cluster the 
hm <- Heatmap(collapse.mat, 
         cluster_columns = F,
         cluster_column_slices = F,
         cluster_row_slices = T,
         show_row_names = F,
         show_column_names = F,
         top_annotation = col_ha,
         split = split,
         column_split = factor(tstrsplit(colnames(collapse.mat), "[_.]")[[1]]),
         col=colorRamp2(colors=c("dodgerblue2", "white", "#E31A1C"),  breaks=c(-2,0,2)), 
         border=T,
         name = "log2(intensity/\nmean(timeZero))",
         na_col = 'darkgrey',
         row_title ="M %s",
         row_title_side = 'right',
         row_title_gp = gpar(fontsize=8),
         column_names_gp = gpar (fontsize= 3),
         column_title_gp = gpar (fontsize= 9))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=12))
hm
BackupAsPDF(hm,  'wgcna.modules.collapsed.alphabeticalOrdered.heatmap', dimensions = c(9,6))
```
Regeenrating the lineplots with the divergent color scale

```{r}
cluster.oi <- c('cluster1', 'cluster6')

g <- ggplot(protQuant[Cluster %in% cluster.oi,], aes(x=timeStr, y=vsTimeZero, group=receptor, color=receptor, label=receptor)) + 
  geom_smooth(stat = 'summary', fun.y = median, se =TRUE, show.legend = T, aes(fill=receptor, alpha=0.2)) +
  scale_color_manual(values = receptorColors) + 
  scale_fill_manual(values = receptorColors) + 
  ggtitle('Median expression vs Time 0') +
  geom_hline(yintercept = 0, linetype=2, alpha=0.6) +
  stat_summary(data=protQuant[Cluster %in% cluster.oi & timeStr == '30',], stat = 'summary', fun.y = median, geom='text', size=3,
                   hjust = -.05, position = position_nudge(x = .05)) +
  #geom_dl()
  labs(x='timepoint (mins)', y="Intensity (log2)") +
  facet_wrap(~cluster_label, scales='free_y', ncol=3) +
  theme_classic() +
  scale_x_discrete(expand=c(0,0.03)) +
  theme(
   panel.spacing.x = unit(2.5, "lines"),
   plot.margin = margin(0.1, 2.6, 0.1, 0.1, "cm"), 
   strip.text = element_text(size = 6)) +
  guides(color=FALSE,
         fill=FALSE,
         alpha=FALSE) +
    coord_cartesian(clip='off') # allow labels to enter canvas boundaries

g 
BackupAsPDF(g, 'clusters.oi.wSE.scales.free.receptorCol.linechart')

```
Figure 2D; median shift modules, reverse clust order so overexpressed at the top of the set

```{r}
receptors=tstrsplit(colnames(collapse.mat), "[_.]")[[1]] %>% 
  unique()

row_ha <- HeatmapAnnotation(receptor=receptors, 
                        col=list(receptor=receptorColors))


# calculate the median shift vs T0 per cluster, per receptor, per timepoint
protQuant[, medianVsTimeZero := median(vsTimeZero, na.rm=T), by=.(Cluster, receptor, timeStr)]

protQuant[,medianVsTimeZero, by=.(Cluster, receptor, timeStr)] %>% 
  unique() %>% 
  .[Cluster == 'cluster8']


# get the maximum shift vs t0 basically taking the max timepoint value
maxShift.mat <- dcast(protQuant, receptor~cluster_label, value.var = 'medianVsTimeZero', fun.aggregate = function(x){x[which.max(abs(x))]}) %>% 
  as.matrix(rownames='receptor') %>% 
  t()

# sanity check; checking in both directions
protQuant[,.N, by=.(Cluster, receptor, timeStr, medianVsTimeZero)][Cluster == 'cluster10' & receptor == 'DOR']
protQuant[,.N, by=.(Cluster, receptor, timeStr, medianVsTimeZero)][Cluster == 'cluster1' & receptor == 'DOR']

# function to cluster w NA values
hc <- clusterWNA(maxShift.mat)
hc$order <- rev(hc$order)

ha <- Heatmap(maxShift.mat, 
              cluster_rows = hc,
              top_annotation =row_ha,
              border=T,
              name = "cluster maximum shift vs T0",
              row_names_gp = gpar(fontsize=8))
ha
BackupAsPDF(ha, 'maxShift.clusterMedian.vsT0.revOrder.heatmap', format='pdf', dimensions = c(8,5))
```
finally plot of the positive controls; just do linecharts and heatmaps
Use a random set of colors

Figure S2E: line charts of positive control proteins (ARRB2, EEA1, and PRKD2) for 11 GPCRs, or heatmap
```{r}
genes.oi <- c('ARRB2', 'EEA1','PRKD2')

protQuant[gene %in% 'PRKD2',]

g <- ggplot(protQuant[gene %in% genes.oi,], aes(x=timeStr, y=vsTimeZero, group=receptor, color=receptor, label=receptor)) + 
 geom_smooth(stat = 'summary', fun.y = mean, se =TRUE, show.legend = T, aes(fill=receptor, alpha=0.2)) +
  scale_color_manual(values = receptorColors) + 
  scale_fill_manual(values = receptorColors) + 
  ggtitle('ARBB2, EEA1 & PKRD2 mean expression vs Time 0') +
  geom_hline(yintercept = 0, linetype=2, alpha=0.6) +
  stat_summary(data=protQuant[gene %in% genes.oi & timeStr == '30',], stat = 'summary', fun.y = median, geom='text', size=3,
                   hjust = -.05, position = position_nudge(x = .05)) +
  labs(x='timepoint (mins)', y="Intensity (log2)") +
  facet_wrap(~gene, scales='free_y', ncol=3) +
  theme_classic() +
  scale_x_discrete(expand=c(0,0.03)) +
  theme(
   panel.spacing.x = unit(2.5, "lines"),
   plot.margin = margin(0.1, 2.6, 0.1, 0.1, "cm"), 
   strip.text = element_text(size = 6)) +
   guides(color=FALSE,
         fill=FALSE,
         alpha=FALSE) +
  coord_cartesian(clip='off') # allow labels to enter canvas boundaries

g 
BackupAsPDF(g, 'posControls.wSE.scalesFree.receptorCol.linechart', dimensions = c(12,7))

# turn off the SE as quite noisy at points
g <- ggplot(protQuant[gene %in% genes.oi,], aes(x=timeStr, y=vsTimeZero, group=receptor, color=receptor, label=receptor)) + 
 geom_smooth(stat = 'summary', fun.y = mean, se =FALSE, show.legend = T, aes(fill=receptor, alpha=0.2)) +
  scale_color_manual(values = receptorColors) + 
  scale_fill_manual(values = receptorColors) + 
  ggtitle('ARBB2, EEA1 & PKRD2 mean expression vs Time 0') +
  geom_hline(yintercept = 0, linetype=2, alpha=0.6) +
  stat_summary(data=protQuant[gene %in% genes.oi & timeStr == '30',], stat = 'summary', fun.y = median, geom='text', size=3,
                   hjust = -.05, position = position_nudge(x = .05)) +
  labs(x='timepoint (mins)', y="Intensity (log2)") +
  facet_wrap(~gene, scales='free_y', ncol=3) +
  theme_classic() +
  scale_x_discrete(expand=c(0,0.03)) +
  theme(
   panel.spacing.x = unit(2.5, "lines"),
   plot.margin = margin(0.1, 2.6, 0.1, 0.1, "cm"), 
   strip.text = element_text(size = 6)) +
   guides(color=FALSE,
         fill=FALSE,
         alpha=FALSE) +
  coord_cartesian(clip='off') # allow labels to enter canvas boundaries

g 
BackupAsPDF(g, 'posControls.woSE.scalesFree.receptorCol.linechart', dimensions = c(12,7))



g <- ggplot(protQuant[gene %in% genes.oi,], aes(x=timeStr, y=vsTimeZero, group=receptor, color=receptor, label=receptor)) + 
  geom_smooth(stat = 'summary', fun.y = mean, se =TRUE, show.legend = T, aes(fill=receptor, alpha=0.2)) +
  scale_color_manual(values = receptorColors) + 
  scale_fill_manual(values = receptorColors) + 
  ggtitle('ARBB2, EEA1 & PKRD2 mean expression vs Time 0') +
  geom_hline(yintercept = 0, linetype=2, alpha=0.6) +
  stat_summary(data=protQuant[gene %in% genes.oi & timeStr == '30',], stat = 'summary', fun.y = median, geom='text', size=1.5,
                   hjust = -.05, position = position_nudge(x = .05)) +
  labs(x='timepoint (mins)', y="Intensity (log2)") +
  facet_wrap(~gene, ncol=3) +
  theme_classic() +
  scale_x_discrete(expand=c(0,0.03)) +
  theme(
   axis.title.x = element_text(size=8),
   axis.text.x=element_text(size=6),
   axis.title.y = element_text(size=8),
   axis.text.y=element_text(size=6),
   plot.title = element_text(size=8, face = 'bold'),
   panel.spacing.x = unit(2.5, "lines"),
   plot.margin = margin(0.1, 2.6, 0.1, 0.1, "cm"), 
   strip.text = element_text(size =6)) +
   guides(color=FALSE,
         fill=FALSE,
         alpha=FALSE) +
  coord_cartesian(clip='off') # allow labels to enter canvas boundaries
g 

BackupAsPDF(g, 'posControls.wSE.scalesFixed.receptorCol.linechart', dimensions = c(7.3,2))


g <- ggplot(protQuant[gene %in% genes.oi,], aes(x=timeStr, y=vsTimeZero, group=receptor, color=receptor, label=receptor)) + 
  geom_smooth(stat = 'summary', fun.y = mean, se =TRUE, show.legend = T, aes(fill=receptor, alpha=0.2)) +
  scale_color_manual(values = receptorColors) + 
  scale_fill_manual(values = receptorColors) + 
  ggtitle('ARBB2, EEA1 & PKRD2 mean expression vs Time 0') +
  geom_hline(yintercept = 0, linetype=2, alpha=0.6) +
  #stat_summary(data=protQuant[gene %in% genes.oi & timeStr == '30',], stat = 'summary', fun.y = median, geom='text', size=1.5,
  #                 hjust = -.05, position = position_nudge(x = .05)) +
  labs(x='timepoint (mins)', y="Intensity (log2)") +
  facet_wrap(~gene, ncol=3) +
  theme_classic() +
  scale_x_discrete(expand=c(0,0.03)) +
  theme(
   legend.title = element_text(size = 6),  # Adjust legend title size
   legend.text = element_text(size = 5),   # Adjust legend text size
   legend.key.size = unit(0.3, "cm"),       # Adjust legend key size
   axis.title.x = element_text(size=8),
   axis.text.x=element_text(size=6),
   axis.title.y = element_text(size=8),
   axis.text.y=element_text(size=6),
   plot.title = element_text(size=8, face = 'bold'),
   panel.spacing.x = unit(2.5, "lines"),
   plot.margin = margin(0.1, 2.6, 0.1, 0.1, "cm"), 
   strip.text = element_text(size =6)) +
   guides(color=guide_legend(override.aes=list(fill='white', size = 1)),
          fill=FALSE,
          alpha=FALSE) +
  coord_cartesian(clip='off') # allow labels to enter canvas boundaries
g 

BackupAsPDF(g, 'posControls.wSE.scalesFixed.receptorCol.wLegend.linechart', dimensions = c(7.3,2))
```

020325- location specific detrended data
Use this to make a paired heatmap with our LFC vs T0 

Readin the vs cyto and filter to this protein set
```{r}
loc.dt <- fread('./output/2025_02_03_LocationSpecificProteinsVsCyto.csv')

bg.mat <- dcast(loc.dt, protein~bg, value.var = "log2FCvsCyto") %>% 
  as.matrix(rownames = "protein")

# read in the collapse matrix 
collapse.mat <- fread('./010725_plotsForKroganWeekly_data/2025_01_22_11.receptor.collapsed.vsT0.csv') %>% 
  dcast(Protein~GROUP, value.var = 'vsTimeZero') %>% 
  as.matrix(rownames='Protein')

# like 93% of proteins in both samples so great!
nrow(collapse.mat[rownames(collapse.mat) %in%  rownames(bg.mat),])/nrow(bg.mat)

# take this set of 
submat <- collapse.mat[rownames(collapse.mat) %in% rownames(bg.mat),]
bg.mat <- bg.mat[rownames(bg.mat) %in% rownames(submat),]


# split the rows by location specific signal
locations <- unique(loc.dt[, .(protein, specificTo)])[rownames(bg.mat), specificTo, on = "protein"]
```

colors 
```{r}
# this is Qiopngyus new palette
receptorColors <- c(
DOR = "#99CCFF",
KOR = "#66CCFF",
MOR = "#3399FF",
MRX2 = "#FF9900",
MRX4 = "#FF6666",
NK1R = "#FF6633",
NPFFR1 = "#9EBCDA",
NPY1R = "#0099CC",
PTGER1 = "#FFCCCC",
PTGER4 = "#85CB9C",
QRFPR = "#FF9999")


col_ha <- HeatmapAnnotation(receptor=tstrsplit(colnames(submat), "[_.]")[[1]],
                            timepoint=tstrsplit(colnames(submat), "[_.]")[[2]],
                            col=list(receptor=receptorColors, timepoint=tp.pal))
```


combine with the collapse matrix plot

```{r}
# using the pca order plot
#pca.order <- c('DOR', 'MOR', 'KOR', 'NPY1R','MRX2', 'NK1R', 'PTGER4', 'PTGER1', 'MRX4', 'QRFPR', 'NPFFR1')

hm1 <- Heatmap (bg.mat, 
                row_split = locations, 
                border=T,
                row_title_rot = 0, 
                show_row_names = FALSE,
                name = "log2FC\nvs Cyto",
                column_names_gp = gpar (fontsize= 8),
                col = circlize::colorRamp2(breaks = seq(from = -2, to = 2
                                                        , length.out = 11),
                                    colors = rev(RColorBrewer::brewer.pal(11, "RdGy"))))



# try to reproduce but cluster the 
hm2 <- Heatmap(submat, 
         cluster_columns = F,
         cluster_column_slices = F,
         cluster_row_slices = T,
         show_row_names = F,
         show_column_names = F,
         top_annotation = col_ha,
         column_split = factor(tstrsplit(colnames(submat), "[_.]")[[1]], levels=pca.order),
         col=colorRamp2(colors=c("dodgerblue2", "white", "#E31A1C"),  breaks=c(-2,0,2)), 
         border=T,
         name = "log2(intensity/\nmean(timeZero))",
         na_col = 'darkgrey',
         row_title ="M %s",
         row_title_side = 'right',
         row_title_gp = gpar(fontsize=8),
         column_names_gp = gpar (fontsize= 3),
         column_title_gp = gpar (fontsize= 9))


BackupAsPDF(draw(hm1 + hm2, ht_gap = unit(c(3), "mm")), 'locationMarkers.LFCvsT0andLFCvsCyto.heatmap', dimensions = c(10,6))

RColorBrewer::display.brewer.all()
```
regenerate with the full dataset to capture the missing proteins
```{r}
# read in DEA results for filtering
mss <- fread('data/2022_08_31_NiceFitsPower3.csv')
# read in exp data
p.quant <- fread('data/2022_08_17_ProteinLevelData.lib.csv.gz')


# fix the groups with an underscore in them MOR_LG and MOR_DAMGO to MOR.DAMGO
p.quant[, GROUP := gsub ("_([A-Z])", ".\\1", GROUP)]
# add some new identifier cols
p.quant[, `:=`(timeStr = str_extract(GROUP,'[0-9]{2}$'),
               receptor = gsub("_[0-9]{2}$", '', p.quant$GROUP)
               )]

group.noi <- p.quant[!GROUP %like% c("DOR|KOR|MOR|MRX2|MRX4|NK1R|NPFFR1|NPY1R|PTGER1|PTGER4|QRFPR|DOR"),unique(GROUP)]
p.quant <- p.quant[!GROUP %in% group.noi, ]
p.quant <- p.quant[!GROUP %like% "MOR.LG94",]
p.quant[GROUP %like% 'MOR[.]DAMGO', GROUP := gsub("[.]DAMGO", "", GROUP)]

# take vs t0 scores
p.quant[, batchControlled := LogIntensities - median(LogIntensities, na.rm = TRUE), by = .(Protein, receptor, SUBJECT)]
p.quant[, vsTimeZero := batchControlled - mean(batchControlled[timeStr == timeStr[!is.na(batchControlled)][1] ]), by = .(Protein, receptor)]

full.mat <- dcast(p.quant, Protein~GROUP, value.var = 'vsTimeZero', fun.aggregate = mean,na.rm=T) %>% 
  as.matrix(rownames='Protein')


# reload location dt
loc.dt <- fread('./output/2025_02_03_LocationSpecificProteinsVsCyto.csv')

bg.mat <- dcast(loc.dt, protein~bg, value.var = "log2FCvsCyto") %>% 
  as.matrix(rownames = "protein")

marker.mat <- full.mat[rownames(full.mat) %in% rownames(bg.mat),]

# split the rows by location specific signal
locations <- unique(loc.dt[, .(protein, specificTo)])[rownames(bg.mat), specificTo, on = "protein"]


hm1 <- Heatmap (bg.mat, 
                row_split = locations, 
                border=T,
                row_title_rot = 0, 
                show_row_names = FALSE,
                name = "log2FC\nvs Cyto",
                column_names_gp = gpar (fontsize= 8),
                col = circlize::colorRamp2(breaks = seq(from = -2, to = 2
                                                        , length.out = 11),
                                    colors = rev(RColorBrewer::brewer.pal(11, "RdGy"))))



# try to reproduce but cluster the 
hm2 <- Heatmap(marker.mat, 
         cluster_columns = F,
         cluster_column_slices = F,
         cluster_row_slices = T,
         show_row_names = F,
         show_column_names = F,
         top_annotation = col_ha,
         column_split = factor(tstrsplit(colnames(marker.mat), "[_.]")[[1]], levels=pca.order),
         col=colorRamp2(colors=c("dodgerblue2", "white", "#E31A1C"),  breaks=c(-2,0,2)), 
         border=T,
         name = "log2(intensity/\nmean(timeZero))",
         na_col = 'darkgrey',
         row_title ="M %s",
         row_title_side = 'right',
         row_title_gp = gpar(fontsize=8),
         column_names_gp = gpar (fontsize= 3),
         column_title_gp = gpar (fontsize= 9))


BackupAsPDF(draw(hm1 + hm2, ht_gap = unit(c(3), "mm")), 'locationMarkers.LFCvsT0andLFCvsCyto.heatmap', dimensions = c(10,6))
```

looks good; 
```{r}
Heatmap(submat, cluster_rows = F, cluster_columns = F) +
  Heatmap(test.mat, cluster_rows = F, cluster_columns = F)


rownames
```

