---
title: "041223.newFig5plots"
author: "Martin Gordon"
date: "2023-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

TODO
---
Double check our top hits with IUPAR classifications
check our summary data.table tn ensure group averages are correctly calculated

## DARPA regenerate the fig 5 plots requested by the group
Outline of work:
Figure 5.
S-curve rank plot with the modified labels of genes (prelim done)

the correlation scores of top proteins for each G protein type heatmap, with clusters of proteins specific to each G protein type

intensity of proteins from clusters specific to each G protein type from heatmap b, across all the receptors.

plots for positive controls: plot the highest log2FC for each receptor and cluster to bars in group of G proteins. (Gq: PRKCA, PACS1, UNC13B; Gi: EYA4, maybe KCTD12).

Venn diagram of candidates significant for YM 5 min vs. 5 min, overlapped with the correlation analysis Gq top correlation candidates. Before we decide to use 5 min or 10 min, we could check the heatmap from a pair-wise comparison for 5min +/- YM and 10min +/- YM.
For figure 2: the WGCNA heatmap, cluster the heatmap for receptors based on the largest log2FC.
For the GO enrichment in the current figure 3: resize it to enable a larger font size of GO terms.


## Todo
Maybe generate the 5th percentile heatmap
For now focus on other requests; complete the Venn Diagram

packages

```{r}
library(data.table)
library(ggplot2)
library(ComplexHeatmap)
library(WGCNA)
library(magrittr)
library(stringr)
library(ggbeeswarm)
library(ggrepel)
library(patchwork)
library(readxl)
library(circlize)

#Ben utils
source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")


# helper scripts for module/network export etc
source ("/Users/martingordon/Documents/utils/mg_utils/HelperFunctions.R")
```

first thing to do is tidy the S-rank plots produced previously
We need to seperate out the ggrepel hits so they are visible 

Read in the results from the correlation produced previously

Correlations with the IUPHAR classification

plot output dir 
```{r}
plot.outdir <- '/Users/martingordon/Documents/projects/082423_QLi_GPR_WGCNA/041223.regen.fig5plots_data/pdfs/'
```


lets first check the raw data 

```{r}
pred <- fread("/Users/martingordon/Documents/projects/310723_QLi_GPCR.correlations/data/predictionScores.txt")
exp <- fread ("/Users/martingordon/Documents/projects/310723_QLi_GPCR.correlations/data/experimentalScores.txt")
iupar <- data.table(read_xlsx("/Users/martingordon/Documents/projects/310723_QLi_GPCR.correlations/data/IUPHAR_11GPCRs_Gprotein_class_primary.xlsx"))

#exp.long <- melt(exp, id.vars = "GPCR", measure.vars = grep ("GNA", colnames(pred)), variable.name = "GProtein", value.name = "score")[, score := as.numeric(score)][]
pred.long <- melt(pred, id.vars = "GPCR", measure.vars = grep ("GNA", colnames(pred)), variable.name = "GProtein", value.name = "score")[, score := as.numeric(score)][]

# convert to long format
iupar.long <- melt(iupar, id.vars = 'GPCR', measure.vars = grep('GPCR', colnames(iupar), invert=T),  variable.name = "GProtein", value.name = "score") %>% 
  .[, score := as.numeric(score)] 
```


```{r scores.list}

scores.list <- list(exp = exp.long,
                    pred = pred.long,
                    iupar = iupar.long)
```
Convert the scores to a matrix

```{r}
scores.mat <- lapply(scores.list, function(x){
  dcast(x, GProtein~GPCR, value.var = 'score') %>% 
    as.matrix(rownames = 'GProtein')
})

lapply(scores.mat, colnames)
```
apply the scores mat to the data
```{r}
scores.mat <- lapply(scores.mat, function(x){
  colnames(x) <- c("MRX2","MRX4","NPFFR1","NPY1R","DOR","KOR","MOR","PTGER1","PTGER4","QRFPR","NK1R")
  return(x)
})
```

reorder the mss columns to follow the order above 

Restrict the analysis to the sig genes from the APEX experiment (pval < 0.005 & abs log2FC(1.5))

```{r}
col.order <-  c("MRX2","MRX4","NPFFR1","NPY1R","DOR","KOR","MOR","PTGER1","PTGER4","QRFPR","NK1R")

mss.out <- fread('./data/2022_08_31_NiceFitsPower3.csv')

exp.mat <- dcast(mss.out,gene~receptor, value.var = 'log2FC') %>% 
  .[, MOR.LG94 := NULL] %>% 
  as.matrix(rownames ='gene')


# filter to sig genes and reorder
sig.genes <- mss.out[pvalue < 0.005 & abs(log2FC) > log2(1.5) & receptor != 'MOR.LG94', unique(gene) ]
exp.mat <- exp.mat[rownames(exp.mat) %in% sig.genes,]

'DIAPH2' %in% sig.genes
# renanme MOR DAMGO to MOR
colnames(exp.mat) <- c('DOR', 'KOR', 'MOR', 'MRX2', 'MRX4', 'NK1R', 'NPFFR1', 'NPY1R','PTGER1', 'PTGER4','QRFPR')

# reorder the cols to match the matrices
col.order <-  c("MRX2","MRX4","NPFFR1","NPY1R","DOR","KOR","MOR","PTGER1","PTGER4","QRFPR","NK1R")

exp.mat[1:10,1:10]
exp.mat <- exp.mat[, col.order]
exp.mat[1:10,1:10]


# sanit check; looks fine
lapply(scores.mat, colnames)
colnames(exp.mat)

exp
```

run the correlations for each of the datasets

```{r}
# cors run colwise with the two vectors
cor.list <- lapply(scores.mat, function(x){
  cor(t(exp.mat), t(x), use= "pairwise.complete.obs")
})

cor.long <- lapply(cor.list, function(x){
  x <- melt(as.data.table(x, keep.rownames = TRUE), id.vars = "rn", variable.name  = "GProtein", value.name = "R")
  setnames(x, old = "rn", new = "gene")
  return(x)
})
```

count number of observations, only want to take results with vals observed for all receptors to reduce FPs

```{r}
cor.long$iupar[, .SD[which.max(R)], by=GProtein]

# most of these have values for all the 11 receptors
obs.counts <- mss.out[!is.na(log2FC) & receptor != 'MOR.LG94' & gene %in% sig.genes, .N, by = gene]
hist(obs.counts$N, breaks=seq(1,11))

```

add this info to the correlation res

```{r}
lapply(cor.long, function(x){
  
  x[obs.counts, numObservations := N, on = 'gene']
  setorder(x, -numObservations, -R, na.last = TRUE) #strongest associations at top
})
```
write these files out to folder for reuse

```{r}
#lapply(seq_along(cor.long), function(x,n, i){
#  fwrite(x[[i]], ScriptAndDatedFileName(paste0(n[[i]], 'gprotein.corscores.csv')) ) 
#}, x=cor.long, n=names(cor.long))
```




read in the iupar result
```{r}
#iupar.cor.dt <- fread(ScriptAndDatedFileName('iupargprotein.corscores.csv'))
iupar.cor.dt <- fread("/Users/martingordon/Documents/projects/082423_QLi_GPR_WGCNA/041223.regen.fig5plots_data/2023_12_04_iupargprotein.corscores.csv")
iupar.cor.dt <- iupar.cor.dt[numObservations == 11,] # only keep those with 11 observations

gprots <-  unique(iupar.cor.dt$GProtein)
```

Aside: create the 5 and 2 percentile data in w format and share

```{r}
snd.pc <- iupar.cor.dt[, lab := ifelse( (R >= quantile(R,0.98) | R <= quantile(R,0.02)), unique(gene), ''), by=GProtein] %>% 
  .[lab != '',unique(gene)]
fifth.pc <- iupar.cor.dt[, lab := ifelse( (R >= quantile(R,0.95) | R <= quantile(R,0.05)), unique(gene), ''), by=GProtein] %>% 
  .[lab != '', unique(gene)]
  
snd.pc %>%  length()
fifth.pc %>%  length() # still few overlaps..

# write out these and also the entire list of genes
snd.dt.wide <- iupar.cor.dt[gene %in% snd.pc,] %>% 
  dcast(., gene~GProtein, value.var='R')

fifth.dt.wide <- iupar.cor.dt[gene %in% fifth.pc,] %>% 
  dcast(., gene~GProtein, value.var='R')


# write out data in wide format
fwrite(snd.dt.wide,  ScriptAndDatedFileName('2perc.gprot.corscores.wide.csv'))
fwrite(fifth.dt.wide, ScriptAndDatedFileName('5perc.gprot.corscores.wide.csv'))

# write out complete dt in wide format 
fwrite(dcast(iupar.cor.dt, gene~GProtein, value.var='R'), ScriptAndDatedFileName('all.gprot.corscores.wide.csv'))

# write out filtered data in long format
fwrite(snd.dt.wide,  ScriptAndDatedFileName('2perc.gprot.corscores.wide.csv'))
fwrite(fifth.dt.wide, ScriptAndDatedFileName('5perc.gprot.corscores.wide.csv'))

# write out long format
snd.dt <- iupar.cor.dt[, lab := ifelse( (R >= quantile(R,0.98) | R <= quantile(R,0.02)), unique(gene), ''), by=GProtein] %>% 
  .[lab != '', ]

fifth.dt <- iupar.cor.dt[, lab := ifelse( (R >= quantile(R,0.95) | R <= quantile(R,0.05)), unique(gene), ''), by=GProtein] %>% 
  .[lab != '',]

fwrite(snd.dt,  ScriptAndDatedFileName('2perc.gprot.corscores.csv'))
fwrite(fifth.dt, ScriptAndDatedFileName('5perc.gprot.corscores.csv'))
```





Request1: s-rank plots
-----

Lets take and label the top two percent
```{r}
iupar.cor.dt

lapply(seq_along(gprots), function(x,i){
  
  g <- ggplot(data=iupar.cor.dt[GProtein == x[[i]]], aes(x=reorder(gene,R), y=R, col=GProtein, label=lab)) +
    geom_point(size=1) +
    geom_hline(aes(alpha=0.6), yintercept = 0, linetype='dotdash') +
    geom_text_repel(data=cor.long$iupar[GProtein == x[[i]]], aes(label=lab),segment.linetype=3, segement.color='grey', max.overlaps = 30,size=1.5, position = 'dodge') +
    xlab('genes (Pearson r ordered)') +
    theme_classic() +
    scale_x_discrete(expand=expansion(add=50)) +
    scale_color_manual(values=c(col.pal[[i]])) +
    ggtitle(paste0(x[[i]], ' Pearson r')) +
    theme(axis.text.x = element_blank()
        )
  g
 #BackupAsPDF(g, 'rankplot.Gprots.R0.7.')
},x=gprots)

```
as a basic first pass; maybe we can extract the top and bottom percentile and plot these as windows to the side?
color by top hits; blue red grey

modify slightly and inlay the two plots; the postively correlated at top and negative at bottom

```{r}
# define the labels we wish to use 
iupar.cor.dt[, sig := '']
iupar.cor.dt[, lab:= '']
iupar.cor.dt[, lab := ifelse( (R >= quantile(R,0.98) | R <= quantile(R,0.02)), unique(gene), ''), by=GProtein]
iupar.cor.dt[lab != '', sig := ifelse(R > 0, 'pos', 'neg')]

iupar.cor.dt[, .N,  by=sig]
iupar.cor.dt$sig

lapply(seq_along(gprots), function(x,i){
  
  
  outlab <- gsub('/', '.', x[[i]])
  
  g <- ggplot(data=iupar.cor.dt[GProtein == x[[i]]], aes(x=reorder(gene,R), y=R,label=lab, color=sig)) +
    geom_point(size=1) +
    geom_hline(aes(alpha=0.7), yintercept = 0, linetype='dotdash') +
    scale_color_manual(values=c('grey'='', 'neg'='blue', 'pos'='red')) +
    #geom_text_repel(data=cor.long$iupar[GProtein == x[[i]]], aes(label=lab),segment.linetype=3, max.overlaps = 20, size=1.5, position = 'dodge') +
    xlab('genes (Pearson r ordered)') +
    theme_classic() +
    guides(color=FALSE) +
    scale_x_discrete(expand=expansion(add=50)) + # just to see both ends of the plot
     theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
  
  # create up.dt
  up.dt <-  iupar.cor.dt[GProtein == x[[i]] & R > 0 & lab != '', ]
  
  up.g <- ggplot(data=up.dt, aes(x=reorder(gene,R), y=R,label=lab, color=sig)) +
    geom_point(size=1) +
    #geom_hline(aes(alpha=0.7), yintercept = 0, linetype='dotdash') +
    scale_color_manual(values=c('grey'='', 'neg'='blue', 'pos'='red')) +
    geom_text_repel(data=up.dt, aes(label=lab),segment.linetype=3, max.overlaps = 30, size=1.5, position = 'dodge') +
    xlab('genes (Pearson r ordered)') +
    theme_classic() +
    guides(color=FALSE) +
    #scale_x_discrete(expand=expansion(add=50)) + # just to see both ends of the plot
    ggtitle('98th Percentile') +
      theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
  
   down.dt <-  iupar.cor.dt[GProtein == x[[i]] & R < 0 & lab != '', ]
  
   down.g <- ggplot(data=down.dt, aes(x=reorder(gene,-R), y=R,label=lab, color=sig)) +
    geom_point(size=1) +
    #geom_hline(aes(alpha=0.7), yintercept = 0, linetype='dotdash') +
    scale_color_manual(values=c('grey'='', 'neg'='blue', 'pos'='red')) +
    geom_text_repel(data=down.dt, aes(label=lab),segment.linetype=3, max.overlaps = 30, size=1.5, position = 'dodge') +
    xlab('genes (Pearson r ordered)') +
    theme_classic() +
    guides(color=FALSE) +
    #scale_x_discrete(expand=expansion(add=50)) + # just to see both ends of the plot
    ggtitle('2nd Percentile') +
      theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
  
  comb.g <- g + (up.g / down.g) +
  plot_annotation(title = paste0(x[[i]], ' Pearson R')) +
  plot_layout(widths = c(1,1))


  #ggsave(paste0(plot.outdir, outlab, 'gprot.cor.faceted.pdf'),comb.g, width = 8, height = 6)
    
 #BackupAsPDF(ggsave(comb.g), paste0(x[[i]], 'gprot.cor.faceted.'))
},x=gprots)

```
I think this actually looks ok and is a better representation of the data.. maybe still abit overcrowded,  but looks ok
retry but in lay the plots

edit: 'flip' the negatively correlated genes so we can see
could also go with the 1D option to save space
```{r}

lapply(seq_along(gprots), function(x,i){
  
  outlab <- gsub('/', '.', x[[i]])
  
  g <- ggplot(data=iupar.cor.dt[GProtein == x[[i]]], aes(x=reorder(gene,R), y=R,label=lab, color=sig)) +
    geom_point(size=1) +
    geom_hline(aes(alpha=0.7), yintercept = 0, linetype='dotdash') +
    scale_color_manual(values=c('grey'='', 'neg'='blue', 'pos'='red')) +
    #geom_text_repel(data=cor.long$iupar[GProtein == x[[i]]], aes(label=lab),segment.linetype=3, max.overlaps = 20, size=1.5, position = 'dodge') +
    xlab('genes (Pearson r ordered)') +
    theme_classic() +
    guides(color=FALSE) +
    scale_x_discrete(expand=expansion(add=50)) + # just to see both ends of the plot
     theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
  
  # create up.dt
  up.dt <-  iupar.cor.dt[GProtein == x[[i]] & R > 0 & lab != '', ]
  
  up.g <- ggplot(data=up.dt, aes(x=reorder(gene,R), y=R,label=lab, color=sig)) +
    geom_point(size=1) +
    #geom_hline(aes(alpha=0.7), yintercept = 0, linetype='dotdash') +
    scale_color_manual(values=c('grey'='', 'neg'='blue', 'pos'='red')) +
    geom_text_repel(data=up.dt, aes(label=lab),segment.linetype=3, max.overlaps = 10, size=1.5, position = 'dodge') +
    #xlab('genes (Pearson r ordered)') +
    theme_bw() +
    guides(color=FALSE) +
    #scale_x_discrete(expand=expansion(add=50)) + # just to see both ends of the plot
    ggtitle('98th Percentile') +
      theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        #axis.text.y=element_blank(),
        #axis.title.y=element_blank(),
        axis.title.x.top = element_blank(),
        axis.ticks.x=element_blank())
  
   down.dt <-  iupar.cor.dt[GProtein == x[[i]] & R < 0 & lab != '', ]
  
   down.g <- ggplot(data=down.dt, aes(x=reorder(gene,R), y=R,label=lab, color=sig)) +
    geom_point(size=1) +
    #geom_hline(aes(alpha=0.7), yintercept = 0, linetype='dotdash') +
    scale_color_manual(values=c('grey'='', 'neg'='blue', 'pos'='red')) +
    geom_text_repel(data=down.dt, aes(label=lab),segment.linetype=3, max.overlaps = 10, size=1.5, position = 'dodge') +
    xlab('genes (Pearson r ordered)') +
    theme_bw() +
    guides(color=FALSE) +
    #scale_x_discrete(expand=expansion(add=50)) + # just to see both ends of the plot
    ggtitle('2nd Percentile') +
      theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        #axis.text.y=element_blank(),
       # axis.title.y=element_blank(),
        axis.title.x.top = element_blank(),
        axis.ticks.x=element_blank())
  
  comb.g <- g + 
    inset_element(up.g, left = 0, bottom = 0.55, right=0.55,  top = 1) +
    inset_element(down.g, left = 0.45,  bottom = 0.01, right = 1, top = 0.45) +
    plot_annotation(title = paste0(x[[i]], ' Pearson R')) 
    

  ggsave(paste0(plot.outdir, outlab, 'gprot.cor.inlaid.reorderneg.pdf'),comb.g, width = 8, height = 6)
},x=gprots)
```
correlation scores of these top scoring prots; visualised as a heatmap
we want to find clusters specific to the receptors we are interested in 

150 proteins that we find strongly (anti)correlated with one of the GProtein receptors

What does the 11 receptor clustering look like? when we take the top 4% of hits
getting 150
```{r}
genes.oi <- sapply(gprots, function(x){
  sub.dt <- iupar.cor.dt[GProtein == x,]
  return(sub.dt[R >= quantile(R,0.98) | R <= quantile(R,0.02), unique(gene)])
}) %>% c() %>% unique()

# each GProtein cor has returned many unique values. Perhaps a good sign!
genes.oi %>%  length()

# dt of gene lists
genes.list <- sapply(gprots, function(x){
  sub.dt <- iupar.cor.dt[GProtein == x,]
  return(sub.dt[R >= quantile(R,0.98) | R <= quantile(R,0.02), unique(gene)])
}) %>%  data.table()
```
v little overlap in sig hits, which is good but no strong signal for Gs

Request 3:
intensity of proteins from clusters specific to each G protein type from heatmap b, across all the receptors.

```{r}
subMat <-  exp.mat[rownames(exp.mat) %in% genes.oi,]

# name and reorder cols to match GProt annotation

# create a different scoring metric

GProtein <- c('Gi/Go & Gq/G11', 'Gq/G11', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gq/G11', 'Gs', 'Gi/Go & Gq/G11', 'Gq/G11 & Gs')

ha <- HeatmapAnnotation(GProtein = GProtein,
                        col=list(GProtein=c('Gi/Go & Gq/G11'='pink', 'Gq/G11'='lightblue',
                                            'Gi/Go'='lightgreen', 'Gs'='#CBC3E3','Gq/G11 & Gs'='#FFD580')) )
                        
hm <- Heatmap(subMat,
              top_annotation = ha,
        cluster_columns = T, 
        clustering_distance_columns = 'pearson',
        clustering_distance_rows = 'pearson',
        show_column_names = T,
        name='log2FC',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 4),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))
        
hm <- draw(hm, column_title='2nd & 98th Percentile R scores')
#BackupAsPDF(hm, ('2.and.98th.percentiles.log2FC.heatmap'), dimensions=c(8,8))
```
lets 'cluster' the log2FC values and see what we get
```{r}
hm <- Heatmap(subMat,
              top_annotation = ha,
        cluster_columns = T, 
        clustering_distance_columns = 'pearson',
        clustering_distance_rows = 'pearson',
        show_column_names = T,
        row_km=3,
        name='log2FC',
        na_col = 'darkgrey',
        row_title = c("", "Gq/G11", "Gi/Go"),
        #row_title =sprintf("%d Proteins",
        #                    nrow(subMat)),
         row_names_gp = gpar (fontsize= 4),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))
     hm   
hm <- draw(hm, column_title='2nd & 98th Percentile R scores')
#BackupAsPDF(hm, ('2.and.98th.percentiles.log2FC.km3.heatmap'), dimensions=c(8,8))
```
extract these cluster genes and visualise

```{r}
lfc.clust.genes <- extractClustersfromHeatmap(hm, subMat)

lfc.clust.genes[, ID := fcase(
  cluster == 1, '',
  cluster == 2, 'Gq/G11',
  cluster == 3, 'Gi/Go'
)]

#fwrite(lfc.clust.genes, ScriptAndDatedFileName('lfc.heatmap.cluster.genes.km3.csv'))
lfc.clust.genes[, enrich_grp := ID]

# load the GO table
gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")

# our background? only genes used for WGCNA analysis
universe <- mss.out$gene %>%  unique()

enrich.dt <- enricherOnGroups(lfc.clust.genes[cluster !=1,], groupColumns = 'enrich_grp', geneColumn = "feature", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')

enrich.dt[p.adjust < 0.05, ]

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, groupColumn = 'enrich_grp', topN = 8,
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 6), upperThreshold = 7,  cluster_columns =F)

BackupAsPDF(ht, 'gprotein.log2FC.cor.GOenrichment.kmgroups.', dimensions=c(8,12))

#save the enrichment results too 
#fwrite(enrich.dt, ScriptAndDatedFileName('gprotein.log2FC.cor.GOenrichment.csv'))
```

Heatmap looks pretty good all in all, cant really see much Gs specific clustering
Is this an artifact of the IUPAR scoring? lots more 0's so maybe a lot of this is noise... enrichment might tell us more

request3: the correlation scores of top proteins for each G protein type heatmap, with clusters of proteins specific to each G protein type

Heatmap of pearson distance of R scores

```{r}
cor.mat <- dcast(iupar.cor.dt, gene~GProtein, value.var = 'R') %>% 
  as.matrix(rownames='gene')

subMat <-  cor.mat[rownames(cor.mat) %in% genes.oi,]

GProtein <- c('Gi/Go', 'Gq/G11', 'Gs')

ha <- HeatmapAnnotation(GProtein = GProtein,
                        col=list(GProtein=c('Gi/Go'='#22A884FF', 'Gq/G11'='#440154FF',
                                            'Gs'='#FDE725FF')) )

hm <- Heatmap(subMat,
        top_annotation = ha,
        cluster_columns = T, 
        row_km = 3,
        row_gap = unit(1.5, "mm"),
        clustering_distance_rows = 'euclidean',
        name='Pearson r',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 4),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))

hm


hm.k3 <- draw(hm, column_title='2nd & 98th Percentile R scores')

BackupAsPDF(hm.k3, 'R.2nd.98thpercentile.gprotein.cor.3k.heatmap', dimensions = c(8,8))
```

try 5-k for more ingrained differences

```{r}
hm <- Heatmap(subMat,
        top_annotation = ha,
        cluster_columns = T, 
        row_km = 5,
        row_gap = unit(1.5, "mm"),
        clustering_distance_rows = 'euclidean',
        name='Pearson r',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 4),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))


hm.k5 <- draw(hm, column_title='2nd & 98th Percentile R scores')
BackupAsPDF(hm.k5, 'R.2nd.98thpercentile.gprotein.cor.5k.heatmap', dimensions = c(8,8))
```
What is going on in these clusters?
extract the hits in k3 and k5 clusters and enrich
```{r}
k3.clust.genes <- extractClustersfromHeatmap(hm.k3, subMat)
k5.clust.genes <- extractClustersfromHeatmap(hm.k5, subMat)

k3.clust.genes[, cluster := paste0('km3.', cluster)]
k5.clust.genes[, cluster := paste0('km5.', cluster)]

comb.clust <-  rbind(k3.clust.genes, k5.clust.genes)

comb.clust[, enrich_grp := cluster]

# load the GO table
gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")

# our background? only genes used for WGCNA analysis
universe <- mss.out$gene %>%  unique()

enrich.dt <- enricherOnGroups(comb.clust, groupColumns = 'enrich_grp', geneColumn = "feature", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')



col.order <- c("km3.1", "km3.2", "km3.3", "km5.1", "km5.2", "km5.3", "km5.4", "km5.5")
col.split.vec <- c('k=3', 'k=3', 'k=3', 'k=5', 'k=5', 'k=5', 'k=5')

enrich.dt[, enrich_grp := factor(enrich_grp, levels = col.order)]

simp.enrich$simplified$enrich_grp %>%  unique() %>%  length()

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, groupColumn = 'enrich_grp', topN = 8,
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 6), upperThreshold = 7,  cluster_columns =F, column_split=col.split.vec)

BackupAsPDF(ht, 'gprotein.cor.GOenrichment.kmgroups.', dimensions=c(8,12))
```
plots for positive controls: plot the highest log2FC for each receptor and cluster to bars in group of G proteins. (Gq: PRKCA, PACS1, UNC13B; Gi: EYA4, maybe KCTD12).
For starters, extract the top hits
For ours, take the top 10 hits (up and down) for each GProtein

```{r}
qiongyu.controls <-  c('PRKCA', 'PACS1', 'UNC13B', 'EYA4', 'KCTD12') #include the top 10 hits in both directions and the other two genes Qiongyu mentioned

# Gi controls are not super positively cor with GProtein group
pos.controls <- iupar.cor.dt[(gene %in% c('PRKCA', 'PACS1', 'UNC13B') & GProtein == 'Gq/G11' | gene %in% c('EYA4', 'KCTD12') & GProtein == 'Gi/Go'),]

# this is just taking top hits from the data; look at Qiongyu data first
pos.cor <- iupar.cor.dt[, .SD[order(-R) %>% head(10), .(gene,R)], by=GProtein]
neg.cor <- iupar.cor.dt[, .SD[order(-R) %>% tail(10), .(gene,R)], by=GProtein]

# take the hits mentioned, plust the top 10 hits up and down in each
data.pos.controls <- rbind(pos.cor, neg.cor, iupar.cor.dt[gene %in% c('EYA4', 'KCTD12'), .(gene,R,GProtein)])
```

subset the de data for proteins included in the correlation analysis
also, duplicate records for members of more than 1 gprotein group

```{r}
col.pal <-  randomcoloR::distinctColorPalette(k=5)

mss.corprots <- mss.out[receptor != 'MOR.LG94' & gene %in% pos.controls$gene,]

mss.corprots[receptor == "MOR.DAMGO", receptor := 'MOR']

# create another label for gprotein
mss.corprots[, iupar := fcase(
  receptor %in% c('PTGER4'), 'Gs',
  receptor %in% c('MOR', 'DOR', 'KOR', 'NPFFR1', 'NPY1R'), 'Gi/Go', 
  receptor %in% c('MRX2', 'QRFPR'), 'Gi/Go & Gq/G11',
  receptor %in% c('PTGER1', 'MRX4'), 'Gq/G11',
  receptor %in% c('NK1R'), 'Gq/G11 & Gs'
)]

# need to duplicate records for receptors assigned to more than one group
dup.rows <- mss.corprots[ grepl("\\&", iupar),] %>% 
  .[, plot.grp := str_extract(iupar, "(?<=&\\s)(Gs|Gq/G11)" )] #positve lookbehind & for matching

# clean the mss.corplots 
mss.corprots[, plot.grp := ifelse(grepl("\\&", iupar), str_extract(iupar, "(Gi/Go|Gq/G11)(?=\\s&)" ), iupar )] #positive lookahead for space & 

mss.corprots <-  rbind(mss.corprots, dup.rows)

#sanity check; looks good
mss.corprots[plot.grp != iupar,] %>% 
  .[order(gene,receptor)]

g <- ggplot(mss.corprots, aes(x=receptor, y=log2FC, fill=plot.grp, label=gene)) + 
  geom_boxplot(alpha=0.6) +
  geom_point(color='black') + 
  geom_hline(yintercept=0, linetype=2, color='grey', alpha=0.5) +
  scale_fill_manual(values=(col.pal)) +
  facet_wrap(~plot.grp, scales="free_x") +
  geom_text_repel(data=mss.corprots, color='black', size=1.5, vjust=-2) +
  theme_bw() +
  guides(
  colour = guide_legend('GProtein')) +
  theme(axis.text.x = element_text(angle=90)) 

g
BackupAsPDF(g, 'positive.control.profiles.boxplot.', dimensions=c(9,5))

g <- ggplot(mss.corprots, aes(x=receptor, y=log2FC, color=plot.grp, label=gene)) + 
  #geom_boxplot(alpha=0.6) +
  geom_point() + 
  geom_hline(yintercept=0, linetype=2, color='grey', alpha=0.5) +
  scale_fill_manual(values=(col.pal)) +
  facet_wrap(~plot.grp, scales="free_x") +
  geom_text_repel(data=mss.corprots, color='black', size=1.5, vjust=-2) +
  theme_bw() +
  guides(
  colour = guide_legend('GProtein')) +
  theme(axis.text.x = element_text(angle=90)) 

g
BackupAsPDF(g, 'positive.control.profiles.noboxplot.', dimensions=c(9,5))
```
This does not look good.... shows the postive controls are also high in other IUPAR groups
Take another view; take each gene and plot calculate their average score across groups

calculate mean of (Gq: PRKCA, PACS1, UNC13B; Gi: EYA4, maybe KCTD12)

```{r}
summary.dt <- mss.corprots[, .(mean.log2FC = mean(log2FC), sd.log2FC=sd(log2FC)), by=.(plot.grp,gene)]
summary.dt[gene == 'UNC13B']

summary.dt %>%  str()
# plot the mean expression of each protein, plus the SD per gprotein
g <- ggplot(summary.dt, aes(x=gene, y=mean.log2FC, fill=gene)) +
  geom_bar(stat = 'Identity', alpha=0.8) + 
  geom_errorbar(aes(ymin=mean.log2FC-sd.log2FC, ymax=mean.log2FC+sd.log2FC, group=gene), width=0.2, lwd=.5, color='darkgrey') +
  facet_wrap(~plot.grp) +
  theme_bw() +
  scale_fill_manual(values=(col.pal)) +
  guides(
  colour = guide_legend('GProtein')) +
  theme(axis.text.x = element_text(angle=90)) 
g
#BackupAsPDF(g, 'positive.controls.geneMeanperGProtgroup.barplot')


mss.corprots[,.N, .(plot.grp, iupar)]


```


intensity of proteins from clusters specific to each G protein type from heatmap b, across all the receptors.
I think this is the log2FC plot covers this...
what do you mean with this? an average log2FC? 
need to consider up and down cor seperately as otherwise will even out..
go with km3 as simplier; look at the average intensity overtime
-----

```{r}
k3.clust.genes
```

plots for positive controls: plot the highest log2FC for each receptor and cluster to bars in group of G proteins. (Gq: PRKCA, PACS1, UNC13B; Gi: EYA4, maybe KCTD12).
----

How does the top 10% of hits look? Think nearly taking top 2 percentiles more specific? less FP results...

```{r}
genes.oi <- sapply(gprots, function(x){
  sub.dt <- iupar.cor.dt[GProtein == x,]
  return(sub.dt[R >= quantile(R,0.95) | R <= quantile(R,0.05), unique(gene)])
}) %>% c() %>% unique()



subMat <-  exp.mat[rownames(exp.mat) %in% genes.oi,]

colnames(subMat)

# create a different scoring metric
GProtein <- c('Gi/Go & Gq/G11', 'Gq/G11', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gq/G11', 'Gs', 'Gi/Go & Gq/G11', 'Gq/G11 & Gs')

ha <- HeatmapAnnotation(GProtein = GProtein,
                        col=list(GProtein=c('Gi/Go & Gq/G11'='pink', 'Gq/G11'='lightblue',
                                            'Gi/Go'='lightgreen', 'Gs'='#CBC3E3','Gq/G11 & Gs'='#FFD580')) )
                        
hm <- Heatmap(subMat,
              top_annotation = ha,
        cluster_columns = T, 
        clustering_distance_columns = 'pearson',
        clustering_distance_rows = 'pearson',
        show_column_names = T,
        col = circlize::colorRamp2(breaks=c(-2,0,2), colors=c('blue', 'white', 'red')),
        name='log2FC',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 4),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))
        
hm <- draw(hm, column_title='5nd & 95th Percentile R scores')
BackupAsPDF(hm, ('5.and.95th.percentiles.log2FC.heatmap'), dimensions=c(8,8))
```
cor heatmap of top 5 percentiles

```{r}
subMat <-  cor.mat[rownames(cor.mat) %in% genes.oi,]

GProtein <- c('Gi/Go', 'Gq/G11', 'Gs')

ha <- HeatmapAnnotation(GProtein = GProtein,
                        col=list(GProtein=c('Gi/Go'='#22A884FF', 'Gq/G11'='#440154FF',
                                            'Gs'='#FDE725FF')) )

hm <- Heatmap(subMat,
        top_annotation = ha,
        cluster_columns = T, 
        clustering_distance_rows = 'pearson',
        name='Pearson r',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 4),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))

hm
hm <- draw(hm, column_title='5nd & 95th Percentile R scores')

BackupAsPDF(hm, 'R.5nd.95thpercentile.gprotein.cor.heatmap', dimensions = c(8,8))
```





Not used...
----


Can we do a better job of highlighting run to run trends?
Maybe we can scale/standaridse the columns; this might more clearly highlight trends
z-score scaling will just put all the data on the same scale
how many SD is the gene from the mean per sample? Better in this case then using abs values due to the differences in range for the sample

```{r}
genes.oi <- sapply(gprots, function(x){
  sub.dt <- iupar.cor.dt[GProtein == x,]
  return(sub.dt[R >= quantile(R,0.98) | R <= quantile(R,0.02), unique(gene)])
}) %>% c() %>% unique()


subMat <-  exp.mat[rownames(exp.mat) %in% genes.oi,]

# create a different scoring metric
GProtein <- c('Gi/Go & Gq/G11', 'Gq/G11', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gq/G11', 'Gs', 'Gi/Go & Gq/G11', 'Gq/G11 & Gs')

ha <- HeatmapAnnotation(GProtein = GProtein,
                        col=list(GProtein=c('Gi/Go & Gq/G11'='pink', 'Gq/G11'='lightblue',
                                            'Gi/Go'='lightgreen', 'Gs'='#CBC3E3','Gq/G11 & Gs'='#FFD580')) )
                        
hm <- Heatmap(subMat,
              top_annotation = ha,
        cluster_columns = T, 
        clustering_distance_columns = 'pearson',
        clustering_distance_rows = 'pearson',
        show_column_names = T,
        col = circlize::colorRamp2(breaks=c(-2,0,2), colors=c('blue', 'white', 'red')),
        name='log2FC',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 4),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))
        
hm <- draw(hm, column_title='2nd & 98th Percentile R scores')
BackupAsPDF(hm, ('2.and.98th.percentiles.log2FC.colscaleadjusted.heatmap'), dimensions=c(8,8))

```
Scale the columns to allow comparisons of trends
First scale by all 'interesting' things used as input for correlation

```{r}
scaledMat <- scale(exp.mat, center=T, scale=T)

subMat <-  scaledMat[rownames(scaledMat) %in% genes.oi,]

# create a different scoring metric
GProtein <- c('Gi/Go & Gq/G11', 'Gq/G11', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gq/G11', 'Gs', 'Gi/Go & Gq/G11', 'Gq/G11 & Gs')

ha <- HeatmapAnnotation(GProtein = GProtein,
                        col=list(GProtein=c('Gi/Go & Gq/G11'='pink', 'Gq/G11'='lightblue',
                                            'Gi/Go'='lightgreen', 'Gs'='#CBC3E3','Gq/G11 & Gs'='#FFD580')) )
                        
hm <- Heatmap(subMat,
              top_annotation = ha,
        cluster_columns = T, 
        clustering_distance_columns = 'pearson',
        clustering_distance_rows = 'pearson',
        show_column_names = T,
        col = circlize::colorRamp2(breaks=c(-3,0,3), colors=c('blue', 'white', 'red')),
        name='log2FC',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 4),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))
        
hm <- draw(hm, column_title='2nd & 98th Percentile R scores')
```
Maybe makes more sense to scale by all the data?

```{r}
# scale by all values
all.mat <- dcast(mss.out,gene~receptor, value.var = 'log2FC') %>% 
  .[, MOR.LG94 := NULL] %>% 
  as.matrix(rownames ='gene')

# renanme MOR DAMGO to MOR
colnames(all.mat) <- c('DOR', 'KOR', 'MOR', 'MRX2', 'MRX4', 'NK1R', 'NPFFR1', 'NPY1R','PTGER1', 'PTGER4','QRFPR')

# reorder the cols to match the matrices
col.order <-  c("MRX2","MRX4","NPFFR1","NPY1R","DOR","KOR","MOR","PTGER1","PTGER4","QRFPR","NK1R")
all.mat <- all.mat[, col.order]
```


```{r}
# scale by the entire dt
scaledMat <- scale(all.mat, center=T, scale=T)
scaledMat %>%  summary()
subMat <-  scaledMat[rownames(scaledMat) %in% genes.oi,]

# create a different scoring metric
GProtein <- c('Gi/Go & Gq/G11', 'Gq/G11', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gq/G11', 'Gs', 'Gi/Go & Gq/G11', 'Gq/G11 & Gs')

ha <- HeatmapAnnotation(GProtein = GProtein,
                        col=list(GProtein=c('Gi/Go & Gq/G11'='pink', 'Gq/G11'='lightblue',
                                            'Gi/Go'='lightgreen', 'Gs'='#CBC3E3','Gq/G11 & Gs'='#FFD580')) )
                        
hm <- Heatmap(subMat,
              top_annotation = ha,
        cluster_columns = T, 
        clustering_distance_columns = 'pearson',
        clustering_distance_rows = 'pearson',
        show_column_names = T,
       # col = circlize::colorRamp2(breaks=c(-2,0,2), colors=c('blue', 'white', 'red')),
        name='log2FC',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 4),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))
        
hm <- draw(hm, column_title='2nd & 98th Percentile R scores')


```



```{r}
subMat <-  exp.mat[rownames(exp.mat) %in% genes.oi,]


subMat %>%  colnames()

scaled.subMat <- scale(subMat)

scaled.subMat %>%  summary()

# create a different scoring metric
GProtein <- c('Gi/Go & Gq/G11', 'Gq/G11', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gi/Go', 'Gq/G11', 'Gs', 'Gi/Go & Gq/G11', 'Gq/G11 & Gs')

ha <- HeatmapAnnotation(GProtein = GProtein,
                        col=list(GProtein=c('Gi/Go & Gq/G11'='pink', 'Gq/G11'='lightblue',
                                            'Gi/Go'='lightgreen', 'Gs'='#CBC3E3','Gq/G11 & Gs'='#FFD580')) )
                        
hm <- Heatmap(scaled.subMat,
              top_annotation = ha,
        cluster_columns = T, 
        clustering_distance_columns = 'pearson',
        clustering_distance_rows = 'pearson',
        show_column_names = T,
        name='log2FC (z-score)',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 4),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))
        
hm <- draw(hm, column_title='2nd & 98th Percentile R scores')

```

```{r}
hm <- Heatmap(subMat,
              top_annotation = ha,
        cluster_columns = T, 
        clustering_distance_columns = 'pearson',
        clustering_distance_rows = 'pearson',
        show_column_names = T,
        name='log2FC',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 4),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))
        
hm <- draw(hm, column_title='2nd & 98th Percentile R scores')
```


Could we look at something like GSEA for enrichment and incorporating the Log2FC and/or R scores in the plot??

```{r}
colnames(subMat)
t(iupar)

iupar
```


cor score plot of the quantiles
```{r}
subMat <-  cor.mat[rownames(cor.mat) %in% genes.oi,]

hm <- Heatmap(subMat,
        cluster_columns = F, 
        show_column_names = F,
        name='Pearson r',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 3),
         column_names_gp = gpar (fontsize= 8),
         column_split = colnames(subMat),
         column_title_gp = gpar (fontsize= 9))

hm <- draw(hm, column_title='5% and 95% quantiles')

BackupAsPDF(hm, 'R.05.95.quantiles.gprotein.heatmap', dimensions = c(6,11))
```
```{r}
# extract the sig genes and reorder columns to match the scoring mat
subMat <-  lfc.mat[rownames(lfc.mat) %in% genes.oi, colnames(scores.mat$iupar)]

hm <- Heatmap(subMat,
              top_annotation = ha,
        cluster_columns = T, 
        show_column_names = T,
        name='log2FC',
        na_col = 'darkgrey',
        row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 3),
         column_names_gp = gpar (fontsize= 8),
         column_title_gp = gpar (fontsize= 9))
hm <- draw(hm, column_title='5% and 95% quantiles')

BackupAsPDF(hm, 'R.05.95.quantiles.log2FC.heatmap', dimensions = c(7,12.5))
```



