---
title: "WGCNA.paperFig5.correlationPlots"
author: "Martin Gordon"
date: "2023-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Generating plots of the correlation scores from figure 5 in the DARPA proteomics paper

```{r}
library(data.table)
library(ggplot2)
library(ComplexHeatmap)
library(WGCNA)
library(magrittr)
library(preprocessCore) # quantile normalisation
library(stringr)
library(matrixStats)
library(simplifyEnrichment)
library(fmsb)
library(readxl)
library(ggbeeswarm)
library(ggrepel)
library(curl)

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")


# helper scripts for module/network export etc
source ("/Users/martingordon/Documents/utils/mg_utils/HelperFunctions.R")
```

plots we wish to generate
- rank plot of the correlation scores
- heatmaps of the top ranking genes [maybe take top 10-20 per group] and plot log2FC and correlation values

lets first look at the data we have; are these cor scores correctly calculated?

06/09/23
-----
Correlations with the IUPHAR classification

lets first check the raw data 

```{r}

pred <- fread("./data/predictionScores.txt")
exp <- fread ("./data/experimentalScores.txt")
iupar <- data.table(read_xlsx("./data/IUPHAR_11GPCRs_Gprotein_class_primary.xlsx"))


exp.long <- melt(exp, id.vars = "GPCR", measure.vars = grep ("GNA", colnames(pred)), variable.name = "GProtein", value.name = "score")[, score := as.numeric(score)][]
pred.long <- melt(pred, id.vars = "GPCR", measure.vars = grep ("GNA", colnames(pred)), variable.name = "GProtein", value.name = "score")[, score := as.numeric(score)][]

# convert to long format
iupar.long <- melt(iupar, id.vars = 'GPCR', measure.vars = grep('GPCR', colnames(iupar), invert=T),  variable.name = "GProtein", value.name = "score") %>% 
  .[, score := as.numeric(score)] 
```


```{r scores.list}
scores.list <- list(exp = exp.long,
                    pred = pred.long,
                    iupar = iupar.long)
```
Convert the scores to a matrix

```{r}
scores.mat <- lapply(scores.list, function(x){
  dcast(x, GProtein~GPCR, value.var = 'score') %>% 
    as.matrix(rownames = 'GProtein')
})

lapply(scores.mat, colnames)
```
apply the scores mat to the data
```{r}
scores.mat <- lapply(scores.mat, function(x){
  colnames(x) <- c("MRX2","MRX4","NPFFR1","NPY1R","DOR","KOR","MOR","PTGER1","PTGER4","QRFPR","NK1R")
  return(x)
})
```



reorder the mss columns to follow the order above 

```{r}
col.order <-  c("MRX2","MRX4","NPFFR1","NPY1R","DOR","KOR","MOR","PTGER1","PTGER4","QRFPR","NK1R")

mss.out <- fread('./data/2022_08_31_NiceFitsPower3.csv')

exp.mat <- dcast(mss.out,gene~receptor, value.var = 'log2FC') %>% 
  .[, MOR.LG94 := NULL] %>% 
  as.matrix(rownames ='gene')


# filter to sig genes and reorder
sig.genes <- mss.out[pvalue < 0.05 & abs(log2FC) > log2(1.5) & receptor != 'MOR.LG94', unique(gene) ]
exp.mat <- exp.mat[rownames(exp.mat) %in% sig.genes,]


# renanme MOR DAMGO to MOR
colnames(exp.mat) <- c('DOR', 'KOR', 'MOR', 'MRX2', 'MRX4', 'NK1R', 'NPFFR1', 'NPY1R','PTGER1', 'PTGER4','QRFPR')

# reorder the cols to match the matrices
col.order <-  c("MRX2","MRX4","NPFFR1","NPY1R","DOR","KOR","MOR","PTGER1","PTGER4","QRFPR","NK1R")

exp.mat[1:10,1:10]
exp.mat <- exp.mat[, col.order]
exp.mat[1:10,1:10]


# sanit check; looks good
lapply(scores.mat, colnames)
```

run the correlations for each of the datasets


```{r}
cor.list <- lapply(scores.mat, function(x){
  cor(t(exp.mat), t(x), use= "pairwise.complete.obs")
})

cor.long <- lapply(cor.list, function(x){
  x <- melt(as.data.table(x, keep.rownames = TRUE), id.vars = "rn", variable.name  = "GProtein", value.name = "R")
  setnames(x, old = "rn", new = "gene")
  return(x)
})
```

count number of observations, only want to take results with vals observed for all receptors to reduce FPs

```{r}
cor.long$iupar[, .SD[which.max(R)], by=GProtein]

# most of these have values for the 11 receptors
obs.counts <- mss.out[!is.na(log2FC) & receptor != 'MOR.LG94' & gene %in% sig.genes, .N, by = gene]
hist(obs.counts$N, breaks=seq(1,11))

```

add this info to the correlation res

```{r}
lapply(cor.long, function(x){
  
  x[obs.counts, numObservations := N, on = 'gene']
  #setorder(x, -numObservations, -R, na.last = TRUE) #strongest associations at top
})

```
```{r}

lapply(cor.long, function(x){
  
  x[, direction := ifelse(R > 0, 'pos', 'neg')]
  x[!is.na(R) & numObservations == 11, percentile.rank := rank(R)/length(R), by=GProtein] # only interested in those found in all
    
})
```
sanity check of the top hits vs the 11 receptor app
```{r}
cor.long$iupar[numObservations == 11 & R > 0.8,]

cor.long$pred
scores.mat$iupar


# sanity check; is ZFYVE21 top Gi/Go correlation?
# yes these seemto be the top scorers
cor(scores.mat$iupar[1,], exp.mat[rownames(exp.mat) %in% 'ZFYVE21',])
cor(scores.mat$iupar[2,], exp.mat[rownames(exp.mat) %in% 'PRKCA',])
```
write out these results for the scores to the WGCNA directory

```{r}
#lapply(seq_along(cor.long), function(x,n,i){
#  fwrite(x[[i]][numObservations == 11], paste0('output/', n[[i]], '.proteinClassificationCorrelations.csv'))
#},x=cor.long, n=names(cor.long))
```

read in this data and plot the correlation ranks, then a heatmap of the top scores per gene




```{r}
cor.long$iupar$GProtein

lapply(cor.long$iupar$GProtein, function(x){
  
  ggplot(cor.long$iupar[ GProtein == x,], aes(x=reorder(gene,-R), y=R)) + geom_point()
  
})
```

