---
title: "WGCNA.clusters.regen"
author: "Martin Gordon"
date: "2023-10-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Regenerating WGCNA clusters with the detrended data (1st) and the whole dataset with NA values removed

```{r}

library(data.table)
library(ggplot2)
library(ComplexHeatmap)
library(WGCNA)
library(magrittr)
library(preprocessCore) # quantile normalisation
library(stringr)
library(matrixStats)
library(simplifyEnrichment)
library(fmsb)
library(readxl)
library(ggbeeswarm)
library(ggrepel)

library(limma)

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")

```


## WGCNA GPCR proteins

```{r}
# read in exp data
p.quant <- fread('data/2022_08_17_ProteinLevelData.lib.csv.gz')

# read in DEA results for filtering
mss <- fread('data/2022_08_31_NiceFitsPower3.csv')

p.quant$GROUP %>%  unique()
```

```{r}
str(p.quant)
```

```{r}

# fix the groups with an underscore in them MOR_LG and MOR_DAMGO to MOR.DAMGO
p.quant[, GROUP := gsub ("_([A-Z])", ".\\1", GROUP)]
# add some new identifier cols
p.quant[, `:=`(timeStr = str_extract(GROUP,'[0-9]{2}$'),
               receptor = gsub("_[0-9]{2}$", '', p.quant$GROUP)
               )]
```

clean up names

```{r}
group.noi <- p.quant[!GROUP %like% c("DOR|KOR|MOR|MRX2|MRX4|NK1R|NPFFR1|NPY1R|PTGER1|PTGER4|QRFPR|DOR"),unique(GROUP)]

p.quant <- p.quant[!GROUP %in% group.noi, ]
p.quant <- p.quant[!GROUP %like% "MOR.LG94",]

p.quant$GROUP %>%  unique()

p.quant[GROUP %like% "MOR", GROUP := gsub(".DAMGO", "", GROUP)]
p.quant$GROUP %>%  unique()

p.quant[, receptor := gsub('_[0-9]{2}', '', GROUP)]
```



```{r}
g <- ggplot(p.quant, aes(x=interaction(GROUP,SUBJECT), y=LogIntensities, fill=receptor)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=90))
g


p.quant
```
# run a lm to check for covariates impacting expression
```{r}
lm.out <- lm(LogIntensities~SUBJECT+receptor+GROUP, data=p.quant)
anova(lm.out)
```
Looking at heatmap there is a clear batch effect.. will need to correct this, but perhaps due to insufficient normalisation?
```{r}
p.mat <- dcast(p.quant, Protein~GROUP+SUBJECT, value.var = 'LogIntensities') %>%  as.matrix(rownames="Protein")


#subMat <- p.mat[complete.cases(p.mat),]
subMat <- p.mat[sample.int(p.mat, 2000),]
sample.int()

clusterwNa <- function(mat, na.val=0, ...){
  
  mat[is.na(mat)] <- na.val
  return(hclust(dist(mat)))
}


batch <- str_extract(colnames(p.mat), 'batch.*')
receptor <- gsub('_batch.[0-9]{2}', '', colnames(p.mat))
timeStr <- str_extract(colnames(p.mat), '00|30|05|01|10')
  
HA <- HeatmapAnnotation(batch = batch, receptor = receptor, time = timeStr)

ComplexHeatmap::Heatmap(p.mat, 
                        cluster_rows = clusterwNa(p.mat),
                        cluster_columns = T,
                        show_row_names = F,
                        top_annotation = HA)


```
Additional round of normalisation first; run TMP 

```{r}

#perform tmp normalisation
tmp <- medpolish(p.mat, na.rm = TRUE) #iteratively subtract row and col median values 
tmp.dt <- data.table(offset = tmp$col, group_subject = names(tmp$col))


p.quant[, group_subject := paste0(GROUP,'_', SUBJECT)]
p.quant[tmp.dt, normalize.offset := i.offset, on = "group_subject"]
p.quant[, adj.LogIntensities := LogIntensities - normalize.offset]

colnames(p.mat)

```
Look at a boxplot of these intensities 
These look much better; write out the adjusted values for now

```{r}
ggplot(p.quant, aes(x=interaction(GROUP,SUBJECT), y=adj.LogIntensities, fill=receptor)) +
  geom_boxplot()


p.quant[, LogIntensities := adj.LogIntensities ]
p.quant[, adj.LogIntensities := NULL]

#fwrite(p.quant, 'data/2022_08_17_ProteinLevelData.lib.tmpNormalized.csv.gz')

p.quant <- fread('data/2022_08_17_ProteinLevelData.lib.tmpNormalized.csv.gz')
```


Now that the data looks better normalised, rerun clustering and heatmap to observe the correlations

```{r}
p.mat <- dcast(p.quant, Protein~GROUP+SUBJECT, value.var = 'LogIntensities') %>%  as.matrix(rownames="Protein")

#subMat<- p.mat[complete.cases(p.mat),]
#subMat <- p.mat[sample.int(nrow(p.mat), 2000),]

batch <- str_extract(colnames(p.mat), 'batch.*')
receptor <- gsub('_[0-9]{2}_batch.[0-9]{2}', '', colnames(p.mat))
timeStr <- str_extract(colnames(p.mat), '00|30|05|01|10')
  
HA <- HeatmapAnnotation(batch = batch, receptor = receptor, time = timeStr)

hm <- ComplexHeatmap::Heatmap(p.mat, 
                        cluster_rows = clusterwNa(p.mat),
                        cluster_columns = T,
                        show_row_names = F,
                        column_names_gp = gpar(fontsize=4),
                        top_annotation = HA)

hm
BackupAsPDF(draw(ha, column_title='all features tmp norm'), '11recept.tmpnorm.heatmap')

```




Look at PCA to assess batch effects; seems samples primarily seperate by receptor. May not need to adjust

```{r}
subMat <- p.mat[complete.cases(p.mat),]

# transpose as want to use genes 
pcaOut <- prcomp(t(subMat))

colInfo <- data.table(colname = colnames(subMat))
colInfo[,c("receptor", "time", "batch") := tstrsplit(colname, "[_]", keep = c(1,2,3)) ] # v useful fucntion

title <- "PCA"

colInfo
#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)


#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color=batch, shape=time)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=receptor), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut)))
p
BackupAsPDF(p, "PCA.complete.proteins.batch")


#plot first two components
# looks like PC5 might respond to batch. enough to worry about removign?
p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color=receptor, shape=time)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=receptor), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut)))
p
BackupAsPDF(p, "PCA.complete.proteins.receptor")
```
Ok so it seems receptor is the biggest driver of  maybe its a lot of the missin values that seem to be driving batch, by removing these we are tec

```{r}
subMat<- p.mat[complete.cases(p.mat),]
#subMat <- p.mat[sample.int(nrow(p.mat), 2000),]

batch <- str_extract(colnames(subMat), 'batch.*')
receptor <- gsub('_[0-9]{2}_batch.[0-9]{2}', '', colnames(subMat))
timeStr <- str_extract(colnames(subMat), '00|30|05|01|10')
  
HA <- HeatmapAnnotation(batch = batch, receptor = receptor, time = timeStr)

hm <- ComplexHeatmap::Heatmap(subMat, 
                        cluster_rows = clusterwNa(subMat),
                        cluster_columns = T,
                        show_row_names = F,
                        column_names_gp = gpar(fontsize=4),
                        top_annotation = HA)

hm
BackupAsPDF(draw(ha, column_title='complete features tmp norm'), '11recept.tmpnorm.noNAs.heatmap')
```

For testing, lets look at batch correction for the data 
Using limma:removebatcheffects to subtract batch cov from the matrix 
batch 6,9,11 not estimated; are all treatments contained within these batches? seems like they are confounded...

```{r}
# use all measured prots to best estimate batch coef
batch <- str_extract(colnames(p.mat), '[0-9]{2}$') %>%  as.factor()

batch
receptor <- gsub('_[0-9]{2}_batch.[0-9]{2}', '', colnames(p.mat)) 
timeStr <- str_extract(colnames(p.mat), '00|30|05|01|10')
condition <- gsub('_batch.[0-9]{2}', '', colnames(p.mat))

# our experimental design mat
design <- model.matrix(~receptor+timeStr)

test.mat <- limma::removeBatchEffect(p.mat, batch=batch,  design=design)
```

plot the heatmap

```{r}
subMat <- test.mat[complete.cases(test.mat),]

# transpose as want to use genes 
pcaOut <- prcomp(t(subMat))

colInfo <- data.table(colname = colnames(subMat))
colInfo[,c("receptor", "time", "batch") := tstrsplit(colname, "[_]", keep = c(1,2,3)) ] # v useful fucntion

title <- "PCA"

colInfo
#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)


#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color=batch, shape=time)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=receptor), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut)))
p
BackupAsPDF(p, "PCA.complete.proteins.batch.batchcorrected")


#plot first two components
# looks like PC5 might respond to batch. enough to worry about removign?
p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color=receptor, shape=time)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=receptor), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut)))
p
BackupAsPDF(p, "PCA.complete.proteins.receptor.batchcorrected")
```


plot the batch corrected results
Looks like better seperation of samples.. check out the heatmap

```{r}
receptor.pal <- randomcoloR::distinctColorPalette(length(unique(receptor)))

batch <- str_extract(colnames(subMat), 'batch.*')
receptor <- gsub('_[0-9]{2}_batch.[0-9]{2}', '', colnames(subMat))
timeStr <- str_extract(colnames(subMat), '00|30|05|01|10')
  

HA <- HeatmapAnnotation(batch = batch, receptor = receptor, time = timeStr, 
                        col=list(receptor = c("DOR"=receptor.pal[1], 'KOR'=receptor.pal[2], 'MOR'=receptor.pal[3],
                                                       'MRX2'=receptor.pal[4], 'MRX4'=receptor.pal[5], 'NK1R'=receptor.pal[6], 
                                                       'NPFFR1'=receptor.pal[7], 'NPY1R'=receptor.pal[8],'PTGER1'=receptor.pal[9], 
                                                       'PTGER4'=receptor.pal[10], 'QRFPR'=receptor.pal[11])))

hm <- ComplexHeatmap::Heatmap(subMat, 
                        cluster_rows = clusterwNa(subMat),
                        cluster_columns = T,
                        show_row_names = F,
                        column_names_gp = gpar(fontsize=4),
                        top_annotation = HA)

BackupAsPDF(draw(hm, column_title='batch corrected complete features'), '11recept.tmpnorm.noNAs.batchcorrected.heatmap')
```
Try entire matrix including missing vals

```{r}
receptor.pal <- randomcoloR::distinctColorPalette(length(unique(receptor)))
#batch.pal <- randomcoloR::distinctColorPalette(length(unique(batch)))

batch <- str_extract(colnames(test.mat), 'batch.*')
receptor <- gsub('_[0-9]{2}_batch.[0-9]{2}', '', colnames(test.mat))
timeStr <- str_extract(colnames(test.mat), '00|30|05|01|10')
  

HA <- HeatmapAnnotation(batch = batch, receptor = receptor, time = timeStr, 
                        col=list(receptor = c("DOR"=receptor.pal[1], 'KOR'=receptor.pal[2], 'MOR'=receptor.pal[3],
                                                       'MRX2'=receptor.pal[4], 'MRX4'=receptor.pal[5], 'NK1R'=receptor.pal[6], 
                                                       'NPFFR1'=receptor.pal[7], 'NPY1R'=receptor.pal[8],'PTGER1'=receptor.pal[9], 
                                                       'PTGER4'=receptor.pal[10], 'QRFPR'=receptor.pal[11])))

hm <- ComplexHeatmap::Heatmap(test.mat, 
                        cluster_rows = clusterwNa(test.mat),
                        cluster_columns = T,
                        show_row_names = F,
                        column_names_gp = gpar(fontsize=4),
                        top_annotation = HA)

hm

BackupAsPDF(draw(hm,column_title='batch corrected all features'), '11recept.tmpnorm.batchcorrected.heatmap')
```
Using the batch-corrected data moving forward to construct the expression matrix; to avoid garbage in our modules, I only want to use proteins with at least 80% completeness across all
- what about biologically relevant missing values? maybe keep proteins missing in all sample/timepoints?

Feature selection
-----

First need to handle missing values... if we just use complete.cases, we lose 45% of our data
Keep prots with 90% completeness

```{r}
# 45% lost with 'full set'
nrow(test.mat)
nrow(test.mat[complete.cases(test.mat),])

# remove rows with more than 33% missing values
p.mat.f <- test.mat[which(rowMeans(!is.na(test.mat)) > 0.90), ] # only keep prots present in 2/3 of the dataset

dim(p.mat.f)

hm <- ComplexHeatmap::Heatmap(p.mat.f, 
                        cluster_rows = clusterwNa(p.mat.f),
                        cluster_columns = T,
                        show_row_names = F,
                        row_title = sprintf('%d proteins', nrow(p.mat.f)),
                        column_names_gp = gpar(fontsize=4),
                        top_annotation = HA)
hm
BackupAsPDF(draw(hm,column_title='wgcna input (batch corrected)'), '11recept.tmpnorm.batchcorrected.wgcna.input.heatmap')

```

WGCNA: 
---
Use the same input params as used previously, enrich output and se what is recovered


input genes look good
```{r}
p.mat <- p.mat.f

gsg <- goodSamplesGenes(t(p.mat))
summary(gsg)
gsg$allOK
```
r ^6 - ^7- ^8 looks good
```{r}
net.in <- t(p.mat)

s.thres <- data.table(pickSoftThreshold(net.in)$fitIndices)

s.thres %>%  
  ggplot(aes(x=Power, y=SFT.R.sq)) + 
  geom_point() +
  geom_hline(yintercept = 0.9, color='darkred') +
  labs(y="Scale Free Topology Model Fit, signed R^2") +
  theme_classic()

s.thres %>%  
  ggplot(aes(x=Power, y=mean.k.)) + 
  geom_point() +
  labs(y='Mean Connectivity') +
  theme_classic() 

s.thres
```
use a power of 7

```{r}
softPower = 7 ##R^2 of .9 and highest connectivity,
adj.mat = adjacency(net.in, power = softPower, corFnc = 'bicor', type = "signed") #Calculating the adjacency matrix
#help(adjacency )

dim(adj.mat)
```
Define the topological overlap matrix

```{r}
TOM <- TOMsimilarity(adj.mat)
```
Generate heatmaps of both modules before proceeding 

```{r}
Heatmap(adj.mat[1:2000,1:2000], cluster_rows = T, cluster_columns=T, column_title = 'adjacency matrix', row_title=sprintf('%s proteins', nrow(p.mat.f)), show_row_names = F, show_column_names = F)
#BackupAsPDF(hm, 'adj.mat.st7.')
Heatmap(TOM[1:2000,1:2000], cluster_rows = T, cluster_columns=T, column_title = 'TOM similiary', row_title=sprintf('%s proteins', nrow(p.mat.f)), show_row_names = F, show_column_names = F)
#BackupAsPDF(hm, 'tom.mat.st7.')
```
Generate the modules

```{r}
TOM.dissimilarity <- 1-TOM
adj.dissimilarity <- 1-adj.mat

#creating the dendrogram 
geneTree <- hclust(as.dist(TOM.dissimilarity), method = "average") 

Modules <- cutreeDynamic(dendro = geneTree, distM =TOM.dissimilarity, deepSplit = 2, pamRespectsDendro = FALSE, minClusterSize = 30)
table(Modules)

```
```{r}
modules.tom <- as.character(Modules)
```

```{r}
col.pal <- randomcoloR::distinctColorPalette(length(unique(Modules)))
col.pal
names(col.pal) <- seq(1,20,by=1) %>%  as.character()
# add wgcna module annotation
row_ha <- rowAnnotation(wgcna.modules=modules.tom, col=list(wgcna.modules = col.pal))

length(split)

seq(1,20,by=1) %>%  as.character()
# add split order
split <- factor(modules.tom, levels=seq(1,20,by=1) %>%  as.character())
split

hm <- Heatmap(p.mat.f, cluster_columns = FALSE, 
         #show_row_names = FALSE,
         split = split,
         right_annotation = row_ha,
         #name = "log2(intensity)",
         column_split = tstrsplit(colnames(p.mat), "[_]")[[1]],
         na_col = 'darkgrey',
         row_title =sprintf("%d Proteins",
                            nrow(prot.mat)),
         row_names_gp = gpar (fontsize= 1),
         column_names_gp = gpar (fontsize= 3),
         cluster_row_slices = F,
         column_title_gp = gpar (fontsize= 7))
hm
hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))

#BackupAsPDF(hm, 'heatmap.wgcna.modules.ordered')
```
Hard to see the trends here, so plot the vsT0 values (just take this list of proteins from p.quant and calculate) to assess
If looks good; check out the enrichment analysis; if oversplitting, adjst the parameters and rerun the analysis

```{r}

```





