---
title: "module1Analysis"
author: "Martin Gordon"
date: "2023-11-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Module 1 Analysis WGCNA

Seems from the inital WGCNA clustering that we see two subclusters in the WGCNA module 1
Lets try clustering submodule and see if we can define subgroups and enrich

Maybe then plot linecharts of the data

```{r}
library(data.table)
library(ggplot2)
library(ComplexHeatmap)
library(WGCNA)
library(magrittr)
library(stringr)
library(ggbeeswarm)
library(ggrepel)

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")

source("/Users/martingordon/Documents/utils/mg_utils/HelperFunctions.R")

clust.genes <- extractClustersfromHeatmap(hm, prot.mat) %>% rbindlist()
clust.genes[,.N, by=Cluster]

```

read in the data 
```{r}
# exp mat
receptor.exp.mat <- readRDS('./output/gpcr.11.receptors.exp.mat.Rdata')

# assignment of WGCNA modules
wgcna.clusters <- fread('./output/wgcna.clustergenes.adjmat.csv')

```

extract module 1 submat
```{r}
clust.oi <-  wgcna.clusters[Cluster == 'cluster1', GeneID]
rownames(receptor.exp.mat) <-  multiUniprots2multiGenes(rownames(receptor.exp.mat))

subMat <-  receptor.exp.mat[rownames(receptor.exp.mat) %in% clust.oi, ]
dim(subMat)
```

clustering of module 1
```{r}
clusterWNA <- function(x,na.val=0){
  
  x[is.na(x)] <- na.val
  return( hclust(dist(x)) )
  
}

hm <- Heatmap(subMat, cluster_columns = FALSE, 
         name = "log2(intensity/\nmean(timeZero))",
         column_split = tstrsplit(colnames(subMat), "[_.]")[[1]],
         na_col = 'darkgrey',
         row_title =sprintf("%d Proteins",
                            nrow(subMat)),
         row_names_gp = gpar (fontsize= 3),
         column_names_gp = gpar (fontsize= 3),
         cluster_row_slices = F,
         show_row_names = T,
         column_title_gp = gpar (fontsize= 9))

hm
```
kmeans clustering to identify subclusters in the data 


convert NA values to 0 (small offset)
```{r}
subMat.rmNA <- subMat
subMat.rmNA[is.na(subMat.rmNA)] <- 1e-10

km.out <- kmeans(subMat.rmNA, iter.max = 10, centers=2)

km.dt <-  data.table(gene=names(km.out$cluster), 
                     km.cluster=km.out$cluster)
```

replot the heatmap with the output of the kmm clustering algorithm

```{r}
hm <- Heatmap(subMat, cluster_columns = FALSE, 
         name = "log2(intensity/\nmean(timeZero))",
         column_split = tstrsplit(colnames(subMat), "[_.]")[[1]],
         na_col = 'darkgrey',
         row_split = paste0("k.means.", km.dt$km.cluster),
        # row_title =sprintf("%d Proteins",
        #                    nrow(subMat)),
         row_names_gp = gpar (fontsize= 5),
         column_names_gp = gpar (fontsize= 3),
         cluster_row_slices = F,
         show_row_names = T,
         column_title_gp = gpar (fontsize= 9))

hm <-  draw(hm)
BackupAsPDF(draw(hm, column_title='Module 1 (Endosome Trafficking)'), 'module1.kmeansclustering.heatmap.', dimensions=c(8,8))
```
Nicely captures the two subclusters; lets extract the cluster genes and run enrichment to identify GO BP processes
Background will be 

```{r}
subclust.genes <- extractClustersfromHeatmap(hm, subMat)

# sanity check
identical(km.dt[order(gene)]$km.cluster,subclust.genes[order(feature)]$cluster)

subclust.genes <-  subclust.genes[, cluster := paste0('k.means.',cluster)] %>% 
  .[, cluster := factor(cluster, levels=c('k.means.1', 'k.means.2'))]

gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")

# our background? only genes used for WGCNA analysis
universe <- rownames(receptor.exp.mat)

enrich.dt <- enricherOnGroups(subclust.genes, groupColumns = 'cluster', geneColumn = "feature", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'cluster')

enrich.dt %>%  View()
#custom ordering
enrich.dt[, cluster := factor(cluster, levels=c('k.means.1', 'k.means.2'))]

simp.enrich$simplified

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, groupColumn = 'cluster.x', topN = 12, title='Module 1 (Endosome Trafficking) subcluster enrichment', 
                                row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 6), upperThreshold = 8, cluster_columns =F)

BackupAsPDF(ht, 'go.enrichment.module1.subclusters', dimensions=c(8,8))
```
plot linecharts of the median expression profiles per receptor
label according to dominant terms

read in the p.quant data 
```{r}
# read in exp data
p.quant <- fread('data/2022_08_17_ProteinLevelData.lib.csv.gz')

p.quant[, gene := multiUniprots2multiGenes(Protein)]
# fix the groups with an underscore in them MOR_LG and MOR_DAMGO to MOR.DAMGO
p.quant[, GROUP := gsub ("_([A-Z])", ".\\1", GROUP)]
p.quant[GROUP %like% "DAMGO", GROUP := sub(".DAMGO","", GROUP)]

# add some new identifier cols
p.quant[, `:=`(timeStr = str_extract(GROUP,'[0-9]{2}$'),
               receptor = gsub("_[0-9]{2}$", '', p.quant$GROUP)
               )]

group.noi <- p.quant[!GROUP %like% c("DOR|KOR|MOR|MRX2|MRX4|NK1R|NPFFR1|NPY1R|PTGER1|PTGER4|QRFPR|DOR"),unique(GROUP)]

p.quant <- p.quant[!GROUP %in% group.noi, ]
p.quant <- p.quant[!GROUP %like% "MOR.LG94",]
#rm DAMGO flag

p.quant[, batchControlled := LogIntensities - median(LogIntensities, na.rm = TRUE), by = .(Protein, receptor, SUBJECT)]
p.quant[, vsTimeZero := batchControlled - mean(batchControlled[timeStr == timeStr[!is.na(batchControlled)][1] ]), by = .(Protein, receptor)]
```

calculate median by subcluster/receptor combo and plot; do these profiles distinguish the two groups

```{r}
p.quant.df <- p.quant[subclust.genes, cluster := i.cluster, on=c('gene','gene')] %>%
  .[, medianVsTimeZero := median(vsTimeZero, na.rm=T), by=.(gene, receptor, cluster)] %>% 
  .[!is.na(cluster),]

# define a useful color pallette
col.pal <- randomcoloR::distinctColorPalette(length(unique(p.quant.df$receptor)))


g <- ggplot(p.quant.df, aes(x=timeStr, y=vsTimeZero, group=cluster, color=cluster)) +
  stat_summary(fun = "median", geom = "line") +
  #stat_summary(fun.data = "mean_cl_normal", geom = "errorbar", width = .2) +
  scale_color_manual(values = col.pal) + 
  ggtitle('Module 1 (Endosome Trafficking) subcluster trends') +
  labs(x='timepoint (mins)', y="Median Intensity vs T0(log2)") +
  facet_wrap(~receptor, scales='free_y', ncol=3) +
  scale_x_discrete(expand=c(0,0)) +
  theme_bw() +
  theme(
   strip.text = element_text(size = 6))
g

BackupAsPDF(g, 'module1.subclusters.linechart')

```

looks really just to be a change in magnitude
```{r}

g <- ggplot(p.quant.df, aes(x=timeStr, y=vsTimeZero, group=receptor, color=receptor)) +
  stat_summary(fun = "median", geom = "line") +
  #stat_summary(fun.data = "mean_cl_normal", geom = "errorbar", width = .2) +
  scale_color_manual(values = col.pal) + 
  ggtitle('Module 1 (Endosome Trafficking) subcluster trends') +
  labs(x='timepoint (mins)', y="Median Intensity vs T0(log2)") +
  facet_wrap(~cluster, scales='free_y', ncol=3) +
  scale_x_discrete(expand=c(0,0)) +
  theme_bw() +
  theme(
   strip.text = element_text(size = 6))
g

BackupAsPDF(g, 'module1.subclusters.receptors.linechart.')
```
To do 
---
Maybe label the subclusters by GO enrichment results
Discuss with BP tomorrow
