---
title: "WGCNA_moduleDSDCytoscape"
author: "Martin Gordon"
date: "2023-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import our DSD modules into cytoscape to visualise
Start with the big modules that are 'interesting' first


# redo the networks, incorrect distance matrix used so retry with 11receptor.STRINGConnections.600.dsdDefaults.dsd

## Packages
```{r}
library(data.table)
library(stringr) #str_extract function 
library(magrittr)
library(ggplot2)
library (ComplexHeatmap)
library(RcppCNPy) #R/W support to import numpy objects into R
library(pbapply) #apply functions with progress bar
library(igraph)
library(RCy3) # cytoscape interface

# load bens util scripts for the analysis
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/LocatePeptidePtmInProteins.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

# netprop scripts
source ("../../utils/bp_utils/STRING_db_utils.R")
source ("../../utils/bp_utils/NetworkPropagation.R")
source("../../utils/bp_utils/PPINetworkModularized.R")
```

## Convert DSD matrix into dt and subset for each module

```{r}
dsd.f <- fread('./output/11receptor.STRINGConnections.600.dsdDefaults.dsd')
#convert to matrix
dsd.mat <- as.matrix(dsd.f, rownames='V1')
```

Filter out connectors not in our data
```{r}
#exp mat
adj.mat <- readRDS('./output/gpcr.11.receptors.adj.mat.Rdata')
selected.prots <- rownames(adj.mat)

# find andremove the one hop connectors form the data
stringMapping <- GetStringIDMapping(selected.prots, stringAliasFile='./data/9606.protein.aliases.v12.0.txt.gz')

#just use the alias of genes with no stringIDs
stringsOI <- c(setdiff(selected.prots, stringMapping$alias), stringMapping$string) 

# remove the 1 hop connectors
stringsInMat <- intersect(rownames(dsd.mat), c(stringsOI))
# sbset to just strings in the datatable
dsd.mat <- dsd.mat[stringsInMat, stringsInMat]


dsd.dt <- reshape2::melt(dsd.mat) %>% 
  setnames(c('STRING.source', 'STRING.target', 'dsd')) %>% 
  as.data.table()

dsd.dt[, interaction := 'interacts with']
```

```{r}
# file with cluster assignment for each gene/protein
cluster.assignment <- fread('./output/wgcna.clustergenes.adjmat.csv')

string.gene.map <- fread('./data/9606.protein.aliases.v12.0.txt.gz')
string.gene.map[,GeneID := alias]
colnames(string.gene.map) <- c('string', 'alias', 'source', 'GeneID')

cluster.assignment[string.gene.map, stringID := i.string, on=c('GeneID', 'GeneID') ]
cluster.assignment[, Cluster := sub('cluster', '', Cluster)]
```

Make our edge tables per module 

```{r}
clusters <- cluster.assignment[, unique(Cluster)]

clusters.edge <- lapply(clusters, function(x){
  #genes of interest
  clust.genes <- cluster.assignment[Cluster == x, stringID]
  # subset to the network of interest
  clust.dt <- dsd.dt[(STRING.source %in% clust.genes & STRING.target %in% clust.genes),]
  
  # remove self-loops
  clust.dt <- clust.dt[ STRING.source != STRING.target , .(STRING.source, interaction, STRING.target, dsd)]
  
  # remove duplicate edges
  
   # cal similarity (inverse transformation)
   clust.dt <- clust.dt[, ds.similarity := 1/(dsd)]
  
  # select edges of interest
  clust.dt <- clust.dt[, c("source", "target", "weight") := .(STRING.source, STRING.target, ds.similarity)] %>% 
    .[,.(source,target,interaction,weight,dsd)]
})


names(clusters.edge) <- clusters

# dont think this is a good metric, but use for now
lapply(clusters.edge, function(x){
  
  hist(x$weight)
  #quantile(x$weight, c(0.8,0.9,0.95))
  #plot(x$weight, scale(x$dsd))
})
```

Make node table per module;
Include some identifier info such 

```{r}
# add ME info to node table
MEs <- fread('./output/top10.K.percluster.csv')

clusters.node <- lapply(seq_along(clusters), function(x,n,i){
  
  node.tab  <- cluster.assignment[(stringID %in% clusters.edge[[i]]$source | stringID %in% clusters.edge[[i]]$target)]
  # node eigengene boolean
  mes <- MEs[i.cluster == i,rn]
  node.tab[GeneID %in% mes, ME := GeneID]
  
  # simplify
  node.tab[, id := stringID] %>% 
    .[,.(id,GeneID,Cluster,ME)]

},x=clusters.edge, n=names(clusters.edge))


names(clusters.node) <- clusters
```

Now import directly to Cytoscape
Issues loading this many edges at once? Should I threshold the network to reduce spurious links 

```{r}

collection.name <- "GPCR.receptors.DSD"

lapply(clusters, function(x){
  RCy3::createNetworkFromDataFrames(nodes = clusters.node[[x]], 
                                    edges = clusters.edge[[x]][,.(source,target,interaction,weight,dsd)],
                                    title = paste0('module',x,'dsd'), 
                                    collection = collection.name
                                    )
})

```
Network Ids for iterating

```{r}
# name of the collection
collection.id <-  getCollectionSuid()

# networks that are part of the collection
network.ids <- getCollectionNetworks(collection.suid = collection.id)

network.ids
```

Create a visual style for the network (think you first need to create a style before modifying)

```{r}
#deleteVisualStyle("gcpr.style")
# color palette for the 11 modules
col.pal <- c("#D57AC7","#8DDDE1","#A1E4AC","#6EE766","#CEDF5F","#DEC5D2","#85A693","#DBBB7C","#7D9AD6","#D96A6E","#A94FDD")  #"#999933" '#DDCC77' "grey"


style.name = "gcpr.style"


defaults <- list(NODE_SHAPE="ellipse",
                 LABEL_FONT_SIZE=30,
                 NODE_SIZE=50,
                 EDGE_TRANSPARENCY=60,
                 EDGE_OPACITY=100,
                 NODE_COLOR="#D3D3D3",
                 NODE_BORDER_WIDTH =2)

nodeLabels      <- mapVisualProperty('node label','GeneID','p')
nodeFills       <- mapVisualProperty('node fill color','Cluster','d', c("1","2", "3","4","5","6","7","8","9", "10","11"), col.pal)
edgeWidth       <- mapVisualProperty('edge width','weight','c', c(0,1), c(1,20))
nodeLabels      <- mapVisualProperty('node label','GeneID','p')

createVisualStyle(style.name, defaults, list(nodeLabels,nodeFills,edgeWidth))
#setVisualStyle(style.name) test

```

Remove duplicate edges from the network doesnt consider undirected edges so need to do manually...

```{r}
lapply(network.ids, function(x){

  deleteDuplicateEdges(network = list(network.ids))

})

```

Apply the visual style to each network 

```{r}
lapply(network.ids, print)

lapply(network.ids, function(x){
  
  setVisualStyle(style.name, network = x)
})
```

## Clustering 

MCL clustering; identify clusters of highly connected nodes in the network using information flow/current through the network
two stages: expansion and contraction of information performed over a number of iterations. Goal to conserve high information edges and trim less informative ones
I parameter for contraction (power matrix product is raised to before thresholding/pruning less informative edges)
I parameter 2-2.5 recommended. Used 2.5 for denser networks

Algorithm is controlled by the inflation

For filtering, check out distributions of DSD
For MCL clustering, exclude edges between nodes with 1/dsd < 0.12 (retaining approx top 10% of edges in each cluster)

```{r}
lapply(clusters.edge, function(x){
  
  #ggplot(x, aes(x=weight)) + geom_histogram() + geom_vline(xintercept=15)
  print(quantile(x$weight))
  print(quantile(x$weight, c(0.75,0.8, 0.9)))
})
```


Enrichment on the different modules detected

For enrichment analysis, we need to adjust the background 
Create two string networks in the collection using i) all genes in clustering and ii) all genes detected in the dataset

```{r}
string.cmd <- paste('string protein query query="', paste(c('ARRB1', 'TP53'), collapse = '\n'), '" cutoff=0.4 limit=0 species="Homo sapiens"',  sep = "")
string.cmd

commandsRun(string.cmd)
```

```{r}
# import the DSD dataframe
RCy3::createNetworkFromDataFrames(edges = dsd.dt[, c('source', 'target', 'interaction') := .(STRING.source, STRING.target, interaction)][, .(source, target, interaction)], 
                                  nodes = dsd.dt[, id := STRING.source][, .(id,interaction)],
                                  title = 'parent.network.wgcna.input', 
                                  collection = collection.name)

```
node size 

```{r}
nodeSize  <- mapVisualProperty('node size','gene','p')
```



```{r}
# now subset the matrix into 11 clusters
modules.mat <- lapply(cluster.vector, function(x){
  
  print(x)
  prots.oi <- cluster.assignment[!is.na(stringID) & Cluster == x, stringID]

  # subset the dsd matrix to each module
  clust.mat <- dsd.mat[rownames(dsd.mat) %in% prots.oi, colnames(dsd.mat) %in% prots.oi]
  return(clust.mat)
  
})



lapply(modules.mat, dim)
cluster.assignment[, .N, by=Cluster]
names(modules.mat) <- sub('cluster','',cluster.vector)
```

