---
title: "WGCNA_detrendedData"
author: "Martin Gordon"
date: "2023-10-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DARPA WGCNA detrended data

Goal is to highlight functional clusters of genes from the detrended data; ideally want to highlight the things interactiing with the receptors rather than just trafficked and detected by the proximity labelling proceedure

```{r}
library(data.table)
library(ggplot2)
library(ComplexHeatmap)
library(WGCNA)
library(magrittr)
library(preprocessCore) # quantile normalisation
library(stringr)
library(readxl)
library(randomcoloR)

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
source ("/Users/martingordon/Documents/utils/mg_utils/HelperFunctions.R")

```

Read in the detrended data and find all the sig sites
Want sites significant before and after detrending

```{r}
spatial.detrended <- read_xlsx('./docs/2023_02_03_ElevenReceptorResults_SpatialDetrendWillChange.xlsx', sheet = 2) %>% 
  as.data.table()
#spatial.detrended[, adj.pvalue := p.adjust(pvalue, method='BH')]

prots.oi <- spatial.detrended[sigBeforeDetrending != 'not' & pvalue < 0.05 & abs(log2FC) > 0.58, unique(Protein)]
```

Read in protein intensity data and process as initally performed

```{r}
# read in exp data
p.quant <- fread('data/2022_08_17_ProteinLevelData.lib.csv.gz')
```


```{r}
# fix the groups with an underscore in them MOR_LG and MOR_DAMGO to MOR.DAMGO
p.quant[, GROUP := gsub ("_([A-Z])", ".\\1", GROUP)]
# add some new identifier cols
p.quant[, `:=`(timeStr = str_extract(GROUP,'[0-9]{2}$'),
               receptor = gsub("_[0-9]{2}$", '', p.quant$GROUP)
               )]
```

Does this data need a second round of normalisation?
Assumption is the data is normalised
```{r}
group.noi <- p.quant[!GROUP %like% c("DOR|KOR|MOR|MRX2|MRX4|NK1R|NPFFR1|NPY1R|PTGER1|PTGER4|QRFPR|DOR"),unique(GROUP)]
group.noi

p.quant <- p.quant[!GROUP %in% group.noi, ]
p.quant <- p.quant[!GROUP %like% "MOR.LG94",]

p.quant$GROUP %>%  unique()
p.quant$GROUP %>%  unique()

# change MOR DAMGO to just MOR
p.quant[GROUP %like% "MOR.DAMGO", GROUP := sub("\\.DAMGO", "", GROUP)]
```

```{r}
g <- ggplot(p.quant, aes(x=interaction(GROUP,SUBJECT), y=LogIntensities, fill=receptor)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=90))
g
#BackupAsPDF(g, 'intensity.boxplots.')
```
generate matrix of sig detrended proteins and remove those missing in more than 33% of the data 
losing about a dozen proteins if filtereing on NA
```{r}
p.mat <- dcast(p.quant, Protein~GROUP+SUBJECT, value.var = 'LogIntensities') %>%  as.matrix(rownames="Protein")

# jsut select the detrended proteins
p.mat <- p.mat[rownames(p.mat) %in% prots.oi,]

p.mat %>%  dim()

# remove proteins with many missing values
p.mat <- p.mat[which(rowMeans(!is.na(p.mat)) > 0.66), ] # only keep prots present in 2/3 of the dataset


# if filtering here, what are the lost proteins
dropped.prots <- prots.oi[!prots.oi %in% rownames(p.mat)]
multiUniprots2multiGenes(dropped.prots)
```

```{r}
clusterwNa <- function(mat, na.val=0, ...){
  
  mat[is.na(mat)] <- na.val
  return(hclust(dist(mat)))
}

hm <- Heatmap(p.mat, cluster_rows = T, cluster_columns = F, na_col = 'darkgrey', column_split = tstrsplit(colnames(p.mat), '_', keep=1), row_title=sprintf('%s sig detrended proteins ', nrow(p.mat)), show_row_names = F, column_names_gp = gpar(fontsize=5), column_title_gp = gpar(fontsize = 8))

hm

BackupAsPDF(draw(hm,column_title=sprintf("%d proteins detrended WGCNA input ", nrow(p.mat)),
           column_title_gp=grid::gpar(fontsize=16)), 'wgcna.input.', dimensions = c(9,12)) 
```
how many proteins have all cases? drop to 508 if we want no missing values of 734 if not allowing missing values..

```{r}
Heatmap(p.mat.f[complete.cases(p.mat),], cluster_rows = T, cluster_columns = F, na_col = 'grey', column_split = tstrsplit(colnames(p.mat), '_', keep=1), row_title=sprintf('%s proteins', nrow(p.mat[complete.cases(p.mat),])))

sum(rownames(p.mat[complete.cases(p.mat),]) %in% prots.oi)
```

For WGCNA input Normalise vs T0 and correct for batch 

```{r}
setorder(p.quant, receptor, Protein, timeStr)

p.quant[, batchControlled := LogIntensities - median(LogIntensities, na.rm = TRUE), by = .(Protein, receptor, SUBJECT)]
p.quant[, vsTimeZero := batchControlled - mean(batchControlled[timeStr == timeStr[!is.na(batchControlled)][1] ]), by = .(Protein, receptor)]

p.mat <- dcast(p.quant, Protein~GROUP+SUBJECT, value.var = 'vsTimeZero') %>%  as.matrix(rownames="Protein")

p.mat <- p.mat[rownames(p.mat) %in% prots.oi & !rownames(p.mat) %in% dropped.prots,]
```

```{r}
hm <- Heatmap(p.mat, cluster_rows = T, cluster_columns = F, na_col = 'darkgrey', column_split = tstrsplit(colnames(p.mat), '_', keep=1), row_title=sprintf('%s proteins', nrow(p.mat)), name='vs T0', show_row_names = F, column_names_gp = gpar(fontsize=3), column_title_gp = gpar(fontsize = 8))
hm

BackupAsPDF(draw(hm,column_title="Detrended sig proteins vs T0 WGCNA input"), dimensions = c(9,12)) 
```
Lets cluster columns and see how the data breaks out
Seems that timepoint and receptor seem to be key drivers of expression

```{r}
batch <- str_extract(colnames(p.mat), 'batch.*')
receptor <- gsub('_[0-9]{2}_batch.[0-9]{2}', '', colnames(p.mat))
timeStr <- str_extract(colnames(p.mat), '00|30|05|01|10')
  
HA <- HeatmapAnnotation(batch = batch, receptor = receptor, time = timeStr)

hm <- ComplexHeatmap::Heatmap(p.mat, 
                        cluster_rows = clusterwNa(p.mat),
                        cluster_columns = T,
                        show_row_names = F,
                        column_names_gp = gpar(fontsize=4),
                        top_annotation = HA)

hm
BackupAsPDF(draw(hm, column_title='clustering of WGCNA input (vs T0, batch corrected)'), '11recept.tmpnorm.noNAs.heatmap', dimensions = c(12,9))
```
PCA plot to assess sample clustering. How does this look vs T0
508 Proteins used as input 

```{r}

log2(10)

subMat <- p.mat[complete.cases(p.mat),]
#colnames(subMat)[!colnames(subMat) %like% "_00_"]
# transpose as want to use genes 
pcaOut <- prcomp(t(subMat))

colInfo <- data.table(colname = colnames(subMat))
colInfo[,c("receptor", "time", "batch") := tstrsplit(colname, "[_]", keep = c(1,2,3)) ] # v useful fucntion

title <- "PCA"

colInfo
#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)


#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color=batch, shape=time)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=receptor), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut)))
p
#BackupAsPDF(p, "PCA.complete.proteins.batch")


#plot first two components
# looks like PC5 might respond to batch. enough to worry about removign?
p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color=receptor, shape=time)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=receptor), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut)))
p
#BackupAsPDF(p, "PCA.complete.proteins.receptor")

```
Remove the T0 points and plot the rest

```{r}
subMat <- subMat[, !colnames(subMat) %like% "_00_"]

pcaOut <- prcomp(t(subMat))

colInfo <- data.table(colname = colnames(subMat))
colInfo[,c("receptor", "time", "batch") := tstrsplit(colname, "[_]", keep = c(1,2,3)) ] # v useful fucntion

title <- "PCA"

colInfo
#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)


#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color=batch, shape=time)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=receptor), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut)))
p
#BackupAsPDF(p, "PCA.complete.proteins.vsT0.batch")


#plot first two components
# looks like PC5 might respond to batch. enough to worry about removign?
p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color=receptor, shape=time)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=receptor), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut)))
p
#BackupAsPDF(p, "PCA.complete.proteins.vsT0.receptor")
```
TODO replot with better shapes and colors


```{r}
receptor.pal <- randomcoloR::distinctColorPalette(length(unique(receptor)))


#plot first two components
# looks like PC5 might respond to batch. enough to worry about removign?
p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color=receptor, shape=time)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=receptor), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  scale_color_manual(values = receptor.pal) +
  scale_shape_manual(values = c(21,22,23,24)) +
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut)))
p
#BackupAsPDF(p, "PCA.complete.proteins.vsT0.receptor")


```

```{r}

#Group data in a dendogram to check outliers
sampleTree = hclust(dist(t(p.mat)), method = "average")


pdf(file = "./GPCR_WGCNA_data/sampleClustering.pdf", width = 12, height = 9); 
par(cex = 0.6)
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering", sub="", xlab="", cex.lab = 1.5, 
     cex.axis = 1.5, cex.main = 2)
#Plot a line showing the cut-off
abline(h = 56, col = "red") #This value of 31000 was chosen based on my data, you need to check the best value to your data

dev.off()
```

Now that we have looked at the samples, lets checkout the genes we have selected
```{r}
gsg <- goodSamplesGenes(t(p.mat))
summary(gsg)
gsg$allOK
```
```{r}
net.in <- t(p.mat)

s.thres <- data.table(pickSoftThreshold(net.in)$fitIndices)

s.thres %>%  
  ggplot(aes(x=Power, y=SFT.R.sq)) + 
  geom_point() +
  geom_hline(yintercept = 0.9, color='darkred') +
  labs(y="Scale Free Topology Model Fit, signed R^2") +
  theme_classic()

s.thres %>%  
  ggplot(aes(x=Power, y=mean.k.)) + 
  geom_point() +
  labs(y='Mean Connectivity') +
  theme_classic() 

s.thres
```
Raise to power of 6 as R^2 > .8 and seems to level off

```{r}
softPower = 6 ##R^2 of .8 and highest connectivity,
adj.mat = adjacency(net.in, power = softPower, corFnc = 'bicor', type = "signed") #Calculating the adjacency matrix
#help(adjacency )

dim(adj.mat)
```
Define the topological overlap matrix

```{r}
TOM <- TOMsimilarity(adj.mat)
```
Generate heatmaps of both modules before proceeding 

```{r}
Heatmap(adj.mat, cluster_rows = T, cluster_columns=T, column_title = 'adjacency matrix', row_title=sprintf('%s proteins', nrow(p.mat)), show_row_names = F, show_column_names = F)
#BackupAsPDF(hm, 'adj.mat.st7.')
Heatmap(TOM, cluster_rows = T, cluster_columns=T, column_title = 'TOM similiary', row_title=sprintf('%s proteins', nrow(p.mat)), show_row_names = F, show_column_names = F)
#BackupAsPDF(hm, 'tom.mat.st7.')
```

Generate the modules
5 Modules recovered 

```{r}
TOM.dissimilarity <- 1-TOM
adj.dissimilarity <- 1-adj.mat

#creating the dendrogram 
geneTree <- hclust(as.dist(TOM.dissimilarity), method = "average") 

Modules <- cutreeDynamic(dendro = geneTree, distM =TOM.dissimilarity, deepSplit = 4, pamRespectsDendro = FALSE, minClusterSize = 20)
table(Modules)

# convert the module genes to a character vector
modules.tom <- as.character(Modules)
modules.tom 
```

Module gene heatmap

```{r}
col.pal <- randomcoloR::distinctColorPalette(length(unique(modules.tom)))

# add wgcna module annotation
row_ha <- rowAnnotation(wgcna.modules=modules.tom, col=list(wgcna.modules = c("1"=col.pal[1], '2'=col.pal[2], '3'=col.pal[3], '4'=col.pal[4], '5'=col.pal[5])))

# add split order
split <- factor(modules.tom, levels=c('1','2','3','4','5'))
split

hm <- Heatmap(p.mat, cluster_columns = FALSE, 
         #show_row_names = FALSE,
         split = split,
         right_annotation = row_ha,
         name = "log2(intensity/\nmean(timeZero))",
         column_split = tstrsplit(colnames(p.mat), "[_.]")[[1]],
         na_col = 'darkgrey',
         row_title =sprintf("%d Proteins",
                            nrow(p.mat)),
         row_names_gp = gpar (fontsize= 1),
         column_names_gp = gpar (fontsize= 3),
         cluster_row_slices = F,
         column_title_gp = gpar (fontsize= 7))

hm <- draw(hm, column_title="WGCNA clusters",
           column_title_gp=grid::gpar(fontsize=10))
hm
BackupAsPDF(hm, 'heatmap.wgcna.modules.ordered', dimensions = c(9,12))
```
The five clusters do look to have distinct expression patterns (maybe not final cluster as a little more noisy...)

Write out the cluster genes to file
```{r}
table(modules.tom)

clust.genes <- extractClustersfromHeatmap(hm, p.mat)

fwrite(clust.genes, './WGCNA_detrendedData_data/wgcna.moduleGenes.csv')
```

Run enrichment on the clusters
Using all prots measured in exp as input
Modules not as well distinguished as previous data
 
```{r}
# helper functions including script to extract heatmap genes
source("/Users/martingordon/Documents/utils/mg_utils/HelperFunctions.R")

# change clusterID to factor and order
clust.genes[, cluster := factor(cluster, levels=c('1', '2', '3', '4', '5'))]
clust.genes[, Gene := multiUniprots2multiGenes(feature)]

# load the GO table
gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")

# our background? only genes used for WGCNA analysis
#universe <- multiUniprots2multiGenes(rownames(p.mat))
universe <- p.quant[, multiUniprots2multiGenes(unique(Protein))]


enrich.dt <- enricherOnGroups(clust.genes, groupColumns = 'cluster', geneColumn = "Gene", term2gene.gmt = gmt.go, universe = universe)
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'cluster')

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, groupColumn = 'cluster.x', topN = 8, title='GO enrichment WGCNA modules', 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 6), upperThreshold = 8, cluster_columns =F)


BackupAsPDF(ht, 'go.enrichment.wgcna.modules.top8', dimensions=c(8,12))
```
validate the enrichment results and that these proteins are within the 5 modules.
ooks quite similiar to earlier run, but maybe more focused? Worth repeating the whole cytoscape enrichment plots?

linecharts of median expression per receptor; also look at Gprotein correlations

```{r}

```












Can we split the clusters more? Seems pretty robust to this cluster size
Try rerun with larger power thresholding and allow smmaller module sizes

Using these thresholds we create another module, assess enrichments and plot

```{r}
softPower = 12 ##R^2 of .8 and highest connectivity,
adj.mat = adjacency(net.in, power = softPower, corFnc = 'bicor', type = "signed") #Calculating the adjacency matrix
#help(adjacency )

```
Define the topological overlap matrix

```{r}
TOM <- TOMsimilarity(adj.mat)
```
```{r}
Heatmap(adj.mat, cluster_rows = T, cluster_columns=T, column_title = 'adjacency matrix', row_title=sprintf('%s proteins', nrow(p.mat.f)), show_row_names = F, show_column_names = F)
#BackupAsPDF(hm, 'adj.mat.st7.')
Heatmap(TOM, cluster_rows = T, cluster_columns=T, column_title = 'TOM similiary', row_title=sprintf('%s proteins', nrow(p.mat.f)), show_row_names = F, show_column_names = F)
#BackupAsPDF(hm, 'tom.mat.st7.')
```
```{r}
TOM.dissimilarity <- 1-TOM
adj.dissimilarity <- 1-adj.mat

#creating the dendrogram 
geneTree <- hclust(as.dist(TOM.dissimilarity), method = "average") 

Modules <- cutreeDynamic(dendro = geneTree, distM =TOM.dissimilarity, deepSplit = 4, pamRespectsDendro = FALSE, minClusterSize = 15)
table(Modules)

# convert the module genes to a character vector
modules.tom <- as.character(Modules)
modules.tom %>%  table()
```

```{r}
col.pal <- randomcoloR::distinctColorPalette(length(unique(modules.tom)))

# add wgcna module annotation
row_ha <- rowAnnotation(wgcna.modules=modules.tom, col=list(wgcna.modules = c("0"='grey',"1"=col.pal[1], '2'=col.pal[2], '3'=col.pal[3], '4'=col.pal[4], '5'=col.pal[5], '6'=col.pal[6])))

# add split order
split <- factor(modules.tom, levels=c('0','1','2','3','4','5','6'))
split

hm <- Heatmap(p.mat, cluster_columns = FALSE, 
         #show_row_names = FALSE,
         split = split,
         right_annotation = row_ha,
         name = "log2(intensity/\nmean(timeZero))",
         column_split = tstrsplit(colnames(p.mat), "[_.]")[[1]],
         na_col = 'darkgrey',
         row_title =sprintf("%d Proteins",
                            nrow(p.mat)),
         row_names_gp = gpar (fontsize= 1),
         column_names_gp = gpar (fontsize= 3),
         cluster_row_slices = F,
         column_title_gp = gpar (fontsize= 7))

hm <- draw(hm, column_title="WGCNA clusters (adj.mat threshold = ^12)",
           column_title_gp=grid::gpar(fontsize=10))
hm
BackupAsPDF(hm, 'heatmap.wgcna.modules.adj.matThresh12.', dimensions = c(9,12))
```
Extract these cluster geens and plot 

```{r}
clust.genes <- extractClustersfromHeatmap(hm, p.mat)

#fwrite(clust.genes, './WGCNA_detrendedData_data/wgcna.moduleGenes.csv')
```
14-11-23
----
Not sure if this is the correct datasets, but prepare for later review
Want to compare the impact of detrending on the data 
For the sig genes we used as input for wgcna network construction, create a scatterplot of log2FC values of detrended vs standard

Taking all significant hits in the detrended and non-detrended data
Look for things that fall off the diagonal

```{r}
detrended.clust.genes <-  fread('./WGCNA_detrendedData_data/wgcna.moduleGenes.csv')
detrended.clust.genes[,.N,by=.(cluster)]

# spatially detrended data
spatial.detrended <- read_xlsx('./docs/2023_02_03_ElevenReceptorResults_SpatialDetrendWillChange.xlsx', sheet = 2) %>% 
  as.data.table()

detrended.prots.oi <- spatial.detrended[sigBeforeDetrending != 'not' & pvalue < 0.05 & abs(log2FC) > 0.58, unique(Protein)]
```

Read in the standard dataset ts results

```{r}
# read in non-detrended
mss <- fread('data/2022_08_31_NiceFitsPower3.csv')
```

what are the number of DEGs per receptor?

```{r}
mss[receptor != 'MOR.LG94' & gene %in% std.prots.oi & pvalue < 0.05 & abs(log2FC) > 0.58, .N, by=.(receptor)] %>% 
  ggplot(., aes(x=reorder(receptor,-N), y=N, fill = receptor)) + 
  geom_bar(stat='Identity') +
  scale_fill_manual(values=col.pal) +
  theme_classic()
  
```
original WGCNA clustering information; extract the sig genes that were used in the clustering, look at their profiles in the detrended data

```{r}
wgcna.clusters <- fread('./output/wgcna.clustergenes.adjmat.csv')
setnames(wgcna.clusters, new=c('gene', 'cluster'))
         
std.prots.oi <- mss[receptor != "MOR.LG94" & gene %in% wgcna.clusters$gene, unique(gene)]

length(detrended.prots.oi)
length(std.prots.oi)
```
per receptor, plot the log2FC of that data from the trended and detrended for the input proteins in WGCNA
- split the two datasets per receptor, iterate through them and add the detrended data to the default dt

```{r}
default.mss.list <- mss[receptor != "MOR.LG94"] %>%  
  .[receptor == 'MOR.DAMGO', receptor := sub(".DAMGO", "", receptor)] %>% 
  split(., .$receptor)

detrended.mss.list <-  spatial.detrended[receptor != "MOR.LG94"] %>%  
  .[receptor == 'MOR.DAMGO', receptor := sub(".DAMGO", "", receptor)] %>% 
  split(., .$receptor)

# sanity check
identical(names(detrended.mss.list),names(default.mss.list))

receptors.oi <-  names(default.mss.list)

# add identifier to the two datasets
comb.mss.list <- lapply(receptors.oi, function(x){
  
 comb.dt  <- default.mss.list[[x]][detrended.mss.list[[x]], `:=`(detrended.log2FC = i.log2FC, 
                                                      detrended.pvalue = i.pvalue), on=c('gene','gene')] 
 return(comb.dt)
  
})

names(comb.mss.list) <-  names(default.mss.list)

# sanity check.. looks good
#comb.mss.list$DOR[gene %in% c('SNX12', 'SNX3', 'EEA1'), .(gene,log2FC,pvalue, detrended.log2FC, detrended.pvalue)]
#default.mss.list$DOR[gene %in% c('SNX12', 'SNX3', 'EEA1'), .(gene,log2FC,pvalue)]
#etrended.mss.list$DOR[gene %in% c('SNX12', 'SNX3', 'EEA1'), .(gene,log2FC,pvalue)]
```

iterate through the receptors, plotting the log2FC for all sig results per receptor

lable things off the diagonal ; biggest movers
Label the 30 things with biggest change in input vs detrended data
These things above the diagonal line are the trafficking-associated proteins, not of interest and want to remove from the dataset

```{r}
lapply(seq_along(receptors.oi), function(x,y,i){
  
  recept <-  x[[i]]
  
  # extract the sig prots per receptor in the non-detrended, we want to see how these profiles change over time
  genes.oi <- y[[recept]][receptor == recept & pvalue < 0.05 & abs(log2FC) > 0.58, gene]
  
  y[[recept]][receptor == recept & gene %in% genes.oi, delta.log2FC := log2FC-detrended.log2FC]
  
  y[[recept]][,lab := '']
  y[[recept]][order(-abs(delta.log2FC))[1:30], lab := gene]
  
  g <- ggplot(data=y[[recept]][receptor == recept & gene %in% genes.oi,], aes(x=detrended.log2FC, y=log2FC, label=gene)) + 
    geom_point() +
    geom_density2d() +
    ggrepel::geom_text_repel(aes(label=lab)) +
    geom_abline(slope=1, intercept = 0, linetype=3, alpha=0.6) +
    ggtitle(paste0(recept, ' all significant results')) +
    theme_bw()
  
},x=receptors.oi, y=comb.mss.list)

```
plot of the significant detrended hits

```{r}
lapply(seq_along(receptors.oi), function(x,y,i){
  
  recept <-  x[[i]]
  
  # extract the sig prots per receptor in the non-detrended, we want to see how these profiles change over time
  genes.oi <-  spatial.detrended[receptor == recept & sigBeforeDetrending != 'not' & pvalue < 0.05 & abs(log2FC) > 0.58, unique(gene)]
  
  y[[recept]][,lab := '']
  y[[recept]][order(-abs(delta.log2FC))[1:20], lab := gene]
  
  ggplot(data=y[[recept]][receptor == recept & gene %in% genes.oi,], aes(x=detrended.log2FC, y=log2FC, label=gene)) + 
    geom_point() +
    geom_density2d() +
    ggrepel::geom_text_repel(aes(label=gene)) +
    geom_abline(slope=1, intercept = 0, linetype=3, alpha=0.6) +
    ggtitle(paste0(recept, ' significant detrended results')) +
    theme_bw()
  
},x=receptors.oi, y=comb.mss.list)
```
of these detrended sig results, how many are still related to trafficking?
lets pull in endosome-trafficking trafficking related genesets for testing

```{r}
library(msigdbr)

sets.oi <-  c('GOBP_EARLY_ENDOSOME_TO_GOLGI_TRANSPORT', 'GOBP_ENDOSOMAL_TRANSPORT', 'GOBP_ENDOSOME_TO_PLASMA_MEMBRANE_PROTEIN_TRANSPORT', 'GOBP_ENDOSOME_TO_LYSOSOME_TRANSPORT')

endosome.transport.genes <- msigdbr(category='C5', subcategory='GO:BP') %>% 
  data.table() %>% 
  .[gs_name %in% sets.oi, unique(human_gene_symbol)]
```

color by trafficking genes
Looking at this, seems distributions of trafficking overlap with other

```{r}

lapply(seq_along(receptors.oi), function(x,y,i){
  
  recept <-  x[[i]]
  
  # extract the sig prots per receptor in the non-detrended, we want to see how these profiles change over time
  genes.oi <-  spatial.detrended[receptor == recept & sigBeforeDetrending != 'not' & pvalue < 0.05 & abs(log2FC) > 0.58, unique(gene)]
  
  # aes for plot
  y[[recept]][,lab := '']
  y[[recept]][order(-abs(delta.log2FC))[1:20], lab := gene]
  y[[recept]][, endo.trafficking := ifelse( gene %in% endosome.transport.genes, 'yes', 'no')]
  
  g <- ggplot(data=y[[recept]][receptor == recept & gene %in% genes.oi,], aes(x=detrended.log2FC, y=log2FC, label=gene, color=endo.trafficking)) + 
    geom_point() +
    geom_density2d() +
    scale_color_manual(values=c('yes'='darkred', 'no'='black')) +
    ggrepel::geom_text_repel(aes(label=gene)) +
    geom_abline(slope=1, intercept = 0, linetype=3, alpha=0.6) +
    ggtitle(paste0(recept, ' significant detrended results')) +
    theme_bw()
  
  #BackupAsPDF(g, paste0(recept,'.log2FC.detrended.vs.ori.endotrafficking.scatterplots.'))
  
},x=receptors.oi, y=comb.mss.list)
```
lets color the points by the location specific APEX results
Find list of location specific protein changes & identify those with a large log2FC in each condition

164 proteins significant in this dataset; presuming this is the 
think(??) this is FC with location specific APEX relative to bg
```{r}
cytoRefs <- fread('https://raw.githubusercontent.com/kroganlab/bp_utils/master/data/2023_08_03_Dar25.28.29_LocationReference_CytoContrasts.csv')
cytoRefs[, gene := multiUniprots2multiGenes(Protein)]
cytoRefs
cols.oi <- c('Cyto - Endo.C19', 'Cyto - GalT', 'Cyto - LAMT', 'Cyto - PM') 

# is this correctly collapsing the remaing parts of the datatable 
cytoRefs.long <-  cytoRefs[, .(Protein, gene, melt(.SD)), .SDcols = cols.oi]


# lets find any location specific large (log FC > 1) movers; 
#location.prots <-  cytoRefs.long[abs(value) > 1, unique(gene)]
location.prots <-  cytoRefs[,unique(gene)]

# dont make threshold
cytoRefs[!gene %in% location.prots, unique(gene)]
```
lets color the plots by these location specific trends; are they moving significantly

```{r}
lapply(seq_along(receptors.oi), function(x,y,i){
  
  recept <-  x[[i]]
  
  # extract the sig prots per receptor in the non-detrended, we want to see how these profiles change over time
  genes.oi <-  spatial.detrended[receptor == recept & sigBeforeDetrending != 'not' & pvalue < 0.05 & abs(log2FC) > 0.58, unique(gene)]
  
  # aes for plot
  y[[recept]][,lab := '']
  y[[recept]][order(-abs(delta.log2FC))[1:20], lab := gene]
  y[[recept]][, location.specific := ifelse( gene %in% location.prots, 'yes', 'no')]
  
  g <- ggplot(data=y[[recept]][receptor == recept & gene %in% genes.oi,], aes(x=detrended.log2FC, y=log2FC, label=gene, color=location.specific)) + 
    geom_point() +
    #geom_density2d() +
    scale_color_manual(values=c('yes'='red', 'no'='darkgrey')) +
    ggrepel::geom_text_repel(aes(label=gene)) +
    geom_abline(slope=1, intercept = 0, linetype=3, alpha=0.6) +
    ggtitle(paste0(recept, ' significant detrended results')) +
    theme_bw()
  
    BackupAsPDF(g, paste0(recept,'.log2FC.detrended.vs.ori.locationSpecificHighlighted.scatterplots.'))
  
},x=receptors.oi, y=comb.mss.list)
```
look at histogram of residuals for the location specific and non-location specific

bland-altman plot: are there systematic biases in the distribution? plot mean log2FC along the x axis and 
the delta between log2FC on the y: is there systematic bias in the values (high or low more noisy?)

```{r}

```



what does the absolute difference distribution look like between the log2FC values per receptor?
For now plot all sig genes
abs view looks like to could also be direction; lets separate
```{r}
lapply(seq_along(receptors.oi), function(x,y,i){

  recept <-  x[[i]]
  
  # extract the sig prots per receptor in the non-detrended, we want to see how these profiles change over time
  genes.oi <- y[[recept]][receptor == recept & pvalue < 0.05 & abs(log2FC) > 0.58, gene]

  ggplot(data=y[[recept]][receptor == recept & gene %in% genes.oi,], aes(x=abs(delta.log2FC))) + 
    geom_histogram(bins=50) +
    ggtitle(paste0(recept, ' detrended log2FC - log2FC')) +
    theme_bw()
  
},x=receptors.oi, y=comb.mss.list)
```
this is the difference in fold change values
mixture of 3 models? + w detrended, - w detrended, unchanged (interesting!)
```{r}
lapply(seq_along(receptors.oi), function(x,y,i){

  recept <-  x[[i]]
  
  # extract the sig prots per receptor in the non-detrended, we want to see how these profiles change over time
  genes.oi <- y[[recept]][receptor == recept & pvalue < 0.05 & abs(log2FC) > 0.58, gene]

  
  ggplot(data=y[[recept]][receptor == recept & gene %in% genes.oi,], aes(x=delta.log2FC)) + 
    geom_histogram(bins=50) +
    ggtitle(paste0(recept, ' detrended log2FC - log2FC')) +
    theme_bw()
  
},x=receptors.oi, y=comb.mss.list)



lapply(seq_along(receptors.oi), function(x,y,i){

  recept <-  x[[i]]
  
  # extract the sig prots per receptor in the non-detrended, we want to see how these profiles change over time
  genes.oi <- y[[recept]][receptor == recept & pvalue < 0.05 & abs(log2FC) > 0.58, gene]

  
  ggplot(data=y[[recept]][receptor == recept & gene %in% genes.oi,], aes(x=delta.log2FC)) + 
    geom_histogram(bins=50) +
    ggtitle(paste0(recept, ' detrended log2FC - log2FC')) +
    theme_bw()
  
},x=receptors.oi, y=comb.mss.list)
```


```{r}
genesOI <-  spatial.detrended[sigBeforeDetrending != 'not' & pvalue < 0.05 & abs(log2FC) > 0.58, unique(gene)]
receptor.col.pal <-  distinctColorPalette(k=length(receptors.oi))

g <-  rbindlist(comb.mss.list, idcol = T) %>% 
  .[gene %in% genesOI,] %>% 
  ggplot(., aes(x=delta.log2FC, fill=.id)) +
  geom_histogram(bins=40) + 
  ggtitle('detrended.log2FC vs ori.log2FC') +
  facet_wrap(~.id, scales='free') +
  scale_fill_manual(values = c(receptor.col.pal) ) +
  theme_classic()

BackupAsPDF(g, 'deltaLog2FC.receptors.histogram.', dimensions=c(10,8))

```

Are these data normally distributed?
kolmogorov-smirnov test for normality - compare the (CDF) of a normal distribution to our data
Using this non-parametric test as interested in shape of the distirbutions comparison as well as location (mean)
None of our data (save perhaps one ) is normally distrbuted..

Due to this, I don't want to take SD of the data directly

```{r}
print(receptors.oi)

lapply(seq_along(receptors.oi), function(x,y,i){

  recept <-  x[[i]]
  
  # extract the sig prots per receptor in the non-detrended, we want to see how these profiles change over time
  genes.oi <- y[[recept]][receptor == recept & pvalue < 0.05 & abs(log2FC) > 0.58, gene]

  test.data <-  y[[recept]][receptor == recept & gene %in% genes.oi,delta.log2FC]

  ks.test(test.data, 'pnorm')

},x=receptors.oi, y=comb.mss.list)
```


lm and look at residuals to identify points worth removing
are we overfitting to points falling off-diagonal? maybe use all shared points or all sig in ori data instead?
Think this is something BP can contribute to, so don't waste more time and park where it is for now

```{r}
lapply(seq_along(receptors.oi), function(x,y,i){
  
  recept <-  x[[i]]
  
  # extract the sig prots per receptor in the non-detrended, we want to see how these profiles change over time
  genes.oi <- y[[recept]][receptor == recept & pvalue < 0.05 & abs(log2FC) > 0.58, gene]
  
  qplot(x=y[[recept]][receptor == recept & gene %in% genes.oi,]$detrended.log2FC, y=y[[recept]][receptor == recept & gene %in% genes.oi,]$log2FC) +
    stat_smooth(method='lm') +
    ggtitle(recept) +
    xlab('detrended log2FC') +
    ylab('log2FC') +
    theme_classic()

},x=receptors.oi, y=comb.mss.list)
```


histogram of log2FC and detrended log2FC for absolute fold change values

```{r}
lapply(seq_along(receptors.oi), function(x,y,i){

  recept <-  x[[i]]
  
  # extract the sig prots per receptor in the non-detrended, we want to see how these profiles change over time
  genes.oi <- y[[recept]][receptor == recept & pvalue < 0.05 & abs(log2FC) > 0.58, gene]

  
  ggplot() + 
    geom_histogram(data=y[[recept]][receptor == recept & gene %in% genes.oi], aes(x=abs(log2FC), fill='pink')) +
    geom_histogram(data=y[[recept]][receptor == recept & gene %in% genes.oi], aes(x=abs(detrended.log2FC), fill='lightblue', alpha=0.4)) +
    scale_fill_manual(values=c('lightblue'='lightblue', 'pink'='pink'), labels=c('lightblue'='detrended', 'pink'='original'), name='source')+
    ggtitle(paste0(recept, ' log2FC histograms')) +
    theme_bw()
  
},x=receptors.oi, y=comb.mss.list)
```


```{r}
lapply(seq_along(receptors.oi), function(x,y,i){

  recept <-  x[[i]]
  
  # extract the sig prots per receptor in the non-detrended, we want to see how these profiles change over time
  genes.oi <- y[[recept]][receptor == recept & pvalue < 0.05 & abs(log2FC) > 0.58, gene]

  
  ggplot(data=y[[recept]][receptor == recept & gene %in% genes.oi,], aes(x=abs(all.delta.log2FC))) + 
    geom_histogram(bins=50) +
    ggtitle(paste0(recept, ' detrended log2FC - log2FC')) +
    theme_bw()
  
},x=receptors.oi, y=comb.mss.list)
```


how to prune results we are not interested in? Find distribution of residuals from line, and drop points > 2+SD from it? Issue is we see that the trafficking related genes are still in the 


finding points distant to line, what about creating a linear model 
```{r}
lapply(seq_along(receptors.oi), function(x,y,i){
  
  recept <-  x[[i]]
  
  # extract the sig prots per receptor in the non-detrended, we want to see how these profiles change over time
  genes.oi <- y[[recept]][receptor == recept & pvalue < 0.05 & abs(log2FC) > 0.58, gene]
  
  qplot(x=y[[recept]][receptor == recept,]$detrended.log2FC, y=y[[recept]][receptor == recept,]$log2FC) +
    stat_smooth(method='lm') +
    geom_abline(slope=1, intercept = 0, linetype=3, alpha=0.6) +
    ggtitle(recept) +
    xlab('detrended log2FC') +
    ylab('log2FC') +
    theme_classic()

},x=receptors.oi, y=comb.mss.list)
```



