---
title: "WGCNA_GProteinCorrelations"
author: "Martin Gordon"
date: "2024-08-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## We have a set of modules from WGCNA that look quite distinct expression patterns
To try connect the modules to putative GProteins specific for these modules, lets take the modules, summarise these values and correlate these modules with the GProtein classification schemes
Definitely an argument to drop cluster 11; looks very noisy
Check the ME plots match the modules



**Notes**
Make sure module assignments match! Doesnt make sense for example that Mod1 is not ver strongly correlated with any group, but 
```{r}
library(data.table)
library(ggplot2)
library(ComplexHeatmap)
library(WGCNA)
library(magrittr)
library(preprocessCore) # quantile normalisation
library(stringr)
library(matrixStats)
library(simplifyEnrichment)
library(fmsb)
library(readxl)
library(ggbeeswarm)
library(ggrepel)

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")

# wgcna mod color pal
col.pal <- c("#D57AC7","#8DDDE1","#A1E4AC","#6EE766","#CEDF5F","#DEC5D2","#85A693","#DBBB7C","#7D9AD6","#D96A6E","#A94FDD")  #"#999933" '#DDCC77' "grey"
```

Save these for later; new names for the modules

```{r}
mod.names <-  data.table(cluster = paste0('cluster', seq(1,11,1)),
                         moduleName = c('C1 Endosome trafficking', 'C2 PM structure/adhesion', 'C3 Actin cytoskeleton', 
                                        'C4 Golgi trafficking', 'C5 G-Protein coupled receptor signalling', 'C6 Protein Kinase C',
                                        'C7 Clathrin', 'C8 Cytoplasmic stress granule','C9 Actin polymerization', 
                                        'C10 Cell morphogenesis', 'C11 RNA processing'))


#fwrite(mod.names, ScriptAndDatedFileName('moduleNamesforUse.txt'))

#mod.names <- fread('~/Documents/projects/082423_QLi_GPR_WGCNA/WGCNA_GProteinCorrelation_data/2024_08_27_moduleNamesforUse.txt')
mod.names <- fread('~/Documents/projects/082423_QLi_GPR_WGCNA/WGCNA_GProteinCorrelation_data/2024_08_27_moduleNamesforUse.txt')
```


From GPCR_WGCNA markdown
15/09/23
The matrix used to generate the 11 receptor data is saved under './output/gpcr.11.receptors.adj.mat.Rdata'
Module assignments saved under './output/gpcr.11.receptors.moduleAssignment.Rdata'


#saveRDS(sig.p.mat,file = './output/gpcr.11.receptors.exp.mat.Rdata') # the expression m

Load in these groups 
```{r}
p.mat <- readRDS('./output/gpcr.11.receptors.exp.mat.Rdata')

module.assignment <- readRDS('./output/gpcr.11.receptors.moduleAssignment.Rdata')
module.assignment


p.mat

iupar.cor <- fread('./output/cluster.cor.med.shiftvsT0.iupar.csv') 
pred.cor <- fread('./data/cluster.cor.med.shiftvsT0.predicted.csv')

submat <- as.matrix(iupar.cor, rownames='cluster')

# matching order int he matrix to the dt
# first argument to match is the order you want to assign
submat <- submat[match(mod.names$cluster,rownames(submat)),]

rownames(submat) <- mod.names$moduleName

hm <- Heatmap(submat, 
        row_km = 2, 
        row_names_gp = gpar(fontsize=10),
        border=T, 
        name='Pearsons r',
        column_title = 'WGCNA Module IUPAR correlations',
        cell_fun = function(j, i, x, y, width, height, fill) {
        if(submat[i, j] >= 0.6 | submat[i, j] <= -0.6){
            grid.text(sprintf("%.2f", submat[i, j]), x, y, gp = gpar(fontface='bold', fontsize = 10, col='white'))
          } else {
            grid.text(sprintf("%.2f", submat[i, j]), x, y, gp = gpar(fontsize = 10, col='black'))
          }})
hm
BackupAsPDF(hm,'IUPAR.medianShiftvsTo.heatmap')
```
Lets look at the prediction correlations

```{r}
submat <- as.matrix(pred.cor, rownames='cluster')

submat <- submat[match(mod.names$cluster,rownames(submat)),]

rownames(submat) <- mod.names$moduleName

hm <- Heatmap(submat, 
        row_km = 2, 
        row_names_gp = gpar(fontsize=10),
        border=T, 
        name='Pearsons r',
        column_title = 'WGCNA Module GProtein classification',
        cell_fun = function(j, i, x, y, width, height, fill) {
        if(submat[i, j] > 0.6 | submat[i, j] < -0.6){
            grid.text(sprintf("%.1f", submat[i, j]), x, y, gp = gpar(fontface='bold', fontsize = 10, col='white'))
          } else {
            grid.text(sprintf("%.1f", submat[i, j]), x, y, gp = gpar(fontsize = 10, col='black'))
          }})

hm
BackupAsPDF(hm,'Predicted.medianShiftvsTo.heatmap', dimensions=c(8,8))
```
I think correlating this median Shift may be an issue; lets look at a different measurement to try capture trends
ME is 1st PC
We want to read in the WGCNA results; identify the module eigengenes for each group ( or just calculate the first PC) and then correlate this with the pred and iupar scores
Next step then is to find the most central genes in each module, or correlate that modules genes with the 

```{r}
MEs <- fread('./output/11.receptors.moduleEigengenes.wrownames.csv')
MEs
ME.mat <- as.matrix(MEs, rownames = 'V1')
colnames(ME.mat) <- paste0('cluster',gsub('ME', '', colnames(ME.mat)))

ME.mat <- ME.mat[, match(mod.names$cluster, colnames(ME.mat))]

colnames(ME.mat) <-  paste('ME', mod.names$moduleName)

```

plot the module eigengene scores for each sample

```{r}
ann.dt <- data.table(rn = rownames(ME.mat))
ann.dt[, c('receptor', 'timepoint') := tstrsplit(rn, '_', keep=c(1,2))]
ann.dt[, batch := str_extract(rn, '[0-9]{1,2}$')]

#set up a color palette
col.pal <- randomcoloR::distinctColorPalette(k=length(unique(ann.dt$timepoint)))
names(col.pal) <- unique(ann.dt$timepoint)

colAnn <- HeatmapAnnotation(timepoint=ann.dt$timepoint,
                            col = list(timepoint=col.pal))


submat <- t(ME.mat)
# also want to add a GPCR classification
hm <- Heatmap(submat,
        column_split = ann.dt$receptor,
        border=T,
        cluster_rows = F,
        name='module Eigengene (1st PC)',
        top_annotation = colAnn,
        cluster_columns = F,
        column_names_gp = gpar(fontsize=5),
        row_names_gp = gpar(fontsize=6))
        
hm

BackupAsPDF(hm, 'moduleEigengenes.sampleCor.noRowClust.heatmap', dimensions=c(14,6))
```
Probably a little too much going on. Get an mean aggregated matrix and plot that
Looks good add some extra annotation

```{r}
# now change the variance
MEs[, c('receptor', 'timepoint') := tstrsplit(V1, '_', keep=c(1,2))]
MEs[, batch := str_extract(V1, '[0-9]{1,2}$')]
MEs

MEs.long.dt <- setDT(reshape2::melt(MEs, id.vars = c('V1', 'receptor', 'timepoint', 'batch')))


ME.sum.mat <- dcast(MEs.long.dt, paste(receptor,timepoint, sep='_')~variable, value.var = 'value', fun.aggregate = mean, na.rm=T)  %>% 
  as.matrix(rownames=1)

# match col order 
ME.sum.mat <- ME.sum.mat[, match(gsub('cluster', '', mod.names$cluster), gsub('ME', '',colnames(ME.sum.mat)))]
colnames(ME.sum.mat) <- paste('ME', mod.names$moduleName)

ann.dt <- data.table(rn = rownames(ME.sum.mat))
ann.dt[, c('receptor', 'timepoint') := tstrsplit(rn, '_', keep=c(1,2))]

# lets prepare the annotation
colAnn <- HeatmapAnnotation(timepoint=ann.dt$timepoint,
                            gprotein=ifelse(grepl('DOR|MOR|KOR|NPY1R|NPFFR1', rownames(ME.sum.mat)), 'Gi/Go',
                                            ifelse(grepl('MRX2|QRFPR', rownames(ME.sum.mat)), 'Gi/Go & Gq/G11',
                                                   ifelse(grepl('MRX4|PTGER1', rownames(ME.sum.mat)), 'Gq/G11',
                                                          ifelse(grepl('NK1R', rownames(ME.sum.mat)), 'Gq/G11 & Gs', 'Gs')))), 
                            col = list(timepoint=col.pal,
                                       gprotein=c('Gi/Go & Gq/G11'='pink', 'Gq/G11'='lightblue', 'Gi/Go'='lightgreen', 'Gs'='#CBC3E3','Gq/G11 & Gs'='#FFD580')))
                                       
submat <- t(ME.sum.mat)
# also want to add a GPCR classification
hm <- Heatmap(submat,
        column_split = gsub('_[0-9]{1,2}','', colnames(submat)),
        border=T,
        name='module Eigengene (1st PC)',
        top_annotation = colAnn,
        cluster_columns = F,
        cluster_rows = F,
        show_column_names = F,
        column_names_gp = gpar(fontsize=5),
        row_names_gp = gpar(fontsize=6))
        
hm
BackupAsPDF(hm, 'moduleEigengenes.sampleCor.noRowClust.heatmap', dimensions=c(14,6))
```
Identifying important genes in out network(s)
First define Gene significance; 
absolute value of correlation between gene and trait (Gi/Go,Gq/G11,Gs) and then define Module membership (correlation between gene and eigengene)


```{r}
pred <- fread("/Users/martingordon/Documents/projects/310723_QLi_GPCR.correlations/data/predictionScores.txt")
iupar <- data.table(read_xlsx("/Users/martingordon/Documents/projects/310723_QLi_GPCR.correlations/data/IUPHAR_11GPCRs_Gprotein_class_primary.xlsx"))


exp.long <- melt(exp, id.vars = "GPCR", measure.vars = grep ("GNA", colnames(pred)), variable.name = "GProtein", value.name = "score")[, score := as.numeric(score)][]
pred.long <- melt(pred, id.vars = "GPCR", measure.vars = grep ("GNA", colnames(pred)), variable.name = "GProtein", value.name = "score")[, score := as.numeric(score)][]

# convert to long format
iupar.long <- melt(iupar, id.vars = 'GPCR', measure.vars = grep('GPCR', colnames(iupar), invert=T),  variable.name = "GProtein", value.name = "score") %>% 
  .[, score := as.numeric(score)] 
```

```{r scores.list}
scores.list <- list(pred = pred.long,
                    iupar = iupar.long)
```

Convert the scores to a matrix

```{r}
scores.mat <- lapply(scores.list, function(x){
  dcast(x, GProtein~GPCR, value.var = 'score') %>% 
    as.matrix(rownames = 'GProtein')
})
```
change the colnames to reflect the mss nomenclature

```{r}
scores.mat <- lapply(scores.mat, function(x){
  colnames(x) <- c("MRX2","MRX4","NPFFR1","NPY1R","DOR","KOR","MOR","PTGER1","PTGER4","QRFPR","NK1R")
  return(x)
})

lapply(scores.mat, colnames)
```
Maybe some kind of module IUPAR classification plot?
Again, to calculate module-trait correlations

```{r}
mss.out <- fread('./data/2022_08_31_NiceFitsPower3.csv')
mss.out[receptor == 'MOR.DAMGO', receptor := 'MOR']

exp.mat <- dcast(mss.out,gene~receptor, value.var = 'log2FC') %>% 
  .[, MOR.LG94 := NULL] %>% 
  as.matrix(rownames ='gene')


# filter to sig genes or strong movers and reorder
sig.genes <- mss.out[pvalue < 0.005 & abs(log2FC) > log2(1.5) & receptor != 'MOR.LG94', unique(gene) ]
exp.mat <- exp.mat[rownames(exp.mat) %in% sig.genes,]

colnames(exp.mat) %in% colnames(scores.mat[[1]])

exp.mat <- exp.mat[, match(colnames(scores.mat[[1]]), colnames(exp.mat))]
colnames(exp.mat) == colnames(scores.mat[[1]])
```

Now run the PW correlations between genes and GProtein group

```{r}
scor.list <- lapply(scores.mat, function(x){
  
  cor(t(exp.mat), t(x), use= "pairwise.complete.obs") # cluster by row, so keep rownames as receptor
})


cor.long <- lapply(scor.list, function(x){
  x <- melt(as.data.table(x, keep.rownames = TRUE), id.vars = "rn", variable.name  = "GProtein", value.name = "R")
  setnames(x, old = "rn", new = "gene")
  return(x)
})

# jsut work with this for now; these are correlations between the genes and proteins
iupar.cor <- cor.long$iupar

#fwrite(cor.long$iupar, ScriptAndDatedFileName('geneSignificance.IUPAR.groups.csv'))
iupar.gs.dt <- fread('~/Documents/projects/082423_QLi_GPR_WGCNA/WGCNA_GProteinCorrelation_data/2024_08_29_geneSignificance.IUPAR.groups.csv')
```

Now correlate the individual genes with the module MEs

```{r}
# check the colnames match
colnames(t(ME.mat)) == colnames(p.mat)

# correlate each gene with the module eigengene 
module.membership <- cor(t(p.mat), ME.mat, use='pairwise.complete.obs')

# convert to DT
mm.dt <- reshape2::melt(module.membership)
setnames(mm.dt, new=c('Protein', 'module', 'module R (pearson)'))
mm.dt[, gene := multiUniprots2multiGenes(Protein, species='HUMAN')]

#fwrite(mm.dt, ScriptAndDatedFileName('moduleMembership.Rscores.csv'))
mm.dt <- fread('~/Documents/projects/082423_QLi_GPR_WGCNA/WGCNA_GProteinCorrelation_data/2024_08_29_moduleMembership.Rscores.csv')
```

create a module assignment parameter
```{r}
module.assignment.dt <- data.table(Protein = rownames(p.mat),
                                   Module = module.assignment)

module.assignment.dt[, gene := multiUniprots2multiGenes(Protein,species='HUMAN')]
# use this from now on for data
#fwrite(module.assignment.dt, ScriptAndDatedFileName('moduleAssignments.csv'))
module.assignment.dt <- fread('~/Documents/projects/082423_QLi_GPR_WGCNA/WGCNA_GProteinCorrelation_data/2024_08_29_moduleAssignments.csv')

comb.dt <- merge(x=mm.dt, y=dcast(iupar.gs.dt, gene~GProtein, value.var='R'), by='gene')
setnames(comb.dt, new=c('gene', 'module', 'module_R', 'protein', 'Gi/Go', 'Gq/G11', 'Gs'))


ggplot(comb.dt[module == 'ME C1 Endosome trafficking', ], aes(x=module_R, y=`GiGo/`)) +
  geom_point()

module.assignment.dt <- data.table(Protein = rownames(p.mat),
                                   Module = module.assignment
                                   )

module.assignment.dt[, gene := multiUniprots2multiGenes(Protein,species='HUMAN')]
```
Combine the module assignments and the gene significance plots and plot the conditions

Calculate pvalues of the MM and GS correlations

```{r}


help("corPvalueStudent")
comb.dt <- merge(x=comb.dt, y=module.assignment.dt[, -c('Protein')], by='gene', all=T)
setnames(comb.dt,  new=c('gene', 'moduleName', 'moduleR', 'protein', 'R_GiGo', 'R_GqG11', 'R_Gs', 'geneModuleAssignment'))

comb.dt[, moduleID := tstrsplit(moduleName, ' ', keep=c(2))]
comb.dt[, moduleID := gsub('C', '', moduleID)]

exp.mat[rownames(exp.mat) %in% comb.dt[is.na(moduleName),]$gene,]

exp.mat[rownames(exp.mat) %in% comb.dt[is.na(moduleName), unique(gene)],]

comb.dt[, moduleName := factor(moduleName, levels=c("ME C1 Endosome trafficking","ME C2 PM structure/adhesion","ME C3 Actin cytoskeleton","ME C4 Golgi trafficking",
                                                    "ME C5 G-Protein coupled receptor signalling","ME C6 Protein Kinase C","ME C7 Clathrin","ME C8 Cytoplasmic stress granule", 
                                                    "ME C9 Actin polymerization","ME C10 Cell morphogenesis","ME C11 RNA processing"))]

comb.dt <- melt(comb.dt, measure.vars= c('R_GiGo','R_GqG11','R_Gs'))
setnames(comb.dt, old=c('variable', 'value'), new=c('IUPAR', 'R score'))
setnames(comb.dt, old=c('R score'), new=c('IUPAR_R'))

g <- ggplot(comb.dt[moduleID == geneModuleAssignment,], aes(x=moduleR, y=IUPAR_R, label=gene, color=moduleName)) +
  geom_point(size=1) +
  ggrepel::geom_text_repel(data=comb.dt[moduleID == geneModuleAssignment & IUPAR_R > 0.5 & moduleR > 0.75,], max.overlaps = 30, segment.alpha=0.3, segment.color='black', size=2) +
  geom_vline(xintercept=0.75, alpha=0.4, linetype=2) +
  geom_hline(yintercept = 0.5, alpha=0.4, linetype=2) +
  xlim(c(0,1)) +
  ggtitle('Module membership vs Gene Significance') +
  xlab('Module Membership (R)') +
  ylab('Gene significance (R )') +
  scale_color_manual(values=col.pal) +
  ggh4x::facet_grid2(moduleName~IUPAR) +
  theme_bw()
g
BackupAsPDF(g, 'GSvsMM.scatterplot', dimensions = c(14,24))
```




```{r}

# coordinator function for generating radar plots: see https://stackoverflow.com/questions/42562128/ggplot2-connecting-points-in-polar-coordinates-with-a-straight-line-2#42572133
coord_radar <- function (theta = "x", start = 0, direction = 1) {
  theta <- match.arg(theta, c("x", "y"))
  r <- if (theta == "x") "y" else "x"
  ggproto("CordRadar", CoordPolar, theta = theta, r = r, start = start, 
          direction = sign(direction),
          is_linear = function(coord) TRUE)
}

radar.data <- protQuant[!is.na(Cluster)][, .(receptor, GO.terms, medianMaxShiftvsT0)] %>%  
  unique()

g <- ggplot(radar.data, aes(x=GO.terms, y=medianMaxShiftvsT0, group=receptor, color=receptor)) + 
  stat_summary(fun = "median", geom = "line") +
  scale_color_manual(values = col.pal) + 
  labs(x='timepoint (mins)', y="Intensity (log2)") +
 # facet_wrap(~receptor, ncol=3) +
  theme_classic() +
  scale_x_discrete(expand=c(0,0.03)) +
  theme(
   strip.text = element_text(size = 6)) +
  coord_radar() +
  
g

```



i) Other idea for plot; summarise the ME per receptor and correlate this with IUPAR classifications.. how to do this
ii) plot gene  significance vs module membershape (trait vs module gene correlation)
iii)

For now try collapse the info in ME
Take the max value in each group/row; biggest deviation to T0 i guess
Summarise to max value in each



Don't bother with this summarisation to one value per module/receptor

```{r}
ME.sum.mat <- dcast(MEs.long.dt, receptor~variable, value.var = 'value', fun.aggregate = max, na.rm=T)  %>% 
  as.matrix(rownames=1)

# match col order 
ME.sum.mat <- ME.sum.mat[, match(gsub('cluster', '', mod.names$cluster), gsub('ME', '',colnames(ME.sum.mat)))]
colnames(ME.sum.mat) <- paste('ME', mod.names$moduleName)

ann.dt <- data.table(rn = rownames(ME.sum.mat))

# lets prepare the annotation
colAnn <- HeatmapAnnotation(
                            gprotein=ifelse(grepl('DOR|MOR|KOR|NPY1R|NPFFR1', rownames(ME.sum.mat)), 'Gi/Go',
                                            ifelse(grepl('MRX2|QRFPR', rownames(ME.sum.mat)), 'Gi/Go & Gq/G11',
                                                   ifelse(grepl('MRX4|PTGER1', rownames(ME.sum.mat)), 'Gq/G11',
                                                          ifelse(grepl('NK1R', rownames(ME.sum.mat)), 'Gq/G11 & Gs', 'Gs')))), 
                            col = list(timepoint=col.pal,
                                       gprotein=c('Gi/Go & Gq/G11'='pink', 'Gq/G11'='lightblue', 'Gi/Go'='lightgreen', 'Gs'='#CBC3E3','Gq/G11 & Gs'='#FFD580')))
                                       
submat <- t(ME.sum.mat)
# also want to add a GPCR classification
hm <- Heatmap(submat,
        border=T,
        clustering_distance_rows = 'pearson',
        clustering_distance_columns = 'pearson',
        name='module Eigengene (1st PC)',
        top_annotation = colAnn,
        cluster_columns = T,
        col=colorRamp2(breaks=c(0,0.4), c('white', 'red')),
        show_column_names = T,
        column_names_gp = gpar(fontsize=5),
        row_names_gp = gpar(fontsize=6))
        
hm
BackupAsPDF(hm, 'moduleEigengenes.sampleCor.heatmap', dimensions=c(14,6))
```
I dont know if this does a good job of capturing the trends... maybe try take max in either direction?
Or take the IUPAR/Pred scores, multiply by 5

Correlate scores with ME
```{r}
scor.list <- lapply(scores.mat, function(x){
  
  mat.x <- submat[, colnames(x)]
  colnames(mat.x) == colnames(x)
  
  print(mat.x)
  print(t(x))
  cor(mat.x, t(x), use= "pairwise.complete.obs") # cluster by row, so keep rownames as receptor
})


lapply(scor.list, function(x)Heatmap(x))

cor.long <- lapply(cor.list, function(x){
  x <- melt(as.data.table(x, keep.rownames = TRUE), id.vars = "rn", variable.name  = "GProtein", value.name = "R")
  setnames(x, old = "rn", new = "cluster")
  return(x)
})
```

Tried creating protein groups for got the predictiion scores, maybe ask Qiongyu, but leave out for now

```{r}
gprotein.grps <- data.table(gprotein     = c('GNAI3', 'GNAI1', 'GNAZ',  'GNAO1', 'GNA12', 'GNA13', 'GNAQ', 'GNA14', 'GNA15', 'GNAS',  'GNAL'),
                            gprotein_family = c('Gi', 'Gi', 'Gi', '', 'G12', 'G13', 'Gq', 'Gq', 'Gq', 'Gs'))
```


Write out boht matrices incase I want to make changes

```{r}
#fwrite(as.data.table(ME.sum.mat, keep.rownames = T), ScriptAndDatedFileName('ME.meanAggregate.matrix.txt'))
#fwrite(as.data.table(ME.mat, keep.rownames = T), ScriptAndDatedFileName('ME.matrix.txt'))
```

```{r}

# now change the variance
MEs[, c('receptor', 'timepoint') := tstrsplit(V1, '_', keep=c(1,2))]
MEs[, batch := str_extract(V1, '[0-9]{1,2}$')]


# now take the average of the receptor, timepoint combos
MEs[receptor == 'DOR' & timepoint == '00',]
MEs
MEs[, grep('V1|ME', colnames(MEs), value=T)]
```

Plot a heatmap of the MEs of each module; these are the PCs of each, so lets see if any MEs associate with particular groups

```{r}
ggplot(MEs, aes(x=ME1, y=ME2, colour = timepoint)) +
  geom_point() +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept=0)

MEs.long.dt <- setDT(reshape2::melt(MEs, id.vars = c('V1', 'receptor', 'timepoint', 'batch')))

ggplot(MEs.long.dt, aes(x=receptor, y=value, color=timepoint)) +
  geom_point() +
  facet_wrap(~variable)
```



```{r}
ann.dt <- data.table(rn = rownames(ME.mat))
ann.dt[, c('receptor', 'timepoint') := tstrsplit(rn, '_', keep=c(1,2))]
ann.dt[, batch := str_extract(rn, '[0-9]{1,2}$')]

#set up a color palette
col.pal <- randomcoloR::distinctColorPalette(k=length(unique(ann.dt$timepoint)))
names(col.pal) <- unique(ann.dt$timepoint)

colAnn <- HeatmapAnnotation(timepoint=ann.dt$timepoint,
                            col = list(timepoint=col.pal))

# also want to add a GPCR classification

hm <- Heatmap(t(ME.mat),
        column_split = ann.dt$receptor,
        border=T,
        name='module Eigengene (1st PC)',
        top_annotation = colAnn,
        column_order = ann.dt$rn,
        column_names_gp = gpar(fontsize=5)
        )
hm


scores.mat
```


Get table of the module assignments

```{r}
modAssignments <- fread('./WGCNA_GProteinCorrelation_data/2024_08_29_moduleAssignments.csv')
modAssignments[, Module:= as.character(Module)]

mod.names[, Module := gsub('cluster', '', cluster)]

modAssignments <- merge(modAssignments, y=mod.names[,.(Module,moduleName)], by='Module')

fwrite(modAssignments[, .(Module,Protein, gene, NProtsInModule = .N), by=moduleName], ScriptAndDatedFileName('modAssignments.csv'))
```






**Not used**

plot the median Shift (not vs T0)

```{r}

-log10(0.005)
iupar.cor <- fread('./output/cluster.cor.med.log2FC.iupar.csv') 
pred.cor <- fread('./data/cluster.cor.med.log2FC.predicted.csv')

submat <- as.matrix(iupar.cor, rownames='cluster')

hm <- Heatmap(submat, 
        row_km = 2, 
        border=T, 
        name='Pearsons r',
        column_title = 'WGCNA Module IUPAR classification',
        cell_fun = function(j, i, x, y, width, height, fill) {
        if(submat[i, j] >= 0.6 | submat[i, j] <= -0.6){
            grid.text(sprintf("%.2f", submat[i, j]), x, y, gp = gpar(fontface='bold', fontsize = 10, col='white'))
          } else {
            grid.text(sprintf("%.2f", submat[i, j]), x, y, gp = gpar(fontsize = 10, col='black'))
          }},
        row_title = "KM cluster %s")
hm
BackupAsPDF(hm,'IUPAR.medianShift.heatmap')


# plot the same for the predicted classification
submat <- as.matrix(pred.cor, rownames='cluster')


hm <- Heatmap(submat, 
        row_km = 2, 
        border=T, 
        name='Pearsons r',
        column_title = 'WGCNA Module GProtein classification',
        cell_fun = function(j, i, x, y, width, height, fill) {
        if(submat[i, j] >= 0.6 | submat[i, j] <= -0.6){
            grid.text(sprintf("%.2f", submat[i, j]), x, y, gp = gpar(fontface='bold', fontsize = 10, col='white'))
          } else {
            grid.text(sprintf("%.2f", submat[i, j]), x, y, gp = gpar(fontsize = 10, col='black'))
          }},
        row_title = "KM cluster %s")
hm
BackupAsPDF(hm,'pred.medianShift.heatmap')

```

Now do the PW correlations to each of the  prediciton scores

```{r}
# correlate t with each
cor.list <- lapply(scores.mat, function(x){
  cor(t(ME.mat), t(x), use= "pairwise.complete.obs") # cluster by row, so keep rownames as receptor
})


cor.long <- lapply(cor.list, function(x){
  x <- melt(as.data.table(x, keep.rownames = TRUE), id.vars = "rn", variable.name  = "GProtein", value.name = "R")
  setnames(x, old = "rn", new = "cluster")
  return(x)
})
```




This is the WGCNA data; lets retry the correlation work to see what we are recovering

```{r}
pred <- fread("/Users/martingordon/Documents/projects/310723_QLi_GPCR.correlations/data/predictionScores.txt")
exp <- fread ("/Users/martingordon/Documents/projects/310723_QLi_GPCR.correlations/data/experimentalScores.txt")
iupar <- data.table(read_xlsx("/Users/martingordon/Documents/projects/310723_QLi_GPCR.correlations/data/IUPHAR_11GPCRs_Gprotein_class_primary.xlsx"))


exp.long <- melt(exp, id.vars = "GPCR", measure.vars = grep ("GNA", colnames(pred)), variable.name = "GProtein", value.name = "score")[, score := as.numeric(score)][]
pred.long <- melt(pred, id.vars = "GPCR", measure.vars = grep ("GNA", colnames(pred)), variable.name = "GProtein", value.name = "score")[, score := as.numeric(score)][]

# convert to long format
iupar.long <- melt(iupar, id.vars = 'GPCR', measure.vars = grep('GPCR', colnames(iupar), invert=T),  variable.name = "GProtein", value.name = "score") %>% 
  .[, score := as.numeric(score)] 
```

```{r scores.list}
scores.list <- list(pred = pred.long,
                    iupar = iupar.long)
```

Convert the scores to a matrix

```{r}
scores.mat <- lapply(scores.list, function(x){
  dcast(x, GProtein~GPCR, value.var = 'score') %>% 
    as.matrix(rownames = 'GProtein')
})
```
change the colnames to reflect the mss nomenclature

```{r}
scores.mat <- lapply(scores.mat, function(x){
  colnames(x) <- c("MRX2","MRX4","NPFFR1","NPY1R","DOR","KOR","MOR","PTGER1","PTGER4","QRFPR","NK1R")
  return(x)
})
```


```{r}
readRDS('./output/wgcna.clustergeens.adjmat.Rdata')
```

