---
title: "WGCNA_clustersSTRING"
author: "Martin Gordon"
date: "2023-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## WGCNA CLusters STRING network 

```{r}
library(data.table)
library(stringr) #str_extract function 
library(magrittr)
library(ggplot2)
library(pbapply) #apply functions with progress bar
library(igraph)
library(RCy3) # cytoscape interface

# load bens util scripts for the analysis
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/LocatePeptidePtmInProteins.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

# netprop scripts
source ("../../utils/bp_utils/STRING_db_utils.R")
source ("../../utils/bp_utils/NetworkPropagation.R")
source("../../utils/bp_utils/PPINetworkModularized.R")

```

Load the cluster assignment and submit each cluster gene list to STRING to generate a PPIs

```{r}
# file with cluster assignment for each gene/protein
cluster.assignment <- fread('./output/wgcna.clustergenes.adjmat.csv')

string.gene.map <- fread('./data/9606.protein.aliases.v12.0.txt.gz')
string.gene.map[,GeneID := alias]
colnames(string.gene.map) <- c('string', 'alias', 'source', 'GeneID')

cluster.assignment[string.gene.map, stringID := i.string, on=c('GeneID', 'GeneID') ]
cluster.assignment[, Cluster := sub('cluster', '', Cluster)]

# add ME info to node table
MEs <- fread('./output/top10.K.percluster.csv')

MEs[, GeneID := rn]

cluster.assignment[MEs, ME := GeneID, on=c('GeneID', 'GeneID')]
cluster.assignment[!is.na(ME),]

cluster.node.table <- split(cluster.assignment, cluster.assignment$Cluster)
#reorder
cluster.node.table <- cluster.node.table[clusters]

```




Create a visual style for the network

```{r}
#deleteVisualStyle("gcpr.style")
# color palette for the 11 modules
col.pal <- c("#D57AC7","#8DDDE1","#A1E4AC","#6EE766","#CEDF5F","#DEC5D2","#85A693","#DBBB7C","#7D9AD6","#D96A6E","#A94FDD")  #"#999933" '#DDCC77' "grey"


style.name = "gcpr.style"


defaults <- list(NODE_SHAPE="ellipse",
                 LABEL_FONT_SIZE=30,
                 NODE_SIZE=50,
                 EDGE_TRANSPARENCY=100,
                 EDGE_OPACITY=100,
                 NODE_COLOR="#D3D3D3",
                 NODE_BORDER_WIDTH =2)

nodeLabels      <- mapVisualProperty('node label','GeneID','p')
nodeFills       <- mapVisualProperty('node fill color','Cluster','d', c("1","2", "3","4","5","6","7","8","9", "10","11"), col.pal)
edgeWidth       <- mapVisualProperty('edge width','stringdb::score','c', c(0.8,1), c(1,5))


createVisualStyle(style.name, defaults, list(nodeLabels,nodeFills,edgeWidth))
```

iterate through the clusters and create a string network from each 

```{r}
clusters <- cluster.assignment[, unique(Cluster)]
```

create a  high-confidence PPI from the STRING network (0.8)
Don't add connectors right now as want to see differences in what we are measuring 
```{r}
str.query <- lapply(clusters, function(x){
  string.cmd <- paste('string protein query query="', paste(cluster.assignment[Cluster == x, GeneID], collapse = '\n'), '" cutoff=0.8 limit=0 species="Homo sapiens"',  sep = "")
  })

names(str.query) <- clusters
```

Create the networks and apply node table data and styles

```{r}
lapply(seq_along(str.query), function(x,n,i){
  # create STRING network
  commandsRun(x[[i]])
  # name network (how to merge to same collections)
  net.name <- paste0(n[[i]],'.module.STRING.800')
  renameNetwork(net.name)
  # load additional node info
  loadTableData(cluster.node.table[[i]], data.key.column=c("GeneID"), table.key.column="query term")
  
  # use above node info to set visuals for each network
  setVisualStyle(style.name)
  }, x=str.query, n=names(str.query))
```

Use MCL clustering and create subnetworks from each of the networks
We cluster on STRING score and inflation value of 2


```{r}

setNodeShapeDefault("ELLIPSE",  style.name)
lockNodeDimensions("TRUE",  style.name)
setNodeSize("65", style.name)
setNodeColorDefault("#D3D3D3", style.name)
setNodeBorderWidthDefault("2", style.name)
setNodeBorderColorDefault("#616060", style.name)
setNodeLabelMapping("display name", style.name)
setNodeSizeDefault("65", style.name)
setNodeFontSizeDefault("15", style.name)
setEdgeOpacityDefault(140, style.name)

setNodeColorMapping('log2FC', colors=paletteColorBrewerRdBu, style.name=style.name)
setNodeSizeMapping('Degree', c(1,12),c(60,110), style.name=style.name)
#setEdgeLineWidthMapping('stringdb::score', c(0.4,1), c(1,3), style.name="avidd.style")
```


Highlight the Hub Nodes; make larger and surround with a boarder
```{r}
setVisualStyle(style.name)

setNodeSizeDefault("80", style.name)
setNodeFontSizeDefault("15", style.name)
setEdgeOpacityDefault(140, style.name)

#save plots
exportImage('STRINGnetwork.Overview.800', 'PDF')
exportImage('STRINGnetwork.Overview.800', 'PNG', zoom=200)
#setNodeSizeMapping('ME', c(1,12),c(60,110), style.name="avidd.style")

```

Generate more plots of the overall network

```{r}
setNodeSizeDefault("100", style.name)
setNodeFontSizeDefault("30", style.name)
setEdgeOpacityDefault(140, style.name)

#save plots
exportImage('STRINGnetwork.module11.Overview.800', 'PDF')
exportImage('STRINGnetwork.module11.Overview.800', 'PNG', zoom=200)
```

