---
title: "WGCNA.moduleGPCRCorrelation"
author: "Martin Gordon"
date: "2023-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(ggplot2)
library(ComplexHeatmap)
library(WGCNA)
library(magrittr)
library(preprocessCore) # quantile normalisation
library(stringr)
library(matrixStats)
library(simplifyEnrichment)
library(fmsb)
library(readxl)
library(ggbeeswarm)
library(ggrepel)
library(curl)

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")


# helper scripts for module/network export etc
source ("/Users/martingordon/Documents/utils/mg_utils/HelperFunctions.R")

```

27/09/23
-----
Notes: why is the VsT0 median shift different to the median value per cluster?
Best approach is probably the eigengene correlation
have the expression and module eigengene values for each of the modules. 
Now correlate these scores with GPCR classifications

To do correlate module eigengene values with each receptor

```{r}

pred <- fread("/Users/martingordon/Documents/projects/310723_QLi_GPCR.correlations/data/predictionScores.txt")
exp <- fread ("/Users/martingordon/Documents/projects/310723_QLi_GPCR.correlations/data/experimentalScores.txt")
iupar <- data.table(read_xlsx("/Users/martingordon/Documents/projects/310723_QLi_GPCR.correlations/data/IUPHAR_11GPCRs_Gprotein_class_primary.xlsx"))


exp.long <- melt(exp, id.vars = "GPCR", measure.vars = grep ("GNA", colnames(pred)), variable.name = "GProtein", value.name = "score")[, score := as.numeric(score)][]
pred.long <- melt(pred, id.vars = "GPCR", measure.vars = grep ("GNA", colnames(pred)), variable.name = "GProtein", value.name = "score")[, score := as.numeric(score)][]

# convert to long format
iupar.long <- melt(iupar, id.vars = 'GPCR', measure.vars = grep('GPCR', colnames(iupar), invert=T),  variable.name = "GProtein", value.name = "score") %>% 
  .[, score := as.numeric(score)] 

```

```{r scores.list}
scores.list <- list(exp = exp.long,
                    pred = pred.long,
                    iupar = iupar.long)
```

Convert the scores to a matrix

```{r}
scores.mat <- lapply(scores.list, function(x){
  dcast(x, GProtein~GPCR, value.var = 'score') %>% 
    as.matrix(rownames = 'GProtein')
})
```

change the  colnames to reflect the mss nomenclature

```{r}
scores.mat <- lapply(scores.mat, function(x){
  colnames(x) <- c("MRX2","MRX4","NPFFR1","NPY1R","DOR","KOR","MOR","PTGER1","PTGER4","QRFPR","NK1R")
  return(x)
})

```

take the max shift vs T0 per cluster per receptor
redo: not very good correlations, look at median value, maybe less dominated by outliers


First reorder the cols to follow above

```{r}
# read in mss results
mss.out <- fread("/Users/martingordon/Documents/projects/310723_QLi_GPCR.correlations/data/2022_08_31_NiceFitsPower3.csv")

exp.mat <- dcast(mss.out,gene~receptor, value.var = 'log2FC') %>% 
  .[, MOR.LG94 := NULL] %>% 
  as.matrix(rownames ='gene')


# filter to sig genes and reorder
sig.genes <- mss.out[pvalue < 0.05 & abs(log2FC) > log2(1.5) & receptor != 'MOR.LG94', unique(gene) ]
exp.mat <- exp.mat[rownames(exp.mat) %in% sig.genes,]

exp.mat[1:10,1:10]

# renanme MOR DAMGO to MOR
colnames(exp.mat) <- c('DOR', 'KOR', 'MOR', 'MRX2', 'MRX4', 'NK1R', 'NPFFR1', 'NPY1R','PTGER1', 'PTGER4','QRFPR')

# reorder the cols to match the scoring mat
col.order <-  c("MRX2","MRX4","NPFFR1","NPY1R","DOR","KOR","MOR","PTGER1","PTGER4","QRFPR","NK1R")
exp.mat <- exp.mat[, col.order]
```

```{r}

# take the max log2FC value per cluster per receptor and correlate these
mss.out[, max.log2FC := max(log2FC, na.rm=T), by=.(cluster,receptor)]
mss.out[, median.log2FC := median(log2FC, na.rm=T), by=.(cluster,receptor)]

# now want a matrix of cluster/receptor combinations
# take the median log2FC value per cluster (of the max log2FC voer time)

clust.mat <- mss.out %>% 
  .[receptor != 'MOR.LG94', .(median.log2FC,cluster,receptor)] %>% 
  unique() %>% 
  dcast(receptor~cluster, value.var = 'median.log2FC')  %>% 
  as.matrix(rownames='receptor') %>% 
  t()

colnames(clust.mat)

colnames(clust.mat) <- c('DOR', 'KOR', 'MOR', 'MRX2', 'MRX4', 'NK1R', 'NPFFR1', 'NPY1R','PTGER1', 'PTGER4','QRFPR')
# reorder to match above
clust.mat <- clust.mat[, col.order]


clust.mat %>%  t() %>%  head()
lapply(scores.mat, function(x) t(x) %>%  head())
```

with the maximum log2FC val per receptor/cluster, correlations were quite poor (max 0.6) but improved when taking median log2FC values per cluster
cluster 6 showed v strong correlation with GqG11 protein profile
save these results
```{r}

# correlate t with each
cor.list <- lapply(scores.mat, function(x){
  cor(t(clust.mat), t(x), use= "pairwise.complete.obs") # cluster by row, so keep rownames as receptor
})


cor.long <- lapply(cor.list, function(x){
  x <- melt(as.data.table(x, keep.rownames = TRUE), id.vars = "rn", variable.name  = "GProtein", value.name = "R")
  setnames(x, old = "rn", new = "cluster")
  return(x)
})
```

generate a heatmap of the 3 results
convert to wide format and a matrix

```{r}
lapply(cor.long, names)
```

plot the enrichment results for the 'interesting' modules

```{r}
mods <- c('1','2','3','4','6','7','8','9','11')


cor.long %>%  names()
names(cor.long) <- c("Experimental","Predicted","IUPAR")

lapply(seq_along(cor.long), function(x,n,i){
  

  # convert to wide
  x.mat <- dcast(x[[i]], str_extract(cluster,'[0-9].*$')~GProtein, value.var = 'R') %>% as.matrix(rownames='cluster')
  x.mat <- x.mat[rownames(x.mat) %in% mods,] #filter out noisy clusters
  
  # annotation cols  for all mods
  col.pal <- c("#D57AC7","#8DDDE1","#A1E4AC","#6EE766","#CEDF5F","#DEC5D2","#85A693","#DBBB7C","#7D9AD6","#D96A6E","#A94FDD")  #"#999933" '#DDCC77' "grey"
  
  # add wgcna module annotation
  row_ha <- rowAnnotation(modules=rownames(x.mat), 
                        #set any annotaiton settings for CH 
                        annotation_legend_param = list(modules = list(
                       	ncol = 2, 
                       	title = "wgcna modules")),
                        #setting annotation cols
                        col=list(modules = c("1"=col.pal[1], '2'=col.pal[2], '3'=col.pal[3],
                                                       '4'=col.pal[4], '6'=col.pal[6], 
                                                       '7'=col.pal[7], '8'=col.pal[8],'9'=col.pal[9], 
                                                       '11'=col.pal[11])) )

 
  hm <- Heatmap(x.mat,
          name='Pearson r',
          right_annotation = row_ha,
          cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(sprintf("%.2f", x.mat[i, j]), x, y, gp = gpar(fontsize = 7)) }
          
          )
  
  BackupAsPDF(draw(hm, column_title=paste0('Module GPCR ', n[[i]], ' Score Correlations')), paste0(n[[i]],'.module.cor.heatmaps.pdf'))

},x=cor.long, n=names(cor.long))

```

write out results first then try regenerate with maximum change from t0

```{r}
# cor table of IUPAR gene scores
cor.tab <- cor.long$IUPAR
cor.table <- dcast(cor.tab, cluster~GProtein, value.var = c("R"))
fwrite(cor.table, './output/cluster.cor.med.log2FC.iupar.csv') 


cor.tab <- cor.long$Predicted
cor.table <- dcast(cor.tab, cluster~GProtein, value.var = c("R"))
fwrite(cor.table, './data/cluster.cor.med.log2FC.predicted.csv') 

cor.tab <- cor.long$Experimental
cor.table <- dcast(cor.tab, cluster~GProtein, value.var = c("R"))
fwrite(cor.table, './data/cluster.cor.med.log2FC.experimental.csv') 
```


Regenerate the correlations using vT0 values (what we used for the clustering) and check for GProtein correlation

```{r}
protQuant <- fread('data/2022_08_17_ProteinLevelData.lib.csv.gz')

group.noi <- protQuant[!GROUP %like% c("DOR|KOR|MOR|MRX2|MRX4|NK1R|NPFFR1|NPY1R|PTGER1|PTGER4|QRFPR|DOR"),unique(GROUP)]

protQuant <- protQuant[!GROUP %in% group.noi, ]
protQuant <- protQuant[!GROUP %like% "MOR_LG94",]

protQuant[, `:=`(timeStr = str_extract(GROUP,'[0-9]{2}$'),
                 receptor = gsub("_[0-9]{2}$", '', protQuant$GROUP)
               )]

setorder(protQuant, receptor, Protein, timeStr)

protQuant[, batchControlled := LogIntensities - median(LogIntensities, na.rm = TRUE), by = .(Protein, receptor, SUBJECT)]
protQuant[, vsTimeZero := batchControlled - mean(batchControlled[timeStr == timeStr[!is.na(batchControlled)][1] ]), by = .(Protein, receptor)]
protQuant[, gene := multiUniprots2multiGenes(Protein)]
```

subset the expression matrix to genes in the clusters and get matrix of max shift vs T0 per cluster,receptor 
max shift an issue due to noise in the clusters... what about max shift per gene in each cluster, then take the median vlaue of that?
maybe ask Ben

```{r}
protQuant[wgcna.clusters, cluster := i.Cluster, on=c(gene = 'GeneID')]

# this takes the max shift; too strongly influenced by outliers; take median instead
prot.mat <- protQuant[!is.na(cluster),] %>% 
  .[, .SD[, vsTimeZero[which.max(abs(vsTimeZero))]], by=.(cluster,receptor)] %>% # extract the max absolute shift vs T0 
  setnames( old=c('cluster', 'receptor', 'V1'), new = c('cluster', 'receptor', 'max.shift')) %>% 
  dcast(receptor~cluster, value.var = 'max.shift') %>% 
  as.matrix(rownames="receptor") %>% 
  t()
  
prot.mat <- protQuant[!is.na(cluster),] %>% 
  .[, .SD[, median(vsTimeZero)], by=.(cluster,receptor)] %>% 
  setnames( old=c('cluster', 'receptor', 'V1'), new = c('cluster', 'receptor', 'median.shift')) %>% 
  dcast(receptor~cluster, value.var = 'median.shift') %>% 
  as.matrix(rownames="receptor") %>% 
  t()
  
```

rename the receptors and reorder to match the scoring matrices

```{r}
colnames(prot.mat) <- c('DOR', 'KOR', 'MOR', 'MRX2', 'MRX4', 'NK1R', 'NPFFR1', 'NPY1R','PTGER1', 'PTGER4','QRFPR')
prot.mat <- prot.mat[, col.order]
```

rerun the correlations with median shift per cluster/receptor
cluster 6 and 9 associated with Gq/G11

```{r}
# correlate t with each
cor.list <- lapply(scores.mat, function(x){
  cor(t(prot.mat), t(x), use= "pairwise.complete.obs")
})

cor.long <- lapply(cor.list, function(x){
  x <- melt(as.data.table(x, keep.rownames = TRUE), id.vars = "rn", variable.name  = "GProtein", value.name = "R")
  setnames(x, old = "rn", new = "cluster")
  return(x)
})
```

plot the enrichment results for the 'interesting' modules

```{r}
mods <- c('1','2','3','4','6','7','8','9','11')


cor.long %>%  names()
names(cor.long) <- c("Experimental","Predicted","IUPAR")

lapply(seq_along(cor.long), function(x,n,i){
  

  # convert to wide
  x.mat <- dcast(x[[i]], str_extract(cluster,'[0-9].*$')~GProtein, value.var = 'R') %>% as.matrix(rownames='cluster')
  x.mat <- x.mat[rownames(x.mat) %in% mods,] #filter out noisy clusters
  
  # annotation cols  for all mods
  col.pal <- c("#D57AC7","#8DDDE1","#A1E4AC","#6EE766","#CEDF5F","#DEC5D2","#85A693","#DBBB7C","#7D9AD6","#D96A6E","#A94FDD")  #"#999933" '#DDCC77' "grey"
  
  # add wgcna module annotation
  row_ha <- rowAnnotation(modules=rownames(x.mat), 
                        #set any annotaiton settings for CH 
                        annotation_legend_param = list(modules = list(
                       	ncol = 2, 
                       	title = "wgcna modules")),
                        #setting annotation cols
                        col=list(modules = c("1"=col.pal[1], '2'=col.pal[2], '3'=col.pal[3],
                                                       '4'=col.pal[4], '6'=col.pal[6], 
                                                       '7'=col.pal[7], '8'=col.pal[8],'9'=col.pal[9], 
                                                       '11'=col.pal[11])) )

 
  hm <- Heatmap(x.mat,
          name='Pearson r',
          right_annotation = row_ha,
          cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(sprintf("%.2f", x.mat[i, j]), x, y, gp = gpar(fontsize = 7)) }
          
          )
  
  draw(hm, column_title=paste0('Module GPCR ', n[[i]], ' Score Correlations'))
  BackupAsPDF(draw(hm, column_title=paste0('Module GPCR ', n[[i]], ' Score Correlations')), paste0(n[[i]],'.module.cor.vsT0.heatmaps.pdf'))

},x=cor.long, n=names(cor.long))

```

Write out median shift per cluster, need to correct for other scoring, think it might actually improve the scores

```{r}
# cor table of IUPAR gene scores
cor.tab <- cor.long$IUPAR
cor.table <- dcast(cor.tab, cluster~GProtein, value.var = c("R"))
fwrite(cor.table, './output/cluster.cor.med.shiftvsT0.iupar.csv') 


cor.tab <- cor.long$Predicted
cor.table <- dcast(cor.tab, cluster~GProtein, value.var = c("R"))
fwrite(cor.table, './data/cluster.cor.med.shiftvsT0.predicted.csv') 

cor.tab <- cor.long$Experimental
cor.table <- dcast(cor.tab, cluster~GProtein, value.var = c("R"))
fwrite(cor.table, './data/cluster.cor.med.shiftvsT0.experimental.csv') 
```

Correlate eigengene meaasurements with the GPCR classification schemes
The moduleEignegene function returns an eigengene matrix; col eigengenes and rows samples


```{r}
mod.EG <- fread('./output/11.receptors.moduleEigengenes.wrownames.csv')
mod.EG %>%  str()

#extract the receptor names and timeStr and calculate avg per receptor
mod.EG[, receptor := gsub('_.*', '', V1)]
mod.EG[receptor == 'MOR.DAMGO', receptor := 'MOR']

```

skewed module eigengene distibution.. use median as summary statistic per receptor

```{r}
as.data.table(mod.EG) %>% 
  melt(., id.vars=c('V1','receptor'), measure.vars=grep('ME', colnames(mod.EG), value = T)) %>% 
  ggplot(., aes(x=value, fill=receptor)) +
  geom_density() +
  facet_wrap(~variable, scales = "free_x") +
  labs(title = "Density Plots of Module Eigengenes per Receptor") #%>% 
  #BackupAsPDF(., 'density.plots.MEperReceptor')
```
Take median ME value per receptor,ME and correlate with GPCR classification 

```{r}
EG.mat <- mod.EG %>% 
  melt(., id.vars=c('V1','receptor'), grep('ME', colnames(mod.EG)) ) %>% 
  setnames(., new=c('sample', 'receptor', 'ME', 'value')) %>% 
  .[, .SD[, median(value)], by=.(receptor, ME)] %>% 
  dcast(., receptor~ME, value.var = 'V1') %>%  as.matrix(rownames='receptor') %>% 
  t()
    
#reorder rows of the matrix  to match
EG.mat <- EG.mat[,col.order]

```
Correlate modEG values with the GPCR classifications

```{r}

# correlate t with each
cor.list <- lapply(scores.mat, function(x){
  cor(t(EG.mat), t(x), use= "pairwise.complete.obs")
})

cor.long <- lapply(cor.list, function(x){
  x <- melt(as.data.table(x, keep.rownames = TRUE), id.vars = "rn", variable.name  = "GProtein", value.name = "R")
  setnames(x, old = "rn", new = "cluster")
  return(x)
})
```


plot the enrichment results for the 'interesting' modules

```{r}
mods <- c('1','2','3','4','6','7','8','9','11')


cor.long %>%  names()
names(cor.long) <- c("Experimental","Predicted","IUPAR")

lapply(seq_along(cor.long), function(x,n,i){
  

  # convert to wide
  x.mat <- dcast(x[[i]], str_extract(cluster,'[0-9].*$')~GProtein, value.var = 'R') %>% as.matrix(rownames='cluster')
  x.mat <- x.mat[rownames(x.mat) %in% mods,] #filter out noisy clusters
  
  # annotation cols  for all mods
  col.pal <- c("#D57AC7","#8DDDE1","#A1E4AC","#6EE766","#CEDF5F","#DEC5D2","#85A693","#DBBB7C","#7D9AD6","#D96A6E","#A94FDD")  #"#999933" '#DDCC77' "grey"
  
  # add wgcna module annotation
  row_ha <- rowAnnotation(modules=rownames(x.mat), 
                        #set any annotaiton settings for CH 
                        annotation_legend_param = list(modules = list(
                       	ncol = 2, 
                       	title = "wgcna modules")),
                        #setting annotation cols
                        col=list(modules = c("1"=col.pal[1], '2'=col.pal[2], '3'=col.pal[3],
                                                       '4'=col.pal[4], '6'=col.pal[6], 
                                                       '7'=col.pal[7], '8'=col.pal[8],'9'=col.pal[9], 
                                                       '11'=col.pal[11])) )

 
  hm <- Heatmap(x.mat,
          name='Pearson r',
          right_annotation = row_ha,
          cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(sprintf("%.2f", x.mat[i, j]), x, y, gp = gpar(fontsize = 7)) }
          
          )
  
  draw(hm, column_title=paste0('Module GPCR ', n[[i]], ' Score Correlations'))
  BackupAsPDF(draw(hm, column_title=paste0('Module GPCR ', n[[i]], ' Score Correlations')), paste0(n[[i]],'.module.cor.moduleEigenGene.heatmaps.pdf'))

},x=cor.long, n=names(cor.long))

```

cor.long

```{r}
cor.long$IUPAR[cluster == 'ME6']

EG.mat['ME6',]

scores.mat$iupar
```




Thoughts
--- 
Need to rerun the wgcna and filter for missing values. Want only proteins present in 90% of data? maybe impute mean values for other members of group then
metric to use? max shift in expression vs t0 per receptor per clust (also try median)



```{r}



# min and max value across all timepoints per gene per receptor
p.quant.clust[, gene.min := min(vsTimeZero), by=.(gene, receptor)]
p.quant.clust[, gene.max := max(vsTimeZero), by=.(gene, receptor)]

# calculate median expression of min and max values ( i think this will be useful for lollipop chart)
p.quant.clust[, cluster.min.median := median(gene.min), by=.(cluster, receptor)]
p.quant.clust[, cluster.max.median := median(gene.max), by=.(cluster, receptor)]

#calculate median value per tp per cluster
p.quant.clust[, cluster.median.vsT0 := median(vsTimeZero), by=.(cluster, receptor, timeStr)]


p.quant.clust[, cluster.med.min := median(vsTimeZero), by=.(cluster, receptor)]
p.quant.clust[, cluster.receptor := paste0(cluster, '.', receptor)]

```
