ls---
title: "AF_scores"
author: "Martin Gordon"
date: "2024-03-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## First pass of the alphafold scores, 
les just quickly plot iPTM and PTM values


packages
```{r}
library(data.table)
library(ggplot2)
library(stringr)
library(magrittr)
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping

source("../../utils/bp_utils/manageScriptData.R")
```



read in the data

```{r}
scores.dt <- fread('./data/032524_allScores.csv')
setnames(scores.dt, new=c('file', 'ptm', 'iptm'))
         
scores.dt[, pair := tstrsplit(file, '[/]', keep=2)]
scores.dt[, pair := sub("SUB", '', pair)]
scores.dt[, model :=  sub("model_","", str_extract(file, 'model_[1-5]'))]

```
5 models produced for 23 of the pairs
Others failed, one of the 3 partially completed, so will rerun and share
```{r}
scores.dt[, .N, by=pair]
```
get gene names 

```{r}
scores.dt[, bait_gene := multiUniprots2multiGenes(bait, species='HUMAN')]
scores.dt[, prey_gene := multiUniprots2multiGenes(prey, species='HUMAN')]
scores.dt[, c('bait', 'prey') := tstrsplit(pair, "__", keep=c(1,2))]

scores.dt[, gene_pair := paste0(bait_gene, "__", prey_gene)]
```


boxplots of the scores

```{r}
scores.dt[, above.threshold := ifelse(iptm > .5, 'yes', 'no')]

g <- ggplot(scores.dt, aes(x=gene_pair, y=iptm, color=above.threshold)) +
  geom_boxplot(alpha=0.4, color='darkgrey') +
  geom_point(alpha=0.8) +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  scale_color_manual(values=c('no'='black', 'yes'='red')) +
  labs(title='AF-multimer Sacsin pair scores') +
  ylim(c(0,1)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90))

g
BackupAsPDF(g, 'AFscores.boxplot', dimensions=c(14,10))
```

What models contribute to the high scores

```{r}
g <- ggplot(scores.dt, aes(x=gene_pair, y=iptm, shape=model, color=above.threshold)) +
  geom_point(alpha=0.8) +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  labs(title='AF-multimer Sacsin pair scores') +
  ylim(c(0,1)) +
  scale_color_manual(values=c('no'='black', 'yes'='red')) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90))

g
BackupAsPDF(g, 'AFscores.boxplot', dimensions=c(14,10))
```
Rerunning:
Q9NZJ4SUB__Q12955
Q9NZJ4SUB__Q03001
Q9NZJ4SUB__O15020 * rerun this one
