---
title: "pLDDT_plots"
author: "Martin Gordon"
date: "2024-04-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## plots of the pLDDT scores from the different subsections of the SACSIN AF run 
Currently formatted as json files, but can convert these to DT in r
```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(jsonlite) # convert json to do
library(stringr)
library(gggenes) # draw gene segements
library(patchwork) # combine plots
library(ComplexHeatmap)
library(circlize)

source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source("../../utils/bp_utils/manageScriptData.R")
```
read in all files in the. directory
```{r}
json.paths <- dir(path='./data/plDDT_jsons', pattern = "*.json", full.names = T)

json.list <- lapply(json.paths, function(x){
  return(fromJSON(x) %>% 
    as.data.table()) })

# add f names
 # only matches b preceded by an a
names(json.list)
names(json.list) <- paste0(str_extract(json.paths, pattern='[A-Za-z0-9_]+(?=_result_model_)'), str_extract(json.paths, "_model[_][1-5]{1}"))
```
Now we want to index each of these tables
```{r}
plDDT.dt <- rbindlist(json.list, idcol='sample')

plDDT.dt[,.N, by=sample]
plDDT.dt[, pos := seq_len(.N), by=sample]

# add the aa residue info to the table
plDDT.dt[, aaResidue := ifelse(sample %like% "SACs_1_2000", pos,
                               ifelse(sample %like% "SACs_1000_3000", pos + 999,
                                      ifelse(sample %like% "SACs_2000_4466", pos + 1999,
                                             ifelse(sample %like% "SACs_domain1", pos,
                                                    ifelse(sample %like% "SACs_domain2", pos + 1470,
                                                           pos + 2922 )))))]
```
Plot of the pLDDT scores individually, then plot all the 

```{r}
plDDT.dt[, subsequence := sub('_model_[1-5]{1}', '', sample)]
plDDT.dt[, model := str_extract(sample, '[1-5]$')]

subseqs <- plDDT.dt$subsequence %>%  unique()

lapply(subseqs, function(x){
  
  min.pos <- plDDT.dt[subsequence == x, min(aaResidue)]
  max.pos <- plDDT.dt[subsequence == x, max(aaResidue)]
  
  g <- ggplot(plDDT.dt[subsequence == x,], aes(x=aaResidue, y=.,color=model)) +
    geom_line(alpha=0.6, size=1) +
    geom_hline(yintercept = c(70,90), linetype=2, alpha=0.6, color='black') +
    stat_summary(aes(group=subsequence), fun.y=mean, geom="line", colour="black") +
    annotate(geom='text',x=min.pos+40, y=72, label='good', color='black') +
    annotate(geom='text',x=min.pos+40, y=92, label='excellent', color='black') +
    scale_y_continuous(breaks=seq(0,110, by=10)) +
    scale_x_continuous(breaks=seq(0, max.pos, by=100)) +
    ggtitle(paste0(x, ' Predicted LDDT residue scores')) +
    theme_bw() +
    guides(alpha=guide_legend(title="mean pLDDT"))
  
  g + ggplot()
  
  ggplot(example_genes, aes(xmin = start, xmax = end, y = molecule)) +
  facet_wrap(~ molecule, scales = "free", ncol = 1) +
  geom_gene_arrow(fill = "white") +
  geom_subgene_arrow(data = example_subgenes,
    aes(xmin = start, xmax = end, y = molecule, fill = gene,
        xsubmin = from, xsubmax = to), color="black", alpha=.7) +
  theme_genes()
  
 BackupAsPDF(g, paste0(x,'.plddt.linechart'), dimensions = c(14,10))
})
```
Lets plot the gene domains (if possible) otherwise we can use geom_segements to highlight

```{r}
col.pal <- randomcoloR::distinctColorPalette(k=10)

# mimic this structure
example_genes %>%  str()
example_subgenes

sacsin.dt <- data.table(molecule=c('SACS','SACS','SACS','SACS','SACS','SACS','SACS','SACS','SACS','SACS'),
                        region=c('Domain1','Domain2','Domain3', 'UBL', 'SRR1', 'SRR2', 'SRR3','XPCB', 'DNAJ', 'HEPN'),
                        start=c(1,1,1,1,1,1,1,1,1,1),
                        end=c(4579,4579,4579,4579,4579,4579,4579,4579,4579,4579),
                        from=c(1, 1471, 2922,1,107,1471,2539,3659,4295,4450),
                        to=c(1471,2922,4466,67,505,1921,2922,3735,4384,4466))


plDDT.dt
lapply(subseqs, function(x){
  
  min.pos <- plDDT.dt[subsequence == x, min(aaResidue)]
  max.pos <- plDDT.dt[subsequence == x, max(aaResidue)]
  
  g <- ggplot(plDDT.dt[subsequence == x,], aes(x=aaResidue, y=plddt,color=model)) +
    geom_line(alpha=0.6, size=1) +
    geom_hline(yintercept = c(70,90), linetype=2, alpha=0.6, color='black') +
    stat_summary(aes(group=subsequence), fun.y=mean, geom="line", colour="black") +
    annotate(geom='text',x=min.pos+40, y=72, label='good', color='black') +
    annotate(geom='text',x=min.pos+40, y=92, label='excellent', color='black') +
    scale_y_continuous(breaks=seq(0,110, by=10)) +
    scale_x_continuous(breaks=seq(0, max.pos, by=100)) +
    ggtitle(paste0(x, ' Predicted LDDT residue scores')) +
    theme_bw() +
    guides(alpha=guide_legend(title="mean pLDDT"))
  
  # I need to modify the anno dt to only extend as far as the region listed
  
  anno.dt <- sacsin.dt[, `:=`(start = min.pos,
                               end = max.pos) ]
  
 # p <- ggplot(anno.dt[region %in% c('Domain1','Domain2','Domain3'),], aes(xmin=start, xmax=end, y=molecule)) +
#    geom_gene_arrow(fill = "white") +
#    geom_subgene_arrow(data = anno.dt[region %in% c('Domain1','Domain2','Domain3'),],
#    aes(xmin = start, xmax = end, y = molecule, fill = region,
#        xsubmin = from, xsubmax = to), color="black", alpha=.7) +
#    scale_fill_brewer(type='qual') +
#    theme_genes()
  
  q <- ggplot(anno.dt[region %in% c('UBL', 'SRR1', 'SRR2', 'SRR3','XPCB', 'DNAJ', 'HEPN'),], aes(xmin=start, xmax=end, y=molecule)) +
    geom_gene_arrow(fill = "white") +
    geom_subgene_arrow(data = anno.dt[region %in% c('UBL', 'SRR1', 'SRR2', 'SRR3','XPCB', 'DNAJ', 'HEPN'),], aes(xmin = start, xmax = end, y = molecule, fill = region, xsubmin = from, xsubmax = to), color="black", alpha=.7) +
    scale_fill_brewer(type='qual', palette=2) +
    theme_genes()
  
  patch <- g/q + plot_layout(heights=c(5,1,1))
  #BackupAsPDF(patch, paste0(x,'.plddt.linechart.genetrack'), dimensions = c(14,10))
})
```
Maybe we want to present this as some kind of matrix
plot all of the residues and do a facet wrap

```{r}
plDDT.dt[subsequence == "SACs_domain2", ][order(aaResidue)]

plDDT.dt[, dataset := ifelse(subsequence %in% c("SACs_1_2000","SACs_1000_3000","SACs_2000_4466"), 'slidingWindow', 'domain')]

g <- ggplot(plDDT.dt, aes(x=aaResidue, y=., color=subsequence)) +
    geom_line(alpha=0.6, size=1) +
    geom_hline(yintercept = c(70,90), linetype=2, alpha=0.6, color='black') +
    stat_summary(aes(group=subsequence), fun.y=mean, geom="line", colour="black") +
    scale_y_continuous(breaks=seq(0,110, by=10)) +
    scale_x_continuous() +
    facet_wrap(~dataset, nrow=2) +
    ggtitle(' Predicted LDDT residue scores') +
    theme_classic() +
    guides(alpha=guide_legend(title="mean pLDDT"))

q <- ggplot(sacsin.dt, 
              aes(xmin=start, xmax=end, y=molecule)) +
    geom_gene_arrow(fill = "white") +
    geom_subgene_arrow(data = sacsin.dt[!region %in% c('Domain1', 'Domain2', 'Domain3'),], aes(xmin = start, xmax = end, y = molecule, fill = region, xsubmin = from, xsubmax = to), color="black", alpha=.7) +
    scale_fill_brewer(type='qual', palette=2) +
    theme_genes() +
     theme(legend.title = element_text( size=8), legend.text=element_text(size=6), legend.position="bottom")
  
patch <- g/q + plot_layout(heights=c(8,1))
BackupAsPDF(patch, paste0('plddt.linechart.genetrack.combined'), dimensions = c(14,10))

```

Adding another annotation

```{r}
sacsin.dt <- data.table(molecule=c('SACS','SACS','SACS','SACS','SACS','SACS','SACS','SACS','SACS','SACS',
                                   'SACS','SACS','SACS','SACS','SACS'),
                        region=c('Domain1','Domain2','Domain3', 'UBL', 'SRR1', 'SRR2', 'SRR3','XPCB', 'DNAJ', 'HEPN',
                                 'SIRPT1','SIRPT2','SIRPT3','Ubl','Hepn'),
                        start=c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1),
                        end=c(4579,4579,4579,4579,4579,4579,4579,4579,4579,4579,4579,4579,4579,4579,4579),
                        from=c(1, 1471, 2922,1,107,1471,2539,3659,4295,4450,84,1444,2512,1,4422),
                        to=c(1471,2922,4466,67,505,1921,2922,3735,4384,4466,1374,2443,4282,84,4579))

# define the datasets
sacsin.dt[,dataset := ifelse(region %in% c('Domain1','Domain2', 'Domain3'), 'domain',
                             ifelse(region %in% c("UBL","SRR1", 'SRR2', 'SRR3','XPCB', 'DNAJ', 'HEPN'), 
                                    'slidingWindow', 'Gentil_2018'))]
```

```{r}

g <- ggplot(plDDT.dt, aes(x=aaResidue, y=., color=subsequence)) +
    geom_line(alpha=0.6, size=1) +
    geom_hline(yintercept = c(70,90), linetype=2, alpha=0.6, color='black') +
    stat_summary(aes(group=subsequence), fun.y=mean, geom="line", colour="black") +
    scale_y_continuous(breaks=seq(0,110, by=10)) +
    scale_x_continuous() +
    facet_wrap(~dataset, nrow=2) +
    ggtitle(' Predicted LDDT residue scores') +
    theme_classic() +
    guides(alpha=guide_legend(title="mean pLDDT"))

q <- ggplot(sacsin.dt, 
              aes(xmin=start, xmax=end, y=dataset)) +
    geom_gene_arrow(fill = "white") +
    geom_subgene_arrow(data = sacsin.dt[region %in% c("UBL","SRR1", 'SRR2', 'SRR3','XPCB', 'DNAJ', 'HEPN'),], aes(xmin = start, xmax = end, y = dataset, fill = region, xsubmin = from, xsubmax = to), color="black", alpha=.7) +
    labs(y='Sacsin domains') +
    scale_fill_brewer(type='qual', palette=2) +
    theme_genes() +
     theme(legend.title = element_text( size=8), legend.text=element_text(size=6))


p <- ggplot() +
    geom_subgene_arrow(data = sacsin.dt[region %in% c('SIRPT1','SIRPT2','SIRPT3','Ubl','Hepn'),], aes(xmin = start, xmax = end, y = dataset, fill = region, xsubmin = from, xsubmax = to), color="black", alpha=.7) +
   geom_subgene_arrow(data = sacsin.dt[region %in% c("UBL","SRR1", 'SRR2', 'SRR3','XPCB', 'DNAJ', 'HEPN'),], aes(xmin = start, xmax = end, y = dataset, fill = region, xsubmin = from, xsubmax = to), color="black", alpha=.7) +
    labs(y='Sacsin annotations') +
    scale_fill_brewer(type='qual', palette=3) +
    theme_genes() +
     theme(legend.title = element_text( size=8), legend.text=element_text(size=6))
  
patch <- g/p + plot_layout(heights=c(8,1,1))
patch
BackupAsPDF(patch, paste0('plddt.linechart.genetrack.combined'), dimensions = c(14,10))

```

Lets do a heatmap of the genes; only color in sections that are covered

```{r}
setnames(plDDT.dt, old='.', new='plddt')

plddt.mat <- dcast(plDDT.dt, aaResidue~sample, value.var ='plddt') %>%  
  as.matrix(rownames='aaResidue')

colnames(plddt.mat)

col.order <- c("SACs_domain1_model_1","SACs_domain1_model_2","SACs_domain1_model_3", "SACs_domain1_model_4","SACs_domain1_model_5",
              "SACs_1_2000_model_1","SACs_1_2000_model_2","SACs_1_2000_model_3","SACs_1_2000_model_4","SACs_1_2000_model_5",
              "SACs_1000_3000_model_1","SACs_1000_3000_model_2","SACs_1000_3000_model_3","SACs_1000_3000_model_4","SACs_1000_3000_model_5",
              "SACs_domain2_model_1","SACs_domain2_model_2","SACs_domain2_model_3","SACs_domain2_model_4","SACs_domain2_model_5",
               "SACs_2000_4466_model_1","SACs_2000_4466_model_2","SACs_2000_4466_model_3","SACs_2000_4466_model_4","SACs_2000_4466_model_5",
               "SACs_domain3_model_1","SACs_domain3_model_2","SACs_domain3_model_3","SACs_domain3_model_4","SACs_domain3_model_5")

plddt.mat <- plddt.mat[, col.order]


# create heatmap annotations of the different domains previ
sacsin.dt %>%  View()

# first annotation 
sacsin.anno <-rep('',4466)
sacsin.anno[1:67] <- 'UBL'
sacsin.anno[107:505] <- 'SRR1'
sacsin.anno[1471:1921] <- 'SRR2'
sacsin.anno[2539:2922] <- 'SRR3'
sacsin.anno[3659:3735] <- 'XPCB'
sacsin.anno[4295:4384] <- 'DNAJ'
sacsin.anno[4450:4466] <- 'HEPN'

#Gentil 
gentil.anno <-rep('',4466)
gentil.anno[1:84] <- 'UBL'
gentil.anno[84:1374] <- 'SIRPT1'
gentil.anno[1444:2443] <- 'SIRPT2'
gentil.anno[2512:4282] <- 'SIRPT3'
gentil.anno[4422:4466] <- 'HEPN'


hm <- Heatmap((plddt.mat), 
        name = 'pLDDT',
        row_names_side = 'left',
        col=colorRamp2(breaks=c(20,60,70,80,90), colors = c('yellow','orange', 'darkorange', 'red', 'darkred')),
        column_names_gp = gpar(fontsize=6),
        row_title = 'aaResidue (descreasing)',
        show_row_names = F,
        column_title_gp = gpar(fontsize=8),
     #   column_split = gsub("_model_[0-9]{1}$",'', colnames(plddt.mat)),
        row_names_gp = gpar(fontsize=2),
        cluster_rows=F, 
        na_col = 'white',
        cluster_columns = F) +
  
  Heatmap((sacsin.anno), 
          name='Domains',
          col=structure(col.pal, names=unique(sacsin.anno)),
          ) +
  
    Heatmap((gentil.anno), 
          name='Gentil et al. 2018',
          col=structure(col.pal, names=unique(gentil.anno)),
          )
hm
BackupAsPDF(hm, 'plddt.residues.noColSplit.heatmap', dimensions = c(12,14))
```
Redo the plot with just the mean line and color according to the avg plddt score

```{r}
lapply(subseqs, function(x){
  
  min.pos <- plDDT.dt[subsequence == x, min(aaResidue)]
  max.pos <- plDDT.dt[subsequence == x, max(aaResidue)]
  
  summary.dt <- plDDT.dt[subsequence == x, .(plddt, model, mean.plddt = mean(plddt)), by=aaResidue]
  
  g <- ggplot(summary.dt, aes(x=aaResidue, y=mean.plddt, color=mean.plddt)) +
    geom_line(alpha=0.6, size=1) +
    geom_hline(yintercept = c(70,90), linetype=2, alpha=0.6) +
    annotate(geom='text',x=min.pos+300, y=32, label=paste0('Best model mean pLDDT: ', summary.dt[model ==1,round(mean(plddt),1)]), size=3) +
    #annotate(geom='text',x=min.pos+40, y=92, label='excellent', color='black') +
    scale_color_gradientn("plDDT", limits = c(0,100), colors = c(red = "#FE0012",
                                                      orange = "#FFA42B",
                                                      yellow = "#FFFD42","#FFFD42",
                                                      palegreen = "palegreen2",
                                                      blue = "lightblue","lightblue",#"#6097E8",
                                                      darkBlue = "#001DF9"),
                        values = c(0, .5, 0.7,0.75,0.8,0.85, 0.9, 1.0)) +
    scale_y_continuous(breaks=seq(0,140, by=10)) +
    scale_x_continuous(breaks=seq(0, max.pos, by=100)) +
    ggtitle(paste0(x, ' Predicted LDDT residue scores')) +
    ylab('mean pLDDT') +
    xlab('Amino acid residue') +
    theme_bw() +
    guides(alpha=guide_legend(title="mean pLDDT"))
  
  # I need to modify the anno dt to only extend as far as the region listed
  anno.dt <- sacsin.dt[, `:=`(start = min.pos,
                               end = max.pos)]
  
  q <- ggplot(anno.dt[region %in% c('UBL', 'SRR1', 'SRR2', 'SRR3','XPCB', 'DNAJ', 'HEPN'),], aes(xmin=start, xmax=end, y=molecule)) +
    geom_gene_arrow(fill = "white") +
    geom_subgene_arrow(data = anno.dt[region %in% c('UBL', 'SRR1', 'SRR2', 'SRR3','XPCB', 'DNAJ', 'HEPN'),], aes(xmin = start, xmax = end, y = molecule, fill = region, xsubmin = from, xsubmax = to), color="black", alpha=.7) +
    scale_fill_brewer(type='qual', palette=2) +
    theme_genes()
  
  patch <- g/q + plot_layout(heights=c(5,1,1))
  BackupAsPDF(patch, paste0(x,'.mean.plddt.linechart.genetrack'), dimensions = c(14,10))
})

```
Now plot the overlaps of all the samples again

```{r}
g <- ggplot(plDDT.dt[, .(plddt, model, mean.plddt = mean(plddt), dataset), by=.(aaResidue,subsequence)], aes(x=aaResidue, y=mean.plddt, color=subsequence, alpha=0.7)) +
    geom_line(alpha=0.6, size=1) +
    geom_hline(yintercept = c(70,90), linetype=2, alpha=0.6, color='black') +
    scale_y_continuous(breaks=seq(0,110, by=10)) +
    scale_x_continuous() +
    xlab('Amino acid residue') +
    ylab('mean pLDDT') +
    facet_wrap(~dataset, nrow=2) +
    ggtitle('Predicted LDDT residue scores') +
    theme_classic() +
    guides(alpha=guide_legend(title="mean pLDDT"))

q <- ggplot(sacsin.dt[dataset == 'slidingWindow',], 
              aes(xmin=start, xmax=end, y=molecule)) +
    geom_gene_arrow(fill = "white") +
    geom_subgene_arrow(data =sacsin.dt[dataset == 'slidingWindow',], aes(xmin = 1, xmax = 4466, y = molecule, fill = region, xsubmin = from, xsubmax = to), color="black", alpha=.7) +
    scale_fill_brewer(type='qual', palette=2) +
    theme_genes() +
     theme(legend.title = element_text( size=8), legend.text=element_text(size=6))

p <- ggplot(sacsin.dt[dataset == 'domain',], 
              aes(xmin=start, xmax=end, y=molecule)) +
    geom_gene_arrow(fill = "white") +
    geom_subgene_arrow(data =sacsin.dt[dataset == 'domain',], aes(xmin = 1, xmax = 4466, y = molecule, fill = region, xsubmin = from, xsubmax = to), color="black", alpha=.7) +
    scale_fill_brewer(type='qual', palette=2) +
    theme_genes() +
     theme(legend.title = element_text( size=8), legend.text=element_text(size=6))
  
patch <- g/q + plot_layout(heights=c(8,1))
patch


BackupAsPDF(patch, paste0('plddt.linechart.genetrack.subseqColor.combined'), dimensions = c(14,10))
```
Now color by pLDDT score

```{r}
g <- ggplot(plDDT.dt[, .(plddt, model, mean.plddt = mean(plddt), dataset), by=.(aaResidue,subsequence)], aes(x=aaResidue, y=mean.plddt, color=mean.plddt, alpha=0.7)) +
    geom_line(alpha=0.6, size=1) +
    geom_hline(yintercept = c(70,90), linetype=2, alpha=0.6, color='black') +
    scale_color_gradientn("plDDT", limits = c(0,100), colors = c(red = "#FE0012",
                                                      orange = "#FFA42B",
                                                      yellow = "#FFFD42","#FFFD42",
                                                      palegreen = "palegreen2",
                                                      blue = "lightblue","lightblue",#"#6097E8",
                                                      darkBlue = "#001DF9"),
                        values = c(0, .5, 0.7,0.75,0.8,0.85, 0.9, 1.0)) +
    scale_y_continuous(breaks=seq(0,110, by=10)) +
    scale_x_continuous() +
    xlab('Amino acid residue') +
    ylab('mean pLDDT') +
    facet_wrap(~dataset, nrow=2) +
    ggtitle('Predicted LDDT residue scores') +
    theme_classic() +
    guides(alpha=guide_legend(title="mean pLDDT"))

q <- ggplot(sacsin.dt[dataset == 'slidingWindow',], 
              aes(xmin=start, xmax=end, y=molecule)) +
    geom_gene_arrow(fill = "white") +
    geom_subgene_arrow(data =sacsin.dt[dataset == 'slidingWindow',], aes(xmin = 1, xmax = 4466, y = molecule, fill = region, xsubmin = from, xsubmax = to), color="black", alpha=.7) +
    scale_fill_brewer(type='qual', palette=2) +
    theme_genes() +
     theme(legend.title = element_text( size=8), legend.text=element_text(size=6))

p <- ggplot(sacsin.dt[dataset == 'domain',], 
              aes(xmin=start, xmax=end, y=molecule)) +
    geom_gene_arrow(fill = "white") +
    geom_subgene_arrow(data =sacsin.dt[dataset == 'domain',], aes(xmin = 1, xmax = 4466, y = molecule, fill = region, xsubmin = from, xsubmax = to), color="black", alpha=.7) +
    scale_fill_brewer(type='qual', palette=2) +
    theme_genes() +
     theme(legend.title = element_text( size=8), legend.text=element_text(size=6))
  
patch <- g/q/p + plot_layout(heights=c(8,1,1))
patch


BackupAsPDF(patch, paste0('plddt.linechart.genetrack.pLDDTColor.combined'), dimensions = c(14,10))


test <- scale_fill_brewer(type='qual', palette=2)

  
library(scales)
g <- brewer_pal(type="qual", palette=2)

```

getting the ggplot colors to color the AF structures in chimeraX

```{r}
library(scales)
g <- brewer_pal(type="qual", palette=2)

```

