---
title: "NP06_contrastofcontrasts"
author: "Martin Gordon"
date: "2023-08-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Running contrast of contrast analysis

#Two way ANOVA: assess effect of two predictor vs one (continuous) outcome variable

Here we are investigating how compound 101 treatment (blocks FBXW11) x NPFF treatment affect expression per gene
3 variables of interest, BT CM, NP, but not comparing BT groups
Samples are treated with compound 101 a FBXW11 blocker and then treated +/- NPFF (00 vs 30)
Must also control for batch effects
BT is the proteomose inhibitor

# A two-way ANOVA test reveals the results of two independent variables on a dependent variable. ANOVA test results can then be used in an F-test, a statistical test used to determine whether two populations with normal distributions share variances or a standard deviation, on the significance of the regression formula overall.

```{r}
library (data.table)
library (ComplexHeatmap)
library (ggplot2)
library (emmeans)
library(stringr)

rotate.x.axis.text <- theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")

WriteSessionInfo()
WriteInstalledPackages()
```

Prepare the data 
```{r}
# including the dropped sample for now
protQuant <- fread ("/Users/martingordon/Documents/projects//071823_QLi_NPo6/NPO6.all.samples.protein.lvl.quant.csv.gz")


# parse GROUP names to define the actual columns that the linear models will work on
# this work depends on how your columsn are defined
protQuant[, np.status :=  as.factor(str_extract(GROUP,"NP[0-9]{2}"))]
protQuant[, cm.status  := as.factor(str_extract(GROUP,"CM[0-9]{2}"))]
protQuant[, bt.status  := as.factor(str_extract(GROUP,"BT[0-9]{2}"))]
protQuant[, batch      := as.factor(SUBJECT)]

```

First perform an three-way ANOVA
(cm status as is does not seem to have a significant effect on expression)
there is a significant interaction between cm status and np status
```{r}
model <- aov(LogIntensities ~ batch + np.status + cm.status +np.status:cm.status, data=protQuant)
summary(model)
```
Analyse the BT00 and BT02 separately as we know this data is quite different

```{r}
protQuant.bt00 <- protQuant[bt.status == 'BT00',]
protQuant.bt02 <- protQuant[bt.status == 'BT02',]

protQuant.list <- list(BT00 = protQuant.bt00,
                       BT02 = protQuant.bt02)

```

contrast formula:

```{r}


# linearModelsAllProteins can this function which computes statistic for a contrast of two different log2FC
#emmeans.contrastOfContrasts <- function (l, factorFormula = ~status|ant){  # how does the treatment effect change per antagonist level? basically compare treatment effect within each antagonist level
emmeans.contrastOfContrasts <- function (l, factorFormula = ~np.status|cm.status){  # how does the NPFF treatment effect change with CM101 treatment status
  emm <- emmeans::emmeans(l, factorFormula)
  contrast1 <- pairs(emm)
  contrast2 <- pairs(contrast1, by = NULL, adjust = "none")
  return (as.data.table(contrast2))
}

# postProcessFunction = emmeans.contrastOfContrasts tells it to use the function above after doing anova.
# it is optional, adds significant time to the computation (~5x), but is nice to
# put a magnitude and p value on specific differences of differences
#
# cl = 6 parameter says to use 6 processes.  Set to NULL if multiprocess is failing, or higher or lower as you wish.
# Numbers near the number of processors available in your machine is probably counterproductive.

lm.out <- lapply(protQuant.list, function(x){
  
  # Numbers near the number of processors available in your machine is probably counterproductive.
  x.lm <- linearModelsAllProteins(x, 
                                  formulaList = list(full   = LogIntensities~batch + np.status + cm.status + cm.status:np.status), # full model; 4 way ANOV
                                  splitColumn = "Protein", postProcessFunction = emmeans.contrastOfContrasts, cl = 6) |> suppressWarnings()
  
  
  return(x.lm)
  
})


lm.out$BT00$anova
```


# inspect anova p values per term
V strong batch effect in the data and little evidence of interaction between NP and CM
```{r}

lapply(seq_along(lm.out), function(x,n,i){

  dt <- x[[i]][['anova']]
  str(dt)

 g <- ggplot(dt, aes(x=dt$p.value)) +
   geom_histogram(boundary=0.0, binwidth = 0.005) +
   facet_wrap(~dt$term)
 
 BackupAsPDF(g, paste0(n[[i]], '.aov.term.pval.historgrams'))
 
},x=lm.out, n=names(lm.out))


```
Looks like there is very few significant hits here except for batch. Zooming in at the 0-0.05 range this also appears to be the case

```{r}
lapply(seq_along(lm.out), function(x,n,i){

  dt <- x[[i]][['anova']]
  str(dt)

 g <- ggplot(dt, aes(x=dt$p.value)) +
   geom_histogram(boundary=0.0) +
   xlim(0, 0.05) +
   facet_wrap(~dt$term) 
   
 
 BackupAsPDF(g, paste0(n[[i]], '.aov.term.pval0-pval0.05.historgrams'))
 
},x=lm.out, n=names(lm.out))

```


## adjust pvalues 

```{r}
lapply(lm.out, function(x){
  
  x[['anova']][, fdr := p.adjust(p.value, method = "BH"), by= term]
  
})

```
Write out files to dir
```{r}
setwd("/Users/martingordon/Documents/projects/071823_QLi_Npo6")

lapply(seq_along(lm.out), function(x,n,i){
  
  fwrite(x[[i]][['anova']], paste0(n[[i]], '.anova.contrasts.csv.gz'))
  
},x=lm.out, n=names(lm.out))
```


# CM101 treatment effects
No significant results at FDR < 0.05
Take genes significant at pval < 0.001 

```{r}
setwd("/Users/martingordon/Documents/projects/071823_QLi_Npo6")

#play with thresholds here
lapply(seq_along(lm.out), function(x,n,i){
  dt <- x[[i]][['anova']]
  
  sigProtein <- dt[term %in% c( "cm.status") & p.value < 0.001,] %>% 
   #  .[order(p.value), head(.SD, 100), .SDcols = c('p.value', 'Protein')] %>% 
    .[,unique(Protein)]

  # filter just to our data
   prot.mat <- dcast (protQuant[Protein %in% sigProtein & bt.status == n[[i]], ], Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "Protein")


  rownames(prot.mat) <- multiUniprotSites2multiGeneSites(rownames(prot.mat), species = "HUMAN")
  
  prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))
  
  hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'cm101 log2 Ints/Median',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (p.val < 0.001)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("NP00", colnames(prot.mat)), "NP00", "NP30"),
                                ifelse(grepl("CM00", colnames(prot.mat)), "CM00", "CM15")))
  
 BackupAsPDF(hm, paste0(n[[i]], '.cm101.pval<0.001.heatmap'), dimensions = c(8,7))
  
},x=lm.out, n=names(lm.out))

```
# NPFF effects

```{r}
setwd("/Users/martingordon/Documents/projects/071823_QLi_Npo6")

#play with thresholds here
lapply(seq_along(lm.out), function(x,n,i){
  dt <- x[[i]][['anova']]
  
  sigProtein <- dt[term %in% c( "np.status") & p.value < 0.001,] %>% 
   #  .[order(p.value), head(.SD, 100), .SDcols = c('p.value', 'Protein')] %>% 
    .[,unique(Protein)]
  
  print(sigProtein)

  # filter just to our data
   prot.mat <- dcast (protQuant[Protein %in% sigProtein & bt.status == n[[i]], ], Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "Protein")


  rownames(prot.mat) <- multiUniprotSites2multiGeneSites(rownames(prot.mat), species = "HUMAN")
  
  prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))
  
  hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'npff log2 Ints/Median',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (p.val < 0.001)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list (ifelse(grepl("NP00", colnames(prot.mat)), "NP00", "NP30"),
                                ifelse(grepl("CM00", colnames(prot.mat)), "CM00", "CM15")))
  
 BackupAsPDF(hm, paste0(n[[i]], '.npff.pval<0.001.heatmap'), dimensions = c(8,9))
  
},x=lm.out, n=names(lm.out))
```
## interaction effects
(different response to NPFF treatment, given CM101 treatment status)



```{r}

setwd("/Users/martingordon/Documents/projects/071823_QLi_Npo6")

#play with thresholds here
lapply(seq_along(lm.out), function(x,n,i){
  dt <- x[[i]][['anova']]
  
  sigProtein <- dt[term %in% c( "np.status:cm.status") & p.value < 0.001,] %>% 
   #  .[order(p.value), head(.SD, 100), .SDcols = c('p.value', 'Protein')] %>% 
    .[,unique(Protein)]
  
  print(sigProtein)

  # filter just to our data
   prot.mat <- dcast (protQuant[Protein %in% sigProtein & bt.status == n[[i]], ], Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "Protein")


  rownames(prot.mat) <- multiUniprotSites2multiGeneSites(rownames(prot.mat), species = "HUMAN")
  
  prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))
  
  hm <- Heatmap(prot.mat, cluster_columns = FALSE, 
                name = 'npff:cm101 Ints/Median',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (p.val < 0.001)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 8),
              column_split = list(ifelse(grepl("NP00", colnames(prot.mat)), "NP00", "NP30"),
                                ifelse(grepl("CM00", colnames(prot.mat)), "CM00", "CM15")))
  
 BackupAsPDF(hm, paste0(n[[i]], '.npff:cm101.pval<0.001.heatmap'), dimensions = c(8,9))
  
},x=lm.out, n=names(lm.out))
```
## interaction effects observed in contrast of contrasts

Value here is we can require a magnitude of difference. 
Signs are confusing here because the loss of a negative effect will show as  a positive effect, and vice versa. 

```{r, fig.width = 8, fig.height = 8}

#play with thresholds here
lapply(seq_along(lm.out), function(x,n,i){
  dt <- x[[i]][['postProcess']]
  head(dt)
  
   sigProtein <- dt[contrast == "(NP00 - NP30 CM00) - (NP00 - NP30 CM15)" & p.value < 0.001 & abs(estimate) > 1, Protein]# %>% 
  

  # filter just to our data
  prot.mat <- dcast (protQuant[Protein %in% sigProtein & bt.status == n[[i]], ], Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "Protein")


  rownames(prot.mat) <- multiUniprotSites2multiGeneSites(rownames(prot.mat), species = "HUMAN")
  
  prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))

  hm <- Heatmap(prot.mat, cluster_columns = FALSE,
                name = 'npff:cm101 Ints/Median',
            #  column_title = paste0(n[[i]], ' CM101 treatment effects'),
              row_title = sprintf('%s Proteins (p.val < 0.001)', nrow(prot.mat)),
              row_names_gp = gpar(fontsize = 7),
              column_names_gp = gpar(fontsize = 8),
              column_split = list(
                                ifelse(grepl("CM00", colnames(prot.mat)), "CM-", "CM+")))

 BackupAsPDF(hm, paste0(n[[i]], '.npff:cm101.contrastofcontrasts.pval<0.001.heatmap'), dimensions = c(8,9))
hm
},x=lm.out, n=names(lm.out))


lm.out$postProcess
#play with thresholds here
# there weren't many hits by fdr, so use a strong p.value
sigProtein <- lm.out$postProcess[ p.value < 0.001 & abs(estimate) > 0.5, Protein]

prot.mat <- dcast (protQuant[Protein %in% sigProtein], Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "Protein")
rownames(prot.mat) <- multiUniprotSites2multiGeneSites(rownames(prot.mat), species = "HUMAN")

prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))


Heatmap (prot.mat, cluster_columns = FALSE, 
         column_split = list (column_split = list (str_extract(colnames(prot.mat),"AT[0-9]{2}"),
                              str_extract(colnames(prot.mat),"PGE[0-9]{2}"))))

View(lm.out$BT00$postProcess)
```