---
title: "AF3_multimer_plots"
author: "Martin Gordon"
date: "2024-05-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## AF3 multimer SACSIN runs

Look at the output of AF3 multimer runs and plot the iPTM for each of the 5 models

```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(jsonlite) # convert json to do
library(stringr)
library(gggenes) # draw gene segements
library(patchwork) # combine plots
library(ComplexHeatmap)
library(circlize)
#library(rjson) # for reading in the R json files
library(jsonlite)
library(RColorBrewer)

source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source("../../utils/bp_utils/manageScriptData.R")
```

Read in the data for each of the models

As these datasets are very large, we will need to delete these after each of the runs
```{r}
getwd()
models <- dir(path='/Users/martingordon/Documents/projects/042224_GJang_ARSACS-AF/output', pattern="*sacs_vdac3_full_data_[01234].json", full.names = T, recursive = T)
models
af.mods <- lapply(models, function(x){jsonlite::fromJSON(x) })
names(af.mods) <- str_extract(models, 'sacs_[a-z0-9]+_full_data_[0-9]')
```

heatmap of pae scores

```{r}
col.pal <- brewer.pal(8, name='Dark2')

lapply(seq_along(af.mods[c(1:2)]), function(x,n,i){
  
  
  # first annotation  
  sacsin.anno <-rep('', length(x[[i]]$token_res_ids))
  sacsin.anno[1:67] <- 'UBL'
  sacsin.anno[107:505] <- 'SRR1'
  sacsin.anno[1471:1921] <- 'SRR2'
  sacsin.anno[2539:2922] <- 'SRR3'
  sacsin.anno[3659:3735] <- 'XPCB'
  sacsin.anno[4295:4384] <- 'DNAJ'
  sacsin.anno[4450:4466] <- 'HEPN'

  # seconda annotation
  chain.anno <- x[[i]]$token_chain_ids
  
  
  # set up matrix
  mat <- x[[i]]$pae
  rownames(mat) <- x[[i]]$token_res_ids
  colnames(mat) <- x[[i]]$token_res_ids

  #simplify dim names; if not in 
  rownames(mat) <- ifelse(rownames(mat) %in% seq(1, nrow(mat), by=100), rownames(mat), '')
  colnames(mat) <- ifelse(rownames(mat) %in% seq(1, nrow(mat), by=100), rownames(mat), '')
  
  hm1 <- Heatmap(mat, 
                cluster_rows = FALSE, 
                cluster_columns = FALSE, 
                column_title = paste0('model ', gsub('full_data_', '', n[[i]])),
                show_row_names = T, 
                show_column_names = T,
                name='Predicted Aligned Error (Angstroms)')
 
   hm2 <- Heatmap(sacsin.anno, 
          name='Domains',
          col=structure(col.pal, names=unique(sacsin.anno)),
          width=unit(0.5, 'cm'))
   
   hm3 <- Heatmap(chain.anno, 
          name='Chain IDs',
          col=structure(col.pal, names=unique(chain.anno)),
          width=unit(0.5, 'cm'))
   
  BackupAsPDF(draw(hm3+hm2+hm1), paste0(n[[i]], '.pae.heatmap'), dimensions=c(14,14))
  
},x=af.mods, n=names(af.mods))
```

Read the summary results and plot the iPTM scores per model

```{r}
summary <- dir(path='/Users/martingordon/Documents/projects/042224_GJang_ARSACS-AF/output', pattern="*_summary_confidences_[01234].json", full.names = T, recursive = T)

summary.dt <- lapply(summary, function(x){
  # read in all info 
  json.dt <- jsonlite::fromJSON(x)
  # just take the iPTM scores
  return(json.dt['iptm'])
  })

names(summary.dt) <- str_extract(summary, 'sacs_[a-z0-9]+_summary_confidences_[01234]')
names(summary.dt) <- gsub('summary_confidences_', '', names(summary.dt))
names(summary.dt)

summary.dt <- rbindlist(summary.dt, idcol='model')

summary.dt[, `:=`(pair = gsub('_[0-9]$','', model),
                  modelRanking = str_extract(model, '[0-4]$'))]
summary.dt[, c('bait','prey') := tstrsplit(model, '_', keep=c(1,2))]
```
plot the output

```{r}
g <- ggplot(summary.dt, aes(x=pair, y=iptm, color=modelRanking)) +
  geom_point(alpha=0.4) +
  geom_hline(yintercept=c(0.5), linetype=c(2), alpha=0.5) +
  #annotate(geom='text', label='excellent prediction', x=1.5, y=0.82, size=3) +
  annotate(geom='text', label='iPTM threshold', x=1.5, y=0.51, size=4) +
  ylab('iPTM') +
  xlab('Interacting Pair') +
  ggtitle('Accuracy of the predicted relative positions of complex subunits') +
  scale_color_brewer(type='qual', palette=2) +
  theme_bw()

BackupAsPDF(g, 'SacsinInteractor.iPTMscores.dotplot')
```
As these datasets are very large, we will need to delete these after each of the runs
Instead, loop through each pair, generate the PAE plots for the top 2 ranked models
```{r}
models <- dir(path='/Users/martingordon/Documents/projects/042224_GJang_ARSACS-AF/output', pattern="*_full_data_0.json", full.names = T, recursive = T)
pairs <- unique(str_extract(models, 'sacs_[a-z0-9]+'))

af.mods <- lapply(models, function(x){jsonlite::fromJSON(x) })
names(af.mods) <- str_extract(models, 'sacs_[a-z0-9]+')
```

heatmap of pae scores

```{r}
col.pal <- brewer.pal(8, name='Dark2')
col.pal


lapply(seq_along(af.mods), function(x,n,i){
  
  
  # first annotation  
  sacsin.anno <-rep('', length(x[[i]]$token_res_ids))
  sacsin.anno[1:67] <- 'UBL'
  sacsin.anno[107:505] <- 'SRR1'
  sacsin.anno[1471:1921] <- 'SRR2'
  sacsin.anno[2539:2922] <- 'SRR3'
  sacsin.anno[3659:3735] <- 'XPCB'
  sacsin.anno[4295:4384] <- 'DNAJ'
  sacsin.anno[4450:4466] <- 'HEPN'

  # seconda annotation
  chain.anno <- x[[i]]$token_chain_ids
  
  
  # set up matrix
  mat <- x[[i]]$pae
  rownames(mat) <- x[[i]]$token_res_ids
  colnames(mat) <- x[[i]]$token_res_ids

  #simplify dim names; if not in 
  rownames(mat) <- ifelse(rownames(mat) %in% seq(1, nrow(mat), by=100), rownames(mat), '')
  colnames(mat) <- ifelse(rownames(mat) %in% seq(1, nrow(mat), by=100), rownames(mat), '')
  
  hm1 <- Heatmap(mat, 
                cluster_rows = FALSE, 
                cluster_columns = FALSE, 
                column_title = paste0('best model ',n[[i]]),
                show_row_names = T, 
                show_column_names = T,
                name='Predicted Aligned Error (Angstroms)')
 
   hm2 <- Heatmap(sacsin.anno, 
          name='Domains',
          col=structure(col.pal, names=unique(sacsin.anno)),
          width=unit(0.5, 'cm'))
   
   hm3 <- Heatmap(chain.anno, 
          name='Chain IDs',
          col=structure(col.pal, names=unique(chain.anno)),
          width=unit(0.5, 'cm'))
   
  BackupAsPDF(draw(hm3+hm2+hm1), paste0(n[[i]], '.pae.heatmap'), dimensions=c(14,14))
  
},x=af.mods, n=names(af.mods))
```



```{r}
col.pal
```

