---
title: "SACSdomains_AF"
author: "Martin Gordon"
date: "2024-04-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introduction
---
Out of 208 jobs in total, we have 5 models produced for 192 inteactions
Seems most of these are not scoring well, although there are a handful of high scorers


packages
```{r}
library(data.table)
library(ggplot2)
library(stringr)
library(magrittr)
library(ComplexHeatmap)
library(circlize)

source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source("../../utils/bp_utils/manageScriptData.R")
```

read in the data

```{r}
scores.dt <- fread('./data/allScores.csv')
setnames(scores.dt, new=c('file', 'ptm', 'iptm'))

scores.dt[, pair := tstrsplit(file, '[/]', keep=2)]
scores.dt[pair %like% "HEPN", pair := sub('HEPN_', 'HEPN-', pair)]

scores.dt[, model :=  sub("model_","", str_extract(file, 'model_[1-5]'))]
scores.dt[, c('Bait', 'Prey') :=  tstrsplit(pair, '__', keep=c(1,2))]


scores.dt[, c('Protein', 'Domain') :=  tstrsplit(Bait, '_', keep=c(1,2))]
# fix the domain name
scores.dt[is.na(Domain), Domain := 'fullProtein']


scores.dt$Domain %>%  unique()
```

Get gene names 
```{r}
scores.dt[, Bait_gene := multiUniprots2multiGenes('Q9NZJ4', species='HUMAN')]
scores.dt[, Prey_gene := multiUniprots2multiGenes(Prey, species='HUMAN')]


scores.dt$Domain %>%  unique()

scores.dt[, gene_pair := ifelse(Domain == 'fullProtein', paste0(Bait_gene, '__', Prey_gene), 
                                paste0(Bait_gene, '_', Domain, "__", Prey_gene)) ]

scores.dt[, mean.iptm := mean(iptm), by=pair]
scores.dt[, above.threshold := ifelse(iptm > .5, 'yes', 'no')]
```

## plots of the domain interactions

```{r}
g <- ggplot(scores.dt[Domain != 'fullProtein',], aes(x=gene_pair, y=iptm, color=above.threshold)) +
  #geom_boxplot(alpha=0.4, color='darkgrey') +
  geom_point(alpha=0.8) +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  scale_color_manual(values=c('no'='black', 'yes'='red')) +
  labs(title='AF-multimer GPCR-GProtein pair scores') +
  ylim(c(0,1)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, size=5))

g
BackupAsPDF(g, 'AFscores.boxplot', dimensions=c(16,10))
```
Again, looks like not an awful lot of these proteins are scoring very well consistently

we have 192 different interactions modelled
```{r}
scores.dt[Domain != 'fullProtein',.N, by=.(gene_pair)]
```
Order by mean IPTM

```{r}
g <- ggplot(scores.dt[Domain != 'fullProtein',], aes(x=reorder(gene_pair, -mean.iptm), y=iptm, color=above.threshold)) +
  #geom_boxplot(alpha=0.4, color='darkgrey') +
  geom_point(alpha=0.8) +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  scale_color_manual(values=c('no'='black', 'yes'='red')) +
  labs(title='AF-multimer pair scores') +
  ylim(c(0,1)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, size=5))

g
BackupAsPDF(g, 'AFscores.ordered.boxplot', dimensions=c(14,10))
```
How many of each of the domains created a full model

```{r}
scores.dt[Domain != 'fullProtein' & model ==5,.N, by=Domain]
```

plot and facet by domain

```{r}
g <- ggplot(scores.dt[Domain != 'fullProtein',], aes(x=reorder(gene_pair, -mean.iptm), y=iptm, color=above.threshold)) +
  #geom_boxplot(alpha=0.4, color='darkgrey') +
  geom_point(alpha=0.8) +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  scale_color_manual(values=c('no'='black', 'yes'='red')) +
  labs(title='AF-multimer Sacsin Prey Score Distributions') +
  ylim(c(0,1)) +
  facet_wrap(~Domain, scales='free_x', ncol=2) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, size=5))

g
BackupAsPDF(g, 'AFscores.ordered.boxplot.faceted', dimensions=c(10,16))


scores.dt[gene_pair %like% 'SACS_HEPN-S',]
```
prey gene facet
```{r}
scores.dt$Prey_gene

g <- ggplot(scores.dt[Domain != 'fullProtein',], aes(x=reorder(gene_pair, -mean.iptm), y=iptm, color=above.threshold)) +
  #geom_boxplot(alpha=0.4, color='darkgrey') +
  geom_point(alpha=0.8) +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  scale_color_manual(values=c('no'='black', 'yes'='red')) +
  labs(title='AF-multimer Sacsin Prey Score Distributions') +
  ylim(c(0,1)) +
  facet_wrap(~Prey_gene, scales='free_x') +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, size=5))

g
BackupAsPDF(g, 'AFscores.ordered.boxplot.preyfaceted', dimensions=c(10,16))
```

Barplots of mean IPTM scores

Barplot of mean IPTM scores

```{r}
dt <- scores.dt[,.(above.threshold = ifelse(mean.iptm > 0.5, 'yes', 'no'), gene_pair, mean.iptm, Bait_gene, Domain, Prey_gene)] %>% 
  unique()


g <- ggplot(dt, aes(x=reorder(gene_pair,-mean.iptm), y=mean.iptm, fill=above.threshold)) +
  geom_bar(alpha=0.8, stat='Identity') +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  scale_fill_manual(values=c('no'='black', 'yes'='red')) +
  labs(title='AF-multimer pair scores barplots') +
  ylim(c(0,1)) +
  facet_wrap(~Prey_gene,scales = 'free_x') +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, size=5))

g
BackupAsPDF(g, 'AFscores.Preyfacet.barplot', dimensions=c(14,14))

g <- ggplot(dt, aes(x=reorder(gene_pair,-mean.iptm), y=mean.iptm, fill=above.threshold)) +
  geom_bar(alpha=0.8, stat='Identity') +
  geom_hline(yintercept = 0.5, linetype=2, alpha=0.4) +
  scale_fill_manual(values=c('no'='black', 'yes'='red')) +
  labs(title='AF-multimer pair scores barplots') +
  ylim(c(0,1)) +
  facet_wrap(~Domain,scales = 'free_x', ncol=2) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, size=5))

g
BackupAsPDF(g, 'AFscores.SacsinDomainfacet.barplot', dimensions=c(14,16))
```
Plot a heatmap of the results so we can see clustering of mean iPTM
```{r}
scores.dt[Domain != 'fullProtein', .(Bait_Domain = paste0(Bait_gene, '_', Domain), Prey_gene,iptm)][, unique(Bait_Domain)]

# just get the mean iptm from each of the groups
scores.mat <- dcast(scores.dt[Domain != 'fullProtein', .(Bait_Domain = paste0(Bait_gene, '_', Domain), Prey_gene,iptm)], Prey_gene~Bait_Domain, value.var = 'iptm', fun.aggregate=mean) %>% 
  as.matrix(., rownames='Prey_gene')


# draw heatmap
hm <- Heatmap(scores.mat, 
              cluster_rows=T,
              show_row_names = T,
              cluster_columns = T, 
             # top_annotation = colAnn, 
              name='Mean iPTM')
hm

BackupAsPDF(hm, 'mean.iPTM.heatmap')
```

is there any kind of row/column effect? Ie do we expect higher scores with certain Proteins and/or domains?
Barplots of average iPTM scores per domain/protein

```{r}
g <- ggplot(scores.dt[Domain != 'fullProtein',], aes(x=Prey_gene, y=iptm, fill=Prey_gene)) +
  geom_boxplot() +
  ggtitle('Distributions of iPTM scores') +
  theme_classic() +
  theme(axis.text.x = element_text(angle=90))
g
BackupAsPDF(g, 'Prey.iPTM.distributions.boxplots.')

g <- ggplot(scores.dt[Domain != 'fullProtein'], aes(x=Domain, y=iptm, fill=Domain)) +
  geom_boxplot() +
  ggtitle('Distributions of iPTM scores') +
  theme_classic() +
  theme(axis.text.x = element_text(angle=90))
g 
BackupAsPDF(g, 'SACS.domains.Distributions.boxplots.')
```
Lets try the TMP proceedure to iteratively subtract row and column effects and see if something stands out
Residual heatmap suggests that for GNG5 and LRCH2, there is some interaction that is not explained by either the column (domain) or row (Protein prey) factor levels; this could indicate a meaningful interaction

```{r}

colorRamp2()
tmp.out <- medpolish(scores.mat)
tmp.out$residuals %>% str()

# draw heatmap
hm <- Heatmap(tmp.out$residuals, 
              cluster_rows=T,
              show_row_names = T,
              cluster_columns = T, 
              name='TMP residuals')

BackupAsPDF(hm, 'TMP.residuals.heatmap')
```
trends in data? Dont want to see anything obvious (want normally distributed residuals..)
```{r}
plot(tmp.out)
```
Maybe we should investigate and view the models for the two tables
For now leave as is. Can share the modesl this evening for any of interest


Llets look at the full protein breakdown and see if we see any of these interesting domain interactions also score highly
```{r}
scores.dt[Domain == 'fullProtein' & !Bait %like% 'SUB',]
```

