---
title: "060625_prepareSECDataForShiny"
author: "Martin Gordon"
date: "2025-06-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview 

One issue with our sec input file for the Shiny app; it appears quite jagged and the fractions are clearly not aligned 
What we may want to do here is to use the aligned (and interpolated fractions) and normalized data as input for our visualization
Read in the SEC data and present 

```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)
library(stringr)
library(viridis)
library(pracma) # needed for the peak finding algorithm
library(magick)
#library(UniProt.ws) #cant load this packge; get naming conventions another way

# source Bens scripts so we can do a quick QC assessment of the input data before running
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/UniprotIDMapping.R") # map hu to mm 
source("~/Documents/utils/bp_utils/enrichmentTestFunctions.R")

# flag consec proteins
source("~/Documents/utils/bp_utils/SEC_MS_utils.localMods.R")
source("~/Documents/utils/bp_utils/SEC_MS_utils.R")
```

```{r}
sec.in <- fread('./data/hiv.donor12.seclong.csv.gz')
sec.in


View(scaledIntensityMatrices)
```

Look at QC plots to confirm these are normalized; they are normalized so look good

```{r}
qc.dt <- qcSummaryTable(sec.long)

g <- qcPlotMedianInt(qc.dt) +
  theme_bw()
g
```
Need to fix the fractions; use the same alignment function as in the differential testing, then use interpolation to fix the 
Read in the allPeaks 

Dont really think aligning peaks across these fractions is going to do a good job...  maybe for now we can just 

```{r, readFiles}
sec.long <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_HIV_CD4T_TwoDonors_data/2025_05_06_sec.long.donor1donor2.normalized.interpolated.scaled.csv.gz')
allPeaks <- fread('~/Documents/projects/050625_RKakke_HIV_DonorNewSearch/050625_peakDetectionAndDifferentialExpression_data/2025_05_06_donor1donor2.detectedPeaks.csv.gz')

peakTables <- split(allPeaks, allPeaks$sample)
```
Align fractions

```{r}
standardizeAllPeakTablesToStandard(peakTables, sec.long, standardIdx = "cd4_uninfected_2", fitPortion = 0.65, minPeaksPerFraction = 45, startFitAtPeak = 20)
```
Now, want to interpolate the fractions 

```{r}
# interpolating log2 values

sec.long[,c('treatment', 'replicate') := tstrsplit(sample, '_', keep=c(2,3))]
sec.long[, interpolatedIntensity := approx(standardFraction, intensity, xout = fraction)$y, by = .(sample, protein)]

sec.long[, newInts := ifelse(interpolatedIntensity == 0, NA, interpolatedIntensity)]
sec.long[, newInts := 2^interpolatedIntensity]

fwrite(sec.long[, .(sample, treatment, replicate, protein, gene, fraction, intensity=interpolatedIntensity)], ScriptAndDatedFileName('hiv.2donor.interpolatedVals.secShiny.csv.gz'))
fwrite(sec.long[, .(sample, treatment, replicate, protein, gene, fraction, intensity)], ScriptAndDatedFileName('hiv.2donor.normalizedVals.secShiny.csv.gz'))




sec.long[, intensity := 2^intensity]
sec.long[, interpolatedIntensity := 2^interpolatedIntensity]

sec.long
sec.long[,  intensity.interpolated := approx(x=standardFraction, y=ifelse(is.na(intensity), 0, intensity), xout=fraction)$y, by=.(protein, sample)]


sec.long[sample == 'cd4_infected_1' & gene == 'E2F8']$fraction
sec.long[sample == 'cd4_infected_1' & gene == 'E2F8']$standardFraction
sec.long[sample == 'cd4_infected_1' & gene == 'E2F8']$intensity_totalScaled
sec.long[sample == 'cd4_infected_1' & gene == 'E2F8']$intensity
sec.long[sample == 'cd4_infected_1' & gene == 'E2F8']$interpolatedIntensity
sec.long[sample == 'cd4_infected_1' & gene == 'E2F8']$newInts
sec.long[sample == 'cd4_infected_1' & gene == 'E2F8']$fraction.standardised
sec.long[sample == 'cd4_infected_1' & gene == 'E2F8']$intensity.standardised

# now we want to interpolate the intensity value per fraction per protein/sample

# x is independent (our stdFraction), y is response norm(intensity) 
approx(xtestStd, yInts, xout=xtest, method = "linear")


# could also use this to take the average but fine as is for now
sec.in
```

To me, the infected and the challenged look good and similiar to the uninfected for donor2
Undo the standardisation for both of these samples as they look good.. not sure what to do for donor1 for now.... proceed as is and speak to BP tomorrow morning.

```{r, fix-peak-alignments}
peakTables$cd4_infected_2[, cofmN.standardized := cofmN]
sec.long[sample == "cd4_infected_2", standardFraction := fraction]

peakTables$cd4_challenged_2[, cofmN.standardized := cofmN]
sec.long[sample == "cd4_challenged_2", standardFraction := fraction]
```

# 06-18-25
## prepare the Donor 123 samples and the differential analysis tables 

Ok the DE contrast set looks good! Compare the uploaded SEC-MS and the sec profiles
```{r}
anova.dt <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_hiv.d123.pwContrasts.rmfrac20peaks.csv.gz')
anova.dt[, donor := tstrsplit(contrast, '_', keep=3)]
fwrite(anova.dt, ScriptAndDatedFileName('hiv.d123.anova.csv.gz'))
```

look at the sec dt data

```{r}
sec.eg <- fread('./data/2025_06_06_hiv.2donor.normalizedVals.secShiny.csv.gz')
head(sec.eg)
```
change to include all 3
```{r}
sec.dt <- fread('~/Documents/projects/061425_MMuralidharan_HIVdonor23_SECMS/061725_D123_BPSECMS_pipeline_data/2025_06_18_hiv.d123.processed.ints.csv.gz')

fwrite(sec.dt[, .(sample, treatment, replicate, protein, gene, fraction, intensity=intensity_interpolated)], ScriptAndDatedFileName('hiv.3donors.interpolatedVals.secShiny.csv.gz'))
fwrite(sec.dt[, .(sample, treatment, replicate, protein, gene, fraction, intensity)], ScriptAndDatedFileName('hiv.3donors.normalizedVals.secShiny.csv.gz'))
```
these are cmds fro the Phagemap server

```{r}
 nano  //bpolacco/ShinyApps/060625_HIVDonor12_SECMS/app.R

sudo nano  /home/bpolacco/shiny_server_logs
sudo nano /home/bpolacco/shiny_server_logs
```

