---
title: "ED.ph.Redo.wBatch"
author: "Martin Gordon"
date: "2023-12-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Rerun the analysis controlling for batch effect outlined by Xiaofang

```{r}
library(ggplot2)
library(data.table)
library(magrittr)
library(ComplexHeatmap)
library(randomcoloR)
library(stringr)
library(OmnipathR)
library(circlize)
library(ggsignif)


#source BP utility scripts
source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/MSstats_Helper_Functions.R")
source("/Users/martingordon/Documents/utils/bp_utils/MSstats_V4_Functions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/linearModels.R")

# kinase enrichment analysis
source("/Users/martingordon/Documents/utils/bp_utils/KinaseActivityScores.R")

# mg utility functions
source("/Users/martingordon/Documents/utils/mg_utils/HelperFunctions.R")
```



```{r}
msinput <- fread("./output/MSstatsInput.PH.csv")

dp.out <- MSstats::dataProcess(setDF(msinput),
                               MBimpute = FALSE,
                               featureSubset = "highQuality",
                               remove_uninformative_feature_outlier = TRUE)

saveRDS(dp.out, ScriptAndDatedFileName('ph.mss.dataproc.rds'))
```

```{r}
contrast.mat <- makeContrast.regEx(dp.out,c( 
                               "^M100-Veh",
                               "^Nbon-Veh",
                               "^M100_Nbon-Veh",
                               "^M100_Nbon-Nbon",
                               "^M100_Nbon-M100"
                               ))
```

run mss pw comparisons between groups w Batch controlled
```{r}
dp.out <- readRDS('~/Documents/projects/112823_XZhong_DARPA2/DE.ph.LMredo.noBatch_data/2023_11_30_ph.mss.dataproc')

p.quant<- setDT(dp.out$ProteinLevelData)

p.quant[, OLD_SUBJECT := paste0(GROUP,'.',SUBJECT)]

p.quant$OLD_SUBJECT %>%  unique()

p.quant[, SUBJECT := fcase(
  OLD_SUBJECT %in% c('Nbon.3', 'Nbon.2', 'M100.4', 'M100_Nbon.1'), 'pseudobatch.1',
  OLD_SUBJECT %in% c('M100_Nbon.2', 'M100_Nbon.4', 'M100.3', 'Veh.1'), 'pseudobatch.2',
  OLD_SUBJECT %in% c('M100_Nbon.3', 'Veh.4','Veh.2','M100.1'), 'pseudobatch.3',
  OLD_SUBJECT %in% c('Veh.3', 'M100.2', 'Nbon.4', 'Nbon.1'), 'psuedobatch.4'
)]

# sanity check.. looks good
p.quant[,.N, by=.(SUBJECT, OLD_SUBJECT)] %>% 
  .[order(SUBJECT)]

dp.out$ProteinLevelData <- p.quant

gc <- MSstats::groupComparison(contrast.mat, dp.out)
```

setting pval threshold at fc +/- 50% and pval< 0.005
```{r}
mss.dt <- setDT(gc$ComparisonResult)

fwrite(mss.dt, ScriptAndDatedFileName('mss.pwcomparisons.newbatch.raw.csv'))

proc.dt <- mss.dt[,gene := multiUniprotSites2multiGeneSites(Protein, species='MOUSE')] %>% #needed as in factor format
  .[!issue %in% c('completeMissing','oneConditionMissing'), ] %>% 
  .[is.finite(log2FC) & !is.na(pvalue), ] %>% 
  .[is.finite(log2FC), ] %>% 
  .[, p.adj := p.adjust(pvalue, method='BH'), by=.(Label)] %>% 
  .[, sig := 'not'] %>% 
  .[abs(log2FC) > 0.58 & pvalue < 0.005, sig := ifelse(log2FC > 0, 'up', 'down')]


fwrite(proc.dt, ScriptAndDatedFileName('mss.pwcomparisons.newbatch.processed.csv'))

# v few sig hits a adj.pval < 0.05 & log2FC +/- 50%, so dropped to pval < 0.001
# 10 more hits removing batch term
proc.dt[sig != 'not',.N, by=.(sig,Label)]$N %>% sum()
```

```{r}
ggplot(mss.dt, aes(x=pvalue)) + 
  geom_histogram() +
  facet_wrap(~Label)
```

```{r}
g <- ggplot(proc.dt, aes(x=log2FC, y=-log10(pvalue), color=sig, label=gene)) + 
      geom_point() +
      geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
      geom_hline(yintercept = -log10(0.005), linetype=2, alpha=0.4) +
      ggrepel::geom_text_repel(data=proc.dt[sig != 'not',],  size = 2, max.overlaps = 20) +
      scale_color_manual(values=c('up'='red', 'down'='blue', 'not'='grey')) +
      ggtitle(' (FC +/ 50% & pval < 0.005)') +
      facet_wrap(~Label, scales = 'free') +
      theme_bw() +
      theme(strip.text.x = element_text(size = 5))
      
BackupAsPDF(g, 'ph.grpcontrasts.pval.0.005.multipanel.volcanoplot.', dimensions=c(13,10))
```
generate individual volcano plots

```{r}
labels.oi <- proc.dt[, unique(Label)]

for (lab in labels.oi){
  
  g <- ggplot(proc.dt[Label == lab,], aes(x=log2FC, y=-log10(pvalue), color=sig, label=gene)) + 
      geom_point() +
      geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
      geom_hline(yintercept = -log10(0.005), linetype=2, alpha=0.4) +
      ggrepel::geom_text_repel(data=proc.dt[Label == lab & sig != 'not',],  size = 2, max.overlaps = 20) +
      scale_color_manual(values=c('up'='red', 'down'='blue', 'not'='grey')) +
      ggtitle(paste0(lab, ' (FC +/ 50% & pval < 0.005)')) +
      theme_bw() +
      theme(strip.text.x = element_text(size = 5))

  BackupAsPDF(g, paste0(lab, '.pval.0.005.volcano'))
}
```
What about kinase enrichment?
```{r}
# modified function to avoid omnipath kinase download and read from file
loadKinaseDataOmniPath <- function(species = "HUMAN", removeNonKinases = TRUE, fixNames = TRUE, omniPathFile = './data/mouse.omnipath.csv', ...){
  species <- toupper(species)
  organismID <- c(HUMAN = 9606, RAT = 10116, MOUSE = 10090)[species]
  if(is.na(organismID))
    stop(species, " not a recognized organism")
  enzsub <- fread(omniPathFile)
  setDT (enzsub)
  # for consistency with loadKinaseData I need columns: "CTRL_GENE_NAME", "TARGET_GENE_NAME", "TARGET_RES", "TARGET_POS", "TARGET_UP_ID"
  
  if (any(grepl("ProtMapper", enzsub$sources))){
    message ("OmniPath includes ProtMapper data which has many non-kinases.")
    
    if(removeNonKinases == TRUE) {
      message ("Removing all enzymes that are not kinases according to org.Hs.eg.db GO:0016301")
      # get GO kinase list
      
      if (species == "HUMAN"){
        library(org.Hs.eg.db)
        kinaseTable <- AnnotationDbi::select (org.Hs.eg.db, get(c("GO:0016301"), org.Hs.egGO2ALLEGS), c("ENTREZID", "GENENAME", "SYMBOL")) 
      } else if (species == "MOUSE"){
        library(org.Mm.eg.db)
        kinaseTable <- AnnotationDbi::select (org.Mm.eg.db, get(c("GO:0016301"), org.Mm.egGO2ALLEGS), c("ENTREZID", "GENENAME", "SYMBOL")) 
      } else{
        stop ("I don't yet know how to find kinases for ", species)
      }
      
      kinases <- unique(kinaseTable$SYMBOL)
      #label and filter into a new table
      enzsub[, is_kinase := enzyme_genesymbol %in% kinases]
      enzsub <- enzsub[is_kinase == TRUE]
    }
  }
  
  if (fixNames){
    setnames(enzsub, 
             old = c("enzyme_genesymbol", "substrate_genesymbol", "residue_type", "residue_offset", "substrate"), 
             new = c("CTRL_GENE_NAME", "TARGET_GENE_NAME", "TARGET_RES", "TARGET_POS", "TARGET_UP_ID"))
  }
  
  return (enzsub[modification == "phosphorylation"])
}

kinaseData  <- loadKinaseDataOmniPath(species='MOUSE')

# just look at enrichment of the complete set
mss.dt <-  proc.dt


mss.dt[,gene := multiUniprotSites2multiGeneSites(Protein, species = "MOUSE")]
singleSiteResults <- prepare_AMSS_ResultsFile(mss.dt, column = "gene")



labels <- unique(singleSiteResults$Label)

kinActList <- lapply (labels, FUN=function(lab){kinaseActivity(singleSiteResults[Label == lab & representative==TRUE],
                                                               plots = FALSE,
                                                               kinaseData = kinaseData, 
                                                               do.sea = TRUE)})


names(kinActList) <- labels

kinActFull.scores <- rbindlist(lapply(kinActList, FUN = function(x)x$scores), idcol="Label")
kinActFull.mapped <- rbindlist(lapply(kinActList, FUN = function(x)x$kinaseMapped)) # Label is already in these tables


fwrite(kinActFull.scores, ScriptAndDatedFileName("kinaseSummaryScores.csv"))
fwrite(kinActFull.mapped, ScriptAndDatedFileName("kinaseSubstrateData.csv"))
message (sprintf("Writing kinase output files to\n\t%s\n\t%s", kinaseSummaryScores.csv, kinaseSubstrateData.csv))
```

## Heatmap of significant kinases
less hits than ori run... leave as is for now...
```{r, fig.width = 6, fig.height= 4}
sigKinases <-  kinActFull.scores[fdr.BH < 0.05 & N > 2, unique(CTRL_GENE_NAME)]

sigKinase.mat.z <- as.matrix(dcast (kinActFull.scores[CTRL_GENE_NAME %in% sigKinases], CTRL_GENE_NAME~Label, value.var = "Z"),
                              rownames = "CTRL_GENE_NAME")

sigKinase.mat.N <- as.matrix(dcast (kinActFull.scores[CTRL_GENE_NAME %in% sigKinases], CTRL_GENE_NAME~Label, value.var = "N"),
                                  rownames = "CTRL_GENE_NAME")


hm <- Heatmap (sigKinase.mat.z, 
         cluster_columns=TRUE, 
         name = "Kinase Z Score",
         #column_split = tstrsplit(colnames(sigKinase.mat.z), split="[-_.]")[[1]],
         col = circlize::colorRamp2(breaks = c(-3, -1, 1, 3), colors = c("blue", "gray", "gray", "red")),  # two midpoints in case you want to have a wider gray bar around 0
         row_title = '',
         cell_fun = function(j, i, x, y, width, height, fill) {
                                  if (!is.na(sigKinase.mat.N[i,j])){
                                    grid.text(sprintf("%.0f", sigKinase.mat.N[i, j]), x, y, gp = gpar(fontsize=10, col="white"))
                                  }
                                })

hm
BackupAsPDF(hm, prefix = "Heatmap_SigKinases")

hm <- Heatmap (sigKinase.mat.z, 
         cluster_columns=FALSE, 
         name = "Kinase Z Score",
        # column_split = tstrsplit(colnames(sigKinase.mat.z), split="[-_.]")[[1]],
         col = circlize::colorRamp2(breaks = c(-3, -1, 1, 3), colors = c("blue", "gray", "gray", "red")),  # two midpoints in case you want to have a wider gray bar around 0
         row_title = '',
         cell_fun = function(j, i, x, y, width, height, fill) {
                                  if (!is.na(sigKinase.mat.N[i,j])){
                                    grid.text(sprintf("%.0f", sigKinase.mat.N[i, j]), x, y, gp = gpar(fontsize=10, col="white"))
                                  }
                                })
hm
BackupAsPDF(hm, prefix = "Heatmap_SigKinases_noColCluster")
```


Plot Go enrichment results of the significant genes

```{r}
# load mouse gmt file 
gmt.go <- loadGmtFromBioconductor(dbName='org.Mm.eg.db', ontology = "ALL", keyType = "UNIPROT")

p.quant <-  fread('./DE.ph.291123_data/2023_11_29_MSstatsInput.csv')
# get matching col
p.quant[, Protein := gsub("_[STY][0-9]+", "",ProteinName)]

universe <-  p.quant$Protein %>%  unique()
length(universe) # ~6.2k

# get a simplifed dt of gene names of the moving sites
gene.dt <-  proc.dt[sig != 'not', ] %>% 
  .[, ProteinName := gsub("_[STY][0-9]+", "",Protein)] %>% 
  .[, .(Label, ProteinName, sig)] %>% 
  .[order(Label,ProteinName)]

# create an ernich grouping col
gene.dt[, enrich_grp := paste0(Label, '.', sig)]


enrich.dt <- enricherOnGroups(gene.dt, groupColumns = 'enrich_grp', geneColumn = "ProteinName", term2gene.gmt = gmt.go, universe = universe)

simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')


ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt, groupColumn = 'enrich_grp', topN = 8, title='GO term enrichment', 
                                  negCols=unique(simp.enrich$simplified$enrich_grp[grep('down', simp.enrich$simplified$enrich_grp)]), 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 8), upperThreshold = 4)
BackupAsPDF(ht, 'go.enrichment.msstats.pwgroups.nobatchCor.heatmap.')
```
Interaction Analysis
----
Rerun the lm model with a batch term

```{r}
dp.out <- readRDS(ScriptAndDatedFileName('mss.pwcomparisons.newbatch.processed.csv'))

p.quant<- setDT(dp.out$ProteinLevelData)

p.quant[, OLD_SUBJECT := paste0(GROUP,'.',SUBJECT)]

p.quant$OLD_SUBJECT %>%  unique()

p.quant[, SUBJECT := fcase(
  OLD_SUBJECT %in% c('Nbon.3', 'Nbon.2', 'M100.4', 'M100_Nbon.1'), 'pseudobatch.1',
  OLD_SUBJECT %in% c('M100_Nbon.2', 'M100_Nbon.4', 'M100.3', 'Veh.1'), 'pseudobatch.2',
  OLD_SUBJECT %in% c('M100_Nbon.3', 'Veh.4','Veh.2','M100.1'), 'pseudobatch.3',
  OLD_SUBJECT %in% c('Veh.3', 'M100.2', 'Nbon.4', 'Nbon.1'), 'psuedobatch.4'
)]

# sanity check.. looks good
p.quant[,.N, by=.(SUBJECT, OLD_SUBJECT)] %>% 
  .[order(SUBJECT)]

# now want 3 factors: batch, treatment, antagonist in model
p.quant[, `:=`(batch = SUBJECT,
               ant = ifelse(GROUP %like% 'M100', 'M100', 'No_Ag'),
               treat = ifelse(GROUP %like% 'Nbon', 'Nbon', 'No_Treat') )]

#sanity.check 
p.quant[,.N, by=.(ant,treat,OLD_SUBJECT,batch)]

# example of single lm controlling for batch (P07759 used as significant in msstats modelling)
lm(LogIntensities~treat+ant+batch+ant:treat, data = p.quant[Protein == "Q9D0E1_S452"]) %>% summary()
model <- aov(LogIntensities ~ treat+ant+batch+ant:treat, data=p.quant[Protein == "Q9D0E1_S452"])  
summary(model)
```

```{r}
# linearModelsAllProteins can this function which computes statistic for a contrast of two different log2FC
# puts pval and log2FC on sig differences of differences
emmeans.contrastOfContrasts <- function (l, factorFormula = ~treat|ant){  # how does treatment effect change per agonist level
  emm <- emmeans::emmeans(l, factorFormula)
  contrast1 <- pairs(emm)
  contrast2 <- pairs(contrast1, by = NULL, adjust = "none")
  return (as.data.table(contrast2))
}


  
  # Numbers near the number of processors available in your machine is probably counterproductive.
lm.out <- linearModelsAllProteins(p.quant, 
                            formulaList = list(full   = (LogIntensities~batch+treat+ant+ ant:treat)),
                            splitColumn = "Protein", postProcessFunction = emmeans.contrastOfContrasts, cl = 6) %>%  suppressWarnings()


#saveRDS(lm.out, ScriptAndDatedFileName('lm.newBatch.obj'))
```

inspect anova table 
not a lot of interaction between the terms

```{r}
ggplot(lm.out$anova, aes(x=p.value)) +
  geom_histogram() +
  facet_wrap(~term)

lm.out$anova[ term %in% c('treat', 'ant', 'treat:ant')] %>% 
  .[, fdr := p.adjust(p.value, method='fdr'), by=term]

fwrite(lm.out$anova, ScriptAndDatedFileName('anova.table.nobatch.csv'))
```

plot of the significant interaction terms

```{r}
# cluster w NA
clusterwNA <- function(x, na.val=0){
  
  x[is.na(x)] <- na.val
  dist(x) %>%  hclust(method='complete')
}

```


```{r}
sig.prots <- lm.out$anova[term %in% 'treat:ant' & p.value < 0.005]
sig.prots %>% length()

sig.prots[, gene := multiUniprotSites2multiGeneSites(Protein, species='MOUSE')]

p.mat <- dcast(p.quant, Protein~OLD_SUBJECT, value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')


column.vec <- sub('.[1-6]$', '', colnames(p.mat))

rownames(p.mat) <- multiUniprotSites2multiGeneSites(rownames(p.mat), species='MOUSE')

subMat <-  p.mat[rownames(p.mat) %in% sig.prots$gene,]
subMat <-  sweep(subMat, 1, FUN='-', STATS = apply(subMat, 1, median, na.rm=T))

subMat

hm <- Heatmap(subMat, 
              cluster_columns = FALSE, 
              name = 'log Ints/Median',
              cluster_rows = clusterwNA(subMat),
              #column_title = '(M100+Nbon vs Veh) - (M100 vs Veh)',
              row_title = paste0(nrow(subMat), ' sites p.val < 0.005)'),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 6),
              col=colorRamp2(breaks=c(-2,0,2), colors=c('blue', 'white', 'red')),
              column_split = column.vec,
              row_title_gp = gpar(fontsize=9),
              column_title_gp = gpar(fontsize=7))

hm
BackupAsPDF(draw(hm, column_title = "Differential Response to NBON w/wo M100"),'m100.nbon.vs.nbon.interaction.newbatch.mediansweep.clean.heatmap', dimensions = c(8,8))
```


Continue running from here
------

```{r}

# droping cols with many missing values
subMat <-  p.mat[rownames(p.mat) %in% sig.prots$gene,]
subMat <-  sweep(subMat, 1, FUN='-', STATS = apply(subMat, 1, median, na.rm=T))

missing.prots <- rownames(subMat)[apply(subMat, 1, function(x){ sum(is.na(x)) >= 5})]
missing.prots

cleanMat <- subMat[!rownames(subMat) %in% missing.prots,]
dim(cleanMat)

hm <- Heatmap(cleanMat, 
              cluster_columns = FALSE, 
              name = 'log Ints/Median',
              cluster_rows = clusterwNA(cleanMat),
              #column_title = '(M100+Nbon vs Veh) - (M100 vs Veh)',
              row_title = paste0(nrow(cleanMat), ' sites p.val < 0.005)'),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 6),
              col=colorRamp2(breaks=c(-2,0,2), colors=c('blue', 'white', 'red')),
              column_split = column.vec,
              row_title_gp = gpar(fontsize=9),
              column_title_gp = gpar(fontsize=10))

hm


BackupAsPDF(draw(hm, column_title = "Differential Response to NBON w/wo M100"),'m100.nbon.vs.nbon.interaction.newbatch.rmnoisyrows.mediansweep.clean.heatmap', dimensions = c(8,8))
```

now generate plots of Htr2a, check if any of these are sig

```{r}
mss.dt <- fread(ScriptAndDatedFileName('mss.pwcomparisons.newbatch.processed.csv'))

# ok, so one sig site in the comparisos w/o correction
mss.dt[gene %like% 'Htr2a' & sig != 'not',]
```
```{r}
p.quant<- setDT(dp.out$ProteinLevelData)
p.quant[, gene := multiUniprotSites2multiGeneSites(Protein, species='MOUSE')]

p.quant %>%  str()
sites.oi <- c('Htr2a_S298','Htr2a_S421')

lab.col.pal <- randomcoloR::distinctColorPalette(5)



p.quant[gene %like% sites.oi,]

g <- ggplot(p.quant[gene %in% sites.oi,], aes(x=GROUP, y=LogIntensities, color=GROUP)) +
  geom_point() + 
  stat_summary(fun.dat='mean_cl_boot', color='grey', size=0.2) +
  scale_color_manual(values=col.pal) +
  ggtitle('HTR2A PH sites') +
  facet_wrap(~gene) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, vjust = 0.7))
g
BackupAsPDF(g, 'htr2a.sites.oi.linechart.fixedxaxis',dimensions=c(8,6))
```
lollipop plot with pval 
```{r}
# annotate with pval if < 0.05
mss.dt[, lab := '']
mss.dt[pvalue < 0.05, lab := ifelse(pvalue < 0.01, '**', '*')]
mss.dt[pvalue < 0.001, lab := '***']

g <- mss.dt[gene %in% sites.oi,] %>% 
  ggplot(., aes(x=Label, y=log2FC, color=Label, label=paste0(lab, ' ', round(pvalue, 3) )) ) +
  geom_segment(aes(x=Label, xend=Label, y=0, yend=log2FC), color='grey', lwd=.5, linetype=1) +
  geom_point(size=3) +
  geom_hline(yintercept=0, alpha=0.4, linetype=2) +
  geom_text(data=mss.dt[gene %in% sites.oi & pvalue < 0.05], color='black', size=3, vjust=-2) +
  facet_wrap(~gene) +
  ylim(c(-1,2)) +
  theme_bw() + 
  #coord_flip() +
  scale_color_manual(values=c(lab.col.pal)) +
  theme(
    axis.text.x = element_text(angle=90, vjust=0.6)
  )

g

BackupAsPDF(g, 'htr2a.lollipop.pvals.')

```

```{r}
g <- mss.dt[gene %in% sites.oi,] %>% 
  ggplot(., aes(x=Label, y=log2FC, color=Label, label=paste0(lab, ' ', round(pvalue, 4) )) ) +
  geom_segment(aes(x=Label, xend=Label, y=0, yend=log2FC), color='grey', lwd=.5, linetype=1) +
  geom_point(size=3) +
  geom_hline(yintercept=0, alpha=0.4, linetype=2) +
  geom_text(data=mss.dt[gene %in% sites.oi & pvalue < 0.05], color='darkgrey', size=3, vjust=-2) +
  facet_wrap(~gene) +
  ylim(c(-1,2)) +
  theme_bw() + 
  coord_flip() +
  scale_color_manual(values=c(lab.col.pal)) +
  theme(
    axis.text.x = element_text(angle=90, vjust=0.6)
  )

g

BackupAsPDF(g, 'htr2a.lollipop.pvals.rotated.')
```
```{r}
mss.dt[gene %in% sites.oi & pvalue < 0.05,]

annotation_data <- data.table(
  GROUP = c('M100', 'M100_Nbon', 'Nbon', 'Veh'),
  xend = c('M100_Nbon', 'M100_Nbon', 'Nbon', 'M100_Nbon'),
  y = c(15.4, 15, 15.5, 15.3),
  annotation = c("**", "", "", "*"),
  gene = rep(sites.oi[1], 4)
)

new.anno <- data.table(
  GROUP = c('M100', 'M100_Nbon', 'Nbon', 'Veh'),
  xend = c('M100_Nbon', 'M100_Nbon','Nbon', 'Veh'),
  y = c(14.8, 15, 15.5, 15.3),
  annotation = c("*", "", "", ""),
  gene = rep(sites.oi[2], 4)
)



annotation_data <-  rbind(annotation_data, new.anno)
annotation_data

g <- ggplot(p.quant[gene %in% sites.oi,], aes(x=GROUP, y=LogIntensities, color=GROUP)) +
  geom_boxplot(color='grey') +
  geom_point() + 
  geom_signif(stat="identity",
             data=annotation_data,
            aes(x=GROUP, y=y, xend=xend, yend=y, annotation=annotation),color='darkgrey', size=4) +
  scale_color_manual(values=col.pal) +
  ggtitle('HTR2A PH sites') +
  facet_wrap(~gene) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, vjust = 0.7))
g

BackupAsPDF(g, 'htr2a.S298.sites.siglabelled.boxplots.')
log2(1.5)
```


041223
----
Load in the data and convert to wide format for sharing
msstats

```{r}
mss.out <- fread('./021223.ph.rereunWBatch_data/2023_12_02_mss.pwcomparisons.newbatch.processed.csv')
anova.out <- fread('./021223.ph.rereunWBatch_data/2023_12_02_anova.table.nobatch.csv')

mss.dt <- mss.out[, .(gene, Protein, Label, log2FC, pvalue, adj.pvalue, SE, Tvalue)]
setnames(mss.dt, new=c('geneSite', 'proteinSite', 'Label', 'log2FC','pvalue', 'adj.pvalue', 'SE', 'Tvalue'))


fwrite(mss.dt, ScriptAndDatedFileName('mss.pwcomparisons.csv'))

# also present in wideformat
fwrite(dcast(mss.dt[,.(geneSite,proteinSite, Label, log2FC, pvalue, adj.pvalue)], geneSite+proteinSite~Label, value.var=c('log2FC', 'pvalue', 'adj.pvalue')),
       ScriptAndDatedFileName('mss.pwcomparisons.wide.csv'))


# anova table 
anova.out %>%  str()
anova.dt <- anova.out[term == 'treat:ant',]
anova.dt[, geneSite := multiUniprotSites2multiGeneSites(Protein, species='MOUSE')]

anova.dt <-  anova.dt[, .(Protein, geneSite, term, p.value)] %>% 
  .[order(p.value)]

setNames(anova.dt, c('proteinSite', 'geneSite', 'term', 'p.value'))
anova.dt[ , term := 'NBON-M100 interaction']
anova.dt[, adj.pval := p.adjust(p.value, method='BH')]

fwrite(anova.dt, ScriptAndDatedFileName('interaction.analysis.csv'))

```


anova results
```{r}


lm.out <-  readRDS('./021223.ph.rereunWBatch_data/2023_12_02_lm.newBatch.obj')

```

