---
title: "de.abundance"
author: "Martin Gordon"
date: "2023-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overivew
----

Differential expression analysis of AB data
Murine origin, so be wary of this when converting between identifiers

Simple experimental set-up: 4 conditions; control (veh), compound (Nbon), antagonist (M100), M100_Nbon (antagonist + compound)
$ replicates per condition

Have both phospho and abundance data available
Goal is to present by Dec 4 DARPA II meeting

libraries and custom scripts
-----

```{r}
library(ggplot2)
library(data.table)
library(magrittr)
library(ComplexHeatmap)
library(randomcoloR)
library(stringr)
library(emmeans)

#source BP utility scripts
source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/MSstats_Helper_Functions.R")
source("/Users/martingordon/Documents/utils/bp_utils/MSstats_V4_Functions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/linearModels.R")

# mg utility functions
source("/Users/martingordon/Documents/utils/mg_utils/HelperFunctions.R")
```

first off read in the data
AB DIA data produced by spectronaut

```{r}
spec <- fread('./data/MB03/AB/20231114_222317_MB03_AB_dirDIA_Report.tsv')
keys <-  fread('./data/MB03/AB/MB03_AB_dirDIA_ConditionSetup.tsv') # not really needed for this, think this file is used for artMS
spec %>%  str()
```

Investigate intensity distributions
Known issue with Spectronaut that it produces low-intensity 'garbage' measurements that can be removed

```{r}
ggplot(spec, aes(x=log2(Intensity))) +
  geom_histogram()


spec <-  spec[log2(Intensity) > 4]

# check clean data
# all looks good so far
ggplot(spec, aes(x=log2(Intensity))) +
  geom_histogram() +
  facet_wrap(~Condition+BioReplicate)
```
Each row in the table should be the intensity from one MS "feature" per run (an elution peak from a single peptide ion). It's possible that there are multiple features per peptide ion and run--if the peptide elutes in two different peaks perhaps.

are there any multiple-feature peptide ions?

```{r}
# FALSE, so we dont need to consider Precursor Charge when counting fragment ions for plotting
any(spec[, .N, by=.(PeptideSequence, Run, PrecursorCharge)]$N) > 1
```

How many fragment ions per run?
Counts per run look consistent

```{r}
# generate a distinct color palette for each condition
col.pal <-  randomcoloR::distinctColorPalette(length(unique(spec$Condition)))

g <- ggplot(spec, aes(x=reorder(interaction(Condition, BioReplicate)), fill=Condition)) +
  geom_bar(stat='Count') +
  scale_fill_manual(values=c(col.pal)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle=90))

BackupAsPDF(g, 'peptideion.counts.barplot.')

```

Peptide Intensity distributions
Raw run intensities look good; consistent between runs
```{r}
g <- ggplot(spec, aes(x=reorder(interaction(Condition, BioReplicate)), y=log2(Intensity), fill=Condition)) +
  geom_boxplot() +
  scale_fill_manual(values=c(col.pal)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle=90))
g
# BP function in manageScritData.R file
# 3 main arguments: the ggplot obj, file name, and dimensions (optional; default is dev.size(units = "in"))
BackupAsPDF(g, 'rawIntensity.boxplots.')
```
How ds sample clustering look based on the raw peptide intensities

Heatmap of peptide intensities
```{r}
# create a intensity matrix
f.mat <- dcast(spec, interaction(PeptideSequence,PrecursorCharge)~Condition+BioReplicate, value.var = 'Intensity') %>% 
  as.matrix(rownames='PeptideSequence')

f.mat <- log2(f.mat)

#66k measured unique peptide ion intensities
dim(f.mat)
#54k measured unique peptide ion intensities
dim(f.mat[complete.cases(f.mat),])

# as we hae missing values in our dt, need to define function handle clustering of these 

clusterwNA <- function(mat, na.val=0,...){
  
  # assign val to missing
  mat[is.na(mat)] <- na.val
  return(hclust(dist(mat)))
}

```

```{r}

#defining our annotation vecotrs for the bar charts
colnames(f.mat)
condition <- sub('_[1-4]{1}$', '', colnames(f.mat))
bioreplicate <- str_extract(colnames(f.mat), '[1-4]$')

ha <- HeatmapAnnotation(condition=condition,
                        bioreplicate=bioreplicate,
                        # set cols for levels
                        col=list(condition=c("M100"='red', "M100_Nbon"='blue',"Nbon"='green', "Veh"='yellow'),
                                  bioreplicate=c('1'='orange', '2'='black', '3'='pink', '4'='purple'))
                        )


mat.size <-  2000

#subsample the matrix to visualise
subMat <-  f.mat[sample(rownames(f.mat), mat.size), ]

hm <- Heatmap(subMat,
        cluster_rows = clusterwNA(subMat), #ComplexHeatmap can accept a function as input for clustering. This handles NA vals by converting to 0
        name='(Log2) Ints.',
        cluster_columns = T,
        show_row_names = F,
        top_annotation = ha,
        row_title = sprintf('%s peptide ions', mat.size)
        )

BackupAsPDF(hm, 'peptideion.rawintensity.heatmap') 

```
looking at the heatmap its clear the raw intensity values are driving the clustering
Subtract the median peptide value across runs (rows) to better see differences between the runs

```{r}
# sweep subtracts a vector of values from either rows (dimension = 1) or columns (dimension = 2)
# apply applies a function (median here) to each row (dimension = 1 )
# used together here we are subtracting row medians
subMat <-  sweep(f.mat, MARGIN=1, FUN='-', STATS = apply(f.mat, 1, median, na.rm=T))


#subsample the matrix to visualise
subMat <-  subMat[sample(rownames(subMat), mat.size), ]

hm <- Heatmap(subMat,
        cluster_rows = clusterwNA(subMat), #ComplexHeatmap can accept a function as input for clustering. This handles NA vals by converting to 0
        name='(Log2) Ints./Median',
        cluster_columns = T,
        show_row_names = F,
        top_annotation = ha,
        row_title = sprintf('%s peptide ions', mat.size)
        )

BackupAsPDF(hm, 'peptideion.rawintensity.mediansweep.heatmap') 
```
PCA plot of samples
No clear clustering of samples on 1st and second PCs

```{r}
# for PCA we need to remove any missing vals from our matrix
f.mat <-  f.mat[complete.cases(f.mat),]


pcaOut <- prcomp(t(f.mat))


colInfo <- data.table(colname = colnames(f.mat))
colInfo[colname %like% 'M100_Nbon', colname := sub('_', '.', colname)]

colInfo[,c("condition", "rep") := tstrsplit(colname, "[_]",  keep = c(1,2)) ] # split the string and transpose into two columns


#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaDT[rn  %like% 'M100_Nbon', rn := sub('_', '.', rn)]

pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = condition, shape = rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, "complete.features.rawintensity.PCA")
```


How much variance is explained by the individual PCs?

```{r}
barplot(pcaPercentVar)
```

```{r}
p <- ggplot (pcaDT, aes(x=PC3, y=PC4,  fill = condition, shape = rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC3, %.1f%%", pcaPercentVar[3])) + 
  ylab (sprintf ("PC4, %.1f%%", pcaPercentVar[4])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
```

protein summarisation and normalisation
```{r}
# MSStats like NA values explicit (assigned row with 'emtpy' val NA etc...) this BP function handles this
mssInput <- specFileToCompleteMSstats(spec)

# colunm necessay to run MSstats and not output by spectronuat
mssInput[, IsotopeLabelType := 'L']
```

```{r}
dp.out <- MSstats::dataProcess(setDF(mssInput), #convert to data.frame
                              MBimpute =  FALSE, # turn off imputation
                              normalization = 'EQUALIZEMEDIANS', # default normalisation method
                              summaryMethod = "TMP", #tukey median polish for protein summarisation
                              featureSubset = "highQuality", # only yse
                              remove_uninformative_feature_outlier = TRUE)

# save the output of msstats data.process 
#saveRDS(dp.out, './output/281123.msstats.dp.rds')
```


```{r}
# read in the data 
dp.out <- readRDS('./output/281123.msstats.dp.rds')
```

QC check of the summarised data; Protein counts, boxplots of logIntensities

```{r}
p.quant <- data.table(dp.out$ProteinLevelData)
f.quant <-data.table(dp.out$FeatureLevelData)

# protein intensities look good
g <- ggplot(p.quant, aes(x=reorder(interaction(GROUP,SUBJECT)), y=LogIntensities, fill=GROUP)) +
  geom_boxplot() +
  scale_fill_manual(values=c(col.pal)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle=90))

BackupAsPDF(g, 'proteinintensity.boxplots.')


g <- ggplot(p.quant, aes(x=reorder(interaction(GROUP,SUBJECT)), fill=GROUP)) +
  geom_bar(stat='Count') +
  scale_fill_manual(values=c(col.pal)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle=90))

BackupAsPDF(g, 'protein.counts.barplot.')
```
Sample clustering at the protein level 

```{r}
p.mat <- dcast(p.quant, Protein~GROUP+SUBJECT, value.var="LogIntensities") %>% 
  as.matrix(rownames='Protein')
```

heatmap of subsample of proteins

```{r}
#defining our annotation vecotrs for the bar charts
condition <- sub('_[1-4]{1}$', '', colnames(p.mat))
bioreplicate <- str_extract(colnames(p.mat), '[1-4]$')

ha <- HeatmapAnnotation(condition=condition,
                        bioreplicate=bioreplicate,
                        # set cols for levels
                        col=list(condition=c("M100"='red', "M100_Nbon"='blue',"Nbon"='green', "Veh"='yellow'),
                                  bioreplicate=c('1'='orange', '2'='black', '3'='pink', '4'='purple'))
                        )


mat.size <-  2000

#subsample the matrix to visualise
subMat <-  p.mat[sample(rownames(p.mat), mat.size), ]

hm <- Heatmap(subMat,
        cluster_rows = clusterwNA(subMat), #ComplexHeatmap can accept a function as input for clustering. This handles NA vals by converting to 0
        name='(Log2) Ints.',
        cluster_columns = T,
        show_row_names = F,
        top_annotation = ha,
        row_title = sprintf('%s Proteins', mat.size)
        )

hm
BackupAsPDF(hm, 'proteinIntensity.heatmap') 
```
sweep median values to visualise run-run differences 

```{r}
#subsample the matrix to visualise
subMat <-  sweep(p.mat, MARGIN=1, FUN='-', STAT=apply(p.mat, 1, median, na.rm=T))

subMat <-  subMat[sample(rownames(subMat), mat.size), ]

hm <- Heatmap(subMat,
        cluster_rows = clusterwNA(subMat), #ComplexHeatmap can accept a function as input for clustering. This handles NA vals by converting to 0
        name='(Log2) Ints./Median',
        cluster_columns = T,
        show_row_names = F,
        top_annotation = ha,
        row_title = sprintf('%s Proteins', mat.size)
        )

BackupAsPDF(hm, 'proteinIntensity.heatmap')
```
PCA plot at the protein level
No strong signal differentiating treatment groups
```{r}
# 'full' matrix
dim(p.mat) #4682 prots
p.mat <- p.mat[complete.cases(p.mat),]
dim(p.mat) #4570 prots


colInfo <- data.table(colname = colnames(p.mat))
colInfo[colname %like% 'M100_Nbon', colname := sub('_', '.', colname)]
colInfo[,c("condition", "rep") := tstrsplit(colname, "[_]",  keep = c(1,2)) ] # split the string and transpose into two columns


#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaDT[rn  %like% 'M100_Nbon', rn := sub('_', '.', rn)]

pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = condition, shape = rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )


BackupAsPDF(p, "complete.proteins.PCA")
```
plot the other PCs
```{r}
barplot(pcaPercentVar)
```
Before performing contrasts between conditions, we want to remove proteins with only 1 identified feature as likely a mislabelling
The `NumMeasuredFeature` column in the msstats `ProteinLevelData`  gives the number of measured features per run/protein

I think for now will record the names of the 'troublesome' proteins and inspect sig results (still issue of MT correction...)
#Drop the proteins with 1 measured feature detected in > 5/16 runs ( 

```{r}
#bar plot of number of measured features per protein/run
# ~6.5 k instances of 1 feature protein
ggplot(p.quant[NumMeasuredFeature %in% seq(1,10,by=1)], aes(x=NumMeasuredFeature)) + 
  geom_bar(stat = 'Count') +
  scale_x_continuous(breaks=seq(0,10, 1), limits=c(0,11)) +
  theme_bw()


# remove all proteins with 1 measured feature
# count of number of proteins with N missing values; we can see that over 300 proteins have only 1 meadured feature in all runs
p.quant[NumMeasuredFeature == 1, .N, by=Protein] %>% 
  ggplot(., aes(x=(N))) + geom_bar(stat='Count')


# what di the proteins with only one measured feature in one run look like?
quick.look <- p.quant[NumMeasuredFeature == 1, .N, by=Protein] %>% .[N %in% c(1),Protein]
# majority of these just have two measured features..
p.quant[Protein %in%  quick.look, NumMeasuredFeature] %>%  summary()

# remove proteins from dt with 1 measured feature in all runs
#Drop the proteins with 1 measured feature detected in 10 or more runs based on hist (increase in )
drop.prots <-  p.quant[NumMeasuredFeature == 1, .N, by=Protein] %>% 
  .[N %in% seq(5,16, by=1), Protein]

# dropping 440 proteins
drop.prots %>%  length()

# inspect the distruibution of measured features in these proteins
# out of near 7k protein/run combos, only 331 have 2 measured features and none more
p.quant[Protein %in% drop.prots, .N,by=NumMeasuredFeature]


# record the other proteins with 1 measured feature per run/protein in up to 4 runs
suspect.prots <-  p.quant[NumMeasuredFeature == 1, .N, by=Protein] %>% 
  .[N %in% seq(1,4, by=1), Protein]

#233 retained in the dataset
suspect.prots %>%  length()

# remove the proteins 
p.quant <-  p.quant[!Protein %in% drop.prots, ]
```

Run MSstats pw comparisons for now

create a contrast matrix for the msstats contrasts
4 conditions, so prob 

ag - ctrl
treat - ctrl
ag_treat - ctrl
ag_treat - treat # key question; what are the things changing 

key question here is...
(ag_treat - ctrl) - (ag_treat - treat) # use emmeans to answer this q

```{r}
# first MRX2 
mat.dimnames <- list(rnames='M100-Veh',
                     cnames= c('M100', 'M100_Nbon', 'Nbon', 'Veh'))

mss.contrasts <- matrix(data=c(1,0,0,-1), nrow = 1, dimnames=mat.dimnames)
mss.contrasts <- rbind(mss.contrasts, c(0,0,1,-1))
rownames(mss.contrasts)[2] <- 'Nbon-Veh'
mss.contrasts <- rbind(mss.contrasts, c(0,1,0,-1))
rownames(mss.contrasts)[3] <- 'M100_Nbon-Veh'
mss.contrasts <- rbind(mss.contrasts, c(0,1,-1,0))
rownames(mss.contrasts)[4] <- 'M100_Nbon-Nbon'

# sanity check
#p.quant$GROUP %>%  unique() %in% colnames(mss.contrasts)
```

```{r}
# use SUBJECT as a batch term to correct for batch effect in linear modelling
p.quant[, SUBJECT := paste0('batch.', SUBJECT)]

# write out p.quant for reuse in contrast of contrasts
#fwrite(p.quant, './output/mss.dp.ProteinLevelData.processed.csv')

# now replace PLD with the modified data.table
dp.out$ProteinLevelData <- setDF(p.quant)

mss <- MSstats::groupComparison(contrast.matrix = mss.contrasts, dp.out)
mss.dt <- setDT(mss$ComparisonResult)

#fwrite(mss.dt, './output/mss.grpcomparison.ori.csv')
```

```{r}
# read in the data
mss.dt <- fread('./output/mss.grpcomparison.ori.csv')
```

plot the DEPs; volcano plots of each contrasts
Get all the DEGs and run GO enrichment; plot heatmaps
write out the results in long format

no significant prots @ log2FC > 50% adj.pval < 0.05, adj.pval < 0.1
pval histograms look un

```{r}
# why is there NA vals in the group comparison result?
mss.dt

proc.dt <- mss.dt[,gene := multiUniprots2multiGenes(as.character(Protein), species='MOUSE')] %>% #needed as in factor format
  .[!issue %in% c('completeMissing','oneConditionMissing'), ] %>% 
  .[!is.na(log2FC), ] %>% 
  .[, p.adj := p.adjust(pvalue, method='BH'), by=.(Label)] %>% 
  .[, sig := 'not'] %>% 
  .[abs(log2FC) > 0.58 & p.adj < 0.1, sig := ifelse(log2FC > 0, 'up', 'down')]

# output the res
fwrite(proc.dt, './output/mss.grpcomparison.processed.csv')

proc.dt[, .N, by=.(sig,Label)]

# what do the pvalue histograms looking like 
# slight spike at low pval for treatment, antagonist vs ctrl. treat+antagonist looks uniform distribution 
ggplot(mss.dt, aes(x=pvalue)) + 
  geom_histogram() +
  facet_wrap(~Label)

proc.dt <- mss.dt[,gene := multiUniprots2multiGenes(as.character(Protein), species='MOUSE')] %>% #needed as in factor format
  .[!issue %in% c('completeMissing','oneConditionMissing'), ] %>% 
  .[!is.na(log2FC), ] %>% 
  .[, p.adj := p.adjust(pvalue, method='BH'), by=.(Label)] %>% 
  .[, sig := 'not'] %>% 
  .[abs(log2FC) > 0.58 & pvalue < 0.01, sig := ifelse(log2FC > 0, 'up', 'down')]

```
generate volcanoplots of the 4 comparisons
nothing at adj.pval < 0.05, so drop threshold to pval < 0.01

```{r}

g <- ggplot(proc.dt, aes(x=log2FC, y=-log10(pvalue), color=sig, label=gene)) + 
      geom_point() +
      geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
      geom_hline(yintercept = -log10(0.05), linetype=2, alpha=0.4) +
      ggrepel::geom_text_repel(data=proc.dt[sig != 'not',],  size = 2, max.overlaps = 20) +
      scale_color_manual(values=c('up'='red', 'down'='blue', 'not'='grey')) +
      ggtitle(' (FC +/ 50% & pval < 0.01)') +
      facet_wrap(~Label, scales = 'free') +
      theme_bw() +
      theme(strip.text.x = element_text(size = 5))
      
g
BackupAsPDF(g, 'grpcontrasts.pval.0.01.multipanel.volcanoplot.')

```
sanity check...
```{r}
ggplot(p.quant[Protein == 'P07759',], aes(x=reorder(GROUP), y=LogIntensities, color=GROUP)) + geom_point() + theme(axis.text.x = element_text(angle=90))
```

Not going to recover much here from the enrichment analysis
lets look at the contrast of contrasts instead and plot these heatmaps per condition

```{r}
p.quant <- fread('./output/mss.dp.ProteinLevelData.processed.csv')

# create a treatment and batch factor for lm 
p.quant[, batch := factor(SUBJECT)]
p.quant[, treatment := factor(GROUP, levels=c('Veh', 'M100','Nbon', 'M100_Nbon'))]


# example of single lm controlling for batch (P07759 used as significant in msstats modelling)
lm(LogIntensities~treatment+batch, data = p.quant[Protein == "P07759"]) %>% summary()

model <- aov(LogIntensities ~ batch+treatment, data=p.quant[Protein == "P07759"])  
summary(model)
```

```{r}
# linearModelsAllProteins can this function which computes statistic for a contrast of two different log2FC
# how does the treatment effect change per antagonist level? basically compare treatment effect within each antagonist level
emmeans.contrastOfContrasts <- function (l, factorFormula = ~treatment){  # how does response overe time change with pretreatment status
  emm <- emmeans::emmeans(l, factorFormula)
  contrast1 <- pairs(emm) # pairwise contrasts of condiitona
  contrast2 <- pairs(contrast1, by = NULL, adjust = "none") # contrasts of the contrasts.
  return (as.data.table(contrast2))
}

# perform pairwise comparisons of each grp
lm.out <- linearModelsAllProteins(p.quant, formulaList = list(basic = LogIntensities~treatment+batch), splitColumn = "Protein", 
                                  cl = 6, emmeansFormula = pairwise~treatment,
                                  postProcessFunction = emmeans.contrastOfContrasts)


```

generally looks uniform, but for Veh-M100_Nbon & Veh-Nbon there seems to be a clear spike at pval < 0.05 although they look very similiar
```{r}
dt <- lm.out[["contrast"]]

g <- ggplot(dt, aes(x=dt$p.t)) +
   geom_histogram(boundary=0.0, binwidth = 0.01) +
   facet_wrap(~dt$contrast)
g
```

write out the significant results of each of the contrasts
again, no significant hits after correction (p.adj < 0.05)
but at pval < 0.01 more significiant results Veh - M100_Nbon compared to Nbon alone vs ctrl

```{r}
dt[, p.adj := p.adjust(p.t, method='BH'), by=contrast]

dt[, gene := multiUniprots2multiGenes(Protein, species = 'MOUSE')]
dt[, sig := 'not']
dt[p.t < 0.001 & abs(estimate) > 0.58, sig := ifelse(estimate > 0, 'up', 'down')]

# only 10 sig hits in total...
dt[sig != 'not',]


dt <- dt[, .(Protein, gene, contrast, estimate, p.t, p.adj, sig)] %>% 
  setnames(new=c('Protein', 'gene', 'contrast', 'log2FC', 'pvalue', 'p.adj', 'sig'))

fwrite(dt, './output/ab.lm.contrasts.csv')
```

Now lets look at interaction analysis for the different hits.

What is the impact of antagonist addition on treatment effect?
"(Veh - Nbon) - (Veh - M100_Nbon)" this is our contrast of contrasts; is there a significant difference between contrasts to control?

```{r}
lm.out$postProcess$contrast %>%  unique()
contrast.oi <- c("(Veh - Nbon) - (Veh - M100_Nbon)", "(Veh - M100) - (Veh - M100_Nbon)" )

# jsut extract the contrast we are interested in
dt <- lm.out$postProcess[contrast %in% contrast.oi,]

dt[, sig := 'not']
dt[, gene := multiUniprots2multiGenes(Protein, species="MOUSE")]
# only 4 genes are significant at pval < 0.001 and thats w/o adjusting log2FC thresholds..
dt[p.value < 0.001, ]

# sanity check 
# addition of M100 agonist leads to reduction
ggplot(p.quant[Protein == 'P23953',], aes(x=(GROUP), y=LogIntensities, color=GROUP)) + geom_point() + theme(axis.text.x = element_text(angle=90))

# this one looks a little more interesting. Seems combo of M100 and Nbon leads to a reduction relative to other groups, but not v large effect size
# Shisa9; involved in neuronal plasticity
ggplot(p.quant[Protein == 'Q9CZN4',], aes(x=(GROUP), y=LogIntensities, color=GROUP)) + geom_point() + theme(axis.text.x = element_text(angle=90))
```



Initial look
----
V few significant changes between conditions 
Few hits (4) found in contrast of contrasts, although Shisa9 may be interesting
Check out the phospho data tomorrow and see if we find any sig hits of interest

