---
title: "DE.ph.redo.lm"
author: "Martin Gordon"
date: "2023-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MSstats 

redo of lm w/o batch
Batch makes no sense in this case as we cannout assume that the mice come from the same batches
Treat each as individual and drop the batch term

want to
---
compare our res with no batch modelling ( n hits)
compare our res with modelling for pseudo-batch

```{r}
library(ggplot2)
library(data.table)
library(magrittr)
library(ComplexHeatmap)
library(randomcoloR)
library(stringr)
library(OmnipathR)
library(circlize)


#source BP utility scripts
source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/MSstats_Helper_Functions.R")
source("/Users/martingordon/Documents/utils/bp_utils/MSstats_V4_Functions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/linearModels.R")

# kinase enrichment analysis
source("/Users/martingordon/Documents/utils/bp_utils/KinaseActivityScores.R")

# mg utility functions
source("/Users/martingordon/Documents/utils/mg_utils/HelperFunctions.R")
```

load in the previous mss data 
```{r}
ori.mss <- fread('./DE.ph.291123_data/2023_11_29_ph.mss.grpcomparison.processed.csv')

ori.mss[sig != 'not',.N, by=.(sig,Label)] 
```
```{r}
msinput <- fread("./output/MSstatsInput.PH.csv")

dp.out <- MSstats::dataProcess(setDF(msinput),
                               MBimpute = FALSE,
                               featureSubset = "highQuality",
                               remove_uninformative_feature_outlier = TRUE)

#aveRDS(dp.out, ScriptAndDatedFileName('ph.mss.dataproc'))
```
```{r}
contrast.mat <- makeContrast.regEx(dp.out,c( 
                               "^M100-Veh",
                               "^Nbon-Veh",
                               "^M100_Nbon-Veh",
                               "^M100_Nbon-Nbon",
                               "^M100_Nbon-M100"
                               ))
```

run mss pw comparisons between groups
```{r}
dp.out <- readRDS('~/Documents/projects/112823_XZhong_DARPA2/DE.ph.LMredo.noBatch_data/2023_11_30_ph.mss.dataproc')

p.quant<- setDT(dp.out$ProteinLevelData)
p.quant[, SUBJECT := paste0(GROUP,'.',SUBJECT)]

dp.out$ProteinLevelData <- p.quant

gc <- MSstats::groupComparison(contrast.mat, dp.out)
```

```{r}
nobatch.mss.dt <- setDT(gc$ComparisonResult)

nobatch.proc.dt <- nobatch.mss.dt[,gene := multiUniprotSites2multiGeneSites(Protein, species='MOUSE')] %>% #needed as in factor format
  .[!issue %in% c('completeMissing','oneConditionMissing'), ] %>% 
  .[is.finite(log2FC) & !is.na(pvalue), ] %>% 
  .[, p.adj := p.adjust(pvalue, method='BH'), by=.(Label)] %>% 
  .[, sig := 'not'] %>% 
  .[abs(log2FC) > 0.58 & pvalue < 0.005, sig := ifelse(log2FC > 0, 'up', 'down')]


fwrite(nobatch.proc.dt, ScriptAndDatedFileName('mss.pwcomparisons.nobatch.csv'))

# v few sig hits a adj.pval < 0.05 & log2FC +/- 50%, so dropped to pval < 0.001
# 10 more hits removing batch term
nobatch.proc.dt[sig != 'not', .N, by=.(sig,Label)]$N %>% sum()
ori.mss[sig != 'not',.N, by=.(sig,Label)]$N %>% sum()

nobatch.proc.dt[sig != 'not', .N, by=.(sig,Label)]
ori.mss[sig != 'not',.N, by=.(sig,Label)]
```

enrichment analysis of the non-batch corrected data
```{r}
# load mouse gmt file 
gmt.go <- loadGmtFromBioconductor(dbName='org.Mm.eg.db', ontology = "ALL", keyType = "UNIPROT")

p.quant <-  fread('./DE.ph.291123_data/2023_11_29_MSstatsInput.csv')
# get matching col
p.quant[, Protein := gsub("_[STY][0-9]+", "",ProteinName)]

universe <-  p.quant$Protein %>%  unique()
length(universe) # ~6.2k

# get a simplifed dt of gene names of the moving sites
gene.dt <-  nobatch.proc.dt[sig != 'not', ] %>% 
  .[, ProteinName := gsub("_[STY][0-9]+", "",Protein)] %>% 
  .[, .(Label, ProteinName, sig)] %>% 
  .[order(Label,ProteinName)]

# create an ernich grouping col
gene.dt[, enrich_grp := paste0(Label, '.', sig)]


enrich.dt <- enricherOnGroups(gene.dt, groupColumns = 'enrich_grp', geneColumn = "ProteinName", term2gene.gmt = gmt.go, universe = universe)

simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')


ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = enrich.dt, groupColumn = 'enrich_grp', topN = 8, title='GO term enrichment', 
                                  negCols=unique(simp.enrich$simplified$enrich_grp[grep('down', simp.enrich$simplified$enrich_grp)]), 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 8), upperThreshold = 4)
BackupAsPDF(ht, 'go.enrichment.msstats.pwgroups.nobatchCor.heatmap.')
```
what about kinase enrichment for this group?
```{r}

# modified function to avoid omnipath kinase download and read from file
loadKinaseDataOmniPath <- function(species = "HUMAN", removeNonKinases = TRUE, fixNames = TRUE, omniPathFile = './data/mouse.omnipath.csv', ...){
  species <- toupper(species)
  organismID <- c(HUMAN = 9606, RAT = 10116, MOUSE = 10090)[species]
  if(is.na(organismID))
    stop(species, " not a recognized organism")
  enzsub <- fread(omniPathFile)
  setDT (enzsub)
  # for consistency with loadKinaseData I need columns: "CTRL_GENE_NAME", "TARGET_GENE_NAME", "TARGET_RES", "TARGET_POS", "TARGET_UP_ID"
  
  if (any(grepl("ProtMapper", enzsub$sources))){
    message ("OmniPath includes ProtMapper data which has many non-kinases.")
    
    if(removeNonKinases == TRUE) {
      message ("Removing all enzymes that are not kinases according to org.Hs.eg.db GO:0016301")
      # get GO kinase list
      
      if (species == "HUMAN"){
        library(org.Hs.eg.db)
        kinaseTable <- AnnotationDbi::select (org.Hs.eg.db, get(c("GO:0016301"), org.Hs.egGO2ALLEGS), c("ENTREZID", "GENENAME", "SYMBOL")) 
      } else if (species == "MOUSE"){
        library(org.Mm.eg.db)
        kinaseTable <- AnnotationDbi::select (org.Mm.eg.db, get(c("GO:0016301"), org.Mm.egGO2ALLEGS), c("ENTREZID", "GENENAME", "SYMBOL")) 
      } else{
        stop ("I don't yet know how to find kinases for ", species)
      }
      
      kinases <- unique(kinaseTable$SYMBOL)
      #label and filter into a new table
      enzsub[, is_kinase := enzyme_genesymbol %in% kinases]
      enzsub <- enzsub[is_kinase == TRUE]
    }
  }
  
  if (fixNames){
    setnames(enzsub, 
             old = c("enzyme_genesymbol", "substrate_genesymbol", "residue_type", "residue_offset", "substrate"), 
             new = c("CTRL_GENE_NAME", "TARGET_GENE_NAME", "TARGET_RES", "TARGET_POS", "TARGET_UP_ID"))
  }
  
  return (enzsub[modification == "phosphorylation"])
}

kinaseData  <- loadKinaseDataOmniPath(species='MOUSE')

# just look at enrichment of the complete set
mss.dt <-  nobatch.proc.dt


mss.dt[,gene := multiUniprotSites2multiGeneSites(Protein, species = "MOUSE")]
singleSiteResults <- prepare_AMSS_ResultsFile(mss.dt, column = "gene")



labels <- unique(singleSiteResults$Label)

kinActList <- lapply (labels, FUN=function(lab){kinaseActivity(singleSiteResults[Label == lab & representative==TRUE],
                                                               plots = FALSE,
                                                               kinaseData = kinaseData, 
                                                               do.sea = TRUE)})


names(kinActList) <- labels

kinActFull.scores <- rbindlist(lapply(kinActList, FUN = function(x)x$scores), idcol="Label")
kinActFull.mapped <- rbindlist(lapply(kinActList, FUN = function(x)x$kinaseMapped)) # Label is already in these tables


fwrite(kinActFull.scores, ScriptAndDatedFileName("kinaseSummaryScores.csv"))
fwrite(kinActFull.mapped, ScriptAndDatedFileName("kinaseSubstrateData.csv"))
message (sprintf("Writing kinase output files to\n\t%s\n\t%s", kinaseSummaryScores.csv, kinaseSubstrateData.csv))
```

## Heatmap of significant kinases
less hits than ori run... leave as is for now...
```{r, fig.width = 6, fig.height= 4}
sigKinases <-  kinActFull.scores[fdr.BH < 0.05 & N > 2, unique(CTRL_GENE_NAME)]

sigKinase.mat.z <- as.matrix(dcast (kinActFull.scores[CTRL_GENE_NAME %in% sigKinases], CTRL_GENE_NAME~Label, value.var = "Z"),
                              rownames = "CTRL_GENE_NAME")

sigKinase.mat.N <- as.matrix(dcast (kinActFull.scores[CTRL_GENE_NAME %in% sigKinases], CTRL_GENE_NAME~Label, value.var = "N"),
                                  rownames = "CTRL_GENE_NAME")


hm <- Heatmap (sigKinase.mat.z, 
         cluster_columns=TRUE, 
         name = "Kinase Z Score",
         #column_split = tstrsplit(colnames(sigKinase.mat.z), split="[-_.]")[[1]],
         col = circlize::colorRamp2(breaks = c(-3, -1, 1, 3), colors = c("blue", "gray", "gray", "red")),  # two midpoints in case you want to have a wider gray bar around 0
         row_title = '',
         cell_fun = function(j, i, x, y, width, height, fill) {
                                  if (!is.na(sigKinase.mat.N[i,j])){
                                    grid.text(sprintf("%.0f", sigKinase.mat.N[i, j]), x, y, gp = gpar(fontsize=10, col="white"))
                                  }
                                })

hm
BackupAsPDF(hm, prefix = "Heatmap_SigKinases")

hm <- Heatmap (sigKinase.mat.z, 
         cluster_columns=FALSE, 
         name = "Kinase Z Score",
        # column_split = tstrsplit(colnames(sigKinase.mat.z), split="[-_.]")[[1]],
         col = circlize::colorRamp2(breaks = c(-3, -1, 1, 3), colors = c("blue", "gray", "gray", "red")),  # two midpoints in case you want to have a wider gray bar around 0
         row_title = '',
         cell_fun = function(j, i, x, y, width, height, fill) {
                                  if (!is.na(sigKinase.mat.N[i,j])){
                                    grid.text(sprintf("%.0f", sigKinase.mat.N[i, j]), x, y, gp = gpar(fontsize=10, col="white"))
                                  }
                                })
hm
BackupAsPDF(hm, prefix = "Heatmap_SigKinases_noColCluster")
```

what if we look at kinase enrichment for the pseudobatch data?

```{r}

# just look at enrichment of the complete set
mss.dt <-  nobatch.proc.dt

mss.dt[,gene := multiUniprotSites2multiGeneSites(Protein, species = "MOUSE")]
singleSiteResults <- prepare_AMSS_ResultsFile(mss.dt, column = "gene")



labels <- unique(singleSiteResults$Label)

kinActList <- lapply (labels, FUN=function(lab){kinaseActivity(singleSiteResults[Label == lab & representative==TRUE],
                                                               plots = FALSE,
                                                               kinaseData = kinaseData, 
                                                               do.sea = TRUE)})


names(kinActList) <- labels

kinActFull.scores <- rbindlist(lapply(kinActList, FUN = function(x)x$scores), idcol="Label")
kinActFull.mapped <- rbindlist(lapply(kinActList, FUN = function(x)x$kinaseMapped)) # Label is already in these tables

fwrite(kinActFull.scores, ScriptAndDatedFileName("kinaseSummaryScores.csv"))
fwrite(kinActFull.mapped, ScriptAndDatedFileName("kinaseSubstrateData.csv"))
```


10 less hits modelling w/o batch term (treat the samples as individuals..). Check enrichment
create pseudobatch corresponding to pca clusters and rerun the lm
```{r}
dp.out <- readRDS('~/Documents/projects/112823_XZhong_DARPA2/DE.ph.LMredo.noBatch_data/2023_11_30_ph.mss.dataproc')

p.quant$OLD_SUBJECT %>%  unique()

p.quant<- setDT(dp.out$ProteinLevelData)
p.quant[, OLD_SUBJECT := paste0(GROUP,'.',SUBJECT)]

p.quant$OLD_SUBJECT %>%  unique()

p.quant[, SUBJECT := fcase(
  OLD_SUBJECT %in% c('Nbon.3', 'Nbon.2', 'M100.4', 'M100_Nbon.1'), 'pseudobatch.1',
  OLD_SUBJECT %in% c('M100_Nbon.2', 'M100_Nbon.4'), 'pseudobatch.2',
  OLD_SUBJECT %in% c('M100_Nbon.3', 'Veh.4','Veh.2','M100.1'), 'pseudobatch.3',
  OLD_SUBJECT %in% c('Veh.1', 'Veh.3', 'M100.3', 'M100.2', 'Nbon.4', 'Nbon.1'), 'psuedobatch.4'
)]

# sanity check
p.quant[,.N, by=.(SUBJECT, OLD_SUBJECT)] %>% 
  .[order(SUBJECT)]

```
rerun lm
```{r}
dp.out$ProteinLevelData <- p.quant
gc <- MSstats::groupComparison(contrast.mat, dp.out)
```

Look at the DEP output after correction

```{r}
pseudobatch.mss.dt <- setDT(gc$ComparisonResult)

pseudobatch.proc.dt <- pseudobatch.mss.dt[,gene := multiUniprotSites2multiGeneSites(Protein, species='MOUSE')] %>% #needed as in factor format
  .[!issue %in% c('completeMissing','oneConditionMissing'), ] %>% 
  .[is.finite(log2FC) & !is.na(pvalue), ] %>% 
  .[, p.adj := p.adjust(pvalue, method='BH'), by=.(Label)] %>% 
  .[, sig := 'not'] %>% 
  .[abs(log2FC) > 0.58 & pvalue < 0.005, sig := ifelse(log2FC > 0, 'up', 'down')]

# v few sig hits a adj.pval < 0.05 & log2FC +/- 50%, so dropped to pval < 0.001
# 10 more hits removing batch term
nobatch.proc.dt[sig != 'not', .N, by=.(sig,Label)]$N %>% sum()
ori.mss[sig != 'not',.N, by=.(sig,Label)]$N %>% sum()
pseudobatch.proc.dt[sig != 'not',.N, by=.(sig,Label)]$N %>% sum()

# some difference in hits, but not many
nobatch.proc.dt[sig != 'not', Protein] %in% ori.mss[sig != 'not', Protein]
pseudobatch.proc.dt[sig != 'not', Protein] %in% ori.mss[sig != 'not', Protein]


pseudobatch.proc.dt
```

enrichment analysis of the non-batch corrected data
no real change in res just leave as is
```{r}
# load mouse gmt file 
gmt.go <- loadGmtFromBioconductor(dbName='org.Mm.eg.db', ontology = "ALL", keyType = "UNIPROT")

p.quant <-  fread('./DE.ph.291123_data/2023_11_29_MSstatsInput.csv')
# get matching col
p.quant[, Protein := gsub("_[STY][0-9]+", "",ProteinName)]

universe <-  p.quant$Protein %>%  unique()
length(universe) # ~6.2k

# get a simplifed dt of gene names of the moving sites
gene.dt <-  pseudobatch.proc.dt[sig != 'not', ] %>% 
  .[, ProteinName := gsub("_[STY][0-9]+", "",Protein)] %>% 
  .[, .(Label, ProteinName, sig)] %>% 
  .[order(Label,ProteinName)]

# create an ernich grouping col
gene.dt[, enrich_grp := paste0(Label, '.', sig)]


enrich.dt <- enricherOnGroups(gene.dt, groupColumns = 'enrich_grp', geneColumn = "ProteinName", term2gene.gmt = gmt.go, universe = universe)
enrich.dt[p.adjust < 0.05, .N, by=enrich_grp]
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')


ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, NULL, groupColumn = 'enrich_grp', topN = 8, title='GO term enrichment', 
                                  negCols=unique(simp.enrich$simplified$enrich_grp[grep('down', simp.enrich$simplified$enrich_grp)]), 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 8), upperThreshold = 4)
#BackupAsPDF(ht, 'go.enrichment.msstats.pwgroups.heatmap.')

```
what about the linear modelling?

Try rerun, but w/o batch term

```{r}
dp.out <- readRDS('~/Documents/projects/112823_XZhong_DARPA2/DE.ph.LMredo.noBatch_data/2023_11_30_ph.mss.dataproc')

p.quant<- setDT(dp.out$ProteinLevelData)
p.quant[, SUBJECT := paste0(GROUP,'.',SUBJECT)]


# now want 3 factors: batch, treatment, antagonist in model
p.quant[, `:=`(ant = ifelse(GROUP %like% 'M100', 'M100', 'No_AG'),
               treat = ifelse(GROUP %like% 'Nbon', 'Nbon', 'No_Treat') )]

# example of single lm controlling for batch (P07759 used as significant in msstats modelling)
lm(LogIntensities~treat+ant+batch+ant:treat, data = p.quant[Protein == "Q9D0E1_S452"]) %>% summary()
model <- aov(LogIntensities ~ treat+ant+batch+ant:treat, data=p.quant[Protein == "Q9D0E1_S452"])  
summary(model)
```

rerun lm function with two models; reduced(no batch) and batch
see if the addition of batch impacts res

```{r}
# linearModelsAllProteins can this function which computes statistic for a contrast of two different log2FC
# puts pval and log2FC on sig differences of differences
emmeans.contrastOfContrasts <- function (l, factorFormula = ~treat|ant){  # how does treatment effect change per agonist level
  emm <- emmeans::emmeans(l, factorFormula)
  contrast1 <- pairs(emm)
  contrast2 <- pairs(contrast1, by = NULL, adjust = "none")
  return (as.data.table(contrast2))
}


  
  # Numbers near the number of processors available in your machine is probably counterproductive.
twoModel.lm <- linearModelsAllProteins(p.quant, 
                            formulaList = list(full   = (LogIntensities~treat+ant+ ant:treat)),
                            splitColumn = "Protein", postProcessFunction = emmeans.contrastOfContrasts, cl = 6) %>%  suppressWarnings()
```

modelling 'pseudobatch' removes this htr2a from the significant results

```{r}
sig.prots <- twoModel.lm$anova[term %in% 'treat:ant' & p.value < 0.005]

sig.prots[, gene := multiUniprotSites2multiGeneSites(Protein, species='MOUSE')]

p.mat <- dcast(p.quant, Protein~GROUP+SUBJECT, value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')

rownames(p.mat) <- multiUniprotSites2multiGeneSites(rownames(p.mat), species='MOUSE')

subMat <-  p.mat[rownames(p.mat) %in% sig.prots$gene,]
subMat <-  sweep(subMat, 1, FUN='-', STATS = apply(subMat, 1, median, na.rm=T))

sig.prots[,gene]

hm <- Heatmap(subMat, 
              cluster_columns = FALSE, 
              name = 'log Ints/Median',
              cluster_rows = clusterwNA(subMat),
              #column_title = '(M100+Nbon vs Veh) - (M100 vs Veh)',
              row_title = paste0(nrow(subMat), ' sites p.val < 0.005)'),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 6),
              col=colorRamp2(breaks=c(-2,0,2), colors=c('blue', 'white', 'red')),
              column_split = column.vec,
              row_title_gp = gpar(fontsize=9),
              column_title_gp = gpar(fontsize=10))

BackupAsPDF(draw(hm, column_title = "Differential Response to NBON w/wo M100"),'m100.nbon.vs.nbon.interaction.nobatch.mediansweep.clean.heatmap', dimensions = c(8,8))


# droping
subMat <-  p.mat[rownames(p.mat) %in% sig.prots$gene,]
subMat <-  sweep(subMat, 1, FUN='-', STATS = apply(subMat, 1, median, na.rm=T))

missing.prots <- rownames(subMat)[apply(subMat, 1, function(x){ sum(is.na(x)) >= 8})]
missing.prots

cleanMat <- subMat[!rownames(subMat) %in% missing.prots,]
dim(cleanMat)

hm <- Heatmap(cleanMat, 
              cluster_columns = FALSE, 
              name = 'log Ints/Median',
              cluster_rows = clusterwNA(cleanMat),
              #column_title = '(M100+Nbon vs Veh) - (M100 vs Veh)',
              row_title = paste0(nrow(cleanMat), ' sites p.val < 0.005)'),
              row_names_gp = gpar(fontsize = 5.5),
              column_names_gp = gpar(fontsize = 6),
              col=colorRamp2(breaks=c(-2,0,2), colors=c('blue', 'white', 'red')),
              column_split = column.vec,
              row_title_gp = gpar(fontsize=9),
              column_title_gp = gpar(fontsize=10))

hm


BackupAsPDF(draw(hm, column_title = "Differential Response to NBON w/wo M100"),'m100.nbon.vs.nbon.interaction.nobatch.cleanMatrmMissingVals.mediansweep.clean.heatmap', dimensions = c(8,8))

#BackupAsPDF(draw(hm, column_title = "Differential Response to NBON w/wo M100"),'m100.nbon.vs.nbon.interaction.metabatch.mediansweep.clean.heatmap', dimensions = c(8,8))

```


plot of HTR2A w 

```{r}
dp.out <- readRDS('~/Documents/projects/112823_XZhong_DARPA2/DE.ph.LMredo.noBatch_data/2023_11_30_ph.mss.dataproc')

p.quant<- setDT(dp.out$ProteinLevelData)
p.quant[, SUBJECT := paste0(GROUP,'.',SUBJECT)]
p.quant[, gene := multiUniprotSites2multiGeneSites(Protein, species='MOUSE')]

sites.oi <- c('Htr2a_S298','Htr2a_S421')


p.quant[gene %like% sites.oi,]


g <- ggplot(p.quant[gene %in% sites.oi,], aes(x=GROUP, y=LogIntensities, color=GROUP)) +
  geom_point() + 
  stat_summary(fun.dat='mean_cl_boot', color='grey', size=0.2) +
  scale_color_manual(values=col.pal) +
  ggtitle('HTR2A PH sites') +
  facet_wrap(~gene, scales = 'free_y') +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, vjust = 0.7))
g
#BackupAsPDF(g, 'htr2a.sites.oi.linechart',dimensions=c(8,6))




```
```{r}
# combine this with the mss out.dt for this data
nobatch.proc.dt[gene %in% sites.oi,]

nobatch.wide <- dcast(nobatch.proc.dt, gene+Protein~Label, value.var=c('log2FC', 'pvalue', 'p.adj', 'sig'))
fwrite(nobatch.wide, 'mss.pwcomparisons.nobatchCorrection.csv')


fwrite(nobatch.wide, ScriptAndDatedFileName('mss.pwcomparisons.nobatchCorrection.wide.csv'))
```











Not used

what about no batch term? Does this rm htr2a?

```{r}
# linearModelsAllProteins can this function which computes statistic for a contrast of two different log2FC
# puts pval and log2FC on sig differences of differences
emmeans.contrastOfContrasts <- function (l, factorFormula = ~treat|ant){  # how does treatment effect change per agonist level
  emm <- emmeans::emmeans(l, factorFormula)
  contrast1 <- pairs(emm)
  contrast2 <- pairs(contrast1, by = NULL, adjust = "none")
  return (as.data.table(contrast2))
}


  
  # Numbers near the number of processors available in your machine is probably counterproductive.
twoModel.lm <- linearModelsAllProteins(p.quant, 
                            formulaList = list(full   = (LogIntensities~treat+ant+ ant:treat)),
                            splitColumn = "Protein", postProcessFunction = emmeans.contrastOfContrasts, cl = 6) %>%  suppressWarnings()
```




```{r}
nobatch.proc.dt[gene %like% 'Htr2a', ]
pseudobatch.proc.dt$gene

nobatch.proc.dt[sig != 'not', gene]  %>% length()
pseudobatch.proc.dt[sig != 'not', gene] %in% nobatch.proc.dt[sig != 'not', gene]



```

Create the plots for the msstats plotting w/o batch modelling

```{r}
nobatch.proc.dt$Label %>%  unique()
```

```{r}
g <- ggplot(nobatch.proc.dt, aes(x=log2FC, y=-log10(pvalue), color=sig, label=gene)) + 
      geom_point() +
      geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
      geom_hline(yintercept = -log10(0.005), linetype=2, alpha=0.4) +
      ggrepel::geom_text_repel(data=nobatch.proc.dt[sig != 'not',],  size = 2, max.overlaps = 20) +
      scale_color_manual(values=c('up'='red', 'down'='blue', 'not'='grey')) +
      ggtitle(' (FC +/ 50% & pval < 0.005)') +
      facet_wrap(~Label, scales = 'free') +
      theme_bw() +
      theme(strip.text.x = element_text(size = 5))
      
BackupAsPDF(g, 'ph.grpcontrasts.nobatchcor.pval.0.005.multipanel.volcanoplot.', dimensions=c(12,8))
```
plot them individually 

```{r}
nobatch.proc.dt$Label %>%  unique()

labels.oi <- nobatch.proc.dt[, unique(Label)]

for (lab in labels.oi){
  
  g <- ggplot(nobatch.proc.dt[Label == lab,], aes(x=log2FC, y=-log10(pvalue), color=sig, label=gene)) + 
      geom_point() +
      geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
      geom_hline(yintercept = -log10(0.005), linetype=2, alpha=0.4) +
      ggrepel::geom_text_repel(data=nobatch.proc.dt[Label == lab & sig != 'not',],  size = 2, max.overlaps = 20) +
      scale_color_manual(values=c('up'='red', 'down'='blue', 'not'='grey')) +
      ggtitle(paste0(lab, ' (FC +/ 50% & pval < 0.005)')) +
      theme_bw() +
      theme(strip.text.x = element_text(size = 5))

  BackupAsPDF(g, paste0(lab, '.nobatchCor.pval.0.005.volcano'))
}
```
```{r}

```


