---
title: "htr2a.plots"
author: "Martin Gordon"
date: "2023-12-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Generate plots of htr2a for DARPA meeting 4/12/23

```{r}
library(ggplot2)
library(data.table)
library(magrittr)
library(ComplexHeatmap)
library(randomcoloR)
library(stringr)
library(OmnipathR)
library(circlize)
library(ggsignif)
library(ggpubr)



#source BP utility scripts
source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/MSstats_Helper_Functions.R")
source("/Users/martingordon/Documents/utils/bp_utils/MSstats_V4_Functions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/linearModels.R")

# kinase enrichment analysis
source("/Users/martingordon/Documents/utils/bp_utils/KinaseActivityScores.R")

# mg utility functions
source("/Users/martingordon/Documents/utils/mg_utils/HelperFunctions.R")
```



```{r}
col.pal <- randomcoloR::distinctColorPalette(5)


mss.dt <- fread('~/Documents/projects/112823_XZhong_DARPA2/DE.ph.LMredo.noBatch_data/2023_11_30_mss.pwcomparisons.nobatch.csv')
  
dp.out <- readRDS('~/Documents/projects/112823_XZhong_DARPA2/DE.ph.LMredo.noBatch_data/2023_11_30_ph.mss.dataproc')
p.quant<- setDT(dp.out$ProteinLevelData)
p.quant[, gene := multiUniprotSites2multiGeneSites(Protein, species='MOUSE')]

sites.oi <- c('Htr2a_S298','Htr2a_S421')

g <- ggplot(p.quant[gene %in% sites.oi,], aes(x=GROUP, y=LogIntensities, color=GROUP)) +
  geom_point() + 
  stat_summary(fun.dat='mean_cl_boot', color='grey', size=0.2) +
  scale_color_manual(values=col.pal) +
  ggtitle('HTR2A PH sites') +
  facet_wrap(~gene) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, vjust = 0.7))
g
BackupAsPDF(g, 'htr2a.sites.oi.linechart.fixedxaxis',dimensions=c(8,6))
```
another option; take the log2FC values per site, combine with SE estimates and overlay
rotate and do a lollipop chart
```{r}
lab.col.pal <- randomcoloR::distinctColorPalette(5)

g <- mss.dt[gene %in% sites.oi,] %>% 
  ggplot(., aes(x=Label, y=log2FC, color=Label)) +
  geom_segment(aes(x=Label, xend=Label, y=0, yend=log2FC), color='grey', lwd=.5, linetype=1) +
  geom_point(size=3) +
  geom_hline(yintercept=0, alpha=0.4, linetype=2) +
  facet_wrap(~gene) +
  ylim(c(-1,2)) +
  theme_bw() + 
  #coord_flip() +
  scale_color_manual(values=c(lab.col.pal)) +
  theme(
    axis.text.x = element_text(angle=90, vjust=0.6)
  )

BackupAsPDF(g, 'htr2a.lollipop.basic.')

# annotate with pval if < 0.05
mss.dt[, lab := '']
mss.dt[pvalue < 0.05, lab := ifelse(pvalue < 0.01, '**', '*')]
mss.dt[pvalue < 0.001, lab := '***']

g <- mss.dt[gene %in% sites.oi,] %>% 
  ggplot(., aes(x=Label, y=log2FC, color=Label, label=paste0(lab, ' ', round(pvalue, 3) )) ) +
  geom_segment(aes(x=Label, xend=Label, y=0, yend=log2FC), color='grey', lwd=.5, linetype=1) +
  geom_point(size=3) +
  geom_hline(yintercept=0, alpha=0.4, linetype=2) +
  geom_text(data=mss.dt[gene %in% sites.oi & pvalue < 0.05], color='darkgrey', size=3, vjust=-2) +
  facet_wrap(~gene) +
  ylim(c(-1,2)) +
  theme_bw() + 
  #coord_flip() +
  scale_color_manual(values=c(lab.col.pal)) +
  theme(
    axis.text.x = element_text(angle=90, vjust=0.6)
  )

g

BackupAsPDF(g, 'htr2a.lollipop.pvals.')

```
include error bars as we want an estimate of SE on our plot

```{r}
g <- mss.dt[gene %in% sites.oi,] %>% 
  ggplot(., aes(x=Label, y=log2FC, color=Label, label=paste0(lab, ' ', round(pvalue, 4) )) ) +
  geom_segment(aes(x=Label, xend=Label, y=0, yend=log2FC), color='grey', lwd=.5, linetype=1) +
  geom_point(size=3) +
  geom_hline(yintercept=0, alpha=0.4, linetype=2) +
  geom_text(data=mss.dt[gene %in% sites.oi & pvalue < 0.05], color='darkgrey', size=3, vjust=-2) +
  facet_wrap(~gene) +
  ylim(c(-1,2)) +
  theme_bw() + 
  coord_flip() +
  scale_color_manual(values=c(lab.col.pal)) +
  theme(
    axis.text.x = element_text(angle=90, vjust=0.6)
  )

g
```


```{r}

g <- mss.dt[gene %in% sites.oi,] %>% 
  ggplot(., aes(x=Label, y=log2FC, color=Label, label=paste0(lab, ' ', round(pvalue, 3) )) ) +
  #geom_segment(aes(x=Label, xend=Label, y=0, yend=log2FC), color='grey', lwd=.5, linetype=1) +
  geom_hline(yintercept=0, alpha=0.4, linetype=2) +
  geom_text(data=mss.dt[gene %in% sites.oi & pvalue < 0.05], color='darkgrey', font='bold', size=3.8, vjust=-4) +
  geom_errorbar(data=mss.dt[gene %in% sites.oi,], aes(ymin=log2FC-SE, ymax=log2FC+SE, group=Label), width=0.2, lwd=.5, color='darkgrey') +
  geom_point(size=3) +
  facet_wrap(~gene) +
  ylim(c(-1.1,2.2)) +
  theme_bw() + 
  #coord_flip() +
  scale_color_manual(values=c(lab.col.pal)) +
  theme(
    axis.text.x = element_text(angle=90, vjust=0.6)
  )
g
```
last attempt; take the original plot and just use geom_
look into why not working...

one option I am thinking is
```{r}
# an annotation dataf 
annotation_data <- data.table(
  GROUP = c('M100', 'M100_Nbon', 'Nbon', 'Veh'),
  xend = c('M100_Nbon', 'M100_Nbon', 'Nbon', 'M100_Nbon'),
  y = c(15.4, 15, 15.5, 15.3),
  annotation = c("**", "", "", "*"),
  gene = rep(sites.oi[1], 4)
)

g <- ggplot(p.quant[gene %in% sites.oi,], aes(x=GROUP, y=LogIntensities, color=GROUP)) +
  geom_boxplot(color='grey') +
  geom_point() + 
  geom_signif(stat="identity",
             data=annotation_data,
            aes(x=GROUP, y=y, xend=xend, yend=y, annotation=annotation),color='darkgrey', size=4) +
  scale_color_manual(values=col.pal) +
  ggtitle('HTR2A PH sites') +
  facet_wrap(~gene) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, vjust = 0.7))
g

BackupAsPDF(g, 'htr2a.S298.sites.siglabelled.boxplots.')


# try modifying slightly the GROUP name
```
```{r}
g <- ggplot(p.quant[gene %in% sites.oi,], aes(x=GROUP, y=LogIntensities, color=GROUP)) +
  geom_point() + 
  stat_summary(fun.dat='mean_cl_boot', color='grey', size=0.2) +
  scale_color_manual(values=col.pal) +
  ggtitle('HTR2A PH sites') +
  facet_wrap(~gene) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, vjust = 0.7))
g
```



show boxplots of the sites with the sig difference
```{r}

stat.test <- data.table(group1=c('M100_Nbon', 'M100_Nbon'),
                        group2=c('M100', 'Veh'),
                        pval=c('** 0.002', '* 0.041'))

g <- ggboxplot(p.quant[gene == 'Htr2a_S421',], x="GROUP", y="LogIntensities", color="GROUP", palette='simpsons', title='Htr2a_S421') 
g
BackupAsPDF(g, 'htr2a_S421.boxplots.pvalsInc.')
```

