---
title: "MOFA-omics-int"
author: "Martin Gordon"
date: "2023-06-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MOFA first run
 big N*P problem; we don't have enough samples (recommended at least 15 per cond for results to be meaningful) and far too many features
For now, we will use the top n (500) features used for the network propagation analysis
Features were ranked on adj.pvalue log2FC geometric mean, take these gene names and subset for each of the datasets and rerun

##Impressions:
Only 3 factors learned, factor 1 explained most of the variance in the data, but only factor 2 seemed to show some correlation with condition
When I ran enrichment on the output found no signnificant hits

## R Markdown

## Packages

```{r}
library(RColorBrewer) 
library(mixOmics)
library(data.table)
library(stringr) #str_extract function 
library(magrittr)
library(ggplot2)
library (ComplexHeatmap)
library(RcppCNPy) #R/W support to import numpy objects into R
library(VennDiagram)
library(pbapply) #apply functions with progress bar
library(eulerr)
library(MOFA2) #this is the tool used for 
library(MOFAdata)

library(DESeq2)

# load bens util scripts for the analysis
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/LocatePeptidePtmInProteins.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

# netprop scripts
source ("../../utils/bp_utils/STRING_db_utils.R")
source ("../../utils/bp_utils/NetworkPropagation.R")


# using pyhton installation with correct architecute
#Sys.setenv(RETICULATE_PYTHON = "/Users/martingordon/anaconda3/bin/python")
#library(reticulate)
#reticulate::py_config()

```

TO DO
---
Create data matricies for the 3 omics technologies (raw data)


First lets view the raw data:
```{r}
utils::data("CLL_data")       
lapply(CLL_data,dim)
```

Lets import our own data and generate a matrix
For now use all samples as we don't have enough per condition

Protein RNA & PH data
```{r}
p.ab.quant <- fread('~/Documents/projects/051623_DSwaney/2023_05_16_DSwaney_AA08_AB_data/2023_05_25_ProteinLevelData.csv.gz') 
p.ph.quant <- fread('~/Documents/projects/060123_DSwaney_AshworthPH/060123_DSwaney_AshworthPH_data/2023_06_01_protein-lvl-data.csv.gz')
rna.quant <- fread('~/Library/CloudStorage/Box-Box/Mac1_omics_data/CellLines_RNAseq/Expression/gene_expression.xls')
p.ph.quant
box.loc <- '~/Library/CloudStorage/Box-Box/Mac1_omics_data/'
rna.ab.core <- fread(paste0(box.loc, 'CellLines_RNAseq/core_table_gene.xls.gz'), fill=T)

cols.OI <- grep("^read_count|gene_symbol",colnames(rna.ab.core), value=T)

rna.mat <- rna.ab.core[, colnames(rna.ab.core) %in% cols.OI, with=F] %>% 
  .[,-25, with=F] #rm dup gene_symbol

```

Convert to matrix and adjust colnames
```{r}
p.ph.mat <- dcast(p.ph.quant, Protein~paste0(GROUP,"_",SUBJECT), value.var = 'LogIntensities')
p.ab.mat <- dcast(p.ab.quant, Protein~paste0(GROUP, "_", SUBJECT), value.var = 'LogIntensities')

colnames(rna.mat) <- gsub('read_count_','',colnames(rna.mat))

# set up colnames to match rna-seq dataset
colnames(p.ph.mat) <- c("Protein", "EV_IFNg_A",  "EV_IFNg_B",   "EV_IFNg_C",  "EV_IFNg_D",  "ND_IFNg_A",   "ND_IFNg_B",   "ND_IFNg_C", "ND_IFNg_D", "Par_IFNg_A", "Par_IFNg_B",  "Par_IFNg_C", "Par_IFNg_D", "Par_ctrl_A",  "Par_ctrl_B",  "Par_ctrl_C",  "Par_ctrl_D", "WT_IFNg_A",   "WT_IFNg_B",   "WT_IFNg_C", "WT_IFNg_C") 
colnames(p.ab.mat) <- c("Protein", "EV_IFNg_A",  "EV_IFNg_B",   "EV_IFNg_C",  "EV_IFNg_D",  "ND_IFNg_A",   "ND_IFNg_B",   "ND_IFNg_C", "ND_IFNg_D", "Par_ctrl_A", "Par_ctrl_B",  "Par_ctrl_C", "Par_ctrl_D", "Par_IFNg_A",  "Par_IFNg_B",  "Par_IFNg_C",  "Par_IFNg_D", "WT_IFNg_A",   "WT_IFNg_B",   "WT_IFNg_C", "WT_IFNg_C") 
```
Drop any columns in RNA DT not in the AB PH datasets

```{r}
cond.OI <- c(grep('_[ABC]', colnames(p.ab.mat), value=T), 'gene_symbol')
rna.mat <- rna.mat[, colnames(rna.mat) %in% cond.OI, with = F ]
```

Normalise RNA input
-----

PH & AB data normalised by msstats.
Need to repeat this for RNAseq; create a design matrix and input
Normalise the RNAseq data and remove low-info features

```{r}
#convert to matrix
rna.mat <- rna.mat %>% 
  as.matrix(rownames='gene_symbol')

design.mat <- data.frame(replicate=as.factor(c(rep(c('A','B','C'),3),'B','C','A','B','C')),
                         condition=as.factor(str_extract(colnames(rna.mat), 'IFNg|ctrl')), 
                         celltype=as.factor(str_extract(colnames(rna.mat),'EV|Par|ND|WT')))

design.mat$contrast <- as.factor(paste0(design.mat$celltype,"_", design.mat$condition))



#assign rownames from the colnames of count matrix
rownames(design.mat) <- colnames(rna.mat)
```

Run DEseq2 to normalize library & extract variance stabilised counts
```{r}
# only accepts integers
rna.mat <- round(rna.mat)

dds <- DESeqDataSetFromMatrix(countData = rna.mat,
                              colData = design.mat,
                              design = ~ contrast)

all(rownames(design.mat) == colnames(rna.mat))


#rna.mat <- assay(vsd)

#using log2 transformed counts instead as more gaussian
rna.mat <- log2(assay(dds))
```



Create 3 matrices and make list
```{r}
p.ph.mat <- p.ph.mat %>% 
  as.matrix(rownames='Protein')

p.ab.mat <- p.ab.mat %>% 
  as.matrix(rownames='Protein')


omics.list <- list(p.ph = p.ph.mat, 
                   p.ab = p.ab.mat,
                   rna = rna.mat)

```

Cant run the analysis where the sample list is different.. subset the samplses but then the dataset will likely be too small
Remove 
```{r}
# minimal set of shared samples
minimal.set <- c("EV_IFNg_A","EV_IFNg_B","EV_IFNg_C","ND_IFNg_A","ND_IFNg_B","ND_IFNg_C","Par_IFNg_A","Par_IFNg_B","Par_IFNg_C","WT_IFNg_A","WT_IFNg_B","WT_IFNg_C", "Par_ctrl_B" ,"Par_ctrl_C")

omics.list[['rna']] <- omics.list[['rna']][, colnames(omics.list[['rna']]) %in% minimal.set]

omics.list[['p.ph']] <- omics.list[['p.ph']][, colnames(omics.list[['p.ph']]) %in% minimal.set]
omics.list[['p.ph']] <- omics.list[['p.ph']][, -15]


omics.list[['p.ab']] <- omics.list[['p.ab']][, colnames(omics.list[['p.ab']]) %in% minimal.set]
omics.list[['p.ab']] <- omics.list[['p.ab']][, -15]


# need to reorder samples in rna.mat to align with other omics 
colnames(omics.list[['rna']])
colnames(omics.list[['p.ab']])
colnames(omics.list[['p.ph']])

# reorder to align with other modes
omics.list[['p.ab']] <- omics.list[['p.ab']][, c(1:6,9:11,7:8,12:14)]
```
Feature selection
----
Need to reduce the number of features in each dataset to limit 
Now lets select the most variable genes in each of the datasets
Will use tools built in mixomics 

Identifiy genes with low variance - unfortunately for most of the datasets there is quite a bit of variance as we are looking at many different conditions
```{r, eval=FALSE}
near0Var.list <- lapply(omics.list, function(x){
    nearZeroVar(as.data.frame(t(x)))
})
  
```

Instead, select the top features that were deg in contrasts of interest
For first pass, take the top 1k differential features (500 per contrast) per dataset and use this as input

```{r}
topn.g <- fread('./output/top_n_genes.csv.gz')

omics.list[['rna']] <- omics.list[['rna']][rownames(omics.list[['rna']]) %in% topn.g[omics == 'rna', gene],]

topn.uniprot <- translateGeneName2Uniprot(topn.g[omics == 'p.ab',gene])
omics.list[['p.ab']] <- omics.list[['p.ab']][rownames(omics.list[['p.ab']]) %in% topn.uniprot,]



# you never changed the protein names to gene.. many replicates of the same proteins...
#rownames(omics.list[['p.ph']]) #%in% topn.g[omics == 'p.ab', ]

# for now drop the ph data
omics.list[['p.ph']] <- NULL

omics.list[['p.ab']]<- omics.list[['p.ab']][!is.na(rownames(omics.list[['p.ab']])),]

#omics.list[['p.ab']] <- omics.list[['p.ab']][!rownames(omics.list[['p.ab']]) %in% c('TMPO', 'CUX1'),]


#topn.g[, .N, by=.(Label,omics)]
```




Remvoe NA & infinite values from the. dataset
```{r}
#remove rows with NA values; don't want these proteins to influence factor generation
omics.list[['p.ab']] <- omics.list[['p.ab']][complete.cases(omics.list[['p.ab']]),]

#remove any infinite values omics.list[['rna']][!is.infinite(omics.list[['rna']])] <- NA
omics.list[['rna']][!is.finite(omics.list[['rna']])] <- NA
omics.list[['rna']] <- omics.list[['rna']][complete.cases(omics.list[['rna']]),]


lapply(omics.list, function(x){
   #any(is.infinite((x)))
   any(is.na(x))
  # dim(x)
})

```



```{r}

MOFAobject <- create_mofa(omics.list)

```

Plot overview of the data
```{r}
plot_data_overview(MOFAobject)
```
for first run get the default options for training options just to visualise

```{r}

data_opts <- get_default_data_options(MOFAobject)
# custom settings
data_opts$scale_views <- TRUE #scale different omics to same total variance


model_opts <- get_default_model_options(MOFAobject)
# custom settings; 3 factors only for this n sample size
model_opts$num_factors <- 3

train_opts <- get_default_training_options(MOFAobject)
# custom
train_opts$convergence_mode <- 'slow'
train_opts$seed <- 42 #set random seed
```

Train the model

```{r}
MOFAobject <- prepare_mofa(
  object = MOFAobject,
  data_options = data_opts,
  model_options = model_opts,
  training_options = train_opts
)
```

```{r}
outfile = file.path(getwd(),"model.hdf5")
MOFAobject.trained <- run_mofa(MOFAobject, outfile, use_basilisk = T) #use_basilisk: R conda installation to maintain a python environment with necessary packages installed

```
Add sample metadata to the MOFA object
```{r}
#col must be named sample
meta <- data.frame(sample=colnames(omics.list[['rna']]),
                   condition=gsub('_[ABC]','',colnames(omics.list[['rna']])))

samples_metadata(MOFAobject.trained) <- meta
```

Look at amt of variance explained by each factor in each modality
Seems factor 1 is capturing most variance in both datasets, can see that facotr 2 also captures some 15-20% so might be interesting


Second plot: is the model a good fit? explains only 40% of variance in p.ab, but 80% in rna
```{r}
plot_variance_explained(MOFAobject.trained, x="view", y="factor")

plot_variance_explained(MOFAobject.trained, plot_total = T)[[2]] #amt of variance explained by each factor in each modality; seems MS prot much noisier
```
Visualise samples in latent space:
Factor 2 looks to distinguish WT_IFNg & EV_IFNg
In all 3 factors WT_INFg and ND_IFng are quite close

```{r}
plot_factor(MOFAobject.trained, 
  factor = 1:3,
  color_by = "condition"
)
```
Can also visualise as violin plots
Don't have the sample N for this type of analysis

```{r}
p <- plot_factor(MOFAobject.trained, 
  factors = c(1,2,3),
  color_by = "condition",
  dot_size = 3,        # change dot size
  dodge = T,           # dodge points with different colors
  legend = F,          # remove legend
  add_violin = T,      # add violin plots,
  violin_alpha = 0.25  # transparency of violin plots
)
p
```
```{r}
plot_factors(MOFAobject.trained, 
  factors = 1:3,
  color_by = "condition"
)
```


What genes/proteins contribute most to each factor?
Most negatively scored; indicates that feature has low levels in cells with positve factor values

```{r}
plot_weights(MOFAobject.trained,
  view = "rna",
  factor = 1,
  nfeatures = 10,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)


```
```{r}
plot_top_weights(MOFAobject.trained,
  view = "rna",
  factor = 2,
  nfeatures = 10
)
```
See how the factors are correlated with condititon:
Looks like only factor 2 may correlate with condition

```{r}
correlate_factors_with_covariates(MOFAobject.trained, 
  covariates = c('condition'), 
  cluster_cols = F,
  plot="log_pval"
)

```
Quickly run enrichment to see if the same information pops out:
No 
```{r}
# for ENSG conversion to gene symbol
library("AnnotationDbi")
library("org.Hs.eg.db")

#convert colnames to 
colnames(MSigDB_v6.0_C2_human) = mapIds(org.Hs.eg.db, 
                                        keys=colnames(MSigDB_v6.0_C2_human), 
                                        column="SYMBOL",
                                        keytype="ENSEMBL",
                                        multiVals="first")


# GSEA on positive weights, with default options
res.positive <- run_enrichment(MOFAobject.trained, 
  feature.sets = MSigDB_v6.0_C2_human, 
  view = "rna",
  factors = 1,
  sign = "positive"
)

# GSEA on negative weights, with default options
res.negative <- run_enrichment(MOFAobject.trained, 
  feature.sets = MSigDB_v6.0_C2_human, 
  view = "rna",
  factors = 1,
  sign = "negative"
)

#convert the entrez ids to 
colnames(MSigDB_v6.0_C2_human) <- translateEntrez2Uniprot(colnames(MSigDB_v6.0_C2_human)) 
  
plot_enrichment(res.negative, factor=2, max.pathways=1)

```



Look at test data to see what is needed:
```{r}
utils::data("CLL_data")       
lapply(CLL_data,dim)

#check this out for metadata formatting
CLL_metadata <- fread("ftp://ftp.ebi.ac.uk/pub/databases/mofa/cll_vignette/sample_metadata.txt")
```