---
title: "netprop-rerun"
author: "Martin Gordon"
date: "2023-06-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Rerun of the network propagation 

For this run, we will apply input heat for the top 500 'features' returned for each dataset.
## Packages

```{r}
library("RColorBrewer") 
library(data.table)
library(stringr) #str_extract function 
library(magrittr)
library(ggplot2)
library (ComplexHeatmap)
library(RcppCNPy) #R/W support to import numpy objects into R
library(VennDiagram)
library(pbapply) #apply functions with progress bar
library(eulerr)

# load bens util scripts for the analysis
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/LocatePeptidePtmInProteins.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

# netprop scripts
source ("../../utils/bp_utils/STRING_db_utils.R")
source ("../../utils/bp_utils/NetworkPropagation.R")
```

Read in clean file

```{r}
rna <- fread('./output/rna_ab.clean.csv.gz')
rna[, adj.pvalue := p.adjust(pvalue, method='BH')]

p.ab <- fread('output/prot_ab.clean.csv.gz')
p.ph <-  fread('output/prot_ph.clean.csv.gz')


heats.input <- list(rna = rna,
                    p.ab = p.ab,
                    p.ph = p.ph)
```

significance score; control for extreme values in p.adjust in rnaseq dataset

```{r}

HeatScore <- function(log2FC, pvalue){
  FC <-  abs(log2FC)
  significance <- -log10(pvalue)
  significance <- ifelse (significance > FC, FC, significance) # control for extreme values but probably not needed
  sqrt(FC * significance)
}

#similiar to above
magSigScore <- function(log2FC, pvalue, magnitudeScale = 2){
  magnitude <- magnitudeScale * abs(log2FC)
  significance <- -log10(pvalue)
  significance <- ifelse (significance > magnitude, magnitude, significance)
  sqrt(magnitude * significance)
}
# apply heats to the network
lapply(heats.input, function(x){
  x[, heat := HeatScore(log2FC, pvalue)]
})

# add the stringdb information to teh gene
heats.input[['rna']][, gene_str := GetStringIDMapping.inOrder(gene, stringAliasFile='./data/9606.protein.aliases.v11.5.txt.gz')]
heats.input[['p.ph']][, gene_str := GetStringIDMapping.inOrder(uniprot, stringAliasFile='./data/9606.protein.aliases.v11.5.txt.gz')]
heats.input[['p.ab']][, gene_str := GetStringIDMapping.inOrder(uniprot, stringAliasFile='./data/9606.protein.aliases.v11.5.txt.gz')]
```

remove the multiprots in the ph data; just keep the highest scoring

```{r}
heats.input[['p.ph']][, uniprot := gsub("_[STY][0-9]+", "", Protein) %>% #strip phosphosite info
          strsplit (";") %>%  #split multiprots; creates a list of vectors of each element
          lapply (unique) %>% #removes any duplicate elements in list
          lapply (paste, collapse = ";") %>% #colapses back to unique prot names
          unlist]

heats.input[['p.ph']] <- heats.input[['p.ph']] [,uniprot := sub("\\;.*","", uniprot)] #keep only the first protein name instead of dropping these rows
#omics_list[['prot_ph']] <- omics_list[['prot_ph']][!grepl(";", uniprot),] #only select rows w/o multiuniprots



#only taking record with max magSig score
heats.input[['p.ph']] <- heats.input[['p.ph']][!is.na(heat), .(heat = max(heat, na.rm = TRUE)), by = .(Label, uniprot, gene_str)]

# using gene symbols as identifiers
heats.input[['p.ph']] <- heats.input[['p.ph']][, .(Label, gene = translateUniprot2GeneName(uniprot, species = 'HUMAN'), gene_str,heat, uniprot) ] # or "MOUSE" or "RAT"
heats.input[['p.ph']] <- heats.input[['p.ph']][!is.na(gene),]



heats.input[['p.ab']] <- heats.input[['p.ab']][,uniprot := sub("\\;.*","",Protein)] #take only first name with multi uniprots
#omics_list[['prot_ab']] <- omics_list[['prot_ab']][,!grepl(";", Protein),] #only select rows w/o multiuniprots


#only taking record with max magSig score (incase of multiple measurements) - no change
heats.input[['p.ab']] <- heats.input[['p.ab']][!is.na(heat), .(heat = max(heat, na.rm = TRUE)),by = .(Label, uniprot, gene_str)]

# using gene symbols as identifiers
heats.input[['p.ab']] <- heats.input[['p.ab']][, .(Label, gene = translateUniprot2GeneName(uniprot), gene_str, heat, uniprot) ] # or "MOUSE" or "RAT"
heats.input[['p.ph']] <- heats.input[['p.ph']][!is.na(gene),]

```

as the heatscore is geometric mean of log2FC and significance,  rank & take the top 500  and use as input for the network propagation (similiar to netcoloc protocol)
```{r}

top_n <- lapply(heats.input, function(x, n=500){
  x <- x[order(-heat),] 
  x.x <- x[Label == 'EV_IFNg-vs-WT_IFNg',][1:n]
  x.y <- x[Label == 'WT_IFNg-vs-ND_IFNg',][1:n]
  
  y <- rbind(x.x,x.y)
  return(y)
  })

topn.genes <- lapply(heats.input, function(x){
  x[ !is.na(gene), gene,Label]
})
  
topn.g <- rbindlist(topn.genes, idcol='omics')

fwrite(topn.g,  './output/top_n_genes.csv.gz')
```


```{r}
heats.input <- top_n
getwd()
```

We are using RWR with pr=.3 for first run

```{r}

# this is precomputed by a python script NetworkPropagation_Build_S_matrix.py (I modified the input to use latest version of string network)
#python3 ../../../utils/bp_utils/NetworkPropagation_Build_S_matrix_mg.py


#using restart prob=50 for now. Hopefully should keep the network more localised and we can find overlaps
S_matrix.30 <- LoadNumPyS_matrix(matrixPath = "./data/S_matrix.human.string.pr0.30.npy", 
                              nodesTablePath = "./data/S_matrix.human.string.pr0.30.nodeNames.csv" )

```

```{r}

#get each unique contrast
contrast.groups <- c('EV_IFNg-vs-WT_IFNg','WT_IFNg-vs-ND_IFNg')


rna.allNP <- pbapply::pblapply(contrast.groups, function(g){
  
   NetworkPropagateS_matrix(S_matrix.30,
                            geneHeats = heats.input[['rna']][Label == g, .(gene = gene_str, heat)], # requires gene and heat column (using string as gene)
                            networkHeatOnly = TRUE,
                            permuteOnlyInObserved=TRUE,
                            numPermutations = 20000)
  
})

rna.allNPdt <- rbindlist(rna.allNP, idcol = "Label")
fwrite (rna.allNPdt,  "./output/AllNetworkPropagation_RNA_STRING_Interactions.rs0.3.n500.csv")


p.ab.allNP <- pbapply::pblapply(contrast.groups, function(g){
  
   NetworkPropagateS_matrix(S_matrix.30,
                            geneHeats = heats.input[['p.ab']][Label == g, .(gene = gene_str, heat)], # requires gene and heat column (using string as gene)
                            networkHeatOnly = TRUE,
                            permuteOnlyInObserved=TRUE,
                            numPermutations = 20000)
  
})

p.ab.allNPdt <- rbindlist(p.ab.allNP, idcol = "Label")
fwrite (p.ab.allNPdt,  "./output/AllNetworkPropagation_PROT_AB_STRING_Interactions.rs0.3.n500.csv")

p.ph.allNP <- pbapply::pblapply(contrast.groups, function(g){
  
   NetworkPropagateS_matrix(S_matrix.30,
                            geneHeats = heats.input[['p.ph']][Label == g, .(gene = gene_str, heat)], # requires gene and heat column (using string as gene)
                            networkHeatOnly = TRUE,
                            permuteOnlyInObserved=TRUE,
                            numPermutations = 20000)
  
})

p.ph.allNP
p.ph.allNPdt <- rbindlist(p.ph.allNP, idcol = "Label")
fwrite (p.ph.allNPdt,  "./output/AllNetworkPropagation_PROT_PH_STRING_Interactions.rs0.3.n500.csv")


```



Review p-values after running the analysis

```{r}
hist(rna.allNPdt[Label == 1]$pvalue)
```


