---
title: "sig.featrures.in.db"
author: "Martin Gordon"
date: "2023-07-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages}
library("RColorBrewer") 
library(data.table)
library(stringr) #str_extract function 
library(magrittr)
library(ggplot2)

library(openxlsx)

# load bens util scripts for the analysis
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/LocatePeptidePtmInProteins.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

```

## Identify sdignificant overlaps in the SAINT and ADPr databases.


Read in sig exp and np results

```{r}
exp.sig.genes <- fread('./output/sig-genes/experimental-omics-sig-genes.csv') %>% 
  split(.$Label)
  
np.sig.genes <- fread('./output/sig-genes/netprop-omics-sig-genes.csv') %>% 
  split(.$Label)

```

Read in the databases. Lets see how these look:
```{r}
saint.db <- read.xlsx('./data/SAINT_Results.xlsx', sheet=1) %>% 
  .[, unique('PreyGene')]

adpr.db <- fread('./data/ADPr_Database_v2.csv') %>% 
  .[, unique(GeneName)] 

#remove empty strings
adpr.db[adpr.db == ''] <- NA
adpr.db <- adpr.db[complete.cases(adpr.db)]

#read in stringdb as the universe for the NP tests
S_matrix.30 <- LoadNumPyS_matrix(matrixPath = "./data/S_matrix.human.string.pr0.30.npy", 
                              nodesTablePath = "./data/S_matrix.human.string.pr0.30.nodeNames.csv" )

uni <- rownames(S_matrix.30) %>% 
  unique() %>% 
  translateString2Gene(species= "HUMAN")
```

test the genesets for over representation

## see here for hypergeometric set-up from clusterprofiler author: https://www.biostars.org/p/485827/#9483835

```{r}

# loop through the conditions in each and run a hypergeometric test for both conditions
lapply(seq_along(np.sig.genes), function(x,n,i){

  #get intersect with pathway & deg
  deg = x[[i]][, unique(gene)]
  intsec = intersect(deg, adpr.db)

  
  # run hypergeometric test for enrichmet (lower.tail=F)
   print(paste0('testing ',n[[i]], ' for enrichment vs ADPR genes' ))
   
 adp <- phyper(q=length(intsec), #size of intersect of genes in pathway & deg genes
       m=length(adpr.db), #genes in pathway/db
       n=length(uni) - length(adpr.db), #universe - genes in db
       k=length(deg), #n degs
       lower.tail = F) #look for enrichment rather than depletion
  
 
 print(adp)
 
 print(paste0('testing ',n[[i]], ' for enrichment vs SAINT genes' ))
 intsec = intersect(deg, saint.db)
 
  saint <- phyper(q=length(intsec), #size of intersect of genes in pathway & deg
       m=length(saint.db), #genes in pathway/db
       n=length(uni) - length(saint.db), #universe - genes in db
       k=length(deg), #n degs
       lower.tail = F) #look for enrichment rather than depletion
 
  print(saint)
},x=np.sig.genes, n=names(np.sig.genes))

```

output for NP results:
[1] "testing EV_IFNg-vs-WT_IFNg for enrichment vs ADPR genes"
[1] 0.1770627
[1] "testing EV_IFNg-vs-WT_IFNg for enrichment vs SAINT genes"
[1] 0.00419192
[1] "testing WT_IFNg-vs-ND_IFNg for enrichment vs ADPR genes"
[1] 0.1004477
[1] "testing WT_IFNg-vs-ND_IFNg for enrichment vs SAINT genes"
[1] 0.001469973


Repeat the test for the experimental results:

Taking all significant genes for both conditions
Background will be all genes detected in any of the prot or rna experiments
```{r}
# read in the clean omics files
rna.ab.clean <- fread('./output/rna_ab.clean.csv.gz') %>% 
  .[, unique(gene)]
prot.ab.clean <- fread('./output/prot_ab.clean.csv.gz') %>% 
  .[, unique(gene)]

uni <- unique(union(rna.ab.clean, prot.ab.clean))
```

```{r}
# loop through the conditions in each and run a hypergeometric test for both conditions
lapply(seq_along(exp.sig.genes), function(x,n,i){

  #get intersect with pathway & deg
  deg = x[[i]][, unique(gene)]
  intsec = intersect(deg, adpr.db)

  
  # run hypergeometric test for enrichmet (lower.tail=F)
   print(paste0('testing ',n[[i]], ' for enrichment vs ADPR genes' ))
   
 adp <- phyper(q=length(intsec), #size of intersect of genes in pathway & deg
       m=length(adpr.db), #genes in pathway/db
       n=length(uni) - length(adpr.db), #universe - genes in db
       k=length(deg), #n degs
       lower.tail = F) #look for enrichment rather than depletion
  
 
 print(adp)
 
 print(paste0('testing ',n[[i]], ' for enrichment vs SAINT genes' ))
 intsec = intersect(deg, saint.db)
 
  saint <- phyper(q=length(intsec), #size of intersect of genes in pathway & deg
       m=length(saint.db), #genes in pathway/db
       n=length(uni) - length(saint.db), #universe - genes in db
       k=length(deg), #n degs
       lower.tail = F) #look for enrichment rather than depletion
 
  print(saint)
},x=exp.sig.genes, n=names(exp.sig.genes))

```
output for the experimentally significant genes:

[1] "testing EV_IFNg-vs-WT_IFNg for enrichment vs ADPR genes"
[1] 0.3782759
[1] "testing EV_IFNg-vs-WT_IFNg for enrichment vs SAINT genes"
[1] 0.0006523889
[1] "testing WT_IFNg-vs-ND_IFNg for enrichment vs ADPR genes"
[1] 0.172597
[1] "testing WT_IFNg-vs-ND_IFNg for enrichment vs SAINT genes"
[1] 3.715225e-05
