---
title: "T-SNE_exp"
author: "Martin Gordon"
date: "2023-07-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## packages 

```{r}
library("RColorBrewer") 
library(data.table)
library(stringr) #str_extract function 
library(magrittr)
library(ggplot2)
library (ComplexHeatmap)
library(RcppCNPy) #R/W support to import numpy objects into R
library(VennDiagram)
library(pbapply) #apply functions with progress bar
library(eulerr)


# load bens util scripts for the analysis
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/LocatePeptidePtmInProteins.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

source ("../../utils/bp_utils/PPINetworkModularized.R")

# netprop scripts
source ("../../utils/bp_utils/STRING_db_utils.R")
source ("../../utils/bp_utils/NetworkPropagation.R")


# Attach the `umap` library
library(umap)

library(cluster) #k-mediod clustering
library(factoextra) #estimate optimal k
```


## Repeat of the method used to visualise the output of the network propagation 
Take signinficant genes in each condition in each dataset and run DSD clusterinfg

Read in the clean data, cn also plot SAINT data and include in the analysis but do this first

```{r}

# read in the clean omics files
rna.ab.clean <- fread('./output/rna_ab.clean.csv.gz')
rna.ab.clean[, omics :='rna']
rna.ab.clean[, adj.pvalue := p.adjust(pvalue, method = 'BH')]

prot.ab.clean <- fread('./output/prot_ab.clean.csv.gz')
prot.ab.clean[, omics := 'p.ab']
prot.ph.clean <- fread('./output/prot_ph.clean.csv.gz')
prot.ph.clean[, omics := 'p.ph']

omics.clean <- list(rna_ab = rna.ab.clean,
                    prot_ab = prot.ab.clean,
                    prot_ph = prot.ph.clean)
```


filter the data to only include significant genes 

```{r}
sig.omics <-lapply(omics.clean, function(x){
  x[adj.pvalue < 0.05 & abs(log2FC) > log2(1.5),]
})
```

For the DSD clustering on the STRING network, we need to map to string ID's. For phospho, collapse to gene level and take the instance with highest scoring geom.mean of log2FC and pvalue

```{r}
 

sig.omics[['prot_ph']][, uniprot := gsub("_[STY][0-9]+", "", Protein) %>% #strip phosphosite info
          strsplit (";") %>%  #split multiprots; creates a list of vectors of each element
          lapply (unique) %>% #removes any duplicate elements in list
          lapply (paste, collapse = ";") %>% #colapses back to unique prot names
          unlist]


sig.omics[['prot_ph']] <- sig.omics[['prot_ph']][!grepl(";", uniprot),] #only select rows w/o multiuniprots


#only taking record with best gm score
sig.omics[['prot_ph']] <- sig.omics[['prot_ph']][,  .SD[which.min(gm.lfc.fdr)], by=.(Label, uniprot)]
sig.omics[['prot_ph']][,gene := translateUniprot2GeneName(uniprot)]
```

Extract significant gene names stringIDs and contrasts/labels
(same format as NP to run functions)

```{r}
sig.omics.genes <- lapply(sig.omics, function(x){
  x <- x[, .(string = translateGene2String(gene, species='HUMAN'),symbol = gene,Label)]
  x[, .(gene = string, Label, symbol)]
})

```


```{r}



#combinethe netprop sig results & take the unique gene list
sig.df <- rbindlist(sig.omics.genes, idcol = 'omics')


sig.WT.ND.df <- rbindlist(sig.omics.genes, idcol = 'omics') %>% 
  .[Label == 'WT_IFNg-vs-ND_IFNg',]

sigfeatures.WT.ND <- unique(sig.WT.ND.df$symbol)



# get string subnetwork for each contrast
string.net.list <- lapply(contrasts.OI, function(i){
  
  sigfeatures <- sig.df[Label == i, unique(symbol)]
  stringMapping <- GetStringIDMapping(sigfeatures, stringAliasFile='./data/9606.protein.aliases.v11.5.txt.gz')
  
  stringsOI <- c(setdiff(sigfeatures, stringMapping$alias), stringMapping$string)
  stringEdges <- GetStringSubNetwork(stringsOI,  oneHopConnections = TRUE, stringFile = './data/9606.protein.links.detailed.v11.5.txt.gz' ) #includes 1 hop connectors

  })
# add names to the dataset
names(string.net.list) <- contrasts.OI



lapply(seq_along(string.net.list), function(x,n,i){
  
  fwrite(x[[i]], paste0("./data/", n[[i]], "expdata.StringSubNetwork.1hop.txt"), sep = "\t", col.names=FALSE)
  },x=string.net.list, n=names(string.net.list))

```
### DSD in external python script

```{r, eval=FALSE}
# run this in CL
#conda activate python2.7
# -c confidence values in ppi as edge weights
# -s 20 steps in random walk

#for i in EV_IFNg-vs-WT_IFNg WT_IFNg-vs-ND_IFNg; do echo " running DSD on $i"; python ./scripts/capDSD-src/DSD.py ./data/${i}expdataStringSubNetwork.1hop.txt ./data/${i}expdataStringSubNetwork.1hop; done

```


```{r, fig.width = 12, fig.height = 12}
dsd.ev.wt <- fread ("./data/EV_IFNg-vs-WT_IFNgexpdata.StringSubNetwork.1hop.dsd")
dsd.wt.nd <-fread('./data/WT_IFNg-vs-ND_IFNgexpdata.StringSubNetwork.1hop.dsd')

dsd.list <- list("EV_IFNg-vs-WT_IFNg" = dsd.ev.wt,
                 "WT_IFNg-vs-ND_IFNg" = dsd.wt.nd)

#convert to matrix
dsd.list <- lapply(dsd.list, function(x){x <- as.matrix(x, rownames='V1')})

# create subnet distance mat 
string.dist.list <- lapply(contrasts.OI, function(x,y,i){
  
  sigfeatures <- y[Label == i, unique(symbol)]
  stringMapping <- GetStringIDMapping(sigfeatures, stringAliasFile='./data/9606.protein.aliases.v11.5.txt.gz') # convert to string
  # get interesting strings
  stringsOI <- c(setdiff(sigfeatures, stringMapping$alias), stringMapping$string) #return elements unique in sigfeatures not in str$aliad
  
  #remove the 1 hop connectors
  stringsInMat <- intersect(rownames(x[[i]]), c(stringsOI))
  submat <- x[[i]][stringsInMat, stringsInMat]
  
  },x=dsd.list, y=sig.df)

names(string.dist.list) <- contrasts.OI


lapply(seq_along(string.dist.list), function(x,n,i){
  
  #get distance matrix for clustering
  dist.mat <- as.dist(x[[i]])
  ddr <- hclust(dist.mat, method= "average") %>% 
    as.dendrogram()

  hm <- Heatmap(x[[i]], show_row_names = FALSE, show_column_names = FALSE, cluster_rows = ddr, cluster_columns = ddr)
  BackupAsPDF(draw(hm),paste0(n[[i]],'dissimilarity-matrix.pdf'))

  
}, x=string.dist.list, n=names(string.dist.list))

```
### tsne based on DSD

For visualizing the modularization later

```{r, fig.width = 6, fig.height =  6}

tsne.list <- lapply(seq_along(string.dist.list), function(x,n,i){

 dist.mat <- as.dist(x[[i]])
 t <- Rtsne::Rtsne(dist.mat, is.distance = TRUE, theta = 0.0 ) # exact tsne when theta = 0.0

 tsne.dt <- as.data.table(t$Y)
 tsne.dt$string <- attr(dist.mat, "Labels")
 return(tsne.dt)

}, x= string.dist.list, n=names(string.dist.list))


names(tsne.list) <- contrasts.OI

lapply(tsne.list, function(x){ 
  ggplot(x, aes(x=V1, y=V2)) + geom_point()})
```
# modularization

Going to use k-medoid clustering (more robust than k-means)


First use the `factoextra` package to estimate otpimal number of clusters
Weighted sum of squares suggests that 5~6 clusters works for both

```{r}

names(string.dist.list)

lapply(string.dist.list, function(x){
  fviz_nbclust(x, kmeans,method='gap_stat') #5-7
})

names(string.dist.list)
```
Plot the T-SNE modules
```{r, fig.width = 6, fig.height =  6}

# cluster.dt <- defineModulesFromDistance("/Users/ben/UCSF/kroganlab/BenPolacco/roche/2023_06_02_RSV_NetProp_HeatDiffusion_data/2023_06_05_StringSubNetwork.Preys.1hop.dsd",geneNamesOI = stringsOI, deepSplit = 0.8)
# tsne.dt[cluster.dt, c("clusterID", "cluster") := .(i.clusterID, i.cluster), on = c(string = "gene")]


lapply(contrasts.OI, function(x,y,i){
  
  pam.out <- cluster::pam(x[[i]],diss = TRUE, k = 9)
  y[[i]][names(pam.out$clustering), clusterID := pam.out$clustering, on = "string"]

  y[[i]][, cluster := sprintf("clust.%02d", clusterID)]
  y[[i]][, c("x", "y") := .(V1, V2)]

  y[[i]][, .N, by=cluster]
  
 p <- ggplot(y[[i]], aes (x = x, y = y,)) +
  #geom_segment(data = edgeView.dt, aes(xend = xend, yend = yend),  alpha = 0.2, linewidth = 0.1) +
 geom_point(aes( color = cluster)) +
 #scale_size_manual(values = c(bait = 4, prey = 0.5)) +
 #scale_shape_manual(values = c(bait = 23, prey = 1)) +
 coord_fixed() + theme_void()
p

}, x=string.dist.list, y=tsne.list)

```

## enrich and label the modules
```{r}
gmt <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")

```

```{r}

lapply(seq_along(tsne.list), function(x,n,i){
  x[[i]][, gene := translateString2Gene(string, species = "HUMAN")]
  geneToCluster <- x[[i]][,.(gene = gene, cluster = cluster)]
  
  #label each of the modules by enrichment results
  enrichNames.out <- labelModulesByEnrichment(cluster.dt = geneToCluster, gmt = gmt, numProcessors = 8)
  
  # choose a naming scheme
 clusterNames.dt <- enrichNames.out$clusterNames.dt[, .(cluster, name = termScore.name, go = termScore.go)]
 enrich.dt <- enrichNames.out$enrich.dt
 
 x[[i]][clusterNames.dt, c("name", "go" ) := .( i.name, i.go), on = "cluster"]
 x[[i]][gmt, nameMatch := TRUE, on = c(gene = "gene", go = "ont") ]
 x[[i]][nameMatch == TRUE, nameMatchCluster := cluster]
  
 
 p <- ggplot(x[[i]], aes (x = x, y = y,)) +
  #geom_segment(data = edgeView.dt, aes(xend = xend, yend = yend),  alpha = 0.2, linewidth = 0.1) +
  geom_point(aes( color = name  )) +
   ggtitle(paste0(n[[i]], ' enrichment' )) +
  #scale_size_manual(values = c(bait = 4, prey = 0.5)) +
  #scale_shape_manual(values = c(bait = 23, prey = 1)) +
  coord_fixed() + theme_void()
p

#BackupAsPDF(p, paste0(n[[i]],'exp.k5.tsne.enrichment.pdf'))

}, x=tsne.list, n=names(tsne.list))


#on plot: gene ratio 19/23 # ratio of the number of genes within a given gene set that are present in the input gene list
#bg ratio 22/215 #null set what you would expect by chance. 22 here respents count in the universe/background set
```

Get the enrichment tables

```{r}
lapply(seq_along(tsne.list), function(x,n,i){
  x[[i]][, gene := translateString2Gene(string, species = "HUMAN")]
  geneToCluster <- x[[i]][,.(gene = gene, cluster = cluster)]
  
  #label each of the modules by enrichment results
  enrichNames.out <- labelModulesByEnrichment(cluster.dt = geneToCluster, gmt = gmt, numProcessors = 8)
  # choose a naming scheme
  clusterNames.dt <- enrichNames.out$clusterNames.dt[, .(cluster, name = termScore.name, go = termScore.go)]
 # print(clusterNames.dt)
 fwrite(clusterNames.dt, paste0('./output/',n[[i]], '-t-SNE.k9.cluster.go.enrichment.csv.gz'))


}, x=tsne.list, n=names(tsne.list))
```

