---
title: "TF.enrichment"
author: "Martin Gordon"
date: "2023-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Takehome 
only three TF Target geensets significantly enirched in our DEG's RNAseq
CREB is induced by a variety of growth factors and inflammatory signals and subsequently mediates the transcription of genes containing a cAMP-responsive element
DLX6: This family has at least six members that encode proteins with roles in forebrain and craniofacial development

Both significantly depleted; means that targets for this TF are down-regulated in WT_IFNg-vs-ND_IFNg comparison (where WT is denominator), meaning expression for these target genes are increased in WT relative to ND mutant

In the other condition raw p-value is significant to 1e3 (1 in 1000) but not after correction
MTC burden here? can we filter out the number of genesets we are testing enrichment for? Find a better geneset maybe

No sig results for the protein abundance data

RNAseq
---
WT_IFNg-vs-ND_IFNg	significantly depleted for DLX6 and CREB3L4 targets


## Transcription Factor enrichment of AB MS and RNAseq data
A549 cell lines

```{r}
library(data.table)
library(stringr) #str_extract function 
library(magrittr)
library(ggplot2)
library (ComplexHeatmap)
library(circlize)
library(ggrepel)
library(clusterProfiler)
library(msigdbr) #MSigDB in tidy format

# load bens util scripts for the analysis
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/LocatePeptidePtmInProteins.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

# enrichment functions and FGSEA 
source("../../utils/bp_utils/Localization_CellMap.R")

```

In RNA res the denominator goes first, opposite in Prot AB

Read in the DEG results for RNAseq and AB MS
Again changing the sign of AB to mimic RNAseq so denominator is the same for both

```{r}
ab.de.res <- fread('/Users/martingordon/Documents/projects/051623_DSwaney/CellLines_Abundance/tables/2023_05_25_msstats_groupcomparisons.csv.gz')
ab.de.res[Label %in% c('WTpos vs N10t20pos', 'WTpos vs EVpos') & gene == 'ASS1',]

ab.de.res <- ab.de.res[Label %in% c('WTpos vs EVpos', 'WTpos vs N10t20pos'),]
ab.de.res[Label == "WTpos vs EVpos", Label := "EV_IFNg-vs-WT_IFNg"]
ab.de.res[Label == "WTpos vs N10t20pos", log2FC := -1 * log2FC]
ab.de.res[Label == "WTpos vs N10t20pos", Label := "WT_IFNg-vs-ND_IFNg"]

#prot.ab[Label == "WTpos vs EVpos", log2FC := -1 * log2FC]
prot.ab[Label == "WTpos vs EVpos", Label := "EV_IFNg-vs-WT_IFNg"]
prot.ab[Label == "WTpos vs N10t20pos", log2FC := -1 * log2FC]
prot.ab[Label == "WTpos vs N10t20pos", Label := "WT_IFNg-vs-ND_IFNg"]

# clean RNAseq
box.loc <- '~/Library/CloudStorage/Box-Box/Mac1_omics_data/'
rna.ab.core <- fread(paste0(box.loc, 'CellLines_RNAseq/core_table_gene.xls.gz'), fill=T)

EV.WT.cols.OI <- c(2,3, grep('EV_IFNg-vs-WT_IFNg', colnames(rna.ab.core)))
WT.ND.cols.OI <- c(2,3, grep('WT_IFNg-vs-ND_IFNg', colnames(rna.ab.core)))
#easiest way to reformat is to perform the split seperately and then 

cols.OI <- c(EV.WT.cols.OI, WT.ND.cols.OI)

  
EV.WT.rna.res <- rna.ab.core[, ..EV.WT.cols.OI] %>% 
    .[,Label := 'EV_IFNg-vs-WT_IFNg'] %>% 
    melt(., id.vars = c('Label','gene_id','gene_symbol'), 
            measure.vars = list(c('diffexp_log2fc_EV_IFNg-vs-WT_IFNg'), 
                                                       c('diffexp_deseq2_pvalue_EV_IFNg-vs-WT_IFNg'),
                                                       c('diffexp_deseq2_qvalue_EV_IFNg-vs-WT_IFNg')),
                                     value.name = c('log2FC', 'pvalue', 'qvalue'))
  
  


WT.ND.rna.res <- rna.ab.core[, ..WT.ND.cols.OI] %>% 
  .[,Label := 'WT_IFNg-vs-ND_IFNg'] %>% 
  melt(., id.vars = c('Label', 'gene_id','gene_symbol'), 
          measure.vars = list(c('diffexp_log2fc_WT_IFNg-vs-ND_IFNg'), 
                                                       c('diffexp_deseq2_pvalue_WT_IFNg-vs-ND_IFNg'),
                                                       c('diffexp_deseq2_qvalue_WT_IFNg-vs-ND_IFNg')),
                                   value.name = c('log2FC', 'pvalue', 'qvalue'))


rna.res <- rbind(EV.WT.rna.res, WT.ND.rna.res) %>% 
  .[,variable := NULL]

#sanity check - we want denominato-numerator naming convention
ab.de.res[gene == 'ASS1',]
rna.res[gene_symbol == 'ASS1',]

```
write out 

```{r}
#write(rna.res, './output/rna.clean.dea.results.csv')
#fwrite(ab.de.res,  './output/prot.ab.clean.dea.results.csv')
```

read in results 

```{r}
prot.ab <- fread('./output/prot.ab.clean.dea.results.csv')
rna <-fread('./output/rna.clean.dea.results.csv')
```

load in the MSigDB dataset for TF targets in humans for GSEA enrichment  
(not sure about curation here; ask )

```{r}
tft_gene_sets = msigdbr(species = "human", category = "C3", subcategory = "TFT:GTRD") %>% 
  as.data.table()

tft_gene_sets %>%  str()
tft_gene_sets[gs_name == 'MYF6_TARGET_GENES', .N, ]

# list of vectors - not needed
#tf.sets <- tft_gene_sets %>%  
#  .[, .(gs_name, gene_symbol)] #%>% 
 # .[, lapply(.SD, paste0, collapse=" "), by=gs_name]

tf.sets <- split(tft_gene_sets$gene_symbol, tft_gene_sets$gs_name)
```

calculate ranked metric for the log2FC, for this we will use signed pvalue
remove infinite values too 

```{r}
rna <- rna[, rank.metric :=  -log10(pvalue)*sign(log2FC), by=Label] %>% 
  .[is.finite(rank.metric)]

prot.ab <- prot.ab[, rank.metric :=  -log10(pvalue)*sign(log2FC), by=Label] %>% 
    .[is.finite(rank.metric)]
```

convert both datasets to a matrix 
```{r}

rna.mat <- dcast(rna, gene_symbol ~ Label, value.var = 'rank.metric') %>%  
  as.matrix(rownames='gene_symbol')

prot.mat <- dcast(prot.ab, gene ~ Label, value.var = 'rank.metric') %>% 
  as.matrix(rownames='gene')


mat.list <- list('rna'=rna.mat,
            'ab.ms'=prot.mat)
```

looked for significant enrichment of results 
In the RNAseq dataset only targets for 3 TFs significantly enriched (actually depleted!)

```{r}
FGSEA.enrich <- lapply(mat.list, function(x){
  matrixFGSEA(x, sets = tf.sets)
})


```
not enough sig results to plot, so jsut provide the enrichment results tables

```{r}
fwrite(FGSEA.enrich[['rna']], './output/TFT.enrichment.rna.csv')
fwrite(FGSEA.enrich[['ab.ms']], './output/TFT.enrichment.ab.ms.csv')
```

```{r}
hist(FGSEA.enrich[['rna']]$pval)
```
Look for enrichment for the IPA pathways
Issues with both files.. speak to Adrian tomorrow

```{r}
TFacts <- fread('./docs/TFactS_sensitive.gmt', sep='\t', header = F, fill=T)
#collapse third column on into genes list

# Define the columns to concatenate
columns_to_concat <- names(TFacts)[3:ncol(TFacts)]
TFacts[, gene.list := do.call(paste, c(.SD, sep = ",")), .SDcols = columns_to_concat]
TFacts[, gene.list := sub(',+$', '', gene.list)]
#TFacts[, gene.list := gsub(' ', '\t', gene.list)]
TFacts <- TFacts[, .(V1,gene.list)]

TFacts
# create a gene list of the two for the enrichment
tfacts.set <- split(TFacts$gene.list, TFacts$V1)
tfacts.set <- lapply(tfacts.set, function(x) unlist(strsplit(x, ',')))
```
run fgsea enrichment on these results
```{r}
FGSEA.enrich <- lapply(mat.list, function(x){
  matrixFGSEA(x, sets = tfacts.set)
})

```
check the results
no significant terms after correction, but many are very enriched 

```{r}
lapply(FGSEA.enrich, function(x) setnames(x, old = 'pathway', new='TF_direction'))

lapply(seq_along(FGSEA.enrich), function(x,n,i){
  
  x[[i]][order(padj)] %>% 
    .[, dataset := n[[i]] ]
  
  fwrite(x[[i]], paste0('./output/', n[[i]], '.TFact.enrichment.table.csv'))
  
},x=FGSEA.enrich, n=names(FGSEA.enrich))

names(FGSEA.enrich)

fread('./output/ab.ms.TFact.enrichment.table.csv')
```

171023
---
Read in the TF feature tables and 

```{r}
input <- dir(path='./output/TFtarget.enrichment', pattern = '.csv$', full.names = T) %>% 
  as.data.table() %>% 
  setnames('.', 'path') %>% 
  .[, table := str_extract(path, 'ab.ms.TFact|rna.TFact|TFT.enrichment.ab.ms|TFT.enrichment.rna')]


enrich.dt <- lapply(input$path, function(x){
  
  e.dt <- fread(x)
  e.dt[pval < 0.05,]
})



names(enrich.dt) <- c('TFTargetS.ABMS', 'TFTargetS.RNA', 'MSigDB.TFT.ABMS', 'MSigDB.TFT.RNA')

lapply(seq_along(enrich.dt), function(x,n,i){
  
  fwrite(x[[i]], paste0('./output/', n[[i]], '.enrichment.csv'))
},x=enrich.dt, n=names(enrich.dt))
```

