---
title: "081524_QCAssessment"
author: "Martin Gordon"
date: "2024-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## libraries 

```{r}
library(data.table)
library(magrittr)
library(MSstats)
library(ggplot2)
library(stringr)
library(ggh4x)
library(MSstats)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

source ("../../utils/mg_utils/r_utils/IDmapping.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```

# read in files 
```{r}

contrasts <- fread('./data/contrasts.txt', header=F)
keys <- fread('./data/keys.txt')
patient.batches <- fread('./data/Patient_batches.txt')

spec <- fread('./data/evidence_new.txt')

# so this format contains no NA but some 0 intensity? what is a zero intensity
# fitler out uninteresting columns
spec <- spec[,.(Protein, Genes, Run, Condition, BioReplicate, IsotopeLabelType, Intensity, PeptideSequence=Stripped.Sequence, ModPeptideSequence=Modified.Sequence, Precursor.Charge)]
spec[, noModPeptideSequence := PeptideSequence]
spec[, PeptideSequence := ModPeptideSequence]
spec[, runNumber := as.numeric(gsub('[.]d', '', str_extract(Run, "[0-9]+[.]d$")))]

# some 0 values in the output
spec <- spec[Intensity != 0,]
```


# merge the files

```{r}
# some mismathcing IDs
spec[!Run %in% unique(patient.batches$RawFile), unique(Run)]
# fix names; place a _ befroe the first D symvol
spec[!Run %in% unique(patient.batches$RawFile), Run := sub('D', '_D', Run)]

spec <- merge(x=spec[,-c('BioReplicate')], y=keys[,.(RawFile, Condition, BioReplicate, RunOrder=Run, Batch)], by.x=c('Run','Condition'), by.y=c('RawFile','Condition'))
```


Count number of features per run
```{r}
# no features detected more than once
any(spec[,.N, by=.(ModPeptideSequence,Precursor.Charge,Run)]$N > 1)
```

PCA and sample clustering
Why is there a huge drop in sample completeness? I think this couold be just poorer coverage in the batch2 samples

```{r}
featureMat <- dcast(spec,paste0(ModPeptideSequence, Precursor.Charge)~interaction(Condition,BioReplicate), value.var = "Intensity") %>% 
  as.matrix(rownames = "ModPeptideSequence")

# over 8k features detected in total
featureMat %>%  dim()

completeness.dt <- apply(featureMat, 1, function(x) sum(!is.na(x))/length(x)) %>% 
  data.table(seqID=names(.),
             completeness=.)

```
For now lets check out the heatmpa

```{r}
submat <- log2(featureMat)

submat <- sweep(submat, 1, apply(submat, 1, median, na.rm=T), '-')
submat <- submat[sample(rownames(submat), 3000),]

ann <- spec[,.N, by=.( Group=paste0(Condition, '.', BioReplicate),Condition,Patient=as.factor(BioReplicate),Batch)][, N := NULL]
# match the anno dt rows to order of the heatmap vector
ann <- ann[match(colnames(featureMat), Group),]


col.pal <- randomcoloR::distinctColorPalette(k=29)
names(col.pal) <- unique(ann$Patient)


colours <- list('Batch' = c('B1' = 'lightpink', 'B2'="slategrey"),
                'Condition' = c('HMT_D0'='azure','HMT_D7' = '#66CCEE','HMT_D21' = '#4477AA','LMT_D0'="#CCBB44", 'LMT_D7'='orange','LMT_D21'='#EE6677'),
                'Patient' = col.pal
                )

colAnn <- HeatmapAnnotation(df = ann[,.(Condition,Batch,Patient)], col = colours)


hm <- Heatmap(submat, 
        top_annotation = colAnn,
        name='Ints./Median',
        cluster_rows = clusterWNA(mat=submat),
        column_names_gp=gpar(fontsize=3),
        show_row_names = F)

BackupAsPDF(hm, 'features.medianScaled.heatmap', dimensions = c(12,6))
```
More missingness in batch 1. look at complete cases

```{r}
submat <- log2(featureMat)

submat <- sweep(submat, 1, apply(submat, 1, median, na.rm=T), '-')
submat <- submat[complete.cases(submat),]


colours <- list('Batch' = c('B1' = 'lightpink', 'B2'="slategrey"),
                'Condition' = c('HMT_D0'='lightblue2','HMT_D7' = '#66CCEE','HMT_D21' = '#4477AA','LMT_D0'="#CCBB44", 'LMT_D7'='orange','LMT_D21'='#EE6677'),
                'Patient' = col.pal
                )

colAnn <- HeatmapAnnotation(df = ann[,.(Condition,Batch,Patient)], col = colours)


hm <- Heatmap(submat, 
        top_annotation = colAnn,
        name='Ints./Median',
        row_title =sprintf('%s complete features', nrow(submat)),
        cluster_rows = clusterWNA(mat=submat),
        column_names_gp=gpar(fontsize=3),
        show_row_names = F)

hm
BackupAsPDF(hm, 'features.medianScaled.completeCases.heatmap', dimensions = c(12,6))


hm <- Heatmap(submat, 
        top_annotation = colAnn,
        name='Ints./Median',
        column_split=ann$Condition,
        row_title =sprintf('%s complete features', nrow(submat)),
        cluster_rows = clusterWNA(mat=submat),
        border = T,
        column_names_gp=gpar(fontsize=3),
        show_row_names = F)

BackupAsPDF(hm, 'features.medianScaled.condSplit.completeCases.heatmap', dimensions = c(12,6))


hm <- Heatmap(submat, 
        top_annotation = colAnn,
        name='Ints./Median',
        column_split=ann$Batch,
        row_title =sprintf('%s complete features', nrow(submat)),
        cluster_rows = clusterWNA(mat=submat),
        border = T,
        column_names_gp=gpar(fontsize=3),
        show_row_names = F)
hm
BackupAsPDF(hm, 'features.medianScaled.batchSplit.completeCases.heatmap', dimensions = c(12,6))
```
PCA 
PC1 clearly driven by batch 
PC2,4 clusters by region which is promising but not much variance explained

```{r}
submat <- log2(featureMat)
submat <- submat[complete.cases(submat),]

pcaOut <- prcomp(t(submat))

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, ann, by.x = "rn", by.y = "Group", all.x = TRUE)
pcaDT[, region := ifelse(Condition %like% 'HMT', 'High', 'Low')]
pcaDT[, timepoint := tstrsplit(Condition, '_', keep=2)]
pcaDT[, .(Condition, rn,timepoint, region,Patient)]

g <- ggplot(data.table(PC=seq(1,length(pcaPercentVar), by=1), var=pcaPercentVar)[PC <= 10,], aes(x=as.factor(PC), y=var)) +
  geom_bar(stat='Identity') +
  xlab('PC') +
  ylab('variance explained')
BackupAsPDF(g,'feature.pca.varExplained.barplot')


#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2, fill = Condition, shape=Batch)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ylab (sprintf ("PC4, %.1f%%", pcaPercentVar[4])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
 # scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_fill_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:22) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, "feature.pca.pc24.regionCol")
```
Summarise to protein and replot
```{r}
mssInput <- spec[, .(ProteinName = Protein, 
         PeptideModifiedSequence = PeptideSequence,
         PeptideSequence = noModPeptideSequence,
         Run,
         Condition,
         BioReplicate,
         Intensity,
         PrecursorCharge = Precursor.Charge,
         IsotopeLabelType,
         ProductCharge = NA,
         FragmentIon = NA)]

dp.out <- dataProcess(mssInput, 
                      MBimpute =  FALSE, 
                      summaryMethod = 'TMP',
                      featureSubset = "all")
```

Plot the standard QC at the protein level

```{r}
saveRDS(dp.out, ScriptAndDatedFileName('dp.out.rds'))

#fwrite(setDT(dp.out$ProteinLevelData), ScriptAndDatedFileName('proteinLvl.quant.csv'))
#fwrite(setDT(dp.out$FeatureLevelData), ScriptAndDatedFileName('featureLvl.quant.csv'))
```

```{r}
p.quant <- fread('~/Documents/projects/081524_MBouhaddou_MalariaPatientSamples_BatchNorm/081524_QCAssessment_data/2024_08_16_proteinLvl.quant.csv')
```

QC heatmaps & PCA
```{r}
p.mat <- dcast(p.quant,Protein~paste0(GROUP,'.', SUBJECT), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')

submat <- sweep(p.mat, 1, apply(p.mat, 1, median, na.rm=T), '-')


colours <- list('Batch' = c('B1' = 'lightpink', 'B2'="slategrey"),
                'Condition' = c('HMT_D0'='lightblue2','HMT_D7' = '#66CCEE','HMT_D21' = '#4477AA','LMT_D0'="#CCBB44", 'LMT_D7'='orange','LMT_D21'='#EE6677'),
                'Patient' = col.pal
                )

colAnn <- HeatmapAnnotation(df = ann[,.(Batch)], col = colours)

hm <- Heatmap(submat,
        top_annotation = colAnn,
        show_row_names = F,
        column_names_gp = gpar(fontsize=4),
        row_title = sprintf('%s Proteins', nrow(submat)),
        column_split = ann$Condition,
        name='log2 Ints./Median',
        cluster_rows = clusterWNA(submat))

BackupAsPDF(hm, 'protens.medianScale.heatmap', dimensions = c(10,6))


# plot a complete cases heatmap
hm <- Heatmap(submat[complete.cases(submat),],
        top_annotation = colAnn,
        show_row_names = F,
        column_names_gp = gpar(fontsize=4),
        row_title = sprintf('%s Proteins (no missing)', nrow(submat[complete.cases(submat),])),
        column_km = 2,
        name='log2 Ints./Median',
        cluster_rows = clusterWNA(submat[complete.cases(submat),]))
hm

BackupAsPDF(hm, 'protens.medianScale.completecases.heatmap', dimensions = c(10,6))
```
Look at PCA at the protein level 

```{r}
submat <- p.mat[complete.cases(p.mat),]

pcaOut <- prcomp(t(submat))

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, ann, by.x = "rn", by.y = "Group", all.x = TRUE)
pcaDT[, region := ifelse(Condition %like% 'HMT', 'High', 'Low')]
pcaDT[, timepoint := tstrsplit(Condition, '_', keep=2)]
pcaDT[, .(Condition, rn,timepoint, region,Patient)]

g <- ggplot(data.table(PC=seq(1,length(pcaPercentVar), by=1), var=pcaPercentVar)[PC <= 10,], aes(x=as.factor(PC), y=var)) +
  geom_bar(stat='Identity') +
  xlab('PC') +
  ylab('variance explained')
BackupAsPDF(g,'protein.pca.varExplained.barplot')


#plot first two components
p <- ggplot (pcaDT, aes(x=PC3, y=PC4, fill = timepoint, shape=Batch)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
 # scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_fill_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:22) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, "protein.pca.pc12.regionCol")
```


**Combat Workflow**

Lets generate metadata and model matrix
batch and patient seem completely confounded, so cant estimate batch covariate at patient level

Edit our p.quant output to reflect the conditions
```{r}
p.quant[,Region := str_extract(GROUP, 'HMT|LMT')]
p.quant[,Patient := SUBJECT]
p.quant[Region == 'HMT' , Patient := as.character(Patient + 29)]
p.quant[, sample := paste0(GROUP,'.', Patient)]

p.mat <- dcast(p.quant, Protein~paste0(GROUP,'.',Patient), value.var = 'LogIntensities') %>% 
  as.matrix(rownames='Protein')
```

```{r}
ann <- spec[,.N, by=.( Group=paste0(Condition, '.', BioReplicate),Condition,Patient=BioReplicate, Batch)][, N := NULL]

ann[, c('Region', 'Timepoint') := tstrsplit(Condition, '_', keep=c(1,2))]
ann[Region == 'HMT' ,Patient := as.character(Patient + 29)]
ann[, c('Condition', 'Patient', 'Timepoint','Batch', 'Region') := lapply(.SD, as.factor), .SDcols = c('Condition', 'Patient', 'Timepoint','Batch', 'Region')]

# create a new group ID
ann[,Group := as.factor(paste0(Condition,'.', Patient))]

# match order of our expression matrix
ann <- ann[match(colnames(p.mat), Group),]

batch <- as.factor(ann$Batch)

# a two factor model; these are the variables of interest
mod <- model.matrix(~Region+Timepoint, data=ann)
rownames(mod) <- ann$Group
```

```{r}
clean.mat <- limma::removeBatchEffect(p.mat, batch, design=mod)

submat <- sweep(clean.mat, 1, apply(clean.mat, 1, median, na.rm=T))

hm <- Heatmap(submat,
        cluster_rows = clusterWNA(submat),
       # column_km=2,
        row_title= sprintf('%s complete proteins', nrow(submat)),
        show_row_names = F,
        column_names_gp = gpar(fontsize=4),
        top_annotation = colAnn)

hm
BackupAsPDF(hm, 'cleanedMat.medianScaled.heatmap')
```
Look at PCA

```{r}
submat <- clean.mat[complete.cases(clean.mat),]

pcaOut <- prcomp(t(submat))

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, ann, by.x = "rn", by.y = "Group", all.x = TRUE)
pcaDT[, region := ifelse(Condition %like% 'HMT', 'High', 'Low')]
pcaDT[, timepoint := tstrsplit(Condition, '_', keep=2)]

g <- ggplot(data.table(PC=seq(1,length(pcaPercentVar), by=1), var=pcaPercentVar)[PC <= 10,], aes(x=as.factor(PC), y=var)) +
  geom_bar(stat='Identity') +
  xlab('PC') +
  ylab('variance explained')
BackupAsPDF(g,'protein.pca.varExplained.cleaned.barplot')


#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2, fill = Batch, shape=Batch)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
 # scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_fill_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:22) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, 'protein.clean.pca.PC12.batchCol')

#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2, fill = Condition, shape=Batch)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
 # scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_fill_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:22) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, 'protein.clean.pca.PC12.conditionCol')


#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2, fill = Timepoint, shape=Batch)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
 # scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_fill_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:22) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
BackupAsPDF(p, 'protein.clean.pca.PC12.conditionCol')

```

Clustering of corrected intensities gives better group seperation. Save output for MSstats DE testing. 

```{r}
clean.dt <- reshape2::melt(clean.mat)
setnames(clean.dt, new=c('Protein', 'sample', 'adjLogIntensities'))

p.quant <- merge(x=p.quant, y=clean.dt, by=c('Protein','sample'))

p.quant[, oldLogIntensities := LogIntensities]
p.quant[, LogIntensities := adjLogIntensities]

fwrite(p.quant, ScriptAndDatedFileName('proteinLvl.quant.adjusted.csv'))
```

