---
title: "AFmultimer.FirstLook"
author: "Martin Gordon"
date: "2024-01-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## AF multimer first look at output results

Check out the results and 


Issue pulling output from the pkl file, only safe to open these files in the same environment they were created
Try opening this in the singularity enviornment
My thinking is this jaxlib environmnt needed to open the file is now depreciated but wasnt built for arm64.. If I try open in a x64_84 conda env will be clashes with 
For some reason I need an older version of jax installed  in env to open the pkl files.. is this an error with this alphafold version?

```{r}
library(data.table)
library(reticulate) # need this to open picklefiles
library(magrittr)
```
```{r}
dataDir <-  './data/'

use_condaenv("python3.9.arm64", conda = "/Users/martingordon/anaconda3/bin/conda")
pd <- import('pandas')
df <- pd$read_pickle(paste0(dataDir, "result_model_1_multimer_v3_pred_0.pkl")) 
```

```{r}
df %>% str()


lapply(df, print)
View(df)
```

