---
title: "022025_prepPPIProphetMatrices"
author: "Martin Gordon"
date: "2025-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## intro 

Preparing the PPIprophet input matrices 
Take the raw data provided by Andrea (mising standards, so invent) and input


```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(stringr)
library(ComplexHeatmap)

source("../../utils/bp_utils/ManageScriptData.R")
```

read in the DIA-NN matrices and check for the 
```{r}
secms <- dir('./data/MASSIVE_data', pattern='*tsv', full.names = T) %>% 
  grep('_psm', ., invert = T, value=T)

secms.list <- lapply(secms, fread)
names(secms.list) <- gsub('\\.\\/data\\/MASSIVE\\_data\\/|\\.tsv', '', secms)

lapply(secms.list, colnames)
```

Conditions in libraries 1-3 not clear; just take the PhiPA3 example and lets see if we can generate a matrix and run the pipeline
If this works and looks similar scores to the original, then create a conda/docker environment
```{r}
lapply(secms.list, head)
phiPA3.dt <- secms.list[[4]]
```

Alright, lets conver thtis to a PCP matrix
Need to look up the DIANN colnames

data already filtered for Q value <= 0.01
Looks like things are already summarised to the protein level, so 

The key columns in DIA-NN are:
 matched proteins ('Protein.Group' column) and quantities for both precursor and proteins ('Precursor.Normalised' and 'PG.MaxLFQ', respectively), for each run in the experiment

```{r}
phiPA3.dt[Q.Value <= 0.01,]
phiPA3.dt[, .(Precursor.Id, Precursor.Charge,Modified.Sequence, Stripped.Sequence, Proteotypic, Protein.Group)]

phiPA3.dt[Run == 'PA01_PhiPA3_DIA_R2_F57_redo_BE9_1_11973' & Protein.Group == 'Q9I1H1']
```

Is this protein level intensity or peptide? Protein level intensity, but the peptides are returned

```{r}
colnames(phiPA3.dt)
# no multi-elute peptides; now want to try a summarisation method..
# I guess as a first pass, we just need to check if this actually works and we pull out any similar complexes;
# for now just use TMP normalization by MSstats
unique(phiPA3.dt[,.N, by=.(Protein.Group, Modified.Sequence, Precursor.Charge, Run)]$N )

# maxLFQ is unique to each protein so easy to save as a matrix and scale before inputting to the algorithm
any(duplicated(phiPA3.dt[,.N, by=.(Protein.Group, PG.MaxLFQ, Run)][,.(Protein.Group, Run)]))
```
Lets tidy the columns and summarise with msstats
```{r}
# same nromalization is turned off
phiPA3.dt[Precursor.Normalised != Precursor.Quantity]
```

Tidy for MSS input
```{r}
phiPA3.dt <- phiPA3.dt[,.(Run, Condition='phiPA3', ProteinName = Protein.Group, gene=Genes, 
                          PeptideSequence=Modified.Sequence, PrecursorCharge=Precursor.Charge,  
                          BioReplicate = factor(gsub('_DIA_R', '', str_extract(Run, '_DIA_R[0-9]')), levels=c('1','2')),
                          fraction = str_extract(Run, '(?<=_F)[0-9]{1,2}(?=_)'),
                          FragmentIon = NA, IsotopeLabelType='L', PG.MaxLFQ, Intensity=Precursor.Quantity)]
```
Summarise at the level of each fraction
```{r}
phiPA3.dt[, Condition := paste0(Condition, '.', fraction)]

phiPA3.dt[Run =='PA01_PhiPA3_DIA_R2_F57_redo_BE9_1_11973' & gene == 'ambC']

phiPA3.dt[,.(Run, fraction, BioReplicate)] %>% 
  unique()
```

```{r}
#fwrite(phiPA3.dt, ScriptAndDatedFileName('phiPA3.clean.long.csv'))
phiPA3.dt <-  fread('~/Documents/projects/022025_IJKim_PPIProphetPilotStudy/022025_preparePPIprophetMats_data/2025_02_21_phiPA3.clean.long.csv')
phiPA3.dt[, ProductCharge := NA]


dp.out <- MSstats::dataProcess(phiPA3.dt, 
                               MBimpute =  FALSE, 
                               normalization = FALSE, # turn off for SECMS
                               summaryMethod = "TMP",
                               featureSubset = 'highQuality',
                               remove_uninformative_feature_outlier=T)


#saveRDS(dp.out, ScriptAndDatedFileName('phiPA3.rds'))
```

```{r}
dp.out <- readRDS('~/Documents/projects/022025_IJKim_PPIProphetPilotStudy/022025_preparePPIprophetMats_data/2025_02_21_phiPA3.rds')
p.quant <- setDT(dp.out$ProteinLevelData)
```

In general the profiles look quite similiar.. I wonder about the papers normalization method
```{r}
g <- ggplot(dp.out$ProteinLevelData, aes(x=GROUP, y=LogIntensities)) +
  geom_boxplot() +
  facet_wrap(~SUBJECT) +
  theme(axis.text.x = element_text(angle=90, size=6))

BackupAsPDF(g, 'phiPA3.tmp.boxplot')

g <- ggplot(phiPA3.dt, aes(x=Condition, y=log2(PG.MaxLFQ))) +
  geom_boxplot() +
  facet_wrap(~BioReplicate) +
  theme(axis.text.x = element_text(angle=90, size=6))

BackupAsPDF(g, 'phiPA3.maxLFQ.boxplot')

# set the column order for the data
col.order <- paste0('phiPA3.', seq(1,72,1))
```

create matrix, populate and upload

```{r}

mat.1 <- dcast(p.quant[SUBJECT == 1, ], Protein~GROUP, value.var='LogIntensities') %>% 
    as.matrix(rownames='Protein')

mat.2 <- dcast(p.quant[SUBJECT == 2, ], Protein~GROUP, value.var='LogIntensities') %>% 
    as.matrix(rownames='Protein')


mat.2
  
colnames(mat.1) <- gsub('[A-Za-z0-9]+[.]', '', colnames(mat.1))
colnames(mat.2) <- gsub('[A-Za-z0-9]+[.]', '', colnames(mat.2))

colorder <- as.character(seq(1,72,1))
  
message('reordering matrix columns and filling in missing fractions...') 
  # fills in missing fractions with NA so easy to populate
mat.1 <- mat.1[,  match(colorder, colnames(mat.1))]
mat.2 <- mat.2[,  match(colorder, colnames(mat.2))]

colnames(mat.1) <- colorder
colnames(mat.2) <- colorder
  
message('converting intensities to linear scale...') 
mat.1 <- 2^mat.1
mat.2 <- 2^mat.2
  
  # convert to 0
mat.1[is.na(mat.1)] <- 0
mat.2[is.na(mat.2)] <- 0
  
mat.list <- list(biorep1=mat.1, biorep2=mat.2)
```

write out the list of intensities to file
```{r}

mat.list
lapply(names(mat.list), function(x,i){

  p.mat <- data.table(x[[i]], keep.rownames = T)
  setnames(p.mat, old='rn', new='GN')
  #p.mat[, GN := multiUniprots2multiGenes(ID, species='HUMAN')]
  
  #col.order <- c('GN', 'ID', as.character(seq(1,72,1)))
  col.order <- c('GN', as.character(seq(1,72,1)))
  p.mat <-  p.mat[, ..col.order]
  
  fwrite(p.mat, sep='\t', ScriptAndDatedFileName(paste0(i, '.phiPA3.ppiProphet.txt')))
  
},x=mat.list)
```
Take the already prepared matrix and use that as input 

```{r}
testmat <- fread('data/MASSIVE_data/PA01_C3_KZ_directDIA.pr_matrix.tsv')
testmat <- fread('data/MASSIVE_data/PA01_KZ_library_SEC.tsv')
testmat
```


