---
title: "AF3_AF2_ASD_scoreComparison"
author: "Martin Gordon"
date: "2025-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(ggplot2)
library(magrittr)
library(ggrepel)
library(stringr)
library(ComplexHeatmap)
library(circlize)
library(viridis)
library(RColorBrewer)
library(hrbrthemes) # visually nice set of themes
library(showtext)
library(readxl)
library(pROC) #ROC curves


# fonts needed for hrbrthemes
library(extrafont)
font_import()
loadfonts()

# directly from google font
sysfonts::font_add_google("Roboto Condensed")
showtext_auto()


customTheme <- theme_ipsum_rc() +
  theme(panel.border = element_rect(color = "lightgrey", 
                                    fill = NA, 
                                    size = 1),
        axis.text.x = element_text(angle=90)
        )

source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/mg_utils/r_utils/plottingHelperFunctions.R")

#get col palette
col.pal <- getQualitativePalette(n=13)
```

Read in the AF2 scores and the AF3 scores for the comparison

```{r}
asd.ppi <- fread('./output/asd_ppi_summaryScores.csv')
asd.random <- fread('./output/asd_random_summaryScores.csv')
asd.af2 <- fread('./output/asd_AF2.ppi.csv')

# subset the AF2 set to bait-interaction and bait-random
asd.af2 <- asd.af2[asdppi_disjoint_group %in% c('bait_random', 'bait_int'),]

asd.ppi[, asdppi_disjoint_group := 'bait_int']
asd.random[, asdppi_disjoint_group := 'bait_random']
asd.af3 <- rbind(asd.ppi, asd.random)

asd.af3[,.N, by=asdppi_disjoint_group]

asd.af3[, pairOrdered := toupper(gsub('[.]seed.+$', '', model))]
summary.asd.af3 <- asd.af3[, .(mean_iptm=mean(iptm, na.rm=T), sd_iptm=sd(iptm, na.rm=T), mean_ptm=mean(ptm, na.rm=T), sd_ptm=sd(ptm, na.rm=T)), by=.(pairOrdered, asdppi_disjoint_group)]
```

join the two
```{r}
summary.asd.af3[, model := 'AF3']
asd.af2[, model := 'AF2']

comb.dt <-  rbind(x=summary.asd.af3[, .(pairOrdered, asdppi_disjoint_group, mean_iptm, sd_iptm, model)], 
                  y=asd.af2[, .(pairOrdered, asdppi_disjoint_group, mean_iptm, sd_iptm, model)])

```

```{r}

# af2 pairs are alphabetically ordered; match this for AF3
summary.asd.af3[pairOrdered == 'CELF4__AUP1']
asd.af2[pairOrdered == 'AUP1__CELF4']

comb.dt[, pair := pairOrdered]
comb.dt[, pairOrdered := paste0(sort(unlist(strsplit(pairOrdered, '__'))), collapse='__'), by=.I]
comb.dt[pair == 'CELF4__AUP1'] # sanity check 


# accidentally modelled one ppip in both directions... scores vs similar.. remove one row
comb.dt <- comb.dt[!pair %in% c('FOXP2__FOXP1', 'TCF7L2__CTNNB1')]

comb.dt[, label := '< 0.5 iPTM']
comb.dt[mean_iptm >= 0.5, label := '> 0.5 iPTM']

# find the pairs that overlap
sharedIDs <- intersect(comb.dt[model=='AF3', pairOrdered], comb.dt[model=='AF2', pairOrdered])

comb.dt[,inBothModels := ifelse(pairOrdered %in% sharedIDs, 'yes', 'no')]
#fwrite(comb.dt, ScriptAndDatedFileName('af3.af2.ASD.iptmScores.csv'))
```

```{r}
comb.dt<- fread('~/Documents/projects/120324_MGordon_MGordon_AF3_pipeline/AF3_AF2_ASD_scoreComparison_data/2025_01_27_af3.af2.ASD.iptmScores.csv')

comb.wide.dt <- dcast(comb.dt, pairOrdered~interaction(model, asdppi_disjoint_group), value.var = c('mean_iptm', 'sd_iptm', 'label'))
```

plot a scatterplot of the random and ppi set

```{r}
# add a sig identifier
comb.wide.dt[, bait_int.sig := ifelse(label_AF2.bait_int ==  '> 0.5 iPTM' & label_AF3.bait_int == '> 0.5 iPTM', 'AF2 & AF3 > 0.5 iPTM',
                                      ifelse(label_AF2.bait_int ==  '> 0.5 iPTM' & label_AF3.bait_int != '> 0.5 iPTM',  'AF2 > 0.5 iPTM',
                                              ifelse(label_AF2.bait_int !=  '> 0.5 iPTM' & label_AF3.bait_int == '> 0.5 iPTM',  'AF3 > 0.5 iPTM',
                                                     'AF2 & AF3 < 0.5 iPTM')))]


comb.wide.dt[, bait_random.sig := ifelse(label_AF2.bait_random ==  '> 0.5 iPTM' & label_AF3.bait_random == '> 0.5 iPTM', 'AF2 & AF3 > 0.5 iPTM',
                                      ifelse(label_AF2.bait_random ==  '> 0.5 iPTM' & label_AF3.bait_random != '> 0.5 iPTM',  'AF2 > 0.5 iPTM',
                                              ifelse(label_AF2.bait_random !=  '> 0.5 iPTM' & label_AF3.bait_random == '> 0.5 iPTM',  'AF3 > 0.5 iPTM',
                                                     'AF2 & AF3 < 0.5 iPTM')))]
                                            

g <- ggplot(comb.wide.dt[!is.na(bait_int.sig)], aes(x=mean_iptm_AF2.bait_int, y=mean_iptm_AF3.bait_int)) +
  geom_point(aes(color=bait_int.sig)) +
  geom_density_2d() +
  geom_hline(yintercept=0.5, linetype=2, alpha=0.4) +
  geom_vline(xintercept=0.5, linetype=2, alpha=0.4) +
  geom_abline(intercept=0, slope=1) +
  scale_color_ipsum() +
  ggtitle('AF2m vs AF3 iptm scores ASD bait-interactor set') +
  customTheme 
g
BackupAsPDF(g, 'asd.baitInt.scatterplot')

g <- ggplot(comb.wide.dt[!is.na(bait_random.sig)], aes(x=mean_iptm_AF2.bait_random, y=mean_iptm_AF3.bait_random)) +
  geom_point(aes(color=bait_random.sig)) +
  geom_density_2d() +
  geom_hline(yintercept=0.5, linetype=2, alpha=0.4) +
  geom_vline(xintercept=0.5, linetype=2, alpha=0.4) +
  geom_abline(intercept=0, slope=1) +
  scale_color_ipsum() +
  ggtitle('AF2m vs AF3 iptm scores ASD bait-random set') +
  customTheme 
g

BackupAsPDF(g, 'asd.baitRandom.scatterplot')
```
barplot of the number of sig things in each set

```{r}
summary.dt <- comb.dt[pairOrdered %in% sharedIDs,.N, by=.(model,asdppi_disjoint_group,label)][,.(prop=N/sum(N),label), by=.(model, asdppi_disjoint_group)][label != '< 0.5 iPTM']

g <- ggplot(summary.dt, aes(x=model, y=prop, fill=asdppi_disjoint_group)) +
  geom_bar(stat='identity',position='dodge') +
  ggtitle('Proportion of pairs mean iPTM > 0.5') +
  scale_fill_manual(values=col.pal) +
  customTheme 
g
BackupAsPDF(g, 'prop.ModelsPassiPTM.barplot')
```
boxplots of the average iptm scores for each
Want a grouped boxplot with bait/random side-by-side

```{r}
library(ggbeeswarm)

g <- ggplot(comb.dt, aes(x=model, y=mean_iptm, color=asdppi_disjoint_group)) +
  geom_hline(yintercept=0.5, linetype=2, alpha=0.3) +
  geom_boxplot() +
  #geom_point(data=comb.dt[mean_iptm > 0.5], aes(fill =  asdppi_disjoint_group), size = 1.25, shape = 21, alpha=0.7, position = position_jitterdodge(jitter.width=0.2, dodge.width = .75, seed=1), color='black') +
  ggtitle('iPTM distributions') +
  scale_color_manual(values=col.pal) +
  scale_fill_manual(values=col.pal) +
  customTheme 
g
BackupAsPDF(g, 'af.iptm.distributions.boxplot')
```
Lets look at the empirical cumulative distribution function and see the curves

```{r}
g <- ggplot(comb.dt, aes(x=mean_iptm, color=interaction(model,asdppi_disjoint_group))) +
         stat_ecdf(pad=T) +
  customTheme +
  scale_fill_ipsum() +
  guides(color=guide_legend("dataset"))

BackupAsPDF(g, 'iptm.iptm.ecdf.linechart')
```

create a ROC curve and plot the two against each other; see which score is better at distinguishing FP from TP 
Take only the shared baits cateogry


```{r}
# first run for af2
subdt <- comb.dt[pairOrdered %in% sharedIDs,]
subdt[, asdppi_disjoint_group := factor(asdppi_disjoint_group, levels=c('bait_random', 'bait_int'))]

roc.af2 <- roc(subdt[model == 'AF2',]$asdppi_disjoint_group, subdt[model == 'AF2',]$mean_iptm)
roc.af3 <- roc(subdt[model == 'AF3',]$asdppi_disjoint_group, subdt[model == 'AF3',]$mean_iptm)


roc.af2.dt <- data.table(tpp=roc.af2$sensitivities*100, ## tpp = true positive percentage
                         fpp=(1 - roc.af2$specificities)*100, ## fpp = false positive percentage
                         thresholds=roc.af2$thresholds,
                         model='AF2')
                        
roc.af3.dt <- data.table(tpp=roc.af3$sensitivities*100, ## tpp = true positive percentage
                         fpp=(1 - roc.af3$specificities)*100, ## fpp = false positive percentage
                          thresholds=roc.af3$thresholds,
                          model='AF3')

roc.comb <- rbind(roc.af2.dt, roc.af3.dt)


g <- ggplot(roc.comb, aes(x=fpp, y=tpp, color=model)) +
  geom_step(linewidth=1, alpha=0.75) +
  ggtitle('Full dataset: iPTM as predictor of PPI/random set') +
  labs(x="False Positive Percentage", y="True Postive Percentage") +
  scale_color_manual(values=col.pal) +
  coord_cartesian(xlim=c(0,100), ylim=c(0,100)) + 
  customTheme

BackupAsPDF(g, 'full.data.roc.linechart')
```
Ok, using the full dataset, we see no seperation; what about looking at > .25?

```{r}
roc.af2 <- roc(subdt[model == 'AF2' & mean_iptm >= 0.25,]$asdppi_disjoint_group, subdt[model == 'AF2' & mean_iptm >= 0.25,]$mean_iptm)
roc.af3 <- roc(subdt[model == 'AF3' & mean_iptm >= 0.25,]$asdppi_disjoint_group, subdt[model == 'AF3' & mean_iptm >= 0.25,]$mean_iptm)

roc.af2.dt <- data.table(tpp=roc.af2$sensitivities*100, ## tpp = true positive percentage
                         fpp=(1 - roc.af2$specificities)*100, ## fpp = false positive percentage
                         thresholds=roc.af2$thresholds,
                         model='AF2')
                        
roc.af3.dt <- data.table(tpp=roc.af3$sensitivities*100, ## tpp = true positive percentage
                         fpp=(1 - roc.af3$specificities)*100, ## fpp = false positive percentage
                          thresholds=roc.af3$thresholds,
                          model='AF3')

roc.comb <- rbind(roc.af2.dt, roc.af3.dt)


g <- ggplot(roc.comb, aes(x=fpp, y=tpp, color=model)) +
  geom_step(linewidth=1, alpha=0.75) +
  ggtitle(' iPTM > .25: iPTM as predictor of PPI/random set') +
  labs(x="False Positive Percentage", y="True Postive Percentage") +
  scale_color_manual(values=col.pal) +
  coord_cartesian(xlim=c(0,100), ylim=c(0,100)) + 
  customTheme
g
BackupAsPDF(g, 'iptm25threshold.data.roc.linechart')
```
```{r}

roc.af2 <- roc(subdt[model == 'AF2' & mean_iptm >= 0.5,]$asdppi_disjoint_group, subdt[model == 'AF2' & mean_iptm >= 0.5,]$mean_iptm)
roc.af3 <- roc(subdt[model == 'AF3' & mean_iptm >= 0.5,]$asdppi_disjoint_group, subdt[model == 'AF3' & mean_iptm >= 0.5,]$mean_iptm)

roc.af2.dt <- data.table(tpp=roc.af2$sensitivities*100, ## tpp = true positive percentage
                         fpp=(1 - roc.af2$specificities)*100, ## fpp = false positive percentage
                         thresholds=roc.af2$thresholds,
                         model='AF2')
                        
roc.af3.dt <- data.table(tpp=roc.af3$sensitivities*100, ## tpp = true positive percentage
                         fpp=(1 - roc.af3$specificities)*100, ## fpp = false positive percentage
                          thresholds=roc.af3$thresholds,
                          model='AF3')

roc.comb <- rbind(roc.af2.dt, roc.af3.dt)


g <- ggplot(roc.comb, aes(x=fpp, y=tpp, color=model)) +
  geom_step( alpha=0.75) +
  geom_point() +
  ggtitle(' iPTM > .50: iPTM as predictor of PPI/random set') +
  labs(x="False Positive Percentage", y="True Postive Percentage") +
  scale_color_manual(values=col.pal) +
  coord_cartesian(xlim=c(0,100), ylim=c(0,100)) + 
  customTheme
g
BackupAsPDF(g, 'iptm50threshold.data.roc.linechart')
```
plot the ratio of the ecdf for af3 and af2

```{r}
af.mat <- dcast(comb.dt[inBothModels == 'yes' & model=='AF2', ], pairOrdered~paste0(model,'.',asdppi_disjoint_group), value.var = 'mean_iptm') %>% 
  as.matrix(rownames='pairOrdered')

comb.dt[asdppi_disjoint_group == 'bait_int' & model=='AF2', unique(pairOrdered)] %>% length()
comb.dt[asdppi_disjoint_group == 'bait_random' & model=='AF3', unique(pairOrdered)] %>% length()

apply(af.mat, 2, ecdf)

decdf <- function(x, baseline, treatment)  ecdf(baseline)(x) - ecdf(treatment)(x)

ecdf(comb.dt[asdppi_disjoint_group == 'bait_random' & model=='AF3', mean_iptm])
ecdf(comb.dt[asdppi_disjoint_group == 'bait_int' & model=='AF3', mean_iptm])
ecdf(comb.dt[asdppi_disjoint_group == 'bait_random' & model=='AF3', mean_iptm]) - ecdf(comb.dt[asdppi_disjoint_group == 'bait_int' & model=='AF3', mean_iptm])



ecdf(comb.dt[asdppi_disjoint_group == 'bait_int' & model=='AF3', mean_iptm])()
```

