---
title: "210224_MultipleSearchesComparison"
author: "Martin Gordon"
date: "2024-02-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Multiple Searches Comparison for the AViDD AB data
To identify issues with previous search, we will look at the four searches together, plot and share the QC results

```{r}
library(magrittr)
library(data.table)
library(circlize)
library(ComplexHeatmap)
library(circlize)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(MSstats)
library(ggrepel)
library(readxl)
library(nichenetr)
library(patchwork) # combine plots
library(viridis)
library(ggvenn)


source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```
metadata on mice available in the docs column; would be nice to use mouse weight or viral load in the analysis
Most simple approach would be to include in the PCA

For now reach out to Jack and proceed with the analysis

```{r}
spec.list <-  list(dda_bg = fread('./data/DDA Lib.tsv'),
                   dda_noBg = fread('./data/DDA Lib No Background.tsv'),
                   dual_bg = fread('./data/Dual Lib .tsv'),
                   dual_noBg = fread('./data/2024.02.21 - Mac1 AB - Dual No Bg_Report_MSStats Format (Normal).tsv'),
                   allSamples = fread('../020524_AVIDD_ABproteomics/data/AB_Report_MSstats_Format.tsv'))

# sample ids 
keys <- fread('./data/keys.txt')
keys
keys[,.N, .(Condition,RawFile)]
# contrasts
contrasts.txt <- fread('./data/contrasts.txt', sep='-', header=F)
```

```{r}
# the dual DIA and DDA is a much bigger library; more things detected, but perhaps also an issue
lapply(spec.list, dim)
lapply(spec.list, str)
```

combine with sample metadata
```{r}
keys$RawFile %>%  unique()
lapply(spec.dt.list, function(x){
  x$Run %>%  unique()
})

spec.dt.list <- lapply(spec.list, function(X){
  
  spec.dt <- merge(x=X[,!c("Condition", "BioReplicate")], y=keys[,.(Condition, BioReplicate, RawFile, Fraction,IsotopeLabelType)], by.x='Run', by.y='RawFile')
  return(spec.dt)
})
```

Read in the MOUSE ID mapping file 

```{r}
# lets map the homologs to the mss output and write the complete and tidied data to file 
mh_mappings <- fread('~/Documents/utils/mg_utils/data/mouseHumanIDConversion.txt')

# gene sets OI
piGenes <- read_xlsx('./docs/1-s2.0-S0092867423009157-mmc4.xlsx', col_names = T, sheet=7) %>% 
  setDT(.) %>% 
  .[category == 'Proinflammatory'] %>% 
  .[,unique(gene)]

isGenes <- fread('./docs/ISGs.txt', header=F) %>% 
  .[,V1]
```


For the searches, fix the SARS-CoV2 protein names



need to check out what is in each of these runs... 
Seems the other treatment conditions are dropped in this; only interested in comparison to mock,WT and vehicle

```{r}
spec.list[[3]][!Run %in% spec.list[[4]]$Run,unique(Run)]
#"ex014889.raw"
spec.list[[4]][Run == 'ex014889.raw',]

#mock biorep 5
keys[RawFile == 'ex014889.raw',]

# sanity test 
tidy.keys <- keys[,.N, by=.(RawFile,Condition,BioReplicate)][,.(RawFile,Condition, BioReplicate)][!Condition %in% c('x3769_7D','x3769_4D', 'x4052_4D','x4052_7D'),][order(RawFile),]

# check the merging 
lapply(spec.dt.list, function(x){
  
  print(x[,.N, by=.(Run,Condition, BioReplicate)][, .(Run,Condition,BioReplicate)][order(Run),])
  tidy.keys
})
```
These are only the samples

```{r}
spec.dt.list <- lapply(spec.dt.list, function(x){
  hist(log2(x$Intensity))
  spec.dt <-  x[Intensity > 2^5, ]
  hist(log2(spec.dt$Intensity))
  return(spec.dt)
})
```
For now drop the sample with the missing rep and process

```{r}
spec.dt.list[[4]] <-  NULL

```

quick qc
----

multi-feature peptide ions?
Seems all the featues are unique

```{r}
# no multiple feature peptide ions detected
spec.dt[,.N, by=.(PeptideSequence,PrecursorCharge,Run)][N >1]


lapply(spec.dt.list, function(x){
  
  x[, .N, by=.(PeptideSequence,PrecursorCharge, Run)][N > 1]
})
```


boxplots of intensities

```{r}
col.pal <-  randomcoloR::distinctColorPalette(length(unique(spec.dt.list[[3]]$Condition)))

lapply(seq_along(spec.dt.list), function(x,n,i){
  
  
  g <- ggplot(x[[i]], aes(x=reorder(interaction(Condition,BioReplicate)), y=log2(Intensity), fill=Condition)) +
    geom_boxplot() +
    theme_classic() +
    labs(title = n[[i]]) +
    scale_fill_manual(values=col.pal) +
    theme(axis.text.x = element_text(angle=90))

  g
  BackupAsPDF(g, paste0(n[[i]],'.raw.intensities.boxplot'))
  
},x=spec.dt.list, n=names(spec.dt.list))

```
n features
---
This bg imputation doesnt seem to make a big difference in the samples

```{r}

lapply(seq_along(spec.dt.list), function(x,n,i){
  
  
  g <- ggplot(x[[i]][,.N, by=.(Condition,BioReplicate)], aes(x=reorder(interaction(Condition,BioReplicate)), y=N, fill=Condition)) +
    geom_bar(stat='Identity') +
    labs(title = n[[i]]) +
    scale_fill_manual(values=col.pal) +
    theme_classic() +
    theme(axis.text.x = element_text(angle=90))

  g
  BackupAsPDF(g, paste0(n[[i]],'.Nfeatures.barplot'))
  
},x=spec.dt.list, n=names(spec.dt.list))

```
PCA of the features 

```{r}
lapply(seq_along(spec.dt.list), function(x,n,i){
  
  
  featureMat <-  dcast(x[[i]], 
                  paste0(PeptideSequence, PrecursorCharge)~interaction(Condition,BioReplicate), value.var = "Intensity") %>% as.matrix(rownames = 'PeptideSequence')
  featureMat <- featureMat[complete.cases(featureMat), ]
  
  # 
  pcaOut <- prcomp(t(log2(featureMat)))
  colInfo <- data.table(colname = colnames(featureMat))
  colInfo[,c("status","rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]

  #PCA
  pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
  pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
  pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
  #plot first two components
  p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = status, shape = rep)) + 
    geom_point(alpha=1.0, size=4) + 
    ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
    theme_bw() + 
    labs(title = n[[i]]) +
    xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
    ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
    ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_shape_manual(values = 21:25) +
    #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
    guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

  p
  BackupAsPDF(p, paste0(n[[i]],'.completeFeatures.pca.'))

  
},x=spec.dt.list, n=names(spec.dt.list))

```
feature matrix 

```{r}
lapply(seq_along(spec.dt.list), function(x,n,i){
  
  featureMat <-  dcast(x[[i]], 
                  paste0(PeptideSequence, PrecursorCharge)~interaction(Condition,BioReplicate), value.var = "Intensity") %>% as.matrix(rownames = 'PeptideSequence')
  
  print(featureMat)
  featureMat <-  log2(featureMat)

  dim(featureMat)
  dim(featureMat[complete.cases(featureMat),])

  # create annotation
  ann <- data.table(replicate=str_extract(colnames(featureMat),'[1-5]$'),
                  treatment=sub("[_.].+", "", colnames(featureMat)),
                  timepoint=ifelse(grepl("[47]D", colnames(featureMat)), str_extract(colnames(featureMat), "[47]D"), 'na'))


  colours <- list('timepoint' = c('na' = '#4477AA', '4D'="#228833", '7D'="#CCBB44"),
                'replicate' = c('1' = '#4477AA', '2'='#66CCEE', '3'="#228833", '4'="#CCBB44", '5'='#EE6677'),
                'treatment' =c('Mock' = col.pal[1], 'Vehicle'=col.pal[2], 'N1062D'=col.pal[3], 'x3769'=col.pal[4], 'x4052'=col.pal[5], 'x4206'=col.pal[6])
                )

  colAnn <- HeatmapAnnotation(df = ann, col = colours)

  # subset the data
  submat <- featureMat[sample(rownames(featureMat), 2000), ]
  


  hm <- Heatmap(submat, cluster_rows=clusterWNA(submat), show_row_names = F, cluster_columns = T, top_annotation = colAnn, name='LogIntensities', column_names_gp = gpar(fontsize=6))
  hm
  BackupAsPDF(hm, paste0(n[[i]],'.features.clustered.heatmap'))

  submat <- sweep(submat,1, STATS=apply(submat, 1, median, na.rm=T))
  hm <- Heatmap(submat, cluster_rows=clusterWNA(submat), show_row_names = F, cluster_columns = T, top_annotation = colAnn, name='LogIntensities/Median',  column_names_gp = gpar(fontsize=6))
  hm
  BackupAsPDF(hm, paste0(n[[i]],'.features.clustered.medianScaled.heatmap'))
  
}, x=spec.dt.list, n=names(spec.dt.list))
```
Protein Summarisation

convert to msstats format
```{r}
mssInput.list <-  lapply(spec.dt.list,  function(x){
  
  mss.dt <- specFileToCompleteMSstats(x)
  mss.dt[,  IsotopeLabelType := 'L']
  
  return(mss.dt)
})

names(mssInput.list) <- names(spec.dt.list)
```
summarise to protein level
-----
- NB dont use the specFiletoCompleteMSstats as it doesnt seem to handle the fractionation in these files (perhaps its a result of the )
```{r}
dp.out.list <-  lapply(spec.dt.list,  function(x){
  
 dp.out <- MSstats::dataProcess(x, 
                              MBimpute =  FALSE, 
                              featureSubset = "highQuality", 
                              remove_uninformative_feature_outlier = TRUE)
 
 return(dp.out)
})

names(dp.out.list) <- names(mssInput.list)

```

Write out the RDS files

```{r}
lapply(seq_along(dp.out.list), function(x,n,i){
  
  saveRDS(x[[i]], ScriptAndDatedFileName(paste0(n[[i]], ".rds")))
},x=dp.out.list, n=names(dp.out.list))
```

Save the individual files

```{r}
lapply(seq_along(dp.out.list), function(x,n,i){
  fwrite (x[[i]]$ProteinLevelData, ScriptAndDatedFileName(paste0(n[[i]], ".ProteinLevelData.csv")))
  fwrite (x[[i]]$FeatureLevelData,  ScriptAndDatedFileName(paste0(n[[i]], ".FeatureLevelData.csv")))
},x=dp.out.list, n=names(dp.out.list))
```

Now read in the protein quant, look at the sample clustering and look at the ISG profiles

```{r}
p.quant.list <-  list(dda_bg = fread('./AB_multiplesearchesComparison_data/2024_02_21_dda_bg.ProteinLevelData.csv'),
                   dda_noBg = fread('./AB_multiplesearchesComparison_data/2024_02_21_dda_noBg.ProteinLevelData.csv'),
                   dual_bg = fread('./AB_multiplesearchesComparison_data/2024_02_21_dual_bg.ProteinLevelData.csv'))


lapply(p.quant.list, head)

names(dp.out.list)
```
QC plots

boxplots of normalised intensities

```{r}

lapply(seq_along(p.quant.list), function(x,n,i){
  
  g <- ggplot(x[[i]], aes (x= interaction(SUBJECT, GROUP), y = LogIntensities, fill = GROUP)) +
  geom_boxplot() +
  labs(title=n[[i]]) +
  theme_classic() +
  scale_fill_manual(values=col.pal) +
  theme(axis.text.x = element_text(angle=90))
  g
  BackupAsPDF(g, paste0(n[[i]],'protein.intensities.boxplot'))
  
}, x=p.quant.list, n=names(p.quant.list))

```

Protein Counts per sample
Weird.. we found a lot more peptides/fragments, but N proteins relatively consistent across samples

```{r}

lapply(seq_along(p.quant.list), function(x,n,i){
  
  g <- ggplot(x[[i]][,.N,by=.(GROUP,SUBJECT)], aes(x=reorder(interaction(GROUP,SUBJECT)), y = N, fill = GROUP)) +
  geom_bar(stat='Identity') +
  labs(title=n[[i]]) +
  theme_classic() +
  scale_fill_manual(values=col.pal) +
  theme(axis.text.x = element_text(angle=90))
  g
  BackupAsPDF(g, paste0(n[[i]],'nProteins.barplot'))
  
}, x=p.quant.list, n=names(p.quant.list))

```
pca and sample clustering 
```{r}
lapply(seq_along(p.quant.list), function(x,n,i){
  
  prot.mat <- dcast (x[[i]], Protein~interaction(GROUP,SUBJECT), value.var = "LogIntensities") |> as.matrix (rownames = "Protein")
  prot.mat <- prot.mat[complete.cases(prot.mat),]
  
  # run pc
  pcaOut <- prcomp(t(prot.mat))
  
  
  colInfo <- data.table(colname = colnames(prot.mat))
  colInfo[,c("status", "rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]
  
  

  #PCA
  pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
  pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
  pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
  pcaDT[, `:=`(timepoint = ifelse(grepl("[47]D", rn), str_extract(rn, "[47]D"), 'na'),
             treat = gsub("[_.].+", "", rn)
             )]
  
  
  # scree plot of the PC var explained
  q <- qplot(1:10,  pcaPercentVar[1:10]) +
  geom_line() +
  xlab('Princpal Component 1:10') +
  ylab(('% Variance explained')) +
  ggtitle(paste0(n[[i]],' scree plot'))

  BackupAsPDF(q, paste0(n[[i]],'pc.screeplot.'))
  

  #plot first two components
  p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = status, shape = rep)) + 
    geom_point(alpha=1.0, size=4) + 
    ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
    theme_bw() + 
    ggtitle(n[[i]]) +
    xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
    ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
    #ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_shape_manual(values = 21:25) +
    #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
    guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ))

  BackupAsPDF(p, paste0(n[[i]], 'completeProteins.pc1pc2.pca.'))
  
  
  # plot pc2 and pc3
  p <- ggplot (pcaDT, aes(x=PC2, y=PC3,  fill = status, shape = rep)) + 
    geom_point(alpha=1.0, size=4) + 
    ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
    theme_bw() + 
    ggtitle(n[[i]]) +
    xlab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
    ylab (sprintf ("PC3, %.1f%%", pcaPercentVar[3])) + 
    #ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_shape_manual(values = 21:25) +
    #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
    guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ))

  BackupAsPDF(p, paste0(n[[i]], 'completeProteins.pc2pc3.pca.'))
  
}, x=p.quant.list, n=names(p.quant.list))
```

Out of curiosity, based solely on ISGs how do the samples cluster?
Look v different with bg turned off compared to others...

```{r}
lapply(seq_along(p.quant.list), function(x,n,i){
  x[[i]][, gene := multiUniprots2multiGenes(Protein, species='MOUSE')]
},x=p.quant.list, n=names.quant.list)
```


```{r}
isGenes
mh_mappings
# map to MOUSE IDs and subset
lapply(seq_along(p.quant.list), function(x,n,i){
  
  prot.mat <- dcast (x[[i]], Protein~interaction(GROUP,SUBJECT), value.var = "LogIntensities") |> as.matrix (rownames = "Protein")
  prot.mat <- prot.mat[complete.cases(prot.mat),]
  
  # subset to ISG set based on mouse mapping
  # less features detected in the second set... w/o imputation, less complete.cases possibly?
  prot.mat <- prot.mat[rownames(prot.mat) %in% mh_mappings[Symbol.human %in% isGenes, SWISS_PROT_IDs.mouse],]
  
  # run pc
  pcaOut <- prcomp(t(prot.mat))
  
  colInfo <- data.table(colname = colnames(prot.mat))
  colInfo[,c("status", "rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]
  
  #PCA
  pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
  pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
  pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
  pcaDT[, `:=`(timepoint = ifelse(grepl("[47]D", rn), str_extract(rn, "[47]D"), 'na'),
             treat = gsub("[_.].+", "", rn)
             )]
  
  # scree plot of the PC var explained
  q <- qplot(1:10,  pcaPercentVar[1:10]) +
  geom_line() +
  xlab('Princpal Component 1:10') +
  ylab(('% Variance explained')) +
  ggtitle(paste0(n[[i]],' scree plot'))
  BackupAsPDF(q, paste0(n[[i]],'pc.ISGsubset.screeplot.'))
  

  #plot first two components
  p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = status, shape = rep)) + 
    geom_point(alpha=1.0, size=4) + 
    ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
    theme_bw() + 
    ggtitle(paste0(n[[i]], ' ISG subset')) +
    xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
    ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
    #ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_shape_manual(values = 21:25) +
    #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
    guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ))

  BackupAsPDF(p, paste0(n[[i]], 'ISGsubset.pc1pc2.pca.'))
  
  
},x=p.quant.list, n=names(p.quant.list))


```

look at sample ISG sample clustering and share with Jack, then do the DEA

```{r}
  
lapply(seq_along(p.quant.list), function(x,n,i){
  
  prot.mat <- dcast (x[[i]], Protein~interaction(GROUP,SUBJECT), value.var = "LogIntensities") |> as.matrix (rownames = "Protein")

  # subset to ISG set based on mouse mapping
  # less features detected in the second set... w/o imputation, less complete.cases possibly?
  prot.mat <- prot.mat[rownames(prot.mat) %in% mh_mappings[Symbol.human %in% isGenes, SWISS_PROT_IDs.mouse],]
  rownames(prot.mat) <- multiUniprots2multiGenes(rownames(prot.mat), species='MOUSE')
  
  # create annotation
  ann <- data.table(replicate=str_extract(colnames(prot.mat),'[1-5]$'),
                  treatment=sub("[_.].+", "", colnames(prot.mat)),
                  timepoint=ifelse(grepl("[47]D", colnames(prot.mat)), str_extract(colnames(prot.mat), "[47]D"), 'na'))

  colours <- list('timepoint' = c('na' = '#4477AA', '4D'="#228833", '7D'="#CCBB44"),
                'replicate' = c('1' = '#4477AA', '2'='#66CCEE', '3'="#228833", '4'="#CCBB44", '5'='#EE6677'),
                'treatment' =c('Mock' = col.pal[1], 'Vehicle'=col.pal[2], 'N1062D'=col.pal[3], 'x3769'=col.pal[4], 'x4052'=col.pal[5], 'x4206'=col.pal[6])
                )

  colAnn <- HeatmapAnnotation(df = ann, col = colours)
  

  hm <- Heatmap(prot.mat, 
                cluster_rows=clusterWNA(prot.mat), 
                show_row_names = T, 
                cluster_columns = T, 
                top_annotation = colAnn,
                name='LogIntensities',
                column_names_gp = gpar(fontsize=6), 
                row_names_gp = gpar(fontsize=6))
  hm
  #BackupAsPDF(hm, 'ISGproteins.clustered.heatmap')

  subMat <- sweep(prot.mat,1, STATS=apply(prot.mat, 1, median, na.rm=T))

  hm <- Heatmap(subMat, 
                cluster_rows=clusterWNA(subMat), 
               show_row_names = T, 
               cluster_columns = F, 
              #top_annotation = colAnn, 
              name='LogIntensities/rowMedian',  
              column_names_gp = gpar(fontsize=6), 
              row_names_gp = gpar(fontsize=8), 
              column_split=sub("[.][1-5]$", "", colnames(subMat)),
              column_title_gp = gpar(fontsize=5))
  draw(hm, column_title=n[[i]])
  
  BackupAsPDF(draw(hm, column_title=paste0(n[[i]], ' ISG subset')), paste0(n[[i]], '.isg.subset.medianScaled.heatmap'))
  
},x=p.quant.list, n=names(p.quant.list))
#BackupAsPDF(hm, 'ISGproteins.medianscaled.heatmap')
  
```
Take a random subset and view the clustering; do the same samples still stand out?

```{r}
  
lapply(seq_along(p.quant.list), function(x,n,i){
  
  prot.mat <- dcast (x[[i]], Protein~interaction(GROUP,SUBJECT), value.var = "LogIntensities") |> as.matrix (rownames = "Protein")
  submat <-  scale(prot.mat, center=T, scale=T)
  submat <- submat[sample(rownames(submat),2000), ]
  
  rownames(submat) <- multiUniprots2multiGenes(rownames(submat), species='MOUSE')
  
  # create annotation
  ann <- data.table(replicate=str_extract(colnames(submat),'[1-5]$'),
                  treatment=sub("[_.].+", "", colnames(submat)),
                  timepoint=ifelse(grepl("[47]D", colnames(submat)), str_extract(colnames(submat), "[47]D"), 'na'))

  colours <- list('timepoint' = c('na' = '#4477AA', '4D'="#228833", '7D'="#CCBB44"),
                'replicate' = c('1' = '#4477AA', '2'='#66CCEE', '3'="#228833", '4'="#CCBB44", '5'='#EE6677'),
                'treatment' =c('Mock' = col.pal[1], 'Vehicle'=col.pal[2], 'N1062D'=col.pal[3], 'x3769'=col.pal[4], 'x4052'=col.pal[5], 'x4206'=col.pal[6])
                )

  colAnn <- HeatmapAnnotation(df = ann, col = colours)
  

  hm <- Heatmap(submat, 
                cluster_rows=clusterWNA(submat), 
                show_row_names = F, 
                cluster_columns = T, 
                top_annotation = colAnn,
                name='z-score(Ints.)',
                column_names_gp = gpar(fontsize=6), 
                row_names_gp = gpar(fontsize=6))
  hm
  #BackupAsPDF(hm, paste0(n[[i]],'.proteins.z-scores.clustered.heatmap'))

  
  submat <- prot.mat[sample(rownames(prot.mat),2000), ]
  subMat <- sweep(prot.mat,1, STATS=apply(prot.mat, 1, median, na.rm=T))

  hm <- Heatmap(subMat, 
                cluster_rows=clusterWNA(subMat), 
               show_row_names = F, 
               cluster_columns = T, 
              #top_annotation = colAnn, 
              name='LogIntensities/rowMedian',  
              column_names_gp = gpar(fontsize=6), 
              row_names_gp = gpar(fontsize=8), 
              cluster_column_slices = T,
              column_split=sub("[.][1-5]$", "", colnames(subMat)),
              column_title_gp = gpar(fontsize=5))
  draw(hm, column_title=n[[i]])
  
  BackupAsPDF(draw(hm, column_title=n[[i]]), paste0(n[[i]], '.proteins.medianScaled.heatmap'))
  
},x=p.quant.list, n=names(p.quant.list))

```
```{r}
lapply(seq_along(p.quant.list), function(x,n,i){
  
  prot.mat <- dcast (x[[i]], Protein~interaction(GROUP,SUBJECT), value.var = "LogIntensities") |> as.matrix (rownames = "Protein")

  submat <- prot.mat[sample(rownames(prot.mat),2000), ]
  subMat <- sweep(prot.mat,1, STATS=apply(prot.mat, 1, median, na.rm=T))
  
    # create annotation
  ann <- data.table(replicate=str_extract(colnames(submat),'[1-5]$'),
                  treatment=sub("[_.].+", "", colnames(submat)),
                  timepoint=ifelse(grepl("[47]D", colnames(submat)), str_extract(colnames(submat), "[47]D"), 'na'))

  colours <- list('timepoint' = c('na' = '#4477AA', '4D'="#228833", '7D'="#CCBB44"),
                'replicate' = c('1' = '#4477AA', '2'='#66CCEE', '3'="#228833", '4'="#CCBB44", '5'='#EE6677'),
                'treatment' =c('Mock' = col.pal[1], 'Vehicle'=col.pal[2], 'N1062D'=col.pal[3], 'x3769'=col.pal[4], 'x4052'=col.pal[5], 'x4206'=col.pal[6])
                )

  colAnn <- HeatmapAnnotation(df = ann, col = colours)

  hm <- Heatmap(subMat, 
                cluster_rows=clusterWNA(subMat), 
               show_row_names = F, 
               cluster_columns = T, 
              top_annotation = colAnn, 
              name='LogIntensities/rowMedian',  
              column_names_gp = gpar(fontsize=6), 
              row_names_gp = gpar(fontsize=8), 
            #  cluster_column_slices = T,
              #column_split=sub("[.][1-5]$", "", colnames(subMat)),
              column_title_gp = gpar(fontsize=5))
  draw(hm, column_title=n[[i]])
  
  #BackupAsPDF(draw(hm, column_title=n[[i]]), paste0(n[[i]], '.proteins.medianScaled.clusteering.heatmap'))
  
},x=p.quant.list, n=names(p.quant.list))
```

PW comparisons of each of the setups
-----

```{r}
contrasts.list <- split(contrasts.txt, seq(nrow(contrasts.txt)))
names(contrasts.list) <- apply(contrasts.txt, 1, function(x){ paste(x[1], x[2], sep='-') }) 

dp.proc.list <- lapply(seq_along(dp.out.list), function(x,n,i){
  
  f.quant <- setDT(x[[i]]$FeatureLevelData)
  p.quant <- setDT(x[[i]]$ProteinLevelData)
  
  p.quant[, SUBJECT := interaction(GROUP,SUBJECT)]
  f.quant[, SUBJECT := interaction(GROUP,SUBJECT)]
  
  x[[i]]$FeatureLevelData <- f.quant
  x[[i]]$ProteinLevelData <- p.quant
  
  return(x[[i]])

},x=dp.out.list, n=names(dp.out.list))


names(dp.proc.list) <- names(dp.out.list)
```
run the pw comparisons
```{r}
contrasts.new.txt <-  fread('./data/contrasts_new.txt', sep='-', header=F)

contrasts.list <- split(contrasts.new.txt, seq(nrow(contrasts.new.txt)))
names(contrasts.list) <- apply(contrasts.new.txt, 1, function(x){ paste(x[1], x[2], sep='-') }) 

ms.out.list <- lapply(seq_along(dp.proc.list), function(x,n,i){
  
  contrasts.mat <- MSstats::MSstatsContrastMatrix(contrasts.list, 
                               conditions = unique(x[[i]]$ProteinLevelData$GROUP),
                               labels = names(contrasts.list))
  
  print(contrasts.mat)
  # run msstats correcting for batch 
  mss <- groupComparison(contrast.matrix=contrasts.mat, data=x[[i]])
  mss.dt <- setDT(mss$ComparisonResult)

  #fwrite(mss.dt, ScriptAndDatedFileName(paste0(n[[i]], 'mss.pwcontrasts.unfiltered.csv')))
  
},x=dp.proc.list, n=names(dp.proc.list))


names(ms.out.list) <- names(dp.proc.list)
```
Write the output of the MSstats analysis

```{r}
mss.proc.list <- lapply(seq_along(ms.out.list), function(x,n,i){
  
  x[[i]][, gene := multiUniprots2multiGenes(as.character(Protein), species='MOUSE')]
  fwrite(x[[i]], ScriptAndDatedFileName(paste0(n[[i]], 'mss.pwcontrasts.unfiltered.csv')))
  
  # now 
  x.dt <- x[[i]][!is.infinite(abs(log2FC)) & !issue %in% c("oneConditionMissing","completeMissing"), ] %>% 
  .[, sig := 'not'] %>% 
  .[abs(log2FC) > 0.58 & adj.pvalue < 0.05, sig := ifelse(log2FC > 0, 'up', 'down')]
  
  fwrite(x.dt, ScriptAndDatedFileName(paste0(n[[i]], 'mss.pwcontrasts.processed.csv')))
  return(x.dt)

},x=ms.out.list, n=names(ms.out.list))

names(mss.proc.list) <-  names(ms.out.list)
```

```{r}
mss.comb <- rbindlist(mss.proc.list, idcol = T)

g <- ggplot(data=mss.comb[sig != 'not', .N, by=.(sig,.id,Label)], aes(x=sig, y=N, fill=sig)) +
  geom_bar(stat='Identity') +
  scale_fill_manual(values=c('up'='red', 'down'='blue')) +
  facet_grid(.id~Label) +
  theme_bw() +
  theme(
  strip.text = element_text(
    size = 5))

BackupAsPDF(g, 'Nsighits.contrastvsearch.barplot', dimensions = c(12,6))
```

plot summary of the sig hits per conditons

```{r}
g <- ggplot(data=mss.comb[sig != 'not', .N, by=.(sig,.id)], aes(x=sig, y=N, fill=sig)) +
  geom_bar(stat='Identity') +
  scale_fill_manual(values=c('up'='red', 'down'='blue')) +
  facet_grid(.~.id) +
  theme_bw() +
  theme(
  strip.text = element_text(
    size = 8))

BackupAsPDF(g, 'Nsighits.search.barplot', dimensions = c(6,4))
```
MOUSE ID mappings 
```{r}

new.mss <- merge.data.table(x=mss.comb, y=mh_mappings[,.(Symbol.mouse, Symbol.human, SWISS_PROT_IDs.mouse, SWISS_PROT_IDs.human)], by.x='Protein', by.y='SWISS_PROT_IDs.mouse', all.x=T)
mss <- unique(new.mss, by=c("Protein", "Label", ".id")) #drop dup records from 1:many mappings

nrow(mss.comb)
nrow(new.mss)
nrow(mss)
```

Volcanoplot vs vehicle 
```{r}
contrasts.oi <-  grep('-Vehicle', unique(mss.comb$Label), value=T)
contrasts.oi <- contrasts.oi[-1]


g <-  ggplot(mss[Symbol.human %in% isGenes & Label %in% contrasts.oi, ], aes(x=log2FC, y=-log10(adj.pvalue), col=sig, label=Symbol.mouse)) +
    geom_point() +
    geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
    geom_hline(yintercept = -log10(0.05), linetype=2, alpha=0.4) +
    scale_color_manual(values=c('up'='#990033', 'down'="#2A788EFF", 'not'='grey')) +
    ggrepel::geom_text_repel(size = 2, max.overlaps = 20, segment.color = 'grey80') +
    ggtitle(paste('ISG Fold Change Comparisons')) + 
    theme_bw() +
    facet_grid(.id~Label)

g
BackupAsPDF(g, 'isg.vsVeh.volcanoplots', dimensions=c(10,8))
```

Volcanoplots of the DEGs in each contrast, 1 row per search

```{r}

g <- ggplot(mss[Label %in% contrasts.oi,], aes(x=log2FC, y=-log10(adj.pvalue), col=sig, label=gene)) +
  geom_point() + 
  ylab('-log10 adjusted p-value') +
  ggrepel::geom_text_repel(data=mss[Label %in% contrasts.oi & sig != 'not',], show.legend = FALSE, size = 2, max.overlaps = 20) +
  geom_vline(xintercept = c(-0.58,0.58), linetype=2, alpha=0.4) +
  geom_hline(yintercept = -log10(0.05), linetype=2, alpha=0.4) +
  # add the points for the SARS-CoV2 proteins
  #geom_point(data=mss[Protein %in% virus.prots,], aes(x=log2FC, y=-log10(adj.pvalue)), col='#669966') +
  #ggrepel::geom_text_repel(data=mss[Protein %in% virus.prots,],  size = 2, max.overlaps = 20, segment.color = 'grey80', colour = "#669966") +
  scale_color_manual(values=c('up'='#990033', 'down'='#336699', 'not'='grey')) +
  facet_wrap(Label~.id, ncol=3) +
  theme_bw()
g
BackupAsPDF(g, 'combined.volcano', dimensions=c(12,18))


mss[Protein %like% "SARS",]

dp.out.list[[1]][Protein %like% "Q9Z0G7",]
spec.list[[1]][grep("P0DTC3", ProteinName),]


grep()

mss[gene == 'NaN',]


spec.dt.list[[1]][ProteinName == 'NaN',]
```
Venn Diagrams of overlaps in protein sig hits 
- compare this to the unfiltered list



```{r}
searches.venn <- list(dual_bg = mss[.id == 'dual_bg' & sig != 'not',  unique(gene)],
                      dda_bg = mss[.id == 'dda_bg' & sig != 'not',  unique(gene)],
                      dda_nobg = mss[.id == 'dda_noBg' & sig != 'not',  unique(gene)]
                      )

g <- ggvenn(searches.venn) +
  labs(title='overlap in sig hits (log2FC +/ 50% & adj.pval < 0.05)')
g

searches.venn <- list(dual_bg = ms.out.list[['dual_bg']][sig != 'not',  unique(gene)],
                      dda_bg = ms.out.list[['dda_bg']][sig != 'not',  unique(gene)],
                      dda_nobg = ms.out.list[['dda_noBg']][ sig != 'not',  unique(gene)]
                      )

g <- ggvenn(searches.venn) +
  labs(title='overlap in sig hits (log2FC +/ 50% & adj.pval < 0.05)')

```
Compare the N sig hits to the unfiltereed set, seems many of these will appear as condition missing

```{r}
mss.raw.comb <- rbindlist(ms.out.list, idcol = T)

g <- ggplot(mss.raw.comb[abs(log2FC) > 0.58 & adj.pvalue < 0.05 & !is.na(issue),.N, by=.(issue,.id)], aes(x=.id, y=N, fill=issue)) +
  geom_bar(stat='Identity') +
  labs(title='N proteins with missing vals in comparisons') +
  theme_bw()
g
BackupAsPDF(g, 'NcomparisonsWConditionsMissing.barplot')
```
Heatmaps of proteins with conditions missing 

```{r}
lapply(seq_along(p.quant.list), function(x,y,n,i){
  
  p.mat <-  dcast(x[[i]], Protein~interaction(GROUP,SUBJECT), value.var = 'LogIntensities') %>% 
    as.matrix(rownames='Protein')
  
  p.mat <-  p.mat[rownames(p.mat) %in% y[.id == n[[i]] & !is.na(issue), unique(Protein)],  ]
  
  rownames(p.mat) <- multiUniprots2multiGenes(rownames(p.mat), species = 'MOUSE')

  
  hm <- Heatmap(p.mat, 
                cluster_rows = clusterWNA(p.mat), 
                cluster_columns=F, 
                col=viridis(100),
                column_title_gp = gpar(fontsize=6),
                column_split = list(gsub("[.][1-5]$", "", colnames(p.mat)), 
                                                                       str_extract(colnames(p.mat), "[47]D")
                                                                       ))
  BackupAsPDF(draw(hm, column_title=paste0(n[[i]], ' oneConditionMissingProteins')), paste0(n[[i]], 'missingProts.heatmap'), dimensions=c(12,8))
  
  
},x=p.quant.list, y=mss.raw.comb, n=names(p.quant.list))

```
Lets look at the enrichment for each of the sets

just run enrichment of the 3;

```{r}

contrasts.oi <- grep("-Mock", unique(mss$Label), invert = T, value=T)

# load the GO table
gmt.go <- loadGmtFromBioconductor(dbName = 'org.Mm.eg.db', ontology = "ALL", keyType = "UNIPROT")

## loop through the 
lapply(seq_along(mss.ou.list), function(x,n,i){
  
  universe <- x[[i]][, unique(Protein)]
  
  x[[i]][,enrich.grp := interaction(Label,sig)]
  
  enrich.dt <- enricherOnGroups(x[[i]][Label %in% contrasts.oi & sig != 'not'], 
                              groupColumns = 'enrich.grp', 
                              geneColumn = "Protein", 
                              term2gene.gmt = gmt.go, 
                              universe = universe)

  fwrite(enrich.dt, ScriptAndDatedFileName(paste0(n[[i]],'.GOenrichments.')))

  simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich.grp')
  fwrite(simp.enrich$simplified, ScriptAndDatedFileName(paste0(n[[i]],'GOenrichments.simplified.')))

  ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, groupColumn = 'enrich.grp', topN = 8, title=paste0(n[[i]],' GO term enrichment'), 
                                  negCols=unique(simp.enrich$simplified$enrich.grp[grep('down', simp.enrich$simplified$enrich.grp)]), 
                                  row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 6), upperThreshold = 8)
  ht
  BackupAsPDF(ht, paste0(n[[i]],'go.heatmap.top8.'), dimensions=c(10,8))
}, x=mss.proc.list, n=names(mss.proc.list))
```
These enrichments dont look great... what happened to all the sig hits from previous work?
Why are the enrichments so different and where are the sig hits?

```{r}
mss.raw.comb[, sig := 'not']
mss.raw.comb[abs(log2FC) > 0.58 & adj.pvalue < 0.05,  sig := ifelse(log2FC > 0, 'up', 'down')]
# not a huge amt of sig hits here...
mss.raw.comb[sig != 'not',.N, by=.(Label,sig,.id)][grep("Mock", Label, invert=T)][grep("4206_4D-", Label)][order(Label)]

# look at the processed... far less hits when we remove missing signals in these groups...
mss.comb[sig != 'not',.N, by=.(Label,sig,.id)][grep("Mock", Label, invert=T)][grep("4206_4D-", Label)][order(Label)]
```
What about our sig hits previously in this comparison? Read in the data, look at a heatmap of the sig hits in the first search
Also, compare the run condition matching with what we have now...

First look at the sig hit profiles from the old data

```{r}
# p.quant 
old.p.quant <- fread("../020524_AVIDD_ABproteomics/output/AB.ProteinLevelData.csv")
old.p.quant[, REP := SUBJECT]
old.p.quant[, SUBJECT := interaction(GROUP,SUBJECT)]
# for plotting group averages
old.p.quant[, mean.LogIntensities := mean(LogIntensities, na.rm=T), by=.(Protein,GROUP)]

#msstats output
old.mss <- fread('../020524_AVIDD_ABproteomics/output/mss.contrasts.ab.csv')

#clean the res
old.mss <- old.mss[!is.infinite(abs(log2FC)) & !issue %in% c("oneConditionMissing","completeMissing"), ] %>% 
  .[, gene := multiUniprots2multiGenes(Protein, species='MOUSE')] %>% 
  .[, sig := 'not'] %>% 
  .[abs(log2FC) > 0.58 & adj.pvalue < 0.05, sig := ifelse(log2FC > 0, 'up', 'down')]
```

```{r}
old.hits <- old.mss[Label == 'x4206_4D-Vehicle_4D' & sig != 'not',  unique(Protein)]


# look at our  new data almost no significant hits detected? Whats going on lost the vast majority of sig hits...
mss.raw.comb[Label == 'x4206_4D-Vehicle_4D' & Protein %in% old.hits & sig != 'not', .N, by=.(sig, .id)] 

# fold changes still quite large, but our pvalues much lower.. why is this?
g <- ggplot(mss.raw.comb[Label == 'x4206_4D-Vehicle_4D', ],
       aes(x=log2FC, y=-log10(adj.pvalue), color=sig)) +
   geom_point() +
  scale_color_manual(values=c('up'='red','down' = 'blue','not'='grey')) +
  geom_point(data=mss.raw.comb[Label == 'x4206_4D-Vehicle_4D' & Protein %in% old.hits, ],  aes(x=log2FC, y=-log10(adj.pvalue), color='green')) +
  facet_wrap(~.id)

BackupAsPDF(g, 'x4206-vsVehD4.comparisontoOlddata')
```
Need to figure out why these are so different... 
- compare output of the ms models. 
- compare number of cols in the p.quant data (were there samples missing in the first run?)
- compare the contrast matrices
```{r}
# 51 rows
old.p.quant[,.N, by=(SUBJECT)][grep('x3769|x4052', SUBJECT, invert=T)]


# lets look at the other data 
lapply(p.quant.list, function(x){
  
  x[, .N, by=.(GROUP,SUBJECT)]
})
```
Look at scatterplots of the p.quant data conditions; make sure these dont vary. IF they are v similar it must be down to the modelling...
combine the four matrices and check
```{r}
old.p.quant[, .id := 'oldSearch']
old.p.quant[, SUBJECT := REP]
p.quant.comb <- rbind(rbindlist(p.quant.list, idcol = T), old.p.quant, fill=T)

# look at group mean scatterplots
wide.dt <- dcast(p.quant.comb, Protein~interaction(GROUP,.id), value.var='LogIntensities', fun.aggregate = mean, na.rm=T)

# create scatterplots of x4206 mean vs other group
g <- ggplot(wide.dt, aes(x=x4206_4D.dual_bg, y=x4206_4D.oldSearch)) + geom_abline(intercept = 0, slope=1) + geom_point() + geom_density_2d()
g
g <- ggplot(wide.dt, aes(x=x4206_4D.dda_noBg, y=x4206_4D.oldSearch)) + geom_abline(intercept = 0, slope=1) + geom_point() + geom_density_2d()
g
g <- ggplot(wide.dt, aes(x=x4206_4D.dda_bg, y=x4206_4D.oldSearch)) + geom_abline(intercept = 0, slope=1) + geom_point() + geom_density_2d()
g
```

compare the raw spectronaut files

```{r}
# from Box folder 060224_AVIDD_AB_DIA/20240206_115609_2024.02.06 - Mac1 Inhibitor Mouse Lung AB/2024.02.06 - Mac1 Inhibitor Mouse Lung AB_Report_MSStats Format (Normal).tsv
old.spec <- fread('../020524_AVIDD_ABproteomics/data/AB_Report_MSstats_Format.tsv')
old.spec[, .id := 'oldSearch']


old.spec[, gene := multiUniprots2multiGenes(ProteinName, species='MOUSE')]

spec.list <-  list(dda_bg = fread('./data/DDA Lib.tsv'),
                   dda_noBg = fread('./data/DDA Lib No Background.tsv'),
                   dual_bg = fread('./data/Dual Lib .tsv'),
                   dual_noBg = fread('./data/240221.dual.lib.nobg.tsv'))

# get shared raw files
shared.runs <- unique(spec.list[[1]]$Run)

lapply(spec.list, function(x){nrow(x)})
nrow(old.spec[Run %in% shared.runs,])

# 32
shared.runs %>% length()

# still 32
old.spec[Run %in% shared.runs, .N, by=.(Run)]
```

Compare the counts in each of the datasets

```{r}
spec.comb <- rbind(rbindlist(spec.list, idcol=T), old.spec, fill=T)
spec.comb[, gene := multiUniprots2multiGenes(ProteinName, species='MOUSE')] 
spec.comb <- spec.comb[Intensity > 2^5,]

spec.wide <- dcast(spec.comb, interaction(PeptideSequence,PrecursorCharge)~interaction(.id,Run), value.var = "Intensity") %>% 
  as.matrix(rownames=1)

# just get complete prots for count comparisons
spec.wide <-  spec.wide[complete.cases(spec.wide),]

# comvert to dt
spec.wide <- data.table(spec.wide, keep.rownames = T)

ggplot(spec.wide, aes(x=log2(dual_bg.ex014876.raw), y=log2(oldSearch.ex014876.raw))) +
  geom_point() +
  geom_density2d()


spec.wide %>%  colnames()
```




Todo
----

how do we compare enrichments. Use all prots detected across runs as bg
What about using only shared features instead?
```{r}

```


enrichment of the set of dual library vs dda only grp

What about the additonal dual lib hits? look at enrichment of proteins not in dual lib and enrich
Also maybe scatterplot of log2FC values of any sig proteins for each of the groups?


```{r}

```



Todo
---
Filter out proteins with low N peptides
Heatmaps?
Ask about the viral proteins.. why are they not recovered in the search?

‚