---
title: "100424.FormatPCPDatabase"
author: "Martin Gordon"
date: "2024-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Steps to produce clustering
i) Took pearson correlation matrix of CRISPR knockouts
ii) Created a distance matrix
iii) Performed hierarchical clustering with complete linkage (largest distanc/dissimilarity between cluster1 & cluster2 is the distance between the clusters) 
  (goal here is to find small, tight clusters as we are looking at complexes
iv) Then we step through the dendogram, find any merges within a certain distance threshold (0.4 quantile) and extract all members of this cluster. This finds redundant clusters (for example, if  A-B-C are a cluster that meet the threshold, A-B, A-C and C-A will also be recovered)
v) as a control randomized labels on the clusters and found ~50% less clusters, so some distinguishing but poor FP recovery. (rerun with more stringent threshold?)

v) Because these are transcriptome responses to CRISPR editing, in most cases pearson similiarity was quite low. 

Format a database in perturbSeq structure so we can use it with PCProphet to detect clusters

```{r library}
library(data.table)
library(magrittr)
library(ggplot2)
library(ComplexHeatmap)
library(viridis)
library(dendextend) # working with dendrogram

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
source("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R")
```
Read in the CORUMDB and compare with the cluster set
```{r}
corum.dt <- fread('./data/humanComplexes.txt')
clusters.dt <-  fread('./data/KOLF2-SrongPerturbation-Clusters.csv')

# 50 clusters in total, ranging in size from 64 to 9 (ranked by size)
clusters.dt[,.N, by=cluster]
```

```{r}
corum.dt %>%  head()
corum.dt %>%  colnames()


  getwd()
```
Create a minimal corum DB format; just complex Name, ID, Gene names, uniproit and entrez
```{r}
clusters.dt[,ComplexID := cluster+1]
clusters.dt[, ComplexName := paste('Cluster', ComplexID, sep='_')]


# get uniprot and geneIDs
clusters.dt[, Entrez := translateGeneName2Entrez(gene_target, species='HUMAN')]
clusters.dt[, Uniprot := translateGeneName2Uniprot(gene_target, species='HUMAN')]


clusters.corum.dt <- clusters.dt[, .(`subunits(Gene name)`=paste(gene_target, collapse=';'), "subunits(UniProt IDs)"=paste(Uniprot, collapse = ';'), "subunits(Entrez IDs)"= paste(Entrez, collapse = ';'), 'Organism'='Human'), by=.(ComplexID, ComplexName)]


clusters.dt[,.N, by=cluster]$N
clusters.corum.dt[, .SD[,length(unlist(strsplit(`subunits(Gene name)`, ';')))]]

# looks good
clusters.corum.dt[, .SD[,length(unlist(strsplit(`subunits(Gene name)`, ';')))], by=ComplexID]$V1 == clusters.dt[,.N, by=cluster]$N


# check colnames 
colnames(clusters.corum.dt) %in% colnames(corum.dt)


head(corum.dt)
head(clusters.corum.dt)


# write out and test if works
fwrite(clusters.corum.dt[,.(ComplexID, ComplexName, Organism, `subunits(Gene name)`, `subunits(UniProt IDs)`, `subunits(Entrez IDs)`)], sep='\t', ScriptAndDatedFileName('peturbSeq.clusters.db.txt'))

fread('~/Documents/projects/100324_AForget_PCProphetPipeline/100724.makePCPDB_data/2024_10_07_peturbSeq.clusters.db.csv')
```

Inspect the cluster matrix that was produced
This is a similarity matrix; convert to a distance matrix
```{r}
cluster.mat <- fread('./data/KOLF2-SrongPerturbation-PearsonCorr.csv')  %>% 
  as.matrix(rownames='V1')

#-1 to 1; dont rescale yet; just leave as is
# most distant possible score of 2, most similar is 1
range(cluster.mat)

# convert to. a distance matrix and perform hierarchical clusterin
dist.mat <- 1 - cluster.mat 

# perform hierarichcial clustering of the distance matr
# coerce to a distance matrix (no calculatons performed)

# I think we want complete linkage as we want all the things in hte same cluster to be very similar as considering as part of same complex
hc <- hclust(as.dist(dist.mat), method='complete')

quantile(hc$height, probs=c(.05,.10,.15,.25))

# ok error is my threshold to pull out the complexes I based on the pw distances rather than the 
hist(hc$height)
abline(v=0.65)

Heatmap(dist.mat, 
        cluster_rows = hc,
        name='1 - Pearsons R',
        col = viridis(3),
        cluster_columns = hc, 
        show_row_names = F,
        show_column_names = F)
```
Create function to detect subclusters in the hierarchical clustering 

```{r}
# give a node id; pull out the genes that are members of this cluster
extractLabelsfromClusters <- function(hc, nodeID){
  
  # if its a negative value, just return the gene
  if(sign(nodeID) == -1){
    
    # this ID is negative, return gene name
    return(hc$labels[[-nodeID]])
    
  } else {
    # if clusterID is positive, it means it has been previously merged
    # the merge matrix points to the row where this took place 
    node1 <- hc$merge[nodeID, 1]
    node2 <- hc$merge[nodeID, 2]
    
    # recursively search for the nodes
    leftSetNodes <- getGenesfromClusters(hc, node1)
    rightSetNodes <- getGenesfromClusters(hc, node2)
    
    return(c(leftSetNodes, rightSetNodes))
  }
}


# now extract sets that meet these thresholds

extractClustersfromDendogram <- function(hc, threshold){
  
  message(paste0('found ',length(unique(hc$labels)), ' leaf nodes. Extracting clusters..'))

  if (missing(threshold)){
    message(paste0('No threshold specified. Extracting all clusters (maximum distance =', threshold, ')...'))
    threshold <- max(hc$height)
  }
  message(paste0('Iterating through the dendogram and pruning groups with distance < ', threshold, '...'))
  
  # iterate through the dendogram
  clusters.dt <- lapply(1:nrow(hc$merge), function(i){
    # if the distance is below the threshold, extract all cluster members
    if (hc$height[i] < threshold){
    
      dt <- data.table(index = i,
                       members =  paste0(extractLabelsfromClusters(hc, i), collapse=';'))
    
    }
  }) %>% rbindlist()
  
  message(paste0('Found ', nrow(clusters.dt), ' clusters with a maximum distance < ', threshold))
  
  return(clusters.dt)
}
# just pulling out sets that are very similar
clus.test <- extractClustersfromDendogram(hc=hc)

clus.test

cluster.mat[rownames(cluster.mat) %in% c("LRPPRC","POLRMT","MTPAP","PNPT1"), colnames(cluster.mat) %in% c("LRPPRC","POLRMT","MTPAP","PNPT1")] %>%  View()
```


100924-Regen the clusters used for PCProphet
Just randomize the genes in the clusters
```{r}
cluster.db <- fread('./100824.PCP.perturbSeq_data/2024_10_08_newPerturbSseq.db.txt')


# identify all the genes used in the clustering
sapply(strsplit(cluster.db$`subunits(Gene name)`, ';'), function(x) unlist(x)) %>% 
  unlist()

# take just the genes from these extracted clusters
genes.oi <- unique(unlist(strsplit(cluster.db$`subunits(Gene name)`, ';')))




# first, count the number of subunits on each row
cluster.db[, Nmembers := as.numeric(length(unlist(strsplit(`subunits(Gene name)`, ';')))), by=ComplexID]
cluster.db
# replace F as dont want two of the same prots in one cluster

# question: why does top one work but bottom fail
cluster.db[, newSubunits := paste(sample(x=genes.oi, size=Nmembers), collapse=';'), by=ComplexID]
#cluster.db[, newSubunits := paste(sample(x=genes.oi, size=Nmembers), collapse=';'),]

# new complex Name
cluster.db[, newName := paste0('randomized_oriCluster__', ComplexName)]

fwrite(cluster.db[, .(ComplexID, ComplexName = newName, Organism,"subunits(Gene name)"=newSubunits)], sep='\t', ScriptAndDatedFileName('perturbSeq.clusters.randomized.csv'))

fread('~/Documents/projects/100324_AForget_PCProphetPipeline/100724.makePCPDB_data/2024_10_09_perturbSeq.clusters.randomized.csv')
```




```{r}
# Create a function to check which clusters are below the distance threshold
find_subclusters <- function(hc, threshold) {
  # Get the merge matrix from hclust object
  merge <- hc$merge
  heights <- hc$height
  
  # Initialize a list to store clusters
  cluster_list <- list()
  
  # Iterate through the merge matrix to find subclusters
  for (i in 1:nrow(merge)) {
    if (heights[i] <= threshold) {
      left <- merge[i, 1]
      right <- merge[i, 2]
      
      # If the left or right are negative, they represent individual points
      cluster <- c()
      if (left < 0) cluster <- c(cluster, -left)
      if (right < 0) cluster <- c(cluster, -right)
      
      # If they are positive, they represent previously merged clusters
      if (left > 0) cluster <- c(cluster, cluster_list[[left]])
      if (right > 0) cluster <- c(cluster, cluster_list[[right]])
      
      cluster_list[[i]] <- cluster
    }
  }
  
  return(cluster_list)
}




# maybe the first funciton should be to perepare the full list of clustered genes

extractClustersfromDendogram <-  function(hc_cluster= hc, threshold=0.5){
  
  
  message('Extracting clusters from dendogram')
  message(paste0(length(unique(hc$labels)), ' leaf nodes detected'))
  # first extract all clusters below a certain height
  message(paste0('Finding clusters with maximum distance < ', threshold, '...'))
  
  # iteerate through the merge matrix rows
  lapply(1:nrow(hc_cluster$merge), function(i){
    
    
    # threshold is our distance threshold
    if (hc_cluster$height[i] <= threshold){
      
      dt <- data.table(cluster = i, 
                       nodes = '')
      
      # get the merge nodes
      node1 <- merge.mat[i, 1]
      node2 <- merge.mat[i, 2]
      
      # hclust clustering identifies individual nodes with '-' prefix, if positive, it represents an earlier cluster
      ifelse(sign(node1) == -1 & sign(node2) == -1, dt[, nodes := paste(node1, node2, collapse=';')],
             ifelse(sign(node1) == -1 & sign(node2) != -1, dt[, nodes := paste(node1, node2, collapse=';')]
             ){
        
        cluster
      }) %>% rbindlist()
      
    }
    
    
  }) 
  
  message(paste0('Found N clusters with maximum distance < ', threshold, '...'))
  
}

extractClustersfromDendogram(hc_cluster = hc, threshold = 0.5)

# give a node id; pull out the genes that are members of this cluster
extractLabelsfromClusters <- function(hc, nodeID){
  
  # if its a negative value, just return the gene
  if(sign(nodeID) == -1){
    
    # this ID is negative, return gene name
    return(hc$labels[[-nodeID]])
    
  } else {
    # if clusterID is positive, it means it has been previously merged
    # the merge matrix points to the row where this took place 
    node1 <- hc$merge[nodeID, 1]
    node2 <- hc$merge[nodeID, 2]
    
    # recursively search for the nodes
    leftSetNodes <- getGenesfromClusters(hc, node1)
    rightSetNodes <- getGenesfromClusters(hc, node2)
    
    return(c(leftSetNodes, rightSetNodes))
  }
}

extractLabelsfromClusters(hc, 26)


# correspond to 26
test <- c(-505,  -631, -556,  -873,-911, -520, -819,  -848, -849, -887, -1054)

# results look good
sapply(test, function(x)getGenesfromClusters(hc, x)) %in% getGenesfromClusters(hc, 26) 
```

```{r}
plot(hc)

dend  %>% get_nodes_attr('members')

library(dynamicTreeCut)

dist.mat %>% str()

clustersTest <- cutreeDynamic(hc, distM=dist.mat,deepSplit = 4)
clustersTest %>% table()
help(cutreeDynamic)
```

