---
title: "DARPA33"
author: "Martin Gordon"
date: "2023-07-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary
Perform Differential Expression Analysis of MS-MS Spectronaut data using MSstats
Data from DARPA project: Single timepoint with 6 conditions in total: 3 ligands (AT23, AT53, AT99) vs PGE, treated vs untreated with 30 min exposure to NPFF
Perform pairwise contrasts for each of the conditions

###: Contact: Qiongyu Li
###: Date: 10-07-23

```{r packages}
library(purrr)
library(tidyr)
library(stringr)
library(dplyr)
library(data.table)
library(MSstats)
library(ggplot2)
library(magrittr)
library(circlize)
library(ComplexHeatmap)
library(parallelDist) #needed for some custom enrichment functions
library(ggvenn)

#source Bens utility scripts
source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source ("/Users/martingordon/Documents/utils/bp_utils/MSstats_Helper_Functions.R")
source("/Users/martingordon/Documents/utils/bp_utils/MSstats_V4_Functions.R")
source ("/Users/martingordon/Documents/utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("/Users/martingordon/Documents/utils/bp_utils/enrichmentTestFunctions.R")
```

Read in spectronaut data & metadata
----
```{r read-in}

input <- fread('data/DAR33_PTGER4/20230705_212758_DAR33_dirDIA_EXc_MSStatsFormatReport.tsv')
meta <- fread('data/DAR33_PTGER4/DAR33_dirDIA_EXc_ConditionSetup.tsv')
```

```{r data-inspection}
input[,BioReplicate := as.factor(BioReplicate)]
input[,Run := as.factor(Run)]
input[,Condition := as.factor(Condition)]
```

Data inspection
---------

Bioreplicate numbers (1-3) are repeated across conditions. Modified as msstats assumes repeated bioreplicate are processed as timeseries data
Asigned each cond/biorep combo a unique ID (1-20) for msstats to model as a case-control experiment.

```{r}

input[, .N, by=.(Condition,BioReplicate)]
# create new ID col for pairwise comparisons
input[, ID:=.GRP, by=.(Condition,BioReplicate)]


#reassign to biorep col and drop ID
input[, BioReplicate:=ID]
input[, ID:=NULL]

#sanity check
input[, .N, by = .(Condition, BioReplicate)]
```
```{r}
ggplot(data=input,  mapping=aes(x=log2(Intensity))) +
  geom_histogram(binwidth = 0.1) + 
  facet_wrap(~Condition)

input <- input[log2(Intensity) > 5] #filter out low abundance 

ggplot(data=input,  mapping=aes(x=log2(Intensity))) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~Condition)


print(input)
```
```{r}
any(input[, .N, by = .(Run, PeptideSequence, PrecursorCharge)]$N) > 1
```
AT00_PGE00_8 low peptide counts compared to others

```{r}
### plot of peptide ions per run
p <- input %>%  
      ggplot(mapping=aes(x=reorder(interaction(Condition,BioReplicate)), fill=Condition)) +
      geom_bar() + 
      theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

BackupAsPDF(p, "BarPlot_NPeptides_")
```
Boxplots of distributions of raw intensity values per condition/replicate. Again abundances are pretty consistent and no real concerns with raw data quality. 

One of the controls does not look good
```{r create contrasts}
p <- ggplot(data = input, mapping=aes(x=reorder(interaction(Condition, BioReplicate)), y=log2(Intensity), fill=Condition)) +
      geom_boxplot() +
      theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

BackupAsPDF(p, "BoxPlot_RawIntensities_")

```
----
Convert spectronaut data to msstats format
```{r msstats-formatting}
ab.mss <- specFileToCompleteMSstats(input)
ab.mss[, IsotopeLabelType := 'L'] #label free
setDF(ab.mss)  # convert to DF for msstats processing

```
MSStats DataProcess: Normalise (equilizemedians), filter & summarize peptides to protein intensities

```{r mss-dataproc}

dp.out <- MSstats::dataProcess(ab.mss,
                               normalization = 'equalizeMedians',
                               summaryMethod="TMP",
                               featureSubset = "highQuality",
                               remove_uninformative_feature_outlier = TRUE,
                               MBimpute = FALSE )
```
Write out protein level data to file (provide feature level data also)
```{r writeout-data}
# the main data of interest is this one:
fwrite(dp.out$ProteinLevelData, ScriptAndDatedFileName("ProteinLevelData.csv.gz"))

#this table is mostly a copy of the input with some additional columns for hte post-normalization intensity (labeled ABUNDANCE), and columns for is_outlier and is_informative
fwrite(dp.out$FeatureLevelData, ScriptAndDatedFileName("FeatureLevelData.csv.gz"))
```
Clear issue with one of the AT00_PGE00 repllicates

Proceed as is but may need to drop

```{r}
prot.quant <- setDT(dp.out$ProteinLevelData)

p <- prot.quant %>% 
      ggplot(mapping=aes(x=reorder(interaction(GROUP,SUBJECT)), y=LogIntensities, fill=GROUP)) +
      geom_boxplot() +
      theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))
p
BackupAsPDF(p, "BoxPlot_NormalizedProteinIntensities_")
```

```{r}

  rowClusterWithNA <- function(mat, na.value = 0, ...){
    mat[is.na(mat)] <- na.value
    hclust(dist(mat), ...)
  }



intensity.mat <- as.matrix(dcast(prot.quant, Protein ~GROUP+SUBJECT, value.var = "LogIntensities"),
                           rownames = "Protein")
  
subMat <- intensity.mat[sample.int(nrow(intensity.mat), 2000),]


colnames(subMat)
unique(sub("_[0-9]{1,2}$", "", colnames(subMat)))
#create heatmap annotation col
type <-sub("_[0-9]{1,2}$", "", colnames(subMat))
ha <- HeatmapAnnotation(type = type, annotation_name_side = "right", col= list(type = c('AT00_PGE00'='red', 'AT00_PGE30' = 'blue', 'AT29_PGE00'='green', 'AT29_PGE30'='orange', 'AT53_PGE00'='cyan', 'AT53_PGE30'='purple', 'AT99_PGE00'='yellow',  "AT99_PGE30"='pink')))

  #generate heatmap
  hm <- Heatmap(subMat,
        #clustering_distance_rows = function(m) dist(m, method = 'euclidean'),
        #clustering_method_rows= "complete",
        top_annotation = ha,
        cluster_rows = rowClusterWithNA(subMat),
        show_row_names = FALSE,
        row_title = sprintf("%d subsampled proteins", nrow(subMat)),
        na_col = 'grey',
       # column_title = paste0(' heatmap'),
        name = "Log2 Int.")
  hm
  BackupAsPDF(hm,"heatmap_subsample_log2Intensities_")

```

Subtract median intensies per run and re-plot

```{r}
intensity.mat <- sweep(intensity.mat, 1, apply(intensity.mat, 1, median, na.rm=T)) 

intensity.mat
  
subMat <- intensity.mat[sample.int(nrow(intensity.mat), 2000),]

#generate heatmap
hm <- Heatmap(subMat,
        #clustering_distance_rows = function(m) dist(m, method = 'euclidean'),
        #clustering_method_rows= "complete",
        top_annotation = ha,
        cluster_rows = rowClusterWithNA(subMat),
        show_row_names = FALSE,
        row_title = sprintf("%d subsampled proteins", nrow(subMat)),
        na_col = 'grey',
       # column_title = paste0( heatmap'),
        name = "Log2 Int.")


BackupAsPDF(hm, "heatmap_subsample_log2Intensities_scaled_")

```
Maybe two samples to remove: AT00_PCG30_23 & AT00_PGE00_24

```{r}
complete.mat <- intensity.mat[complete.cases(intensity.mat),]

pcaOut <- prcomp(t(complete.mat))
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
#pcaDT[, mainGroup := tstrsplit(rn, "_")[[1]]] #transpose & split
pcaDT[, mainGroup := sub("_[0-9]{1,2}$", "", rn)]

pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)

p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color = mainGroup )) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA using %d proteins (log intensity)", nrow(complete.mat))) 
p

BackupAsPDF(p, "PCA_Complete_Proteins")
```
Drop AT00_PGE00_24

```{r}
out.samples <- c('AT00_PGE00_24')
complete.mat <- intensity.mat[complete.cases(intensity.mat), !colnames(intensity.mat) %in% out.samples ]

pcaOut <- prcomp(t(complete.mat))
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
#pcaDT[, mainGroup := tstrsplit(rn, "_")[[1]]] #transpose & split
pcaDT[, mainGroup := sub("_[0-9]{1,2}$", "", rn)]

pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)

p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color = mainGroup )) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA using %d proteins (log intensity)", nrow(complete.mat))) 
p

BackupAsPDF(p, "PCA_RmOutliers_Proteins")
```
Color by timepoint
Samples clustering by  

```{r}
pcaDT[, mainGroup := str_extract(rn, "PGE[0-9]{2}")]
pcaDT

pcaDT
p <- ggplot (pcaDT, aes(x=PC1, y=PC2, color = mainGroup )) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA using %d proteins (log intensity)", nrow(complete.mat))) 
p

BackupAsPDF(p, "PCA_RmOutliers_PGEcol_Proteins")
```

Differential Expression Analysis
------

Perform DEA with provided contrasts
Generate a contrast matrix for requested pairwise comparisons of all groups:

contrast is log2(A) - log2(B) == 1 -1 == 'A vs B'
```{r make contrasts}

dp.contrasts <- NULL

mat.dimnames <- list(rnames='AT00_PGE30 vs AT00_PGE00',
                     cnames=c("AT00_PGE00","AT00_PGE30","AT29_PGE00", "AT29_PGE30","AT53_PGE00","AT53_PGE30","AT99_PGE00","AT99_PGE30"))


#compare t30 to t0
dp.contrasts <- matrix(data= c(-1, 1,0,0,0,0,0,0), nrow = 1, dimnames=mat.dimnames)
dp.contrasts <- rbind(dp.contrasts, c(0, 0,-1,1,0,0,0,0))
rownames(dp.contrasts)[2] <- 'AT29_PGE30 vs AT29_PGE00'
dp.contrasts <- rbind(dp.contrasts, c(0, 0,0,0,-1,1,0,0))
rownames(dp.contrasts)[3] <- 'AT53_PGE30 vs AT53_PGE00'
dp.contrasts <- rbind(dp.contrasts, c(0, 0,0,0,0,0,-1,1))
rownames(dp.contrasts)[4] <- 'AT99_PGE30 vs AT99_PGE00'

# compare each of the ligands vs control at t30
dp.contrasts <- rbind(dp.contrasts, c(0,-1,0,1,0,0,0,0))
rownames(dp.contrasts)[5] <- 'AT29_PGE30 vs AT00_PGE30'
dp.contrasts <- rbind(dp.contrasts, c(0,-1,0,0,0,1,0,0))
rownames(dp.contrasts)[6] <- 'AT53_PGE30 vs AT00_PGE30'
dp.contrasts <- rbind(dp.contrasts, c(0,-1,0,0,0,0,0,1))
rownames(dp.contrasts)[7] <- 'AT99_PGE30 vs AT00_PGE30'

# compare ligands at t30
dp.contrasts <- rbind(dp.contrasts, c(0,0,0,1,0,-1,0,0))
rownames(dp.contrasts)[8] <- 'AT29_PGE30 vs AT53_PGE30'
dp.contrasts <- rbind(dp.contrasts, c(0,0,0,1,0,0,0,-1))
rownames(dp.contrasts)[9] <- 'AT29_PGE30 vs AT99_PGE30'
dp.contrasts <- rbind(dp.contrasts, c(0,0,0,0,0,1,0,-1))
rownames(dp.contrasts)[10] <- 'AT53_PGE30 vs AT99_PGE30'

#compare ligands to control @ t0
dp.contrasts <- rbind(dp.contrasts, c(-1,0,1,0,0,0,0,0))
rownames(dp.contrasts)[11] <- 'AT29_PGE00 vs AT00_PGE00'
dp.contrasts <- rbind(dp.contrasts, c(-1,0,0,0,1,0,0,0))
rownames(dp.contrasts)[12] <- 'AT53_PGE00 vs AT00_PGE00'
dp.contrasts <- rbind(dp.contrasts, c(-1,0,0,0,0,0,1,0))
rownames(dp.contrasts)[13] <- 'AT99_PGE00 vs AT00_PGE00'

```
Run the pairwise comparisons
```{r mss-pairwise}
#rm the oulier
dp.out$ProteinLevelData <- prot.quant[paste0(GROUP,'_',SUBJECT) != out.samples,]

dp.out$ProteinLevelData[,.N, by=.(GROUP,SUBJECT)] %>% 
  .[order(SUBJECT,GROUP)]


test.mss <- groupComparison(contrast.matrix=dp.contrasts, data=dp.out)
test.mss.all <- setDT(test.mss$ComparisonResult)

# adjust p values using the Benjamini Hochberg method, which estimates False Discovery Rate
test.mss.all[, fdr := p.adjust(pvalue, method = "BH")] 
#get gene names
test.mss.all[, gene := multiUniprots2multiGenes(as.character(Protein), simplify = TRUE, allowDups = FALSE)]


fwrite(test.mss.all, ScriptAndDatedFileName("mss.groupcomparisons.csv.gz"))
```

Generate Volcano plots for each of the pairwise comparisons
Again 
```{r multi-volcanoplot}
## volcano plots

# define significant proteins
test.mss.all[, sig := "Not"]
test.mss.all[fdr < 0.05 & abs(log2FC) > log2(1.5),sig := ifelse (log2FC  > 0, "Up", "Down")]

p <- ggplot (test.mss.all, aes (x = log2FC, y = -log10(fdr), color = sig, label = gene)) +
  geom_point() +
  scale_color_manual(values = c(Not = "gray", Down = "blue", Up = "red")) +
  ggrepel::geom_text_repel(data = test.mss.all[sig != "Not"]  ,  size = 2, max.overlaps = 20) +
  theme_bw() +
  facet_wrap(~ Label,)
p <- p + theme(strip.text.x.top = element_text(size=6))
BackupAsPDF(p, ".volcanoplot.multipanel.")

```


Over plotting. Generate each volcano plot separately.
```{r split-dt}
#split datatable by comparison

test.mss.list <- split(test.mss.all, test.mss.all$Label)
```

```{r cal-geommean}
#calculate geomemtric mean
test.mss.list <- test.mss.list %>% 
  map( ~{
    dt <- .x
    dt$gm.pval.log2fc <- sqrt(dt[,'fdr']*abs(dt[,'log2FC']))
    
    return(dt)
  })

```

Again looks like the NPFF addition has the largest effect
```{r volcanoplots}
# use purr to plot each list element 


  imap(test.mss.list, ~ {
    dt <- .x
    dt.name <- .y
    
    p <- ggplot (dt, aes (x = log2FC, y = -log2(fdr), color = sig, label = gene)) +
      geom_point() +
      scale_color_manual(values = c(Not = "gray", Down = "blue", Up = "red")) +
      ggrepel::geom_text_repel(data = dt[sig != "Not"] ,  size = 2, max.overlaps = 20) +
      ggtitle(paste0(dt.name, ' Volcano plot')) +
      theme_minimal()
    p
   BackupAsPDF(p, paste0(dt.name,".volcanoplot.adjpval<0.05&log2FC>.58"))
  })
```


Enrichment Heatmaps for the different comparisons

Perform enrichment analysis for each of the comparisons
```{r}
#creat new factor col of combined label sig info
test.mss.all[, enrich_grp:= interaction(Label,sig)]

# all go ontology genes (biological process, cellular componetn & metabolic function) for enrichment analysis
gmt.go <- loadGmtFromBioconductor(ontology = "ALL", keyType = "SYMBOL")

# all identified genes as the background set
universe <- unique(test.mss.all$gene)

# perform enrichment on significant genes for all groups
enrich.dt <- enricherOnGroups(test.mss.all[test.mss.all$sig != 'Not'], groupColumns = 'enrich_grp', term2gene.gmt = gmt.go, universe = universe)

#simplify enrich results to remove redundant GO terms
simp.enrich <- simplifyEnrichBySimilarUniverseMembership(enrichResultsTable = enrich.dt, gmt=gmt.go, groupColumn = 'enrich_grp')


# failed.. troubleshoot
#enrichmentOnGroupsPL(test.mss.all[test.mss.all$sig != 'Not'], geneColumn = 'gene', groupColumns = 'enrich_grp', gmt = gmt.go, universe = universe)

ht <- enrichHeatmapBestPerGroup(simplifiedEnrichTable = simp.enrich$simplified, NULL, groupColumn = 'enrich_grp', topN = 8, negCols =unique(simp.enrich$simplified$enrich_grp[grep('Down', simp.enrich$simplified$enrich_grp)]), row_names_gp = gpar(fontsize = 7), column_names_gp= gpar(fontsize = 6), upperThreshold = 8) 
ht# compare enrichment in the top 10 terms per group
BackupAsPDF(ht, "heatmap_GO")
```

Interestingly, it seems based on the enrichment plot a number of genes are enriched in the AT29_30 vs AT29_00 comparison relative to the other antagonists.

Find the genes enriched and plot these for the other T30_vsT00 groups
Barplots of interesting genes

```{r}
#look at SCAR complex, actin cytoskeletion organisation, protien domain specific binding & cell adhesion molecule binding
go.OI <-  c('actin cytoskeleton organization', 'SCAR complex', 'protein domain specific binding', 'cell adhesion molecule binding')


genes.OI <- simp.enrich$simplified[enrich_grp == 'AT29_PGE30 vs AT29_PGE00.Down' & ID %in% go.OI,  str_split(geneID, '/')] %>% 
  .[,unique(c(V1,V2,V3,V4))]

contrasts.IO <- c('AT00_PGE30 vs AT00_PGE00', 'AT29_PGE30 vs AT29_PGE00','AT53_PGE30 vs AT53_PGE00','AT99_PGE30 vs AT99_PGE00')

p <- ggplot(test.mss.all[gene %in% genes.OI & Label %in% contrasts.IO], aes(x=Label, y=abs(log2FC), fill=Label, alpha=adj.pvalue < '0.05')) +
         geom_col() +
         geom_errorbar( aes(x=Label, ymin=abs(log2FC)-SE, ymax=abs(log2FC)+SE), width=0.2, colour="black", alpha=0.5, size=0.5) +
         geom_hline(yintercept = 0.58, linetype='dotted' ) +
         theme_classic() +
          scale_alpha_discrete(range = c(0.1, 0.9)) +
         coord_flip() +
         theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 6 ), 
               axis.text.y = element_text(size=4),
               legend.title = element_text(size = 3), 
               legend.text = element_text(size = 3)) +
         facet_wrap(~gene, nrow = 3)
p
BackupAsPDF(p, 'log2FC_barplot_GOsig_T0-T30contrasts')


test.mss.all
```


Interesingly we can see there are a number of proteins significantly differentially expressed in AT29 t30 vs other antagonists
Regenerate the plot for the other contrasts AT29 vs other groups

THinking here is that the bioreps for non controls are a little noisy

```{r}
contrasts.IO <- c( "AT29_NP30 vs AT99_NP30","AT53_NP30 vs AT99_NP30" , 'AT29_NP30 vs AT53_NP30', "AT29_NP30 vs AT00_NP30","AT53_NP30 vs AT00_NP30","AT99_NP30 vs AT00_NP30")


test.mss.all[Label %in% contrasts.IO & gene == 'ZNF121',]

p <- ggplot(test.mss.all[gene %in% genes.OI & Label %in% contrasts.IO], aes(x=Label, y=abs(log2FC), fill=Label, alpha=adj.pvalue < '0.05')) +
         geom_col() +
         geom_errorbar( aes(x=Label, ymin=abs(log2FC)-SE, ymax=abs(log2FC)+SE), width=0.2, colour="black", alpha=0.5, size=0.5) +
         geom_hline(yintercept = 0.58, linetype='dotted' ) +
         theme_classic() +
          scale_alpha_discrete(range = c(0.1, 0.9)) +
         coord_flip() +
         theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 6 ), 
               axis.text.y = element_text(size=4),
               legend.title = element_text(size = 3), 
               legend.text = element_text(size = 3)) +
         facet_wrap(~gene, nrow = 3)
p
BackupAsPDF(p, 'log2FC_barplot_GOsig_ATcontrasts')

unique(test.mss.all$Label)

```
dotplot of genes of interest for the comparisons
```{r}
# check for others of interest
#genesOI <- c('BTRC', 'SKP1', 'ARRB2', 'FBXW11', 'STAM2', 'HGS')


prot.quant[, gene := multiUniprots2multiGenes(as.character(Protein) , species='HUMAN', simplify = T)]

p <- ggplot(prot.quant[gene %in% genes.OI,], aes(x=GROUP, y=LogIntensities, color=GROUP)) + 
  geom_point(size=1.5) +
  facet_wrap(~gene, scales = 'free') + 
  stat_summary(fun.data = "mean_cl_normal",
               geom = "errorbar",
               color = 'black',
               width = .4, 
               alpha=0.4) +
  stat_summary(fun = "mean", geom = "point", color='black', alpha=0.4) +
  #stat_summary(fun.y = mean, geom='point') +#  fun.args = list(mult = 1), geom = "pointrange", colour = "red")
  theme_bw() +
  theme(
    axis.text.x = element_text(angle=90, hjust=1, size= 6)
  )

p
BackupAsPDF(p, 'geneIO.log2intensities.dotplot.pdf', dimensions = c(10,8))
```

Plot each of these individually

```{r}
for (g in genesOI){
  
 p <- ggplot(prot.quant[gene %in% g,], aes(x=GROUP, y=LogIntensities, color=GROUP, alpha=0.2)) + 
  geom_point() +
  geom_hline(aes(yintercept = mean.Ints,col=GROUP))
 # facet_wrap(~gene, scales = 'free') + 
 # stat_summary(fun.data = 'mean_cl_normal', geom='errorbar') +
  ggtitle(paste0(g, ' Log2Intensity'))
  theme_classic() +
  theme(
    axis.text.x = element_text(angle=90, hjust=1, size= 6)
  )
 
 print(p)
}

```




