---
title: "DARPA33_contrastsofcontrast"
author: "Martin Gordon"
date: "2023-07-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Running contrast of contrast analysis

#Two way ANOVA: assess effect of two predictor vs one (continuous) outcome variable 
# A two-way ANOVA test reveals the results of two independent variables on a dependent variable. ANOVA test results can then be used in an F-test, a statistical test used to determine whether two populations with normal distributions share variances or a standard deviation, on the significance of the regression formula overall.

Issue with groups having the same sample size: should I include the discarded sample?

```{r}
library (data.table)
library (ComplexHeatmap)
library (ggplot2)
library (emmeans)
rotate.x.axis.text <- theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))

source("/Users/martingordon/Documents/utils/bp_utils/ManageScriptData.R")
source("/Users/martingordon/Documents/utils//bp_utils/LinearModels.R")
WriteSessionInfo()
WriteInstalledPackages()
```
Samples are pre-treated with antagonists (AT29, AT53 and AT99), and then treated +/- PGE00 PGE_30. Observed time is t0 and t30


The two way anova is agonist x time post-treatment

Main goal is to detect psilocin-dependent phosphorylation that is unique to the antagonist case. To answer that, these statistics look for significant differences in the time-post-treatment effect (TPT) with and without antagonist, and then we inspect the data to see if the differences are in the direction of interest.

# Fit linear models, one per "protein"
protein is actually a phopho-site-combination
```{r}
# including the dropped sample for now
tprotQuant <- fread ("/Users/martingordon/Documents/projects/071023_QLi_DARPA/DARPA_33_data/2023_07_10_ProteinLevelData.csv.gz")

str(tprotQuant)

str_extract("AT[0-9]{2}", protQuant$GROUP)
# parse GROUP names to define the actual columns that the linear models will work on
# this work depends on how your columsn are defined
tprotQuant[, ant :=  as.factor(str_extract(GROUP,"AT[0-9]{2}"))]
tprotQuant[, status := as.factor(ifelse(grepl("00", GROUP), "00", "30"))]
tprotQuant

# linearModelsAllProteins can this function which computes statistic for a contrast of two different log2FC
emmeans.contrastOfContrasts <- function (l, factorFormula = ~status|ant){  # how does the treatment effect change per antagonist level? basically compare treatment effect within each antagonist level
  emm <- emmeans::emmeans(l, factorFormula)
  contrast1 <- pairs(emm)
  contrast2 <- pairs(contrast1, by = NULL, adjust = "none")
  return (as.data.table(contrast2))
}

# postProcessFunction = emmeans.contrastOfContrasts tells it to use the function above after doing anova.
# it is optional, adds significant time to the computation (~5x), but is nice to
# put a magnitude and p value on specific differences of differences
#
# cl = 6 parameter says to use 6 processes.  Set to NULL if multiprocess is failing, or higher or lower as you wish.
# Numbers near the number of processors available in your machine is probably counterproductive.
lm.out <- linearModelsAllProteins(protQuant, 
                                  formulaList = list(interaction = LogIntensities~status + ant + ant:status ), # you could define other models here, but we name and define one model. 
                                  splitColumn = "Protein", postProcessFunction = emmeans.contrastOfContrasts, cl = 6) |> suppressWarnings()
```

# inspect anova p values per term
V weak evidence of antagonist effect: looks to be mainly treatment
```{r}

ggplot (lm.out$anova, aes(x = p.value)) + geom_histogram(boundary = c(0.0), 
                                                         # adjust binwidth to as narrow as is legible
                                                         binwidth = 0.005)+
  facet_wrap(~term)


```

## adjusted p values in anova
```{r}
lm.out$anova[, fdr := p.adjust(p.value, method = "BH"), by = term]
```


## T0 v T30 effects viewed as a heatmap
All look v similiar

```{r}
#play with thresholds here
sigProtein <- lm.out$anova[term %in% c( "status") & fdr < 0.05, Protein]


prot.mat <- dcast (protQuant[Protein %in% sigProtein], Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "Protein")
rownames(prot.mat) <- multiUniprotSites2multiGeneSites(rownames(prot.mat), species = "HUMAN")

prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))


Heatmap (prot.mat, cluster_columns = FALSE, 
         column_split = list (str_extract(colnames(prot.mat),"AT[0-9]{2}"),
                              str_extract(colnames(prot.mat),"PGE[0-9]{2}")))
```




## Antagonist effects


```{r}
#play with thresholds here
# no hits by fdr so reduce threshold
sigProtein <- lm.out$anova[term %in% c( "ant") & p.value < 0.001, Protein]
sigProtein

prot.mat <- dcast (protQuant[Protein %in% sigProtein], Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "Protein")
rownames(prot.mat)
rownames(prot.mat) <- multiUniprotSites2multiGeneSites(rownames(prot.mat), species = "HUMAN")
rownames(prot.mat)

prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))

Heatmap (prot.mat, cluster_columns = FALSE, 
         column_split = list (str_extract(colnames(prot.mat),"AT[0-9]{2}"),
                              str_extract(colnames(prot.mat),"PGE[0-9]{2}")))

```

## interaction effects 
(different antagonist post-treatment response)
Plotting logIntes

```{r, fig.width = 8, fig.height = 8}
#play with thresholds here
# there weren't many hits by fdr, so use a strong p.value
sigProtein <- lm.out$anova[term %in% c( "status:ant") & p.value < 0.001, Protein]


prot.mat <- dcast (protQuant[Protein %in% sigProtein], Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "Protein")
rownames(prot.mat) <- multiUniprotSites2multiGeneSites(rownames(prot.mat), species = "HUMAN")

prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))

Heatmap (prot.mat, cluster_columns = FALSE, 
         column_split = list ( column_split = list (str_extract(colnames(prot.mat),"AT[0-9]{2}"),
                              str_extract(colnames(prot.mat),"PGE[0-9]{2}"))))

```
## interaction effects observed in contrast of contrasts

Value here is we can require a magnitude of difference.  See below where I require abs(estimate) > 1.0.  Signs are confusing here because the loss of a negative effect will show as  a positive effect, and vice versa. 

```{r, fig.width = 8, fig.height = 8}

lm.out$postProcess
#play with thresholds here
# there weren't many hits by fdr, so use a strong p.value
sigProtein <- lm.out$postProcess[ p.value < 0.001 & abs(estimate) > 0.5, Protein]

prot.mat <- dcast (protQuant[Protein %in% sigProtein], Protein~GROUP+SUBJECT, value.var = "LogIntensities") |> as.matrix(rownames = "Protein")
rownames(prot.mat) <- multiUniprotSites2multiGeneSites(rownames(prot.mat), species = "HUMAN")

prot.mat <- sweep(prot.mat, 1, apply(prot.mat, 1, median, na.rm = TRUE))


Heatmap (prot.mat, cluster_columns = FALSE, 
         column_split = list (column_split = list (str_extract(colnames(prot.mat),"AT[0-9]{2}"),
                              str_extract(colnames(prot.mat),"PGE[0-9]{2}"))))
```

