---
title: "060224_AB_DIASpectronaut"
author: "Martin Gordon"
date: "2024-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup of spec Run for the new data

```{r}
library(data.table)
library (ComplexHeatmap)
library (circlize)
library(ggplot2)
library(stringr)
library(randomcoloR)
library(magrittr)

source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source("../../utils/bp_utils/ManageScriptData.R")
source("../../utils/bp_utils/MSstats_Helper_Functions.R")
source ("../../utils/bp_utils/MSstats_V4_Functions.R")
source ("../../utils/bp_utils/UniprotIDMapping.R") #uniprot to gene id mapping
source ("../../utils/bp_utils/enrichmentTestFunctions.R")

# function to cluster w NA values
clusterWNA <-  function(mat, na.value=0){
  
  mat[is.na(mat)] <- na.value
  return(hclust(dist(mat)))
}
```

```{r}
spec <- fread('./data/MSstats.txt')
p.quant <- fread('./data/results-mss-ProteinLevelData.txt')
keys <- fread('docs/keys.txt')
contrasts <- fread('docs/contrasts.txt')

spec.dt <- merge(x=spec[,!c("Condition", "BioReplicate")], y=keys[,.(Condition, BioReplicate, RawFile, Fraction,IsotopeLabelType)], by.x='Run', by.y='RawFile')
spec.dt

spec$BioReplicate %>%  unique()
keys$BioReplicate %>%  unique()
```

Each row in the table should be the intensity from one MS "feature" per run (an elution peak from a single peptide ion). It's possible that there are multiple features per peptide ion and run--if the peptide elutes in two different peaks perhaps.

```{r}
# no multiple feature peptide ions detected
spec.dt[,.N, by=.(PeptideSequence,PrecursorCharge,Run)][N >1]


hist(log2(spec.dt$Intensity))
spec.dt <-  spec.dt[Intensity > 2^5, ]
hist(log2(spec.dt$Intensity))
```

```{r}
col.pal <-  randomcoloR::distinctColorPalette(length(unique(spec.dt$Condition)))

g <- ggplot(spec.dt, aes(x=reorder(interaction(Condition,BioReplicate)), y=log2(Intensity), fill=Condition)) +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values=col.pal) +
  theme(axis.text.x = element_text(angle=90))

BackupAsPDF(g, 'raw.intensities.boxplot')
```

```{r}
g <- ggplot(spec.dt[,.N, by=.(Condition,BioReplicate)], aes(x=reorder(interaction(Condition,BioReplicate)), y=N, fill=Condition)) +
  geom_bar(stat='Identity') +
  ggtitle('N features per sample') +
  scale_fill_manual(values= col.pal) +
  theme(axis.text.x = element_text(angle=90))

BackupAsPDF(g, 'rawFeatureCounts.barplot')


spec.dt$Condition %>%  unique()
```

```{r}
wide <- dcast(spec.dt, ProteinName+PeptideSequence+PrecursorCharge~Condition+BioReplicate, value.var = "Intensity")

# good agreement between replicates it looks like 

# one of the mocks may be an issue but generally v simialr 
ggplot (wide, aes (x =log2(Vehicle_7D_1), y = log2(Vehicle_7D_2))) + geom_point(shape = ".") + geom_density_2d() + coord_fixed()
ggplot (wide, aes (x =log2(x4206_7D_2), y = log2(x4206_7D_4))) + geom_point(shape = ".") + geom_density_2d() + coord_fixed()
```

 PCA features 
 ----
 
```{r}
featureMat <- dcast (spec.dt,
                     paste0(PeptideSequence, PrecursorCharge)~interaction(Condition,BioReplicate), value.var = "Intensity") |> as.matrix(rownames = "PeptideSequence")

featureMat <- featureMat[complete.cases(featureMat), ]

```


```{r}
pcaOut <- prcomp(t(log2(featureMat)))

colInfo <- data.table(colname = colnames(featureMat))

colInfo
colInfo[,c("status","rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]

title <- "PCA"

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = status, shape = rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )
BackupAsPDF(p, "PCA_Complete_Features")
```

```{r}
featureMat <- dcast (spec.dt,
                     paste0(PeptideSequence, PrecursorCharge)~interaction(Condition,BioReplicate), value.var = "Intensity") |> as.matrix(rownames = "PeptideSequence")

#  log transform
featureMat <-  log2(featureMat)

dim(featureMat)
dim(featureMat[complete.cases(featureMat),])
colnames(featureMat)

# create annotation
ann <- data.table(replicate=str_extract(colnames(featureMat),'[1-5]$'),
                  treatment=sub("[_.].+", "", colnames(featureMat)),
                  timepoint=ifelse(grepl("[47]D", colnames(featureMat)), str_extract(colnames(featureMat), "[47]D"), 'na'))


colours <- list('timepoint' = c('na' = '#4477AA', '4D'="#228833", '7D'="#CCBB44"),
                'replicate' = c('1' = '#4477AA', '2'='#66CCEE', '3'="#228833", '4'="#CCBB44", '5'='#EE6677'),
                'treatment' =c('Mock' = col.pal[1], 'Vehicle'=col.pal[2], 'N1062D'=col.pal[3], 'x3769'=col.pal[4], 'x4052'=col.pal[5], 'x4206'=col.pal[6])
                )

colAnn <- HeatmapAnnotation(df = ann, col = colours)



# subset the data
submat <- featureMat[sample(rownames(featureMat), 2000), ]
  


head(log2(submat))
hm <- Heatmap(submat, cluster_rows=clusterWNA(submat), show_row_names = F, cluster_columns = T, top_annotation = colAnn, name='LogIntensities', column_names_gp = gpar(fontsize=6))
hm
#BackupAsPDF(hm, 'proteins.clustered.heatmap')

submat <- sweep(submat,1, STATS=apply(submat, 1, median, na.rm=T))
hm <- Heatmap(submat, cluster_rows=clusterWNA(submat), show_row_names = F, cluster_columns = T, top_annotation = colAnn, name='LogIntensities/Median',  column_names_gp = gpar(fontsize=6))
hm#BackupAsPDF(hm, 'proteins.clustered.medianscaled.heatmap')

```
## Summarisation & Normalisation
Running MSstats processing

# protein summarization

```{r}
mssInput <- specFileToCompleteMSstats(spec.dt)
mssInput[, IsotopeLabelType := 'L']
```

```{r}
dp.out <- MSstats::dataProcess(mssInput, 
                              MBimpute =  FALSE, 
                              featureSubset = "highQuality", 
                              remove_uninformative_feature_outlier = TRUE)

```

```{r}
fwrite (dp.out$ProteinLevelData, "./output/AB.ProteinLevelData.csv")
fwrite (dp.out$FeatureLevelData, "./output/AB.FeatureLevelData.csv.gz")
```

## MSstats processing QC
----------------

```{r}
p.quant <- fread ("./output/AB.ProteinLevelData.csv")
```

boxplots of normalised intensities

```{r}
g <- ggplot(p.quant, aes (x= interaction(SUBJECT, GROUP), y = LogIntensities, fill = GROUP)) +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values=col.pal) +
  theme(axis.text.x = element_text(angle=90))
g
BackupAsPDF(g, 'protein.intensities.boxplot')
```

Protein Counts per sample

```{r}

g <- ggplot(p.quant[,.N,by=.(GROUP,SUBJECT)], aes(x=SUBJECT, y = N, fill = GROUP)) +
  geom_bar(stat='Identity') +
  theme_classic() +
  scale_fill_manual(values=col.pal) +
  theme(axis.text.x = element_text(angle=90))
g
BackupAsPDF(g, 'nProteins.barplot')
```

# pca of the protein summarised data
 
```{r}
prot.mat <- dcast (p.quant, Protein~SUBJECT, value.var = "LogIntensities") |> as.matrix (rownames = "Protein")
colnames(prot.mat)

prot.mat <- prot.mat[complete.cases(prot.mat),]
```


```{r}
pcaOut <- prcomp(t(prot.mat))

colInfo <- data.table(colname = colnames(prot.mat))
colInfo[,c("status", "rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]

title <- "PCA"

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
pcaDT[, `:=`(timepoint = ifelse(grepl("[47]D", rn), str_extract(rn, "[47]D"), 'na'),
             treat = gsub("[_.].+", "", rn)
             )]

#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = status, shape = rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
#BackupAsPDF(p, "PCA_Complete_Proteins")
```

variance explained:

```{r}
q <- qplot(1:10,  pcaPercentVar[1:10]) +
  geom_line() +
  xlab('Princpal Component 1:10') +
  ylab(('% Variance explained')) +
  ggtitle('Mock + Day7 samples scree plot')

#BackupAsPDF(q, 'screeplot.')
```

Look at the PCAs of the timepoints seperately

Day 4
```{r}
prot.mat <- dcast (p.quant, Protein~interaction(SUBJECT), value.var = "LogIntensities") |> as.matrix (rownames = "Protein")
prot.mat <- prot.mat[complete.cases(prot.mat),]


prot.mat <- prot.mat[, grepl("Mock|4D", colnames(prot.mat))]
colnames(prot.mat)


pcaOut <- prcomp(t(prot.mat))

colInfo <- data.table(colname = colnames(prot.mat))
colInfo[,c("status", "rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]

title <- "PCA"

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)

#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = status, shape = rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
#BackupAsPDF(p, "completeProteins.day4.pca.RunID.")


q <- qplot(1:10,  pcaPercentVar[1:10]) +
  geom_line() +
  xlab('Princpal Component 1:10') +
  ylab(('% Variance explained')) +
  ggtitle('Mock + Day4 samples scree plot')

BackupAsPDF(q, 'screeplot.Day4plusMock.')


#BackupAsPDF(p, "completeProteins.day4.pca.RunID.")

```

day7
```{r}

prot.mat <- dcast (p.quant, Protein~SUBJECT, value.var = "LogIntensities") |> as.matrix (rownames = "Protein")
prot.mat <- prot.mat[complete.cases(prot.mat),]

prot.mat <- prot.mat[, grepl("Mock|7D", colnames(prot.mat))]
colnames(prot.mat)


pcaOut <- prcomp(t(prot.mat))

colInfo <- data.table(colname = colnames(prot.mat))
colInfo[,c("status", "rep") := tstrsplit(colname, "[.]", keep = c(1,2)) ]

title <- "PCA"

#PCA
pcaDT <- as.data.table(pcaOut$x, keep.rownames=TRUE)
pcaPercentVar <- round(100 * (pcaOut$sdev^2)/sum(pcaOut$sdev^2), 1)
pcaDT <- merge (pcaDT, colInfo, by.x = "rn", by.y = "colname", all.x = TRUE)
#plot first two components
p <- ggplot (pcaDT, aes(x=PC1, y=PC2,  fill = status, shape = rep)) + 
  geom_point(alpha=1.0, size=4) + 
  ggrepel::geom_text_repel(aes(label=rn), show.legend = FALSE, size = 3) +
  theme_bw() + 
  xlab (sprintf ("PC1, %.1f%%", pcaPercentVar[1])) + 
  ylab (sprintf ("PC2, %.1f%%", pcaPercentVar[2])) + 
  ggtitle (sprintf ("PCA %s using %d features (log intensity)", title, nrow(pcaOut))) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_shape_manual(values = 21:25) +
  #scale_fill_manual(values = c(`05` = "gray", `30` = "black")) +
  guides(fill = guide_legend(override.aes = list(shape =21) ) ,
         color = guide_legend(override.aes = list(shape =21) ) )

p
#BackupAsPDF(p, "completeProteins.day7.pca.")

q <- qplot(1:10,  pcaPercentVar[1:10]) +
  geom_line() +
  xlab('Princpal Component 1:10') +
  ylab(('% Variance explained')) +
  ggtitle('Mock + Day7 samples scree plot')

#BackupAsPDF(q, 'screeplot.Day7plusMock.')

```

sample clusteting plot 

```{r}
prot.mat <- dcast (p.quant, Protein~SUBJECT, value.var = "LogIntensities") |> as.matrix (rownames = "Protein")

dim(prot.mat)
dim(prot.mat[complete.cases(prot.mat),])
colnames(prot.mat)

# create annotation
ann <- data.table(replicate=str_extract(colnames(prot.mat),'[1-5]$'),
                  treatment=sub("[_.].+", "", colnames(prot.mat)),
                  timepoint=ifelse(grepl("[47]D", colnames(prot.mat)), str_extract(colnames(prot.mat), "[47]D"), 'na'))

colours <- list('timepoint' = c('na' = '#4477AA', '4D'="#228833", '7D'="#CCBB44"),
                'replicate' = c('1' = '#4477AA', '2'='#66CCEE', '3'="#228833", '4'="#CCBB44", '5'='#EE6677'),
                'treatment' =c('Mock' = col.pal[1], 'Vehicle'=col.pal[2], 'N1062D'=col.pal[3], 'x3769'=col.pal[4], 'x4052'=col.pal[5], 'x4206'=col.pal[6])
                )

colAnn <- HeatmapAnnotation(df = ann, col = colours)

hm <- Heatmap(prot.mat, cluster_rows=T, show_row_names = F, cluster_columns = T, top_annotation = colAnn, name='LogIntensities', column_names_gp = gpar(fontsize=6))
#BackupAsPDF(hm, 'proteins.clustered.heatmap')

submat <- sweep(prot.mat,1, STATS=apply(prot.mat, 1, median, na.rm=T))
hm <- Heatmap(submat, cluster_rows=T, show_row_names = F, cluster_columns = T, top_annotation = colAnn, name='LogIntensities/Median',  column_names_gp = gpar(fontsize=6))
#BackupAsPDF(hm, 'proteins.clustered.medianscaled.heatmap')
```

MSstats differential expression analysis

```{r}
```


```{r}
#read back feature data and remove the poor quality sample
f.quant <- fread("./output/AB.FeatureLevelData.csv.gz")

# add batch to SUBJECT header and include in the design matrix
p.quant[, SUBJECT := interaction(GROUP,SUBJECT)]
f.quant[, SUBJECT := interaction(GROUP,SUBJECT)]

dp.out$FeatureLevelData <- f.quant
dp.out$ProteinLevelData <- p.quant

# run msstats correcting for batch 
mss <- groupComparison(contrast.matrix=dp.contrasts, data=dp.out)
mss.dt <- setDT(mss$ComparisonResult)

fwrite(mss.dt, './output/mss.contrasts.ab.csv')
```

